### 2012-10-26 doc/rotation/abstract.tex
Added 'rotation' project document

\begin{abstract}
This is the abstract for this article.
\end{abstract}

### 2012-10-26 doc/rotation/perturbation-flow.tex
Added 'rotation' project document

\section{Perturbation Flow Diagrams}
\label{sec:perturbation-flow-diagrams}

Each of the spectral algorithms summarised in this report perform
a sequence of basic linear algebraic operations like a matrix inverse or
SVD transformations to arrive solve a matrix equation. These algorithms
can be represented as a flow charts, a representation that makes it
easier to trace the dependencies in the algorithm and more importantly
the manner in which error is propagated.

\tableref{tab:operators} lists the basic building blocks of these
algorithmic flow charts, along with absolute and relative errors of the
output {\em to first order}. The exact bounds and proofs for these can
be found in \appendixref{sec:perturbation-theory}.

\begin{table}[htbp]
\floatconts
  {tab:operators}%
  {\caption{Upper Bounds on Absolute Error of Basic Operators and their
  Diagrammatic Representation}}%
  {\begin{tabular}{l | c | c }
  \toprule
  \bfseries Operation & \bfseries Diagram & \bfseries Abs. Error (Upper Bound) \\
  \midrule
  $A+B$         & \pdiagadd & $\aerr{A} + \aerr{B}$ \\
  $A*B$         & \pdiagmul & $\aerr{A} \|B\| + \|A\| \aerr{B} $ \\
  $A^{-1}$      & \pdiaginv & $\|A^{-1}\| \frac{\cnd{A}}{\|A\| - \aerr{A} \cnd{A}} \aerr{A}$ \\
  $\lambda(A)$  & \pdiageig & $ (2n-1) \cnd{A} \aerr{A}$ \\
  $\lambda(A_1, \cdots, A_{n})$  & \pdiageign & $O(n^3 \cnd{R}^2 \|\inv{R}\|^2 \inv{\gap{\lambda}}) \aerr{A}$ \\
  $\xi(A)$      & \pdiagevec & $n^3 \cnd{R} \|\inv{R}\| \gap{\lambda} \aerr{A}$ \\
  $v(A)$      & \pdiagsvec & $\frac{2 \sqrt{k} \aerr{A}}{\gap(A) - \aerr(A)}$ \\
  $\hat{U}^T A \hat{V}$     & \pdiagrot & $\aerr{A}$ \\
  $W(A)$        & \pdiagwhite & $\frac{2}{\sigma_k(A)^2} (\frac{\alpha_A}{1-\alpha_A})$ \\
  $T[A,A,A]$    &             & $\|\Ap\|\aerr{T} + \|T\| (\|A\|^2 + \|A\|\|\Ap\| + \|\Ap\|^2 ) \aerr{W} $ \\
  \bottomrule
  \end{tabular}
  }
\end{table}

\begin{table}[htbp]
\floatconts
  {tab:rel-err}%
  {\caption{Relative Errors of Basic Operators}}%
  {\begin{tabular}{l | c | c c | l}
  \toprule
  \bfseries Operation & \bfseries Diagram & \bfseries Rel. Err. \\
  \midrule
  $A+B$         & \pdiagadd & $\frac{\|A\|\rerr{A} + \|B\|\rerr{B}}{\|A+B\|}$ \\
  $A*B$         & \pdiagmul & $\frac{\|A\|\|B\|}{\|AB\|} (\rerr{A} + \rerr{B})$ \\
  $A^{-1}$      & \pdiaginv & $\frac{\cnd{A}}{1 - \rerr{A}\cnd{A}} \rerr{A}$ \\
  $\lambda(A)$  & \pdiageig & $(2n-1) \cnd{A} \rerr{A}$ \\
  $\lambda(A_1, \cdots, A_{n})$  & \pdiageign & $O(n^3 \cnd{R}^2 \|\inv{R}\|^2 \inv{\gap{\lambda}}) \rerr{A}$ \\
  $\xi(A)$      & \pdiagevec & $n^3 \cnd{R} \|\inv{R}\| \gap{\lambda} \rerr{A}$ \\
  $\nu(A)$      & \pdiagsvec & $\rerr{A} + \rerr{B}$ \\
  $\hat{U}^T A \hat{V}$     & \pdiagrot & $\rerr{A}$ \\
  $W(A)$        & \pdiagwhite & \\
  $T[A,A,A]$    &             & \\
  \bottomrule
  \end{tabular}
  }
\end{table}

Describe the flow diagrams for each model and the computed complexity. (3hrs)

### 2012-11-17 doc/rotation/abstract.tex
Added a rough introduction and fixed some typos

@@
 \begin{abstract}
-This is the abstract for this article.
+
+There has been a series of results recently on recovering hidden parameters
+(inference) from probabilistic models using the method of moments. While this
+problem is in general hard, use of certain spectral techniques like
+eigen-decomposition have resulted in tractable inference using only the first
+few moments (3-4) of the data. This article will survey these results, attempt
+to describe the approach in general and discuss possible issues with their
+generality.
+
 \end{abstract}

### 2012-11-17 doc/rotation/perturbation-flow.tex
Added a rough introduction and fixed some typos

@@
 \tableref{tab:operators} lists the basic building blocks of these
-algorithmic flow charts, along with absolute and relative errors of the
-output {\em to first order}. The exact bounds and proofs for these can
-be found in \appendixref{sec:perturbation-theory}.
+algorithmic flow charts, along with absolute errors of the output {\em
+to first order}. The exact bounds and proofs for these can be found in
+\appendixref{sec:perturbation-theory}.
 
@@
 
-\begin{table}[htbp]
-\floatconts
-  {tab:rel-err}%
-  {\caption{Relative Errors of Basic Operators}}%
-  {\begin{tabular}{l | c | c c | l}
-  \toprule
-  \bfseries Operation & \bfseries Diagram & \bfseries Rel. Err. \\
-  \midrule
-  $A+B$         & \pdiagadd & $\frac{\|A\|\rerr{A} + \|B\|\rerr{B}}{\|A+B\|}$ \\
-  $A*B$         & \pdiagmul & $\frac{\|A\|\|B\|}{\|AB\|} (\rerr{A} + \rerr{B})$ \\
-  $A^{-1}$      & \pdiaginv & $\frac{\cnd{A}}{1 - \rerr{A}\cnd{A}} \rerr{A}$ \\
-  $\lambda(A)$  & \pdiageig & $(2n-1) \cnd{A} \rerr{A}$ \\
-  $\lambda(A_1, \cdots, A_{n})$  & \pdiageign & $O(n^3 \cnd{R}^2 \|\inv{R}\|^2 \inv{\gap{\lambda}}) \rerr{A}$ \\
-  $\xi(A)$      & \pdiagevec & $n^3 \cnd{R} \|\inv{R}\| \gap{\lambda} \rerr{A}$ \\
-  $\nu(A)$      & \pdiagsvec & $\rerr{A} + \rerr{B}$ \\
-  $\hat{U}^T A \hat{V}$     & \pdiagrot & $\rerr{A}$ \\
-  $W(A)$        & \pdiagwhite & \\
-  $T[A,A,A]$    &             & \\
-  \bottomrule
-  \end{tabular}
-  }
-\end{table}
-
-Describe the flow diagrams for each model and the computed complexity. (3hrs)
+% \begin{table}[htbp]
+% \floatconts
+%   {tab:rel-err}%
+%   {\caption{Relative Errors of Basic Operators}}%
+%   {\begin{tabular}{l | c | c c | l}
+%   \toprule
+%   \bfseries Operation & \bfseries Diagram & \bfseries Rel. Err. \\
+%   \midrule
+%   $A+B$         & \pdiagadd & $\frac{\|A\|\rerr{A} + \|B\|\rerr{B}}{\|A+B\|}$ \\
+%   $A*B$         & \pdiagmul & $\frac{\|A\|\|B\|}{\|AB\|} (\rerr{A} + \rerr{B})$ \\
+%   $A^{-1}$      & \pdiaginv & $\frac{\cnd{A}}{1 - \rerr{A}\cnd{A}} \rerr{A}$ \\
+%   $\lambda(A)$  & \pdiageig & $(2n-1) \cnd{A} \rerr{A}$ \\
+%   $\lambda(A_1, \cdots, A_{n})$  & \pdiageign & $O(n^3 \cnd{R}^2 \|\inv{R}\|^2 \inv{\gap{\lambda}}) \rerr{A}$ \\
+%   $\xi(A)$      & \pdiagevec & $n^3 \cnd{R} \|\inv{R}\| \gap{\lambda} \rerr{A}$ \\
+%   $\nu(A)$      & \pdiagsvec & $\rerr{A} + \rerr{B}$ \\
+%   $\hat{U}^T A \hat{V}$     & \pdiagrot & $\rerr{A}$ \\
+%   $W(A)$        & \pdiagwhite & \\
+%   $T[A,A,A]$    &             & \\
+%   \bottomrule
+%   \end{tabular}
+%   }
+% \end{table}

### 2013-01-17 doc/icml2013/abstract.tex
Added ICML 2013 draft

\begin{abstract} 
  Spectral algorithms for latent variable models have seen considerable
  recent interest for being efficient consistent estimators of model
  parameters. These algorithms make few if any assumptions about the
  generative process of the data, while providing a polynomial sample
  and computational complexity. We present a new spectral algorithm for
  a discriminative model, a mixture of linear regressions and show that
  it can recover the regression coefficients with similar polynomial
  guarantees. We evaluate the algorithm on linearly and non-linearly
  generated data, as well as on a motion tracking task and compare it's
  characteristics with an E-M algorithm.
\end{abstract}

### 2013-01-17 doc/icml2013/algo.tex
Added ICML 2013 draft

\section{Algorithm}
\label{sec:algo}

Recovery of $B$, $\mathcal{B}$.

Use AHK2012

Lemma error of $B$.

Lemma AHK2012 error

Theorem on polynomial recovery

### 2013-01-17 doc/icml2013/discussion.tex
Added ICML 2013 draft

\section{Discussion}
\label{sec:discussion}

Role of spectral algorithms in discriminative tasks.

Challenges and ways forward?

### 2013-01-17 doc/icml2013/evaluation.tex
Added ICML 2013 draft

\section{Evaluation}
\label{sec:evaluation}

Mixture of linear regressions,

linear 
quadratic
motion tracking data

### 2013-01-17 doc/icml2013/intro.tex
Added ICML 2013 draft

\section{Introduction}
\label{sec:intro}

Local optima are a common problem faced when learning parameters in
latent variable models. Recently, there has been an interest in spectral
algorithms that provide statistically consistent estimators, i.e. given
sufficient data they are guaranteed to converge to the global optimum,
given that the model is identifiable. Furthermore, these algorithms have
polynomial sample and computational complexities for several such
models, e.g.  hidden Markov models \cite{anandkumar-hsu-kakade2012},
probabilistic context free grammars \cite{hsu-kakade-liang2012}, etc. In
this work, we provide a spectral algorithm for a discriminative model,
a mixture of linear regressions.

\todo{Where are mixtures of experts used? Why are they interesting?}

Our approach uses regression to recover the moments of the data from the
scalar observable output variable. Once we recover the second and third
moments of the regression coefficients, we use a slight modification of
the method outlined in \citet{anandkumar-hsu-kakade2012}.

We evaluate our algorithm on simulated linear and non-linear data to
understand how well the algorithm scales with the number of components,
dimensions and separation. To test how the algorithms perform on a more
real dataset, we study their performance on a motion tracking dataset.

In \autoref{sec:model}, we formally define the mixture of linear
regressions model and discuss when it is identifiable. We describe our
algorithm and provide some theoretical results in \autoref{sec:algo}. We
present the results of our evaluations in \autoref{sec:evaluation}.
Finally, we discuss the merits and demerits of spectral algorithms in
discriminative models and conclude in \autoref{sec:discussion}. The
details of some proofs can be found in \autoref{sec:proofs}.

### 2013-01-17 doc/icml2013/model.tex
Added ICML 2013 draft

\section{Model}
\label{sec:model}

Mixture of Experts

\subsection{Identifiability}

### 2013-02-06 java/src/learning/linalg/FullTensor.java
Added tensor power-method code

@@ -192,11 +192,19 @@ public class FullTensor implements Tensor {
    * @return
    */
   public FullTensor rotate(SimpleMatrix M1, SimpleMatrix M2, SimpleMatrix M3) {
-    FullTensor Y = this;
-    if(M1 != null) Y = Y.rotate(0, M1);
-    if(M2 != null) Y = Y.rotate(1, M1);
-    if(M3 != null) Y = Y.rotate(2, M1);
-    return Y;
+    double[][][] Y = X.clone();
+    for(int d1 = 0; d1 < D1; d1++) {
+      for(int d2 = 0; d2 < D2; d2++) {
+        for(int d3 = 0; d3 < D3; d3++) {
+          Y[d1][d2][d3] = 0;
+          for(int l1 = 0; l1 < D1; l1++)
+            for(int l2 = 0; l2 < D2; l2++)
+              for(int l3 = 0; l3 < D3; l3++)
+                Y[d1][d2][d3] += X[l1][l2][l3] * M1.get(l1, d1) * M2.get(l2, d2) * M3.get(l3, d3);
+        }
+      }
+    }
+    return new FullTensor(Y);
   }
 
   @Override

### 2013-02-06 java/src/learning/linalg/MatrixOps.java

@@ -714,6 +714,27 @@ public class MatrixOps {
   }
 
   /**
+   * Get the sqrt of a diagonal matrix
+   * @param X
+   * @return
+   */
+  public static SimpleMatrix sqrt( SimpleMatrix X ) {
+    double[] diagonal = MatrixFactory.toVector(X.extractDiag());
+    for(int i = 0; i < diagonal.length; i++) {
+      diagonal[i] = Math.sqrt(diagonal[i]);
+    }
+    return MatrixFactory.diag(MatrixFactory.fromVector(diagonal));
+  }
+
+
+  public static SimpleMatrix whitener( SimpleMatrix X ) {
+    SimpleSVD UWV = X.svd();
+    SimpleMatrix U = UWV.getU();
+    SimpleMatrix W = UWV.getW();
+    return U.mult( sqrt( W ).invert() );
+  }
+
+  /**
    * Align the rows of matrix X so that the rows/columns are matched with the
    * columns of Y
    */

### 2013-02-06 java/src/learning/linalg/TensorFactorization.java

@@ -1,0 +1,131 @@
+package learning.linalg;
+
+import org.ejml.simple.SimpleMatrix;
+import org.ejml.simple.SimpleSVD;
+import org.javatuples.Pair;
+
+/**
+ * Implements Anandkumar/Ge/Hsu/Kakade/Telgarsky, 2012.
+ * Method of moments for learning latent-variable models.
+ */
+public class TensorFactorization {
+  static final double EPS_CLOSE = 1e-10;
+
+  /**
+   * Perform a single eigen-decomposition step
+   * @param T
+   * @return
+   */
+  protected static Pair<Double,SimpleMatrix> eigendecomposeStep( Tensor T, int iters, int attempts ) {
+    int N = iters;
+    int D = T.getDim(0);
+
+    double maxEigenvalue = Double.NEGATIVE_INFINITY;
+    SimpleMatrix maxEigenvector = null;
+
+    for( int attempt = 0; attempt < attempts; attempt++ ) {
+      SimpleMatrix theta = RandomFactory.randn(1, D);
+      theta.scale(1.0/MatrixOps.norm(theta));
+
+      // Hit the tensor with this vector and repeat till you converge
+      for(int n = 0; n < N; n++ ) {
+        SimpleMatrix theta_ = T.project2(1, 2, theta, theta);
+        // Normalize
+        theta_ = theta_.scale(1.0 / MatrixOps.norm(theta_));
+        if( MatrixOps.norm(theta_.minus(theta)) < EPS_CLOSE ) {
+          theta = theta_; break;
+        } else {
+          theta = theta_;
+        }
+      }
+      double eigenvalue = T.project3(theta, theta, theta);
+      if(eigenvalue > maxEigenvalue) {
+        maxEigenvalue = eigenvalue;
+        maxEigenvector = theta;
+      }
+    }
+
+    return new Pair<>(maxEigenvalue, maxEigenvector);
+  }
+
+  /**
+   * Return T - scale vector^{\otimes 3}
+   * "Efficient" inplace deflation
+   * @param scale
+   * @param v
+   * @return
+   */
+  protected static void deflate(FullTensor T, double scale, double[] v) {
+    int D = T.getDim(0);
+    for( int d1 = 0; d1 < D; d1++ )
+      for( int d2 = 0; d2 < D; d2++ )
+        for( int d3 = 0; d3 < D; d3++ )
[…]

### 2013-02-06 java/tests/learning/linalg/MatrixOpsTest.java

@@ -448,4 +448,20 @@ public class MatrixOpsTest {
     Assert.assertTrue( MatrixOps.allclose( X, Y ) );
     Assert.assertTrue( MatrixOps.allclose( X.transpose(), Z ) );
   }
+
+  @Test
+  public void whitener() {
+    double[][] X_ = {
+            { 0.2740856 ,  0.61844862,  0.84960229},
+            { 0.76229792,  0.78220896, -0.32681905},
+            { 0.58632667, -0.07530232,  0.41396288}};
+    SimpleMatrix X = new SimpleMatrix(X_);
+    SimpleMatrix Y = X.mult(X.transpose());
+
+    SimpleMatrix W = MatrixOps.whitener(Y);
+    SimpleMatrix z = W.mult(X);
+    SimpleMatrix Z = z.mult(z.transpose());
+
+    Assert.assertTrue( MatrixOps.allclose( W.transpose().mult(Y).mult(W), SimpleMatrix.identity(3) ) );
+  }
 }

### 2013-02-06 java/tests/learning/linalg/TensorFactorizationTest.java

@@ -1,0 +1,116 @@
+package learning.linalg;
+
+import junit.framework.Assert;
+import org.ejml.simple.SimpleMatrix;
+import org.javatuples.Pair;
+import org.junit.Test;
+
+/**
+ * Test the tensor factorization methods
+ */
+public class TensorFactorizationTest {
+
+  /**
+   * Test a single step of the eigen decomposition
+   */
+  @Test
+  public void eigendecompositionStep() {
+    // Construct a tensor to have just one principal direction
+    SimpleMatrix x = MatrixFactory.fromVector( new double[] {1.0, 0.0, 0.0});
+
+    FullTensor T = FullTensor.fromUnitVector(x);
+    Pair<Double, SimpleMatrix> pair = TensorFactorization.eigendecomposeStep(T, 10, 1);
+    double value = pair.getValue0();
+    SimpleMatrix y = pair.getValue1();
+    Assert.assertEquals(value, 1.0);
+    Assert.assertTrue(MatrixOps.allclose(x, y));
+
+    // Construct a tensor with two principal directions
+    SimpleMatrix w = new SimpleMatrix( new double[][] {{ 1.1, 0.8, 0.7 }} );
+    SimpleMatrix X = new SimpleMatrix( new double[][] {
+            {1.0, 0.0, 0.0},
+            {0.0, 1.0, 0.0},
+            {0.0, 0.0, 1.0}}
+            );
+
+    T = FullTensor.fromDecomposition(w, X);
+    pair = TensorFactorization.eigendecomposeStep(T, 10, 10);
+    value = pair.getValue0();
+    y = pair.getValue1();
+    Assert.assertEquals(value, w.get(0));
+    Assert.assertTrue(MatrixOps.allclose(MatrixOps.col(X,0), y));
+  }
+
+  @Test
+  public void eigendecompositionIdentity() {
+    // Construct a tensor with two principal directions
+    SimpleMatrix w = new SimpleMatrix( new double[][] {{ 1.1, 0.8, 0.7 }} );
+    SimpleMatrix X = new SimpleMatrix( new double[][] {
+            {1.0, 0.0, 0.0},
+            {0.0, 1.0, 0.0},
+            {0.0, 0.0, 1.0}}
+    );
+
+    FullTensor T = FullTensor.fromDecomposition(w, X);
[…]

### 2013-02-07 doc/icml2013/intro.tex
Updated the outline for the paper

@@
 
-Local optima are a common problem faced when learning parameters in
+Local optima are a common problem plaguing the learning of parameters in
 latent variable models. Recently, there has been an interest in spectral
@@
 polynomial sample and computational complexities for several such
-models, e.g.  hidden Markov models \cite{anandkumar-hsu-kakade2012},
-probabilistic context free grammars \cite{hsu-kakade-liang2012}, etc. In
+models, e.g.  hidden Markov models \cite{AnandkumarHsuKakade2012},
+probabilistic context free grammars \cite{HsuKakadeLiang2012}, etc. In
 this work, we provide a spectral algorithm for a discriminative model,
-a mixture of linear regressions.
+the mixture of linear regressions.
 
@@
 moments of the regression coefficients, we use a slight modification of
-the method outlined in \citet{anandkumar-hsu-kakade2012}.
+the method outlined in \citet{AnandkumarHsuKakade2012}.
 
@@
 
-In \autoref{sec:model}, we formally define the mixture of linear
+In \sectionref{sec:model}, we formally define the mixture of linear
 regressions model and discuss when it is identifiable. We describe our
-algorithm and provide some theoretical results in \autoref{sec:algo}. We
-present the results of our evaluations in \autoref{sec:evaluation}.
+algorithm and provide some theoretical results in \sectionref{sec:algo}. We
+present the results of our evaluations in \sectionref{sec:evaluation}.
 Finally, we discuss the merits and demerits of spectral algorithms in
-discriminative models and conclude in \autoref{sec:discussion}. The
-details of some proofs can be found in \autoref{sec:proofs}.
+discriminative models and conclude in \sectionref{sec:discussion}. The
+details of some proofs can be found in \sectionref{sec:proofs}.

### 2013-02-08 java/src/learning/em/MixtureOfExperts.java
Log intermediate $\beta$ for EM

@@ -240,6 +240,7 @@ public class MixtureOfExperts implements Runnable {
       // epsilion.
       if( Math.abs(lhood - old_lhood) < eps )
         break;
+      Execution.putOutput( "betas" + i, (new SimpleMatrix( state.betas )).transpose() );
 
       old_lhood = lhood;
       old_state = state;
@@ -291,7 +292,10 @@ public class MixtureOfExperts implements Runnable {
       betas_ = MatrixOps.alignMatrix( betas_, betas, true );
       Execution.putOutput( "betas", betas );
       Execution.putOutput( "betas_", betas_ );
+
       double err = MatrixOps.norm( betas.minus( betas_ ) );
+      Execution.putOutput( "betasErr", err );
+
       System.out.printf( "%.4f\n", err );
     } catch( ClassNotFoundException | IOException e ) {
       LogInfo.error( e.getMessage() );

### 2013-02-08 doc/icml2013/algo.tex

@@
 
-Recovery of $B$, $\mathcal{B}$.
+An overview of the algorithm is as follows; use low-rank/ridge regression recovery techniques to recover the $B$ and $\mathcal{B}$ from moments of the data.
+Use the tensor factorization to recover the $\pi$, $\beta$ which are eigenvalues/eigenvectors.
 
-Use AHK2012
+\begin{theorem}
+  Recovery analysis of mixture of linear regressions.
+\end{theorem}
+\begin{proof}
+\end{proof}
 
-Lemma error of $B$.
+\subsection{Recovering $B$ and $\mathcal{B}$}
 
-Lemma AHK2012 error
+Writing out the regression problem.
 
-Theorem on polynomial recovery
+\begin{lemma}
+  Recovery analysis of ridge regression is $\le \Omega(\frac{d^2}{n})$.
+  Try and figure out dependence on condition numbers. It's similarly
+  $\le \Omega(\frac{d^3}{n})$.
+\end{lemma}
+\begin{proof}
+  Theorem X \cite{HsuKakadeZhang2011}.
+\end{proof}
+
+Using low-rank gradient methods. Overview SVD for
+
+\begin{lemma}
+  Recovery analysis of nuclear-norm regression is $\le
+  \Omega(\frac{kd}{n})$. Try and figure out dependence on condition
+  numbers.
+\end{lemma}
+\begin{proof}
+  Theorem X \cite{NegahbanWainwright2009}.
+  Theorem Y \cite{Tomioka2011}.
+\end{proof}
+
+\subsection{Recovering Regression Coefficients}
+
+Write out the form of $B$ and $\mathcal{B}$ in symmetric tensor form.
+
+Use Tensor Power Method to recovery $\beta_k$.
+
+\begin{lemma}
+  Bound error in the $\beta$ based on the errors of $B$ and $\mathcal{B}$.
+\end{lemma}
+\begin{proof}
+  \cite{AnandkumarGeHsu2012}
+\end{proof}

### 2013-02-08 doc/icml2013/model.tex
minor changes

@@
 
-Mixture of Experts
+\newcommand{\xn}[1]{x^{(#1)}}
+\newcommand{\xni}{\xn{i}}
+\newcommand{\yn}[1]{y^{(#1)}}
+\newcommand{\yni}{\yn{i}}
+
+The mixture of linear regressions model describes response variables $y
+\in \Re$ dependent on covariates $x \in \Re^d$ through $K$ different
+modes. In particular, each mode occurs with a fixed probability of
+$\pi_k$, and each mode is independently described by regression
+coefficients $\beta_k$, with a common variance $\sigma^2$. Formally, the
+generative process is,
+\begin{eqnarray}
+  z &\sim& \mult(\pi) \\
+  y | z = k &\sim& \normal{ \beta_{k}^T x, \sigma^2 }.
+\end{eqnarray}
+
+We would like to recover $B = [\beta_1 | \beta_2 | \cdots | \beta_K]$
+and the mixture probabilities $\pi = [\pi_1 | \pi_2 | \cdots | \pi_K]$.
+
+\subsection{Non-linearities}
+
+Approach to handle non-linearities.
+
+Note about dependencies in the n-th moments.
 
 \subsection{Identifiability}
+
+The model is in general identifiable, except for a pathological case
+with how data points are placed.
+
+Plot of the above scenario.
+
+Interesting question; how do these extend in the non-linear case (which
+would be ``new'').

### 2013-02-10 java/src/learning/Misc.java
Added serialVersionID's to classes and added a reader that can ignore version incompatibilities and run

@@ -6,6 +6,13 @@
 package learning;
 
 import java.util.Comparator;
+import java.io.IOException;
+import java.io.InputStream;
+import java.io.InvalidClassException;
+import java.io.ObjectInputStream;
+import java.io.ObjectStreamClass;
+
+import fig.basic.LogInfo;
 
 /**
  * Miscellaneous utilities
@@ -150,6 +157,40 @@ public class Misc {
   public static void traverseChoices( int N, int K, TraversalFunction fn ) {
     int[] accum = new int[ N ];
     traverseChoices( N, K, 0, accum, fn );
+  }
+
+  /**
+   * A hack to read objects of different version numbers
+   */
+  public static class DecompressibleInputStream extends ObjectInputStream {
+      public DecompressibleInputStream(InputStream in) throws IOException {
+          super(in);
+      }
+
+      protected ObjectStreamClass readClassDescriptor() throws IOException, ClassNotFoundException {
+          ObjectStreamClass resultClassDescriptor = super.readClassDescriptor(); // initially streams descriptor
+          Class localClass; // the class in the local JVM that this descriptor represents.
+          try {
+              localClass = Class.forName(resultClassDescriptor.getName()); 
+          } catch (ClassNotFoundException e) {
+              LogInfo.error("No local class for " + resultClassDescriptor.getName(), e);
+              return resultClassDescriptor;
+          }
+          ObjectStreamClass localClassDescriptor = ObjectStreamClass.lookup(localClass);
+          if (localClassDescriptor != null) { // only if class implements serializable
+              final long localSUID = localClassDescriptor.getSerialVersionUID();
+              final long streamSUID = resultClassDescriptor.getSerialVersionUID();
+              if (streamSUID != localSUID) { // check for serialVersionUID mismatch.
+                  final StringBuffer s = new StringBuffer("Overriding serialized class version mismatch: ");
+                  s.append("local serialVersionUID = ").append(localSUID);
+                  s.append(" stream serialVersionUID = ").append(streamSUID);
+                  Exception e = new InvalidClassException(s.toString());
+                  LogInfo.error(e);
+                  resultClassDescriptor = localClassDescriptor; // Use local class descriptor for deserialization
+              }
+          }
+          return resultClassDescriptor;
+      }
   }
 
 }

### 2013-02-10 java/src/learning/data/ProjectedCorpus.java

@@ -27,6 +27,7 @@ import org.ejml.simple.SimpleMatrix;
  * Stores a corpus in an integer array
  */
 public class ProjectedCorpus extends Corpus implements Serializable {
+  private static final long serialVersionUID = 2L;
   public int projectionDim;
   protected long[] seeds;
 

### 2013-02-10 java/src/learning/models/RealHiddenMarkovModel.java

@@ -30,6 +30,8 @@ import java.util.Arrays;
 public class RealHiddenMarkovModel extends HiddenMarkovModel {
   // Maps discrete feature indicies into a feature map
 	public static class Features implements Serializable {
+    private static final long serialVersionUID = 2L;
+
     int emissionCount;
     int dimension;
     public double[][] features; 

### 2013-02-10 java/src/learning/models/transforms/NonLinearity.java

@@ -6,6 +6,8 @@ import java.io.Serializable;
  * Represents a generic non-linear transform of the data
  */
 public class NonLinearity implements Serializable {
+  private static final long serialVersionUID = 2L;
+
   public int getLinearDimension( int dimension  ) {return dimension;}
   public double[] getLinearEmbedding( final double[] x ) {return x;}
   public double[][] getExponents() {return null;}

### 2013-02-10 doc/icml2013/model.tex

@@
 
-We would like to recover $B = [\beta_1 | \beta_2 | \cdots | \beta_K]$
-and the mixture probabilities $\pi = [\pi_1 | \pi_2 | \cdots | \pi_K]$.
+We would like to recover $B = [\beta_1 \mid \beta_2 \mid \cdots \mid \beta_K]$
+and the mixture probabilities $\pi = (\pi_1, \pi_2, \dots, \pi_K)$.
 
-\subsection{Non-linearities}
-
-Approach to handle non-linearities.
-
-Note about dependencies in the n-th moments.
-
-\subsection{Identifiability}
-
-The model is in general identifiable, except for a pathological case
-with how data points are placed.
-
-Plot of the above scenario.
-
-Interesting question; how do these extend in the non-linear case (which
-would be ``new'').
+Learning framework:
+Get $(x^{(1)},y^{(1)})$, ...
+don't observe $h$

### 2013-02-10 python/models/LinearRegressions.py
Initializing the low-rank recovery with ridge regression.

@@ -4,13 +4,13 @@ Generate data from a mixture of linear regressions
 
 import scipy as sc
 import scipy.linalg
-from scipy import array, zeros, ones, eye
+from scipy import array, zeros, ones, eye, rand
 from scipy.linalg import inv
 from models.Model import Model
 from util import chunked_update #, ProgressBar
 
 multinomial = sc.random.multinomial
-multivariate_normal = sc.random.multivariate_normal 
+#multivariate_normal = sc.random.multivariate_normal 
 dirichlet = sc.random.dirichlet
 
 from spectral.rand import permutation, wishart
@@ -65,7 +65,8 @@ class LinearRegressionsMixture( Model ):
             def update( start, stop ):
                 """Sample random vectors and then assign them to X in
                 order"""
-                Z = multivariate_normal( mean, sigma, int(stop - start) )
+                #Z = multivariate_normal( mean, sigma, int(stop - start) )
+                Z = 2*rand( int(stop - start), self.d ) - 1
                 # Insert into X in a shuffled order
                 p = perm[ start:stop ]
                 perm_ = p[ p < n ]

### 2013-02-10 python/optim/PhaseRecovery.py

@@ -26,17 +26,13 @@ class PhaseRecovery( ProximalGradient ):
         $\diff{L}{B_{ij}} = \sum_i (x^i' B x^i - y^i^2) x^i_i x^i_j
         """ 
 
-        Q2 = self.Q2
         d, _ = B.shape
         N = len(y)
 
         # Compute x^T B x - y
         dB = zeros( (d, d) )
         # Forgive me father for I have multiplied two large matrices.
-        if Q2 is None:
-            Z = ( xMy( B, X, X ) - y)
-        else:
-            Z = ( xMy( B, X, X ) - y).dot( Q2 )
+        Z = ( xMy( B, X, X ) - y)
 
         for i in xrange( N ):
             x_i = X[i]
@@ -59,11 +55,8 @@ class PhaseRecovery( ProximalGradient ):
             tot += (x_i.dot( B ).dot( x_i ) - y_i)**2
         return tot/N
 
-    def solve( self, y, X, Q2 = None, *args, **kwargs ):
+    def solve( self, y, X, *args, **kwargs ):
         """Solve using a Q value"""
-        N = len(y)
-
-        self.Q2 = Q2
 
         return ProximalGradient.solve( self, y, X, *args, **kwargs )
 

### 2013-02-10 python/optim/TensorRecovery.py

@@ -25,16 +25,12 @@ class TensorRecovery( ProximalGradient ):
         $\diff{L}{B_{ij}} = \sum_i (B(x^i, x^i, x^i)' - y^i^2) x^i \otimes x^i \otimes x^i 
         """ 
 
-        Q2 = self.Q2
         d = B.shape[0]
         N = len(y)
 
         # Compute x^T B x - y
         dB = zeros( (d, d, d) )
-        if Q2 is None:
-            Z = (Txyz( B, X, X, X ) - y)
-        else:
-            Z = (Txyz( B, X, X, X ) - y).dot( Q2 )
+        Z = (Txyz( B, X, X, X ) - y)
 
         for i in xrange( N ):
             x_i = X[i]
@@ -61,11 +57,8 @@ class TensorRecovery( ProximalGradient ):
             tot += (einsum( "ijk,i,j,k", B, x_i, x_i, x_i) - y_i)**2
         return tot/N
 
-    def solve( self, y, X, Q2 = None, *args, **kwargs ):
+    def solve( self, y, X, *args, **kwargs ):
         """Solve using a Q value"""
-        N = len(y)
-
-        self.Q2 = Q2
 
         return ProximalGradient.solve( self, y, X, *args, **kwargs )
 

### 2013-02-10 python/spectral/LinearRegressions.py

@@ -3,13 +3,14 @@ Quick implementation of the mixture of linear regressions code
 """
 
 import shutil, tempfile 
+import ipdb
 
 import scipy as sc
 import scipy.spatial
 import scipy.linalg
 from scipy import diag, array, ndim, outer, eye, ones,\
         log, sqrt, zeros, floor, exp
-from scipy.linalg import norm, svd, svdvals, eig, eigvals, inv, pinv, cholesky
+from scipy.linalg import norm, svd, svdvals, eig, eigvals, inv, pinv, cholesky, solve
 from scipy.spatial.distance import cdist
 
 permutation, rand, dirichlet = scipy.random.permutation, scipy.random.rand, scipy.random.dirichlet
@@ -24,13 +25,17 @@ from spectral.MultiView import recover_M3
 
 from optim import PhaseRecovery, TensorRecovery
 
-def recover_B( k, y, X, iters = 50, alpha0 = 0.1, Q = None, B2B3 = (None,None) ):
+def recover_B( k, y, X, iters = 50, alpha0 = 0.1, reg = 0, B20B30 = (None,None), B2B3 = (None, None) ):
     """Recover the mixture weights B"""
-    B20, B30 = B2B3
+    B20, B30 = B20B30
+    B2, B3 = B2B3
 
     # Use convex optimisation to recover B2 and B3
-    B2 = PhaseRecovery().solve( y**2, X, Q, B20 = B20, alpha = "1/sqrt(T)", alpha0 = alpha0, iters = iters, reg = 1e-3, verbose = False )
-    B3 = TensorRecovery().solve( y**3, X, Q, B30 = B30, alpha = "1/sqrt(T)", alpha0 = alpha0, iters = iters, reg = 1e-4, verbose = False )
+    B2_ = PhaseRecovery().solve( y**2, X, B0 = B20, alpha = "1/sqrt(T)", alpha0 = alpha0, iters = iters, reg = reg, verbose = False )
+    print norm( B20 - B2 ), norm( B2_ - B2 ) 
+
+    B3_ = TensorRecovery().solve( y**3, X, B0 = B30, alpha = "1/sqrt(T)", alpha0 = alpha0, iters = iters, reg = reg, verbose = False )
+    print norm( B30 - B3 ), norm( B3_ - B3 ) 
 
     B3_ = lambda theta: sc.einsum( 'abj,j->ab', B3, theta )
 
[…]

### 2013-02-10 doc/icml2013/algo.tex

@@
-\section{Algorithm}
+\section{Method of Moments}
 \label{sec:algo}
 
-An overview of the algorithm is as follows; use low-rank/ridge regression recovery techniques to recover the $B$ and $\mathcal{B}$ from moments of the data.
-Use the tensor factorization to recover the $\pi$, $\beta$ which are eigenvalues/eigenvectors.
+\subsection{Identifiability}
+
+The model is in general identifiable, except for a pathological case
+with how data points are placed.
+
+Plot of the above scenario.
+
+Interesting question; how do these extend in the non-linear case (which
+would be ``new'').
+
+Linear independence to give intuition
+
+Quadratic and cubic independence
+
+necessary if only get low lower moments up to $p$
+
+Note about dependencies in the n-th moments.
+
+Method of moments
+argue why naive thing doesn't work.
+
+Note: we haven't shown identifiability from all moments.
+
+\subsection{Algorithm}
+
+If knew $M_2$ and $M_3$, then can use existing techniques.
+
+We can do low-rank regression to recover $M_2$ and $M_3$.
+
+An overview of the algorithm is as follows; use low-rank/ridge regression
+recovery techniques to recover the $B$ and $\mathcal{B}$ from moments of the
+data. Use the tensor factorization to recover the $\pi$, $\beta$ which are
+eigenvalues/eigenvectors.
+
+% State algorithm
 

### 2013-02-11 doc/icml2013/algo.tex
Updated draft; intro, models, algo

--- 2013-01-17 doc/icml2013/algo.tex
@@
-\section{Algorithm}
+\section{A Method of Moments Algorithm}
 \label{sec:algo}
 
-Recovery of $B$, $\mathcal{B}$.
+In this section, we will outline some shortcomings of the standard
+method of moments formulation for the mixture of linear regressions
+model. We then describe how we can work around this limitation and
+reduce our problem to a known one, which can be solved using symmetric
+tensor decomposition. Finally, we will state some implications this
+reduction has regarding identifiability.
 
-Use AHK2012
+\subsection{Notation}
 
-Lemma error of $B$.
+We have already defined $y$ to be the response variable and $x \in
+\Re^d$ to be the covariates it depends on. Note that the matrix of
+regressors $B$ is a $d \times K$ matrix, where the $k$-th column is
+$\beta_k$.
 
-Lemma AHK2012 error
+We will use the notation $x\tp{p}$ to describe the $p$-th order tensor
+formed by taking the outer product of $x$; i.e. $x\tp{p}_{i_1 i_2 \ldots
+i_p} = x_{i_1} x_{i_2} \cdots x_{i_p}$. We will use $\odot$ to denote
+the general dot product between $p$-th order tensors, i.e. $X \odot
+Y = \sum_{i_1, i_2, \ldots i_p} X_{i_1, i_2, \ldots i_p} Y_{i_1, i_2,
+\ldots i_p}.$ Similar to how $x^T A$ is the projection of a vector $x$
+onto the row space of a matrix $A$, when $X \in \Re^{d^p}$ and $Y \in
+\Re^{d^{p+q}}$,  with $p, q > 0$, $X \odot Y \in \Re^{d^q}$ is the
+projection of $X$ onto the first $p$ dimensions of $Y$, i.e. $(X \odot
+Y)_{i_1, \ldots, i_q} = \sum_{j_1, \ldots, j_p} X_{j_1, \ldots, j_p}
+Y_{j_1, \ldots, j_p, i_1, \ldots, i_q}$.
 
-Theorem on polynomial recovery
+The expected moments of the parameters $\beta_1, \dots, \beta_k$ play an
+important role in our algorithm and we will use $M_p = \E[\beta\tp{p}]
+= \sum_{k=1}^{K} \pi_k \beta_k^{\otimes p}$ to denote them.
+
+\subsection{Identifiability}
+
+The model is in general identifiable, except for a pathological case
+concerning how data points are placed \citationneeded.
+
+\subsection{Overview}
+
+Conventionally, when using the method of moments, the moments of
+observed variables are expressed algebraically in terms of the model
+parameters. Using these equations, we attempt to solve for the model
+parameters using empirical estimates of the moments.
+
+In the case of the mixture of linear regressions, we have two observed
+quantities, $x$ and $y$. Unfortunately, because $y$ is a scalar
+quantity, we have only one equation for every moment of $y$, and with
+this naive approach, would have to consider $O( (K+1) D)$ moments to
+recover $B \in \Re{K \times D}$ and $\pi \in \Re^K$.
+
+% To keep the discussion simple, we will consider the case where there is
+% no noise, i.e.  $\sigma^2 = 0$. In this case, the moments of $y$ are,
+% \begin{eqnarray*}
+%   \E[ y ] &=& (\sum_{k=1}^{K} \pi_{k} \beta_k)^T x \\
+%           &=& M_1^T \E[x] \\
+%   \E[ y^2 ] &=& \sum_{k=1}^{K} \pi_{k} \E[(\beta_k^T x)]^2 \\
+%   &=& \sum_{k=1}^{K} \pi_{k} (\beta_k\tp{2} \odot \E[x\tp{2}]) \\
+%   &=& M_2 \odot \E[x\tp{2}] \\
+%   \E[ y^p ] &=& M_p \odot \E[x\tp{p}].
+% \end{eqnarray*}
+% 
+% We note that joint moments with $x$ will not help either; consider
+% \begin{eqnarray}
+%   \E[ y^p x^q ] &=& M_p \odot E[ x\tp{p+q} ].
+% \end{eqnarray}
+% The elements of $\E[ y^p x^q ]$ are linear combinations of $\E[y^p]$,
+% and hence we do not get any new equations; \todo{This isn't exactly
+% correct; should we even be presenting this line of reasoning?}. This is
+% to be expected, since any information about the $\beta$s is really
+% encoded in $y$, and not in the $X$.
+
+To work around this problem, we will use regression on the conditional
+moments, $\E[y | x]$ to recover the expected moments of the regressors,
+$M_2$ and $M_3$. Recovering the $\beta_k$ and $\pi_k$ given these
+symmetric moments has been addressed in prior work
+\citep{AnandkumarHsuKakade2012, AnandkumarGeHsu2012}.  This algorithm as
+been summarized in \algorithmref{algo:spectral-experts}.
+
+\begin{algorithm}[t]
+  \caption{Spectral Experts}
+  \label{algo:spectral-experts}
+  \begin{algorithmic}[1]
+    \REQUIRE $\mathcal{D} = \{ (\xn{1}, \yn{1}), \cdots, (\xn{N}, \yn{N}) \}$
+    \REQUIRE $K$, the number of clusters, $\sigma^2$ 
+    \STATE Recover $M_1$ using least-squares regression on $\mathcal{D}$.
+    \STATE Recover $M_2, M_3$ using low-rank regression on $\mathcal{D}$.
+    \STATE Return the eigenvectors $\beta_k$ and eigenvalues $\pi_k$ of $M_3$ using $M_2$.
+  \end{algorithmic}
+\end{algorithm}
+
+\subsection{Recovering $M_2$ and $M_3$}
+
+Let $\epsilon \sim \normal{0, \sigma^2}$ be independent Gaussian noise.
+we have,
+\begin{eqnarray*}
+  y &=& \sum_{k=1}^{K} \delta_{k} \beta_k^T x + \epsilon \\
+  \E[ y | x ] &=& M_1^T x \\
+  \E[ y^2 | x ] &=& \sum_{k=1}^{K} \pi_{k} (\beta_k^T x)^2 + \E[ \epsilon^2 ] \\
+  &=& M_2 \odot x^{\otimes 2} + \sigma^2 \\
+  \E[ y^3 | x ] &=& \sum_{k=1}^{K} \pi_{k} (\beta_k^T x)^3
+            + 3 \E[\epsilon^2] (\sum_{k=1}^{K} \pi_{k} \beta_k^T) x \\
+  &=& M_3 \odot x^{\otimes 3} + 3 \sigma^2 M_1.
+\end{eqnarray*}
+
+% Details about low rank regression.
+The symmetric structure of $M_2$ and $M_3$ implies that there are
+$\frac{d (d-1)}{2}$ and $\frac{d (d-1) (d-2)}{6}$ degrees of freedom
+respectively. However, the moments are further constrained to have
+a rank of $K$. To exploit this property, we will use trace norm
+regularization with a proximal subgradient descent
+algorithm\citationneeded to recover $M_2$,
+\begin{eqnarray*}
+  M_2 &=& \arg\min_{M} \half \sum_{i=1}^N ((\xn{i})\tp{2} \odot M - (\yn{i})^2)^2 + \lambda \|M\|_*.
+\end{eqnarray*}
+
+The analogue of trace norm regularization for higher order tensors
+corresponds to the sum of the trace norms of the mode-k unfolding of the
+tensor \cite{Tomioka2011}, $X_{(k)}$, is a $d_k \times (\prod_{k' \neq
+k} d_{k'})$ matrix obtained by concatenating the entries for all
+dimensions other than $k$. For example, the 1-mode unfolding of a 3rd
+order tensor has entries, $X_{(1)}^{i_1, (i_2, i_3)} = X_{i_1, i_2,
+i_3}$. Thus, we employ a proximal subgradient descent algorithm to solve
+the following optimization problem to recovery $M_3$.
+\begin{eqnarray*}
+  M_3 &=& \arg\min_{M} \half \sum_{i=1}^N ((\xn{i})\tp{3} \odot M - (\yn{i})^3)^2 
+          + \frac{\lambda}{3} \sum_{k=1}^{3} \|M_{(k)}\|_*.
+\end{eqnarray*}
+
+In \sectionref{sec:theory}, we will show that recovery of matrices so
+constrained requires only $O(\frac{\sqrt{K} d}{\epsilon})$ samples as
+opposed to $O(\frac{d^3}{\epsilon})$ samples to recover moments within
+$\epsilon$ from their true values.
+
+\subsubsection{Implications for Identifiability}
+
+Our approach employs regression on $x\tp{2}$ and $x\tp{3}$, and thus
+requires $\mathcal{X} \otimes \mathcal{X}$ and $\mathcal{X} \otimes
+\mathcal{X} \otimes \mathcal{X}$ to be linearly independent modulo
+symmetry. We call these condition {\em quadratic} and {\em cubic}
+independence.
+
+This condition trivially holds when the space $\mathcal{X}$ consists of
+independent vectors $\{ x_1, x_2, \cdots, x_d \}$. Unfortunately, it
+places some constraints on non-linear featurizations. For example, the
+common polynomial basis expansion, $1, x_1, x_1^2, \dots, x_1^p$ is
+neither quadratically nor cubically independent because the product
+space is strictly a subset of $\Re^{\frac{d (d-1)}{2}}$ since it
+contains $x_1 \times x_1 = x_1^2 \times 1$, and $x_1 \times x_1 \times
+x_1 = x_1^2 \times x_1 \times 1 = x_1^3 \times 1 \times 1$. In fact, the
+number of unique terms in $x\tp{p}$ is only of the order of $p \times
+d$, and hence we will not have enough equations to solve for $M_2$ or
+$M_3$ unless we consider $p = d^2$ moments.
+
+{\todo Note: we haven't shown identifiability from all moments. What
+does this mean?}
+
+\subsection{Recovering $B$ and $\pi$}
+
+Having recovered the moments $M_2$ and $M_3$, we must now solve for the
+actual regressors $\beta_k$ and mixture weights $\pi_k$, which are the
+eigenvectors and eigenvalues of the symmetric moments respectively,
+\begin{eqnarray*}
+  M_2 &=&  \sum_{k=1}^{K} \pi_k \beta_k^{\otimes 2}\\
+  M_3 &=& \sum_{k=1}^{K} \pi_k \beta_k^{\otimes 3}.
+\end{eqnarray*}
+
+We note that the eigendecomposition of $M_2$ is not sufficient to
+recover the $\beta_k$ because they are not necessarily orthogonal to
+each other. \citet{AnandkumarGeHsu2012} describes an approach that uses
+rotates $M_3$ to an orthogonal basis by using the whitening transform of
+$M_2$. The eigenvectors and eigenvalues recovered from the
+eigendecomposition of $M_3(W, W, W)$ can be de-whitened to recover the
+$\beta_k$ and $\pi_k$.

### 2013-02-11 doc/icml2013/intro.tex
rough changes
Updated draft; intro, models, algo

@@
 
-Local optima are a common problem plaguing the learning of parameters in
-latent variable models. Recently, there has been an interest in spectral
-algorithms that provide statistically consistent estimators, i.e. given
-sufficient data they are guaranteed to converge to the global optimum,
-given that the model is identifiable. Furthermore, these algorithms have
-polynomial sample and computational complexities for several such
-models, e.g.  hidden Markov models \cite{AnandkumarHsuKakade2012},
-probabilistic context free grammars \cite{HsuKakadeLiang2012}, etc. In
-this work, we provide a spectral algorithm for a discriminative model,
-the mixture of linear regressions.
+% Latent variables are important
+% Local optima is problem
+% Recent method of moments spectral methods sidestep local optima
 
-\todo{Where are mixtures of experts used? Why are they interesting?}
+Latent variable models have routinely been used in probabilistic
+modelling, (examples). Parameters for these models have been learnt
+through a variety of methods, EM, variational methods, etc. The local
+nature of these algorithms lead them to converge to local optima.
+Recently, there has been an interest in moment of methods estimators
+with polynomial sample and computational complexities. These estimators
+are statistically consistent, i.e. given sufficient data they are
+guaranteed to converge to the global optimum in an identifiable model,
+thus sidestepping the local optima issue altogether.
 
-Our approach uses regression to recover the moments of the data from the
-scalar observable output variable. Once we recover the second and third
-moments of the regression coefficients, we use a slight modification of
-the method outlined in \citet{AnandkumarHsuKakade2012}.
+% Discriminative models work better in practice
+% Discriminative models with latent variables add expressive power (examples)
+This approach has been applied to a variety of generative models, e.g.
+hidden Markov models \cite{AnandkumarHsuKakade2012}, probabilistic
+context free grammars \cite{HsuKakadeLiang2012}, etc. However, in
+practice it has been observed that discriminative models work better in
+terms of parameter recovery\citationneeded, particularly in the large
+sample regime. \todo{Examples! CRFs, etc.}
+
+% This paper: first step in establishing provably correct estimation
+% with simple mixture of linear regression
+In this work, we provide a first step in establishing provably correct
+estimation for such discriminative models by proposing an algorithm for
+the simple mixture of linear regressions. The mixture of linear
+regressions model \todo{models ?}. We describe the model formally in more
+detail in \sectionref{sec:model}.
+
+% Previous work: moment equations solvable by eigendecomposition or
+% tensor factorization
+While method of moment estimators were \todo{described by Fisher(?)},
+they required as many moments as there were parameters, and hence deemed
+impractical. Recently, \citet{AnandkumarHsuKakade2012} showed how using
+eigendecomposition, the moment equations could be solved using just the
+first three moments of data.
+
+% Challenge: E[X^m Y^n] doesn't work
+In the discriminative setting, we are trying to learn the parameters for
+a model for $y | X$. However, as the response variables are just
+scalars, it is not possible to recover the moments simply from $\E[ X^m
+y^n ].$ At the same time, knowing the $X$ should give us more
+information than the scalar $y$.
+
+% Approach: Derive moment functions based on linear regression on Y^2
+% and Y^3 Algorithm: trace norm regularization to recover tensor
+% + tensor factorization
+We propose an algorithm in \sectionref{sec:algo} that derives moment
+functions based on linear regression on the moments of $y$, i.e.
+$\E[y^2]$ and $\E[y^3]$. We exploit the low rank properties of the
+second and third order moments to efficiently recover parameters, and
+then recover the individual regressors, $\beta_k$, using the tensor
+factorization method \cite{AnandkumarGeHsu2012}.
+
+% Show some identifiability results; require quadratic and cubic
+% independence
+We prove that this approach is indeed statistically consistent with
+polynomial sample complexity in \sectionref{sec:theory}. Aside from
+standard identifiability results \citationneeded, we also show that
+recovery using moments places some independence requirements on the
+variables $X$.
+
+% Empirical evaluation
+% It works
+%  EM is sensitive to initialization (local procdure)
+%  Spectral provides good initialization
+% Intuition: basin of attraction, local methods like EM improve
 
@@
 understand how well the algorithm scales with the number of components,
-dimensions and separation. To test how the algorithms perform on a more
-real dataset, we study their performance on a motion tracking dataset.
+dimensions and separation. To show how robust the algorithm is to model
+misspecification, we evaluate it's performance on \todo{a motion
+tracking dataset(?)}.
 
-In \sectionref{sec:model}, we formally define the mixture of linear
-regressions model and discuss when it is identifiable. We describe our
-algorithm and provide some theoretical results in \sectionref{sec:algo}. We
-present the results of our evaluations in \sectionref{sec:evaluation}.
-Finally, we discuss the merits and demerits of spectral algorithms in
-discriminative models and conclude in \sectionref{sec:discussion}. The
-details of some proofs can be found in \sectionref{sec:proofs}.
+% What properties of the problem make it hard on EM versus spectral?
+%  Separation, dimensionality, number of clusters, etc.
+% Real dataset?
+
+We compare our approach with expectation-maximization; our results,
+presented in \sectionref{sec:evaluation} shows that spectral methods do
+indeed provide an answer close enough to the global optima that using
+a local method from there recovers the true parameters. \todo{Describe
+some of our other high-level findings.}
+
+\todo{What about talking about initialization? Generalizing and kmeans++}

### 2013-02-11 doc/icml2013/model.tex
Updated draft; intro, models, algo

--- 2013-01-17 doc/icml2013/model.tex
@@
 
-Mixture of Experts
+\newcommand{\xn}[1]{x^{(#1)}}
+\newcommand{\xni}{\xn{i}}
+\newcommand{\yn}[1]{y^{(#1)}}
+\newcommand{\yni}{\yn{i}}
 
-\subsection{Identifiability}
+The mixture of linear regressions model describes response variables $y
+\in \Re$ dependent on covariates $x \in \Re^d$ through $K$ different
+modes which occur with probability $\pi_1, \dots, \pi_K$ respectively.
+Each mode is independently described by regression coefficients
+$\beta_k$ and we consider a common variance term $\sigma^2$. Given $x$,
+the generative process for the response variables is,
+\begin{eqnarray*}
+  h &\sim& \mult(\pi) \\
+  y | h = k &\sim& \normal{ \beta_{k}^T x}{ \sigma^2 },
+\end{eqnarray*}
+where $h$ is a latent variable corresponding to the mode.
+
+Our learning framework can be described as follows; we observe $N$
+samples, $(\xn{1}, \yn{1}), \dots, (\xn{N}, \yn{N})$ and would like to
+recover the regressors, $B = [\beta_1 \mid \beta_2 \mid \cdots \mid
+\beta_K]$, and the mixture probabilities, $\pi = (\pi_1, \pi_2, \dots,
+\pi_K)$. Note that we do not observe $h$. If we did, the optimal
+solution for $\beta_k$ would be the regressor found using least squares
+on the set of points, $\{(x_i, y_i) | h_i = k \}$.

### 2013-02-11 doc/icml2013/theory.tex
Updated draft; intro, models, algo

\section{Theoretical Properties}
\label{sec:theory}

\begin{theorem}
  Recovery analysis of mixture of linear regressions.
\end{theorem}
\begin{proof}
\end{proof}

To show: We satisfy RSC.

\subsection{Recovering $M_2$ and $M_3$}

Using low-rank gradient methods. Overview SVD for

\begin{lemma}
  Recovery analysis of nuclear-norm regression is $\le
  \Omega(\frac{kd}{n})$. Try and figure out dependence on condition
  numbers.
\end{lemma}
\begin{proof}
  Theorem X \cite{NegahbanWainwright2009}.
  Theorem Y \cite{Tomioka2011}.
\end{proof}

\subsection{Recovering Regression Coefficients}

Write out the form of $B$ and $\mathcal{B}$ in symmetric tensor form.

Use Tensor Power Method to recovery $\beta_k$.

\begin{lemma}
  Bound error in the $\beta$ based on the errors of $B$ and $\mathcal{B}$.
\end{lemma}
\begin{proof}
  \cite{AnandkumarGeHsu2012}
\end{proof}

### 2013-02-12 doc/icml2013/algo.tex
minor updates

@@
 \label{sec:algo}
+
+The goal is to recover the parameters $(\pi, B)$.
+What should the moments be?
+
+OUTLINE
+
+\subsection{Recovering $B$ and $\pi$ from $M_2,M_3$}
+
+If we knew the following quantities, then we could apply tensor factorization
+\cite{AnandkumarGeHsu2012}.
+
+Having recovered the compound parameters $M_2$ and $M_3$, we must now solve for the
+actual parameters $\{\beta_k\}$ and $\{\pi_k\}$, which are the
+eigenvectors and eigenvalues of the symmetric moments respectively,
+\begin{eqnarray*}
+  M_2 &=&  \sum_{k=1}^{K} \pi_k \beta_k^{\otimes 2}\\
+  M_3 &=& \sum_{k=1}^{K} \pi_k \beta_k^{\otimes 3}.
+\end{eqnarray*}
+
+We note that the eigendecomposition of $M_2$ is not sufficient to
+recover the $\beta_k$ because they are not necessarily orthogonal to
+each other. \citet{AnandkumarGeHsu2012} describes an approach that uses
+rotates $M_3$ to an orthogonal basis by using the whitening transform of
+$M_2$. The eigenvectors and eigenvalues recovered from the
+eigendecomposition of $M_3(W, W, W)$ can be de-whitened to recover the
+$\beta_k$ and $\pi_k$.
+
+Tensor power method
 
@@
 
-\subsection{Notation}
+We want to estimate from low-order moments.
 
-We have already defined $y$ to be the response variable and $x \in
-\Re^d$ to be the covariates it depends on. Note that the matrix of
-regressors $B$ is a $d \times K$ matrix, where the $k$-th column is
-$\beta_k$.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-We will use the notation $x\tp{p}$ to describe the $p$-th order tensor
-formed by taking the outer product of $x$; i.e. $x\tp{p}_{i_1 i_2 \ldots
-i_p} = x_{i_1} x_{i_2} \cdots x_{i_p}$. We will use $\odot$ to denote
-the general dot product between $p$-th order tensors, i.e. $X \odot
-Y = \sum_{i_1, i_2, \ldots i_p} X_{i_1, i_2, \ldots i_p} Y_{i_1, i_2,
-\ldots i_p}.$ Similar to how $x^T A$ is the projection of a vector $x$
-onto the row space of a matrix $A$, when $X \in \Re^{d^p}$ and $Y \in
-\Re^{d^{p+q}}$,  with $p, q > 0$, $X \odot Y \in \Re^{d^q}$ is the
-projection of $X$ onto the first $p$ dimensions of $Y$, i.e. $(X \odot
-Y)_{i_1, \ldots, i_q} = \sum_{j_1, \ldots, j_p} X_{j_1, \ldots, j_p}
-Y_{j_1, \ldots, j_p, i_1, \ldots, i_q}$.
-
-The expected moments of the parameters $\beta_1, \dots, \beta_k$ play an
-important role in our algorithm and we will use $M_p = \E[\beta\tp{p}]
-= \sum_{k=1}^{K} \pi_k \beta_k^{\otimes p}$ to denote them.
-
-\subsection{Identifiability}
-
-The model is in general identifiable, except for a pathological case
-concerning how data points are placed \citationneeded.
-
-\subsection{Overview}
+\subsection{Recovering $M_2$ and $M_3$}
 
@@
 To work around this problem, we will use regression on the conditional
-moments, $\E[y | x]$ to recover the expected moments of the regressors,
+moments, $\E[y \mid x]$ to recover the expected moments of the regressors,
 $M_2$ and $M_3$. Recovering the $\beta_k$ and $\pi_k$ given these
@@
 
-\subsection{Recovering $M_2$ and $M_3$}
-
-Let $\epsilon \sim \normal{0, \sigma^2}$ be independent Gaussian noise.
+Let $\epsilon \sim \normal{0}{\sigma^2}$ be independent Gaussian noise.
 we have,
 \begin{eqnarray*}
-  y &=& \sum_{k=1}^{K} \delta_{k} \beta_k^T x + \epsilon \\
+  %y &=& \sum_{k=1}^{K} \delta_{k} \beta_k^T x + \epsilon \\
   \E[ y | x ] &=& M_1^T x \\
@@
 % Details about low rank regression.
-The symmetric structure of $M_2$ and $M_3$ implies that there are
-$\frac{d (d-1)}{2}$ and $\frac{d (d-1) (d-2)}{6}$ degrees of freedom
-respectively. However, the moments are further constrained to have
+%The symmetric structure of $M_2$ and $M_3$ implies that there are
+%$\frac{d (d-1)}{2}$ and $\frac{d (d-1) (d-2)}{6}$ degrees of freedom
+%respectively.
+However, the moments are further constrained to have
 a rank of $K$. To exploit this property, we will use trace norm
@@
 algorithm\citationneeded to recover $M_2$,
-\begin{eqnarray*}
-  M_2 &=& \arg\min_{M} \half \sum_{i=1}^N ((\xn{i})\tp{2} \odot M - (\yn{i})^2)^2 + \lambda \|M\|_*.
-\end{eqnarray*}
+\begin{align}
+  %\hat M_2 &= \arg\min_{M_2} \frac{1}{2 N} \sum_{i=1}^N ((\xn{i})\tp{2} \odot M_2 - (\yn{i})^2)^2 + \lambda \|M\|_*.
+  \hat M_2 &= \arg\min_{M_2} \frac{1}{2 N} \sum_{(x,y) \in D} (x\tp{2} \odot M_2 - y^2)^2 + \lambda \|M\|_*.
+\end{align}
 
@@
 
-\subsubsection{Implications for Identifiability}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Identifiability from moments}
+
+The model is in general identifiable, except for a pathological case
+concerning how data points are placed \citationneeded.
 
@@
 does this mean?}
-
-\subsection{Recovering $B$ and $\pi$}
-
-Having recovered the moments $M_2$ and $M_3$, we must now solve for the
-actual regressors $\beta_k$ and mixture weights $\pi_k$, which are the
-eigenvectors and eigenvalues of the symmetric moments respectively,
-\begin{eqnarray*}
-  M_2 &=&  \sum_{k=1}^{K} \pi_k \beta_k^{\otimes 2}\\
-  M_3 &=& \sum_{k=1}^{K} \pi_k \beta_k^{\otimes 3}.
-\end{eqnarray*}
-
-We note that the eigendecomposition of $M_2$ is not sufficient to
-recover the $\beta_k$ because they are not necessarily orthogonal to
-each other. \citet{AnandkumarGeHsu2012} describes an approach that uses
-rotates $M_3$ to an orthogonal basis by using the whitening transform of
-$M_2$. The eigenvectors and eigenvalues recovered from the
-eigendecomposition of $M_3(W, W, W)$ can be de-whitened to recover the
-$\beta_k$ and $\pi_k$.

### 2013-02-12 doc/icml2013/model.tex
minor updates

@@
+\subsection{Notation}
+
+%We have already defined $y$ to be the response variable and $x \in
+%\Re^d$ to be the covariates it depends on. Note that the matrix of
+%regressors $B$ is a $d \times K$ matrix, where the $k$-th column is
+%$\beta_k$.
+
+We will use the notation $x\tp{p}$ to describe the $p$-th order tensor
+formed by taking the outer product of $x$; i.e. $x\tp{p}_{i_1 i_2 \ldots
+i_p} = x_{i_1} x_{i_2} \cdots x_{i_p}$. We will use $\odot$ to denote
+the general dot product between $p$-th order tensors, i.e. $X \odot
+Y = \sum_{i_1, i_2, \ldots i_p} X_{i_1, i_2, \ldots i_p} Y_{i_1, i_2,
+\ldots i_p}.$ Similar to how $x^T A$ is the projection of a vector $x$
+onto the row space of a matrix $A$, when $X \in \Re^{d^p}$ and $Y \in
+\Re^{d^{p+q}}$,  with $p, q > 0$, $X \odot Y \in \Re^{d^q}$ is the
+projection of $X$ onto the first $p$ dimensions of $Y$, i.e. $(X \odot
+Y)_{i_1, \ldots, i_q} = \sum_{j_1, \ldots, j_p} X_{j_1, \ldots, j_p}
+Y_{j_1, \ldots, j_p, i_1, \ldots, i_q}$.
+
+The expected moments of the parameters $\beta_1, \dots, \beta_k$ play an
+important role in our algorithm and we will use $M_p = \E[\beta\tp{p}]
+= \sum_{k=1}^{K} \pi_k \beta_k^{\otimes p}$ to denote them.
+
 \section{Model}
@@
   h &\sim& \mult(\pi) \\
-  y | h = k &\sim& \normal{ \beta_{k}^T x}{ \sigma^2 },
+  y \mid h &\sim& \normal{ \beta_{h}^T x}{ \sigma^2 },
 \end{eqnarray*}
@@
 on the set of points, $\{(x_i, y_i) | h_i = k \}$.
+
+Typically people use EM, but the maximum marginal likelihood is non-convex.
+
+Reference mixture of experts
+
+Acknowledge that mixing proporitions can't depend on $x$

### 2013-02-12 doc/icml2013/algo.tex
Spruced up algo section and added details for the regression and tensor power method.

@@
 
-The goal is to recover the parameters $(\pi, B)$.
-What should the moments be?
+In this section, we will outline an algorithm to recover the parameters
+$\pi$ and $B$ of the mixture of linear regressions model.
 
-OUTLINE
+% Motivate how we can recover $B$ from regression by lower order moments
+Our goal is to recover the parameters $(\pi, B)$, ideally from lower
+order moments. Conventionally, when using the method of moments, the
+moments of observed variables are expressed algebraically in terms of
+the model parameters. Using these equations, we attempt to solve for the
+model parameters using empirical estimates of the moments.
 
-\subsection{Recovering $B$ and $\pi$ from $M_2,M_3$}
+Let us consider the response variable $y$, 
+\begin{align}
+  y &=& \innerp{\beta_h}{x} + \epsilon,
+\end{align}
+where $\epsilon \sim \normal{0}{\sigma^2}$ is the observation noise.  It
+is natural to consider regression on the data $\{\xni,
+\yni\}_{i=1}^{N}$, but it would only help us find the mean regression
+coefficients, $M_1 = \E[\beta_h] = \sum_h \pi_h \beta_h$, which are insufficient to
+identify the model. Let us consider some more low powers of $y$, noting
+that taking the expectation gives us the corresponding low-order
+moments,
+\begin{align}
+  y^2 &= \innerp{\beta_h\tp{2}}{x\tp{2}} + \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}} + \epsilon^2 \label{eq:y2} \\ 
+  y^3 &= \innerp{\beta_h\tp{3}}{x\tp{3}} + 3 \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}} \notag \\
+  &+ 3 \epsilon^2 \innerp{\beta_h}{x} + \epsilon^3. \label{eq:y3} 
+\end{align}
+Regression with the second order moments would give us $M_2 = \E[\beta_h
+\tp{2}] = \sum_h \pi_h \beta_h\tp{2}$ which can identify $\pi$ and the
+subspace spanned the $\beta_h$, but not the $\beta_h$ themselves. To see
+this, note that the eigen-decomposition of $M_2$ is not unique.
+\todo{explain better}.
 
-If we knew the following quantities, then we could apply tensor factorization
-\cite{AnandkumarGeHsu2012}.
-
-Having recovered the compound parameters $M_2$ and $M_3$, we must now solve for the
-actual parameters $\{\beta_k\}$ and $\{\pi_k\}$, which are the
-eigenvectors and eigenvalues of the symmetric moments respectively,
-\begin{eqnarray*}
-  M_2 &=&  \sum_{k=1}^{K} \pi_k \beta_k^{\otimes 2}\\
-  M_3 &=& \sum_{k=1}^{K} \pi_k \beta_k^{\otimes 3}.
-\end{eqnarray*}
+This brings us to regression with the third order moments, giving us the
+tensor $M_3 = \E[\beta_h \tp{3}] = \sum_h \pi_h \beta_h\tp{3}$.
+\citet{AnandkumarGeHsu} show that these three moments suffice to
+identify $\pi$ and $B$ and provide an efficient algorithm to recover the
+parameters.
 
@@
 
-Tensor power method
+% Describe the rest of the algorithm.
+This description completes a sketch of the algorithm, described in
+\algorithmref{algo:spectral-experts}. Going ahead, we have yet to show
+that the regression is well-behaved which we will do in
+\sectionref{sec:regression}. This is of concern because the regression
+problem we have has variance introduced from component selection,
+independent of any observation noise. We will show that we can indeed
+efficiently recover $M_2$ and $M_3$ using ideas from low-rank
+regression. Finally, we will outline the tensor power method to recover
+$B$ and $\pi$ given these two quantities, $M_2$ and $M_3$ in
+\sectionref{sec:tensor-power}.
 
-In this section, we will outline some shortcomings of the standard
-method of moments formulation for the mixture of linear regressions
-model. We then describe how we can work around this limitation and
-reduce our problem to a known one, which can be solved using symmetric
-tensor decomposition. Finally, we will state some implications this
-reduction has regarding identifiability.
+\begin{theorem}
+  Recovery analysis of mixture of linear regressions.
+\end{theorem}
+\begin{proof}
+\end{proof}
 
-We want to estimate from low-order moments.
+\begin{algorithm}[t]
+  \caption{Spectral Experts}
+  \label{algo:spectral-experts}
+  \begin{algorithmic}[1]
+    \REQUIRE $\mathcal{D} = \{ (\xn{1}, \yn{1}), \cdots, (\xn{N}, \yn{N}) \}$
+    \REQUIRE $K$, the number of clusters, $\sigma^2$ 
+    \STATE Recover $M_1$ using least-squares regression on $\mathcal{D}$.
+    \STATE Recover $M_2, M_3$ using low-rank regression on $\mathcal{D}$.
+    \STATE Return the eigenvectors $\beta_k$ and eigenvalues $\pi_k$ of $M_3$ using $M_2$.
+  \end{algorithmic}
+\end{algorithm}
+
+We will conclude this section by
+discussing some identifiability concerns for our algorithm.
 
@@
 
-\subsection{Recovering $M_2$ and $M_3$}
+\subsection{Identifiability from moments}
 
-Conventionally, when using the method of moments, the moments of
-observed variables are expressed algebraically in terms of the model
-parameters. Using these equations, we attempt to solve for the model
-parameters using empirical estimates of the moments.
+The mixture of linear regressions model is identifiable subject to
+permutations of the components\citationneeded. However, our approach
+employs regression on $x\tp{2}$ and $x\tp{3}$, and thus requires
+$\mathcal{X} \otimes \mathcal{X}$ and $\mathcal{X} \otimes \mathcal{X}
+\otimes \mathcal{X}$ to be linearly independent modulo symmetry. We call
+these condition {\em quadratic} and {\em cubic} independence.
 
-In the case of the mixture of linear regressions, we have two observed
-quantities, $x$ and $y$. Unfortunately, because $y$ is a scalar
-quantity, we have only one equation for every moment of $y$, and with
-this naive approach, would have to consider $O( (K+1) D)$ moments to
-recover $B \in \Re{K \times D}$ and $\pi \in \Re^K$.
+This condition trivially holds when the space $\mathcal{X}$ consists of
+independent vectors $\{ x_1, x_2, \cdots, x_d \}$. Unfortunately, it
+places some constraints on non-linear featurizations. For example, the
+common polynomial basis expansion, $1, x_1, x_1^2, \dots, x_1^p$ is
+neither quadratically nor cubically independent because the product
+space is strictly a subset of $\Re^{\frac{d (d-1)}{2}}$ since it
+contains $x_1 \times x_1 = x_1^2 \times 1$, and $x_1 \times x_1 \times
+x_1 = x_1^2 \times x_1 \times 1 = x_1^3 \times 1 \times 1$. In fact, the
+number of unique terms in $x\tp{p}$ is only of the order of $p \times
+d$, and hence we will not have enough equations to solve for $M_2$ or
+$M_3$ unless we consider $p = d^2$ moments.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% \subsection{Recovering $M_2$ and $M_3$}
+
+% In the case of the mixture of linear regressions, we have two observed
+% quantities, $x$ and $y$. Unfortunately, because $y$ is a scalar
+% quantity, we have only one equation for every moment of $y$, and with
+% this naive approach, would have to consider $O( (K+1) D)$ moments to
+% recover $B \in \Re{K \times D}$ and $\pi \in \Re^K$.
 
@@
 % encoded in $y$, and not in the $X$.
-
-To work around this problem, we will use regression on the conditional
-moments, $\E[y \mid x]$ to recover the expected moments of the regressors,
-$M_2$ and $M_3$. Recovering the $\beta_k$ and $\pi_k$ given these
-symmetric moments has been addressed in prior work
-\citep{AnandkumarHsuKakade2012, AnandkumarGeHsu2012}.  This algorithm as
-been summarized in \algorithmref{algo:spectral-experts}.
-
-\begin{algorithm}[t]
-  \caption{Spectral Experts}
-  \label{algo:spectral-experts}
-  \begin{algorithmic}[1]
-    \REQUIRE $\mathcal{D} = \{ (\xn{1}, \yn{1}), \cdots, (\xn{N}, \yn{N}) \}$
-    \REQUIRE $K$, the number of clusters, $\sigma^2$ 
-    \STATE Recover $M_1$ using least-squares regression on $\mathcal{D}$.
-    \STATE Recover $M_2, M_3$ using low-rank regression on $\mathcal{D}$.
-    \STATE Return the eigenvectors $\beta_k$ and eigenvalues $\pi_k$ of $M_3$ using $M_2$.
-  \end{algorithmic}
-\end{algorithm}
-
-Let $\epsilon \sim \normal{0}{\sigma^2}$ be independent Gaussian noise.
-we have,
-\begin{eqnarray*}
-  %y &=& \sum_{k=1}^{K} \delta_{k} \beta_k^T x + \epsilon \\
-  \E[ y | x ] &=& M_1^T x \\
-  \E[ y^2 | x ] &=& \sum_{k=1}^{K} \pi_{k} (\beta_k^T x)^2 + \E[ \epsilon^2 ] \\
-  &=& M_2 \odot x^{\otimes 2} + \sigma^2 \\
-  \E[ y^3 | x ] &=& \sum_{k=1}^{K} \pi_{k} (\beta_k^T x)^3
-            + 3 \E[\epsilon^2] (\sum_{k=1}^{K} \pi_{k} \beta_k^T) x \\
-  &=& M_3 \odot x^{\otimes 3} + 3 \sigma^2 M_1.
-\end{eqnarray*}
-
-% Details about low rank regression.
-%The symmetric structure of $M_2$ and $M_3$ implies that there are
-%$\frac{d (d-1)}{2}$ and $\frac{d (d-1) (d-2)}{6}$ degrees of freedom
-%respectively.
-However, the moments are further constrained to have
-a rank of $K$. To exploit this property, we will use trace norm
-regularization with a proximal subgradient descent
-algorithm\citationneeded to recover $M_2$,
-\begin{align}
-  %\hat M_2 &= \arg\min_{M_2} \frac{1}{2 N} \sum_{i=1}^N ((\xn{i})\tp{2} \odot M_2 - (\yn{i})^2)^2 + \lambda \|M\|_*.
-  \hat M_2 &= \arg\min_{M_2} \frac{1}{2 N} \sum_{(x,y) \in D} (x\tp{2} \odot M_2 - y^2)^2 + \lambda \|M\|_*.
-\end{align}
-
-The analogue of trace norm regularization for higher order tensors
-corresponds to the sum of the trace norms of the mode-k unfolding of the
-tensor \cite{Tomioka2011}, $X_{(k)}$, is a $d_k \times (\prod_{k' \neq
-k} d_{k'})$ matrix obtained by concatenating the entries for all
-dimensions other than $k$. For example, the 1-mode unfolding of a 3rd
-order tensor has entries, $X_{(1)}^{i_1, (i_2, i_3)} = X_{i_1, i_2,
-i_3}$. Thus, we employ a proximal subgradient descent algorithm to solve
-the following optimization problem to recovery $M_3$.
-\begin{eqnarray*}
-  M_3 &=& \arg\min_{M} \half \sum_{i=1}^N ((\xn{i})\tp{3} \odot M - (\yn{i})^3)^2 
-          + \frac{\lambda}{3} \sum_{k=1}^{3} \|M_{(k)}\|_*.
-\end{eqnarray*}
-
-In \sectionref{sec:theory}, we will show that recovery of matrices so
-constrained requires only $O(\frac{\sqrt{K} d}{\epsilon})$ samples as
-opposed to $O(\frac{d^3}{\epsilon})$ samples to recover moments within
-$\epsilon$ from their true values.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Identifiability from moments}
-
-The model is in general identifiable, except for a pathological case
-concerning how data points are placed \citationneeded.
-
-Our approach employs regression on $x\tp{2}$ and $x\tp{3}$, and thus
-requires $\mathcal{X} \otimes \mathcal{X}$ and $\mathcal{X} \otimes
-\mathcal{X} \otimes \mathcal{X}$ to be linearly independent modulo
-symmetry. We call these condition {\em quadratic} and {\em cubic}
-independence.
-
-This condition trivially holds when the space $\mathcal{X}$ consists of
-independent vectors $\{ x_1, x_2, \cdots, x_d \}$. Unfortunately, it
-places some constraints on non-linear featurizations. For example, the
-common polynomial basis expansion, $1, x_1, x_1^2, \dots, x_1^p$ is
-neither quadratically nor cubically independent because the product
-space is strictly a subset of $\Re^{\frac{d (d-1)}{2}}$ since it
-contains $x_1 \times x_1 = x_1^2 \times 1$, and $x_1 \times x_1 \times
-x_1 = x_1^2 \times x_1 \times 1 = x_1^3 \times 1 \times 1$. In fact, the
-number of unique terms in $x\tp{p}$ is only of the order of $p \times
-d$, and hence we will not have enough equations to solve for $M_2$ or
-$M_3$ unless we consider $p = d^2$ moments.
-
-{\todo Note: we haven't shown identifiability from all moments. What
-does this mean?}

### 2013-02-12 doc/icml2013/intro.tex
Spruced up algo section and added details for the regression and tensor power method.

@@
 nature of these algorithms lead them to converge to local optima.
-Recently, there has been an interest in moment of methods estimators
-with polynomial sample and computational complexities. These estimators
-are statistically consistent, i.e. given sufficient data they are
-guaranteed to converge to the global optimum in an identifiable model,
-thus sidestepping the local optima issue altogether.
+Recently, there has been an interest in method of moments estimators for
+generative models such as the hidden Markov
+model\cite{AnandkumarHsuKakade2012} or probabilistic context free
+grammars\cite{HsuKakadeLiang2012}.These estimators feature polynomial
+sample and computational complexities while being statistically
+consistent, i.e. given sufficient data they are guaranteed to converge
+to the global optimum in an identifiable model, thus sidestepping the
+local optima issue altogether.
 
-% Discriminative models work better in practice
-% Discriminative models with latent variables add expressive power (examples)
-This approach has been applied to a variety of generative models, e.g.
-hidden Markov models \cite{AnandkumarHsuKakade2012}, probabilistic
-context free grammars \cite{HsuKakadeLiang2012}, etc. However, in
-practice it has been observed that discriminative models work better in
-terms of parameter recovery\citationneeded, particularly in the large
+% Discriminative models work better in practice Discriminative models
+% with latent variables add expressive power (examples)
+In practice it has been observed that discriminative models work better
+in terms of parameter recovery\citationneeded, particularly in the large
 sample regime. \todo{Examples! CRFs, etc.}
-
 % This paper: first step in establishing provably correct estimation
@@
 estimation for such discriminative models by proposing an algorithm for
-the simple mixture of linear regressions. The mixture of linear
-regressions model \todo{models ?}. We describe the model formally in more
-detail in \sectionref{sec:model}.
+the simple mixture of linear regressions. We describe the model formally
+in more detail in \sectionref{sec:model}.
 
@@
 % tensor factorization
-While method of moment estimators were \todo{described by Fisher(?)},
-they required as many moments as there were parameters, and hence deemed
-impractical. Recently, \citet{AnandkumarHsuKakade2012} showed how using
-eigendecomposition, the moment equations could be solved using just the
-first three moments of data.
+%While method of moment estimators were \todo{described by Fisher(?)},
+%they required as many moments as there were parameters, and hence deemed
+%impractical. Recently, \citet{AnandkumarHsuKakade2012} showed how using
+%eigendecomposition, the moment equations could be solved using just the
+%first three moments of data.
 
 % Challenge: E[X^m Y^n] doesn't work
-In the discriminative setting, we are trying to learn the parameters for
-a model for $y | X$. However, as the response variables are just
-scalars, it is not possible to recover the moments simply from $\E[ X^m
-y^n ].$ At the same time, knowing the $X$ should give us more
-information than the scalar $y$.
+%In the discriminative setting, we are trying to learn the parameters for
+%a model for $y | X$. However, as the response variables are just
+%scalars, it is not possible to recover the moments simply from $\E[ X^m
+%y^n ].$ At the same time, knowing the $X$ should give us more
+%information than the scalar $y$.
 
@@
 We propose an algorithm in \sectionref{sec:algo} that derives moment
-functions based on linear regression on the moments of $y$, i.e.
-$\E[y^2]$ and $\E[y^3]$. We exploit the low rank properties of the
-second and third order moments to efficiently recover parameters, and
-then recover the individual regressors, $\beta_k$, using the tensor
+functions based on linear regression on the conditional moments of $y$,
+i.e. $\E[y^2 | X]$ and $\E[y^3 | X]$. We exploit the low rank properties
+of the second and third order moments to efficiently recover parameters,
+and then recover the individual regressors, $\beta_k$, using the tensor
 factorization method \cite{AnandkumarGeHsu2012}.
@@
 \todo{What about talking about initialization? Generalizing and kmeans++}
+
+\subsection{Notation}
+
+%We have already defined $y$ to be the response variable and $x \in
+%\Re^d$ to be the covariates it depends on. Note that the matrix of
+%regressors $B$ is a $d \times K$ matrix, where the $k$-th column is
+%$\beta_k$.
+
+We will use the notation $x\tp{p}$ to describe the $p$-th order tensor
+formed by taking the outer product of $x$; i.e. $x\tp{p}_{i_1 i_2 \ldots
+i_p} = x_{i_1} x_{i_2} \cdots x_{i_p}$. We will use
+$\innerp{\cdot}{\cdot}$ to denote the general dot product between $p$-th
+order tensors, i.e. $\innerp{X}{Y} = \sum_{i_1, i_2, \ldots i_p} X_{i_1,
+i_2, \ldots i_p} Y_{i_1, i_2, \ldots i_p}.$ Similar to how $x^T A$ is
+the projection of a vector $x$ onto the row space of a matrix $A$, when
+$X \in \Re^{d^p}$ and $Y \in \Re^{d^{p+q}}$,  with $p, q > 0$,
+$\innerp{X}{Y} \in \Re^{d^q}$ is the projection of $X$ onto the first
+$p$ dimensions of $Y$, i.e. $(\innerp{X}{Y})_{i_1, \ldots, i_q}
+= \sum_{j_1, \ldots, j_p} X_{j_1, \ldots, j_p} Y_{j_1, \ldots, j_p, i_1,
+\ldots, i_q}$.
+
+The expected moments of the parameters $\beta_1, \dots, \beta_k$ play an
+important role in our algorithm and we will use $M_p = \E[\beta\tp{p}]
+= \sum_{k=1}^{K} \pi_k \beta_k^{\otimes p}$ to denote them.

### 2013-02-12 doc/icml2013/model.tex
Spruced up algo section and added details for the regression and tensor power method.

--- 2013-02-11 doc/icml2013/model.tex
@@
 
-The mixture of linear regressions model describes response variables $y
-\in \Re$ dependent on covariates $x \in \Re^d$ through $K$ different
-modes which occur with probability $\pi_1, \dots, \pi_K$ respectively.
-Each mode is independently described by regression coefficients
-$\beta_k$ and we consider a common variance term $\sigma^2$. Given $x$,
-the generative process for the response variables is,
+The mixture of linear regressions model\citep{VieleTong} describes
+response variables $y \in \Re$ dependent on covariates $x \in \Re^d$
+through $K$ different modes which occur with probability $\pi_1, \dots,
+\pi_K$ respectively. Each mixture component is independently described
+by regression coefficients $\beta_k$ and we consider a common variance
+term $\sigma^2$. Given $x$, the generative process for the response
+variables is,
 \begin{eqnarray*}
   h &\sim& \mult(\pi) \\
-  y | h = k &\sim& \normal{ \beta_{k}^T x}{ \sigma^2 },
+  y \mid h &\sim& \normal{ \beta_{h}^T x}{ \sigma^2 },
 \end{eqnarray*}
@@
 on the set of points, $\{(x_i, y_i) | h_i = k \}$.
+
+% Typically people use EM, but the maximum marginal likelihood is
+% non-convex.
+Typically, the parameters $\pi, B$ are learnt usig
+expectation-maximization\cite{VieleTong}, but the maximum marginal
+likelihood is non-convex. We provide several instances where this leads
+to poor local optima in \sectionref{sec:evaluation}.
+
+% Reference mixture of experts. 
+% Acknowledge that mixing proportions % can't depend on $x$
+The mixture of linear regressions model is closely related to the
+mixture of experts model\citationneeded, wherein the components can be
+SVMs, GPs, regression trees and other estimators. In practice, it is
+desirable to have mixture proportions that are dependent on $x$. We have
+not considered these extensions in our treatment of the model.

### 2013-02-12 doc/icml2013/regression.tex
Spruced up algo section and added details for the regression and tensor power method.

\section{Recovering $M_2, M_3$ with low-rank regression}
\label{sec:regression}

In the previous section, we described an algorithm for the mixture of
linear regressions using regression to recover $M_2$ and $M_3$,
described by \equationref{eq:y2} and \equationref{eq:y3}, as
a subroutine. In this section, we will characterize the rate of
convergence of regression.

Analysis for regression in the fixed and random design settings have
been studied before\citep{HsuKakadeZhang}, however our setup differs
substantially from the noise models assumed in the literature. In our
scenario, the variance in the estimation comes not only from the
Gaussian observation noise (which has been studied before), but also
from the variance in the latent variable $h$.

Let us now formally define the class of regression problems we wish to
analyze, i.e. regression on the set $(x\tp{p}, y^p)$,
\begin{align*}
  y^p &= \innerp{x\tp{p}}{M_p} + (\innerp{x\tp{p}}{M_p - \beta_h\tp{p}} + \varepsilon).
\end{align*}
Note that the term in brackets is the error in our estimates and has
zero mean.  A more succinct way of representing this problem is to use
the operator $\opX(M_p) : \Re^{d\tp{p}} \to \Re^{N}$, which uses the
input $\{ x_i \}_{i=1}^{N}$ to map $M_p$ to a vector of responses $\hat
y_i = \opX(M_p)_i = \innerp{x_i\tp{p}}{M_p}$. We can also represent the
errors with the vector $\vec \epsilon$. Rewriting the equation with this notation, 
\begin{align*}
  \vec y^p &= \opX(M_p) + \vec \epsilon.
\end{align*}

\todo{Describe/define the convex tensor stuff.} We would like to exploit
the property that $M_p$ is low rank (as typically $K \ll D$). It has
been shown that a convex relaxation for this problem regression with
trace norm regularization, which can be solved using a proximal
subgradient descent algorithm\citationneeded. The analogue of trace norm
regularization for higher order tensors corresponds to the sum of the
trace norms of the mode-k unfolding of the tensor \cite{Tomioka2011},
$X_{(k)}$, is a $d_k \times (\prod_{k' \neq k} d_{k'})$ matrix obtained
by concatenating the entries for all dimensions other than $k$. For
example, the 1-mode unfolding of a 3rd order tensor has entries,
$X_{(1)}^{i_1, (i_2, i_3)} = X_{i_1, i_2, i_3}$.

In general, the optimization problem we'd like to solve is,
\begin{align}
  \hat M_p &= 
  \arg\min_{M_p}& \frac{1}{2N} \| \vec y - \opX(M_p) \|_2^2 + \frac{\lambda_N}{K} \sum_{h=1}^K \| (M_p)_{(h)} \|_* \label{eq:regression}.
\end{align}

We begin our analysis by stating a generic theorem about convex
optimization problem in \equationref{eq:regression}, 
\begin{lemma}(Convergence of Trace-Norm Regularized Regression)
  Let $\hat M_p$ be the solution of the optimization problem
  \eqref{eq:regression} with $\lambda_N \ge 2 \frac{\| \opX^*(\epsilon)
  \|}{N}$. Suppose $\opX$ satisfies the restricted strong convexity
  condition, $$\frac{1}{2N} \| \opX( \Delta )\|_2^2 \ge \kappa(\opX)
  \|\Delta\|^2_F$$, then
  $\| \hat M_p - M_p^* \|_F \le \frac{32 \lambda_N}{\kappa(\opX)} \frac{1}{K} \sum_{k=1}^{K} \sqrt{r_k}$.
\end{lemma}
\begin{proof}
  \citet{Tomioka2011}, Theorem 1.

  This proof follows from convex properties \todo{elaborate}.
  \citet{NegahbanWainwright2009} prove this theorem for matrices, i.e.
  when $p = 2$. \citet{Tomioka2011} generalized it for arbitrary $p$.
\end{proof}

Going forward, we need to lower bound the restricted strong convexity
parameter $\kappa(\opX)$ and upper bound the adjoint operator
$\|\opX^*(\epsilon)\|_{2}^2$, which corresponds to a gradient update. We
will appeal to the random design framework that models the input $x$ as
random and show bounds that hold with high probability.

\begin{lemma}(Restricted Strong Convexity)
  Let $\opX$ be the linear operator previously defined. Then,
  $$\kappa{\opX} \ge Y$$.
\end{lemma}
\begin{proof}
  \todo{Do I need to transform rewrite $M_p$ in terms of orthogonal components?}

  \begin{align*}
    \frac{\|\opX(M_p)\|_1}{N} &= \frac{1}{N}\sum_{i=1}^{N} |\innerp{x_i\tp{p}}{M_p}| \\
    &\ge \frac{1}{N}\sum_{i=1}^{N} \sigma_{min}( M_p ) \|x_i\|^p \\
    &\ge \frac{\kappa(M_p)}{\sqrt{D}} \frac{\sum_{i=1}^{N} \|x_i\|^p}{N} \| M_p \|_F.
  \end{align*}

  While $\kappa(M_p)$ depends on $M_p$ we can lower bound it by
  restricting the convex space to $\{ M_p | \kappa(M_p) \ge \kappa \}$.
  Further, using concentration bounds, we can lower bound $\|x_i\|^p$ as
  well. \todo{What is $\E[\|x_i\|^3]$? It is some quantity greater than
  zero}.

  Thus, 
  \begin{align*}
  \frac{ \|\opX(M_p)\|_2 }{2 N} 
  &\ge
  \frac{ \|\opX(M_p)\|_1 }{2 \sqrt{D}N} 
  &\ge \frac{\kappa}{2 D} (\E[\|x_i\|^p] - \frac{t}{N}) \| M_p \|_F.
  \end{align*}
  with high probability $1 - \delta$. \todo{The bound needs to be shown
  for $\| \opX(M_p) \|_2^2$.}

  In other words, $\kappa(\opX)
  = \frac{\kappa}{2 D} (\E[\|x_i\|^p] - \frac{t}{N})$ is a valid
  lower bound.
\end{proof}

\begin{lemma}(Bounded Error)
  Let $\opX$ be the linear operator previously defined. Then,
  $$\opX^*(\epsilon) \le Y$$.
\end{lemma}
\begin{proof}
  \begin{align*}
    \frac{\opX^*(\epsilon)}{N} 
    &= \frac{1}{N}\sum_{i=1}^{N} \epsilon_i \innerp{x_i\tp{p}}{M_p} \\
    &= \frac{1}{N}\sum_{i=1}^{N} (\innerp{x_i\tp{p}}{M_p - \beta_i\tp{p}} + \varepsilon)  \innerp{x_i\tp{p}}{M_p}.
  \end{align*}
  Let $M_p - \beta_i\tp{p} = \tilde M_i$ and consider it to be a random
  variable.
  \begin{align*}
    \frac{\opX^*(\epsilon)}{N} 
    &= \frac{1}{N}\sum_{i=1}^{N} \innerp{\tilde M_i}{x_i\tp{2p}} + \varepsilon \innerp{M_p}{x_i\tp{p}} \\
    &\le \frac{1}{N}\sum_{i=1}^{N} \|\tilde M_i\|_{op} \|x_i\|^{2p} + \varepsilon \|M_p\|_{op}\|x_i\|^{p} \\
    &\le \|M_p\|_{op} ( \frac{1}{N}\sum_{i=1}^{N} \|x_i\|^{2p} + \varepsilon \|x_i\|^{p} ).
  \end{align*}

  We can probabilistically bound the above by considering the union
  bound of $\varepsilon$ and $\|x_i\|^p$.

  Finally,
  \begin{align*}
    \frac{\opX^*(\epsilon)}{N} 
    &\le \|M_p\|_{op} (\E[\|x_i\|^{2p}] + t/N + (t/N) (\E[ \|x_i\|^{p} ] + t/N) ),
  \end{align*}
  with high probability. Note that this does not converge to $0$ as $N
  \to \infty$ because of the variance of the selection probability.
\end{proof}

### 2013-02-12 doc/icml2013/theory.tex
Spruced up algo section and added details for the regression and tensor power method.

@@
 \label{sec:theory}
-
-\begin{theorem}
-  Recovery analysis of mixture of linear regressions.
-\end{theorem}
-\begin{proof}
-\end{proof}
-
-To show: We satisfy RSC.
-
-\subsection{Recovering $M_2$ and $M_3$}
-
-Using low-rank gradient methods. Overview SVD for
-
-\begin{lemma}
-  Recovery analysis of nuclear-norm regression is $\le
-  \Omega(\frac{kd}{n})$. Try and figure out dependence on condition
-  numbers.
-\end{lemma}
-\begin{proof}
-  Theorem X \cite{NegahbanWainwright2009}.
-  Theorem Y \cite{Tomioka2011}.
-\end{proof}
-
-\subsection{Recovering Regression Coefficients}
-
-Write out the form of $B$ and $\mathcal{B}$ in symmetric tensor form.
-
-Use Tensor Power Method to recovery $\beta_k$.
-
-\begin{lemma}
-  Bound error in the $\beta$ based on the errors of $B$ and $\mathcal{B}$.
-\end{lemma}
-\begin{proof}
-  \cite{AnandkumarGeHsu2012}
-\end{proof}

### 2013-02-13 java/src/learning/linalg/MatrixOps.java
Fixed serious bug wherein 'y' is overwritten in Pairs and doom is unleashed

@@ -56,6 +56,7 @@ public class MatrixOps {
    * Print entries of a arrays
    */
   public static String arrayToString( double[][] X ) {
+    if( X == null ) return "null";
     String out = "";
     out += "{\n";
     for( int i = 0; i < X.length; i++ ) {
@@ -69,6 +70,7 @@ public class MatrixOps {
     return out;
   }
   public static String arrayToString( int[][] X ) {
+    if( X == null ) return "null";
     String out = "";
     out += "{\n";
     for( int i = 0; i < X.length; i++ ) {

### 2013-02-13 java/src/learning/spectral/applications/SpectralExperts.java

@@ -142,7 +142,7 @@ public class SpectralExperts implements Runnable {
       weightsErr = MatrixOps.norm( weights.minus( weights_ ) );
 
       if( saveToExecution ) {
-        Execution.putOutput( "weights", betas );
+        Execution.putOutput( "weights", weights );
         Execution.putOutput( "weights_", weights_ );
         Execution.putOutput( "weightsErr", weightsErr);
       }
@@ -339,8 +339,7 @@ public class SpectralExperts implements Runnable {
 
       ProximalGradientSolver solver = new ProximalGradientSolver();
       // Tweak the response variables to give an unbiased estimate
-      DenseMatrix64F y_ = y.getMatrix();
-      CommonOps.elementMult(y_, y_);
+      DenseMatrix64F y_ = y.elementMult(y).getMatrix();
       if(analysis != null)
         CommonOps.add(y_, -analysis.model.getSigma2());
       y = SimpleMatrix.wrap(y_);
@@ -368,7 +367,7 @@ public class SpectralExperts implements Runnable {
 
     int N = X.numRows();
     int D = X.numCols();
-    int D_ = D * (D+1) * (D+2) / 3;
+    int D_ = D * (D+1) * (D+2) / 6;
 
     SimpleMatrix xScaling = MatrixFactory.ones(D);
     double yScaling = 1.0;
@@ -385,20 +384,22 @@ public class SpectralExperts implements Runnable {
     int idx;
     for( int n = 0; n < N; n++ ) {
       idx = 0;
-      for( int d = 0; d < D; d++ ) {
-        for( int d_ = 0; d_ <= d; d_++ ) {
-          for( int d__ = 0; d__ <= d_; d__++ ) {
+      for( int d1 = 0; d1 < D; d1++ ) {
+        for( int d2 = 0; d2 <= d1; d2++ ) {
+          for( int d3 = 0; d3 <= d2; d3++ ) {
             double multiplicity =
-                    ( d == d_ && d_ == d__ ) ? 1 :
-                            ( d == d_ || d_ == d__ || d == d__) ? 3 : 6;
-            A_[n][idx++] = X.get(n, d) * X.get(n,d_) * X.get(n,d__) * multiplicity;
+                    ( d1 == d2 && d2 == d3 ) ? 1 :
+                            ( d1 == d2 || d2 == d3 || d1 == d3) ? 3 : 6;
[…]

### 2013-02-13 java/scripts/generateData.scala
Script now generates a number of files to test on

@@ -10,13 +10,13 @@ val CLASSPATH = "bin:" + (new File("deps/")).listFiles().mkString( ":" )
 
 // Check arguments
 def checkArgs(args:Array[String]) {
-  if( args.length != 3 ) {
-    println("Usage: N nlType <out-directory> ");
+  if( args.length != 4 ) {
+    println("Usage: N E nlType <out-directory> ");
     System.exit(1);
   }
 }
 
-def generate( N:String, outputPath : File, d:Int, nld:Int, nlType:String, k:Int ){
+def generate( N:String, outputPath : File, d:Int, nld:Int, nlType:String, k:Int, e:Int ){
   submit("java"
     -('cp, CLASSPATH)
     -"Xmx2g"
@@ -32,7 +32,7 @@ def generate( N:String, outputPath : File, d:Int, nld:Int, nlType:String, k:Int 
     -('sigma2, 0.1)
     -('bias, false)
     -('nlType, nlType)
-    -('outputPath, outputPath.getAbsolutePath() + File.separator + d + "-" + nld + "-" + k + ".dat")
+    -('outputPath, outputPath.getAbsolutePath() + File.separator + d + "-" + nld + "-" + k + "." + e + ".dat")
     -('D, d)
     -('nlDegree, nld)
     -('K, k)
@@ -42,13 +42,14 @@ def main() {
   checkArgs(args)
 
   val N = args(0);
-  val nlType = args(1);
-  val outputPath = new File( args(2) ++ File.separator ++ nlType);
+  val E = args(1).toInt;
+  val nlType = args(2);
+  val outputPath = new File( args(3) ++ File.separator ++ nlType);
   // Make output directory
   outputPath.mkdirs();
 
   // Construct a list of D, nD, K
-  val D = List(1,2)
+  val D = List(1,2,3)
   val nlD = List(3,4,5)
   val K = List(2,3,5,7)
   val candidateTriples = 
@@ -64,12 +65,13 @@ def main() {
               case "random-fractional" => new FractionalPolynomial(nd, d)
               case "fourier" => new FourierNonLinearity(nd)
             }
-          k <= nl.getLinearDimension(d) 
+          k <= nl.getLinearDimension(d) && nl.getLinearDimension(d) <= 20
       })
   // Spawn creation jobs
   for (( d, nd, k ) <- candidateTriples ) {
     println( d, nd, k );
-    generate( N, outputPath, d, nd, nlType, k );
[…]

### 2013-02-13 doc/icml2013/evaluation.tex
Updated the outline for the paper
rough changes

@@
-\section{Evaluation}
+\section{Empirical Evaluation}
 \label{sec:evaluation}
 
-Mixture of linear regressions,
+\subsection{Setup}
 
-linear 
-quadratic
-motion tracking data
+\paragraph{Algorithms}
+
+Describe EM (talk about initialization, number of iterations / stopping
+condition).
+
+Describe spectral (briefly pointing to algorithm);
+initialization, clipping.
+
+Describe spectral+EM
+
+\paragraph{Datasets}
+Describe the Curves dataset.
+$x = (t, t^4, t^9)$
+$d = b^p$
+
+Distribution over $x$.
+
+Canonical settings:
+
+Default values:
+$n = 10^5$
+$k = 5$
+
+Dataset 1: visualization ($b = 1, p = 3$)
+
+Dataset 2: $b = 30$, $p = 1$
+
+\paragraph{Motion tracking data?}
+
+\subsection{Results}
+
+\paragraph{Basic results}
+
+Punchline: spectral by itself is medicore, offers a good initialization to EM.
+EM with random initialization is much less stable.
+
+Using canonical settings, on the various datasets
+
+Plot the parameters: the curves, three algorithms
+
+Plot: the histogram over errors
+
+\paragraph{Effect on number of data points}
+
+Curves dataset 2.
+Plot: $n$ versus error for various algorithms
+
+Point: tradeoff between statistical error and computational error;
+not enough data, spectral is actually a lot worse
+given enough data, spectral+EM is much better
+
+\paragraph{Effect of dimensionality}
+
+\paragraph{Effect of noise}
+
+\paragraph{Effect of separation between components}

### 2013-02-13 doc/icml2013/evaluation.tex
Added EM Iters as a flag for spectral experts

@@
 
-Describe EM (talk about initialization, number of iterations / stopping
-condition).
+We ran expectation-maximization until the log likelihood converged
+$10^-5$. Initialized the entries of $\beta$ with a random Gaussian, $\normal{0}{1}$.
+
+In all our experiments, EM converged.
 
@@
 Describe the Curves dataset.
-$x = (t, t^4, t^9)$
+$x = (1, t, t^4, t^9)$
 $d = b^p$
@@
 
-Dataset 1: visualization ($b = 1, p = 3$)
+Dataset 1: visualization ($b = 1, p = 3$) (DONE)
+Plot for visualization
 
 Dataset 2: $b = 30$, $p = 1$
+
+Table on different settings: 
+  (To show that we work comparably or better to EM always)
 

### 2013-02-13 java/src/learning/spectral/applications/SpectralExperts.java

@@ -58,6 +58,8 @@ public class SpectralExperts implements Runnable {
   public boolean scaleData = false;
   @Option(gloss = "Run EM on spectral output?")
   public boolean runEM = false;
+  @Option(gloss = "Run EM on spectral output?")
+  public int emIters = 100;
   @Option(gloss = "Use tensor power method")
   public boolean useTensorPowerMethod = true;
   @Option(gloss = "Use ridge regression")
@@ -552,7 +554,7 @@ public class SpectralExperts implements Runnable {
       if( runEM ) {
         learning.em.MixtureOfExperts.Parameters initState = new learning.em.MixtureOfExperts.Parameters(
                 this.K, MatrixFactory.ones(K).scale(1.0/K), betas_.transpose(), 0.1);
-        learning.em.MixtureOfExperts emAlgo = new learning.em.MixtureOfExperts(K);
+        learning.em.MixtureOfExperts emAlgo = new learning.em.MixtureOfExperts(K); emAlgo.iters = emIters;
         learning.em.MixtureOfExperts.Parameters params = emAlgo.run( y, X, initState);
         SimpleMatrix betasEM = (new SimpleMatrix( params.betas )).transpose();
         SimpleMatrix weightsEM = MatrixFactory.fromVector(params.weights);

### 2013-02-14 doc/icml2013/abstract.tex
updates

@@
 \begin{abstract}
-  Spectral algorithms for latent variable models have seen considerable
-  recent interest for being efficient consistent estimators of model
-  parameters. These algorithms make few if any assumptions about the
-  generative process of the data, while providing a polynomial sample
-  and computational complexity. We present a new spectral algorithm for
-  a discriminative model, a mixture of linear regressions and show that
-  it can recover the regression coefficients with similar polynomial
-  guarantees. We evaluate the algorithm on linearly and non-linearly
-  generated data, as well as on a motion tracking task and compare it's
-  characteristics with an E-M algorithm.
+
+Discriminative latent-variable models are typically learned using
+EM or gradient-based optimization, which suffer from local optima.
+In this paper, we develop a new computationally efficient
+and provably consistent estimator for the mixture of linear regressions,
+a simple instance of discriminative latent-variable models.
+Our approach relies on a low-rank linear regression to recover
+a symmetric tensor, which can be factorized into the parameters
+using the tensor power method.
+We prove rates of convergence for our estimator
+and provide an empirical evaluation illustrating
+its strengths relative to local optimization (EM).
+
+%  Spectral algorithms for latent variable models have seen considerable
+%  recent interest for being efficient consistent estimators of model
+%  parameters. These algorithms make few if any assumptions about the
+%  generative process of the data, while providing a polynomial sample
+%  and computational complexity. We present a new spectral algorithm for
+%  a discriminative model, a mixture of linear regressions and show that
+%  it can recover the regression coefficients with similar polynomial
+%  guarantees. We evaluate the algorithm on linearly and non-linearly
+%  generated data, as well as on a motion tracking task and compare it's
+%  characteristics with an E-M algorithm.
 \end{abstract}

### 2013-02-14 doc/icml2013/model.tex
updates

--- 2013-02-11 doc/icml2013/model.tex
@@
 
-The mixture of linear regressions model describes response variables $y
-\in \Re$ dependent on covariates $x \in \Re^d$ through $K$ different
-modes which occur with probability $\pi_1, \dots, \pi_K$ respectively.
-Each mode is independently described by regression coefficients
-$\beta_k$ and we consider a common variance term $\sigma^2$. Given $x$,
-the generative process for the response variables is,
+The mixture of linear regressions model \citep{VieleTong2002} defines
+a conditional distribution over a response $y \in \Re$
+given covariates $x \in \Re^d$.
+The generation of $y$ given $x$ involves two steps:
+(i) draw the mixture component $h \in [K]$ according to mixture proportions
+$\pi = (\pi_1, \dots, \pi_K)$;
+and (ii) generate $y$ from a Gaussian centered at $\beta_h^\top x$ with
+variance $\sigma^2$.
+%Given $x$, the generative process for the response
+%variables is,
 \begin{eqnarray*}
-  h &\sim& \mult(\pi) \\
-  y | h = k &\sim& \normal{ \beta_{k}^T x}{ \sigma^2 },
+  h &\sim& \mult(\pi), \\
+  y \mid h &\sim& \normal{ \beta_{h}^T x}{ \sigma^2 }.
 \end{eqnarray*}
-where $h$ is a latent variable corresponding to the mode.
+%where $h$ is a latent variable corresponding to the mode.
+The parameters of the model are $\theta = (\pi, B, \sigma^2)$,
+where $\pi \in \Re^d$ are the mixture proportions,
+$B = [\beta_1 \mid \dots \mid \beta_K] \in \Re^{d \times K}$,
+are the regression coefficients,
+and $\sigma^2$ is the variance of the observation noise.
+The learning problem is stated as follows:
+given $N$ i.i.d. samples $(\xn{1}, \yn{1}), \dots, (\xn{N}, \yn{N})$
+drawn from the model with some unknown parameters $\theta^*$,
+return an estimate of the parameters $\hat\theta$.
 
-Our learning framework can be described as follows; we observe $N$
-samples, $(\xn{1}, \yn{1}), \dots, (\xn{N}, \yn{N})$ and would like to
-recover the regressors, $B = [\beta_1 \mid \beta_2 \mid \cdots \mid
-\beta_K]$, and the mixture probabilities, $\pi = (\pi_1, \pi_2, \dots,
-\pi_K)$. Note that we do not observe $h$. If we did, the optimal
-solution for $\beta_k$ would be the regressor found using least squares
-on the set of points, $\{(x_i, y_i) | h_i = k \}$.
+%and would like to
+%recover the regressors, $B = [\beta_1 \mid \beta_2 \mid \cdots \mid
+%\beta_K]$, and the mixture probabilities, $\pi = (\pi_1, \pi_2, \dots,
+%\pi_K)$.
+
+%Note that we do not observe $h$. If we did, the optimal
+%solution for $\beta_k$ would be the regressor found using least squares
+%on the set of points, $\{(x_i, y_i) | h_i = k \}$.
+
+% Typically people use EM, but the maximum marginal likelihood is
+% non-convex.
+The estimation problem is difficult primarily because the mixture
+components $h$ are unobserved, rendering the marginal likelihood
+a non-convex function of $\theta$.
+Typically, the parameters are learned using
+expectation maximization (EM) \cite{VieleTong2002},
+which suffers from local optima
+(we show this empirically in \sectionref{sec:evaluation}).
+%We provide several instances where this leads
+%to poor local optima in \sectionref{sec:evaluation}.
+
+This model was used for XXX.
+
+TODO: move to later
+% Reference mixture of experts.
+% Acknowledge that mixture proportions % can't depend on $x$
+The mixture of linear regressions model is an instance of
+mixture of experts \cite{jacobs91},
+%wherein the components can be SVMs, GPs, regression trees and other estimators.
+%In practice, it is
+in which the mixture proportions are allowed to depend on $x$.
+We have not considered these extensions in our treatment of the model.

### 2013-02-14 doc/icml2013/evaluation.tex
Minor fixes; printing tensor recovery error

@@
 
-Dataset 1: visualization ($b = 1, p = 3$) (DONE)
+Dataset 1: visualization ($b = 1, p = 3$) 
 Plot for visualization
-
-Dataset 2: $b = 30$, $p = 1$
 
@@
   (To show that we work comparably or better to EM always)
+  (Spawned!)
+
+Dataset 2: $b = 30$, $p = 1$
+(Spawned many tasks to find spec performance. Will run EM on interesting cases)
 
@@
 Punchline: spectral by itself is medicore, offers a good initialization to EM.
-EM with random initialization is much less stable.
+EM with random initialization is much less stable. (by table)
 
@@
 
-Plot the parameters: the curves, three algorithms
+Plot the parameters: the curves, three algorithms (DONE)
 
-Plot: the histogram over errors
+Plot: the histogram over errors (for EM?)
 

### 2013-02-14 java/src/learning/linalg/FullTensor.java

@@ -468,4 +468,23 @@ public class FullTensor implements Tensor {
     return fold( axis, in.getMatrix(), D );
   }
 
+  public static double testOrthogonality (FullTensor T, SimpleMatrix P) {
+    SimpleMatrix U = P.svd().getU();
+    SimpleMatrix W = MatrixOps.whitener(P);
+
+    T = T.rotate(W, W, W);
+
+    int K = U.numCols();
+    SimpleMatrix Ui[] = new SimpleMatrix[K];
+    for( int i = 0; i < K; i++ )
+      Ui[i] = MatrixOps.col(U, i);
+
+    double diff = 0.0;
+    for( int i = 0; i < K; i++ )
+      diff += Math.pow( T.project3( Ui[i], Ui[i], Ui[(i+1) % K] ), 2);
+
+    return Math.sqrt(diff);
+  }
+
+
 }

### 2013-02-14 java/src/learning/linalg/MatrixOps.java

@@ -19,6 +19,8 @@ import org.ejml.simple.SimpleSVD;
 import org.ejml.simple.SimpleEVD;
 import org.javatuples.*;
 
+import java.util.ArrayList;
+
 public class MatrixOps {
 
   public static double EPS_CLOSE = 1e-4;
@@ -951,11 +953,30 @@ public class MatrixOps {
       CommonOps.scale(w, dX);
     CommonOps.subEquals(dX, X);
     CommonOps.scale(1.0 / (n + 1), dX);
-    CommonOps.addEquals( X, dX );
+    CommonOps.addEquals(X, dX);
   }
   public static void incrementalAverageUpdate( int n, DenseMatrix64F dX, DenseMatrix64F X) {
     incrementalAverageUpdate(1.0, n, dX, X);
   }
+
+  public static double[][] removeInRange( double[][] X, double lbound, double ubound ) {
+    ArrayList<double[]> entries = new ArrayList<>();
+    for( double[] x : X ) {
+      boolean inRange = false;
+      for( double x_i : x ) if( x_i  > lbound && x_i < ubound ) inRange = true;
+      if(!inRange)
+        entries.add(x);
+    }
+
+    double[][] Y = entries.toArray(new double[][]{});
+    return Y;
+  }
+  public static SimpleMatrix removeInRange( SimpleMatrix X, double lbound, double ubound ) {
+    return new SimpleMatrix( removeInRange( MatrixFactory.toArray(X), lbound, ubound ) );
+  }
+
+
+
 }
 
 

### 2013-02-14 java/src/learning/linalg/TensorFactorization.java

@@ -116,10 +116,12 @@ public class TensorFactorization {
    * @return
    */
   public static Pair<SimpleMatrix, SimpleMatrix> eigendecompose( FullTensor T, SimpleMatrix P, int K, int attempts, int iters ) {
+
     // Whiten
     SimpleMatrix W = MatrixOps.whitener(P);
     SimpleMatrix Winv = MatrixOps.colorer(P);
     FullTensor Tw = T.rotate(W,W,W);
+
     Pair<SimpleMatrix, SimpleMatrix> pair = eigendecompose(Tw, K, attempts, iters);
 
     // Color them in again
@@ -130,11 +132,10 @@ public class TensorFactorization {
     int D = Tw.getDim(0);
 
     // TODO: Remove
-    // Make sure these are eigenvectors of T
-//    for( int i = 0; i < K; i++ ) {
-//      SimpleMatrix v = MatrixOps.col(eigenvectors, i);
-//      assert( isEigenvector(Tw, v) );
-//    }
+    // Make sure this was a factorization
+    FullTensor Tw_ = FullTensor.fromDecomposition( eigenvalues, eigenvectors );
+    LogInfo.logs( "Tw_: " + MatrixOps.diff(Tw, Tw_) );
+
 
     // Scale the vectors by 1/sqrt(eigenvalues);
     {
@@ -147,6 +148,9 @@ public class TensorFactorization {
     // Eigenvalues are w^{-1/2}; w is what we want.
     for(int i = 0; i < K; i++)
       eigenvalues.set( i, Math.pow(eigenvalues.get(i), -2) );
+
+    FullTensor T_ = FullTensor.fromDecomposition( eigenvalues, eigenvectors );
+    LogInfo.logs( "T_: " + MatrixOps.diff(T, T_) );
 
     return new Pair<>(eigenvalues, eigenvectors);
   }

### 2013-02-14 java/src/learning/models/MixtureOfExperts.java

@@ -23,6 +23,7 @@ import fig.exec.Execution;
 
 import java.io.*;
 
+import java.util.ArrayList;
 import java.util.Random;
 
 /**
@@ -45,6 +46,8 @@ public class MixtureOfExperts implements Serializable {
 
   protected boolean bias;
   protected NonLinearity nl;
+
+  protected boolean removeThirds = false;
 
   Random rnd = new Random();
 
@@ -101,8 +104,17 @@ public class MixtureOfExperts implements Serializable {
 
     // Add a bias term
     double[][] X_ = MatrixFactory.toArray( X );
+    // Remove thirds
+    if(removeThirds) {
+      X_ = MatrixOps.removeInRange(X_, -0.5, -0.25);
+      X_ = MatrixOps.removeInRange(X_, 0.25, 0.5);
+      N = X_.length;
+    }
+
+
     if( bias ) {
         for( int n = 0; n < N; n++ ) {
+          // If x in a range, just remove it
           double[] x = X_[n];
           X_[n] = new double[ D + 1 ];
           X_[n][0] = 1.0;
@@ -304,7 +316,10 @@ public class MixtureOfExperts implements Serializable {
         throw new NoSuchMethodError();
     }
 
-    return generate( options.K, options.D, options.sigma2, wDistribution, bDistribution, mDistribution, SDistribution, options.pointSigma, options.bias, nl );
+    MixtureOfExperts model =  generate( options.K, options.D, options.sigma2, wDistribution, bDistribution, mDistribution, SDistribution, options.pointSigma, options.bias, nl );
+    model.removeThirds = options.removeThirds;
+
+    return model;
   }
 
   public static class GenerationOptions {
@@ -335,6 +350,8 @@ public class MixtureOfExperts implements Serializable {
     public int nlDegree = 1;
     @Option(gloss="Non-linearity type")
     public String nlType = "poly";
+    @Option(gloss="RemoveThirds")
+    public boolean removeThirds = false;
   }
   public static class OutputOptions {
     @Option(gloss="Print binary file in plain text and exit")

### 2013-02-14 java/src/learning/optimization/ProximalGradientSolver.java

@@ -27,7 +27,7 @@ public class ProximalGradientSolver {
     public double getRate(double t) {
       switch(type) {
         case CONSTANT: return rate;
-        case BY_T: return rate/t+1;
+        case BY_T: return rate/(t+1);
         case BY_SQRT_T: return rate/Math.sqrt(t + 1);
       }
       return rate;
@@ -72,10 +72,10 @@ public class ProximalGradientSolver {
 
       // Compute lhood
       double lhood_ = problem.loss(state);
-      //LogInfo.logs("Iteration " + i + ": " + lhood_ );
+      LogInfo.logs("Iteration " + i + ": " + lhood_ );
 
       // Check convergence
-      if(Math.abs(lhood - lhood_) < eps) {
+      if(lhood - lhood_ < eps) {
         LogInfo.logsForce("Converged.");
         break;
       }
@@ -86,7 +86,7 @@ public class ProximalGradientSolver {
     return state;
   }
   public DenseMatrix64F optimize( ProximalOptimizable problem, DenseMatrix64F initial, LearningRate rate, int maxIters) {
-    return optimize(problem, initial, rate, maxIters, 1e-7);
+    return optimize(problem, initial, rate, maxIters, 1e-3);
   }
   public DenseMatrix64F optimize( ProximalOptimizable problem, DenseMatrix64F initial, LearningRate rate) {
     return optimize(problem, initial, rate, 1000);

### 2013-02-14 java/src/learning/spectral/TensorMethod.java

@@ -14,7 +14,7 @@ import fig.basic.LogInfo;
 public class TensorMethod {
 
   @Option
-  int iters = 100;
+  int iters = 1000;
   @Option
   int attempts = 10;
 

### 2013-02-14 java/src/learning/spectral/applications/SpectralExperts.java

@@ -69,7 +69,7 @@ public class SpectralExperts implements Runnable {
   @Option(gloss = "Use low rank recovery")
   public boolean useLowRankRecovery = true;
   @Option(gloss = "Use low rank recovery")
-  public int lowRankIters = 20;
+  public int lowRankIters = 1000;
 	@Option(gloss = "Number of Threads to use")
 	public int nThreads = 1;
 	@Option(gloss = "Random seed")
@@ -131,6 +131,7 @@ public class SpectralExperts implements Runnable {
       TriplesErr = MatrixOps.diff(Triples, Triples_ );
       if( saveToExecution ) {
         Execution.putOutput("Triples0Err", TriplesErr);
+        Execution.putOutput( "Triples0", Triples_ );
       }
       LogInfo.logsForce("Triples0: " + TriplesErr);
     }
@@ -139,6 +140,8 @@ public class SpectralExperts implements Runnable {
       TriplesErr = MatrixOps.diff(Triples, Triples_ );
       if( saveToExecution ) {
         Execution.putOutput("TriplesErr", TriplesErr);
+        Execution.putOutput( "Triples", Triples );
+        Execution.putOutput( "Triples_", Triples_ );
       }
       LogInfo.logsForce("Triples: " + TriplesErr);
     }
@@ -500,6 +503,8 @@ public class SpectralExperts implements Runnable {
     if( useTensorPowerMethod ) {
       // Use the tensor power method to recover $\betas$.
       TensorMethod algo = new TensorMethod();
+      // Use exact moments god dammit
+//      Pairs = analysis.Pairs; Triples = analysis.Triples;
       Pair<SimpleMatrix, SimpleMatrix> pair = algo.recoverParameters( K, Pairs, Triples );
       // Somewhat of a "hack" to try and rescale the weights to sum to 1
       SimpleMatrix weights = pair.getValue0();
@@ -507,10 +512,10 @@ public class SpectralExperts implements Runnable {
       analysis.reportWeights(weights);
       analysis.reportBetas(betas);
 
-      double sum = weights.elementSum();
-      // Weird
-      //betas = betas.scale( sum );
-      weights = weights.scale( 1/sum );
+//      double sum = weights.elementSum();
+//      Weird
+//      betas = betas.scale( Math.pow(sum,1.0/3) );
+//      weights = weights.scale( 1/sum );
 
       return new Pair<>(weights, betas);
     } else {
[…]

### 2013-02-14 java/tests/learning/spectral/applications/SpectralExpertsTest.java

@@ -23,10 +23,18 @@ import fig.basic.OptionsParser;
  */
 public class SpectralExpertsTest {
 
+  @Before
+  public void setup(){
+    LogInfo.writeToStdout = false;
+    LogInfo.init();
+  }
+
   public void testMomentRunner(MixtureOfExperts model, int N, double reg) {
     int K = model.getK(); int D = model.getD();
     SpectralExperts algo = new SpectralExperts();
+
     algo.enableAnalysis(model);
+
 
     // Compute the empirical moments
     Pair<SimpleMatrix, SimpleMatrix> yX = model.sample(N);

### 2013-02-14 doc/icml2013/algo.tex

@@
-\section{A Method of Moments Algorithm}
+\section{Spectral Experts Algorithm}
 \label{sec:algo}
 
-In this section, we will outline an algorithm to recover the parameters
-$\pi$ and $B$ of the mixture of linear regressions model.
+In this section, we outline our Spectral Experts algorithm
+for recover the model parameters $\theta = (\pi, B, \sigma^2)$.
+Our algorithm consists of two parts:
+(i) low-rank regression to estimate certain symmetric tensors;
+and (ii) tensor factorzation to recover the parameters.
+% Motivate how we can recover $B$ from regression by lower order moments
+%ideally from lower order moments.
+%Conventionally, when using the method of moments, the
+%moments of observed variables are expressed algebraically in terms of
+%the model parameters. Using these equations, we attempt to solve for the
+%model parameters using empirical estimates of the moments.
 
-% Motivate how we can recover $B$ from regression by lower order moments
-Our goal is to recover the parameters $(\pi, B)$, ideally from lower
-order moments. Conventionally, when using the method of moments, the
-moments of observed variables are expressed algebraically in terms of
-the model parameters. Using these equations, we attempt to solve for the
-model parameters using empirical estimates of the moments.
+%%% first moment
+To warm-up, let us consider performing linear regression
+on the response $y$ given $x$:
+\begin{align}
+  y &= \innerp{\bar \beta}{x} + (\innerp{\bar\beta - \beta_h}{x} + \epsilon),
+\end{align}
+where $\bar\beta \eqdef \sum_{h=1}^K \pi_h \beta_h$ is the average regression coefficient.
+Here, there are two noise terms:
+(i) the \emph{mixing noise} $\innerp{\bar\beta - \beta_h}{x}$
+due to the choice of the mixture component $h$
+and (ii) the \emph{observation noise} $\epsilon \sim \normal{0}{\sigma^2}$.
+Performing regression on the data $\{\xni,
+\yni\}_{i=1}^{N}$ would provide a consistent estimate of $\bar\beta$,
+but this is clearly insufficient for identifying all the parameters $\theta$.
 
-Let us consider the response variable $y$, 
+%%% second moment
+Regressing on $x$ provides only first-order
+information.  The key insight is that to get more information about the
+parameters, we need to consider regressing on higher order powers,
+analogous to consider higher order moments in method of moments estimators.
+Let us then consider regressing $y^2$ on $x^\otimes$:
 \begin{align}
-  y &=& \innerp{\beta_h}{x} + \epsilon,
+  y^2 &= \innerp{\beta_h\tp{2}}{x\tp{2}} + \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}} + \epsilon^2 \label{eq:y2}.
 \end{align}
-where $\epsilon \sim \normal{0}{\sigma^2}$ is the observation noise.  It
-is natural to consider regression on the data $\{\xni,
-\yni\}_{i=1}^{N}$, but it would only help us find the mean regression
-coefficients, $M_1 = \E[\beta_h] = \sum_h \pi_h \beta_h$, which are insufficient to
-identify the model. Let us consider some more low powers of $y$, noting
+
+Let us consider some more low powers of $y$, noting
 that taking the expectation gives us the corresponding low-order
@@
 \sectionref{sec:tensor-power}.
+
+\cite{candesPhaseLift} regresses on the square of the response,
+because that's the observation.
 

### 2013-02-14 doc/icml2013/intro.tex

@@
 
-% Latent variables are important
-% Local optima is problem
-% Recent method of moments spectral methods sidestep local optima
-
-Latent variable models have routinely been used in probabilistic
-modelling, (examples). Parameters for these models have been learnt
-through a variety of methods, EM, variational methods, etc. The local
-nature of these algorithms lead them to converge to local optima.
-Recently, there has been an interest in method of moments estimators for
-generative models such as the hidden Markov
-model\cite{AnandkumarHsuKakade2012} or probabilistic context free
-grammars\cite{HsuKakadeLiang2012}.These estimators feature polynomial
-sample and computational complexities while being statistically
-consistent, i.e. given sufficient data they are guaranteed to converge
-to the global optimum in an identifiable model, thus sidestepping the
-local optima issue altogether.
-
 % Discriminative models work better in practice Discriminative models
 % with latent variables add expressive power (examples)
-In practice it has been observed that discriminative models work better
-in terms of parameter recovery\citationneeded, particularly in the large
-sample regime. \todo{Examples! CRFs, etc.}
+Discriminative latent-variable models,
+which combine the high accuracy of discriminative models
+with the compact expressiveness of latent-variable models,
+have been widely applied to many tasks, including
+object recognition \cite{quattoni04nips},
+human action recognition \cite{wang09},
+syntactic parsing \cite{petrov08discriminative},
+and machine translation \cite{liang06discrimative}.
+% http://www.cs.sfu.ca/~mori/research/papers/wang_cvpr09.pdf
+%Slav Petrov, Discriminative Log-Linear Grammars with Latent Variables
+However, parameter estimation in these models is difficult;
+past approaches have relied exclusively on local optimization (EM, 
+gradient descent) and are vulerable to local optima.
+%Our goal is to develop consistent estimators which sidestep
+
+% Goal of this paper
 % This paper: first step in establishing provably correct estimation
 % with simple mixture of linear regression
-In this work, we provide a first step in establishing provably correct
-estimation for such discriminative models by proposing an algorithm for
-the simple mixture of linear regressions. We describe the model formally
-in more detail in \sectionref{sec:model}.
+Our broad goal is to develop efficient provably consistent estimators for
+discriminative latent-variable models.
+In this paper, we provide a first step in this 
+direction by proposing a new algorithm for a simple model,
+\emph{mixture of linear regressions} (MLG) \cite{VieleTong2002}.
+%mixture of linear regressions (\sectionref{sec:model}).
+
+% Recent method of moments spectral methods sidestep local optima
+Recently, method of moments estimators have been developed for
+\emph{generative} latent-variable models, including
+mixture models, HMMs \cite{AnandkumarHsuKakade2012},
+Latent Dirichlet Allocation \cite{AnandkumarLDA},
+and parsing models \cite{HsuKakadeLiang2012}.
+The basic idea of this work is to express
+the unknown model parameters as a tensor factorization
+of the third-order moments of the model distribution, which
+can be estimated from data.
+The moments have a special symmetric structure
+which permits the factorization to be computed efficiently using the robust
+tensor power method \cite{AnandkumarGeHsu2012}.
+%\cite{hsu09spectral}
+
+% Twist
+In mixture of linear regressions, using third-order moments does not directly
+reveal the tensor structure of the problem, so we cannot directly
+apply the above tensor factorization techniques.
+Our solution is to employ low-rank linear regression \cite{NegahbanWainwright2009,Tomioka2011} on second
+and third powers of the response to first recover the appropriate symmetric
+tensors, from which we can then apply the tensor power method to retrieve the
+final parameters.
+
+% Approach: Derive moment functions based on linear regression on Y^2
+% and Y^3 Algorithm: trace norm regularization to recover tensor
+% + tensor factorization
+%We propose an algorithm in \sectionref{sec:algo} that derives moment
+%functions based on linear regression on the conditional moments of $y$,
+%i.e. $\E[y^2 | X]$ and $\E[y^3 | X]$. We exploit the low rank properties
+%of the second and third order moments to efficiently recover parameters,
+%and then recover the individual regressors, $\beta_k$, using the tensor
+%factorization method \cite{AnandkumarGeHsu2012}.
+
+%These estimators feature polynomial
+%sample and computational complexities while being statistically
+%consistent, i.e. given sufficient data they are guaranteed to converge
+%to the global optimum in an identifiable model, thus sidestepping the
+%local optima issue altogether.
 
@@
 
-% Approach: Derive moment functions based on linear regression on Y^2
-% and Y^3 Algorithm: trace norm regularization to recover tensor
-% + tensor factorization
-We propose an algorithm in \sectionref{sec:algo} that derives moment
-functions based on linear regression on the conditional moments of $y$,
-i.e. $\E[y^2 | X]$ and $\E[y^3 | X]$. We exploit the low rank properties
-of the second and third order moments to efficiently recover parameters,
-and then recover the individual regressors, $\beta_k$, using the tensor
-factorization method \cite{AnandkumarGeHsu2012}.
-
 % Show some identifiability results; require quadratic and cubic
 % independence
-We prove that this approach is indeed statistically consistent with
-polynomial sample complexity in \sectionref{sec:theory}. Aside from
-standard identifiability results \citationneeded, we also show that
-recovery using moments places some independence requirements on the
-variables $X$.
+%We prove that this approach is indeed statistically consistent with
+%polynomial sample complexity.
+% PSL: leave out for now (too detailed)
+%Aside from
+%standard identifiability results \citationneeded, we also show that
+%recovery using moments places some independence requirements on the
+%variables $X$.
 
@@
 
-We evaluate our algorithm on simulated linear and non-linear data to
-understand how well the algorithm scales with the number of components,
-dimensions and separation. To show how robust the algorithm is to model
-misspecification, we evaluate it's performance on \todo{a motion
-tracking dataset(?)}.
+% Results
+The result is a simple and efficient two-stage algorithm,
+which we call Spectral Experts.
+We prove that our algorithm yields consistent parameter estimates under certain
+new identifiability conditions.  We also conduct an empirical evaluation
+of our technique to understand its statistical properties (\sectionref{sec:evaluation}).
+In particular, we find that in the low data regime,
+Spectral Experts underperforms EM due to its weaker statistical efficiency,
+but in the high data regime,
+Spectral Experts outperforms EM.
+We get the best results by using Spectral Experts as an initialization
+to EM.
+
+%We evaluate our algorithm on simulated linear and non-linear data to
+%understand how well the algorithm scales with the number of components,
+%dimensions and separation.
+
+%To show how robust the algorithm is to model
+%misspecification, we evaluate it's performance on \todo{a motion
+%tracking dataset(?)}.
 
@@
 
-We compare our approach with expectation-maximization; our results,
-presented in \sectionref{sec:evaluation} shows that spectral methods do
-indeed provide an answer close enough to the global optima that using
-a local method from there recovers the true parameters. \todo{Describe
-some of our other high-level findings.}
+%We compare our approach with expectation-maximization; our results,
+%presented in \sectionref{sec:evaluation} shows that spectral methods do
+%indeed provide an answer close enough to the global optima that using
+%a local method from there recovers the true parameters. \todo{Describe
+%some of our other high-level findings.}
 
-\todo{What about talking about initialization? Generalizing and kmeans++}
+%PSL: not for this paper
+%\todo{What about talking about initialization? Generalizing and kmeans++}
 
@@
 
-We will use the notation $x\tp{p}$ to describe the $p$-th order tensor
-formed by taking the outer product of $x$; i.e. $x\tp{p}_{i_1 i_2 \ldots
+We use $x\tp{p}$ to represent the $p$-th order tensor
+formed by taking the outer product of $x \in \Re^d$; i.e. $x\tp{p}_{i_1 i_2 \ldots
 i_p} = x_{i_1} x_{i_2} \cdots x_{i_p}$. We will use
-$\innerp{\cdot}{\cdot}$ to denote the general dot product between $p$-th
-order tensors, i.e. $\innerp{X}{Y} = \sum_{i_1, i_2, \ldots i_p} X_{i_1,
-i_2, \ldots i_p} Y_{i_1, i_2, \ldots i_p}.$ Similar to how $x^T A$ is
-the projection of a vector $x$ onto the row space of a matrix $A$, when
-$X \in \Re^{d^p}$ and $Y \in \Re^{d^{p+q}}$,  with $p, q > 0$,
+$\innerp{\cdot}{\cdot}$ to denote the generalized dot product between two $p$-th
+order tensors: $\innerp{X}{Y} = \sum_{i_1, i_2, \ldots i_p} X_{i_1,
+i_2, \ldots i_p} Y_{i_1, i_2, \ldots i_p}.$
+For
+$X \in (\Re^d)\tp{p}$ and $Y \in (\Re^d)\tp{(p+q)}$,% $p, q > 0$,
 $\innerp{X}{Y} \in \Re^{d^q}$ is the projection of $X$ onto the first
-$p$ dimensions of $Y$, i.e. $(\innerp{X}{Y})_{i_1, \ldots, i_q}
+$p$ modes of $Y$, i.e. $\innerp{X}{Y}_{i_1, \ldots, i_q}
 = \sum_{j_1, \ldots, j_p} X_{j_1, \ldots, j_p} Y_{j_1, \ldots, j_p, i_1,
 \ldots, i_q}$.
+Let $[n] = \{ 1, \dots, n \}$ denote the first $n$ positive integers.
+%Similar to how $x^T A$ is
+%the projection of a vector $x$ onto the row space of a matrix $A$
 
-The expected moments of the parameters $\beta_1, \dots, \beta_k$ play an
-important role in our algorithm and we will use $M_p = \E[\beta\tp{p}]
-= \sum_{k=1}^{K} \pi_k \beta_k^{\otimes p}$ to denote them.
+%The expected moments of the parameters $\beta_1, \dots, \beta_k$ play an
+%important role in our algorithm and we will use $M_p = \E[\beta\tp{p}]
+%= \sum_{k=1}^{K} \pi_k \beta_k^{\otimes p}$ to denote them.

### 2013-02-14 doc/icml2013/algo.tex
add bib

@@
 Regressing on $x$ provides only first-order
-information.  The key insight is that to get more information about the
-parameters, we need to consider regressing on higher order powers,
-analogous to consider higher order moments in method of moments estimators.
-Let us then consider regressing $y^2$ on $x^\otimes$:
+information.  The key insight is that
+since higher-order moments usually yield more information,
+let us consider regressing $y^2$ on $x^{\otimes 2}$:
 \begin{align}
-  y^2 &= \innerp{\beta_h\tp{2}}{x\tp{2}} + \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}} + \epsilon^2 \label{eq:y2}.
+  y^2 &= \innerp{M_2}{x\tp{2}} + \epsilon \innerp{\beta_h\tp{2} - M_2}{x\tp{2}} + \epsilon^2 \label{eq:y2}.
 \end{align}

### 2013-02-15 doc/icml2013/algo.tex

--- 2013-02-12 doc/icml2013/algo.tex
@@
-\section{A Method of Moments Algorithm}
+\section{Spectral Experts Algorithm}
 \label{sec:algo}
 
-In this section, we will outline an algorithm to recover the parameters
-$\pi$ and $B$ of the mixture of linear regressions model.
+In this section, we describe our Spectral Experts algorithm
+for estimating model parameters $\theta = (\pi, B, \sigma^2)$.
+Our algorithm consists of two steps:
+(i) low-rank regression to estimate certain symmetric tensors;
+and (ii) tensor factorzation to recover the parameters.
+% Motivate how we can recover $B$ from regression by lower order moments
+%ideally from lower order moments.
+%Conventionally, when using the method of moments, the
+%moments of observed variables are expressed algebraically in terms of
+%the model parameters. Using these equations, we attempt to solve for the
+%model parameters using empirical estimates of the moments.
 
-% Motivate how we can recover $B$ from regression by lower order moments
-Our goal is to recover the parameters $(\pi, B)$, ideally from lower
-order moments. Conventionally, when using the method of moments, the
-moments of observed variables are expressed algebraically in terms of
-the model parameters. Using these equations, we attempt to solve for the
-model parameters using empirical estimates of the moments.
+%%% first moment
+To warm up, let us consider linear regression
+on the response $y$ given $x$.
+From the model definition, we have $y = \beta_h^\top x + \epsilon$ where
+$\epsilon \sim \normal{0}{\sigma^2}$, and $h$ is a random quantity
+independent of $x$.
+We can average over this randomness by defining
+the average regression coefficients
+$M_1 \eqdef \sum_{h=1}^k \pi_h \beta_h$.
+Now we can express $y$ as a linear function of $x$ with coefficients $M_1$
+plus some noise $\eta_1(x)$:
+\begin{align}
+  y &= \innerp{M_1}{x} +
+  \underbrace{(\innerp{M_1 - \beta_h}{x} + \epsilon)}_{\eqdef \eta_1(x)}.
+\end{align}
+The noise $\eta_1(x)$ is the sum of two terms:
+(i) the \emph{mixing noise} $\innerp{M_1 - \beta_h}{x}$
+due to the random choice of the mixture component $h$,
+and (ii) the \emph{observation noise} $\epsilon \sim \normal{0}{\sigma^2}$.
+Although the noise depends on $x$,
+it still has zero mean conditioned on $x$.
+We will later show that we can
+perform linear regression on the data $\{\xni,
+\yni\}_{i=1}^{n}$ to produce a consistent estimate of $M_1$
+under identifiability conditions.
+But clearly, knowing $M_1$ is insufficient
+for identifying all the parameters $\theta$.
 
-Let us consider the response variable $y$, 
+%%% second moments
+Intuitively, performing regression on only $y$ given $x$ provides only first-order
+information.  The key insight is that we can perform regression
+on higher order powers to obtain more information about the parameters.
+Specifically, for an integer $p \ge 1$, let us define the average
+$m$-th order tensor power of the parameters as follows:
 \begin{align}
-  y &=& \innerp{\beta_h}{x} + \epsilon,
+M_p &\eqdef \sum_{h=1}^k \pi_h \beta_h\tp{p}. \label{eqn:Mp} % \in (\Re^{d})\tp{p}.
 \end{align}
-where $\epsilon \sim \normal{0}{\sigma^2}$ is the observation noise.  It
-is natural to consider regression on the data $\{\xni,
-\yni\}_{i=1}^{N}$, but it would only help us find the mean regression
-coefficients, $M_1 = \E[\beta_h] = \sum_h \pi_h \beta_h$, which are insufficient to
-identify the model. Let us consider some more low powers of $y$, noting
-that taking the expectation gives us the corresponding low-order
-moments,
+Now consider performing regression on $y^2$ given $x\tp{2}$.
+Expanding $y^2 = (\innerp{\beta_h}{x} + \epsilon)^2$,
+using the fact that $\innerp{\beta_h}{x}^p = \innerp{\beta_h\tp{p}}{x\tp{p}}$,
+we have:
 \begin{align}
-  y^2 &= \innerp{\beta_h\tp{2}}{x\tp{2}} + \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}} + \epsilon^2 \label{eq:y2} \\ 
-  y^3 &= \innerp{\beta_h\tp{3}}{x\tp{3}} + 3 \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}} \notag \\
-  &+ 3 \epsilon^2 \innerp{\beta_h}{x} + \epsilon^3. \label{eq:y3} 
+y^2 &= \innerp{M_2}{x\tp{2}} + \sigma^2 + \eta_2(x), \label{eqn:y2} \\
+\eta_2(x) &= \innerp{\beta_h\tp{2} - M_2}{x\tp{2}} + \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}} + (\epsilon^2 - \sigma^2). \nonumber
+%y^2 &= \innerp{M_2}{x\tp{2}} + \sigma^2 + \underbrace{\innerp{\beta_h\tp{2} - M_2}{x\tp{2}} + \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}}}_{\eqdef \eta_2(x)} + (\epsilon^2 - \sigma^2) \label{eq:y2}, \\
 \end{align}
-Regression with the second order moments would give us $M_2 = \E[\beta_h
-\tp{2}] = \sum_h \pi_h \beta_h\tp{2}$ which can identify $\pi$ and the
-subspace spanned the $\beta_h$, but not the $\beta_h$ themselves. To see
-this, note that the eigen-decomposition of $M_2$ is not unique.
-\todo{explain better}.
+Again, we have expressed $y^2$ has a linear function of $[x\tp{2}, 1]$
+with regression coefficients $[M_2, \sigma^2]$, plus noise.
+Importantly, the noise has mean zero; 
+in fact each of the three terms has mean zero
+by (i) definition of $M_2$, (ii) independence of $\epsilon$ and $h$,
+and (iii) the fact that $\E[\epsilon^2] = \sigma^2$.
 
-This brings us to regression with the third order moments, giving us the
-tensor $M_3 = \E[\beta_h \tp{3}] = \sum_h \pi_h \beta_h\tp{3}$.
-\citet{AnandkumarGeHsu} show that these three moments suffice to
-identify $\pi$ and $B$ and provide an efficient algorithm to recover the
-parameters.
+Performing regression yields a consistent estimate of $[M_2, \sigma^2]$,
+but does not identify all the parameters $\theta$.
+In particular, $B$ is only identified up to rotation:
+if $B = [\beta_1 \mid \cdots \mid \beta_k]$ satisfies
+$B \diag(\pi) B^\top = M_2$, then $(B Q) \diag(\pi) (Q B^\top)$
+for any orthogonal matrix $Q$.
+%Regression with the second order moments would give us $M_2 = \E[\beta_h
+%\tp{2}] = \sum_h \pi_h \beta_h\tp{2}$ which can identify $\pi$ and the
+%subspace spanned the $\beta_h$, but not the $\beta_h$ themselves. To see
+%this, note that the eigen-decomposition of $M_2$ is not unique.
+%\todo{explain better}.
 
-We note that the eigendecomposition of $M_2$ is not sufficient to
-recover the $\beta_k$ because they are not necessarily orthogonal to
-each other. \citet{AnandkumarGeHsu2012} describes an approach that uses
-rotates $M_3$ to an orthogonal basis by using the whitening transform of
-$M_2$. The eigenvectors and eigenvalues recovered from the
-eigendecomposition of $M_3(W, W, W)$ can be de-whitened to recover the
-$\beta_k$ and $\pi_k$.
+%%% third moment
+Let us now proceed to the third moment with the hope that it will
+provide additional information.
+We can write $y^3$ as a linear function of $x\tp{3}$ plus zero mean noise:
+\begin{align}
+y^3 &= \innerp{M_3}{x\tp{3}} + 3\sigma^2 \innerp{M_1}{x} + \eta_3(x), \label{eqn:y3} \\
+\eta_3(x) &= \innerp{\beta_h\tp{3} - M_3}{x\tp{3}}
++ 3 \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}} \nonumber \\
+&+ 3(\epsilon^2 \innerp{\beta_h}{x} - \sigma^2 \innerp{M_1}{x})
++ \epsilon^3, \nonumber
+\end{align}
+where $\E[\eta_3(x) \mid x] = 0$.
+Performing regression on our data yields estimates $M_3$ and $\sigma^2 M_1$.
+It turns out that knowledge of $M_2$ and $M_3$ are sufficient to recover
+all the parameters.
+
+%We note that the eigendecomposition of $M_2$ is not sufficient to
+%recover the $\beta_k$ because they are not necessarily orthogonal to
+%each other.
+
+% Full algorithm
+Now we are ready to state our full algorithm, which we call Spectral Experts.
+The full algorithm is described in \algorithmref{sec:algo}.
+First, we perform two regressions according to \refeqn{y2}
+and \refeqn{y3} to recover the compound parameters $M_2$ and $M_3$.
+The first regression \refeqn{estimateM2}
+produces estimates of $\hat M_2$ and $\hat\sigma^2$;
+the second regression \refeqn{estimateM3} produces $\hat M_3$ (we ignore $\hat S_1$).
+Since $M_2$ and $M_3$ have low rank structure,
+we use nuclear norm regularization \cite{Tomioka2011,NegahbanWainwright2009}
+to exploit this structure.
+The regularization strengths $\lambda_n^{(2)}$ and $\lambda_n^{(2)}$
+are set to $O(\frac{1}{\sqrt{n}})$.
+See \sectionref{sec:regression} for more details.
+
+Second, given a matrix $M_2$ and tensor $M_3$ of the form in \refeqn{Mp},
+\citet{AnandkumarGeHsu2012} showed that due to symmetry of $M_2$ and $M_3$,
+we can efficiently perform tensor factorization of the whitened
+version of $M_3$.  The eigenvalues and eigenvectors
+directly yield the parameters $\pi$ and
+$\{\beta_h\}$.
+See \sectionref{sec:tensor-power} for details.
+%\citet{AnandkumarGeHsu2012} showed that given $M_2$ and $M_3$
+%of the form in \refeqn{Mp},
+%the robust tensor power method can be employed to recover $\pi$ and
+%$\{\beta_h\}$.
+
+%\citet{AnandkumarGeHsu2012} describes an approach that uses
+%rotates $M_3$ to an orthogonal basis by using the whitening transform of
+%$M_2$. The eigenvectors and eigenvalues recovered from the
+%eigendecomposition of $M_3(W, W, W)$ can be de-whitened to recover the
+%$\beta_k$ and $\pi_k$.
 
 % Describe the rest of the algorithm.
-This description completes a sketch of the algorithm, described in
-\algorithmref{algo:spectral-experts}. Going ahead, we have yet to show
-that the regression is well-behaved which we will do in
-\sectionref{sec:regression}. This is of concern because the regression
-problem we have has variance introduced from component selection,
-independent of any observation noise. We will show that we can indeed
-efficiently recover $M_2$ and $M_3$ using ideas from low-rank
-regression. Finally, we will outline the tensor power method to recover
-$B$ and $\pi$ given these two quantities, $M_2$ and $M_3$ in
-\sectionref{sec:tensor-power}.
-
-\begin{theorem}
-  Recovery analysis of mixture of linear regressions.
-\end{theorem}
-\begin{proof}
-\end{proof}
+%This description completes a sketch of the algorithm, described in
+%\algorithmref{algo:spectral-experts}. Going ahead, we have yet to show
+%that the regression is well-behaved which we will do in
+%\sectionref{sec:regression}. This is of concern because the regression
+%problem we have has variance introduced from component selection,
+%independent of any observation noise. We will show that we can indeed
+%efficiently recover $M_2$ and $M_3$ using ideas from low-rank
+%regression. Finally, we will outline the tensor power method to recover
+%$B$ and $\pi$ given these two quantities, $M_2$ and $M_3$ in
+%\sectionref{sec:tensor-power}.
 
@@
   \begin{algorithmic}[1]
-    \REQUIRE $\mathcal{D} = \{ (\xn{1}, \yn{1}), \cdots, (\xn{N}, \yn{N}) \}$
-    \REQUIRE $K$, the number of clusters, $\sigma^2$ 
-    \STATE Recover $M_1$ using least-squares regression on $\mathcal{D}$.
-    \STATE Recover $M_2, M_3$ using low-rank regression on $\mathcal{D}$.
-    \STATE Return the eigenvectors $\beta_k$ and eigenvalues $\pi_k$ of $M_3$ using $M_2$.
+    \INPUT Data $\mathcal{D} = \{ (\xn{1}, \yn{1}), \cdots, (\xn{n}, \yn{n}) \}$.
+    \OUTPUT Parameters $\hat\theta = (\hat \pi, [\hat \beta_1 \mid \cdots \mid \hat \beta_k], \hat\sigma^2)$.
+    \STATE Estimate compound parameters $M_2, M_3$ using \textbf{low-rank regression}:
+    \begin{align}
+      &[\hat M_2, \hat \sigma^2] = \arg\min_{M_2,\sigma^2} \label{eqn:estimateM2} \\
+      &\quad\frac{1}{2n} \sum_{(x,y) \in \sD} (\innerp{M_2}{x\tp{2}} + \sigma^2 - y^2)^2 + \lambda_n^{(2)} \|M_2\|_*, \nonumber \\
+       &[\hat M_3, \hat S_1] = \arg\min_{M_3,S_1} \label{eqn:estimateM3} \\
+       &\quad\frac{1}{2n} \sum_{(x,y) \in \sD} (\innerp{M_3}{x\tp{3}} + \innerp{S_1}{x} - y^3)^2 + \lambda_n^{(3)} \|M_3\|_*. \nonumber
+    \end{align}
+    \STATE Estimate the coefficients $\{\beta_h\}$ using \textbf{tensor factorzation}:
+    \begin{enumerate}
+      \item [(a)] Compute whitening matrix $W \in \Re^{d \times k}$ (such that $W^\top
+      \hat M_2 W = I$) using SVD.
+      \item [(b)] Compute the eigenvalues $\{\lambda_h\}_{h=1}^k$
+      and eigenvectors $\{v_h\}_{h=1}^k$
+      of the whitened tensor $M_3(W, W, W) \in \Re^{k \times k \times k}$
+      by using the robust tensor power method.
+    \item [(c)] Set parameter estimates $\hat\pi_h = \frac{1}{\lambda_h^2}$
+    and $\hat\beta_h = (W^{\top})^\dagger (\lambda_h v_h)$.
+    \end{enumerate}
   \end{algorithmic}
@@
 
-We will conclude this section by
-discussing some identifiability concerns for our algorithm.
+Our main theoretical result shows that Spectral Experts returns
+consistent parameter estimates (see Appendix
+\begin{theorem}[Convergence of Spectral Experts]
+Let $\sD$ consist of $n$ i.i.d. points drawn from the mixture
+of linear regressions model with parameter $\theta^*$.
+Assume that $\E[[1,\unfold(x\tp{2})]\tp{2}] \succ 0$ 
+and $\E[[x,\unfold(x\tp{3})]\tp{2}] \succ 0$.
+Then the parameter estimates $\hat\theta$ returned by
+\algorithmref{algo:spectral-experts}, with the columns appropriate permuted,
+satisfies $\sigma^2 \le \epsilon$,
+$\|\hat\pi - \pi^*\|_{\infty} \le \epsilon$
+and for all $h \in [k]$,
+$\|\hat\beta_h - \beta^*_h\| \le \epsilon$,
+for some error tolerance $\epsilon = O(\frac{kd^2}{n (\min_h \pi_h^*)})$.
+\end{theorem}
+\begin{proof}
+Let $T^* = M_3^*(W^*, W^*, W^*)$.
+By Theorem 5.1 of \cite{AnandkumarGeHsu2012},
+\end{proof}
+
+% Method of moments interpretation
+%Though we have motivated Spectral Experts using regression on powers
+%of $y$ given powers of $x$,
+%we can also think of the algorithm in terms of
+%the method of moments.  Specifically,
+%the optimality conditions for the regression of $M_2$ and $M_3$ are:
+%\begin{align}
+%\E[x\tp{2} (\innerp{M_2}{x\tp{2}} + \sigma^2 - y^2)] = 0, \\
+%\E[x\tp{3} (\innerp{M_3}{x\tp{3}} + 3 \sigma^2 \innerp{M_1}{x} - y^3)] = 0.
+%\end{align}
 
@@
 % quantity, we have only one equation for every moment of $y$, and with
-% this naive approach, would have to consider $O( (K+1) D)$ moments to
-% recover $B \in \Re{K \times D}$ and $\pi \in \Re^K$.
+% this naive approach, would have to consider $O( (k+1) D)$ moments to
+% recover $B \in \Re{k \times D}$ and $\pi \in \Re^k$.
 
@@
 % \begin{eqnarray*}
-%   \E[ y ] &=& (\sum_{k=1}^{K} \pi_{k} \beta_k)^T x \\
+%   \E[ y ] &=& (\sum_{k=1}^{k} \pi_{k} \beta_k)^T x \\
 %           &=& M_1^T \E[x] \\
-%   \E[ y^2 ] &=& \sum_{k=1}^{K} \pi_{k} \E[(\beta_k^T x)]^2 \\
-%   &=& \sum_{k=1}^{K} \pi_{k} (\beta_k\tp{2} \odot \E[x\tp{2}]) \\
+%   \E[ y^2 ] &=& \sum_{k=1}^{k} \pi_{k} \E[(\beta_k^T x)]^2 \\
+%   &=& \sum_{k=1}^{k} \pi_{k} (\beta_k\tp{2} \odot \E[x\tp{2}]) \\
 %   &=& M_2 \odot \E[x\tp{2}] \\

### 2013-02-15 doc/icml2013/regression.tex

@@
 
-In the previous section, we described an algorithm for the mixture of
-linear regressions using regression to recover $M_2$ and $M_3$,
-described by \equationref{eq:y2} and \equationref{eq:y3}, as
-a subroutine. In this section, we will characterize the rate of
-convergence of regression.
+In this section, we analyze step 1 of the Spectral Experts
+algorithm: that of recovering the compound parameters $M_2,M_3$
+from data.
+
+%In the previous section, we described an algorithm for the mixture of
+%linear regressions using regression to recover $M_2$ and $M_3$,
+%described by \equationref{eqn:y2} and \equationref{eq:y3}, as
+%a subroutine. In this section, we will characterize the rate of
+%convergence of regression.
+
+\cite{candesPhaseLift} regresses on the square of the response,
+because that's the observation.
 
 Analysis for regression in the fixed and random design settings have
-been studied before\citep{HsuKakadeZhang}, however our setup differs
+been studied before \citep{HsuKakadeZhang}, however our setup differs
 substantially from the noise models assumed in the literature. In our
@@
   \hat M_p &= 
-  \arg\min_{M_p}& \frac{1}{2N} \| \vec y - \opX(M_p) \|_2^2 + \frac{\lambda_N}{K} \sum_{h=1}^K \| (M_p)_{(h)} \|_* \label{eq:regression}.
+  \arg\min_{M_p}& \frac{1}{2N} \| \vec y - \opX(M_p) \|_2^2 + \frac{\lambda_n}{K} \sum_{h=1}^K \| (M_p)_{(h)} \|_* \label{eq:regression}.
 \end{align}
@@
 
-\begin{lemma}(Restricted Strong Convexity)
+\begin{lemma}[restricted strong convexity]
 Let $\opX$ be the linear operator previously defined. Then,
-  $$\kappa{\opX} \ge Y$$.
+$$\kappa(\opX) \ge Y$$.
 \end{lemma}
@@
 
-\begin{lemma}(Bounded Error)
+\begin{lemma}[bounded error]
   Let $\opX$ be the linear operator previously defined. Then,

### 2013-02-15 doc/icml2013/model.tex

@@
 where $\pi \in \Re^d$ are the mixture proportions,
-$B = [\beta_1 \mid \dots \mid \beta_K] \in \Re^{d \times K}$,
+$B = [\beta_1 \mid \dots \mid \beta_K] \in \Re^{d \times K}$
 are the regression coefficients,

### 2013-02-15 doc/icml2013/algo.tex
checked in

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Theoretical results}
+
 Our main theoretical result shows that Spectral Experts returns
@@
 % Method of moments interpretation
+% This is garbage
 %Though we have motivated Spectral Experts using regression on powers
@@
 
-\subsection{Identifiability from moments}
+\subsection{Identifiability}
+
+In ordinary linear regression, the regression coefficients are identifiable
+only if the data has full rank: $\E[x\tp{2}] \succ 0$.
+However, in mixture of linear regressions.
 
 The mixture of linear regressions model is identifiable subject to
-permutations of the components\citationneeded. However, our approach
+permutations of the components. However, our approach
 employs regression on $x\tp{2}$ and $x\tp{3}$, and thus requires

### 2013-02-15 doc/icml2013/evaluation.tex

@@
 
-\subsection{Setup}
+\subsection{Experimental Setup}
 
@@
 
-We ran expectation-maximization until the log likelihood converged
-$10^-5$. Initialized the entries of $\beta$ with a random Gaussian, $\normal{0}{1}$.
+We implemented the Spectral Experts algorithm using a proximal
+subgradient descent algorithm\citationneeded for the low rank recovery
+of $M_2$ and $M_3$. The algorithm was initialized with the regularized
+least squares solution and was run to convergence in each instance. Both
+regularization parameters were annealed as $\frac{1}{\sqrt{T}}$.
 
-In all our experiments, EM converged.
-
-Describe spectral (briefly pointing to algorithm);
-initialization, clipping.
-
-Describe spectral+EM
+We also implemented EM for this problem. In one set of experiments, we
+initialized the parameters $\beta$ with random Gaussian entries with
+mean zero and variance 1 and the mixture probabilities $\pi$ were
+initialized to be uniform. We also compared the performance of EM when
+initialized with the starting points given by Spectral Experts. We ran
+EM for at most 1000 iterations, but in practice the experiments always
+converged before this.
 
 \paragraph{Datasets}
+
+We ran our experiments on synthetically generated non-linear data. To
+conform to the identifiability criteria discussed in
+\sectionref{sec:algo}, we considered only features that would guarantee
+this condition.
+
+\begin{figure*}[t]
+  \centering
+  \subfigure{}{
+    \includegraphics[width=0.30\textwidth]{figures/curves-vis/1-8-3-spec.png}}
+  \subfigure{}{
+    \includegraphics[width=0.30\textwidth]{figures/curves-vis/1-8-3-em.png}}
+  \subfigure{}{
+    \includegraphics[width=0.30\textwidth]{figures/curves-vis/1-8-3-spem.png}}
+  \label{fig:curves}
+\end{figure*}
+
+As an example, in the one dimensional scenario, one set of features we
+considered were $\{1, x, x^4, x^7\}$. The data and the curves fit using
+the spectral algorithm, EM with random initializations and EM
+initialized with the parameters recovered using spectral experts are
+plotted in \figureref{fig:curves}.
+
+We generated $x$ uniformly from the unit hypercube.
+
 Describe the Curves dataset.

### 2013-02-15 doc/icml2013/algo.tex
many changes

--- 2013-02-15 doc/icml2013/algo.tex
@@
 In this section, we describe our Spectral Experts algorithm
-for estimating model parameters $\theta = (\pi, B, \sigma^2)$.
-Our algorithm consists of two steps:
+for estimating model parameters $\theta = (\pi, B)$.
+The algorithm consists of two steps:
 (i) low-rank regression to estimate certain symmetric tensors;
 and (ii) tensor factorzation to recover the parameters.
+The two steps can be performed efficiently using
+convex optimization and tensor power method, respectively.
 % Motivate how we can recover $B$ from regression by lower order moments
@@
   y &= \innerp{M_1}{x} +
-  \underbrace{(\innerp{M_1 - \beta_h}{x} + \epsilon)}_{\eqdef \eta_1(x)}.
+  \underbrace{(\innerp{M_1 - \beta_h}{x} + \epsilon)}_{\eqdef \eta_1(x)}. \label{eqn:y1}
 \end{align}
@@
 y^2 &= \innerp{M_2}{x\tp{2}} + \sigma^2 + \eta_2(x), \label{eqn:y2} \\
-\eta_2(x) &= \innerp{\beta_h\tp{2} - M_2}{x\tp{2}} + \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}} + (\epsilon^2 - \sigma^2). \nonumber
-%y^2 &= \innerp{M_2}{x\tp{2}} + \sigma^2 + \underbrace{\innerp{\beta_h\tp{2} - M_2}{x\tp{2}} + \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}}}_{\eqdef \eta_2(x)} + (\epsilon^2 - \sigma^2) \label{eq:y2}, \\
+\eta_2(x) &= \innerp{\beta_h\tp{2} - M_2}{x\tp{2}} + 2 \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}} + (\epsilon^2 - \sigma^2). \nonumber
 \end{align}
-Again, we have expressed $y^2$ has a linear function of $[x\tp{2}, 1]$
-with regression coefficients $[M_2, \sigma^2]$, plus noise.
+Again, we have expressed $y^2$ has a linear function of $x\tp{2}$
+with regression coefficients $M_2$, plus a known bias $\sigma^2$ and noise.
 Importantly, the noise has mean zero; 
@@
 
-Performing regression yields a consistent estimate of $[M_2, \sigma^2]$,
+Performing regression yields a consistent estimate of $M_2$,
 but does not identify all the parameters $\theta$.
@@
 for any orthogonal matrix $Q$.
-%Regression with the second order moments would give us $M_2 = \E[\beta_h
-%\tp{2}] = \sum_h \pi_h \beta_h\tp{2}$ which can identify $\pi$ and the
-%subspace spanned the $\beta_h$, but not the $\beta_h$ themselves. To see
-%this, note that the eigen-decomposition of $M_2$ is not unique.
-%\todo{explain better}.
 
 %%% third moment
-Let us now proceed to the third moment with the hope that it will
-provide additional information.
-We can write $y^3$ as a linear function of $x\tp{3}$ plus zero mean noise:
+Let us look to the third moment for additional information.
+We can write $y^3$ as a linear function of $x\tp{3}$ with coefficients $M_3$
+plus a bias $3 \sigma^2 \innerp{M_1}{x}$ that can be estimated from \refeqn{y1} and zero mean noise $\eta_3(x)$:
 \begin{align}
@@
 + 3 \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}} \nonumber \\
-&+ 3(\epsilon^2 \innerp{\beta_h}{x} - \sigma^2 \innerp{M_1}{x})
-+ \epsilon^3, \nonumber
+&\quad + 3(\epsilon^2 \innerp{\beta_h}{x} - \sigma^2 \innerp{M_1}{x})
++ \epsilon^3. \nonumber
 \end{align}
-where $\E[\eta_3(x) \mid x] = 0$.
-Performing regression on our data yields estimates $M_3$ and $\sigma^2 M_1$.
+Performing regression yields estimates for $M_3$.
 It turns out that knowledge of $M_2$ and $M_3$ are sufficient to recover
@@
 
-%We note that the eigendecomposition of $M_2$ is not sufficient to
-%recover the $\beta_k$ because they are not necessarily orthogonal to
-%each other.
+% Full algorithm
+Now we are ready to state our full algorithm, which we call Spectral Experts
+(\algorithmref{sec:algo}).
+First, we perform three regressions to recover the compound parameters
+$M_1$ \refeqn{y1},
+$M_2$ \refeqn{y2}, and
+$M_3$ \refeqn{y3}.
+Since $M_2$ and $M_3$ both only have rank $k$,
+we can use nuclear norm regularization
+\cite{Tomioka2011,NegahbanWainwright2009}
+to exploit this low-rank structure and improve our compound parameter estimates.
+In the algorithm, the regularization strengths $\lambda_n^{(2)}$ and $\lambda_n^{(2)}$
+are set to $\frac{1}{\sqrt{n}}$.
+The resulting semidefinite program is a standard one which has received
+much attention in recent years.
+We use a rather simple gradient-based approach,
+where the nuclear norm is handled in closed form by taking an SVD
+and soft-thresholding the eigenvalues \cite{donoho95soft,cai10soft}.
 
-% Full algorithm
-Now we are ready to state our full algorithm, which we call Spectral Experts.
-The full algorithm is described in \algorithmref{sec:algo}.
-First, we perform two regressions according to \refeqn{y2}
-and \refeqn{y3} to recover the compound parameters $M_2$ and $M_3$.
-The first regression \refeqn{estimateM2}
-produces estimates of $\hat M_2$ and $\hat\sigma^2$;
-the second regression \refeqn{estimateM3} produces $\hat M_3$ (we ignore $\hat S_1$).
-Since $M_2$ and $M_3$ have low rank structure,
-we use nuclear norm regularization \cite{Tomioka2011,NegahbanWainwright2009}
-to exploit this structure.
-The regularization strengths $\lambda_n^{(2)}$ and $\lambda_n^{(2)}$
-are set to $O(\frac{1}{\sqrt{n}})$.
-See \sectionref{sec:regression} for more details.
+% Tensor factorization
+Having estimated the compound parameters $M_2$ and $M_3$,
+it remains to recover the original parameters $\theta$.
+\citet{AnandkumarGeHsu2012} showed that for $M_2$ and $M_3$ of
+the forms in \refeqn{Mp}, it is possible to efficiently accomplish this.
+Specifically, we first compute a whitening matrix based on the SVD of $M_2$
+and use that to construct a tensor $T = M_3(W, W, W)$ whose factors are orthogonal.
+We can use the robust tensor power method to compute all the
+eigenvalues and eigenvectors of $T$, from which it is easy to recover
+the parameters $\pi$ and $\{\beta_h\}$.
 
-Second, given a matrix $M_2$ and tensor $M_3$ of the form in \refeqn{Mp},
-\citet{AnandkumarGeHsu2012} showed that due to symmetry of $M_2$ and $M_3$,
-we can efficiently perform tensor factorization of the whitened
-version of $M_3$.  The eigenvalues and eigenvectors
-directly yield the parameters $\pi$ and
-$\{\beta_h\}$.
-See \sectionref{sec:tensor-power} for details.
-%\citet{AnandkumarGeHsu2012} showed that given $M_2$ and $M_3$
-%of the form in \refeqn{Mp},
-%the robust tensor power method can be employed to recover $\pi$ and
-%$\{\beta_h\}$.
+% Unmixing
+At a conceptual level, Spectral Experts bears some resemblence to the unmixing
+algorithm for estimation of restricted PCFGs
+\cite{hsu12identifiability}.
+In that work, the observations (moments) provided a linear combination over
+the compound parameters.  ``Unmixing'' involves solving for the compound
+parameters by inverting a mixing matrix.  In this work,
+each data point (appropriately transformed) provides a noisy slice of the
+compound parameters; retrieving the compound parameters is more involved.
+
+% Signal
+The idea of performing low-rank regression on $y^2$ has been explored
+in the context of signal recovery from magnitude measurements
+\cite{candes11phaselift,ohlsson12phase}.
+There, the actual observed response was $y^2$,
+whereas in our case, we delibrately construct powers $y,y^2,y^3$
+to identify the parameters.
 
@@
   \begin{algorithmic}[1]
-    \INPUT Data $\mathcal{D} = \{ (\xn{1}, \yn{1}), \cdots, (\xn{n}, \yn{n}) \}$.
-    \OUTPUT Parameters $\hat\theta = (\hat \pi, [\hat \beta_1 \mid \cdots \mid \hat \beta_k], \hat\sigma^2)$.
+    \INPUT Data $\mathcal{D}_r = \{ (\xn{1}, \yn{1}), \cdots, (\xn{n}, \yn{n}) \}$, $r = 1, 2, 3$.
+    \OUTPUT Parameters $\hat\theta = (\hat \pi, [\hat \beta_1 \mid \cdots \mid \hat \beta_k])$.
     \STATE Estimate compound parameters $M_2, M_3$ using \textbf{low-rank regression}:
     \begin{align}
-      &[\hat M_2, \hat \sigma^2] = \arg\min_{M_2,\sigma^2} \label{eqn:estimateM2} \\
-      &\quad\frac{1}{2n} \sum_{(x,y) \in \sD} (\innerp{M_2}{x\tp{2}} + \sigma^2 - y^2)^2 + \lambda_n^{(2)} \|M_2\|_*, \nonumber \\
-       &[\hat M_3, \hat S_1] = \arg\min_{M_3,S_1} \label{eqn:estimateM3} \\
-       &\quad\frac{1}{2n} \sum_{(x,y) \in \sD} (\innerp{M_3}{x\tp{3}} + \innerp{S_1}{x} - y^3)^2 + \lambda_n^{(3)} \|M_3\|_*. \nonumber
+      &\hat M_1 = \arg\min_{M_1} \label{eqn:estimateM1} \\
+      &\quad\frac{1}{2n} \sum_{(x,y) \in \sD_1} (\innerp{M_1}{x\tp{2}} - y)^2, \nonumber \\
+      &\hat M_2 = \arg\min_{M_2} \label{eqn:estimateM2} \\
+      &\quad\frac{1}{2n} \sum_{(x,y) \in \sD_2} (\innerp{M_2}{x\tp{2}} + \sigma^2 - y^2)^2 + \lambda_n^{(2)} \|M_2\|_*, \nonumber \\
+      &\hat M_3 = \arg\min_{M_3} \label{eqn:estimateM3} \\
+      &\quad\frac{1}{2n} \sum_{(x,y) \in \sD_3} (\innerp{M_3}{x\tp{3}} + 3 \sigma^2\innerp{\hat M_1}{x} - y^3)^2 + \lambda_n^{(3)} \|M_3\|_*. \nonumber
     \end{align}
-    \STATE Estimate the coefficients $\{\beta_h\}$ using \textbf{tensor factorzation}:
+    \STATE Estimate the parameters $\theta = (\pi, B)$ using \textbf{tensor factorzation}:
     \begin{enumerate}
@@
       \hat M_2 W = I$) using SVD.
-      \item [(b)] Compute the eigenvalues $\{\lambda_h\}_{h=1}^k$
+      \item [(b)] Compute the eigenvalues $\{a_h\}_{h=1}^k$
       and eigenvectors $\{v_h\}_{h=1}^k$
@@
       by using the robust tensor power method.
-    \item [(c)] Set parameter estimates $\hat\pi_h = \frac{1}{\lambda_h^2}$
-    and $\hat\beta_h = (W^{\top})^\dagger (\lambda_h v_h)$.
+    \item [(c)] Set parameter estimates $\hat\pi_h = \frac{1}{a_h^2}$
+    and $\hat\beta_h = (W^{\top})^\dagger (a_h v_h)$.
     \end{enumerate}
@@
 
-Our main theoretical result shows that Spectral Experts returns
-consistent parameter estimates (see Appendix
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Theoretical results}
+
+In this section, we conduct a theoretical analysis of Spectral Experts.
+In particular, we prove that it enjoys consistent parameter estimates.
 \begin{theorem}[Convergence of Spectral Experts]
-Let $\sD$ consist of $n$ i.i.d. points drawn from the mixture
-of linear regressions model with parameter $\theta^*$.
-Assume that $\E[[1,\unfold(x\tp{2})]\tp{2}] \succ 0$ 
-and $\E[[x,\unfold(x\tp{3})]\tp{2}] \succ 0$.
-Then the parameter estimates $\hat\theta$ returned by
-\algorithmref{algo:spectral-experts}, with the columns appropriate permuted,
-satisfies $\sigma^2 \le \epsilon$,
-$\|\hat\pi - \pi^*\|_{\infty} \le \epsilon$
+Each $\sD_p$ (for $p = 1, 2, 3$) consists of $n$ i.i.d. points drawn from the mixture
+of linear regressions model with parameter $\theta^*$
+(having three independent copies simplifies the analysis).
+Assume that $\E[\unfold(x\tp{2})\tp{2}] \succ 0$ 
+and $\E[\unfold(x\tp{3})\tp{2}] \succ 0$.
+Let $\|x\|_2 \le R$ with probability $1$
+and let $\|\beta_h\|_2 \le L$ for all $h \in [k]$.
+There exists universal constants $c_1,c_2$ such that
+if the number of samples is
+$n \ge \frac{c_1 k d^2 L R}{\epsilon}$
+and $\lambda_n^{(2)} \ge \frac{c_2 L R}{\sqrt{n}}$,
+and $\lambda_n^{(3)} \ge \frac{c_3 L R}{\sqrt{n}}$,
+then the parameter estimates $\hat\theta$ returned by
+\algorithmref{algo:spectral-experts} (with the columns appropriate permuted)
+satisfies 
+$\|\hat\pi - \pi^*\|_{\infty} \le \frac{\epsilon}{???}$
 and for all $h \in [k]$,
-$\|\hat\beta_h - \beta^*_h\| \le \epsilon$,
-for some error tolerance $\epsilon = O(\frac{kd^2}{n (\min_h \pi_h^*)})$.
+$\|\hat\beta_h - \beta^*_h\|_2 \le \epsilon ???$,
 \end{theorem}
@@
 By Theorem 5.1 of \cite{AnandkumarGeHsu2012},
+Need to convert error from $\epsilon$
+hit by $(W^\top)^\dagger$
 \end{proof}
-
-% Method of moments interpretation
-%Though we have motivated Spectral Experts using regression on powers
-%of $y$ given powers of $x$,
-%we can also think of the algorithm in terms of
-%the method of moments.  Specifically,
-%the optimality conditions for the regression of $M_2$ and $M_3$ are:
-%\begin{align}
-%\E[x\tp{2} (\innerp{M_2}{x\tp{2}} + \sigma^2 - y^2)] = 0, \\
-%\E[x\tp{3} (\innerp{M_3}{x\tp{3}} + 3 \sigma^2 \innerp{M_1}{x} - y^3)] = 0.
-%\end{align}
 
@@
 
-\subsection{Identifiability from moments}
+\subsection{Identifiability}
 
-The mixture of linear regressions model is identifiable subject to
-permutations of the components\citationneeded. However, our approach
-employs regression on $x\tp{2}$ and $x\tp{3}$, and thus requires
-$\mathcal{X} \otimes \mathcal{X}$ and $\mathcal{X} \otimes \mathcal{X}
-\otimes \mathcal{X}$ to be linearly independent modulo symmetry. We call
-these condition {\em quadratic} and {\em cubic} independence.
+In ordinary linear regression, the regression coefficients are identifiable
+if and only if the data has full rank: $\E[x\tp{2}] \succ 0$.
+However, in mixture of linear regressions, this is insufficient.  
+For example, suppose $x = (1, t, \cdots, t^p)$, 
+the common polynomial basis expansion.
+Regressing on all powers up to $p$ yield at most $3$.
 
-This condition trivially holds when the space $\mathcal{X}$ consists of
-independent vectors $\{ x_1, x_2, \cdots, x_d \}$. Unfortunately, it
-places some constraints on non-linear featurizations. For example, the
-common polynomial basis expansion, $1, x_1, x_1^2, \dots, x_1^p$ is
-neither quadratically nor cubically independent because the product
-space is strictly a subset of $\Re^{\frac{d (d-1)}{2}}$ since it
-contains $x_1 \times x_1 = x_1^2 \times 1$, and $x_1 \times x_1 \times
-x_1 = x_1^2 \times x_1 \times 1 = x_1^3 \times 1 \times 1$. In fact, the
-number of unique terms in $x\tp{p}$ is only of the order of $p \times
-d$, and hence we will not have enough equations to solve for $M_2$ or
-$M_3$ unless we consider $p = d^2$ moments.
+%$1, x_1, x_1^2, \dots, x_1^p$ is
+%neither quadratically nor cubically independent because the product
+%space is strictly a subset of $\Re^{\frac{d (d-1)}{2}}$ since it
+%contains $x_1 \times x_1 = x_1^2 \times 1$, and $x_1 \times x_1 \times
+%x_1 = x_1^2 \times x_1 \times 1 = x_1^3 \times 1 \times 1$.
+%In fact, the
+%number of unique terms in $x\tp{p}$ is only of the order of $p \times
+%d$, and hence we will not have enough equations to solve for $M_2$ or
+%$M_3$ unless we consider $p = d^2$ moments.
+
+%The mixture of linear regressions model is identifiable subject to
+%permutations of the components. However, our approach
+%employs regression on $x\tp{2}$ and $x\tp{3}$, and thus requires
+%$\mathcal{X} \otimes \mathcal{X}$ and $\mathcal{X} \otimes \mathcal{X}
+%\otimes \mathcal{X}$ to be linearly independent modulo symmetry. We call
+%these condition {\em quadratic} and {\em cubic} independence.
+
+%This condition trivially holds when the space $\mathcal{X}$ consists of
+%independent vectors $\{ x_1, x_2, \cdots, x_d \}$. Unfortunately, it
+%places some constraints on non-linear featurizations.
 

### 2013-02-15 doc/icml2013/regression.tex
checked in
many changes

@@
-\section{Recovering $M_2, M_3$ with low-rank regression}
+\subsection{Low-rank Regression}
 \label{sec:regression}
 
-In this section, we analyze step 1 of the Spectral Experts
-algorithm: that of recovering the compound parameters $M_2,M_3$
-from data.
+The goal of this section is to bound the error of
+the compound parameter estimates,
+specifically $\|\Delta_2\|_F^2$ and $\|\Delta_3\|_F^2$,
+where $\Delta_2 \eqdef \hat M_2 - M_2$
+and $\Delta_3 \eqdef \hat M_3 - M_3$.
+
+Our analysis is based on the low-rank regression framework of
+\citet{Tomioka2011} for tensors, which builds off of
+\citet{NegahbanWainwright2009} for matrices.
+Define the observation operator $\opX(M_p) : \Re^{d\tp{p}} \to \Re^{n}$
+mapping compound parameters $M_p$:
+\begin{align}
+  \opX(M_p; \sD)_i &\eqdef \frac12 \sum_{(x,y) \in \sD} \innerp{M_p}{x\tp{p}}_i, \quad i  \in [n].
+\end{align}
+
+Let $\kappa(\opX)$ be the restricted strong constant,
+and let $\opX^*(\eta_p; \sD) = \sum_{(x,y) \in \sD} \eta_p(x) x\tp{p}$
+be the adjoint.
+
+\paragraph{Restricted strong convexity}
+
+In this section, we compute the restricted strong convexity constant
+$\kappa(\opX)$ for our problem.
 
@@
 
-\cite{candesPhaseLift} regresses on the square of the response,
-because that's the observation.
+%Analysis for regression in the fixed and random design settings have
+%been studied before \citep{HsuKakadeZhang}, however our setup differs
+%substantially from the noise models assumed in the literature. In our
+%scenario, the variance in the estimation comes not only from the
+%Gaussian observation noise (which has been studied before), but also
+%from the variance in the latent variable $h$.
 
-Analysis for regression in the fixed and random design settings have
-been studied before \citep{HsuKakadeZhang}, however our setup differs
-substantially from the noise models assumed in the literature. In our
-scenario, the variance in the estimation comes not only from the
-Gaussian observation noise (which has been studied before), but also
-from the variance in the latent variable $h$.
+%Let us now formally define the class of regression problems we wish to
+%analyze, i.e. regression on the set $(x\tp{p}, y^p)$,
+%\begin{align*}
+%  y^p &= \innerp{x\tp{p}}{M_p} + (\innerp{x\tp{p}}{M_p - \beta_h\tp{p}} + \varepsilon).
+%\end{align*}
 
-Let us now formally define the class of regression problems we wish to
-analyze, i.e. regression on the set $(x\tp{p}, y^p)$,
-\begin{align*}
-  y^p &= \innerp{x\tp{p}}{M_p} + (\innerp{x\tp{p}}{M_p - \beta_h\tp{p}} + \varepsilon).
-\end{align*}
-Note that the term in brackets is the error in our estimates and has
-zero mean.  A more succinct way of representing this problem is to use
-the operator $\opX(M_p) : \Re^{d\tp{p}} \to \Re^{N}$, which uses the
-input $\{ x_i \}_{i=1}^{N}$ to map $M_p$ to a vector of responses $\hat
-y_i = \opX(M_p)_i = \innerp{x_i\tp{p}}{M_p}$. We can also represent the
-errors with the vector $\vec \epsilon$. Rewriting the equation with this notation, 
-\begin{align*}
-  \vec y^p &= \opX(M_p) + \vec \epsilon.
-\end{align*}
+%\todo{Describe/define the convex tensor stuff.}
+%We would like to exploit
+%the property that $M_p$ is low rank (as typically $K \ll D$). It has
+%been shown that a convex relaxation for this problem regression with
+%trace norm regularization, which can be solved using a proximal
+%subgradient descent algorithm\citationneeded.
+%The analogue of trace norm
+%regularization for higher order tensors corresponds to the sum of the
+%trace norms of the mode-k unfolding of the tensor \cite{Tomioka2011},
+%$X_{(k)}$, is a $d_k \times (\prod_{k' \neq k} d_{k'})$ matrix obtained
+%by concatenating the entries for all dimensions other than $k$. For
+%example, the 1-mode unfolding of a 3rd order tensor has entries,
+%$X_{(1)}^{i_1, (i_2, i_3)} = X_{i_1, i_2, i_3}$.
 
-\todo{Describe/define the convex tensor stuff.} We would like to exploit
-the property that $M_p$ is low rank (as typically $K \ll D$). It has
-been shown that a convex relaxation for this problem regression with
-trace norm regularization, which can be solved using a proximal
-subgradient descent algorithm\citationneeded. The analogue of trace norm
-regularization for higher order tensors corresponds to the sum of the
-trace norms of the mode-k unfolding of the tensor \cite{Tomioka2011},
-$X_{(k)}$, is a $d_k \times (\prod_{k' \neq k} d_{k'})$ matrix obtained
-by concatenating the entries for all dimensions other than $k$. For
-example, the 1-mode unfolding of a 3rd order tensor has entries,
-$X_{(1)}^{i_1, (i_2, i_3)} = X_{i_1, i_2, i_3}$.
+%In general, the optimization problem we'd like to solve is,
+%\begin{align}
+%  \hat M_p &= 
+%  \arg\min_{M_p}& \frac{1}{2N} \| \vec y - \opX(M_p) \|_2^2 + \frac{\lambda_n}{K} \sum_{h=1}^K \| (M_p)_{(h)} \|_* \label{eq:regression}.
+%\end{align}
 
-In general, the optimization problem we'd like to solve is,
-\begin{align}
-  \hat M_p &= 
-  \arg\min_{M_p}& \frac{1}{2N} \| \vec y - \opX(M_p) \|_2^2 + \frac{\lambda_n}{K} \sum_{h=1}^K \| (M_p)_{(h)} \|_* \label{eq:regression}.
-\end{align}
-
-We begin our analysis by stating a generic theorem about convex
-optimization problem in \equationref{eq:regression}, 
-\begin{lemma}(Convergence of Trace-Norm Regularized Regression)
-  Let $\hat M_p$ be the solution of the optimization problem
-  \eqref{eq:regression} with $\lambda_N \ge 2 \frac{\| \opX^*(\epsilon)
-  \|}{N}$. Suppose $\opX$ satisfies the restricted strong convexity
-  condition, $$\frac{1}{2N} \| \opX( \Delta )\|_2^2 \ge \kappa(\opX)
-  \|\Delta\|^2_F$$, then
-  $\| \hat M_p - M_p^* \|_F \le \frac{32 \lambda_N}{\kappa(\opX)} \frac{1}{K} \sum_{k=1}^{K} \sqrt{r_k}$.
+%We begin our analysis by stating a generic theorem about convex
+%optimization problem in \equationref{eq:regression}, 
+\begin{lemma}[convergence of low-rank regression]
+If restricted strong convexity holds
+$$\frac{1}{2n} \| \opX( \Delta )\|_2^2 \ge \kappa(\opX) \|\Delta\|^2_F \quad \text{and} \quad
+\lambda_n \ge \frac{|\opX^*(\eta_p)|}{n}.$$
+Then the error of $\hat M_p$ is bounded as follows:
+$\| \hat M_p - M_p^* \|_F \le \frac{\lambda_n \sqrt{k}}{\kappa(\opX)}$.
 \end{lemma}
@@
 
+By Lemma 19 of \cite{hsu13spherical},
+we have that $\frac{1}{n} \sum_{i=1}^n \epsilon_i^3 = O(\sqrt{\frac{\log^3(1/\delta)}{n}})$.
+with probability at least $1-\delta$.
+
 \begin{lemma}[bounded error]

### 2013-02-15 doc/icml2013/evaluation.tex
Added some figures and a table

--- 2013-02-14 doc/icml2013/evaluation.tex
@@
 
-\subsection{Setup}
+\todo{Fill it up with juicy bits once we actually have them graphs.}
+
+\subsection{Experimental Setup}
 
@@
 
-We ran expectation-maximization until the log likelihood converged
-$10^-5$. Initialized the entries of $\beta$ with a random Gaussian, $\normal{0}{1}$.
+We implemented the low rank recovery optimization routines for $M_2$ and
+$M_3$ using a proximal subgradient descent algorithm\citationneeded. The
+algorithm was initialized with the regularized least squares solution
+and was run to convergence in each instance. Both regularization
+parameters were annealed as $\frac{1}{T}$.
 
-In all our experiments, EM converged.
-
-Describe spectral (briefly pointing to algorithm);
-initialization, clipping.
-
-Describe spectral+EM
+As a baseline, we also implemented EM with two different
+initializations. For the first, we used $\beta$s with random standard
+Gaussian entries and uniform mixture probabilities, $\pi$. In the other,
+we initialized EM with the starting points given by Spectral Experts.
 
 \paragraph{Datasets}
-Describe the Curves dataset.
-$x = (1, t, t^4, t^9)$
-$d = b^p$
 
-Distribution over $x$.
+We generated synthetic data using covariates from the unit hyper cube
+and non-linear responses.  To conform to the identifiability criteria
+discussed in \sectionref{sec:algo}, we considered only features that
+would guarantee this condition.
 
-Canonical settings:
+As an example, one set of features we considered in the uni-dimensional
+setting were $\{1, x, x^4, x^7\}$. The data and the curves fit using
+Spectral Experts, EM with random initializations and EM initialized with
+the parameters recovered using spectral experts are shown in
+\figureref{fig:curves}.
 
-Default values:
-$n = 10^5$
-$k = 5$
+\begin{figure*}[t]
+  \centering
+  \subfigure[Spectral]{
+    \includegraphics[width=0.50\textwidth]{figures/curves-vis/1-8-3-specm.pdf}}
+    \hspace{-2em}
+  \subfigure[EM]{
+    \includegraphics[width=0.50\textwidth]{figures/curves-vis/1-8-3-em.pdf}}
+    \hspace{-2em}
+%  \subfigure[Spectral + EM]{
+%    \includegraphics[width=0.34\textwidth]{figures/curves-vis/1-8-3-spem.png}}
+  \caption{A comparison between SpectralExperts and EM. The dashed lines
+  on the left denote the solution recovered by Spectral Experts. Whilst
+  not a perfect fit, it provides an good initialization for EM to find
+  the true global minima. The dotted lines on the right show different
+  local optima found by EM on this example.}
+  \label{fig:curves}
+\end{figure*}
 
-Dataset 1: visualization ($b = 1, p = 3$) 
-Plot for visualization
+%Default values:
+%$n = 10^6$
+%$k = 5$
 
-Table on different settings: 
-  (To show that we work comparably or better to EM always)
-  (Spawned!)
+\begin{table*}[t]
+\caption{Parameter Recovery ($N = 500,000$).}
+\label{tbl:parameter-recovery}
+\vskip 0.15in
+\begin{center}
+\begin{small}
+\begin{sc}
 
-Dataset 2: $b = 30$, $p = 1$
-(Spawned many tasks to find spec performance. Will run EM on interesting cases)
+  \begin{tabular}{ >{\centering}p{5cm}<{\centering} r c c c }
+\hline
+\abovespace\belowspace
+Features  & $K$ & Spectral & EM & Spectral + EM \\
+\hline
+\abovespace
+$\{1$, $ x_1$, $ x_1^4\}$ 
+  & 2 & 2.54 $\pm$ 2.81 & 0.28 $\pm$ 0.82 & 0.13 $\pm$ 0.55 \\
+$\{1$, $ x_1$, $ x_2$, $ x_1^2 x_2^2\}$ 
+  & 3 & 3.93 $\pm$ 5.82 & 0.33 $\pm$ 0.96 & 0.35 $\pm$ 1.23 \\
+$\{1$, $ x_1$, $ x_2$, $ x_1 x_2^3$, $ x_1^2 x_2^2$, $ x_1^3 x_2 \}$ 
+  & 5 & 5.27 $\pm$ 2.32 & 1.80 $\pm$ 1.80 & 1.51 $\pm$ 1.77 \\
+$\{1$, $ x_1$, $ x_2$, $ x_1 x_2^3$, $ x_1^2 x_2^2$, $ x_1^3 x_2$, $ x_1^3 x_2^4$, $ x_1^4 x_2^3 \}$ 
+  & 7 & 8.42 $\pm$ 1.66 & 7.41 $\pm$ 2.99 & 7.31 $\pm$ 2.47 \\
+$\{1$, $ x_1$, $ x_2$, $ x_3$, $ x_1 x_2 x_3^2$, $ x_1 x_2^2 x_3$, $ x_1^2 x_2 x_3$, $ x_1^2 x_2^2$, $ x_1^2 x_3^2$, $ x_2^2 x_3^2$, $ x_3^2\}$
+  & 3 & 3.78 $\pm$ 0.90 & 0.67 $\pm$ 1.49 & 0.51 $\pm$ 1.12 \\
+$\{1$, $ x_1$, $ x_2$, $ x_3$, $ x_1 x_2 x_3^2$, $ x_1 x_2^2 x_3$, $ x_1^2 x_2 x_3$, $ x_1^2 x_2^2$, $ x_1^2 x_3^2$, $ x_2^2 x_3^2$, $ x_3^2\}$
+  & 7 & 9.97 $\pm$ 3.22 & 1.43 $\pm$ 1.97 & 1.42 $\pm$ 2.17 \\
+\hline
+\end{tabular}
+\end{sc}
+\end{small}
+\end{center}
+\vskip -0.1in
+\end{table*}
+
+\todo{Dataset 2}: $b = 30$, $p = 1$ [TODO]
 
@@
 
+\paragraph{Misspecified Data}
+
+To show how well spectral methods can handle mis-specification, we
+compared the performance of the different algorithms with entire
+segments removed from the generated data. \todo{Table X shows figures}.
+
 \paragraph{Effect on number of data points}
 
-Curves dataset 2.
-Plot: $n$ versus error for various algorithms
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=0.45\textwidth]{figures/vs-n/1-8-3-3.pdf}
+  \caption{}
+  \label{fig:vs-n}
+\end{figure}
+
+\todo{We see that the error of spectral is actually less than EM after a point.}
 
@@
 
-\paragraph{Effect of dimensionality}
-
 \paragraph{Effect of noise}

### 2013-02-15 doc/icml2013/intro.tex

@@
 have been widely applied to many tasks, including
-object recognition \cite{quattoni04nips},
-human action recognition \cite{wang09},
+object recognition \cite{quattoni04crf},
+human action recognition \cite{wang09crf},
 syntactic parsing \cite{petrov08discriminative},
 and machine translation \cite{liang06discrimative}.
-% http://www.cs.sfu.ca/~mori/research/papers/wang_cvpr09.pdf
-%Slav Petrov, Discriminative Log-Linear Grammars with Latent Variables
 However, parameter estimation in these models is difficult;
@@
 gradient descent) and are vulerable to local optima.
-%Our goal is to develop consistent estimators which sidestep
 
@@
 \emph{mixture of linear regressions} (MLG) \cite{VieleTong2002}.
-%mixture of linear regressions (\sectionref{sec:model}).
 
@@
 mixture models, HMMs \cite{AnandkumarHsuKakade2012},
-Latent Dirichlet Allocation \cite{AnandkumarLDA},
-and parsing models \cite{HsuKakadeLiang2012}.
+Latent Dirichlet Allocation \cite{anandkumar12lda},
+and parsing models \cite{hsu12identifiability}.
 The basic idea of this work is to express
@@
 tensor power method \cite{AnandkumarGeHsu2012}.
-%\cite{hsu09spectral}
 
@@
 apply the above tensor factorization techniques.
-Our solution is to employ low-rank linear regression \cite{NegahbanWainwright2009,Tomioka2011} on second
+Our solution is to employ low-rank linear regression
+\cite{NegahbanWainwright2009,Tomioka2011} on second
 and third powers of the response to first recover the appropriate symmetric
@@
 i_2, \ldots i_p} Y_{i_1, i_2, \ldots i_p}.$
-For
-$X \in (\Re^d)\tp{p}$ and $Y \in (\Re^d)\tp{(p+q)}$,% $p, q > 0$,
-$\innerp{X}{Y} \in \Re^{d^q}$ is the projection of $X$ onto the first
-$p$ modes of $Y$, i.e. $\innerp{X}{Y}_{i_1, \ldots, i_q}
-= \sum_{j_1, \ldots, j_p} X_{j_1, \ldots, j_p} Y_{j_1, \ldots, j_p, i_1,
-\ldots, i_q}$.
+For a $p$-th order tensor $X \in (\Re^d)\tp{p}$, let $X_{(i)} \in \Re^{d \times d^{p-1}}$
+be the mode-$i$ unfolding of $X$.
+%For $X \in (\Re^d)\tp{p}$ and $Y \in (\Re^d)\tp{(p+q)}$,% $p, q > 0$,
+%$\innerp{X}{Y} \in \Re^{d^q}$ is the projection of $X$ onto the first
+%$p$ modes of $Y$, i.e. $\innerp{X}{Y}_{i_1, \ldots, i_q}
+%= \sum_{j_1, \ldots, j_p} X_{j_1, \ldots, j_p} Y_{j_1, \ldots, j_p, i_1,
+%\ldots, i_q}$.
+
+% Norms
+For a matrix $X$,
+let $\|X\|_*$ denote the nuclear (trace) norm (sum of singular values),
+let $\|X\|_F$ denote the Frobenius norm (square root of sum of squares of singular values),
+let $\|X\|_\text{op}$ denote the operator norm (largest singular value).
+For a rank-$K$ tensor $X$,
+let $\|X\|_* = \frac{1}{p} \sum_{i=1}^p \|X_{(i)}\|_*$ denote
+the average nuclear norm over all $p$ unfoldings,
+and let $\|X\|_\text{op} = \frac{1}{p} \sum_{i=1}^p \|X_{(i)}\|_\text{op}$
+denote the average operator norm over all $p$ unfoldings.
+Let $\unfold(X)$ be the vectorized version of $X$.
+
 Let $[n] = \{ 1, \dots, n \}$ denote the first $n$ positive integers.
-%Similar to how $x^T A$ is
-%the projection of a vector $x$ onto the row space of a matrix $A$
-
-%The expected moments of the parameters $\beta_1, \dots, \beta_k$ play an
-%important role in our algorithm and we will use $M_p = \E[\beta\tp{p}]
-%= \sum_{k=1}^{K} \pi_k \beta_k^{\otimes p}$ to denote them.

### 2013-02-15 doc/icml2013/model.tex

@@
 variance $\sigma^2$.
-%Given $x$, the generative process for the response
-%variables is,
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-%where $h$ is a latent variable corresponding to the mode.
-The parameters of the model are $\theta = (\pi, B, \sigma^2)$,
-where $\pi \in \Re^d$ are the mixture proportions,
+The parameters of the model are $\theta = (\pi, B)$,
+where $\pi \in \Re^d$ are the mixture proportions and
 $B = [\beta_1 \mid \dots \mid \beta_K] \in \Re^{d \times K}$
-are the regression coefficients,
-and $\sigma^2$ is the variance of the observation noise.
+are the regression coefficients.
+We assume the variance of the observation noise $\sigma^2$ is known.
 The learning problem is stated as follows:
-given $N$ i.i.d. samples $(\xn{1}, \yn{1}), \dots, (\xn{N}, \yn{N})$
+given $n$ i.i.d. samples $(\xn{1}, \yn{1}), \dots, (\xn{n}, \yn{n})$
 drawn from the model with some unknown parameters $\theta^*$,
@@
 
-%and would like to
-%recover the regressors, $B = [\beta_1 \mid \beta_2 \mid \cdots \mid
-%\beta_K]$, and the mixture probabilities, $\pi = (\pi_1, \pi_2, \dots,
-%\pi_K)$.
-
-%Note that we do not observe $h$. If we did, the optimal
-%solution for $\beta_k$ would be the regressor found using least squares
-%on the set of points, $\{(x_i, y_i) | h_i = k \}$.
+% History of model
+The mixture of linear regressions model has been applied
+in the statistics literature for modeling music perception, where $x$ is the
+actual tone and $y$ is the tone perceived by a musician \cite{VieleTong2002}.
+% Reference mixture of experts.
+% Acknowledge that mixture proportions % can't depend on $x$
+The model is an instance of hierarchical mixture of experts
+\cite{jacobs91experts}, in which the mixture proportions are allowed to depend
+on $x$, known as a gating function.
+This dependence allow the experts to be localized in input space,
+providing more flexbility, but we do not consider this dependence in our model.
 
@@
 % non-convex.
-The estimation problem is difficult primarily because the mixture
-components $h$ are unobserved, rendering the marginal likelihood
-a non-convex function of $\theta$.
-Typically, the parameters are learned using
-expectation maximization (EM) \cite{VieleTong2002},
-which suffers from local optima
-(we show this empirically in \sectionref{sec:evaluation}).
+The estimation problem for mixture of linear regressions is difficult because
+the mixture components $h$ are unobserved,
+resulting in a non-convex log marginal likelihood.
+The parameters are typically learned using
+expectation maximization (EM) or Gibbs sampling \cite{VieleTong2002},
+which suffers from local optima.
+In the next section, we present a new algorithm
+that sidesteps the local optima problem entirely.
+%(we show this empirically in \sectionref{sec:evaluation}).
 %We provide several instances where this leads
 %to poor local optima in \sectionref{sec:evaluation}.
-
-This model was used for XXX.
-
-TODO: move to later
-% Reference mixture of experts.
-% Acknowledge that mixture proportions % can't depend on $x$
-The mixture of linear regressions model is an instance of
-mixture of experts \cite{jacobs91},
-%wherein the components can be SVMs, GPs, regression trees and other estimators.
-%In practice, it is
-in which the mixture proportions are allowed to depend on $x$.
-We have not considered these extensions in our treatment of the model.

### 2013-02-15 doc/icml2013/regression.tex

@@
 
-In this section, we compute the restricted strong convexity constant
-$\kappa(\opX)$ for our problem.
+In this section, we lower bound the restricted strong convexity constant
+$\kappa(\opX)$.
+
+\paragraph{Adjoint operator}
+
+In this section, we upper bound the operator norm of the adjoint
+$\|\opX(\eta_p)\|_\text{op}$.
 
@@
 
-%We begin our analysis by stating a generic theorem about convex
-%optimization problem in \equationref{eq:regression}, 
-\begin{lemma}[convergence of low-rank regression]
-If restricted strong convexity holds
+\begin{lemma}[\citet{Tomioka2011}, Theorem 1]
+Suppose there exists a restricted strong convexity constant $\kappa(\opX)$ such that
 $$\frac{1}{2n} \| \opX( \Delta )\|_2^2 \ge \kappa(\opX) \|\Delta\|^2_F \quad \text{and} \quad
@@
 \end{lemma}
-\begin{proof}
-  \citet{Tomioka2011}, Theorem 1.
-
-  This proof follows from convex properties \todo{elaborate}.
-  \citet{NegahbanWainwright2009} prove this theorem for matrices, i.e.
-  when $p = 2$. \citet{Tomioka2011} generalized it for arbitrary $p$.
-\end{proof}
 
@@
 parameter $\kappa(\opX)$ and upper bound the adjoint operator
-$\|\opX^*(\epsilon)\|_{2}^2$, which corresponds to a gradient update. We
-will appeal to the random design framework that models the input $x$ as
-random and show bounds that hold with high probability.
+$\|\opX^*(\epsilon)\|_{2}^2$.
+%We
+%will appeal to the random design framework that models the input $x$ as
+%random and show bounds that hold with high probability.
 
-\begin{lemma}[restricted strong convexity]
+\begin{lemma}[lower bound on restricted strong convexity]
+Let $\Sigma_p \eqdef \E[\unfold(x\tp{p})\tp{2}]$.
+With high probability at least $1-\delta$,
+for some constant $c$,
+$$\kappa(\opX) \ge c \sigma_\text{min}^p(\Sigma_p)$$,
+where
+%$\sigma_\text{min}^p(\Sigma_p) = \min_{M_p \in \sM_p : \|M_p\|_F \le 1} \unfold(M_p)^\top \Sigma_p \unfold(M_p)$.
+$\sigma_\text{min}^p(\Sigma_p) = \min_{M_p \in \sM_p : \|M_p\|_F \le 1} \trace(\unfold(M_p)^{\otimes 2} \Sigma_p)$.
+\end{lemma}
+Note that $\sigma_\text{min}$ 
+\begin{proof}
+Expanding the definition of the observation operator:
+$\frac{1}{2n} \|\opX(\Delta)\|_2^2
+= \frac{1}{2n} \sum_{(x,y) \in \sD} \innerp{\Delta}{x\tp{p}}^2$.
+Unfolding the tensor and letting $\hat\Sigma_p \eqdef \frac{1}{n} \sum_{(x,y) \in \sD} \unfold(x\tp{p})\tp{2}$,
+we have 
+$\frac{1}{2n} \|\opX(\Delta)\|_2^2
+\ge \frac{1}{2} \trace(\unfold(\Delta)\tp{2} \hat\Sigma_p)
+\ge \frac{1}{2} \|\Delta\|_F^2 \sigma_\text{min}^p(\hat\Sigma_p)$.
+
+%  \begin{align*}
+%    \frac{\|\opX(M_p)\|_1}{N} &= \frac{1}{N}\sum_{i=1}^{N} |\innerp{x_i\tp{p}}{M_p}| \\
+%    &\ge \frac{1}{N}\sum_{i=1}^{N} \sigma_{min}( M_p ) \|x_i\|^p \\
+%    &\ge \frac{\kappa(M_p)}{\sqrt{D}} \frac{\sum_{i=1}^{N} \|x_i\|^p}{N} \| M_p \|_F.
+%  \end{align*}
+%
+%  While $\kappa(M_p)$ depends on $M_p$ we can lower bound it by
+%  restricting the convex space to $\{ M_p | \kappa(M_p) \ge \kappa \}$.
+%  Further, using concentration bounds, we can lower bound $\|x_i\|^p$ as
+%  well. \todo{What is $\E[\|x_i\|^3]$? It is some quantity greater than
+%  zero}.
+%  
+%  Thus, 
+%  \begin{align*}
+%  \frac{ \|\opX(M_p)\|_2 }{2 N} 
+%  &\ge
+%  \frac{ \|\opX(M_p)\|_1 }{2 \sqrt{D}N} 
+%  &\ge \frac{\kappa}{2 D} (\E[\|x_i\|^p] - \frac{t}{N}) \| M_p \|_F.
+%  \end{align*}
+%  with high probability $1 - \delta$. \todo{The bound needs to be shown
+%  for $\| \opX(M_p) \|_2^2$.}
+%  
+%  In other words, $\kappa(\opX)
+%  = \frac{\kappa}{2 D} (\E[\|x_i\|^p] - \frac{t}{N})$ is a valid
+%  lower bound.
+\end{proof}
+
+\begin{lemma}[bounded error]
   Let $\opX$ be the linear operator previously defined. Then,
-$$\kappa(\opX) \ge Y$$.
+  $$\|\opX^*(\epsilon)\|_\text{op} \le Y$$.
 \end{lemma}
 \begin{proof}
-  \todo{Do I need to transform rewrite $M_p$ in terms of orthogonal components?}
-
-  \begin{align*}
-    \frac{\|\opX(M_p)\|_1}{N} &= \frac{1}{N}\sum_{i=1}^{N} |\innerp{x_i\tp{p}}{M_p}| \\
-    &\ge \frac{1}{N}\sum_{i=1}^{N} \sigma_{min}( M_p ) \|x_i\|^p \\
-    &\ge \frac{\kappa(M_p)}{\sqrt{D}} \frac{\sum_{i=1}^{N} \|x_i\|^p}{N} \| M_p \|_F.
-  \end{align*}
-
-  While $\kappa(M_p)$ depends on $M_p$ we can lower bound it by
-  restricting the convex space to $\{ M_p | \kappa(M_p) \ge \kappa \}$.
-  Further, using concentration bounds, we can lower bound $\|x_i\|^p$ as
-  well. \todo{What is $\E[\|x_i\|^3]$? It is some quantity greater than
-  zero}.
-
-  Thus, 
-  \begin{align*}
-  \frac{ \|\opX(M_p)\|_2 }{2 N} 
-  &\ge
-  \frac{ \|\opX(M_p)\|_1 }{2 \sqrt{D}N} 
-  &\ge \frac{\kappa}{2 D} (\E[\|x_i\|^p] - \frac{t}{N}) \| M_p \|_F.
-  \end{align*}
-  with high probability $1 - \delta$. \todo{The bound needs to be shown
-  for $\| \opX(M_p) \|_2^2$.}
-
-  In other words, $\kappa(\opX)
-  = \frac{\kappa}{2 D} (\E[\|x_i\|^p] - \frac{t}{N})$ is a valid
-  lower bound.
-\end{proof}
+%  \begin{align*}
+%    \frac{\opX^*(\epsilon)}{N} 
+%    &= \frac{1}{N}\sum_{i=1}^{N} \epsilon_i \innerp{x_i\tp{p}}{M_p} \\
+%    &= \frac{1}{N}\sum_{i=1}^{N} (\innerp{x_i\tp{p}}{M_p - \beta_i\tp{p}} + \varepsilon)  \innerp{x_i\tp{p}}{M_p}.
+%  \end{align*}
+%  Let $M_p - \beta_i\tp{p} = \tilde M_i$ and consider it to be a random
+%  variable.
+%  \begin{align*}
+%    \frac{\opX^*(\epsilon)}{N} 
+%    &= \frac{1}{N}\sum_{i=1}^{N} \innerp{\tilde M_i}{x_i\tp{2p}} + \varepsilon \innerp{M_p}{x_i\tp{p}} \\
+%    &\le \frac{1}{N}\sum_{i=1}^{N} \|\tilde M_i\|_{op} \|x_i\|^{2p} + \varepsilon \|M_p\|_{op}\|x_i\|^{p} \\
+%    &\le \|M_p\|_{op} ( \frac{1}{N}\sum_{i=1}^{N} \|x_i\|^{2p} + \varepsilon \|x_i\|^{p} ).
+%  \end{align*}
+%
+%  We can probabilistically bound the above by considering the union
+%  bound of $\varepsilon$ and $\|x_i\|^p$.
+%
+%
+%  Finally,
+%  \begin{align*}
+%    \frac{\opX^*(\epsilon)}{N} 
+%    &\le \|M_p\|_{op} (\E[\|x_i\|^{2p}] + t/N + (t/N) (\E[ \|x_i\|^{p} ] + t/N) ),
+%  \end{align*}
+%  with high probability. Note that this does not converge to $0$ as $N
+%  \to \infty$ because of the variance of the selection probability.
 
@@
 
-\begin{lemma}[bounded error]
-  Let $\opX$ be the linear operator previously defined. Then,
-  $$\opX^*(\epsilon) \le Y$$.
-\end{lemma}
-\begin{proof}
-  \begin{align*}
-    \frac{\opX^*(\epsilon)}{N} 
-    &= \frac{1}{N}\sum_{i=1}^{N} \epsilon_i \innerp{x_i\tp{p}}{M_p} \\
-    &= \frac{1}{N}\sum_{i=1}^{N} (\innerp{x_i\tp{p}}{M_p - \beta_i\tp{p}} + \varepsilon)  \innerp{x_i\tp{p}}{M_p}.
-  \end{align*}
-  Let $M_p - \beta_i\tp{p} = \tilde M_i$ and consider it to be a random
-  variable.
-  \begin{align*}
-    \frac{\opX^*(\epsilon)}{N} 
-    &= \frac{1}{N}\sum_{i=1}^{N} \innerp{\tilde M_i}{x_i\tp{2p}} + \varepsilon \innerp{M_p}{x_i\tp{p}} \\
-    &\le \frac{1}{N}\sum_{i=1}^{N} \|\tilde M_i\|_{op} \|x_i\|^{2p} + \varepsilon \|M_p\|_{op}\|x_i\|^{p} \\
-    &\le \|M_p\|_{op} ( \frac{1}{N}\sum_{i=1}^{N} \|x_i\|^{2p} + \varepsilon \|x_i\|^{p} ).
-  \end{align*}
-
-  We can probabilistically bound the above by considering the union
-  bound of $\varepsilon$ and $\|x_i\|^p$.
-
-  Finally,
-  \begin{align*}
-    \frac{\opX^*(\epsilon)}{N} 
-    &\le \|M_p\|_{op} (\E[\|x_i\|^{2p}] + t/N + (t/N) (\E[ \|x_i\|^{p} ] + t/N) ),
-  \end{align*}
-  with high probability. Note that this does not converge to $0$ as $N
-  \to \infty$ because of the variance of the selection probability.
 \end{proof}

### 2013-02-15 doc/icml2013/evaluation.tex
Now featuring more graphs

@@
     \includegraphics[width=0.50\textwidth]{figures/curves-vis/1-8-3-em.pdf}}
-    \hspace{-2em}
 %  \subfigure[Spectral + EM]{
@@
 
-\todo{Dataset 2}: $b = 30$, $p = 1$ [TODO]
-
-\paragraph{Motion tracking data?}
+\todo{Dataset 2: $b = 30$, $p = 1$. I'm unclear as to what we can show
+on this sort of data set. EM works extremely well, and the spectral
+methods do not converge easily.}
 
@@
 
-Using canonical settings, on the various datasets
+Using canonical settings, on the various datasets: \tableref{tbl:parameter-recovery}.
 
-Plot the parameters: the curves, three algorithms (DONE)
-
-Plot: the histogram over errors (for EM?)
+Plot: the histogram over errors (for EM?) \todo{Some plots I have doesn't really show much}.
 
@@
 
-\begin{figure}[t]
+\begin{figure*}[t]
   \centering
-  \includegraphics[width=0.45\textwidth]{figures/vs-n/1-8-3-3.pdf}
+  \subfigure[Complete Data]{
+    \includegraphics[width=0.50\textwidth]{figures/vs-n/1-8-3-3.pdf}
+  }
+    \hspace{-2em}
+  \subfigure[Incomplete Data]{
+    \includegraphics[width=0.50\textwidth]{figures/vs-n/1-8-3-3-rm.pdf}
+  }
   \caption{}
   \label{fig:vs-n}
-\end{figure}
+\end{figure*}
 

### 2013-02-15 doc/icml2013/regression.tex
cvec notation

@@
 \citet{NegahbanWainwright2009} for matrices.
+The main challenge is to control the noise $\eta_p(x)$,
+which in addition to containing the usual Gaussian part, is also dependent on
+$x$, involves the mixing noise and various cross terms.
+We first set up some notation to all three regressions (\refeqn{estimateM1}, \refeqn{estimateM2}, and \refeqn{estimateM3}).
 Define the observation operator $\opX(M_p) : \Re^{d\tp{p}} \to \Re^{n}$
@@
 $$\frac{1}{2n} \| \opX( \Delta )\|_2^2 \ge \kappa(\opX) \|\Delta\|^2_F \quad \text{and} \quad
-\lambda_n \ge \frac{|\opX^*(\eta_p)|}{n}.$$
+\lambda_n \ge \frac{\|\opX^*(\eta_p)\|_2}{n}.$$
 Then the error of $\hat M_p$ is bounded as follows:
@@
 \begin{lemma}[lower bound on restricted strong convexity]
-Let $\Sigma_p \eqdef \E[\unfold(x\tp{p})\tp{2}]$.
+Let $\Sigma_p \eqdef \E[\cvec(x\tp{p})\tp{2}]$.
 With high probability at least $1-\delta$,
@@
 where
-%$\sigma_\text{min}^p(\Sigma_p) = \min_{M_p \in \sM_p : \|M_p\|_F \le 1} \unfold(M_p)^\top \Sigma_p \unfold(M_p)$.
-$\sigma_\text{min}^p(\Sigma_p) = \min_{M_p \in \sM_p : \|M_p\|_F \le 1} \trace(\unfold(M_p)^{\otimes 2} \Sigma_p)$.
+%$\sigma_\text{min}^p(\Sigma_p) = \min_{M_p \in \sM_p : \|M_p\|_F \le 1} \cvec(M_p)^\top \Sigma_p \cvec(M_p)$.
+$\sigma_\text{min}^p(\Sigma_p) = \min_{M_p \in \sM_p : \|M_p\|_F \le 1} \trace(\cvec(M_p)^{\otimes 2} \Sigma_p)$.
 \end{lemma}
@@
 = \frac{1}{2n} \sum_{(x,y) \in \sD} \innerp{\Delta}{x\tp{p}}^2$.
-Unfolding the tensor and letting $\hat\Sigma_p \eqdef \frac{1}{n} \sum_{(x,y) \in \sD} \unfold(x\tp{p})\tp{2}$,
+Unfolding the tensor and letting $\hat\Sigma_p \eqdef \frac{1}{n} \sum_{(x,y) \in \sD} \cvec(x\tp{p})\tp{2}$,
 we have 
 $\frac{1}{2n} \|\opX(\Delta)\|_2^2
-\ge \frac{1}{2} \trace(\unfold(\Delta)\tp{2} \hat\Sigma_p)
+\ge \frac{1}{2} \trace(\cvec(\Delta)\tp{2} \hat\Sigma_p)
 \ge \frac{1}{2} \|\Delta\|_F^2 \sigma_\text{min}^p(\hat\Sigma_p)$.

### 2013-02-15 doc/icml2013/evaluation.tex
Updated experiments

@@
 and was run to convergence in each instance. Both regularization
-parameters were annealed as $\frac{1}{T}$.
+parameters, $\lambda_n^{(2)}$ and, $\lambda_n^{(3)}$ were annealed as
+$\frac{1}{n}$.
 
 As a baseline, we also implemented EM with two different
-initializations. For the first, we used $\beta$s with random standard
+initializations. In the first, we used $\beta$s with random standard
 Gaussian entries and uniform mixture probabilities, $\pi$. In the other,
@@
 $\{1$, $ x_1$, $ x_1^4\}$ 
-  & 2 & 2.54 $\pm$ 2.81 & 0.28 $\pm$ 0.82 & 0.13 $\pm$ 0.55 \\
+  & 2 & 1.52 $\pm$ 0.71 & 0.28 $\pm$ 0.82 & {\bf 0.13 $\pm$ 0.55} \\
 $\{1$, $ x_1$, $ x_2$, $ x_1^2 x_2^2\}$ 
-  & 3 & 3.93 $\pm$ 5.82 & 0.33 $\pm$ 0.96 & 0.35 $\pm$ 1.23 \\
+  & 3 & 1.87 $\pm$ 1.20 & {\bf 0.33 $\pm$ 0.96} & 0.35 $\pm$ 1.23 \\
 $\{1$, $ x_1$, $ x_2$, $ x_1 x_2^3$, $ x_1^2 x_2^2$, $ x_1^3 x_2 \}$ 
-  & 5 & 5.27 $\pm$ 2.32 & 1.80 $\pm$ 1.80 & 1.51 $\pm$ 1.77 \\
+& 5 & 5.27 $\pm$ 2.32 & 1.80 $\pm$ 1.80 & {\bf 1.51 $\pm$ 1.77} \\
 $\{1$, $ x_1$, $ x_2$, $ x_1 x_2^3$, $ x_1^2 x_2^2$, $ x_1^3 x_2$, $ x_1^3 x_2^4$, $ x_1^4 x_2^3 \}$ 
-  & 7 & 8.42 $\pm$ 1.66 & 7.41 $\pm$ 2.99 & 7.31 $\pm$ 2.47 \\
+& 7 & 8.42 $\pm$ 1.66 & 7.41 $\pm$ 2.99 & {\bf 7.31 $\pm$ 2.47} \\
 $\{1$, $ x_1$, $ x_2$, $ x_3$, $ x_1 x_2 x_3^2$, $ x_1 x_2^2 x_3$, $ x_1^2 x_2 x_3$, $ x_1^2 x_2^2$, $ x_1^2 x_3^2$, $ x_2^2 x_3^2$, $ x_3^2\}$
-  & 3 & 3.78 $\pm$ 0.90 & 0.67 $\pm$ 1.49 & 0.51 $\pm$ 1.12 \\
+& 3 & 3.78 $\pm$ 0.90 & 0.67 $\pm$ 1.49 & {\bf 0.51 $\pm$ 1.12} \\
 $\{1$, $ x_1$, $ x_2$, $ x_3$, $ x_1 x_2 x_3^2$, $ x_1 x_2^2 x_3$, $ x_1^2 x_2 x_3$, $ x_1^2 x_2^2$, $ x_1^2 x_3^2$, $ x_2^2 x_3^2$, $ x_3^2\}$
-  & 7 & 9.97 $\pm$ 3.22 & 1.43 $\pm$ 1.97 & 1.42 $\pm$ 2.17 \\
+  & 7 & 9.97 $\pm$ 3.22 & 1.43 $\pm$ 1.97 & {\bf 1.42 $\pm$ 2.17} \\
 \hline
@@
 
-\todo{Dataset 2: $b = 30$, $p = 1$. I'm unclear as to what we can show
-on this sort of data set. EM works extremely well, and the spectral
-methods do not converge easily.}
+% \todo{Dataset 2: $b = 30$, $p = 1$. I'm unclear as to what we can show
+% on this sort of data set. EM works extremely well, and the spectral
+% methods do not converge easily.}
 
@@
 
-\paragraph{Basic results}
+\tableref{tbl:parameter-recovery} presents the Fro\"ebenius norm of the
+difference between true and estimated parameters for the model, averaged
+over 20 different random instances for each feature set and 10 attempts
+for each instance. The experiments were run using $N = 500,000$ samples.
 
-Punchline: spectral by itself is medicore, offers a good initialization to EM.
-EM with random initialization is much less stable. (by table)
+There is a lot of variation across random instances; some are easy for
+EM to find the global minima and others more difficult. Moreover, simply
+increasing the number of samples did not necessarily improve EM
+performance. In general, we found that while our spectral algorithm did
+not recover parameters extremely well, it provided a good initialization
+for EM.
 
-Using canonical settings, on the various datasets: \tableref{tbl:parameter-recovery}.
-
-Plot: the histogram over errors (for EM?) \todo{Some plots I have doesn't really show much}.
+\todo{Plot: the histogram over errors (for EM?) Some plots I have
+doesn't really show much}.
 
@@
 
-To show how well spectral methods can handle mis-specification, we
-compared the performance of the different algorithms with entire
-segments removed from the generated data. \todo{Table X shows figures}.
+To evaluate how robust the algorithm was to model mis-specification, we
+removed large contiguous sections from $x \in [-0.5,-0.25] \cup
+[0.25,0.5]$. The lower half of \tableref{tbl:parameter-recovery} reports
+recovery errors in this scenario.
 
-\paragraph{Effect on number of data points}
+\paragraph{Effect of number of data points}
 
@@
 
+Highlight
+
 \todo{We see that the error of spectral is actually less than EM after a point.}

### 2013-02-15 doc/icml2013/algo.tex

@@
 Again, we have expressed $y^2$ has a linear function of $x\tp{2}$
-with regression coefficients $M_2$, plus a known bias $\sigma^2$ and noise.
+with regression coefficients $M_2$, plus a known bias $\sigma^2$ and noise.\footnote{If $\sigma^2$ were not known,
+we could treat it as another coefficient
+to be estimated.  The coefficients $M_2$ and $\sigma^2$ can be estimated jointly
+provided that $x$ does not already contain a bias ($x_j$ must be non-constant for every $j \in [d]$).}
 Importantly, the noise has mean zero; 
@@
 
+\paragraph{Remarks}
+
+% Spectral
+In recent years, there has a been substantial interest in ``spectral'' methods
+for learning latent-variable models.  One line of work has
+focused on observerable operator models \cite{hsu09spectral}
+in which the true parameters are not recovered, but a reparametrization is used
+which enables prediction.
+Another line of work is based on the method of moments and uses eigendecomposition of a certain tensor
+to recover the parameters \cite{anandkumar12moments,anandkumar12svd,AnandkumarHsuKakade2012}.
+Our work extends this line of work into the space of discriminative models,
+which requires regression to obtain the desired tensor.
+
 % Unmixing
-At a conceptual level, Spectral Experts bears some resemblence to the unmixing
+In spirit, Spectral Experts also bears some resemblence to the unmixing
 algorithm for estimation of restricted PCFGs
@@
   \begin{algorithmic}[1]
-    \INPUT Data $\mathcal{D}_r = \{ (\xn{1}, \yn{1}), \cdots, (\xn{n}, \yn{n}) \}$, $r = 1, 2, 3$.
+    \INPUT Three independent datasets $\mathcal{D}_r = \{ (\xn{1}, \yn{1}), \cdots, (\xn{n}, \yn{n}) \}$ for $r = 1, 2, 3$;
+    regularization strengths $\lambda_n^{(2)} = \frac{c_2}{\sqrt{n}}$, $\lambda_n^{(3)} = \frac{c_3}{\sqrt{n}}$;
+    observation noise variance $\sigma^2$.
     \OUTPUT Parameters $\hat\theta = (\hat \pi, [\hat \beta_1 \mid \cdots \mid \hat \beta_k])$.
@@
       &\hat M_1 = \arg\min_{M_1} \label{eqn:estimateM1} \\
-      &\quad\frac{1}{2n} \sum_{(x,y) \in \sD_1} (\innerp{M_1}{x\tp{2}} - y)^2, \nonumber \\
-      &\hat M_2 = \arg\min_{M_2} \label{eqn:estimateM2} \\
-      &\quad\frac{1}{2n} \sum_{(x,y) \in \sD_2} (\innerp{M_2}{x\tp{2}} + \sigma^2 - y^2)^2 + \lambda_n^{(2)} \|M_2\|_*, \nonumber \\
-      &\hat M_3 = \arg\min_{M_3} \label{eqn:estimateM3} \\
-      &\quad\frac{1}{2n} \sum_{(x,y) \in \sD_3} (\innerp{M_3}{x\tp{3}} + 3 \sigma^2\innerp{\hat M_1}{x} - y^3)^2 + \lambda_n^{(3)} \|M_3\|_*. \nonumber
+      &\quad\quad\frac{1}{2n} \sum_{(x,y) \in \sD_1} (\innerp{M_1}{x} - y)^2, \nonumber \\
+      &\hat M_2 = \arg\min_{M_2} \quad \lambda_n^{(2)} \|M_2\|_* + \label{eqn:estimateM2} \\
+      &\quad\quad\frac{1}{2n} \sum_{(x,y) \in \sD_2} (\innerp{M_2}{x\tp{2}} + \sigma^2 - y^2)^2, \nonumber \\
+      &\hat M_3 = \arg\min_{M_3} \quad \lambda_n^{(3)} \|M_3\|_* + \label{eqn:estimateM3} \\
+      &\quad\quad\frac{1}{2n} \sum_{(x,y) \in \sD_3} (\innerp{M_3}{x\tp{3}} + 3 \sigma^2\innerp{\hat M_1}{x} - y^3)^2. \nonumber
     \end{align}
@@
     \begin{enumerate}
-      \item [(a)] Compute whitening matrix $W \in \Re^{d \times k}$ (such that $W^\top
-      \hat M_2 W = I$) using SVD.
-      \item [(b)] Compute the eigenvalues $\{a_h\}_{h=1}^k$
-      and eigenvectors $\{v_h\}_{h=1}^k$
-      of the whitened tensor $M_3(W, W, W) \in \Re^{k \times k \times k}$
+      \item [(a)] Compute whitening matrix $\hat W \in \Re^{d \times k}$ (such that $\hat W^\top
+      \hat M_2 \hat W = I$) using SVD.
+      \item [(b)] Compute the eigenvalues $\{\hat a_h\}_{h=1}^k$
+      and eigenvectors $\{\hat v_h\}_{h=1}^k$
+      of the whitened tensor $\hat M_3(\hat W, \hat W, \hat W) \in \Re^{k \times k \times k}$
       by using the robust tensor power method.
-    \item [(c)] Set parameter estimates $\hat\pi_h = \frac{1}{a_h^2}$
-    and $\hat\beta_h = (W^{\top})^\dagger (a_h v_h)$.
+    \item [(c)] Set parameter estimates $\hat\pi_h = \frac{1}{\hat a_h^2}$
+    and $\hat\beta_h = (\hat W^{\top})^\dagger (\hat a_h \hat v_h)$.
     \end{enumerate}
@@
 
+%Define $\sM_p = \{ \sum_{h=1}^k \pi_h \beta_h\tp{p} : \pi_h \in \R^k, \beta_h
+%\in \R^d \}$ be the set of all $p$-th order compound parameters.
+%Let $\bM_p$ be the set of
+
+%Define the symmetric rank
+%\begin{align}
+%\E[\cvec(x\tp{p})\tp{2}] \succ 0.
+%\end{align}
+
 In this section, we conduct a theoretical analysis of Spectral Experts.
@@
 \begin{theorem}[Convergence of Spectral Experts]
+  \label{thm:convergence}
 Each $\sD_p$ (for $p = 1, 2, 3$) consists of $n$ i.i.d. points drawn from the mixture
@@
 (having three independent copies simplifies the analysis).
-Assume that $\E[\unfold(x\tp{2})\tp{2}] \succ 0$ 
-and $\E[\unfold(x\tp{3})\tp{2}] \succ 0$.
+Assume that $\E[\cvec(x\tp{p})\tp{2}] \succ 0$ for each $p \in \{1,2,3\}$.
+Assume $d \ge k$.
 Let $\|x\|_2 \le R$ with probability $1$
@@
 
-\subsection{Identifiability}
+\subsection{Identifiability from moments}
 
-In ordinary linear regression, the regression coefficients are identifiable
-if and only if the data has full rank: $\E[x\tp{2}] \succ 0$.
-However, in mixture of linear regressions, this is insufficient.  
-For example, suppose $x = (1, t, \cdots, t^p)$, 
-the common polynomial basis expansion.
-Regressing on all powers up to $p$ yield at most $3$.
+In ordinary linear regression, the regression coefficients $\beta \in \Re^d$ are
+identifiable if and only if the data has full
+rank: $\E[x\tp{2}] \succ 0$,
+and furthermore, identifying $\beta$ requires only moments
+$\E[xy]$ and $\E[x\tp{2}]$ (by observing the optimality conditions for \refeqn{y1}).
+However, in mixture of linear regressions, these two moments only allow us to recover $M_1$.
+\refthm{convergence} shows that if we have the higher order analogues,
+$\E[x\tp{p}y\tp{p}]$ and $\E[x\tp{2p}]$ for $p \in \{1,2,3\}$,
+we can then identify the parameters $\theta = (\pi, B)$,
+provided $\E[\cvec(x\tp{p})\tp{2}] \succ 0$ for $p \in \{1,2,3\}$.
+
+We note that the positive definite constraints are more stringent.
+For example, suppose $x = (1, t, t^2)$,
+the common polynomial basis expansion, so that all the coordinates are deterministically related.
+Then $\E[x\tp{2}] \succ 0$ can easily be satisfied, say with $x \sim \normal{0}{I}$.
+However, one can verify that $\E[\cvec(x\tp{2})\tp{2}]$ is singular,
+since $\cvec(x\tp{2}) = [1 \cdot 1, t\cdot t, 2(1 \cdot t^2), 2(t \cdot t^2), (t^2 \cdot t^2)]$ contains
+components $t \cdot t$ and $2(1 \cdot t^2)$, which are linearly dependent.
+So our method would not be able to identify the parameters of a mixture of
+linear regressions for this distribution over $x$.
+
+We now show that some amount of unidentifiability is intrinsic to the model,
+not just our estimation procedure.
+Suppose $x = (t, \dots, t^d)$.
+Even if we observed all moments
+$\E[x\tp{p}y\tp{p}]$ and $\E[x\tp{2p}]$ for $p \in [r]$,
+all the coordinates would be monomials up to degree $2dr$,
+the therefore the moments live in a $2(dr)$-dimensional subspace.
+The parameters live in a subspace of at least dimension $dk$.
+Therefore, at least $r \ge k/2$ moments are required for identifiability of any
+algorithm for this monomial example.
+
+%In fact, considering moments up to power $p
+%Regressing on all powers up to $p$ yield at most $3$.
 

### 2013-02-15 doc/icml2013/intro.tex

@@
 
+Let $[n] = \{ 1, \dots, n \}$ denote the first $n$ positive integers.
+
 We use $x\tp{p}$ to represent the $p$-th order tensor
-formed by taking the outer product of $x \in \Re^d$; i.e. $x\tp{p}_{i_1 i_2 \ldots
-i_p} = x_{i_1} x_{i_2} \cdots x_{i_p}$. We will use
+formed by taking the outer product of $x \in \Re^d$; i.e. $x\tp{p}_{i_1 \ldots
+i_p} = x_{i_1} \cdots x_{i_p}$. We will use
 $\innerp{\cdot}{\cdot}$ to denote the generalized dot product between two $p$-th
-order tensors: $\innerp{X}{Y} = \sum_{i_1, i_2, \ldots i_p} X_{i_1,
-i_2, \ldots i_p} Y_{i_1, i_2, \ldots i_p}.$
+order tensors: $\innerp{X}{Y} = \sum_{i_1, \ldots i_p} X_{i_1,
+\ldots i_p} Y_{i_1, \ldots i_p}.$
+A tensor $X$ is symmetric if for all $i,j \in [d]^p$ which are permutations,
+$X_{i_1 \cdots i_p}$ = $X_{j_1 \cdots j_p}$ (all tensors in this paper will be symmetric).
 For a $p$-th order tensor $X \in (\Re^d)\tp{p}$, let $X_{(i)} \in \Re^{d \times d^{p-1}}$
@@
 denote the average operator norm over all $p$ unfoldings.
-Let $\unfold(X)$ be the vectorized version of $X$.
 
-Let $[n] = \{ 1, \dots, n \}$ denote the first $n$ positive integers.
+For a tensor $X \in (\Re^d)\tp{p}$,
+let $\cvec(X) \in \Re^{\sum_{i=1}^p \binom{n}{i}}$ be
+the collapsed vectorization of $X$: each component of $\cvec(X)$
+is indexed by a vector of counts $(c_1, \dots, c_d)$ with total sum $\sum_i c_i = p$.
+The value of that component is
+$\sum_{k \in K(c)} X_{k_1 \cdots k_p}$,
+where $K(c) = \{ k \in [d]^p : \forall i \in [d], c_i = |\{ j \in [p] : k_j = i \}| \}$ are the set of index vectors $k$ with that count profile.
+For example, for $X \in \Re^{d \times d}$,
+$\cvec(X) = (X_{ii} : i \in [d]; X_{ij} + X_{ji} : i,j \in [d], i<j)$.
+For symmetric tensors, $\innerp{X}{Y} = \innerp{\cvec(X)}{\cvec(Y)}$.

### 2013-02-15 doc/icml2013/algo.tex
theory done

@@
 y^2 &= \innerp{M_2}{x\tp{2}} + \sigma^2 + \eta_2(x), \label{eqn:y2} \\
-\eta_2(x) &= \innerp{\beta_h\tp{2} - M_2}{x\tp{2}} + 2 \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}} + (\epsilon^2 - \sigma^2). \nonumber
+\eta_2(x) &= \innerp{\beta_h\tp{2} - M_2}{x\tp{2}} + 2 \epsilon \innerp{\beta_h}{x} + (\epsilon^2 - \sigma^2). \nonumber
 \end{align}
@@
 (having three independent copies simplifies the analysis).
-Assume that $\E[\cvec(x\tp{p})\tp{2}] \succ 0$ for each $p \in \{1,2,3\}$.
-Assume $d \ge k$.
+Assume that $\E[\cvec(x\tp{p})\tp{2}] \succ 0$ for each $p \in \{1,2,3\}$
+and that $d \ge k$.
 Let $\|x\|_2 \le R$ with probability $1$
 and let $\|\beta_h\|_2 \le L$ for all $h \in [k]$.
-There exists universal constants $c_1,c_2$ such that
-if the number of samples is
-$n \ge \frac{c_1 k d^2 L R}{\epsilon}$
-and $\lambda_n^{(2)} \ge \frac{c_2 L R}{\sqrt{n}}$,
-and $\lambda_n^{(3)} \ge \frac{c_3 L R}{\sqrt{n}}$,
+Suppose the number of samples is
+$n = \Omega(d^p R^p \sigmamin(\Sigma_p)^{-1} \sqrt{p \log(d) \log(1/\delta)})$ and
+$n = \Omega\left( \sigma^6 L^6 R^12 k \sigmamin(\Sigma_p)^{-2} \frac{\log^3 (1/\delta)}{\epsilon^2} \right)$,
+and that the regularization strengths $\lambda_n^{(2)}$ and $\lambda_n^{(3)}$ are
+$\Omega(\sigma^3 L^3 R^6 \sqrt{\frac{\log(1/\delta)}{n}})$,
 then the parameter estimates $\hat\theta$ returned by
@@
 satisfies 
-$\|\hat\pi - \pi^*\|_{\infty} \le \frac{\epsilon}{???}$
+$\|\hat\pi - \pi^*\|_{\infty} \le \epsilon$
 and for all $h \in [k]$,
-$\|\hat\beta_h - \beta^*_h\|_2 \le \epsilon ???$,
+$\|\hat\beta_h - \beta^*_h\|_2 \le \frac{\epsilon}{\pi_h^*}$.
 \end{theorem}
+
 \begin{proof}
-Let $T^* = M_3^*(W^*, W^*, W^*)$.
-By Theorem 5.1 of \cite{AnandkumarGeHsu2012},
-Need to convert error from $\epsilon$
-hit by $(W^\top)^\dagger$
+By \reflem{lowRank} along with \reflem{lowRankLower} and \reflem{lowRankUpper} (discussed below),
+we can control the Frobenius norm of the parameter estimates:
+If $n = \Omega(d^p R^p \sigmamin(\Sigma_p)^{-1} \sqrt{p \log(d) \log(1/\delta)})$, then
+\begin{align}
+  \|\hat M_p - M_p\|_F = O\left( \sigma^3 L^3 R^6 k^{\frac12} \sigmamin(\Sigma_p)^{-1} \sqrt{\frac{\log^3 (1/\delta)}{n}} \right).
+  %\|\hat M_p - M_p\|_F = O\left( \frac{\sigma^3 L^3 R^6 \sqrt{k} \log^3 (1/\delta)}{\sqrt{n} (\sigmamin(\Sigma_p) - d^p R^p \sqrt{\frac{p \log(d) \log(1/\delta)}{n}}} \right).
+\end{align}
+The Frobenius norm upper bounds the operator norm,
+so we can directly apply Theorem 5.1 of \cite{AnandkumarGeHsu2012}.
+%Let $T^* = M_3^*(W^*, W^*, W^*)$.
+%Need to convert error from $\epsilon$
+%hit by $(W^\top)^\dagger$
 \end{proof}

### 2013-02-15 doc/icml2013/regression.tex
theory done

@@
-\subsection{Low-rank Regression}
-\label{sec:regression}
+\paragraph{Low-rank Regression}
+%\label{sec:regression}
 
@@
 We first set up some notation to all three regressions (\refeqn{estimateM1}, \refeqn{estimateM2}, and \refeqn{estimateM3}).
-Define the observation operator $\opX(M_p) : \Re^{d\tp{p}} \to \Re^{n}$
+Define the observation operator $\opX_p(M_p) : \Re^{d\tp{p}} \to \Re^{n}$
 mapping compound parameters $M_p$:
 \begin{align}
-  \opX(M_p; \sD)_i &\eqdef \frac12 \sum_{(x,y) \in \sD} \innerp{M_p}{x\tp{p}}_i, \quad i  \in [n].
+\opX_p(M_p; \sD)_i &\eqdef \frac12 \sum_{(x,y) \in \sD} \innerp{M_p}{x\tp{p}}_i, \quad i  \in [n].
 \end{align}
 
-Let $\kappa(\opX)$ be the restricted strong constant,
-and let $\opX^*(\eta_p; \sD) = \sum_{(x,y) \in \sD} \eta_p(x) x\tp{p}$
+Let $\kappa(\opX_p)$ be the restricted strong constant,
+and let $\opX^*_p(\eta_p; \sD) = \sum_{(x,y) \in \sD} \eta_p(x) x\tp{p}$
 be the adjoint.
@@
 In this section, we lower bound the restricted strong convexity constant
-$\kappa(\opX)$.
-
-\paragraph{Adjoint operator}
-
-In this section, we upper bound the operator norm of the adjoint
-$\|\opX(\eta_p)\|_\text{op}$.
+$\kappa(\opX_p)$.
 
@@
 %  \hat M_p &= 
-%  \arg\min_{M_p}& \frac{1}{2N} \| \vec y - \opX(M_p) \|_2^2 + \frac{\lambda_n}{K} \sum_{h=1}^K \| (M_p)_{(h)} \|_* \label{eq:regression}.
+%  \arg\min_{M_p}& \frac{1}{2N} \| \vec y - \opX_p(M_p) \|_2^2 + \frac{\lambda_n}{K} \sum_{h=1}^K \| (M_p)_{(h)} \|_* \label{eq:regression}.
 %\end{align}
@@
 \begin{lemma}[\citet{Tomioka2011}, Theorem 1]
+\label{lem:lowRank}
 Suppose there exists a restricted strong convexity constant $\kappa(\opX)$ such that
-$$\frac{1}{2n} \| \opX( \Delta )\|_2^2 \ge \kappa(\opX) \|\Delta\|^2_F \quad \text{and} \quad
+$$\frac{1}{2n} \| \opX_p( \Delta )\|_2^2 \ge \kappa(\opX) \|\Delta\|^2_F \quad \text{and} \quad
 \lambda_n \ge \frac{\|\opX^*(\eta_p)\|_2}{n}.$$
@@
 parameter $\kappa(\opX)$ and upper bound the adjoint operator
-$\|\opX^*(\epsilon)\|_{2}^2$.
+$\|\opX^*(\eta_p)\|_{2}^2$.
 %We
@@
 
+%\paragraph{Adjoint operator}
+
+%In this section, we upper bound the operator norm of the adjoint
+%$\|\opX_p(\eta_p)\|_\text{op}$.
+
+First, let us lower bound the restricted strong convexity parameter $\kappa(\opX_p)$:
+
 \begin{lemma}[lower bound on restricted strong convexity]
+\label{lem:lowRankLower}
 Let $\Sigma_p \eqdef \E[\cvec(x\tp{p})\tp{2}]$.
-With high probability at least $1-\delta$,
-for some constant $c$,
-$$\kappa(\opX) \ge c \sigma_\text{min}^p(\Sigma_p)$$,
-where
-%$\sigma_\text{min}^p(\Sigma_p) = \min_{M_p \in \sM_p : \|M_p\|_F \le 1} \cvec(M_p)^\top \Sigma_p \cvec(M_p)$.
-$\sigma_\text{min}^p(\Sigma_p) = \min_{M_p \in \sM_p : \|M_p\|_F \le 1} \trace(\cvec(M_p)^{\otimes 2} \Sigma_p)$.
+With probability at least $1-\delta$,
+$$\kappa(\opX) = \Omega\left(\sigmamin(\Sigma_p) - d^p R^{p} \sqrt{\frac{p \log(d) \log (1/\delta)}{n}}\right).$$
 \end{lemma}
-Note that $\sigma_\text{min}$ 
 \begin{proof}
 Expanding the definition of the observation operator:
-$\frac{1}{2n} \|\opX(\Delta)\|_2^2
+$\frac{1}{2n} \|\opX_p(\Delta)\|_2^2
 = \frac{1}{2n} \sum_{(x,y) \in \sD} \innerp{\Delta}{x\tp{p}}^2$.
@@
 we have 
-$\frac{1}{2n} \|\opX(\Delta)\|_2^2
+$\frac{1}{2n} \|\opX_p(\Delta)\|_2^2
 \ge \frac{1}{2} \trace(\cvec(\Delta)\tp{2} \hat\Sigma_p)
-\ge \frac{1}{2} \|\Delta\|_F^2 \sigma_\text{min}^p(\hat\Sigma_p)$.
-
-%  \begin{align*}
-%    \frac{\|\opX(M_p)\|_1}{N} &= \frac{1}{N}\sum_{i=1}^{N} |\innerp{x_i\tp{p}}{M_p}| \\
-%    &\ge \frac{1}{N}\sum_{i=1}^{N} \sigma_{min}( M_p ) \|x_i\|^p \\
-%    &\ge \frac{\kappa(M_p)}{\sqrt{D}} \frac{\sum_{i=1}^{N} \|x_i\|^p}{N} \| M_p \|_F.
-%  \end{align*}
-%
-%  While $\kappa(M_p)$ depends on $M_p$ we can lower bound it by
-%  restricting the convex space to $\{ M_p | \kappa(M_p) \ge \kappa \}$.
-%  Further, using concentration bounds, we can lower bound $\|x_i\|^p$ as
-%  well. \todo{What is $\E[\|x_i\|^3]$? It is some quantity greater than
-%  zero}.
-%  
-%  Thus, 
-%  \begin{align*}
-%  \frac{ \|\opX(M_p)\|_2 }{2 N} 
-%  &\ge
-%  \frac{ \|\opX(M_p)\|_1 }{2 \sqrt{D}N} 
-%  &\ge \frac{\kappa}{2 D} (\E[\|x_i\|^p] - \frac{t}{N}) \| M_p \|_F.
-%  \end{align*}
-%  with high probability $1 - \delta$. \todo{The bound needs to be shown
-%  for $\| \opX(M_p) \|_2^2$.}
-%  
-%  In other words, $\kappa(\opX)
-%  = \frac{\kappa}{2 D} (\E[\|x_i\|^p] - \frac{t}{N})$ is a valid
-%  lower bound.
+\ge \frac{1}{2} \|\Delta\|_F^2 \sigmamin(\hat\Sigma_p)$.
+It remains to show that $\sigmamin(\hat\Sigma_p)$ and $\sigmamin(\Sigma_p)$ are close.
+To do this, we first apply Hoeffding's inequality elementwise.
+Since $\|x\| \le R$, we have that for each element $(i,j)$,
+$|(\hat\Sigma_p)_{ij} - (\Sigma_p)_{ij}| = O(R^{p}\sqrt{\frac{\log (1/\delta)}{n}})$.
+Applying the union bound over the $d^{2p}$ elements of $\hat\Sigma_p - \Sigma_p$,
+we have that the max norm is bounded:
+$\|\hat\Sigma_p - \Sigma_p\|_\text{\rm max} = O(R^{p} \sqrt{\frac{p \log(d) \log (1/\delta)}{n}})$.
+The max norm times $d^p$ upper bounds the Frobenius norm, which upper bounds the operator norm, so we have that
+$\|\hat\Sigma_p - \Sigma_p\|_\text{\rm op} = O(d^p R^{p} \sqrt{\frac{p \log(d) \log (1/\delta)}{n}})$.
+Using the fact that $\sigmamin(\hat\Sigma_p) = \sigmamin(\Sigma_p) + \|\hat\Sigma_p - \Sigma_p\|_\text{\rm op}$
+yields the result.
 \end{proof}
 
-\begin{lemma}[bounded error]
-  Let $\opX$ be the linear operator previously defined. Then,
-  $$\|\opX^*(\epsilon)\|_\text{op} \le Y$$.
+Next, let us upper bound the operator norm of the observation adjoint:
+
+\begin{lemma}[upper bound on adjoint noise]
+\label{lem:lowRankUpper}
+Let $\opX_p$ be the linear operator previously defined. Then with probability at least $1-\delta$,
+$$\frac1{n} \|\opX_p^*(\eta_p)\|_\op \le O\left(\sigma^3 L^3 R^6 \sqrt{\frac{\log^3(1/\delta)}{n}}\right).$$
+for each $p \in \{1,2,3\}$.
 \end{lemma}
 \begin{proof}
-%  \begin{align*}
-%    \frac{\opX^*(\epsilon)}{N} 
-%    &= \frac{1}{N}\sum_{i=1}^{N} \epsilon_i \innerp{x_i\tp{p}}{M_p} \\
-%    &= \frac{1}{N}\sum_{i=1}^{N} (\innerp{x_i\tp{p}}{M_p - \beta_i\tp{p}} + \varepsilon)  \innerp{x_i\tp{p}}{M_p}.
-%  \end{align*}
-%  Let $M_p - \beta_i\tp{p} = \tilde M_i$ and consider it to be a random
-%  variable.
-%  \begin{align*}
-%    \frac{\opX^*(\epsilon)}{N} 
-%    &= \frac{1}{N}\sum_{i=1}^{N} \innerp{\tilde M_i}{x_i\tp{2p}} + \varepsilon \innerp{M_p}{x_i\tp{p}} \\
-%    &\le \frac{1}{N}\sum_{i=1}^{N} \|\tilde M_i\|_{op} \|x_i\|^{2p} + \varepsilon \|M_p\|_{op}\|x_i\|^{p} \\
-%    &\le \|M_p\|_{op} ( \frac{1}{N}\sum_{i=1}^{N} \|x_i\|^{2p} + \varepsilon \|x_i\|^{p} ).
-%  \end{align*}
-%
-%  We can probabilistically bound the above by considering the union
-%  bound of $\varepsilon$ and $\|x_i\|^p$.
-%
-%
-%  Finally,
-%  \begin{align*}
-%    \frac{\opX^*(\epsilon)}{N} 
-%    &\le \|M_p\|_{op} (\E[\|x_i\|^{2p}] + t/N + (t/N) (\E[ \|x_i\|^{p} ] + t/N) ),
-%  \end{align*}
-%  with high probability. Note that this does not converge to $0$ as $N
-%  \to \infty$ because of the variance of the selection probability.
+Let $\hat\E_p[f(x,\epsilon,h)]$ denote the empirical expectation over the examples in dataset $\sD_p$
+(recall the $\sD_p$'s are independent to simplify the analysis).
+By definition,
+$\frac1n \|\opX_p^*(\eta_p)\|_\op = \hat\E_p[\eta_p(x) x\tp{p}]$
+for $p \in \{1,2,3\}$ defined in \refeqn{y1}, \refeqn{y2}, and \refeqn{y3}.
+Each $\eta_p(x)$ is composed of several zero-mean random variables,
+and since $\|A + B\|_\op \le \|A\|_\op + \|B\|_\op$,
+it suffices to consider each term in turn.
+For each term, we will condition on $x$, use a standard concentration inequality to
+bound the noise.
 
-By Lemma 19 of \cite{hsu13spherical},
-we have that $\frac{1}{n} \sum_{i=1}^n \epsilon_i^3 = O(\sqrt{\frac{\log^3(1/\delta)}{n}})$.
-with probability at least $1-\delta$.
+First, consider $\eta_1(x)$, which contains two terms.
+The first term is $\innerp{M_1 - \beta_h}{x}$.
+We have that $\|M_1 - \beta_h\|_2 \le L$,
+so by Hoeffding's inequality, and using the fact that $\|x\|_2 \le R$,
+with probability $1-\delta_1$,
+$\|\hat\E_1[\innerp{M_1 - \beta_h}{x} x]\|_\op = O(R^2 L \sqrt{\frac{\log(1/\delta_1)}{n}})$.
+The second term is $\epsilon \sim \normal{0}{\sigma^2}$,
+with probability $1-\delta_1$,
+$\|\hat\E_1[\epsilon x]\|_\op = O(R \sigma \sqrt{\frac{\log(1/\delta_1)}{n}})$.
 
+Now, consider $\eta_2(x)$, which contains three terms.
+Using similar techniques, we have that
+$\|\hat\E_2[\eta_2(x) x\tp{2}]\|_\op = O(R^2 (L^2 R^2 + \sigma L R + \sigma^2) \sqrt{\frac{\log(1/\delta_1)}{n}})$.
+
+Finally, $\eta_3(x)$ contains five terms plus an additional bias since the estimator uses $\hat M_1$ rather than $M_1$.
+There are two additional considerations.
+First, to bound $\hat\E_3[\epsilon^3]$, we use Lemma 19 of \cite{hsu13spherical}.
+Second, the additional bias is small by the previous bound (and it is constructed on $\sD_1$, which is independent of $\sD_3$
+used to contruct $\hat M_3$.
+Therefore, we incur another $O(\sigma^2 \cdot R \sigma \sqrt{\frac{\log(1/\delta_1)}{n}})$.
+Handling the other terms using conventional means
+$\|\hat\E_3[\eta_3(x) x\tp{3}]\|_\op = O(R^3 (L^3 R^3 + \sigma L^2 R^2 + \sigma^2 L R + \sigma^3 + \sigma^3 R) \sqrt{\frac{\log^3(1/\delta_1)}{n}})$.
+Finally, taking $\delta_1 = \delta/3$, and taking the union bound,
+we get our result, where we are bounding quantities fairly crudely for the sake of simplicity.
 \end{proof}

### 2013-02-15 doc/icml2013/evaluation.tex

@@
 \label{sec:evaluation}
-
-\todo{Fill it up with juicy bits once we actually have them graphs.}
 
@@
 We implemented the low rank recovery optimization routines for $M_2$ and
-$M_3$ using a proximal subgradient descent algorithm\citationneeded. The
-algorithm was initialized with the regularized least squares solution
-and was run to convergence in each instance. Both regularization
-parameters, $\lambda_n^{(2)}$ and, $\lambda_n^{(3)}$ were annealed as
-$\frac{1}{n}$.
+$M_3$ using a proximal subgradient descent
+algorithm.%\cite{candes11phaselift,tomioka2010estimation}. (Note sure about this citation and not required?)
+The algorithm
+was initialized with the regularized least squares solution and was run
+for 500 steps or till convergence. Both regularization parameters,
+$\lambda_n^{(2)}$ and, $\lambda_n^{(3)}$ were annealed as $\frac{1}{n}$.
 
@@
 
-We generated synthetic data using covariates from the unit hyper cube
-and non-linear responses.  To conform to the identifiability criteria
-discussed in \sectionref{sec:algo}, we considered only features that
-would guarantee this condition.
+We generated synthetic data using covariates from the $b$-dimensional
+unit hyper cube $\{-1,1\}^b$ and considered non-linear featurizations of
+this data that conformed to the identifiability criteria discussed in
+\sectionref{sec:algo}. The $\beta$ were drawn from a standard Gaussian.
 
@@
 the parameters recovered using spectral experts are shown in
-\figureref{fig:curves}.
+\figureref{fig:curves}. We note that even on well separated data such as
+this, EM converged to the global optima only 13\% of the time.
 
@@
 \begin{table*}[t]
-\caption{Parameter Recovery ($N = 500,000$).}
+\caption{Parameter Recovery Error $\|\theta^* - \hat \theta\|_F$ ($N = 500,000$).}
 \label{tbl:parameter-recovery}
@@
 
-  \begin{tabular}{ >{\centering}p{5cm}<{\centering} r c c c }
+  \begin{tabular}{ r r r c c c }
 \hline
 \abovespace\belowspace
-Features  & $K$ & Spectral & EM & Spectral + EM \\
+Variables ($b$) & Features ($D$) & Components ($K$) & Spectral & EM & Spectral + EM \\
 \hline
 \abovespace
-$\{1$, $ x_1$, $ x_1^4\}$ 
-  & 2 & 1.52 $\pm$ 0.71 & 0.28 $\pm$ 0.82 & {\bf 0.13 $\pm$ 0.55} \\
-$\{1$, $ x_1$, $ x_2$, $ x_1^2 x_2^2\}$ 
-  & 3 & 1.87 $\pm$ 1.20 & {\bf 0.33 $\pm$ 0.96} & 0.35 $\pm$ 1.23 \\
-$\{1$, $ x_1$, $ x_2$, $ x_1 x_2^3$, $ x_1^2 x_2^2$, $ x_1^3 x_2 \}$ 
-& 5 & 5.27 $\pm$ 2.32 & 1.80 $\pm$ 1.80 & {\bf 1.51 $\pm$ 1.77} \\
-$\{1$, $ x_1$, $ x_2$, $ x_1 x_2^3$, $ x_1^2 x_2^2$, $ x_1^3 x_2$, $ x_1^3 x_2^4$, $ x_1^4 x_2^3 \}$ 
-& 7 & 8.42 $\pm$ 1.66 & 7.41 $\pm$ 2.99 & {\bf 7.31 $\pm$ 2.47} \\
-$\{1$, $ x_1$, $ x_2$, $ x_3$, $ x_1 x_2 x_3^2$, $ x_1 x_2^2 x_3$, $ x_1^2 x_2 x_3$, $ x_1^2 x_2^2$, $ x_1^2 x_3^2$, $ x_2^2 x_3^2$, $ x_3^2\}$
-& 3 & 3.78 $\pm$ 0.90 & 0.67 $\pm$ 1.49 & {\bf 0.51 $\pm$ 1.12} \\
-$\{1$, $ x_1$, $ x_2$, $ x_3$, $ x_1 x_2 x_3^2$, $ x_1 x_2^2 x_3$, $ x_1^2 x_2 x_3$, $ x_1^2 x_2^2$, $ x_1^2 x_3^2$, $ x_2^2 x_3^2$, $ x_3^2\}$
-  & 7 & 9.97 $\pm$ 3.22 & 1.43 $\pm$ 1.97 & {\bf 1.42 $\pm$ 2.17} \\
+%$\{1$, $ x_1$, $ x_1^4\}$ 
+  1 & 4 & 2 & 1.52 $\pm$ 0.71 & 0.28 $\pm$ 0.82 & {\bf 0.13 $\pm$ 0.55} \\
+%$\{1$, $ x_1$, $ x_2$, $ x_1^2 x_2^2\}$ 
+  2 & 5 & 2 & 2.92 $\pm$ 1.10 & 0.00 $\pm$ 0.00 & {\bf 0.00 $\pm$ 0.00} \\
+  2 & 5 & 3 & 3.76 $\pm$ 1.93 & 0.43 $\pm$ 1.07 & {\bf 0.23 $\pm$ 0.76} \\
+%$\{1$, $ x_1$, $ x_2$, $ x_1 x_2^3$, $ x_1^2 x_2^2$, $ x_1^3 x_2 \}$ 
+  2 & 6 & 2 & 4.52 $\pm$ 1.67 & 0.63 $\pm$ 1.29 & {\bf 0.18 $\pm$ 0.66} \\
+\belowspace
+  2 & 6 & 5 & 6.78 $\pm$ 2.18 & 2.29 $\pm$ 1.79 & {\bf 1.77 $\pm$ 1.89} \\
+
+%  2 & 5 & 3 & 1.87 $\pm$ 1.20 & {\bf 0.33 $\pm$ 0.96} & 0.35 $\pm$ 1.23 \\
+% 2 & 6 & 5 & 5.27 $\pm$ 2.32 & 1.80 $\pm$ 1.80 & {\bf 1.51 $\pm$ 1.77} \\
+%$\{1$, $ x_1$, $ x_2$, $ x_1 x_2^3$, $ x_1^2 x_2^2$, $ x_1^3 x_2$, $ x_1^3 x_2^4$, $ x_1^4 x_2^3 \}$ 
+% 2 & 8 & 7 & 8.42 $\pm$ 1.66 & 7.41 $\pm$ 2.99 & {\bf 7.31 $\pm$ 2.47} \\
+%$\{1$, $ x_1$, $ x_2$, $ x_3$, $ x_1 x_2 x_3^2$, $ x_1 x_2^2 x_3$, $ x_1^2 x_2 x_3$, $ x_1^2 x_2^2$, $ x_1^2 x_3^2$, $ x_2^2 x_3^2$, $ x_3^2\}$
+% 3 & 10 & 3 & 3.78 $\pm$ 0.90 & 0.67 $\pm$ 1.49 & {\bf 0.51 $\pm$ 1.12} \\
+%$\{1$, $ x_1$, $ x_2$, $ x_3$, $ x_1 x_2 x_3^2$, $ x_1 x_2^2 x_3$, $ x_1^2 x_2 x_3$, $ x_1^2 x_2^2$, $ x_1^2 x_3^2$, $ x_2^2 x_3^2$, $ x_3^2\}$
+% 3 & 10  & 7 & 9.97 $\pm$ 3.22 & 1.43 $\pm$ 1.97 & {\bf 1.42 $\pm$ 2.17} \\
+
 \hline
+
 \end{tabular}
@@
 
-There is a lot of variation across random instances; some are easy for
-EM to find the global minima and others more difficult. Moreover, simply
-increasing the number of samples did not necessarily improve EM
-performance. In general, we found that while our spectral algorithm did
-not recover parameters extremely well, it provided a good initialization
-for EM.
+One of the main reasons for the high variance is the variation across
+random instances; some are easy for EM to find the global minima and
+others more difficult. In general, while the spectral algorithm did not
+recover parameters extremely well, it provided a good initialization for
+EM.
 
-\todo{Plot: the histogram over errors (for EM?) Some plots I have
-doesn't really show much}.
+\begin{figure}[tb]
+  \centering
+  \includegraphics[width=0.50\textwidth]{figures/hist.pdf}
+  \caption{Histogram over recovery errors between the three algorithms. $b = 1, d = 4, k = 3, N = 500,000$}
+  \label{fig:hist}
+\end{figure}
+
+To study the stability of the solutions returned by the spectral method,
+consider the histogram in \figureref{fig:hist}, which shows the recovery
+errors of the algorithms over 170 attempts a dataset with $b = 1, d = 4,
+k = 3$. Typically, the spectral algorithm returned a stable solution.
+When these parameters were close enough to the true parameters, we found
+that EM almost always converged to the global optima. Randomly
+initialized EM only finds the true parameters a little over 10\% of the
+time and shows considerably higher variance.
+
+\paragraph{Effect of number of data points}
+
+\begin{figure*}[tb]
+  \centering
+  \subfigure[Well Data]{
+    \includegraphics[width=0.50\textwidth]{figures/vs-n/1833-decay.pdf}
+  }
+    \hspace{-2em}
+  \subfigure[Misspecified Data]{
+    \includegraphics[width=0.50\textwidth]{figures/vs-n/1-8-3-3-rm.pdf}
+  }
+  \caption{Recovey error vs $n$  ($b = 1, d = 5, K = 3, N = 500,000$).}
+  \label{fig:vs-n}
+\end{figure*}
+
+In \figureref{fig:vs-n}, we show how the recovery error varies as we get
+more data. Each data point shows the mean error over 10 attempts, with
+error bars. We note that the recovery performance of EM does not
+particularly improve; this is expected because of the problem of local
+optima. The spectral algorithm's error decays slowly, and as it gets
+closer to zero, EM initialized at the spectral parameters finds the true
+parameters more often as well. This behaviour highlights the trade-off
+between statistical and computational error.
 
@@
 
+\begin{table*}[tb]
+\caption{Parameter Recovery Error $\|\theta^* - \hat \theta\|_F$ with misspecified data ($N = 500,000$).}
+\label{tbl:parameter-recovery-mis}
+\vskip 0.15in
+\begin{center}
+\begin{small}
+\begin{sc}
+
+  \begin{tabular}{ r r r c c c }
+\hline
+\abovespace\belowspace
+Variables ($b$) & Features ($d$) & Components ($K$) & Spectral & EM & Spectral + EM \\
+\hline
+\abovespace
+ 1 & 4 & 2 &  2.94 $\pm$ 1.52 & 0.29 $\pm$ 0.85 &  {\bf 0.22 $\pm$ 0.72} \\
+\belowspace
+ 2 & 5 & 3 &  3.82 $\pm$ 2.00 & 0.44 $\pm$ 1.12 &  {\bf 0.35 $\pm$ 1.00} \\
+ 2 & 6 & 5 &  9.89 $\pm$ 4.46 & {\bf 2.53 $\pm$ 1.77} &  2.69 $\pm$ 1.83 \\
+ 2 & 8 & 7 & 23.07 $\pm$ 7.10 & 9.62 $\pm$ 1.03 &  {\bf 8.16 $\pm$ 2.31}  \\
+\hline
+
+\end{tabular}
+\end{sc}
+\end{small}
+\end{center}
+\vskip -0.1in
+\end{table*}
+
 To evaluate how robust the algorithm was to model mis-specification, we
 removed large contiguous sections from $x \in [-0.5,-0.25] \cup
-[0.25,0.5]$. The lower half of \tableref{tbl:parameter-recovery} reports
-recovery errors in this scenario.
-
-\paragraph{Effect of number of data points}
-
-\begin{figure*}[t]
-  \centering
-  \subfigure[Complete Data]{
-    \includegraphics[width=0.50\textwidth]{figures/vs-n/1-8-3-3.pdf}
-  }
-    \hspace{-2em}
-  \subfigure[Incomplete Data]{
-    \includegraphics[width=0.50\textwidth]{figures/vs-n/1-8-3-3-rm.pdf}
-  }
-  \caption{}
-  \label{fig:vs-n}
-\end{figure*}
-
-Highlight
-
-\todo{We see that the error of spectral is actually less than EM after a point.}
-
-Point: tradeoff between statistical error and computational error;
-not enough data, spectral is actually a lot worse
-given enough data, spectral+EM is much better
-
-\paragraph{Effect of noise}
-
-\paragraph{Effect of separation between components}
+[0.25,0.5]$ and ran the algorithms again.
+\tableref{tbl:parameter-recovery-mis} reports recovery errors in this
+scenario. The error in the estimates grows larger for higher $d$.

### 2013-02-15 doc/icml2013/algo.tex
Minor fixes

--- 2013-02-15 doc/icml2013/algo.tex
@@
 Specifically, for an integer $p \ge 1$, let us define the average
-$m$-th order tensor power of the parameters as follows:
+$p$-th order tensor power of the parameters as follows:
 \begin{align}
@@
 Let us look to the third moment for additional information.
-We can write $y^3$ as a linear function of $x\tp{3}$ with coefficients $M_3$
-plus a bias $3 \sigma^2 \innerp{M_1}{x}$ that can be estimated from \refeqn{y1} and zero mean noise $\eta_3(x)$:
+We can write $y^3$ as a linear function of $x\tp{3}$ with coefficients $M_3$,
+a bias $3 \sigma^2 \innerp{M_1}{x}$ that can be estimated from \refeqn{y1}, and zero mean noise $\eta_3(x)$:
 \begin{align}
@@
 to exploit this low-rank structure and improve our compound parameter estimates.
-In the algorithm, the regularization strengths $\lambda_n^{(2)}$ and $\lambda_n^{(2)}$
+In the algorithm, the regularization strengths $\lambda_n^{(2)}$ and $\lambda_n^{(3)}$
 are set to $\frac{1}{\sqrt{n}}$.
@@
 the forms in \refeqn{Mp}, it is possible to efficiently accomplish this.
-Specifically, we first compute a whitening matrix based on the SVD of $M_2$
+Specifically, we first compute a whitening matrix $W$ based on the SVD of $M_2$
 and use that to construct a tensor $T = M_3(W, W, W)$ whose factors are orthogonal.
@@
   \label{thm:convergence}
-Each $\sD_p$ (for $p = 1, 2, 3$) consists of $n$ i.i.d. points drawn from the mixture
+Each dataset $\sD_p$ (for $p = 1, 2, 3$) consists of $n$ i.i.d. points drawn from the mixture
 of linear regressions model with parameter $\theta^*$

### 2013-02-15 doc/icml2013/intro.tex
theory done
Minor fixes

@@
 % Twist
-In mixture of linear regressions, using third-order moments does not directly
-reveal the tensor structure of the problem, so we cannot directly
-apply the above tensor factorization techniques.
-Our solution is to employ low-rank linear regression
-\cite{NegahbanWainwright2009,Tomioka2011} on second
-and third powers of the response to first recover the appropriate symmetric
-tensors, from which we can then apply the tensor power method to retrieve the
-final parameters.
+In mixture of linear regressions, using third-order moments does not
+directly reveal the tensor structure of the problem, so we cannot
+directly apply the above tensor factorization techniques.  Our solution
+is to employ low-rank linear regression
+\cite{NegahbanWainwright2009,Tomioka2011} on second and third powers of
+the response to recover the appropriate symmetric tensors, from which we
+can then apply the tensor power method to retrieve the final parameters.
 
@@
 \ldots i_p} Y_{i_1, \ldots i_p}.$
-A tensor $X$ is symmetric if for all $i,j \in [d]^p$ which are permutations,
+A tensor $X$ is symmetric if for all $i,j \in [d]^p$ which are permutations of each other,
 $X_{i_1 \cdots i_p}$ = $X_{j_1 \cdots j_p}$ (all tensors in this paper will be symmetric).
@@
 $\cvec(X) = (X_{ii} : i \in [d]; X_{ij} + X_{ji} : i,j \in [d], i<j)$.
-For symmetric tensors, $\innerp{X}{Y} = \innerp{\cvec(X)}{\cvec(Y)}$.
+%For symmetric tensors, $\innerp{X}{Y} = \innerp{\cvec(X)}{\cvec(Y)}$.

### 2013-02-15 doc/icml2013/regression.tex
More fixes

@@
 $x$, involves the mixing noise and various cross terms.
-We first set up some notation to all three regressions (\refeqn{estimateM1}, \refeqn{estimateM2}, and \refeqn{estimateM3}).
+We first set up some notation that unifies all three regressions (\refeqn{estimateM1}, \refeqn{estimateM2}, and \refeqn{estimateM3}).
 Define the observation operator $\opX_p(M_p) : \Re^{d\tp{p}} \to \Re^{n}$
@@
 
-Let $\kappa(\opX_p)$ be the restricted strong constant,
+Let $\kappa(\opX_p)$ be the restricted strong convexity constant,
 and let $\opX^*_p(\eta_p; \sD) = \sum_{(x,y) \in \sD} \eta_p(x) x\tp{p}$

### 2013-02-15 doc/icml2013/discussion.tex
conclusion

@@
-\section{Discussion}
-\label{sec:discussion}
+\section{Conclusion}
+\label{sec:conclusion}
 
-Role of spectral algorithms in discriminative tasks.
-
-Challenges and ways forward?
+In this paper, we take a first step at developing computationally efficient and
+statistically consistent estimators for discriminative latent-variable models,
+by focusing on mixture of linear regressions.  We showed that by regressing on
+higher order powers of the data with a regularizer that encourages low rank
+structure, coupled with an efficient tensor power method, we obtain a very natural algorithm.
+Empirically, our algorithm can outperform EM, and when used to initialize EM,
+yields significantly more accurate and stable results.

### 2013-02-15 doc/icml2013/algo.tex
edit

@@
 (i) low-rank regression to estimate certain symmetric tensors;
-and (ii) tensor factorzation to recover the parameters.
+and (ii) tensor factorization to recover the parameters.
 The two steps can be performed efficiently using
@@
 for learning latent-variable models.  One line of work has
-focused on observerable operator models \cite{hsu09spectral}
+focused on observable operator models \cite{hsu09spectral}
 in which the true parameters are not recovered, but a reparametrization is used
@@
 % Unmixing
-In spirit, Spectral Experts also bears some resemblence to the unmixing
+In spirit, Spectral Experts also bears some resemblance to the unmixing
 algorithm for estimation of restricted PCFGs
@@
 There, the actual observed response was $y^2$,
-whereas in our case, we delibrately construct powers $y,y^2,y^3$
+whereas in our case, we deliberately construct powers $y,y^2,y^3$
 to identify the parameters.
@@
     \end{align}
-    \STATE Estimate the parameters $\theta = (\pi, B)$ using \textbf{tensor factorzation}:
+    \STATE Estimate the parameters $\theta = (\pi, B)$ using \textbf{tensor factorization}:
     \begin{enumerate}

### 2013-02-15 doc/icml2013/intro.tex
edit

@@
 past approaches have relied exclusively on local optimization (EM, 
-gradient descent) and are vulerable to local optima.
+gradient descent) and are vulnerable to local optima.
 
@@
 Let $[n] = \{ 1, \dots, n \}$ denote the first $n$ positive integers.
+We use $O(f(n))$ to denote a function $g(n)$ such that $\lim_{n \to\infty} g(n)/f(n) < \infty$.
+We will only hide universal constants, exposing all dependence on the number of samples,
+dimension, norms, noise variance, etc.
 
@@
 denote the average operator norm over all $p$ unfoldings.
+Let $\sigma_\text{min}(X)$ be the smallest singular value of $X$.
 

### 2013-02-15 doc/icml2013/model.tex
Minor fixes
edit

@@
 are the regression coefficients.
-We assume the variance of the observation noise $\sigma^2$ is known.
+We assume the variance of the observation noise $\sigma^2$ is known and that the mixture components are independent.
 The learning problem is stated as follows:
@@
 This dependence allow the experts to be localized in input space,
-providing more flexbility, but we do not consider this dependence in our model.
+providing more flexibility, but we do not consider this dependence in our model.
 

### 2013-02-15 doc/icml2013/regression.tex
edit

@@
 Second, the additional bias is small by the previous bound (and it is constructed on $\sD_1$, which is independent of $\sD_3$
-used to contruct $\hat M_3$.
+used to construct $\hat M_3$.
 Therefore, we incur another $O(\sigma^2 \cdot R \sigma \sqrt{\frac{\log(1/\delta_1)}{n}})$.

### 2013-02-16 doc/icml2013/algo.tex
edit

@@
 
-%Define $\sM_p = \{ \sum_{h=1}^k \pi_h \beta_h\tp{p} : \pi_h \in \R^k, \beta_h
-%\in \R^d \}$ be the set of all $p$-th order compound parameters.
-%Let $\bM_p$ be the set of
+In this section, we provide theoretical guarantees for our Spectral Experts algorithm.
+In particular, the following theorem provides a rate of convergence:
 
-%Define the symmetric rank
-%\begin{align}
-%\E[\cvec(x\tp{p})\tp{2}] \succ 0.
-%\end{align}
-
-In this section, we conduct a theoretical analysis of Spectral Experts.
-In particular, we prove that it enjoys consistent parameter estimates.
 \begin{theorem}[Convergence of Spectral Experts]
@@
 (having three independent copies simplifies the analysis).
-Assume that $\E[\cvec(x\tp{p})\tp{2}] \succ 0$ for each $p \in \{1,2,3\}$
-and that $d \ge k$.
 Let $\|x\|_2 \le R$ with probability $1$
 and let $\|\beta_h\|_2 \le L$ for all $h \in [k]$.
+Let $\Sigma_p \eqdef \E[\cvec(x\tp{p})\tp{2}]$, 
+and assume $\Sigma_p \succ 0$ for each $p \in \{1,2,3\}$.
+Assume $d \ge k$.
 Suppose the number of samples is

### 2013-02-16 doc/icml2013/algo.tex
edit

@@
 Let $\|x\|_2 \le R$ with probability $1$
-and let $\|\beta_h\|_2 \le L$ for all $h \in [k]$.
+and let $\|\beta_h^*\|_2 \le L$ for all $h \in [k]$.
 Let $\Sigma_p \eqdef \E[\cvec(x\tp{p})\tp{2}]$, 
@@
 Suppose the number of samples is
-$n = \Omega(d^p R^p \sigmamin(\Sigma_p)^{-1} \sqrt{p \log(d) \log(1/\delta)})$ and
-$n = \Omega\left( \sigma^6 L^6 R^12 k \sigmamin(\Sigma_p)^{-2} \frac{\log^3 (1/\delta)}{\epsilon^2} \right)$,
-and that the regularization strengths $\lambda_n^{(2)}$ and $\lambda_n^{(3)}$ are
-$\Omega(\sigma^3 L^3 R^6 \sqrt{\frac{\log(1/\delta)}{n}})$,
-then the parameter estimates $\hat\theta$ returned by
+$n = \max(n_1,n_2)$
+where $n_1 = \Omega(d^p R^p \sigmamin(\Sigma_p)^{-1} \sqrt{p \log(d) \log(1/\delta)})$ and
+$n_2 = \Omega\left( \sigma^6 L^6 R^12 k \sigmamin(\Sigma_p)^{-2} \frac{\log^3 (1/\delta)}{\epsilon^2} \right)$.
+If the regularization strengths $\lambda_n^{(2)}$ and $\lambda_n^{(3)}$ are
+set to $\Omega(\sigma^3 L^3 R^6 \sqrt{\frac{\log(1/\delta)}{n}})$,
+then the parameter estimates $\hat\theta = (\hat\pi, \hat B)$ returned by
 \algorithmref{algo:spectral-experts} (with the columns appropriate permuted)
@@
 we can control the Frobenius norm of the parameter estimates:
-If $n = \Omega(d^p R^p \sigmamin(\Sigma_p)^{-1} \sqrt{p \log(d) \log(1/\delta)})$, then
+If $n \ge n_1$, then
 \begin{align}

### 2013-02-16 doc/icml2013/regression.tex
edit

@@
 
-The goal of this section is to bound the error of
+We now bound the error of
 the compound parameter estimates,
-specifically $\|\Delta_2\|_F^2$ and $\|\Delta_3\|_F^2$,
+$\|\Delta_2\|_F^2$ and $\|\Delta_3\|_F^2$,
 where $\Delta_2 \eqdef \hat M_2 - M_2$

### 2013-02-16 doc/icml2013/discussion.tex
changes

--- 2013-01-17 doc/icml2013/discussion.tex
@@
-\section{Discussion}
-\label{sec:discussion}
+\section{Conclusion}
+\label{sec:conclusion}
 
-Role of spectral algorithms in discriminative tasks.
-
-Challenges and ways forward?
+In this paper, we take a first step at developing computationally efficient and
+statistically consistent estimators for discriminative latent-variable models
+by focusing on mixture of linear regressions. We showed that by regressing on
+higher-order powers of the data with a regularizer that encourages low rank
+structure, coupled with an efficient tensor power method, we obtain a very natural algorithm.
+Empirically, our algorithm can outperform EM, and when used to initialize EM,
+yields significantly more accurate and stable results.

### 2013-02-16 doc/icml2013/evaluation.tex
edit
changes

@@
 
-\begin{figure*}[t]
+\begin{figure*}[p]
   \centering
@@
 
-\begin{table*}[t]
+\begin{table*}[p]
 \caption{Parameter Recovery Error $\|\theta^* - \hat \theta\|_F$ ($N = 500,000$).}
@@
 
-\begin{figure}[tb]
+\begin{figure}[th]
   \centering
@@
 consider the histogram in \figureref{fig:hist}, which shows the recovery
-errors of the algorithms over 170 attempts a dataset with $b = 1, d = 4,
+errors of the algorithms over 170 attempts on a dataset with $b = 1, d = 4,
 k = 3$. Typically, the spectral algorithm returned a stable solution.
@@
 
-\begin{figure*}[tb]
+\begin{figure*}[p]
   \centering
@@
   }
-  \caption{Recovey error vs $n$  ($b = 1, d = 5, K = 3, N = 500,000$).}
+  \caption{Recovery error vs $n$  ($b = 1, d = 5, K = 3, N = 500,000$).}
   \label{fig:vs-n}
@@
 error bars. We note that the recovery performance of EM does not
-particularly improve; this is expected because of the problem of local
-optima. The spectral algorithm's error decays slowly, and as it gets
-closer to zero, EM initialized at the spectral parameters finds the true
-parameters more often as well. This behaviour highlights the trade-off
-between statistical and computational error.
+particularly improve; this suggests that EM continues to get stuck in
+a local optima. The spectral algorithm's error decays slowly, and as it
+gets closer to zero, EM initialized at the spectral parameters finds the
+true parameters more often as well. This behavior highlights the
+trade-off between statistical and computational error.
 
@@
 
-\begin{table*}[tb]
+\begin{table*}[t]
 \caption{Parameter Recovery Error $\|\theta^* - \hat \theta\|_F$ with misspecified data ($N = 500,000$).}

### 2013-02-16 doc/icml2013/evaluation.tex
edit

--- 2013-02-15 doc/icml2013/evaluation.tex
@@
 was initialized with the regularized least squares solution and was run
-for 500 steps or till convergence. Both regularization parameters,
+for 500 iterations or until convergence up to a tolerance of $10^{-3}$. Both regularization parameters,
 $\lambda_n^{(2)}$ and, $\lambda_n^{(3)}$ were annealed as $\frac{1}{n}$.
@@
 
-We generated synthetic data using covariates from the $b$-dimensional
-unit hyper cube $\{-1,1\}^b$ and considered non-linear featurizations of
+We generated synthetic data using covariates $x$ sampled uniformly over the $b$-dimensional
+unit hypercube $[-1,1]^b$ and considered non-linear featurizations of
 this data that conformed to the identifiability criteria discussed in
-\sectionref{sec:algo}. The $\beta$ were drawn from a standard Gaussian.
+\sectionref{sec:algo}. The coefficients $\{\beta_h\}$ were drawn from a standard Gaussian.
 
@@
 %    \includegraphics[width=0.34\textwidth]{figures/curves-vis/1-8-3-spem.png}}
-  \caption{A comparison between SpectralExperts and EM. The dashed lines
+  \caption{A comparison between Spectral Experts and EM. The dashed lines
   on the left denote the solution recovered by Spectral Experts. Whilst
@@
 closer to zero, EM initialized at the spectral parameters finds the true
-parameters more often as well. This behaviour highlights the trade-off
+parameters more often as well. This behavior highlights the trade-off
 between statistical and computational error.

### 2013-02-16 doc/icml2013/algo.tex

@@
 where $n_1 = \Omega(d^p R^p \sigmamin(\Sigma_p)^{-1} \sqrt{p \log(d) \log(1/\delta)})$ and
-$n_2 = \Omega\left( \sigma^6 L^6 R^12 k \sigmamin(\Sigma_p)^{-2} \frac{\log^3 (1/\delta)}{\epsilon^2} \right)$.
+$n_2 = \Omega\left( \sigma^6 L^6 R^{12} k \sigmamin(\Sigma_p)^{-2} \frac{\log^3 (1/\delta)}{\epsilon^2} \right)$.
 If the regularization strengths $\lambda_n^{(2)}$ and $\lambda_n^{(3)}$ are

### 2013-02-16 doc/icml2013/evaluation.tex

@@
 error bars. We note that the recovery performance of EM does not
-particularly improve; this is expected because of the problem of local
-optima. The spectral algorithm's error decays slowly, and as it gets
-closer to zero, EM initialized at the spectral parameters finds the true
-parameters more often as well. This behavior highlights the trade-off
-between statistical and computational error.
+particularly improve; this suggests that EM continues to get stuck in
+a local optima. The spectral algorithm's error decays slowly, and as it
+gets closer to zero, EM initialized at the spectral parameters finds the
+true parameters more often as well. This behavior highlights the
+trade-off between statistical and computational error.
 

### 2013-02-16 doc/icml2013/evaluation.tex
edit

--- 2013-02-16 doc/icml2013/evaluation.tex
@@
 
-We generated synthetic data using covariates $x$ sampled uniformly over the $b$-dimensional
-unit hypercube $[-1,1]^b$ and considered non-linear featurizations of
-this data that conformed to the identifiability criteria discussed in
-\sectionref{sec:algo}. The coefficients $\{\beta_h\}$ were drawn from a standard Gaussian.
+We generated synthetic data as follows:
+First, we generated a vector $t$ sampled uniformly over the $b$-dimensional
+unit hypercube $[-1,1]^b$.
+Then, to get the actual covariates $x$, we applied a non-linear function of $t$
+that conformed to the identifiability criteria discussed in
+\sectionref{sec:algo}.
+The coefficients $\{\beta_h\}$ were drawn from a standard Gaussian.
 
-As an example, one set of features we considered in the uni-dimensional
-setting were $\{1, x, x^4, x^7\}$. The data and the curves fit using
+As an example, one feature map we considered in the one-dimensional
+setting was $x = (1, t, t^4, t^7)$. The data and the curves fit using
 Spectral Experts, EM with random initializations and EM initialized with
-the parameters recovered using spectral experts are shown in
-\figureref{fig:curves}. We note that even on well separated data such as
-this, EM converged to the global optima only 13\% of the time.
+the parameters recovered using Spectral Experts are shown in
+\figureref{fig:curves}. We note that even on well-separated data such as
+this, EM converged to the correct basin of attraction only 13\% of the time.
 
@@
   \caption{A comparison between Spectral Experts and EM. The dashed lines
-  on the left denote the solution recovered by Spectral Experts. Whilst
+  on the left denote the solution recovered by Spectral Experts. While
   not a perfect fit, it provides an good initialization for EM to find
@@
 
-\tableref{tbl:parameter-recovery} presents the Fro\"ebenius norm of the
+\tableref{tbl:parameter-recovery} presents the Frobenius norm of the
 difference between true and estimated parameters for the model, averaged

### 2013-02-16 doc/icml2013/evaluation.tex
edit

@@
 \begin{table*}[p]
-\caption{Parameter Recovery Error $\|\theta^* - \hat \theta\|_F$ ($N = 500,000$).}
+\caption{Parameter Recovery Error $\|\theta^* - \hat \theta\|_F$ ($n = 500,000$).}
 \label{tbl:parameter-recovery}
@@
 \abovespace\belowspace
-Variables ($b$) & Features ($D$) & Components ($K$) & Spectral & EM & Spectral + EM \\
+Variables ($b$) & Features ($d$) & Components ($k$) & Spectral & EM & Spectral + EM \\
 \hline
@@
 over 20 different random instances for each feature set and 10 attempts
-for each instance. The experiments were run using $N = 500,000$ samples.
+for each instance. The experiments were run using $n = 500,000$ samples.
 
@@
   \includegraphics[width=0.50\textwidth]{figures/hist.pdf}
-  \caption{Histogram over recovery errors between the three algorithms. $b = 1, d = 4, k = 3, N = 500,000$}
+  \caption{Histogram over recovery errors between the three algorithms. $b = 1, d = 4, k = 3, n = 500,000$}
   \label{fig:hist}
@@
   }
-  \caption{Recovery error vs $n$  ($b = 1, d = 5, K = 3, N = 500,000$).}
+  \caption{Recovery error vs $n$  ($b = 1, d = 5, k = 3, n = 500,000$).}
   \label{fig:vs-n}
@@
 \begin{table*}[t]
-\caption{Parameter Recovery Error $\|\theta^* - \hat \theta\|_F$ with misspecified data ($N = 500,000$).}
+\caption{Parameter Recovery Error $\|\theta^* - \hat \theta\|_F$ with misspecified data ($n = 500,000$).}
 \label{tbl:parameter-recovery-mis}
@@
 \abovespace\belowspace
-Variables ($b$) & Features ($d$) & Components ($K$) & Spectral & EM & Spectral + EM \\
+Variables ($b$) & Features ($d$) & Components ($k$) & Spectral & EM & Spectral + EM \\
 \hline

### 2013-02-16 doc/icml2013/intro.tex
Updated figures

@@
 \cite{NegahbanWainwright2009,Tomioka2011} on second and third powers of
-the response to recover the appropriate symmetric tensors, from which we
+the response to recover the appropriate symmetric tensors, using which we
 can then apply the tensor power method to retrieve the final parameters.

### 2013-02-16 doc/icml2013/regression.tex
Updated figures

@@
 %\label{sec:regression}
-
+\vspace{-0.5em}
 We now bound the error of

### 2013-02-16 doc/icml2013/discussion.tex
edit

--- 2013-01-17 doc/icml2013/discussion.tex
@@
-\section{Discussion}
-\label{sec:discussion}
+\section{Conclusion}
+\label{sec:conclusion}
 
-Role of spectral algorithms in discriminative tasks.
-
-Challenges and ways forward?
+In this paper, we developed a computationally efficient and statistically
+consistent estimator for mixture of linear regressions, a first step at
+tackling discriminative latent-variable models.  Our algorithm, Specral
+Experts, regresses on higher-order powers of the data with a regularizer that
+encourages low rank structure, followed by tensor factorization to recover the
+actual parameters.  Empirically, Spectral Experts generally does not outperform
+EM, but when used to initialize EM, yields significantly more accurate and
+stable results than EM.

### 2013-02-16 doc/icml2013/intro.tex
edit

--- 2013-02-15 doc/icml2013/intro.tex
@@
 of our technique to understand its statistical properties (\sectionref{sec:evaluation}).
-In particular, we find that in the low data regime,
-Spectral Experts underperforms EM due to its weaker statistical efficiency,
-but in the high data regime,
-Spectral Experts outperforms EM.
-We get the best results by using Spectral Experts as an initialization
-to EM.
+%In particular, we find that in the low data regime,
+%but in the high data regime,
+%Spectral Experts outperforms EM.
+While Spectral Experts generally does not outperform EM, presumably due to its
+weaker statistical efficiency, it serves as an effective initialization for EM,
+significantly outperforming EM with random initialization.
 

### 2013-02-16 doc/icml2013/algo.tex
edit

@@
 In the algorithm, the regularization strengths $\lambda_n^{(2)}$ and $\lambda_n^{(3)}$
-are set to $\frac{1}{\sqrt{n}}$.
+are set to $\frac{c}{\sqrt{n}}$ for some constant $c$.
 The resulting semidefinite program is a standard one which has received
 much attention in recent years.
-We use a rather simple gradient-based approach,
-where the nuclear norm is handled in closed form by taking an SVD
-and soft-thresholding the eigenvalues \cite{donoho95soft,cai10soft}.
+We use a rather simple proximal gradient-based approach,
+in which the nuclear norm is handled in closed form by taking an SVD
+and soft-thresholding the singular values \cite{donoho95soft,cai10soft}.
 
@@
 to recover the parameters \cite{anandkumar12moments,anandkumar12svd,AnandkumarHsuKakade2012}.
-Our work extends this line of work into the space of discriminative models,
-which requires regression to obtain the desired tensor.
+Our work extends this second line into the space of discriminative models,
+which requires regression to obtain the desired tensor to factorize.
 
 % Unmixing
-In spirit, Spectral Experts also bears some resemblance to the unmixing
+In spirit, Spectral Experts bears some resemblance to the unmixing
 algorithm for estimation of restricted PCFGs
@@
 parameters by inverting a mixing matrix.  In this work,
-each data point (appropriately transformed) provides a noisy slice of the
-compound parameters; retrieving the compound parameters is more involved.
+each data point (appropriately transformed) provides a different noisy projection of
+the compound parameters.
 
@@
 whereas in our case, we deliberately construct powers $y,y^2,y^3$
-to identify the parameters.
+to identify the underlying parameters.
 

### 2013-02-16 doc/icml2013/algo.tex

--- 2013-02-16 doc/icml2013/algo.tex
@@
 Another line of work is based on the method of moments and uses eigendecomposition of a certain tensor
-to recover the parameters \cite{anandkumar12moments,anandkumar12svd,AnandkumarHsuKakade2012}.
+to recover the parameters \cite{anandkumar12svd,anandkumar12moments,AnandkumarHsuKakade2012}.
 Our work extends this line of work into the space of discriminative models,

### 2013-02-16 doc/icml2013/evaluation.tex
Updated figures

@@
 was initialized with the regularized least squares solution and was run
-for 500 iterations or until convergence up to a tolerance of $10^{-3}$. Both regularization parameters,
-$\lambda_n^{(2)}$ and, $\lambda_n^{(3)}$ were annealed as $\frac{1}{n}$.
+for 500 iterations or until convergence up to a tolerance of $10^{-3}$.
+%Both regularization parameters, $\lambda_n^{(2)}$ and, $\lambda_n^{(3)}$ were annealed as $\frac{1}{n}$.
 
@@
   \centering
-  \subfigure[Spectral]{
+  \subfigure[Spectral, Spectral+EM]{
     \includegraphics[width=0.50\textwidth]{figures/curves-vis/1-8-3-specm.pdf}}
@@
 %    \includegraphics[width=0.34\textwidth]{figures/curves-vis/1-8-3-spem.png}}
-  \caption{A comparison between Spectral Experts and EM. The dashed lines
-  on the left denote the solution recovered by Spectral Experts. While
-  not a perfect fit, it provides an good initialization for EM to find
-  the true global minima. The dotted lines on the right show different
-  local optima found by EM on this example.}
+  \caption{Visualization of the parameters estimated by Spectral Experts versus EM.
+  (a) The dashed lines denote the solution recovered by Spectral Experts. While
+  not a perfect fit, it provides an good initialization for EM to further improve the solution (solid lines).
+  (b) The dotted lines show different local optima found by EM.}
   \label{fig:curves}
@@
 \begin{table*}[p]
-\caption{Parameter Recovery Error $\|\theta^* - \hat \theta\|_F$ ($n = 500,000$).}
+\caption{Parameter error $\|\theta^* - \hat \theta\|_F$ ($n = 500,000$)
+as the number of variables $b$, number of features $d$ and the number of components $k$ increases.
+While Spectral by itself does not produce good parameter estimates, Spectral+EM improves
+over EM significantly.
+}
 \label{tbl:parameter-recovery}
@@
   \centering
-  \subfigure[Well Data]{
+  \subfigure[Well-specified data]{
     \includegraphics[width=0.50\textwidth]{figures/vs-n/1833-decay.pdf}
@@
     \hspace{-2em}
-  \subfigure[Misspecified Data]{
+  \subfigure[Misspecified data]{
     \includegraphics[width=0.50\textwidth]{figures/vs-n/1-8-3-3-rm.pdf}
   }
-  \caption{Recovery error vs $n$  ($b = 1, d = 5, k = 3, n = 500,000$).}
+  \caption{Learning curves: parameter error as a function of the number of samples $n$ ($b = 1, d = 5, k = 3$).}
   \label{fig:vs-n}
@@
 \begin{table*}[t]
-\caption{Parameter Recovery Error $\|\theta^* - \hat \theta\|_F$ with misspecified data ($n = 500,000$).}
+\caption{Parameter rrror $\|\theta^* - \hat \theta\|_F$ when the data is misspecified ($n = 500,000$).
+Spectral+EM degrades slightly, but still outperforms EM overall.
+}
 \label{tbl:parameter-recovery-mis}

### 2013-02-16 doc/icml2013/algo.tex

--- 2013-02-16 doc/icml2013/algo.tex
@@
 Another line of work is based on the method of moments and uses eigendecomposition of a certain tensor
-to recover the parameters \cite{anandkumar12moments,anandkumar12svd,AnandkumarHsuKakade2012}.
-Our work extends this second line into the space of discriminative models,
-which requires regression to obtain the desired tensor to factorize.
+to recover the parameters \cite{anandkumar12svd,anandkumar12moments,AnandkumarHsuKakade2012}.
+Our work extends this line of work into the space of discriminative models,
+which requires regression to obtain the desired tensor.
 

### 2013-02-16 doc/icml2013/algo.tex
edit

--- 2013-02-16 doc/icml2013/algo.tex
@@
 convex optimization and tensor power method, respectively.
-% Motivate how we can recover $B$ from regression by lower order moments
-%ideally from lower order moments.
-%Conventionally, when using the method of moments, the
-%moments of observed variables are expressed algebraically in terms of
-%the model parameters. Using these equations, we attempt to solve for the
-%model parameters using empirical estimates of the moments.
 
@@
     \INPUT Three independent datasets $\mathcal{D}_r = \{ (\xn{1}, \yn{1}), \cdots, (\xn{n}, \yn{n}) \}$ for $r = 1, 2, 3$;
-    regularization strengths $\lambda_n^{(2)} = \frac{c_2}{\sqrt{n}}$, $\lambda_n^{(3)} = \frac{c_3}{\sqrt{n}}$;
+    %regularization strengths $\lambda_n^{(2)} = \frac{c_2}{\sqrt{n}}$, $\lambda_n^{(3)} = \frac{c_3}{\sqrt{n}}$;
+    regularization strengths $\lambda_n^{(2)}$, $\lambda_n^{(3)}$;
     observation noise variance $\sigma^2$.
@@
 \label{thm:convergence}
-Each dataset $\sD_p$ (for $p = 1, 2, 3$) consists of $n$ i.i.d. points drawn from the mixture
+Each dataset $\sD_p$ (for $p = 1, 2, 3$) consists of $n$ i.i.d. points drawn from a mixture
 of linear regressions model with parameter $\theta^*$
@@
 and for all $h \in [k]$,
-$\|\hat\beta_h - \beta^*_h\|_2 \le \frac{\epsilon}{\pi_h^*}$.
+$\|\hat\beta_h - \beta^*_h\|_2 \le \frac{\epsilon}{\sqrt{\pi_h^*}}$.
 \end{theorem}
 
-\begin{proof}
-By \reflem{lowRank} along with \reflem{lowRankLower} and \reflem{lowRankUpper} (discussed below),
-we can control the Frobenius norm of the parameter estimates:
-If $n \ge n_1$, then
-\begin{align}
-  \|\hat M_p - M_p\|_F = O\left( \sigma^3 L^3 R^6 k^{\frac12} \sigmamin(\Sigma_p)^{-1} \sqrt{\frac{\log^3 (1/\delta)}{n}} \right).
-  %\|\hat M_p - M_p\|_F = O\left( \frac{\sigma^3 L^3 R^6 \sqrt{k} \log^3 (1/\delta)}{\sqrt{n} (\sigmamin(\Sigma_p) - d^p R^p \sqrt{\frac{p \log(d) \log(1/\delta)}{n}}} \right).
-\end{align}
-The Frobenius norm upper bounds the operator norm,
-so we can directly apply Theorem 5.1 of \cite{AnandkumarGeHsu2012}.
-%Let $T^* = M_3^*(W^*, W^*, W^*)$.
-%Need to convert error from $\epsilon$
-%hit by $(W^\top)^\dagger$
-\end{proof}
+The proof of the theorem has two parts.
+First, we bound the error in the compound parameters estimates $\hat M_2,\hat M_3$
+using results from \citet{Tomioka2011}.
+Then, we use results from \cite{AnandkumarGeHsu2012} to convert this error
+into a bound on the actual parameter estimates $\hat\theta = (\hat\pi, \hat B)$
+derived from the robust tensor power method.
+But first, let us study a more basic property: identifiability.
 
@@
 components $t \cdot t$ and $2(1 \cdot t^2)$, which are linearly dependent.
-So our method would not be able to identify the parameters of a mixture of
+Therefore, Spectral Experts would not be able to identify the parameters of a mixture of
 linear regressions for this distribution over $x$.
 
-We now show that some amount of unidentifiability is intrinsic to the model,
-not just our estimation procedure.
+We now show that some amount of unidentifiability is intrinsic to estimation from low-order moments,
+not just an artifact of our estimation procedure.
 Suppose $x = (t, \dots, t^d)$.
@@
 $\E[x\tp{p}y\tp{p}]$ and $\E[x\tp{2p}]$ for $p \in [r]$,
-all the coordinates would be monomials up to degree $2dr$,
-the therefore the moments live in a $2(dr)$-dimensional subspace.
-The parameters live in a subspace of at least dimension $dk$.
+all the resulting coordinates would be monomials up to degree $2dr$,
+and thus the moments live in a $2(dr)$-dimensional subspace.
+On the other hand, the parameters $\theta$ live in a subspace of at least dimension $dk$.
 Therefore, at least $r \ge k/2$ moments are required for identifiability of any
 algorithm for this monomial example.
-
-%In fact, considering moments up to power $p
-%Regressing on all powers up to $p$ yield at most $3$.
-
-%$1, x_1, x_1^2, \dots, x_1^p$ is
-%neither quadratically nor cubically independent because the product
-%space is strictly a subset of $\Re^{\frac{d (d-1)}{2}}$ since it
-%contains $x_1 \times x_1 = x_1^2 \times 1$, and $x_1 \times x_1 \times
-%x_1 = x_1^2 \times x_1 \times 1 = x_1^3 \times 1 \times 1$.
-%In fact, the
-%number of unique terms in $x\tp{p}$ is only of the order of $p \times
-%d$, and hence we will not have enough equations to solve for $M_2$ or
-%$M_3$ unless we consider $p = d^2$ moments.
-
-%The mixture of linear regressions model is identifiable subject to
-%permutations of the components. However, our approach
-%employs regression on $x\tp{2}$ and $x\tp{3}$, and thus requires
-%$\mathcal{X} \otimes \mathcal{X}$ and $\mathcal{X} \otimes \mathcal{X}
-%\otimes \mathcal{X}$ to be linearly independent modulo symmetry. We call
-%these condition {\em quadratic} and {\em cubic} independence.
-
-%This condition trivially holds when the space $\mathcal{X}$ consists of
-%independent vectors $\{ x_1, x_2, \cdots, x_d \}$. Unfortunately, it
-%places some constraints on non-linear featurizations.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \subsection{Recovering $M_2$ and $M_3$}
-
-% In the case of the mixture of linear regressions, we have two observed
-% quantities, $x$ and $y$. Unfortunately, because $y$ is a scalar
-% quantity, we have only one equation for every moment of $y$, and with
-% this naive approach, would have to consider $O( (k+1) D)$ moments to
-% recover $B \in \Re{k \times D}$ and $\pi \in \Re^k$.
-
-% To keep the discussion simple, we will consider the case where there is
-% no noise, i.e.  $\sigma^2 = 0$. In this case, the moments of $y$ are,
-% \begin{eqnarray*}
-%   \E[ y ] &=& (\sum_{k=1}^{k} \pi_{k} \beta_k)^T x \\
-%           &=& M_1^T \E[x] \\
-%   \E[ y^2 ] &=& \sum_{k=1}^{k} \pi_{k} \E[(\beta_k^T x)]^2 \\
-%   &=& \sum_{k=1}^{k} \pi_{k} (\beta_k\tp{2} \odot \E[x\tp{2}]) \\
-%   &=& M_2 \odot \E[x\tp{2}] \\
-%   \E[ y^p ] &=& M_p \odot \E[x\tp{p}].
-% \end{eqnarray*}
-% 
-% We note that joint moments with $x$ will not help either; consider
-% \begin{eqnarray}
-%   \E[ y^p x^q ] &=& M_p \odot E[ x\tp{p+q} ].
-% \end{eqnarray}
-% The elements of $\E[ y^p x^q ]$ are linear combinations of $\E[y^p]$,
-% and hence we do not get any new equations; \todo{This isn't exactly
-% correct; should we even be presenting this line of reasoning?}. This is
-% to be expected, since any information about the $\beta$s is really
-% encoded in $y$, and not in the $X$.

### 2013-02-16 doc/icml2013/regression.tex
edit

--- 2013-02-16 doc/icml2013/regression.tex
@@
-\paragraph{Low-rank Regression}
+\paragraph{Low-rank regression}
 %\label{sec:regression}
@@
 \end{proof}
+
+With these lemmas in hand, we now give the proof of \refthm{convergence}.
+\begin{proof}
+By \reflem{lowRank} along with \reflem{lowRankLower} and \reflem{lowRankUpper} (discussed below),
+we can control the Frobenius norm of the parameter estimates:
+If $n \ge n_1$, then
+\begin{align}
+  \|\hat M_p - M_p\|_F = O\left( \sigma^3 L^3 R^6 k^{\frac12} \sigmamin(\Sigma_p)^{-1} \sqrt{\frac{\log^3 (1/\delta)}{n}} \right).
+  %\|\hat M_p - M_p\|_F = O\left( \frac{\sigma^3 L^3 R^6 \sqrt{k} \log^3 (1/\delta)}{\sqrt{n} (\sigmamin(\Sigma_p) - d^p R^p \sqrt{\frac{p \log(d) \log(1/\delta)}{n}}} \right).
+\end{align}
+The Frobenius norm upper bounds the operator norm,
+so we can directly apply Theorem 5.1 of \cite{AnandkumarGeHsu2012}.
+%Let $T^* = M_3^*(W^*, W^*, W^*)$.
+%Need to convert error from $\epsilon$
+%hit by $(W^\top)^\dagger$
+\end{proof}

### 2013-02-16 doc/icml2013/regression.tex
edit

@@
 Therefore, we incur another $O(\sigma^2 \cdot R \sigma \sqrt{\frac{\log(1/\delta_1)}{n}})$.
-Handling the other terms using conventional means
+We can handle the other terms using conventional means.
+Combining everything, we obtain
 $\|\hat\E_3[\eta_3(x) x\tp{3}]\|_\op = O(R^3 (L^3 R^3 + \sigma L^2 R^2 + \sigma^2 L R + \sigma^3 + \sigma^3 R) \sqrt{\frac{\log^3(1/\delta_1)}{n}})$.
-Finally, taking $\delta_1 = \delta/3$, and taking the union bound,
-we get our result, where we are bounding quantities fairly crudely for the sake of simplicity.
+
+Finally, taking $\delta_1 = \delta/3$, and taking the union bound over the bounds for $p \in \{1,2,3\}$,
+we get our result.
+Note that we are bounding quantities fairly crudely for the sake of simplicity.
 \end{proof}

### 2013-02-16 doc/icml2013/regression.tex
edit

@@
 
-\paragraph{Restricted strong convexity}
+%\paragraph{Restricted strong convexity}
 
-In this section, we lower bound the restricted strong convexity constant
-$\kappa(\opX_p)$.
+%Let us first lower bound the restricted strong convexity constant
+%$\kappa(\opX_p)$:
 
@@
 
-With these lemmas in hand, we now give the proof of \refthm{convergence}.
+These lemmas allow us to control the compound parameter error.
+We now apply them in the proof of \refthm{convergence}:
 \begin{proof}
-By \reflem{lowRank} along with \reflem{lowRankLower} and \reflem{lowRankUpper} (discussed below),
+By \reflem{lowRank} along with \reflem{lowRankLower} and \reflem{lowRankUpper},
 we can control the Frobenius norm of the parameter estimates:
@@
 The Frobenius norm upper bounds the operator norm,
-so we can directly apply Theorem 5.1 of \cite{AnandkumarGeHsu2012}.
+so we can directly apply Theorem 5.1 of \citet{AnandkumarGeHsu2012}.
 %Let $T^* = M_3^*(W^*, W^*, W^*)$.

### 2013-02-16 doc/icml2013/intro.tex
Changes to Lemma 2

@@
 let $\|X\|_F$ denote the Frobenius norm (square root of sum of squares of singular values),
+let $\|X\|_{\max}$ denote the max norm (elementwise maximum),
 let $\|X\|_\text{op}$ denote the operator norm (largest singular value).

### 2013-02-16 doc/icml2013/regression.tex
Changes to Lemma 2

--- 2013-02-16 doc/icml2013/regression.tex
@@
-\paragraph{Low-rank regression}
+\paragraph{Low-rank Regression}
 %\label{sec:regression}
@@
 Let $\Sigma_p \eqdef \E[\cvec(x\tp{p})\tp{2}]$.
-With probability at least $1-\delta$,
-$$\kappa(\opX) = \Omega\left(\sigmamin(\Sigma_p) - d^p R^{p} \sqrt{\frac{p \log(d) \log (1/\delta)}{n}}\right).$$
+If $$n = \Omega \left( p ~ d^{2p} \log(d) \frac{R^{2p} \log(1/\delta)}{\sigmamin(\Sigma_p)^2} \right),$$
+with probability at least $1-\delta$,
+$$\kappa(\opX) \ge \frac{\sigmamin(\Sigma_p)}{2}.$$
 \end{lemma}
@@
 To do this, we first apply Hoeffding's inequality elementwise.
-Since $\|x\| \le R$, we have that for each element $(i,j)$,
-$|(\hat\Sigma_p)_{ij} - (\Sigma_p)_{ij}| = O(R^{p}\sqrt{\frac{\log (1/\delta)}{n}})$.
+Since $\|x\|_2 \le R$, we have that for each element $(i,j)$,
+$|(\hat\Sigma_p)_{ij} - (\Sigma_p)_{ij}| = O(R^{2p}\sqrt{\frac{\log (1/\delta)}{n}})$.
 Applying the union bound over the $d^{2p}$ elements of $\hat\Sigma_p - \Sigma_p$,
 we have that the max norm is bounded:
-$\|\hat\Sigma_p - \Sigma_p\|_\text{\rm max} = O(R^{p} \sqrt{\frac{p \log(d) \log (1/\delta)}{n}})$.
+$\|\hat\Sigma_p - \Sigma_p\|_\text{\rm max} = O(R^{2p} \sqrt{\frac{p \log(d) \log (1/\delta)}{n}})$.
 The max norm times $d^p$ upper bounds the Frobenius norm, which upper bounds the operator norm, so we have that
-$\|\hat\Sigma_p - \Sigma_p\|_\text{\rm op} = O(d^p R^{p} \sqrt{\frac{p \log(d) \log (1/\delta)}{n}})$.
-Using the fact that $\sigmamin(\hat\Sigma_p) = \sigmamin(\Sigma_p) + \|\hat\Sigma_p - \Sigma_p\|_\text{\rm op}$
+$\|\hat\Sigma_p - \Sigma_p\|_\text{\rm op} = O(d^p R^{2p} \sqrt{\frac{p \log(d) \log (1/\delta)}{n}})$.
+Using the fact that $\sigmamin(\hat\Sigma_p) \ge \sigmamin(\Sigma_p) - \|\hat\Sigma_p - \Sigma_p\|_\text{\rm op}$
 yields the result.
@@
 \end{proof}
-
-With these lemmas in hand, we now give the proof of \refthm{convergence}.
-\begin{proof}
-By \reflem{lowRank} along with \reflem{lowRankLower} and \reflem{lowRankUpper} (discussed below),
-we can control the Frobenius norm of the parameter estimates:
-If $n \ge n_1$, then
-\begin{align}
-  \|\hat M_p - M_p\|_F = O\left( \sigma^3 L^3 R^6 k^{\frac12} \sigmamin(\Sigma_p)^{-1} \sqrt{\frac{\log^3 (1/\delta)}{n}} \right).
-  %\|\hat M_p - M_p\|_F = O\left( \frac{\sigma^3 L^3 R^6 \sqrt{k} \log^3 (1/\delta)}{\sqrt{n} (\sigmamin(\Sigma_p) - d^p R^p \sqrt{\frac{p \log(d) \log(1/\delta)}{n}}} \right).
-\end{align}
-The Frobenius norm upper bounds the operator norm,
-so we can directly apply Theorem 5.1 of \cite{AnandkumarGeHsu2012}.
-%Let $T^* = M_3^*(W^*, W^*, W^*)$.
-%Need to convert error from $\epsilon$
-%hit by $(W^\top)^\dagger$
-\end{proof}

### 2013-03-30 doc/icml2013/algo.tex
Several typo/math fixes for the paper

@@
 Now we are ready to state our full algorithm, which we call Spectral Experts
-(\algorithmref{sec:algo}).
+(\algorithmref{algo:spectral-experts}).
 First, we perform three regressions to recover the compound parameters
@@
 $n = \max(n_1,n_2)$
-where $n_1 = \Omega(d^p R^p \sigmamin(\Sigma_p)^{-1} \sqrt{p \log(d) \log(1/\delta)})$ and
-$n_2 = \Omega\left( \sigma^6 L^6 R^{12} k \sigmamin(\Sigma_p)^{-2} \frac{\log^3 (1/\delta)}{\epsilon^2} \right)$.
+where $n_1 = \Omega\left(d^6 R^12 (\max_{p \in [3]} \sigmamin(\Sigma_p)^{-2}) \log(d) \log(1/\delta)\right)$ and
+$n_2 = \Omega\left( \sigma^6 L^6 R^{12} k (\max_{p \in [3]} \sigmamin(\Sigma_p)^{-1}) \frac{\log^3 (1/\delta)}{\epsilon^2} \right)$.
 If the regularization strengths $\lambda_n^{(2)}$ and $\lambda_n^{(3)}$ are

### 2013-03-30 doc/icml2013/regression.tex
Several typo/math fixes for the paper

@@
 \begin{align}
-\opX_p(M_p; \sD)_i &\eqdef \frac12 \sum_{(x,y) \in \sD} \innerp{M_p}{x\tp{p}}_i, \quad i  \in [n].
+\opX_p(M_p; \sD) &\eqdef \frac12 \sum_{(x,y) \in \sD} \innerp{M_p}{x\tp{p}}
 \end{align}
@@
 Suppose there exists a restricted strong convexity constant $\kappa(\opX)$ such that
-$$\frac{1}{2n} \| \opX_p( \Delta )\|_2^2 \ge \kappa(\opX) \|\Delta\|^2_F \quad \text{and} \quad
-\lambda_n \ge \frac{\|\opX^*(\eta_p)\|_2}{n}.$$
+$$\frac{1}{2n} \| \opX_p( \Delta )\|_2^2 \ge \kappa(\opX_p) \|\Delta\|^2_F \quad \text{and} \quad
+\lambda_n \ge \frac{\|\opX_p^*(\eta_p)\|_2}{n}.$$
 Then the error of $\hat M_p$ is bounded as follows:
-$\| \hat M_p - M_p^* \|_F \le \frac{\lambda_n \sqrt{k}}{\kappa(\opX)}$.
+$\| \hat M_p - M_p^* \|_F \le \frac{\lambda_n \sqrt{k}}{\kappa(\opX_p)}$.
 \end{lemma}
@@
 Going forward, we need to lower bound the restricted strong convexity
-parameter $\kappa(\opX)$ and upper bound the adjoint operator
-$\|\opX^*(\eta_p)\|_{2}^2$.
+parameter $\kappa(\opX_p)$ and upper bound the adjoint operator
+$\|\opX_p^*(\eta_p)\|_{2}^2$.
 %We
@@
 Let $\Sigma_p \eqdef \E[\cvec(x\tp{p})\tp{2}]$.
-If $$n = \Omega \left( p ~ d^{2p} \log(d) \frac{R^{2p} \log(1/\delta)}{\sigmamin(\Sigma_p)^2} \right),$$
+If $$n = \Omega \left( p ~ d^{2p} \log(d) \frac{R^{4p} \log(1/\delta)}{\sigmamin(\Sigma_p)^2} \right),$$
 with probability at least $1-\delta$,
-$$\kappa(\opX) \ge \frac{\sigmamin(\Sigma_p)}{2}.$$
+$$\kappa(\opX_p) \ge \frac{\sigmamin(\Sigma_p)}{4}.$$
 \end{lemma}
@@
 $\frac{1}{2n} \|\opX_p(\Delta)\|_2^2
-\ge \frac{1}{2} \trace(\cvec(\Delta)\tp{2} \hat\Sigma_p)
+= \frac{1}{2} \trace(\cvec(\Delta)\tp{2} \hat\Sigma_p)
 \ge \frac{1}{2} \|\Delta\|_F^2 \sigmamin(\hat\Sigma_p)$.
@@
 The first term is $\innerp{M_1 - \beta_h}{x}$.
-We have that $\|M_1 - \beta_h\|_2 \le L$,
+We have that $\|M_1 - \beta_h\|_2 \le 2L$,
 so by Hoeffding's inequality, and using the fact that $\|x\|_2 \le R$,
@@
 
-These lemmas allow us to control the compound parameter error.
-We now apply them in the proof of \refthm{convergence}:
+\paragraph{Method of Moments}
+
+Now that we have bounds on the errors of recovering $M_2$ and $M_3$ from
+the data, we will extend the analysis of the robust tensor method in
+\citet{AnandkumarGeHsu2012} to bound the error on the parameters
+$\theta$. 
+\begin{lemma}
+  \label{lem:tensorPower} Let $\|\hat M_2 - M_2\|_{op} \le \epsilon$ and
+  $\|\hat M_3 - M_3\|_{op} \le \epsilon$, for some $\epsilon$ such that 
+  $$\epsilon < \min\{\sigmamin(M_2)/2,O(k \sigmamin(M_2)^{1.5} \|M_3\|_{op}^{-1})\}.$$ 
+  Then, for some permutation of indices, the parameter estimates found
+  in part 2 of \algorithmref{algo:spectral-experts} satisfy the
+  following with probability atleast $1 - \delta$,
+  \begin{align*}
+  \|\hat \pi - \pi \|_{\infty}
+    &= O( k \pi_{max}^{2.5} \pi_{min}^{-0.5} \sigmamin(M_2)^{-1.5} \|M_3\|_{op} \epsilon ) \\
+  \|\hat \beta_h - \beta_h\|_2
+    &= O( k \pi_{max} \sigmamin(M_2)^{-1.5} \|M_2\|_{op}^{0.5} \|M_3\|_{op} \epsilon ).
+  \end{align*}
+  for all $h \in [k]$.
+\end{lemma}
+The proof of the above lemma has been deferred to
+\appendixref{sec:proofs}.
+
+\paragraph{Synthesis}
+Together, these lemmas allow us to control the compound parameter error
+and the recovery error. We now apply them in the proof of
+\refthm{convergence}:
+
 \begin{proof}
-By \reflem{lowRank} along with \reflem{lowRankLower} and \reflem{lowRankUpper},
-we can control the Frobenius norm of the parameter estimates:
-If $n \ge n_1$, then
+By \reflem{lowRank} along with \reflem{lowRankLower} and
+\reflem{lowRankUpper}, we can control the operator norm of the parameter
+estimates which are directly upper bounded by the Frobenius norm: If $n
+\ge n_1$, then
 \begin{align}
-  \|\hat M_p - M_p\|_F = O\left( \sigma^3 L^3 R^6 k^{\frac12} \sigmamin(\Sigma_p)^{-1} \sqrt{\frac{\log^3 (1/\delta)}{n}} \right).
+  \|\hat M_p - M_p\|_{op} = O\left( \sigma^3 L^3 R^6 k^{\frac12} \sigmamin(\Sigma_p)^{-1} \sqrt{\frac{\log^3 (1/\delta)}{n}} \right).
   %\|\hat M_p - M_p\|_F = O\left( \frac{\sigma^3 L^3 R^6 \sqrt{k} \log^3 (1/\delta)}{\sqrt{n} (\sigmamin(\Sigma_p) - d^p R^p \sqrt{\frac{p \log(d) \log(1/\delta)}{n}}} \right).
 \end{align}
-The Frobenius norm upper bounds the operator norm,
-so we can directly apply Theorem 5.1 of \citet{AnandkumarGeHsu2012}.
-%Let $T^* = M_3^*(W^*, W^*, W^*)$.
-%Need to convert error from $\epsilon$
-%hit by $(W^\top)^\dagger$
+
+We complete the proof by applying \reflem{tensorPower} with the above
+bound on $\|\hat M_p - M_p\|_{op}$.
+
 \end{proof}

### 2013-04-01 doc/icml2013/rebuttal.txt
Inclusions for the rebuttal round

Thank you for your constructive comments and suggestions; we will try to
respond to your questions below.

# Major Comments:

## Reviewer 30 (Regarding Lemma 3):
The $O(1/\sqrt{n})$ convergence follows from applying Hoeffding's
inequality on each term of the random vectors $\eta_p(x) x$, which have
zero mean when taking expectation over $h$ and $\epsilon$. For example,
in $\eta_2(x) x$, we have three terms (Equation 3), which each have
expectation zero by independence of $x$, $h$ and $\epsilon$.

This does not require that the $x$ have zero mean, and note that we do
not take expectations over $x$ at any point in our analysis.

## Reviewer 4 (Regarding analysis with whitening):
Your point is well noted; results from standard techniques (Hsu
& Kakade, 2013) show that the whitening step introduces an additional
factor of $O(\sigma_{min}(M_2)^{-1.5} \|M_2\|_{op}^{0.5} \|M_3\|_{op})$.
We will include this as a lemma in the final version as we can not
upload an updated version in the rebuttal phase.

## Reviewer 30 (Regarding convergence rate in experiments):

One of the obstacles in applying the bounds we derived is that it
requires the error of the empirical moment estimates to be less than
$\sigma_{\min}(M_2)/2$, a condition that did not hold true for our
experiments. For the data in Figure 3, this condition required the
moment errors to be less than 0.15, while the error in estimates we
recovered were at least 0.9. We have also identified that one major
reason for this was that the proximal gradient algorithm we used did not
converge in the specified number of iterations (i.e. 500).

# Minor Comments:

## Reviewer 30 (Regarding Identifiability):
Identifiability is only a problem when the input features are not
algebraically independent, i.e. each coordinate can not be expressed as
a deterministic analytic function of the remaining coordinates. Thus, it
is identifiable when the coordinates are generated from a random
Gaussian (or any other distribution).

## Reviewer 4 (Regarding Lemma 3):
We do not need to know $h$; in fact we are trying to bound the
difference between mean, $M_1$ which we recover by regression, and the
actual value of $\beta_h$ at each data point; it is useful to think of
it as bounding the variance of the zero-mean term $M_1 - \beta_h$.

## Reviewer 30 (Regarding operator norm for vectors):
This was a typo, and we meant the 2-norm.

## Reviewer 4 (Regarding references):
Thank you, we will include them in the final version.

## Reviewer 4 (Regarding $\sigma$):
We used $\sigma = 0.1$ for the results we presented, but found that they
were not very sensitive to the noise level.

## Reviewer 4 (Regarding Regularization):
We set the regularization parameters by grid search for $n = 10^5$ and
used the $1/\sqrt{n}$ rate to extrapolate for other $n$.

## Reviewer 7 (Regarding Generality of Experiments):
We found that robust initialization happened for almost all inputs. The
cases where we did not get robust initialization occurred when we did
not have good enough moment estimates and thus poor parameter estimates.
Furthermore, this problem went away when we used more samples.

## Reviewer 7 (Regarding Plot of Convergence):
We have compared the convergence of spectral experts and EM to the exact
parameters for increasing values of $n$ in Figure 3 in terms of the parameter
error.

### 2013-04-01 java/scripts/plotAnimation.py
added some plotting scripts

@@ -1,0 +1,104 @@
+import numpy as np
+import operator
+import matplotlib
+matplotlib.use("Agg")
+
+import matplotlib.pyplot as plt
+from matplotlib import animation 
+
+def parseBetasFile(fname):
+  betas = []
+
+  beta = []
+  for line in open(fname).xreadlines():
+    if line.startswith("betas"):
+      if( len(beta) > 0 ):
+        betas.append(np.array(beta))
+        beta = []
+    else:
+      try:
+        beta.append( map(float, line.split()) )
+      except Exception as e:
+        print "[Warning]: Couldn't parse line", line
+        print e
+        continue
+  if( len(beta) > 0 ):
+    betas.append(np.array(beta))
+  return betas
+
+def fs153(x):
+    return np.array([1.0, x, x**4])
+
+def getLine( beta, fn, W ):
+    """Gets a surface with X, Y, Z = f(X,Y) to be plotted"""
+    X = np.linspace( -1, 1, W)
+    # Grid of X's
+    Y = map(lambda (x): beta.dot(fn(x)), X)
+
+    return X, Y
+
+def plotLine(ax, X, Y, color):
+    ax.plot(X, Y, color)
+
+def plotLines(ax, betas, colors = None, W = 100):
+    if colors is None:
+        colors = ['b','r','g']
+    for beta, color in zip(betas.T, colors):
+        X, Y = getLine( beta, fs153, W )
+        plotLine(ax, X, Y, color)
+
+def plotAnimation( data, trueBetas, stepsBetas, out = "animation.mp4" ):
+    """Plot an animation with betas being fixed, and betas_ coming from
+    EM"""
+    D, K = trueBetas.shape
+
+    fig = plt.figure()
+
+    ax = plt.axes(xlim=(-1, 1), ylim=(-3, 3))
+
+    # Draw data
+    ax.scatter( data[0], data[1], alpha=0.3 )
+    # Draw the true betas
+    plotLines( ax, trueBetas )
+
+    # Draw dashed lines for the steps
+    lines = reduce( operator.add, 
+            [ax.plot([], [], color, lw=2) for (_, color) in zip(range(K), ["b--","r--","g--"])])
+
+    # initialization function: plot the background of each frame
+    def init():
+        for line in lines:
+            line.set_data([], [])
+        return lines
+
+    # animation function.  This is called sequentially
+    def animate(i):
[…]

### 2013-04-01 java/scripts/plotLines.py

@@ -1,0 +1,103 @@
+import numpy as np
+import operator
+import matplotlib
+#matplotlib.use("Agg")
+
+import matplotlib.pyplot as plt
+
+def parseBetasFile(fname):
+  betas = []
+
+  beta = []
+  for line in open(fname).xreadlines():
+    if line.startswith("betas"):
+      if( len(beta) > 0 ):
+        betas.append(np.array(beta))
+        beta = []
+    else:
+      try:
+        beta.append( map(float, line.split()) )
+      except Exception as e:
+        print "[Warning]: Couldn't parse line", line
+        print e
+        continue
+  if( len(beta) > 0 ):
+    betas.append(np.array(beta))
+  return betas
+
+def fs153(x):
+    return np.array([1.0, x, x**4])
+
+def getLine( beta, fn, W ):
+    """Gets a surface with X, Y, Z = f(X,Y) to be plotted"""
+    X = np.linspace( -1, 1, W)
+    # Grid of X's
+    Y = map(lambda (x): beta.dot(fn(x)), X)
+
+    return X, Y
+
+def plotLine(ax, X, Y, color):
+    ax.plot(X, Y, color)
+
+def plotLines(ax, betas, modifier = "-", colors = None, W = 100):
+    if colors is None:
+        colors = ['b','r','g']
+    colors = [ color + modifier for color in colors ]
+    for beta, color in zip(betas.T, colors):
+        X, Y = getLine( beta, fs153, W )
+        plotLine(ax, X, Y, color)
+
+def makePlot( data, trueBetas, specBetas, spemBetas, em0Betas, emBetas ):
+    fig = plt.figure()
+    ax = plt.axes(xlim=(-1, 1), ylim=(-3, 3))
+
+    # Plot the data
+    ax.scatter( data[0], data[1], alpha = 0.3 )
+
+    # Now draw each line
+    plotLines( ax, trueBetas, "-" )
+
+    # Spectral is dashed
+    plotLines( ax, specBetas, "--" )
+    plotLines( ax, spemBetas, "--" )
+
+    # EM is dash-dot
+    plotLines( ax, em0Betas, ".." )
+    plotLines( ax, emBetas, ".." )
+
+    plt.show()
+
+
+def getBeta( specifier ):
+  """ Get a particular beta"""
+  if ":" in specifier:
+      fname, idx = specifier.split(':')
+  else:
[…]

### 2013-04-01 java/scripts/plotPlanes.py

@@ -1,0 +1,84 @@
+import numpy as np
+from mpl_toolkits.mplot3d import Axes3D
+import matplotlib.pyplot as plt
+
+def parseBetasFile(fname):
+  betas = []
+
+  beta = []
+  for line in open(fname).xreadlines():
+    if line.startswith("betas"):
+      if( len(beta) > 0 ):
+        betas.append(np.array(beta))
+        beta = []
+    else:
+      try:
+        beta.append( map(float, line.split()) )
+      except Exception as e:
+        print "[Warning]: Couldn't parse line", line
+        print e
+        continue
+  if( len(beta) > 0 ):
+    betas.append(np.array(beta))
+  return betas
+
+def fs(x,y):
+    return np.array([1.0, x, y, x * y**3, x**2 * y**2, x**3 * y])
+
+def getSurface( beta, fn, W ):
+    """Gets a surface with X, Y, Z = f(X,Y) to be plotted"""
+    u = np.linspace( -1, 1, W)
+    # Grid of X's
+    X, Y = np.meshgrid(u,u)
+    Z = map(lambda (x,y): beta.dot(fn(x,y)), zip(X.flatten(),Y.flatten()))
+    Z = np.array(Z).reshape( W, W )
+
+    return X, Y, Z
+
+def plotSurface(ax, X, Y, Z, color ):
+    surf = ax.plot_surface(X, Y, Z, color = color, alpha = 0.7 )
+def plotWireframe(ax, X, Y, Z, color ):
+    surf = ax.plot_wireframe(X, Y, Z, color = color, alpha = 0.7 )
+
+def plotSurfaces(ax, betas, colors = None, W = 100):
+    if colors is None:
+        colors = ['b','r','g']
+    for beta, color in zip(betas.T, colors):
+        X, Y, Z = getSurface( beta, W )
+        plotSurface(ax, X, Y, Z, color)
+
+def plotWireframes(ax, betas, colors = None, W = 100):
+    if colors is None:
+        colors = ['b','r','g']
+    for beta, color in zip(betas.T, colors):
+        X, Y, Z = getSurface( beta, W )
+        plotWireframe(ax, X, Y, Z, color)
+
+def getAxes():
+    fig = plt.figure()
+    ax = fig.gca(projection='3d')
+    return ax
+
+def plotDifferences( ax, betas, betas_, colors = None, W = 100):
+    if colors is None:
[…]

### 2013-04-23 doc/icml2013/algo.tex
Couple of corrections to the math in the theory and introduction section. Included a line about mode-k tensor unfoldings. Added details about $\sigma^2$ to the experiments

@@
   y &= \innerp{M_1}{x} +
-  \underbrace{(\innerp{M_1 - \beta_h}{x} + \epsilon)}_{\eqdef \eta_1(x)}. \label{eqn:y1}
+  \underbrace{(\innerp{\beta_h - M_1}{x} + \epsilon)}_{\eqdef \eta_1(x)}. \label{eqn:y1}
 \end{align}
@@
 \end{algorithm}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Theoretical results}
-
-In this section, we provide theoretical guarantees for our Spectral Experts algorithm.
-In particular, the following theorem provides a rate of convergence:
-
-\begin{theorem}[Convergence of Spectral Experts]
-\label{thm:convergence}
-Each dataset $\sD_p$ (for $p = 1, 2, 3$) consists of $n$ i.i.d. points drawn from a mixture
-of linear regressions model with parameter $\theta^*$
-(having three independent copies simplifies the analysis).
-Let $\|x\|_2 \le R$ with probability $1$
-and let $\|\beta_h^*\|_2 \le L$ for all $h \in [k]$.
-Let $\Sigma_p \eqdef \E[\cvec(x\tp{p})\tp{2}]$, 
-and assume $\Sigma_p \succ 0$ for each $p \in \{1,2,3\}$.
-Assume $d \ge k$.
-Suppose the number of samples is
-$n = \max(n_1,n_2)$
-where $n_1 = \Omega\left(d^6 R^12 (\max_{p \in [3]} \sigmamin(\Sigma_p)^{-2}) \log(d) \log(1/\delta)\right)$ and
-$n_2 = \Omega\left( \sigma^6 L^6 R^{12} k (\max_{p \in [3]} \sigmamin(\Sigma_p)^{-1}) \frac{\log^3 (1/\delta)}{\epsilon^2} \right)$.
-If the regularization strengths $\lambda_n^{(2)}$ and $\lambda_n^{(3)}$ are
-set to $\Omega(\sigma^3 L^3 R^6 \sqrt{\frac{\log(1/\delta)}{n}})$,
-then the parameter estimates $\hat\theta = (\hat\pi, \hat B)$ returned by
-\algorithmref{algo:spectral-experts} (with the columns appropriate permuted)
-satisfies 
-$\|\hat\pi - \pi^*\|_{\infty} \le \epsilon$
-and for all $h \in [k]$,
-$\|\hat\beta_h - \beta^*_h\|_2 \le \frac{\epsilon}{\sqrt{\pi_h^*}}$.
-\end{theorem}
-
-The proof of the theorem has two parts.
-First, we bound the error in the compound parameters estimates $\hat M_2,\hat M_3$
-using results from \citet{Tomioka2011}.
-Then, we use results from \cite{AnandkumarGeHsu2012} to convert this error
-into a bound on the actual parameter estimates $\hat\theta = (\hat\pi, \hat B)$
-derived from the robust tensor power method.
-But first, let us study a more basic property: identifiability.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Identifiability from moments}
-
-In ordinary linear regression, the regression coefficients $\beta \in \Re^d$ are
-identifiable if and only if the data has full
-rank: $\E[x\tp{2}] \succ 0$,
-and furthermore, identifying $\beta$ requires only moments
-$\E[xy]$ and $\E[x\tp{2}]$ (by observing the optimality conditions for \refeqn{y1}).
-However, in mixture of linear regressions, these two moments only allow us to recover $M_1$.
-\refthm{convergence} shows that if we have the higher order analogues,
-$\E[x\tp{p}y\tp{p}]$ and $\E[x\tp{2p}]$ for $p \in \{1,2,3\}$,
-we can then identify the parameters $\theta = (\pi, B)$,
-provided $\E[\cvec(x\tp{p})\tp{2}] \succ 0$ for $p \in \{1,2,3\}$.
-
-We note that the positive definite constraints are more stringent.
-For example, suppose $x = (1, t, t^2)$,
-the common polynomial basis expansion, so that all the coordinates are deterministically related.
-Then $\E[x\tp{2}] \succ 0$ can easily be satisfied, say with $x \sim \normal{0}{I}$.
-However, one can verify that $\E[\cvec(x\tp{2})\tp{2}]$ is singular,
-since $\cvec(x\tp{2}) = [1 \cdot 1, t\cdot t, 2(1 \cdot t^2), 2(t \cdot t^2), (t^2 \cdot t^2)]$ contains
-components $t \cdot t$ and $2(1 \cdot t^2)$, which are linearly dependent.
-Therefore, Spectral Experts would not be able to identify the parameters of a mixture of
-linear regressions for this distribution over $x$.
-
-We now show that some amount of unidentifiability is intrinsic to estimation from low-order moments,
-not just an artifact of our estimation procedure.
-Suppose $x = (t, \dots, t^d)$.
-Even if we observed all moments
-$\E[x\tp{p}y\tp{p}]$ and $\E[x\tp{2p}]$ for $p \in [r]$,
-all the resulting coordinates would be monomials up to degree $2dr$,
-and thus the moments live in a $2(dr)$-dimensional subspace.
-On the other hand, the parameters $\theta$ live in a subspace of at least dimension $dk$.
-Therefore, at least $r \ge k/2$ moments are required for identifiability of any
-algorithm for this monomial example.

### 2013-04-23 doc/icml2013/evaluation.tex
Several typo/math fixes for the paper
Couple of corrections to the math in the theory and introduction section. Included a line about mode-k tensor unfoldings. Added details about $\sigma^2$ to the experiments

@@
 \sectionref{sec:algo}.
-The coefficients $\{\beta_h\}$ were drawn from a standard Gaussian.
+The coefficients $\{\beta_h\}$ were drawn from a standard Gaussian. The
+results presented below use $\sigma^2 = 0.1$; we did not observe any
+qualitatively different behaviour for choices of $\sigma^2$ in the range
+$[0.01, 0.4]$.
 
@@
 
-\begin{table*}[p]
+\begin{table*}[tbhp]
 \caption{Parameter error $\|\theta^* - \hat \theta\|_F$ ($n = 500,000$)
@@
 
-\begin{figure}[th]
+\begin{figure}[tbhp]
   \centering
@@
 
-\begin{figure*}[p]
+\begin{figure*}[tbhp]
   \centering
@@
 
-\paragraph{Misspecified Data}
+\paragraph{Misspecified data}
 
-\begin{table*}[t]
-\caption{Parameter rrror $\|\theta^* - \hat \theta\|_F$ when the data is misspecified ($n = 500,000$).
+\begin{table*}[tbhp]
+\caption{Parameter error $\|\theta^* - \hat \theta\|_F$ when the data is misspecified ($n = 500,000$).
 Spectral+EM degrades slightly, but still outperforms EM overall.

### 2013-04-23 doc/icml2013/intro.tex
Couple of corrections to the math in the theory and introduction section. Included a line about mode-k tensor unfoldings. Added details about $\sigma^2$ to the experiments

@@
 
-We use $x\tp{p}$ to represent the $p$-th order tensor
-formed by taking the outer product of $x \in \Re^d$; i.e. $x\tp{p}_{i_1 \ldots
-i_p} = x_{i_1} \cdots x_{i_p}$. We will use
-$\innerp{\cdot}{\cdot}$ to denote the generalized dot product between two $p$-th
-order tensors: $\innerp{X}{Y} = \sum_{i_1, \ldots i_p} X_{i_1,
-\ldots i_p} Y_{i_1, \ldots i_p}.$
-A tensor $X$ is symmetric if for all $i,j \in [d]^p$ which are permutations of each other,
-$X_{i_1 \cdots i_p}$ = $X_{j_1 \cdots j_p}$ (all tensors in this paper will be symmetric).
-For a $p$-th order tensor $X \in (\Re^d)\tp{p}$, let $X_{(i)} \in \Re^{d \times d^{p-1}}$
+We use $x\tp{p}$ to represent the $p$-th order tensor formed by taking
+the outer product of $x \in \Re^d$; i.e. $x\tp{p}_{i_1 \ldots i_p}
+= x_{i_1} \cdots x_{i_p}$. We will use $\innerp{\cdot}{\cdot}$ to denote
+the generalized dot product between two $p$-th order tensors:
+$\innerp{X}{Y} = \sum_{i_1, \ldots i_p} X_{i_1, \ldots i_p} Y_{i_1,
+\ldots i_p}$.  A tensor $X$ is symmetric if for all $i,j \in [d]^p$
+which are permutations of each other, $X_{i_1 \cdots i_p}$ = $X_{j_1
+\cdots j_p}$ (all tensors in this paper will be symmetric).  For
+a $p$-th order tensor $X \in (\Re^d)\tp{p}$, the mode-$i$ unfolding of
+$X$ is a matrix whose $j$-th row contains all the elements of $X$ whose
+$i$-th index is equal to $j$. Let $X_{(i)} \in \Re^{d \times d^{p-1}}$
 be the mode-$i$ unfolding of $X$.
@@
 %$\innerp{X}{Y} \in \Re^{d^q}$ is the projection of $X$ onto the first
-%$p$ modes of $Y$, i.e. $\innerp{X}{Y}_{i_1, \ldots, i_q}
-%= \sum_{j_1, \ldots, j_p} X_{j_1, \ldots, j_p} Y_{j_1, \ldots, j_p, i_1,
-%\ldots, i_q}$.
+%$p$ modes of $Y$, i.e. $\innerp{X}{Y}_{i_1, \ldots, i_q} = \sum_{j_1,
+%\ldots, j_p} X_{j_1, \ldots, j_p} Y_{j_1, \ldots, j_p, i_1, \ldots,
+%i_q}$.
 
@@
 let $\|X\|_\text{op}$ denote the operator norm (largest singular value).
-For a rank-$K$ tensor $X$,
+For a tensor $X$,
 let $\|X\|_* = \frac{1}{p} \sum_{i=1}^p \|X_{(i)}\|_*$ denote
@@
 
-For a tensor $X \in (\Re^d)\tp{p}$,
-let $\cvec(X) \in \Re^{\sum_{i=1}^p \binom{n}{i}}$ be
-the collapsed vectorization of $X$: each component of $\cvec(X)$
-is indexed by a vector of counts $(c_1, \dots, c_d)$ with total sum $\sum_i c_i = p$.
-The value of that component is
-$\sum_{k \in K(c)} X_{k_1 \cdots k_p}$,
-where $K(c) = \{ k \in [d]^p : \forall i \in [d], c_i = |\{ j \in [p] : k_j = i \}| \}$ are the set of index vectors $k$ with that count profile.
-For example, for $X \in \Re^{d \times d}$,
-$\cvec(X) = (X_{ii} : i \in [d]; X_{ij} + X_{ji} : i,j \in [d], i<j)$.
+For a symmetric tensor $X \in (\Re^d)\tp{p}$, let $\cvec(X) \in
+\Re^{\binom{d + p + 1}{p}}$ be the collapsed vectorization of distinct
+elements in $X$: each component of $\cvec(X)$ is indexed by a vector of
+counts $(c_1, \dots, c_d)$ with total sum $\sum_i c_i = p$.  The value
+of that component is $\sum_{k \in K(c)} X_{k_1 \cdots k_p}$, where $K(c)
+= \{ k \in [d]^p : \forall i \in [d], c_i = |\{ j \in [p] : k_j = i \}|
+\}$ are the set of index vectors $k$ with that count profile.  For
+example, for $X \in \Re^{d \times d}$, $\cvec(X) = (X_{ii} : i \in [d];
+X_{ij} + X_{ji} : i,j \in [d], i<j)$.
 %For symmetric tensors, $\innerp{X}{Y} = \innerp{\cvec(X)}{\cvec(Y)}$.

### 2013-04-23 doc/icml2013/theory.tex
Couple of corrections to the math in the theory and introduction section. Included a line about mode-k tensor unfoldings. Added details about $\sigma^2$ to the experiments

@@
-\section{Theoretical Properties}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Theoretical results}
 \label{sec:theory}
+
+In this section, we provide theoretical guarantees for our Spectral Experts algorithm.
+In particular, the following theorem provides a rate of convergence:
+
+\begin{theorem}[Convergence of Spectral Experts]
+\label{thm:convergence}
+Each dataset $\sD_p$ (for $p = 1, 2, 3$) consists of $n$ i.i.d.\ points drawn from a mixture
+of linear regressions model with parameter $\theta^*$
+(having three independent copies simplifies the analysis).
+Let $\|x\|_2 \le R$ with probability $1$
+and let $\|\beta_h^*\|_2 \le L$ for all $h \in [k]$.
+Let $\Sigma_p \eqdef \E[\cvec(x\tp{p})\tp{2}]$, 
+and assume $\Sigma_p \succ 0$ for each $p \in \{1,2,3\}$.
+Assume $d \ge k$.
+Suppose the number of samples is
+$n = \max(n_1,n_2)$
+where $n_1 = \Omega\left(d^6 R^{12} (\max_{p \in [3]} \sigmamin(\Sigma_p)^{-2}) \log(d) \log(1/\delta)\right)$ and
+$n_2 = \Omega\left( \sigma^6 L^6 R^{12} k (\max_{p \in [3]} \sigmamin(\Sigma_p)^{-1}) \frac{\log^3 (1/\delta)}{\epsilon^2} \right)$.
+If the regularization strengths $\lambda_n^{(2)}$ and $\lambda_n^{(3)}$ are
+set to $\Omega\left(\sigma^3 L^3 R^6 \sqrt{\frac{\log(1/\delta)}{n}}\right)$,
+then the parameter estimates $\hat\theta = (\hat\pi, \hat B)$ returned by
+\algorithmref{algo:spectral-experts} (with the columns appropriate permuted)
+satisfies 
+$\|\hat\pi - \pi^*\|_{\infty} \le \epsilon$
+and for all $h \in [k]$,
+$\|\hat\beta_h - \beta^*_h\|_2 \le \frac{\epsilon}{\sqrt{\pi_h^*}}$.
+\end{theorem}
+
+The proof of the theorem has two parts.
+First, we bound the error in the compound parameters estimates $\hat M_2,\hat M_3$
+using results from \citet{Tomioka2011}.
+Then, we use results from \cite{AnandkumarGeHsu2012} to convert this error
+into a bound on the actual parameter estimates $\hat\theta = (\hat\pi, \hat B)$
+derived from the robust tensor power method.
+But first, let us study a more basic property: identifiability.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Identifiability from moments}
+
+In ordinary linear regression, the regression coefficients $\beta \in
+\Re^d$ are identifiable if and only if the data has full rank:
+$\E[x\tp{2}] \succ 0$, and furthermore, identifying $\beta$ requires
+only moments $\E[xy]$ and $\E[x\tp{2}]$ (by observing the optimality
+conditions for \refeqn{y1}).  However, in mixture of linear regressions,
+these two moments only allow us to recover $M_1$.  \refthm{convergence}
+shows that if we have the higher order analogues, $\E[x\tp{p}y\tp{p}]$
+and $\E[x\tp{2p}]$ for $p \in \{1,2,3\}$, we can then identify the
+parameters $\theta = (\pi, B)$, provided $\E[\cvec(x\tp{p})\tp{2}] \succ
+0$ for $p \in \{1,2,3\}$.
+
+We note that this requires more stringent positive definite constraints
+when using features. For example, suppose $x = (1, t, t^2)$, the common
+polynomial basis expansion, so that all the coordinates are
+deterministically related.  Then $\E[x\tp{2}] \succ 0$ can easily be
+satisfied. 
+%, say with $x \sim \normal{0}{I}$.
+However, one can verify that $\E[\cvec(x\tp{2})\tp{2}]$ is singular for
+any data distribution, since $\cvec(x\tp{2}) = [1 \cdot 1, t\cdot t, 2(1
+\cdot t^2), 2(t \cdot t^2), (t^2 \cdot t^2)]$ contains components $t
+\cdot t$ and $2(1 \cdot t^2)$, which are linearly dependent.  Therefore,
+Spectral Experts would not be able to identify the parameters of
+a mixture of linear regressions for this distribution over $x$.
+
+We now show that some amount of unidentifiability is intrinsic to
+estimation from low-order moments, not just an artifact of our
+estimation procedure.  Suppose $x = (t, \dots, t^d)$.  Even if we
+observed all moments $\E[x\tp{p}y\tp{p}]$ and $\E[x\tp{2p}]$ for $p \in
+[r]$, all the resulting coordinates would be monomials up to degree
+$2dr$, and thus the moments live in a $2(dr)$-dimensional subspace.  On
+the other hand, the parameters $\theta$ live in a subspace of at least
+dimension $dk$.  Therefore, at least $r \ge k/2$ moments are required
+for identifiability of any algorithm for this monomial example.
+
+\paragraph{Low-rank regression}
+%\label{sec:regression}
+\vspace{-0.5em}
+We now bound the error of
+the compound parameter estimates,
+$\|\Delta_2\|_F^2$ and $\|\Delta_3\|_F^2$,
+where $\Delta_2 \eqdef \hat M_2 - M_2$
+and $\Delta_3 \eqdef \hat M_3 - M_3$.
+
+Our analysis is based on the low-rank regression framework of
+\citet{Tomioka2011} for tensors, which builds off of
+\citet{NegahbanWainwright2009} for matrices.
+The main challenge is to control the noise $\eta_p(x)$,
+which in addition to containing the usual Gaussian part, is also dependent on
+$x$, involves the mixing noise and various cross terms.
+We first set up some notation that unifies all three regressions (\refeqn{estimateM1}, \refeqn{estimateM2}, and \refeqn{estimateM3}).
+Define the observation operator $\opX_p(M_p) : \Re^{d\tp{p}} \to \Re^{n}$
+mapping compound parameters $M_p$:
+\begin{align}
+\opX_p(M_p; \sD)_i &\eqdef \innerp{M_p}{x\tp{p}_i}, & i \in [n].
+\end{align}
+
+Let $\kappa(\opX_p)$ be the restricted strong convexity constant,
+and let $\opX^*_p(\eta_p; \sD) = \sum_{(x,y) \in \sD} \eta_p(x) x\tp{p}$
+be the adjoint.
+
+%\paragraph{Restricted strong convexity}
+
+%Let us first lower bound the restricted strong convexity constant
+%$\kappa(\opX_p)$:
+
+%In the previous section, we described an algorithm for the mixture of
+%linear regressions using regression to recover $M_2$ and $M_3$,
+%described by \equationref{eqn:y2} and \equationref{eq:y3}, as
+%a subroutine. In this section, we will characterize the rate of
+%convergence of regression.
+
+%Analysis for regression in the fixed and random design settings have
+%been studied before \citep{HsuKakadeZhang}, however our setup differs
+%substantially from the noise models assumed in the literature. In our
+%scenario, the variance in the estimation comes not only from the
+%Gaussian observation noise (which has been studied before), but also
+%from the variance in the latent variable $h$.
+
+%Let us now formally define the class of regression problems we wish to
+%analyze, i.e. regression on the set $(x\tp{p}, y^p)$,
+%\begin{align*}
+%  y^p &= \innerp{x\tp{p}}{M_p} + (\innerp{x\tp{p}}{M_p - \beta_h\tp{p}} + \varepsilon).
+%\end{align*}
+
+%\todo{Describe/define the convex tensor stuff.}
+%We would like to exploit
+%the property that $M_p$ is low rank (as typically $K \ll D$). It has
+%been shown that a convex relaxation for this problem regression with
+%trace norm regularization, which can be solved using a proximal
+%subgradient descent algorithm\citationneeded.
+%The analogue of trace norm
+%regularization for higher order tensors corresponds to the sum of the
+%trace norms of the mode-k unfolding of the tensor \cite{Tomioka2011},
+%$X_{(k)}$, is a $d_k \times (\prod_{k' \neq k} d_{k'})$ matrix obtained
+%by concatenating the entries for all dimensions other than $k$. For
+%example, the 1-mode unfolding of a 3rd order tensor has entries,
+%$X_{(1)}^{i_1, (i_2, i_3)} = X_{i_1, i_2, i_3}$.
+
+%In general, the optimization problem we'd like to solve is,
+%\begin{align}
+%  \hat M_p &= 
+%  \arg\min_{M_p}& \frac{1}{2N} \| \vec y - \opX_p(M_p) \|_2^2 + \frac{\lambda_n}{K} \sum_{h=1}^K \| (M_p)_{(h)} \|_* \label{eq:regression}.
+%\end{align}
+
+\begin{lemma}[\citet{Tomioka2011}, Theorem 1]
+\label{lem:lowRank}
+Suppose there exists a restricted strong convexity constant $\kappa(\opX_p)$ such that
+$$\frac{1}{2n} \| \opX_p( \Delta )\|_2^2 \ge \kappa(\opX_p) \|\Delta\|^2_F \quad \text{and} \quad
+\lambda_n \ge \frac{\|\opX_p^*(\eta_p)\|_{op}}{n}.$$
+Then the error of $\hat M_p$ is bounded as follows:
+$\| \hat M_p - M_p^* \|_F \le \frac{\lambda_n \sqrt{k}}{\kappa(\opX_p)}$.
+\end{lemma}
+
+Going forward, we need to lower bound the restricted strong convexity
+parameter $\kappa(\opX_p)$ and upper bound the adjoint operator
+$\|\opX_p^*(\eta_p)\|_{op}$.
+%We
+%will appeal to the random design framework that models the input $x$ as
+%random and show bounds that hold with high probability.
+
+%\paragraph{Adjoint operator}
+
+%In this section, we upper bound the operator norm of the adjoint
+%$\|\opX_p(\eta_p)\|_\text{op}$.
+
+First, let us lower bound the restricted strong convexity parameter $\kappa(\opX_p)$:
+
+\begin{lemma}[lower bound on restricted strong convexity]
+\label{lem:lowRankLower}
+Let $\Sigma_p \eqdef \E[\cvec(x\tp{p})\tp{2}]$.
+If $$n = \Omega \left( p ~ d^{2p} \log(d) \frac{R^{4p} \log(1/\delta)}{\sigmamin(\Sigma_p)^2} \right),$$
+with probability at least $1-\delta$,
+$$\kappa(\opX_p) \ge \frac{\sigmamin(\Sigma_p)}{4}.$$
+\end{lemma}
+\begin{proof}
+Expanding the definition of the observation operator:
+$\frac{1}{2n} \|\opX_p(\Delta)\|_2^2
+= \frac{1}{2n} \sum_{(x,y) \in \sD} \innerp{\Delta}{x\tp{p}}^2$.
+Unfolding the tensor and letting $\hat\Sigma_p \eqdef \frac{1}{n} \sum_{(x,y) \in \sD} \cvec(x\tp{p})\tp{2}$,
+we have 
+$\frac{1}{2n} \|\opX_p(\Delta)\|_2^2
+= \frac{1}{2} \trace(\cvec(\Delta)\tp{2} \hat\Sigma_p)
+\ge \frac{1}{2} \|\Delta\|_F^2 \sigmamin(\hat\Sigma_p)$.
+It remains to show that $\sigmamin(\hat\Sigma_p)$ and $\sigmamin(\Sigma_p)$ are close.
+To do this, we first apply Hoeffding's inequality elementwise.
+Since $\|x\|_2 \le R$, we have that for each element $(i,j)$,
+$|(\hat\Sigma_p)_{ij} - (\Sigma_p)_{ij}| = O(R^{2p}\sqrt{\frac{\log (1/\delta)}{n}})$.
+Applying the union bound over the $d^{2p}$ elements of $\hat\Sigma_p - \Sigma_p$,
+we have that the max norm is bounded:
+$\|\hat\Sigma_p - \Sigma_p\|_\text{\rm max} = O(R^{2p} \sqrt{\frac{p \log(d) \log (1/\delta)}{n}})$.
+The max norm times $d^p$ upper bounds the Frobenius norm, which upper bounds the operator norm, so we have that
+$\|\hat\Sigma_p - \Sigma_p\|_\text{\rm op} = O(d^p R^{2p} \sqrt{\frac{p \log(d) \log (1/\delta)}{n}})$.
+Using the fact that $\sigmamin(\hat\Sigma_p) \ge \sigmamin(\Sigma_p) - \|\hat\Sigma_p - \Sigma_p\|_\text{\rm op}$
+yields the result.
+\end{proof}
+
+Next, let us upper bound the operator norm of the observation adjoint:
+
+\begin{lemma}[upper bound on adjoint noise]
+\label{lem:lowRankUpper}
+Let $\opX_p$ be the linear operator previously defined. Then with
+probability at least $1-\delta$,
+$$\frac1{n} \|\opX_p^*(\eta_p)\|_\op \le O\left(\sigma^3 L^3 R^6 \sqrt{\frac{\log^3(1/\delta)}{n}}\right).$$
+for each $p \in \{1,2,3\}$.
+\end{lemma}
+\begin{proof}
+Let $\hat\E_p[f(x,\epsilon,h)]$ denote the empirical expectation over
+the examples in dataset $\sD_p$ (recall the $\sD_p$'s are independent to
+simplify the analysis).  By definition,
+$\frac1n \|\opX_p^*(\eta_p)\|_\op = \hat\E_p[\eta_p(x) x\tp{p}]$
+for $p \in \{1,2,3\}$ defined in \refeqn{y1}, \refeqn{y2}, and
+\refeqn{y3}. Each $\eta_p(x)$ is composed of several zero-mean random
+variables, and since $\|A + B\|_\op \le \|A\|_\op + \|B\|_\op$, it
+suffices to consider each term in turn. For each term, we will
+condition on $x$ and use a standard concentration inequality to bound
+the noise.
+
+%\begin{align}
+%  \| \hat\E_1[ \eta_1(x) x ] \|_{op} &\le 
+%  \underbrace{ \|\hat\E_1[\innerp{M_1 - \beta_h}{x} x]\|_\op }_{ O(R^2 L \sqrt{\frac{\log(1/\delta_1)}{n}}) } +
+%  \underbrace{ \|\hat\E_1[\epsilon x]\|_\op }_{ O(R^2 L \sqrt{\frac{\log(1/\delta_1)}{n}}) } +
+%\end{align}
+
+First, consider $\eta_1(x)$, which contains two terms.
+The first term is $\innerp{M_1 - \beta_h}{x}$.
+We have that $\|M_1 - \beta_h\|_2 \le 2L$,
+so by Hoeffding's inequality, and using the fact that $\|x\|_2 \le R$,
+with probability $1-\delta_1$,
+$\|\hat\E_1[\innerp{M_1 - \beta_h}{x} x]\|_\op = O(R^2 L \sqrt{\frac{\log(1/\delta_1)}{n}})$.
+The second term is $\epsilon \sim \normal{0}{\sigma^2}$,
+with probability $1-\delta_1$,
+$\|\hat\E_1[\epsilon x]\|_\op = O(R \sigma \sqrt{\frac{\log(1/\delta_1)}{n}})$.
+
+Now, consider $\eta_2(x)$, which contains three terms.
+Using similar techniques, we have that
+$\|\hat\E_2[\eta_2(x) x\tp{2}]\|_\op = O(R^2 (L^2 R^2 + \sigma L R + \sigma^2) \sqrt{\frac{\log(1/\delta_1)}{n}})$.
+
+Finally, $\eta_3(x)$ contains five terms plus an additional bias since the estimator uses $\hat M_1$ rather than $M_1$.
+There are two additional considerations.
+First, to bound $\hat\E_3[\epsilon^3]$, we use Lemma 19 of \cite{hsu13spherical}.
+Second, the additional bias is small by the previous bound (and it is constructed on $\sD_1$, which is independent of $\sD_3$
+used to construct $\hat M_3$.
+Therefore, we incur another $O(\sigma^2 \cdot R \sigma \sqrt{\frac{\log(1/\delta_1)}{n}})$.
+We can handle the other terms using conventional means.
+Combining everything, we obtain
+$\|\hat\E_3[\eta_3(x) x\tp{3}]\|_\op = O(R^3 (L^3 R^3 + \sigma L^2 R^2 + \sigma^2 L R + \sigma^3 + \sigma^3 R) \sqrt{\frac{\log^3(1/\delta_1)}{n}})$.
+
+Finally, taking $\delta_1 = \delta/3$, and taking the union bound over the bounds for $p \in \{1,2,3\}$,
+we get our result.
+Note that we are bounding quantities fairly crudely for the sake of simplicity.
+\end{proof}
+
+\paragraph{Method of Moments}
+
+Now that we have bounds on the errors of recovering $M_2$ and $M_3$ from
+the data, we will extend the analysis of the robust tensor method in
+\citet{AnandkumarGeHsu2012} to bound the error on the parameters
+$\theta$. 
+\begin{lemma}
+  \label{lem:tensorPower} Let $\|\hat M_2 - M_2\|_{op} \le \epsilon$ and
+  $\|\hat M_3 - M_3\|_{op} \le \epsilon$, for some $\epsilon$ such that 
+  $$\epsilon < \min\{\sigmamin(M_2)/2,O(k \sigmamin(M_2)^{1.5} \|M_3\|_{op}^{-1})\}.$$ 
+  Then, for some permutation of indices, the parameter estimates found
+  in part 2 of \algorithmref{algo:spectral-experts} satisfy the
+  following with probability atleast $1 - \delta$,
+  \begin{align*}
+  \|\hat \pi - \pi \|_{\infty}
+    &= O( k \pi_{max}^{2.5} \pi_{min}^{-0.5} \sigmamin(M_2)^{-1.5} \|M_3\|_{op} \epsilon ) \\
+  \|\hat \beta_h - \beta_h\|_2
+    &= O( k \pi_{max} \sigmamin(M_2)^{-1.5} \|M_2\|_{op}^{0.5} \|M_3\|_{op} \epsilon ).
+  \end{align*}
+  for all $h \in [k]$.
+\end{lemma}
+The proof of the above lemma has been deferred to
+\appendixref{sec:proofs}.
+
+\paragraph{Synthesis}
+Together, these lemmas allow us to control the compound parameter error
+and the recovery error. We now apply them in the proof of
+\refthm{convergence}:
+
+\begin{proof}
+By \reflem{lowRank} along with \reflem{lowRankLower} and
+\reflem{lowRankUpper}, we can control the operator norm of the parameter
+estimates which are directly upper bounded by the Frobenius norm: If $n
+\ge n_1$, then
+\begin{align}
+  \|\hat M_p - M_p\|_{op} = O\left( \sigma^3 L^3 R^6 k^{\frac12} \sigmamin(\Sigma_p)^{-1} \sqrt{\frac{\log^3 (1/\delta)}{n}} \right).
+  %\|\hat M_p - M_p\|_F = O\left( \frac{\sigma^3 L^3 R^6 \sqrt{k} \log^3 (1/\delta)}{\sqrt{n} (\sigmamin(\Sigma_p) - d^p R^p \sqrt{\frac{p \log(d) \log(1/\delta)}{n}}} \right).
+\end{align}
+
+We complete the proof by applying \reflem{tensorPower} with the above
+bound on $\|\hat M_p - M_p\|_{op}$.
+
+\end{proof}

### 2013-04-23 doc/icml2013/theory.tex
Cleaned proof of Lemma 3, and fixed the figure placements

@@
 suffices to consider each term in turn. For each term, we will
-condition on $x$ and use a standard concentration inequality to bound
-the noise.
+condition on $x$ and use standard concentration inequalities to bound
+the noise with probability atleast $1-\delta_1$,
+{\scriptsize
+\begin{align*}
+  \| \hat\E_1[ \eta_1(x) x ] \|_{op} &\le 
+            \underbrace{ \|\hat\E_1[\innerp{M_1 - \beta_h}{x} x]\|_\op  }_{ O\left( L R^2     \sqrt{\frac{\log(1/\delta_1)}{n}} \right) } +
+            \underbrace{ \|\hat\E_1[\epsilon x]\|_\op                   }_{ O\left( R \sigma  \sqrt{\frac{\log(1/\delta_1)}{n}} \right) } \\
+  %  &= O\left( \sigma L R^2 \sqrt{\frac{\log(1/\delta_1)}{n}} \right) \\
+    \| \hat\E_2[ \eta_2(x) x\tp{2} ] \|_{op} &\le 
+            \underbrace{ \|\hat\E_2[\innerp{\beta_h\tp{2} - M_2}{x\tp{2}} x\tp{2}]\|_\op }_{ O\left( L^2 R^4      \sqrt{\frac{\log^2(1/\delta_1)}{n}} \right) } \\
+   &\quad + \underbrace{ \|\hat\E_2[2 \epsilon \innerp{\beta_h}{x} x\tp{2}]\|_\op        }_{ O\left( \sigma L R^3 \sqrt{\frac{\log  (1/\delta_1)}{n}} \right) } 
+          + \underbrace{ \|\hat\E_2[(\epsilon^2 - \sigma^2) x\tp{2}]\|_\op               }_{ O\left( \sigma^2 R^2 \sqrt{\frac{\log^2(1/\delta_1)}{n}} \right) } \\
+  %&= O\left( \sigma^2 L^2 R^4 \sqrt{\frac{\log^2(1/\delta_1)}{n}} \right) \\
+  \| \hat\E_3[ \eta_3(x) x\tp{3} ] \|_{op} &\le 
+            \underbrace{ \|\hat\E_3[\innerp{\beta_h\tp{3} - M_3}{x\tp{3}}                      x\tp{3}]\|_\op }_{O\left( L^3 R^6        \sqrt{\frac{\log  (1/\delta_1)}{n}} \right) } \\
+   &\quad + \underbrace{ \|\hat\E_3[3 \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}}                 x\tp{3}]\|_\op }_{O\left( \sigma L^2 R^5 \sqrt{\frac{\log^2(1/\delta_1)}{n}} \right) } 
+          + \underbrace{ \|\hat\E_3[\epsilon^3                                  x\tp{3}]\|_\op }_{O\left( \sigma^3 R^3   \sqrt{\frac{\log  (1/\delta_1)}{n}} \right) } \\
+   &\quad + \underbrace{ \|\hat\E_3[3(\epsilon^2 \innerp{\beta_h}{x} -\sigma^2 \innerp{\hat M_1}{x}) x\tp{3}]\|_\op }_{O\left( (\sigma^3 L R^4 + \sigma^3 R^4) \sqrt{\frac{\log^3(1/\delta_1)}{n}} \right) } \\
+  %&= O\left( \sigma^3 L^3 R^6 \sqrt{\frac{\log^3(1/\delta_1)}{n}} \right).
+\end{align*}
+}
+There are two additional considerations when bounding the terms in $\eta_3$.
+First, to bound $\hat\E_3[\epsilon^3]$, we use Lemma 19 of \cite{hsu13spherical}.
+Second, there is an additional bias since the estimator uses $\hat M_1$ instead of $M_1$; this bias is small by the previous bound (and is constructed on an independent dataset, $\sD_1$).
+Therefore, we incur another $O(\sigma^2 \cdot R \sigma \sqrt{\frac{\log(1/\delta_1)}{n}})$.
+We can handle the other terms using conventional means, the details of which can be found in \todo{Appendix B}.
 
-%\begin{align}
-%  \| \hat\E_1[ \eta_1(x) x ] \|_{op} &\le 
-%  \underbrace{ \|\hat\E_1[\innerp{M_1 - \beta_h}{x} x]\|_\op }_{ O(R^2 L \sqrt{\frac{\log(1/\delta_1)}{n}}) } +
-%  \underbrace{ \|\hat\E_1[\epsilon x]\|_\op }_{ O(R^2 L \sqrt{\frac{\log(1/\delta_1)}{n}}) } +
-%\end{align}
+%First, consider $\eta_1(x)$, which contains two terms.
+%The first term is $\innerp{M_1 - \beta_h}{x}$.
+%We have that $\|M_1 - \beta_h\|_2 \le 2L$,
+%so by Hoeffding's inequality, and using the fact that $\|x\|_2 \le R$,
+%with probability $1-\delta_1$,
+%$\|\hat\E_1[\innerp{M_1 - \beta_h}{x} x]\|_\op = O(R^2 L \sqrt{\frac{\log(1/\delta_1)}{n}})$.
+%The second term is $\epsilon \sim \normal{0}{\sigma^2}$,
+%with probability $1-\delta_1$,
+%$\|\hat\E_1[\epsilon x]\|_\op = O(R \sigma \sqrt{\frac{\log(1/\delta_1)}{n}})$.
+%
+%Now, consider $\eta_2(x)$, which contains three terms.
+%Using similar techniques, we have that
+%$\|\hat\E_2[\eta_2(x) x\tp{2}]\|_\op = O(R^2 (L^2 R^2 + \sigma L R + \sigma^2) \sqrt{\frac{\log(1/\delta_1)}{n}})$.
 
-First, consider $\eta_1(x)$, which contains two terms.
-The first term is $\innerp{M_1 - \beta_h}{x}$.
-We have that $\|M_1 - \beta_h\|_2 \le 2L$,
-so by Hoeffding's inequality, and using the fact that $\|x\|_2 \le R$,
-with probability $1-\delta_1$,
-$\|\hat\E_1[\innerp{M_1 - \beta_h}{x} x]\|_\op = O(R^2 L \sqrt{\frac{\log(1/\delta_1)}{n}})$.
-The second term is $\epsilon \sim \normal{0}{\sigma^2}$,
-with probability $1-\delta_1$,
-$\|\hat\E_1[\epsilon x]\|_\op = O(R \sigma \sqrt{\frac{\log(1/\delta_1)}{n}})$.
-
-Now, consider $\eta_2(x)$, which contains three terms.
-Using similar techniques, we have that
-$\|\hat\E_2[\eta_2(x) x\tp{2}]\|_\op = O(R^2 (L^2 R^2 + \sigma L R + \sigma^2) \sqrt{\frac{\log(1/\delta_1)}{n}})$.
-
-Finally, $\eta_3(x)$ contains five terms plus an additional bias since the estimator uses $\hat M_1$ rather than $M_1$.
-There are two additional considerations.
-First, to bound $\hat\E_3[\epsilon^3]$, we use Lemma 19 of \cite{hsu13spherical}.
-Second, the additional bias is small by the previous bound (and it is constructed on $\sD_1$, which is independent of $\sD_3$
-used to construct $\hat M_3$.
-Therefore, we incur another $O(\sigma^2 \cdot R \sigma \sqrt{\frac{\log(1/\delta_1)}{n}})$.
-We can handle the other terms using conventional means.
-Combining everything, we obtain
-$\|\hat\E_3[\eta_3(x) x\tp{3}]\|_\op = O(R^3 (L^3 R^3 + \sigma L^2 R^2 + \sigma^2 L R + \sigma^3 + \sigma^3 R) \sqrt{\frac{\log^3(1/\delta_1)}{n}})$.
+% Finally, $\eta_3(x)$ contains five terms plus an additional bias since the estimator uses $\hat M_1$ rather than $M_1$.
+% There are two additional considerations.
+% First, to bound $\hat\E_3[\epsilon^3]$, we use Lemma 19 of \cite{hsu13spherical}.
+% Second, the additional bias is small by the previous bound (and it is constructed on $\sD_1$, which is independent of $\sD_3$
+% used to construct $\hat M_3$.
+% Therefore, we incur another $O(\sigma^2 \cdot R \sigma \sqrt{\frac{\log(1/\delta_1)}{n}})$.
+% We can handle the other terms using conventional means.
+% Combining everything, we obtain
+% $\|\hat\E_3[\eta_3(x) x\tp{3}]\|_\op = O(R^3 (L^3 R^3 + \sigma L^2 R^2 + \sigma^2 L R + \sigma^3 + \sigma^3 R) \sqrt{\frac{\log^3(1/\delta_1)}{n}})$.
 
@@
 the data, we will extend the analysis of the robust tensor method in
-\citet{AnandkumarGeHsu2012} to bound the error on the parameters
-$\theta$. 
+\citet{AnandkumarGeHsu2012} to include the whitening and unwhitening
+steps, 2(a) and 2(c) respectively, of
+\algorithmref{algo:spectral-experts}.
 \begin{lemma}
@@
   $$\epsilon < \min\{\sigmamin(M_2)/2,O(k \sigmamin(M_2)^{1.5} \|M_3\|_{op}^{-1})\}.$$ 
-  Then, for some permutation of indices, the parameter estimates found
-  in part 2 of \algorithmref{algo:spectral-experts} satisfy the
-  following with probability atleast $1 - \delta$,
+  Then, there exists a permutation of indices such that  the parameter
+  estimates found in step 2 of \algorithmref{algo:spectral-experts}
+  satisfy the following with probability atleast $1 - \delta$,
   \begin{align*}
@@
 \end{lemma}
-The proof of the above lemma has been deferred to
+\begin{proof}
+  The proof follows by applying standard matrix perturbation results for
+  the whitening and unwhitening operators, and has been deferred
   \appendixref{sec:proofs}. 
+\end{proof}
 

### 2013-04-24 doc/icml2013/theory.tex
Improved bounds for lemma 2 and some proof adjustment

@@
 $n = \max(n_1,n_2)$
-where $n_1 = \Omega\left(d^6 R^{12} (\max_{p \in [3]} \sigmamin(\Sigma_p)^{-2}) \log(d) \log(1/\delta)\right)$ and
+where $n_1 = \Omega\left(R^{12} (\max_{p \in [3]} \sigmamin(\Sigma_p)^{-2}) \log(1/\delta)\right)$ and
 $n_2 = \Omega\left( \sigma^6 L^6 R^{12} k (\max_{p \in [3]} \sigmamin(\Sigma_p)^{-1}) \frac{\log^3 (1/\delta)}{\epsilon^2} \right)$.
@@
 Let $\Sigma_p \eqdef \E[\cvec(x\tp{p})\tp{2}]$.
-If $$n = \Omega \left( p ~ d^{2p} \log(d) \frac{R^{4p} \log(1/\delta)}{\sigmamin(\Sigma_p)^2} \right),$$
+If $$n = \Omega \left( \frac{R^{4p} \log(1/\delta)}{\sigmamin(\Sigma_p)^2} \right),$$
 with probability at least $1-\delta$,
@@
 \ge \frac{1}{2} \|\Delta\|_F^2 \sigmamin(\hat\Sigma_p)$.
-It remains to show that $\sigmamin(\hat\Sigma_p)$ and $\sigmamin(\Sigma_p)$ are close.
-To do this, we first apply Hoeffding's inequality elementwise.
-Since $\|x\|_2 \le R$, we have that for each element $(i,j)$,
-$|(\hat\Sigma_p)_{ij} - (\Sigma_p)_{ij}| = O(R^{2p}\sqrt{\frac{\log (1/\delta)}{n}})$.
-Applying the union bound over the $d^{2p}$ elements of $\hat\Sigma_p - \Sigma_p$,
-we have that the max norm is bounded:
-$\|\hat\Sigma_p - \Sigma_p\|_\text{\rm max} = O(R^{2p} \sqrt{\frac{p \log(d) \log (1/\delta)}{n}})$.
-The max norm times $d^p$ upper bounds the Frobenius norm, which upper bounds the operator norm, so we have that
-$\|\hat\Sigma_p - \Sigma_p\|_\text{\rm op} = O(d^p R^{2p} \sqrt{\frac{p \log(d) \log (1/\delta)}{n}})$.
-Using the fact that $\sigmamin(\hat\Sigma_p) \ge \sigmamin(\Sigma_p) - \|\hat\Sigma_p - \Sigma_p\|_\text{\rm op}$
-yields the result.
+By Weyl's theorem, $\sigmamin(\hat\Sigma_p) \ge \sigmamin(\Sigma_p) - \|\hat\Sigma_p - \Sigma_p\|_{op}$. 
+Thus, it remains to show that $\|\hat\Sigma_p - \Sigma_p\|_{op} \le \|\hat\Sigma_p - \Sigma_p\|_{F}$ is small. 
+To do so, we apply a matrix concentration result for the Frobenius norm\citationneeded; with probability at least $1 - \delta$, 
+$\| \hat\Sigma_p - \Sigma_p \|_F = O\left( M \sqrt{\frac{\log(1/\delta)}{n}} \right)$,
+where $M$ is an upper bound on $\| \Sigma_p \|_F$. Since $\|x\|_2 \le R$, we can use the bound $\| \Sigma_p \|_F \le R^{2p}$ in the above to get the stated result.
+
+%To do this, we first apply Hoeffding's inequality elementwise.
+%Since $\|x\|_2 \le R$, we have that for each element $(i,j)$,
+%$|(\hat\Sigma_p)_{ij} - (\Sigma_p)_{ij}| = O(R^{2p}\sqrt{\frac{\log (1/\delta)}{n}})$.
+%Applying the union bound over the $d^{2p}$ elements of $\hat\Sigma_p - \Sigma_p$,
+%we have that the max norm is bounded:
+%$\|\hat\Sigma_p - \Sigma_p\|_\text{\rm max} = O(R^{2p} \sqrt{\frac{p \log(d) \log (1/\delta)}{n}})$.
+%The max norm times $d^p$ upper bounds the Frobenius norm, which upper bounds the operator norm, so we have that
+%$\|\hat\Sigma_p - \Sigma_p\|_\text{\rm op} = O(d^p R^{2p} \sqrt{\frac{p \log(d) \log (1/\delta)}{n}})$.
+%Using the fact that $\sigmamin(\hat\Sigma_p) \ge \sigmamin(\Sigma_p) - \|\hat\Sigma_p - \Sigma_p\|_\text{\rm op}$
+%yields the result.
 \end{proof}
@@
 variables, and since $\|A + B\|_\op \le \|A\|_\op + \|B\|_\op$, it
-suffices to consider each term in turn. For each term, we will
+suffices to consider each term in turn. 
+For each term, we will
 condition on $x$ and use standard concentration inequalities to bound
-the noise with probability atleast $1-\delta_1$,
+the noise.
+
+For example, consider $\eta_1(x)$, which has two terms. 
+The first term is $\innerp{M_1 - \beta_h}{x}$.
+We have that $\|M_1 - \beta_h\|_2 \le 2L$, so by Hoeffding's inequality, and using the fact that $\|x\|_2 \le R$, with probability $1-\delta_1$, $\|\hat\E_1[\innerp{M_1 - \beta_h}{x} x]\|_2 = O(R^2 L \sqrt{\frac{\log(1/\delta_1)}{n}})$.
+The second term is $\epsilon \sim \normal{0}{\sigma^2}$, thus with probability $1-\delta_1$, $\|\hat\E_1[\epsilon x]\|_2 = O(R \sigma \sqrt{\frac{\log(1/\delta_1)}{n}})$. 
+The bounds on the terms presented below hold with probability atleast $1-\delta_1$ and are proved similarly.
 {\scriptsize
 \begin{align*}
-  \| \hat\E_1[ \eta_1(x) x ] \|_{op} &\le 
-            \underbrace{ \|\hat\E_1[\innerp{M_1 - \beta_h}{x} x]\|_\op  }_{ O\left( L R^2     \sqrt{\frac{\log(1/\delta_1)}{n}} \right) } +
-            \underbrace{ \|\hat\E_1[\epsilon x]\|_\op                   }_{ O\left( R \sigma  \sqrt{\frac{\log(1/\delta_1)}{n}} \right) } \\
+%   \| \hat\E_1[ \eta_1(x) x ] \|_{op} &\le 
+%             \underbrace{ \|\hat\E_1[\innerp{M_1 - \beta_h}{x} x]\|_\op  }_{ O\left( L R^2     \sqrt{\frac{\log(1/\delta_1)}{n}} \right) } +
+%             \underbrace{ \|\hat\E_1[\epsilon x]\|_\op                   }_{ O\left( R \sigma  \sqrt{\frac{\log(1/\delta_1)}{n}} \right) } \\
   %  &= O\left( \sigma L R^2 \sqrt{\frac{\log(1/\delta_1)}{n}} \right) \\
@@
 Therefore, we incur another $O(\sigma^2 \cdot R \sigma \sqrt{\frac{\log(1/\delta_1)}{n}})$.
-We can handle the other terms using conventional means, the details of which can be found in \todo{Appendix B}.
-
-%First, consider $\eta_1(x)$, which contains two terms.
-%The first term is $\innerp{M_1 - \beta_h}{x}$.
-%We have that $\|M_1 - \beta_h\|_2 \le 2L$,
-%so by Hoeffding's inequality, and using the fact that $\|x\|_2 \le R$,
-%with probability $1-\delta_1$,
-%$\|\hat\E_1[\innerp{M_1 - \beta_h}{x} x]\|_\op = O(R^2 L \sqrt{\frac{\log(1/\delta_1)}{n}})$.
-%The second term is $\epsilon \sim \normal{0}{\sigma^2}$,
-%with probability $1-\delta_1$,
-%$\|\hat\E_1[\epsilon x]\|_\op = O(R \sigma \sqrt{\frac{\log(1/\delta_1)}{n}})$.
-%
-%Now, consider $\eta_2(x)$, which contains three terms.
-%Using similar techniques, we have that
-%$\|\hat\E_2[\eta_2(x) x\tp{2}]\|_\op = O(R^2 (L^2 R^2 + \sigma L R + \sigma^2) \sqrt{\frac{\log(1/\delta_1)}{n}})$.
-
-% Finally, $\eta_3(x)$ contains five terms plus an additional bias since the estimator uses $\hat M_1$ rather than $M_1$.
-% There are two additional considerations.
-% First, to bound $\hat\E_3[\epsilon^3]$, we use Lemma 19 of \cite{hsu13spherical}.
-% Second, the additional bias is small by the previous bound (and it is constructed on $\sD_1$, which is independent of $\sD_3$
-% used to construct $\hat M_3$.
-% Therefore, we incur another $O(\sigma^2 \cdot R \sigma \sqrt{\frac{\log(1/\delta_1)}{n}})$.
-% We can handle the other terms using conventional means.
-% Combining everything, we obtain
-% $\|\hat\E_3[\eta_3(x) x\tp{3}]\|_\op = O(R^3 (L^3 R^3 + \sigma L^2 R^2 + \sigma^2 L R + \sigma^3 + \sigma^3 R) \sqrt{\frac{\log^3(1/\delta_1)}{n}})$.
+%We can handle the other terms using conventional means, the details of which can be found in \appendixref{sec:proofs}.
 

### 2013-04-24 doc/icml2013/evaluation.tex
Cleaned proof of Lemma 3, and fixed the figure placements
Added comments from Percy

@@
 
-We implemented the low rank recovery optimization routines for $M_2$ and
-$M_3$ using a proximal subgradient descent
-algorithm.%\cite{candes11phaselift,tomioka2010estimation}. (Note sure about this citation and not required?)
-The algorithm
-was initialized with the regularized least squares solution and was run
-for 500 iterations or until convergence up to a tolerance of $10^{-3}$.
+We solved the low rank recovery optimization routines for $M_2$ and
+$M_3$ using \texttt{cvx}\citationneeded. 
+%ainterior point method an interior proximal subgradient descent
+%algorithm.%\cite{candes11phaselift,tomioka2010estimation}. (Note sure about this citation and not required?)
+%The algorithm was initialized with the regularized least squares
+%solution and was run for 500 iterations or until convergence up to
+%a tolerance of $10^{-3}$.
 %Both regularization parameters, $\lambda_n^{(2)}$ and, $\lambda_n^{(3)}$ were annealed as $\frac{1}{n}$. 
-
 As a baseline, we also implemented EM with two different
@@
 this, EM converged to the correct basin of attraction only 13\% of the time.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=0.50\textwidth]{figures/hist.pdf}
+  \caption{Histogram over recovery errors between the three algorithms. $b = 1, d = 4, k = 3, n = 500,000$}
+  \label{fig:hist}
+\end{figure}
 
@@
 
-\subsection{Results}
-
-\tableref{tbl:parameter-recovery} presents the Frobenius norm of the
-difference between true and estimated parameters for the model, averaged
-over 20 different random instances for each feature set and 10 attempts
-for each instance. The experiments were run using $n = 500,000$ samples.
-
-One of the main reasons for the high variance is the variation across
-random instances; some are easy for EM to find the global minima and
-others more difficult. In general, while the spectral algorithm did not
-recover parameters extremely well, it provided a good initialization for
-EM.
-
-\begin{figure}[tbhp]
-  \centering
-  \includegraphics[width=0.50\textwidth]{figures/hist.pdf}
-  \caption{Histogram over recovery errors between the three algorithms. $b = 1, d = 4, k = 3, n = 500,000$}
-  \label{fig:hist}
-\end{figure}
-
-To study the stability of the solutions returned by the spectral method,
-consider the histogram in \figureref{fig:hist}, which shows the recovery
-errors of the algorithms over 170 attempts on a dataset with $b = 1, d = 4,
-k = 3$. Typically, the spectral algorithm returned a stable solution.
-When these parameters were close enough to the true parameters, we found
-that EM almost always converged to the global optima. Randomly
-initialized EM only finds the true parameters a little over 10\% of the
-time and shows considerably higher variance.
-
-\paragraph{Effect of number of data points}
-
 \begin{figure*}[tbhp]
@@
 
-In \figureref{fig:vs-n}, we show how the recovery error varies as we get
-more data. Each data point shows the mean error over 10 attempts, with
-error bars. We note that the recovery performance of EM does not
-particularly improve; this suggests that EM continues to get stuck in
-a local optima. The spectral algorithm's error decays slowly, and as it
-gets closer to zero, EM initialized at the spectral parameters finds the
-true parameters more often as well. This behavior highlights the
-trade-off between statistical and computational error.
-
-\paragraph{Misspecified data}
+\subsection{Results}
 
@@
 
+\tableref{tbl:parameter-recovery} presents the Frobenius norm of the
+difference between true and estimated parameters for the model, averaged
+over 20 different random instances for each feature set and 10 attempts
+for each instance. The experiments were run using $n = 500,000$ samples.
+
+One of the main reasons for the high variance is the variation across
+random instances; some are easy for EM to find the global minima and
+others more difficult. In general, while the spectral algorithm did not
+recover parameters extremely well, it provided a good initialization for
+EM.
+
+To study the stability of the solutions returned by the spectral method,
+consider the histogram in \figureref{fig:hist}, which shows the recovery
+errors of the algorithms over 170 attempts on a dataset with $b = 1, d = 4,
+k = 3$. Typically, the spectral algorithm returned a stable solution.
+When these parameters were close enough to the true parameters, we found
+that EM almost always converged to the global optima. Randomly
+initialized EM only finds the true parameters a little over 10\% of the
+time and shows considerably higher variance.
+
+\paragraph{Effect of number of data points}
+
+In \figureref{fig:vs-n}, we show how the recovery error varies as we get
+more data. Each data point shows the mean error over 10 attempts, with
+error bars. We note that the recovery performance of EM does not
+particularly improve; this suggests that EM continues to get stuck in
+a local optima. The spectral algorithm's error decays slowly, and as it
+gets closer to zero, EM initialized at the spectral parameters finds the
+true parameters more often as well. This behavior highlights the
+trade-off between statistical and computational error.
+
+\paragraph{Misspecified data}
+
 To evaluate how robust the algorithm was to model mis-specification, we

### 2013-04-24 doc/icml2013/intro.tex
Added comments from Percy

@@
 direction by proposing a new algorithm for a simple model,
-\emph{mixture of linear regressions} (MLG) \cite{VieleTong2002}.
+\emph{a mixture of linear regressions} (MLG) \cite{VieleTong2002}.
 
@@
 % Twist
-In mixture of linear regressions, using third-order moments does not
+In a mixture of linear regressions, using third-order moments does not
 directly reveal the tensor structure of the problem, so we cannot
@@
 %\Re^d$ to be the covariates it depends on. Note that the matrix of
-%regressors $B$ is a $d \times K$ matrix, where the $k$-th column is
+%regressors $B$ is a $d \times k$ matrix, where the $k$-th column is
 %$\beta_k$.
@@
 a $p$-th order tensor $X \in (\Re^d)\tp{p}$, the mode-$i$ unfolding of
-$X$ is a matrix whose $j$-th row contains all the elements of $X$ whose
-$i$-th index is equal to $j$. Let $X_{(i)} \in \Re^{d \times d^{p-1}}$
-be the mode-$i$ unfolding of $X$.
+$X$ is a matrix, $X_{(i)} \in \Re^{d \times d^{p-1}}$, whose $j$-th row contains all the elements of $X$ whose
+$i$-th index is equal to $j$. 
 %For $X \in (\Re^d)\tp{p}$ and $Y \in (\Re^d)\tp{(p+q)}$,% $p, q > 0$,
@@
 let $\|X\|_{\max}$ denote the max norm (elementwise maximum),
-let $\|X\|_\text{op}$ denote the operator norm (largest singular value).
+let $\|X\|_\Lop$ denote the operator norm (largest singular value),
+let $\sigma_\text{min}(X)$ be the smallest singular value of $X$.
 For a tensor $X$,
@@
 the average nuclear norm over all $p$ unfoldings,
-and let $\|X\|_\text{op} = \frac{1}{p} \sum_{i=1}^p \|X_{(i)}\|_\text{op}$
+and let $\|X\|_\Lop = \frac{1}{p} \sum_{i=1}^p \|X_{(i)}\|_\Lop$
 denote the average operator norm over all $p$ unfoldings.
-Let $\sigma_\text{min}(X)$ be the smallest singular value of $X$.
 
 For a symmetric tensor $X \in (\Re^d)\tp{p}$, let $\cvec(X) \in
-\Re^{\binom{d + p + 1}{p}}$ be the collapsed vectorization of distinct
-elements in $X$: each component of $\cvec(X)$ is indexed by a vector of
+\Re^{D_p}, D_p = \binom{d + p + 1}{p}$ be the collapsed vectorization of distinct
+elements in $X$, for example, for $X \in \Re^{d \times d}$, $\cvec(X)
+= (X_{ii} : i \in [d]; X_{ij} + X_{ji} : i,j \in [d], i<j)$.
+In general, each component of $\cvec(X)$ is indexed by a vector of
 counts $(c_1, \dots, c_d)$ with total sum $\sum_i c_i = p$.  The value
@@
 = \{ k \in [d]^p : \forall i \in [d], c_i = |\{ j \in [p] : k_j = i \}|
-\}$ are the set of index vectors $k$ with that count profile.  For
-example, for $X \in \Re^{d \times d}$, $\cvec(X) = (X_{ii} : i \in [d];
-X_{ij} + X_{ji} : i,j \in [d], i<j)$.
+\}$ are the set of index vectors $k$ with that count profile.  
 %For symmetric tensors, $\innerp{X}{Y} = \innerp{\cvec(X)}{\cvec(Y)}$.

### 2013-04-24 doc/icml2013/theory.tex
Added comments from Percy

--- 2013-04-23 doc/icml2013/theory.tex
@@
 using results from \citet{Tomioka2011}.
-Then, we use results from \cite{AnandkumarGeHsu2012} to convert this error
+Then, we use results from \citet{AnandkumarGeHsu2012} to convert this error
 into a bound on the actual parameter estimates $\hat\theta = (\hat\pi, \hat B)$
@@
 [r]$, all the resulting coordinates would be monomials up to degree
-$2dr$, and thus the moments live in a $2(dr)$-dimensional subspace.  On
+$2dr$, and thus the moments live in a $2dr$-dimensional subspace.  On
 the other hand, the parameters $\theta$ live in a subspace of at least
@@
 \begin{align}
-\opX_p(M_p; \sD)_i &\eqdef \innerp{M_p}{x\tp{p}_i}, & i \in [n].
+\opX_p(M_p; \sD)_i &\eqdef \innerp{M_p}{x\tp{p}_i}, & (x_i, y_i) \in \sD.
 \end{align}
@@
 %We would like to exploit
-%the property that $M_p$ is low rank (as typically $K \ll D$). It has
+%the property that $M_p$ is low rank (as typically $k \ll D$). It has
 %been shown that a convex relaxation for this problem regression with
@@
 %  \hat M_p &= 
-%  \arg\min_{M_p}& \frac{1}{2N} \| \vec y - \opX_p(M_p) \|_2^2 + \frac{\lambda_n}{K} \sum_{h=1}^K \| (M_p)_{(h)} \|_* \label{eq:regression}.
+%  \arg\min_{M_p}& \frac{1}{2N} \| \vec y - \opX_p(M_p) \|_2^2 + \frac{\lambda_n}{k} \sum_{h=1}^k \| (M_p)_{(h)} \|_* \label{eq:regression}.
 %\end{align}
@@
 $$\frac{1}{2n} \| \opX_p( \Delta )\|_2^2 \ge \kappa(\opX_p) \|\Delta\|^2_F \quad \text{and} \quad
-\lambda_n \ge \frac{\|\opX_p^*(\eta_p)\|_{op}}{n}.$$
+\lambda_n \ge \frac{\|\opX_p^*(\eta_p)\|_\Lop}{n}.$$
 Then the error of $\hat M_p$ is bounded as follows:
@@
 parameter $\kappa(\opX_p)$ and upper bound the adjoint operator
-$\|\opX_p^*(\eta_p)\|_{op}$.
-%We
-%will appeal to the random design framework that models the input $x$ as
-%random and show bounds that hold with high probability.
+$\|\opX_p^*(\eta_p)\|_\Lop$. The proofs of the following lemmas follow
+from standard concentration inequalities, and have been deferred to
+\appendixref{sec:proofs}.
 
+%We will appeal to the random design framework that models the input $x$
+%as random and show bounds that hold with high probability.
 %\paragraph{Adjoint operator}
+%In this section, we upper bound the operator norm of the adjoint
+%$\|\opX_p(\eta_p)\|_\Lop$.
 
-%In this section, we upper bound the operator norm of the adjoint
-%$\|\opX_p(\eta_p)\|_\text{op}$.
-
-First, let us lower bound the restricted strong convexity parameter $\kappa(\opX_p)$:
+% First, let us lower bound the restricted strong convexity parameter $\kappa(\opX_p)$:
 
@@
 \end{lemma}
-\begin{proof}
-Expanding the definition of the observation operator:
-$\frac{1}{2n} \|\opX_p(\Delta)\|_2^2
-= \frac{1}{2n} \sum_{(x,y) \in \sD} \innerp{\Delta}{x\tp{p}}^2$.
-Unfolding the tensor and letting $\hat\Sigma_p \eqdef \frac{1}{n} \sum_{(x,y) \in \sD} \cvec(x\tp{p})\tp{2}$,
-we have 
-$\frac{1}{2n} \|\opX_p(\Delta)\|_2^2
-= \frac{1}{2} \trace(\cvec(\Delta)\tp{2} \hat\Sigma_p)
-\ge \frac{1}{2} \|\Delta\|_F^2 \sigmamin(\hat\Sigma_p)$.
-It remains to show that $\sigmamin(\hat\Sigma_p)$ and $\sigmamin(\Sigma_p)$ are close.
-To do this, we first apply Hoeffding's inequality elementwise.
-Since $\|x\|_2 \le R$, we have that for each element $(i,j)$,
-$|(\hat\Sigma_p)_{ij} - (\Sigma_p)_{ij}| = O(R^{2p}\sqrt{\frac{\log (1/\delta)}{n}})$.
-Applying the union bound over the $d^{2p}$ elements of $\hat\Sigma_p - \Sigma_p$,
-we have that the max norm is bounded:
-$\|\hat\Sigma_p - \Sigma_p\|_\text{\rm max} = O(R^{2p} \sqrt{\frac{p \log(d) \log (1/\delta)}{n}})$.
-The max norm times $d^p$ upper bounds the Frobenius norm, which upper bounds the operator norm, so we have that
-$\|\hat\Sigma_p - \Sigma_p\|_\text{\rm op} = O(d^p R^{2p} \sqrt{\frac{p \log(d) \log (1/\delta)}{n}})$.
-Using the fact that $\sigmamin(\hat\Sigma_p) \ge \sigmamin(\Sigma_p) - \|\hat\Sigma_p - \Sigma_p\|_\text{\rm op}$
-yields the result.
-\end{proof}
 
-Next, let us upper bound the operator norm of the observation adjoint:
+% Next, let us upper bound the operator norm of the observation adjoint:
 
@@
 \end{lemma}
-\begin{proof}
-Let $\hat\E_p[f(x,\epsilon,h)]$ denote the empirical expectation over
-the examples in dataset $\sD_p$ (recall the $\sD_p$'s are independent to
-simplify the analysis).  By definition,
-$\frac1n \|\opX_p^*(\eta_p)\|_\op = \hat\E_p[\eta_p(x) x\tp{p}]$
-for $p \in \{1,2,3\}$ defined in \refeqn{y1}, \refeqn{y2}, and
-\refeqn{y3}. Each $\eta_p(x)$ is composed of several zero-mean random
-variables, and since $\|A + B\|_\op \le \|A\|_\op + \|B\|_\op$, it
-suffices to consider each term in turn. For each term, we will
-condition on $x$ and use a standard concentration inequality to bound
-the noise.
-
-%\begin{align}
-%  \| \hat\E_1[ \eta_1(x) x ] \|_{op} &\le 
-%  \underbrace{ \|\hat\E_1[\innerp{M_1 - \beta_h}{x} x]\|_\op }_{ O(R^2 L \sqrt{\frac{\log(1/\delta_1)}{n}}) } +
-%  \underbrace{ \|\hat\E_1[\epsilon x]\|_\op }_{ O(R^2 L \sqrt{\frac{\log(1/\delta_1)}{n}}) } +
-%\end{align}
-
-First, consider $\eta_1(x)$, which contains two terms.
-The first term is $\innerp{M_1 - \beta_h}{x}$.
-We have that $\|M_1 - \beta_h\|_2 \le 2L$,
-so by Hoeffding's inequality, and using the fact that $\|x\|_2 \le R$,
-with probability $1-\delta_1$,
-$\|\hat\E_1[\innerp{M_1 - \beta_h}{x} x]\|_\op = O(R^2 L \sqrt{\frac{\log(1/\delta_1)}{n}})$.
-The second term is $\epsilon \sim \normal{0}{\sigma^2}$,
-with probability $1-\delta_1$,
-$\|\hat\E_1[\epsilon x]\|_\op = O(R \sigma \sqrt{\frac{\log(1/\delta_1)}{n}})$.
-
-Now, consider $\eta_2(x)$, which contains three terms.
-Using similar techniques, we have that
-$\|\hat\E_2[\eta_2(x) x\tp{2}]\|_\op = O(R^2 (L^2 R^2 + \sigma L R + \sigma^2) \sqrt{\frac{\log(1/\delta_1)}{n}})$.
-
-Finally, $\eta_3(x)$ contains five terms plus an additional bias since the estimator uses $\hat M_1$ rather than $M_1$.
-There are two additional considerations.
-First, to bound $\hat\E_3[\epsilon^3]$, we use Lemma 19 of \cite{hsu13spherical}.
-Second, the additional bias is small by the previous bound (and it is constructed on $\sD_1$, which is independent of $\sD_3$
-used to construct $\hat M_3$.
-Therefore, we incur another $O(\sigma^2 \cdot R \sigma \sqrt{\frac{\log(1/\delta_1)}{n}})$.
-We can handle the other terms using conventional means.
-Combining everything, we obtain
-$\|\hat\E_3[\eta_3(x) x\tp{3}]\|_\op = O(R^3 (L^3 R^3 + \sigma L^2 R^2 + \sigma^2 L R + \sigma^3 + \sigma^3 R) \sqrt{\frac{\log^3(1/\delta_1)}{n}})$.
-
-Finally, taking $\delta_1 = \delta/3$, and taking the union bound over the bounds for $p \in \{1,2,3\}$,
-we get our result.
-Note that we are bounding quantities fairly crudely for the sake of simplicity.
-\end{proof}
 
@@
 the data, we will extend the analysis of the robust tensor method in
-\citet{AnandkumarGeHsu2012} to bound the error on the parameters
-$\theta$. 
+\citet{AnandkumarGeHsu2012} to include the whitening and unwhitening
+steps, 2(a) and 2(c) respectively, of
+\algorithmref{algo:spectral-experts}.
 \begin{lemma}
-  \label{lem:tensorPower} Let $\|\hat M_2 - M_2\|_{op} \le \epsilon$ and
-  $\|\hat M_3 - M_3\|_{op} \le \epsilon$, for some $\epsilon$ such that 
-  $$\epsilon < \min\{\sigmamin(M_2)/2,O(k \sigmamin(M_2)^{1.5} \|M_3\|_{op}^{-1})\}.$$ 
-  Then, for some permutation of indices, the parameter estimates found
-  in part 2 of \algorithmref{algo:spectral-experts} satisfy the
-  following with probability atleast $1 - \delta$,
+  \label{lem:tensorPower} Let $\|\hat M_2 - M_2\|_\Lop \le \epsilon$ and
+  $\|\hat M_3 - M_3\|_\Lop \le \epsilon$, for some $\epsilon$ such that 
+  $$\epsilon < \min\{\sigmamin(M_2)/2,O(k \sigmamin(M_2)^{1.5} \|M_3\|_\Lop^{-1})\}.$$ 
+  Then, there exists a permutation of indices such that  the parameter
+  estimates found in step 2 of \algorithmref{algo:spectral-experts}
+  satisfy the following with probability at least $1 - \delta$,
   \begin{align*}
   \|\hat \pi - \pi \|_{\infty}
-    &= O( k \pi_{max}^{2.5} \pi_{min}^{-0.5} \sigmamin(M_2)^{-1.5} \|M_3\|_{op} \epsilon ) \\
+    &= O( k \pi_{max}^{2.5} \pi_{min}^{-0.5} \sigmamin(M_2)^{-1.5} \|M_3\|_\Lop \epsilon ) \\
   \|\hat \beta_h - \beta_h\|_2
-    &= O( k \pi_{max} \sigmamin(M_2)^{-1.5} \|M_2\|_{op}^{0.5} \|M_3\|_{op} \epsilon ).
+    &= O( k \pi_{max} \sigmamin(M_2)^{-1.5} \|M_2\|_\Lop^{0.5} \|M_3\|_\Lop \epsilon ).
   \end{align*}
@@
 \end{lemma}
-The proof of the above lemma has been deferred to
+
+The proof follows by applying standard matrix perturbation results for
+the whitening and unwhitening operators, and has again been deferred
 \appendixref{sec:proofs}.
@@
 
-\begin{proof}
+\begin{proof}(Theorem 1)
 By \reflem{lowRank} along with \reflem{lowRankLower} and
@@
 \begin{align}
-  \|\hat M_p - M_p\|_{op} = O\left( \sigma^3 L^3 R^6 k^{\frac12} \sigmamin(\Sigma_p)^{-1} \sqrt{\frac{\log^3 (1/\delta)}{n}} \right).
+  \|\hat M_p - M_p\|_\Lop = O\left( \sigma^3 L^3 R^6 k^{\frac12} \sigmamin(\Sigma_p)^{-1} \sqrt{\frac{\log^3 (1/\delta)}{n}} \right).
   %\|\hat M_p - M_p\|_F = O\left( \frac{\sigma^3 L^3 R^6 \sqrt{k} \log^3 (1/\delta)}{\sqrt{n} (\sigmamin(\Sigma_p) - d^p R^p \sqrt{\frac{p \log(d) \log(1/\delta)}{n}}} \right).
@@
 We complete the proof by applying \reflem{tensorPower} with the above
-bound on $\|\hat M_p - M_p\|_{op}$.
+bound on $\|\hat M_p - M_p\|_\Lop$.
 

### 2013-04-26 doc/icml2013/theory.tex
Minor caps changes and math changes

@@
 \label{thm:convergence}
-Each dataset $\sD_p$ (for $p = 1, 2, 3$) consists of $n$ i.i.d.\ points drawn from a mixture
-of linear regressions model with parameter $\theta^*$
-(having three independent copies simplifies the analysis).
-Let $\|x\|_2 \le R$ with probability $1$
-and let $\|\beta_h^*\|_2 \le L$ for all $h \in [k]$.
+Assume each dataset $\sD_p$ (for $p = 1, 2, 3$) consists of $n$ i.i.d.\ points independently drawn from a mixture
+of linear regressions model with parameter $\theta^*$\footnote{Having three independent copies simplifies the analysis}.
+Further, assume $\|x\|_2 \le R$ with probability $1$, $\|\beta_h^*\|_2
+\le L$ for all $h \in [k]$ and $B$ is full rank.
 Let $\Sigma_p \eqdef \E[\cvec(x\tp{p})\tp{2}]$, 
 and assume $\Sigma_p \succ 0$ for each $p \in \{1,2,3\}$.
-Assume $d \ge k$.
 Suppose the number of samples is

### 2013-05-06 doc/icml2013/algo.tex
Added comments from Percy
Minor caps changes and math changes
Updated sections adding a brief discussion after Theorem 1 and before the experiments

@@
-\section{Spectral Experts Algorithm}
+\section{Spectral Experts algorithm}
 \label{sec:algo}
@@
 
+\begin{algorithm}[t]
+  \caption{Spectral Experts}
+  \label{algo:spectral-experts}
+  \begin{algorithmic}[1]
+    \INPUT Three datasets $\mathcal{D}_r = \{ (\xn{1}, \yn{1}), \cdots, (\xn{n}, \yn{n}) \}$ for $r = 1, 2, 3$;
+    %regularization strengths $\lambda_n^{(2)} = \frac{c_2}{\sqrt{n}}$, $\lambda_n^{(3)} = \frac{c_3}{\sqrt{n}}$;
+    regularization strengths $\lambda_n^{(2)}$, $\lambda_n^{(3)}$;
+    observation noise variance $\sigma^2$.
+    \OUTPUT Parameters $\hat\theta = (\hat \pi, [\hat \beta_1 \mid \cdots \mid \hat \beta_k])$.
+    \STATE Estimate compound parameters $M_2, M_3$ using \textbf{low-rank regression}:
+    \begin{align}
+      &\hat M_1 = \arg\min_{M_1} \label{eqn:estimateM1} \\
+      &\quad\quad\frac{1}{2n} \sum_{(x,y) \in \sD_1} (\innerp{M_1}{x} - y)^2, \nonumber \\
+      &\hat M_2 = \arg\min_{M_2} \quad \lambda_n^{(2)} \|M_2\|_* + \label{eqn:estimateM2} \\
+      &\quad\quad\frac{1}{2n} \sum_{(x,y) \in \sD_2} (\innerp{M_2}{x\tp{2}} + \sigma^2 - y^2)^2, \nonumber \\
+      &\hat M_3 = \arg\min_{M_3} \quad \lambda_n^{(3)} \|M_3\|_* + \label{eqn:estimateM3} \\
+      &\quad\quad\frac{1}{2n} \sum_{(x,y) \in \sD_3} (\innerp{M_3}{x\tp{3}} + 3 \sigma^2\innerp{\hat M_1}{x} - y^3)^2. \nonumber
+    \end{align}
+    \STATE Estimate the parameters $\theta = (\pi, B)$ using \textbf{tensor factorization}:
+    \begin{enumerate}
+      \item [(a)] Compute whitening matrix $\hat W \in \Re^{d \times k}$ (such that $\hat W^\top
+      \hat M_2 \hat W = I$) using SVD.
+      \item [(b)] Compute the eigenvalues $\{\hat a_h\}_{h=1}^k$
+      and eigenvectors $\{\hat v_h\}_{h=1}^k$
+      of the whitened tensor $\hat M_3(\hat W, \hat W, \hat W) \in \Re^{k \times k \times k}$
+      by using the robust tensor power method.
+    \item [(c)] Set parameter estimates $\hat\pi_h = \frac{1}{\hat a_h^2}$
+    and $\hat\beta_h = (\hat W^{\top})^\dagger (\hat a_h \hat v_h)$.
+    \end{enumerate}
+  \end{algorithmic}
+\end{algorithm}
+
 % Full algorithm
@@
 much attention in recent years.
-We use a rather simple proximal gradient-based approach,
-in which the nuclear norm is handled in closed form by taking an SVD
-and soft-thresholding the singular values \cite{donoho95soft,cai10soft}.
+% We use a standard off-the-shelf SDP solver, CVX, to solve .
+% We use a rather simple proximal gradient-based approach,
+% in which the nuclear norm is handled in closed form by taking an SVD
+% and soft-thresholding the singular values \cite{donoho95soft,cai10soft}.
 
 % Tensor factorization
-Having estimated the compound parameters $M_2$ and $M_3$,
-it remains to recover the original parameters $\theta$.
+Having estimated the compound parameters $M_1$, $M_2$ and $M_3$, it
+remains to recover the original parameters $\theta$.
 \citet{AnandkumarGeHsu2012} showed that for $M_2$ and $M_3$ of
@@
 %\sectionref{sec:tensor-power}.
-
-\begin{algorithm}[t]
-  \caption{Spectral Experts}
-  \label{algo:spectral-experts}
-  \begin{algorithmic}[1]
-    \INPUT Three independent datasets $\mathcal{D}_r = \{ (\xn{1}, \yn{1}), \cdots, (\xn{n}, \yn{n}) \}$ for $r = 1, 2, 3$;
-    %regularization strengths $\lambda_n^{(2)} = \frac{c_2}{\sqrt{n}}$, $\lambda_n^{(3)} = \frac{c_3}{\sqrt{n}}$;
-    regularization strengths $\lambda_n^{(2)}$, $\lambda_n^{(3)}$;
-    observation noise variance $\sigma^2$.
-    \OUTPUT Parameters $\hat\theta = (\hat \pi, [\hat \beta_1 \mid \cdots \mid \hat \beta_k])$.
-    \STATE Estimate compound parameters $M_2, M_3$ using \textbf{low-rank regression}:
-    \begin{align}
-      &\hat M_1 = \arg\min_{M_1} \label{eqn:estimateM1} \\
-      &\quad\quad\frac{1}{2n} \sum_{(x,y) \in \sD_1} (\innerp{M_1}{x} - y)^2, \nonumber \\
-      &\hat M_2 = \arg\min_{M_2} \quad \lambda_n^{(2)} \|M_2\|_* + \label{eqn:estimateM2} \\
-      &\quad\quad\frac{1}{2n} \sum_{(x,y) \in \sD_2} (\innerp{M_2}{x\tp{2}} + \sigma^2 - y^2)^2, \nonumber \\
-      &\hat M_3 = \arg\min_{M_3} \quad \lambda_n^{(3)} \|M_3\|_* + \label{eqn:estimateM3} \\
-      &\quad\quad\frac{1}{2n} \sum_{(x,y) \in \sD_3} (\innerp{M_3}{x\tp{3}} + 3 \sigma^2\innerp{\hat M_1}{x} - y^3)^2. \nonumber
-    \end{align}
-    \STATE Estimate the parameters $\theta = (\pi, B)$ using \textbf{tensor factorization}:
-    \begin{enumerate}
-      \item [(a)] Compute whitening matrix $\hat W \in \Re^{d \times k}$ (such that $\hat W^\top
-      \hat M_2 \hat W = I$) using SVD.
-      \item [(b)] Compute the eigenvalues $\{\hat a_h\}_{h=1}^k$
-      and eigenvectors $\{\hat v_h\}_{h=1}^k$
-      of the whitened tensor $\hat M_3(\hat W, \hat W, \hat W) \in \Re^{k \times k \times k}$
-      by using the robust tensor power method.
-    \item [(c)] Set parameter estimates $\hat\pi_h = \frac{1}{\hat a_h^2}$
-    and $\hat\beta_h = (\hat W^{\top})^\dagger (\hat a_h \hat v_h)$.
-    \end{enumerate}
-  \end{algorithmic}
-\end{algorithm}

### 2013-05-07 java/src/learning/linalg/RandomFactory.java
Fixed mucho bug using \sigma2 instead of \sqrt{sigma2} to scale Gaussian

@@ -48,8 +48,8 @@ public class RandomFactory {
    * @param sigma - noise
    * @return
    */
-  public static double randn(double sigma) {
-    return rand.nextGaussian() * sigma;
+  public static double randn(double sigma2) {
+    return rand.nextGaussian() * Math.sqrt(sigma2);
   }
 
   /**

### 2013-05-07 java/src/learning/linalg/MatrixOps.java
Added small math helper routines

@@ -44,6 +44,9 @@ public class MatrixOps {
   public static void printSize( SimpleMatrix X ) {
     System.out.printf( "(%d, %d)\n", X.numRows(), X.numCols() );
   }
+  public static void printSize( DenseMatrix64F X ) {
+    System.out.printf( "(%d, %d)\n", X.numRows, X.numCols );
+  }
   public static void printSize( Tensor X ) {
     System.out.printf( "(%d, %d, %d)\n", X.getDim(0), X.getDim(1), X.getDim(2) );
   }
@@ -471,7 +474,13 @@ public class MatrixOps {
       sum += x[i];
     return sum;
   }
-
+  public static double sum(double[][] x ) {
+    double sum = 0.0;
+    for( int i = 0; i < x.length; i++ )
+      for( int j = 0; j < x[i].length; j++ )
+        sum += x[i][j];
+    return sum;
+  }
   /**
    * Find the sum of a vector
    */
@@ -857,6 +866,12 @@ public class MatrixOps {
   }
   public static SimpleMatrix colorer( SimpleMatrix X ) {
     Triplet<SimpleMatrix, SimpleMatrix, SimpleMatrix> UDV = svdk(X);
+    SimpleMatrix U = UDV.getValue0();
+    SimpleMatrix D = UDV.getValue1();
+    return U.mult(sqrt(D));
+  }
+  public static SimpleMatrix colorer( SimpleMatrix X, int K ) {
+    Triplet<SimpleMatrix, SimpleMatrix, SimpleMatrix> UDV = svdk(X, K);
     SimpleMatrix U = UDV.getValue0();
     SimpleMatrix D = UDV.getValue1();
     return U.mult(sqrt(D));

### 2013-05-07 java/src/learning/em/MixtureOfExperts.java
Added validity check for EM

@@ -89,6 +89,18 @@ public class MixtureOfExperts implements Runnable {
 
       return new Parameters( K, weights, betas, sigma2 );
     }
+
+    public boolean isValid() {
+      boolean weightsInSimplex = MatrixOps.equal( MatrixOps.sum( weights ), 1.0 );
+      boolean betasFinite = (MatrixOps.sum( betas ) != Double.POSITIVE_INFINITY) && (MatrixOps.sum( betas ) != Double.NEGATIVE_INFINITY);
+      boolean sigma2Positive = (sigma2 >= 0.0);
+      LogInfo.logsForce( "Valid: " + weightsInSimplex + " " + betasFinite + " " + sigma2Positive );
+      if ( !weightsInSimplex ) return false;
+      if ( !betasFinite ) return false;
+      if ( !sigma2Positive  ) return false;
+
+      return true;
+    }
   }
 
   public int K = 2;
@@ -129,6 +141,7 @@ public class MixtureOfExperts implements Runnable {
         MatrixOps.minus( R[n], Z );
         MatrixOps.exp( R[n] );
       }
+      MatrixOps.normalize( R[n] );
       assert( MatrixOps.equal( MatrixOps.sum( R[n] ), 1.0 ) );
     }
 
@@ -224,6 +237,8 @@ public class MixtureOfExperts implements Runnable {
   public Parameters run( double[] y, double[][] X, Parameters state ) {
     LogInfo.begin_track( "MixtureOfExperts-em" );
     double old_lhood = Double.NEGATIVE_INFINITY;
+
+    assert( state.isValid() );
 
     for( int i = 0; i < iters; i++ ) {
       double[][] responsibilities = computeResponsibilities( y, X, state );

### 2013-05-07 java/src/learning/spectral/applications/SpectralExperts.java
Support for CVX, Finer grain control for TensorPowerMethod

@@ -8,6 +8,7 @@ package learning.spectral.applications;
 import learning.exceptions.RecoveryFailure;
 import learning.exceptions.NumericalException;
 
+import learning.Misc;
 import learning.linalg.*;
 import learning.models.MixtureOfExperts;
 
@@ -15,6 +16,7 @@ import learning.optimization.PhaseRecovery;
 import learning.optimization.ProximalGradientSolver;
 import learning.optimization.ProximalGradientSolver.LearningRate;
 import learning.optimization.TensorRecovery;
+import learning.optimization.MatlabProxy;
 import learning.spectral.MultiViewMixture;
 
 import learning.data.MomentComputer;
@@ -36,13 +38,8 @@ import fig.exec.Execution;
 
 import java.lang.ref.SoftReference;
 import java.util.Date;
-import java.io.IOException;
-import java.io.FileNotFoundException;
-import java.io.FileInputStream;
-import java.io.FileOutputStream;
-import java.io.ObjectInputStream;
-import java.io.ObjectOutputStream;
 import java.lang.ClassNotFoundException;
+import java.io.*;
 
 /**
  * Spectral Experts.
@@ -54,6 +51,8 @@ public class SpectralExperts implements Runnable {
 	public String inputPath;
   @Option(gloss = "Number of samples to use (0 for all)")
   public double subsampleN = 0;
+  @Option(gloss = "Forced Resample")
+  public boolean forceResample = false;
 
   @Option(gloss = "Adjust the regularizer to be 1/sqrt(N) of the given number")
   public boolean adjustReg = true;
@@ -62,25 +61,38 @@ public class SpectralExperts implements Runnable {
   @Option(gloss = "Ridge Regression Regularization (Triples) (<0 => ridgeReg2*10)")
   public double ridgeReg3 = -1;
   @Option(gloss = "Trace Norm Regularization (Pairs)")
-  public double traceReg2 = 1e0;
+  public double traceReg2 = 1e-4;
   @Option(gloss = "Trace Norm Regularization (Triples)")
   public double traceReg3 = -1;
   @Option(gloss = "Scale Data?")
   public boolean scaleData = false;
   @Option(gloss = "Run spectral? (if false, just compute the moments)")
   public boolean runSpectral = true;
+
   @Option(gloss = "Run EM on spectral output?")
   public boolean runEM = false;
   @Option(gloss = "Run EM on spectral output?")
[…]

### 2013-05-07 java/tests/learning/spectral/applications/SpectralExpertsTest.java
Added test for regression

@@ -10,6 +10,7 @@ import learning.models.MixtureOfExperts;
 
 import org.javatuples.*;
 import org.ejml.simple.SimpleMatrix;
+import org.ejml.data.DenseMatrix64F;
 import org.junit.Assert;
 import org.junit.Test;
 import org.junit.Before;
@@ -27,6 +28,92 @@ public class SpectralExpertsTest {
   public void setup(){
     LogInfo.writeToStdout = false;
     LogInfo.init();
+  }
+
+  public void testRegressionRunner(MixtureOfExperts model, int N, double reg) {
+    int K = model.getK(); int D = model.getD();
+    SpectralExperts algo = new SpectralExperts();
+    algo.K = K;
+    algo.enableAnalysis(model);
+
+    // Set up the regression problem.
+    SimpleMatrix B = model.getBetas();
+    Pair<SimpleMatrix, FullTensor> moments = SpectralExperts.computeExactMoments( model );
+    SimpleMatrix Pairs = moments.getValue0();
+    FullTensor Triples = moments.getValue1();
+
+    // Compute X's 
+    SimpleMatrix X = RandomFactory.rand( N, D );
+
+    // Compute y2, y3's
+    SimpleMatrix y2 = MatrixFactory.zeros(N);
+    SimpleMatrix y3 = MatrixFactory.zeros(N);
+    for( int n = 0; n < N; n++ ) {
+      DenseMatrix64F Xn = MatrixOps.row(X.getMatrix(), n );
+      y2.set( n, MatrixOps.xMy( Xn, Pairs.getMatrix(), Xn ) );
+      y3.set( n, Triples.project3( Xn, Xn, Xn ) );
+    }
+
+    // Regress to get Pairs_ and Triples_
+    SimpleMatrix Pairs_ = algo.recoverPairsByRidgeRegression( y2, X );
+    FullTensor Triples_ = algo.recoverTriplesByRegression( y3, X );
+
+    System.out.println( algo.analysis.Pairs );
+    System.out.println( Pairs_ );
+    algo.analysis.reportPairs(Pairs_);
+    algo.analysis.reportTriples(Triples_);
+
+    Assert.assertTrue( algo.analysis.PairsErr < 1e-2 );
+    Assert.assertTrue( algo.analysis.TriplesErr < 1e-1 );
+  }
+
+  public void arbitraryRegressionTest(SpectralExperts algo, MixtureOfExperts.GenerationOptions options) throws NumericalException {
[…]

### 2013-05-07 doc/icml2013/abstract.tex

@@
 In this paper, we develop a new computationally efficient
-and provably consistent estimator for the mixture of linear regressions,
+and provably consistent estimator for a mixture of linear regressions,
 a simple instance of discriminative latent-variable models.

### 2013-05-07 doc/icml2013/evaluation.tex

@@
 \label{sec:evaluation}
+
+In the previous section, we showed that Spectral Experts is a consistent
+algorithm that converges to the global solution at a $\frac{1}{\sqrt}$
+rate. In this section, we will support these theoretical results with
+experiments on simulated data. We will see that in the finite sample
+regime, even when Spectral Experts does not converge to the optimal
+answer, it provides robust initialization for local methods like EM.
 

### 2013-05-07 doc/icml2013/model.tex

@@
 The generation of $y$ given $x$ involves two steps:
-(i) draw the mixture component $h \in [K]$ according to mixture proportions
-$\pi = (\pi_1, \dots, \pi_K)$;
+(i) draw the mixture component $h \in [k]$ according to mixture proportions
+$\pi = (\pi_1, \dots, \pi_k)$;
 and (ii) generate $y$ from a Gaussian centered at $\beta_h^\top x$ with
@@
 where $\pi \in \Re^d$ are the mixture proportions and
-$B = [\beta_1 \mid \dots \mid \beta_K] \in \Re^{d \times K}$
+$B = [\beta_1 \mid \dots \mid \beta_k] \in \Re^{d \times k}$
 are the regression coefficients.
@@
 % Acknowledge that mixture proportions % can't depend on $x$
-The model is an instance of hierarchical mixture of experts
+The model is an instance of the hierarchical mixture of experts
 \cite{jacobs91experts}, in which the mixture proportions are allowed to depend
@@
 % non-convex.
-The estimation problem for mixture of linear regressions is difficult because
+The estimation problem for a mixture of linear regressions is difficult because
 the mixture components $h$ are unobserved,

### 2013-05-07 doc/icml2013/theory.tex

@@
 using results from \citet{Tomioka2011}.
-Then, we use results from \citet{AnandkumarGeHsu2012} to convert this error
+While the $R^{12}$ term looks formidable, it is unavoidable since we
+perform regression on the third moment of the data. Classically, the
+number of samples required is norm squared of the covariance matrix,
+which itself is bounded by the norm squared of the data, $R^3$. This
+third-order dependence also shows up in the rate the regression
+coefficients are annealed; the cubic terms bound each of $\epsilon^3$,
+$\beta_h^3$ and $\|(x\tp{3})\tp{2}\|_F$ with high probability.
+
+Next, we use results from \citet{AnandkumarGeHsu2012} to convert this error
 into a bound on the actual parameter estimates $\hat\theta = (\hat\pi, \hat B)$

### 2013-05-07 doc/icml2013/evaluation.tex
Added bit of discussion, and minor changes

@@
-\section{Empirical Evaluation}
+\section{Empirical evaluation}
 \label{sec:evaluation}
@@
 In the previous section, we showed that Spectral Experts is a consistent
-algorithm that converges to the global solution at a $\frac{1}{\sqrt}$
+algorithm that converges to the global solution at a $\frac{1}{\sqrt n}$
 rate. In this section, we will support these theoretical results with
@@
 
-\subsection{Experimental Setup}
+\subsection{Experimental setup}
 
@@
 We solved the low rank recovery optimization routines for $M_2$ and
-$M_3$ using \texttt{cvx}\citationneeded. 
+$M_3$ using an off-the-shelf solver for convex programs, CVX\cite{cvx}.
 %ainterior point method an interior proximal subgradient descent

### 2013-05-07 doc/icml2013/intro.tex
Added bit of discussion, and minor changes

@@
 let $\|X\|_{\max}$ denote the max norm (elementwise maximum),
-let $\|X\|_\Lop$ denote the operator norm (largest singular value),
+let $\|X\|_\op$ denote the operator norm (largest singular value),
 let $\sigma_\text{min}(X)$ be the smallest singular value of $X$.
@@
 the average nuclear norm over all $p$ unfoldings,
-and let $\|X\|_\Lop = \frac{1}{p} \sum_{i=1}^p \|X_{(i)}\|_\Lop$
+and let $\|X\|_\op = \frac{1}{p} \sum_{i=1}^p \|X_{(i)}\|_\op$
 denote the average operator norm over all $p$ unfoldings.
@@
 For a symmetric tensor $X \in (\Re^d)\tp{p}$, let $\cvec(X) \in
-\Re^{D_p}, D_p = \binom{d + p + 1}{p}$ be the collapsed vectorization of distinct
+\Re^{C_{d,p}}, C_{d,p} = \binom{d + p + 1}{p}$ be the collapsed vectorization of distinct
 elements in $X$, for example, for $X \in \Re^{d \times d}$, $\cvec(X)

### 2013-05-07 doc/icml2013/theory.tex
Added bit of discussion, and minor changes

@@
 
-We note that this requires more stringent positive definite constraints
-when using features. For example, suppose $x = (1, t, t^2)$, the common
-polynomial basis expansion, so that all the coordinates are
-deterministically related.  Then $\E[x\tp{2}] \succ 0$ can easily be
-satisfied. 
+While this is not a concern for independent dimensions, we note that
+this requires more stringent positive definite constraints when using
+features. For example, suppose $x = (1, t, t^2)$, the common polynomial
+basis expansion, so that all the coordinates are deterministically
+related.  Then $\E[x\tp{2}] \succ 0$ can easily be satisfied. 
 %, say with $x \sim \normal{0}{I}$.
@@
 $$\frac{1}{2n} \| \opX_p( \Delta )\|_2^2 \ge \kappa(\opX_p) \|\Delta\|^2_F \quad \text{and} \quad
-\lambda_n \ge \frac{\|\opX_p^*(\eta_p)\|_\Lop}{n}.$$
+\lambda_n \ge \frac{\|\opX_p^*(\eta_p)\|_\op}{n}.$$
 Then the error of $\hat M_p$ is bounded as follows:
@@
 parameter $\kappa(\opX_p)$ and upper bound the adjoint operator
-$\|\opX_p^*(\eta_p)\|_\Lop$. The proofs of the following lemmas follow
-from standard concentration inequalities, and have been deferred to
-\appendixref{sec:proofs}.
+$\|\opX_p^*(\eta_p)\|_\op$. The proofs of the following lemmas follow
+from standard concentration inequalities and appear in detail in the
+supplementary material.
+%have been deferred to \appendixref{sec:proofs}.
 
@@
 %In this section, we upper bound the operator norm of the adjoint
-%$\|\opX_p(\eta_p)\|_\Lop$.
+%$\|\opX_p(\eta_p)\|_\op$.
 
@@
 \begin{lemma}
-  \label{lem:tensorPower} Let $\|\hat M_2 - M_2\|_\Lop \le \epsilon$ and
-  $\|\hat M_3 - M_3\|_\Lop \le \epsilon$, for some $\epsilon$ such that 
-  $$\epsilon < \min\{\sigmamin(M_2)/2,O(k \sigmamin(M_2)^{1.5} \|M_3\|_\Lop^{-1})\}.$$ 
+  \label{lem:tensorPower} Let $\|\hat M_2 - M_2\|_\op \le \epsilon$ and
+  $\|\hat M_3 - M_3\|_\op \le \epsilon$, for some $\epsilon$ such that 
+  $$\epsilon < \min\{\sigmamin(M_2)/2,O(k \sigmamin(M_2)^{1.5} \|M_3\|_\op^{-1})\}.$$ 
   Then, there exists a permutation of indices such that  the parameter
@@
   \|\hat \pi - \pi \|_{\infty}
-    &= O( k \pi_{max}^{2.5} \pi_{min}^{-0.5} \sigmamin(M_2)^{-1.5} \|M_3\|_\Lop \epsilon ) \\
+    &= O( k \pi_{max}^{2.5} \pi_{min}^{-0.5} \sigmamin(M_2)^{-1.5} \|M_3\|_\op \epsilon ) \\
   \|\hat \beta_h - \beta_h\|_2
-    &= O( k \pi_{max} \sigmamin(M_2)^{-1.5} \|M_2\|_\Lop^{0.5} \|M_3\|_\Lop \epsilon ).
+    &= O( k \pi_{max} \sigmamin(M_2)^{-1.5} \|M_2\|_\op^{0.5} \|M_3\|_\op \epsilon ).
   \end{align*}
@@
 The proof follows by applying standard matrix perturbation results for
-the whitening and unwhitening operators, and has again been deferred
-\appendixref{sec:proofs}.
+the whitening and unwhitening operators and has again been deferred
+to the supplementary material.
+%\appendixref{sec:proofs}.
 
@@
 
-\begin{proof}(Theorem 1)
-By \reflem{lowRank} along with \reflem{lowRankLower} and
-\reflem{lowRankUpper}, we can control the operator norm of the parameter
-estimates which are directly upper bounded by the Frobenius norm: If $n
-\ge n_1$, then
+\begin{proof}[Proof of Theorem 1]
+By \reflem{lowRank}, \reflem{lowRankLower} and \reflem{lowRankUpper}, we
+can control the Frobenius norm of the error in the moments, which
+directly upper bounds the operator norm: If $n \ge \max\{n_1, n_2\}$,
+then
 \begin{align}
-  \|\hat M_p - M_p\|_\Lop = O\left( \sigma^3 L^3 R^6 k^{\frac12} \sigmamin(\Sigma_p)^{-1} \sqrt{\frac{\log^3 (1/\delta)}{n}} \right).
+  \|\hat M_p - M_p\|_\op = O\left( \sigma^3 L^3 R^6 k^{\frac12} \sigmamin(\Sigma_p)^{-1} \sqrt{\frac{\log^3 (1/\delta)}{n}} \right).
   %\|\hat M_p - M_p\|_F = O\left( \frac{\sigma^3 L^3 R^6 \sqrt{k} \log^3 (1/\delta)}{\sqrt{n} (\sigmamin(\Sigma_p) - d^p R^p \sqrt{\frac{p \log(d) \log(1/\delta)}{n}}} \right).
@@
 We complete the proof by applying \reflem{tensorPower} with the above
-bound on $\|\hat M_p - M_p\|_\Lop$.
+bound on $\|\hat M_p - M_p\|_\op$.
 

### 2013-05-08 doc/icml2013/algo.tex
Expanded the whole of the whitening theorem

@@
 Another line of work is based on the method of moments and uses eigendecomposition of a certain tensor
-to recover the parameters \cite{anandkumar12svd,anandkumar12moments,AnandkumarHsuKakade2012}.
+to recover the parameters \cite{anandkumar12lda,anandkumar12moments}.
 Our work extends this line of work into the space of discriminative models,

### 2013-05-08 doc/icml2013/intro.tex
Expanded the whole of the whitening theorem

@@
 \emph{generative} latent-variable models, including
-mixture models, HMMs \cite{AnandkumarHsuKakade2012},
+mixture models, HMMs \cite{anandkumar12moments},
 Latent Dirichlet Allocation \cite{anandkumar12lda},
@@
 %they required as many moments as there were parameters, and hence deemed
-%impractical. Recently, \citet{AnandkumarHsuKakade2012} showed how using
+%impractical. Recently, \citet{anandkumar12moments} showed how using
 %eigendecomposition, the moment equations could be solved using just the

### 2013-05-08 doc/icml2013/abstract.tex
updated intro and model

@@
 and provably consistent estimator for a mixture of linear regressions,
-a simple instance of discriminative latent-variable models.
+a simple instance of a discriminative latent-variable model.
 Our approach relies on a low-rank linear regression to recover
 a symmetric tensor, which can be factorized into the parameters
-using the tensor power method.
+using a tensor power method.
 We prove rates of convergence for our estimator

### 2013-05-08 doc/icml2013/intro.tex
updated intro and model

@@
 However, parameter estimation in these models is difficult;
-past approaches have relied exclusively on local optimization (EM, 
+past approaches rely on local optimization (EM, 
 gradient descent) and are vulnerable to local optima.
@@
 and parsing models \cite{hsu12identifiability}.
-The basic idea of this work is to express
+The basic idea of these methods is to express
 the unknown model parameters as a tensor factorization
-of the third-order moments of the model distribution, which
-can be estimated from data.
+of the third-order moments of the model distribution, a quantity
+which can be estimated from data.
 The moments have a special symmetric structure
@@
 directly reveal the tensor structure of the problem, so we cannot
-directly apply the above tensor factorization techniques.  Our solution
+simply apply the above tensor factorization techniques.  Our approach
 is to employ low-rank linear regression
-\cite{NegahbanWainwright2009,Tomioka2011} on second and third powers of
-the response to recover the appropriate symmetric tensors, using which we
-can then apply the tensor power method to retrieve the final parameters.
+\cite{NegahbanWainwright2009,Tomioka2011} to predict the second and third powers of
+the response.  The solution to these regression problems provide the appropriate symmetric tensors,
+on which we can then apply the tensor power method to retrieve the final parameters.
 
@@
 We prove that our algorithm yields consistent parameter estimates under certain
-new identifiability conditions.  We also conduct an empirical evaluation
+identifiability conditions.  We also conduct an empirical evaluation
 of our technique to understand its statistical properties (\sectionref{sec:evaluation}).
@@
 We use $O(f(n))$ to denote a function $g(n)$ such that $\lim_{n \to\infty} g(n)/f(n) < \infty$.
-We will only hide universal constants, exposing all dependence on the number of samples,
-dimension, norms, noise variance, etc.
+%We will only hide universal constants, exposing all dependence on the number of samples,
+%dimensionality, norms, etc.
 
 We use $x\tp{p}$ to represent the $p$-th order tensor formed by taking
-the outer product of $x \in \Re^d$; i.e. $x\tp{p}_{i_1 \ldots i_p}
+the tensor product of $x \in \Re^d$; i.e. $x\tp{p}_{i_1 \ldots i_p}
 = x_{i_1} \cdots x_{i_p}$. We will use $\innerp{\cdot}{\cdot}$ to denote
@@
 a $p$-th order tensor $X \in (\Re^d)\tp{p}$, the mode-$i$ unfolding of
-$X$ is a matrix, $X_{(i)} \in \Re^{d \times d^{p-1}}$, whose $j$-th row contains all the elements of $X$ whose
+$X$ is a matrix $X_{(i)} \in \Re^{d \times d^{p-1}}$, whose $j$-th row contains all the elements of $X$ whose
 $i$-th index is equal to $j$. 
@@
 let $\|X\|_*$ denote the nuclear (trace) norm (sum of singular values),
-let $\|X\|_F$ denote the Frobenius norm (square root of sum of squares of singular values),
-let $\|X\|_{\max}$ denote the max norm (elementwise maximum),
-let $\|X\|_\op$ denote the operator norm (largest singular value),
-let $\sigma_\text{min}(X)$ be the smallest singular value of $X$.
-For a tensor $X$,
+$\|X\|_F$ denote the Frobenius norm (square root of sum of squares of singular values),
+$\|X\|_{\max}$ denote the max norm (elementwise maximum),
+$\|X\|_\op$ denote the operator norm (largest singular value), and
+$\sigma_k(X)$ be the $k$-th largest singular value of $X$.
+For a $p$-th order tensor $X$,
 let $\|X\|_* = \frac{1}{p} \sum_{i=1}^p \|X_{(i)}\|_*$ denote
@@
 
-For a symmetric tensor $X \in (\Re^d)\tp{p}$, let $\cvec(X) \in
-\Re^{C_{d,p}}, C_{d,p} = \binom{d + p + 1}{p}$ be the collapsed vectorization of distinct
-elements in $X$, for example, for $X \in \Re^{d \times d}$, $\cvec(X)
+For a tensor $X \in (\Re^d)\tp{p}$, let $\cvec(X) \in
+\Re^{N(d,p)}, N(d,p) = \binom{d + p - 1}{p}$ be the collapsed vectorization of
+$X$. For example, if $X \in \Re^{d \times d}$, $\cvec(X)
 = (X_{ii} : i \in [d]; X_{ij} + X_{ji} : i,j \in [d], i<j)$.
@@
 = \{ k \in [d]^p : \forall i \in [d], c_i = |\{ j \in [p] : k_j = i \}|
-\}$ are the set of index vectors $k$ with that count profile.  
-%For symmetric tensors, $\innerp{X}{Y} = \innerp{\cvec(X)}{\cvec(Y)}$.
+\}$ are the set of index vectors $k$ whose count profile is $c$.
+For symmetric tensors $X$ and $Y$, $\innerp{X}{Y} = \innerp{\cvec(X)}{\cvec(Y)}$.
+Later, we'll see that vectorization allow us to perform regression on tensors,
+and collapsing simplifies our identifiability condition.

### 2013-05-08 doc/icml2013/model.tex
updated intro and model

@@
 given covariates $x \in \Re^d$.
+Let $k$ be the number of mixture components.
 The generation of $y$ given $x$ involves two steps:
-(i) draw the mixture component $h \in [k]$ according to mixture proportions
+(i) draw a mixture component $h \in [k]$ according to mixture proportions
 $\pi = (\pi_1, \dots, \pi_k)$;
-and (ii) generate $y$ from a Gaussian centered at $\beta_h^\top x$ with
-variance $\sigma^2$.
-\begin{eqnarray*}
+and (ii) set $y$ to $\beta_h^\top x$ plus some observation noise $\epsilon$.
+\begin{eqnarray}
   h &\sim& \mult(\pi), \\
-  y \mid h &\sim& \normal{ \beta_{h}^T x}{ \sigma^2 }.
-\end{eqnarray*}
+  y \mid h &=& \beta_{h}^T x + \epsilon.
+\end{eqnarray}
 The parameters of the model are $\theta = (\pi, B)$,
@@
 are the regression coefficients.
-We assume the variance of the observation noise $\sigma^2$ is known and that the mixture components are independent.
+Note that the choice of mixture component $h$ and the noise $\epsilon$ are independent.
+We also assume the distribution of the observation noise is known and has bounded support.
 The learning problem is stated as follows:

### 2013-05-08 doc/icml2013/algo.tex
algorithm updated

@@
 on the response $y$ given $x$.
-From the model definition, we have $y = \beta_h^\top x + \epsilon$ where
-$\epsilon \sim \normal{0}{\sigma^2}$, and $h$ is a random quantity
-independent of $x$.
-We can average over this randomness by defining
-the average regression coefficients
+From the model definition, we have $y = \beta_h^\top x + \epsilon$.
+The challenge is that the regression coefficients $\beta_h$ depend on the random $h$.
+%$\epsilon \sim \normal{0}{\E[\epsilon^2]}$, and $h$ is a random quantity
+%independent of $x$.
+The first key step is to average over this randomness by defining
+average regression coefficients
 $M_1 \eqdef \sum_{h=1}^k \pi_h \beta_h$.
-Now we can express $y$ as a linear function of $x$ with coefficients $M_1$
-plus some noise $\eta_1(x)$:
+Now we can express $y$ as a linear function of $x$ with non-random coefficients $M_1$
+plus some a new noise term $\eta_1(x)$:
 \begin{align}
@@
 due to the random choice of the mixture component $h$,
-and (ii) the \emph{observation noise} $\epsilon \sim \normal{0}{\sigma^2}$.
+and (ii) the \emph{observation noise} $\epsilon \sim \sE$.
 Although the noise depends on $x$,
@@
 perform linear regression on the data $\{\xni,
-\yni\}_{i=1}^{n}$ to produce a consistent estimate of $M_1$
-under identifiability conditions.
+\yni\}_{i=1}^{n}$ to produce a consistent estimate of $M_1$.
+%under identifiability conditions.
 But clearly, knowing $M_1$ is insufficient
-for identifying all the parameters $\theta$.
+for identifying all the parameters $\theta$,
+as
+$M_1$ only contains $d$ degrees of freedom whereas $\theta$ contains $O(kd)$.
 
 %%% second moments
-Intuitively, performing regression on only $y$ given $x$ provides only first-order
-information.  The key insight is that we can perform regression
-on higher order powers to obtain more information about the parameters.
+Intuitively, performing regression on $y$ given $x$ provides only first-order
+information.  The second key insight is that we can perform regression
+on higher-order powers to obtain more information about the parameters.
 Specifically, for an integer $p \ge 1$, let us define the average
@@
 \begin{align}
-y^2 &= \innerp{M_2}{x\tp{2}} + \sigma^2 + \eta_2(x), \label{eqn:y2} \\
-\eta_2(x) &= \innerp{\beta_h\tp{2} - M_2}{x\tp{2}} + 2 \epsilon \innerp{\beta_h}{x} + (\epsilon^2 - \sigma^2). \nonumber
+  y^2 &= \innerp{M_2}{x\tp{2}} + \E[\epsilon^2] + \eta_2(x), \label{eqn:y2} \\
+\eta_2(x) &= \innerp{\beta_h\tp{2} - M_2}{x\tp{2}} + 2 \epsilon \innerp{\beta_h}{x} + (\epsilon^2 - \E[\epsilon^2]). \nonumber
 \end{align}
 Again, we have expressed $y^2$ has a linear function of $x\tp{2}$
-with regression coefficients $M_2$, plus a known bias $\sigma^2$ and noise.\footnote{If $\sigma^2$ were not known,
+with regression coefficients $M_2$, plus a known bias $\E[\epsilon^2]$ and noise.\footnote{If $\E[\epsilon^2]$ were not known,
 we could treat it as another coefficient
-to be estimated.  The coefficients $M_2$ and $\sigma^2$ can be estimated jointly
+to be estimated.  The coefficients $M_2$ and $\E[\epsilon^2]$ can be estimated jointly
 provided that $x$ does not already contain a bias ($x_j$ must be non-constant for every $j \in [d]$).}
 Importantly, the noise has mean zero; 
-in fact each of the three terms has mean zero
-by (i) definition of $M_2$, (ii) independence of $\epsilon$ and $h$,
-and (iii) the fact that $\E[\epsilon^2] = \sigma^2$.
+in fact each of the three terms has zero mean
+by definition of $M_2$ and independence of $\epsilon$ and $h$.
 
 Performing regression yields a consistent estimate of $M_2$,
-but does not identify all the parameters $\theta$.
+but still does not identify all the parameters $\theta$.
 In particular, $B$ is only identified up to rotation:
 if $B = [\beta_1 \mid \cdots \mid \beta_k]$ satisfies
-$B \diag(\pi) B^\top = M_2$, then $(B Q) \diag(\pi) (Q B^\top)$
+$B \diag(\pi) B^\top = M_2$ and $\pi$ is uniform, then $(B Q) \diag(\pi) (Q^\top B^\top) = M_2$
 for any orthogonal matrix $Q$.
@@
 %%% third moment
-Let us look to the third moment for additional information.
+Let us now look to the third moment for additional information.
 We can write $y^3$ as a linear function of $x\tp{3}$ with coefficients $M_3$,
-a bias $3 \sigma^2 \innerp{M_1}{x}$ that can be estimated from \refeqn{y1}, and zero mean noise $\eta_3(x)$:
+a bias $3 \E[\epsilon^2] \innerp{M_1}{x} + \E[\epsilon^3]$ that can be estimated from \refeqn{y1}, and zero mean noise $\eta_3(x)$:
 \begin{align}
-y^3 &= \innerp{M_3}{x\tp{3}} + 3\sigma^2 \innerp{M_1}{x} + \eta_3(x), \label{eqn:y3} \\
+  y^3 &= \innerp{M_3}{x\tp{3}} + 3\E[\epsilon^2] \innerp{M_1}{x} + \E[\epsilon^3] + \eta_3(x), \nonumber \\
 \eta_3(x) &= \innerp{\beta_h\tp{3} - M_3}{x\tp{3}}
-+ 3 \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}} \nonumber \\
-&\quad + 3(\epsilon^2 \innerp{\beta_h}{x} - \sigma^2 \innerp{M_1}{x})
-+ \epsilon^3. \nonumber
++ 3 \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}} \label{eqn:y3} \\
+&\quad + 3(\epsilon^2 \innerp{\beta_h}{x} - \E[\epsilon^2] \innerp{M_1}{x})
++ (\epsilon^3 - \E[\epsilon^3]). \nonumber
 \end{align}
-Performing regression yields estimates for $M_3$.
-It turns out that knowledge of $M_2$ and $M_3$ are sufficient to recover
+Performing this regression yields an estimate of $M_3$.
+We will see shortly that knowledge of $M_2$ and $M_3$ are sufficient to recover
 all the parameters.
@@
     regularization strengths $\lambda_n^{(2)}$, $\lambda_n^{(3)}$;
-    observation noise variance $\sigma^2$.
+    observation noise variance $\E[\epsilon^2]$.
     \OUTPUT Parameters $\hat\theta = (\hat \pi, [\hat \beta_1 \mid \cdots \mid \hat \beta_k])$.
@@
       &\hat M_2 = \arg\min_{M_2} \quad \lambda_n^{(2)} \|M_2\|_* + \label{eqn:estimateM2} \\
-      &\quad\quad\frac{1}{2n} \sum_{(x,y) \in \sD_2} (\innerp{M_2}{x\tp{2}} + \sigma^2 - y^2)^2, \nonumber \\
+      &\quad\quad\frac{1}{2n} \sum_{(x,y) \in \sD_2} (\innerp{M_2}{x\tp{2}} + \E[\epsilon^2] - y^2)^2, \nonumber \\
       &\hat M_3 = \arg\min_{M_3} \quad \lambda_n^{(3)} \|M_3\|_* + \label{eqn:estimateM3} \\
-      &\quad\quad\frac{1}{2n} \sum_{(x,y) \in \sD_3} (\innerp{M_3}{x\tp{3}} + 3 \sigma^2\innerp{\hat M_1}{x} - y^3)^2. \nonumber
+      &\quad\quad\frac{1}{2n} \sum_{(x,y) \in \sD_3} (\innerp{M_3}{x\tp{3}} + 3 \E[\epsilon^2]\innerp{\hat M_1}{x} - y^3)^2. \nonumber
     \end{align}
@@
 (\algorithmref{algo:spectral-experts}).
-First, we perform three regressions to recover the compound parameters
+First, we perform three regressions to recover the \emph{compound parameters}
 $M_1$ \refeqn{y1},
@@
 are set to $\frac{c}{\sqrt{n}}$ for some constant $c$.
-The resulting semidefinite program is a standard one which has received
-much attention in recent years.
+%The resulting semidefinite program is a standard one which has received
+%much attention in recent years.
 % We use a standard off-the-shelf SDP solver, CVX, to solve .

### 2013-05-08 doc/icml2013/model.tex
algorithm updated

@@
 Let $k$ be the number of mixture components.
-The generation of $y$ given $x$ involves two steps:
+The generation of $y$ given $x$ involves three steps:
 (i) draw a mixture component $h \in [k]$ according to mixture proportions
 $\pi = (\pi_1, \dots, \pi_k)$;
-and (ii) set $y$ to $\beta_h^\top x$ plus some observation noise $\epsilon$.
+(ii) draw observation noise $\epsilon$ from a known zero-mean noise distribution $\sE$,
+and (ii) set $y$ deterministically based on $h$ and $\epsilon$.
+More compactly: %to $\beta_h^\top x$ plus some observation noise $\epsilon$.
 \begin{eqnarray}
   h &\sim& \mult(\pi), \\
-  y \mid h &=& \beta_{h}^T x + \epsilon.
+  \epsilon &\sim& \sE, \\
+  y &=& \beta_{h}^T x + \epsilon.
 \end{eqnarray}
@@
 are the regression coefficients.
-Note that the choice of mixture component $h$ and the noise $\epsilon$ are independent.
-We also assume the distribution of the observation noise is known and has bounded support.
+Note that the choice of mixture component $h$ and the observation noise $\epsilon$ are independent.
+%We also assume the distribution of the observation noise is known and has bounded support.
 The learning problem is stated as follows:

### 2013-05-08 doc/icml2013/algo.tex
related work

@@
   \begin{algorithmic}[1]
-    \INPUT Three datasets $\mathcal{D}_r = \{ (\xn{1}, \yn{1}), \cdots, (\xn{n}, \yn{n}) \}$ for $r = 1, 2, 3$;
+    \INPUT Datasets $\mathcal{D}_p = \{ (\xn{1}, \yn{1}), \cdots, (\xn{n}, \yn{n}) \}$ for $p = 1, 2, 3$;
     %regularization strengths $\lambda_n^{(2)} = \frac{c_2}{\sqrt{n}}$, $\lambda_n^{(3)} = \frac{c_3}{\sqrt{n}}$;
     regularization strengths $\lambda_n^{(2)}$, $\lambda_n^{(3)}$;
-    observation noise variance $\E[\epsilon^2]$.
+    observation noise moments $\E[\epsilon^2], \E[\epsilon^3]$.
     \OUTPUT Parameters $\hat\theta = (\hat \pi, [\hat \beta_1 \mid \cdots \mid \hat \beta_k])$.
@@
       &\hat M_1 = \arg\min_{M_1} \label{eqn:estimateM1} \\
-      &\quad\quad\frac{1}{2n} \sum_{(x,y) \in \sD_1} (\innerp{M_1}{x} - y)^2, \nonumber \\
+      &\quad\sum_{(x,y) \in \sD_1} (\innerp{M_1}{x} - y)^2, \nonumber \\
       &\hat M_2 = \arg\min_{M_2} \quad \lambda_n^{(2)} \|M_2\|_* + \label{eqn:estimateM2} \\
-      &\quad\quad\frac{1}{2n} \sum_{(x,y) \in \sD_2} (\innerp{M_2}{x\tp{2}} + \E[\epsilon^2] - y^2)^2, \nonumber \\
+      &\quad\sum_{(x,y) \in \sD_2} (\innerp{M_2}{x\tp{2}} + \E[\epsilon^2] - y^2)^2, \nonumber \\
       &\hat M_3 = \arg\min_{M_3} \quad \lambda_n^{(3)} \|M_3\|_* + \label{eqn:estimateM3} \\
-      &\quad\quad\frac{1}{2n} \sum_{(x,y) \in \sD_3} (\innerp{M_3}{x\tp{3}} + 3 \E[\epsilon^2]\innerp{\hat M_1}{x} - y^3)^2. \nonumber
+      &\quad\sum_{(x,y) \in \sD_3} (\innerp{M_3}{x\tp{3}} + 3 \E[\epsilon^2]\innerp{\hat M_1}{x} + \E[\epsilon^3] - y^3)^2. \nonumber
     \end{align}
-    \STATE Estimate the parameters $\theta = (\pi, B)$ using \textbf{tensor factorization}:
+    \STATE Estimate parameters $\theta = (\pi, B)$ using \textbf{tensor factorization}:
     \begin{enumerate}
@@
       \hat M_2 \hat W = I$) using SVD.
-      \item [(b)] Compute the eigenvalues $\{\hat a_h\}_{h=1}^k$
+      \item [(b)] Compute eigenvalues $\{\hat a_h\}_{h=1}^k$
       and eigenvectors $\{\hat v_h\}_{h=1}^k$
@@
       by using the robust tensor power method.
-    \item [(c)] Set parameter estimates $\hat\pi_h = \frac{1}{\hat a_h^2}$
+    \item [(c)] Return parameter estimates $\hat\pi_h = \hat a_h^{-2}$
     and $\hat\beta_h = (\hat W^{\top})^\dagger (\hat a_h \hat v_h)$.
@@
 
-\paragraph{Remarks}
+\paragraph{Related work}
 
 % Spectral
-In recent years, there has a been substantial interest in ``spectral'' methods
+In recent years, there has a been a surge of interest in ``spectral'' methods
 for learning latent-variable models.  One line of work has
-focused on observable operator models \cite{hsu09spectral}
-in which the true parameters are not recovered, but a reparametrization is used
-which enables prediction.
+focused on observable operator models \cite{hsu09spectral,song10kernel,parikh12spectral,cohen12pcfg,balle11transducer,balle12automata}
+in which a reparametrization of the true parameters are recovered,
+which suffices for prediction and density estimation.
 Another line of work is based on the method of moments and uses eigendecomposition of a certain tensor
-to recover the parameters \cite{anandkumar12lda,anandkumar12moments}.
-Our work extends this line of work into the space of discriminative models,
-which requires regression to obtain the desired tensor.
+to recover the parameters \cite{anandkumar12moments,anandkumar12lda,hsu12identifiability,hsu13spherical}.
+Our work extends this second line of work to models that
+require regression to obtain the desired tensor.
 
@@
 the compound parameters.  ``Unmixing'' involves solving for the compound
-parameters by inverting a mixing matrix.  In this work,
+parameters by inverting a mixing matrix.
+In this work,
 each data point (appropriately transformed) provides a different noisy projection of
 the compound parameters.
+
+% 
+%Previous work has focused on using spectral methods to learn finite state automata and transducers.
+Other work has focused on learning discriminative models,
+notably \citet{balle11transducer} for finite state transducers (functions from strings to strings),
+and \citet{balle12automata} for weighted finite state automata (functions from strings to real numbers).
+Similar to Spectral Experts,
+\citet{balle12automata} used a two-step approach,
+where convex optimization is first used to estimate moments (the Hankel matrix in their case),
+after which these moments are subjected to spectral decomposition.  
+However, these methods are developed in the observable operator framework, whereas we consider parameter estimation.
 

### 2013-05-08 doc/icml2013/theory.tex

@@
 of linear regressions model with parameter $\theta^*$\footnote{Having three independent copies simplifies the analysis}.
-Further, assume $\|x\|_2 \le R$ with probability $1$, $\|\beta_h^*\|_2
-\le L$ for all $h \in [k]$ and $B$ is full rank.
+Further, assume 
+$\|x\|_2 \le R$ with probability $1$, 
+$\|\beta_h^*\|_2 \le L$ for all $h \in [k]$,
+$|\epsilon| \le S$
+and $B$ is full rank.
 Let $\Sigma_p \eqdef \E[\cvec(x\tp{p})\tp{2}]$, 
@@
 $n = \max(n_1,n_2)$
-where $n_1 = \Omega\left(R^{12} (\max_{p \in [3]} \sigmamin(\Sigma_p)^{-2}) \log(1/\delta)\right)$ and
-$n_2 = \Omega\left( \sigma^6 L^6 R^{12} k (\max_{p \in [3]} \sigmamin(\Sigma_p)^{-1}) \frac{\log^3 (1/\delta)}{\epsilon^2} \right)$.
+where 
+%\begin{align*}
+$$n_1 = \Omega \left(\max_{p \in [3]} \frac{R^{4p} (p!)^2 \log(1/\delta)}{\sigmamin(\Sigma_p)^2} \right),$$
+$$n_2 = \Omega \left(\max_{p\in [3]} \frac{L^{2p} S^{2p} R^{4p} \log(1/\delta)}{\left(\lambda_n^{(p)}\right)^2} \right).$$
+%\end{align*}
+Let $\epsilon$ be 
+$$O\left( \min\left\{ 
+    \frac{\sigma_k(M_2)}{2}, 
+    \frac{\sigma_k(M_2)^{3/2}}{k \pi_{\max} \|M_3\|_\op},
+    \frac{\sigma_k(M_2)^{5/2}}{k \pi_{\max} \|M_3\|_\op} 
+    \right\}\right).$$ 
 If the regularization strengths $\lambda_n^{(2)}$ and $\lambda_n^{(3)}$ are
-set to $\Omega\left(\sigma^3 L^3 R^6 \sqrt{\frac{\log(1/\delta)}{n}}\right)$,
+set to $O\left(\frac{\sigmamin(\Sigma_p)}{\sqrt{k}}~ \epsilon \right)$, 
+%$\Omega\left(\sigma^3 L^3 R^6 \sqrt{\frac{\log(1/\delta)}{n}}\right)$,
 then the parameter estimates $\hat\theta = (\hat\pi, \hat B)$ returned by
-\algorithmref{algo:spectral-experts} (with the columns appropriate permuted)
+\algorithmref{algo:spectral-experts} (with the columns appropriately permuted)
 satisfies 
-$\|\hat\pi - \pi^*\|_{\infty} \le \epsilon$
-and for all $h \in [k]$,
-$\|\hat\beta_h - \beta^*_h\|_2 \le \frac{\epsilon}{\sqrt{\pi_h^*}}$.
+  \begin{align*}
+  \|\hat \pi - \pi \|_{\infty}
+  &= O\left(\frac{k \pi_1^{5/2}\| {M_3} \|_\op}{\min\{ \sigma_k(M_2)^{3/2}, \sigma_k(M_2)^{5/2} \}} ~ \epsilon \right) \\
+  \|\hat \beta_h - \beta_h\|_2
+  &= O\left( \frac{k \pi_1 \|M_2\|_\op^{1/2} \| {M_3} \|_\op}{\min\{ \sigma_k(M_2), \sigma_k(M_2)^{5/2} \}}~ \epsilon \right),
+  \end{align*}
+  for all $h \in [k]$.
+%$\|\hat\pi - \pi^*\|_{\infty} \le \epsilon$
+%and for all $h \in [k]$,
+%$\|\hat\beta_h - \beta^*_h\|_2 \le \frac{\epsilon}{\sqrt{\pi_h^*}}$.
 \end{theorem}
@@
 Suppose there exists a restricted strong convexity constant $\kappa(\opX_p)$ such that
-$$\frac{1}{2n} \| \opX_p( \Delta )\|_2^2 \ge \kappa(\opX_p) \|\Delta\|^2_F \quad \text{and} \quad
-\lambda_n \ge \frac{\|\opX_p^*(\eta_p)\|_\op}{n}.$$
+$$\frac{1}{n} \| \opX_p( \Delta )\|_2^2 \ge \kappa(\opX_p) \|\Delta\|^2_F \quad \text{and} \quad
+\lambda^{(p)}_n \ge \frac{\|\opX_p^*(\eta_p)\|_\op}{n}.$$
 Then the error of $\hat M_p$ is bounded as follows:
-$\| \hat M_p - M_p^* \|_F \le \frac{\lambda_n \sqrt{k}}{\kappa(\opX_p)}$.
+$\| \hat M_p - M_p^* \|_F \le \frac{\lambda^{(p)}_n \sqrt{k}}{\kappa(\opX_p)}$.
 \end{lemma}
@@
 Let $\Sigma_p \eqdef \E[\cvec(x\tp{p})\tp{2}]$.
-If $$n = \Omega \left( \frac{R^{4p} \log(1/\delta)}{\sigmamin(\Sigma_p)^2} \right),$$
+If $$n_1 = \Omega \left(\frac{R^{4p} (p!)^2 \log(1/\delta)}{\sigmamin(\Sigma_p)^2} \right),$$
 with probability at least $1-\delta$,
-$$\kappa(\opX_p) \ge \frac{\sigmamin(\Sigma_p)}{4}.$$
+$$\kappa(\opX_p) \ge \frac{\sigmamin(\Sigma_p)}{2}.$$
 \end{lemma}
@@
 \label{lem:lowRankUpper}
-Let $\opX_p$ be the linear operator previously defined. Then with
-probability at least $1-\delta$,
-$$\frac1{n} \|\opX_p^*(\eta_p)\|_\op \le O\left(\sigma^3 L^3 R^6 \sqrt{\frac{\log^3(1/\delta)}{n}}\right).$$
+% Let $\opX_p$ be the linear operator previously defined. 
+If $$n_2 = \Omega \left(\max_{p\in [3]} \frac{L^{2p} S^{2p} R^{4p} \log(1/\delta)}{\left(\lambda_n^{(p)}\right)^2} \right),$$
+then with probability at least $1-\delta$,
+$$\lambda_n^{(p)} \ge \frac1{n} \|\opX_p^*(\eta_p)\|_\op,$$
 for each $p \in \{1,2,3\}$.
@@
   \label{lem:tensorPower} Let $\|\hat M_2 - M_2\|_\op \le \epsilon$ and
-  $\|\hat M_3 - M_3\|_\op \le \epsilon$, for some $\epsilon$ such that 
-  $$\epsilon < \min\{\sigmamin(M_2)/2,O(k \sigmamin(M_2)^{1.5} \|M_3\|_\op^{-1})\}.$$ 
+  $\|\hat M_3 - M_3\|_\op \le \epsilon$, for some $\epsilon$ in
+  $$O\left( \min\left\{ 
+    \frac{\sigma_k(M_2)}{2}, 
+    \frac{\sigma_k(M_2)^{3/2}}{k \pi_{\max} \|M_3\|_\op},
+    \frac{\sigma_k(M_2)^{5/2}}{k \pi_{\max} \|M_3\|_\op} 
+    \right\}\right).$$ 
   Then, there exists a permutation of indices such that  the parameter
@@
   \|\hat \pi - \pi \|_{\infty}
-    &= O( k \pi_{max}^{2.5} \pi_{min}^{-0.5} \sigmamin(M_2)^{-1.5} \|M_3\|_\op \epsilon ) \\
+  &= O\left(\frac{k \pi_1^{5/2}\| {M_3} \|_\op}{\min\{ \sigma_k(M_2)^{3/2}, \sigma_k(M_2)^{5/2} \}} ~ \epsilon \right) \\
   \|\hat \beta_h - \beta_h\|_2
-    &= O( k \pi_{max} \sigmamin(M_2)^{-1.5} \|M_2\|_\op^{0.5} \|M_3\|_\op \epsilon ).
+  &= O\left( \frac{k \pi_1 \|M_2\|_\op^{1/2} \| {M_3} \|_\op}{\min\{ \sigma_k(M_2), \sigma_k(M_2)^{5/2} \}}~ \epsilon \right),
   \end{align*}
@@
 \begin{align}
-  \|\hat M_p - M_p\|_\op = O\left( \sigma^3 L^3 R^6 k^{\frac12} \sigmamin(\Sigma_p)^{-1} \sqrt{\frac{\log^3 (1/\delta)}{n}} \right).
+  \|\hat M_p - M_p\|_\op = O\left( \lambda_n^{(p)} \sqrt{k} \sigmamin(\Sigma_p)^{-1} \right).
+  %\|\hat M_p - M_p\|_\op = O\left( \sigma^3 L^3 R^6 k^{\frac12} \sigmamin(\Sigma_p)^{-1} \sqrt{\frac{\log^3 (1/\delta)}{n}} \right).
   %\|\hat M_p - M_p\|_F = O\left( \frac{\sigma^3 L^3 R^6 \sqrt{k} \log^3 (1/\delta)}{\sqrt{n} (\sigmamin(\Sigma_p) - d^p R^p \sqrt{\frac{p \log(d) \log(1/\delta)}{n}}} \right).

### 2013-05-08 doc/icml2013/algo.tex
fix algorithm pseudocode, some random changes

@@
       &\hat M_1 = \arg\min_{M_1} \label{eqn:estimateM1} \\
-      &\quad\sum_{(x,y) \in \sD_1} (\innerp{M_1}{x} - y)^2, \nonumber \\
+      &\quad\frac{1}{2n}\sum_{(x,y) \in \sD_1} (\innerp{M_1}{x} - y)^2, \nonumber \\
       &\hat M_2 = \arg\min_{M_2} \quad \lambda_n^{(2)} \|M_2\|_* + \label{eqn:estimateM2} \\
-      &\quad\sum_{(x,y) \in \sD_2} (\innerp{M_2}{x\tp{2}} + \E[\epsilon^2] - y^2)^2, \nonumber \\
+      &\quad\frac{1}{2n}\sum_{(x,y) \in \sD_2} (\innerp{M_2}{x\tp{2}} + \E[\epsilon^2] - y^2)^2, \nonumber \\
       &\hat M_3 = \arg\min_{M_3} \quad \lambda_n^{(3)} \|M_3\|_* + \label{eqn:estimateM3} \\
-      &\quad\sum_{(x,y) \in \sD_3} (\innerp{M_3}{x\tp{3}} + 3 \E[\epsilon^2]\innerp{\hat M_1}{x} + \E[\epsilon^3] - y^3)^2. \nonumber
+      &\quad\frac{1}{2n}\sum_{(x,y) \in \sD_3} (\innerp{M_3}{x\tp{3}} + 3 \E[\epsilon^2]\innerp{\hat M_1}{x} + \E[\epsilon^3] - y^3)^2. \nonumber
     \end{align}

### 2013-05-08 doc/icml2013/theory.tex
fix algorithm pseudocode, some random changes

@@
 
-In this section, we provide theoretical guarantees for our Spectral Experts algorithm.
+In this section, we provide theoretical guarantees for the Spectral Experts algorithm.
+Our main result shows that our parameter estimates $\hat\theta$ converge to $\theta$
+
 In particular, the following theorem provides a rate of convergence:
 
-\begin{theorem}[Convergence of Spectral Experts]
+\begin{theorem}[convergence of Spectral Experts]
 \label{thm:convergence}
 Assume each dataset $\sD_p$ (for $p = 1, 2, 3$) consists of $n$ i.i.d.\ points independently drawn from a mixture
-of linear regressions model with parameter $\theta^*$\footnote{Having three independent copies simplifies the analysis}.
+of linear regressions model with parameter $\theta^*$.\footnote{Having three independent copies simplifies the analysis.}
 Further, assume 
@@
 $|\epsilon| \le S$
-and $B$ is full rank.
+and $B$ is rank $k$.
 Let $\Sigma_p \eqdef \E[\cvec(x\tp{p})\tp{2}]$, 
@@
 and $\E[x\tp{2p}]$ for $p \in \{1,2,3\}$, we can then identify the
-parameters $\theta = (\pi, B)$, provided $\E[\cvec(x\tp{p})\tp{2}] \succ
+parameters $\theta = (\pi, B)$,
+provided the following \emph{identifiability condition} holds: $\E[\cvec(x\tp{p})\tp{2}] \succ
 0$ for $p \in \{1,2,3\}$.
 
-While this is not a concern for independent dimensions, we note that
-this requires more stringent positive definite constraints when using
-features. For example, suppose $x = (1, t, t^2)$, the common polynomial
+This identifiability condition warrants a little care,
+as we can run into trouble when components of $x$ are dependent on each other
+in a particular algebraic way.
+For example, suppose $x = (1, t, t^2)$, the common polynomial
 basis expansion, so that all the coordinates are deterministically
-related.  Then $\E[x\tp{2}] \succ 0$ can easily be satisfied. 
-%, say with $x \sim \normal{0}{I}$.
-However, one can verify that $\E[\cvec(x\tp{2})\tp{2}]$ is singular for
-any data distribution, since $\cvec(x\tp{2}) = [1 \cdot 1, t\cdot t, 2(1
+related.  While $\E[x\tp{2}] \succ 0$ might be satisfied (sufficient for ordinary linear regression),
+$\E[\cvec(x\tp{2})\tp{2}]$ is singular for
+any data distribution.
+To see this, note that $\cvec(x\tp{2}) = [1 \cdot 1, t\cdot t, 2(1
 \cdot t^2), 2(t \cdot t^2), (t^2 \cdot t^2)]$ contains components $t
@@
 Spectral Experts would not be able to identify the parameters of
-a mixture of linear regressions for this distribution over $x$.
+a mixture of linear regressions for this data distribution.
 
-We now show that some amount of unidentifiability is intrinsic to
+We can show that some amount of unidentifiability is intrinsic to
 estimation from low-order moments, not just an artifact of our
@@
 observed all moments $\E[x\tp{p}y\tp{p}]$ and $\E[x\tp{2p}]$ for $p \in
-[r]$, all the resulting coordinates would be monomials up to degree
+[r]$ for some $r$, all the resulting coordinates would be monomials of $t$ up to only degree
 $2dr$, and thus the moments live in a $2dr$-dimensional subspace.  On
@@
 
-\paragraph{Low-rank regression}
-%\label{sec:regression}
-\vspace{-0.5em}
-We now bound the error of
-the compound parameter estimates,
-$\|\Delta_2\|_F^2$ and $\|\Delta_3\|_F^2$,
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Analysis of low-rank regression}
+\label{sec:regression}
+
+In this section, we will bound the error of
+the compound parameter estimates $\|\Delta_2\|_F^2$ and $\|\Delta_3\|_F^2$,
 where $\Delta_2 \eqdef \hat M_2 - M_2$
 and $\Delta_3 \eqdef \hat M_3 - M_3$.
+Our analysis is based on the low-rank regression framework of
+\citet{Tomioka2011} for tensors, which builds on
+\citet{NegahbanWainwright2009} for matrices.
+The main calculation involved is controling the noise $\eta_p(x)$,
+which involves various polynomial combinations of the mixing noise and observation noise.
 
-Our analysis is based on the low-rank regression framework of
-\citet{Tomioka2011} for tensors, which builds off of
-\citet{NegahbanWainwright2009} for matrices.
-The main challenge is to control the noise $\eta_p(x)$,
-which in addition to containing the usual Gaussian part, is also dependent on
-$x$, involves the mixing noise and various cross terms.
-We first set up some notation that unifies all three regressions (\refeqn{estimateM1}, \refeqn{estimateM2}, and \refeqn{estimateM3}).
+Let us first establish some notation that unifies the three regressions (\refeqn{estimateM1}, \refeqn{estimateM2}, and \refeqn{estimateM3}).
 Define the observation operator $\opX_p(M_p) : \Re^{d\tp{p}} \to \Re^{n}$
@@
 $$\frac{1}{n} \| \opX_p( \Delta )\|_2^2 \ge \kappa(\opX_p) \|\Delta\|^2_F \quad \text{and} \quad
-\lambda^{(p)}_n \ge \frac{\|\opX_p^*(\eta_p)\|_\op}{n}.$$
+\lambda^{(p)}_n \ge \frac{2 \|\opX_p^*(\eta_p)\|_\op}{n}.$$
 Then the error of $\hat M_p$ is bounded as follows:
-$\| \hat M_p - M_p^* \|_F \le \frac{\lambda^{(p)}_n \sqrt{k}}{\kappa(\opX_p)}$.
+$$\| \hat M_p - M_p \|_F \le \frac{32 \lambda^{(p)}_n \sqrt{k}}{\kappa(\opX_p)}.$$
 \end{lemma}
@@
 Going forward, we need to lower bound the restricted strong convexity
-parameter $\kappa(\opX_p)$ and upper bound the adjoint operator
+constant $\kappa(\opX_p)$ and upper bound the operator norm of the adjoint operator
 $\|\opX_p^*(\eta_p)\|_\op$. The proofs of the following lemmas follow
@@
 
-\begin{lemma}[lower bound on restricted strong convexity]
+\begin{lemma}[lower bound on restricted strong convexity constant]
 \label{lem:lowRankLower}
-Let $\Sigma_p \eqdef \E[\cvec(x\tp{p})\tp{2}]$.
-If $$n_1 = \Omega \left(\frac{R^{4p} (p!)^2 \log(1/\delta)}{\sigmamin(\Sigma_p)^2} \right),$$
-with probability at least $1-\delta$,
+%Let $\Sigma_p \eqdef \E[\cvec(x\tp{p})\tp{2}]$.
+If $$n = \Omega \left(\frac{R^{4p} (p!)^2 \log(1/\delta)}{\sigmamin(\Sigma_p)^2} \right),$$
+then with probability at least $1-\delta$:
 $$\kappa(\opX_p) \ge \frac{\sigmamin(\Sigma_p)}{2}.$$
@@
 
-\begin{lemma}[upper bound on adjoint noise]
+\begin{lemma}[upper bound on adjoint operator]
 \label{lem:lowRankUpper}
 % Let $\opX_p$ be the linear operator previously defined. 
-If $$n_2 = \Omega \left(\max_{p\in [3]} \frac{L^{2p} S^{2p} R^{4p} \log(1/\delta)}{\left(\lambda_n^{(p)}\right)^2} \right),$$
-then with probability at least $1-\delta$,
+If $$n = \Omega \left(\max_{p\in [3]} \frac{L^{2p} S^{2p} R^{4p} \log(1/\delta)}{\left(\lambda_n^{(p)}\right)^2} \right),$$
+then with probability at least $1-\delta$:
 $$\lambda_n^{(p)} \ge \frac1{n} \|\opX_p^*(\eta_p)\|_\op,$$
@@
 
-\paragraph{Method of Moments}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Analysis of the tensor factorization} 
+\label{sec:tensorError}
 
-Now that we have bounds on the errors of recovering $M_2$ and $M_3$ from
-the data, we will extend the analysis of the robust tensor method in
-\citet{AnandkumarGeHsu2012} to include the whitening and unwhitening
-steps, 2(a) and 2(c) respectively, of
-\algorithmref{algo:spectral-experts}.
+Having bounded the error of the compound parameter estimates $\hat M_2$ and $\hat M_3$,
+we will now study how this error propagates through the tensor factorization step of 
+\algorithmref{algo:spectral-experts},
+which includes whitening, applying the robust tensor power method \cite{AnandkumarGeHsu2012},
+and unwhitening.
 \begin{lemma}
@@
   estimates found in step 2 of \algorithmref{algo:spectral-experts}
-  satisfy the following with probability at least $1 - \delta$,
+  satisfy the following with probability at least $1 - \delta$:
   \begin{align*}
@@
 
-\paragraph{Synthesis}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Synthesis}
 Together, these lemmas allow us to control the compound parameter error

### 2013-05-08 doc/icml2013/theory.tex
Minor changes + experiments

@@
 %Let $\Sigma_p \eqdef \E[\cvec(x\tp{p})\tp{2}]$.
-If $$n = \Omega \left(\frac{R^{4p} (p!)^2 \log(1/\delta)}{\sigmamin(\Sigma_p)^2} \right),$$
+If $$n = \Omega \left(\max_{p\in [3]} \frac{R^{4p} (p!)^2 \log(1/\delta)}{\sigmamin(\Sigma_p)^2} \right),$$
 then with probability at least $1-\delta$:
-$$\kappa(\opX_p) \ge \frac{\sigmamin(\Sigma_p)}{2}.$$
+$$\kappa(\opX_p) \ge \frac{\sigmamin(\Sigma_p)}{2},$$
+for each $p \in [3]$.
 \end{lemma}
@@
 $$\lambda_n^{(p)} \ge \frac1{n} \|\opX_p^*(\eta_p)\|_\op,$$
-for each $p \in \{1,2,3\}$.
+for each $p \in [3]$.
 \end{lemma}

### 2013-05-08 doc/icml2013/algo.tex
bias

@@
 We can write $y^3$ as a linear function of $x\tp{3}$ with coefficients $M_3$,
-a bias $3 \E[\epsilon^2] \innerp{M_1}{x} + \E[\epsilon^3]$ that can be estimated from \refeqn{y1}, and zero mean noise $\eta_3(x)$:
+a known bias $3 \E[\epsilon^2] \innerp{\hat M_1}{x} + \E[\epsilon^3]$ and some noise $\eta_3(x)$:
 \begin{align}
-  y^3 &= \innerp{M_3}{x\tp{3}} + 3\E[\epsilon^2] \innerp{M_1}{x} + \E[\epsilon^3] + \eta_3(x), \nonumber \\
+  y^3 &= \innerp{M_3}{x\tp{3}} + 3\E[\epsilon^2] \innerp{\hat M_1}{x} + \E[\epsilon^3] + \eta_3(x), \nonumber \\
 \eta_3(x) &= \innerp{\beta_h\tp{3} - M_3}{x\tp{3}}
 + 3 \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}} \label{eqn:y3} \\
-&\quad + 3(\epsilon^2 \innerp{\beta_h}{x} - \E[\epsilon^2] \innerp{M_1}{x})
+&\quad + 3(\epsilon^2 \innerp{\beta_h}{x} - \E[\epsilon^2] \innerp{\hat M_1}{x})
 + (\epsilon^3 - \E[\epsilon^3]). \nonumber
 \end{align}
+The only wrinkle here is that $\eta_3(x)$ does not quite have zero mean.
+It would if $\hat M_1$ were replaced with $M_1$, but $M_1$ is not available to us.
+Nonetheless, as $\hat M_1$ concentrates around $M_1$, the noise bias will go to zero.
 Performing this regression yields an estimate of $M_3$.

### 2013-05-08 doc/icml2013/theory.tex
Changed theorem 1 to directly bound the parameters.

@@
 
-\begin{theorem}[convergence of Spectral Experts]
+\begin{theorem}[Convergence of Spectral Experts]
 \label{thm:convergence}
@@
 and assume $\Sigma_p \succ 0$ for each $p \in \{1,2,3\}$.
+Let $\epsilon$ be such that, 
+$$\epsilon = O\left( \min\left\{ \frac{k\|M_3\|_\op \pi_{\max}}{\sigma_k(M_2)^{3/2}}, \half \right\}\right).$$ 
 Suppose the number of samples is
@@
 where 
-%\begin{align*}
-$$n_1 = \Omega \left(\max_{p \in [3]} \frac{R^{4p} (p!)^2 \log(1/\delta)}{\sigmamin(\Sigma_p)^2} \right),$$
-$$n_2 = \Omega \left(\max_{p\in [3]} \frac{L^{2p} S^{2p} R^{4p} \log(1/\delta)}{\left(\lambda_n^{(p)}\right)^2} \right).$$
-%\end{align*}
-Let $\epsilon$ be 
-$$O\left( \min\left\{ 
-    \frac{\sigma_k(M_2)}{2}, 
-    \frac{\sigma_k(M_2)^{3/2}}{k \pi_{\max} \|M_3\|_\op},
-    \frac{\sigma_k(M_2)^{5/2}}{k \pi_{\max} \|M_3\|_\op} 
-    \right\}\right).$$ 
-If the regularization strengths $\lambda_n^{(2)}$ and $\lambda_n^{(3)}$ are
-set to $O\left(\frac{\sigmamin(\Sigma_p)}{\sqrt{k}}~ \epsilon \right)$, 
+\begin{align*}
+n_1 &= \Omega \left(\frac{R^{12} \log(1/\delta)}{\min_{p \in [3]} \sigmamin(\Sigma_p)^2} \right) \\
+n_2 &= \Omega \left(\epsilon^{-2}~ \frac{k^2 \pi^2_{\max} \|M_3\|_\op^2}{\sigma_k(M_2)^{5}} L^{6} S^{6} R^{12} \log(1/\delta) \right).
+\end{align*}
+If the regularization strengths $\lambda_n^{(p)}$ are
+set to 
+$$\Theta\left( L^p S^p R^{2p} \sqrt{\frac{\log(1/\delta)}{n}} \right),$$
+for $p \in 2, 3$,
+%$O\left(\frac{\sigmamin(\Sigma_p)}{\sqrt{k}}~ \epsilon \right)$, 
 %$\Omega\left(\sigma^3 L^3 R^6 \sqrt{\frac{\log(1/\delta)}{n}}\right)$,
@@
   \begin{align*}
-  \|\hat \pi - \pi \|_{\infty}
-  &= O\left(\frac{k \pi_1^{5/2}\| {M_3} \|_\op}{\min\{ \sigma_k(M_2)^{3/2}, \sigma_k(M_2)^{5/2} \}} ~ \epsilon \right) \\
-  \|\hat \beta_h - \beta_h\|_2
-  &= O\left( \frac{k \pi_1 \|M_2\|_\op^{1/2} \| {M_3} \|_\op}{\min\{ \sigma_k(M_2), \sigma_k(M_2)^{5/2} \}}~ \epsilon \right),
+  \|\hat \pi - \pi \|_{\infty} &\le \epsilon \\
+  %&= O\left(\frac{k \pi_{\max}^{5/2}\| {M_3} \|_\op}{\sigma_k(M_2)^{5/2}} ~ \epsilon \right) \\
+  \|\hat \beta_h - \beta_h\|_2 &\le \epsilon \\
+  %&= O\left( \frac{k \pi_{\max} \|M_2\|_\op^{1/2} \| {M_3} \|_\op}{\sigma_k(M_2)^{5/2}}~ \epsilon \right),
   \end{align*}
@@
     \frac{\sigma_k(M_2)}{2}, 
-    \frac{\sigma_k(M_2)^{3/2}}{k \pi_{\max} \|M_3\|_\op},
     \frac{\sigma_k(M_2)^{5/2}}{k \pi_{\max} \|M_3\|_\op} 
@@
   \|\hat \pi - \pi \|_{\infty}
-  &= O\left(\frac{k \pi_1^{5/2}\| {M_3} \|_\op}{\min\{ \sigma_k(M_2)^{3/2}, \sigma_k(M_2)^{5/2} \}} ~ \epsilon \right) \\
+  &= O\left(\frac{k \pi_1^{5/2}\| {M_3} \|_\op}{\sigma_k(M_2)^{5/2}} ~ \epsilon \right) \\
   \|\hat \beta_h - \beta_h\|_2
-  &= O\left( \frac{k \pi_1 \|M_2\|_\op^{1/2} \| {M_3} \|_\op}{\min\{ \sigma_k(M_2), \sigma_k(M_2)^{5/2} \}}~ \epsilon \right),
+  &= O\left( \frac{k \pi_1 \|M_2\|_\op^{1/2} \| {M_3} \|_\op}{ \sigma_k(M_2)^{5/2} }~ \epsilon \right),
   \end{align*}

### 2013-05-08 doc/icml2013/algo.tex
Various fixes to make it submittable

@@
       &\hat M_3 = \arg\min_{M_3} \quad \lambda_n^{(3)} \|M_3\|_* + \label{eqn:estimateM3} \\
-      &\quad\frac{1}{2n}\sum_{(x,y) \in \sD_3} (\innerp{M_3}{x\tp{3}} + 3 \E[\epsilon^2]\innerp{\hat M_1}{x} + \E[\epsilon^3] - y^3)^2. \nonumber
+      % NOTE: hspace added to make ICML accept our paper.
+      &{\frac{1}{2n} \hspace{-0.5em}\sum_{(x,y) \in \sD_3} \hspace{-1em}(\innerp{M_3}{x\tp{3}} + 3 \E[\epsilon^2]\innerp{\hat M_1}{x} + \E[\epsilon^3] - y^3)^2}. \nonumber
     \end{align}

### 2013-05-08 doc/icml2013/theory.tex
rewording

@@
 In this section, we provide theoretical guarantees for the Spectral Experts algorithm.
-Our main result shows that our parameter estimates $\hat\theta$ converge to $\theta$
-
-In particular, the following theorem provides a rate of convergence:
+Our main result shows that the parameter estimates $\hat\theta$ converge to $\theta$
+at a $\frac{1}{\sqrt{n}}$ rate that depends polynomially on the bounds on the
+parameters, covariates, and noise, as well the $k$-th smallest singular values
+of the compound parameters and various covariance matrices.
 
@@
 Further, assume 
-$\|x\|_2 \le R$ with probability $1$, 
+$\|x\|_2 \le R$, 
 $\|\beta_h^*\|_2 \le L$ for all $h \in [k]$,
@@
 \end{align*}
-If the regularization strengths $\lambda_n^{(p)}$ are
-set to 
+If each regularization strength $\lambda_n^{(p)}$ is set to 
 $$\Theta\left( L^p S^p R^{2p} \sqrt{\frac{\log(1/\delta)}{n}} \right),$$
@@
   \begin{align*}
-  \|\hat \pi - \pi \|_{\infty} &\le \epsilon \\
+  \|\hat \pi - \pi \|_{\infty} \le \epsilon \quad\quad 
   %&= O\left(\frac{k \pi_{\max}^{5/2}\| {M_3} \|_\op}{\sigma_k(M_2)^{5/2}} ~ \epsilon \right) \\
-  \|\hat \beta_h - \beta_h\|_2 &\le \epsilon \\
+  \|\hat \beta_h - \beta_h\|_2 \le \epsilon
   %&= O\left( \frac{k \pi_{\max} \|M_2\|_\op^{1/2} \| {M_3} \|_\op}{\sigma_k(M_2)^{5/2}}~ \epsilon \right),
@@
 
+While the dependence on some of the norms ($L^6,S^6,R^{12}$) looks formidable,
+it is in some sense unavoidable, since we
+need to perform regression on third-order moments.
+Classically, the number of samples required is squared norm of the covariance matrix,
+which itself is bounded by the squared norm of the data, $R^3$. This
+third-order dependence also shows up in the regularization strengths;
+the cubic terms bound each of $\epsilon^3$,
+$\beta_h^3$ and $\|(x\tp{3})\tp{2}\|_F$ with high probability.
+
 The proof of the theorem has two parts.
@@
 using results from \citet{Tomioka2011}.
-While the $R^{12}$ term looks formidable, it is unavoidable since we
-perform regression on the third moment of the data. Classically, the
-number of samples required is norm squared of the covariance matrix,
-which itself is bounded by the norm squared of the data, $R^3$. This
-third-order dependence also shows up in the rate the regression
-coefficients are annealed; the cubic terms bound each of $\epsilon^3$,
-$\beta_h^3$ and $\|(x\tp{3})\tp{2}\|_F$ with high probability.
-
-Next, we use results from \citet{AnandkumarGeHsu2012} to convert this error
+Then we use results from \citet{AnandkumarGeHsu2012} to convert this error
 into a bound on the actual parameter estimates $\hat\theta = (\hat\pi, \hat B)$
@@
 $\|\opX_p^*(\eta_p)\|_\op$. The proofs of the following lemmas follow
-from standard concentration inequalities and appear in detail in the
+from standard concentration inequalities and are detailed in the
 supplementary material.

### 2013-05-08 doc/icml2013/theory.tex
Various minor fixes

--- 2013-05-08 doc/icml2013/theory.tex
@@
 n_1 &= \Omega \left(\frac{R^{12} \log(1/\delta)}{\min_{p \in [3]} \sigmamin(\Sigma_p)^2} \right) \\
-n_2 &= \Omega \left(\epsilon^{-2}~ \frac{k^2 \pi^2_{\max} \|M_3\|_\op^2}{\sigma_k(M_2)^{5}} L^{6} S^{6} R^{12} \log(1/\delta) \right).
+n_2 &= \Omega \left(\epsilon^{-2}~ \frac{k^2 \pi^2_{\max} \|M_3\|_\op^2}{\sigma_k(M_2)^{5} }\frac{ L^{6} S^{6} R^{12}}{\sigmamin(\Sigma_1)^2} \log(1/\delta) \right).
 \end{align*}
@@
 % Let $\opX_p$ be the linear operator previously defined. 
-If $$n = \Omega \left(\max_{p\in [3]} \frac{L^{2p} S^{2p} R^{4p} \log(1/\delta)}{\left(\lambda_n^{(p)}\right)^2} \right),$$
+If $$n = \Omega \left(\max_{p\in [3]} \frac{L^{2p} S^{2p} R^{4p} \log(1/\delta)}{ \sigmamin(\Sigma_1)^2 \left(\lambda_n^{(p)}\right)^2} \right),$$
 then with probability at least $1-\delta$:
@@
   \|\hat \pi - \pi \|_{\infty}
-  &= O\left(\frac{k \pi_1^{5/2}\| {M_3} \|_\op}{\sigma_k(M_2)^{5/2}} ~ \epsilon \right) \\
+  &= O\left(\frac{k \pi_{\max}^{5/2}\| {M_3} \|_\op}{\sigma_k(M_2)^{5/2}} ~ \epsilon \right) \\
   \|\hat \beta_h - \beta_h\|_2
-  &= O\left( \frac{k \pi_1 \|M_2\|_\op^{1/2} \| {M_3} \|_\op}{ \sigma_k(M_2)^{5/2} }~ \epsilon \right),
+  &= O\left( \frac{k \pi_{\max} \|M_2\|_\op^{1/2} \| {M_3} \|_\op}{ \sigma_k(M_2)^{5/2} }~ \epsilon \right),
   \end{align*}

### 2013-05-08 doc/icml2013/theory.tex
Changed whitening lemma to express parameter error as epsilon. Also minor fixes

@@
 and assume $\Sigma_p \succ 0$ for each $p \in \{1,2,3\}$.
-Let $\epsilon$ be such that, 
-$$\epsilon = O\left( \min\left\{ \frac{k\|M_3\|_\op \pi_{\max}}{\sigma_k(M_2)^{3/2}}, \half \right\}\right).$$ 
+Let $\epsilon < \half$.
 Suppose the number of samples is
@@
 n_1 &= \Omega \left(\frac{R^{12} \log(1/\delta)}{\min_{p \in [3]} \sigmamin(\Sigma_p)^2} \right) \\
-n_2 &= \Omega \left(\epsilon^{-2}~ \frac{k^2 \pi^2_{\max} \|M_3\|_\op^2}{\sigma_k(M_2)^{5} }\frac{ L^{6} S^{6} R^{12}}{\sigmamin(\Sigma_1)^2} \log(1/\delta) \right).
+n_2 &= \Omega \left(\epsilon^{-2}~ \frac{k^2 \pi^2_{\max} \|M_2\|_\op^{1/2} \|M_3\|_\op^2 { L^{6} S^{6} R^{12}}}{\sigma_k(M_2)^{5} {\sigmamin(\Sigma_1)^2}} \log(1/\delta) \right).
 \end{align*}
 If each regularization strength $\lambda_n^{(p)}$ is set to 
-$$\Theta\left( L^p S^p R^{2p} \sqrt{\frac{\log(1/\delta)}{n}} \right),$$
+$$\Theta\left( \frac{L^p S^p R^{2p}}{\sigmamin(\Sigma_1)^2} \sqrt{\frac{\log(1/\delta)}{n}} \right),$$
 for $p \in 2, 3$,
@@
 \begin{lemma}
-  \label{lem:tensorPower} Let $\|\hat M_2 - M_2\|_\op \le \epsilon$ and
-  $\|\hat M_3 - M_3\|_\op \le \epsilon$, for some $\epsilon$ in
-  $$O\left( \min\left\{ 
-    \frac{\sigma_k(M_2)}{2}, 
-    \frac{\sigma_k(M_2)^{5/2}}{k \pi_{\max} \|M_3\|_\op} 
-    \right\}\right).$$ 
+  \label{lem:tensorPower} Let $\|\hat M_2 - M_2\|_\op$ and $\|\hat M_3 - M_3\|_\op$ both be less than
+  \vspace{-0.5em}
+  $$\frac{\sigma_k(M_2)^{5/2}}{k \pi_{\max} \|M_2\|_\op^{1/2} \| {M_3} \|_\op}~ \epsilon,$$
+  for some $\epsilon < \half$. 
   Then, there exists a permutation of indices such that  the parameter
@@
   \begin{align*}
-  \|\hat \pi - \pi \|_{\infty}
-  &= O\left(\frac{k \pi_{\max}^{5/2}\| {M_3} \|_\op}{\sigma_k(M_2)^{5/2}} ~ \epsilon \right) \\
-  \|\hat \beta_h - \beta_h\|_2
-  &= O\left( \frac{k \pi_{\max} \|M_2\|_\op^{1/2} \| {M_3} \|_\op}{ \sigma_k(M_2)^{5/2} }~ \epsilon \right),
+  \|\hat \pi - \pi \|_{\infty} &\le \epsilon \\
+  %&= O\left(\frac{k \pi_{\max}^{5/2}\| {M_3} \|_\op}{\sigma_k(M_2)^{5/2}} ~ \epsilon \right) \\
+  \|\hat \beta_h - \beta_h\|_2 &\le \epsilon.
+  %&= O\left( \frac{k \pi_{\max} \|M_2\|_\op^{1/2} \| {M_3} \|_\op}{ \sigma_k(M_2)^{5/2} }~ \epsilon \right),
   \end{align*}

### 2013-05-08 doc/icml2013/evaluation.tex
Minor changes + experiments
Various fixes to make it submittable
experiments

--- 2013-04-24 doc/icml2013/evaluation.tex
@@
-\section{Empirical Evaluation}
+\section{Empirical evaluation}
 \label{sec:evaluation}
 
-\subsection{Experimental Setup}
+In the previous section, we showed that Spectral Experts provides a consistent
+estimator.  %that converges a $\frac{1}{\sqrt n}$
+In this section, we explore the empirical properties of our algorithm on simulated data.
+Our main finding is that Spectral Experts alone attains higher parameter error than EM,
+but this is not the complete story.
+If we initialize EM with the estimates returned by Spectral Experts,
+then we end up with much better estimates than EM from a random initialization.
+%In this section, we will support these theoretical results with
+%experiments on simulated data.
+%We will see that in the finite sample
+%regime, even when Spectral Experts does not converge to the optimal
+%answer, it provides robust initialization for local methods like EM.
+
+\subsection{Experimental setup}
 
@@
 
-We solved the low rank recovery optimization routines for $M_2$ and
-$M_3$ using \texttt{cvx}\citationneeded. 
+We experimented with three algorithms.
+The first algorithm (Spectral) is simply the Spectral Experts.
+We set the regularization strengths $\lambda_n^{(2)} = \frac{1}{10^{5} \sqrt{n}}$
+and $\lambda_n^{(3)} = \frac{1}{10^{3} \sqrt{n}}$;
+the algorithm was not very sensitive to these choices.
+We solved the low-rank regression to estimate $M_2$ and
+$M_3$ using an off-the-shelf convex optimizer, CVX~\cite{cvx}.
 %ainterior point method an interior proximal subgradient descent
@@
 %Both regularization parameters, $\lambda_n^{(2)}$ and, $\lambda_n^{(3)}$ were annealed as $\frac{1}{n}$. 
-As a baseline, we also implemented EM with two different
-initializations. In the first, we used $\beta$s with random standard
-Gaussian entries and uniform mixture probabilities, $\pi$. In the other,
-we initialized EM with the starting points given by Spectral Experts.
+The second algorithm (EM) is EM where the $\beta$'s are initialized from a standard normal
+and $\pi$ was set to the uniform distribution plus some small perturbations.
+We ran EM for 1000 iterations.
+In the final algorithm (Spectral+EM),
+we initialized EM with the output of Spectral Experts.
 
-\paragraph{Datasets}
+\paragraph{Data}
 
@@
 \sectionref{sec:algo}.
-The coefficients $\{\beta_h\}$ were drawn from a standard Gaussian. The
-results presented below use $\sigma^2 = 0.1$; we did not observe any
+The true regression coefficients $\{\beta_h\}$ were drawn from a standard normal
+and $\pi$ is set to the uniform distribution.
+The observation noise $\epsilon$ is drawn from a normal with variance $\sigma^2$.
+Results are presented below for $\sigma^2 = 0.1$, but we did not observe any
 qualitatively different behaviour for choices of $\sigma^2$ in the range
@@
 As an example, one feature map we considered in the one-dimensional
-setting was $x = (1, t, t^4, t^7)$. The data and the curves fit using
+setting $(b=1)$ was $x = (1, t, t^4, t^7)$. The data and the curves fit using
 Spectral Experts, EM with random initializations and EM initialized with
@@
   \centering
-  \includegraphics[width=0.50\textwidth]{figures/hist.pdf}
+  \includegraphics[width=0.50\textwidth]{figures/hist.png}
   \caption{Histogram over recovery errors between the three algorithms. $b = 1, d = 4, k = 3, n = 500,000$}
@@
   \subfigure[Spectral, Spectral+EM]{
-    \includegraphics[width=0.50\textwidth]{figures/curves-vis/1-8-3-specm.pdf}}
+    \includegraphics[width=0.50\textwidth]{figures/curves-vis/1-8-3-specm.png}}
     \hspace{-2em}
   \subfigure[EM]{
-    \includegraphics[width=0.50\textwidth]{figures/curves-vis/1-8-3-em.pdf}}
+    \includegraphics[width=0.50\textwidth]{figures/curves-vis/1-8-3-em.png}}
 %  \subfigure[Spectral + EM]{
@@
 %$\{1$, $ x_1$, $ x_1^4\}$ 
-  1 & 4 & 2 & 1.52 $\pm$ 0.71 & 0.28 $\pm$ 0.82 & {\bf 0.13 $\pm$ 0.55} \\
+  1 & 4 & 2 & 2.45 $\pm$ 3.68 & 0.28 $\pm$ 0.82 & {\bf 0.17 $\pm$ 0.57} \\
+% 1-5-2 2.451423 (+/- 3.686348) 0.176685 (+/- 0.579108) 0.007634 (+/- 0.001706) 0.341620 (+/- 0.584178) 0.010296 (+/- 0.007839)
 %$\{1$, $ x_1$, $ x_2$, $ x_1^2 x_2^2\}$ 
-  2 & 5 & 2 & 2.92 $\pm$ 1.10 & 0.00 $\pm$ 0.00 & {\bf 0.00 $\pm$ 0.00} \\
-  2 & 5 & 3 & 3.76 $\pm$ 1.93 & 0.43 $\pm$ 1.07 & {\bf 0.23 $\pm$ 0.76} \\
+2 & 5 & 2 & 1.38 $\pm$ 0.84 & {\bf 0.00 $\pm$ 0.00} & {\bf 0.00 $\pm$ 0.00} \\
+% 2-3-2 1.383626 (+/- 0.840325) 0.001416 (+/- 0.000083) 0.003312 (+/- 0.000399) 0.454076 (+/- 0.758689) 0.003340 (+/- 0.000397)
+  2 & 5 & 3 & 2.92 $\pm$ 1.71 & 0.43 $\pm$ 1.07 & {\bf 0.31 $\pm$ 1.02} \\
+% 2-3-3 2.917581 (+/- 1.714054) 0.308649 (+/- 1.024135) 0.003885 (+/- 0.000442) 0.661009 (+/- 0.543162) 0.012455 (+/- 0.033185)
 %$\{1$, $ x_1$, $ x_2$, $ x_1 x_2^3$, $ x_1^2 x_2^2$, $ x_1^3 x_2 \}$ 
-  2 & 6 & 2 & 4.52 $\pm$ 1.67 & 0.63 $\pm$ 1.29 & {\bf 0.18 $\pm$ 0.66} \\
-\belowspace
-  2 & 6 & 5 & 6.78 $\pm$ 2.18 & 2.29 $\pm$ 1.79 & {\bf 1.77 $\pm$ 1.89} \\
+  2 & 6 & 2 & 2.33 $\pm$ 0.67 & 0.63 $\pm$ 1.29 & {\bf 0.01 $\pm$ 0.01} \\
+% 2-4-2 2.333000 (+/- 0.674578) 0.004067 (+/- 0.001391) 0.001167 (+/- 0.001096) 0.592667 (+/- 0.478518) 0.001200 (+/- 0.001143)
+%\belowspace
+%  2 & 6 & 5 & 6.78 $\pm$ 2.18 & 2.29 $\pm$ 1.79 & {\bf 1.77 $\pm$ 1.89} \\
 
@@
   \subfigure[Well-specified data]{
-    \includegraphics[width=0.50\textwidth]{figures/vs-n/1833-decay.pdf}
+    \includegraphics[width=0.50\textwidth]{figures/vs-n/1833-decay.png}
   }
@@
   \subfigure[Misspecified data]{
-    \includegraphics[width=0.50\textwidth]{figures/vs-n/1-8-3-3-rm.pdf}
+    \includegraphics[width=0.50\textwidth]{figures/vs-n/1-8-3-3-rm.png}
   }
@@
 \abovespace
- 1 & 4 & 2 &  2.94 $\pm$ 1.52 & 0.29 $\pm$ 0.85 &  {\bf 0.22 $\pm$ 0.72} \\
-\belowspace
- 2 & 5 & 3 &  3.82 $\pm$ 2.00 & 0.44 $\pm$ 1.12 &  {\bf 0.35 $\pm$ 1.00} \\
+ 1 & 4 & 2 &  1.55 $\pm$ 0.84 & 0.29 $\pm$ 0.85 &  {\bf 0.01 $\pm$ 0.07} \\
+ % 1-5-2 1.553269 (+/- 0.841002) 0.008463 (+/- 0.067282) 0.006283 (+/- 0.007291) 0.700833 (+/- 1.650770) 0.006344 (+/- 0.007351) 0.124876 (+/- 0.145539) 2.304894 (+/- 2.457321)
+%\belowspace
+ 2 & 5 & 3 &  1.37 $\pm$ 0.85 & 0.44 $\pm$ 1.12 &  {\bf 0.00 $\pm$ 0.00} \\
+% 2-3-2 1.373689 (+/- 0.849721) 0.002074 (+/- 0.000116) 0.004720 (+/- 0.000000) 0.501348 (+/- 0.656840) 0.004702 (+/- 0.000009)
  2 & 6 & 5 &  9.89 $\pm$ 4.46 & {\bf 2.53 $\pm$ 1.77} &  2.69 $\pm$ 1.83 \\

### 2013-05-08 doc/icml2013/theory.tex
experiments

--- 2013-05-08 doc/icml2013/theory.tex
@@
 n_1 &= \Omega \left(\frac{R^{12} \log(1/\delta)}{\min_{p \in [3]} \sigmamin(\Sigma_p)^2} \right) \\
-n_2 &= \Omega \left(\epsilon^{-2}~ \frac{k^2 \pi^2_{\max} \|M_3\|_\op^2}{\sigma_k(M_2)^{5} }\frac{ L^{6} S^{6} R^{12}}{\sigmamin(\Sigma_1)^2} \log(1/\delta) \right).
+n_2 &= \Omega \left(\epsilon^{-2}~ \frac{k^2 \pi^2_{\max} \|M_3\|_\op^2}{\sigma_k(M_2)^{5}} L^{6} S^{6} R^{12} \log(1/\delta) \right).
 \end{align*}
@@
 % Let $\opX_p$ be the linear operator previously defined. 
-If $$n = \Omega \left(\max_{p\in [3]} \frac{L^{2p} S^{2p} R^{4p} \log(1/\delta)}{ \sigmamin(\Sigma_1)^2 \left(\lambda_n^{(p)}\right)^2} \right),$$
+If $$n = \Omega \left(\max_{p\in [3]} \frac{L^{2p} S^{2p} R^{4p} \log(1/\delta)}{\left(\lambda_n^{(p)}\right)^2} \right),$$
 then with probability at least $1-\delta$:
@@
   \|\hat \pi - \pi \|_{\infty}
-  &= O\left(\frac{k \pi_{\max}^{5/2}\| {M_3} \|_\op}{\sigma_k(M_2)^{5/2}} ~ \epsilon \right) \\
+  &= O\left(\frac{k \pi_1^{5/2}\| {M_3} \|_\op}{\sigma_k(M_2)^{5/2}} ~ \epsilon \right) \\
   \|\hat \beta_h - \beta_h\|_2
-  &= O\left( \frac{k \pi_{\max} \|M_2\|_\op^{1/2} \| {M_3} \|_\op}{ \sigma_k(M_2)^{5/2} }~ \epsilon \right),
+  &= O\left( \frac{k \pi_1 \|M_2\|_\op^{1/2} \| {M_3} \|_\op}{ \sigma_k(M_2)^{5/2} }~ \epsilon \right),
   \end{align*}
@@
 
-\begin{proof}[Proof of Theorem 1]
+\begin{proof}[Proof of Theorem 1 (sketch)]
 By \reflem{lowRank}, \reflem{lowRankLower} and \reflem{lowRankUpper}, we

### 2013-05-09 doc/icml2013/discussion.tex
finish experiments

--- 2013-02-16 doc/icml2013/discussion.tex
@@
 
-In this paper, we take a first step at developing computationally efficient and
-statistically consistent estimators for discriminative latent-variable models
-by focusing on mixture of linear regressions. We showed that by regressing on
-higher-order powers of the data with a regularizer that encourages low rank
-structure, coupled with an efficient tensor power method, we obtain a very natural algorithm.
-Empirically, our algorithm can outperform EM, and when used to initialize EM,
-yields significantly more accurate and stable results.
+In this paper, we developed a computationally efficient and statistically
+consistent estimator for mixture of linear regressions.
+Our algorithm, Spectral
+Experts, regresses on higher-order powers of the data with a regularizer that
+encourages low rank structure, followed by tensor factorization to recover the
+actual parameters.  Empirically, we found Spectral Experts 
+to be an excellent initializer for EM.
+%yielding significantly more accurate and stable results than EM.

### 2013-05-09 doc/icml2013/evaluation.tex

@@
 setting $(b=1)$ was $x = (1, t, t^4, t^7)$. The data and the curves fit using
-Spectral Experts, EM with random initializations and EM initialized with
+Spectral Experts, EM with random initialization and EM initialized with
 the parameters recovered using Spectral Experts are shown in
@@
   \includegraphics[width=0.50\textwidth]{figures/hist.png}
-  \caption{Histogram over recovery errors between the three algorithms. $b = 1, d = 4, k = 3, n = 500,000$}
+  \caption{Histogram over recovery errors between the three algorithms when $b = 1, d = 4, k = 3, n = 500,000$.}
   \label{fig:hist}
@@
 \caption{Parameter error $\|\theta^* - \hat \theta\|_F$ ($n = 500,000$)
-as the number of variables $b$, number of features $d$ and the number of components $k$ increases.
+as the number of base variables $b$, number of features $d$ and the number of components $k$ increases.
 While Spectral by itself does not produce good parameter estimates, Spectral+EM improves
@@
 random instances; some are easy for EM to find the global minima and
-others more difficult. In general, while the spectral algorithm did not
-recover parameters extremely well, it provided a good initialization for
+others more difficult. In general, while Spectral Experts did not
+recover parameters by itself extremely well, it provided a good initialization for
 EM.
 
-To study the stability of the solutions returned by the spectral method,
+To study the stability of the solutions returned by Spectral Experts,
 consider the histogram in \figureref{fig:hist}, which shows the recovery
 errors of the algorithms over 170 attempts on a dataset with $b = 1, d = 4,
-k = 3$. Typically, the spectral algorithm returned a stable solution.
+k = 3$. Typically, Spectral Experts returned a stable solution.
 When these parameters were close enough to the true parameters, we found

### 2013-05-09 doc/icml2013/evaluation.tex
Made figure legend black/white, minor language fixes

@@
 Results are presented below for $\sigma^2 = 0.1$, but we did not observe any
-qualitatively different behaviour for choices of $\sigma^2$ in the range
+qualitatively different behavior for choices of $\sigma^2$ in the range
 $[0.01, 0.4]$.
@@
 \abovespace
- 1 & 4 & 2 &  1.55 $\pm$ 0.84 & 0.29 $\pm$ 0.85 &  {\bf 0.01 $\pm$ 0.07} \\
- % 1-5-2 1.553269 (+/- 0.841002) 0.008463 (+/- 0.067282) 0.006283 (+/- 0.007291) 0.700833 (+/- 1.650770) 0.006344 (+/- 0.007351) 0.124876 (+/- 0.145539) 2.304894 (+/- 2.457321)
+ 1 & 4 & 2 &  1.70 $\pm$ 0.85 & 0.29 $\pm$ 0.85 &  {\bf 0.03 $\pm$ 0.09} \\
+% 1-5-2 1.707638 (+/- 0.849695) 0.028238 (+/- 0.093263) 0.008125 (+/- 0.002117) 0.657658 (+/- 0.769400) 0.008112 (+/- 0.002153)
 %\belowspace

### 2013-05-09 doc/icml2013/theory.tex
Made figure legend black/white, minor language fixes

@@
 We can show that some amount of unidentifiability is intrinsic to
-estimation from low-order moments, not just an artifact of our
+estimation from low-order moments, not just an artefact of our
 estimation procedure.  Suppose $x = (t, \dots, t^d)$.  Even if we
@@
 \citet{NegahbanWainwright2009} for matrices.
-The main calculation involved is controling the noise $\eta_p(x)$,
+The main calculation involved is controlling the noise $\eta_p(x)$,
 which involves various polynomial combinations of the mixing noise and observation noise.

### 2013-05-09 doc/icml2013/theory.tex
Corrections to the tensor whitening lemma
Added flag to optionally include the appendix, with some corresponding fixes

@@
 $\|\opX_p^*(\eta_p)\|_\op$. The proofs of the following lemmas follow
-from standard concentration inequalities and are detailed in the
-supplementary material.
+from standard concentration inequalities and are detailed in 
+\iftoggle{withappendix}{
+\appendixref{sec:proofs:regression}.
+}{
+the supplementary material.
+}
 %have been deferred to \appendixref{sec:proofs}.
@@
 \begin{lemma}
-  \label{lem:tensorPower} Let $\|\hat M_2 - M_2\|_\op$ and $\|\hat M_3 - M_3\|_\op$ both be less than
+  \label{lem:tensorPower}
+  Let $M_3 = \sum_{h=1}^{k} \pi_h \beta_h\tp{3}$.
+  Let $\|\hat M_2 - M_2\|_\op$ and $\|\hat M_3 - M_3\|_\op$ both be less than
   \vspace{-0.5em}
@@
 The proof follows by applying standard matrix perturbation results for
-the whitening and unwhitening operators and has again been deferred
-to the supplementary material.
+the whitening and unwhitening operators and 
+\iftoggle{withappendix}{
+can be found in \appendixref{sec:proofs:tensors}.
+}{
+has again been deferred to the supplementary material.
+}
 %\appendixref{sec:proofs}.

### 2013-05-21 doc/nips2013/abstract.tex
first draft of NIPS paper

\begin{abstract}
Latent-variable models.
\end{abstract}

### 2013-05-21 doc/nips2013/utils.js

@@ -1,0 +1,160 @@
+var G = sfig.serverSide ? global : this;
+sfig.importAllMethods(G);
+sfig.latexMacro('cl', 1, '\\textsf{#1}');  // Category labels
+sfig.latexMacro('il', 1, '\\texttt{#1}');  // Identifier labels
+sfig.latexMacro('E', 0, '\\mathbb{E}');
+sfig.latexMacro('phix', 0, 't');
+sfig.initialize();
+
+G.prez = sfig.presentation();
+
+G.node = function(x, shaded) { return overlay(circle(25).fillColor(shaded ? 'lightgray' : 'white') , std(x).orphan(true)).center(); }
+
+G.moveLeftOf = function(a, b, offset) { return transform(a).pivot(1, 0).shift(b.left().sub(offset == null ? 5 : offset), b.ymiddle()); }
+G.moveRightOf = function(a, b, offset) { return transform(a).pivot(-1, 0).shift(b.right().add(offset == null ? 5 : offset), b.ymiddle()); }
+G.moveTopOf = function(a, b, offset) { return transform(a).pivot(0, 1).shift(b.xmiddle(), b.top().up(offset == null ? 5 : offset)); }
+G.moveBottomOf = function(a, b, offset) { return transform(a).pivot(0, -1).shift(b.xmiddle(), b.bottom().down(offset == null ? 5 : offset)); }
+G.moveCenterOf = function(a, b) { return transform(a).pivot(0, 0).shift(b.xmiddle(), b.ymiddle()); }
+G.moveBottomLeftOf = function(a, b, offset) { return transform(a).pivot(1, -1).shift(b.left().sub(offset == null ? 5 : offset), b.bottom().down(offset == null ? 5 : offset)); }
+G.moveBottomRightOf = function(a, b, offset) { return transform(a).pivot(-1, -1).shift(b.right().add(offset == null ? 5 : offset), b.bottom().down(offset == null ? 5 : offset)); }
+
+G.red = function(x) { return x.fontcolor('red'); }
+G.green = function(x) { return x.fontcolor('green'); }
+G.blue = function(x) { return x.fontcolor('blue'); }
+G.darkblue = function(x) { return x.fontcolor('darkblue'); }
+
+G.wholeNumbers = function(n) {
+  var result = [];
+  for (var i = 0; i < n; i++) result[i] = i;
[…]

### 2013-05-21 doc/icml2013/intro.tex
Finished lemma 5 and minor changes in supplementary material
Made figure legend black/white, minor language fixes

@@
 direction by proposing a new algorithm for a simple model,
-\emph{a mixture of linear regressions} (MLG) \cite{VieleTong2002}.
+\emph{a mixture of linear regressions} \cite{VieleTong2002}. % (MLG) ?
 
@@
 % Norms
+For a vector $X$,
+let $\|X\|_\op$ denote the 2-norm.
 For a matrix $X$,
@@
 
+Let $\vvec(X)$ be the vectorization of a $p$-th order tensor. For
+example, if $X \in (\Re^{2})\tp{3}$, $\vvec(X) = (X_{111}, X_{112},
+\cdots, X_{222})$.
 For a tensor $X \in (\Re^d)\tp{p}$, let $\cvec(X) \in
@@
 $X$. For example, if $X \in \Re^{d \times d}$, $\cvec(X)
-= (X_{ii} : i \in [d]; X_{ij} + X_{ji} : i,j \in [d], i<j)$.
+= (X_{ii} : i \in [d]; \frac{X_{ij} + X_{ji}}{\sqrt{2}} : i,j \in [d], i<j)$.
 In general, each component of $\cvec(X)$ is indexed by a vector of
 counts $(c_1, \dots, c_d)$ with total sum $\sum_i c_i = p$.  The value
-of that component is $\sum_{k \in K(c)} X_{k_1 \cdots k_p}$, where $K(c)
+of that component is $\frac{1}{\sqrt{|K(c)|}} \sum_{k \in K(c)} X_{k_1 \cdots k_p}$, where $K(c)
 = \{ k \in [d]^p : \forall i \in [d], c_i = |\{ j \in [p] : k_j = i \}|
 \}$ are the set of index vectors $k$ whose count profile is $c$.
-For symmetric tensors $X$ and $Y$, $\innerp{X}{Y} = \innerp{\cvec(X)}{\cvec(Y)}$.
+We note that for a symmetric tensor $X$ and any tensor $Y$,
+$\innerp{X}{Y} = \innerp{\cvec(X)}{\cvec(Y)}$; this property is not true
+in general though.
 Later, we'll see that vectorization allow us to perform regression on tensors,

### 2013-05-22 java/tests/learning/models/MixtureOfGaussiansTest.java
Abstracted MixtureOfGaussian generation to a separate routine

@@ -20,6 +20,38 @@ import org.junit.Before;
  */
 public class MixtureOfGaussiansTest {
 
+  // Generation routines
+  public static MixtureOfGaussians generateMultiView(int K, int D, int V, MeanDistribution means) {
+    return MixtureOfGaussians.generate(K, D, V, 
+        WeightDistribution.Uniform, 
+        means, 
+        CovarianceDistribution.Spherical, 1.0);
+  }
+  public static MixtureOfGaussians generateSmallEye() {
+		int K = 2;
+		int D = 3;
+		int V = 3;
+    return generateMultiView( K, D, V, MeanDistribution.Hypercube );
+  }
+  public static MixtureOfGaussians generateSmallRandom() {
+		int K = 2;
+		int D = 3;
+		int V = 3;
+    return generateMultiView( K, D, V, MeanDistribution.Random );
+  }
+  public static MixtureOfGaussians generateMediumEye() {
+		int K = 4;
+		int D = 6;
+		int V = 3;
+    return generateMultiView( K, D, V, MeanDistribution.Hypercube );
+  }
+  public static MixtureOfGaussians generateMediumRandom() {
+		int K = 4;
+		int D = 6;
+		int V = 3;
+    return generateMultiView( K, D, V, MeanDistribution.Random );
+  }
+
   public void testModel( int N, GenerationOptions options ) {
     int D = (int) options.D;
     int K = (int) options.K;

### 2013-05-22 java/tests/learning/spectral/TensorMethodTest.java

@@ -12,6 +12,7 @@ import learning.exceptions.RecoveryFailure;
 
 import learning.models.MixtureOfGaussians;
 import learning.models.MixtureOfGaussians.*;
+import static learning.models.MixtureOfGaussiansTest.*;
 
 import org.javatuples.*;
 
@@ -34,39 +35,6 @@ public class TensorMethodTest {
     //LogInfo.writeToStdout = false;
     LogInfo.init();
   }
-
-  // Generation routines
-  public MixtureOfGaussians generateMultiView(int K, int D, int V, MeanDistribution means) {
-    return MixtureOfGaussians.generate(K, D, V, 
-        WeightDistribution.Uniform, 
-        means, 
-        CovarianceDistribution.Spherical, 1.0);
-  }
-  public MixtureOfGaussians generateSmallEye() {
-		int K = 2;
-		int D = 3;
-		int V = 3;
-    return generateMultiView( K, D, V, MeanDistribution.Hypercube );
-  }
-  public MixtureOfGaussians generateSmallRandom() {
-		int K = 2;
-		int D = 3;
-		int V = 3;
-    return generateMultiView( K, D, V, MeanDistribution.Random );
-  }
-  public MixtureOfGaussians generateMediumEye() {
-		int K = 4;
-		int D = 6;
-		int V = 3;
-    return generateMultiView( K, D, V, MeanDistribution.Hypercube );
-  }
-  public MixtureOfGaussians generateMediumRandom() {
-		int K = 4;
-		int D = 6;
-		int V = 3;
-    return generateMultiView( K, D, V, MeanDistribution.Random );
-  }
-
   // Actual tests
 
   public void testSymmetrization( MixtureOfGaussians model ) {

### 2013-05-22 doc/nips2013/factorial.tex

\section{Factorial models} \label{sec:factorialModels}

So far, we have relied on the existence of bottleneck hidden variables $h_j$ on which
we have three conditionally independent views.  Such bottlenecks do not always
exist; for example, in \reffig{factorialModels}(a), neither $h_1$ or $h_2$ is a bottleneck.
Our solution is to create a composite latent variable out of existing latent
variables.  We can invoke $\TensorFactorize$ to obtain composite marginals and
conditional means.  However, the identity of $h_1$ and $h_2$ becomes lost in the composite,
and must be recovered.  To do this, we develop a new algorithm $\UnshuffleFactorize$
to produce constraints on the original parameters.

\Fig{figures/factorialModels}{0.3}{factorialModels}{Examples of factorial models.
There do not exist three-view bottlenecks, so we first identify composite bottlenecks,
and then perform an \emph{unshuffling factorization}.}

%restricted Boltzmann machines \cite{salakhutdinov09softmax}

We focus on the factorial mixture model.
Let $h = (h_1, \dots, h_s) \in [\nh]^s$ be the latent variables (sources)
and $x = (x_1, x_2, x_3) \in [\nphix]^3$ be the observed variables.
\begin{align}
p_\theta(x, h) = \prod_{i=1}^s p_\theta(h_i) \prod_{j=1}^3 p_\theta(x_j \mid h), \quad
p_\theta(x_j \mid h) = \exp\left\{ \sum_{i=1}^s \theta_i(x_j, h_i) - A(\theta; h) \right\},
\end{align}
where each $\theta_i \in \R^{\nphix \times \nh}$ are the parameters for the $i$-th source.
%Let $\theta = (\theta_1, \dots, \theta_b) \in \R^{\nphix \times b \nh}$.

Let $B \eqdef (p_\theta(x_j \mid h))_{x_j \in [\nphix],h \in [\nh]^s} \in
\R^{\nphix \times \nh^s}$ be the conditional mean.

\begin{assumption}
  Assume that $B$ has full column rank ($k^s$).\footnote{Note that this implies
  $d \ge k^s$, which is a rather stringent assumption, but is necessary for the
  three-view method of moments technique to work.}
\end{assumption}

We will measure $M_2 \eqdef \E[x_1 \otimes x_2]$ and
$M_3 \eqdef \E[x_1 \otimes x_2 \otimes x_3]$.
We run $\TensorFactorize$, which will produce $B = (\E[x_j \mid c])$, where the
cluster $c$ ranges over $k^s$ possible clusters.  Next, we still have to
extract the actual parameters $\theta$ from $B$.  In the simple mixture model,
the clusters produced by $\TensorFactorize$ were in exact one-to-one correspondence
with the actual mixture components, but in the factorial case, we must solve
the \emph{credit assignment problem} to the $s$ sources.

Let $C$ be the elementwise logarithm of $B$; that is
$C(x_j, h) = \log B(x_j, h)$.

Let $v_{ia} = \theta_i(\cdot, a) \in \R^\ell$ be the collection of parameter vectors that we seek to recover.
What we observe are normalized sums of these vectors.
Specifically, we observe the unordered collection of vectors $\{ \sum_{i=1}^s v_{ia_i} - Z_{\ba} : \ba = (a_1, \dots, a_s) \in [\nh]^s \}$,
where $Z_{\ba} = \log \sum_{e=1}^\ell (\sum_{i=1}^s \exp(v_{ia_i}(e)))$ is the normalization constant.

We say two vectors are \emph{comparable} if they differ by a constant times $\bone$.

Here's the algorithm:
\begin{itemize}
  \item For each pair of vectors $(i,j) \in [\nh^s]^2$, consider the difference $B_i - B_j$.
  \item Put the pairs into bins, where each bin has differences which are comparable.
  \item Keep only the bins with the maximum number of pairs (which should be $\nh$).
  Now, each bin $b$ corresponds to a signature $v_{ia} - v_{ia'}$ for some source $i$ and two values $a$ and $a'$.
  \item Group the bins into sources.
  Construct a graph with the bins as nodes as follows.
  Let two bins be connected if the difference of their signatures is the signature of some other bin.
  For example, bin $b_1$ with signature $v_{ia} - v_{ia'}$
  is connected to bin $b_2$ with signature $v_{ia''} - v_{ia'}$
  because some bin $b_3$ has signature $v_{ia} - v_{ia''}$ by construction of the bins.
  The connected components of graph are exactly the sources.
  \item For each connected component, identify a subset of the signatures which all have a common difference:
  $v_{ia_1} - v_{ia'}$, $v_{ia_2} - v_{ia'}$, $v_{ia_3} - v_{ia'}$, etc.
  \item Construct a linear program.  It will have $\nh^s \cdot d$ constraints, one for each entry of $B$.
  It will have $s\nh \cdot d + \nh^s$ variables,
  one for each of $s\nh$ vectors $v_{ia}$ and $\nh^s$ for the normalization constant
  (these are actually deterministic functions of the vectors, but non-linear, so let's linearize).
  TODO: finish explaining this.
\end{itemize}

### 2013-05-22 doc/nips2013/mixture.tex

\section{Three-view mixture models}
\label{sec:threeViewMixtureModel}

\Fig{figures/simpleModels}{0.3}{simpleModels}{Three-view mixture models: conditioned on a latent variable $h \in [\nh]$,
we have three conditionally independent observed variables.  We consider
estimation for both directed (left) and undirected (right) versions.
}

In this section, we consider the simplest model to illustrate the form of our method.
Consider the undirected mixture model model in \reffig{simpleModels}(a).
We have one discrete latent variable $h \in [\nh]$ and
three observed variables
$x = (x_1, x_2, x_3) \in \sX_1 \times \sX_2 \times \sX_3 = \sX$
which are conditionally independent given $h$.
Formally, the model is as follows:\footnote{We assume for simplicity that the $x_j$'s are identically distributed conditioned on $h$,
but it is possible to use different distributions (see \citet{anandkumar12moments}).}
\begin{align}
  \label{eqn:mixtureModel}
  p_\theta(x, h) \eqdef \exp \{ \phi(x, h)^\top\theta - A(\theta) \}, \quad\text{where} \, \phi(x, h) \eqdef \phiprior(h) + \sum_{j=1}^3 \philocal(x_j, h) \in \R^d.
\end{align}

In the following, we will detail the algorithm for producing a parameter
estimate $D$.  For simplicity of exposition, we will assume we have an infinite
amount of data and work with the population quantities.  Actual parameter
estimates are constructed by simply plugging in empirical quantities.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Step 1: Compute observed moments (aggregation)}

Rather than directly trying to connect the parameters $\theta$ to the data, we start
bottom-up by considering what properties of the data we can measure.
Let $\phix : \sX \mapsto \R^\ell$ denote a \emph{transformation function} (the requirement will be specified later),
which will extract the properties of the observed variables $x$ relevant to $h$.
Define the \emph{observed moments} (analogous to those used in \cite{anandkumar12moments}):
\begin{align}
  M_2 \eqdef \E[\phix(x_1) \otimes \phix(x_2)] \in \R^{\nphix \times \nphix}, \quad\quad
  M_3 \eqdef \E[\phix(x_1) \otimes \phix(x_2) \otimes \phix(x_3)] \in \R^{\nphix \times \nphix \times \nphix}.
\end{align}
Importantly, these quantities are easily estimated from data by taking empirical averages.

\subsection{Step 2: Recover hidden moments (factorization)}

By exploiting the conditional independence structure of the model, the observed
moments $M_2,M_3$ (which only reference the observed variables) admit a
factorization in terms of the \emph{hidden moments}
(which provides a window into the hidden variables).

Define
$\pi \eqdef (\BP[h=1], \dots, \BP[h=\nh])$ to be the marginal probabilities over the latent variable $h$,
and define the matrix $B \eqdef (\E[\phix(x_j) \mid h])_{h \in [\nh]} \in \R^{\ell \times k}$,
where each column is the expected transformated data conditioned on $h$.

Now, we can express the observed moments $(M_2,M_3)$ in terms of the hidden moments $(\pi,B)$:
\begin{align}
  M_2 = \sum_h \pi_h B_h \otimes B_h \in \R^{\nphix \times \nphix}, \quad\quad
  M_3 = \sum_h \pi_h B_h \otimes B_h \otimes B_h \in \R^{\nphix \times \nphix \times \nphix}.
\end{align}
\begin{assumption}
  The hidden moments matrix $B \in \R^{\nphix \times \nh}$ has full column rank $\nh$.
\end{assumption}

\citep{anandkumar13tensor} presented a robust tensor power method, which we call $\TensorFactorize$,
which converts consistent estimates of $(M_2,M_3)$ into consistent estimates of $(\pi, B)$ (up to permutation of the columns).
In brief, algorithm first computes a whitening matrix $W \in \R^{\nphix \times \nh}$ such that $W^\top M_2 W = I$,
and uses it to whiten $M_3$.  Then a robust tensor power method is applied to extract the eigenvectors of the whitened $M_3$,
which correspond to columns of $B$.
%\begin{lemma}[tensor factorization \citep{anandkumar13tensor}]
%Let $B \in \R^{\nphix \times \nh}$ be a rank $k$ matrix.
%Let $M_2 = \sum_h \pi_h B_h \otimes B_h \in \R^{\nphix \times \nphix}$ be a rank $k$ symmetric matrix.
%and $M_3 = \sum_h \pi_h B_h \otimes B_h \otimes B_h \in \R^{\nphix \times \nphix \times \nphix}$ be a rank $k$ symmetric tensor.
%There exists an polynomial time algorithm $\TensorFactorize$ that takes consistent estimates of $M_2$ and $M_3$
%and returns consistent estimates of $(\pi, B)$ (up to permutation of the columns).
%\end{lemma}

We need to specify some conditions on the transformation function $\phix$
to ensure that enough information is retained to pinpoint the parameters later.
The following condition characterizes $\phix$:
\begin{property}
  \label{property:transformation}
  The transformation $\phix$ (chosen by us) is such that $C \eqdef
  (\E[\philocal(x_j, h) \mid h])_{h \in [\nh]} \in \R^{d \times h}$ is a
  function of $B$.
\end{property}
Here, $C$ provide a view on the local features $\phi_L(x_j,h)$ which is more
directly related to the model parameters \refeqn{mixtureModel}, whereas $B$ is
what we have recovered so far.  Note that we couldn't recover $C$ directly because it depends on $h$.

However, it is not difficult to construct $\phix$ such that Property~\ref{property:transformation} holds.
Since $h$ is discrete, we can always take $\phix(x) = \flatten((\philocal(x, h))_{h \in [\nh]}) \in \R^{dk}$,
which contains $k$ blocks, one for each latent state.  We can recover a column $C_h = \E[\philocal(x_j, h) \mid h]$
by taking the block of column $B_h = \E[\flatten(\philocal(x_j, h'))_{h' \in [\nh]} \mid h]$
corresponding to $h = h'$.
Often, the local features are simply the tensor product of some function $g(x)$ and an indicator on $h$:
$\philocal(x,h) = \flatten(g(x) \otimes e_h)$.
In this case, it suffices to take $\phix(x) = g(x)$.

Finally, having retrieved $C_h$, it is easy to compute the \emph{hidden moments} $\mu$:
\begin{align}
  \mu \eqdef \E[\phi(x,h)] = \E[\phiprior(h)] + \sum_{j=1}^3 \E[\philocal(x_j,h)] = \sum_{h \in [\nh]} \pi_h (\phi_p(h) + 3 C_h).
\end{align}
In \cite{anandkumar12moments}, these hidden moments would be exactly the parameters (multinomial probabilities), but
for log-linear models, we still need to do more work to extract the parameters $\theta$.

\subsection{Step 3: Recover parameters (optimization)}

For the mixture model, the hidden moments $\E[\phi(x,h)]$ we recover are actually
are the sufficient statistics of the log-linear model.
To elaborate, let us suppose for the moment that we have infinite examples of $(x,h)$.
Let us consider the undirected mixture model.
Then we would seek to maximize the joint log-likelihood:
\begin{align}
L(\theta) \eqdef \E[\log p_\theta(x, h \mid \theta)] = {\underbrace{\E[\phi(x,h)]}_{\mu}}^\top\theta - A(\theta).
\end{align}
Note that the optimization problem here only depends on the data via $\mu$, which is usually estimated from examples of $(x,h)$.
Since $h$ is latent in our latent-variable setting, we must recover $\mu$ from latent observations.
By the moment-generating properties of the log-partition function,
the optimality condition is that $\mu = \E_\theta[\phi(x, h)]$.

We thus obtain a consistent estimator for $\theta$ assuming the exponential family is identifiable:
\begin{assumption}
The log-linear model is identifiable from fully-observed data.
Equivalently, the log-partition function $A(\theta)$ is strongly convex.
\end{assumption}
Identifiability means that the moments $\mu$ uniquely determine $\theta^*$.
In summary, factorization yields the mean parameters of the exponential family,
and we can use convex optimization to retrieve the (canonical) parameters $\theta$.

% Directed mixture model
\paragraph{Directed mixture model}

The directed mixture model, we have $p_\theta(x,h) = p_\theta(h) \prod_{j=1}^3 p_\theta(x_j \mid h)$,
where
$p_\theta(h) \eqdef \exp\{\phiprior(h)^\top\theta - \Aprior(\theta)\}$
and
$p_\theta(x_j \mid h) \eqdef \exp\{\philocal(x_j, h)^\top\theta - \Alocal(\theta; h)\}$.
Consequently, the joint log-likelihood is:
\begin{align}
  L(\theta) = {\underbrace{\E[\phi(x,h)]}_{\mu}}^\top\theta - \E[A(\theta; z)], \quad
  \E[A(\theta, z)] = \Aprior(\theta) + 3 \sum_{h \in [\nh]} \pi_h \Alocal(\theta;h).
\end{align}
Note that in directed models, the log-partition functions depends on the parent variables (in this case, $h$).
Having recovered the marginal distribution over $\pi$, we can compute the desired expectation.
Again, this problem is convex.

\paragraph{Perturbation analysis.}
We have so far worked with exact population values $M_2,M_3,\pi,B,\mu,\theta^*$.
With finite samples, we must work the corresponding estimated quantities $\hat M_2,\hat M_3,\hat\pi,\hat B,\hat\mu,\hat\theta$.
How the error in the moments $(\hat M_2,\hat M_3)$ affects error in $(\hat\pi,\hat B)$ is well studied
(see \citet{anandkumar12moments,anandkumar13tensor}).
The hidden moments $\hat\mu$ are simply a quadratic combination of $(\hat\pi,\hat B)$,
and $\hat\theta = \nabla^{-1} A(\hat\mu)$, which is Lipschitz.
Therefore, we have a consistent estimator with polynomial sample complexity.

### 2013-05-22 doc/nips2013/abstract.tex
Adding files to nips that should've already been there

@@
 \begin{abstract}
-Latent-variable models.
+  Latent variable models are at the core of machine learning today, yet
+  the local estimation methods in use, like EM, are neither well
+  understood nor are do they come with any guarantees. In this paper, we
+  study asymptotic rates of consistent estimators, based on the method
+  of moments, for a large class of latent variable models. We do so by
+  deriving rates for several base cases like the moments of vectors and
+  conditional moments and composing results in a modular way using
+  combinators like the mixture combinator and the logistic operator.  We
+  compare our asymptotic bounds with those of recently proposed method
+  of moments algorithms for the mixture of Gaussians and linear
+  regressions. Finally, we formalize the intuition that estimates from
+  method of moment estimators are good initializations for local methods
+  like EM by characterizing the ``local convexity'' region for the above
+  two models.
 \end{abstract}

### 2013-05-22 doc/nips2013/discussion.tex

\section{Related work}

% Observable operator models
HMMs \cite{hsu09spectral},
low-treewidth graphs \cite{parikh12spectral},
probabilistic context-free grammars \cite{cohen12pcfg},
weighted finite state automata \cite{balle12automata},
finite state transducers \cite{balle11transducer}.

% Method of moments
Latent Dirichlet Allocation \cite{anandkumar12lda},
mixture of spherical Gaussians \cite{hsu13spherical},
latent parse tree models \cite{hsu12identifiability}
mixture of linear regressions \cite{chaganty13regression}.
linear Bayesian networks \cite{anandkumar12linear}.

% Structure recovery
structure recovery
\cite{anandkumar11tree}

% Alternate estimators
There are multiple consistent estimators.
Composite likelihood \cite{besag75pseudo,lindsay88composite,liang08asymptotics}
We are assuming inference is tractable.
Our main problem is non-convexity.

\section{Conclusion}

That's it.

### 2013-05-22 doc/nips2013/measurements.tex

\section{Measurements: partial information from moments} \label{sec:generalModels}

Our learning algorithm for three-view mixture models relies on the fact that the hidden moments we can recover
are exactly the sufficient statistics of the exponential family model.
Of course, these conditions do not hold for general log-linear models.
Rather than falling back on local optimization of the likelihood in these cases, we would still like
to use the \emph{partial information} offered by the hidden moments provide we
\emph{can} recover in some way.  We will show how to incorporate these moments
into a likelihood-based objective.

Partial information is available when the following holds:
\begin{property}
  There exists some latent variable $h_{j_0} \in [\nh]$ (the bottleneck)
  and three conditionally independent observed variables $x_{j_1},x_{j_2},x_{j_3}$ (views) such that
  each conditional mean matrix $\E[\phix(x_{j}) \mid h_{j_0}] \in \R^{\nphix \times \nh}$ ($j \in \{ j_1, j_2, j_3 \}$) has full column rank $\nh$,
  for some transformation $\phix$.
\end{property}

% Examples
\reffig{generalModels} shows examples of log-linear models.
For simplicitly, assume each hidden variable $h_j \in [\nh]$ and each observed
variable $x_j \in \R^\ell$ are indicator vectors encoding one of $\ell$ possible values.

\Fig{figures/generalModels}{0.3}{generalModels}{
Three examples of undirected log-linear models.  In each of these cases,
we can recover hidden moments, which are themselves not sufficient to identify the parameters,
but provide partial information which leads to a more well-behaved likelihood function.
}

\paragraph{(a) HMM}
Following the reduction in \citet{anandkumar12moments},
take $h_2$ as the bottleneck and $x_1, x_2, x_3$ as the three views.
Define $\pi = (\BP[h_1]) \in \R^{\nh}$,
transition matrix $T = (\BP[h_3 \mid h_2]) \in \R^{\nh \times \nh}$
and observation matrix $O = (\E[x_2 \mid h_2]) \in \R^{\nphix \times \nh}$.
$\TensorFactorize$ recovers the marginal probability of the bottleneck
$T \pi$ and conditional means of the three views
$O T^\top \diag(T \pi)^{-1}$,
$O$,
and $O T$.
Note that we can recover $T = O^\dagger(OT)$ and $\pi = T^{-1} (T\pi)$,
from which we can compute the hidden moments/sufficient statistics $\mu = \E[\phi(x,h)]$.

However, this model only uses the first three observations,
which doesn't use all the information available.
Let us consider a new reduction:
Let $h_j$ be the bottleneck where $j$ is chosen uniformly at random over $\{2,
\dots, T-1\}$, where $L$ is the number of time steps; let $x_{j-1},x_j,x_{j+1}$ be the three views.\footnote{
When we actually estimate the observed moments, we simply average over the choice of $j$.}
Now, note that we can recover $O = \inv{L-2} \sum_{j=2}^{L-1} \E[x_j \mid h_j]$ from the second view,
but using the data more fully.  The first and third views are unfortunately now
quite complex due to marginalization over the rest of the variables.

\paragraph{(b) Tall mixture model}
Let $h_0$ be the bottleneck and $x_1,x_2,x_3$ be the three
views.  The conditional means that $\TensorFactorize$ reveals are $\E[x_j \mid h_0]$, which are not even
expectations over features in the original model, let alone sufficient
statistics.\footnote{Note that if the features were arbitrary, then the
parameters would not be identifiable, but because log-linear models permit
parameter sharing, then identifiability is still possible and therefore
parameter estimation in this model is still meaningful.}

\paragraph{(c) Grid model}
As in the HMM example, choose a random $h_{ij}$ as the bottleneck,
$x_{ij}^a,x_{ij}^b$ as two of the views, and any other observed variable as the third view.
$\TensorFactorize$ would reveal conditional means $\E[x_{ij}^a \mid h_{ij}]$ and $\E[x_{ij}^b \mid h_{ij}]$,
and a third quantity that involves the hidden-to-hidden features and the observation features.
Unfortunately, this quantity involves marginalizing out the rest of the graph,
and therefore impossible to extract something useful like $\BP[h_{ij}, h_{i(j+1)}]$
in closed form.
Note that the analog of this was still possible in the first HMM reduction that used only the first three observations,
but not in the second one.

In each of these three models, we could extract marginals $\pi = \BP[h_j] \in \R^{\nh}$
and some conditional mean matrix $B = \E[x_{j'} \mid h_j] \in \R^{\nphix \times \nh}$,
but that it was insufficient to identify the parameters.  How can we still use this partial information $(\pi,B)$?
To this end, we adopt the measurements framework of \citet{liang09measurements},
which allows us to incorporate moment constraints into a likelihood-based objective in a coherent way.
The basic idea is to define two families of distributions: $\sQ$ which are
consistent with all information and $\sP$ the exponential family.  The
framework essentially tries to choose $q \in \sQ$ and $p \in \sP$ to minimize
$\KL{q}{p}$.

Specifically,
let $\bx = (x^{(1)}, \dots, x^{(n)})$ be the vector of observed variables
$\bh = (h^{(1)}, \dots, h^{(n)})$ be the vector of latent variables.
Our model family $\sP$ is defined by product distributions
$p_\theta(\bx,\bh) = \prod_{i=1}^n p_\theta(x^{(i)}, h^{(i)})$ as $\theta \in \R^d$.
Suppose our observations are $(\tilde x^{(1)}, \dots, \tilde x^{(n)})$
and we have recovered $(\pi,B)$, where $\pi$ is the marginal of the bottleneck $h_j$ and $B$ is the conditional mean of $x_{j'}$.

We introduce the following constraints on each $q \in \sQ$:\footnote{We use a hard constraint
for expository clarity, but in practice, we use a quadratic penalty.  See
\citet{liang09measurements} for details.}
\begin{align}
  q(x^{(i)} = \tilde x^{(i)}) &= 1, \quad i \in [n] \label{eqn:exConstraint} \\
  \inv{n} \sum_{i \in [n]} q(x_{j'}^{(i)} = v, h_j^{(i)} = u) &= \pi_u B_{v u}, \quad u \in [\nh], v \in [\nphix]. \label{eqn:momentConstraint}
\end{align}
The first constraint $\sQ$ encodes our observed examples
and the second puts additional constraints on the latent variables.
Our final objective function is as follows:
\begin{align}
  \min_{q \in \sQ} \min_{\theta \in \R^d} \KL{q(\bx, \bh)}{p_\theta(\bx, \bh)}. \label{eqn:minKL}
\end{align}
\paragraph{Intuitions.}
With only the first constraint \refeqn{exConstraint}, the objective
\refeqn{minKL} is equivalent to our original maximum (empirical) marginal
likelihood problem, which is non-convex.  If the second constraint
\refeqn{momentConstraint} are exactly the features/sufficient statistics of the
model (as in the three-view mixture model from \refsec{threeViewMixtureModel}),
then \refeqn{minKL} is always convex and
converges to the (expected) maximum likelihood objective function as the number of examples $n \to\infty$.

% Optimization
We optimize \refeqn{minKL} using a variant of EM.
Briefly, the E-step is equivalent to a generalized maximum entropy problem,
which we solve by taking the dual, resulting in a parameter for each moment constraint \refeqn{momentConstraint}.
In the M-step, we maximize the expected log-likelihood with respect to $q$.

%$\beta = \argmax_{q} H(q) \tau^\top\beta - B(\beta, \theta) + r(\beta)$
%E-step: $\beta = \arg\max_{\beta} \tau^\top\beta - B(\beta, \theta) - r(\beta)$
%M-step: $\theta = \arg\max_\theta \E_q[\phi(\bx,\bh)^\top - A(\theta)]$.

### 2013-05-22 doc/nips2013/setup.tex

\section{Setup} \label{sec:setup}

Let $z$ be a collection of variables indexed by $V$ (that is, $z = \{z_j\}_{j \in V}$).
Suppose that some of the variables are observed ($x = \{ z_j \}_{j \in O}$, where $O \subset V$) 
and the rest are latent ($h = \{ z_j \}_{j \in H}$, where $H = V \backslash O$).
Let $C \subset 2^V$ be a collection of subset of variables,
and let $\phi_c(z_c) \in \R^d$ be the feature map for each $c \in C$.
In this paper, we will consider both directed and undirected models over $z$.
In the undirected case, we define the standard globally-normalized exponential family model:
\begin{align}
  p_\theta(z) = \exp\{ \phi(z)^\top\theta - A(\theta) \}, \aside{undirected}
\end{align}
where $\phi(z) \eqdef \sum_{c \in C} \phi_c(z_c) \in \R^d$ is the global feature vector, $\theta \in \R^d$ is the parameter vector,
and $A(\theta) = \log \int \exp\{\phi(z)^\top\theta\} dz$ is the global log-partition function.

In the directed case, we assume a directed acyclic graph over $V$,
and define $\pi(j) \subset V$ to be the parents of each node $j \in V$.
Next, define a model as a product of locally-normalized models:
\begin{align}
  p_\theta(z) = \prod_{j \in V} p_\theta(z_j \mid z_{\pi(j)}), \quad\quad
  p_\theta(z_j \mid z_{\pi(j)}) = \exp \{ \phi_j(z_j, z_{\pi(j)})^\top\theta - A_j(\theta; z_{\pi(j)}) \}.
\end{align}
Here, each local model has its own local features $\phi_j(z_j, z_{\pi(j)})$
and its own local log-partition function $A(\theta; z_{\pi(j)})$, which depends on values of its parent.
To make the directed and undirected models more comparable,
define $\phi(z) = \sum_{j \in V} \phi(z_j, z_{\pi(j)})$
and $A(\theta; z) = \sum_{j \in V} A_j(\theta; z_{\pi(j)})$,
so that:
\begin{align}
  p_\theta(z) \eqdef \exp\{ \phi(z)^\top\theta - A(\theta; z) \}. \aside{directed}
\end{align}

%To some extent, our methods only depend on the conditional independence
%structure of the models, not on whether the model is directed or undirected.
%Later, we will show how to estimate

%\paragraph{Example.}
%For example, \reffig{simpleModels} Markov random fields are an instance.

\paragraph{Problem statement.}

% Statement
This paper focuses on the problem of parameter estimation:
We are given $n$ i.i.d.~examples of the observed variables $D = (x^{(1)}, \dots, x^{(n)})$
where each $x^{(i)} \sim p_{\theta^*}$ for some true parameters $\theta^*$.
Our goal is to produce a parameter estimate $\hat\theta$ which is hopefully
close to $\theta^*$.

% Maximum likelihood
The standard estimation procedure is maximum (marginal) likelihood,
$\hatthetaml \eqdef \max_{\theta \in \R^d} \sum_{x \in D} \log p_\theta(x)$,
which is statistically efficient but computationally intractable.
In practice, one uses EM or gradient-based optimization procedure on the marginal likelihood,
which can get stuck in local optima.

%Since both models have the same sufficient statistics lead to convex
%likelihoods (assuming all variables are observed), in many cases our methods
%will not be particular to which.  In some cases, however (e.g., the factorial models),
%we will rely on marginalization properties of the directed models.

% Principles
%By insisting on developing a consistent estimator and considering identifiability,
%The key is that a variable in the model must correspond to a bottleneck.

### 2013-05-22 java/src/learning/data/CachedProjectedCorpus.java
Added tests for corpus and projections.

@@ -18,9 +18,12 @@ import com.spaceprogram.kittycache.KittyCache;
 public class CachedProjectedCorpus extends ProjectedCorpus {
   protected ProjectedCorpus PC;
   protected KittyCache<Integer,double[]> featureCache;
+  public CachedProjectedCorpus( ProjectedCorpus PC, int cacheSize ) {
+    super( PC );
+    this.featureCache = new KittyCache<>( cacheSize );
+  }
   public CachedProjectedCorpus( ProjectedCorpus PC ) {
-    this.PC = PC;
-    this.featureCache = new KittyCache<>( 1000 );
+    this( PC, 1000 );
   }
 
   /**
@@ -31,7 +34,7 @@ public class CachedProjectedCorpus extends ProjectedCorpus {
     double[] x = featureCache.get(i);
     // Otherwise call PC to get it
     if( x == null ) {
-      x = PC.featurize( i );
+      x = featurize( i );
       featureCache.put(i, x, -1);
     }
 

### 2013-05-22 java/src/learning/data/Corpus.java

@@ -25,13 +25,13 @@ import org.ejml.simple.SimpleMatrix;
 /**
  * Stores a corpus in an integer array
  */
-public class Corpus implements Serializable {
+public class Corpus {
   private static final long serialVersionUID = 2L;
   /// Stores a list of unique vocabulary
   public String[] dict;
 
   /// Stores an integer representation of the corpus, with each
-  //sentence as a sequence of integer indices into words.
+  /// sentence as a sequence of integer indices into words.
   public int[][] C; 
 
   /// Keys for catchall classes
@@ -39,49 +39,19 @@ public class Corpus implements Serializable {
   public static final String LOWER_CLASS = "@LOWER@";
   public static final String UPPER_CLASS = "@UPPER@";
   public static final String MISC_CLASS = "@MISC@";
+  // TODO: Add support here for auxillary features defined in
+  // KirkpatrikBouchard2010
 
   protected Corpus() {}
+
+  public Corpus(Corpus C) {
+    this.dict = C.dict;
+    this.C = C.C;
+  }
 
   public Corpus( String[] dict, int[][] C ) {
     this.dict = dict;
     this.C = C;
-  }
-
-  /**
-   * Classify a word into the following categories: DIGIT, LOWER, UPPER, MISC
-   * @param word
-   * @return
-   */
-  protected static String classify( String word ) {
-    // If the word contains a digit, throw it in the digit bin
-    if( word.matches(".*[0-9].*") )
-      return DIGIT_CLASS;
-    else if( word.matches("^[A-Z][a-z]+") )
-      return UPPER_CLASS;
-    else if( word.matches("[a-z]+") )
-      return LOWER_CLASS;
-    else 
-      return MISC_CLASS;
-  }
-
-  /**
-   * Remove words which occur less than 'count' number of times,
-   */
-  protected static String[] pruneDictionary( HashMap<String,Integer> counts, int cutoff ) {
-    Vector<String> keys = new Vector<>();
-    for (Map.Entry<String, Integer> entry : counts.entrySet()) {
-      if( entry.getValue() > cutoff )
-        keys.add( entry.getKey() );
-    }
-    Collections.sort( keys );
-
-    // Add the sentinel words
-    keys.add( DIGIT_CLASS );
-    keys.add( LOWER_CLASS );
-    keys.add( UPPER_CLASS );
-    keys.add( MISC_CLASS );
-
-    return keys.toArray(new String[0]);
   }
 
[…]

### 2013-05-22 java/src/learning/data/CachedProjectedCorpus.java
Minor fixes to have MomentComputer working

@@ -34,7 +34,7 @@ public class CachedProjectedCorpus extends ProjectedCorpus {
     double[] x = featureCache.get(i);
     // Otherwise call PC to get it
     if( x == null ) {
-      x = featurize( i );
+      x = super.featurize( i );
       featureCache.put(i, x, -1);
     }
 

### 2013-05-22 java/src/learning/data/MomentComputer.java

@@ -163,7 +163,7 @@ public class MomentComputer implements Runnable {
     // Read corpus
     try {
       ProjectedCorpus PC = ProjectedCorpus.fromCorpus( 
-        Corpus.parseText( opts.mapPath, opts.dataPath ),
+        Corpus.parseText( opts.dataPath, opts.mapPath ),
         opts.randomProjDim, opts.randomProjSeed );
       
       // Compute moments

### 2013-05-22 java/src/learning/linalg/FullTensor.java

@@ -5,10 +5,12 @@ import org.ejml.simple.SimpleMatrix;
 import org.ejml.data.DenseMatrix64F;
 import org.javatuples.Triplet;
 
+import java.io.Serializable;
+
 /**
   A Tensor constructed with a full DxDxD matrix
  */
-public class FullTensor implements Tensor {
+public class FullTensor implements Tensor, Serializable {
   public double[][][] X;
   public int D1, D2, D3;
 

### 2013-05-22 java/src/learning/spectral/applications/HiddenMarkovModel.java

@@ -5,8 +5,7 @@
  */
 package learning.spectral.applications;
 
-import learning.exceptions.RecoveryFailure;
-import learning.spectral.MultiViewMixture;
+import learning.spectral.TensorMethod;
 import learning.data.Corpus;
 import learning.data.ProjectedCorpus;
 import learning.data.MomentComputer;
@@ -39,11 +38,14 @@ import java.lang.ClassNotFoundException;
 public class HiddenMarkovModel implements Runnable {
 
   public Triplet<SimpleMatrix, SimpleMatrix, SimpleMatrix>
-      run( SimpleMatrix M12, SimpleMatrix M23, SimpleMatrix M13, 
-          FullTensor M123 ) {
+      run( int K, 
+          SimpleMatrix M12, SimpleMatrix M13, SimpleMatrix M23, FullTensor M123 ) {
     // Tensor Factorize
-    //
+    TensorMethod algo = new TensorMethod();
+    Quartet<SimpleMatrix, SimpleMatrix, SimpleMatrix, SimpleMatrix> params = algo.recoverParameters( K, M12, M13, M23, M123 );
+
     // Return emission and transition probabilities.
+
     return null;
   }
 

### 2013-05-22 java/run.sh
BIG: Made several bugfixes to FullTensor (problem was with the indicing

@@ -7,5 +7,5 @@ for f in deps/*.jar; do
   deps=$f:$deps;
 done;
 
-$JAVA_PATH/java -ea -Xms20g -Xmx25g -cp $deps:bin $@
+$JAVA_PATH/java -ea -Xms6g -Xmx8g -cp $deps:bin $@
 

### 2013-05-22 java/src/learning/linalg/FullTensor.java

@@ -42,10 +42,11 @@ public class FullTensor implements Tensor {
 
     double[][][] X = new double[D1][D2][D3];
 
+    int idx = 0;
     for(int d1 = 0; d1 < D1; d1++) {
       for(int d2 = 0; d2 < D2; d2++) {
         for(int d3 = 0; d3 < D3; d3++) {
-          X[d1][d2][d3] = V.get( d1*(D2 + D3) + d2*(D3) + d3 );
+          X[d1][d2][d3] = V.get( idx++ );
         }
       }
     }
@@ -62,7 +63,7 @@ public class FullTensor implements Tensor {
       for(int d2 = 0; d2 < D2; d2++) {
         builder.append( "  { " );
         for(int d3 = 0; d3 < D3; d3++) {
-          builder.append( X[d1][d2][d3] + ", " );
+          builder.append( String.format( "%.4f", X[d1][d2][d3]) + ", " );
         }
         builder.append( " }\n" );
       }

### 2013-05-22 java/src/learning/models/MixtureOfExperts.java

@@ -380,16 +380,19 @@ public class MixtureOfExperts implements Serializable {
   }
 
   public static void printData( SimpleMatrix y, SimpleMatrix X, MixtureOfExperts model ) {
-    double[][] exponents = model.getNonLinearity().getExponents();
-
-    // Pretty print the exponents
-    System.out.printf( "# " );
-    for( double[] exp : exponents ) {
-      for( int d = 0; d < exp.length; d++ )
-        System.out.printf( "x_" + d + "^" + exp[d] );
-      System.out.printf( " " );
+    double[][] exponents = 
+      ( model != null ) ? model.getNonLinearity().getExponents()
+        : null;
+    if( exponents != null ) {
+      // Pretty print the exponents
+      System.out.printf( "# " );
+      for( double[] exp : exponents ) {
+        for( int d = 0; d < exp.length; d++ )
+          System.out.printf( "x_" + d + "^" + exp[d] );
+        System.out.printf( " " );
+      }
+      System.out.printf( "\n" );
     }
-    System.out.printf( "\n" );
 
     int N = X.numRows();
     for( int i = 0; i < N; i++ ) {
@@ -397,7 +400,6 @@ public class MixtureOfExperts implements Serializable {
         System.out.printf( "%f ", X.get(i,d) );
       System.out.printf( "%f\n", y.get(i) );
     }
-
   }
 
   /**

### 2013-05-22 java/tests/learning/spectral/applications/SpectralMomentsTests.java

@@ -1,0 +1,98 @@
+package learning.spectral.applications;
+
+import learning.exceptions.RecoveryFailure;
+import learning.linalg.*;
+import learning.exceptions.NumericalException;
+
+import learning.spectral.applications.SpectralExperts;
+
+import learning.models.MixtureOfExperts;
+
+import org.javatuples.*;
+import org.ejml.simple.SimpleMatrix;
+import org.ejml.data.DenseMatrix64F;
+import org.junit.Assert;
+import org.junit.Test;
+import org.junit.Before;
+
+import fig.basic.LogInfo;
+import fig.basic.Option;
+import fig.basic.OptionsParser;
+
+/**
+ * Tests for spectral experts
+ */
+public class SpectralMomentsTests {
+
+  @Before
+  public void setup(){
+    LogInfo.writeToStdout = false;
+    LogInfo.init();
+  }
+
+  public void arbitraryOptimizationMatchTest(SpectralExperts algo, MixtureOfExperts.GenerationOptions options) throws NumericalException {
+    LogInfo.writeToStdout = true;
+    LogInfo.init();
+
+    MixtureOfExperts model = MixtureOfExperts.generate(options);
+    int K = model.getK(); int D = model.getD();
+    algo.K = K;
+    algo.enableAnalysis(model);
+    if( algo.ridgeReg3 < 0 ) algo.ridgeReg3 = algo.ridgeReg2 * 10;
+    if( algo.traceReg3 < 0 ) algo.traceReg3 = algo.traceReg2 * 10;
+    if (algo.adjustReg) {
+      algo.traceReg2 /= Math.sqrt(N); //N; //Math.pow(N, 1.0/3);
+      algo.traceReg3 /= Math.sqrt(N); //N; //Math.pow(N, 1.0/3);
+    }
+
+    Pair<SimpleMatrix, FullTensor> moments = SpectralExperts.computeExactMoments( model );
+    SimpleMatrix Pairs = moments.getValue0();
+    FullTensor Triples = moments.getValue1();
+
+    Pair<SimpleMatrix, SimpleMatrix> yX = model.sample((int) N);
+    SimpleMatrix y = yX.getValue0();
+    SimpleMatrix X = yX.getValue1();
+
+    SimpleMatrix PairsM, PairsR;
+    FullTensor TriplesM, TriplesR;
+
+    // Do regression with CVX
+    {
[…]

### 2013-05-22 matlab/recoverB2.m

@@ -3,18 +3,24 @@ function [B2] = recoverB2( y, X, sigma2, lambda )
   [N, d] = size( X );
   y = y.^2 - sigma2;
 
-  cvx_begin sdp quiet;
+  % Construct the tensor form of X
+  X_ = zeros( N, d.^2 );
+  for n = 1:N
+    Xn = X(n,:);
+    X_(n,:) = vec(kron(Xn, Xn));
+  end
+
+  cvx_begin sdp;
     variables B2(d, d);
     variables W1(d, d);
     variables W2(d, d);
     variables t;
 
-    minimize ( lambda * t + 0.5/N * norm( y - diag( X * B2 * X' ) ) );
+    minimize ( lambda * t + 0.5/N * norm( y - X_ * vec(B2) ) );
     subject to 
-      B2 == semidefinite( d );
       0.5 * trace(W1) + 0.5 * trace(W2) <= t;
-      [W1, B2; B2', W2] == semidefinite(2*d);
-      sigma2 >= 0;
+      [W1, B2; B2', W2] == semidefinite(d + d);
   cvx_end;
+
 end
 

### 2013-05-22 matlab/recoverB3.m

@@ -1,8 +1,8 @@
 % Recover the second moments, B3.
 function B = recoverB3( y, X, sigma2, lambda )
   [N, d] = size( X );
-  avgBetas = X\y;
-  y = y.^3 - 3 * sigma2 * X * avgBetas;
+  avgBetas = X\y; % y_i = x_i^T \beta
+  y = y.^3 - 3 * sigma2 * (X * avgBetas);
 
   % Construct the tensor form of X
   X_ = zeros( N, d.^3 );
@@ -10,6 +10,7 @@ function B = recoverB3( y, X, sigma2, lambda )
     Xn = X(n,:);
     X_(n,:) = vec( tensor([d,d,d], 'unit', Xn ) );
   end
+  X = X_;
 
   cvx_begin sdp;
     variables B(d, d, d);
@@ -24,7 +25,12 @@ function B = recoverB3( y, X, sigma2, lambda )
     variables t2;
     variables t3;
 
-    minimize ( lambda * (t1+t2+t3)/3 + 0.5/N * norm( y - X_ * vec(B) ) );
+    %minimize ( 0.5/N * norm( y - X_ * vec(B) ) );
+    %subject to
+    %  B == shiftdim( B, 1 );
+    %  B == shiftdim( B, 2 );
+
+    minimize ( lambda * (t1 + t2 + t3)/3 + 0.5/N * norm( y - X_ * vec(B) ) );
     subject to
       0.5 * trace(W11) + 0.5 * trace(W12) <= t1;
       [W11, mode_unfold(B,1); mode_unfold(B,1)', W12] == semidefinite(d.^2 + d);

### 2013-05-22 matlab/sdpB2.m

@@ -5,30 +5,8 @@ function sdpB2(path)
   X = load( strcat(path,'/X.txt'), '-ascii' );
   sigma2 = load( strcat(path,'/sigma2.txt'), '-ascii' );
   lambda = load( strcat(path,'/lambda2.txt'), '-ascii' );
-  
-  [N, d] = size( X );
 
-  y = y.^2 - sigma2;
-  % Construct the tensor form of X
-  X_ = zeros( N, d.^2 );
-  for n = 1:N
-    Xn = X(n,:);
-    X_(n,:) = vec(kron(Xn, Xn));
-  end
-  X = X_;
-
-  cvx_begin sdp;
-    variables B2(d, d);
-    variables W1(d, d);
-    variables W2(d, d);
-    variables t;
-
-    minimize ( lambda * t + 0.5/N * norm( y - X * vec(B2) ) );
-    subject to 
-      B2 == semidefinite( d );
-      0.5 * trace(W1) + 0.5 * trace(W2) <= t;
-      [W1, B2; B2', W2] == semidefinite(d + d);
-  cvx_end;
+  B2 = recoverB2(y, X, sigma2, lambda);
 
   save( strcat(path,'/B2.txt'), '-ascii', 'B2' );
 end

### 2013-05-22 matlab/uncvec.m

@@ -1,0 +1,21 @@
+% Recover the second moments, B3.
+function B = uncvec(v, d)
+  B = zeros(d,d,d);
+  D = d * (d+1) * (d+2)/6;
+
+  idx = 1;
+  for d1 = 1:d
+    for d2 = 1:d1
+      for d3 = 1:d2
+        B(d1,d2,d3) = v(idx);
+        B(d1,d3,d2) = v(idx);
+        B(d2,d1,d3) = v(idx);
+        B(d2,d3,d1) = v(idx);
+        B(d3,d2,d1) = v(idx);
+        B(d3,d1,d2) = v(idx);
+        idx = idx + 1;
+      end
+    end
+  end
+end
+

### 2013-05-22 doc/icml2013/algo.tex

@@
 Now we can express $y$ as a linear function of $x$ with non-random coefficients $M_1$
-plus some a new noise term $\eta_1(x)$:
+plus a noise term $\eta_1(x)$:
 \begin{align}
@@
 focused on observable operator models \cite{hsu09spectral,song10kernel,parikh12spectral,cohen12pcfg,balle11transducer,balle12automata}
-in which a reparametrization of the true parameters are recovered,
+in which a re-parametrization of the true parameters are recovered,
 which suffices for prediction and density estimation.

### 2013-05-22 doc/icml2013/model.tex

@@
 (ii) draw observation noise $\epsilon$ from a known zero-mean noise distribution $\sE$,
-and (ii) set $y$ deterministically based on $h$ and $\epsilon$.
+and (iii) set $y$ deterministically based on $h$ and $\epsilon$.
 More compactly: %to $\beta_h^\top x$ plus some observation noise $\epsilon$.
@@
 The learning problem is stated as follows:
-given $n$ i.i.d. samples $(\xn{1}, \yn{1}), \dots, (\xn{n}, \yn{n})$
+given $n$ i.i.d.\ samples $(\xn{1}, \yn{1}), \dots, (\xn{n}, \yn{n})$
 drawn from the model with some unknown parameters $\theta^*$,
@@
 The mixture of linear regressions model has been applied
-in the statistics literature for modeling music perception, where $x$ is the
+in the statistics literature for modelling music perception, where $x$ is the
 actual tone and $y$ is the tone perceived by a musician \cite{VieleTong2002}.

### 2013-05-22 doc/nips2013/introduction.tex

\section{Introduction} \label{sec:introduction}

% Goal: parameter estimation in latent-variable models
Latent-variable log-linear models provide two complementary strengths: (i) allowing
latent representations to be automatically learned from data, and (ii)
allowing domain knowledge to be encoded in the form of features
to influence learning when appropriate.
Despite their success across many fields \cite{quattoni04crf,haghighi06prototype,liang06discrimative,kirkpatrick10painless,XXX},
learning these models remains a difficult problem due to the non-convexity of the likelihood.
Local optimization is the standard approach but is susceptible to local optima.

% Method of moments
Recently, unsupervised learning techniques based on the method of moments and
spectral decomposition have offered a refreshing and promising perspective on
this learning problem \citep{hsu09spectral,anandkumar11tree,anandkumar12moments,anandkumar12lda,hsu12identifiability,balle11transducer,balle12automata}.
These methods exploit the linear algebraic properties of the model and
factorize the moments into parameters, providing strong theoretical guarantees.
However, these methods are not as easy to apply as EM, and have yet to be
developed for log-linear models.

% This paper
In this work, we develop a technique for parameter estimation in log-linear
models, shown in \reffig{schema}.  The key idea in
\citet{anandkumar12moments,anandkumar13tensor} is to study the conditional
independence factorization structure implied by the model and identify
bottlenecks.  We construct third-order tensors around these bottlenecks, which
are factorized in our case into moments involving the latent variables.

\Fig{figures/schema}{0.3}{schema}{Schema.}

This paper contributes three ideas:
First, in certain cases, these latent moments are exactly the sufficient statistics of
the log-linear model, in which we case use convex optimization directly to obtain the parameters,
resulting in a consistent parameter estimation procedure.
Second, in other cases, the latent moments only provide partial information about the parameters,
but we show that this information can be naturally incorporated into the measurements framework of \citet{liang09measurements}.
This provides a natural way to leverage the method of moments when providing an
exact consistent estimator is difficult.
Finally, for factorial models, we introduce a new \emph{unshuffling
factorization} procedure to recover parameter estimates.

### 2013-05-31 doc/nips2013/abstract.tex
added pdf figures

--- 2013-05-21 doc/nips2013/abstract.tex
@@
 \begin{abstract}
-Latent-variable models.
+Parameter estimation in log-linear latent-variable models is typically
+performed using local optimization, which is prone to local optima.
+In this paper,
+%we show how to leverage recent advances in the method of moments
+%to alleviate these problems.
+we leverage recent advances in the method of moments to estimate latent
+moments of the target distribution,
+which we incorporate as constraints on the original optimization
+problem.
+In favorable cases, these constraints provide full information and render the
+original optimization problem
+convex; in other cases, the constraints offer partial information,
+but still provide significant aid to the original optimization problem.
+Factorial models do require special attention,
+so we develop a new unshuffling factorization technique to induce
+constraints for these models.
+Empirically, we show that our methods provide substantial gains
+across several models.
 \end{abstract}

### 2013-05-31 doc/nips2013/utils.js

@@ -38,7 +38,7 @@ G.schema = function() {
   style = function(x) { return x.strokeWidth(10).color('brown'); }
   return overlay(
     table(
-      [boxed('Data $x$'), nil(), boxed('Parameters $\\theta$')],
+      [boxed('Data $x^{(1)}, \\dots, x^{(n)}$'), nil(), boxed('Parameters $\\theta$')],
       [a1 = style(downArrow(s=70)), nil(), a3 = style(upArrow(s))],
       [boxed('Observed moments $\\E[\\phix(x)]$'), a2 = style(rightArrow(200)), boxed('Latent moments $\\E[\\phi(x,h) \\mid h]$')],
     _).margin(50, 30).center(),

### 2013-05-31 doc/nips2013/factorial.tex

@@
-\section{Factorial models} \label{sec:factorialModels}
+\section{Unshuffling factorization for factorial mixture models} \label{sec:factorialModels}
 
-So far, we have relied on the existence of bottleneck hidden variables $h_j$ on which
+So far, we have relied on the existence of a bottleneck hidden variable $h_j$ on which
 we have three conditionally independent views.  Such bottlenecks do not always
 exist; for example, in \reffig{factorialModels}(a), neither $h_1$ or $h_2$ is a bottleneck.
-Our solution is to create a composite latent variable out of existing latent
-variables.  We can invoke $\TensorFactorize$ to obtain composite marginals and
+Our solution is to create a composite latent variable $c$ which is in
+one-to-one correspondence with the existing latent variables $(h_1,h_2)$.
+We can invoke $\TensorFactorize$ to obtain composite marginals and
 conditional means.  However, the identity of $h_1$ and $h_2$ becomes lost in the composite,
@@
 
-\Fig{figures/factorialModels}{0.3}{factorialModels}{Examples of factorial models.
-There do not exist three-view bottlenecks, so we first identify composite bottlenecks,
-and then perform an \emph{unshuffling factorization}.}
+\Fig{figures/factorialModels}{0.3}{factorialModels}{The factorial mixture model (a)
+does not have three-view bottlenecks, so we first identify composite bottlenecks,
+and then perform an \emph{unshuffling factorization} (b):
+Given known composite conditional means $B_c = \BP[x_j \mid c] \in \R^\ell$,
+which are derived from combinations of unknown parameters $f_a = \theta_1(\cdot, a), g_b = \theta_2(\cdot, b)$,
+we group differences into bins (dashed boxes), and maximal bins into sources (red rounded boxes),
+producing a linear program with variables $f_a,g_b,Z_{ab}$ and constraints for each $B_c$.
+}
 
@@
 
-Let $B \eqdef (p_\theta(x_j \mid h))_{x_j \in [\nphix],h \in [\nh]^s} \in
-\R^{\nphix \times \nh^s}$ be the conditional mean.
+Let $B \eqdef (\BP[x_j \mid c])_{x_j \in [\nphix],c \in [\nh]^s} \in
+\R^{\nphix \times \nh^s}$ be the conditional mean of $x_j$ given composite cluster $c$.
 
@@
 
-We will measure $M_2 \eqdef \E[x_1 \otimes x_2]$ and
-$M_3 \eqdef \E[x_1 \otimes x_2 \otimes x_3]$.
-We run $\TensorFactorize$, which will produce $B = (\E[x_j \mid c])$, where the
-cluster $c$ ranges over $k^s$ possible clusters.  Next, we still have to
-extract the actual parameters $\theta$ from $B$.  In the simple mixture model,
-the clusters produced by $\TensorFactorize$ were in exact one-to-one correspondence
+First, we run $\TensorFactorize$ on estimates of
+$M_2 \eqdef \E[x_1 \otimes x_2]$ and $M_3 \eqdef \E[x_1 \otimes x_2 \otimes x_3]$,
+which produces $B = (\E[x_j \mid c])_{c \in [\nh]^s}$.
+%Next, we still have to extract the actual parameters $\theta$ from $B$. 
+In the simple mixture model,
+the clusters produced by $\TensorFactorize$ are in exact one-to-one correspondence
 with the actual mixture components, but in the factorial case, we must solve
-the \emph{credit assignment problem} to the $s$ sources.
+a \emph{credit assignment problem} to match combinations of $s$ sources
+with columns of $B$.
 
-Let $C$ be the elementwise logarithm of $B$; that is
+Let $\{ v_{ia} \}_{i \in [s], a \in [k]}$, where $v_{ia} = \theta_i(\cdot, a) \in \R^\ell$,
+be the collection of parameter vectors that we seek to recover.
+Let $C \in \R^{\ell \times k^s}$ be the elementwise logarithm of $B$; that is
 $C(x_j, h) = \log B(x_j, h)$.
+The columns of $C$ are exactly the unordered collection of vectors
+$\{ \sum_{i=1}^s v_{ia_i} - Z_{\ba} : \ba = (a_1, \dots, a_s) \in [\nh]^s \}$,
+where $Z_{\ba} = \log \sum_{e=1}^\ell (\sum_{i=1}^s \exp(v_{ia_i}(e)))$ is the normalization constant (a scalar).
 
-Let $v_{ia} = \theta_i(\cdot, a) \in \R^\ell$ be the collection of parameter vectors that we seek to recover.
-What we observe are normalized sums of these vectors.
-Specifically, we observe the unordered collection of vectors $\{ \sum_{i=1}^s v_{ia_i} - Z_{\ba} : \ba = (a_1, \dots, a_s) \in [\nh]^s \}$,
-where $Z_{\ba} = \log \sum_{e=1}^\ell (\sum_{i=1}^s \exp(v_{ia_i}(e)))$ is the normalization constant.
+Note that there is a combinatorial dimension to this problem since the association between
+columns of $C$ and combinations of $\{ v_{ia} \}$ is unknown.
+As a running example (\reffig{factorialModels}(b)),
+consider $s=2$ sources, each taking on $k=3$ values (let $f_a = v_{1a}$ and $g_b = v_{2b}$).
+The following algorithm creates a linear program which puts constraints on variables $\{v_{ia}\}$ and $Z_{\ba}$:
+\begin{enumerate}
 
-We say two vectors are \emph{comparable} if they differ by a constant times $\bone$.
+\item Compute the difference vector $B_i - B_j$ for every distinct pair of columns $c,d \in [\nh^s]$.
 
-Here's the algorithm:
-\begin{itemize}
-  \item For each pair of vectors $(i,j) \in [\nh^s]^2$, consider the difference $B_i - B_j$.
-  \item Put the pairs into bins, where each bin has differences which are comparable.
-  \item Keep only the bins with the maximum number of pairs (which should be $\nh$).
-  Now, each bin $b$ corresponds to a signature $v_{ia} - v_{ia'}$ for some source $i$ and two values $a$ and $a'$.
-  \item Group the bins into sources.
-  Construct a graph with the bins as nodes as follows.
-  Let two bins be connected if the difference of their signatures is the signature of some other bin.
-  For example, bin $b_1$ with signature $v_{ia} - v_{ia'}$
-  is connected to bin $b_2$ with signature $v_{ia''} - v_{ia'}$
-  because some bin $b_3$ has signature $v_{ia} - v_{ia''}$ by construction of the bins.
-  The connected components of graph are exactly the sources.
-  \item For each connected component, identify a subset of the signatures which all have a common difference:
-  $v_{ia_1} - v_{ia'}$, $v_{ia_2} - v_{ia'}$, $v_{ia_3} - v_{ia'}$, etc.
-  \item Construct a linear program.  It will have $\nh^s \cdot d$ constraints, one for each entry of $B$.
+\item Put the pairs $(c,d)$ into bins, where the difference vectors $B_c-B_d$
+in each bin are \emph{comparable}, by which we mean that they
+differ by a constant scalar times $\bone$ (for the normalization constant).
+
+\item Keep only the maximal bins.
+These bins should contain $\nh^{s-1}$ pairs, each corresponding to a fixed setting of the $s-1$ sources
+and changing the value of one source.
+Specifically, the difference vector is equal to $v_{ia} - v_{ia'}$ up to a
+constant scalar for some source $i \in [k]$ and distinct values $a, a' \in
+[\nh]$.
+
+\item Group the bins into sources:
+  For each source $i = 1, \dots, s$,
+  chose any bin, declare it as $v_{i1} - v_{i2}$, and add it to the source.
+  Let $\gamma$ be the set of $c$ such that that bin contains some difference vector $B_c - B_d$,
+  and include any other bin with the same $\gamma$; these should correspond to $v_{i1} - v_{ia}$ for $a \ge 3$.
+  Then throw away any bin whose difference vector is $-(B_c - B_d)$
+  or is comparable to the difference of two existing difference vectors (e.g., $v_{i2} - v_{i1}$ or $v_{i2} - v_{i3}$).
+  The resulting source should have $\nh-1$ bins, corresponding to $\{ v_{i1} - v_{ia} : 2 \le a \le \nh \}$.
+  At this point, we have identified which vectors contribute to each $B_c$,
+  taking care of the combinatorial aspect of the problem,
+  but have yet to recover the actual vectors.
+
+  \item To do this last step, construct a linear program with $\nh^s \cdot d$ constraints, one for each entry of $B$.
   It will have $s\nh \cdot d + \nh^s$ variables,
-  one for each of $s\nh$ vectors $v_{ia}$ and $\nh^s$ for the normalization constant
-  (these are actually deterministic functions of the vectors, but non-linear, so let's linearize).
-  TODO: finish explaining this.
-\end{itemize}
+  one for each of $s\nh$ vectors $v_{ia}$ and $\nh^s$ for the normalization constants $Z_\ba$.
+  Note that $Z_\ba$ is actually a determinstic function of the the vectors, but since it is a non-linear one,
+  it is difficult to work with; therefore, we create a separate variable to keep the problem linear.
+
+  It turns out that this is LP is rank deficient for two reasons, the first intrinsic to our model,
+  and the second due to loose linearization.
+  First, the vectors $v_{ia}$ are non-identifiable since we can always add any vector $u \in \R^\ell$ to
+  all the vectors of a source $i_1$ ($v_{i_1 a}$ for all $a \in [\nh]$)
+  subtract $u$ from all the vectors of another source $v_{i_2 a}$ for all $a \in [\nh]$.
+  Second, if we knew the normalization constants $\{Z_\ba\}$, then this is only rank deficiency.
+  However, treating $\{Z_\ba\}$ as independent variables creates additional rank deficiency.
+
+  Despite this, in the spirit of providing partial information,
+  $\UnshuffleFactorize$ still gives us linear constraints on the parameters.
+  In particular, we propose adding to the likelihood objective a strong regularization term
+  $R(\theta, Z) = \frac{\lambda}{2} (\bA_1 \theta - \bA_2 Z - B)^2$, where $\bA_1$ and $\bA_2$ are coefficients
+  recovered by unshuffling.
+  In particular, we propose optimizing $\min_{\theta, Z} \sum_{i=1}^n -\log p_\theta(x^{(i)}) + R(\theta, Z)$.
+  Although the resulting objective is still non-convex, the regularization imposes quite strong constraints,
+  and the optimization would happen effectively over a much lower-dimensional subspace.
+\end{enumerate}

### 2013-05-31 doc/nips2013/introduction.tex

@@
 % Goal: parameter estimation in latent-variable models
-Latent-variable log-linear models provide two complementary strengths: (i) allowing
+Latent-variable log-linear models provide two complementary strengths: (i) they allow
 latent representations to be automatically learned from data, and (ii)
-allowing domain knowledge to be encoded in the form of features
-to influence learning when appropriate.
-Despite their success across many fields \cite{quattoni04crf,haghighi06prototype,liang06discrimative,kirkpatrick10painless,XXX},
+they allow domain knowledge to be encoded in the form of features to influence learning when appropriate.
+Despite their success across many fields \cite{quattoni04crf,haghighi06prototype,liang06discrimative,kirkpatrick10painless,deselaers12latent},
 learning these models remains a difficult problem due to the non-convexity of the likelihood.

### 2013-05-31 doc/nips2013/abstract.tex
made a pass up through section 4 (measurements)

--- 2013-05-21 doc/nips2013/abstract.tex
@@
 \begin{abstract}
-Latent-variable models.
+Parameter estimation in log-linear latent-variable models is typically
+performed using gradient-based optimization on the log-likelihood, which is prone to local optima.
+In this paper,
+%we show how to leverage recent advances in the method of moments
+%to alleviate these problems.
+we leverage recent advances in the method of moments to estimate latent
+moments of the target distribution,
+which we incorporate as constraints on the original optimization
+problem.
+In favorable cases, these constraints provide full information and render the
+original optimization problem
+convex; in other cases, the constraints offer only partial information,
+but still appear to provide significant aid to the original optimization
+problem.
+Factorial models do require special attention;
+to handle these models,
+we introduce a new technique called unshuffling factorization to induce
+linear constraints on the parameters.
+Empirically, our approach provides improvements on several models.
 \end{abstract}

### 2013-05-31 doc/nips2013/introduction.tex
made a pass up through section 4 (measurements)

@@
 factorize the moments into parameters, providing strong theoretical guarantees.
-However, these methods are not as easy to apply as EM, and have yet to be
-developed for log-linear models.
+However, these methods are not as universally applicable as EM, and have yet to
+be developed for log-linear models.
+
+\Fig{figures/schema}{0.3}{schema}{Schema of our approach.}
 
@@
 
-\Fig{figures/schema}{0.3}{schema}{Schema.}
-
 This paper contributes three ideas:
@@
 the log-linear model, in which we case use convex optimization directly to obtain the parameters,
-resulting in a consistent parameter estimation procedure.
+resulting in a consistent parameter estimation procedure (\refsec{threeViewMixtureModel}).
 Second, in other cases, the latent moments only provide partial information about the parameters,
-but we show that this information can be naturally incorporated into the measurements framework of \citet{liang09measurements}.
-This provides a natural way to leverage the method of moments when providing an
-exact consistent estimator is difficult.
+but we show that this information can be naturally incorporated
+using the measurements framework of \citet{liang09measurements} (\refsec{generalModels}).
+This provides a natural way to leverage the method of moments when constructing
+a consistent estimator is difficult.
 Finally, for factorial models, we introduce a new \emph{unshuffling
-factorization} procedure to recover parameter estimates.
+factorization} procedure to recover parameter estimates (\refsec{factorialModels}).

### 2013-05-31 doc/nips2013/measurements.tex
made a pass up through section 4 (measurements)

@@
 
-Our learning algorithm for three-view mixture models relies on the fact that the hidden moments we can recover
-are exactly the sufficient statistics of the exponential family model.
+Our learning algorithm for three-view mixture models relies on the fact that
+the sufficient statistics of the exponential family
+can be expressed as a function of the latent moments.
 Of course, these conditions do not hold for general log-linear models.
-Rather than falling back on local optimization of the likelihood in these cases, we would still like
-to use the \emph{partial information} offered by the hidden moments provide we
-\emph{can} recover in some way.  We will show how to incorporate these moments
-into a likelihood-based objective.
-
+Rather than falling back on plain local optimization of the likelihood in these
+cases, we would still like
+to somehow use the \emph{partial information} offered by the latent moments.
+%We will show how to incorporate these moments into a likelihood-based
+%objective.
 Partial information is available when the following holds:
@@
   and three conditionally independent observed variables $x_{j_1},x_{j_2},x_{j_3}$ (views) such that
-  each conditional mean matrix $\E[\phix(x_{j}) \mid h_{j_0}] \in \R^{\nphix \times \nh}$ ($j \in \{ j_1, j_2, j_3 \}$) has full column rank $\nh$,
+  each conditional mean matrix $\E[\phix(x_{j}) \mid h_{j_0}] \in \R^{\nphix \times \nh}$ ($j \in \{ j_1, j_2, j_3 \}$) has full column rank $\nh$
   for some transformation $\phix$.
@@
 \reffig{generalModels} shows examples of log-linear models.
-For simplicitly, assume each hidden variable $h_j \in [\nh]$ and each observed
-variable $x_j \in \R^\ell$ are indicator vectors encoding one of $\ell$ possible values.
+For notational simplicitly, assume each latent variable $h_j \in [\nh]$ is an integer
+and each observed variable $x_j \in \R^\ell$ is an indicator vector encoding
+one of $\ell$
+possible values.
 
@@
 Three examples of undirected log-linear models.  In each of these cases,
-we can recover hidden moments, which are themselves not sufficient to identify the parameters,
+we can recover latent moments, which are themselves not sufficient to identify the parameters,
 but provide partial information which leads to a more well-behaved likelihood function.
@@
 Following the reduction in \citet{anandkumar12moments},
-take $h_2$ as the bottleneck and $x_1, x_2, x_3$ as the three views.
+we could take $h_2$ as the bottleneck and $x_1, x_2, x_3$ as the three views.
 Define $\pi = (\BP[h_1]) \in \R^{\nh}$,
@@
 Note that we can recover $T = O^\dagger(OT)$ and $\pi = T^{-1} (T\pi)$,
-from which we can compute the hidden moments/sufficient statistics $\mu = \E[\phi(x,h)]$.
+from which we can compute the latent moments/sufficient statistics $\mu = \E[\phi(x,h)]$.
 
@@
 Let us consider a new reduction:
-Let $h_j$ be the bottleneck where $j$ is chosen uniformly at random over $\{2,
-\dots, T-1\}$, where $L$ is the number of time steps; let $x_{j-1},x_j,x_{j+1}$ be the three views.\footnote{
+Let $h_j$ be the bottleneck where $j$ is a \emph{random} variable
+taking on values uniformly from $\{2,
+\dots, L-1\}$, where $L$ is the number of time steps.
+Let $x_{j-1},x_j,x_{j+1}$ be the three views.\footnote{
 When we actually estimate the observed moments, we simply average over the choice of $j$.}
 Now, note that we can recover $O = \inv{L-2} \sum_{j=2}^{L-1} \E[x_j \mid h_j]$ from the second view,
-but using the data more fully.  The first and third views are unfortunately now
-quite complex due to marginalization over the rest of the variables.
+but using the data more fully.  The conditional means of the first and third
+views are now a complex function of the parameters.
+%due to marginalization over the rest of the variables,
+However, the point is that we can get most of our utility out of the partial
+information offered by $O$ and ignore information which is hard to ingest.
 
-\paragraph{(b) Tall mixture model}
+\paragraph{(b) Grid model}
+Analogous to the second HMM reduction, let $h_{ij}$ be the bottleneck,
+where $(i,j)$ is chosen randomly.
+Let $x_{ij}^a,x_{ij}^b$ be two of the views, and any other observed variable as the third view.
+$\TensorFactorize$ would reveal conditional means $\E[x_{ij}^a \mid h_{ij}]$
+and $\E[x_{ij}^b \mid h_{ij}]$,
+and a third quantity that involves the latent-to-latent features and the observation features.
+This quantity involves marginalizing out the rest of the graph,
+and therefore difficult to use.
+%to extract something useful like $\BP[h_{ij}, h_{i(j+1)}]$
+%in closed form.
+%Note that the analog of this was still possible in the first HMM reduction that
+%used only the first three observations, but not in the second one.
+
+\paragraph{(c) Tall mixture model}
 Let $h_0$ be the bottleneck and $x_1,x_2,x_3$ be the three
-views.  The conditional means that $\TensorFactorize$ reveals are $\E[x_j \mid h_0]$, which are not even
-expectations over features in the original model, let alone sufficient
-statistics.\footnote{Note that if the features were arbitrary, then the
-parameters would not be identifiable, but because log-linear models permit
-parameter sharing, then identifiability is still possible and therefore
+views.
+$\TensorFactorize$ reveals the conditional means $\E[x_j \mid h_0]$, which are
+not even expectations over features in the original model, let alone sufficient
+statistics.\footnote{Note that
+the model might not be identifiable in general \cite{anandkumar11tree}.
+However, the model is identifiable with adequate parameter sharing,
+and therefore
 parameter estimation in this model is still meaningful.}
-
-\paragraph{(c) Grid model}
-As in the HMM example, choose a random $h_{ij}$ as the bottleneck,
-$x_{ij}^a,x_{ij}^b$ as two of the views, and any other observed variable as the third view.
-$\TensorFactorize$ would reveal conditional means $\E[x_{ij}^a \mid h_{ij}]$ and $\E[x_{ij}^b \mid h_{ij}]$,
-and a third quantity that involves the hidden-to-hidden features and the observation features.
-Unfortunately, this quantity involves marginalizing out the rest of the graph,
-and therefore impossible to extract something useful like $\BP[h_{ij}, h_{i(j+1)}]$
-in closed form.
-Note that the analog of this was still possible in the first HMM reduction that used only the first three observations,
-but not in the second one.
 
@@
 and some conditional mean matrix $B = \E[x_{j'} \mid h_j] \in \R^{\nphix \times \nh}$,
-but that it was insufficient to identify the parameters.  How can we still use this partial information $(\pi,B)$?
+but these are insufficient to identify the parameters.
+How can we still use this partial information $(\pi,B)$?
 To this end, we adopt the measurements framework of \citet{liang09measurements},
 which allows us to incorporate moment constraints into a likelihood-based objective in a coherent way.
-The basic idea is to define two families of distributions: $\sQ$ which are
-consistent with all information and $\sP$ the exponential family.  The
-framework essentially tries to choose $q \in \sQ$ and $p \in \sP$ to minimize
+The basic idea is to define two families of distributions:
+$\sQ$, which are consistent with all information; and
+$\sP$, the exponential family.  The
+framework tries to choose $q \in \sQ$ and $p \in \sP$ to minimize
 $\KL{q}{p}$.
@@
 let $\bx = (x^{(1)}, \dots, x^{(n)})$ be the vector of observed variables
+and let
 $\bh = (h^{(1)}, \dots, h^{(n)})$ be the vector of latent variables.
@@
 
-We introduce the following constraints on each $q \in \sQ$:\footnote{We use a hard constraint
+Define $\sQ$ to be $q$ satisfying the following constraints:\footnote{We use a hard constraint
 for expository clarity, but in practice, we use a quadratic penalty.  See

### 2013-05-31 doc/nips2013/mixture.tex
First cut at the experiments section
made a pass up through section 4 (measurements)

@@
 
-In this section, we consider the simplest model to illustrate the form of our method.
-Consider the undirected mixture model model in \reffig{simpleModels}(a).
+In this section, we consider the simplest model to illustrate the form of our
+method.  Consider the undirected mixture model model in \reffig{simpleModels}(b).
 We have one discrete latent variable $h \in [\nh]$ and
@@
 
-In the following, we will detail the algorithm for producing a parameter
-estimate $D$.  For simplicity of exposition, we will assume we have an infinite
-amount of data and work with the population quantities.  Actual parameter
-estimates are constructed by simply plugging in empirical quantities.
+We will now describe the algorithm for producing a parameter
+estimate $\hat\theta$.  For simplicity of exposition, we will assume we have an infinite
+amount of data and work with population quantities.  Actual parameter
+estimates are constructed by simply plugging in the empirical versions.
 
@@
 
-Rather than directly trying to connect the parameters $\theta$ to the data, we start
-bottom-up by considering what properties of the data we can measure.
-Let $\phix : \sX \mapsto \R^\ell$ denote a \emph{transformation function} (the requirement will be specified later),
-which will extract the properties of the observed variables $x$ relevant to $h$.
+%Rather than directly trying to connect the parameters $\theta$ to the data $D$,
+Let us start bottom-up by simply computing properties of the data.
+Let $\phix : \sX \mapsto \R^\ell$ denote a \emph{transformation function}
+(the requirement on $t$ will be specified later),
+which aims to extract properties of $x$ relevant to $h$.
 Define the \emph{observed moments} (analogous to those used in \cite{anandkumar12moments}):
@@
 
-\subsection{Step 2: Recover hidden moments (factorization)}
+\subsection{Step 2: Recover latent moments (factorization)}
 
@@
 moments $M_2,M_3$ (which only reference the observed variables) admit a
-factorization in terms of the \emph{hidden moments}
-(which provides a window into the hidden variables).
+factorization in terms of the \emph{latent moments}
+(which provides a window into the latent variables) \cite{anandkumar12moments}.
 
 Define
-$\pi \eqdef (\BP[h=1], \dots, \BP[h=\nh])$ to be the marginal probabilities over the latent variable $h$,
+$\pi \eqdef (\BP[h=1], \dots, \BP[h=\nh])$ to be the marginal probabilities
+over $h$,
 and define the matrix $B \eqdef (\E[\phix(x_j) \mid h])_{h \in [\nh]} \in \R^{\ell \times k}$,
@@
 
-Now, we can express the observed moments $(M_2,M_3)$ in terms of the hidden moments $(\pi,B)$:
+Now we can express the observed moments $(M_2,M_3)$ in terms of the latent moments $(\pi,B)$:
 \begin{align}
@@
 \begin{assumption}
-  The hidden moments matrix $B \in \R^{\nphix \times \nh}$ has full column rank $\nh$.
+  The latent moments matrix $B \in \R^{\nphix \times \nh}$ has full column rank $\nh$.
 \end{assumption}
@@
 which converts consistent estimates of $(M_2,M_3)$ into consistent estimates of $(\pi, B)$ (up to permutation of the columns).
-In brief, algorithm first computes a whitening matrix $W \in \R^{\nphix \times \nh}$ such that $W^\top M_2 W = I$,
-and uses it to whiten $M_3$.  Then a robust tensor power method is applied to extract the eigenvectors of the whitened $M_3$,
-which correspond to columns of $B$.
+In brief, the algorithm first computes a whitening matrix $W \in \R^{\nphix \times \nh}$ such that $W^\top M_2 W = I_{k \times k}$,
+and uses it to whiten $M_3$.  Then a robust tensor power method is applied to
+extract the eigenvectors of the whitened $M_3$;
+unwhitening yields the columns of $B$.
 %\begin{lemma}[tensor factorization \citep{anandkumar13tensor}]
@@
 
-We need to specify some conditions on the transformation function $\phix$
+We need to specify conditions on the transformation function $\phix$
 to ensure that enough information is retained to pinpoint the parameters later.
@@
 \end{property}
-Here, $C$ provide a view on the local features $\phi_L(x_j,h)$ which is more
-directly related to the model parameters \refeqn{mixtureModel}, whereas $B$ is
-what we have recovered so far.  Note that we couldn't recover $C$ directly because it depends on $h$.
+Here, $C$ provide a view on the local features $\philocal(x_j,h)$,
+which are
+closely related to the model parameters \refeqn{mixtureModel},
+and $B$ is what we have recovered so far.
+%Note that we couldn't recover $C$ directly because it depends on $h$.
 
-However, it is not difficult to construct $\phix$ such that Property~\ref{property:transformation} holds.
-Since $h$ is discrete, we can always take $\phix(x) = \flatten((\philocal(x, h))_{h \in [\nh]}) \in \R^{dk}$,
-which contains $k$ blocks, one for each latent state.  We can recover a column $C_h = \E[\philocal(x_j, h) \mid h]$
-by taking the block of column $B_h = \E[\flatten(\philocal(x_j, h'))_{h' \in [\nh]} \mid h]$
-corresponding to $h = h'$.
-Often, the local features are simply the tensor product of some function $g(x)$ and an indicator on $h$:
-$\philocal(x,h) = \flatten(g(x) \otimes e_h)$.
-In this case, it suffices to take $\phix(x) = g(x)$.
-
-Finally, having retrieved $C_h$, it is easy to compute the \emph{hidden moments} $\mu$:
-\begin{align}
-  \mu \eqdef \E[\phi(x,h)] = \E[\phiprior(h)] + \sum_{j=1}^3 \E[\philocal(x_j,h)] = \sum_{h \in [\nh]} \pi_h (\phi_p(h) + 3 C_h).
-\end{align}
-In \cite{anandkumar12moments}, these hidden moments would be exactly the parameters (multinomial probabilities), but
-for log-linear models, we still need to do more work to extract the parameters $\theta$.
+It is easy to construct $\phix$ such that Property~\ref{property:transformation} holds.
+Since $h$ is discrete,
+we can always take $\phix(x) = \flatten((\philocal(x, h))_{h \in [\nh]}) \in \R^{dk}$,
+which contains $k$ blocks, one for each latent state.
+Then each column $C_h \in \R^d$ is simply the $h$-th block
+of the column $B_h \in \R^{kd}$.
+%We can recover a column $C_h = \E[\philocal(x_j, h) \mid h]$
+%Each column $C_h = \E[\philocal(x_j, h) \mid h]$
+%by taking the block of column $B_h = \E[\flatten(\philocal(x_j, h'))_{h' \in [\nh]} \mid h]$
+%corresponding to $h = h'$.
+More simply, if
+the local features are simply the tensor product of some function $g(x)$ and an
+indicator on $h$ ($\philocal(x,h) = \flatten(g(x) \otimes e_h)$),
+then $\phix(x) = g(x)$ suffices.
 
@@
 
-For the mixture model, the hidden moments $\E[\phi(x,h)]$ we recover are actually
-are the sufficient statistics of the log-linear model.
-To elaborate, let us suppose for the moment that we have infinite examples of $(x,h)$.
-Let us consider the undirected mixture model.
-Then we would seek to maximize the joint log-likelihood:
+In \cite{anandkumar12moments}, our latent moments $(\pi, B)$
+would be the parameters, but for log-linear models,
+we still need to do a bit more work.
+First, note that we can express the sufficient statistics $\E[\phi(x,h)]$
+of the model in terms of the quantities recovered in Step 2:
+\begin{align}
+  \mu \eqdef \E[\phi(x,h)] = \E[\phiprior(h)] + \sum_{j=1}^3 \E[\philocal(x_j,h)] = \sum_{h \in [\nh]} \pi_h (\phiprior(h) + 3 C_h).
+\end{align}
+By properties of the exponential family, $\mu$ suffices to recover $\theta$.
+To elaborate, if we had infinite \emph{labeled} examples $(x,h)$,
+then we would seek to maximize the joint log-likelihood:
 \begin{align}
@@
 \end{align}
-Note that the optimization problem here only depends on the data via $\mu$, which is usually estimated from examples of $(x,h)$.
-Since $h$ is latent in our latent-variable setting, we must recover $\mu$ from latent observations.
-By the moment-generating properties of the log-partition function,
-the optimality condition is that $\mu = \E_\theta[\phi(x, h)]$.
+Note that the optimization problem here only depends on the data via $\mu$.
+Instead of estimating $\mu$ from labeled examples,
+we estimate $\mu$ from tensor factorization applied to unlabeled examples.
 
-We thus obtain a consistent estimator for $\theta$ assuming the exponential family is identifiable:
-\begin{assumption}
-The log-linear model is identifiable from fully-observed data.
-Equivalently, the log-partition function $A(\theta)$ is strongly convex.
-\end{assumption}
-Identifiability means that the moments $\mu$ uniquely determine $\theta^*$.
-In summary, factorization yields the mean parameters of the exponential family,
-and we can use convex optimization to retrieve the (canonical) parameters $\theta$.
+We thus obtain a consistent estimator for $\theta$ assuming the exponential
+family is identifiable from fully-observed data (that is
+the log-artition function $A(\theta)$ is strongly convex).
+%\begin{assumption}
+%The log-linear model is identifiable from fully-observed data.
+%Equivalently, the log-partition function $A(\theta)$ is strongly convex.
+%\end{assumption}
+%Identifiability means that the moments $\mu$ uniquely determine $\theta^*$.
+%By the moment-generating properties of the log-partition function,
+%the optimality condition is that $\mu = \nabla A(\theta) = \E_\theta[\phi(x, h)]$.
+In other words,
+tensor factorization (Step 2) yields the mean parameters of the exponential family
+($\mu = \nabla A(\theta) = \E_\theta[\phi(x, h)]$),
+which can be mapped to (canonical) parameters $\theta$ via convex optimization.
 
@@
 
-\paragraph{Perturbation analysis.}
+\paragraph{Perturbation analysis (sketch)}
 We have so far worked with exact population values $M_2,M_3,\pi,B,\mu,\theta^*$.
@@
 (see \citet{anandkumar12moments,anandkumar13tensor}).
-The hidden moments $\hat\mu$ are simply a quadratic combination of $(\hat\pi,\hat B)$,
+The latent moments $\hat\mu$ are simply a quadratic combination of $(\hat\pi,\hat B)$,
 and $\hat\theta = \nabla^{-1} A(\hat\mu)$, which is Lipschitz.

### 2013-05-31 doc/nips2013/setup.tex
made a pass up through section 4 (measurements)

@@
 and its own local log-partition function $A(\theta; z_{\pi(j)})$, which depends on values of its parent.
-To make the directed and undirected models more comparable,
+To consolidate notation,
 define $\phi(z) = \sum_{j \in V} \phi(z_j, z_{\pi(j)})$
@@
 where each $x^{(i)} \sim p_{\theta^*}$ for some true parameters $\theta^*$.
-Our goal is to produce a parameter estimate $\hat\theta$ which is hopefully
-close to $\theta^*$.
+Our goal is to produce a parameter estimate $\hat\theta$ that approximates $\theta^*$.
 

### 2013-05-31 doc/nips2013/figures.js

@@ -14,8 +14,16 @@ _).margin(100, 20).center();
 prez.addSlide(simpleModels.id('simpleModels'));
 
 generalModels = table(
-  [hmm({directed: false, len: 4}), tallMixtureModel(false), gridModel({directed: false, numRows: 3, numCols: 2})],
-  ['(a) Hidden Markov model', '(b) Tall mixture model', '(c) Grid model'],
+  [
+    hmm({directed: false, len: 4}),
+    gridModel({directed: false, numRows: 3, numCols: 2}),
+    tallMixtureModel(false),
+  _],
+  [
+    '(a) Hidden Markov model',
+    '(b) Grid model',
+    '(c) Tall mixture model',
+  _],
 _).margin(100, 20).center();
 prez.addSlide(generalModels.id('generalModels'));
 

### 2013-05-31 doc/nips2013/utils.js

@@ -34,14 +34,14 @@ G.indent = function(x, n) { return frame(x).xpadding(n != null ? n : 20).xpivot(
 ////////////////////////////////////////////////////////////
 
 G.schema = function() {
-  boxed = function(a) { return frame(a).padding(10).bg.strokeWidth(2).strokeColor('gray').end; }
+  boxed = function(a) { return frame(a).padding(5).bg.strokeWidth(2).strokeColor('gray').end; }
   style = function(x) { return x.strokeWidth(10).color('brown'); }
   return overlay(
     table(
       [boxed('Data $x^{(1)}, \\dots, x^{(n)}$'), nil(), boxed('Parameters $\\theta$')],
       [a1 = style(downArrow(s=70)), nil(), a3 = style(upArrow(s))],
-      [boxed('Observed moments $\\E[\\phix(x)]$'), a2 = style(rightArrow(200)), boxed('Latent moments $\\E[\\phi(x,h) \\mid h]$')],
-    _).margin(50, 30).center(),
+      [boxed('Observed moments $\\E[\\phix(x)]$'), a2 = style(rightArrow(200)), boxed('Latent moments $\\E[t(x) \\mid h]$')],
+    _).margin(50, 20).center(),
     moveRightOf('(1) aggregation'.bold(), a1),
     moveBottomOf('(2) factorization'.bold(), a2),
     moveRightOf('(3) optimization'.bold(), a3),

### 2013-05-31 doc/nips2013/appendix.tex

\appendix

\section{Directed models}
\label{sec:directedModels}

In the directed case, we assume a directed acyclic graph over $V$,
and define $\pi(j) \subset V$ to be the parents of each node $j \in V$.
Next, define a model as a product of locally-normalized models:
\begin{align}
  p_\theta(z) = \prod_{j \in V} p_\theta(z_j \mid z_{\pi(j)}), \quad\quad
  p_\theta(z_j \mid z_{\pi(j)}) = \exp \{ \phi_j(z_j, z_{\pi(j)})^\top\theta - A_j(\theta; z_{\pi(j)}) \}.
\end{align}
Here, each local model has its own local features $\phi_j(z_j, z_{\pi(j)})$
and its own local log-partition function $A(\theta; z_{\pi(j)})$, which depends on values of its parent.
To consolidate notation,
define $\phi(z) = \sum_{j \in V} \phi(z_j, z_{\pi(j)})$
and $A(\theta; z) = \sum_{j \in V} A_j(\theta; z_{\pi(j)})$,
so that:
\begin{align}
  p_\theta(z) \eqdef \exp\{ \phi(z)^\top\theta - A(\theta; z) \}. \aside{directed}
\end{align}

In this section, we describe parameter estimation in other models using our
framework.
a two-layer sigmoid belief network \cite{saul96sigmoid}.

\subsection{Directed mixture model}

In \refsecs{threeViewMixtureModel}{generalModels}, we focused for simplicity on
undirected models, but we parameter estimation in directed models is also
possible.
In the directed mixture model, we have
\begin{align}
p_\theta(x,h) = p_\theta(h) \prod_{j=1}^3 p_\theta(x_j \mid h),
\end{align}
where
$p_\theta(h) \eqdef \exp\{\phiprior(h)^\top\theta - \Aprior(\theta)\}$
and
$p_\theta(x_j \mid h) \eqdef \exp\{\philocal(x_j, h)^\top\theta - \Alocal(\theta; h)\}$.

Consequently, the joint log-likelihood is:
\begin{align}
  L(\theta) = {\underbrace{\E[\phi(x,h)]}_{\mu}}^\top\theta - \E[A(\theta; z)], \quad
  \E[A(\theta, z)] = \Aprior(\theta) + 3 \sum_{h \in [\nh]} \pi_h \Alocal(\theta;h).
\end{align}

Note that in directed models, the log-partition functions depends on the parent variables (in this case, $h$).
Having recovered the marginal distribution over $\pi$, we can compute the desired expectation.
Again, this problem is convex.

\subsection{Factorial hidden Markov models}

\Fig{figures/factorialHMM}{0.7}{factorialHMM}{
Parameter in a factorial HMM.
}

\cite{ghahramani97fhmm}

### 2013-05-31 doc/nips2013/factorial.tex

@@
-\section{Unshuffling factorization for factorial mixture models} \label{sec:factorialModels}
+\section{Unshuffling factorization for factorial mixture models}
+\label{sec:factorialModels}
 
 So far, we have relied on the existence of a bottleneck hidden variable $h_j$ on which
-we have three conditionally independent views.  Such bottlenecks do not always
-exist; for example, in \reffig{factorialModels}(a), neither $h_1$ or $h_2$ is a bottleneck.
-Our solution is to create a composite latent variable $c$ which is in
+we have three conditionally independent views $x_{j_1},x_{j_2},x_{j_3}$.  Such bottlenecks do not always
+exist; for example, in \reffig{factorialModels}(a), neither $h_1$ nor $h_2$ is
+a bottleneck.
+In this section, we present a technique to handle such models.
+We start by creating a composite latent variable $c$ which is in
 one-to-one correspondence with the existing latent variables $(h_1,h_2)$.
 We can invoke $\TensorFactorize$ to obtain composite marginals and
-conditional means.  However, the identity of $h_1$ and $h_2$ becomes lost in the composite,
-and must be recovered.  To do this, we develop a new algorithm $\UnshuffleFactorize$
-to produce constraints on the original parameters.
+conditional means, but the identities of $h_1$ and $h_2$ becomes lost in the
+composite.  We develop a new algorithm $\UnshuffleFactorize$
+to restore the identities and place constraints on the original
+parameters.
 
@@
 and then perform an \emph{unshuffling factorization} (b):
-Given known composite conditional means $B_c = \BP[x_j \mid c] \in \R^\ell$,
-which are derived from combinations of unknown parameters $f_a = \theta_1(\cdot, a), g_b = \theta_2(\cdot, b)$,
-we group differences into bins (dashed boxes), and maximal bins into sources (red rounded boxes),
-producing a linear program with variables $f_a,g_b,Z_{ab}$ and constraints for each $B_c$.
+Given known composite conditional means
+$L_c = \log \BP[x_j \mid c] \in \R^\ell$,
+which are derived from combinations of unknown parameters
+$f_a = \theta_1(\cdot, a), g_b = \theta_2(\cdot, b)$ and log-normalization scalars $Z_{ab}$,
+we group difference vectors into bins (dashed boxes), and maximal bins into sources (red rounded boxes),
+producing a linear program with variables $\{f_a\},\{g_b\},\{Z_{ab}\}$ and
+a constraint for each $B_c$.
 }
 
-%restricted Boltzmann machines \cite{salakhutdinov09softmax}
-
-We focus on the factorial mixture model.
+We focus on the restricted Boltzmann machine
+\cite{smolensky86rbm,salakhutdinov09softmax},
+which we refer to as an undirected factorial mixture model.
 Let $h = (h_1, \dots, h_s) \in [\nh]^s$ be the latent variables (sources)
 and $x = (x_1, x_2, x_3) \in [\nphix]^3$ be the observed variables.
+The joint model probability is as follows:
 \begin{align}
-p_\theta(x, h) = \prod_{i=1}^s p_\theta(h_i) \prod_{j=1}^3 p_\theta(x_j \mid h), \quad
-p_\theta(x_j \mid h) = \exp\left\{ \sum_{i=1}^s \theta_i(x_j, h_i) - A(\theta; h) \right\},
+%p_\theta(x, h) = \prod_{i=1}^s p_\theta(h_i) \prod_{j=1}^3 p_\theta(x_j \mid h), \quad
+%p_\theta(x_j \mid h) = \exp\left\{ \sum_{i=1}^s \theta_i(x_j, h_i) - A(\theta; h) \right\},
+p_\theta(x, h) = \exp \left\{ \sum_{j=1}^3 \sum_{i=1}^s \theta_i(x_j, h_i) - A(\theta) \right\},
 \end{align}
-where each $\theta_i \in \R^{\nphix \times \nh}$ are the parameters for the $i$-th source.
-%Let $\theta = (\theta_1, \dots, \theta_b) \in \R^{\nphix \times b \nh}$.
+where each $\theta_i \in \R^{\nphix \times \nh}$ is the parameter matrix for the $i$-th source.
 
@@
 with the actual mixture components, but in the factorial case, we must solve
-a \emph{credit assignment problem} to match combinations of $s$ sources
+a \emph{credit assignment problem} to match combinations of the $s$ sources
 with columns of $B$.
@@
 be the collection of parameter vectors that we seek to recover.
-Let $C \in \R^{\ell \times k^s}$ be the elementwise logarithm of $B$; that is
-$C(x_j, h) = \log B(x_j, h)$.
-The columns of $C$ are exactly the unordered collection of vectors
+Let $L \in \R^{\ell \times k^s}$ be the elementwise logarithm of $B$; that is
+$L(x_j, h) = \log B(x_j, h)$.
+The key observation is that the columns of $L$ are exactly the unordered collection of vectors
 $\{ \sum_{i=1}^s v_{ia_i} - Z_{\ba} : \ba = (a_1, \dots, a_s) \in [\nh]^s \}$,
-where $Z_{\ba} = \log \sum_{e=1}^\ell (\sum_{i=1}^s \exp(v_{ia_i}(e)))$ is the normalization constant (a scalar).
+where $Z_{\ba} = \log \sum_{e=1}^\ell (\sum_{i=1}^s \exp(v_{ia_i}(e)))$ is the
+log-normalization constant (a scalar) that sums over possible values $e$ of $x_j$.
 
 Note that there is a combinatorial dimension to this problem since the association between
-columns of $C$ and combinations of $\{ v_{ia} \}$ is unknown.
-As a running example (\reffig{factorialModels}(b)),
-consider $s=2$ sources, each taking on $k=3$ values (let $f_a = v_{1a}$ and $g_b = v_{2b}$).
-The following algorithm creates a linear program which puts constraints on variables $\{v_{ia}\}$ and $Z_{\ba}$:
+columns of $L$ and combinations of $\{ v_{ia} \}$ is unknown.
+For example, example (\reffig{factorialModels}(b)),
+consider $s=2$ sources, each taking on $k=3$ values
+(for convenience, let $f_a = v_{1a}$ and $g_b = v_{2b}$).
+The following algorithm, $\UnshuffleFactorize$,
+creates a linear program that places constraints on
+variables $\{v_{ia}\}$ and $Z_{\ba}$:
 \begin{enumerate}
 
-\item Compute the difference vector $B_i - B_j$ for every distinct pair of columns $c,d \in [\nh^s]$.
+\item Compute the difference vector $L_c - L_d$ for every distinct pair of
+columns $c,d \in [\nh^s]$.
 
-\item Put the pairs $(c,d)$ into bins, where the difference vectors $B_c-B_d$
+\item Put the pairs $(c,d)$ into bins such that the difference vectors $L_c-L_d$
 in each bin are \emph{comparable}, by which we mean that they
@@
 \item Keep only the maximal bins.
-These bins should contain $\nh^{s-1}$ pairs, each corresponding to a fixed setting of the $s-1$ sources
-and changing the value of one source.
+Such bins should contain $\nh^{s-1}$ pairs, each corresponding to a fixed
+setting of the $s-1$ sources
+and changing the value of only one source.
 Specifically, the difference vector is equal to $v_{ia} - v_{ia'}$ up to a
@@
   For each source $i = 1, \dots, s$,
-  chose any bin, declare it as $v_{i1} - v_{i2}$, and add it to the source.
-  Let $\gamma$ be the set of $c$ such that that bin contains some difference vector $B_c - B_d$,
-  and include any other bin with the same $\gamma$; these should correspond to $v_{i1} - v_{ia}$ for $a \ge 3$.
-  Then throw away any bin whose difference vector is $-(B_c - B_d)$
-  or is comparable to the difference of two existing difference vectors (e.g., $v_{i2} - v_{i1}$ or $v_{i2} - v_{i3}$).
-  The resulting source should have $\nh-1$ bins, corresponding to $\{ v_{i1} - v_{ia} : 2 \le a \le \nh \}$.
-  At this point, we have identified which vectors contribute to each $B_c$,
+  chose any bin $\omega$, declare it as $v_{i1} - v_{i2}$,
+  and add $\omega$ to the source.
+  Let $\gamma$ be the set of $c$ such that $\omega$
+  contains some difference vector $L_c - L_d$,
+  and include any other bin $\omega'$ with the same $\gamma$;
+  $\omega$' should correspond to $v_{i1} - v_{ia}$ for $a \ge 3$.
+
+  Then throw away any bin whose difference vector is $-(L_c - L_d)$
+  or is comparable to the difference of two existing difference vectors
+  (we want to discard bins corresponding to $v_{i2} - v_{i1}$ and $v_{i2} - v_{i3}$).
+  Finally, the resulting source should have $\nh-1$ bins, corresponding to $\{ v_{i1} - v_{ia} : 2 \le a \le \nh \}$.
+  At this point, we have identified which vectors contribute to each $L_c$,
   taking care of the combinatorial aspect of the problem,
@@
 
-  It turns out that this is LP is rank deficient for two reasons, the first intrinsic to our model,
+\end{enumerate}
+
+It turns out that this is LP is rank deficient for two reasons,
+the first intrinsic to our model,
 and the second due to loose linearization.
@@
 all the vectors of a source $i_1$ ($v_{i_1 a}$ for all $a \in [\nh]$)
-  subtract $u$ from all the vectors of another source $v_{i_2 a}$ for all $a \in [\nh]$.
-  Second, if we knew the normalization constants $\{Z_\ba\}$, then this is only rank deficiency.
-  However, treating $\{Z_\ba\}$ as independent variables creates additional rank deficiency.
+and subtract $u$ from all the vectors of another source $v_{i_2 a}$ for all $a \in [\nh]$.
+Second, if we treated log-normalization constants $\{Z_\ba\}$ as a deterministic
+function of $v_{ia}$, then the resulting program would have no other sources
+of rank deficiency, but would also be non-linear.
+Treating $\{Z_\ba\}$ as independent variables creates additional rank
+deficiency.
 
 Despite this, in the spirit of providing partial information,
-  $\UnshuffleFactorize$ still gives us linear constraints on the parameters.
-  In particular, we propose adding to the likelihood objective a strong regularization term
-  $R(\theta, Z) = \frac{\lambda}{2} (\bA_1 \theta - \bA_2 Z - B)^2$, where $\bA_1$ and $\bA_2$ are coefficients
+we can still use the linear constraints provided by $\UnshuffleFactorize$.
+In particular, define a regularization term
+$R(\theta, Z) = \frac{1}{2} (\bA_1 \theta - \bA_2 Z - B)^2$, where $\bA_1$ and $\bA_2$ are coefficients
 recovered by unshuffling.
-  In particular, we propose optimizing $\min_{\theta, Z} \sum_{i=1}^n -\log p_\theta(x^{(i)}) + R(\theta, Z)$.
-  Although the resulting objective is still non-convex, the regularization imposes quite strong constraints,
-  and the optimization would happen effectively over a much lower-dimensional subspace.
-\end{enumerate}
+In particular, we propose optimizing $\min_{\theta, Z} \sum_{i=1}^n -\log
+p_\theta(x^{(i)}) + \lambda R(\theta, Z)$,
+where $\lambda$ will be relatively large.
+Although the resulting objective is still non-convex,
+the regularization imposes quite strong constraints,
+resulting in an optimization problem
+that effectively operates over a much lower-dimensional subspace.

### 2013-05-31 doc/nips2013/mixture.tex

@@
 
-\Fig{figures/simpleModels}{0.3}{simpleModels}{Three-view mixture models: conditioned on a latent variable $h \in [\nh]$,
-we have three conditionally independent observed variables.  We consider
-estimation for both directed (left) and undirected (right) versions.
-}
+\Fig{figures/mixtureModel}{0.3}{mixtureModel}{An (undirected) three-view
+mixture model:
+conditioned on a latent variable $h \in [\nh]$,
+the observed variables $x_1,x_2,x_3$ are conditionally independent.}
 
 In this section, we consider the simplest model to illustrate the form of our
-method.  Consider the undirected mixture model model in \reffig{simpleModels}(b).
+method.  Consider the undirected mixture model model in \reffig{mixtureModel}.
 We have one discrete latent variable $h \in [\nh]$ and
@@
 \subsection{Step 2: Recover latent moments (factorization)}
+\label{sec:factorization}
 
@@
 
-% Directed mixture model
-\paragraph{Directed mixture model}
-
-The directed mixture model, we have $p_\theta(x,h) = p_\theta(h) \prod_{j=1}^3 p_\theta(x_j \mid h)$,
-where
-$p_\theta(h) \eqdef \exp\{\phiprior(h)^\top\theta - \Aprior(\theta)\}$
-and
-$p_\theta(x_j \mid h) \eqdef \exp\{\philocal(x_j, h)^\top\theta - \Alocal(\theta; h)\}$.
-Consequently, the joint log-likelihood is:
-\begin{align}
-  L(\theta) = {\underbrace{\E[\phi(x,h)]}_{\mu}}^\top\theta - \E[A(\theta; z)], \quad
-  \E[A(\theta, z)] = \Aprior(\theta) + 3 \sum_{h \in [\nh]} \pi_h \Alocal(\theta;h).
-\end{align}
-Note that in directed models, the log-partition functions depends on the parent variables (in this case, $h$).
-Having recovered the marginal distribution over $\pi$, we can compute the desired expectation.
-Again, this problem is convex.
-
 \paragraph{Perturbation analysis (sketch)}

### 2013-05-31 doc/nips2013/setup.tex

@@
 and $A(\theta) = \log \int \exp\{\phi(z)^\top\theta\} dz$ is the global log-partition function.
+To some extent, our methods only depend on the conditional independence
+structure of the models.
+We focus on undirected models for notational simplicity;
+directed models are discussed in \refapp{directedModels}.
 
-In the directed case, we assume a directed acyclic graph over $V$,
-and define $\pi(j) \subset V$ to be the parents of each node $j \in V$.
-Next, define a model as a product of locally-normalized models:
-\begin{align}
-  p_\theta(z) = \prod_{j \in V} p_\theta(z_j \mid z_{\pi(j)}), \quad\quad
-  p_\theta(z_j \mid z_{\pi(j)}) = \exp \{ \phi_j(z_j, z_{\pi(j)})^\top\theta - A_j(\theta; z_{\pi(j)}) \}.
-\end{align}
-Here, each local model has its own local features $\phi_j(z_j, z_{\pi(j)})$
-and its own local log-partition function $A(\theta; z_{\pi(j)})$, which depends on values of its parent.
-To consolidate notation,
-define $\phi(z) = \sum_{j \in V} \phi(z_j, z_{\pi(j)})$
-and $A(\theta; z) = \sum_{j \in V} A_j(\theta; z_{\pi(j)})$,
-so that:
-\begin{align}
-  p_\theta(z) \eqdef \exp\{ \phi(z)^\top\theta - A(\theta; z) \}. \aside{directed}
-\end{align}
-
-%To some extent, our methods only depend on the conditional independence
-%structure of the models, not on whether the model is directed or undirected.
-%Later, we will show how to estimate
-
-%\paragraph{Example.}
-%For example, \reffig{simpleModels} Markov random fields are an instance.
-
-\paragraph{Problem statement.}
+\paragraph{Problem statement}
 
@@
 which can get stuck in local optima.
-
-%Since both models have the same sufficient statistics lead to convex
-%likelihoods (assuming all variables are observed), in many cases our methods
-%will not be particular to which.  In some cases, however (e.g., the factorial models),
-%we will rely on marginalization properties of the directed models.
-
-% Principles
-%By insisting on developing a consistent estimator and considering identifiability,
-%The key is that a variable in the model must correspond to a bottleneck.

### 2013-05-31 doc/nips2013/factorial.tex
done with appendix

@@
 we have three conditionally independent views $x_{j_1},x_{j_2},x_{j_3}$.  Such bottlenecks do not always
-exist; for example, in \reffig{factorialModels}(a), neither $h_1$ nor $h_2$ is
+exist; for example, in \reffig{factorialMixtureModel}(a), neither $h_1$ nor $h_2$ is
 a bottleneck.
@@
 
-\Fig{figures/factorialModels}{0.3}{factorialModels}{The factorial mixture model (a)
+\Fig{figures/factorialMixtureModel}{0.3}{factorialMixtureModel}{The factorial mixture model (a)
 does not have three-view bottlenecks, so we first identify composite bottlenecks,
@@
 \cite{smolensky86rbm,salakhutdinov09softmax},
-which we refer to as an undirected factorial mixture model.
+which we refer to as an undirected factorial mixture model
+(the directed version is a two-layer sigmoid belief network \cite{saul96sigmoid}).
 Let $h = (h_1, \dots, h_s) \in [\nh]^s$ be the latent variables (sources)
@@
 columns of $L$ and combinations of $\{ v_{ia} \}$ is unknown.
-For example, example (\reffig{factorialModels}(b)),
+For example, example (\reffig{factorialMixtureModel}(b)),
 consider $s=2$ sources, each taking on $k=3$ values

### 2013-05-31 doc/nips2013/figures.js

@@ -19,8 +19,8 @@ _).margin(100, 20).center();
 prez.addSlide(simpleModels.id('simpleModels'));*/
 
 directedModels = table(
-  [mixtureModel(true), hmm({directed: true, len: 4}), factorialMixtureModel(true)],
-  ['(a) Mixture model', '(b) Hidden Markov model', '(c) Factorial mixture model'],
+  [mixtureModel(true), hmm({directed: true, len: 4})],
+  ['(a) Mixture model', '(b) Hidden Markov model'],
 _).margin(100, 20).center();
 prez.addSlide(directedModels.id('directedModels'));
 
@@ -38,22 +38,16 @@ generalModels = table(
 _).margin(100, 20).center();
 prez.addSlide(generalModels.id('generalModels'));
 
-/*factorialModels = table(
-  [factorialMixtureModel(), factorialHMM()],
-  ['(a) Factorial mixture model', '(b) Factorial HMM'],
-_).margin(50, 20).center();
-prez.addSlide(factorialModels.id('factorialModels'));*/
-
-factorialModels = table(
+model = table(
   [factorialMixtureModel(false), unshuffle()],
   ['(a) Factorial mixture model', '(b) Unshuffling factorization'],
 _).margin(50, 20).center();
-prez.addSlide(factorialModels.id('factorialModels'));
+prez.addSlide(model.id('factorialMixtureModel'));
 
-factorialHMM = table(
-  [factorialHMM()],
-  ['(a) Factorial HMM'],
+factorialModels = table(
+  [factorialMixtureModel(true), factorialHMM()],
+  ['(a) Factorial mixture model', '(b) Factorial HMM'],
 _).margin(50, 20).center();
-prez.addSlide(factorialHMM.id('factorialHMM'));
+prez.addSlide(factorialModels.id('factorialModels'));
 
 prez.writePdf({outPrefix: 'figures', combine: false});

### 2013-05-31 doc/nips2013/discussion.tex
spell check, mods to experiments

@@
-\section{Related work}
+\section{Discussion}
 
-% Observable operator models
+In recent years, there has been much interest
+on using the method of moments and spectral decomposition
+for learning latent-variable models.
+One line of work uses observable operator models to learn
 HMMs \cite{hsu09spectral},
@@
 weighted finite state automata \cite{balle12automata},
-finite state transducers \cite{balle11transducer}.
+finite state transducers \cite{balle11transducer},
+but do not produce parameter estimates.
 
 % Method of moments
+Our work builds on a second line of work which focuses
+on parameter estimation.  These methods have been used to learn
+hidden Markov models \cite{anandkumar12moments},
 Latent Dirichlet Allocation \cite{anandkumar12lda},
+linear Bayesian networks \cite{anandkumar12linear},
 mixture of spherical Gaussians \cite{hsu13spherical},
-latent parse tree models \cite{hsu12identifiability}
+latent parse tree models \cite{hsu12identifiability}, and
 mixture of linear regressions \cite{chaganty13regression}.
-linear Bayesian networks \cite{anandkumar12linear}.
+Not only can actual parameter estimates be valuable for scientific
+applications, they also provide a natural initialization point
+for further local optimization, which has been found to improve accuracy
+\cite{chaganty13regression}.
+
+Extending this latter perspective,
+our work departs from the standard goal of
+insisting on constructing consistent parameter estimators, but rather strives
+to use spectral decomposition to obtain constraints to improve
+parameter estimation.  Using these techniques,
+we found empirically that our approach provides
+improvements in various models such as the grid model and factorial HMM
+for which we do not have a consistent parameter estimate.
+We believe using partial information from factorized moments
+in conjunction with likelihood optimization
+broadens the toolbox of techniques for learning
+latent-variable models.
+Finally, our technique still relies fundamentally on
+having three-view bottlenecks \cite{anandkumar12moments}.
+How to extract partial information without this assumption
+remains an interesting open problem.
 
 % Structure recovery
-structure recovery
-\cite{anandkumar11tree}
+%structure recovery
+%\cite{anandkumar11tree}
 
 % Alternate estimators
-There are multiple consistent estimators.
-Composite likelihood \cite{besag75pseudo,lindsay88composite,liang08asymptotics}
-We are assuming inference is tractable.
-Our main problem is non-convexity.
-
-\section{Conclusion}
-
-That's it.
+%There are multiple consistent estimators.
+%Composite likelihood \cite{besag75pseudo,lindsay88composite,liang08asymptotics}

### 2013-05-31 doc/nips2013/measurements.tex
spell check, mods to experiments

@@
 \reffig{generalModels} shows examples of log-linear models.
-For notational simplicitly, assume each latent variable $h_j \in [\nh]$ is an integer
+For notational simplicity, assume each latent variable $h_j \in [\nh]$ is an integer
 and each observed variable $x_j \in \R^\ell$ is an indicator vector encoding
@@
 and therefore difficult to use.
+Note that if each latent variable had only one observation,
+there would be no bottlenecks, and our method would not be applicable.
 %to extract something useful like $\BP[h_{ij}, h_{i(j+1)}]$
@@
 How can we still use this partial information $(\pi,B)$?
-To this end, we adopt the measurements framework of \citet{liang09measurements},
+To this end, we adopt the measurements framework of \citet{liang09measurements}
+(also, see \cite{graca08em}),
 which allows us to incorporate moment constraints into a likelihood-based objective in a coherent way.

### 2013-05-31 doc/nips2013/mixture.tex
spell check, mods to experiments

@@
 and define the matrix $B \eqdef (\E[\phix(x_j) \mid h])_{h \in [\nh]} \in \R^{\ell \times k}$,
-where each column is the expected transformated data conditioned on $h$.
+where each column is the expected transformed data conditioned on $h$.
 
@@
 family is identifiable from fully-observed data (that is
-the log-artition function $A(\theta)$ is strongly convex).
+the log-partition function $A(\theta)$ is strongly convex).
 %\begin{assumption}

### 2013-05-31 doc/nips2013/appendix.tex

@@
 
-In the directed case, we assume a directed acyclic graph over $V$,
+We have so far focused on undirected models for simplicity,
+but our techniques also apply to directed models in which each
+local conditional distribution is a log-linear model.
+
+The setting is as follows:~Assume we have a directed acyclic graph over
+nodes $V$,
 and define $\pi(j) \subset V$ to be the parents of each node $j \in V$.
-Next, define a model as a product of locally-normalized models:
+Define the joint distribution as a product of locally-normalized models:
 \begin{align}
@@
 and its own local log-partition function $A(\theta; z_{\pi(j)})$, which depends on values of its parent.
-To consolidate notation,
-define $\phi(z) = \sum_{j \in V} \phi(z_j, z_{\pi(j)})$
+To consolidate notation with \refeqn{undirectedSetup},
+define the sum of all the feature vectors
+$\phi(z) = \sum_{j \in V} \phi(z_j, z_{\pi(j)})$
 and $A(\theta; z) = \sum_{j \in V} A_j(\theta; z_{\pi(j)})$,
@@
 \begin{align}
-  p_\theta(z) \eqdef \exp\{ \phi(z)^\top\theta - A(\theta; z) \}. \aside{directed}
+  p_\theta(z) \eqdef \exp\{ \phi(z)^\top\theta - A(\theta; z) \}.
 \end{align}
-
-In this section, we describe parameter estimation in other models using our
-framework.
-a two-layer sigmoid belief network \cite{saul96sigmoid}.
 
@@
 
-In \refsecs{threeViewMixtureModel}{generalModels}, we focused for simplicity on
-undirected models, but we parameter estimation in directed models is also
-possible.
-In the directed mixture model, we have
+In this section, we demonstrate parameter estimation for the directed mixture
+model (\reffig{directedModels}(a)).  This is the usual mixture model, defined
+as follows:
 \begin{align}
@@
 
-\subsection{Factorial hidden Markov models}
-
-\Fig{figures/factorialHMM}{0.7}{factorialHMM}{
-Parameter in a factorial HMM.
+\Fig{figures/directedModels}{0.3}{directedModels}{
+Various directed versions of the undirected models.
 }
 
-\cite{ghahramani97fhmm}
+\Fig{figures/factorialModels}{0.3}{factorialModels}{
+Directed factorial models.
+}
+
+\subsection{Other models}
+
+\reffig{directedModels} shows examples of the directed models
+for which we have already derived undirected estimates.
+In general, the aggregation
+and factorization steps (Steps 1 and 2) in \refsec{threeViewMixtureModel}
+remain the same as the undirected cases since these steps
+only rely on the conditional independence structure of the models,
+not particular parametrizations.
+
+\subsection{Factorial models}
+
+For factorial models, the directed and undirected version do differ in their
+conditional independence statements.
+For example, consider the 
+the directed factorial mixture model,
+also known as a sigmoid belief network \citep{saul96sigmoid},
+in \reffig{factorialModels}(a).
+Here ,$h_1$ and $h_2$ are marginally independent in the directed
+case but conditionally indepdent in the undirected case.
+However, since our algorithm treats $(h_1,h_2)$ as a composite variable $c$
+anyway, it is not sensitive to these differences.
+The factorial HMM \cite{ghahramani97fhmm} is similar and can be handled
+similar to our treatment of HMMs plus unshuffling factorization.

### 2013-05-31 doc/nips2013/factorial.tex

@@
   one for each of $s\nh$ vectors $v_{ia}$ and $\nh^s$ for the normalization constants $Z_\ba$.
-  Note that $Z_\ba$ is actually a determinstic function of the the vectors, but since it is a non-linear one,
+  Note that $Z_\ba$ is actually a deterministic function of the the vectors, but since it is a non-linear one,
   it is difficult to work with; therefore, we create a separate variable to keep the problem linear.

### 2013-05-31 doc/nips2013/introduction.tex
typos

@@
 First, in certain cases, these latent moments are exactly the sufficient statistics of
-the log-linear model, in which we case use convex optimization directly to obtain the parameters,
+the log-linear model, in which case we use convex optimization directly to obtain the parameters,
 resulting in a consistent parameter estimation procedure (\refsec{threeViewMixtureModel}).

### 2013-05-31 doc/nips2013/setup.tex
done with appendix

@@
 and the rest are latent ($h = \{ z_j \}_{j \in H}$, where $H = V \backslash O$).
-Let $C \subset 2^V$ be a collection of subset of variables,
+Let $C \subset 2^V$ be a collection of subsets of variables,
 and let $\phi_c(z_c) \in \R^d$ be the feature map for each $c \in C$.
@@
 \begin{align}
-  p_\theta(z) = \exp\{ \phi(z)^\top\theta - A(\theta) \}, \aside{undirected}
+  \label{eqn:undirectedSetup}
+  p_\theta(z) = \exp\{ \phi(z)^\top\theta - A(\theta) \},
 \end{align}
@@
 which is statistically efficient but computationally intractable.
-In practice, one uses EM or gradient-based optimization procedure on the marginal likelihood,
-which can get stuck in local optima.
+In practice, one uses gradient-based optimization procedures (e.g., EM or L-BFGS)
+on the marginal likelihood, which can get stuck in local optima.

### 2013-05-31 doc/nips2013/abstract.tex
small

@@
 In this paper,
-%we show how to leverage recent advances in the method of moments
-%to alleviate these problems.
 we leverage recent advances in the method of moments to estimate latent

### 2013-05-31 doc/nips2013/appendix.tex
a

@@
 similar to our treatment of HMMs plus unshuffling factorization.
+
+\subsection{Experiments}
+
+Table~\ref{tab:errorsa} shows the full set of results for the mixture
+model and HMM.
+
+\begin{table}
+\begin{center}
+    \begin{tabular}{l | l l | l l l }
+        Algorithm & $k$ & $\ell$ & Accuracy & $\Delta \hat \E[\phi(x,h)]$ & Log Likelihood \\ \hline
+        & \multicolumn{2}{|c|}{Mixture model} & & & \\ \hline
+        \multirow{3}{*}{EM} 
+        & 3 & 2 & 0.65 (+/- 0.01) & 1.23 & -3.72 \\
+        & 3 & 3 & 0.51 (+/- 0.06) & 1.81 & -4.21\\
+        & 5 & 2 & 0.72 (+/- 0.07) & 1.13 & -5.21\\ \hline
+        \multirow{3}{*}{Spectral} 
+        & 3 & 2 & 0.80 (+/- 0.09) & 0.23 & -3.54 \\
+        & 3 & 3 & 0.68 (+/- 0.10) & 1.01 & -3.76 \\
+        & 5 & 2 & 0.80 (+/- 0.05) & 0.18 & -5.04 \\ \hline
+        \multirow{3}{*}{Oracle} 
+        & 3 & 2 & 0.80 (+/- 0.08) & 0.01& -3.54 \\
+        & 3 & 3 & 0.70 (+/- 0.11) & 0.02& -3.87 \\
+        & 5 & 2 & 0.80 (+/- 0.05) & 0.01& -5.03 \\ \hline
+        & \multicolumn{2}{|c|}{HMM} & & & \\ \hline
+        \multirow{3}{*}{EM} 
+& 3 & 2 & 0.64 (+/- 0.10) & 1.71 & -4.51 \\  
+& 3 & 3 & 0.42 (+/- 0.05) & 2.41 & -6.09 \\  
+& 5 & 2 & 0.62 (+/- 0.09) & 1.60 & -6.47 \\  \hline
+        \multirow{3}{*}{Spectral} 
+& 3 & 2 & 0.67 (+/- 0.07) & 2.59 & -3.71 \\ 
+& 3 & 3 & 0.45 (+/- 0.08) & 4.96 & -4.36 \\ 
+& 5 & 2 & 0.64 (+/- 0.10) & 3.43 & -5.31 \\ \hline
+        \multirow{3}{*}{Oracle} 
+& 3 & 2 & 0.71 (+/- 0.07) & 0.39 & -4.64 \\ 
+& 3 & 3 & 0.47 (+/- 0.07) & 1.53 & -5.85 \\ 
+& 5 & 2 & 0.73 (+/- 0.07) & 0.39 & -6.16 \\ \hline
+    \end{tabular}
+    \caption{Micro-averaged results for various models, methods.}
+    \label{tab:errorsa}
+\end{center}
+\end{table}

### 2013-05-31 doc/nips2013/factorial.tex
a

@@
   For each source $i = 1, \dots, s$,
-  chose any bin $\omega$, declare it as $v_{i1} - v_{i2}$,
+  choose any bin $\omega$, declare it as $v_{i1} - v_{i2}$,
   and add $\omega$ to the source.

### 2013-06-03 doc/icml-presentation/presentation.tex
Skeletal ICML presentation

\documentclass[xcolor={svgnames}]{beamer}

\setbeameroption{hide notes}

%\usetheme{NLP}
\usepackage{beamerthemesplit}
\usetheme{default}
\useoutertheme{infolines}

\usepackage{graphicx}
\usepackage{lmodern}

\usepackage{soul}

\usepackage{amsmath,amsthm,amssymb}

\usepackage{listings}

\usepackage{algorithm,algorithmic}

\usepackage{tikz}
\usetikzlibrary{positioning,shapes,shadows,arrows}

\input{macros}
\input{slide-macros}
\input{spectral-macros}

% these will be used later in the title page
\title[Spectral Experts]{Spectral Experts for Estimating Mixtures of Linear Regressions}
\author[Chaganty, Liang]{%
    Arun Tejasvi Chaganty\\
    Percy Liang
}
\institute{Stanford University}

\begin{document}

% "Beamer, do the following at the start of every section"
\AtBeginSection[] 
{%
\begin{frame}<beamer> 
\frametitle{Outline} % make a frame titled "Outline"
\tableofcontents[currentsection]  % show TOC and highlight current section
\end{frame}
}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}
  \frametitle{Discriminative Latent Variable Models}

  \begin{itemize}
    \item {\bf Discriminative Latent Variable Models are useful, but hard to learn.}
    \item Discriminative LVMs combine the predictive accuracy with compact expressiveness.
    \item They have been used in several applications, e.g.\ object recognition, syntactic parsing, machine translation, etc.
    \item Learning parameters for these models is hard because they have non-convex likelihood functions.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Goal: Consistent Parameter Estimation.}

  \begin{itemize}
    \item {\bf We would like to develop efficient consistent estimators for discriminative LVMS.}
    \item Past approaches rely on local optimization, which are susceptible to local optima.
    \item And they don't just go away with more data.
    \item Recently, there has been work on consistent parameter
      estimation for generative LVMS\@. We'd like to extend their work to
      the discriminative case.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Mixture of Linear Regressions}

  \begin{itemize}
    \item {\bf First step, consider a simple discriminative model, mixture of linear regressions.}
    \item The mixture of linear regressions model defines a conditional distribution over a response $y \in \Re$ given covariates $x \in \Re^d$.
     \item The generative procedure is as follows,
       \begin{enumerate}
        \item Draw a mixture component $h \in [k] \sim Mult(\pi)$, where $\pi = [\pi_1 | \dots | \pi_k]$ defines the mixture proportions.
        \item Draw the noise $\epsilon \sim \mathcal{E}$, where $\mathcal{E}$ is the noise distribution.
        \item Set $y = \beta_h^T x + \epsilon$, where $\{\beta_h\}_{h=1}^{k}$ are the conditional means of the regression coefficients.
       \end{enumerate}
    \item The parameters that we would like to learn from this model are $\pi$ and $B = [ \beta_1 | \cdots | \beta_k ]$.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Method of Moments.}

  \begin{itemize}
    \item {\bf We approach the problem by using the method of moments on three conditional views. }
    \item Consider a moment map $\mathcal{M}$ that maps the parameters $\theta$ to the moments $m$. For a Gaussian, we have that $\mathcal{M} = (\mu, \sigma^2)$.
    \item In general, we will compute the inverse of the moment map to learn the parameters from the sample estimates.
    \item By the central limit theorem, our sample estimates of the moments converge at a $1/\sqrt{n}$ rate, so we expect that our parameters will also converge at this rate.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Method of Moments for Generative LVMs.}

  \begin{itemize}
    \item {\bf We can recover the means of a generative LVM by exploiting their symmetric structure. }
    \item 
    \begin{align*}
    M_2 &= \sum_{h=1}^k \pi_h \beta_h\tp{2} & M_3 &= \sum_{h=1}^k \pi_h \beta_h\tp{3}.
  \end{align*}
    \item In general, $M_2$ is insufficient to identify the model, because it is invariant to rotations of $B$. 
    \item If the $\beta_h$ were orthogonal, then the eigenvectors and eigenvalues of $M_3$ would give us the answer. 
    \item In general, we need both $M_2$ and $M_3$. We can use the whitening transformation for $M_2$ to whiten $M_3$ such that it has a orthogonal decomposition.
\item The robust tensor power method by \cite{AnandkumarHsuGe2012} find stable eigenvectors.
  \item In the discriminative case, we do not observe $M_2$ directly from the data, but rather, as we'll see, linear projections of them.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Recovering the Moments.}
  \begin{itemize}
    \item {\bf The powers of the response variables $y^p$ are linear measurements of the moments of the parameters $M_p$!}
    \item As noted earlier, the first problem we run into is that we can't observe the moments of the parameters $B$ and $\pi$ directly!
\item However, observe that 
\begin{align}
  y &= \innerp{\beta_h}{x} + \epsilon \\
    &= \innerp{M_1}{x} + \underbrace{{\beta_h - M_1}{x} + \epsilon }_{\textrm{noise}},
\end{align}
  where $M_1 = \sum_{h=1}^k \pi_h \beta_h$, the mean regression coefficient. 
  \item We note that while the noise term is dependent on $x$, it has a zero-mean. Thus, we could potentially recover $M_1$ through regression.
\item Similarly, we can cast the remaining two terms as regression problems.

\item An additional fact that we can exploit is that both $M_2$ and $M_3$ are low rank, so we can use low-rank regression to recover estimates $\hat M_2$ and $\hat M_3$ efficiently from data.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Rates for Recovery}
  \begin{itemize}
    \item {\bf The algorithm is consistent, and requires $O(?)$ samples.}
\item The rate of convergence for the spectral experts algorithm to the
true parameters breaks into two parts; the rates for learning the
moments, which feeds into the rates for learning the parameters.
\item For low rank regression, we have the following bound on recovery by (Tomioka2011);
  $$ \| \hat M_p - M_p \|_F \le \frac{32 \lambda^{(p)}_n \sqrt{k}}{\kappa(\opX_p)}, $$
where $\kappa(\opX_p)$ is the (restricted) strong convexity constant, and $\lambda^{(p)} > \|\opX^*_p(\eta)\|$. 
\item Because we assume our noise is bounded, it is easy to show that the error concentrates. 
\item In the tensor recovery case, we will need to whiten $M_3$ before
applying the tensor decomposition and unwhiten it afterwards; this
modifies the error bounds slightly.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Experimental Results}
  \begin{itemize}
    \item {\bf Spectral Experts provides a good initialization to EM.}
      \item We simulated the performance of spectral experts and compared it to EM. 
\item In a non-trivial number of cases, EM did not converge to the right parameters and got stuck in local optima.
\item The parameters estimated by the spectral method were not very good even with $O(10^5)$ samples.
\item However, EM initialized with these parameters did extremely well. 
\item This finding that spectral methods should be a good initialization for EM is not surprising.
  \begin{itemize}
    \item The biggest sell for spectral methods is that they give a global guarantee on where the parameters are.
    \item The parameters might not be at the global optima, but will hopefully lie in the potential well around it.
    \item EM will then converge to the global optima.
  \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Conclusions}
  \begin{itemize}
    \item {\bf We can learn the parameters of conditional models where the observed moments on their own have such sparse information.}
    \item The key intuition is that we could construct a regression problem to learn the moments of the parameters from their projections onto the data.
    \item We found that while the parameters learned this way are usually not better than those learned via local methods like EM, etc., they are a good way to initialize these local methods.
  \end{itemize}
\end{frame}

\end{document}

% TikZ
%\begin{tikzpicture}
%  \node (img1) at (0,0) {\includegraphics[height=0.6\textheight,width=0.4\linewidth,keepaspectratio]{}};
%\end{tikzpicture}

% Notes 
% \note[item]{}

% 2-column
% \begin{columns}
%   \begin{column}{0.48\textwidth}
%     \begin{itemize}
%         \item
%     \end{itemize}
%   \end{column}
%   \hfill
%   \begin{column}{0.48\textwidth}
%     \begin{itemize}
%         \item
%     \end{itemize}
%   \end{column}
% \end{columns}

\end{document}

% TikZ
%\begin{tikzpicture}
%  \node (img1) at (0,0) {\includegraphics[height=0.6\textheight,width=0.4\linewidth,keepaspectratio]{}};
%\end{tikzpicture}

% Notes 
% \note[item]{}

% 2-column
% \begin{columns}
%   \begin{column}{0.48\textwidth}
%     \begin{itemize}
%         \item
%     \end{itemize}
%   \end{column}
%   \hfill
%   \begin{column}{0.48\textwidth}
%     \begin{itemize}
%         \item
%     \end{itemize}
%   \end{column}
% \end{columns}

### 2013-06-03 doc/nips2013/mixture.tex

@@
 \end{property}
-Here, $C$ provide a view on the local features $\philocal(x_j,h)$,
+Here, $C$ provides a view on the local features $\philocal(x_j,h)$,
 which are

### 2013-06-04 doc/icml-presentation/presentation.tex
Marked out the flow of the presentation

@@
 \begin{frame}
-  \frametitle{Discriminative Latent Variable Models}
+  \frametitle{Generative vs. Discriminative Latent Variable Models}
 
-  \begin{itemize}
-    \item {\bf Discriminative Latent Variable Models are useful, but hard to learn.}
-    \item Discriminative LVMs combine the predictive accuracy with compact expressiveness.
-    \item They have been used in several applications, e.g.\ object recognition, syntactic parsing, machine translation, etc.
-    \item Learning parameters for these models is hard because they have non-convex likelihood functions.
-  \end{itemize}
+  \begin{centering}
+  \begin{tikzpicture}
+    % x, y
+    \node at (0,0) {Generative Models};
+    \node[style=box] at (0,-1.5) {\obj{%
+      \vspace{4em}
+      \hfill CRFs
+      }};
+    \node[style=box] at (0,-4) {\obj{%
+      \vspace{4em}
+      \hfill PCFGs
+      }};
+
+    \node<2-> at (6,0) {Discriminative Models};
+    \node<2->[style=box] at (6,-1.5) {\obj{%
+      \vspace{4em}
+      \hfill Liang et.\ al (2006)
+      }};
+    \node<2->[style=box] at (6,-4) {\obj{%
+      \vspace{4em}
+      \hfill Wang \& Mori (2009)
+      }};
+
+    \node<2-> at (7,-5.5) {\obj{%
+      Easier to include features, \\
+      tend to be more accuracy.
+      }};
+  \end{tikzpicture}
+  \end{centering}
+
+    \note[item]{Discriminative Latent Variable Models are useful, but hard to learn.}
+    \note[item]{Split models into generative and discriminative models.}
+    \note[item]{Discriminative LVMs combine the predictive accuracy with compact expressiveness.}
+    \note[item]{They have been used in several applications, e.g.\ object recognition, syntactic parsing, machine translation, etc.}
+    \note[item]{Learning parameters for these models is hard because they have non-convex likelihood functions.}
 \end{frame}
@@
 \begin{frame}
-  \frametitle{Goal: Consistent Parameter Estimation.}
+  \frametitle{Parameter Estimation is Hard}
+
+  \begin{centering}
+  \begin{tikzpicture}
+    % x, y
+    \node[style=box] at (0,-1.5) {\objw{36em}{%
+      \vspace{10em}
+      {\em log-likelihood picture}
+      }};
+  \end{tikzpicture}
+  \end{centering}
 
   \begin{itemize}
-    \item {\bf We would like to develop efficient consistent estimators for discriminative LVMS.}
-    \item Past approaches rely on local optimization, which are susceptible to local optima.
-    \item And they don't just go away with more data.
-    \item Recently, there has been work on consistent parameter
+    \item<1-> Log-likelihood function is non-convex.
+    \item<2-> This is a problem that doesn't go away even with infinite data.
+    \item<3-> Can we build {\bf consistent estimators} that get better with more data?
+  \end{itemize}
+
+  \note[item]{We would like to develop efficient consistent estimators for discriminative LVMS.}
+  \note[item]{Past approaches rely on local optimization, which are susceptible to local optima.}
+  \note[item]{And they don't just go away with more data.}
+  \note[item]{Recently, there has been work on consistent parameter
       estimation for generative LVMS\@. We'd like to extend their work to
-      the discriminative case.
-  \end{itemize}
+      the discriminative case.}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Spectral Methods for Generative Models.}
+  \begin{centering}
+  \begin{tikzpicture}
+    % x, y
+    \node at (0,0) {Generative Models};
+    \node[style=box] at (0,-1.5) {\obj{%
+      \vspace{4em}
+      \hfill CRFs
+      }};
+    \node[style=box] at (0,-4) {\obj{%
+      \vspace{4em}
+      \hfill PCFGs
+      }};
+
+    %\node<2-> at (6,0) {Discriminative Models};
+    \node<2->[style=box] at (6,-1.5) {\obj{%
+      Solve the general multi-view case using the tensor factorization method.\\
+      \hfill Anandkumar et.\ al (2012) 
+      }};
+    \node<3->[style=box] at (6,-4) {\obj{%
+      Reduce the problem to multi-view learning via tree-unmixing.\\
+      \hfill Hsu et.\ al (2012) 
+      }};
+  \end{tikzpicture}
+  \end{centering}
+
+   \todo{Talk about observable operator stuff.}
 \end{frame}
@@
 
+  \begin{centering}
+  \begin{tikzpicture}
+    % x, y
+    \node<1-3> at (0,0){Generative Process};
+    \node<1-3>[style=box] at (0,-2.5) {\obj{%
+      Graphical model \vspace{3em}\\
+      Given parameters $\pi$ and $B = [ \beta_1 | \cdots | \beta_k ]$,
       \begin{itemize}
-    \item {\bf First step, consider a simple discriminative model, mixture of linear regressions.}
-    \item The mixture of linear regressions model defines a conditional distribution over a response $y \in \Re$ given covariates $x \in \Re^d$.
-     \item The generative procedure is as follows,
+        \item<1-> $h \in [k] \sim Mult(\pi)$.
+        \item<2-> $\epsilon \sim \mathcal{E}$.
+        \item<3-> $y = \beta_h^T x + \epsilon$.
+      \end{itemize}
+      }};
+    \node<4>[style=box] at (0,0) {\obj{%
+    {\bf Given:} $\mathcal{D} = (x_i, y_i)_{i=1}^{n}$,\\
+    can we {\bf estimate parameters} $\pi$ and $B = [ \beta_1 | \cdots | \beta_k ]$?
+      }};
+
+    %\node<2-> at (6,0) {Discriminative Models};
+    \node[style=box] at (6,-1.5) {\obj{%
+      \vspace{12em}
+      Diagram
+      }};
+  \end{tikzpicture}
+  \end{centering}
+
+  \note[item]{First step, consider a simple discriminative model, mixture of linear regressions.}
+  \note[item]{The mixture of linear regressions model defines a conditional distribution over a response $y \in \Re$ given covariates $x \in \Re^d$.}
+     \note[item]{The generative procedure is as follows,
        \begin{enumerate}
@@
         \item Set $y = \beta_h^T x + \epsilon$, where $\{\beta_h\}_{h=1}^{k}$ are the conditional means of the regression coefficients.
-       \end{enumerate}
-    \item The parameters that we would like to learn from this model are $\pi$ and $B = [ \beta_1 | \cdots | \beta_k ]$.
-  \end{itemize}
+      \end{enumerate}}
+   \note[item]{The parameters that we would like to learn from this model are $\pi$ and $B = [ \beta_1 | \cdots | \beta_k ]$.}
 \end{frame}
@@
 
+  \begin{centering}
+  \begin{tikzpicture}
+    % x, y
+    \node[style=box] at (0,0) {\objw{36em}{%
+      \vspace{10em}
+      {\em moment-map picture}
+      }};
+  \end{tikzpicture}
+  \end{centering}
   \begin{itemize}
-    \item {\bf We approach the problem by using the method of moments on three conditional views. }
-    \item Consider a moment map $\mathcal{M}$ that maps the parameters $\theta$ to the moments $m$. For a Gaussian, we have that $\mathcal{M} = (\mu, \sigma^2)$.
-    \item In general, we will compute the inverse of the moment map to learn the parameters from the sample estimates.
-    \item By the central limit theorem, our sample estimates of the moments converge at a $1/\sqrt{n}$ rate, so we expect that our parameters will also converge at this rate.
+    \item<1-> Moments are expected to converge at a $1/\sqrt{n}$ rate, by CLT.
+    \item<2-> A continuous mapping $\mathcal{M}^{-1}$ will also converge at a $1/\sqrt{n}$ rate.
   \end{itemize}
+
+    \note[item]{We approach the problem by using the method of moments on three conditional views. }
+    \note[item]{Consider a moment map $\mathcal{M}$ that maps the parameters $\theta$ to the moments $m$. For a Gaussian, we have that $\mathcal{M} = (\mu, \sigma^2)$.}
+    \note[item]{In general, we will compute the inverse of the moment map to learn the parameters from the sample estimates.}
+    \note[item]{By the central limit theorem, our sample estimates of the moments converge at a $1/\sqrt{n}$ rate, so we expect that our parameters will also converge at this rate.}
 \end{frame}
@@
 
-  \begin{itemize}
-    \item {\bf We can recover the means of a generative LVM by exploiting their symmetric structure. }
-    \item 
-    \begin{align*}
-    M_2 &= \sum_{h=1}^k \pi_h \beta_h\tp{2} & M_3 &= \sum_{h=1}^k \pi_h \beta_h\tp{3}.
-  \end{align*}
-    \item In general, $M_2$ is insufficient to identify the model, because it is invariant to rotations of $B$. 
-    \item If the $\beta_h$ were orthogonal, then the eigenvectors and eigenvalues of $M_3$ would give us the answer. 
-    \item In general, we need both $M_2$ and $M_3$. We can use the whitening transformation for $M_2$ to whiten $M_3$ such that it has a orthogonal decomposition.
-\item The robust tensor power method by \cite{AnandkumarHsuGe2012} find stable eigenvectors.
-  \item In the discriminative case, we do not observe $M_2$ directly from the data, but rather, as we'll see, linear projections of them.
-  \end{itemize}
+  \begin{centering}
+  \begin{tikzpicture}
+    % x, y
+    \node[style=box] at (0,0) {\objw{36em}{%
+      \vspace{10em}
+      {\em Work out the moments in the generative case, with indicative diagrams for matrices, tensors, etc.}
+      }};
+  \end{tikzpicture}
+  \end{centering}
+
+    \note[item]{\bf We can recover the means of a generative LVM by exploiting their symmetric structure.}
+    \note[item]{\begin{align*} M_2 &= \sum_{h=1}^k \pi_h \beta_h\tp{2} & M_3 &= \sum_{h=1}^k \pi_h \beta_h\tp{3}.  \end{align*}}
+      \note[item]{In general, $M_2$ is insufficient to identify the model, because it is invariant to rotations of $B$. }
+      \note[item]{If the $\beta_h$ were orthogonal, then the eigenvectors and eigenvalues of $M_3$ would give us the answer. }
+      \note[item]{In general, we need both $M_2$ and $M_3$. We can use the whitening transformation for $M_2$ to whiten $M_3$ such that it has a orthogonal decomposition.}
+      \note[item]{The robust tensor power method by \cite{AnandkumarHsuGe2012} find stable eigenvectors.}
 \end{frame}
@@
 \begin{frame}
-  \frametitle{Recovering the Moments.}
-  \begin{itemize}
-    \item {\bf The powers of the response variables $y^p$ are linear measurements of the moments of the parameters $M_p$!}
-    \item As noted earlier, the first problem we run into is that we can't observe the moments of the parameters $B$ and $\pi$ directly!
-\item However, observe that 
+  \frametitle{Method of Moments for the Mixture of Linear Regressions.}
+
+  \begin{centering}
+  \begin{tikzpicture}
+    % x, y
+    \node[style=box] at (0,0) {\objw{36em}{%
+      \vspace{10em}
+      {\em Try to work out the moments in our case.}
+      }};
+  \end{tikzpicture}
+  \end{centering}
+      \note[item]{In the discriminative case, we do not observe $M_2$ directly from the data, but rather, as we'll see, linear measurements of them.}
+      \note[item]{Try to work out the moments for this case and see how they {\tt \#fail} epically.}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Method of Moments for the Mixture of Linear Regressions.}
+
+  \begin{centering}
+  \begin{tikzpicture}
+    % x, y
+    \node[style=box] at (0,0) {\objw{18em}{%
+      \vspace{10em}
+      {\em Different tactic: show that $y^p$ are linear measurements of the moments of the parameters $M_p$! }
+      }};
+
+    \node[style=box] at (6,0) {\objw{8em}{%
+      {$y^p$ are linear measurements of the moments of the parameters $M_p$!}
+      }};
+  \end{tikzpicture}
+  \end{centering}
+
+    \note[item]{The powers of the response variables $y^p$ are linear measurements of the moments of the parameters $M_p$!}
+    \note[item]{As noted earlier, the first problem we run into is that we can't observe the moments of the parameters $B$ and $\pi$ directly!}
+    \note[item]{However, observe that 
     \begin{align}
@@
     \end{align}
-  where $M_1 = \sum_{h=1}^k \pi_h \beta_h$, the mean regression coefficient. 
-  \item We note that while the noise term is dependent on $x$, it has a zero-mean. Thus, we could potentially recover $M_1$ through regression.
-\item Similarly, we can cast the remaining two terms as regression problems.
-
-\item An additional fact that we can exploit is that both $M_2$ and $M_3$ are low rank, so we can use low-rank regression to recover estimates $\hat M_2$ and $\hat M_3$ efficiently from data.
-  \end{itemize}
+    where $M_1 = \sum_{h=1}^k \pi_h \beta_h$, the mean regression coefficient.}
+    \note[item]{We note that while the noise term is dependent on $x$, it has a zero-mean. Thus, we could potentially recover $M_1$ through regression.}
+    \note[item]{Similarly, we can cast the remaining two terms as regression problems.}
+    \note[item]{An additional fact that we can exploit is that both $M_2$ and $M_3$ are low rank, so we can use low-rank regression to recover estimates $\hat M_2$ and $\hat M_3$ efficiently from data.}
 \end{frame}
@@
 \begin{frame}
-  \frametitle{Rates for Recovery}
-  \begin{itemize}
-    \item {\bf The algorithm is consistent, and requires $O(?)$ samples.}
-\item The rate of convergence for the spectral experts algorithm to the
-true parameters breaks into two parts; the rates for learning the
-moments, which feeds into the rates for learning the parameters.
-\item For low rank regression, we have the following bound on recovery by (Tomioka2011);
-  $$ \| \hat M_p - M_p \|_F \le \frac{32 \lambda^{(p)}_n \sqrt{k}}{\kappa(\opX_p)}, $$
-where $\kappa(\opX_p)$ is the (restricted) strong convexity constant, and $\lambda^{(p)} > \|\opX^*_p(\eta)\|$. 
-\item Because we assume our noise is bounded, it is easy to show that the error concentrates. 
-\item In the tensor recovery case, we will need to whiten $M_3$ before
-applying the tensor decomposition and unwhiten it afterwards; this
-modifies the error bounds slightly.
-  \end{itemize}
+  \frametitle{Efficient Consistent Recovery}
+
+  \begin{centering}
+  \begin{tikzpicture}
+    % x, y
+    \node[style=box] at (0,0) {\objw{12em}{%
+      \vspace{10em}
+      Regression step
+      }};
+
+    \node[style=box] at (6,0) {\objw{12em}{%
+      The tensor factorization method.
+      }};
+  \end{tikzpicture}
+  \end{centering}
+
+  \note[item]{The algorithm is consistent, and requires $O(?)$ samples.}
+  \note[item]{The rate of convergence for the spectral experts algorithm to the true parameters breaks into two parts; the rates for learning the moments, which feeds into the rates for learning the parameters.}
+  \note[item]{For low rank regression, we have the following bound on recovery by (Tomioka2011); $$ \| \hat M_p - M_p \|_F \le \frac{32 \lambda^{(p)}_n \sqrt{k}}{\kappa(\opX_p)}, $$ where $\kappa(\opX_p)$ is the (restricted) strong convexity constant, and $\lambda^{(p)} > \|\opX^*_p(\eta)\|$.}
+  \note[item]{Because we assume our noise is bounded, it is easy to show that the error concentrates.}
+  \note[item]{In the tensor recovery case, we will need to whiten $M_3$ before applying the tensor decomposition and unwhiten it afterwards; this modifies the error bounds slightly.}
 \end{frame}
@@
 \begin{frame}
-  \frametitle{Experimental Results}
+  \frametitle{Experimental Insights}
   \begin{itemize}

### 2013-06-05 doc/icml-presentation/presentation.tex

@@
   \begin{tikzpicture}
-    % x, y
-    \node at (0,0) {Generative Models};
-    \node[style=box] at (0,-1.5) {\obj{%
-      \vspace{4em}
-      \hfill CRFs
+    % Tasks.
+    \node[style=box] (pos) at (0,0) {\objw{8cm}{%
+      \includegraphics[width=\textwidth]{figures/pos.png}\\
+      \hfill Part-of-Speech Tagging
       }};
-    \node[style=box] at (0,-4) {\obj{%
-      \vspace{4em}
-      \hfill PCFGs
+      \node[style=box,below=0.5cm of pos] (dep) {\objw{8cm}{%
+      \includegraphics[width=\textwidth]{figures/dep-parse.png}\\
+      \hfill Dependency Parsing
+      }};
+      \node[style=box, below=0.5cm of dep] (act) {\objw{8cm}{%
+      \includegraphics[width=\textwidth,height=3cm,keepaspectratio]{figures/activity-recognition.png}\\
+      \hfill Activity Recognition
       }};
 
-    \node<2-> at (6,0) {Discriminative Models};
-    \node<2->[style=box] at (6,-1.5) {\obj{%
-      \vspace{4em}
-      \hfill Liang et.\ al (2006)
+    \node<1->[style=box] (gen) at (6,0) {\obj{%
+      {\bf Generative Models}
+      \begin{itemize}
+        \item Hidden Markov Models
+        \item Conditional Random Fields
+        \item PCFGs
+        \item \dots
+      \end{itemize}
       }};
-    \node<2->[style=box] at (6,-4) {\obj{%
-      \vspace{4em}
-      \hfill Wang \& Mori (2009)
-      }};
-
-    \node<2-> at (7,-5.5) {\obj{%
-      Easier to include features, \\
-      tend to be more accuracy.
+    %\node<1->[below=0.5cm of gen-models] (disc) {Discriminative Models};
+    \node<1->[style=box,below=1.0cm of gen] (disc) {\obj{%
+    {\bf Discriminative Models}
+      \begin{itemize}
+        \item Max-Entropy Markov Model (Toutanova and Manning, 2000)
+        \item Discriminative Parsers (Liang et.\ al, 2006)
+        \item Semi-Latent Topic Models (Wang and Mori, 2009)
+        \item Easy to include features and tend to be more accurate.
+      \end{itemize}
       }};
@@
 
+    \note[item]{Would like to start by talking about the bigger picture of parameter estimation.}
     \note[item]{Discriminative Latent Variable Models are useful, but hard to learn.}
@@
     % x, y
-    \node[style=box] at (0,-1.5) {\objw{36em}{%
-      \vspace{10em}
-      {\em log-likelihood picture}
+    \node[style=box] at (0,-1.5) {\objw{12cm}{%
+    \includegraphics<1>[width=\textwidth,height=8cm,keepaspectratio]{figures/likelihood.png}
+    \includegraphics<2>[width=\textwidth,height=8cm,keepaspectratio]{figures/likelihood-em1.png}
+    \includegraphics<3-4>[width=\textwidth,height=8cm,keepaspectratio]{figures/likelihood-em2.png}
+    \includegraphics<5>[width=\textwidth,height=8cm,keepaspectratio]{figures/likelihood-mom.png}
+    \includegraphics<6>[width=\textwidth,height=8cm,keepaspectratio]{figures/likelihood-mom1.png}
+      \hfill{\em expected log-likelihood}
       }};
@@
     \item<1-> Log-likelihood function is non-convex.
-    \item<2-> This is a problem that doesn't go away even with infinite data.
-    \item<3-> Can we build {\bf consistent estimators} that get better with more data?
+    \item<4-> This is a problem that doesn't go away even with infinite data.
+    \item<5-> Can we build {\bf consistent estimators} that get better with more data?
   \end{itemize}
@@
     % x, y
-    \node at (0,0) {Generative Models};
-    \node[style=box] at (0,-1.5) {\obj{%
-      \vspace{4em}
-      \hfill CRFs
+    \node<1-2>[style=box] (pos) at (0,0) {\objw{6cm}{%
+      \includegraphics[width=\textwidth]{figures/pos.png}\\
+      \hfill Part-of-Speech Tagging
       }};
-    \node[style=box] at (0,-4) {\obj{%
-      \vspace{4em}
-      \hfill PCFGs
+    \node<3->[style=box] (dep) at (0,0) {\objw{6cm}{%
+      \includegraphics[width=\textwidth]{figures/dep-parse.png}\\
+      \hfill Dependency Parsing
       }};
@@
     %\node<2-> at (6,0) {Discriminative Models};
-    \node<2->[style=box] at (6,-1.5) {\obj{%
-      Solve the general multi-view case using the tensor factorization method.\\
-      \hfill Anandkumar et.\ al (2012) 
+    \node<1-2>[style=box] (ahk) at (6,0) {\objw{8cm}{%
+      \includegraphics[width=\textwidth,height=4cm,keepaspectratio]{figures/three-view.png}\\
+      \begin{itemize}
+        \item Solution for multi-view case using tensor factorization (Anandkumar et.\ al, 2012).
+        \item<2> HMMs (Anandkumar et.\ al, 2012) 
+        \item<2> LDA (Anandkumar et.\ al, 2012) 
+      \end{itemize}
+      \hfill 
       }};
-    \node<3->[style=box] at (6,-4) {\obj{%
-      Reduce the problem to multi-view learning via tree-unmixing.\\
-      \hfill Hsu et.\ al (2012) 
+      \node<3->[style=box] at (6,0) {\objw{8cm}{%
+      \includegraphics[width=\textwidth,height=4cm,keepaspectratio]{figures/unmixing.png}\\
+      \begin{itemize}
+        \item Reduce the problem to multi-view learning via tree-unmixing (Hsu et.\ al, 2012). 
+      \end{itemize}
+      }};
+      \node<4->[style=box] at (4,-3) {\objw{8cm}{%
+      {\bf How can we apply these methods for discriminative models?}
       }};
@@
   \end{centering}
-
-   \todo{Talk about observable operator stuff.}
+  \note[item]{\todo{Talk about observable operator stuff.}}
 \end{frame}
@@
     % x, y
-    \node<1-3> at (0,0){Generative Process};
-    \node<1-3>[style=box] at (0,-2.5) {\obj{%
-      Graphical model \vspace{3em}\\
-      Given parameters $\pi$ and $B = [ \beta_1 | \cdots | \beta_k ]$,
+    \node<1-10>[style=box] at (0,0) {\obj{%
+    {\bf Parameters:} $\pi = [\pi_1, \cdots, \pi_k]$ and $B = [ \beta_1 | \cdots | \beta_k ]$.
+    \includegraphics[width=\textwidth,height=4cm,keepaspectratio]{figures/disc-model.png}\\
+    For a given $x$,
       \begin{itemize}
-        \item<1-> $h \in [k] \sim Mult(\pi)$.
-        \item<2-> $\epsilon \sim \mathcal{E}$.
-        \item<3-> $y = \beta_h^T x + \epsilon$.
+        \item<3-> $h \in [k] \sim Mult(\pi)$.
+        \item<4-> $\epsilon \sim \mathcal{E}$.
+        \item<5-> $y = \beta_h^T x + \epsilon$.
       \end{itemize}
       }};
-    \node<4>[style=box] at (0,0) {\obj{%
+    \node<11>[style=box] at (0,0) {\obj{%
+    \includegraphics[width=\textwidth,height=4cm,keepaspectratio]{figures/disc-model.png}\\
     {\bf Given:} $\mathcal{D} = (x_i, y_i)_{i=1}^{n}$,\\
@@
     %\node<2-> at (6,0) {Discriminative Models};
-    \node[style=box] at (6,-1.5) {\obj{%
-      \vspace{12em}
-      Diagram
+    \node[style=box] at (6,0) {\obj{%
+    \includegraphics<1>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-1.png}
+    \includegraphics<2>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-2.png}
+    \includegraphics<3-4>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-3a.png}
+    \includegraphics<5>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-3b.png}
+    \includegraphics<6>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-4a.png}
+    \includegraphics<7-8>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-4b.png}
+    \includegraphics<9>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-4c.png}
+    \includegraphics<10>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-5.png}
+    \includegraphics<11>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-6.png}
+      \\{\em Data}
       }};
@@
     % x, y
-    \node[style=box] at (0,0) {\objw{36em}{%
-      \vspace{10em}
-      {\em moment-map picture}
+    \node[style=box] at (0,0) {\objw{12cm}{%
+      \includegraphics<1>[width=\textwidth,height=8cm,keepaspectratio]{figures/mom-1.png}
+      \includegraphics<2>[width=\textwidth,height=8cm,keepaspectratio]{figures/mom-2.png}
+      \includegraphics<3>[width=\textwidth,height=8cm,keepaspectratio]{figures/mom-3.png}
+      \includegraphics<4>[width=\textwidth,height=8cm,keepaspectratio]{figures/mom-4.png}
+      \includegraphics<5>[width=\textwidth,height=8cm,keepaspectratio]{figures/mom-5.png}
+      %\\{\em moment-map}
       }};
+      \node<4->[style=box] at (3,-2) {$(\hat m - m^*) \to \normal{0}{\frac{1}{n} \Sigma}$};
+      \node<5->[style=box] at (-2,-2) {$(\hat \theta - \theta^*) \to \normal{0}{\frac{1}{n} (\Delta \mathcal{M}^{-1})^T \Sigma \Delta \mathcal{M}^{-1}}$};
   \end{tikzpicture}
   \end{centering}
-  \begin{itemize}
-    \item<1-> Moments are expected to converge at a $1/\sqrt{n}$ rate, by CLT.
-    \item<2-> A continuous mapping $\mathcal{M}^{-1}$ will also converge at a $1/\sqrt{n}$ rate.
-  \end{itemize}
+  %\begin{itemize}
+  %  \item<4-> Moments are expected to converge at a $1/\sqrt{n}$ rate, by CLT.
+  %  \item<5-> A continuous mapping $\mathcal{M}^{-1}$ will also converge at a $1/\sqrt{n}$ rate.
+  %\end{itemize}
 
+    \note[item]{I'm unsure whether this stuff really belongs here.}
     \note[item]{We approach the problem by using the method of moments on three conditional views. }
@@
     % x, y
-    \node[style=box] at (0,0) {\objw{36em}{%
-      \vspace{10em}
-      {\em Work out the moments in the generative case, with indicative diagrams for matrices, tensors, etc.}
+    \node<1>[style=box] (moments) at (0,0) {\objw{12cm}{%
+    \includegraphics[width=\textwidth,height=6cm,keepaspectratio]{figures/moments.png}
+      }};
+    \node<2->[style=box] (moments) at (0,0) {\objw{12cm}{%
+      \begin{itemize}
+        \item<2-> {\bf Tensor Power Method} for orthonormal $\beta_h$,
+          \begin{align*}
+            M_3(\beta_h,\beta_h) 
+              &= \sum_{h'=1}^k \pi_{h'} (\beta_{h}^T \beta_{h'})^2 \beta_{h'} \\
+              &= \sum_{h'=1}^k \pi_{h'} \delta_{hh'} \beta_{h'} \\
+              &= \pi_{h'} \beta_{h}.
+          \end{align*}
+        \item<3-> Use $M_2$ to whiten $M_3$.
+      \end{itemize}
+      }};
+    \node<1->[style=box,below=0.1cm of moments]  {\objw{12cm}{%
+      \begin{align*}
+        \underbrace{\E[x]}_{M_1} &= \sum_{h=1}^k \pi_h \beta_h & 
+        \underbrace{\E[x\tp{2}]}_{M_2} &= \sum_{h=1}^k \pi_h \beta_h\tp{2} & 
+        \underbrace{\E[x\tp{3}]}_{M_3} &= \sum_{h=1}^k \pi_h \beta_h\tp{3}. 
+      \end{align*}
     }};
@@
     \node[style=box] at (0,0) {\objw{36em}{%
-      \vspace{10em}
-      {\em Try to work out the moments in our case.}
+        \begin{align*}
+          \action<1->{
+            \E[y] &= \E[\beta_h^T x + \epsilon] \\
+                  &= M_1^T \E[x]\\
+          }
+          \action<2->{ 
+            \E[y x] &= \E[(\beta_h^T x) x + \epsilon x] \\
+              &= M_1^T \E[x\tp{2}]\\
+          }
+          \action<3->{ 
+            \E[y^2] &= \E[(\beta_h^T x)^2 + 2 \epsilon (\beta_h^T x) + \epsilon^2 ] \\
+                    &= \E[\trace(\beta_h \beta_h^T x x^T)] + \E[\epsilon^2 ] \\
+                    &= \trace(M_2 \E[x\tp{2}]) + \E[\epsilon^2 ] \\
+          }
+          \action<4->{ 
+            \E[y^p x\tp{q}] 
+                    &= \trace(M_p \E[x\tp{p+q}]) + \cdots.
+          }
+        \end{align*}
       }};
@@
     % x, y
-    \node[style=box] at (0,0) {\objw{18em}{%
-      \vspace{10em}
-      {\em Different tactic: show that $y^p$ are linear measurements of the moments of the parameters $M_p$! }
+    \node[style=box] at (0,0) {\objw{12cm}{%
+        \begin{align*}
+          \action<+->{
+            y &= \beta_h^T x + \epsilon \\
+          }
+          \action<+->{
+            &= \underbrace{M_1^T x}_{\textrm{linear measurement}} + \underbrace{(\beta_h - M_1)^T x + \epsilon}_{\textrm{noise}} \\
+          }
+          \action<+->{
+            y^2 
+              &= (\beta_h^T x)^2 + 2 (\beta_h^T x) \epsilon + \epsilon^2 \\
+              &= \trace(\beta_h\tp{2} x\tp{2}) + 2 (\beta_h^T x) \epsilon + \epsilon^2 \\
+              &= \trace(M_2 x\tp{2}) + \E[\epsilon^2] + \trace((\beta_h\tp{2} - M_2) x\tp{2}) + 2 (\beta_h^T x) \epsilon \\
+              &= \trace(M_2 x\tp{2}) + \underbrace{\E[\epsilon^2]}_{\textrm{bias}} + \underbrace{\trace((\beta_h\tp{2} - M_2) x\tp{2}) + 2 (\beta_h^T x) \epsilon}_{\textrm{noise}} \\
+          }
+          \action<+->{ 
+          y^3 &= \innerp{M_3}{x\tp{3}} + \textrm{bias} + \textrm{noise}.
+          }
+        \end{align*}
       }};
-
-    \node[style=box] at (6,0) {\objw{8em}{%
-      {$y^p$ are linear measurements of the moments of the parameters $M_p$!}
-      }};
+    %\node[style=box] at (6,0) {\objw{8em}{%
+    %  {$y^p$ are linear measurements of the moments of the parameters $M_p$!}
+    %  }};
   \end{tikzpicture}
@@
 \begin{frame}
-  \frametitle{Efficient Consistent Recovery}
+  \frametitle{Method of Moments for the Mixture of Linear Regressions.}
 
@@
     % x, y
-    \node[style=box] at (0,0) {\objw{12em}{%
-      \vspace{10em}
-      Regression step
+    \node[style=box] (note) at (0,0) {\objw{12cm}{%
+      {$y^p$ are linear measurements of the moments of the parameters $M_p$!}
       }};
+    \node[style=box,below=0.1cm of note] {\objw{12cm}{%
+      \includegraphics<1>[width=\textwidth,height=8cm,keepaspectratio]{figures/regression-1.png}
+      \includegraphics<2>[width=\textwidth,height=8cm,keepaspectratio]{figures/regression-2.png}
+      \includegraphics<3>[width=\textwidth,height=8cm,keepaspectratio]{figures/regression-3.png}
+      }};
+  \end{tikzpicture}
+  \end{centering}
+\end{frame}
 
-    \node[style=box] at (6,0) {\objw{12em}{%
-      The tensor factorization method.
+\begin{frame}
+  \frametitle{Perturbation Analysis}
+
+  \begin{centering}
+  \begin{tikzpicture}
+    % x, y
+    \node[style=box] at (0,0) {\objw{12cm}{%
+      \includegraphics<1>[width=\textwidth,height=8cm,keepaspectratio]{figures/err-1.png}
+      \includegraphics<2>[width=\textwidth,height=8cm,keepaspectratio]{figures/err-2.png}
+      \includegraphics<3>[width=\textwidth,height=8cm,keepaspectratio]{figures/err-3.png}
+      \includegraphics<4>[width=\textwidth,height=8cm,keepaspectratio]{figures/err-4.png}
+      \includegraphics<5>[width=\textwidth,height=8cm,keepaspectratio]{figures/err-5.png}
       }};

### 2013-06-05 doc/icml-presentation/presentation.tex
Added experiments stuff too

@@
     %\node<1->[below=0.5cm of gen-models] (disc) {Discriminative Models};
-    \node<1->[style=box,below=1.0cm of gen] (disc) {\obj{%
+    \node<2->[style=box,below=1.0cm of gen] (disc) {\obj{%
     {\bf Discriminative Models}
@@
       }};
-      \node<4->[style=box] at (4,-3) {\objw{8cm}{%
+      \node<4->[style=box] at (4,-3) {\objw{12cm}{%
       {\bf How can we apply these methods for discriminative models?}
@@
     % x, y
-    \node<1>[style=box] (moments) at (0,0) {\objw{12cm}{%
+    \node<1->[style=box]  (moments) at (0,0) {\objw{12cm}{%
+      \begin{align*}
+        \underbrace{\E[x]}_{M_1} &= \sum_{h=1}^k \pi_h \beta_h & 
+        \underbrace{\E[x\tp{2}]}_{M_2} &= \sum_{h=1}^k \pi_h \beta_h\tp{2} & 
+        \underbrace{\E[x\tp{3}]}_{M_3} &= \sum_{h=1}^k \pi_h \beta_h\tp{3}. 
+      \end{align*}
+    }};
+    \node<1>[style=box,below=0.1cm of moments] {\objw{12cm}{%
     \includegraphics[width=\textwidth,height=6cm,keepaspectratio]{figures/moments.png}
       }};
-    \node<2->[style=box] (moments) at (0,0) {\objw{12cm}{%
+    \node<2->[style=box,below=0.1cm of moments] {\objw{12cm}{%
       \begin{itemize}
@@
       \end{itemize}
-      }};
-    \node<1->[style=box,below=0.1cm of moments]  {\objw{12cm}{%
-      \begin{align*}
-        \underbrace{\E[x]}_{M_1} &= \sum_{h=1}^k \pi_h \beta_h & 
-        \underbrace{\E[x\tp{2}]}_{M_2} &= \sum_{h=1}^k \pi_h \beta_h\tp{2} & 
-        \underbrace{\E[x\tp{3}]}_{M_3} &= \sum_{h=1}^k \pi_h \beta_h\tp{3}. 
-      \end{align*}
       }};
@@
         \begin{align*}
-          \action<1->{
+          \action<1->{%
             \E[y] &= \E[\beta_h^T x + \epsilon] \\
@@
           }
-          \action<2->{ 
+          \action<2->{%
             \E[y x] &= \E[(\beta_h^T x) x + \epsilon x] \\
@@
           }
-          \action<3->{ 
+          \action<3->{% 
             \E[y^2] &= \E[(\beta_h^T x)^2 + 2 \epsilon (\beta_h^T x) + \epsilon^2 ] \\
@@
           }
-          \action<4->{ 
+          \action<4->{% 
             \E[y^p x\tp{q}] 
@@
   \frametitle{Experimental Insights}
+
+  \begin{tikzpicture}
+    % x, y
+    \node<1-4> at (0,0) {%
+      \includegraphics<1>[width=\textwidth,height=5cm,keepaspectratio]{figures/1-8-3.png}
+      \includegraphics<2->[width=\textwidth,height=5cm,keepaspectratio]{figures/1-8-3-em.png}
+      };
+    % x, y
+    \node<1-4> at (6,0) {%
+      \includegraphics<3>[width=\textwidth,height=5cm,keepaspectratio]{figures/1-8-3-spec.png}
+      \includegraphics<4->[width=\textwidth,height=5cm,keepaspectratio]{figures/1-8-3-specm.png}
+      };
+      \node<5-> at (0,0) {\objw{12cm}{%
+      \begin{small}
+  \begin{tabular}{ r r r c c c }
+\hline
+$b$ & $d$ & $k$ & Spectral & EM & Spectral + EM \\
+\hline
+  1 & 4 & 2 & 2.45 $\pm$ 3.68 & 0.28 $\pm$ 0.82 & {\bf 0.17 $\pm$ 0.57} \\
+  2 & 5 & 2 & 1.38 $\pm$ 0.84 & {\bf 0.00 $\pm$ 0.00} & {\bf 0.00 $\pm$ 0.00} \\
+  2 & 5 & 3 & 2.92 $\pm$ 1.71 & 0.43 $\pm$ 1.07 & {\bf 0.31 $\pm$ 1.02} \\
+  2 & 6 & 2 & 2.33 $\pm$ 0.67 & 0.63 $\pm$ 1.29 & {\bf 0.01 $\pm$ 0.01} \\
+\hline
+\end{tabular}
+      \end{small}
+    }};
+  \end{tikzpicture}
+
   \begin{itemize}
-    \item {\bf Spectral Experts provides a good initialization to EM.}
-      \item We simulated the performance of spectral experts and compared it to EM. 
-\item In a non-trivial number of cases, EM did not converge to the right parameters and got stuck in local optima.
-\item The parameters estimated by the spectral method were not very good even with $O(10^5)$ samples.
-\item However, EM initialized with these parameters did extremely well. 
-\item This finding that spectral methods should be a good initialization for EM is not surprising.
+      \item Spectral methods can provide a good initializaton.
+  \end{itemize}
+
+    \note[item]{Spectral Experts provides a good initialization to EM.}
+    \note[item]{We simulated the performance of spectral experts and compared it to EM.}
+    \note[item]{In a non-trivial number of cases, EM did not converge to the right parameters and got stuck in local optima.}
+    \note[item]{The parameters estimated by the spectral method were not very good even with $O(10^5)$ samples.}
+    \note[item]{However, EM initialized with these parameters did extremely well.}
+\note[item]{This finding that spectral methods should be a good initialization for EM is not surprising.}
+\note[item]{%
   \begin{itemize}
@@
   \end{itemize}
-  \end{itemize}
+  }
 \end{frame}
@@
   \begin{itemize}
-    \item {\bf We can learn the parameters of conditional models where the observed moments on their own have such sparse information.}
-    \item The key intuition is that we could construct a regression problem to learn the moments of the parameters from their projections onto the data.
-    \item We found that while the parameters learned this way are usually not better than those learned via local methods like EM, etc., they are a good way to initialize these local methods.
+    \item Consistent estimator for the mixture of linear regressions with polynomial sample complexity.
+    \item Tackle conditional moments using regression while exploiting structure.
+    \item Method of moment estimates can be a good initialization for EM.
   \end{itemize}
@@
 
-\end{document}
-
 % TikZ

### 2013-06-06 doc/icml-presentation/presentation.tex
Minor change
Updated to Percy's notes

--- 2013-06-05 doc/icml-presentation/presentation.tex
@@
 \begin{frame}
+  \frametitle{Spectral Methods}
+  \begin{itemize}
+    \item Subspace identification [Ljung, 1987]
+    \item Observable operator models [Jaeger, 2000]
+    \item {\bf Hidden Markov models [Hsu et al, 2009]}
+    \item Low-treewidth graphs [Parikh et al., 2012]
+    \item Weighted finite state automata [Balle \& Mohri, 2012]
+  \end{itemize}
+  \begin{itemize}
+    \item Mixture of Gaussians [Kalai/Moitra/Valiant, 2010]
+    \item {\bf Mixture models, HMMs [Anandkumar/Hsu/Kakade, 2012]}
+    \item Latent Dirichlet Allocation [Anandkumar/Hsu/Kakade, 2012]
+    \item Stochastic block models [Anandkumar/Ge/Hsu/Kakade, 2012]
+    \item Linear Bayesian networks [Anandkumar/Hsu/Javanmard/Kakade, 2012]
+  \end{itemize}
+  \note[item]{\todo{Talk about observable operator stuff.}}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Spectral Methods: Tensor Factorization}
+  \begin{centering}
+  \begin{tikzpicture}
+    % x, y
+    \node<1>[style=box] (moments) at (0,0) {\objw{12cm}{%
+    \includegraphics[width=\textwidth,height=6cm,keepaspectratio]{figures/moments.png}
+      }};
+    \node<2->[style=box] (moments) at (0,0) {\objw{12cm}{%
+      \begin{itemize}
+        \item<2-> {\bf Tensor Power Method} for orthonormal $\beta_h$,
+          \begin{align*}
+            M_3(\beta_h,\beta_h) 
+              &= \sum_{h'=1}^k \pi_{h'} (\beta_{h}^T \beta_{h'})^2 \beta_{h'} \\
+              &= \sum_{h'=1}^k \pi_{h'} \delta_{hh'} \beta_{h'} \\
+              &= \pi_{h'} \beta_{h}.
+          \end{align*}
+        \item<3-> Use $M_2$ to whiten $M_3$.
+      \end{itemize}
+      }};
+    \node<1->[style=box,below=0.1cm of moments]  {\objw{12cm}{%
+      \begin{align*}
+        \underbrace{\E[x]}_{M_1} &= \sum_{h=1}^k \pi_h \beta_h & 
+        \underbrace{\E[x\tp{2}]}_{M_2} &= \sum_{h=1}^k \pi_h \beta_h\tp{2} & 
+        \underbrace{\E[x\tp{3}]}_{M_3} &= \sum_{h=1}^k \pi_h \beta_h\tp{3}. 
+      \end{align*}
+    }};
+  \end{tikzpicture}
+  \end{centering}
+
+    \note[item]{\bf We can recover the means of a generative LVM by exploiting their symmetric structure.}
+    \note[item]{\begin{align*} M_2 &= \sum_{h=1}^k \pi_h \beta_h\tp{2} & M_3 &= \sum_{h=1}^k \pi_h \beta_h\tp{3}.  \end{align*}}
+      \note[item]{In general, $M_2$ is insufficient to identify the model, because it is invariant to rotations of $B$. }
+      \note[item]{If the $\beta_h$ were orthogonal, then the eigenvectors and eigenvalues of $M_3$ would give us the answer. }
+      \note[item]{In general, we need both $M_2$ and $M_3$. We can use the whitening transformation for $M_2$ to whiten $M_3$ such that it has a orthogonal decomposition.}
+      \note[item]{The robust tensor power method by \cite{AnandkumarHsuGe2012} find stable eigenvectors.}
+\end{frame}
+
+\begin{comment}
+\begin{frame}
   \frametitle{Spectral Methods for Generative Models.}
@@
 \end{frame}
+\end{comment}
 
@@
 
+\begin{comment}
 \begin{frame}
@@
 \end{frame}
+\end{comment}
 
-\begin{frame}
-  \frametitle{Method of Moments for Generative LVMs.}
-
-  \begin{centering}
-  \begin{tikzpicture}
-    % x, y
-    \node<1>[style=box] (moments) at (0,0) {\objw{12cm}{%
-    \includegraphics[width=\textwidth,height=6cm,keepaspectratio]{figures/moments.png}
-      }};
-    \node<2->[style=box] (moments) at (0,0) {\objw{12cm}{%
-      \begin{itemize}
-        \item<2-> {\bf Tensor Power Method} for orthonormal $\beta_h$,
-          \begin{align*}
-            M_3(\beta_h,\beta_h) 
-              &= \sum_{h'=1}^k \pi_{h'} (\beta_{h}^T \beta_{h'})^2 \beta_{h'} \\
-              &= \sum_{h'=1}^k \pi_{h'} \delta_{hh'} \beta_{h'} \\
-              &= \pi_{h'} \beta_{h}.
-          \end{align*}
-        \item<3-> Use $M_2$ to whiten $M_3$.
-      \end{itemize}
-      }};
-    \node<1->[style=box,below=0.1cm of moments]  {\objw{12cm}{%
-      \begin{align*}
-        \underbrace{\E[x]}_{M_1} &= \sum_{h=1}^k \pi_h \beta_h & 
-        \underbrace{\E[x\tp{2}]}_{M_2} &= \sum_{h=1}^k \pi_h \beta_h\tp{2} & 
-        \underbrace{\E[x\tp{3}]}_{M_3} &= \sum_{h=1}^k \pi_h \beta_h\tp{3}. 
-      \end{align*}
-    }};
-  \end{tikzpicture}
-  \end{centering}
-
-    \note[item]{\bf We can recover the means of a generative LVM by exploiting their symmetric structure.}
-    \note[item]{\begin{align*} M_2 &= \sum_{h=1}^k \pi_h \beta_h\tp{2} & M_3 &= \sum_{h=1}^k \pi_h \beta_h\tp{3}.  \end{align*}}
-      \note[item]{In general, $M_2$ is insufficient to identify the model, because it is invariant to rotations of $B$. }
-      \note[item]{If the $\beta_h$ were orthogonal, then the eigenvectors and eigenvalues of $M_3$ would give us the answer. }
-      \note[item]{In general, we need both $M_2$ and $M_3$. We can use the whitening transformation for $M_2$ to whiten $M_3$ such that it has a orthogonal decomposition.}
-      \note[item]{The robust tensor power method by \cite{AnandkumarHsuGe2012} find stable eigenvectors.}
-\end{frame}
-
+\begin{comment}
 \begin{frame}
@@
 
+\begin{frame}
+  \frametitle{}
+  Thank you.
+\end{frame}
+
 \end{document}

### 2013-06-07 doc/icml-presentation/presentation.tex

--- 2013-06-05 doc/icml-presentation/presentation.tex
@@
         \item Semi-Latent Topic Models (Wang and Mori, 2009)
-        \item Easy to include features and tend to be more accurate.
+        \item {\bf Easy to include features and tend to be more accurate.}
       \end{itemize}
@@
 \end{frame}
+\end{comment}
 
-\begin{frame}
-  \frametitle{Method of Moments for Generative LVMs.}
+% \begin{frame}
+%   \frametitle{Method of Moments for Generative LVMs.}
+% 
+%   \begin{centering}
+%   \begin{tikzpicture}
+%     % x, y
+%     \node<1->[style=box]  (moments) at (0,0) {\objw{12cm}{%
+%       \begin{align*}
+%         \underbrace{\E[x]}_{M_1} &= \sum_{h=1}^k \pi_h \beta_h & 
+%         \underbrace{\E[x\tp{2}]}_{M_2} &= \sum_{h=1}^k \pi_h \beta_h\tp{2} & 
+%         \underbrace{\E[x\tp{3}]}_{M_3} &= \sum_{h=1}^k \pi_h \beta_h\tp{3}. 
+%       \end{align*}
+%     }};
+%     \node<1>[style=box,below=0.1cm of moments] {\objw{12cm}{%
+%     \includegraphics[width=\textwidth,height=6cm,keepaspectratio]{figures/moments.png}
+%       }};
+%     \node<2->[style=box,below=0.1cm of moments] {\objw{12cm}{%
+%       \begin{itemize}
+%         \item<2-> {\bf Tensor Power Method} for orthonormal $\beta_h$,
+%           \begin{align*}
+%             M_3(\beta_h,\beta_h) 
+%               &= \sum_{h'=1}^k \pi_{h'} (\beta_{h}^T \beta_{h'})^2 \beta_{h'} \\
+%               &= \sum_{h'=1}^k \pi_{h'} \delta_{hh'} \beta_{h'} \\
+%               &= \pi_{h'} \beta_{h}.
+%           \end{align*}
+%         \item<3-> Use $M_2$ to whiten $M_3$.
+%       \end{itemize}
+%       }};
+%   \end{tikzpicture}
+%   \end{centering}
+% 
+%     \note[item]{\bf We can recover the means of a generative LVM by exploiting their symmetric structure.}
+%     \note[item]{\begin{align*} M_2 &= \sum_{h=1}^k \pi_h \beta_h\tp{2} & M_3 &= \sum_{h=1}^k \pi_h \beta_h\tp{3}.  \end{align*}}
+%       \note[item]{In general, $M_2$ is insufficient to identify the model, because it is invariant to rotations of $B$. }
+%       \note[item]{If the $\beta_h$ were orthogonal, then the eigenvectors and eigenvalues of $M_3$ would give us the answer. }
+%       \note[item]{In general, we need both $M_2$ and $M_3$. We can use the whitening transformation for $M_2$ to whiten $M_3$ such that it has a orthogonal decomposition.}
+%       \note[item]{The robust tensor power method by \cite{AnandkumarHsuGe2012} find stable eigenvectors.}
+% \end{frame}
 
-  \begin{centering}
-  \begin{tikzpicture}
-    % x, y
-    \node<1->[style=box]  (moments) at (0,0) {\objw{12cm}{%
-      \begin{align*}
-        \underbrace{\E[x]}_{M_1} &= \sum_{h=1}^k \pi_h \beta_h & 
-        \underbrace{\E[x\tp{2}]}_{M_2} &= \sum_{h=1}^k \pi_h \beta_h\tp{2} & 
-        \underbrace{\E[x\tp{3}]}_{M_3} &= \sum_{h=1}^k \pi_h \beta_h\tp{3}. 
-      \end{align*}
-    }};
-    \node<1>[style=box,below=0.1cm of moments] {\objw{12cm}{%
-    \includegraphics[width=\textwidth,height=6cm,keepaspectratio]{figures/moments.png}
-      }};
-    \node<2->[style=box,below=0.1cm of moments] {\objw{12cm}{%
-      \begin{itemize}
-        \item<2-> {\bf Tensor Power Method} for orthonormal $\beta_h$,
-          \begin{align*}
-            M_3(\beta_h,\beta_h) 
-              &= \sum_{h'=1}^k \pi_{h'} (\beta_{h}^T \beta_{h'})^2 \beta_{h'} \\
-              &= \sum_{h'=1}^k \pi_{h'} \delta_{hh'} \beta_{h'} \\
-              &= \pi_{h'} \beta_{h}.
-          \end{align*}
-        \item<3-> Use $M_2$ to whiten $M_3$.
-      \end{itemize}
-      }};
-  \end{tikzpicture}
-  \end{centering}
-
-    \note[item]{\bf We can recover the means of a generative LVM by exploiting their symmetric structure.}
-    \note[item]{\begin{align*} M_2 &= \sum_{h=1}^k \pi_h \beta_h\tp{2} & M_3 &= \sum_{h=1}^k \pi_h \beta_h\tp{3}.  \end{align*}}
-      \note[item]{In general, $M_2$ is insufficient to identify the model, because it is invariant to rotations of $B$. }
-      \note[item]{If the $\beta_h$ were orthogonal, then the eigenvectors and eigenvalues of $M_3$ would give us the answer. }
-      \note[item]{In general, we need both $M_2$ and $M_3$. We can use the whitening transformation for $M_2$ to whiten $M_3$ such that it has a orthogonal decomposition.}
-      \note[item]{The robust tensor power method by \cite{AnandkumarHsuGe2012} find stable eigenvectors.}
-\end{frame}
-
+\begin{comment}
 \begin{frame}

### 2013-06-07 doc/icml-presentation/presentation.tex
Updates to the presentation

@@
     % Tasks.
-    \node[style=box] (pos) at (0,0) {\objw{8cm}{%
-      \includegraphics[width=\textwidth]{figures/pos.png}\\
-      \hfill Part-of-Speech Tagging
-      }};
-      \node[style=box,below=0.5cm of pos] (dep) {\objw{8cm}{%
-      \includegraphics[width=\textwidth]{figures/dep-parse.png}\\
-      \hfill Dependency Parsing
-      }};
-      \node[style=box, below=0.5cm of dep] (act) {\objw{8cm}{%
-      \includegraphics[width=\textwidth,height=3cm,keepaspectratio]{figures/activity-recognition.png}\\
-      \hfill Activity Recognition
+    \node[style=box] (gen) at (0,0) {\objw{6cm}{%
+      \includegraphics[width=\textwidth,height=3cm,keepaspectratio]{figures/gen.png}\\
+      \hfill Generative Models
       }};
 
-    \node<1->[style=box] (gen) at (6,0) {\obj{%
+    \node<1->[style=box,below=0.1cm of gen] {\obj{%
       {\bf Generative Models}
       \begin{itemize}
+        \item Gaussian Mixture Models
         \item Hidden Markov Models
-        \item Conditional Random Fields
+        \item Latent Dirichlet Allocation
         \item PCFGs
@@
       }};
-    %\node<1->[below=0.5cm of gen-models] (disc) {Discriminative Models};
-    \node<2->[style=box,below=1.0cm of gen] (disc) {\obj{%
+    \node<2->[style=box] (disc) at (6,0) {\objw{6cm}{%
+      \includegraphics[width=\textwidth,height=3cm,keepaspectratio]{figures/disc.png}\\
+      \hfill Discriminative Models
+      }};
+    \node<2->[style=box,below=0.1cm of disc] (disc-models) {\obj{%
     {\bf Discriminative Models}
       \begin{itemize}
-        \item Max-Entropy Markov Model (Toutanova and Manning, 2000)
-        \item Discriminative Parsers (Liang et.\ al, 2006)
+        \item Logistic Regression
+        \item Conditional Random Fields
         \item Semi-Latent Topic Models (Wang and Mori, 2009)
-        \item {\bf Easy to include features and tend to be more accurate.}
+        %\item Discriminative Parsers (Liang et.\ al, 2006)
+        \item \dots
       \end{itemize}
       }};
+    \node<3->[style=box,below=0.1cm of disc-models] {\obj{%
+      {\bf Easy to include features and tend to be more accurate.}
+    }};
+
   \end{tikzpicture}
   \end{centering}
+
+  % Really do known Gen (HMMs, ) vs Disc (CRFs) models
 
@@
     \node[style=box] at (0,-1.5) {\objw{12cm}{%
-    \includegraphics<1>[width=\textwidth,height=8cm,keepaspectratio]{figures/likelihood.png}
-    \includegraphics<2>[width=\textwidth,height=8cm,keepaspectratio]{figures/likelihood-em1.png}
-    \includegraphics<3-4>[width=\textwidth,height=8cm,keepaspectratio]{figures/likelihood-em2.png}
-    \includegraphics<5>[width=\textwidth,height=8cm,keepaspectratio]{figures/likelihood-mom.png}
-    \includegraphics<6>[width=\textwidth,height=8cm,keepaspectratio]{figures/likelihood-mom1.png}
+    \includegraphics<1>[width=\textwidth,height=6cm,keepaspectratio]{figures/likelihood.png}
+    \includegraphics<2>[width=\textwidth,height=6cm,keepaspectratio]{figures/likelihood-mle.png}
+    \includegraphics<3>[width=\textwidth,height=6cm,keepaspectratio]{figures/likelihood-em.png}
+    \includegraphics<4>[width=\textwidth,height=6cm,keepaspectratio]{figures/likelihood-mom.png}
       \hfill{\em expected log-likelihood}
@@
 
+  % Simple message: MLE is consistent but intractable, EM is efficient not but consistent. Can we get something in between.
+
   \begin{itemize}
     \item<1-> Log-likelihood function is non-convex.
-    \item<4-> This is a problem that doesn't go away even with infinite data.
-    \item<5-> Can we build {\bf consistent estimators} that get better with more data?
+    \item<2-> MLE is consistent but intractable.
+    \item<3-> EM is efficient but inconsistent.
+    \item<4-> Can we build {\bf efficient yet consistent estimators} that get better with more data?
   \end{itemize}
@@
   \begin{itemize}
+     \item Representation Learning
+    \begin{itemize}
     \item Subspace identification [Ljung, 1987]
@@
   \end{itemize}
+     \item<2-> Parameter Estimation
   \begin{itemize}
@@
   \end{itemize}
+  \end{itemize}
   \note[item]{\todo{Talk about observable operator stuff.}}
@@
 \begin{frame}
-  \frametitle{Spectral Methods: Tensor Factorization}
+  \frametitle{Spectral Methods via Tensor Factorization}
   \begin{centering}
@@
     \node<1>[style=box] (moments) at (0,0) {\objw{12cm}{%
-    \includegraphics[width=\textwidth,height=6cm,keepaspectratio]{figures/moments.png}
+      \begin{align*}
+        %\underbrace{\E[x]}_{M_1} &= \sum_{h=1}^k \pi_h \beta_h & 
+        %\underbrace{\E[x\tp{2}]}_{M_2} &= \sum_{h=1}^k \pi_h \beta_h\tp{2} & 
+        \underbrace{\E[x\tp{3}]}_{M_3} &= \sum_{h=1}^k \pi_h \beta_h\tp{3}. 
+      \end{align*}
     }};
-    \node<2->[style=box] (moments) at (0,0) {\objw{12cm}{%
-      \begin{itemize}
-        \item<2-> {\bf Tensor Power Method} for orthonormal $\beta_h$,
+    \node<1>[style=box,below=0.4cm of moments] {\objw{12cm}{%
+    \includegraphics[width=\textwidth,height=6cm,keepaspectratio]{figures/tensor.png}
+      }};
+    \node<2>[style=box] (tensorf) at (0,0) {\objw{12cm}{%
+      Given orthonormal $\beta_h$,
           \begin{align*}
             M_3(\beta_h,\beta_h) 
-              &= \sum_{h'=1}^k \pi_{h'} (\beta_{h}^T \beta_{h'})^2 \beta_{h'} \\
-              &= \sum_{h'=1}^k \pi_{h'} \delta_{hh'} \beta_{h'} \\
+              %&= \sum_{h'=1}^k \pi_{h'} (\beta_{h}^T \beta_{h'})^2 \beta_{h'} \\
+              %&= \sum_{h'=1}^k \pi_{h'} \delta_{hh'} \beta_{h'} \\
               &= \pi_{h'} \beta_{h}.
           \end{align*}
-        \item<3-> Use $M_2$ to whiten $M_3$.
-      \end{itemize}
       }};
-    \node<1->[style=box,below=0.1cm of moments]  {\objw{12cm}{%
-      \begin{align*}
-        \underbrace{\E[x]}_{M_1} &= \sum_{h=1}^k \pi_h \beta_h & 
-        \underbrace{\E[x\tp{2}]}_{M_2} &= \sum_{h=1}^k \pi_h \beta_h\tp{2} & 
-        \underbrace{\E[x\tp{3}]}_{M_3} &= \sum_{h=1}^k \pi_h \beta_h\tp{3}. 
-      \end{align*}
+
+    \node<2>[style=box,below=0.1cm of tensorf] {\objw{12cm}{%
+    \includegraphics[width=\textwidth,height=6cm,keepaspectratio]{figures/tensor-decomp.png}
+      \hfill {Tensor Power Method}
       }};
@@
 
-\begin{comment}
-\begin{frame}
-  \frametitle{Spectral Methods for Generative Models.}
-  \begin{centering}
-  \begin{tikzpicture}
-    % x, y
-    \node<1-2>[style=box] (pos) at (0,0) {\objw{6cm}{%
-      \includegraphics[width=\textwidth]{figures/pos.png}\\
-      \hfill Part-of-Speech Tagging
-      }};
-    \node<3->[style=box] (dep) at (0,0) {\objw{6cm}{%
-      \includegraphics[width=\textwidth]{figures/dep-parse.png}\\
-      \hfill Dependency Parsing
-      }};
-
-    %\node<2-> at (6,0) {Discriminative Models};
-    \node<1-2>[style=box] (ahk) at (6,0) {\objw{8cm}{%
-      \includegraphics[width=\textwidth,height=4cm,keepaspectratio]{figures/three-view.png}\\
-      \begin{itemize}
-        \item Solution for multi-view case using tensor factorization (Anandkumar et.\ al, 2012).
-        \item<2> HMMs (Anandkumar et.\ al, 2012) 
-        \item<2> LDA (Anandkumar et.\ al, 2012) 
-      \end{itemize}
-      \hfill 
-      }};
-      \node<3->[style=box] at (6,0) {\objw{8cm}{%
-      \includegraphics[width=\textwidth,height=4cm,keepaspectratio]{figures/unmixing.png}\\
-      \begin{itemize}
-        \item Reduce the problem to multi-view learning via tree-unmixing (Hsu et.\ al, 2012). 
-      \end{itemize}
-      }};
-      \node<4->[style=box] at (4,-3) {\objw{12cm}{%
-      {\bf How can we apply these methods for discriminative models?}
-      }};
-  \end{tikzpicture}
-  \end{centering}
-  \note[item]{\todo{Talk about observable operator stuff.}}
-\end{frame}
-\end{comment}
-
 \begin{frame}
@@
     % x, y
-    \node<1-10>[style=box] at (0,0) {\obj{%
+    \node<1-9>[style=box] at (0,0) {\obj{%
+    \includegraphics[width=\textwidth,height=4cm,keepaspectratio]{figures/disc-model.png}\\
     {\bf Parameters:} $\pi = [\pi_1, \cdots, \pi_k]$ and $B = [ \beta_1 | \cdots | \beta_k ]$.
-    \includegraphics[width=\textwidth,height=4cm,keepaspectratio]{figures/disc-model.png}\\
     For a given $x$,
       \begin{itemize}
-        \item<3-> $h \in [k] \sim Mult(\pi)$.
-        \item<4-> $\epsilon \sim \mathcal{E}$.
-        \item<5-> $y = \beta_h^T x + \epsilon$.
+        \item<2-> $h \in [k] \sim Mult(\pi)$.
+        \item<3-> $\epsilon \sim \mathcal{E}$.
+        \item<4-> $y = \beta_h^T x + \epsilon$.
       \end{itemize}
       }};
-    \node<11>[style=box] at (0,0) {\obj{%
+    \node<10>[style=box] at (0,0) {\obj{%
     \includegraphics[width=\textwidth,height=4cm,keepaspectratio]{figures/disc-model.png}\\
@@
     \node[style=box] at (6,0) {\obj{%
-    \includegraphics<1>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-1.png}
-    \includegraphics<2>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-2.png}
-    \includegraphics<3-4>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-3a.png}
-    \includegraphics<5>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-3b.png}
-    \includegraphics<6>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-4a.png}
-    \includegraphics<7-8>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-4b.png}
-    \includegraphics<9>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-4c.png}
-    \includegraphics<10>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-5.png}
-    \includegraphics<11>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-6.png}
+    \includegraphics<1>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-2.png}
+    \includegraphics<2-3>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-3a.png}
+    \includegraphics<4>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-3b.png}
+    \includegraphics<5>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-4a.png}
+    \includegraphics<6-7>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-4b.png}
+    \includegraphics<8>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-4c.png}
+    \includegraphics<9>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-5.png}
+    \includegraphics<10>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-6.png}
       \\{\em Data}
@@
 \end{frame}
-
-\begin{comment}
-\begin{frame}
-  \frametitle{Method of Moments.}
-
-  \begin{centering}
-  \begin{tikzpicture}
-    % x, y
-    \node[style=box] at (0,0) {\objw{12cm}{%
-      \includegraphics<1>[width=\textwidth,height=8cm,keepaspectratio]{figures/mom-1.png}
-      \includegraphics<2>[width=\textwidth,height=8cm,keepaspectratio]{figures/mom-2.png}
-      \includegraphics<3>[width=\textwidth,height=8cm,keepaspectratio]{figures/mom-3.png}
-      \includegraphics<4>[width=\textwidth,height=8cm,keepaspectratio]{figures/mom-4.png}
-      \includegraphics<5>[width=\textwidth,height=8cm,keepaspectratio]{figures/mom-5.png}
-      %\\{\em moment-map}
-      }};
-      \node<4->[style=box] at (3,-2) {$(\hat m - m^*) \to \normal{0}{\frac{1}{n} \Sigma}$};
-      \node<5->[style=box] at (-2,-2) {$(\hat \theta - \theta^*) \to \normal{0}{\frac{1}{n} (\Delta \mathcal{M}^{-1})^T \Sigma \Delta \mathcal{M}^{-1}}$};
-  \end{tikzpicture}
-  \end{centering}
-  %\begin{itemize}
-  %  \item<4-> Moments are expected to converge at a $1/\sqrt{n}$ rate, by CLT.
-  %  \item<5-> A continuous mapping $\mathcal{M}^{-1}$ will also converge at a $1/\sqrt{n}$ rate.
-  %\end{itemize}
-
-    \note[item]{I'm unsure whether this stuff really belongs here.}
-    \note[item]{We approach the problem by using the method of moments on three conditional views. }
-    \note[item]{Consider a moment map $\mathcal{M}$ that maps the parameters $\theta$ to the moments $m$. For a Gaussian, we have that $\mathcal{M} = (\mu, \sigma^2)$.}
-    \note[item]{In general, we will compute the inverse of the moment map to learn the parameters from the sample estimates.}
-    \note[item]{By the central limit theorem, our sample estimates of the moments converge at a $1/\sqrt{n}$ rate, so we expect that our parameters will also converge at this rate.}
-\end{frame}
-\end{comment}
 
@@
 
-\begin{comment}
-\begin{frame}
-  \frametitle{Method of Moments for the Mixture of Linear Regressions.}
-
-  \begin{centering}
-  \begin{tikzpicture}
-    % x, y
-    \node[style=box] at (0,0) {\objw{36em}{%
-        \begin{align*}
-          \action<1->{%
-            \E[y] &= \E[\beta_h^T x + \epsilon] \\
-                  &= M_1^T \E[x]\\
-          }
-          \action<2->{%
-            \E[y x] &= \E[(\beta_h^T x) x + \epsilon x] \\
-              &= M_1^T \E[x\tp{2}]\\
-          }
-          \action<3->{% 
-            \E[y^2] &= \E[(\beta_h^T x)^2 + 2 \epsilon (\beta_h^T x) + \epsilon^2 ] \\
-                    &= \E[\trace(\beta_h \beta_h^T x x^T)] + \E[\epsilon^2 ] \\
-                    &= \trace(M_2 \E[x\tp{2}]) + \E[\epsilon^2 ] \\
-          }
-          \action<4->{% 
-            \E[y^p x\tp{q}] 
-                    &= \trace(M_p \E[x\tp{p+q}]) + \cdots.
-          }
-        \end{align*}
-      }};
-  \end{tikzpicture}
-  \end{centering}
-      \note[item]{In the discriminative case, we do not observe $M_2$ directly from the data, but rather, as we'll see, linear measurements of them.}
-      \note[item]{Try to work out the moments for this case and see how they {\tt \#fail} epically.}
-\end{frame}
-\end{comment}
-
 \begin{frame}
@@
           \action<+->{
-            y &= \beta_h^T x + \epsilon \\
+          y &= \underbrace{\beta_h^T}_{\textrm{random}} x + \epsilon \\
           }
           \action<+->{
-            &= \underbrace{M_1^T x}_{\textrm{linear measurement}} + \underbrace{(\beta_h - M_1)^T x + \epsilon}_{\textrm{noise}} \\
-          }
-          \action<+->{
-            y^2 
-              &= (\beta_h^T x)^2 + 2 (\beta_h^T x) \epsilon + \epsilon^2 \\
-              &= \trace(\beta_h\tp{2} x\tp{2}) + 2 (\beta_h^T x) \epsilon + \epsilon^2 \\
-              &= \trace(M_2 x\tp{2}) + \E[\epsilon^2] + \trace((\beta_h\tp{2} - M_2) x\tp{2}) + 2 (\beta_h^T x) \epsilon \\
-              &= \trace(M_2 x\tp{2}) + \underbrace{\E[\epsilon^2]}_{\textrm{bias}} + \underbrace{\trace((\beta_h\tp{2} - M_2) x\tp{2}) + 2 (\beta_h^T x) \epsilon}_{\textrm{noise}} \\
-          }
-          \action<+->{ 
-          y^3 &= \innerp{M_3}{x\tp{3}} + \textrm{bias} + \textrm{noise}.
+            &= \underbrace{\E[\beta_h]^T x}_{\textrm{linear measurement}} + \underbrace{(\beta_h - M_1)^T x + \epsilon}_{\textrm{noise}} \\
           }
@@
     % x, y
-    \node[style=box] (note) at (0,0) {\objw{12cm}{%
-      {$y^p$ are linear measurements of the moments of the parameters $M_p$!}
+    \node[style=box] (measurements) at (0,0) {\objw{12cm}{%
+        \begin{align*}
+          \action<+->{%
+            y &= \underbrace{M_1^T x}_{\textrm{linear measurement}} + \underbrace{(\beta_h - M_1)^T x + \epsilon}_{\textrm{noise}} \\
+            M_1 &= \E[\beta_h] \\
+          }
+          \action<+->{%
+            y^2 
+              %&= (\beta_h^T x)^2 + 2 (\beta_h^T x) \epsilon + \epsilon^2 \\
+              %&= \trace(\beta_h\tp{2} x\tp{2}) + 2 (\beta_h^T x) \epsilon + \epsilon^2 \\
+              %&= \trace(M_2 x\tp{2}) + \E[\epsilon^2] + \trace((\beta_h\tp{2} - M_2) x\tp{2}) + 2 (\beta_h^T x) \epsilon \\
+              &= \trace(M_2 x\tp{2}) + \underbrace{\E[\epsilon^2]}_{\textrm{bias}} + \underbrace{\trace((\beta_h\tp{2} - M_2) x\tp{2}) + 2 (\beta_h^T x) \epsilon}_{\textrm{noise}} \\
+              M_2 &= \E[\beta_h\tp{2}] \\
+          }
+          \action<+->{%
+          y^3 &= \innerp{M_3}{x\tp{3}} + \textrm{bias} + \textrm{noise} \\
+          \innerp{A}{B} &= \sum_{ijk} A_{ijk} B_{ijk}.
+          }
+        \end{align*}
       }};
-    \node[style=box,below=0.1cm of note] {\objw{12cm}{%
-      \includegraphics<1>[width=\textwidth,height=8cm,keepaspectratio]{figures/regression-1.png}
-      \includegraphics<2>[width=\textwidth,height=8cm,keepaspectratio]{figures/regression-2.png}
-      \includegraphics<3>[width=\textwidth,height=8cm,keepaspectratio]{figures/regression-3.png}
+    \node<4->[style=box,below=0.1cm of measurements]  {\objw{12cm}{%
+    We can use \uncover<5->{{\color{blue}{(low-rank)}}} regression to recover $M_p$ from linear measurements.\\
+    \uncover<5->{\hfill {\color{blue}{{[Negabhan and Wainwright (2009); Tomioka (2011)]}}}}
       }};
+    %\node[style=box] at (6,0) {\objw{8em}{%
+    %  {$y^p$ are linear measurements of the moments of the parameters $M_p$!}
+    %  }};
   \end{tikzpicture}
   \end{centering}
+
+    \note[item]{The powers of the response variables $y^p$ are linear measurements of the moments of the parameters $M_p$!}
+    \note[item]{As noted earlier, the first problem we run into is that we can't observe the moments of the parameters $B$ and $\pi$ directly!}
+    \note[item]{However, observe that 
+    \begin{align}
+      y &= \innerp{\beta_h}{x} + \epsilon \\
+        &= \innerp{M_1}{x} + \underbrace{{\beta_h - M_1}{x} + \epsilon }_{\textrm{noise}},
+    \end{align}
+    where $M_1 = \sum_{h=1}^k \pi_h \beta_h$, the mean regression coefficient.}
+    \note[item]{We note that while the noise term is dependent on $x$, it has a zero-mean. Thus, we could potentially recover $M_1$ through regression.}
+    \note[item]{Similarly, we can cast the remaining two terms as regression problems.}
+    \note[item]{An additional fact that we can exploit is that both $M_2$ and $M_3$ are low rank, so we can use low-rank regression to recover estimates $\hat M_2$ and $\hat M_3$ efficiently from data.}
 \end{frame}
+
+% \begin{frame}
+%   \frametitle{Method of Moments for the Mixture of Linear Regressions.}
+% 
+%   \begin{centering}
+%   \begin{tikzpicture}
+%     % x, y
+%     \node[style=box] (note) at (0,0) {\objw{12cm}{%
+%       {$y^p$ are linear measurements of the moments of the parameters $M_p$!}
+%       }};
+%     \node[style=box,below=0.1cm of note] {\objw{12cm}{%
+%       \includegraphics<1>[width=\textwidth,height=8cm,keepaspectratio]{figures/regression-1.png}
+%       \includegraphics<2>[width=\textwidth,height=8cm,keepaspectratio]{figures/regression-2.png}
+%       \includegraphics<3>[width=\textwidth,height=8cm,keepaspectratio]{figures/regression-3.png}
+%       }};
+%   \end{tikzpicture}
+%   \end{centering}
+% \end{frame}
 
@@
       \includegraphics<1>[width=\textwidth,height=8cm,keepaspectratio]{figures/err-1.png}
-      \includegraphics<2>[width=\textwidth,height=8cm,keepaspectratio]{figures/err-2.png}
-      \includegraphics<3>[width=\textwidth,height=8cm,keepaspectratio]{figures/err-3.png}
-      \includegraphics<4>[width=\textwidth,height=8cm,keepaspectratio]{figures/err-4.png}
-      \includegraphics<5>[width=\textwidth,height=8cm,keepaspectratio]{figures/err-5.png}
+      %\includegraphics<2>[width=\textwidth,height=8cm,keepaspectratio]{figures/err-2.png}
+      \includegraphics<2>[width=\textwidth,height=8cm,keepaspectratio]{figures/err-3.png}
+      %\includegraphics<4>[width=\textwidth,height=8cm,keepaspectratio]{figures/err-4.png}
+      \includegraphics<3>[width=\textwidth,height=8cm,keepaspectratio]{figures/err-5.png}
       }};
@@
     % x, y
-    \node<1-4> at (0,0) {%
+    \node<1-4> (em) at (-3,-0.5) {%
       \includegraphics<1>[width=\textwidth,height=5cm,keepaspectratio]{figures/1-8-3.png}
@@
     % x, y
-    \node<1-4> at (6,0) {%
+    \node<1-4> at (3,-0.5) {%
       \includegraphics<3>[width=\textwidth,height=5cm,keepaspectratio]{figures/1-8-3-spec.png}
@@
       };
-      \node<5-> at (0,0) {\objw{12cm}{%
-      \begin{small}
-  \begin{tabular}{ r r r c c c }
-\hline
-$b$ & $d$ & $k$ & Spectral & EM & Spectral + EM \\
-\hline
-  1 & 4 & 2 & 2.45 $\pm$ 3.68 & 0.28 $\pm$ 0.82 & {\bf 0.17 $\pm$ 0.57} \\
-  2 & 5 & 2 & 1.38 $\pm$ 0.84 & {\bf 0.00 $\pm$ 0.00} & {\bf 0.00 $\pm$ 0.00} \\
-  2 & 5 & 3 & 2.92 $\pm$ 1.71 & 0.43 $\pm$ 1.07 & {\bf 0.31 $\pm$ 1.02} \\
-  2 & 6 & 2 & 2.33 $\pm$ 0.67 & 0.63 $\pm$ 1.29 & {\bf 0.01 $\pm$ 0.01} \\
-\hline
-\end{tabular}
-      \end{small}
-    }};
+    \node<5-> at (0,-0.5) {%
+      \includegraphics[width=\textwidth,height=5cm,keepaspectratio]{figures/hist.png}
+    };
+
+    \node[above=0.1 cm of em] {%
+        $y = x + x^4 + x^7 + \epsilon$
+    };
+
+%      \begin{small}
+%  \begin{tabular}{ r r r c c c }
+%\hline
+%$b$ & $d$ & $k$ & Spectral & EM & Spectral + EM \\
+%\hline
+%  1 & 4 & 2 & 2.45 $\pm$ 3.68 & 0.28 $\pm$ 0.82 & {\bf 0.17 $\pm$ 0.57} \\
+%  2 & 5 & 2 & 1.38 $\pm$ 0.84 & {\bf 0.00 $\pm$ 0.00} & {\bf 0.00 $\pm$ 0.00} \\
+%  2 & 5 & 3 & 2.92 $\pm$ 1.71 & 0.43 $\pm$ 1.07 & {\bf 0.31 $\pm$ 1.02} \\
+%  2 & 6 & 2 & 2.33 $\pm$ 0.67 & 0.63 $\pm$ 1.29 & {\bf 0.01 $\pm$ 0.01} \\
+%\hline
+%\end{tabular}
+%      \end{small}
+%    }};
   \end{tikzpicture}
 
-  \begin{itemize}
-      \item Spectral methods can provide a good initializaton.
-  \end{itemize}
-
+  \note[item]{Stable!}
     \note[item]{Spectral Experts provides a good initialization to EM.}
@@
 \begin{frame}
+  \frametitle{Experimental Insights}
+
+  \begin{centering}
+  \begin{tikzpicture}
+    \node[style=box] at (0,-1.5) {\objw{12cm}{%
+      \includegraphics<1>[width=\textwidth,height=6cm,keepaspectratio]{figures/likelihood-mom.png}
+      \includegraphics<2>[width=\textwidth,height=6cm,keepaspectratio]{figures/likelihood-mom1.png}
+        \hfill{\em expected log-likelihood}
+      }};
+  \end{tikzpicture}
+  \end{centering}
+
+\end{frame}
+
+\begin{frame}
   \frametitle{Conclusions}
   \begin{itemize}
-    \item Consistent estimator for the mixture of linear regressions with polynomial sample complexity.
-    \item Tackle conditional moments using regression while exploiting structure.
-    \item Method of moment estimates can be a good initialization for EM.
+    \item {\bf Consistent estimator} for the mixture of linear regressions with {\bf polynomial sample and computational complexity}.
+    \item<2-> Derive conditional moments using regression.
+    \item<3-> Method of moment estimates can be a good initialization for EM.
   \end{itemize}

### 2013-06-10 doc/icml-presentation/presentation.tex

@@
 
-  \begin{centering}
   \begin{tikzpicture}
@@
       }};
+    \node[scale=0.7] at (0.5,0.5) {Latent variable};
 
-    \node<1->[style=box,below=0.1cm of gen] {\obj{%
+    \node<2->[style=box,below=0.1cm of gen] {\obj{%
       {\bf Generative Models}
@@
       }};
-    \node<2->[style=box] (disc) at (6,0) {\objw{6cm}{%
+    \node<3->[style=box] (disc) at (6,0) {\objw{6cm}{%
       \includegraphics[width=\textwidth,height=3cm,keepaspectratio]{figures/disc.png}\\
@@
       }};
-    \node<2->[style=box,below=0.1cm of disc] (disc-models) {\obj{%
+    \node<4->[style=box,below=0.1cm of disc] (disc-models) {\obj{%
     {\bf Discriminative Models}
@@
         \item Logistic Regression
-        \item Conditional Random Fields
-        \item Semi-Latent Topic Models (Wang and Mori, 2009)
+        \item Latent CRFs
+        \item Discriminative LDA
+        %\item Semi-Latent Topic Models (Wang and Mori, 2009)
         %\item Discriminative Parsers (Liang et.\ al, 2006)
@@
       }};
-    \node<3->[style=box,below=0.1cm of disc-models] {\obj{%
+    \node<5->[style=box,below=0.1cm of disc-models] {\obj{%
       {\bf Easy to include features and tend to be more accurate.}
@@
   \end{tikzpicture}
-  \end{centering}
 
@@
 \begin{frame}
-  \frametitle{Spectral Methods}
+  \frametitle{Method of Moments a.k.a.\ Spectral Methods}
   \begin{itemize}
-     \item Representation Learning
+    \item<+-> Method of Moments [Pearson, 1894]
+    \item<+-> Observable operator models
     \begin{itemize}
@@
       \item Observable operator models [Jaeger, 2000]
+        \note[item]{Entry point into machine learning}
       \item {\bf Hidden Markov models [Hsu et al, 2009]}
+        \note[item]{Entry point into machine learning}
       \item Low-treewidth graphs [Parikh et al., 2012]
@@
     \end{itemize}
-     \item<2-> Parameter Estimation
+     \item<+-> Parameter Estimation
   \begin{itemize}
@@
   \end{itemize}
-  \note[item]{\todo{Talk about observable operator stuff.}}
 \end{frame}
@@
 \begin{frame}
-  \frametitle{Spectral Methods via Tensor Factorization}
-  \begin{centering}
+  \frametitle{Example: Gaussian Mixture Model}
   \begin{tikzpicture}
     % x, y
-    \node<1>[style=box] (moments) at (0,0) {\objw{12cm}{%
+    \node<1->[style=box] (defn) at (0,0) {\objw{6cm}{%
+    \begin{itemize} 
+      \item Generative process:
       \begin{align*}
-        %\underbrace{\E[x]}_{M_1} &= \sum_{h=1}^k \pi_h \beta_h & 
-        %\underbrace{\E[x\tp{2}]}_{M_2} &= \sum_{h=1}^k \pi_h \beta_h\tp{2} & 
-        \underbrace{\E[x\tp{3}]}_{M_3} &= \sum_{h=1}^k \pi_h \beta_h\tp{3}. 
+        h \in [k] &\sim \Mult(\pi) \\
+        x &\sim \normal{\beta_h}{\sigma^2}.
+      \end{align*}
+    \end{itemize}
+    }};
+    \node<1->[style=box] (models) at (6,0) {\objw{6cm}{%
+      \includegraphics[width=0.45\textwidth,height=3cm,keepaspectratio]{figures/gen.png}
+      \includegraphics[width=0.45\textwidth,height=3cm,keepaspectratio]{figures/mog.png}
+    }};
+
+    \node<2->[style=box,below=0.1cm of defn] {\objw{6cm}{%
+    \begin{itemize} 
+      \item Moments:
+    \begin{align*}
+        \action<2->{%
+        \E[x|h] &= \beta_h \\
+        }
+        \action<3->{%
+        \E[x] &= \sum_h \pi_h \beta_h \\
+        }
+        \action<4->{%
+        \E[x\tp{2}] &= \sum_h \pi_h \beta_h{\hlmath{\tp{2}}} \\
+                    &= \sum_h \pi_h (\beta_h \beta_h^T) \\
+                    }
+        \action<5->{%
+        \E[x\tp{3}] &= \sum_h \pi_h \beta_h\tp{3}.
+        }
+      \end{align*}
+    \end{itemize}
+    }};
+    %\tikzrect{1}{1};
+    \uncover<4->{%
+    \tikzrect{black,fill=white}{(6,-2)}{1}{1};
+    \node at (7,-2.5) {$\E[x\tp{2}]$};
+    \node at (4.5,-2.5) {$d$};
+    }
+    \uncover<5->{%
+    \tikzcube{black,fill=white}{(6,-4,0)}{1}{1}{1};
+    \node at (7,-4.5,0) {$\E[x\tp{3}]$};
+    \node at (4.5,-4.5) {$d$};
+    \node at (4.5,-4.5) {$d$};
+    }
+  \end{tikzpicture}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{Aside: Tensor Operations}
+  \begin{tikzpicture}
+    % x, y
+    \node<1->[style=box] (defn) at (0,0) {\objw{6cm}{%
+    \begin{align*}
+        \action<1->{%
+        x\tp{3}_{ijk} &= x_i x_j x_k \comment{indexing}\\
+        }
+        \action<2->{%
+        A \cdot B &= \sum_{ijk} A_{ijk} B_{ijk} \comment{inner product}\\
+        }
+        \action<3->{%
+        A(u)_{ij} &= \sum_{k} A_{ijk} u_k \comment{projection}\\
+        }
+        \action<4->{%
+        A(u,v)_{i} &= \sum_{jk} A_{ijk} u_j v_k \comment{projection} \\
+        }
       \end{align*}
     }};
-    \node<1>[style=box,below=0.4cm of moments] {\objw{12cm}{%
+    \tikzcube{black,fill=white}{(6,2,0)}{1}{1}{1};
+    \node (pt) at (5.75,1.75,0) {$\hlmath{\times}$};
+    \uncover<3->{%
+    \tikzrect{black,fill=white}{(6,0)}{1}{1};
+    }
+    \uncover<4->{%
+    \tikzrect{black,fill=white}{(6,-2)}{0.1}{1};
+    }
+  \end{tikzpicture}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{Example: Gaussian Mixture Model}
+  \begin{tikzpicture}
+    % x, y
+    \node<1->[style=box] (moments) at (0,0) {\objw{6cm}{%
+      Assume $\beta_h$ to be orthogonal;
+      \begin{align*}
+        \E[x\tp{3}] &= \sum_{h=1}^k \pi_h \beta_h\tp{3}. 
+      \end{align*}
+    }};
+    \node<1->[style=box] (models) at (6,0) {\objw{6cm}{%
+      \includegraphics[width=0.45\textwidth,height=3cm,keepaspectratio]{figures/gen.png}
+      \includegraphics[width=0.45\textwidth,height=3cm,keepaspectratio]{figures/mog.png}
+    }};
+
+    \node<1-> at (0,-3) {\objw{6cm}{%
     \includegraphics[width=\textwidth,height=6cm,keepaspectratio]{figures/tensor.png}
       }};
-    \node<2>[style=box] (tensorf) at (0,0) {\objw{12cm}{%
-      Given orthonormal $\beta_h$,
+    \node<2->[style=box] at (6,-3) {\objw{6cm}{%
+      $\beta_h$ are eigenvectors!
         \begin{align*}
-            M_3(\beta_h,\beta_h) 
+          \E[x\tp{3}](\beta_h,\beta_h) 
             %&= \sum_{h'=1}^k \pi_{h'} (\beta_{h}^T \beta_{h'})^2 \beta_{h'} \\
             %&= \sum_{h'=1}^k \pi_{h'} \delta_{hh'} \beta_{h'} \\
-              &= \pi_{h'} \beta_{h}.
+            &= \pi_{h} \beta_{h}.
         \end{align*}
+
+        \hfill{\em Anandkumar, Hsu, et.\ al (2012)}
       }};
-
-    \node<2>[style=box,below=0.1cm of tensorf] {\objw{12cm}{%
-    \includegraphics[width=\textwidth,height=6cm,keepaspectratio]{figures/tensor-decomp.png}
-      \hfill {Tensor Power Method}
-      }};
+%
+%    \node<2>[style=box,below=0.1cm of tensorf] {\objw{12cm}{%
+%    \includegraphics[width=\textwidth,height=6cm,keepaspectratio]{figures/tensor-decomp.png}
+%      \hfill {Tensor Power Method}
+%      }};
   \end{tikzpicture}
-  \end{centering}
 
@@
 
-  \begin{centering}
   \begin{tikzpicture}
     % x, y
-    \node<1-9>[style=box] at (0,0) {\obj{%
-    \includegraphics[width=\textwidth,height=4cm,keepaspectratio]{figures/disc-model.png}\\
-    {\bf Parameters:} $\pi = [\pi_1, \cdots, \pi_k]$ and $B = [ \beta_1 | \cdots | \beta_k ]$.
-    For a given $x$,
+    \node[style=box] at (0,0) {\objw{4cm}{%
+    \includegraphics[width=\textwidth,height=4cm,keepaspectratio]{figures/disc.png}
       \begin{itemize}
-        \item<2-> $h \in [k] \sim Mult(\pi)$.
-        \item<3-> $\epsilon \sim \mathcal{E}$.
+        \item<2-> $h \in [k] \sim \Mult(\pi)$.
         \item<4-> $y = \beta_h^T x + \epsilon$.
@@
       }};
-    \node<10>[style=box] at (0,0) {\obj{%
-    \includegraphics[width=\textwidth,height=4cm,keepaspectratio]{figures/disc-model.png}\\
-    {\bf Given:} $\mathcal{D} = (x_i, y_i)_{i=1}^{n}$,\\
-    can we {\bf estimate parameters} $\pi$ and $B = [ \beta_1 | \cdots | \beta_k ]$?
+
+    \node[style=box] at (6,0) {\obj{%
+    \includegraphics<1>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-2.png}
+    \includegraphics<2>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-3a.png}
+    \includegraphics<3>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-3b.png}
+    \includegraphics<4>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-4a.png}
+    \includegraphics<5>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-4b.png}
+    %\includegraphics<6>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-4c.png}
+    \includegraphics<6>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-5.png}
+      }};
+  \end{tikzpicture}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Mixture of Linear Regressions}
+
+  \begin{tikzpicture}
+    \node[style=box] (data) at (0,0) {\obj{%
+    \includegraphics[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-6.png}
+      \\{\em Data}
       }};
 
-    %\node<2-> at (6,0) {Discriminative Models};
-    \node[style=box] at (6,0) {\obj{%
-    \includegraphics<1>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-2.png}
-    \includegraphics<2-3>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-3a.png}
-    \includegraphics<4>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-3b.png}
-    \includegraphics<5>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-4a.png}
-    \includegraphics<6-7>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-4b.png}
-    \includegraphics<8>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-4c.png}
-    \includegraphics<9>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-5.png}
-    \includegraphics<10>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-6.png}
-      \\{\em Data}
+    % x, y
+    \node[style=box] (params) at (7,0) {\objw{4cm}{%
+    \includegraphics[width=\textwidth,height=4cm,keepaspectratio]{figures/disc.png}
+    \\
+    {\bf Parameters: $\pi, B$}
     }};
+
+    \draw[->,dashed] (data) -- node[above]{?} (params);
+
   \end{tikzpicture}
-  \end{centering}
-
-  \note[item]{First step, consider a simple discriminative model, mixture of linear regressions.}
-  \note[item]{The mixture of linear regressions model defines a conditional distribution over a response $y \in \Re$ given covariates $x \in \Re^d$.}
-     \note[item]{The generative procedure is as follows,
-       \begin{enumerate}
-        \item Draw a mixture component $h \in [k] \sim Mult(\pi)$, where $\pi = [\pi_1 | \dots | \pi_k]$ defines the mixture proportions.
-        \item Draw the noise $\epsilon \sim \mathcal{E}$, where $\mathcal{E}$ is the noise distribution.
-        \item Set $y = \beta_h^T x + \epsilon$, where $\{\beta_h\}_{h=1}^{k}$ are the conditional means of the regression coefficients.
-      \end{enumerate}}
-   \note[item]{The parameters that we would like to learn from this model are $\pi$ and $B = [ \beta_1 | \cdots | \beta_k ]$.}
 \end{frame}
@@
 % 
-%   \begin{centering}
 %   \begin{tikzpicture}
@@
 %   \end{tikzpicture}
-%   \end{centering}
 % 
@@
           \action<+->{
-            &= \underbrace{\E[\beta_h]^T x}_{\textrm{linear measurement}} + \underbrace{(\beta_h - M_1)^T x + \epsilon}_{\textrm{noise}} \\
+            &= \underbrace{\E[\beta_h]^T x}_{\textrm{linear measurement}} + \underbrace{(\beta_h - \E[\beta_h])^T x + \epsilon}_{\textrm{noise}} \\
           }
@@
       y &= \innerp{\beta_h}{x} + \epsilon \\
-        &= \innerp{M_1}{x} + \underbrace{{\beta_h - M_1}{x} + \epsilon }_{\textrm{noise}},
+        &= \innerp{\E[\beta_h]}{x} + \underbrace{{\beta_h - \E[\beta_h]}{x} + \epsilon }_{\textrm{noise}},
     \end{align}
@@
           \action<+->{%
-            y &= \underbrace{M_1^T x}_{\textrm{linear measurement}} + \underbrace{(\beta_h - M_1)^T x + \epsilon}_{\textrm{noise}} \\
-            M_1 &= \E[\beta_h] \\
+            y &= \underbrace{\E[\beta_h]^T \odot x}_{\textrm{linear measurement}} &\quad & + \underbrace{(\beta_h - \E[\beta_h])^T x + \epsilon}_{\textrm{noise}} \\
           }
@@
             y^2 
-              %&= (\beta_h^T x)^2 + 2 (\beta_h^T x) \epsilon + \epsilon^2 \\
-              %&= \trace(\beta_h\tp{2} x\tp{2}) + 2 (\beta_h^T x) \epsilon + \epsilon^2 \\
-              %&= \trace(M_2 x\tp{2}) + \E[\epsilon^2] + \trace((\beta_h\tp{2} - M_2) x\tp{2}) + 2 (\beta_h^T x) \epsilon \\
-              &= \trace(M_2 x\tp{2}) + \underbrace{\E[\epsilon^2]}_{\textrm{bias}} + \underbrace{\trace((\beta_h\tp{2} - M_2) x\tp{2}) + 2 (\beta_h^T x) \epsilon}_{\textrm{noise}} \\
-              M_2 &= \E[\beta_h\tp{2}] \\
+              &= \E[\beta_h\tp{2}] \odot x\tp{2} &+ \textrm{bias} &+ \textrm{noise} \\
           }
           \action<+->{%
-          y^3 &= \innerp{M_3}{x\tp{3}} + \textrm{bias} + \textrm{noise} \\
-          \innerp{A}{B} &= \sum_{ijk} A_{ijk} B_{ijk}.
+          y^3 &= \E[\beta_h\tp{2}] \odot x\tp{3} &+ \textrm{bias} &+ \textrm{noise} \\
           }
         \end{align*}
+
+        \uncover<4->{%
+          We can use regression to recover $M_p$ from linear measurements.\\
+          }
       }};
-    \node<4->[style=box,below=0.1cm of measurements]  {\objw{12cm}{%
-    We can use \uncover<5->{{\color{blue}{(low-rank)}}} regression to recover $M_p$ from linear measurements.\\
-    \uncover<5->{\hfill {\color{blue}{{[Negabhan and Wainwright (2009); Tomioka (2011)]}}}}
+  \end{tikzpicture}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Method of Moments for the Mixture of Linear Regressions.}
+
+  \begin{tikzpicture}
+    % x, y
+    \node[style=box] (measurements) at (0,0) {\objw{12cm}{%
+      \includegraphics[width=\textwidth,height=6cm,keepaspectratio]{figures/tensor.png}
       }};
-    %\node[style=box] at (6,0) {\objw{8em}{%
-    %  {$y^p$ are linear measurements of the moments of the parameters $M_p$!}
-    %  }};
+    \node<2->[style=box,below=0.1cm of measurements] {\objw{12cm}{%
+      We can use low-rank regression techniques [Negabhan \& Wainwright 2009; Tomioka 2011].
+      \begin{align*}
+        \hat\E[\beta_h\tp{2}] &= \arg\min_{M} \left(y^2 - M] \odot x\tp{2}\right)^2_{(x,y)\in\mathcal{D}} + \|M\|_{*} \\
+        \|M\|_{*} &= \sum_i \sigma_i(M).
+      \end{align*}
+      }};
   \end{tikzpicture}
-  \end{centering}
-
-    \note[item]{The powers of the response variables $y^p$ are linear measurements of the moments of the parameters $M_p$!}
-    \note[item]{As noted earlier, the first problem we run into is that we can't observe the moments of the parameters $B$ and $\pi$ directly!}
-    \note[item]{However, observe that 
-    \begin{align}
-      y &= \innerp{\beta_h}{x} + \epsilon \\
-        &= \innerp{M_1}{x} + \underbrace{{\beta_h - M_1}{x} + \epsilon }_{\textrm{noise}},
-    \end{align}
-    where $M_1 = \sum_{h=1}^k \pi_h \beta_h$, the mean regression coefficient.}
-    \note[item]{We note that while the noise term is dependent on $x$, it has a zero-mean. Thus, we could potentially recover $M_1$ through regression.}
-    \note[item]{Similarly, we can cast the remaining two terms as regression problems.}
-    \note[item]{An additional fact that we can exploit is that both $M_2$ and $M_3$ are low rank, so we can use low-rank regression to recover estimates $\hat M_2$ and $\hat M_3$ efficiently from data.}
 \end{frame}
@@
 %   \end{tikzpicture}
-%   \end{centering}
 % \end{frame}
+
+\begin{frame}
+  \frametitle{Overview: Spectral Experts}
+
+  \begin{tikzpicture}
+    % x, y
+    \node[style=box] at (0,0) {\objw{12cm}{%
+      \includegraphics<1>[width=\textwidth,height=8cm,keepaspectratio]{figures/err-1.png}
+      }};
+  \end{tikzpicture}
+\end{frame}
 
@@
     % x, y
-    \node<1-4> (em) at (-3,-0.5) {%
+    \node<-4> (em) at (-3,-0.5) {%
       \includegraphics<1>[width=\textwidth,height=5cm,keepaspectratio]{figures/1-8-3.png}
@@
     % x, y
-    \node<1-4> at (3,-0.5) {%
+    \node<-4> at (3,-0.5) {%
       \includegraphics<3>[width=\textwidth,height=5cm,keepaspectratio]{figures/1-8-3-spec.png}
@@
     \node[above=0.1 cm of em] {%
-        $y = x + x^4 + x^7 + \epsilon$
+    $y = \underbrace{t + t^4 + t^7}_{x} + \epsilon$
     };
@@
       \includegraphics<1>[width=\textwidth,height=6cm,keepaspectratio]{figures/likelihood-mom.png}
-      \includegraphics<2>[width=\textwidth,height=6cm,keepaspectratio]{figures/likelihood-mom1.png}
+      \includegraphics<2>[width=\textwidth,height=6cm,keepaspectratio]{figures/likelihood-mom-em1.png}
+      \includegraphics<3>[width=\textwidth,height=6cm,keepaspectratio]{figures/likelihood-mom-em2.png}
         \hfill{\em expected log-likelihood}

### 2013-06-10 doc/icml-presentation/presentation.tex
Fixed regression figure, and slight reordering of presentation

@@
 \usepackage{tikz}
-\usetikzlibrary{positioning,shapes,shadows,arrows}
+\usetikzlibrary{calc,positioning,shapes,shadows,arrows}
 
@@
       \begin{itemize}
-        \item Logistic Regression
         \item Latent CRFs
         \item Discriminative LDA
+        \item Mixture of Experts
         %\item Semi-Latent Topic Models (Wang and Mori, 2009)
@@
 
-  % Really do known Gen (HMMs, ) vs Disc (CRFs) models
-
-    \note[item]{Would like to start by talking about the bigger picture of parameter estimation.}
-    \note[item]{Discriminative Latent Variable Models are useful, but hard to learn.}
-    \note[item]{Split models into generative and discriminative models.}
-    \note[item]{Discriminative LVMs combine the predictive accuracy with compact expressiveness.}
-    \note[item]{They have been used in several applications, e.g.\ object recognition, syntactic parsing, machine translation, etc.}
-    \note[item]{Learning parameters for these models is hard because they have non-convex likelihood functions.}
 \end{frame}
@@
     \includegraphics<4>[width=\textwidth,height=6cm,keepaspectratio]{figures/likelihood-mom.png}
-      \hfill{\em expected log-likelihood}
+      \hfill{\em expected negative log-likelihood}
       }};
@@
     \item<2-> MLE is consistent but intractable.
-    \item<3-> EM is efficient but inconsistent.
+    \item<3-> Local methods (EM, gradient descent, etc.) are tractable but inconsistent.
     \item<4-> Can we build {\bf efficient yet consistent estimators} that get better with more data?
@@
 
-  \note[item]{We would like to develop efficient consistent estimators for discriminative LVMS.}
-  \note[item]{Past approaches rely on local optimization, which are susceptible to local optima.}
-  \note[item]{And they don't just go away with more data.}
-  \note[item]{Recently, there has been work on consistent parameter
-      estimation for generative LVMS\@. We'd like to extend their work to
-      the discriminative case.}
 \end{frame}
@@
 \begin{frame}
-  \frametitle{Method of Moments a.k.a.\ Spectral Methods}
+  \frametitle{Related Work}
   \begin{itemize}
@@
       \item Observable operator models [Jaeger, 2000]
-        \note[item]{Entry point into machine learning}
-      \item {\bf Hidden Markov models [Hsu et al, 2009]}
-        \note[item]{Entry point into machine learning}
+      \item {\em Hidden Markov models [Hsu et al, 2009]}
       \item Low-treewidth graphs [Parikh et al., 2012]
@@
 \begin{frame}
+  \frametitle{Aside: Tensor Operations}
+  \begin{tikzpicture}
+    % x, y
+    \node<1->[style=box] (defn) at (0,0) {\objw{6cm}{%
+    \begin{align*}
+        \action<1->{%
+        x\tp{3}_{ijk} &= x_i x_j x_k \comment{tensoring}\\
+        }
+        \action<2->{%
+        \innerp{A}{B} &= \sum_{ijk} A_{ijk} B_{ijk} \comment{inner product}\\
+        }
+        \action<3->{%
+        A(u)_{ij} &= \sum_{k} A_{ijk} u_k \comment{projection}\\
+        }
+        \action<4->{%
+        A(u,v)_{i} &= \sum_{jk} A_{ijk} u_j v_k \comment{projection} \\
+        }
+      \end{align*}
+    }};
+    \tikzcube{black,fill=white}{(6,2,0)}{1}{1}{1};
+    %\node (pt) at (5.75,1.75,0) {$\hlmath{\times}$};
+    \uncover<3->{%
+    \tikzrect{black,fill=white}{(6,0)}{1}{1};
+    }
+    \uncover<4->{%
+    \tikzrect{black,fill=white}{(6,-2)}{0.1}{1};
+    }
+  \end{tikzpicture}
+
+\end{frame}
+
+\begin{frame}
   \frametitle{Example: Gaussian Mixture Model}
@@
       \begin{align*}
-        h \in [k] &\sim \Mult(\pi) \\
+        h &\sim \Mult(\pi) \\
         x &\sim \normal{\beta_h}{\sigma^2}.
@@
         \action<4->{%
-        \E[x\tp{2}] &= \sum_h \pi_h \beta_h{\hlmath{\tp{2}}} \\
-                    &= \sum_h \pi_h (\beta_h \beta_h^T) \\
+        \E[x\tp{2}] &= \sum_h \pi_h (\beta_h \beta_h^T) \\
+                    &= \sum_h \pi_h \beta_h{\tp{2}} \\
                     }
@@
 \begin{frame}
-  \frametitle{Aside: Tensor Operations}
-  \begin{tikzpicture}
-    % x, y
-    \node<1->[style=box] (defn) at (0,0) {\objw{6cm}{%
-    \begin{align*}
-        \action<1->{%
-        x\tp{3}_{ijk} &= x_i x_j x_k \comment{indexing}\\
-        }
-        \action<2->{%
-        A \cdot B &= \sum_{ijk} A_{ijk} B_{ijk} \comment{inner product}\\
-        }
-        \action<3->{%
-        A(u)_{ij} &= \sum_{k} A_{ijk} u_k \comment{projection}\\
-        }
-        \action<4->{%
-        A(u,v)_{i} &= \sum_{jk} A_{ijk} u_j v_k \comment{projection} \\
-        }
-      \end{align*}
-    }};
-    \tikzcube{black,fill=white}{(6,2,0)}{1}{1}{1};
-    \node (pt) at (5.75,1.75,0) {$\hlmath{\times}$};
-    \uncover<3->{%
-    \tikzrect{black,fill=white}{(6,0)}{1}{1};
-    }
-    \uncover<4->{%
-    \tikzrect{black,fill=white}{(6,-2)}{0.1}{1};
-    }
-  \end{tikzpicture}
-
-\end{frame}
-
-\begin{frame}
   \frametitle{Example: Gaussian Mixture Model}
@@
     \node<1->[style=box] (moments) at (0,0) {\objw{6cm}{%
-      Assume $\beta_h$ to be orthogonal;
+      If $\beta_h$ are orthogonal;
       \begin{align*}
@@
 
-    \node<1-> at (0,-3) {\objw{6cm}{%
+    \node<1-> at (6,-3) {\objw{6cm}{%
     \includegraphics[width=\textwidth,height=6cm,keepaspectratio]{figures/tensor.png}
       }};
-    \node<2->[style=box] at (6,-3) {\objw{6cm}{%
+    \node<2->[style=box] at (0,-3) {\objw{6cm}{%
       $\beta_h$ are eigenvectors!
@@
   \end{tikzpicture}
+\end{frame}
 
-    \note[item]{\bf We can recover the means of a generative LVM by exploiting their symmetric structure.}
-    \note[item]{\begin{align*} M_2 &= \sum_{h=1}^k \pi_h \beta_h\tp{2} & M_3 &= \sum_{h=1}^k \pi_h \beta_h\tp{3}.  \end{align*}}
-      \note[item]{In general, $M_2$ is insufficient to identify the model, because it is invariant to rotations of $B$. }
-      \note[item]{If the $\beta_h$ were orthogonal, then the eigenvectors and eigenvalues of $M_3$ would give us the answer. }
-      \note[item]{In general, we need both $M_2$ and $M_3$. We can use the whitening transformation for $M_2$ to whiten $M_3$ such that it has a orthogonal decomposition.}
-      \note[item]{The robust tensor power method by \cite{AnandkumarHsuGe2012} find stable eigenvectors.}
+\begin{frame}
+  \frametitle{Generative vs. Discriminative Latent Variable Models}
+
+  \begin{tikzpicture}
+    % Tasks.
+    \node[style=box] (gen) at (0,0) {\objw{6cm}{%
+      \includegraphics[width=\textwidth,height=3cm,keepaspectratio]{figures/gen.png}\\
+      \hfill Generative Models
+      }};
+    \node[scale=0.7] at (0.5,0.5) {Latent variable};
+
+    \node[style=box] (disc) at (6,0) {\objw{6cm}{%
+      \includegraphics[width=\textwidth,height=3cm,keepaspectratio]{figures/disc.png}\\
+      \hfill Discriminative Models
+      }};
+  \end{tikzpicture}
+
 \end{frame}
@@
       \begin{itemize}
-        \item<2-> $h \in [k] \sim \Mult(\pi)$.
+        \item<2-> $h \sim \Mult(\pi)$.
         \item<4-> $y = \beta_h^T x + \epsilon$.
@@
 % 
-%     \note[item]{\bf We can recover the means of a generative LVM by exploiting their symmetric structure.}
-%     \note[item]{\begin{align*} M_2 &= \sum_{h=1}^k \pi_h \beta_h\tp{2} & M_3 &= \sum_{h=1}^k \pi_h \beta_h\tp{3}.  \end{align*}}
-%       \note[item]{In general, $M_2$ is insufficient to identify the model, because it is invariant to rotations of $B$. }
-%       \note[item]{If the $\beta_h$ were orthogonal, then the eigenvectors and eigenvalues of $M_3$ would give us the answer. }
-%       \note[item]{In general, we need both $M_2$ and $M_3$. We can use the whitening transformation for $M_2$ to whiten $M_3$ such that it has a orthogonal decomposition.}
-%       \note[item]{The robust tensor power method by \cite{AnandkumarHsuGe2012} find stable eigenvectors.}
 % \end{frame}
@@
       }};
-    %\node[style=box] at (6,0) {\objw{8em}{%
-    %  {$y^p$ are linear measurements of the moments of the parameters $M_p$!}
-    %  }};
   \end{tikzpicture}
 
-    \note[item]{The powers of the response variables $y^p$ are linear measurements of the moments of the parameters $M_p$!}
-    \note[item]{As noted earlier, the first problem we run into is that we can't observe the moments of the parameters $B$ and $\pi$ directly!}
-    \note[item]{However, observe that 
-    \begin{align}
-      y &= \innerp{\beta_h}{x} + \epsilon \\
-        &= \innerp{\E[\beta_h]}{x} + \underbrace{{\beta_h - \E[\beta_h]}{x} + \epsilon }_{\textrm{noise}},
-    \end{align}
-    where $M_1 = \sum_{h=1}^k \pi_h \beta_h$, the mean regression coefficient.}
-    \note[item]{We note that while the noise term is dependent on $x$, it has a zero-mean. Thus, we could potentially recover $M_1$ through regression.}
-    \note[item]{Similarly, we can cast the remaining two terms as regression problems.}
-    \note[item]{An additional fact that we can exploit is that both $M_2$ and $M_3$ are low rank, so we can use low-rank regression to recover estimates $\hat M_2$ and $\hat M_3$ efficiently from data.}
 \end{frame}
@@
           \action<+->{%
-            y &= \underbrace{\E[\beta_h]^T \odot x}_{\textrm{linear measurement}} &\quad & + \underbrace{(\beta_h - \E[\beta_h])^T x + \epsilon}_{\textrm{noise}} \\
+          y &= \underbrace{\innerp{\E[\beta_h]^T}{x}}_{\textrm{linear measurement}} &\quad & + \underbrace{(\beta_h - \E[\beta_h])^T x + \epsilon}_{\textrm{noise}} \\
           }
@@
             y^2 
-              &= \E[\beta_h\tp{2}] \odot x\tp{2} &+ \textrm{bias} &+ \textrm{noise} \\
+            &= \innerp{\E[\beta_h\tp{2}]}{x\tp{2}} &+ \textrm{bias} &+ \textrm{noise} \\
           }
           \action<+->{%
-          y^3 &= \E[\beta_h\tp{2}] \odot x\tp{3} &+ \textrm{bias} &+ \textrm{noise} \\
+          y^3 &= \innerp{\E[\beta_h\tp{3}]}{x\tp{3}} &+ \textrm{bias} &+ \textrm{noise} \\
           }
@@
         \uncover<4->{%
-          We can use regression to recover $M_p$ from linear measurements.\\
+        We can use regression to recover $\E[\beta_h\tp{p}]$ from linear measurements.\\
           }
       }};
+
+    \node<5->[style=box,below=0.1cm of measurements.south] (tensor) {\objw{12cm}{%
+      \includegraphics[width=\textwidth,height=3cm,keepaspectratio]{figures/tensor.png}
+      }};
+
+  \end{tikzpicture}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Overview: Spectral Experts}
+
+  \begin{tikzpicture}
+    % x, y
+    % Nodes
+    \node[style=txt] (data1) at (0,0) {$\left\{ x, y \right\}_{(x,y) \in \sD}$};
+    \node[style=txt,below=1cm of data1.center] (data2) {$\left\{ x\tp{2}, y^2 \right\}_{(x,y) \in \sD}$};
+    \node[style=txt,below=1cm of data2.center] (data3) {$\left\{ x\tp{3}, y^3 \right\}_{(x,y) \in \sD}$};
+
+    \node[style=txt] (m1) at (4,0) {$\E_h\left[\beta_h\right]$};
+    \node[style=txt,below=1cm of m1.center] (m2) {$\E_h\left[\beta_h\tp{2}\right]$};
+    \node[style=txt,below=1cm of m2.center] (m3) {$\E_h\left[\beta_h\tp{3}\right]$};
+
+    \node[below=0.5cm of m2.center] (params-pos) {};
+    \node[style=txt,right=3.0cm of params-pos] (params) {$\pi, B$};
+
+    % Arrows
+    \draw[-latex] (data1) -- (m1);
+    \draw[-latex] (data2) -- (m2);
+    \draw[-latex] (data3) -- (m3);
+
+    % Bias influence
+    \draw[-latex,dashed,gray] (m1.east) -- ++(0.1cm,0) -- ++(0,-0.50cm) -- ++(-3.00cm,0) node (bias-1) {} -- ( bias-1 |- m3) -- (m3.west);
+
+    % Mean
+    \draw[-latex] (m2.east) -- ++(0.1cm,0) node (tf-1) {} -- (tf-1 |- params-pos) -- node[above,scale=0.5] {tensor factorization} (params);
+    \draw[-latex] (m3.east) -- ++(0.1cm,0) node (tf-2) {} -- (tf-2 |- params-pos) -- (params);
+
+    % Box for regression
+    \begin{pgfonlayer}{bg}
+    \draw[fill=blue,opacity=0.1,dashed] ($(data1.north east) + (-0.1cm,0.1cm)$) rectangle ($(m3.south west) + (0.1cm,-0.1cm)$);
+    \end{pgfonlayer}
+    \node[style=txt] (reg-label) at ($(data3) !.5! (m3) + (0, -1.0cm)$) {regression};
+
+    % tensor factorization
+    \begin{pgfonlayer}{bg}
+    \draw[fill=green,opacity=0.1,dashed] ($(m2.north east) + (-0.1cm,0.1cm)$) rectangle ($(params.south west) + (0.1cm,-1.0cm)$);
+    \end{pgfonlayer}
+    \node (params-mid) at ($(params-pos) !.5! (params)$) {};
+    \node[style=txt] at (params-mid |- reg-label) {tensor factorization};
   \end{tikzpicture}
@@
     % x, y
-    \node[style=box] (measurements) at (0,0) {\objw{12cm}{%
+    \node[style=box] (reg) at (0,0) {\objw{12cm}{%
+      \begin{align*}
+        \hat\E[\beta_h\tp{2}] &= \arg\min_{M} \left(y^2 - \innerp{M}{x\tp{2}}\right)^2_{(x,y)\in\mathcal{D}} \action<3->{+ \hlmath{\|M\|_{*}}} \\
+        \action<3->{%
+        \|M\|_{*} &= \sum_i \sigma_i(M).
+        }
+      \end{align*}
+      }};
+    \node<2->[style=box,below=0.1 of reg] (measurements) {\objw{12cm}{%
       \includegraphics[width=\textwidth,height=6cm,keepaspectratio]{figures/tensor.png}
-      }};
-    \node<2->[style=box,below=0.1cm of measurements] {\objw{12cm}{%
-      We can use low-rank regression techniques [Negabhan \& Wainwright 2009; Tomioka 2011].
-      \begin{align*}
-        \hat\E[\beta_h\tp{2}] &= \arg\min_{M} \left(y^2 - M] \odot x\tp{2}\right)^2_{(x,y)\in\mathcal{D}} + \|M\|_{*} \\
-        \|M\|_{*} &= \sum_i \sigma_i(M).
-      \end{align*}
       }};
@@
 \end{frame}
-
-% \begin{frame}
-%   \frametitle{Method of Moments for the Mixture of Linear Regressions.}
-% 
-%   \begin{tikzpicture}
-%     % x, y
-%     \node[style=box] (note) at (0,0) {\objw{12cm}{%
-%       {$y^p$ are linear measurements of the moments of the parameters $M_p$!}
-%       }};
-%     \node[style=box,below=0.1cm of note] {\objw{12cm}{%
-%       \includegraphics<1>[width=\textwidth,height=8cm,keepaspectratio]{figures/regression-1.png}
-%       \includegraphics<2>[width=\textwidth,height=8cm,keepaspectratio]{figures/regression-2.png}
-%       \includegraphics<3>[width=\textwidth,height=8cm,keepaspectratio]{figures/regression-3.png}
-%       }};
-%   \end{tikzpicture}
-% \end{frame}
 
@@
     % x, y
-    \node[style=box] at (0,0) {\objw{12cm}{%
-      \includegraphics<1>[width=\textwidth,height=8cm,keepaspectratio]{figures/err-1.png}
-      }};
+    % Nodes
+    \node[style=txt] (data1) at (0,0) {$\left\{ x, y \right\}_{(x,y) \in \sD}$};
+    \node[style=txt,below=1cm of data1.center] (data2) {$\left\{ x\tp{2}, y^2 \right\}_{(x,y) \in \sD}$};
+    \node[style=txt,below=1cm of data2.center] (data3) {$\left\{ x\tp{3}, y^3 \right\}_{(x,y) \in \sD}$};
+
+    \node[style=txt] (m1) at (4,0) {$\E_h\left[\beta_h\right]$};
+    \node[style=txt,below=1cm of m1.center] (m2) {$\E_h\left[\beta_h\tp{2}\right]$};
+    \node[style=txt,below=1cm of m2.center] (m3) {$\E_h\left[\beta_h\tp{3}\right]$};
+
+    \node[below=0.5cm of m2.center] (params-pos) {};
+    \node[style=txt,right=3.0cm of params-pos] (params) {$\pi, B$};
+
+    % Arrows
+    \draw[-latex] (data1) -- (m1);
+    \draw[-latex] (data2) -- node[above,scale=0.5] {low rank} (m2);
+    \draw[-latex] (data3) -- node[above,scale=0.5] {low rank} (m3);
+
+    % Bias influence
+    \draw[-latex,dashed,gray] (m1.east) -- ++(0.1cm,0) -- ++(0,-0.50cm) -- ++(-3.00cm,0) node (bias-1) {} -- ( bias-1 |- m3) -- (m3.west);
+
+    % Mean
+    \draw[-latex] (m2.east) -- ++(0.1cm,0) node (tf-1) {} -- (tf-1 |- params-pos) -- node[above,scale=0.5] {tensor factorization} (params);
+    \draw[-latex] (m3.east) -- ++(0.1cm,0) node (tf-2) {} -- (tf-2 |- params-pos) -- (params);
+
+    % Box for regression
+    \begin{pgfonlayer}{bg}
+    \draw[fill=blue,opacity=0.1,dashed] ($(data1.north east) + (-0.1cm,0.1cm)$) rectangle ($(m3.south west) + (0.1cm,-0.1cm)$);
+    \end{pgfonlayer}
+    \node[style=txt] (reg-label) at ($(data3) !.5! (m3) + (0, -1.0cm)$) {regression};
+
+    % tensor factorization
+    \begin{pgfonlayer}{bg}
+    \draw[fill=green,opacity=0.1,dashed] ($(m2.north east) + (-0.1cm,0.1cm)$) rectangle ($(params.south west) + (0.1cm,-1.0cm)$);
+    \end{pgfonlayer}
+    \node (params-mid) at ($(params-pos) !.5! (params)$) {};
+    \node[style=txt] at (params-mid |- reg-label) {tensor factorization};
   \end{tikzpicture}
@@
 \begin{frame}
-  \frametitle{Perturbation Analysis}
+  \frametitle{Sample Complexity}
 
@@
     % x, y
-    \node[style=box] at (0,0) {\objw{12cm}{%
-      \includegraphics<1>[width=\textwidth,height=8cm,keepaspectratio]{figures/err-1.png}
-      %\includegraphics<2>[width=\textwidth,height=8cm,keepaspectratio]{figures/err-2.png}
-      \includegraphics<2>[width=\textwidth,height=8cm,keepaspectratio]{figures/err-3.png}
-      %\includegraphics<4>[width=\textwidth,height=8cm,keepaspectratio]{figures/err-4.png}
-      \includegraphics<3>[width=\textwidth,height=8cm,keepaspectratio]{figures/err-5.png}
-      }};
+    % Nodes
+    \node[style=txt] (data1) at (0,0) {$\left\{ x, y \right\}_{(x,y) \in \sD}$};
+    \node[style=txt,below=1cm of data1.center] (data2) {$\left\{ x\tp{2}, y^2 \right\}_{(x,y) \in \sD}$};
+    \node[style=txt,below=1cm of data2.center] (data3) {$\left\{ x\tp{3}, y^3 \right\}_{(x,y) \in \sD}$};
+
+    \node[style=txt] (m1) at (4,0) {$\E_h\left[\beta_h\right]$};
+    \node[style=txt,below=1cm of m1.center] (m2) {$\E_h\left[\beta_h\tp{2}\right]$};
+    \node[style=txt,below=1cm of m2.center] (m3) {$\E_h\left[\beta_h\tp{3}\right]$};
+
+    \node[below=0.5cm of m2.center] (params-pos) {};
+    \node[style=txt,right=3.0cm of params-pos] (params) {$\pi, B$};
+
+    % Arrows
+    \draw[-latex] (data1) -- (m1);
+    \draw[-latex] (data2) -- node[above,scale=0.5] {low rank} (m2);
+    \draw[-latex] (data3) -- node[above,scale=0.5] {low rank} (m3);
+
+    % Bias influence
+    \draw[-latex,dashed,gray] (m1.east) -- ++(0.1cm,0) -- ++(0,-0.50cm) -- ++(-3.00cm,0) node (bias-1) {} -- ( bias-1 |- m3) -- (m3.west);
+
+    % Mean
+    \draw[-latex] (m2.east) -- ++(0.1cm,0) node (tf-1) {} -- (tf-1 |- params-pos) -- node[above,scale=0.5] {tensor factorization} (params);
+    \draw[-latex] (m3.east) -- ++(0.1cm,0) node (tf-2) {} -- (tf-2 |- params-pos) -- (params);
+
+    % Box for regression
+    \begin{pgfonlayer}{bg}
+    \draw[fill=blue,opacity=0.1,dashed] ($(data1.north east) + (-0.1cm,0.1cm)$) rectangle ($(m3.south west) + (0.1cm,-0.1cm)$);
+    \end{pgfonlayer}
+    \node[style=txt] (reg-label) at ($(data3) !.5! (m3) + (0, -1.0cm)$) {regression};
+
+    % tensor factorization
+    \begin{pgfonlayer}{bg}
+    \draw[fill=green,opacity=0.1,dashed] ($(m2.north east) + (-0.1cm,0.1cm)$) rectangle ($(params.south west) + (0.1cm,-1.0cm)$);
+    \end{pgfonlayer}
+    \node (params-mid) at ($(params-pos) !.5! (params)$) {};
+    \node[style=txt] (tf-label) at (params-mid |- reg-label) {tensor factorization};
+
+    % Error bounds
+    \node<2->[style=txt,below=0.1em of reg-label] {$O\left( k \|x\|^{12} \|\beta\|^{6} \|\E[\epsilon^2]\|^{6} \right)$};
+    \node<3->[style=txt,below=0.1em of tf-label] {$O\left( \frac{k \pi_{\max}^2}{\sigma_k(M_2)^5} \right)$};
+
   \end{tikzpicture}
-
-  \note[item]{The algorithm is consistent, and requires $O(?)$ samples.}
-  \note[item]{The rate of convergence for the spectral experts algorithm to the true parameters breaks into two parts; the rates for learning the moments, which feeds into the rates for learning the parameters.}
-  \note[item]{For low rank regression, we have the following bound on recovery by (Tomioka2011); $$ \| \hat M_p - M_p \|_F \le \frac{32 \lambda^{(p)}_n \sqrt{k}}{\kappa(\opX_p)}, $$ where $\kappa(\opX_p)$ is the (restricted) strong convexity constant, and $\lambda^{(p)} > \|\opX^*_p(\eta)\|$.}
-  \note[item]{Because we assume our noise is bounded, it is easy to show that the error concentrates.}
-  \note[item]{In the tensor recovery case, we will need to whiten $M_3$ before applying the tensor decomposition and unwhiten it afterwards; this modifies the error bounds slightly.}
 \end{frame}
@@
     % x, y
-    \node<-4> (em) at (-3,-0.5) {%
+    \node (em) at (-3,-0.5) {%
       \includegraphics<1>[width=\textwidth,height=5cm,keepaspectratio]{figures/1-8-3.png}
-      \includegraphics<2->[width=\textwidth,height=5cm,keepaspectratio]{figures/1-8-3-em.png}
-      };
-    % x, y
-    \node<-4> at (3,-0.5) {%
+      \includegraphics<2>[width=\textwidth,height=5cm,keepaspectratio]{figures/1-8-3-em.png}
       \includegraphics<3>[width=\textwidth,height=5cm,keepaspectratio]{figures/1-8-3-spec.png}
@@
       };
-    \node<5-> at (0,-0.5) {%
+    % x, y
+    \node<2-> at (3,-0.5) {%
       \includegraphics[width=\textwidth,height=5cm,keepaspectratio]{figures/hist.png}
@@
 
-    \node[above=0.1 cm of em] {%
-    $y = \underbrace{t + t^4 + t^7}_{x} + \epsilon$
-    };
+    \node[above=0.1 cm of em] {\objw{3cm}{%
+      \begin{align*}
+        y &= \beta^T 
+            \begin{bmatrix} 
+              1 \\
+              t \\
+              t^4 \\
+              t^7
+            \end{bmatrix} + \epsilon
+      \end{align*}
+      }};
+  \end{tikzpicture}
+\end{frame}
 
-%      \begin{small}
-%  \begin{tabular}{ r r r c c c }
-%\hline
-%$b$ & $d$ & $k$ & Spectral & EM & Spectral + EM \\
-%\hline
-%  1 & 4 & 2 & 2.45 $\pm$ 3.68 & 0.28 $\pm$ 0.82 & {\bf 0.17 $\pm$ 0.57} \\
-%  2 & 5 & 2 & 1.38 $\pm$ 0.84 & {\bf 0.00 $\pm$ 0.00} & {\bf 0.00 $\pm$ 0.00} \\
-%  2 & 5 & 3 & 2.92 $\pm$ 1.71 & 0.43 $\pm$ 1.07 & {\bf 0.31 $\pm$ 1.02} \\
-%  2 & 6 & 2 & 2.33 $\pm$ 0.67 & 0.63 $\pm$ 1.29 & {\bf 0.01 $\pm$ 0.01} \\
-%\hline
-%\end{tabular}
-%      \end{small}
-%    }};
-  \end{tikzpicture}
+\begin{frame}
+  \frametitle{Experimental Insights}
 
-  \note[item]{Stable!}
-    \note[item]{Spectral Experts provides a good initialization to EM.}
-    \note[item]{We simulated the performance of spectral experts and compared it to EM.}
-    \note[item]{In a non-trivial number of cases, EM did not converge to the right parameters and got stuck in local optima.}
-    \note[item]{The parameters estimated by the spectral method were not very good even with $O(10^5)$ samples.}
-    \note[item]{However, EM initialized with these parameters did extremely well.}
-\note[item]{This finding that spectral methods should be a good initialization for EM is not surprising.}
-\note[item]{%
-  \begin{itemize}
-    \item The biggest sell for spectral methods is that they give a global guarantee on where the parameters are.
-    \item The parameters might not be at the global optima, but will hopefully lie in the potential well around it.
-    \item EM will then converge to the global optima.
-  \end{itemize}
-  }
+\begin{small}
+  \begin{tabular}{r r r c c c}
+\hline
+%$b$ & 
+$d$ & $k$ & Spectral & EM & Spectral + EM \\
+\hline
+  %1 & 
+  4 & 2 & 2.45 $\pm$ 3.68 & 0.28 $\pm$ 0.82 & {\bf 0.17 $\pm$ 0.57} \\
+  %2 & 
+  5 & 2 & 1.38 $\pm$ 0.84 & {\bf 0.00 $\pm$ 0.00} & {\bf 0.00 $\pm$ 0.00} \\
+  %2 & 
+  5 & 3 & 2.92 $\pm$ 1.71 & 0.43 $\pm$ 1.07 & {\bf 0.31 $\pm$ 1.02} \\
+  %2 & 
+  6 & 2 & 2.33 $\pm$ 0.67 & 0.63 $\pm$ 1.29 & {\bf 0.01 $\pm$ 0.01} \\
+\hline
+\end{tabular}
+      \end{small}
+
 \end{frame}
@@
       \includegraphics<3>[width=\textwidth,height=6cm,keepaspectratio]{figures/likelihood-mom-em2.png}
-        \hfill{\em expected log-likelihood}
+        \hfill{\em expected negative log-likelihood}
       }};

### 2013-06-11 doc/icml-presentation/presentation.tex
Started fixing slides

@@
 %\usetheme{NLP}
-\usepackage{beamerthemesplit}
-\usetheme{default}
+\usetheme{boxes}
 \useoutertheme{infolines}
+\setbeamertemplate{background}[grid][step=10pt]
 
@@
 
-\usepackage{algorithm,algorithmic}
+%\usepackage{algorithm,algorithmic}
 
 \usepackage{tikz}
-\usetikzlibrary{calc,positioning,shapes,shadows,arrows}
+\usetikzlibrary{calc,positioning,shapes,shadows,arrows,tikzmark}
 
@@
 \begin{frame}
-  \frametitle{Generative vs. Discriminative Latent Variable Models}
+  \frametitle{Latent Variable Models}
 
-  \begin{tikzpicture}
-    % Tasks.
-    \node[style=box] (gen) at (0,0) {\objw{6cm}{%
-      \includegraphics[width=\textwidth,height=3cm,keepaspectratio]{figures/gen.png}\\
-      \hfill Generative Models
-      }};
-    \node[scale=0.7] at (0.5,0.5) {Latent variable};
-
-    \node<2->[style=box,below=0.1cm of gen] {\obj{%
-      {\bf Generative Models}
+  \splitcolumn{%
+    \begin{itemize} 
+      \item {\bf Generative Models} \tikzmark{gen}
+        \uncover<2->{%
         \begin{itemize}
@@
         \end{itemize}
-      }};
-    \node<3->[style=box] (disc) at (6,0) {\objw{6cm}{%
-      \includegraphics[width=\textwidth,height=3cm,keepaspectratio]{figures/disc.png}\\
-      \hfill Discriminative Models
-      }};
-    \node<4->[style=box,below=0.1cm of disc] (disc-models) {\obj{%
-    {\bf Discriminative Models}
+        }
+      \item<3-> {\bf Discriminative Models} \tikzmark{disc}
+        \uncover<4->{%
         \begin{itemize}
+          \item Mixture of Experts
           \item Latent CRFs
           \item Discriminative LDA
-        \item Mixture of Experts
-        %\item Semi-Latent Topic Models (Wang and Mori, 2009)
-        %\item Discriminative Parsers (Liang et.\ al, 2006)
           \item \dots
+          \item {\em Easy to include features and tend to be more accurate.}
         \end{itemize}
-      }};
-    \node<5->[style=box,below=0.1cm of disc-models] {\obj{%
-      {\bf Easy to include features and tend to be more accurate.}
-    }};
-
+        }
+      \end{itemize}
+  }{%
+  \begin{tikzpicture}[remember picture, overlay]
+    \draw[use as bounding box] (current page.north west) rectangle (current page.south east);
+    \point{mark}{(2cm,0)};
+    \point{gen}{(pic cs:gen)};
+    \point{disc}{(pic cs:disc)};
+    \point{a}{((pic cs:gen) -| mark)};
+    \point{b}{((pic cs:disc) -| mark)};
+    %\node<1-> at (a) {\includegraphics[width=\textwidth,height=3cm,keepaspectratio]{figures/gen.png}};
+    %\node<1-> at (b) {\includegraphics[width=\textwidth,height=3cm,keepaspectratio]{figures/disc.png}};
   \end{tikzpicture}
+  }
 

### 2013-06-11 doc/icml-presentation/presentation.tex
Solved jittery and citation problems

@@
 \usepackage{listings}
+\usepackage[style=verbose,autocite=footnote,maxnames=10,babel=hyphen,hyperref=true,abbreviate=false,mcite]{biblatex}
+\addbibresource{ref.bib}
 
@@
     \draw[use as bounding box] (current page.north west) rectangle (current page.south east);
-    \point{mark}{(2cm,0)};
+    \point{mark}{(3cm,0)};
     \point{gen}{(pic cs:gen)};
     \point{disc}{(pic cs:disc)};
-    \point{a}{((pic cs:gen) -| mark)};
-    \point{b}{((pic cs:disc) -| mark)};
-    %\node<1-> at (a) {\includegraphics[width=\textwidth,height=3cm,keepaspectratio]{figures/gen.png}};
-    %\node<1-> at (b) {\includegraphics[width=\textwidth,height=3cm,keepaspectratio]{figures/disc.png}};
+    \point{a}{(gen -| mark)};
+    \point{b}{(disc -| mark)};
+    \node<1->[anchor=north east] at (a) {\includegraphics[width=\textwidth,height=2cm,keepaspectratio]{figures/gen.png}};
+    \node<3->[anchor=north east] at (b) {\includegraphics[width=\textwidth,height=2cm,keepaspectratio]{figures/disc.png}};
   \end{tikzpicture}
@@
     % x, y
-    \node[style=box] at (0,-1.5) {\objw{12cm}{%
+    \node at (0,-1.5) {%
     \includegraphics<1>[width=\textwidth,height=6cm,keepaspectratio]{figures/likelihood.png}
@@
     \includegraphics<4>[width=\textwidth,height=6cm,keepaspectratio]{figures/likelihood-mom.png}
-      \hfill{\em expected negative log-likelihood}
-      }};
+      };
   \end{tikzpicture}
@@
   \frametitle{Aside: Tensor Operations}
-  \begin{tikzpicture}
-    % x, y
-    \node<1->[style=box] (defn) at (0,0) {\objw{6cm}{%
+  \twocolumn{%
     \begin{align*}
         \action<1->{%
-        x\tp{3}_{ijk} &= x_i x_j x_k \comment{tensoring}\\
+        \tikzmark{tensoring}x\tp{3}_{ijk} &= x_i x_j x_k \comment{tensoring}\\
         }
         \action<2->{%
-        \innerp{A}{B} &= \sum_{ijk} A_{ijk} B_{ijk} \comment{inner product}\\
-        }
-        \action<3->{%
-        A(u)_{ij} &= \sum_{k} A_{ijk} u_k \comment{projection}\\
-        }
-        \action<4->{%
-        A(u,v)_{i} &= \sum_{jk} A_{ijk} u_j v_k \comment{projection} \\
+        \tikzmark{innerp}\innerp{A}{B} &= \sum_{ijk} A_{ijk} B_{ijk} \comment{inner product}\\
         }
     \end{align*}
-    }};
-    \tikzcube{black,fill=white}{(6,2,0)}{1}{1}{1};
-    %\node (pt) at (5.75,1.75,0) {$\hlmath{\times}$};
-    \uncover<3->{%
-    \tikzrect{black,fill=white}{(6,0)}{1}{1};
+  }{%
+  \begin{tikzpicture}[remember picture, overlay]
+    \draw[use as bounding box] (current page.north west) rectangle (current page.south east);
+    \point{mark}{(4cm,0)};
+    \point{tensoring}{(pic cs:tensoring)};
+    \point{innerp}{(pic cs:innerp)};
+    \point{a}{(tensoring -| mark)};
+    \point{b}{(innerp -| mark)};
+      \tikzcube{black,fill=white}{($(a) + (0,0,0)$)}{1}{1}{1};
+      \tikzcube{black,fill=white}{($(b) + (0,0,0)$)}{1}{1}{1};
+    \end{tikzpicture}
   }
-    \uncover<4->{%
-    \tikzrect{black,fill=white}{(6,-2)}{0.1}{1};
-    }
-  \end{tikzpicture}
-
 \end{frame}
@@
   \frametitle{Example: Gaussian Mixture Model}
-  \begin{tikzpicture}
-    % x, y
-    \node<1->[style=box] (moments) at (0,0) {\objw{6cm}{%
-      If $\beta_h$ are orthogonal;
+  \twocolumn{%
+    \begin{itemize}
+      \item If $\beta_h$ are orthogonal;
       \begin{align*}
@@
       \end{align*}
-    }};
-    \node<1->[style=box] (models) at (6,0) {\objw{6cm}{%
-      \includegraphics[width=0.45\textwidth,height=3cm,keepaspectratio]{figures/gen.png}
-      \includegraphics[width=0.45\textwidth,height=3cm,keepaspectratio]{figures/mog.png}
-    }};
-
-    \node<1-> at (6,-3) {\objw{6cm}{%
-    \includegraphics[width=\textwidth,height=6cm,keepaspectratio]{figures/tensor.png}
-      }};
-    \node<2->[style=box] at (0,-3) {\objw{6cm}{%
-      $\beta_h$ are eigenvectors!
+      \item<2-> $\beta_h$ are eigenvectors \footcite{AnandkumarGeHsu2012}!
         \begin{align*}
@@
         \end{align*}
-
-        \hfill{\em Anandkumar, Hsu, et.\ al (2012)}
-      }};
-%
-%    \node<2>[style=box,below=0.1cm of tensorf] {\objw{12cm}{%
-%    \includegraphics[width=\textwidth,height=6cm,keepaspectratio]{figures/tensor-decomp.png}
-%      \hfill {Tensor Power Method}
+    \end{itemize}
+  }{%
+  %\begin{tikzpicture}
+  %  % x, y
+  %  \node<1->[style=box] (models) at (6,0) {\objw{6cm}{%
+  %    \includegraphics[width=0.45\textwidth,height=3cm,keepaspectratio]{figures/gen.png}
+  %    \includegraphics[width=0.45\textwidth,height=3cm,keepaspectratio]{figures/mog.png}
   %  }};
-  \end{tikzpicture}
+  %  \node<1-> at (6,-3) {\objw{6cm}{%
+  %  \includegraphics[width=\textwidth,height=6cm,keepaspectratio]{figures/tensor.png}
+  %    }};
+  %\end{tikzpicture}
+  }
 \end{frame}

### 2013-06-11 doc/icml-presentation/presentation.tex
Fixed figures, references

@@
 \useoutertheme{infolines}
-\setbeamertemplate{background}[grid][step=10pt]
 
@@
 \usepackage{lmodern}
+\usepackage{calc}
 
@@
 \usepackage{listings}
-\usepackage[style=verbose,autocite=footnote,maxnames=10,babel=hyphen,hyperref=true,abbreviate=false,mcite]{biblatex}
+\usepackage[style=authoryear,babel=hyphen]{biblatex}
 \addbibresource{ref.bib}
+\addbibresource{pliang.bib}
 
@@
 \usepackage{tikz}
-\usetikzlibrary{calc,positioning,shapes,shadows,arrows,tikzmark}
+%\usepackage[debug]{scabslides}
+\usepackage{scabslides}
+
+\usepackage{mathtools}
 
 \input{macros}
-\input{slide-macros}
 \input{spectral-macros}
+\input{diagrams}
 
@@
   }{%
-  \begin{tikzpicture}[remember picture, overlay]
-    \draw[use as bounding box] (current page.north west) rectangle (current page.south east);
+  \begin{canvas}
     \point{mark}{(3cm,0)};
-    \point{gen}{(pic cs:gen)};
-    \point{disc}{(pic cs:disc)};
-    \point{a}{(gen -| mark)};
-    \point{b}{(disc -| mark)};
-    \node<1->[anchor=north east] at (a) {\includegraphics[width=\textwidth,height=2cm,keepaspectratio]{figures/gen.png}};
-    \node<3->[anchor=north east] at (b) {\includegraphics[width=\textwidth,height=2cm,keepaspectratio]{figures/disc.png}};
-  \end{tikzpicture}
+    \point{gen}{({pic cs:gen} -| mark)}
+    \point{disc}{({pic cs:disc} -| mark)}
+    \node<1->[anchor=north east] at ($(gen) +(0,1.0cm)$) {\includegraphics[width=\textwidth,height=2cm,keepaspectratio]{figures/gen.png}};
+    \node<3->[anchor=north east] at ($(disc) +(0,-0.5cm)$) {\includegraphics[width=\textwidth,height=2cm,keepaspectratio]{figures/disc.png}};
+  \end{canvas}
   }
@@
     % x, y
-    \node at (0,-1.5) {%
+    \node at (0,0) {%
     \includegraphics<1>[width=\textwidth,height=6cm,keepaspectratio]{figures/likelihood.png}
@@
       };
+    %\point{opt}{(0.4cm,-1.1cm)};
+    %\node[style=lbl,below=0.0cm of opt,scale=1.2] {$\hlmath{blue}{\theta^*}$};
+    % Axes
+    \draw[-latex] (-6cm,-2.25cm) -- node[right,at end]{$-\log(p_{\theta}(x))$}  (-6cm,2cm);
+    \draw[-latex] (-6cm,-2.25cm) -- node[above,at end]{$\theta$}  (5cm,-2.25cm);
   \end{tikzpicture}
@@
   \begin{itemize}
-    \item<+-> Method of Moments [Pearson, 1894]
-    \item<+-> Observable operator models
+    \item<1-> Method of Moments [Pearson, 1894]
+    \item<2-> Observable operator models
     \begin{itemize}
-      \item Subspace identification [Ljung, 1987]
-      \item Observable operator models [Jaeger, 2000]
-      \item {\em Hidden Markov models [Hsu et al, 2009]}
+      \item Control Theory [Ljung, 1987]
+      \item Observable operator models [Jaeger, 2000; Singh et al., 2004]
+      \item \alert<2>{Hidden Markov models [Hsu et al, 2009]}
       \item Low-treewidth graphs [Parikh et al., 2012]
@@
     \end{itemize}
-     \item<+-> Parameter Estimation
+     \item<3-> Parameter Estimation
   \begin{itemize}
     \item Mixture of Gaussians [Kalai/Moitra/Valiant, 2010]
-    \item {\bf Mixture models, HMMs [Anandkumar/Hsu/Kakade, 2012]}
+    \item \alert{Mixture models, HMMs [Anandkumar/Hsu/Kakade, 2012]}
     \item Latent Dirichlet Allocation [Anandkumar/Hsu/Kakade, 2012]
@@
   \frametitle{Aside: Tensor Operations}
-  \twocolumn{%
+  \splitcolumn{%
+    \begin{itemize}
+      \item \tikzmark{tensoring} Tensoring
+        $$x\tp{3}_{ijk} = x_i x_j x_k$$
+      \item \tikzmark{innerp}Inner product
         \begin{align*}
-        \action<1->{%
-        \tikzmark{tensoring}x\tp{3}_{ijk} &= x_i x_j x_k \comment{tensoring}\\
-        }
-        \action<2->{%
-        \tikzmark{innerp}\innerp{A}{B} &= \sum_{ijk} A_{ijk} B_{ijk} \comment{inner product}\\
-        }
-    \end{align*}
-  }{%
-  \begin{tikzpicture}[remember picture, overlay]
-    \draw[use as bounding box] (current page.north west) rectangle (current page.south east);
-    \point{mark}{(4cm,0)};
-    \point{tensoring}{(pic cs:tensoring)};
-    \point{innerp}{(pic cs:innerp)};
-    \point{a}{(tensoring -| mark)};
-    \point{b}{(innerp -| mark)};
-      \tikzcube{black,fill=white}{($(a) + (0,0,0)$)}{1}{1}{1};
-      \tikzcube{black,fill=white}{($(b) + (0,0,0)$)}{1}{1}{1};
-    \end{tikzpicture}
-  }
-\end{frame}
-
-\begin{frame}
-  \frametitle{Example: Gaussian Mixture Model}
-  \begin{tikzpicture}
-    % x, y
-    \node<1->[style=box] (defn) at (0,0) {\objw{6cm}{%
-    \begin{itemize} 
-      \item Generative process:
-      \begin{align*}
-        h &\sim \Mult(\pi) \\
-        x &\sim \normal{\beta_h}{\sigma^2}.
+          \innerp{A}{B} &= \sum_{ijk} A_{ijk} B_{ijk} \\
+          \pause
+            &= \innerp{\vvec{A}}{\vvec{B}} 
         \end{align*}
     \end{itemize}
-    }};
-    \node<1->[style=box] (models) at (6,0) {\objw{6cm}{%
-      \includegraphics[width=0.45\textwidth,height=3cm,keepaspectratio]{figures/gen.png}
-      \includegraphics[width=0.45\textwidth,height=3cm,keepaspectratio]{figures/mog.png}
-    }};
+  }{%
+  }
+  \begin{canvas}
+    \point{mark}{(6.8cm,0)};
+    % Tensoring
+    \tikzcube{tensoring}{black,fill=white}{($({pic cs:tensoring} -| mark) + (0,0,0)$)}{1}{1}{1};
+    \node at ($(tensoring) + (1.0cm,-0.3cm)$) {$=$};
+    \tikzcube{v1}{black,fill=white}{($(tensoring) + (2.5cm,-0.3cm)$)}{1}{0.3}{0.3};
+    \node (lbl1) at ($(v1) + (0.40cm,-0.1cm)$) {$\times$};
+    \tikzcube{v2}{black,fill=white}{($(v1) + (1cm,0.3cm)$)}{0.3}{1}{0.3};
+    \node at ($(lbl1) + (1cm,0.0cm)$) {$\times$};
+    \tikzcube{v3}{black,fill=white}{($(v2) + (1cm,-0.3cm)$)}{0.3}{0.3}{1};
 
-    \node<2->[style=box,below=0.1cm of defn] {\objw{6cm}{%
-    \begin{itemize} 
-      \item Moments:
-    \begin{align*}
-        \action<2->{%
-        \E[x|h] &= \beta_h \\
+    % Inner product
+    \uncover<2->{%
+    \point{ptA}{($({pic cs:innerp} -| mark) + (0,0.2cm)$)}
+    \point{ptB}{($(ptA) + (2cm,0)$)}
+    \tikzcube{innerpA}{black,fill=yellow}{($(ptA) + (0,0,0)$)}{1}{1}{1};
+    \tikzcube{innerpB}{black,fill=blue,opacity=0.7}{($(ptB) + (0,0,0)$)}{1}{1}{1};
+    \node[scale=2.0] (op1a) at ($(ptA) + (-1.3cm,-0.3cm)$) {$\langle$};
+    \node (op1b) at ($(ptA) + (0.7cm,-0.3cm)$) {$,$};
+    \node[scale=2.0] (op1c) at ($(ptB) + (0.8cm,-0.3cm)$) {$\rangle$};
+    \node[scale=2.0] (op2) at ($(op1c) + (1.5cm,0.0cm)$) {$= \hlmath{green}{0.5}$};
     }
-        \action<3->{%
-        \E[x] &= \sum_h \pi_h \beta_h \\
+
+    % Inner product 2
+    \uncover<3->{%
+    \tikzrect{innerpAv}{black,fill=yellow}{($(ptA) + (0,-1.5cm)$)}{0.3}{1.5};
+    \tikzrect{innerpBv}{black,fill=blue,opacity=0.7}{($(ptB) + (0,-1.5cm)$)}{0.3}{1.5};
+    \node[scale=2.0] (op1a) at ($(innerpAv) + (-1.3cm,-0.6cm)$) {$\langle$};
+    \node (op1b) at ($(innerpAv) + (0.7cm,-0.6cm)$) {$,$};
+    \node[scale=2.0] (op1c) at ($(innerpBv) + (0.8cm,-0.6cm)$) {$\rangle$};
+    \node[scale=2.0] (op2) at ($(op1c) + (1.5cm,0.0cm)$) {$= \hlmath{green}{0.5}$};
     }
-        \action<4->{%
-        \E[x\tp{2}] &= \sum_h \pi_h (\beta_h \beta_h^T) \\
-                    &= \sum_h \pi_h \beta_h{\tp{2}} \\
-                    }
-        \action<5->{%
-        \E[x\tp{3}] &= \sum_h \pi_h \beta_h\tp{3}.
-        }
-      \end{align*}
-    \end{itemize}
-    }};
-    %\tikzrect{1}{1};
-    \uncover<4->{%
-    \tikzrect{black,fill=white}{(6,-2)}{1}{1};
-    \node at (7,-2.5) {$\E[x\tp{2}]$};
-    \node at (4.5,-2.5) {$d$};
-    }
-    \uncover<5->{%
-    \tikzcube{black,fill=white}{(6,-4,0)}{1}{1}{1};
-    \node at (7,-4.5,0) {$\E[x\tp{3}]$};
-    \node at (4.5,-4.5) {$d$};
-    \node at (4.5,-4.5) {$d$};
-    }
-  \end{tikzpicture}
+  \end{canvas}
 
@@
   \frametitle{Example: Gaussian Mixture Model}
-  \twocolumn{%
+
+  \splitcolumn{%
     \begin{itemize}
-      \item If $\beta_h$ are orthogonal;
+      \item \tikzmark{gen} Generative process:
       \begin{align*}
-        \E[x\tp{3}] &= \sum_{h=1}^k \pi_h \beta_h\tp{3}. 
+        h &\sim \Mult(\pi) \\
+        x &\sim \normal{\beta_h}{\sigma^2}.
       \end{align*}
-      \item<2-> $\beta_h$ are eigenvectors \footcite{AnandkumarGeHsu2012}!
+      \item Moments:
+      \begin{align*}
+          \uncover<2->{%
+          \E[x|h] &= \beta_h \\
+          }
+          \uncover<3->{%
+          \E[x] &= \sum_h \pi_h \beta_h \\
+          }
+          \uncover<4->{%
+          \tikzmark{m2}
+          \E[x\tp{2}] &= \sum_h \pi_h (\beta_h \beta_h^T) \\
+                      &= \sum_h \pi_h \beta_h{\tp{2}} \\
+                      }
+          \uncover<5->{%
+          \tikzmark{m3}
+          \E[x\tp{3}] &= \sum_h \pi_h \beta_h\tp{3}.
+          }
+        \end{align*}
+    \end{itemize}
+  }{%
+  }
+  \begin{canvas}
+    % The model
+    \point{mark}{(6.5cm,0)};
+    \point{start}{({pic cs:gen} -| mark)};
+    \node[anchor=west] (diag1) at ($(start)$) {%
+      \includegraphics[width=0.45\textwidth,height=2cm,keepaspectratio]{figures/gen.png}
+    };
+    \node[anchor=west] (diag) at ($(start) + (1cm,0.0cm)$) {%
+      \includegraphics[width=0.45\textwidth,height=3cm,keepaspectratio]{figures/mog.png}
+    };
+
+    \uncover<4->{%
+      \point{m2}{({pic cs:m2} -| mark)};
+      \tikzrect{M2}{black,fill=white}{($(m2) + (1cm,1cm)$)}{1}{1};
+      \node[anchor=west] at ($(m2) + (1.1cm, 0.5cm)$) {$\E[x\tp{2}]$};
+      \node[anchor=east] at ($(m2) + (-0.1cm, 0.5cm)$) {$d$};
+      \node[anchor=south] at ($(m2) + (0.5cm, 1.0cm)$) {$d$};
+    }
+    \uncover<5->{%
+      \point{m3}{($({pic cs:m3} -| mark) + (1cm,1cm)$)};
+      \tikzcube{M3}{black,fill=white}{($(m3) + (0,0,0)$)}{1}{1}{1};
+      \node[anchor=west] at ($(m3) + (0.5cm, -0.5cm)$) {$\E[x\tp{3}]$};
+      \node[anchor=east] at ($(m3) + (-1.1cm, -0.5cm)$) {$d$};
+    }
+  \end{canvas}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{Solution: Tensor Factorization}
+  \splitcolumn{%
+    \begin{itemize}
+      \item \tikzmark{gen}$\E[x\tp{3}] = \sum_{h=1}^k \pi_h \beta_h\tp{3}$.
+      \item<2-> If $\beta_h$ are orthogonal, the are eigenvectors! \footcite{AnandkumarGeHsu2012}
         \begin{align*}
@@
         \end{align*}
+      \item<3-> In general, whiten $\E[x\tp{3}]$ first.
     \end{itemize}
   }{%
-  %\begin{tikzpicture}
-  %  % x, y
-  %  \node<1->[style=box] (models) at (6,0) {\objw{6cm}{%
-  %    \includegraphics[width=0.45\textwidth,height=3cm,keepaspectratio]{figures/gen.png}
+  }
+  \begin{canvas}
+    \point{mark}{(6.5cm,0)};
+    \point{start}{({pic cs:gen} -| mark)};
+
+    \node[anchor=west] (diag1) at ($(start)$) {%
+      \includegraphics[width=0.45\textwidth,height=2cm,keepaspectratio]{figures/gen.png}
+    };
+    \node[anchor=west] (diag) at ($(start) + (1cm,0.0cm)$) {%
+      \includegraphics[width=0.45\textwidth,height=3cm,keepaspectratio]{figures/mog.png}
+    };
+    %\drawgen{(start)};
+    %\node[anchor=west] (diag) at ($(start) + (1cm,-0.5cm)$) {%
     %  \includegraphics[width=0.45\textwidth,height=3cm,keepaspectratio]{figures/mog.png}
-  %  }};
-  %  \node<1-> at (6,-3) {\objw{6cm}{%
-  %  \includegraphics[width=\textwidth,height=6cm,keepaspectratio]{figures/tensor.png}
-  %    }};
-  %\end{tikzpicture}
-  }
+    %};
+    % x, y
+    %\node[anchor=north west] at ($(start) - (1cm,2cm)$) {%
+    %    \includegraphics[width=6cm,height=2cm,keepaspectratio]{figures/tensor.png}
+    %};
+  \end{canvas}
 \end{frame}
@@
 \begin{frame}
-  \frametitle{Generative vs. Discriminative Latent Variable Models}
+  \frametitle{}
 
-  \begin{tikzpicture}
+  \begin{canvas}
     % Tasks.
-    \node[style=box] (gen) at (0,0) {\objw{6cm}{%
-      \includegraphics[width=\textwidth,height=3cm,keepaspectratio]{figures/gen.png}\\
-      \hfill Generative Models
-      }};
-    \node[scale=0.7] at (0.5,0.5) {Latent variable};
+    \node[anchor=west] (gen) at (1cm,0) {%
+      \includegraphics[width=\textwidth,height=2.5cm,keepaspectratio]{figures/gen.png}
+     };
+     \node[below=0.1cm of gen.south] {Generative Models};
 
-    \node[style=box] (disc) at (6,0) {\objw{6cm}{%
-      \includegraphics[width=\textwidth,height=3cm,keepaspectratio]{figures/disc.png}\\
-      \hfill Discriminative Models
-      }};
-  \end{tikzpicture}
+    \draw[scale=0.8,fill=green,opacity=0.4,dashed] (6.5cm,2cm) rectangle (12.5cm,-3cm);
+    \node[style=box,anchor=west] (disc) at (6cm,0) {%
+      \includegraphics[width=\textwidth,height=3cm,keepaspectratio]{figures/disc.png}
+     };
+     \node[below=0.1cm of disc.south] {Discriminative Models};
+     % Highlight
+  \end{canvas}
 
@@
 
-  \begin{tikzpicture}
-    % x, y
-    \node[style=box] at (0,0) {\objw{4cm}{%
-    \includegraphics[width=\textwidth,height=4cm,keepaspectratio]{figures/disc.png}
+  \splitcolumn{%
+    \tikzmark{model}
+    \includegraphics[width=3cm,height=4cm,keepaspectratio]{figures/disc.png}
     \begin{itemize}
@@
     \end{itemize}
-      }};
-
-    \node[style=box] at (6,0) {\obj{%
-    \includegraphics<1>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-2.png}
-    \includegraphics<2>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-3a.png}
-    \includegraphics<3>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-3b.png}
-    \includegraphics<4>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-4a.png}
-    \includegraphics<5>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-4b.png}
-    %\includegraphics<6>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-4c.png}
-    \includegraphics<6>[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-5.png}
-      }};
-  \end{tikzpicture}
+  }{%
+  }
+    \begin{canvas}
+      % x, y
+      \point{mark}{(5.5cm,0)}
+      \point{model}{({pic cs:model} -| mark)}
+      \node[anchor=west] at ($(model) + (0,0.5cm)$) {%
+      \includegraphics<1>[width=5cm,height=6cm,keepaspectratio]{figures/mlr-data-2.png}
+      \includegraphics<2>[width=5cm,height=6cm,keepaspectratio]{figures/mlr-data-3a.png}
+      \includegraphics<3>[width=5cm,height=6cm,keepaspectratio]{figures/mlr-data-3b.png}
+      \includegraphics<4>[width=5cm,height=6cm,keepaspectratio]{figures/mlr-data-4a.png}
+      \includegraphics<5>[width=5cm,height=6cm,keepaspectratio]{figures/mlr-data-4b.png}
+      %\includegraphics<6>[width54cm,height=6cm,keepaspectratio]{figures/mlr-data-4c.png}
+      \includegraphics<6>[width=5cm,height=6cm,keepaspectratio]{figures/mlr-data-5.png}
+        };
+    \end{canvas}
 \end{frame}
@@
 
-  \begin{tikzpicture}
-    \node[style=box] (data) at (0,0) {\obj{%
-    \includegraphics[width=\textwidth,height=6cm,keepaspectratio]{figures/mlr-data-6.png}
-      \\{\em Data}
-      }};
+  \begin{canvas}
+    \node[anchor=west] (data) at (0,0) {%
+    \includegraphics[width=4cm,height=6cm,keepaspectratio]{figures/mlr-data.png}
+    };
 
     % x, y
-    \node[style=box] (params) at (7,0) {\objw{4cm}{%
-    \includegraphics[width=\textwidth,height=4cm,keepaspectratio]{figures/disc.png}
-    \\
-    {\bf Parameters: $\pi, B$}
-    }};
+    \node[anchor=west,scale=2.0] (params) at (7cm,0) {%
+      $\begin{bmatrix} \pi \\ B \end{bmatrix}$
+      };
+    \node[anchor=west,right=0.1cm of params] {%
+    \includegraphics[width=4cm,height=2cm,keepaspectratio]{figures/disc.png}
+    };
 
-    \draw[->,dashed] (data) -- node[above]{?} (params);
+    \draw[-latex] (data) -- node[above]{?} (params);
 
-  \end{tikzpicture}
+  \end{canvas}
 \end{frame}
@@
 
-\begin{frame}
+\begin{frame}[c]
   \frametitle{Method of Moments for the Mixture of Linear Regressions.}
-
-  \begin{tikzpicture}
-    % x, y
-    \node[style=box] at (0,0) {\objw{12cm}{%
   \begin{align*}
-          \action<+->{
-          y &= \underbrace{\beta_h^T}_{\textrm{random}} x + \epsilon \\
+    y &= \alt<1>{\innerp{\beta_h}{x}}{\underbrace{\beta_h^T}_{\textrm{\color{DarkGreen}random}} x} + \epsilon \\
+    \action<3->{%
+    &= \sum_h \pi_h \innerp{\beta_h}{x} + \innerp{\left(\beta_h - \sum_h \pi_h \beta_h\right)}{x} + \epsilon \\
     }
-          \action<+->{
-            &= \underbrace{\E[\beta_h]^T x}_{\textrm{linear measurement}} + \underbrace{(\beta_h - \E[\beta_h])^T x + \epsilon}_{\textrm{noise}} \\
+    \action<4-6>{%
+    &= \alt<4>{\E[\beta_h]^T x}{\underbrace{\E[\beta_h]^T x}_{\textrm{\color{blue}linear measurement}}} 
+    + \alt<4-5>{\innerp{(\beta_h - \E[\beta_h])}{x} + \epsilon}{\underbrace{\innerp{(\beta_h - \E[\beta_h])}{x} + \epsilon}_{\textrm{\color{red} noise}}} \\
     }
   \end{align*}
-      }};
-  \end{tikzpicture}
-
 \end{frame}
@@
 
-  \begin{tikzpicture}
-    % x, y
-    \node[style=box] (measurements) at (0,0) {\objw{12cm}{%
   \begin{align*}
-          \action<+->{%
-          y &= \underbrace{\innerp{\E[\beta_h]^T}{x}}_{\textrm{linear measurement}} &\quad & + \underbrace{(\beta_h - \E[\beta_h])^T x + \epsilon}_{\textrm{noise}} \\
+    \action<1->{%
+    y &= \overbrace{\innerp{\E[\beta_h]}{x}}^{\textrm{\color{blue} linear measurement}} & & + \overbrace{(\beta_h - \E[\beta_h])^T x + \epsilon}^{\textrm{\color{red} noise}} \\
     }
-          \action<+->{%
+    \action<2->{%
       y^2 
-            &= \innerp{\E[\beta_h\tp{2}]}{x\tp{2}} &+ \textrm{bias} &+ \textrm{noise} \\
+      &= \innerp{\beta_h}{x}^2 &+ \epsilon^2 &+ 2 \epsilon \innerp{\beta_h}{x}  \\
     }
-          \action<+->{%
-          y^3 &= \innerp{\E[\beta_h\tp{3}]}{x\tp{3}} &+ \textrm{bias} &+ \textrm{noise} \\
+    \action<3->{%
+      &= \color{blue}\innerp{\E[\beta_h\tp{2}]}{x\tp{2}} &+ \color{DarkGreen} \textrm{bias} &+ \color{red} \textrm{noise} \\
+    }
+    \action<4->{%
+    y^3 &= \color{blue} \innerp{\E[\beta_h\tp{3}]}{x\tp{3}} &+ \color{DarkGreen} \textrm{bias} &+ \color{red} \textrm{noise} \\
     }
@@
 
-        \uncover<4->{%
-        We can use regression to recover $\E[\beta_h\tp{p}]$ from linear measurements.\\
+  %\begin{tikzpicture}
+  %  % x, y
+  %  \node<5->[style=box] (tensor) {\objw{12cm}{%
+  %    \includegraphics[width=\textwidth,height=3cm,keepaspectratio]{figures/tensor.png}
+  %    }};
+  %\end{tikzpicture}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Method of Moments for the Mixture of Linear Regressions.}
+  \centering
+
+  \begin{itemize}
+  {\Large
+    \item 
+    $\E[\beta_h\tp{3}] = \sum_{h=1}^k \pi_h \beta_h\tp{3}$
   }
-      }};
-
-    \node<5->[style=box,below=0.1cm of measurements.south] (tensor) {\objw{12cm}{%
-      \includegraphics[width=\textwidth,height=3cm,keepaspectratio]{figures/tensor.png}
-      }};
-
-  \end{tikzpicture}
+    \item<2-> Apply tensor factorization!
+  \end{itemize}
 \end{frame}
@@
 
+  \begin{canvas}
+    \input{spectral-experts.tikz}
+    \node[style=txt] at (reg-label) {\alert<1>{regression}};
+    \node[style=txt] at (tf-label) {\alert<2>{tensor factorization}};
+  \end{canvas}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Exploiting Low-rank Structure.}
+
+    \begin{align*}
+      \hat\E[\beta_h\tp{2}] &= \arg\min_{M} \left(y^2 - \innerp{M}{x\tp{2}}\right)^2_{(x,y)\in\mathcal{D}} 
+        \alt<1>{+ \|M\|_{2}}
+          {+ \hlmath{blue}{\|M\|_{*}}} \\
+      \action<2->{%
+      \|M\|_{*} &= \sum_i \sigma_i(M).
+      }
+    \end{align*}
+
   \begin{tikzpicture}
     % x, y
-    % Nodes
-    \node[style=txt] (data1) at (0,0) {$\left\{ x, y \right\}_{(x,y) \in \sD}$};
-    \node[style=txt,below=1cm of data1.center] (data2) {$\left\{ x\tp{2}, y^2 \right\}_{(x,y) \in \sD}$};
-    \node[style=txt,below=1cm of data2.center] (data3) {$\left\{ x\tp{3}, y^3 \right\}_{(x,y) \in \sD}$};
-
-    \node[style=txt] (m1) at (4,0) {$\E_h\left[\beta_h\right]$};
-    \node[style=txt,below=1cm of m1.center] (m2) {$\E_h\left[\beta_h\tp{2}\right]$};
-    \node[style=txt,below=1cm of m2.center] (m3) {$\E_h\left[\beta_h\tp{3}\right]$};
-
-    \node[below=0.5cm of m2.center] (params-pos) {};
-    \node[style=txt,right=3.0cm of params-pos] (params) {$\pi, B$};
-
-    % Arrows
-    \draw[-latex] (data1) -- (m1);
-    \draw[-latex] (data2) -- (m2);
-    \draw[-latex] (data3) -- (m3);
-
-    % Bias influence
-    \draw[-latex,dashed,gray] (m1.east) -- ++(0.1cm,0) -- ++(0,-0.50cm) -- ++(-3.00cm,0) node (bias-1) {} -- ( bias-1 |- m3) -- (m3.west);
-
-    % Mean
-    \draw[-latex] (m2.east) -- ++(0.1cm,0) node (tf-1) {} -- (tf-1 |- params-pos) -- node[above,scale=0.5] {tensor factorization} (params);
-    \draw[-latex] (m3.east) -- ++(0.1cm,0) node (tf-2) {} -- (tf-2 |- params-pos) -- (params);
-
-    % Box for regression
-    \begin{pgfonlayer}{bg}
-    \draw[fill=blue,opacity=0.1,dashed] ($(data1.north east) + (-0.1cm,0.1cm)$) rectangle ($(m3.south west) + (0.1cm,-0.1cm)$);
-    \end{pgfonlayer}
-    \node[style=txt] (reg-label) at ($(data3) !.5! (m3) + (0, -1.0cm)$) {regression};
-
-    % tensor factorization
-    \begin{pgfonlayer}{bg}
-    \draw[fill=green,opacity=0.1,dashed] ($(m2.north east) + (-0.1cm,0.1cm)$) rectangle ($(params.south west) + (0.1cm,-1.0cm)$);
-    \end{pgfonlayer}
-    \node (params-mid) at ($(params-pos) !.5! (params)$) {};
-    \node[style=txt] at (params-mid |- reg-label) {tensor factorization};
+    \node (measurements) {%
+      \includegraphics[width=\textwidth,height=6cm,keepaspectratio]{figures/tensor.png}
+      };
   \end{tikzpicture}
@@
 \begin{frame}
-  \frametitle{Method of Moments for the Mixture of Linear Regressions.}
+  \frametitle{Spectral Experts}
 
-  \begin{tikzpicture}
-    % x, y
-    \node[style=box] (reg) at (0,0) {\objw{12cm}{%
-      \begin{align*}
-        \hat\E[\beta_h\tp{2}] &= \arg\min_{M} \left(y^2 - \innerp{M}{x\tp{2}}\right)^2_{(x,y)\in\mathcal{D}} \action<3->{+ \hlmath{\|M\|_{*}}} \\
-        \action<3->{%
-        \|M\|_{*} &= \sum_i \sigma_i(M).
-        }
-      \end{align*}
-      }};
-    \node<2->[style=box,below=0.1 of reg] (measurements) {\objw{12cm}{%
-      \includegraphics[width=\textwidth,height=6cm,keepaspectratio]{figures/tensor.png}
-      }};
-  \end{tikzpicture}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Overview: Spectral Experts}
-
-  \begin{tikzpicture}
-    % x, y
-    % Nodes
-    \node[style=txt] (data1) at (0,0) {$\left\{ x, y \right\}_{(x,y) \in \sD}$};
-    \node[style=txt,below=1cm of data1.center] (data2) {$\left\{ x\tp{2}, y^2 \right\}_{(x,y) \in \sD}$};
-    \node[style=txt,below=1cm of data2.center] (data3) {$\left\{ x\tp{3}, y^3 \right\}_{(x,y) \in \sD}$};
-
-    \node[style=txt] (m1) at (4,0) {$\E_h\left[\beta_h\right]$};
-    \node[style=txt,below=1cm of m1.center] (m2) {$\E_h\left[\beta_h\tp{2}\right]$};
-    \node[style=txt,below=1cm of m2.center] (m3) {$\E_h\left[\beta_h\tp{3}\right]$};
-
-    \node[below=0.5cm of m2.center] (params-pos) {};
-    \node[style=txt,right=3.0cm of params-pos] (params) {$\pi, B$};
-
-    % Arrows
-    \draw[-latex] (data1) -- (m1);
-    \draw[-latex] (data2) -- node[above,scale=0.5] {low rank} (m2);
-    \draw[-latex] (data3) -- node[above,scale=0.5] {low rank} (m3);
-
-    % Bias influence
-    \draw[-latex,dashed,gray] (m1.east) -- ++(0.1cm,0) -- ++(0,-0.50cm) -- ++(-3.00cm,0) node (bias-1) {} -- ( bias-1 |- m3) -- (m3.west);
-
-    % Mean
-    \draw[-latex] (m2.east) -- ++(0.1cm,0) node (tf-1) {} -- (tf-1 |- params-pos) -- node[above,scale=0.5] {tensor factorization} (params);
-    \draw[-latex] (m3.east) -- ++(0.1cm,0) node (tf-2) {} -- (tf-2 |- params-pos) -- (params);
-
-    % Box for regression
-    \begin{pgfonlayer}{bg}
-    \draw[fill=blue,opacity=0.1,dashed] ($(data1.north east) + (-0.1cm,0.1cm)$) rectangle ($(m3.south west) + (0.1cm,-0.1cm)$);
-    \end{pgfonlayer}
-    \node[style=txt] (reg-label) at ($(data3) !.5! (m3) + (0, -1.0cm)$) {regression};
-
-    % tensor factorization
-    \begin{pgfonlayer}{bg}
-    \draw[fill=green,opacity=0.1,dashed] ($(m2.north east) + (-0.1cm,0.1cm)$) rectangle ($(params.south west) + (0.1cm,-1.0cm)$);
-    \end{pgfonlayer}
-    \node (params-mid) at ($(params-pos) !.5! (params)$) {};
-    \node[style=txt] at (params-mid |- reg-label) {tensor factorization};
-  \end{tikzpicture}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Sample Complexity}
-
-  \begin{tikzpicture}
-    % x, y
-    % Nodes
-    \node[style=txt] (data1) at (0,0) {$\left\{ x, y \right\}_{(x,y) \in \sD}$};
-    \node[style=txt,below=1cm of data1.center] (data2) {$\left\{ x\tp{2}, y^2 \right\}_{(x,y) \in \sD}$};
-    \node[style=txt,below=1cm of data2.center] (data3) {$\left\{ x\tp{3}, y^3 \right\}_{(x,y) \in \sD}$};
-
-    \node[style=txt] (m1) at (4,0) {$\E_h\left[\beta_h\right]$};
-    \node[style=txt,below=1cm of m1.center] (m2) {$\E_h\left[\beta_h\tp{2}\right]$};
-    \node[style=txt,below=1cm of m2.center] (m3) {$\E_h\left[\beta_h\tp{3}\right]$};
-
-    \node[below=0.5cm of m2.center] (params-pos) {};
-    \node[style=txt,right=3.0cm of params-pos] (params) {$\pi, B$};
-
-    % Arrows
-    \draw[-latex] (data1) -- (m1);
-    \draw[-latex] (data2) -- node[above,scale=0.5] {low rank} (m2);
-    \draw[-latex] (data3) -- node[above,scale=0.5] {low rank} (m3);
-
-    % Bias influence
-    \draw[-latex,dashed,gray] (m1.east) -- ++(0.1cm,0) -- ++(0,-0.50cm) -- ++(-3.00cm,0) node (bias-1) {} -- ( bias-1 |- m3) -- (m3.west);
-
-    % Mean
-    \draw[-latex] (m2.east) -- ++(0.1cm,0) node (tf-1) {} -- (tf-1 |- params-pos) -- node[above,scale=0.5] {tensor factorization} (params);
-    \draw[-latex] (m3.east) -- ++(0.1cm,0) node (tf-2) {} -- (tf-2 |- params-pos) -- (params);
-
-    % Box for regression
-    \begin{pgfonlayer}{bg}
-    \draw[fill=blue,opacity=0.1,dashed] ($(data1.north east) + (-0.1cm,0.1cm)$) rectangle ($(m3.south west) + (0.1cm,-0.1cm)$);
-    \end{pgfonlayer}
-    \node[style=txt] (reg-label) at ($(data3) !.5! (m3) + (0, -1.0cm)$) {regression};
-
-    % tensor factorization
-    \begin{pgfonlayer}{bg}
-    \draw[fill=green,opacity=0.1,dashed] ($(m2.north east) + (-0.1cm,0.1cm)$) rectangle ($(params.south west) + (0.1cm,-1.0cm)$);
-    \end{pgfonlayer}
-    \node (params-mid) at ($(params-pos) !.5! (params)$) {};
-    \node[style=txt] (tf-label) at (params-mid |- reg-label) {tensor factorization};
+  \begin{canvas}
+    \input{spectral-experts.tikz}
+    \node[style=txt] at (reg-label) {{\color{blue}\bf low-rank} regression};
+    \node[style=txt] at (tf-label) {tensor factorization};
 
     % Error bounds
-    \node<2->[style=txt,below=0.1em of reg-label] {$O\left( k \|x\|^{12} \|\beta\|^{6} \|\E[\epsilon^2]\|^{6} \right)$};
-    \node<3->[style=txt,below=0.1em of tf-label] {$O\left( \frac{k \pi_{\max}^2}{\sigma_k(M_2)^5} \right)$};
-
-  \end{tikzpicture}
+    \node<2->[style=txt,below=0.1em of reg-label]{%
+      $O\left( k\ 
+      \hlmath{blue}{\|x\|^{12}}\
+      \hlmath{DarkGreen}{\|\beta\|^{6}}\
+      \hlmath{red}{\|\E[\epsilon^2]\|^{6}}
+      \right)$};
+    \node<3->[style=txt,below=0.1em of tf-label] {%
+    $O\left( \frac{k \pi_{\max}^2}{\sigma_k(M_2)^5} \right)$};
+  \end{canvas}
 \end{frame}
@@
 
-  \begin{tikzpicture}
+  \begin{canvas}
     % x, y
-    \node (em) at (-3,-0.5) {%
+    \node[anchor=west] (em) at (-0.5cm,-1.5cm) {%
       \includegraphics<1>[width=\textwidth,height=5cm,keepaspectratio]{figures/1-8-3.png}
@@
     % x, y
-    \node<2-> at (3,-0.5) {%
+    \node[anchor=west, right=6cm of em] at (em.west) {%
       \includegraphics[width=\textwidth,height=5cm,keepaspectratio]{figures/hist.png}
@@
 
-    \node[above=0.1 cm of em] {\objw{3cm}{%
-      \begin{align*}
-        y &= \beta^T 
+    \node[above=0.1 cm of em] {%
+        $y = \beta^T 
             \begin{bmatrix} 
@@
               t^7
-            \end{bmatrix} + \epsilon
-      \end{align*}
-      }};
-  \end{tikzpicture}
+            \end{bmatrix} + \epsilon$
+      };
+  \end{canvas}
 \end{frame}
@@
   \begin{tikzpicture}
-    \node[style=box] at (0,-1.5) {\objw{12cm}{%
+    \node at (0,0) {%
       \includegraphics<1>[width=\textwidth,height=6cm,keepaspectratio]{figures/likelihood-mom.png}
@@
       \includegraphics<3>[width=\textwidth,height=6cm,keepaspectratio]{figures/likelihood-mom-em2.png}
-        \hfill{\em expected negative log-likelihood}
-      }};
+    };
+    \draw[-latex] (-6cm,-2.25cm) -- node[right,at end]{$-\log(p_{\theta}(x))$}  (-6cm,2cm);
+    \draw[-latex] (-6cm,-2.25cm) -- node[above,at end]{$\theta$}  (5cm,-2.25cm);
+
+    \node<2-> at (1.8cm,0.6cm) {$\hlmath{DarkGreen}{\theta_{spec}}$};
+    \node<3-> at (0.3cm,-0.8cm) {$\hlmath{blue}{\theta^{*}}$};
   \end{tikzpicture}
@@
 % TikZ
-%\begin{tikzpicture}
+%\begin{canvas}
 %  \node (img1) at (0,0) {\includegraphics[height=0.6\textheight,width=0.4\linewidth,keepaspectratio]{}};
-%\end{tikzpicture}
+%\end{canvas}
 

### 2013-06-12 doc/icml-presentation/presentation.tex
Added scabby, robustalt and fixed mat jumping
Added tensor factorization and prob models

@@
 \usepackage{tikz}
-%\usepackage[debug]{scabslides}
-\usepackage{scabslides}
+\usepackage[debug]{scabby}
+%\usepackage{scabby}
+
+\usepackage[customcolors]{hf-tikz}
 
@@
     \begin{itemize} 
-      \item {\bf Generative Models} \tikzmark{gen}
+      \item  \tikzmark{gen} {\bf Generative Models}
         \uncover<2->{%
@@
         }
-      \item<3-> {\bf Discriminative Models} \tikzmark{disc}
+      \item<3-> \tikzmark{disc} {\bf Discriminative Models} 
         \uncover<4->{%
@@
           \item \dots
-          \item {\em Easy to include features and tend to be more accurate.}
         \end{itemize}
         }
+      \item<4-> {\em Easy to include features and tend to be more accurate.}
       \end{itemize}
@@
     \point{disc}{({pic cs:disc} -| mark)}
-    \node<1->[anchor=north east] at ($(gen) +(0,1.0cm)$) {\includegraphics[width=\textwidth,height=2cm,keepaspectratio]{figures/gen.png}};
-    \node<3->[anchor=north east] at ($(disc) +(0,-0.5cm)$) {\includegraphics[width=\textwidth,height=2cm,keepaspectratio]{figures/disc.png}};
+    \drawgen<1->{($(gen) + (0,1.5cm)$)};
+    \drawdisc<3->{($(disc) + (0,0.0cm)$)};
   \end{canvas}
@@
     \item<3-> Local methods (EM, gradient descent, etc.) are tractable but inconsistent.
-    \item<4-> Can we build {\bf efficient yet consistent estimators} that get better with more data?
+    \item<4-> Can we build an {\bf efficient and consistent estimator}?
   \end{itemize}
@@
         $$x\tp{3}_{ijk} = x_i x_j x_k$$
-      \item \tikzmark{innerp}Inner product
+      \item<2-> \tikzmark{innerp}Inner product
         \begin{align*}
           \innerp{A}{B} &= \sum_{ijk} A_{ijk} B_{ijk} \\
-          \pause
+          \uncover<3->{%
             &= \innerp{\vvec{A}}{\vvec{B}} 
+          }
         \end{align*}
@@
     \node[scale=2.0] (op1c) at ($(ptB) + (0.8cm,-0.3cm)$) {$\rangle$};
-    \node[scale=2.0] (op2) at ($(op1c) + (1.5cm,0.0cm)$) {$= \hlmath{green}{0.5}$};
+    \node[scale=1.0] (op2) at ($(op1c) + (1.5cm,0.0cm)$) {$= \mathmg{0.5}$};
     }
@@
     \uncover<3->{%
-    \tikzrect{innerpAv}{black,fill=yellow}{($(ptA) + (0,-1.5cm)$)}{0.3}{1.5};
-    \tikzrect{innerpBv}{black,fill=blue,opacity=0.7}{($(ptB) + (0,-1.5cm)$)}{0.3}{1.5};
-    \node[scale=2.0] (op1a) at ($(innerpAv) + (-1.3cm,-0.6cm)$) {$\langle$};
-    \node (op1b) at ($(innerpAv) + (0.7cm,-0.6cm)$) {$,$};
-    \node[scale=2.0] (op1c) at ($(innerpBv) + (0.8cm,-0.6cm)$) {$\rangle$};
-    \node[scale=2.0] (op2) at ($(op1c) + (1.5cm,0.0cm)$) {$= \hlmath{green}{0.5}$};
+    \tikzrect{innerpAv}{black,fill=yellow}{($(ptA) + (0,-1.5cm)$)}{0.3}{2.0};
+    \tikzrect{innerpBv}{black,fill=blue,opacity=0.7}{($(ptB) + (0,-1.5cm)$)}{0.3}{2.0};
+    \node[scale=2.0] (op1a) at ($(innerpAv) + (-1.3cm,-1.1cm)$) {$\langle$};
+    \node (op1b) at ($(innerpAv) + (0.7cm,-1.1cm)$) {$,$};
+    \node[scale=2.0] (op1c) at ($(innerpBv) + (0.8cm,-1.1cm)$) {$\rangle$};
+    \node[scale=1.0] (op2) at ($(op1c) + (1.5cm,0.0cm)$) {$= \mathmg{0.5}$};
     }
@@
     \begin{itemize}
-      \item \tikzmark{gen} Generative process:
+      \item Generative process\tikzmark{gen}:
       \begin{align*}
-        h &\sim \Mult(\pi) \\
+        h &\sim \Mult([\pi_1, \pi_2, \cdots, \pi_k]) \\
         x &\sim \normal{\beta_h}{\sigma^2}.
@@
     \point{start}{({pic cs:gen} -| mark)};
-    \node[anchor=west] (diag1) at ($(start)$) {%
-      \includegraphics[width=0.45\textwidth,height=2cm,keepaspectratio]{figures/gen.png}
-    };
-    \node[anchor=west] (diag) at ($(start) + (1cm,0.0cm)$) {%
+    \drawgen{($(start) + (0,2cm)$)};
+    %\node[anchor=west] (diag1) at ($(start)$) {%
+    %  \includegraphics[width=0.45\textwidth,height=2cm,keepaspectratio]{figures/gen.png}
+    %};
+    \node[anchor=west] (diag) at ($(start) + (1cm,1.0cm)$) {%
       \includegraphics[width=0.45\textwidth,height=3cm,keepaspectratio]{figures/mog.png}
@@
   \frametitle{Solution: Tensor Factorization}
+  \cornertext<3->{\cite{AnandkumarGeHsu2012}}
+
   \splitcolumn{%
@@
       \item \tikzmark{gen}$\E[x\tp{3}] = \sum_{h=1}^k \pi_h \beta_h\tp{3}$.
-      \item<2-> If $\beta_h$ are orthogonal, the are eigenvectors! \footcite{AnandkumarGeHsu2012}
+      \item<3-> If $\beta_h$ are orthogonal, the are eigenvectors!
         \begin{align*}
@@
         \end{align*}
-      \item<3-> In general, whiten $\E[x\tp{3}]$ first.
+      \item<4-> In general, whiten $\E[x\tp{3}]$ first.
     \end{itemize}
@@
 
-    \node[anchor=west] (diag1) at ($(start)$) {%
-      \includegraphics[width=0.45\textwidth,height=2cm,keepaspectratio]{figures/gen.png}
-    };
+    \drawgen{($(start) + (0,2cm)$)};
     \node[anchor=west] (diag) at ($(start) + (1cm,0.0cm)$) {%
@@
     };
-    %\drawgen{(start)};
-    %\node[anchor=west] (diag) at ($(start) + (1cm,-0.5cm)$) {%
-    %  \includegraphics[width=0.45\textwidth,height=3cm,keepaspectratio]{figures/mog.png}
-    %};
-    % x, y
-    %\node[anchor=north west] at ($(start) - (1cm,2cm)$) {%
-    %    \includegraphics[width=6cm,height=2cm,keepaspectratio]{figures/tensor.png}
-    %};
+    \uncover<2->{%
+      \tensorfactorization{(2cm,-1cm)};
+    }
   \end{canvas}
@@
     % Tasks.
-    \node[anchor=west] (gen) at (1cm,0) {%
-      \includegraphics[width=\textwidth,height=2.5cm,keepaspectratio]{figures/gen.png}
-     };
-     \node[below=0.1cm of gen.south] {Generative Models};
+    \drawgen{(3cm,0)}
+    \node[below=0.1cm of x.south] {Generative Models};
 
-    \draw[scale=0.8,fill=green,opacity=0.4,dashed] (6.5cm,2cm) rectangle (12.5cm,-3cm);
-    \node[style=box,anchor=west] (disc) at (6cm,0) {%
-      \includegraphics[width=\textwidth,height=3cm,keepaspectratio]{figures/disc.png}
-     };
-     \node[below=0.1cm of disc.south] {Discriminative Models};
     % Highlight
+    \draw[scale=0.8,fill=green,opacity=0.4,dashed] (8cm,1cm) rectangle (14cm,-4cm);
+      \drawdisc{(9cm,0)}
+      \node[below=0.1cm of y.south] {Discriminative Models};
+
   \end{canvas}
@@
     \tikzmark{model}
-    \includegraphics[width=3cm,height=4cm,keepaspectratio]{figures/disc.png}
+    \begin{tikzpicture}
+      \drawdisc{(0,0)}
+    \end{tikzpicture}
     \begin{itemize}
-      \item<2-> $h \sim \Mult(\pi)$.
+      \item<2-> $h \sim \Mult([\pi_1, \pi_2, \cdots, \pi_k])$.
       \item<4-> $y = \beta_h^T x + \epsilon$.
@@
       \includegraphics<6>[width=5cm,height=6cm,keepaspectratio]{figures/mlr-data-5.png}
+      \includegraphics<7>[width=5cm,height=6cm,keepaspectratio]{figures/mlr-data-6.png}
         };
@@
     % x, y
-    \node[anchor=west,scale=2.0] (params) at (7cm,0) {%
-      $\begin{bmatrix} \pi \\ B \end{bmatrix}$
+    \node[anchor=west,scale=1.0] (params) at (7cm,0) {%
+      $\begin{bmatrix} \pi_1 \\ \pi_2 \\ \vdots \\ \pi_k \end{bmatrix}  
+       \begin{bmatrix} 
+                 &         &       &         \\
+                 &         &       &         \\
+         \beta_1 & \beta_2 & \dots & \beta_k \\
+                 &         &       &         \\
+                 &         &       &         
+       \end{bmatrix} $
       };
-    \node[anchor=west,right=0.1cm of params] {%
-    \includegraphics[width=4cm,height=2cm,keepaspectratio]{figures/disc.png}
-    };
+    %\node[anchor=west,right=0.1cm of params] {%
+    %\includegraphics[width=4cm,height=2cm,keepaspectratio]{figures/disc.png}
+    %};
 
@@
 \begin{frame}[c]
-  \frametitle{Method of Moments for the Mixture of Linear Regressions.}
+  \frametitle{Finding Tensor Structure}
+  \withrmargin{%
   \begin{align*}
-    y &= \alt<1>{\innerp{\beta_h}{x}}{\underbrace{\beta_h^T}_{\textrm{\color{DarkGreen}random}} x} + \epsilon \\
-    \action<3->{%
-    &= \sum_h \pi_h \innerp{\beta_h}{x} + \innerp{\left(\beta_h - \sum_h \pi_h \beta_h\right)}{x} + \epsilon \\
-    }
-    \action<4-6>{%
-    &= \alt<4>{\E[\beta_h]^T x}{\underbrace{\E[\beta_h]^T x}_{\textrm{\color{blue}linear measurement}}} 
-    + \alt<4-5>{\innerp{(\beta_h - \E[\beta_h])}{x} + \epsilon}{\underbrace{\innerp{(\beta_h - \E[\beta_h])}{x} + \epsilon}_{\textrm{\color{red} noise}}} \\
+    y &= \innerpp{\robustaltm<1>{\beta_h}{\underbrace{\beta_h}_{\textmg{random}}}}
+                {x} + \epsilon \\
+    \uncover<4-6>{%
+    &= \robustaltm<4>{\innerp{\E[\beta_h]}{x}}
+      {\mathmb{\ub{\innerp{\E[\beta_h]}{x}}_{\textrm{linear measurement}}}} 
+    + \robustaltm<4-5>{\innerp{(\beta_h - \E[\beta_h])}{x} + \epsilon}
+    {\mathmr{\ub{\innerp{(\beta_h - \E[\beta_h])}{x} + \epsilon}_{\textrm{noise}}}} \\
     }
   \end{align*}
+  }{%
+    \uncover<3->{%
+        $\mboxg{mom}{\E[\beta_h] = \sum_h \pi_h \beta_h}$.
+    }
+  }
 \end{frame}
@@
 \begin{frame}
-  \frametitle{Method of Moments for the Mixture of Linear Regressions.}
+  \frametitle{Finding Tensor Structure}
 
+  \splitcolumn{%
   \begin{align*}
     \action<1->{%
-    y &= \overbrace{\innerp{\E[\beta_h]}{x}}^{\textrm{\color{blue} linear measurement}} & & + \overbrace{(\beta_h - \E[\beta_h])^T x + \epsilon}^{\textrm{\color{red} noise}} \\
+    y &= \mathmb{\ob{\innerp{\E[\beta_h]}{x}}^{\textrm{linear measurement}}} &&+ \mathmr{\ob{(\beta_h - \E[\beta_h])^T x + \epsilon}^{\textrm{noise}}} \\[4ex]
     }
@@
       y^2 
-      &= \innerp{\beta_h}{x}^2 &+ \epsilon^2 &+ 2 \epsilon \innerp{\beta_h}{x}  \\
+      &= \left(\innerp{\beta_h}{x} + \epsilon\right)^2 && \\
     }
     \action<3->{%
-      &= \color{blue}\innerp{\E[\beta_h\tp{2}]}{x\tp{2}} &+ \color{DarkGreen} \textrm{bias} &+ \color{red} \textrm{noise} \\
+    &= 
+    \color{blue}
+      \robustaltm<3>{%
+            \innerp{\E[\beta_h\tp{2}]}{x\tp{2}}
+        }{%
+          \ub{\innerp{\E[\beta_h\tp{2}]}{x\tp{2}}}_{M_2} 
         }
-    \action<4->{%
-    y^3 &= \color{blue} \innerp{\E[\beta_h\tp{3}]}{x\tp{3}} &+ \color{DarkGreen} \textrm{bias} &+ \color{red} \textrm{noise} \\
+        &&+ \color{DarkGreen} \textrm{bias} + \color{red} \textrm{noise} \\[4ex]
+    }
+    \action<5->{%
+    y^3 &= \color{blue} \ub{\innerp{\E[\beta_h\tp{3}]}{x\tp{3}}}_{M_3} &&+ \color{DarkGreen} \textrm{bias} + \color{red} \textrm{noise} 
     }
   \end{align*}
+  }{%
+  }
 
@@
     \item 
-    $\E[\beta_h\tp{3}] = \sum_{h=1}^k \pi_h \beta_h\tp{3}$
+    $M_3 \eqdef \E[\beta_h\tp{3}] = \sum_{h=1}^k \pi_h \beta_h\tp{3}$
   }
-    \item<2-> Apply tensor factorization!
+    \item<3-> Apply tensor factorization!
   \end{itemize}
+
+  \begin{canvas}
+    \uncover<2->{%
+      \tensorfactorization{(-3cm,-1cm)};
+    }
+  \end{canvas}
+
 \end{frame}
@@
     \input{spectral-experts.tikz}
-    \node[style=txt] at (reg-label) {\alert<1>{regression}};
-    \node[style=txt] at (tf-label) {\alert<2>{tensor factorization}};
+    \node[style=txt] at (reg-label) {\alert<2>{regression}};
+    \node[style=txt] at (tf-label) {\alert<3>{tensor factorization}};
   \end{canvas}
@@
 
+  \cornertext{%
+  \uncover<2->{\cite{fazel2002matrix}}\\
+  \uncover<3->{\cite{tomioka2010estimation}} 
+  }
+
     \begin{align*}
-      \hat\E[\beta_h\tp{2}] &= \arg\min_{M} \left(y^2 - \innerp{M}{x\tp{2}}\right)^2_{(x,y)\in\mathcal{D}} 
-        \alt<1>{+ \|M\|_{2}}
-          {+ \hlmath{blue}{\|M\|_{*}}} \\
-      \action<2->{%
-      \|M\|_{*} &= \sum_i \sigma_i(M).
-      }
+      \hat M_{\robustaltm<-2>{2}{\mathmb{3}}} &= \arg\min_{M} \sum_{(x,y)\in\mathcal{D}} \left(y^2 - 
+      \robustaltm<-2>{\innerp{M_2}{x\tp{2}}}{\mathmb{\innerp{M_3}{x\tp{3}}}}
+      - \textrm{bias}\right)^2 
+      \uncover<2->{+
+      \robustaltm<2>{\mathmb{\ub{\|M\|_{*}}_{\sum_i \sigma_i(M)}}}
+        {\|M\|_{*}}
+      } \\
     \end{align*}
 
-  \begin{tikzpicture}
+  \begin{canvas}
     % x, y
-    \node (measurements) {%
-      \includegraphics[width=\textwidth,height=6cm,keepaspectratio]{figures/tensor.png}
-      };
-  \end{tikzpicture}
+    \alt<-2>{%
+      \matrixfactorization{(2cm,0)}
+    }{%
+      \tensorfactorization{(2cm,0)}
+    }
+  \end{canvas}
 \end{frame}
@@
   \frametitle{Spectral Experts}
+  \cornertext{%
+  \uncover<2->{\cite{NegahbanWainwright2009,Tomioka2011}} \\
+  \uncover<3->{\cite{AnandkumarGeHsu2012}} 
+  }
 
@@
   \begin{itemize}
-    \item {\bf Consistent estimator} for the mixture of linear regressions with {\bf polynomial sample and computational complexity}.
-    \item<2-> Derive conditional moments using regression.
-    \item<3-> Method of moment estimates can be a good initialization for EM.
+    \item<1-> Expose tensor factorization structure through regression.
+    \item<2-> {\bf Consistent estimator} for the mixture of linear regressions with {\bf polynomial sample and computational complexity}.
+    \item<3-> Empirically, method of moment estimates can be a good initialization for EM.
+    \item<4-> How can we handle other discriminative models?
+      \begin{itemize}
+          \item<5-> Dependencies between $h$ and $x$ (mixture of experts).
+          \item<6-> Non-linear link functions (logistic regression).
+      \end{itemize}
   \end{itemize}

### 2013-06-13 doc/icml-presentation/presentation.tex
Added lots of pretty graphs and improved the presentation quatlity.

@@
 \usepackage{tikz}
-\usepackage[debug]{scabby}
-%\usepackage{scabby}
+%\usepackage[debug,debugmarks]{scabby}
+\usepackage{scabby}
 
@@
         }
-      \item<4-> {\em Easy to include features and tend to be more accurate.}
+      \item<5-> {\em Easy to include features and tend to be more accurate.}
       \end{itemize}
@@
     \point{disc}{({pic cs:disc} -| mark)}
-    \drawgen<1->{($(gen) + (0,1.5cm)$)};
-    \drawdisc<3->{($(disc) + (0,0.0cm)$)};
+    \drawgen<1->{($(gen) + (0,0.0cm)$)};
+    \drawdisc<3->{($(disc) + (0,-0.5cm)$)};
   \end{canvas}
@@
     % x, y
-    \node at (0,0) {%
-    \includegraphics<1>[width=\textwidth,height=6cm,keepaspectratio]{figures/likelihood.png}
-    \includegraphics<2>[width=\textwidth,height=6cm,keepaspectratio]{figures/likelihood-mle.png}
-    \includegraphics<3>[width=\textwidth,height=6cm,keepaspectratio]{figures/likelihood-em.png}
-    \includegraphics<4>[width=\textwidth,height=6cm,keepaspectratio]{figures/likelihood-mom.png}
-      };
-    %\point{opt}{(0.4cm,-1.1cm)};
-    %\node[style=lbl,below=0.0cm of opt,scale=1.2] {$\hlmath{blue}{\theta^*}$};
-    % Axes
-    \draw[-latex] (-6cm,-2.25cm) -- node[right,at end]{$-\log(p_{\theta}(x))$}  (-6cm,2cm);
-    \draw[-latex] (-6cm,-2.25cm) -- node[above,at end]{$\theta$}  (5cm,-2.25cm);
+    \llhood{0}{0};
+    \node[scale=0.7] at (mle) {x};
+    \node<2-> at ($(mle) + (0.6cm,0)$) {$\mathmb{\theta^*_{\textrm{MLE}}}$};
+    \node[scale=0.7] at (em1) {x};
+    \node<3-> at ($(em1) + (0.5cm,0)$) {$\mathmr{\hat\theta_{\textrm{EM}}}$};
+    \node[scale=0.7] at (em2) {x};
+    \node<3-> at ($(em2) + (0.5cm,0)$) {$\mathmr{\hat\theta_{\textrm{EM}}}$};
+
+    \draw<4->[latex-latex,DarkGreen,line width=1pt] ($(mle) + (-0.8cm,0.8cm)$) -- node[above]{$\mathmg{\epsilon}$} ($(mle) + (+0.8cm,0.8cm)$);
   \end{tikzpicture}
@@
   \begin{canvas}
-    \point{mark}{(6.8cm,0)};
+    \point{mark}{(1.0cm,0)};
+    \point{tensoring-pos}{({pic cs:tensoring} -| mark)}
     % Tensoring
-    \tikzcube{tensoring}{black,fill=white}{($({pic cs:tensoring} -| mark) + (0,0,0)$)}{1}{1}{1};
+    \tikzcube{tensoring}{black,fill=white}{($(tensoring-pos) + (0,-3cm) + (0,0,0)$)}{1}{1}{1};
     \node at ($(tensoring) + (1.0cm,-0.3cm)$) {$=$};
@@
     \uncover<2->{%
-    \point{ptA}{($({pic cs:innerp} -| mark) + (0,0.2cm)$)}
+    \point{ptA}{($({pic cs:innerp} -| mark) + (0,-3.0cm)$)}
     \point{ptB}{($(ptA) + (2cm,0)$)}
+
     \tikzcube{innerpA}{black,fill=yellow}{($(ptA) + (0,0,0)$)}{1}{1}{1};
-    \tikzcube{innerpB}{black,fill=blue,opacity=0.7}{($(ptB) + (0,0,0)$)}{1}{1}{1};
-    \node[scale=2.0] (op1a) at ($(ptA) + (-1.3cm,-0.3cm)$) {$\langle$};
-    \node (op1b) at ($(ptA) + (0.7cm,-0.3cm)$) {$,$};
-    \node[scale=2.0] (op1c) at ($(ptB) + (0.8cm,-0.3cm)$) {$\rangle$};
-    \node[scale=1.0] (op2) at ($(op1c) + (1.5cm,0.0cm)$) {$= \mathmg{0.5}$};
+    \tikzcube{innerpB}{black,fill=blue!70}{($(ptB) + (0,0,0)$)}{1}{1}{1};
+
+    \innerpdiagt{(ptA)}{(ptB)};
+    \node[scale=1.0] at ($(ptB) + (1.5cm,-0.3cm)$) {$= \mathmg{0.5}$};
     }
@@
     \uncover<3->{%
-    \tikzrect{innerpAv}{black,fill=yellow}{($(ptA) + (0,-1.5cm)$)}{0.3}{2.0};
-    \tikzrect{innerpBv}{black,fill=blue,opacity=0.7}{($(ptB) + (0,-1.5cm)$)}{0.3}{2.0};
-    \node[scale=2.0] (op1a) at ($(innerpAv) + (-1.3cm,-1.1cm)$) {$\langle$};
-    \node (op1b) at ($(innerpAv) + (0.7cm,-1.1cm)$) {$,$};
-    \node[scale=2.0] (op1c) at ($(innerpBv) + (0.8cm,-1.1cm)$) {$\rangle$};
-    \node[scale=1.0] (op2) at ($(op1c) + (1.5cm,0.0cm)$) {$= \mathmg{0.5}$};
+    \point{ptvA}{($(ptA) + (0,-2cm)$)}
+    \point{ptvB}{($(ptB) + (0,-2cm)$)}
+
+    \tikzrect{innerpAv}{black,fill=yellow}{($(ptvA) + (0,0.5cm)$)}{0.3}{2.0};
+    \tikzrect{innerpBv}{black,fill=blue!70}{($(ptvB) + (0,0.5cm)$)}{0.3}{2.0};
+
+    \innerpdiagt{(ptvA)}{(ptvB)};
+    \node[scale=1.0] at ($(ptvB) + (1.5cm,-0.3cm)$) {$= \mathmg{0.5}$};
     }
@@
   \frametitle{Example: Gaussian Mixture Model}
-
   \splitcolumn{%
@@
     % The model
-    \point{mark}{(6.5cm,0)};
-    \point{start}{({pic cs:gen} -| mark)};
-    \drawgen{($(start) + (0,2cm)$)};
+    \point{mark}{(1cm,0)};
+    \point{start}{(1cm,1cm)}; %{pic cs:gen} -| mark)};
+    \drawgen{($(start) + (0,1cm)$)};
     %\node[anchor=west] (diag1) at ($(start)$) {%
@@
     %};
-    \node[anchor=west] (diag) at ($(start) + (1cm,1.0cm)$) {%
-      \includegraphics[width=0.45\textwidth,height=3cm,keepaspectratio]{figures/mog.png}
+    \node[anchor=west] (diag) at ($(start) + (1cm,0.0cm)$) {%
+      \includegraphics[width=0.45\textwidth,height=3cm,keepaspectratio]{figures/mog.pdf}
     };
@@
     \uncover<4->{%
-      \point{m2}{({pic cs:m2} -| mark)};
+      \point{m2-pos}{({pic cs:m2} -| mark)};
+      \point{m2}{($(m2-pos) - (0,4.5cm)$)};
       \tikzrect{M2}{black,fill=white}{($(m2) + (1cm,1cm)$)}{1}{1};
@@
     \uncover<5->{%
-      \point{m3}{($({pic cs:m3} -| mark) + (1cm,1cm)$)};
+      \point{m3-pos}{({pic cs:m3} -| mark)};
+      \point{m3}{($(m3-pos) - (-1cm,3.5cm)$)};
       \tikzcube{M3}{black,fill=white}{($(m3) + (0,0,0)$)}{1}{1}{1};
@@
 
-\begin{frame}
+\begin{frame}[t]
   \frametitle{Solution: Tensor Factorization}
@@
   \begin{canvas}
-    \point{mark}{(6.5cm,0)};
-    \point{start}{({pic cs:gen} -| mark)};
+    \point{mark}{(1cm,0)};
+    \point{start}{(1cm,1cm)}; %{pic cs:gen} -| mark)};
 
-    \drawgen{($(start) + (0,2cm)$)};
+    \drawgen{($(start) + (0,1cm)$)};
     \node[anchor=west] (diag) at ($(start) + (1cm,0.0cm)$) {%
-      \includegraphics[width=0.45\textwidth,height=3cm,keepaspectratio]{figures/mog.png}
+      \includegraphics[width=0.45\textwidth,height=3cm,keepaspectratio]{figures/mog.pdf}
     };
     \uncover<2->{%
-      \tensorfactorization{(2cm,-1cm)};
+      \tensorfactorization{(-3cm,-2cm)};
     }
@@
     % Tasks.
-    \drawgen{(3cm,0)}
+    \drawgen{(-3cm,1cm)}
     \node[below=0.1cm of x.south] {Generative Models};
@@
     % Highlight
-    \draw[scale=0.8,fill=green,opacity=0.4,dashed] (8cm,1cm) rectangle (14cm,-4cm);
-      \drawdisc{(9cm,0)}
+    \draw<2>[scale=0.8,fill=green,opacity=0.4,dashed] (1cm,2.5cm) rectangle (6.5cm,-2.5cm);
+      \drawdisc{(3cm,1cm)}
       \node[below=0.1cm of y.south] {Discriminative Models};
@@
       % x, y
-      \point{mark}{(5.5cm,0)}
+      \point{mark}{(0cm,0cm)}
       \point{model}{({pic cs:model} -| mark)}
-      \node[anchor=west] at ($(model) + (0,0.5cm)$) {%
-      \includegraphics<1>[width=5cm,height=6cm,keepaspectratio]{figures/mlr-data-2.png}
-      \includegraphics<2>[width=5cm,height=6cm,keepaspectratio]{figures/mlr-data-3a.png}
-      \includegraphics<3>[width=5cm,height=6cm,keepaspectratio]{figures/mlr-data-3b.png}
-      \includegraphics<4>[width=5cm,height=6cm,keepaspectratio]{figures/mlr-data-4a.png}
-      \includegraphics<5>[width=5cm,height=6cm,keepaspectratio]{figures/mlr-data-4b.png}
-      %\includegraphics<6>[width54cm,height=6cm,keepaspectratio]{figures/mlr-data-4c.png}
-      \includegraphics<6>[width=5cm,height=6cm,keepaspectratio]{figures/mlr-data-5.png}
-      \includegraphics<7>[width=5cm,height=6cm,keepaspectratio]{figures/mlr-data-6.png}
+      \node[anchor=west] at ($(model) + (0,-1cm)$) {%
+      \includegraphics<1>[width=5cm,height=6cm,keepaspectratio]{figures/mlr-0.pdf}
+      \includegraphics<2>[width=5cm,height=6cm,keepaspectratio]{figures/mlr-1.pdf}
+      \includegraphics<3>[width=5cm,height=6cm,keepaspectratio]{figures/mlr-2.pdf}
+      \includegraphics<4>[width=5cm,height=6cm,keepaspectratio]{figures/mlr-3.pdf}
+      \includegraphics<5>[width=5cm,height=6cm,keepaspectratio]{figures/mlr-4.pdf}
+      \includegraphics<6>[width=5cm,height=6cm,keepaspectratio]{figures/mlr-5.pdf}
+      \includegraphics<7>[width=5cm,height=6cm,keepaspectratio]{figures/mlr-6.pdf}
         };
@@
   \begin{canvas}
-    \node[anchor=west] (data) at (0,0) {%
-    \includegraphics[width=4cm,height=6cm,keepaspectratio]{figures/mlr-data.png}
+    \node[anchor=east] (data) at (-2cm,0) {%
+    \includegraphics[width=4cm,height=6cm,keepaspectratio]{figures/mlr-6.pdf}
     };
@@
     % x, y
-    \node[anchor=west,scale=1.0] (params) at (7cm,0) {%
+    \node[anchor=west,scale=1.0] (params) at (1.5cm,0) {%
       $\begin{bmatrix} \pi_1 \\ \pi_2 \\ \vdots \\ \pi_k \end{bmatrix}  
@@
     \action<1->{%
-    y &= \mathmb{\ob{\innerp{\E[\beta_h]}{x}}^{\textrm{linear measurement}}} &&+ \mathmr{\ob{(\beta_h - \E[\beta_h])^T x + \epsilon}^{\textrm{noise}}} \\[4ex]
+    y\tikzmark{regA} &= \mathmb{\ob{\innerp{\E[\beta_h]}{x}}^{\textrm{linear measurement}}} &&+ \mathmr{\ob{(\beta_h - \E[\beta_h])^T x + \epsilon}^{\textrm{noise}}} \\[4ex]
     }
     \action<2->{%
-      y^2 
+    y^2 \tikzmark{regB}
       &= \left(\innerp{\beta_h}{x} + \epsilon\right)^2 && \\
@@
     \action<5->{%
-    y^3 &= \color{blue} \ub{\innerp{\E[\beta_h\tp{3}]}{x\tp{3}}}_{M_3} &&+ \color{DarkGreen} \textrm{bias} + \color{red} \textrm{noise} 
+    y^3 \tikzmark{regC}
+    &= \color{blue} \ub{\innerp{\E[\beta_h\tp{3}]}{x\tp{3}}}_{M_3} &&+ \color{DarkGreen} \textrm{bias} + \color{red} \textrm{noise} 
     }
@@
   }
+  \begin{canvas}
+    \point{mark}{(3cm,0)};
+    \uncover<1-> {%
+      \regressionA{($({pic cs:regA} -| mark) + (0,-2.5cm)$)};
+    }
+    \uncover<3-> {%
+      \regressionB{($({pic cs:regB} -| mark) + (0,-3cm)$)};
+    }
+    \uncover<5-> {%
+      \regressionC{($({pic cs:regC} -| mark) + (0,-3cm)$)};
+    }
+  \end{canvas}
 
@@
 \begin{frame}
-  \frametitle{Method of Moments for the Mixture of Linear Regressions.}
+  \frametitle{Recovering Parameters.}
   \centering
@@
     \uncover<2->{%
-      \tensorfactorization{(-3cm,-1cm)};
+      \tensorfactorization{(-3cm,-2cm)};
     }
@@
     \begin{align*}
-      \hat M_{\robustaltm<-2>{2}{\mathmb{3}}} &= \arg\min_{M} \sum_{(x,y)\in\mathcal{D}} \left(y^2 - 
-      \robustaltm<-2>{\innerp{M_2}{x\tp{2}}}{\mathmb{\innerp{M_3}{x\tp{3}}}}
+      \robustaltm<-2>{\hat M_2}{\mathmb{\hat M_3}} &= \arg\min_{M} \sum_{(x,y)\in\mathcal{D}} \left(y^2 - 
+      \robustaltm<-2>{\innerp{M}{x\tp{2}}}{\mathmb{\innerp{M}{x\tp{3}}}}
       - \textrm{bias}\right)^2 
@@
     \alt<-2>{%
-      \matrixfactorization{(2cm,0)}
+      \matrixfactorization{(-3cm,-2cm)}
     }{%
-      \tensorfactorization{(2cm,0)}
+      \tensorfactorization{(-3cm,-2cm)}
     }
@@
     % x, y
-    \node[anchor=west] (em) at (-0.5cm,-1.5cm) {%
+    \node[anchor=west] (em) at (-6.5cm,-1.5cm) {%
       \includegraphics<1>[width=\textwidth,height=5cm,keepaspectratio]{figures/1-8-3.png}
@@
 \begin{frame}
-  \frametitle{Experimental Insights}
+  \frametitle{On Initialization (Cartoon)}
 
   \begin{tikzpicture}
-    \node at (0,0) {%
-      \includegraphics<1>[width=\textwidth,height=6cm,keepaspectratio]{figures/likelihood-mom.png}
-      \includegraphics<2>[width=\textwidth,height=6cm,keepaspectratio]{figures/likelihood-mom-em1.png}
-      \includegraphics<3>[width=\textwidth,height=6cm,keepaspectratio]{figures/likelihood-mom-em2.png}
-    };
-    \draw[-latex] (-6cm,-2.25cm) -- node[right,at end]{$-\log(p_{\theta}(x))$}  (-6cm,2cm);
-    \draw[-latex] (-6cm,-2.25cm) -- node[above,at end]{$\theta$}  (5cm,-2.25cm);
+    % x, y
+    \llhood{0}{0};
 
-    \node<2-> at (1.8cm,0.6cm) {$\hlmath{DarkGreen}{\theta_{spec}}$};
-    \node<3-> at (0.3cm,-0.8cm) {$\hlmath{blue}{\theta^{*}}$};
+    \node<2->[scale=0.7] at (em2) {x};
+    \node<2-> at ($(em2) + (0.6cm,0)$) {$\mathmr{\hat\theta_{\textrm{EM}}}$};
+    \draw<2->[-latex,smooth,line width=1pt,red] ($(em2-start) + (+0.1cm,+0.05cm)$) -- ($(em2) + (+0.15cm,0.00cm)$);
+    \draw<2->[dashed,red,line width=0.7pt] ($(em2)-(3.5cm,0)$) -- ($(em2)+(0.5cm,0)$);
+
+    \draw<3>[latex-latex,DarkGreen,line width=1pt,dashed] ($(mle) + (-1.2cm,0.8cm)$) -- node[above]{$\mathmg{\epsilon}$} ($(mle) + (+1.2cm,0.8cm)$);
+    \node<4->[scale=0.7] at (spec) {x};
+    \node<4-> at ($(spec) + (0.5cm,0.3cm)$) {$\mathmg{\hat\theta_{\textrm{spec}}}$};
+    \draw<4->[dashed,DarkGreen,line width=0.7pt] ($(spec)-(0.5cm,0)$) -- ($(spec)+(3.5cm,0)$);
+
+    \draw<5->[-latex,smooth,line width=1pt,DarkGreen] ($(spec) + (-0.1cm,+0.05cm)$) -- ($(mle) + (-0.30cm,+0.10cm)$);
+    %\node[scale=0.7] at (mle) {x};
+    \node<5->[anchor=west] at ($(mle) + (0.3cm,0)$) {$\mathmb{\theta^*}_{\textrm{spec + EM}}$};
+
   \end{tikzpicture}

### 2013-06-13 doc/icml-presentation/presentation.tex
Added crisp figures for experiments. Good version of slideS

@@
           \tikzmark{m2}
-          \E[x\tp{2}] &= \sum_h \pi_h (\beta_h \beta_h^T) \\
-                      &= \sum_h \pi_h \beta_h{\tp{2}} \\
+          \E[x\tp{2}] &= \sum_h \pi_h (\beta_h \beta_h^T) + \sigma^2 \\
+                      &= \sum_h \pi_h \beta_h{\tp{2}} + \sigma^2 \\
                       }
@@
           \tikzmark{m3}
-          \E[x\tp{3}] &= \sum_h \pi_h \beta_h\tp{3}.
+          \E[x\tp{3}] &= \sum_h \pi_h \beta_h\tp{3} + \textrm{bias}.
           }
@@
 \begin{frame}
-  \frametitle{Recovering Parameters.}
+  \frametitle{Recovering Parameters}
   \centering
@@
 \begin{frame}
-  \frametitle{Spectral Experts}
+  \frametitle{Overview: Spectral Experts}
   \cornertext{%
@@
     % x, y
-    \node[anchor=west] (em) at (-6.5cm,-1.5cm) {%
-      \includegraphics<1>[width=\textwidth,height=5cm,keepaspectratio]{figures/1-8-3.png}
-      \includegraphics<2>[width=\textwidth,height=5cm,keepaspectratio]{figures/1-8-3-em.png}
-      \includegraphics<3>[width=\textwidth,height=5cm,keepaspectratio]{figures/1-8-3-spec.png}
-      \includegraphics<4->[width=\textwidth,height=5cm,keepaspectratio]{figures/1-8-3-specm.png}
-      };
-    % x, y
-    \node[anchor=west, right=6cm of em] at (em.west) {%
-      \includegraphics[width=\textwidth,height=5cm,keepaspectratio]{figures/hist.png}
+    \node[anchor=west] (em) at (-6.5cm,-0.0cm) {%
+    \includegraphics<1>[width=\textwidth,height=5cm,keepaspectratio]{figures/1833.pdf}
+      \includegraphics<2-3>[width=\textwidth,height=5cm,keepaspectratio]{figures/EM-1833.pdf}
+      \includegraphics<4-5>[width=\textwidth,height=5cm,keepaspectratio]{figures/Spectral-1833.pdf}
+      \includegraphics<6->[width=\textwidth,height=5cm,keepaspectratio]{figures/Spectral-Spectral+EM-1833.pdf}
       };
 
-    \node[above=0.1 cm of em] {%
+    \node<1-2>[anchor=west, right=1 cm of em.east] {%
         $y = \beta^T 
@@
       };
+    % x, y
+    \node[anchor=west, right=6.2cm of em] at (em.west) {%
+      \includegraphics<3-4>[width=\textwidth,height=5cm,keepaspectratio]{figures/EM-hist.pdf}
+      \includegraphics<5>[width=\textwidth,height=5cm,keepaspectratio]{figures/EM-Spectral-hist.pdf}
+      \includegraphics<6>[width=\textwidth,height=5cm,keepaspectratio]{figures/EM-Spectral-Spectral+EM-hist.pdf}
+    };
   \end{canvas}
@@
 
-\begin{small}
-  \begin{tabular}{r r r c c c}
-\hline
-%$b$ & 
-$d$ & $k$ & Spectral & EM & Spectral + EM \\
-\hline
-  %1 & 
-  4 & 2 & 2.45 $\pm$ 3.68 & 0.28 $\pm$ 0.82 & {\bf 0.17 $\pm$ 0.57} \\
-  %2 & 
-  5 & 2 & 1.38 $\pm$ 0.84 & {\bf 0.00 $\pm$ 0.00} & {\bf 0.00 $\pm$ 0.00} \\
-  %2 & 
-  5 & 3 & 2.92 $\pm$ 1.71 & 0.43 $\pm$ 1.07 & {\bf 0.31 $\pm$ 1.02} \\
-  %2 & 
-  6 & 2 & 2.33 $\pm$ 0.67 & 0.63 $\pm$ 1.29 & {\bf 0.01 $\pm$ 0.01} \\
-\hline
-\end{tabular}
-      \end{small}
+  \begin{canvas}
+    \node (exp1) at (-3cm,2cm) {\includegraphics[width=4.5cm]{figures/err-hist-0.pdf}};
+    \node[below=-0.2cm of exp1] {$d=4, k=2$};
+    \node (exp2) at (3cm,2cm) {\includegraphics[width=4.5cm]{figures/err-hist-1.pdf}};
+    \node[below=-0.2cm of exp2] {$d=5, k=2$};
+    \node (exp3) at (-3cm,-2cm) {\includegraphics[width=4.5cm]{figures/err-hist-2.pdf}};
+    \node[below=-0.2cm of exp3] {$d=5, k=3$};
+    \node (exp4) at (3cm,-2cm) {\includegraphics[width=4.5cm]{figures/err-hist-3.pdf}};
+    \node[below=-0.2cm of exp4] {$d=6, k=2$};
+  \end{canvas}
+
+%\begin{small}
+%  \begin{tabular}{r r r c c c}
+%\hline
+%%$b$ & 
+%$d$ & $k$ & Spectral & EM & Spectral + EM \\
+%\hline
+%  %1 & 
+%  4 & 2 & 2.45 $\pm$ 3.68 & 0.28 $\pm$ 0.82 & {\bf 0.17 $\pm$ 0.57} \\
+%  %2 & 
+%  5 & 2 & 1.38 $\pm$ 0.84 & {\bf 0.00 $\pm$ 0.00} & {\bf 0.00 $\pm$ 0.00} \\
+%  %2 & 
+%  5 & 3 & 2.92 $\pm$ 1.71 & 0.43 $\pm$ 1.07 & {\bf 0.31 $\pm$ 1.02} \\
+%  %2 & 
+%  6 & 2 & 2.33 $\pm$ 0.67 & 0.63 $\pm$ 1.29 & {\bf 0.01 $\pm$ 0.01} \\
+%\hline
+%\end{tabular}
+%      \end{small}
 
@@
     \draw<5->[-latex,smooth,line width=1pt,DarkGreen] ($(spec) + (-0.1cm,+0.05cm)$) -- ($(mle) + (-0.30cm,+0.10cm)$);
-    %\node[scale=0.7] at (mle) {x};
+    \node<5->[scale=0.7] at (mle) {x};
     \node<5->[anchor=west] at ($(mle) + (0.3cm,0)$) {$\mathmb{\theta^*}_{\textrm{spec + EM}}$};
-
+    \draw<5->[dashed,blue,line width=0.7pt] ($(mle)-(0.4cm,0)$) -- ($(mle)+(0.4cm,0)$);
   \end{tikzpicture}
@@
     \item<1-> Expose tensor factorization structure through regression.
-    \item<2-> {\bf Consistent estimator} for the mixture of linear regressions with {\bf polynomial sample and computational complexity}.
+    \item<2-> Consistent estimator for the mixture of linear regressions with polynomial sample and computational complexity.
     \item<3-> Empirically, method of moment estimates can be a good initialization for EM.

### 2013-06-13 doc/icml-presentation/data/1833/plotLines.py

@@ -1,0 +1,138 @@
+import numpy as np
+import operator
+import matplotlib
+#matplotlib.use("Agg")
+
+import matplotlib.pyplot as plt
+
+def parseBetasFile(fname):
+  betas = []
+
+  beta = []
+  for line in open(fname).xreadlines():
+    if line.startswith("betas"):
+      if( len(beta) > 0 ):
+        betas.append(np.array(beta))
+        beta = []
+    else:
+      try:
+        beta.append( map(float, line.split()) )
+      except Exception as e:
+        print "[Warning]: Couldn't parse line", line
+        print e
+        continue
+  if( len(beta) > 0 ):
+    betas.append(np.array(beta))
+  return betas
+
+def fs153(x):
+    return np.array([1.0, x, x**4, x**7])
+
+def getLine( beta, fn, W ):
+    """Gets a surface with X, Y, Z = f(X,Y) to be plotted"""
+    X = np.linspace( -1, 1, W)
+    # Grid of X's
+    Y = map(lambda (x): beta.dot(fn(x)), X)
+
+    return X, Y
+
+def plotLines(ax, betas, modifier = "-", colors = None, W = 100):
+    if colors is None:
+        colors = ['b','r','g']
+    colors = [ color + modifier for color in colors ]
+    lines = []
+    for beta, color in zip(betas.T, colors):
+        X, Y = getLine( beta, fs153, W )
+        lines += ax.plot(X, Y, color, linewidth=5)
+    return lines
+
+def makePlot( data, trueBetas, specBetas, spemBetas, em0Betas, emBetas, outFile ):
+
+    plt.rc("font",size=20)
+
+    for setting in [ [], ["EM"],["Spectral"],["Spectral","Spectral+EM"]]:
+        fig = plt.figure()
+        ax = plt.axes(xlim=(-1, 1), ylim=(-3, 5))
+        ax.set_xlabel("x")
+        ax.set_ylabel("y")
+
+        # Plot the data
+        if( len(setting) == 0 ):
+            ax.scatter( data[0], data[1], alpha = 0.4 )
+        else:
+            ax.scatter( data[0], data[1], alpha = 0.1 )
+
+        if "True" in setting:
+            # Now draw each line
+            plotLines( ax, trueBetas, "-" )
+
+        # Legend 
+        plottables = []
+        # Spectral is dashed
+        if( "Spectral" in setting ):
+            lines = plotLines( ax, specBetas, "--" )
[…]

### 2013-06-13 doc/icml-presentation/presentation.tex
Fixed a couple of typos and bugs

@@
       \item Control Theory [Ljung, 1987]
-      \item Observable operator models [Jaeger, 2000; Singh et al., 2004]
-      \item \alert<2>{Hidden Markov models [Hsu et al, 2009]}
+      \item Observable operator models [Jaeger, 2000; Littman / Sutton / Singh, 2004]
+      \item \alert<2>{Hidden Markov models [Hsu/Kakade/Zhang]}
       \item Low-treewidth graphs [Parikh et al., 2012]
@@
       \item \tikzmark{tensoring} Tensoring
-        $$x\tp{3}_{ijk} = x_i x_j x_k$$
+        \begin{align*}
+          x\tp{3}       &= x \otimes x \otimes x \\
+          x\tp{3}_{ijk} &= x_i x_j x_k
+        \end{align*}
       \item<2-> \tikzmark{innerp}Inner product
@@
       \end{align*}
-      \item Moments:
+      \item<2-> Moments:
       \begin{align*}
@@
       \item \tikzmark{gen}$\E[x\tp{3}] = \sum_{h=1}^k \pi_h \beta_h\tp{3}$.
-      \item<3-> If $\beta_h$ are orthogonal, the are eigenvectors!
+      \item<3-> If $\beta_h$ are orthogonal, they are eigenvectors!
         \begin{align*}
@@
     \begin{itemize}
+      \item<2-> Given x
+      \begin{itemize}
         \item<2-> $h \sim \Mult([\pi_1, \pi_2, \cdots, \pi_k])$.
-      \item<4-> $y = \beta_h^T x + \epsilon$.
+        \item<3-> $y = \beta_h^T x + \epsilon$.
+      \end{itemize}
     \end{itemize}
@@
                 {x} + \epsilon \\
-    \uncover<4-6>{%
-    &= \robustaltm<4>{\innerp{\E[\beta_h]}{x}}
+    \uncover<3-5>{%
+    &= \robustaltm<3>{\innerp{\E[\beta_h]}{x}}
       {\mathmb{\ub{\innerp{\E[\beta_h]}{x}}_{\textrm{linear measurement}}}} 
-    + \robustaltm<4-5>{\innerp{(\beta_h - \E[\beta_h])}{x} + \epsilon}
+    + \robustaltm<3-4>{\innerp{(\beta_h - \E[\beta_h])}{x} + \epsilon}
     {\mathmr{\ub{\innerp{(\beta_h - \E[\beta_h])}{x} + \epsilon}_{\textrm{noise}}}} \\
@@
       \robustaltm<3>{%
-            \innerp{\E[\beta_h\tp{2}]}{x\tp{2}}
+            \innerpp{\E[\beta_h\tp{2}]}{x\tp{2}}
         }{%
-          \ub{\innerp{\E[\beta_h\tp{2}]}{x\tp{2}}}_{M_2} 
+          \innerpp{\ub{\E[\beta_h\tp{2}]}_{M_2}}{x\tp{2}}
         }
@@
     y^3 \tikzmark{regC}
-    &= \color{blue} \ub{\innerp{\E[\beta_h\tp{3}]}{x\tp{3}}}_{M_3} &&+ \color{DarkGreen} \textrm{bias} + \color{red} \textrm{noise} 
+    &= \color{blue} \innerpp{\ub{\E[\beta_h\tp{3}]}_{M_3}}{x\tp{3}} &&+ \color{DarkGreen} \textrm{bias} + \color{red} \textrm{noise} 
     }
@@
     \begin{align*}
-      \robustaltm<-2>{\hat M_2}{\mathmb{\hat M_3}} &= \arg\min_{M} \sum_{(x,y)\in\mathcal{D}} \left(y^2 - 
-      \robustaltm<-2>{\innerp{M}{x\tp{2}}}{\mathmb{\innerp{M}{x\tp{3}}}}
+      \robustaltm<-3>{\hat M_2}{\mathmb{\hat M_3}} &= \arg\min_{M} \sum_{(x,y)\in\mathcal{D}} \left(y^2 - 
+      \robustaltm<-3>{\innerp{M}{x\tp{2}}}{\mathmb{\innerp{M}{x\tp{3}}}}
       - \textrm{bias}\right)^2 
-      \uncover<2->{+
+      \uncover<3->{+
       \robustaltm<2>{\mathmb{\ub{\|M\|_{*}}_{\sum_i \sigma_i(M)}}}
@@
     % x, y
-    \alt<-2>{%
+    \uncover<2-3>{
       \matrixfactorization{(-3cm,-2cm)}
-    }{%
+    }
+    \only<4>{
       \tensorfactorization{(-3cm,-2cm)}
@@
     \input{spectral-experts.tikz}
-    \node[style=txt] at (reg-label) {{\color{blue}\bf low-rank} regression};
+    \node[style=txt] at (reg-label) {\robustalt<1>{\color{blue}\bf low-rank}{low-rank} regression};
     \node[style=txt] at (tf-label) {tensor factorization};
@@
       \includegraphics<2-3>[width=\textwidth,height=5cm,keepaspectratio]{figures/EM-1833.pdf}
-      \includegraphics<4-5>[width=\textwidth,height=5cm,keepaspectratio]{figures/Spectral-1833.pdf}
-      \includegraphics<6->[width=\textwidth,height=5cm,keepaspectratio]{figures/Spectral-Spectral+EM-1833.pdf}
+      \includegraphics<4>[width=\textwidth,height=5cm,keepaspectratio]{figures/Spectral-1833.pdf}
+      \includegraphics<5->[width=\textwidth,height=5cm,keepaspectratio]{figures/Spectral-Spectral+EM-1833.pdf}
       };
@@
         $y = \beta^T 
-            \begin{bmatrix} 
+            \ub{\begin{bmatrix} 
               1 \\
@@
               t^7
-            \end{bmatrix} + \epsilon$
+            \end{bmatrix}}_{x} + \epsilon$
       };
@@
     \node[anchor=west, right=6.2cm of em] at (em.west) {%
-      \includegraphics<3-4>[width=\textwidth,height=5cm,keepaspectratio]{figures/EM-hist.pdf}
-      \includegraphics<5>[width=\textwidth,height=5cm,keepaspectratio]{figures/EM-Spectral-hist.pdf}
-      \includegraphics<6>[width=\textwidth,height=5cm,keepaspectratio]{figures/EM-Spectral-Spectral+EM-hist.pdf}
+      \includegraphics<3>[width=\textwidth,height=5cm,keepaspectratio]{figures/EM-hist.pdf}
+      \includegraphics<4>[width=\textwidth,height=5cm,keepaspectratio]{figures/EM-Spectral-hist.pdf}
+      \includegraphics<5>[width=\textwidth,height=5cm,keepaspectratio]{figures/EM-Spectral-Spectral+EM-hist.pdf}
     };

### 2013-06-13 doc/icml-presentation/data/1833/plotLines.py

@@ -53,7 +53,7 @@ def makePlot( data, trueBetas, specBetas, spemBetas, em0Betas, emBetas, outFile 
     for setting in [ [], ["EM"],["Spectral"],["Spectral","Spectral+EM"]]:
         fig = plt.figure()
         ax = plt.axes(xlim=(-1, 1), ylim=(-3, 5))
-        ax.set_xlabel("x")
+        ax.set_xlabel("t")
         ax.set_ylabel("y")
 
         # Plot the data

### 2013-06-13 doc/icml-presentation/presentation.tex
Minor fixes here and there

@@
       \item Observable operator models [Jaeger, 2000; Littman / Sutton / Singh, 2004]
-      \item \alert<2>{Hidden Markov models [Hsu/Kakade/Zhang]}
+      \item \alert<2>{Hidden Markov models [Hsu/Kakade/Zhang, 2009]}
       \item Low-treewidth graphs [Parikh et al., 2012]
@@
     \begin{align*}
-      \robustaltm<-3>{\hat M_2}{\mathmb{\hat M_3}} &= \arg\min_{M} \sum_{(x,y)\in\mathcal{D}} \left(y^2 - 
-      \robustaltm<-3>{\innerp{M}{x\tp{2}}}{\mathmb{\innerp{M}{x\tp{3}}}}
+      \robustaltm<-2>{\hat M_2}{\mathmb{\hat M_3}} &= \arg\min_{M} \sum_{(x,y)\in\mathcal{D}} \left(y^2 - 
+      \robustaltm<-2>{\innerp{M}{x\tp{2}}}{\mathmb{\innerp{M}{x\tp{3}}}}
       - \textrm{bias}\right)^2 
-      \uncover<3->{+
+      \uncover<2->{+
       \robustaltm<2>{\mathmb{\ub{\|M\|_{*}}_{\sum_i \sigma_i(M)}}}
@@
     % x, y
-    \uncover<2-3>{
+    \uncover<1-2>{
       \matrixfactorization{(-3cm,-2cm)}
     }
-    \only<4>{
+    \only<3>{
       \tensorfactorization{(-3cm,-2cm)}
@@
 \begin{frame}
-  \frametitle{Overview: Spectral Experts}
+  \frametitle{Sample Complexity}
   \cornertext{%
@@
     \input{spectral-experts.tikz}
-    \node[style=txt] at (reg-label) {\robustalt<1>{\color{blue}\bf low-rank}{low-rank} regression};
+    \node[style=txt] at (reg-label) {\robustalt<1>{\color{blue}\textbf{low-rank}}{low-rank}regression};
     \node[style=txt] at (tf-label) {tensor factorization};
@@
 
-    \node<1-2>[anchor=west, right=1 cm of em.east] {%
+    \node<1-2>[anchor=west, right=1 cm of em.east] (eq) {%
         $y = \beta^T 
@@
             \end{bmatrix}}_{x} + \epsilon$
+      };
+    \node<1-2>[anchor=west, below=0.1 cm of eq] {%
+        $k = 3, d = 4, n = 10^5$
       };
@@
     \node<5->[scale=0.7] at (mle) {x};
-    \node<5->[anchor=west] at ($(mle) + (0.3cm,0)$) {$\mathmb{\theta^*}_{\textrm{spec + EM}}$};
+    \node<5->[anchor=west] at ($(mle) + (0.3cm,0)$) {$\mathmb{\hat\theta}_{\textrm{spec + EM}}$};
     \draw<5->[dashed,blue,line width=0.7pt] ($(mle)-(0.4cm,0)$) -- ($(mle)+(0.4cm,0)$);

### 2013-06-14 doc/icml-presentation/presentation.tex
Included a note about assumptions and the rest as usual

@@
     \llhood{0}{0};
-    \node[scale=0.7] at (mle) {x};
-    \node<2-> at ($(mle) + (0.6cm,0)$) {$\mathmb{\theta^*_{\textrm{MLE}}}$};
-    \node[scale=0.7] at (em1) {x};
-    \node<3-> at ($(em1) + (0.5cm,0)$) {$\mathmr{\hat\theta_{\textrm{EM}}}$};
-    \node[scale=0.7] at (em2) {x};
-    \node<3-> at ($(em2) + (0.5cm,0)$) {$\mathmr{\hat\theta_{\textrm{EM}}}$};
+    \node<2->[scale=0.3,circle,fill=black] at (mle) {};
+    \node<2-> at ($(mle) + (0.6cm,0)$) {$\mathmb{\theta_{\textrm{MLE}}}$};
+    \node<3->[scale=0.3,circle,fill=black] at (em1) {};
+    \node<3-> at ($(em1) + (0.5cm,0)$) {$\mathmr{\theta_{\textrm{EM}}}$};
+    \node<3->[scale=0.3,circle,fill=black] at (em2) {};
+    \node<3-> at ($(em2) + (0.5cm,0)$) {$\mathmr{\theta_{\textrm{EM}}}$};
 
-    \draw<4->[latex-latex,DarkGreen,line width=1pt] ($(mle) + (-0.8cm,0.8cm)$) -- node[above]{$\mathmg{\epsilon}$} ($(mle) + (+0.8cm,0.8cm)$);
+    %\draw<4->[latex-latex,DarkGreen,line width=1pt] ($(mle) + (-0.8cm,0.8cm)$) -- node[above]{$\mathmg{\epsilon}$} ($(mle) + (+0.8cm,0.8cm)$);
   \end{tikzpicture}
@@
     \item<1-> Method of Moments [Pearson, 1894]
-    \item<2-> Observable operator models
+    \item<2-> Observable operators
     \begin{itemize}
       \item Control Theory [Ljung, 1987]
-      \item Observable operator models [Jaeger, 2000; Littman / Sutton / Singh, 2004]
-      \item \alert<2>{Hidden Markov models [Hsu/Kakade/Zhang, 2009]}
+      \item Observable operator models [Jaeger, 2000; {\small{Littman/Sutton/Singh, 2004}}]
+      \item Hidden Markov models [Hsu/Kakade/Zhang, 2009]
       \item Low-treewidth graphs [Parikh et al., 2012]
@@
     \begin{itemize}
-      \item \tikzmark{tensoring} Tensoring
+      \item \tikzmark{tensoring} Tensor Product
         \begin{align*}
@@
       $\begin{bmatrix} \pi_1 \\ \pi_2 \\ \vdots \\ \pi_k \end{bmatrix}  
+        \ub{
        \begin{bmatrix} 
@@
                  &         &       &         
-       \end{bmatrix} $
+               \end{bmatrix}}_{B} $
       };
@@
   \begin{canvas}
-    \input{spectral-experts.tikz}
-    \node[style=txt] at (reg-label) {\alert<2>{regression}};
-    \node[style=txt] at (tf-label) {\alert<3>{tensor factorization}};
+    %\input{spectral-experts.tikz}
+    \spectralexpertsdiag
+    \node[style=txt] at (reg-label) {regression};
+    \node[style=txt] at (tf-label) {tensor factorization};
+
+    \point{reg-assumptions}{($(reg-label) - (0,1.0cm)$)};
+    \node<2->[style=txt] at (data3 |- reg-assumptions) {
+      {\bf Assumptions:}
+    };
+    \node<3->[style=txt] at (reg-assumptions) {
+        $\begin{aligned}
+          \hat\E[\vvec(x\tp{2})\tp{2}] &\succ 0\\
+          \hat\E[\vvec(x\tp{3})\tp{2}] &\succ 0.
+        \end{aligned}$
+        };
+    \node<4->[style=txt] at (reg-assumptions -| tf-label) {
+        $\begin{aligned}
+          \pi &\succ 0\\
+          rank(B) &= k \le d
+        \end{aligned}$
+        };
   \end{canvas}
@@
     \begin{align*}
-      \robustaltm<-2>{\hat M_2}{\mathmb{\hat M_3}} &= \arg\min_{M} \sum_{(x,y)\in\mathcal{D}} \left(y^2 - 
-      \robustaltm<-2>{\innerp{M}{x\tp{2}}}{\mathmb{\innerp{M}{x\tp{3}}}}
+      \robustaltm<-2>{\hat M_2}{\mathmb{\hat M_3}} &= \arg\min_{M} \sum_{(x,y)\in\mathcal{D}} \left( 
+      \robustaltm<-2>{y^2 - \innerp{M}{x\tp{2}}}{\mathmb{y^3 - \innerp{M}{x\tp{3}}}}
       - \textrm{bias}\right)^2 
@@
   \begin{canvas}
-    \input{spectral-experts.tikz}
+    \spectralexpertsdiag
     \node[style=txt] at (reg-label) {\robustalt<1>{\color{blue}\textbf{low-rank}}{low-rank}regression};
@@
       \includegraphics<4>[width=\textwidth,height=5cm,keepaspectratio]{figures/Spectral-1833.pdf}
-      \includegraphics<5->[width=\textwidth,height=5cm,keepaspectratio]{figures/Spectral-Spectral+EM-1833.pdf}
+      \includegraphics<5->[width=\textwidth,height=5cm,keepaspectratio]{figures/Spectral+EM-1833.pdf}
       };
@@
 
-    \draw<3>[latex-latex,DarkGreen,line width=1pt,dashed] ($(mle) + (-1.2cm,0.8cm)$) -- node[above]{$\mathmg{\epsilon}$} ($(mle) + (+1.2cm,0.8cm)$);
-    \node<4->[scale=0.7] at (spec) {x};
-    \node<4-> at ($(spec) + (0.5cm,0.3cm)$) {$\mathmg{\hat\theta_{\textrm{spec}}}$};
-    \draw<4->[dashed,DarkGreen,line width=0.7pt] ($(spec)-(0.5cm,0)$) -- ($(spec)+(3.5cm,0)$);
+%    \draw<3>[latex-latex,DarkGreen,line width=1pt,dashed] ($(mle) + (-1.2cm,0.8cm)$) -- node[above]{$\mathmg{\epsilon}$} ($(mle) + (+1.2cm,0.8cm)$);
+    \node<3->[scale=0.7] at (spec) {x};
+    \node<3-> at ($(spec) + (0.5cm,0.3cm)$) {$\mathmg{\hat\theta_{\textrm{spec}}}$};
+    \draw<3->[dashed,DarkGreen,line width=0.7pt] ($(spec)-(0.5cm,0)$) -- ($(spec)+(3.5cm,0)$);
 
-    \draw<5->[-latex,smooth,line width=1pt,DarkGreen] ($(spec) + (-0.1cm,+0.05cm)$) -- ($(mle) + (-0.30cm,+0.10cm)$);
-    \node<5->[scale=0.7] at (mle) {x};
-    \node<5->[anchor=west] at ($(mle) + (0.3cm,0)$) {$\mathmb{\hat\theta}_{\textrm{spec + EM}}$};
-    \draw<5->[dashed,blue,line width=0.7pt] ($(mle)-(0.4cm,0)$) -- ($(mle)+(0.4cm,0)$);
+    \draw<4->[-latex,smooth,line width=1pt,DarkGreen] ($(spec) + (-0.1cm,+0.05cm)$) -- ($(mle) + (-0.30cm,+0.10cm)$);
+    \node<4->[scale=0.7] at (mle) {x};
+    \node<4->[anchor=west] at ($(mle) + (0.3cm,0)$) {$\mathmb{\hat\theta}_{\textrm{spec + EM}}$};
+    \draw<4->[dashed,blue,line width=0.7pt] ($(mle)-(0.4cm,0)$) -- ($(mle)+(0.4cm,0)$);
   \end{tikzpicture}
@@
   \begin{itemize}
-    \item<1-> Expose tensor factorization structure through regression.
-    \item<2-> Consistent estimator for the mixture of linear regressions with polynomial sample and computational complexity.
-    \item<3-> Empirically, method of moment estimates can be a good initialization for EM.
-    \item<4-> How can we handle other discriminative models?
+    \item<+-> Consistent estimator for the mixture of linear regressions
+    \item<+-> Expose tensor factorization structure through regression.
+    \item<+-> {\bf Theory:} Polynomial sample and computational complexity.
+    \item<+-> {\bf Experiments:} Method of moment estimates can be a good initialization for EM.
+    \item<+-> {\bf Future Work:} How can we handle other discriminative models?
       \begin{itemize}
-          \item<5-> Dependencies between $h$ and $x$ (mixture of experts).
-          \item<6-> Non-linear link functions (logistic regression).
+          \item<+-> Dependencies between $h$ and $x$ (mixture of experts).
+          \item<+-> Non-linear link functions (hidden variable logistic regression).
       \end{itemize}
@@
 \end{document}
-
-% TikZ
-%\begin{canvas}
-%  \node (img1) at (0,0) {\includegraphics[height=0.6\textheight,width=0.4\linewidth,keepaspectratio]{}};
-%\end{canvas}
-
-% Notes 
-% \note[item]{}
-
-% 2-column
-% \begin{columns}
-%   \begin{column}{0.48\textwidth}
-%     \begin{itemize}
-%         \item
-%     \end{itemize}
-%   \end{column}
-%   \hfill
-%   \begin{column}{0.48\textwidth}
-%     \begin{itemize}
-%         \item
-%     \end{itemize}
-%   \end{column}
-% \end{columns}
-
-% TikZ
-%\begin{canvas}
-%  \node (img1) at (0,0) {\includegraphics[height=0.6\textheight,width=0.4\linewidth,keepaspectratio]{}};
-%\end{canvas}
-
-% Notes 
-% \note[item]{}
-
-% 2-column
-% \begin{columns}
-%   \begin{column}{0.48\textwidth}
-%     \begin{itemize}
-%         \item
-%     \end{itemize}
-%   \end{column}
-%   \hfill
-%   \begin{column}{0.48\textwidth}
-%     \begin{itemize}
-%         \item
-%     \end{itemize}
-%   \end{column}
-% \end{columns}

### 2013-06-15 doc/icml2013/presentation/render/talk-notes.md/paper.tex
Moved presentation to the icml2013 folder

\documentclass[tablecaption=bottom]{jmlr}

\usepackage[cm]{fullpage}
\usepackage{booktabs}
\usepackage{ctable}

\title{Spectral Experts}

\author{\Name{Arun Tejasvi Chaganty} \Email{chaganty@stanford.edu}}

\input{macros}

\begin{document}

\maketitle

\newcommand{\tp}

{[}1{]}\{\^{}\{\otimes \#1\}\}
\newcommand{\opX}{\mathfrak{X}}

\newcommand{\cvec}{\textrm{cvec}}

\section{Introduction}

\begin{itemize}
\item
  Latent variable models (HMM, mixture models, CRFs, etc.) are extremely
  powerful models of data we've developed.
\item
  However, learning parameters in these models is typically hard because
  their likelihood functions are not convex in the parameters.
\item
  Most common approach is to use local methods like EM, variational
  techniques, etc.
\item
  Local optima are a serious concern when using these methods.
\item
  Importantly, this is a problem that does not necessarily go away with
  more and more data!
\item
  Recently, there have been a number of \emph{consistent estimators}
  proposed based on the method of moments \todo{Include citations}.
  \begin{itemize}
  \item
    These are called ``spectral methods'' for how they utilize spectral
    decompositions to recover the parameters.
  \end{itemize}
\item
  Our work extends this approach of learning parameters to the
  discriminative setting wherein the moments of the parameters are not
  directly observed.
  \begin{itemize}
  \item
    The crux of our approach will be to use regression to first learn
    these moments, followed by application of tensor decomposition to
    learn the parameters.
  \end{itemize}
\end{itemize}
\section{Background}

\subsection{Method of Moments}

\begin{itemize}
\item
  Let us study how method of moment estimators work in general.
\item
  Consider a moment map $\mathcal{M}$ that maps the parameters $\theta$
  to the moments $m$. For a Gaussian, we have that
  $\mathcal{M} = (\mu, \sigma^2)$.
\item
  In general, we will compute the inverse of the moment map to learn the
  parameters from the sample estimates.
\item
  By the central limit theorem, our sample estimates of the moments
  converge at a $1/\sqrt{n}$ rate, so we expect that our parameters will
  also converge at this rate.
\end{itemize}
\subsection{Mixture of Regressions}

\begin{itemize}
\item
  The mixture of linear regressions model defines a conditional
  distribution over a response $y \in \Re$ given covariates
  $x \in \Re^d$.
\item
  The generative procedure is as follows,
  \begin{enumerate}[i)]
  \item
    Draw a mixture component $h \in [k] \sim Mult(\pi)$, where
    $\pi = [\pi_1 | ... | \pi_k]$ defines the mixture proportions.
  \item
    Draw the noise $\epsilon \sim \mathcal{E}$, where $\mathcal{E}$ is
    the noise distribution.
  \item
    Set $y = \beta_h^T x + \epsilon$, where $\{\beta_h\}_{h=1}^{k}$ are
    the conditional means of the regression coefficients.
  \end{enumerate}
\item
  The parameters that we would like to learn from this model are $\pi$
  and $B = [ \beta_1 | ... | \beta_k ]$.
\item
  The challenge in our scenario is that the moments of the data give us
  very filtered information of the parameters.
\end{itemize}
\section{Algorithm}

\subsection{Recovering the moments}

\begin{itemize}
\item
  As noted earlier, the first problem we run into is that we can't
  observe the moments of the parameters $B$ and $\pi$ directly!
\item
  However, observe that
  \begin{align}
    y &= \innerp{\beta_h}{x} + \epsilon \\
      &= \innerp{M_1}{x} + \underbrace{ {\beta_h - M_1}{x} + \epsilon }_{\textrm{noise}},
  \end{align}

  where $M_1 = \sum_{h=1}^k \pi_h \beta_h$, the mean regression
  coefficent. We note that while the noise term is dependent on $x$, it
  has a zero-mean. Thus, we could potentially recover $M_1$ through
  regression.
\item
  However, the first moments are insufficient to learn this model.
\item
  Let's look at the second and third moments of the data,
  \begin{align*}
    y^2 &= (\innerp{\beta_h}{x} + \epsilon)^2 \\
      &= \innerp{M_2}{x\tp{2}} + \E[\epsilon^2] +
       \underbrace{ \innerp{\beta_h\tp{2} - M_2}{x\tp{2}} + \innerp{\beta_h\tp{1} - M_1}{x\tp{1}} + (\epsilon^2 - \E[\epsilon^2]) }_{\textrm{noise}}, \\
    y^3 &= (\innerp{\beta_h}{x} + \epsilon)^3 \\
      &= \innerp{M_3}{x\tp{3}} + 3\E[\epsilon^2] \innerp{M_1}{x}  + \E[\epsilon^3] + {\textrm{noise}},
  \end{align*}
\item
  On it's own, $M_2$ is insufficient to identify the model, because it
  is invariant to rotations of $B$.
\item
  However, it turns out that $M_2$ and $M_3$ are sufficient to identify
  the model, via tensor decomposition, if $k < d$.
\item
  An additional fact that we can exploit is that both $M_2$ and $M_3$
  are low rank, so we can use low-rank regression to recover estimates
  $\hat M_2$ and $\hat M_3$ efficiently from data.
\end{itemize}
\subsubsection{Caveat: Requirements for Identifiability}

\begin{itemize}
\item
  In ordinary linear regression, the regression coefficients
  $\beta \in \Re^d$ are identifiable if and only if the data has full
  rank: $\E[x\tp{2}] \succ 0$.
\item
  However, because we need regression on higher moments to recover $M_2$
  and $M_3$, we also need that $\E[\cvec(x\tp{p})\tp{2}] \succ 0$ for
  $p \in \{1,2,3\}$.
\item
  This has some subtle implications when the features are completely
  independent of each other.
\item
  For example, if $x = (1, t, t^2)$, then $\E[\cvec(x\tp{2})\tp{2}]$ is
  singular for any data distribution.
\end{itemize}
\subsection{Recovering the parameters}

\begin{itemize}
\item
  Now that we've the moments, let us review how the tensor decomposition
  technique can be used to learn the parameters.
\item
  The method exploits the fact that $M_2$ and $M_3$ share a basis,
  \begin{align*}
      M_2 &= \sum_{h=1}^k \pi_h \beta_h\tp{2} \\
      M_3 &= \sum_{h=1}^k \pi_h \beta_h\tp{3}.
    \end{align*}
\item
  Decompositions are not unique however, but we can use the whitening
  transformation for $M_2$ to give \emph{whiten} $M_3$ such that it has
  a orthogonal decomposition.
  \begin{align*}
      I &= W^T M_2 W  \\
        &= \sum_{h=1}^k (\underbrace{\sqrt{\pi_h} W^T \beta_h}_{v_h})\tp{2} \\
      M_3(W,W,W) &= \sum_{h=1}^k \pi_h (W^T \beta_h)\tp{3} \\
                 &= \sum_{h=1}^k \frac{1}{\sqrt{\pi_h}} v_h\tp{3}.
    \end{align*}
\item
  The robust tensor power method by \cite{AnandkumarHsuGe2012} find
  stable eigenvectors using the following iterative algorithm,
  $v_h \to \frac{T(I, v_h, v_h)}{||T(I, v_h, v_h)||_2}.$
\end{itemize}
\section{Theorem: Rates of Recovery}

\begin{itemize}
\item
  The rate of convergence for the spectral experts algorithm to the true
  parameters breaks into two parts; the rates for learning the moments,
  which feeds into the rates for learning the parameters.
\item
  \todo{Diagram showing how the error breaks down.}
\item
  For low rank regression, we have the following bound on recovery by
  (Tomioka2011);
  \[ || \hat M_p - M_p ||_F \le \frac{32 \lambda^{(p)}_n \sqrt{k}}{\kappa(\opX_p)}, \]
  where $\kappa(\opX_p)$ is the (restricted) strong convexity constant,
  and $\lambda^{(p)} > ||\opX^*_p(\eta)||$.
\item
  Because we assume our noise is bounded, it is easy to show that the
  error concentrates.
\item
  In the tensor recovery case, we will need to whiten $M_3$ before
  applying the tensor decomposition and unwhiten it afterwards; this
  modifies the error bounds slightly.
\end{itemize}
\section{Spectral Experts in Practice}

\begin{itemize}
\item
  We simulated the performance of spectral experts and compared it to
  EM.
  \begin{itemize}
  \item
    We generated data from the mixtures of linear regressions model,
    with $x$ drawn uniformly in $[-1,1]^d$.
  \end{itemize}
\item
  In a non-trivial number of cases, EM did not converge to the right
  parameters and got stuck in local optima.
\item
  As a specific instance, we studied this example,
  \todo{Include diagram}.
  \begin{itemize}
  \item
    Only 13 of a 100 attempts with EM successfully identified the true
    parameters.
  \item
    On the other hand, even with $O(10^5)$ samples, the parameters by
    the spectral method weren't great.
  \item
    However, EM when initialized with these parameters did extremely
    well.
  \end{itemize}
\item
  This finding that spectral methods should be a good initialization for
  EM is not surprising.
  \begin{itemize}
  \item
    The biggest sell for spectral methods is that they give a global
    guarantee on where the parameters are.
  \item
    The parameters might not be at the global optima, but will hopefully
    lie in the potential well around it.
  \item
    EM will then converge to the global optima.
  \item
    \todo{Include diagram.}
  \end{itemize}
\item
  \todo{Summarize Other experiments.}
  \begin{itemize}
  \item
    Be frank, this isn't going to replace EM on it's own, but perhaps it
    highlights a principled approach to initializing EM.
  \end{itemize}
\end{itemize}
\section{Conclusions}

\begin{itemize}
\item
  We can learn the parameters of conditional models where the observed
  moments on their own have such sparse information.
\item
  The key intuition is that we could construct a regression problem to
  learn the moments of the parameters from their projections onto the
  data.
\item
  We found that while the parameters learned this way are usually not
  better than those learned via local methods like EM, etc., they are a
  good way to initialize these local methods.
\end{itemize}

\end{document}

### 2013-06-15 doc/icml2013/presentation/render/transcript.md/paper.tex
Moved presentation to the icml2013 folder

\documentclass[tablecaption=bottom]{jmlr}

\usepackage[cm]{fullpage}
\usepackage{booktabs}
\usepackage{ctable}

\title{Talk Transcript}

\author{\Name{Arun Tejasvi Chaganty} \Email{chaganty@stanford.edu}}

\input{macros}

\begin{document}

\maketitle

\section{Preamble (0:14 = 0:14) {[}0:15{]}}

Hi, I'm going to present Spectral Experts, a consistent estimator for
the mixture of linear regressions we developed. First, however, I'd like
to introduce the general context and the underlying goal for this work;
parameter estimation in discriminative latent variable models.

\section{Generative vs.~Discriminative Latent Variable Models (1:20 =
1:34) {[}1:00{]}}

Latent variable models are very powerful tools for inferential tasks in
machine learning. We can categorize these models into two broad classes;
generative and discriminative.

In the first case of generative models, we describe the generative
distribution for the observed variables, $x$, conditioned on some latent
variables, $h$.

Gaussian mixture models, hidden Markov models, and so on are examples of
this class.

In contrast, in discriminative models the generative distribution for
some observed variables, for example the class labels, $y$, are
described conditioned not only on the latent variables, but also the
other observed variables, $x$.

Latent CRFs, discriminative LDA, the mixture of experts, etc. are
examples in this class.

There has been a lot of prior work by the spectral learning community in
addressing consistent parameter estimation for the generative family
(citations?), but not much work has been done in the discriminative
setting.

We think this is important because discriminative models are often
preferred in practice because they make it easier to include arbitrary
features and tend to be more accurate.

\{Highlight box\}

\section{Parameter estimation is hard (1:34 = 3:06) {[}1:20{]}}

\{Draw axes in the cartoon\}

Parameter estimation for latent variable models is in general hard. One
of the main culprits is that the log-likelihood functions in these
models are non-convex.

The maximum likelihood estimator will get us the true parameters,
$\theta^*$ in the limit of infinite data. In fact, statistically, it is
the best estimator for a given amount of data, but it is in general
intractable.

In practice, we might use a local estimator, like EM or a gradient based
method. This is computationally tractable, but can be very dependent on
our initialization. With the right initialization, we might find the
true parameters, but for some other initialization we could get stuck in
local optima, like the $\hat \theta$ shown here. Such local optima often
occur in practice.

Our goal, like that many other spectral methods, is to develop an
algorithm with polynomial sample and computational complexity, and we
will do so using the method of moments. In the limit of infinite data,
we are guaranteed to recover the true parameters. Moreover, with finite
samples, we can still guarantee our estimates to be within $\epsilon$ of
the true parameters.

\section{Method of Moments / Related Work (0:54 = 4:10) {[}0:45{]}}

The method of moments itself has a rich history. It was first introduced
by Pearson in 1894, and found applications in control theory.

Closer to home, there have been several applications in the observable
operator space, for example, predictive state representations, and more
recently for the hidden Markov model by Daniel Hsu et al. in 2009.

We build on a line of work that is focussed on parameter estimation. In
particular, we are influenced by the treatment of multi-view mixture
models, by Anandkumar et al., which was published recently in 2012. I'd
like to review the tensor factorization technique used by Anandkumar et
al. which can be used to recover parameters for the multi-view mixture
model; we'll shortly see how it can be applied to our discriminative
model.

\section{Aside: Tensor Operations (0:54 = 5:06) {[}0:30{]}}

Before I can begin, I'd like to introduce some notation for the tensor
operations I'll be using. I'll use $\otimes$ to refer to the outer or
tensor product. $x\tp{3}$, which I'll just call $x$ cubed, is a tensor
whose $(i,j,k)$-th component is $x_i x_j x_k$. Next, I'll use the
angular brackets for a generalized inner product. Finally, I'll consider
projections $A(u)$ for the matrix formed by summing out the third-index
dot-producted with the vector $u$. We can also project along two axes of
the tensor to get a vector.

\section{Example: Multi-view Gaussian Mixture Model (2:44 = 7:50)
{[}2:00{]}}

\{Use a two-column format instead\} Alright, let's consider the
generative process for the multi-view mixture, with the assumption that
$k \le d$. We have a single latent variable, $h$, which chooses the
mixture component, (one of these three). For each mixture component,
we'll generate three views, $x_1, x_2, x_3$ that, for simplicity, are
identically distributed as Gaussian with mean $\beta_h$.

Let's write out the moments for this model. Note that the conditional
means are the $\beta_h$ by definition. Now the first moment of the data
is the weighted average of the component means. Clearly just observing
this is insufficient to identify our parameters.

Next, we'll look at the cross moments of $x_1$ and $x_2$; the cross
moments lets us ignore variance terms since $x_1$ and $x_2$ are
independent, and hence this is just composed of $K$ rank-one matrices.

The second moments aren't also insufficient to identify the parameters
though; an orthogonal transformation of the $\beta_h$ would give us the
same moments. Finally, let's look at the third order moments for this
model; it's similarly composed of $K$ rank-one third-order tensors.

\{Swap image and eigenvector notation\} The key idea is that if
component means $\beta_h$ were orthogonal, then they are
\emph{eigenvectors} of the third order tensor; if we projected this
tensor along $\beta_h$ twice, we get $\pi_h \beta_h$. We can use the
tensor power method to recover the eigenvectors from the third moments.

More generally, we can use the second order moments to whiten the third
order tensor; for details, I refer you to a prepint by Anandkumar, Hsu,
et al. (2013).

\section{Interlude: Discriminative vs.~Generative (0:25 = 8:15)
{[}0:20{]}}

\{Improve slide\}

In the beginning of this talk, I described two broad classes of latent
variable models, generative and discriminative, and motivated why we
would like to develop a consistent estimator for the latter class. I
will now introduce the particular instance that we developed an
algorithm for, the mixture of linear regressions.

\section{Mixture of Linear of Regressions (1:00 = 9:15) {[}1:00{]}}

Again, let's start with the generative process for this model. We'll
assume that the $h$ and $x$ are independent of each other. The mixture
of linear regressions models our data as a mixtures of lines plus noise.

For any given $x$, we'll select a single line based on the latent
variable $h$. Then, we'll observe the coordinate $y$ along the line at
that $x$, but maybe with some noise. For technical reasons, we'll assume
that the noise is bounded.

For another $x$, we pick a line at random again, and then observe the
coordinate $y$ with some noise. We can repeat this process till we get a
sizable dataset.

Our objective is then, given this data, can we develop a procedure to
recover the parameters; the mixture probabilities, $\pi$, and the set of
regression coefficients, the slopes of the lines
$B = [\beta_1| ... | \beta_h]$?

\section{Method of Moments for the Mixture of Linear Regressions (2:35 =
11:50) {[}2:30-3:00{]}}

Our plan is to reduce the problem to the method of moments, where we
observed the moments of the data. Can we get something similar?

\{Include the messy moments picture?\}

Let's meditate a bit on the generative process. This process looks
awfully like conventional regression, but we have to be careful; the
$\beta_h$ are actually random variables. It's not immediately obvious
what we would get if did do regression here.

So, to make it obvious, let's add and subtract the expected value of the
$\beta$s, with respect to $h$. Now, we have two terms; a constant vector
dot the data plus some zero-mean term, which we can consider to be
``noise''. What this tells us is that if we did regression here, we
would get the expected value of the regression coefficients.

Can we generalize this process to get the expected value of higher
powers of the parameters $\beta_h$?

Here, I've replaced the inner product with this angular bracket
notation, which, remember, was the generalized inner product.

\{Include manipulation?\}

Now, if we wrote out the expression for the second power of $y$, and did
a little bit of arithmetic manipulation, we're going to see the same
structure; an inner product of the expected value of the second power of
the $\beta$s and the second power of the data $x$, plus some noise and
this time, we also have some known bias. This means we can use
regression here to get the second powers of $\beta$s.

We can extend this approach to the third power and so on.

\{Include this structure picture on a slide\}

Now, let's look at the structure of the third power of $\beta$; it's
identical to the third moment of the data in the multi-view mixture
case. That means we can use the tensor factorization algorithm here too!

\section{Algorithm overview (1:15 = 13:05) {[}1:30{]}}

\{Move to the algorithm overview\}

So that's the rough skeleton of the algorithm, regression to get these
powers and tensor factorization to get the parameters from there.

\{Talk about the assumptions\}

For regression to work, we need the assumption that $\E[\beta_h\tp{p}]$
has rank at least $k$, and for the tensor factorization step, we need
that $k < d$ and that $B$ is full rank.

\{Return to the regression picture\}

We have a statistically consistent estimator now, since each step is
consistent, but doing vanilla regression on these higher powers could
potentially be very statistically inefficient. We know that we need
$O(d^3)$ samples to do regression, but we should be able to exploit the
fact that even for $\beta_h\tp{3}$, there are only $kd + k$ degrees of
freedom; we know that these matrices and tensors have a rank of only
$k$. To exploit this, we can use low rank-regression, which adds a
nuclear norm regularizer to the optimization objective.

\{return to algorithm overview\}

This is our complete algorithm. As you can see, each part has a
polynomial computational complexity, so we have a polynomial
computational complexity.

\section{Perturbation Analysis (1:00 = 14:05) {[}1:00{]}}

\{Fix horrendous picture\} Now, I'll briefly talk about the sample
complexity of our algorithm. We adapt results for low-rank regression
from Negahban and Wainwright (2009) and Tomioka (2012) to our noise
setting. We can show that we need a number of samples polynomial in all
our parameters to get estimates of the power of $\beta_h$ with an
$\epsilon$ error.

Next, we apply the analysis of the tensor factorization algorithm to
give us bounds on the recover of the parameters in terms of error of the
inputs, $\beta_h^3$. This step also requires just polynomial number of
samples to get the parameters within $\epsilon$ of the true value.

Thus, our algorithm has a polynomial sample complexity as well.

\section{Experimental insights (2:45 = 16:50) {[}2:30{]}}

\{Fix layout\} Finally, I'd like to talk about some insights we found on
simulated experiments. We can consider featurizations of our input data;
so in this example, where with data $x$ and response $y$, we consider a
polynomial featurization that gave use these curves.

The dimension of our data $d$ is $4$, and the number of components $k$
is $3$. For our experiments, we used $10,000$ examples generated from
the model.

When we ran EM with random initializations on this algorithm, we found
that it easily got stuck in a local optima. This in fact occurred on a
large percentage of the random initializations we tried. The histogram
counts the number of initializations out of 200 random initializations
which achieve a particular parameter error. In this example, only about
13\% actually converged to the true parameters.

Next, we tried the spectral method. We didn't exactly recover the
parameters, but we are operating in the finite sample regime and this
can occur. Importantly though, it seems like the parameters we recover
sufficiently separate the different mixture components. We tried EM
initialized with these parameters, and find that we recover the true
parameters with much higher frequency.

\{Include table\} We ran similar experiments on a number of simulated
datasets with a different number of dimensions and components and
generally observed this trend. I've presented the parameter errors
averaged over 10 different datasets within each configuration of $d$ and
$k$, and over 10 initializations for each dataset.

\{Include parameter annotations\} These observations fit with the
intuition painted by the log-likelihood picture we began with. Though
the log-likelihood function is non-convex, the method of moments
estimator guarantees that we'll recover a parameter estimate within
$\epsilon$ of the true parameters. This might have a lower likelihood
than EM, but if we can get $\epsilon$ small enough, it can be in the
right ball park of the true parameters that running EM initialized with
these parameters will converge to the true parameters.

\section{Conclusions (0:45 = 17:35) {[}0:45{]}}

In conclusion, I've presented a consistent estimator we developed for
the mixture of linear regressions with a polynomial sample and
computational complexity.

The key idea was that we can use regression to recover the powers of
$\beta_h$ which could then be factorized using the tensor factorization
technique.

Finally, our experiments support the intuition that in the finite sample
regime, method of moment estimators can be good initialization for EM.

\section{Thank you.}

\section{Question: Isn't it better to run EM with many random
initializations then use your method?}

We found that the number of initializations that led to the true
parameters were fairly small even on our toy data; it's unclear how many
initializations you would need to reasonable find good parameters for
larger settings. The spectral method is pretty efficient and in our
experiments, took on the same order of time as running full EM, so I
don't think there is a downside to using the method, provided you have
enough sufficient data.

\section{Question: In principle your method is supposed to converge.
Does it?}

The convergence results we present only hold when we have sufficient
samples for some conditions to hold; these conditions depend on the
condition number of the powers of $\beta$. On experiments without any
featurization, the condition numbers were usually quite reasonable and
we observed convergence. On the polynomial featurizations we
experimented with, we found the condition numbers to be much worse and
it appeared we required more data than we could run our code for to
observe convergence.

\section{Question: Isn't it of concern that your method doesn't scale
well for very large $d$?}

The primary regime for the method is when $k << d$ but making this
method work better with larger $d$ is an important next step.

\end{document}

### 2013-06-15 doc/icml2013/presentation/data/1833/plotLines.py

@@ -1,0 +1,139 @@
+import numpy as np
+import operator
+import matplotlib
+#matplotlib.use("Agg")
+
+import matplotlib.pyplot as plt
+
+def parseBetasFile(fname):
+  betas = []
+
+  beta = []
+  for line in open(fname).xreadlines():
+    if line.startswith("betas"):
+      if( len(beta) > 0 ):
+        betas.append(np.array(beta))
+        beta = []
+    else:
+      try:
+        beta.append( map(float, line.split()) )
+      except Exception as e:
+        print "[Warning]: Couldn't parse line", line
+        print e
+        continue
+  if( len(beta) > 0 ):
+    betas.append(np.array(beta))
+  return betas
+
+def fs153(x):
+    return np.array([1.0, x, x**4, x**7])
+
+def getLine( beta, fn, W ):
+    """Gets a surface with X, Y, Z = f(X,Y) to be plotted"""
+    X = np.linspace( -1, 1, W)
+    # Grid of X's
+    Y = map(lambda (x): beta.dot(fn(x)), X)
+
+    return X, Y
+
+def plotLines(ax, betas, modifier = "-", colors = None, W = 100):
+    if colors is None:
+        colors = ['b','r','g']
+    colors = [ color + modifier for color in colors ]
+    lines = []
+    for beta, color in zip(betas.T, colors):
+        X, Y = getLine( beta, fs153, W )
+        lines += ax.plot(X, Y, color, linewidth=5)
+    return lines
+
+def makePlot( data, trueBetas, specBetas, spemBetas, em0Betas, emBetas, outFile ):
+
+    plt.rc("font",size=20)
+
+    for setting in [ [], ["EM"],["Spectral"],["Spectral+EM"],["Spectral","Spectral+EM"]]:
+        fig = plt.figure()
+        ax = plt.axes(xlim=(-1, 1), ylim=(-3, 5))
+        ax.set_xlabel("t")
+        ax.set_ylabel("y")
+
+        # Plot the data
+        if( len(setting) == 0 ):
+            ax.scatter( data[0], data[1], alpha = 0.4 )
+        else:
+            ax.scatter( data[0], data[1], alpha = 0.4 )
+
+        if "True" in setting:
+            # Now draw each line
+            plotLines( ax, trueBetas, "-" )
+
+        # Legend 
+        plottables = []
+        # Spectral is dashed
+        if( "Spectral" in setting ):
[…]

### 2013-06-16 doc/icml-presentation/data/1833/plotLines.py
Moved icml-presentation to icml2013/presentation

@@ -1,139 +1,0 @@
-import numpy as np
-import operator
-import matplotlib
-#matplotlib.use("Agg")
-
-import matplotlib.pyplot as plt
-
-def parseBetasFile(fname):
-  betas = []
-
-  beta = []
-  for line in open(fname).xreadlines():
-    if line.startswith("betas"):
-      if( len(beta) > 0 ):
-        betas.append(np.array(beta))
-        beta = []
-    else:
-      try:
-        beta.append( map(float, line.split()) )
-      except Exception as e:
-        print "[Warning]: Couldn't parse line", line
-        print e
-        continue
-  if( len(beta) > 0 ):
-    betas.append(np.array(beta))
-  return betas
-
-def fs153(x):
-    return np.array([1.0, x, x**4, x**7])
-
-def getLine( beta, fn, W ):
-    """Gets a surface with X, Y, Z = f(X,Y) to be plotted"""
-    X = np.linspace( -1, 1, W)
-    # Grid of X's
-    Y = map(lambda (x): beta.dot(fn(x)), X)
-
-    return X, Y
-
-def plotLines(ax, betas, modifier = "-", colors = None, W = 100):
-    if colors is None:
-        colors = ['b','r','g']
-    colors = [ color + modifier for color in colors ]
-    lines = []
-    for beta, color in zip(betas.T, colors):
-        X, Y = getLine( beta, fs153, W )
-        lines += ax.plot(X, Y, color, linewidth=5)
-    return lines
-
-def makePlot( data, trueBetas, specBetas, spemBetas, em0Betas, emBetas, outFile ):
-
-    plt.rc("font",size=20)
-
-    for setting in [ [], ["EM"],["Spectral"],["Spectral+EM"],["Spectral","Spectral+EM"]]:
-        fig = plt.figure()
-        ax = plt.axes(xlim=(-1, 1), ylim=(-3, 5))
-        ax.set_xlabel("t")
-        ax.set_ylabel("y")
-
-        # Plot the data
-        if( len(setting) == 0 ):
-            ax.scatter( data[0], data[1], alpha = 0.4 )
-        else:
-            ax.scatter( data[0], data[1], alpha = 0.4 )
-
-        if "True" in setting:
-            # Now draw each line
-            plotLines( ax, trueBetas, "-" )
-
-        # Legend 
-        plottables = []
-        # Spectral is dashed
-        if( "Spectral" in setting ):
[…]

### 2013-06-16 doc/icml2013/presentation/data/1833/plotLines.py

@@ -43,7 +43,7 @@ def plotLines(ax, betas, modifier = "-", colors = None, W = 100):
     lines = []
     for beta, color in zip(betas.T, colors):
         X, Y = getLine( beta, fs153, W )
-        lines += ax.plot(X, Y, color, linewidth=5)
+        lines += ax.plot(X, Y, color, linewidth=4)
     return lines
 
 def makePlot( data, trueBetas, specBetas, spemBetas, em0Betas, emBetas, outFile ):
@@ -84,15 +84,9 @@ def makePlot( data, trueBetas, specBetas, spemBetas, em0Betas, emBetas, outFile 
                 lines = plotLines( ax, emBeta, "-", colors=['r']*3 )
             plottables.append( lines[0] )
 
-        #for plot in plottables:
-        #    plot.set_color( "black" )
-
         if( len(plottables) > 0):
-            legend = plt.legend( plottables[::-1], setting[::-1] )
-        #legend = plt.legend( plottables, ["Spectral", "Spectral + EM", "EM"] )
-        #legend = plt.legend( plottables, ["EM"] )
-        #legend = plt.legend( plottables, ["Spectral"] )
-        #legend = plt.legend( plottables, ["Spectral", "Spectral + EM"] )
+            plt.legend( plottables, setting )
+            #legend = plt.legend( plottables[::-1], setting[::-1] )
 
 
         prefix = "-".join(setting + [outFile])

### 2013-06-16 doc/icml2013/poster/column1.tex
Added poster

\begin{block}{Parameter Estimation in Latent Variable Models}
  \begin{center}
  \begin{tikzpicture}[scale=2.0]
    \drawgen{(0,0)};
    \node[below=0.1cm of x.south] {Generative Models};
    \drawdisc{(4,0)};
    \node[below=0.1cm of y.south] {Discriminative Models};
  \end{tikzpicture}
  \end{center}
  \vspace{-2ex}
  \begin{itemize}
    \item Latent variable models (LVMs) are hard to learn because latent
      variables introduce non-convexities in the log-likelihood
      function.
    \item In practice, local methods (EM, gradient descent, etc.) are
      employed, but these can stuck in local optima.
    \item {\bf Can we develop efficient consistent estimators for discriminative latent variable models?}
      \begin{itemize}
        \item Why discriminative LVMs? Easy to add features, often more accurate.
        \item The method of moments has been used for consistent parameter estimation in several generative LVMs, e.g. HMMs${}^1$, LDA${}^1$, and stochastic block models${}^2$.
        \item Can we extend these techniques to discriminative LVMs?
      \end{itemize}
    \item {\bf Main result}: Consistent estimator for a simple discriminative model; the mixture of linear regressions.
      \begin{itemize}
        \item {\bf Key Idea:} Expose tensor factorization structure using regression.
        \item {\bf Theory:} We prove polynomial sample and computational complexity.
      \end{itemize}
  \end{itemize}              
  \footnotesize{%
    \hfill
    [1] Anandkumar, Hsu, Kakade, 2012; [2] Anandkumar, Ge, Hsu, Kakade, 2012
    }
\end{block}

%\vfill

\begin{block}{Aside: Tensor Operations}
  \vspace{-3ex}
  \splitcolumn{%
    \begin{itemize}
      \item Tensor Product
        \begin{align*}
          x\tp{3}       &= x \otimes x \otimes x \\
          x\tp{3}_{ijk} &= x_i x_j x_k
        \end{align*}
      \item Inner product
        \begin{align*}
          \innerp{A}{B} &= \sum_{ijk} A_{ijk} B_{ijk} \\
            &= \innerp{\vvec{A}}{\vvec{B}} 
        \end{align*}
    \end{itemize}
  }{%
  \begin{tikzpicture}[scale=1.3]
    \point{tensoring-pos}{(0,0)}
    % Tensoring
    \tikzcube{tensoring}{black,fill=white}{($(tensoring-pos) + (0cm,-1.0cm) + (0,0,0)$)}{1}{1}{1};
    \node at ($(tensoring) + (1.0cm,-0.3cm)$) {$=$};
    \tikzcube{v1}{black,fill=white}{($(tensoring) + (2.5cm,-0.3cm)$)}{1}{0.3}{0.3};
    \node (lbl1) at ($(v1) + (0.40cm,-0.1cm)$) {$\times$};
    \tikzcube{v2}{black,fill=white}{($(v1) + (1cm,0.3cm)$)}{0.3}{1}{0.3};
    \node at ($(lbl1) + (1cm,0.0cm)$) {$\times$};
    \tikzcube{v3}{black,fill=white}{($(v2) + (1cm,-0.3cm)$)}{0.3}{0.3}{1};

    % Inner product
    \point{ptA}{(0cm,-3.0cm)}
    \point{ptB}{($(ptA) + (2cm,0)$)}

    \tikzcube{innerpA}{black,fill=yellow}{($(ptA) + (0,0,0)$)}{1}{1}{1};
    \tikzcube{innerpB}{black,fill=blue!70}{($(ptB) + (0,0,0)$)}{1}{1}{1};

    \innerpdiagt{(ptA)}{(ptB)};
    \node[scale=1.0] at ($(ptB) + (1.5cm,-0.3cm)$) {$= \mathmg{0.5}$};

    % Inner product 2
    \point{ptvA}{($(ptA) + (0,-2cm)$)}
    \point{ptvB}{($(ptB) + (0,-2cm)$)}

    \tikzrect{innerpAv}{black,fill=yellow}{($(ptvA) + (0,0.5cm)$)}{0.3}{2.0};
    \tikzrect{innerpBv}{black,fill=blue!70}{($(ptvB) + (0,0.5cm)$)}{0.3}{2.0};

    \innerpdiagt{(ptvA)}{(ptvB)};
    \node[scale=1.0] at ($(ptvB) + (1.5cm,-0.3cm)$) {$= \mathmg{0.5}$};
  \end{tikzpicture}
  }
\end{block}

%\vfill

\begin{block}{Mixture of Linear Regressions}
  \begin{columns}[onlytextwidth]
    \begin{column}{0.55\textwidth}
      \begin{tikzpicture}
        \point{mark}{(0,0)}
        \drawdisc{(5cm,0)}
      \end{tikzpicture}
      \begin{itemize}
        \item For a particular $x$, we draw $y$ as follows,
        \begin{itemize}
          \item $h \sim \Mult([\pi_1, \pi_2, \cdots, \pi_k])$.
          \item $y = \beta_h^T x + \epsilon$.
        \end{itemize}
      \end{itemize}
    \end{column}
    \hfill
    \begin{column}{0.45\textwidth}
        \includegraphics[width=7cm,height=8cm,keepaspectratio]{figures/mlr-5.pdf}
    \end{column}
  \end{columns}

  \begin{itemize}
    \item Given $\mathcal{D} = \{(x_i,y_i)\}_{i=1}^n$, we want to recover the parameters $\pi$ and $B$.
  \begin{tikzpicture}[scale=0.7]
    \node[anchor=east] (data) at (-2cm,0) {%
    \includegraphics[width=6cm,height=7cm,keepaspectratio]{figures/mlr-6.pdf}
    };

    % x, y
    \node[anchor=west,scale=1.0] (params) at (1.5cm,0) {%
      $\begin{bmatrix} \pi_1 \\ \pi_2 \\ \vdots \\ \pi_k \end{bmatrix}  
        \ub{
       \begin{bmatrix} 
                 &         &       &         \\
                 &         &       &         \\
         \beta_1 & \beta_2 & \dots & \beta_k \\
                 &         &       &         \\
                 &         &       &         
               \end{bmatrix}}_{B} $
      };
    %\node[anchor=west,right=0.1cm of params] {%
    %\includegraphics[width=4cm,height=2cm,keepaspectratio]{figures/disc.png}
    %};

    \draw[-latex] (data) -- node[above]{?} (params);
  \end{tikzpicture}

  \item Our approach uses low-rank regression to reduce the problem to tensor eigendecomposition. 
  \begin{tikzpicture}[scale=1.8]
    \spectralexpertsdiag
    \node[style=txt] at (reg-label) {low-rank regression};
    \node[style=txt] at (tf-label) {tensor factorization};

    %\point{reg-assumptions}{($(reg-label) - (0cm,1.0cm)$)};
    %\node[style=txt] (assumptions) at (data3 |- reg-assumptions) {%
    %  %{\bf Assumptions:}
    %};
    %\node[style=txt] at (reg-assumptions) {%
    %    %$\begin{aligned}
    %    %  \hat\E[\vvec(x\tp{2})\tp{2}] &\succ 0\\
    %    %  \hat\E[\vvec(x\tp{3})\tp{2}] &\succ 0.
    %    %\end{aligned}$
    %    };
    %\node[style=txt] (tf-assumptions) at ($(reg-assumptions.west) + (1em,0)$) {%
    %    %$\begin{aligned}
    %    %  \pi &\succ 0\\
    %    %  rank(B) &= k \le d
    %    %\end{aligned}$
    %    };
    \node[style=txt,below=0.1em of reg-label] (reg-compl) {%
      $O\left( k\ 
      \hlmath{blue}{\|x\|^{12}}
      \hlmath{DarkGreen}{\|\beta\|^{6}}
      \hlmath{red}{\|\E[\epsilon^2]\|^{6}}
      \right)$};
    \node[style=txt,below=0.1em of tf-label] {%
    $O\left( \frac{k \pi_{\max}^2}{\sigma_k(M_2)^5} \right)$};

    \node[scale=0.8,style=txt, left=0.1em of reg-compl] {%
      {\bf Sample Complexity:}
    };

  \end{tikzpicture}
  \end{itemize}

  \vspace{-1ex}

\end{block}

### 2013-06-16 doc/icml2013/poster/column2.tex
Added poster

\begin{block}{Step 1: Finding Tensor Structure via Regression}
  \begin{itemize}
    \item {\bf Key Observation:} Regression on the powers of
        $(y,x)$ gives us the expected powers of the regression
        coefficients $\beta$. 
    \begin{columns}
      \begin{column}{0.65\textwidth}
  \begin{align*}
    y\tikzmark{regA} 
      &= \innerpp{\beta_h}{x} + \epsilon \\
      &= \mathmb{\ub{\innerp{\E[\beta_h]}{x}}_{\textrm{linear measurement}}} &&+ \mathmr{\ub{(\beta_h - \E[\beta_h])^T x + \epsilon}_{\textrm{noise}}} \\
    y^2 \tikzmark{regB}
      &= \left(\innerp{\beta_h}{x} + \epsilon\right)^2 && \\
      &= \mathmb{\innerpp{\ub{\E[\beta_h\tp{2}]}_{M_2}}{x\tp{2}}}
      &&+ \mathmg{\textrm{bias}_2} + \mathmr{\textrm{noise}_2} \\
    y^3 \tikzmark{regC}
    &= \mathmb{\innerpp{\ub{\E[\beta_h\tp{3}]}_{M_3}}{x\tp{3}}} &&+ \mathmg{\textrm{bias}_3} + \mathmr{\textrm{noise}_3}
  \end{align*}
    \end{column}
      \begin{column}{0.25\textwidth}
  \begin{tikzpicture}
    \point{mark}{(0,0)};
      \regressionA{($(mark) + (0,0cm)$)};
      \regressionB{($(mark) + (0,-2.5cm)$)};
      \regressionC{($(mark) + (0,-4.5cm)$)};
  \end{tikzpicture}
    \end{column}
  \end{columns}

\item $M_2$ and $M_3$ are both of rank $k$, so we can use low rank regression$^{3,4}$!
  \begin{align*}
    \hat M_2 
      &= \arg\min_{M} \sum_{(x,y)\in\mathcal{D}} 
        \left( y^2 - \innerp{M}{x\tp{2}} - \textrm{bias}_2\right)^2 
          + \lambda_2 \mathmb{\ub{\|M\|_{*}}_{\sum_i \sigma_i(M)}}\\
    \hat M_3 
      &= \arg\min_{M} \sum_{(x,y)\in\mathcal{D}} 
        \left( y^3 - \innerp{M}{x\tp{3}} - \textrm{bias}_3\right)^2 
        + \lambda_3 \|M\|_{*}
  \end{align*}
  \end{itemize}

  \footnotesize{%
  \hfill
    [3] Fazel, 2002; [4] Tomoika, Hayashi and Kashima, 2010
  }

\end{block}

\begin{block}{Step 2: Parameter Recovery via Tensor Factorization}
  \begin{itemize}
      \item $M_3$ has a low-rank tensor decomposition:
      $M_3 = \sum_{h=1}^k \pi_h \beta_h\tp{3}$
      \vspace{1ex}\\
      \begin{tikzpicture}[scale=1.5]
          \tensorfactorization{(0cm,0cm)};
      \end{tikzpicture}

    \item {\bf Key Observation:} If $\beta_h$ are orthogonal, they are eigenvectors$^{5}$; 
      $M_3(\beta_h,\beta_h) = \pi_{h} \beta_{h}$.
    \item In general, we can whiten $M_3$ first.
  \end{itemize}
  \footnotesize{%
  \hfill
  [5]: Anandkumar, Ge, Hsu, Kakade, Telgarksy, 2012.
  }

\end{block}

\begin{block}{Experiments}
  \begin{itemize}
      \item With finite samples, Spectral Experts seems to find parameters
        that sufficiently separate components that EM initialized with
        these parameters recovers true parameters more often than EM with random initializations.
  \item In this example, $y = \beta^T [ 1, t, t^4 ,t^7 ]^T + \epsilon$. $k = 3, d = 4, n = 10^5$,
  \begin{tikzpicture}
    \node (graphs) {%
      \includegraphics[width=0.3\textwidth,height=5cm,keepaspectratio]{figures/EM-1833.pdf}
      \includegraphics[width=0.3\textwidth,height=5cm,keepaspectratio]{figures/Spectral-Spectral+EM-1833.pdf}
      \includegraphics[width=0.3\textwidth,height=5cm,keepaspectratio]{figures/EM-Spectral-Spectral+EM-hist.pdf}
      };
  \end{tikzpicture}
  \item Below are parameter errors averaged over 10 initializations on
    10 different simulated datasets with the specified parameter
    configurations,
    \begin{tikzpicture}
      \node (exp1) {\includegraphics[width=0.2\textwidth]{figures/err-hist-0.pdf}};
      \node[scale=0.8,below=-0.2cm of exp1] {$d=4, k=2$};
      \node[right=0cm of exp1] (exp2) {\includegraphics[width=0.2\textwidth]{figures/err-hist-1.pdf}};
      \node[scale=0.8,below=-0.2cm of exp2] {$d=5, k=2$};
      \node[right=0cm of exp2] (exp3) {\includegraphics[width=0.2\textwidth]{figures/err-hist-2.pdf}};
      \node[scale=0.8,below=-0.2cm of exp3] {$d=5, k=3$};
      \node[right=0cm of exp3] (exp4) {\includegraphics[width=0.2\textwidth]{figures/err-hist-3.pdf}};
      \node[scale=0.8,below=-0.2cm of exp4] {$d=6, k=2$};
    \end{tikzpicture}

  \item {\bf Log-likelihood cartoon:} It seems that our parameter estimates fall in the right basin of attraction for EM.\\
      \begin{tikzpicture}
        \point{mark}{(0,0)}
        % x, y
        \llhood{4cm}{0};

        \node[scale=0.3,circle,fill=red] at (em2) {};
        \node at ($(em2) + (0.7cm,0)$) {$\mathmr{\hat\theta_{\textrm{EM}}}$};
        \draw[-latex,smooth,line width=1pt,red] ($(em2-start) + (+0.1cm,+0.05cm)$) -- ($(em2) + (+0.15cm,0.00cm)$);
        %\draw[dashed,red,line width=0.7pt] ($(em2)-(3.5cm,0)$) -- ($(em2)+(0.5cm,0)$);

    %    \draw<3>[latex-latex,DarkGreen,line width=1pt,dashed] ($(mle) + (-1.2cm,0.8cm)$) -- node[above]{$\mathmg{\epsilon}$} ($(mle) + (+1.2cm,0.8cm)$);
        \node[scale=0.3,circle,fill=DarkGreen] at (spec) {};
        \node at ($(spec) + (0.6cm,0.3cm)$) {$\mathmg{\hat\theta_{\textrm{spec}}}$};
        %\draw[dashed,DarkGreen,line width=0.7pt] ($(spec)-(0.5cm,0)$) -- ($(spec)+(3.5cm,0)$);

        \draw[-latex,smooth,line width=1pt,DarkGreen] ($(spec) + (-0.1cm,+0.05cm)$) -- ($(mle) + (-0.30cm,+0.10cm)$);
        \node[scale=0.3,circle,fill=blue] at (mle) {};
        \node[anchor=west] at ($(mle) + (0.2cm,0)$) {$\mathmb{\hat\theta}_{\textrm{spec + EM}}$};

        %\draw[dashed,blue,line width=0.7pt] ($(mle)-(0.4cm,0)$) -- ($(mle)+(0.4cm,0)$);
      \end{tikzpicture}
  \end{itemize}
  \vspace{-1ex}

\end{block}

\begin{block}{Future Work}
  \begin{itemize}
    \item How can we handle other discriminative models?
      \begin{itemize}
          \item Non-linear link functions (hidden variable logistic regression).
          \item Dependencies between $h$ and $x$ (mixture of experts).
      \end{itemize}
  \end{itemize}

\end{block}

\vfill

### 2013-06-16 doc/icml2013/evaluation.tex
Added new plots

@@
   \centering
-  \includegraphics[width=0.50\textwidth]{figures/hist.png}
+  %\includegraphics[width=0.50\textwidth]{figures/hist.png}
+  \includegraphics[width=0.50\textwidth]{figures/EM-Spectral-Spectral+EM-hist.pdf}
   \caption{Histogram over recovery errors between the three algorithms when $b = 1, d = 4, k = 3, n = 500,000$.}
@@
   \subfigure[Spectral, Spectral+EM]{
-    \includegraphics[width=0.50\textwidth]{figures/curves-vis/1-8-3-specm.png}}
+    %\includegraphics[width=0.50\textwidth]{figures/curves-vis/1-8-3-specm.png}}
+    \includegraphics[width=0.50\textwidth]{figures/Spectral-Spectral+EM-1833.pdf}}
     \hspace{-2em}
   \subfigure[EM]{
-    \includegraphics[width=0.50\textwidth]{figures/curves-vis/1-8-3-em.png}}
+    %\includegraphics[width=0.50\textwidth]{figures/curves-vis/1-8-3-em.png}}
+    \includegraphics[width=0.50\textwidth]{figures/EM-1833.pdf}}
 %  \subfigure[Spectral + EM]{
@@
   \subfigure[Well-specified data]{
-    \includegraphics[width=0.50\textwidth]{figures/vs-n/1833-decay.png}
+    %\includegraphics[width=0.50\textwidth]{figures/vs-n/1833-decay.png}
+    \includegraphics[width=0.50\textwidth]{figures/vs-n-1833.pdf}
   }
@@
   \subfigure[Misspecified data]{
-    \includegraphics[width=0.50\textwidth]{figures/vs-n/1-8-3-3-rm.png}
+    \includegraphics[width=0.50\textwidth]{figures/vs-n-1833-rm.pdf}
+    %\includegraphics[width=0.50\textwidth]{figures/vs-n/1-8-3-3-rm.png}
   }

### 2013-06-16 doc/icml2013/presentation/data/1833/plotLines.py

@@ -50,7 +50,7 @@ def makePlot( data, trueBetas, specBetas, spemBetas, em0Betas, emBetas, outFile 
 
     plt.rc("font",size=20)
 
-    for setting in [ [], ["EM"],["Spectral"],["Spectral+EM"],["Spectral","Spectral+EM"]]:
+    for setting in [ [], ["EM"],["Spectral","Spectral+EM"]]: # ["Spectral"],["Spectral+EM"],
         fig = plt.figure()
         ax = plt.axes(xlim=(-1, 1), ylim=(-3, 5))
         ax.set_xlabel("t")
@@ -60,7 +60,7 @@ def makePlot( data, trueBetas, specBetas, spemBetas, em0Betas, emBetas, outFile 
         if( len(setting) == 0 ):
             ax.scatter( data[0], data[1], alpha = 0.4 )
         else:
-            ax.scatter( data[0], data[1], alpha = 0.4 )
+            ax.scatter( data[0], data[1], alpha = 0.1 )
 
         if "True" in setting:
             # Now draw each line
@@ -70,7 +70,7 @@ def makePlot( data, trueBetas, specBetas, spemBetas, em0Betas, emBetas, outFile 
         plottables = []
         # Spectral is dashed
         if( "Spectral" in setting ):
-            lines = plotLines( ax, specBetas, "-", colors=['g']*3 )
+            lines = plotLines( ax, specBetas, "--", colors=['g']*3 )
             plottables.append( lines[0] )
         if( "Spectral+EM" in setting ):
             lines = plotLines( ax, spemBetas, "-", colors=['b']*3 )
@@ -80,8 +80,8 @@ def makePlot( data, trueBetas, specBetas, spemBetas, em0Betas, emBetas, outFile 
         if( "EM0" in setting ):
             plotLines( ax, em0Betas, ":" )
         if( "EM" in setting ):
-            for emBeta in (emBetas[0],):
-                lines = plotLines( ax, emBeta, "-", colors=['r']*3 )
+            for emBeta in emBetas:
+                lines = plotLines( ax, emBeta, ":", colors=['r']*3 )
             plottables.append( lines[0] )
 
         if( len(plottables) > 0):

### 2013-08-14 java/src/learning/linalg/MatrixOps.java
Fixed projectOntoSimplex

@@ -815,7 +815,13 @@ public class MatrixOps {
 
     for( int col = 0; col < nCols; col++ ) {
       // For each column, normalize the vector and zero out negative values.
-      columnNormalize( X, col );
+      // Get the majority sign
+      int positives = 0;
+      for( int row = 0; row < nRows; row++ ) positives += (X.get(row,col) > 0) ? 1 : 0;
+
+      // Flip sign
+      if( positives < nRows / 2 ) for( int row = 0; row < nRows; row++ )
+        X.set(row, col, -X.get(row,col));
 
       for( int row = 0; row < nRows; row++ ) {
         double x  = X_[ X.getIndex(row, col) ];

### 2013-08-14 java/src/learning/applications/bottleneck/BottleneckSpectralEstimation.scala
Added some scala code for bottlenecks

@@ -1,0 +1,25 @@
+package learning.applications.bottleneck
+
+import breeze.linalg.DenseVector
+import breeze.stats.distributions.Rand
+import scala.util.Random
+
+/**
+ * Parameter estimation for general graphical models by estimating conditional moments of "bottlenecks".
+ *     - Uses optimization to solve for parameters.
+ *     - Uses the tensor power method to recover these conditional moments.
+ */
+class BottleneckSpectralEstimation {
+  val rnd = new Random(1)
+
+  def optimize (model : LogLinearModel, data : List[Vector[Double]]) = {
+    val beta = DenseVector.rand(model.size, rnd) // Note that in our realization, beta and theta share features
+    val theta = DenseVector.rand(model.size, rnd)
+
+    // Optimize L(\theta, \beta) with respect to \beta first.
+
+
+    ()
+  }
+
+}

### 2013-08-14 java/src/learning/applications/bottleneck/LoglinearModel.scala

@@ -1,0 +1,20 @@
+package learning.applications.bottleneck
+
+/**
+ * A log-linear model
+ *    - has features phi(x,y)
+ *    - has parameters theta
+ */
+abstract class LogLinearModel extends Optimizable {
+  val features : Map[String,Int]
+//  val parameters : Vector[Double]
+
+  def likelihood : Vector[Option[Double]] => Double
+  def size = features.size
+}
+
+trait Optimizable {
+  def value : Vector[Double] => Double
+  def gradient : Vector[Double] => Vector[Double]
+}
+

### 2013-08-14 java/src/learning/applications/bottleneck/MixtureModel.scala

@@ -1,0 +1,34 @@
+package learning.applications.bottleneck
+import breeze.linalg._
+
+/**
+ * Implements a simple (undirected) mixture model as a log-linear model
+ *  - has features I[h=?] and I[h=?,x=?]
+ */
+class MixtureModel(k:Int, d: Int) extends LogLinearModel {
+  val features : Map[String,Int] = {
+    def idx_h(i : Int) = s"y=$i" -> i
+    def idx_hx(tpl : (Int,Int)) = tpl match { case (i:Int,j:Int) => s"y=$i,x=$j" -> (k + d*i + j) }
+
+    val hidden = List.range[Int](0,k)
+    val observed = List.range[Int](0,d)
+    (hidden.map(idx_h) ++ hidden.cross(observed).map(idx_hx)).toMap
+  }
+
+  implicit class Crossable[X](xs: Traversable[X]) {
+    def cross[Y](ys: Traversable[Y]) = for { x <- xs; y <- ys } yield (x, y)
+  }
+
+  val parameters: Vector[Double] = DenseVector.zeros[Double](features.size)
+
+  def likelihood: (Vector[Option[Double]]) => Double = params => 0.0
+  def value: Vector[Double] => Double = params => this.likelihood(params.map(v => Option(v)))
+  def gradient: Vector[Double] => Vector[Double] = params => DenseVector.zeros(features.size)
+}
+
+object testObject {
+  def main() = {
+    val model = new MixtureModel(2,2)
+    model.features
+  }
+}

### 2013-08-14 java/src/learning/models/loglinear/Main.java

@@ -169,14 +169,15 @@ class LikelihoodFunctionState implements Maximizer.FunctionState {
   public void performGradientCheck() {
     double epsilon = 1e-4;
     // Save point
-    double[] currentGradient = gradient();
-    double[] currentPoint = point();
+    double[] currentGradient = Arrays.copyOf(gradient(), params.weights.length);
+    double[] currentPoint = Arrays.copyOf(point(), params.weights.length);
 
     // Set point to be +/- gradient
     for( int i = 0; i < currentPoint.length; i++ ) {
-      params.weights[i] = currentPoint[i] + epsilon * currentGradient[i];
+      if( !measuredFeatures[i] ) continue;
+      params.weights[i] = currentPoint[i] + epsilon;
       double valuePlus = value();
-      params.weights[i] = currentPoint[i] - epsilon * currentGradient[i];
+      params.weights[i] = currentPoint[i] - epsilon;
       double valueMinus = value();
       params.weights[i] = currentPoint[i];
 

### 2014-01-18 java/src/learning/applications/POSInduction.java
Minor fixes to caching and diagnostic mode

@@ -116,14 +116,14 @@ public class POSInduction implements Runnable {
     int TOP_K = 20;
     for( int k = 0; k < K; k++ ) {
       StringBuilder topk = new StringBuilder();
-      topk.append(C.tagDict[k]).append(": ");
-
       int k_ = perm[k];
+
+      topk.append(C.tagDict[k_]).append(": ");
 
       Parameters parameters = (Parameters) params;
       double[] options = new double[model.D];
       for(int d = 0; d < model.D; d++)
-        options[d] = parameters.weights[parameters.o(k_, d)];
+        options[d] = parameters.weights[parameters.o(k, d)];
       Integer[] sortedWords = MatrixOps.argsort(options);
       for(int i = 0; i < TOP_K; i++ ) {
         topk.append(C.dict[sortedWords[i]]).append(", ");
@@ -238,8 +238,6 @@ public class POSInduction implements Runnable {
     }
     logsForce(table);
     LogInfo.end_track("Confusion matrix");
-
-    printTopK(model, params, C, perm);
 
     end_track("vs-all accuracy");
     return acc;
@@ -446,9 +444,9 @@ public class POSInduction implements Runnable {
       switch(mode) {
         case EM: {
           ExpectationMaximization solver = new ExpectationMaximization();
-          solver.backtrack.tolerance = 1e-4;
-          solver.mIters = 3;
-          solver.iters = 40;
+//          solver.backtrack.tolerance = 1e-4;
+          solver.mIters = 2;
+          solver.iters = 100;
           solver.thetaRegularization = 1e-3;
           solver.solveEM(hmm, data, params);
         } break;

### 2014-01-18 java/src/learning/models/loglinear/ExpectationMaximization.java

@@ -20,7 +20,7 @@ public class ExpectationMaximization implements Runnable {
   @Option(gloss="Regularization for theta") public double thetaRegularization = 1e-5; //0; //1e-3;
 
   @Option(gloss="Number of iterations") public int iters = 1000;
-  @Option(gloss="Number of iterations") public int mIters = 3;
+  @Option(gloss="Number of iterations") public int mIters = 1;
 
   @Option(gloss="Diagnostic mode") public boolean diagnosticMode = false;
 
@@ -29,6 +29,7 @@ public class ExpectationMaximization implements Runnable {
   @OptionSet(name="backtrack") public BacktrackingLineSearch.Options backtrack = new BacktrackingLineSearch.Options();
 
   Maximizer newMaximizer() {
+    backtrack.verbose = 5;
     if (useLBFGS) return new LBFGSMaximizer(backtrack, lbfgs);
     return new GradientMaximizer(backtrack);
   }
@@ -61,7 +62,7 @@ public class ExpectationMaximization implements Runnable {
     }
 
     @Override
-    public void invalidate() { objectiveValid = gradientValid = false; }
+    public void invalidate() { objectiveValid = gradientValid = false; theta.invalidateCache(); }
 
     @Override
     public double[] point() { return theta.toArray(); }
@@ -75,7 +76,6 @@ public class ExpectationMaximization implements Runnable {
   public double value() {
       if( objectiveValid ) return (objective);
       objective = 0.;
-      theta.invalidateCache();
       theta.cache();
 
 
@@ -94,8 +94,7 @@ public class ExpectationMaximization implements Runnable {
     @Override
     public double[] gradient() {
       if( gradientValid ) return gradient.toArray();
-      theta.invalidateCache();
-      theta.cache();
+        theta.cache();
 
       gradient.clear();
 
@@ -164,6 +163,12 @@ public class ExpectationMaximization implements Runnable {
     }
     for( int i = 0; i < iters && !done; i++ ) {
       theta.cache();
+
+      // Optimize each one-by-one
+      // Get marginals
+      marginals.clear();
+      modelA.updateMarginals(theta, data, 1.0, marginals);
+
       List<String> items = new ArrayList<>();
       items.add("iter = " + i);
       items.add("mObjective = " + mObjective.value());
@@ -174,12 +179,10 @@ public class ExpectationMaximization implements Runnable {
         out.flush();
       }
 
-      // Optimize each one-by-one
[…]

### 2014-01-18 java/src/learning/models/loglinear/ExpectationMaximizationTest.java

@@ -21,11 +21,12 @@ public class ExpectationMaximizationTest {
     solver.backtrack.tolerance = 1e-3;
     solver.mIters = 1;
     solver.iters = 500;
+    solver.diagnosticMode = true;
 
     if(testBetterThanTrue)
       solver.thetaRegularization = 0.0;
     else
-      solver.thetaRegularization = 1e-5;
+      solver.thetaRegularization = 0.;//1e-2;
     // Generate examples from the model
     Counter<Example> data =  model.drawSamples(trueParams, rnd, (int) 1e5);
 
@@ -219,6 +220,18 @@ public class ExpectationMaximizationTest {
   @Test
   public void testUndirectedHiddenMarkovModelEstimationLarge() {
     int K = 5; int D = 100; int L = 3;
+    UndirectedHiddenMarkovModel modelA = new UndirectedHiddenMarkovModel(K, D, L);
+    // Set parameters to be sin
+    Params trueParams = modelA.newParams();
+    trueParams.initRandom(rnd, 1.0);
+    Params initParams = trueParams.newParams();
+    initParams.initRandom(rnd, 1.0);
+
+    testRecovery(modelA, trueParams, initParams, false, true);
+  }
+  @Test
+  public void testUndirectedHiddenMarkovModelEstimationLarge1e3() {
+    int K = 5; int D = 1000; int L = 3;
     UndirectedHiddenMarkovModel modelA = new UndirectedHiddenMarkovModel(K, D, L);
     // Set parameters to be sin
     Params trueParams = modelA.newParams();

### 2014-01-18 java/src/learning/models/loglinear/UndirectedHiddenMarkovModel.java

@@ -38,15 +38,13 @@ public class UndirectedHiddenMarkovModel extends ExponentialFamilyModel<Example>
     final public int K;
     final public int D;
     final public double[] weights;
-
-    final double[] expWeights;
-    boolean cacheValid = false;
+    final double[] expWeight;
 
     public Parameters(int K, int D) {
       this.K = K;
       this.D = D;
       this.weights = new double[K*D + K*K];
-      this.expWeights = new double[K*D + K*K];
+      this.expWeight = new double[K*D + K*K];
     }
 
     @Override
@@ -123,15 +121,14 @@ public class UndirectedHiddenMarkovModel extends ExponentialFamilyModel<Example>
       }
     }
 
+    boolean cacheValid = false;
+    double[][][] cachedG;
     @Override
     public void cache() {
-      for(int i = 0; i < weights.length; i++){
-        expWeights[i] = Math.exp(weights[i]);
-      }
+      if(cacheValid) return;
       precomputeG();
       cacheValid = true;
     }
-    double[][][] cachedG;
     void precomputeG() {
       // Use a special index for t = 0 and for examples = 0
       Example ex = new Example(new int[]{0, 0});
@@ -141,15 +138,19 @@ public class UndirectedHiddenMarkovModel extends ExponentialFamilyModel<Example>
           for(int x = 0; x < D; x++) {
             if(y_ == -1) {
               ex.x[0] = x;
-              cachedG[y_+1][y][x+1] = G(0, y_, y, ex);
+              cachedG[y_+1][y][x+1] = G_(0, y_, y, ex);
             } else {
               ex.x[1] = x;
-              cachedG[y_+1][y][x+1] = G(1, y_, y, ex);
+              cachedG[y_+1][y][x+1] = G_(1, y_, y, ex);
             }
           }
           // Now compute the aggregate for 0
           cachedG[y_+1][y][0] = MatrixOps.sum(cachedG[y_+1][y]);
         }
+      }
+
+      for(int i = 0; i < weights.length; i++) {
+        expWeight[i] = weights[i];
       }
     }
 
@@ -158,37 +159,29 @@ public class UndirectedHiddenMarkovModel extends ExponentialFamilyModel<Example>
     public void invalidateCache() {
       cacheValid = false;
     }
-
     /**
      * Return \theta^T [\phi(y_{t-1}, y_{t}), \phi(y_{t}, x_{t}) ]
      */
[…]

### 2014-01-18 java/src/learning/models/loglinear/UndirectedHiddenMarkovModelTest.java

@@ -1,6 +1,7 @@
 package learning.models.loglinear;
 
 import fig.basic.Pair;
+import learning.models.Params;
 import learning.models.loglinear.UndirectedHiddenMarkovModel.Parameters;
 import learning.utils.Counter;
 import org.junit.Assert;
@@ -303,6 +304,37 @@ public class UndirectedHiddenMarkovModelTest {
     for(int i = 0; i < marginal.weights.length; i++) {
       Assert.assertTrue(Math.abs(marginal.weights[i] - marginal_.weights[i]) < 1e-1);
     }
+  }
+
+  @Test
+  public void testCaching() {
+    int D = model.D; int K = model.K; int T = model.L;
+    Example ex = new Example(new int[T], new int[T]);
+
+    for(int t = 0; t < T; t++) {
+      for(int y_ = 0; y_ < K; y_++) {
+        for(int y = 0; y < K; y++) {
+          for(int x = 0; x < K; x++) {
+            ex.x[t] = x;
+            Assert.assertEquals( model1.G(t,y_,y,ex), model1.G_(t,y_,y,ex), 1e-2);
+          }
+        }
+      }
+    }
+    model1.cache();
+    for(int t = 0; t < T; t++) {
+      for(int y_ = 0; y_ < K; y_++) {
+        for(int y = 0; y < K; y++) {
+          for(int x = 0; x < K; x++) {
+            ex.x[t] = x;
+            Assert.assertEquals( model1.G_(t,y_,y,ex), model1.G(t,y_,y,ex), 1e-2);
+          }
+        }
+      }
+    }
+    model1.invalidateCache();
+
+
   }
 
 }

### 2014-01-18 java/src/learning/utils/UtilsJ.java

@@ -113,6 +113,9 @@ public class UtilsJ {
     double oldObjective = Double.NEGATIVE_INFINITY;
     int iter;
     for (iter = 0; iter < numIters && !done; iter++) {
+      if(diagnosticMode) {
+        doGradientCheck(state);
+      }
       // Logging stuff
       List<String> items = new ArrayList<>();
       items.add("iter = " + iter);

### 2014-01-18 doc/local-optima/paper.tex
Added document about local-optima

\documentclass[tablecaption=bottom]{jmlr}

\jmlrproceedings{}{}
\jmlrpages{}{}
\usepackage[cm]{fullpage}
\usepackage{booktabs}
\usepackage{ctable}

\title{Counting local-optima in a log-linear model}

\author{Arun Tejasvi Chaganty \Email{chaganty@cs.stanford.edu}}

\input{macros}

\begin{document}

\maketitle

\section{Introduction}

At the high-level, we're looking to count the number of the local optima
of a log-linear model,
\begin{align}
p_{\theta}(x) &= \sum_z \exp(\theta^T \phi(x,z) - A(\theta)).
\end{align}

This procedure does not exactly cover the mixture of Gaussians though.

\section{Preliminaries}

\section{Mixture of Gaussians}

\section{Log-linear models}

\todo{Write down log-likelihood}

\begin{definition}(Trapping Region)
  $\sT(\theta)$ is a closed subset of $\Re^n$ such that for every
  $\theta' \in \partial \sT(\theta)$, $\sL(\theta', q(\theta')) \le \sL(
  \theta, q(\theta) )$.
\end{definition}

\begin{lemma}(Trapping property of $\sT(\theta)$)
  For any $\theta$, EM converges to a fixed point in $\sT(\theta)$.
\end{lemma}
\begin{proof}
  Let $\theta', q(\theta')$ be the new set of parameters after one
  iteration of EM. We will prove by contradiction that $\theta'$ must
  lie within $\sT(\theta)$.

  Consider a trajectory $\sP$ from $\theta$ to $\theta'$, defined as follows,
  \begin{align}
    \theta^\alpha = \alpha \theta + (1-\alpha) \theta'.
  \end{align}

\end{proof}

\end{document}

### 2014-01-18 doc/local-optima/paper.tex
Updated proof

@@
 
-This procedure does not exactly cover the mixture of Gaussians though.
+\todo{Describe mixtures of Gaussians}
 
@@
 
-\todo{Write down log-likelihood}
+In EM, we maximize a lower bound on the log-likelihood of model which is
+derived as follows,
+\begin{align}
+  \sL(\theta ; \sD) 
+    &\defeq \sum_{x_n \in \sD} \log p_{\theta}(x_n) \\
+    &= \sum_{x_n \in \sD} \log \sum_z \exp(\theta^T \phi(x,z) - A(\theta)) \\
+    &= \sum_{x_n \in \sD} \log \sum_z \frac{q(z|x_n)}{q(z|x_n)} \exp(\theta^T \phi(x,z) - A(\theta)) \\
+    &= \sum_{x_n \in \sD} \sum_z q(z|x_n) \log \frac{1}{q(z|x_n)} \exp(\theta^T \phi(x,z) - A(\theta)) \\
+    &\le \sum_{x_n \in \sD} \sum_z q(z|x_n) [\theta^T \phi(x,z) - A(\theta) - \log q(z|x_n)].
+\end{align}
+
+We use the notation $\sL(\theta, q) \defeq \sum_{x_n \in \sD} \sum_z
+q(z|x_n) [\theta^T \phi(x,z) - A(\theta) - \log q(z|x_n)]$. The solution
+to EM is $\max_\theta \max_q \sL(\theta,q)$ which we do via alternating
+maximization between $\theta$ and $q$.
+
+Furthermore, it can be shown that local optima of $\sL(\theta, q)$ are
+also local optima of the original function, $\sL(\theta ; \sD)$.
+
+Finally, we derive closed form solutions to the E and M steps,
+\begin{align}
+  q^*(\theta) 
+    &\defeq \max_{q \in \Delta_{n-1}} \sL(\theta,q)\\
+  \diff{\sL(\theta,q)}{q(z|x_n)} 
+    &= \theta^T \phi(x_n,z) - A(\theta) - \log q(z|x_n) - 1 + \lambda \\
+  q^*(z|x_n) &\propto \exp(\theta^T \phi(x_n,z) - A(\theta)) \\
+  q^*(z|x_n) &= \frac{exp(\theta^T \phi(x_n,z) - A(\theta))}{\sum_{z'} exp(\theta^T \phi(x_n,z') - A(\theta))}.
+\end{align}
+
+\begin{align}
+  \theta^*(q)
+    &\defeq \max_{\theta \in \Re^{n}} \sL(\theta,q)\\
+  \diff{\sL(\theta,q)}{\theta} 
+  &= \sum_{x_n \sD}\sum_{z} q(z|x_n) \phi(x_n,z) - \diff{A(\theta)}{\theta} \\
+  &= \sum_{x_n \sD}\sum_{z} q(z|x_n) \phi(x_n,z) - \E_{\theta}(\phi(x,z)) \\
+  &= \hat\E_{q}[\phi(x_n,z)] - \E_{\theta}(\phi(x,z)) \\
+  \E_{\theta^*}(\phi(x,z)) &= \hat\E_{q}[\phi(x_n,z)].
+\end{align}
 
 \begin{definition}(Trapping Region)
-  $\sT(\theta)$ is a closed subset of $\Re^n$ such that for every
-  $\theta' \in \partial \sT(\theta)$, $\sL(\theta', q(\theta')) \le \sL(
-  \theta, q(\theta) )$.
+  A closed subset of $\Re^n$ is said to be trapping for $\theta \in
+  \Re^n$ if for every $\theta' \in \del R$, $\sL(\theta',
+  q(\theta')) \le \sL( \theta, q(\theta) )$.
 \end{definition}
 
-\begin{lemma}(Trapping property of $\sT(\theta)$)
-  For any $\theta$, EM converges to a fixed point in $\sT(\theta)$.
+\begin{lemma}(Trapping property)
+  For any $\theta$ and a trapping region $R$ containing $\theta$, expectation
+  maximization initialized at $\theta$ converges to a fixed point in $R$.
 \end{lemma}
@@
   Let $\theta', q(\theta')$ be the new set of parameters after one
-  iteration of EM. We will prove by contradiction that $\theta'$ must
-  lie within $\sT(\theta)$.
+  iteration of EM. The proof proceeds by constructing a path $\sP$ from
+  $\theta$ to $\theta'$ such that every point along the path
+  $\theta^\alpha \in \sP$ has a higher likelihood than $\theta$ and thus
+  could not cross the boundary $\del R$.
 
-  Consider a trajectory $\sP$ from $\theta$ to $\theta'$, defined as follows,
+  Let us define a trajectory $\sP : \Re \to \Re^n$ from $\theta$ to
+  $\theta'$ as follows,
   \begin{align}
-    \theta^\alpha = \alpha \theta + (1-\alpha) \theta'.
+    \sP(\alpha) &\defeq \alpha \theta + (1-\alpha) \theta'.
   \end{align}
+  For notational convenience, we use $\theta_{\alpha} \defeq \sP(\alpha)$.
 
+  Now, note that with $q$ fixed, $\sL(\theta,q)$ is {\em convex} in $\theta$. Thus, 
+  $\sL(\theta^\alpha, q) = (1-\alpha)\sL(\theta, q) + \alpha
+  \sL(\theta', q) > \sL(\theta, q)$. Furthermore, $\sL(\theta^\alpha,
+  q^*(\theta^\alpha)) \ge \sL(\theta^\alpha, q) > \sL(\theta, q)$.
+
+  \todo{Refine with a path for $q$ too}
 \end{proof}
 
+Next, we will define a local region around some parameter $\theta$ that is trapping.
+
 \end{document}

### 2014-01-26 doc/icml2014/paper.tex
Added icml2014 draft

\documentclass{article}

% use Times
\usepackage{times}
% For figures
\usepackage{graphicx} % more modern
\usepackage{subfigure} 
\usepackage{amsmath,amssymb,amsthm}

% For citations
\usepackage{natbib}

% For algorithms
\usepackage{algorithm}
\usepackage{algorithmic}

% As of 2011, we use the hyperref package to produce hyperlinks in the
% resulting PDF.  If this breaks your system, please commend out the
% following usepackage line and replace \usepackage{icml2014} with
% \usepackage[nohyperref]{icml2014} above.
\usepackage{hyperref}

% Packages hyperref and algorithmic misbehave sometimes.  We can fix
% this with the following command.
\newcommand{\theHalgorithm}{\arabic{algorithm}}

% Employ the following version of the ``usepackage'' statement for
% submitting the draft version of the paper for review.  This will set
% the note in the first column to ``Under review.  Do not distribute.''
\usepackage{icml2014} 
% Employ this version of the ``usepackage'' statement after the paper has
% been accepted, when creating the final version.  This will set the
% note in the first column to ``Proceedings of the...''
%\usepackage[accepted]{icml2014}

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
%\icmltitlerunning{Submission and Formatting Instructions for ICML 2014}

\input{macros}

\begin{document}

\twocolumn[
\icmltitle{Moment Constraints Make Learning Latent-Variable Models Easier}

% It is OKAY to include author information, even for blind
% submissions: the style file will automatically remove it for you
% unless you've provided the [accepted] option to the icml2014
% package.
\icmlauthor{Arun Tejasvi Chaganty}{chaganty@cs.stanford.edu}
\icmlauthor{Percy Liang}{pliang@cs.stanford.edu}
\icmladdress{Stanford University,
Stanford, CA, USA}

% You may provide any keywords that you 
% find helpful for describing your paper; these are used to populate 
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{machine learning}

\vskip 0.3in
]

\begin{abstract} 
\end{abstract}

\section{Introduction}
\label{sec:introduction}

% 1. Latent variable models are good.
Latent variable models offer a succinct representation of a rich model
family. 
% 2. Learning them is hard.
Despite their success across many fields
\cite{quattoni04crf,haghighi06prototype,liang06discrimative,kirkpatrick10painless},
learning these models remains a difficult problem due to the
non-convexity of the likelihood. Local optimization (e.g.
expectation-maximization) is the standard approach, but is susceptible
to local optima.

% 3. People have approached unsupervised learning with the MoM magic sauce, but the sauce is limited.
Recently, unsupervised learning techniques based on the method of moments and
spectral decomposition have offered a refreshing and promising perspective on
this learning problem \citep{hsu09spectral,anandkumar11tree,anandkumar12moments,anandkumar12lda,hsu12identifiability,balle11transducer,balle12automata}.
These methods exploit the linear algebraic properties of the model and
factorize the moments into parameters, providing strong theoretical guarantees.
However, these methods are not as universally applicable as EM.

% 4. State what we do: exploit moment constraints to make the problem easier.
In this work, we exploit the spectral method to learn moment constraints
on the observed variables and show how these constraints can greatly
simplify the learning problem.
% 5. We get moments from third-order tensors from bottlenecks and factorize them into marginals.
The key idea in
\citet{anandkumar12moments,anandkumar13tensor} is to study the conditional
independence factorization structure implied by the model in the form of 
three-view bottlenecks. These bottlenecks which can be factorized into
moments involving the latent variables\vague.

% 6. Provide some intuitive examples - Can I haz HMMs?
Consider a hidden Markov model, with initial probabilities $\pi$,
  transitions $T$ and emissions $O$. 
While \citet{hsu09spectral,anandkumar12moments} provide a consistent estimator
  for the parameters of this model that inverts $O$ to recover $T$ and
  $\pi$. We would like to avoid this potentially sensitive procedure via optimization.
The log-likelihood of this model, 
  \begin{align}
    \sL(\pi, T, O) &= \sum_{x\in\sD} \log( \sum_{h} \pi(h_1) O(x_1 | h_1) T(h_2 | h_1) O(x_2 | h_2) \cdots ),
  \end{align}
  is not convex, even if we knew any one of the parameters.
However, intuitively, if we knew $O$ and found that each symbol was
  produced by exactly one state, learning $T$ and $\pi$ could be trivial
  accomplished via counting. 
In general, $O$ still informs us about the states in the observation
  sequence. If we considered instead the piecewise likelihood, 
  \begin{align}
    \sL_{\textrm{piecewise}}(\pi, T, O) &= \sum_{x\in\sD} \log( \sum_{h_1, h_2} \underbrace{\pi(h_1) T(h_2 | h_1)}_{P(h_1,h_2)} O(x_1 | h_1) O(x_2 | h_2) ),
  \end{align}
  then we observe that the likelihood is convex in the marginal
  distribution $P(h_1,h_2)$. This allows us to recover $P$, from which $T$
  can be easily reconstructed.

% 7. Now outline the rest of our results.
In \sectionref{piecewise} we generalize this observation to a directed grid
  model and a directed tree model and show that the {\em piecewise
  likelihood} is strictly convex, providing consistent parameter
  estimators for these models (which is new)\reword.
The approach can be extended to log-linear
  models (\sectionref{log-linear}), where the problem remains non-convex,
  however, we expect a reduction in the number of local optima. 
In this regime, we show empirically that constraining the moments of the
  model via posterior-regularization allows us to learn better models.

\todo{Related work on spectral methods and measurements.}

\section{Parameter Estimation with Convex Piecewise Likelihoods}
\label{sec:piecewise}

\section{Parameter Estimation for Log-linear Latent-Variable Models}
\label{sec:log-linear}

\paragraph{Setup}
Let $z$ be a collection of variables indexed by $V$ (that is, $z = \{z_j\}_{j \in V}$).
Suppose that some of the variables are observed ($x = \{ z_j \}_{j \in O}$, where $O \subset V$) 
and the rest are latent ($h = \{ z_j \}_{j \in H}$, where $H = V \backslash O$).
Let $C \subset 2^V$ be a collection of subsets of variables,
and let $\phi_c(z_c) \in \Re^d$ be the feature map for each $c \in C$.
In this paper, we will consider both directed and undirected models over $z$.
In the undirected case, we define the standard globally-normalized exponential family model:
\begin{align}
  \label{eqn:undirectedSetup}
  p_\theta(z) = \exp\{ \phi(z)^\top\theta - A(\theta) \},
\end{align}
where $\phi(z) \eqdef \sum_{c \in C} \phi_c(z_c) \in \Re^d$ is the global feature vector, $\theta \in \Re^d$ is the parameter vector,
and $A(\theta) = \log \int \exp\{\phi(z)^\top\theta\} dz$ is the global log-partition function.
To some extent, our methods only depend on the conditional independence
structure of the models.

\paragraph{Problem statement}

% Statement
This paper focuses on the problem of parameter estimation:
We are given $n$ i.i.d.~examples of the observed variables $D = (x^{(1)}, \dots, x^{(n)})$
where each $x^{(i)} \sim p_{\theta^*}$ for some true parameters $\theta^*$.
Our goal is to produce a parameter estimate $\hat\theta$ that approximates $\theta^*$.

% Maximum likelihood
The standard estimation procedure is maximum (marginal) likelihood,
$\sL \eqdef \max_{\theta \in \Re^d} \sum_{x \in D} \log p_\theta(x)$,
which is statistically efficient but computationally intractable.
In practice, one uses gradient-based optimization procedures (e.g., EM or L-BFGS)
on the marginal likelihood, which can get stuck in local optima.

\paragraph{Primer: mixture of Bernoulli}

\section{Experiments}
\label{sec:experiments}

\section{Discussion}
\label{sec:discussion}

\bibliography{pliang}
\bibliographystyle{icml2014}

\end{document}

### 2014-01-26 doc/icml2014/measurements.tex
Updated paper

--- 2014-01-26 doc/icml2014/paper.tex
@@
-\documentclass{article}
-
-% use Times
-\usepackage{times}
-% For figures
-\usepackage{graphicx} % more modern
-\usepackage{subfigure} 
-\usepackage{amsmath,amssymb,amsthm}
-
-% For citations
-\usepackage{natbib}
-
-% For algorithms
-\usepackage{algorithm}
-\usepackage{algorithmic}
-
-% As of 2011, we use the hyperref package to produce hyperlinks in the
-% resulting PDF.  If this breaks your system, please commend out the
-% following usepackage line and replace \usepackage{icml2014} with
-% \usepackage[nohyperref]{icml2014} above.
-\usepackage{hyperref}
-
-% Packages hyperref and algorithmic misbehave sometimes.  We can fix
-% this with the following command.
-\newcommand{\theHalgorithm}{\arabic{algorithm}}
-
-% Employ the following version of the ``usepackage'' statement for
-% submitting the draft version of the paper for review.  This will set
-% the note in the first column to ``Under review.  Do not distribute.''
-\usepackage{icml2014} 
-% Employ this version of the ``usepackage'' statement after the paper has
-% been accepted, when creating the final version.  This will set the
-% note in the first column to ``Proceedings of the...''
-%\usepackage[accepted]{icml2014}
-
-% The \icmltitle you define below is probably too long as a header.
-% Therefore, a short form for the running title is supplied here:
-%\icmltitlerunning{Submission and Formatting Instructions for ICML 2014}
-
-\input{macros}
-
-\begin{document}
-
-\twocolumn[
-\icmltitle{Moment Constraints Make Learning Latent-Variable Models Easier}
-
-% It is OKAY to include author information, even for blind
-% submissions: the style file will automatically remove it for you
-% unless you've provided the [accepted] option to the icml2014
-% package.
-\icmlauthor{Arun Tejasvi Chaganty}{chaganty@cs.stanford.edu}
-\icmlauthor{Percy Liang}{pliang@cs.stanford.edu}
-\icmladdress{Stanford University,
-Stanford, CA, USA}
-
-% You may provide any keywords that you 
-% find helpful for describing your paper; these are used to populate 
-% the "keywords" metadata in the PDF but will not be shown in the document
-\icmlkeywords{machine learning}
-
-\vskip 0.3in
-]
-
-\begin{abstract} 
-\end{abstract}
-
-\section{Introduction}
-\label{sec:introduction}
-
-% 1. Latent variable models are good.
-Latent variable models offer a succinct representation of a rich model
-family. 
-% 2. Learning them is hard.
-Despite their success across many fields
-\cite{quattoni04crf,haghighi06prototype,liang06discrimative,kirkpatrick10painless},
-learning these models remains a difficult problem due to the
-non-convexity of the likelihood. Local optimization (e.g.
-expectation-maximization) is the standard approach, but is susceptible
-to local optima.
-
-% 3. People have approached unsupervised learning with the MoM magic sauce, but the sauce is limited.
-Recently, unsupervised learning techniques based on the method of moments and
-spectral decomposition have offered a refreshing and promising perspective on
-this learning problem \citep{hsu09spectral,anandkumar11tree,anandkumar12moments,anandkumar12lda,hsu12identifiability,balle11transducer,balle12automata}.
-These methods exploit the linear algebraic properties of the model and
-factorize the moments into parameters, providing strong theoretical guarantees.
-However, these methods are not as universally applicable as EM.
-
-% 4. State what we do: exploit moment constraints to make the problem easier.
-In this work, we exploit the spectral method to learn moment constraints
-on the observed variables and show how these constraints can greatly
-simplify the learning problem.
-% 5. We get moments from third-order tensors from bottlenecks and factorize them into marginals.
-The key idea in
-\citet{anandkumar12moments,anandkumar13tensor} is to study the conditional
-independence factorization structure implied by the model in the form of 
-three-view bottlenecks. These bottlenecks which can be factorized into
-moments involving the latent variables\vague.
-
-% 6. Provide some intuitive examples - Can I haz HMMs?
-Consider a hidden Markov model, with initial probabilities $\pi$,
-  transitions $T$ and emissions $O$. 
-While \citet{hsu09spectral,anandkumar12moments} provide a consistent estimator
-  for the parameters of this model that inverts $O$ to recover $T$ and
-  $\pi$. We would like to avoid this potentially sensitive procedure via optimization.
-The log-likelihood of this model, 
-  \begin{align}
-    \sL(\pi, T, O) &= \sum_{x\in\sD} \log( \sum_{h} \pi(h_1) O(x_1 | h_1) T(h_2 | h_1) O(x_2 | h_2) \cdots ),
-  \end{align}
-  is not convex, even if we knew any one of the parameters.
-However, intuitively, if we knew $O$ and found that each symbol was
-  produced by exactly one state, learning $T$ and $\pi$ could be trivial
-  accomplished via counting. 
-In general, $O$ still informs us about the states in the observation
-  sequence. If we considered instead the piecewise likelihood, 
-  \begin{align}
-    \sL_{\textrm{piecewise}}(\pi, T, O) &= \sum_{x\in\sD} \log( \sum_{h_1, h_2} \underbrace{\pi(h_1) T(h_2 | h_1)}_{P(h_1,h_2)} O(x_1 | h_1) O(x_2 | h_2) ),
-  \end{align}
-  then we observe that the likelihood is convex in the marginal
-  distribution $P(h_1,h_2)$. This allows us to recover $P$, from which $T$
-  can be easily reconstructed.
-
-% 7. Now outline the rest of our results.
-In \sectionref{piecewise} we generalize this observation to a directed grid
-  model and a directed tree model and show that the {\em piecewise
-  likelihood} is strictly convex, providing consistent parameter
-  estimators for these models (which is new)\reword.
-The approach can be extended to log-linear
-  models (\sectionref{log-linear}), where the problem remains non-convex,
-  however, we expect a reduction in the number of local optima. 
-In this regime, we show empirically that constraining the moments of the
-  model via posterior-regularization allows us to learn better models.
-
-\todo{Related work on spectral methods and measurements.}
-
-\section{Parameter Estimation with Convex Piecewise Likelihoods}
-\label{sec:piecewise}
-
 \section{Parameter Estimation for Log-linear Latent-Variable Models}
@@
 
-\paragraph{Primer: mixture of Bernoulli}
+\paragraph{Primer: Three-view log-linear mixture models}
 
-\section{Experiments}
-\label{sec:experiments}
+\citet{anandkumar12moments} describe an algorithm to recover parameters
+  for a directed three-view mixture model. 
+We will consider an undirected log-linear model variant of this model,
+  shown in \figureref{three-view}.
+The model has features $\phi(x,y) = \{\bI(x_1,y_1), \bI(x_2,y_1), \bI(x_2,y_1)\}$.
 
-\section{Discussion}
-\label{sec:discussion}
+Recover moments with tensor factorization.
+Show that the constrained optimization problem is convex.
 
-\bibliography{pliang}
-\bibliographystyle{icml2014}
+Statement: consistency.
 
-\end{document}
+\paragraph{General latent-variable models with bottlenecks}
+
+Our learning algorithm for three-view mixture models relies on the fact that
+the sufficient statistics of the exponential family
+can be expressed as a function of the latent moments.
+Of course, these conditions do not hold for general log-linear models.
+Rather than falling back on plain local optimization of the likelihood in these
+cases, we would still like
+to somehow use the \emph{partial information} offered by the latent moments.
+
+Partial information is available when the following holds:
+%\begin{property}
+%  There exists some latent variable $h_{j_0} \in [\nh]$ (the bottleneck)
+%  and three conditionally independent observed variables $x_{j_1},x_{j_2},x_{j_3}$ (views) such that
+%  each conditional mean matrix $\E[\phix(x_{j}) \mid h_{j_0}] \in \R^{\nphix \times \nh}$ ($j \in \{ j_1, j_2, j_3 \}$) has full column rank $\nh$
+%  for some transformation $\phix$.
+%\end{property}
+
+Optimization.

### 2014-01-26 doc/icml2014/paper.tex
Updated paper

@@
 
-\section{Introduction}
-\label{sec:introduction}
-
-% 1. Latent variable models are good.
-Latent variable models offer a succinct representation of a rich model
-family. 
-% 2. Learning them is hard.
-Despite their success across many fields
-\cite{quattoni04crf,haghighi06prototype,liang06discrimative,kirkpatrick10painless},
-learning these models remains a difficult problem due to the
-non-convexity of the likelihood. Local optimization (e.g.
-expectation-maximization) is the standard approach, but is susceptible
-to local optima.
-
-% 3. People have approached unsupervised learning with the MoM magic sauce, but the sauce is limited.
-Recently, unsupervised learning techniques based on the method of moments and
-spectral decomposition have offered a refreshing and promising perspective on
-this learning problem \citep{hsu09spectral,anandkumar11tree,anandkumar12moments,anandkumar12lda,hsu12identifiability,balle11transducer,balle12automata}.
-These methods exploit the linear algebraic properties of the model and
-factorize the moments into parameters, providing strong theoretical guarantees.
-However, these methods are not as universally applicable as EM.
-
-% 4. State what we do: exploit moment constraints to make the problem easier.
-In this work, we exploit the spectral method to learn moment constraints
-on the observed variables and show how these constraints can greatly
-simplify the learning problem.
-% 5. We get moments from third-order tensors from bottlenecks and factorize them into marginals.
-The key idea in
-\citet{anandkumar12moments,anandkumar13tensor} is to study the conditional
-independence factorization structure implied by the model in the form of 
-three-view bottlenecks. These bottlenecks which can be factorized into
-moments involving the latent variables\vague.
-
-% 6. Provide some intuitive examples - Can I haz HMMs?
-Consider a hidden Markov model, with initial probabilities $\pi$,
-  transitions $T$ and emissions $O$. 
-While \citet{hsu09spectral,anandkumar12moments} provide a consistent estimator
-  for the parameters of this model that inverts $O$ to recover $T$ and
-  $\pi$. We would like to avoid this potentially sensitive procedure via optimization.
-The log-likelihood of this model, 
-  \begin{align}
-    \sL(\pi, T, O) &= \sum_{x\in\sD} \log( \sum_{h} \pi(h_1) O(x_1 | h_1) T(h_2 | h_1) O(x_2 | h_2) \cdots ),
-  \end{align}
-  is not convex, even if we knew any one of the parameters.
-However, intuitively, if we knew $O$ and found that each symbol was
-  produced by exactly one state, learning $T$ and $\pi$ could be trivial
-  accomplished via counting. 
-In general, $O$ still informs us about the states in the observation
-  sequence. If we considered instead the piecewise likelihood, 
-  \begin{align}
-    \sL_{\textrm{piecewise}}(\pi, T, O) &= \sum_{x\in\sD} \log( \sum_{h_1, h_2} \underbrace{\pi(h_1) T(h_2 | h_1)}_{P(h_1,h_2)} O(x_1 | h_1) O(x_2 | h_2) ),
-  \end{align}
-  then we observe that the likelihood is convex in the marginal
-  distribution $P(h_1,h_2)$. This allows us to recover $P$, from which $T$
-  can be easily reconstructed.
-
-% 7. Now outline the rest of our results.
-In \sectionref{piecewise} we generalize this observation to a directed grid
-  model and a directed tree model and show that the {\em piecewise
-  likelihood} is strictly convex, providing consistent parameter
-  estimators for these models (which is new)\reword.
-The approach can be extended to log-linear
-  models (\sectionref{log-linear}), where the problem remains non-convex,
-  however, we expect a reduction in the number of local optima. 
-In this regime, we show empirically that constraining the moments of the
-  model via posterior-regularization allows us to learn better models.
-
-\todo{Related work on spectral methods and measurements.}
-
-\section{Parameter Estimation with Convex Piecewise Likelihoods}
-\label{sec:piecewise}
-
-\section{Parameter Estimation for Log-linear Latent-Variable Models}
-\label{sec:log-linear}
-
-\paragraph{Setup}
-Let $z$ be a collection of variables indexed by $V$ (that is, $z = \{z_j\}_{j \in V}$).
-Suppose that some of the variables are observed ($x = \{ z_j \}_{j \in O}$, where $O \subset V$) 
-and the rest are latent ($h = \{ z_j \}_{j \in H}$, where $H = V \backslash O$).
-Let $C \subset 2^V$ be a collection of subsets of variables,
-and let $\phi_c(z_c) \in \Re^d$ be the feature map for each $c \in C$.
-In this paper, we will consider both directed and undirected models over $z$.
-In the undirected case, we define the standard globally-normalized exponential family model:
-\begin{align}
-  \label{eqn:undirectedSetup}
-  p_\theta(z) = \exp\{ \phi(z)^\top\theta - A(\theta) \},
-\end{align}
-where $\phi(z) \eqdef \sum_{c \in C} \phi_c(z_c) \in \Re^d$ is the global feature vector, $\theta \in \Re^d$ is the parameter vector,
-and $A(\theta) = \log \int \exp\{\phi(z)^\top\theta\} dz$ is the global log-partition function.
-To some extent, our methods only depend on the conditional independence
-structure of the models.
-
-\paragraph{Problem statement}
-
-% Statement
-This paper focuses on the problem of parameter estimation:
-We are given $n$ i.i.d.~examples of the observed variables $D = (x^{(1)}, \dots, x^{(n)})$
-where each $x^{(i)} \sim p_{\theta^*}$ for some true parameters $\theta^*$.
-Our goal is to produce a parameter estimate $\hat\theta$ that approximates $\theta^*$.
-
-% Maximum likelihood
-The standard estimation procedure is maximum (marginal) likelihood,
-$\sL \eqdef \max_{\theta \in \Re^d} \sum_{x \in D} \log p_\theta(x)$,
-which is statistically efficient but computationally intractable.
-In practice, one uses gradient-based optimization procedures (e.g., EM or L-BFGS)
-on the marginal likelihood, which can get stuck in local optima.
-
-\paragraph{Primer: mixture of Bernoulli}
-
-\section{Experiments}
-\label{sec:experiments}
-
-\section{Discussion}
-\label{sec:discussion}
+\input{intro}
+\input{background}
+\input{piecewise}
+\input{measurements}
+\input{experiments}
+\input{discussion}
 

### 2014-01-26 doc/icml2014/background.tex

\section{Background}
\label{sec:background}

\paragraph{Bottlenecks: Three-view mixture models}

\citet{anandkumar12moments} present an algorithm to solve three-view mixture models.

Assume means $M_1, M_2, M_3$. Observe that the moments of the model are $M_{123}$.

Applying a whitening transformation and get a orthogonal tensor decomposition.

Solved using the robust tensor power method \citet{anandkumar13tensor}.

\todo{Related work on spectral methods and measurements.}

\paragraph{Constrained EM}

Posterior regularization
Measurements
Generalized expectation criterion

Kuzman Ganchev.*

### 2014-01-26 doc/icml2014/intro.tex

\section{Introduction}
\label{sec:introduction}

% 1. Latent variable models are good.
Latent variable models offer a succinct representation of a rich model
family. 
% 2. Learning them is hard.
Despite their success across many fields
\cite{quattoni04crf,haghighi06prototype,liang06discrimative,kirkpatrick10painless},
learning these models remains a difficult problem due to the
non-convexity of the likelihood. Local optimization (e.g.
expectation-maximization) is the standard approach, but is susceptible
to local optima.

% 3. People have approached unsupervised learning with the MoM magic sauce, but the sauce is limited.
Recently, unsupervised learning techniques based on the method of moments and
spectral decomposition have offered a refreshing and promising perspective on
this learning problem \citep{hsu09spectral,anandkumar11tree,anandkumar12moments,anandkumar12lda,hsu12identifiability,balle11transducer,balle12automata}.
These methods exploit the linear algebraic properties of the model and
factorize the moments into parameters, providing strong theoretical guarantees.
However, these methods are not as universally applicable as EM.

% 4. State what we do: exploit moment constraints to make the problem easier.
In this work, we exploit the spectral method to learn moment constraints
on the observed variables and show how these constraints can greatly
simplify the learning problem.
% 5. We get moments from third-order tensors from bottlenecks and factorize them into marginals.
The key idea in
\citet{anandkumar12moments,anandkumar13tensor} is to study the conditional
independence factorization structure implied by the model in the form of 
three-view bottlenecks. These bottlenecks which can be factorized into
moments involving the latent variables\vague.

% 6. Provide some intuitive examples - Can I haz HMMs?
Consider a hidden Markov model, with initial probabilities $\pi$,
  transitions $T$ and emissions $O$. 
While \citet{hsu09spectral,anandkumar12moments} provide a consistent estimator
  for the parameters of this model that inverts $O$ to recover $T$ and
  $\pi$. We would like to avoid this potentially sensitive procedure via optimization.
The log-likelihood of this model, 
  \begin{align}
    \sL(\pi, T, O) &= \sum_{x\in\sD} \log( \sum_{h} \pi(h_1) O(x_1 | h_1) T(h_2 | h_1) O(x_2 | h_2) \cdots ),
  \end{align}
  is not convex, even if we knew any one of the parameters.
However, intuitively, if we knew $O$ and found that each symbol was
  produced by exactly one state, learning $T$ and $\pi$ could be trivial
  accomplished via counting. 
In general, $O$ still informs us about the states in the observation
  sequence. If we considered instead the piecewise likelihood, 
  \begin{align}
    \sL_{\textrm{p}}(\pi, T, O) &= \sum_{x\in\sD} \log( \sum_{h_1, h_2} \underbrace{\pi(h_1) T(h_2 | h_1)}_{P(h_1,h_2)} O(x_1 | h_1) O(x_2 | h_2) ),
  \end{align}
  then we observe that the likelihood is convex in the marginal
  distribution $P(h_1,h_2)$. This allows us to recover $P$, from which $T$
  can be easily reconstructed.

% 7. Now outline the rest of our results.
In \sectionref{piecewise} we generalize this observation to a directed grid
  model and a directed tree model and show that the {\em piecewise
  likelihood} is strictly convex, providing consistent parameter
  estimators for these models (which is new)\reword.
The approach can be extended to log-linear
  models (\sectionref{log-linear}), where the problem remains non-convex,
  however, we expect a reduction in the number of local optima. 
In this regime, we show empirically that constraining the moments of the
  model via posterior-regularization allows us to learn better models.

### 2014-01-26 java/tests/learning/models/HiddenMarkovModelsTest.java
Updated HMM tests.

@@ -336,14 +336,62 @@ public class HiddenMarkovModelsTest {
     log("Fit likelihood: " + model.getLogLikelihood(params, data));
 
     log(marginals);
-
-//    Assert.assertTrue(
-//        MatrixOps.allclose( model.params.pi, hmm3.pi ) );
-//    Assert.assertTrue(
-//        MatrixOps.allclose( model.params.T, hmm3.T ) );
-//    Assert.assertTrue(
-//        MatrixOps.allclose( model.params.O, hmm3.O ) );
   }
+
+  @Test
+  public void testEMTrueFixpoint() {
+    Params hmm3 = model.newParams();
+
+    hmm3.set(piFeature(0), 0.846 );
+    hmm3.set(piFeature(1), 0.154 );
+
+    hmm3.set(tFeature(0, 0), 0.298);
+    hmm3.set(tFeature(0, 1), 0.702);
+    hmm3.set(tFeature(1, 0), 0.106);
+    hmm3.set(tFeature(1, 1), 0.894);
+
+    hmm3.set(oFeature(0,0), 0.357 );
+    hmm3.set(oFeature(0,1), 0.643 );
+    hmm3.set(oFeature(1,0), 0.4292);
+    hmm3.set(oFeature(1,1), 0.5708);
+
+//    Counter<Example> data = model.drawSamples(hmm1, testRandom, (int) 1e6);
+    Counter<Example> data = model.getDistribution(hmm3);
+    Params marginals = model.newParams();
+
+    Params params = model.newParams();
+    params.copyOver(hmm3);
+    double oldLhood = Double.NEGATIVE_INFINITY;
+    for(int i = 0; i < 1000; i++) {
+      // Simple EM
+      marginals.clear();
+      model.updateMarginals(params, data, 1.0, marginals);
+      double diff = params.computeDiff(marginals, null);
+      params.copyOver(marginals);
+
+      double lhood = model.getLogLikelihood(params,data);
+
+      log(outputList(
+              "iter", i,
+              "likelihood", lhood,
+              "diff", diff
+      ));
+      Assert.assertTrue( lhood - oldLhood > -1e-3); // Numerical error.
+      oldLhood = lhood;
+
+      if( diff < 1e-3 ) break;
+    }
+
+    double lhood_true = model.getLogLikelihood(hmm1, data);
+    double lhood_fit = model.getLogLikelihood(params, data);
+
[…]

### 2014-01-27 doc/icml2014/abstract.tex
Updated so many things to figures

\begin{abstract} 
\end{abstract}

\paragraph{Errata}
\begin{enumerate}
  \item The likelihoods as defined are concave, not convex.
  \item Define parameters $\theta$ more uniformly.
\end{enumerate}

### 2014-01-27 doc/icml2014/background.tex
Updated so many things to figures

@@
-\section{Background}
-\label{sec:background}
+\section{Setup}
+\label{sec:setup}
 
-\paragraph{Bottlenecks: Three-view mixture models}
+Let $\sG$ be a discrete graphical model with variables $V$, of which
+  $O \subseteq V$ be observed and $H = V \setminus O$ be hidden.
+Let $\Pa(x) \subset V$ be the parents of a variable $x$ in a directed
+  model, and $\sN(x) \subset V$ be the neighbours of $x$ in an undirected
+  model.
+We will use $o_1, o_2, \cdots \in O$ to denote observed variables and
+  $h_1, h_2, \cdots \in H$ to denote hidden variables.
+For ease of exposition, we assume that the domain of all the observed
+  variables is $[D]$ and the domain of all the hidden variables is
+  $[K]$.
+
+Define $\otimes$ to be the tensor product. In matrix representation, it
+  is the Kronecker product.\expand
+
+\paragraph{Problem statement}
+
+% Statement
+This paper focuses on the problem of parameter estimation:
+We are given $n$ i.i.d.~examples of the observed variables $D = (x^{(1)}, \dots, x^{(n)})$
+where each $x^{(i)} \sim p_{\theta^*}$ for some true parameters $\theta^*$.
+Our goal is to produce a parameter estimate $\hat\theta$ that approximates $\theta^*$.
+
+% Maximum likelihood
+The standard estimation procedure is maximum (marginal) likelihood,
+  \begin{align*}
+    \sL &\eqdef \max_{\theta \in \Re^d} \E_{x} \log p_\theta(x) \\
+        &=      \max_{\theta \in \Re^d} \E_{x} \log \sum_z p_\theta(z),
+  \end{align*}
+  which is statistically efficient but computationally intractable
+  because we must marginalize over latent variables $z$.
+In practice, one uses gradient-based optimization procedures (e.g., EM
+  or L-BFGS) on the marginal likelihood, which can get stuck in local
+  optima.
+
+\paragraph{Estimating three-view mixture models}
 
@@
 
-\todo{Related work on spectral methods and measurements.}
+In general, we call hidden variables with this property ``bottlenecks'',
+  and define them as follows.
+\begin{definition}(Bottleneck)
+  A hidden variable $h$ is said to be a bottleneck if there exists at
+  least three conditionally independent observed variables $o_1, o_2,
+  o_3$. Let $\sB(h)$ be the set of conditionally independent observed
+  variables for $h$.
+\end{definition}
 
-\paragraph{Constrained EM}
-
-Posterior regularization
-Measurements
-Generalized expectation criterion
-
-Kuzman Ganchev.*
+\begin{assumption}
+  We assume that $M_1, M_2, M_3$ are each full rank, and that $\pi \succ
+  0$\reword.
+\end{assumption}

### 2014-01-27 doc/icml2014/intro.tex
Updated so many things to figures

@@
 factorize the moments into parameters, providing strong theoretical guarantees.
-However, these methods are not as universally applicable as EM.
+However, these methods are not as universally applicable as EM, applying to a limited set of models.
+
+\begin{figure*}
+  \centering
+  \input{figures/approach.tikz}
+  \caption{Overview of approach}
+\end{figure*}
 
@@
 In this work, we exploit the spectral method to learn moment constraints
-on the observed variables and show how these constraints can greatly
-simplify the learning problem.
+on the observed variables and show how these constraints lead to consistent parameter estimates for a broad class of graphical models.
+
 % 5. We get moments from third-order tensors from bottlenecks and factorize them into marginals.
-The key idea in
-\citet{anandkumar12moments,anandkumar13tensor} is to study the conditional
-independence factorization structure implied by the model in the form of 
-three-view bottlenecks. These bottlenecks which can be factorized into
-moments involving the latent variables\vague.
+Our approach is illustrated in \figureref{approach}. 
+First, we identify conditionally independent observed variables $x_1,
+  x_2, x_3$ for each hidden variable $h$ (bottlenecks). 
+We use the tensor factorization algorithm of
+  \citet{anandkumar12moments,anandkumar13tensor} to identify the
+  conditional moments.
+Given these factors, we show that for every clique of hidden variables
+  $\sC$, the piecewise likelihood of the observed variables is convex in
+  the marginal distribution. 
+This guarantees that we can recover the true marginal distributions
+  using EM; finally, with some post-processing, we are able to
+  consistently recover the true parameters of the model.
+In \sectionref{piecewise}, we detail our algorithm and describe the
+  settings and technical assumptions in which this approach is guaranteed
+  to work.
+As a particular example, we are able to recover parameters for the
+  grid model depicted in \figureref{approach} under the condition that
+  $O$ is full-rank.
 
-% 6. Provide some intuitive examples - Can I haz HMMs?
-Consider a hidden Markov model, with initial probabilities $\pi$,
-  transitions $T$ and emissions $O$. 
-While \citet{hsu09spectral,anandkumar12moments} provide a consistent estimator
-  for the parameters of this model that inverts $O$ to recover $T$ and
-  $\pi$. We would like to avoid this potentially sensitive procedure via optimization.
-The log-likelihood of this model, 
-  \begin{align}
-    \sL(\pi, T, O) &= \sum_{x\in\sD} \log( \sum_{h} \pi(h_1) O(x_1 | h_1) T(h_2 | h_1) O(x_2 | h_2) \cdots ),
-  \end{align}
-  is not convex, even if we knew any one of the parameters.
-However, intuitively, if we knew $O$ and found that each symbol was
-  produced by exactly one state, learning $T$ and $\pi$ could be trivial
-  accomplished via counting. 
-In general, $O$ still informs us about the states in the observation
-  sequence. If we considered instead the piecewise likelihood, 
-  \begin{align}
-    \sL_{\textrm{p}}(\pi, T, O) &= \sum_{x\in\sD} \log( \sum_{h_1, h_2} \underbrace{\pi(h_1) T(h_2 | h_1)}_{P(h_1,h_2)} O(x_1 | h_1) O(x_2 | h_2) ),
-  \end{align}
-  then we observe that the likelihood is convex in the marginal
-  distribution $P(h_1,h_2)$. This allows us to recover $P$, from which $T$
-  can be easily reconstructed.
-
-% 7. Now outline the rest of our results.
-In \sectionref{piecewise} we generalize this observation to a directed grid
-  model and a directed tree model and show that the {\em piecewise
-  likelihood} is strictly convex, providing consistent parameter
-  estimators for these models (which is new)\reword.
-The approach can be extended to log-linear
-  models (\sectionref{log-linear}), where the problem remains non-convex,
-  however, we expect a reduction in the number of local optima. 
-In this regime, we show empirically that constraining the moments of the
-  model via posterior-regularization allows us to learn better models.
+For undirected log-linear latent variable models, this approach allows us to 
+  recover the marginals for each clique, given which the learning
+  problem is convex (\sectionref{undirected}).
+Finally, our approach gracefully extends to more general model families
+  in which we are unable to recover the marginal distributions for all
+  cliques; in such scenarios, we use posterior regularization to constrain
+  EM. 
+Empirically, we find that this approach leads to better solutions
+  (\sectionref{experiments}).

### 2014-01-27 doc/icml2014/measurements.tex
Updated so many things to figures

@@
 Statement: consistency.
-
-\paragraph{General latent-variable models with bottlenecks}
-
-Our learning algorithm for three-view mixture models relies on the fact that
-the sufficient statistics of the exponential family
-can be expressed as a function of the latent moments.
-Of course, these conditions do not hold for general log-linear models.
-Rather than falling back on plain local optimization of the likelihood in these
-cases, we would still like
-to somehow use the \emph{partial information} offered by the latent moments.
-
-Partial information is available when the following holds:
-%\begin{property}
-%  There exists some latent variable $h_{j_0} \in [\nh]$ (the bottleneck)
-%  and three conditionally independent observed variables $x_{j_1},x_{j_2},x_{j_3}$ (views) such that
-%  each conditional mean matrix $\E[\phix(x_{j}) \mid h_{j_0}] \in \R^{\nphix \times \nh}$ ($j \in \{ j_1, j_2, j_3 \}$) has full column rank $\nh$
-%  for some transformation $\phix$.
-%\end{property}
-
-Optimization.

### 2014-01-27 doc/icml2014/piecewise_notes.tex
Updated so many things to figures

\section{Parameter Estimation with Convex Piecewise Likelihoods}
\label{sec:piecewise}

Key idea: don't ride the MoM train all the way, rather let it do some pre-processing to make your job easier.

\paragraph{Example 1: Improving statistical efficiency by averaging over trigrams}

The algorithm described in \citet{anandkumar12moments} looks only at the
  first three symbols (words) in a sequence (sentence). 
The algorithm can trivially be made to use the data more efficiently
  by using every trigram in a sentence, i.e. $\hat M_{123}
  = \sum_{x\in\sD} \sum_{i = 1}^{|x| - 2} x_{i} \otimes x_{i+1} \otimes
  x_{i+2}$. However, in this case, $\hat M_{123} \to M_{123}$, where 
\begin{align}
  M_{123} &= \sum_{h} \E[(\pi + T \pi + T^2 \pi \cdots T^{|x|-2} \pi)]_h M_1 \otimes O \otimes OT.
\end{align}

Using the tensor factorization algorithm, we recover $O$ and $OT$,
  ignoring $M_1$. From $O$, it is easy to estimate $\pi$ using maximum
  likelihood:
\begin{align}
  \hat \pi &= \frac{1}{|\sD|} \sum_{x\in\sD} O^T_{x_1}.
\end{align}

\paragraph{Example 2: Recovering $T$ from piecewise likelihood}

Inverting $O$ to get $T$ in the above procedure can be dangerous when $O$ is poorly conditioned. 
Alternative: solving the piecewise likelihood,
\begin{align}
  \sL_{p}(\pi, T, O) &= \sum_{x\in\sD} \log( \sum_{h_1, h_2} \underbrace{\pi(h_1) T(h_2 | h_1)}_{P(h_1,h_2)} O(x_1 | h_1) O(x_2 | h_2) ),
\end{align}
Because we know $O$, the likelihood is convex in $P$. Once we have recovered $P^*$ from the above optimization problem, we can recover $T$ as follows:
$T(h_1, h_2) = \frac{P(h_1, h_2}{\sum_{h_2} P(h_1, h_2)}$

It is necessary that $\sL_p$ be {\em strictly convex} to for $P^*$ to be unique \findcite{IBM-1 problem}, as we will prove below,

\begin{theorem}
  $\sL_p$ is strictly convex. The hessian $\grad^2 \sL_p$ is full rank. 
\end{theorem}
\begin{proof}

  We can write out the Hessian as,
\begin{align}
  \grad^2 \sL_p &= \E[ \sum_{x_1, x_2} \frac{ (O_{x_1} \otimes O_{x_2}) (O_{x_1} \otimes O_{x_2})^T }{(O_{x_1}^T P O_{x_2})^2} ].
\end{align}

  Incorporating expectations (complicated expression),
\begin{align}
  \grad^2 \sL_p &= \E[ \sum_{h_1, h_2} \frac{ (O_{x_1} \otimes O_{x_2}) (O_{x_1} \otimes O_{x_2})^T }{(O_{x_1}^T P O_{x_2})^2} ].
\end{align}

Firstly, this is full-rank. Secondly, we can show a lower bound on the singular value as follows.
\end{proof}

\begin{corollary}(Asymptotic efficiency)
  The estimator has so and so efficiency.
\end{corollary}
\begin{proof}
  From the delta method: depends on $\Sigma^-1$, which is basically $\det \Sigma$.
\end{proof}

\paragraph{Example 3: Grid model}

Next, consider the grid model shown in \figureref{grid}. Let us
assume every emission has same parameters $O$ and every transition has
$T$.

Each hidden
node has two observations; conditioning on the hidden node, the two
observations and any third observation are independent, and thus we can
get a consistent estimate for $O$.

Once again, we can write down a piecewise likelihood,
\begin{align}
  \sL_{p}(\pi, T, O) &= \sum_{x\in\sD} \log( \sum_{h_1, h_2, h_3, h_4} P(h_1, h_2, h_3, h_4).
\end{align}

The same approach as above holds; all that remains to be shown is that the model is indeed identifiable,

\begin{theorem}(Identifiability of Grid Model)
  $\sL_p$ is strictly convex. The hessian $\grad^2 \sL_p$ is full rank. 
\end{theorem}
\begin{proof}
  We can write out the Hessian as,
\begin{align}
  \grad^2 \sL_p &= \E[ \sum_{x_1, x_2} \frac{ (O_{x_1} \otimes O_{x_2}) (O_{x_1} \otimes O_{x_2})^T }{(O_{x_1}^T P O_{x_2})^2} ].
\end{align}

  Incorporating expectations (complicated expression),
\begin{align}
  \grad^2 \sL_p &= \E[ \sum_{h_1, h_2} \frac{ (O_{x_1} \otimes O_{x_2}) (O_{x_1} \otimes O_{x_2})^T }{(O_{x_1}^T P O_{x_2})^2} ].
\end{align}

  Firstly, this is full-rank. Secondly, we can show a lower bound on the singular value as follows.
\end{proof}

\paragraph{Example 5: Tree model}

Finally, we consider a tree model as shown in \figureref{tree}. Let us
assume every emission has same parameters $O$ and every transition has
$T$.

\todo{Showing identifiability is a bit trickier in this model. Also, the
algorithm is unlikely to scale.}

### 2014-01-27 doc/icml2014/undirected.tex
Updated so many things to figures

\section{Consistent Parameter Estimation for Undirected Latent Variable Log-linear Models}
\label{sec:undirected}

In this section, we will show that we can extend the results to undirected graphical models.

Let $z$ be a collection of variables indexed by $V$ (that is, $z = \{z_j\}_{j \in V}$).
Suppose that some of the variables are observed ($x = \{ z_j \}_{j \in O}$, where $O \subset V$) 
and the rest are latent ($h = \{ z_j \}_{j \in H}$, where $H = V \backslash O$).
Let $C \subset 2^V$ be a collection of subsets of variables,
and let $\phi_c(z_c) \in \Re^d$ be the feature map for each $c \in C$.
In this paper, we will consider both directed and undirected models over $z$.
In the undirected case, we define the standard globally-normalized exponential family model:
\begin{align}
  \label{eqn:undirectedSetup}
  p_\theta(z) = \exp\{ \phi(z)^\top\theta - A(\theta) \},
\end{align}
where $\phi(z) \eqdef \sum_{c \in C} \phi_c(z_c) \in \Re^d$ is the global feature vector, $\theta \in \Re^d$ is the parameter vector,
and $A(\theta) = \log \int \exp\{\phi(z)^\top\theta\} dz$ is the global log-partition function.
To some extent, our methods only depend on the conditional independence
structure of the models.

\paragraph{Moment Constraints}

In the previous section, we saw that we could learn moments $P(x1,x2)$ given conditions. This makes the problem convex, boom.

\paragraph{Examples}

\paragraph{General latent-variable models with bottlenecks}

Our learning algorithm for three-view mixture models relies on the fact that
the sufficient statistics of the exponential family
can be expressed as a function of the latent moments.
Of course, these conditions do not hold for general log-linear models.
Rather than falling back on plain local optimization of the likelihood in these
cases, we would still like
to somehow use the \emph{partial information} offered by the latent moments.

Partial information is available when the following holds:
%\begin{property}
%  There exists some latent variable $h_{j_0} \in [\nh]$ (the bottleneck)
%  and three conditionally independent observed variables $x_{j_1},x_{j_2},x_{j_3}$ (views) such that
%  each conditional mean matrix $\E[\phix(x_{j}) \mid h_{j_0}] \in \R^{\nphix \times \nh}$ ($j \in \{ j_1, j_2, j_3 \}$) has full column rank $\nh$
%  for some transformation $\phix$.
%\end{property}

Optimization.

### 2014-01-27 java/src/learning/experiments/PiecewiseGridRecovery.java
Now solve grid in a piecewise fashion

@@ -64,6 +64,35 @@ public class PiecewiseGridRecovery implements Runnable {
     return P;
   }
 
+  public double[][] getEdgeParameters(DirectedGridModel model, DirectedGridModel.Parameters params) {
+    double[][] P = new double[K][K];
+
+    for(int h1 = 0; h1 < K; h1++)
+      for(int h2 = 0; h2 < K; h2++)
+          // Sum over the h1
+          P[h1][h2] += params.weights[model.pi(h1)] * params.weights[model.t(h1, h2)];
+
+    assert MatrixOps.equal(MatrixOps.sum(P), 1.0);
+
+    return P;
+  }
+
+  public double[][][] getVStructureParameters(DirectedGridModel model, DirectedGridModel.Parameters params) {
+    double[][][] P = new double[K][K][K];
+
+    for(int h1 = 0; h1 < K; h1++)
+      for(int h2 = 0; h2 < K; h2++)
+        for(int h3 = 0; h3 < K; h3++)
+          for(int h4 = 0; h4 < K; h4++)
+            // Sum over the h1
+            P[h2][h3][h4] += params.weights[model.pi(h1)] * params.weights[model.t(h1, h2)]
+                    * params.weights[model.t(h1, h3)] * params.weights[model.tC(h2, h3, h4)];
+
+    assert MatrixOps.equal(MatrixOps.sum(P), 1.0);
+
+    return P;
+  }
+
   /**
    * grad2 = o1 x o2 x o3 x o4
    * @param model
@@ -136,6 +165,112 @@ public class PiecewiseGridRecovery implements Runnable {
     return new SimpleMatrix(H);
   }
 
+  public SimpleMatrix getVStructureHessian(DirectedGridModel model, DirectedGridModel.Parameters trueParams,
+                                          double[][][] P, Counter<Example> data) {
+    double[][] H = new double[K*K*K][K*K*K];
+
+    double[][] O = trueParams.getO();
+
+    for(Example ex: data) {
+      double weight = data.getFraction(ex);
+
+      int x2 =  ex.x[model.oIdx(0,1,0)];
+      int x2_ = ex.x[model.oIdx(0,1,1)];
+      int x3 =  ex.x[model.oIdx(1,0,0)];
+      int x3_ = ex.x[model.oIdx(1,0,1)];
+      int x4 =  ex.x[model.oIdx(1,1,0)];
+      int x4_ = ex.x[model.oIdx(1,1,1)];
+
+      double z = 0.;
[…]

### 2014-01-28 doc/icml2014/abstract.tex
Updated background

@@
 \begin{abstract}
+
+Parameter estimation for latent variable models is typically
+  performed by optimizing the log-likelihood, which is prone to local
+  optima due to its non-convexity. 
+In this paper, we extend recent advances in the method of moments 
+  to provide consistent parameter estimates for a general family of
+  directed and undirected graphical models, including grid models
+  (\figureref{approach}) and latent tree models.
+The key idea in our approach is that, given the observation potentials
+  between observed and latent variables, the marginal distribution of
+  every clique is recoverable from the corresponding partial moments.
+We characterize the polynomial sample complexity for models in this
+  family.
+For models outside our family, we show how our approach can constrain
+  parameter search by providing partial information via posterior
+  regularization. 
+Finally, we provide empirical intuition via experiments.
+
 \end{abstract}

### 2014-01-28 doc/icml2014/background.tex
Updated background

@@
 For ease of exposition, we assume that the domain of all the observed
-  variables is $[D]$ and the domain of all the hidden variables is
-  $[K]$.
+  variables is $[d]$ and the domain of all the hidden variables is
+  $[k]$.
 
-Define $\otimes$ to be the tensor product. In matrix representation, it
-  is the Kronecker product.\expand
+Define $\otimes$ to be the tensor product.
+To denote the moments of observed variables, we will use the notation
+\begin{align*}
+  M_i &\eqdef \Pr(x_i), &
+  M_{ij} &\eqdef \Pr(x_i, x_j), &
+  M_{ijk} &\eqdef \Pr(x_i, x_j, x_k),
+\end{align*}
+and so on.
+We will view $M_i, M_{ij}, M_{ijk}, \cdots$ as tensors of
+  order 1, 2, 3, \ldots in $\Re^D, \Re^{D\times D}, \Re^{D \times
+  D \times D}$.
+For the moments of hidden variables, we will use analogous notation
+\begin{align*}
+  P_i &\eqdef \Pr(h_i), &
+  P_{ij} &\eqdef \Pr(h_i, h_j), &
+  P_{ijk} &\eqdef \Pr(h_i, h_j, h_k).
+\end{align*}
 
@@
 This paper focuses on the problem of parameter estimation:
-We are given $n$ i.i.d.~examples of the observed variables $D = (x^{(1)}, \dots, x^{(n)})$
+We are given $n$ i.i.d.~examples of the observed variables $\sD = (x^{(1)}, \dots, x^{(n)})$
 where each $x^{(i)} \sim p_{\theta^*}$ for some true parameters $\theta^*$.
@@
 
-\citet{anandkumar12moments} present an algorithm to solve three-view mixture models.
+\begin{figure}[t]
+  \label{fig:three-view}
+  \centering
+  \input{figures/three-view.tikz}
+  \caption{A three-view mixture model}
+\end{figure}
 
-Assume means $M_1, M_2, M_3$. Observe that the moments of the model are $M_{123}$.
+Recently, \citet{anandkumar12moments} presented a consistent algorithm
+  to learn the parameters of a three-view mixture model
+  (\figureref{three-view}) that sidesteps the local optima problem
+  altogether. 
+In the model, we have three observed variables $x_1, x_2, x_3 \in [d]$
+  that are independent conditioned on the latent variable $h \in [k]$. 
+  The parameters of this model are the conditional means $O^{(i)} \eqdef
+  \E[x_i | h] \in \Re^{D \times K}$ for $i \in \{1,2,3\}$ and the mixture probabilities $\pi
+  \in \Delta_{k-1}$.
+  The key idea is that the observed moments $M_{12}, M_{23}, M_{13}$ and
+  $M_{123}$ admit a factorization in terms of parameters;
+\begin{align*}
+  M_{ij} &= \sum_h \pi_h O^{(i)}_h \otimes O^{(j)}_h & \forall i,j \in \{1,2,3\} \\
+  M_{123} &= \sum_h \pi_h O^{(1)}_h \otimes O^{(2)}_h \otimes O^{(3)}_h.
+\end{align*}
 
-Applying a whitening transformation and get a orthogonal tensor decomposition.
-
-Solved using the robust tensor power method \citet{anandkumar13tensor}.
-
-In general, we call hidden variables with this property ``bottlenecks'',
-  and define them as follows.
-\begin{definition}(Bottleneck)
-  A hidden variable $h$ is said to be a bottleneck if there exists at
-  least three conditionally independent observed variables $o_1, o_2,
-  o_3$. Let $\sB(h)$ be the set of conditionally independent observed
-  variables for $h$.
-\end{definition}
+\citet{anandkumar13tensor} presented a robust tensor power method, which
+  we call $\TensorFactorize$, which converts consistent estimates of the observed moments 
+  into consistent estimates of the parameters,
+  $\pi, M_1, M_2, M_3$ (up to permutation of the columns), provided
+  the following assumption holds;
 
 \begin{assumption}
-  We assume that $M_1, M_2, M_3$ are each full rank, and that $\pi \succ
-  0$\reword.
+Assume the conditional moments $M_1, M_2, M_3$ have full column rank
+  $k$, and that $\pi \succ 0$.
 \end{assumption}
+
+In brief, the algorithm first computes a whitening matrix $W \in
+  \Re^{d \times k}$ such that $W^T M_{12} W = I_{k \times k}$,
+  and uses it to transform $M_{123}$ into a symmetric orthogonal tensor.  
+Then a robust tensor power method is applied to extract the eigenvectors
+  of the whitened $M_{123}$; unwhitening yields the columns of $M_{3}$.
+$M_1$ and $M_2$ can then be recovered with some algebra.
+
+\todo{Mention {\em method of moments} somewhere?}

### 2014-01-28 doc/icml2014/directed.tex
Fleshed out a bunch for linear algebra in directed

\section{Consistent Parameter Estimation in Directed Models}
\label{sec:directed}

The tensor factorization method attacks the heart of the non-convexity
  in latent variable models.
The technique is central in solving several other problems with more
  complex correlation structure, for example latent Dirichlet allocation
  \citep{anandkumar12lda}, mixture of Gaussians \citep{hsu13spherical},
  etc.\findcite{more things}.

We now show that once we recover the observation
  potentials, identifying other clique potentials in a directed graphical
  model can be consistently estimated in a systematic manner.
In section \sectionref{undirected} we generalize our results to
  undirected log-linear models.

\subsection{Example: Directed grid model}

\begin{figure}
  \centering
  \input{figures/grid.tikz}
  \caption{A directed grid model.}
  \label{fig:grid}
\end{figure}

Let us illustrate our approach with an example. 
Consider the directed grid model shown in \figureref{grid} which
  captures the different dependency structures possible in a Bayesian network.
The model has eight observed variables $x^a_1, x^b_1 \cdots, x^a_4, x^b_4$ and four
  hidden variables $h_1, \cdots, h_4$.
The parameters of this model are the conditional probability tables
  $\pi, T, V$ and $O$; for the sake of simplicity, we assume that all
  the observation potentials are the same, i.e. $O \eqdef \Pr(x^\cdot_i
  | h_i)$ for all $i$, and that $\Pr(h_2 | h_1) = \Pr(h_3 | h_1) = T$.
Finally, we assume $O$ and $T$ have full column rank.

\paragraph{Estimating $O$}

Note that the observed variables $x^a_1, x^b_1, x^a_2$ are
  conditionally independent given $h_1$, allowing us to use
  \TensorFactorize as reviewed in \sectionref{setup}, to recover
  the moments $O$.
We can aggregate statistics from additional sets of independent observed
  variables, e.g. $\{x^a_2, x^b_2, x^a_3\}$, to arrive at a better
  estimate of $O$.

%Intuitively, we expect the observations to tell us about the underlying
%  hidden state, which should make learning each factor easier. 
%For example, when $O = I$ the observed variables
%  tell us exactly which hidden state we are in; the optimal procedure to
%  estimate factors is simply counting the co-occurrences of the hidden
%  states.

\paragraph{Estimating $\pi$}

We can relate the state of $h_1$ to the moments of $x^a_1$; $M_1
  \eqdef \Pr(x^a_1) = O \pi$. From \assumptionref{full-rank}, $O$ has full
  column rank and thus can be inverted to recover $\pi$; $\pi = \pinv O
  M_1$.

\paragraph{Estimating $T$}

Similarly, the we can write down the moments of $x^a_1, x^a_2$ in terms
of the pairwise marginals $P_{12} \eqdef \Pr(h_1, h_2)$ and solve for
$P_{12}$;
\begin{align*}
  M_{12} &= O P_{12} O^T \\
  P_{12} &= \pinv O M_{12} (\pinv O)^T.
\end{align*}

$T$ can be recovered from the $P_{12}$ by suitable normalization.

\paragraph{Estimating $V$}

Finally, we can estimate $V$ from the third-order moments $M_{234}$ as
follows,
\begin{align*}
  M_{234} &= P_{234}(O, O, O) \\
  P_{234} &= M_{234}(\pinv O, \pinv O, \pinv O).
\end{align*}

\subsection{General algorithm}

\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\begin{algorithm}
  \caption{\LearnFactors}
  \label{algo:directed}
  \begin{algorithmic}
    \REQUIRE A graphical model $\sG$ satisfying \textbf{(P1)}, \textbf{(P2)}, data $\sD$
    \ENSURE Parameters $\theta$ for $\sG$

      \FOR{$h \in H$} 
        \STATE Apply \TensorFactorize for each bottleneck $h$ and learn parameters $\Pr(x \mid h) ~ \forall x \in \sB(h)$.

%    \COMMENT{Recover observation potentials $O$ using bottlenecks}
      \ENDFOR
%      \COMMENT{\textbf{Step 2:} Recover clique potentials from the piecewise likelihood.}
      \FOR{every clique $\sC = \{h_1, \cdots, h_m\} \in \sG$} 
      \STATE Let $x_\sC = \{x_1, \cdots, x_m\}$.
      \STATE $P_\sC = M_{x_\sC}( \pinv O_{x_1}, \cdots, \pinv O_{x_m} )$.
%      Run expectation-maximization to convergence on the piecewise likelihood \eqref{eqn:piecewise}, over data $\{\vec x_\sC : x \in \sD\}$
      \ENDFOR
  \end{algorithmic}
\end{algorithm}

\algorithmref{directed} presents the natural extension of this approach.
Let us now describe the general properties for which the algorithm,
  $\LearnFactors$ works.

For the first step of the algorithm, we wish to identify conditional
  moments between every hidden variable and atleast one observed
  variable.
We call hidden variables with this property ``bottlenecks'',
  and define them as follows.
\begin{definition}(Bottleneck)
  A hidden variable $h$ is said to be a bottleneck if there exists at
    least three conditionally independent observed variables $o_1, o_2,
    o_3$. 
  Let $\vec x_\sB(h)$ be the set of conditionally independent observed
    variables for $h$.
\end{definition}

We will further assume that each observed variable has a single parent. 
This leads to the following two properties.
\begin{property}
  \label{prop:unique-parent}
  Every observed variable $o$ has exactly one parent, $|\Pa(o)| = 1$.
\end{property}
\begin{property}
  \label{prop:bottleneck}
  Every hidden variable is a bottleneck.
\end{property}

We note that \propertyref{unique-parent} can be relaxed if another
  algorithm is used to recover the conditional moments. For example,
  \citet{halpern13noisyor} propose an algorithm that can recover the
  conditional moments for a broad, but less general class of noisy-or
  networks.
\propertyref{bottleneck} can also be relaxed if factors share the same
  parameters. 
For example, because we can identify the observation potential from just
  $x^a_1, x^b_1$ and $x^a_2$, we do not require that $h_2, h_3$ or $h_4$
  be bottlenecks, and can correspondingly omit the observations $x^b_2,
  x^b_3$ and $x^b_4$.

Given \propertyref{unique-parent} and \propertyref{bottleneck}, step
  1 of \algorithmref{directed} is guaranteed to recover potentials
  $O^{(h)}_{x}$ for every hidden variable $h$ and observed variable $x \in
  \vec x_\sB(h)$.
\propertyref{bottleneck} guarantees that for every clique $\sC
  = \{h_1, \cdots, h_m\}$ in the graphical model $\sG$, there exists at
  least one bottleneck observable that depends on exactly one $h_i \in
  \sC$; let $x_{h_i}$ be that observed variable\verify. 
Finally, define $\vec x_\sC = \{x_{h_1}, \cdots x_{h_m}\}$ to be the
  participating observed variables in the clique $\sC$.

Finally, we have the following moment condition,
\begin{align*}
  M_{x_\sC} &= P_{\sC}(O_{x_{h_1}}, \cdots, O_{x_{h_m}}) \\
  P_{\sC} &= M_{x_\sC}(\pinv O_{x_{h_1}}, \cdots, \pinv O_{x_{h_m}}).
\end{align*}

This algorithm is simple and intuitive. Next, we characterize its sample complexity.

\begin{lemma}(Sample complexity for $\sC$)
We require some number of samples to learn the factor for $\sC$.
\end{lemma}

This gives us our final result for a graphical model in our model
family.
\begin{theorem}(Learnability of Bottleneck Graphs)
The parameters for a directed graphical $\sG$ with tree-width $k$
satisfying \propertyref{unique-parent}, \propertyref{bottleneck} and
\assumptionref{full-rank} can be learned with $O(\cdot)$ samples.
\end{theorem}

\paragraph{Aggregating observations}

When we have multiple observed variables for a hidden variable, it is
  always beneficial to aggregate over them.
Note that $O_h$ will have rank atleast that of $O$; this encodes the
  intuitive fact that we might be able to identify $h$ from multiple
  observed variables even if we can not identify it from a single
  observed variable. 
Suppose $x_\sB(h) = \{x_1, x_2, \cdots, x_l\}$, let $O_h \in
  \Re^{D^l \times K}$ be an
  aggregated observation potential, such that $O_h(x_1, x_2, \cdots,
  x_l) = P(x_1|h) P(x_2|h) \cdots P(x_l|h)$.
Given \assumptionref{full-rank}, $O_h$ have rank $K$.
After step 1, $O_h$ is a known quantity.

\subsection{More examples}

We will now instantiate our algorithm for several examples.

\paragraph{Hidden Markov Model}

\paragraph{Latent Tree Structure}

\paragraph{Directed Grid Model}

### 2014-01-28 java/scabby.py

@@ -172,6 +172,9 @@ def plot_many(lines, xlabel='x', ylabels=['y']):
 def start_plot(**kwargs):
     import matplotlib.pyplot as plt 
     fig = plt.figure()
+    plt.rc('text', usetex = True)
+    plt.rc('font', family = "Serif")
+    plt.rc('axes', labelsize = 24)
     ax = fig.gca()
     ax.set( **kwargs )
     return fig, ax

### 2014-01-28 doc/icml2014/intro.tex
Updated background
Updates to paper

@@
 
-\begin{figure*}
+\begin{figure}[t]
+  \label{fig:approach}
   \centering
-  \input{figures/approach.tikz}
+\begin{tikzpicture}
+
+% Import grid.
+%\node[scale=0.9] (step0) at (0,0) {\input{figures/grid.tikz}};
+\node[scale=0.9] (step1) at (0cm,0cm) {\input{figures/grid-outline.tikz}};
+\node[scale=0.75, below=0.5 cm of step1.south] { \textbf{Step 1:} Identify bottlenecks};
+\node[scale=0.9] (step2) at (5cm,0cm) {\input{figures/factors.tikz}};
+\node[scale=0.75, below=0.5 cm of step2.south] { \textbf{Step 2:} Optimize piecewise likelihoods};
+
+\draw[-latex] (step1) -- (step2);
+
+\end{tikzpicture}
   \caption{Overview of approach}
-\end{figure*}
+\end{figure}
 

### 2014-01-29 doc/icml2014/background.tex
Fleshed out a bunch for linear algebra in directed
Added words and random mumbo-jumbo

@@
 \begin{align*}
-  M_i &\eqdef \Pr(x_i), &
-  M_{ij} &\eqdef \Pr(x_i, x_j), &
-  M_{ijk} &\eqdef \Pr(x_i, x_j, x_k),
+  \mO_i &\eqdef \Pr(x_i), &
+  \mO_{ij} &\eqdef \Pr(x_i, x_j), &
+  \mO_{ijk} &\eqdef \Pr(x_i, x_j, x_k),
 \end{align*}
 and so on.
-We will view $M_i, M_{ij}, M_{ijk}, \cdots$ as tensors of
+We will view $\mO_i, \mO_{ij}, \mO_{ijk}, \cdots$ as tensors of
   order 1, 2, 3, \ldots in $\Re^D, \Re^{D\times D}, \Re^{D \times
@@
 \begin{align*}
-  P_i &\eqdef \Pr(h_i), &
-  P_{ij} &\eqdef \Pr(h_i, h_j), &
-  P_{ijk} &\eqdef \Pr(h_i, h_j, h_k).
+  \mH_i &\eqdef \Pr(h_i), &
+  \mH_{ij} &\eqdef \Pr(h_i, h_j), &
+  \mH_{ijk} &\eqdef \Pr(h_i, h_j, h_k).
 \end{align*}
@@
   \in \Delta_{k-1}$.
-  The key idea is that the observed moments $M_{12}, M_{23}, M_{13}$ and
-  $M_{123}$ admit a factorization in terms of parameters;
+  The key idea is that the observed moments $\mO_{12}, \mO_{23}, \mO_{13}$ and
+  $\mO_{123}$ admit a factorization in terms of parameters;
 \begin{align*}
-  M_{ij} &= \sum_h \pi_h O^{(i)}_h \otimes O^{(j)}_h & \forall i,j \in \{1,2,3\} \\
-  M_{123} &= \sum_h \pi_h O^{(1)}_h \otimes O^{(2)}_h \otimes O^{(3)}_h.
+  \mO_{ij} &= \sum_h \pi_h O^{(i)}_h \otimes O^{(j)}_h & \forall i,j \in \{1,2,3\} \\
+  \mO_{123} &= \sum_h \pi_h O^{(1)}_h \otimes O^{(2)}_h \otimes O^{(3)}_h.
 \end{align*}
@@
   into consistent estimates of the parameters,
-  $\pi, M_1, M_2, M_3$ (up to permutation of the columns), provided
+  $\pi, \mO_1, \mO_2, \mO_3$ (up to permutation of the columns), provided
   the following assumption holds;
@@
 \begin{assumption}
-Assume the conditional moments $M_1, M_2, M_3$ have full column rank
+\label{asm:full-rank}
+Assume the conditional moments $\mO_1, \mO_2, \mO_3$ have full column rank
   $k$, and that $\pi \succ 0$.
@@
 In brief, the algorithm first computes a whitening matrix $W \in
-  \Re^{d \times k}$ such that $W^T M_{12} W = I_{k \times k}$,
-  and uses it to transform $M_{123}$ into a symmetric orthogonal tensor.  
+  \Re^{d \times k}$ such that $W^T \mO_{12} W = I_{k \times k}$,
+  and uses it to transform $\mO_{123}$ into a symmetric orthogonal tensor.  
 Then a robust tensor power method is applied to extract the eigenvectors
-  of the whitened $M_{123}$; unwhitening yields the columns of $M_{3}$.
-$M_1$ and $M_2$ can then be recovered with some algebra.
+  of the whitened $\mO_{123}$; unwhitening yields the columns of $\mO_{3}$.
+$\mO_1$ and $\mO_2$ can then be recovered with some algebra.
 

### 2014-01-29 doc/icml2014/directed.tex
Updates to paper
Added words and random mumbo-jumbo

@@
 
-We can relate the state of $h_1$ to the moments of $x^a_1$; $M_1
+We can relate the state of $h_1$ to the moments of $x^a_1$; $\mO_1
   \eqdef \Pr(x^a_1) = O \pi$. From \assumptionref{full-rank}, $O$ has full
   column rank and thus can be inverted to recover $\pi$; $\pi = \pinv O
-  M_1$.
+  \mO_1$.
 
@@
 Similarly, the we can write down the moments of $x^a_1, x^a_2$ in terms
-of the pairwise marginals $P_{12} \eqdef \Pr(h_1, h_2)$ and solve for
-$P_{12}$;
+of the pairwise marginals $\mH_{12} \eqdef \Pr(h_1, h_2)$ and solve for
+$\mH_{12}$;
 \begin{align*}
-  M_{12} &= O P_{12} O^T \\
-  P_{12} &= \pinv O M_{12} (\pinv O)^T.
+  \mO_{12} &= O \mH_{12} O^T \\
+  \mH_{12} &= \pinv O \mO_{12} (\pinv O)^T.
 \end{align*}
 
-$T$ can be recovered from the $P_{12}$ by suitable normalization.
+$T$ can be recovered from the $\mH_{12}$ by suitable normalization.
 
@@
 
-Finally, we can estimate $V$ from the third-order moments $M_{234}$ as
+Finally, we can estimate $V$ from the third-order moments $\mO_{234}$ as
 follows,
 \begin{align*}
-  M_{234} &= P_{234}(O, O, O) \\
-  P_{234} &= M_{234}(\pinv O, \pinv O, \pinv O).
+  \mO_{234} &= \mH_{234}(O, O, O) \\
+  \mH_{234} &= \mO_{234}(\pinv O, \pinv O, \pinv O).
 \end{align*}
@@
       \STATE Let $x_\sC = \{x_1, \cdots, x_m\}$.
-      \STATE $P_\sC = M_{x_\sC}( \pinv O_{x_1}, \cdots, \pinv O_{x_m} )$.
+      \STATE $\mH_\sC = \mO_{x_\sC}( \pinv O_{x_1}, \cdots, \pinv O_{x_m} )$.
 %      Run expectation-maximization to convergence on the piecewise likelihood \eqref{eqn:piecewise}, over data $\{\vec x_\sC : x \in \sD\}$
@@
 \begin{align*}
-  M_{x_\sC} &= P_{\sC}(O_{x_{h_1}}, \cdots, O_{x_{h_m}}) \\
-  P_{\sC} &= M_{x_\sC}(\pinv O_{x_{h_1}}, \cdots, \pinv O_{x_{h_m}}).
+  \mO_{x_\sC} &= \mH_{\sC}(O_{x_{h_1}}, \cdots, O_{x_{h_m}}) \\
+  \mH_{\sC} &= \mO_{x_\sC}(\pinv O_{x_{h_1}}, \cdots, \pinv O_{x_{h_m}}).
 \end{align*}
@@
   aggregated observation potential, such that $O_h(x_1, x_2, \cdots,
-  x_l) = P(x_1|h) P(x_2|h) \cdots P(x_l|h)$.
+  x_l) = \mH(x_1|h) \mH(x_2|h) \cdots \mH(x_l|h)$.
 Given \assumptionref{full-rank}, $O_h$ have rank $K$.
 After step 1, $O_h$ is a known quantity.
-
-\subsection{More examples}
-
-We will now instantiate our algorithm for several examples.
-
-\paragraph{Hidden Markov Model}
-
-\paragraph{Latent Tree Structure}
-
-\paragraph{Directed Grid Model}

### 2014-01-29 doc/icml2014/background.tex
Updated directed section

@@
 
-Let $\sG$ be a discrete graphical model with variables $V$, of which
-  $O \subseteq V$ be observed and $H = V \setminus O$ be hidden.
-Let $\Pa(x) \subset V$ be the parents of a variable $x$ in a directed
-  model, and $\sN(x) \subset V$ be the neighbours of $x$ in an undirected
+Let $\sG$ be a discrete graphical model with variables $V$, 
+  let the subset $X \subseteq V$ be observed and 
+  let $H = V \setminus O$ be hidden.
+Let $\Pa(v) \subset V$ be the parents of a variable $v$ in a directed
+  model, and $\sN(v) \subset V$ be the neighbours of $v$ in an undirected
   model.
-We will use $o_1, o_2, \cdots \in O$ to denote observed variables and
-  $h_1, h_2, \cdots \in H$ to denote hidden variables.
-For ease of exposition, we assume that the domain of all the observed
-  variables is $[d]$ and the domain of all the hidden variables is
-  $[k]$.
+We will use $x_1, x_2, \cdots, x_N \in X$ to denote observed variables and
+  $h_1, h_2, \cdots, h_M \in H$ to denote hidden variables.
+For ease of exposition, we assume that $x_i \in \{1, \cdots, d\}$ for
+  all $i$, and that $h_i \in \{1, \cdots, k\}$ for all i.
 
-Define $\otimes$ to be the tensor product.
+We assume that the model $\sG$ is parameterized at every clique $\sC$.
+If $\sG$ is directed, the parameters correspond to
+  the conditional probability tables for each clique.
+If $\sG$ is undirected, the parameters correspond to the clique
+  potentials.
+
+\paragraph{Problem statement}
+
+% Statement
+This paper focuses on the problem of parameter estimation:
+We are given $n$ i.i.d.~examples of the observed variables $\sD
+  = (\vx^{(1)}, \dots, \vx^{(n)})$ where each $\vx^{(i)} \sim p_{\theta^*}$ for
+  some true parameters $\theta^*$.
+Our goal is to produce a parameter estimate $\hat\theta$ that
+  approximates $\theta^*$.
+
+% Maximum likelihood
+The standard estimation procedure is maximum (marginal) likelihood,
+  \begin{align*}
+    \sL &\eqdef \max_{\theta} \sum_{\vx \in \sD} \log p_\theta(\vx) \\
+        &=      \max_{\theta} \sum_{\vx \in \sD} \log \sum_{\vh} p_\theta(\vx,\vh)
+  \end{align*}
+  which is statistically efficient but computationally intractable
+  because we must marginalize over latent variables $z$.
+In practice, one uses gradient-based optimization procedures (e.g., EM
+  or L-BFGS) on the marginal likelihood, which can get stuck in local
+  optima.
+
+\paragraph{Notation}
+
+We use $\otimes$ to denote the tensor product, e.g if $u \in \Re^k,
+  v \in \Re^d$, then $u \otimes v \in \Re^{k \times d}$.
+For a $\ell$-th order tensor $T \in \Re^{d \times \ldots \times
+  d}$ and vectors $x_1, \cdots, x_\ell \in \Re^{d},$, we define 
+\begin{align*}
+  T(x_1, \cdots, x_\ell) 
+      &= \sum_{i_1 = 1}^{d} \cdots \sum_{i_\ell = 1}^{d} 
+            T_{i_1 i_2 \ldots i_l} ({x_1})_{i_1} \cdots ({x_\ell})_{i_\ell}.
+\end{align*}
+Analogously, for a matrices $M_1 \in \Re^{d \times k}, \cdots,
+  M_\ell \in \Re^{d \times k}$, we define
+\begin{align*}
+  T(M_1, \cdots, M_\ell)_{j_1, \ldots, j_l} 
+      &= \sum_{i_1 = 1}^{d} \cdots \sum_{i_\ell = 1}^{d} 
+            T_{i_1 i_2 \ldots i_l} ({M_1})_{i_1j_1} \cdots ({M_\ell})_{i_\ell j_\ell}.
+\end{align*}
+
 To denote the moments of observed variables, we will use the notation
@@
 We will view $\mO_i, \mO_{ij}, \mO_{ijk}, \cdots$ as tensors of
-  order 1, 2, 3, \ldots in $\Re^D, \Re^{D\times D}, \Re^{D \times
-  D \times D}$.
-For the moments of hidden variables, we will use analogous notation
+  order 1, 2, 3, \ldots in $\Re^d, \Re^{d\times d}, \Re^{d \times
+  d \times d},\ldots$.
+For the moments of hidden variables, we will use 
 \begin{align*}
@@
 \end{align*}
-
-\paragraph{Problem statement}
-
-% Statement
-This paper focuses on the problem of parameter estimation:
-We are given $n$ i.i.d.~examples of the observed variables $\sD = (x^{(1)}, \dots, x^{(n)})$
-where each $x^{(i)} \sim p_{\theta^*}$ for some true parameters $\theta^*$.
-Our goal is to produce a parameter estimate $\hat\theta$ that approximates $\theta^*$.
-
-% Maximum likelihood
-The standard estimation procedure is maximum (marginal) likelihood,
-  \begin{align*}
-    \sL &\eqdef \max_{\theta \in \Re^d} \E_{x} \log p_\theta(x) \\
-        &=      \max_{\theta \in \Re^d} \E_{x} \log \sum_z p_\theta(z),
-  \end{align*}
-  which is statistically efficient but computationally intractable
-  because we must marginalize over latent variables $z$.
-In practice, one uses gradient-based optimization procedures (e.g., EM
-  or L-BFGS) on the marginal likelihood, which can get stuck in local
-  optima.
 
@@
 Recently, \citet{anandkumar12moments} presented a consistent algorithm
-  to learn the parameters of a three-view mixture model
-  (\figureref{three-view}) that sidesteps the local optima problem
-  altogether. 
-In the model, we have three observed variables $x_1, x_2, x_3 \in [d]$
-  that are independent conditioned on the latent variable $h \in [k]$. 
-  The parameters of this model are the conditional means $O^{(i)} \eqdef
-  \E[x_i | h] \in \Re^{D \times K}$ for $i \in \{1,2,3\}$ and the mixture probabilities $\pi
-  \in \Delta_{k-1}$.
+  using the method of moments to learn the parameters of a three-view
+  mixture model (\figureref{three-view}), an important case of directed
+  graphical models, that sidesteps the local optima problem altogether. 
+In the model, we have three observed variables $x_1, x_2, x_3$
+  that are independent conditioned on a latent variable $h$.  
+The parameters of this model are the mixture probabilities $\pi \eqdef
+  \Pr(h) \in \Delta_{k-1}$ and the conditional moments $O^{(i)} \eqdef
+  \Pr(x_i \given h) \in \Re^{d \times k}$ for $i \in \{1,2,3\}$.
+
   The key idea is that the observed moments $\mO_{12}, \mO_{23}, \mO_{13}$ and
@@
 \begin{align*}
-  \mO_{ij} &= \sum_h \pi_h O^{(i)}_h \otimes O^{(j)}_h & \forall i,j \in \{1,2,3\} \\
-  \mO_{123} &= \sum_h \pi_h O^{(1)}_h \otimes O^{(2)}_h \otimes O^{(3)}_h.
+  \mO_{ij} &= \sum_h \pi_h {O\oft i}^T_h \otimes {O\oft j}^T_h & \forall i,j \in \{1,2,3\} \\
+  \mO_{123} &= \sum_h \pi_h {O\oft 1}^T_h \otimes {O\oft 2}^T_h \otimes  {O\oft 3}^T_h.
 \end{align*}
@@
   we call $\TensorFactorize$, which converts consistent estimates of the observed moments 
-  into consistent estimates of the parameters,
-  $\pi, \mO_1, \mO_2, \mO_3$ (up to permutation of the columns), provided
-  the following assumption holds;
+  into consistent estimates of the parameters, $\pi, O\oft{1},
+  O\oft{2}, O\oft{3}$ (up to permutation of the columns), provided
+  the following assumption holds:
 
@@
 \label{asm:full-rank}
-Assume the conditional moments $\mO_1, \mO_2, \mO_3$ have full column rank
+Assume the conditional moments $O\oft{1}, O\oft{2}, O\oft{3}$ have full column rank
   $k$, and that $\pi \succ 0$.
@@
 $\mO_1$ and $\mO_2$ can then be recovered with some algebra.
-
-\todo{Mention {\em method of moments} somewhere?}

### 2014-01-29 doc/icml2014/directed.tex
Updated directed section

@@
 
-We now show that once we recover the observation
-  potentials, identifying other clique potentials in a directed graphical
-  model can be consistently estimated in a systematic manner.
-In section \sectionref{undirected} we generalize our results to
-  undirected log-linear models.
+Once we recover the conditional moments $O\oft{i,j} \eqdef \Pr(x_i
+  | h_j)$ for some $i,j$, we present a systematic approach to learn the
+  conditional probability tables for every clique for a class of directed
+  graphical models.
+We begin by illustrating our approach with an example, followed by
+  describing the simple conditions for which the approach works and some
+  theoretical properties.
 
@@
 
-Let us illustrate our approach with an example. 
 Consider the directed grid model shown in \figureref{grid} which
-  captures the different dependency structures possible in a Bayesian network.
+  captures all the different dependency structures possible in
+  a Bayesian network and is hence a good example for intuition\reword.
 The model has eight observed variables $x^a_1, x^b_1 \cdots, x^a_4, x^b_4$ and four
@@
 The parameters of this model are the conditional probability tables
-  $\pi, T, V$ and $O$; for the sake of simplicity, we assume that all
-  the observation potentials are the same, i.e. $O \eqdef \Pr(x^\cdot_i
-  | h_i)$ for all $i$, and that $\Pr(h_2 | h_1) = \Pr(h_3 | h_1) = T$.
+  $\pi \eqdef \Pr(h_1), T \eqdef \Pr(h_2 | h_1) = \Pr(h_3 | h_1),
+  V \eqdef \Pr(h_4 | h_2, h_3)$ and $O \eqdef \Pr(x^a_i | h_i)
+  =  \Pr(x^b_i | h_i)$. 
 Finally, we assume $O$ and $T$ have full column rank.
 
-\paragraph{Estimating $O$}
+\paragraph{Estimating $~O$}
 
@@
   conditionally independent given $h_1$, allowing us to use
-  \TensorFactorize as reviewed in \sectionref{setup}, to recover
-  the moments $O$.
-We can aggregate statistics from additional sets of independent observed
-  variables, e.g. $\{x^a_2, x^b_2, x^a_3\}$, to arrive at a better
-  estimate of $O$.
+  \TensorFactorize, as reviewed in \sectionref{setup}, to recover
+  the $O$.
+We can exploit parameter sharing between the different $x$ to aggregate
+  statistics from additional sets of independent
+  observed variables, or {\em views}, e.g. $\{x^a_2, x^b_2, x^a_3\}$, to
+  arrive at a better estimate of $O$.
 
@@
 
-\paragraph{Estimating $\pi$}
+\paragraph{Estimating $~\pi$}
 
-We can relate the state of $h_1$ to the moments of $x^a_1$; $\mO_1
-  \eqdef \Pr(x^a_1) = O \pi$. From \assumptionref{full-rank}, $O$ has full
-  column rank and thus can be inverted to recover $\pi$; $\pi = \pinv O
-  \mO_1$.
+The moments of $x^a_1$,$\mO_1 \eqdef \Pr(x^a_1)$ are directly related to
+  $\pi$; $\mO_1 = O \pi$. 
+By \assumptionref{full-rank}, $O$ has full column rank and thus can be
+  inverted to recover $\pi$; $\pi = \pinv O \mO_1$.
 
-\paragraph{Estimating $T$}
+\paragraph{Estimating $~T$}
 
-Similarly, the we can write down the moments of $x^a_1, x^a_2$ in terms
-of the pairwise marginals $\mH_{12} \eqdef \Pr(h_1, h_2)$ and solve for
-$\mH_{12}$;
+Similarly, we can write down the moments of $x^a_1, x^a_2$, $\mO_{12}
+  \eqdef \Pr(x^a_1, x^a_2)$ in terms of the pairwise marginals $\mH_{12}
+  \eqdef \Pr(h_1, h_2)$ and solve for $\mH_{12}$;
 \begin{align*}
   \mO_{12} &= O \mH_{12} O^T \\
-  \mH_{12} &= \pinv O \mO_{12} (\pinv O)^T.
+  \mH_{12} &= \pinv O \mO_{12} \pinvt O.
 \end{align*}
@@
 
-Finally, we can estimate $V$ from the third-order moments $\mO_{234}$ as
+Finally, we can estimate $V$ from the third-order moments $\mO_{234} \eqdef \Pr(x^a_2, x^a_3, x^a_4)$ as
 follows,
@@
 
-\renewcommand{\algorithmicrequire}{\textbf{Input:}}
-\renewcommand{\algorithmicensure}{\textbf{Output:}}
-\begin{algorithm}
-  \caption{\LearnFactors}
-  \label{algo:directed}
-  \begin{algorithmic}
-    \REQUIRE A graphical model $\sG$ satisfying \textbf{(P1)}, \textbf{(P2)}, data $\sD$
-    \ENSURE Parameters $\theta$ for $\sG$
+Fundamentally, our algorithm takes advantage of having a view for every
+  hidden variable in a clique to help identify its hidden state. The
+  ideal property we wish our graphical models to possess is the
+  following:
+\begin{property}(Exclusive views)
+  \label{prop:exclusive-views}
+For every clique $\sC$ of $\sG$ and every hidden variable $h \in \sC$,
+  there is some observed variable, $x_{v_h}$, which is conditionally
+  independent of the rest of the clique given $h$ and whose conditional
+  moment $\mOpp{v_h}{h} \eqdef \Pr(x_{v_h} | h)$ is known. 
+We call $x_{v_h}$ the exclusive view of $h$ in $\sC$.
+\end{property}
 
-      \FOR{$h \in H$} 
-        \STATE Apply \TensorFactorize for each bottleneck $h$ and learn parameters $\Pr(x \mid h) ~ \forall x \in \sB(h)$.
+Before we identify a family of models which entail this property, let us
+  show that it is sufficient to learn the parameters of $\sG$.
 
-%    \COMMENT{Recover observation potentials $O$ using bottlenecks}
-      \ENDFOR
-%      \COMMENT{\textbf{Step 2:} Recover clique potentials from the piecewise likelihood.}
-      \FOR{every clique $\sC = \{h_1, \cdots, h_m\} \in \sG$} 
-      \STATE Let $x_\sC = \{x_1, \cdots, x_m\}$.
-      \STATE $\mH_\sC = \mO_{x_\sC}( \pinv O_{x_1}, \cdots, \pinv O_{x_m} )$.
-%      Run expectation-maximization to convergence on the piecewise likelihood \eqref{eqn:piecewise}, over data $\{\vec x_\sC : x \in \sD\}$
-      \ENDFOR
-  \end{algorithmic}
-\end{algorithm}
+Consider any clique $\sC = \{h_{i_1}, \cdots, h_{i_m}\} \in \sG$. Let $x_{v_j}$ be the exclusive
+  view for $h_{i_j}$ in $\sC$ and $\sV = \{x_{v_1}, \cdots, x_{v_m}\}$
+  be the set of views for the clique $\sC$.
+We can write down the moments $\mO_\sV$,
+\begin{align*}
+  \mO_\sV 
+  &\eqdef \Pr(\Sx{\sV}) \\
+  &= \sum_{h_{i_1} = 1}^{d} \cdots \sum_{h_{i_1} = 1}^{d} 
+      \Pr(\Sh{\sC}) \Pr(\Sx{\sV} \given \Sh{\sC}) \\
+      &= \sum_{h_{i_1} = 1}^{d} \cdots \sum_{h_{i_1} = 1}^{d} \Pr(\Sh{\sC}) 
+          \Pr(x_{i_1} | h_{v_1}) \cdots \Pr(x_{i_m} | h_{v_m}) \\
+    &= Z_{\sC}(\mOpp{v_1}{i_1},\cdots,\mOpp{v_m}{i_m}).
+\end{align*}
 
-\algorithmref{directed} presents the natural extension of this approach.
-Let us now describe the general properties for which the algorithm,
-  $\LearnFactors$ works.
+If each $\mOpp{v_j}{i_j}$ has full column rank, then we can recover the
+hidden marginals $Z_\sC$ by inverting the conditional moments,
+\begin{align*}
+  Z_{\sC} &= \mO_\sV(\pinv{\mOpp{v_1}{i_1}},\cdots,\pinv{\mOpp{v_m}{i_m}}).
+\end{align*}
+Finally, the conditional probability tables for $\sC$ can easily be gotten from
+  $Z_\sC$ via conditioning.
 
-For the first step of the algorithm, we wish to identify conditional
-  moments between every hidden variable and atleast one observed
-  variable.
-We call hidden variables with this property ``bottlenecks'',
-  and define them as follows.
+For each $\mOpp{v_j}{i_j}$ to have full column rank, it is equivalent
+  to the following generalized assumption of
+  \assumptionref{full-rank},
+
+\begin{assumption} 
+  \label{asm:full-rank-plus}
+  For every clique $\sC = \{h_{i_1}, \cdots, h_{i_m}\} \in \sG$, the
+  marginal distribution $Z_\sC$ derived by the parameters is full rank
+  for every mode-unfolding of $Z_\sC$.
+\end{assumption}
+
+Intuitively, this assumption says that the marginal probability of every state for any hidden variable is non-zero.
+
+\begin{lemma}
+  \label{lem:goodness}
+
+  Given \assumptionref{full-rank-plus}, (a) the marginal distribution
+  $\Pr(h_i) \succ 0$ for every $h_i$ in the graph and (b)
+  $\mOpp{v_j}{i_j}$ has full column rank.
+\end{lemma}
+\begin{proof}
+  By variable elimination.
+\end{proof}
+
+\paragraph{Required structural properties}
+
+We will now provide a sufficient condition for
+\propertyref{exclusive-views} to hold. 
+Recall that given three observed
+variables, $x_1, x_2, x_3$ that are independent conditioned on $h$,
+\TensorFactorize allows us to recover the conditional moments $\mOp{1},
+\mOp{2}, \mOp{3}$. We define a hidden variable that has this capability to be a bottleneck.
 \begin{definition}(Bottleneck)
   A hidden variable $h$ is said to be a bottleneck if there exists at
-    least three conditionally independent observed variables $o_1, o_2,
-    o_3$. 
-  Let $\vec x_\sB(h)$ be the set of conditionally independent observed
-    variables for $h$.
+    least three conditionally independent observed variables, or views,
+    $x_1, x_2, x_3$. 
+  Let $\xB{h}$ be the set of views for $h$.
 \end{definition}
 
-We will further assume that each observed variable has a single parent. 
-This leads to the following two properties.
-\begin{property}
-  \label{prop:unique-parent}
-  Every observed variable $o$ has exactly one parent, $|\Pa(o)| = 1$.
-\end{property}
+The only property a graphical models must have is the following,
 \begin{property}
@@
 
-We note that \propertyref{unique-parent} can be relaxed if another
-  algorithm is used to recover the conditional moments. For example,
-  \citet{halpern13noisyor} propose an algorithm that can recover the
-  conditional moments for a broad, but less general class of noisy-or
-  networks.
-\propertyref{bottleneck} can also be relaxed if factors share the same
-  parameters. 
+We note that \propertyref{bottleneck} can be relaxed if some cliques
+  share parameters;
 For example, because we can identify the observation potential from just
@@
   x^b_3$ and $x^b_4$.
+In this case, we only require that the hidden variables in distinct
+  cliques share parameters.
 
-Given \propertyref{unique-parent} and \propertyref{bottleneck}, step
-  1 of \algorithmref{directed} is guaranteed to recover potentials
-  $O^{(h)}_{x}$ for every hidden variable $h$ and observed variable $x \in
-  \vec x_\sB(h)$.
-\propertyref{bottleneck} guarantees that for every clique $\sC
-  = \{h_1, \cdots, h_m\}$ in the graphical model $\sG$, there exists at
-  least one bottleneck observable that depends on exactly one $h_i \in
-  \sC$; let $x_{h_i}$ be that observed variable\verify. 
-Finally, define $\vec x_\sC = \{x_{h_1}, \cdots x_{h_m}\}$ to be the
-  participating observed variables in the clique $\sC$.
+It is not immediately obvious that \propertyref{bottleneck} implies
+  \propertyref{exclusive-views}. 
+Another consideration is that while
+  \TensorFactorize gives us the conditional moments for every
+  bottleneck, it can not recover the conditional probability table when
+  there are multiple parents for an observed variable.
+The following lemma shows us how to handle this case.
 
-Finally, we have the following moment condition,
+\begin{lemma}(Canonical form) 
+  Every directed graphical model can be transformed into one in which
+  the observed variables are leaves with exactly one parent.
+\end{lemma}
+\begin{proof}
+  \begin{figure}
+    \label{fig:reduction}
+    \centering
+    \subimport{figures/}{reduction.tikz}
+    \caption{Reduction to canonical form.}
+  \end{figure}
+
+  Let $x$ be an observed variable with parents $h_1, \cdots, h_p$ and
+  children $h_{p+1}, \cdots, h_c$. Replace $x$ with a phantom hidden
+  variable $h_x$, and make $x$ the child of $h_x$ (\figureref{reduction}). The domain of $h_x$
+  has the same as $x$, and $\mOpp{x}{h_x} = I$.
+
+  The parameters for $h_x$ are the same as $x$.
+\end{proof}
+
+Applying \TensorFactorize to every bottleneck gives us candidate views
+  for every hidden variable. Now we show that every clique also has
+  exclusive views.
+
+\begin{lemma}(Bottlenecks guarantee exclusive views)
+  Provided \propertyref{bottleneck} holds, every hidden variable
+    $h_{i_1}, \cdots, h_{i_m}$ in a clique $\sC$ has a exclusive view.
+\end{lemma}
+\begin{proof}
+  Let $h$ be a hidden variable in $\sC$, let $\sV_{h}$ be the set of
+    views for it and let $\sC^- \eqdef \sC \setminus \{h\}$.
+  By definition, the views of $h$ are conditionally independent, viz.
+    every d-connecting path between two views $x_1, x_2 \in \sV_h$ must pass through
+    $h$.
+  A view $x \in \sV_{h}$ is exclusive for $h$ iff $x \not\in \sV_{h'} ~ \forall h'
+  \in \sC^-$
+
+  Suppose $h$ had no unique view, then $\sV_{h} \subseteq \Union_{h' \in \sC^-} \sV_{h'}$. 
+  However, this would imply that for every $x_1, x_2 \in \sV_{h}$, there
+    exists some $h_1, h_2 \in \sC^-$ (not necessarily distinct) such that
+    $x_1$ is d-connected to $h_1$ and $x_2$ is d-connected $h_2$.
+  By participation in the clique, either $h_1 = h_2$ or $h_1$ is
+    d-connected to $h_2$, implying that $x_1$ and $x_2$ are d-connected
+    via a path that does not pass through $h$, contradicting the
+    statement that $x_1, x_2 \in \sV$.
+
+  Thus it must be the case that $\sV_{h} \subsetneq \Union_{h' \neq
+    h \in \sC} \sV_{h'}$ and hence there exists a unique view for $h$.
+
+  Algorithmically, finding the unique views can be done by subtracting
+    out the remaining set, i.e. $x_{v_h} \in \sV_h \setminus \Union_{h'
+    \in \sC} \sV_{h'}$
+\end{proof}
+
+\paragraph{Aligning the views}
+\todo{How do we make sure all the recovered $O$ are align? Shouldn't be Hard}
+
+With this lemma in place, we present the full algorithm, \LearnMarginals
+in \algorithmref{directed}.
+
+\renewcommand{\algorithmicrequire}{\textbf{Input:}}
+\renewcommand{\algorithmicensure}{\textbf{Output:}}
+\begin{algorithm}
+  \caption{\LearnMarginals}
+  \label{algo:directed}
+  \begin{algorithmic}
+    \REQUIRE A graphical model $\sG$ satisfying \textbf{(P1)}, \textbf{(P2)}, data $\sD$
+    \ENSURE Marginals $Z_\sC$ for every clique $\sC \in \sG$
+
+      \FOR{$h_i \in H$} 
+        \STATE Apply \TensorFactorize to learn conditional moments
+        $\mOpp{v}{i}$ for every $x_v \in \xB h$.
+
+%    \COMMENT{Recover observation potentials $O$ using bottlenecks}
+      \ENDFOR
+%      \COMMENT{\textbf{Step 2:} Recover clique potentials from the piecewise likelihood.}
+\FOR{every clique $\sC = \{h_{i_1}, \cdots, h_{i_m}\} \in \sG$} 
+\STATE Identify unique views $\Sx{\sV} = \{x_{v_1}, \cdots, x_{v_m}\}$.
+\STATE $\mH_\sC \gets \mO_{\Sx{\sV}}( \pinv{\mOpp{v_1}{i_i}}, \cdots, \pinv{\mOpp{v_m}{i_m}} )$.
+      \ENDFOR
+  \end{algorithmic}
+\end{algorithm}
+
+\paragraph{Theoretical properties}
+
+As each step of \LearnMarginals is a consistent estimator, it is
+  straightforward that as whole it is a consistent estimator for $\sG$.
+We describe its asymptotic complexity to see what the dependence on the
+  parameters of the problem are.
+
+\begin{lemma}(Sample complexity for $\sC$)
+
+We require some number of samples to learn the factor for $\sC$.
+\end{lemma}
+\begin{proof}
+
+For the sake of simplicity, let us assume that our dataset of $n$
+  samples is partitioned into two sets.
+  The first $\bytwo{n}$ samples are used to estimate $\mOpphat{v}{i}$ for
+  every $h_i \in H$ and $x_v \in \sV_{h}$ in step 1 of \algorithmref{directed}.
+\citet{anandkumar13tensor} show that 
 \begin{align*}
-  \mO_{x_\sC} &= \mH_{\sC}(O_{x_{h_1}}, \cdots, O_{x_{h_m}}) \\
-  \mH_{\sC} &= \mO_{x_\sC}(\pinv O_{x_{h_1}}, \cdots, \pinv O_{x_{h_m}}).
+  \|\mOpphat{v}{i} - \mOpp{v}{i}\|_F &= O( \frac{k \pi_{\max}^2}{\sigma_k(M_2)^5} ). 
 \end{align*}
 
-This algorithm is simple and intuitive. Next, we characterize its sample complexity.
+The next $\bytwo{n}$ samples are used to compute the moments $M_\sV$ for
+  every clique in step 2.
 
-\begin{lemma}(Sample complexity for $\sC$)
-We require some number of samples to learn the factor for $\sC$.
-\end{lemma}
+We have for each clique,
+\begin{align*}
+  Z_{\sC} &= \mO_\sV(\pinv{\mOpp{v_1}{i_1}},\cdots,\pinv{\mOpp{v_m}{i_m}}) \\
+\end{align*}
 
-This gives us our final result for a graphical model in our model
-family.
-\begin{theorem}(Learnability of Bottleneck Graphs)
-The parameters for a directed graphical $\sG$ with tree-width $k$
-satisfying \propertyref{unique-parent}, \propertyref{bottleneck} and
-\assumptionref{full-rank} can be learned with $O(\cdot)$ samples.
-\end{theorem}
+Choosing to represent $Z_\sC$ as a vector and $\mOpp{\cdot}{\cdot}$ as a matrix,
+let $\pinv{\mOpp{\sV}{\sC}} \eqdef \pinv{\mOpp{v_1}{i_1}} \otimes \cdots \otimes \pinv{\mOpp{v_m}{i_m}}$,
+\begin{align*}
+  Z_{\sC} &= \pinv{\mOpp{\sV}{\sC}} \mO_\sV.
+\end{align*}
 
-\paragraph{Aggregating observations}
+We know the asymptotic distribution of $M_\sV$ goes as,
+\begin{align*}
+  \sqrt{n}(\hat M_\sV - M_\sV) \convind \sN(0, \Sigma_\sV),
+\end{align*}
+where $\Sigma_\sV$ is the variance. Because $M_\sV$ describes the discrete marginal distribution, we have 
+\begin{align*}
+  \Sigma_\sV &= \diag  M_\sV  - M_\sV M_\sV^T,
+\end{align*}
 
-When we have multiple observed variables for a hidden variable, it is
-  always beneficial to aggregate over them.
-Note that $O_h$ will have rank atleast that of $O$; this encodes the
-  intuitive fact that we might be able to identify $h$ from multiple
-  observed variables even if we can not identify it from a single
-  observed variable. 
-Suppose $x_\sB(h) = \{x_1, x_2, \cdots, x_l\}$, let $O_h \in
-  \Re^{D^l \times K}$ be an
-  aggregated observation potential, such that $O_h(x_1, x_2, \cdots,
-  x_l) = \mH(x_1|h) \mH(x_2|h) \cdots \mH(x_l|h)$.
-Given \assumptionref{full-rank}, $O_h$ have rank $K$.
-After step 1, $O_h$ is a known quantity.
+Thus, $Z_\sC$ goes as
+\begin{align*}
+  \sqrt{n}(\hat Z_\sV - Z_\sV) \convind \sN(0, \pinvt{\mOpp{\sV}{\sC}} \Sigma_\sV \pinv{\mOpp{\sV}{\sC}}).
+\end{align*}
+
+\todo{Convert from $Z_\sC$ to the actual CPTs}.
+
+\end{proof}

### 2014-01-30 doc/icml2014/related.tex
Updated undirected section

\section{Related work}

Anandkumar - linear bayesian nets

Le song - trees

David sontag - noisy or

### 2014-01-30 doc/icml2014/undirected.tex
Updated undirected section

@@
 
-In this section, we will show that we can extend the results to undirected graphical models.
+In this section, we will show that we can extend the results to
+  undirected graphical models parameterized as log-linear latent variable
+  models.
+We exploit a connection between the moments of the clique potentials and
+  the log-partition function to show that learning parameters is convex
+  given the clique marginals $Z_\sC$. 
+The end result is a consistent estimator for log-linear
+  latent-variable models which satisfy \propertyref{exclusive-views}.
+Finally, we generalize the result to classes when only a subset of the
+  clique marginals can be learned.
 
-Let $z$ be a collection of variables indexed by $V$ (that is, $z = \{z_j\}_{j \in V}$).
-Suppose that some of the variables are observed ($x = \{ z_j \}_{j \in O}$, where $O \subset V$) 
-and the rest are latent ($h = \{ z_j \}_{j \in H}$, where $H = V \backslash O$).
-Let $C \subset 2^V$ be a collection of subsets of variables,
-and let $\phi_c(z_c) \in \Re^d$ be the feature map for each $c \in C$.
-In this paper, we will consider both directed and undirected models over $z$.
-In the undirected case, we define the standard globally-normalized exponential family model:
+As before, let $\sG$ be a graphical model with variables $V$, with $X
+  \subseteq V$ being observed and $H = V \setminus X$ hidden.
+We assume that the undirected graphical model is a log-linear latent
+  variable model with parameters $\theta$,
+\begin{align*}
+  p_\theta(\vx, \vh) &= \exp( \theta^T\phi(x,z) - A(\theta) ),
+\end{align*}
+where $\phi(x,z)$ is a feature vector with components $\phi_\sC(x,z)$
+  for every clique and $A(\theta) = \sum_{\vx, \vh}  \exp(
+  \theta^T\phi(x,z) )$ is the log-partition function.
+While $\phi_\sC(x,z)$ can be an arbitrary featurization, we assume
+  without loss of generality that $\phi_\sC(x,z)$ are the indicator
+  features $\BI[x,z]$\footnote{Note that $\BI[x,z]$ are sufficient
+  statistics for the model, and hence we do not lose any statistical
+  efficiency in using them}.
+
+As before, our objective is to find a set of parameters $\hat \theta$
+  that maximize the likelihood of data $\sD$.
 \begin{align}
-  \label{eqn:undirectedSetup}
-  p_\theta(z) = \exp\{ \phi(z)^\top\theta - A(\theta) \},
+  \hat \theta 
+      &= \arg\max_{\theta} \sum_{\vx \in \sD} \log p_\theta(\vx) \nonumber \\
+      &= \arg\max_{\theta} \sum_{\vx \in \sD} \log \sum_{\vh} p_\theta(\vx,\vh) \nonumber \\
+      &= \arg\max_{\theta} \sum_{\vx \in \sD} \log \sum_{\vh} \exp( \theta^T\phi(\vx,\vh) - A(\theta) ). \label{eqn:obj-ml}
 \end{align}
-where $\phi(z) \eqdef \sum_{c \in C} \phi_c(z_c) \in \Re^d$ is the global feature vector, $\theta \in \Re^d$ is the parameter vector,
-and $A(\theta) = \log \int \exp\{\phi(z)^\top\theta\} dz$ is the global log-partition function.
-To some extent, our methods only depend on the conditional independence
-structure of the models.
+Note that this problem is not concave in $\theta$.
 
-\paragraph{Moment Constraints}
+An alternate objective is to optimize the following lower-bound on the log-likelihood, 
+\begin{align}
+  \sL_p(q,\theta) &= \sum_{\vx \in \sD} \sum_{\vh} q(\vh \given \vx) \log\frac{p_\theta(\vx,\vh)}{q(\vh \given \vx)} \\
+    &= \KL( q(\vh \given \vx) || p_{\theta}(\vh \given \vx)) \label{eqn:obj-kl}
+\end{align}
 
-In the previous section, we saw that we could learn moments $P(x1,x2)$ given conditions. This makes the problem convex, boom.
+Importantly, given $q$, the above objective is convex in $\theta$ and vice versa.
+It can be shown\cite{} that the local-optima of \equationref{obj-kl}
+are also local-optima of \equationref{obj-ml}. Usually, one
+iteratively optimizes $q$ and $p_\theta$ from some initial parameters
+$\theta_0$ to arrive at a local optimum.
 
-\paragraph{Examples}
+This reduces to,
+\begin{align}
+  \sL_p(\theta) &= \theta^T Z - A(\theta),
+\end{align}
+which is a convex problem.
 
-\paragraph{General latent-variable models with bottlenecks}
+However, from the previous section, for any model satisfying
+  \property{exclusive-views}, we have a consistent estimator for the
+  marginals $q \equiv Z$. 
+Therefore, we can optimize for $\theta$ given this estimate of $q$,
+  $\hat q$.
 
-Our learning algorithm for three-view mixture models relies on the fact that
-the sufficient statistics of the exponential family
-can be expressed as a function of the latent moments.
-Of course, these conditions do not hold for general log-linear models.
-Rather than falling back on plain local optimization of the likelihood in these
-cases, we would still like
-to somehow use the \emph{partial information} offered by the latent moments.
+\renewcommand{\algorithmicrequire}{\textbf{Input:}}
+\renewcommand{\algorithmicensure}{\textbf{Output:}}
+\begin{algorithm}
+  \caption{\LearnLogLinear}
+  \label{algo:undirected}
+  \begin{algorithmic}
+    \REQUIRE A graphical model $\sG$ satisfying \textbf{(P1)}, \textbf{(P2)}, data $\sD$
+    \ENSURE A consistent estimate $\hat \theta$ of the parameters of the model.
+      \STATE Apply \LearnMarginals to $\sG$ to learn clique marginals $Z_\sC$.
+      \STATE Optimize \equationref{obj-kl} with respect to $\theta$.
+  \end{algorithmic}
+\end{algorithm}
 
-Partial information is available when the following holds:
-%\begin{property}
-%  There exists some latent variable $h_{j_0} \in [\nh]$ (the bottleneck)
-%  and three conditionally independent observed variables $x_{j_1},x_{j_2},x_{j_3}$ (views) such that
-%  each conditional mean matrix $\E[\phix(x_{j}) \mid h_{j_0}] \in \R^{\nphix \times \nh}$ ($j \in \{ j_1, j_2, j_3 \}$) has full column rank $\nh$
-%  for some transformation $\phix$.
-%\end{property}
+This approach is outlined in \algorithmref{undirected}.
 
-Optimization.
+\todo{Do we have to worry about the fact that this is constrained?}
+\todo{What about model un-identifiability?}
+
+\paragraph{Learning with partial moment constraints}
+
+In more general models, \propertyref{exclusive-views} may not hold for
+  every clique in the graph, however we can still estimate the marginals
+  $Z_\sC$ for cliques in which it does.
+In this section, we will describe a natural method to use these
+  constraints to simplify the parameter search problem.
+
+Constrain the KL to be in $Q$.
+
+\subsection{Examples}
+
+\paragraph{Latent MRFs}
+
+Image segmentation
+
+Separate features into views
+
+\todo{Some challenge example requiring measurements em}

### 2014-01-31 doc/icml2014/abstract.tex
Added words and random mumbo-jumbo
Updated directed section
Updated draft

--- 2014-01-27 doc/icml2014/abstract.tex
@@
 \begin{abstract}
+
+Parameter estimation in latent variable models is prone to local
+  optima due to the non-convexity of the likelihood. 
+In this paper, we extend recent advances in the method of moments 
+  to provide consistent parameter estimates for a general class of
+  directed and undirected graphical models, thus avoiding the issue of
+  local optima.
+In particular, our algorithm applies to any graphical model wherein each
+  hidden variable has at least three conditionally independent observed
+  variables.
+This family includes loopy graphs with high treewidth.
+We use the method of moments to tie each hidden to some observed
+  variables, and show that the rest of the problem can be cast as convex
+  optimization of the likelihood.
+This approach gracefully extends to models outside this family by
+  providing partial information via posterior regulraization.
 \end{abstract}
 
-\paragraph{Errata}
+\paragraph{Last updated:} \today, \currenttime
+
+\paragraph{Errata / Incomplete}
 \begin{enumerate}
-  \item The likelihoods as defined are concave, not convex.
-  \item Define parameters $\theta$ more uniformly.
+  \item Use \texttt{\textbackslash left(}.
 \end{enumerate}

### 2014-01-31 doc/icml2014/background.tex
Updated draft

@@
 
-Let $\sG$ be a discrete graphical model with variables $V$, 
-  let the subset $X \subseteq V$ be observed and 
-  let $H = V \setminus O$ be hidden.
-Let $\Pa(v) \subset V$ be the parents of a variable $v$ in a directed
-  model, and $\sN(v) \subset V$ be the neighbours of $v$ in an undirected
-  model.
-We will use $x_1, x_2, \cdots, x_N \in X$ to denote observed variables and
+Let $\sG$ be a discrete graphical model with observed variables $X$ and
+  hidden variables $H$.
+Let $\Pa(a) \subset X \union H$ be the parents of a variable $a$ in a directed
+  model, and $\sN(a) \subset X \union H$ be the neighbours of $a$ in an
+  undirected model.
+We will use $x_1, x_2, \cdots, x_L \in X$ to denote observed variables and
   $h_1, h_2, \cdots, h_M \in H$ to denote hidden variables.
 For ease of exposition, we assume that $x_i \in \{1, \cdots, d\}$ for
-  all $i$, and that $h_i \in \{1, \cdots, k\}$ for all i.
+  all $i$, and that $h_i \in \{1, \cdots, k\}$ for all $i$.
+Let $\sX$ be the joint domain of the observed variables, i.e. $\sX \eqdef
+  \underbrace{d \times d \cdots \times d}_{L ~\textrm{times}}$ and
+let $\sH$ be the joint domain of the hidden variables, i.e. $\sH \eqdef
+  \underbrace{k \times k \cdots \times k}_{M ~\textrm{times}}$.
 
@@
     \sL &\eqdef \max_{\theta} \sum_{\vx \in \sD} \log p_\theta(\vx) \\
-        &=      \max_{\theta} \sum_{\vx \in \sD} \log \sum_{\vh} p_\theta(\vx,\vh)
+        &=      \max_{\theta} \sum_{\vx \in \sD} \log \sum_{\vh \in \sH} p_\theta(\vx,\vh)
   \end{align*}
   which is statistically efficient but computationally intractable
-  because we must marginalize over latent variables $z$.
+  because we must marginalize over latent variables $\vh$.
 In practice, one uses gradient-based optimization procedures (e.g., EM
@@
 \paragraph{Notation}
+
+We use the notation $[\cdot]$ to indicate indexing; for example $M[i]$
+  is the $i$-th row of a matrix $M$ and $M[i,j]$ is the $i,j$-th element
+  of $M$.
+For a tensor $T \in \Re^{d \times \ldots \times d}$ and a vector $\vec
+  i = (i_1, \cdots, i_l)$, $T[\vec i] = T_{i_1 \ldots i_l}$.
+Similarly, for a set $\sC$ of size $k$ and vector $\vec x \in \Re^d$, we
+  will use the notation $\vec x_\sC$ to get a vector $\Re^k$ which selects
+  the appropriate elements of $\vec x$
 
@@
       &= \sum_{i_1 = 1}^{d} \cdots \sum_{i_\ell = 1}^{d} 
-            T_{i_1 i_2 \ldots i_l} ({x_1})_{i_1} \cdots ({x_\ell})_{i_\ell}.
+            T[i_1, i_2, \ldots, i_l] x_1[i_1] \cdots x_\ell[i_\ell].
 \end{align*}
@@
 \begin{align*}
-  T(M_1, \cdots, M_\ell)_{j_1, \ldots, j_l} 
+  T(M_1, \cdots, M_\ell)[\vec j]
       &= \sum_{i_1 = 1}^{d} \cdots \sum_{i_\ell = 1}^{d} 
-            T_{i_1 i_2 \ldots i_l} ({M_1})_{i_1j_1} \cdots ({M_\ell})_{i_\ell j_\ell}.
+            T[\vec i] {M_1}[i_1,j_1] \cdots M_\ell[i_\ell, j_\ell],
 \end{align*}
+where $\vec i = (i_1, \cdots, i_\ell)$ and $\vec j = (i_1, \cdots, j_\ell)$
 
@@
   that are independent conditioned on a latent variable $h$.  
-The parameters of this model are the mixture probabilities $\pi \eqdef
-  \Pr(h) \in \Delta_{k-1}$ and the conditional moments $O^{(i)} \eqdef
+The parameters of this model are the mixture probabilities $\pi\oft 1 \eqdef
+  \Pr(h) \in \Delta_{k-1}$ and the conditional moments $\mOpp{i}{1} \eqdef
   \Pr(x_i \given h) \in \Re^{d \times k}$ for $i \in \{1,2,3\}$.
@@
 \begin{align*}
-  \mO_{ij} &= \sum_h \pi_h {O\oft i}^T_h \otimes {O\oft j}^T_h & \forall i,j \in \{1,2,3\} \\
-  \mO_{123} &= \sum_h \pi_h {O\oft 1}^T_h \otimes {O\oft 2}^T_h \otimes  {O\oft 3}^T_h.
+  \mO_{ij} &= \sum_h {\pi \oft 1}[h] {\mOppt{i}{1}}[h] \otimes {\mOppt{j}{1}}[h] \quad \forall i,j \in \{1,2,3\} \\
+  \mO_{123} &= \sum_h {\pi \oft 1}[h] {\mOppt{i}{1}}[h] \otimes {\mOppt{2}{1}}[h] \otimes  {\mOppt{2}{1}}[h].
 \end{align*}
@@
   we call $\TensorFactorize$, which converts consistent estimates of the observed moments 
-  into consistent estimates of the parameters, $\pi, O\oft{1},
-  O\oft{2}, O\oft{3}$ (up to permutation of the columns), provided
+  into consistent estimates of the parameters, $\pi\oft 1, \mOpp{1}{1},
+  \mOpp{2}{1},\mOpp{3}{1}$ (up to permutation of the columns), provided
   the following assumption holds:
@@
 \label{asm:full-rank}
-Assume the conditional moments $O\oft{1}, O\oft{2}, O\oft{3}$ have full column rank
-  $k$, and that $\pi \succ 0$.
+Assume the conditional moments $\mOpp{1}{1}, \mOpp{2}{1}, \mOpp{3}{1}$ have full column rank
+  $k$, and that $\pi\oft 1 \succ 0$.
 \end{assumption}
@@
 Then a robust tensor power method is applied to extract the eigenvectors
-  of the whitened $\mO_{123}$; unwhitening yields the columns of $\mO_{3}$.
-$\mO_1$ and $\mO_2$ can then be recovered with some algebra.
+of the whitened $\mO_{123}$; unwhitening yields the columns of $\mOpp{3}{1}$.
+  $\mOpp{1}{1}$ and $\mOpp{2}{1}$ can then be recovered with some algebra.

### 2014-01-31 doc/icml2014/directed.tex
Updated undirected section
Updated draft

@@
   in latent variable models.
-The technique is central in solving several other problems with more
-  complex correlation structure, for example latent Dirichlet allocation
-  \citep{anandkumar12lda}, mixture of Gaussians \citep{hsu13spherical},
-  etc.\findcite{more things}.
-
-Once we recover the conditional moments $O\oft{i,j} \eqdef \Pr(x_i
-  | h_j)$ for some $i,j$, we present a systematic approach to learn the
+  Once we recover the conditional moments $\mOpp{v}{j} \eqdef \Pr(x_v
+  | h_j)$ for some $v, j$, we present a systematic approach to learn the
   conditional probability tables for every clique for a class of directed
@@
 We begin by illustrating our approach with an example, followed by
-  describing the simple conditions for which the approach works and some
-  theoretical properties.
+  describing the algorithm in full generality.
+For ease of exposition, we will describe our algorithm entirely in the
+  context of directed graphical models, though it generalizes to factor
+  graphs.
+We also present the algorithm solely in terms of the method of moments; in
+  \sectionref{piecewise}, we show how we can get a statistically more
+  efficient estimator using convex optimization.
 
@@
   \input{figures/grid.tikz}
-  \caption{A directed grid model.}
+  \caption{A directed grid model.\todo{Annotate features}}
   \label{fig:grid}
@@
   captures all the different dependency structures possible in
-  a Bayesian network and is hence a good example for intuition\reword.
+  a Bayesian network.
 The model has eight observed variables $x^a_1, x^b_1 \cdots, x^a_4, x^b_4$ and four
@@
 
-\paragraph{Estimating $~O$}
-
+\paragraph{Estimating $O$}
 Note that the observed variables $x^a_1, x^b_1, x^a_2$ are
@@
   the $O$.
-We can exploit parameter sharing between the different $x$ to aggregate
-  statistics from additional sets of independent
-  observed variables, or {\em views}, e.g. $\{x^a_2, x^b_2, x^a_3\}$, to
-  arrive at a better estimate of $O$.
 
-%Intuitively, we expect the observations to tell us about the underlying
-%  hidden state, which should make learning each factor easier. 
-%For example, when $O = I$ the observed variables
-%  tell us exactly which hidden state we are in; the optimal procedure to
-%  estimate factors is simply counting the co-occurrences of the hidden
-%  states.
-
-\paragraph{Estimating $~\pi$}
-
+\paragraph{Estimating $\pi$}
 The moments of $x^a_1$,$\mO_1 \eqdef \Pr(x^a_1)$ are directly related to
@@
 
-\paragraph{Estimating $~T$}
-
+\paragraph{Estimating $T$}
 Similarly, we can write down the moments of $x^a_1, x^a_2$, $\mO_{12}
@@
 \end{align*}
-
 $T$ can be recovered from the $\mH_{12}$ by suitable normalization.
@@
 \paragraph{Estimating $V$}
-
 Finally, we can estimate $V$ from the third-order moments $\mO_{234} \eqdef \Pr(x^a_2, x^a_3, x^a_4)$ as
@@
 Fundamentally, our algorithm takes advantage of having a view for every
-  hidden variable in a clique to help identify its hidden state. The
-  ideal property we wish our graphical models to possess is the
-  following:
+  hidden variable in a clique to help identify its hidden state. We define this condition as follows,
 \begin{property}(Exclusive views)
   \label{prop:exclusive-views}
-For every clique $\sC$ of $\sG$ and every hidden variable $h \in \sC$,
-  there is some observed variable, $x_{v_h}$, which is conditionally
-  independent of the rest of the clique given $h$ and whose conditional
-  moment $\mOpp{v_h}{h} \eqdef \Pr(x_{v_h} | h)$ is known. 
-We call $x_{v_h}$ the exclusive view of $h$ in $\sC$.
+A clique $\sC$ of $\sG$ is said to posses exclusive views if for every hidden variable $h_i \in \sC$,
+  there is some observed variable, $x_{v}$, which is conditionally
+  independent of the rest of the clique given $h_i$ and whose conditional
+  moment $\mOpp{v}{i} \eqdef \Pr(x_{v} | h_i)$ is known. 
+We call $x_v$ the exclusive view of $h_i$ in $\sC$.
 \end{property}
 
-Before we identify a family of models which entail this property, let us
-  show that it is sufficient to learn the parameters of $\sG$.
+\paragraph{Learning cliques}
 
-Consider any clique $\sC = \{h_{i_1}, \cdots, h_{i_m}\} \in \sG$. Let $x_{v_j}$ be the exclusive
-  view for $h_{i_j}$ in $\sC$ and $\sV = \{x_{v_1}, \cdots, x_{v_m}\}$
-  be the set of views for the clique $\sC$.
+We now show this condition is sufficient to learn the marginal
+  distribution of the clique.
+Consider any clique $\sC = \{h_{i_1}, \cdots, h_{i_m}\} \in \sG$ for which this property holds. Let
+  $x_{v_j}$ be the exclusive view for $h_{i_j}$ in $\sC$ and $\sV
+  = \{x_{v_1}, \cdots, x_{v_m}\}$ be the set of views for the clique
+  $\sC$.
 We can write down the moments $\mO_\sV$,
@@
   &\eqdef \Pr(\Sx{\sV}) \\
-  &= \sum_{h_{i_1} = 1}^{d} \cdots \sum_{h_{i_1} = 1}^{d} 
+  &= \sum_{\vh \in \sH}
       \Pr(\Sh{\sC}) \Pr(\Sx{\sV} \given \Sh{\sC}) \\
-      &= \sum_{h_{i_1} = 1}^{d} \cdots \sum_{h_{i_1} = 1}^{d} \Pr(\Sh{\sC}) 
+      &= \sum_{\vh \in sH} \Pr(\Sh{\sC}) 
           \Pr(x_{i_1} | h_{v_1}) \cdots \Pr(x_{i_m} | h_{v_m}) \\
@@
   $Z_\sC$ via conditioning.
+\algorithmref{learnclique} summarizes the procedure, \LearnClique.
 
-For each $\mOpp{v_j}{i_j}$ to have full column rank, it is equivalent
-  to the following generalized assumption of
-  \assumptionref{full-rank},
+\renewcommand{\algorithmicrequire}{\textbf{Input:}}
+\renewcommand{\algorithmicensure}{\textbf{Output:}}
+\begin{algorithm}
+  \caption{\LearnClique}
+  \label{algo:learnclique}
+  \begin{algorithmic}
+    \REQUIRE A $\sC$ with \propertyref{exclusive-views}.
+    \ENSURE The marginal distribution of the clique, $Z_\sC$.
+      \STATE Identify unique views $\Sx{\sV} = \{x_{v_1}, \cdots, x_{v_m}\}$.
+      \STATE $\mH_\sC \gets \mO_{\Sx{\sV}}( \pinv{\mOpp{v_1}{i_i}}, \cdots, \pinv{\mOpp{v_m}{i_m}} )$.
+  \end{algorithmic}
+\end{algorithm}
 
-\begin{assumption} 
-  \label{asm:full-rank-plus}
-  For every clique $\sC = \{h_{i_1}, \cdots, h_{i_m}\} \in \sG$, the
-  marginal distribution $Z_\sC$ derived by the parameters is full rank
-  for every mode-unfolding of $Z_\sC$.
-\end{assumption}
+\paragraph{Structural properties}
 
-Intuitively, this assumption says that the marginal probability of every state for any hidden variable is non-zero.
-
-\begin{lemma}
-  \label{lem:goodness}
-
-  Given \assumptionref{full-rank-plus}, (a) the marginal distribution
-  $\Pr(h_i) \succ 0$ for every $h_i$ in the graph and (b)
-  $\mOpp{v_j}{i_j}$ has full column rank.
-\end{lemma}
-\begin{proof}
-  By variable elimination.
-\end{proof}
-
-\paragraph{Required structural properties}
-
-We will now provide a sufficient condition for
-\propertyref{exclusive-views} to hold. 
-Recall that given three observed
-variables, $x_1, x_2, x_3$ that are independent conditioned on $h$,
-\TensorFactorize allows us to recover the conditional moments $\mOp{1},
-\mOp{2}, \mOp{3}$. We define a hidden variable that has this capability to be a bottleneck.
+If we can guarantee that every clique in a graph possesses exclusive
+  views, then we can apply \LearnClique iteratively to learn the entire
+  model.
+In the example, we were able to guarantee \propertyref{exclusive-views}
+  by identifying a hidden variable, $h^1$, with three conditionally
+  independent observed variables, $x^a_1, x^b_1, x^a_2$, followed by
+  using \TensorFactorize to learn the conditional moments.
+We define such a hidden variable to be a bottleneck;
 \begin{definition}(Bottleneck)
@@
     $x_1, x_2, x_3$. 
-  Let $\xB{h}$ be the set of views for $h$.
+  Let $\sV_{h}$ be the set of views for $h$.
 \end{definition}
-
-The only property a graphical models must have is the following,
+Finally, the claim we make is that we can recover parameters for any
+  graphical model where every hidden variable is a bottleneck;
 \begin{property}
@@
 \end{property}
-
 We note that \propertyref{bottleneck} can be relaxed if some cliques
@@
 
-It is not immediately obvious that \propertyref{bottleneck} implies
-  \propertyref{exclusive-views}. 
-Another consideration is that while
-  \TensorFactorize gives us the conditional moments for every
-  bottleneck, it can not recover the conditional probability table when
-  there are multiple parents for an observed variable.
-The following lemma shows us how to handle this case.
+Applying \TensorFactorize to every bottleneck gives us candidate views
+  for every hidden variable. 
+To proceed, we need to show two things; (i) under what assumptions
+  does \assumptionref{full-rank} hold for each bottleneck and (ii)
+  that the candidate views actually imply that \propertyref{exclusive-views}
+  holds for every clique in the graph.
+
+To address the first, we will need that the following assumption on the
+  parameters of the model:
+\begin{assumption} 
+  \label{asm:full-rank-plus}
+  For every clique $\sC = \{h_{i_1}, \cdots, h_{i_m}\} \in \sG$, let
+    $Z_\sC$ be the marginal distribution for that clique. 
+  Then every mode-$i$ unfolding of $Z_\sC$ has full column rank.
+\end{assumption}
+Intuitively, the assumption guarantees that the marginal probability of
+  any hidden state is non-zero.
+
+Before we establish the second condition, we describe a reduction to
+  a canonical form that will allow us to handle the case where an observed
+  variable has multiple parents.
 
 \begin{lemma}(Canonical form) 
+  \label{lem:reduction}
   Every directed graphical model can be transformed into one in which
@@
 
-Applying \TensorFactorize to every bottleneck gives us candidate views
-  for every hidden variable. Now we show that every clique also has
-  exclusive views.
-
 \begin{lemma}(Bottlenecks guarantee exclusive views)
@@
 \begin{proof}
+  By \lemmaref{reduction}, we can assume w.l.o.g. that every observed
+  variable is a leaf in the graph, causing any v-structures in the graph
+  to be active. This allows us to reason about independence simply by
+  checking whether or not two nodes are connected or not; note that this
+  proof applies for undirected graphs as well.
+
   Let $h$ be a hidden variable in $\sC$, let $\sV_{h}$ be the set of
@@
   By definition, the views of $h$ are conditionally independent, viz.
-    every d-connecting path between two views $x_1, x_2 \in \sV_h$ must pass through
+    every connecting path between two views $x_1, x_2 \in \sV_h$ must pass through
     $h$.
@@
     exists some $h_1, h_2 \in \sC^-$ (not necessarily distinct) such that
-    $x_1$ is d-connected to $h_1$ and $x_2$ is d-connected $h_2$.
+    $x_1$ is connected to $h_1$ and $x_2$ is connected $h_2$.
   By participation in the clique, either $h_1 = h_2$ or $h_1$ is
-    d-connected to $h_2$, implying that $x_1$ and $x_2$ are d-connected
+    connected to $h_2$, implying that $x_1$ and $x_2$ are connected
     via a path that does not pass through $h$, contradicting the
@@
 
-\paragraph{Aligning the views}
-\todo{How do we make sure all the recovered $O$ are align? Shouldn't be Hard}
-
-With this lemma in place, we present the full algorithm, \LearnMarginals
+With this lemma in place, we present the full algorithm, \LearnMarginals~
 in \algorithmref{directed}.
@@
   \begin{algorithmic}
-    \REQUIRE A graphical model $\sG$ satisfying \textbf{(P1)}, \textbf{(P2)}, data $\sD$
+    \REQUIRE A graphical model $\sG$ satisfying \propertyref{bottleneck}, data $\sD$
     \ENSURE Marginals $Z_\sC$ for every clique $\sC \in \sG$
@@
         \STATE Apply \TensorFactorize to learn conditional moments
-        $\mOpp{v}{i}$ for every $x_v \in \xB h$.
+        $\mOpp{v}{i}$ for every $\sV_h$.
 
@@
 \FOR{every clique $\sC = \{h_{i_1}, \cdots, h_{i_m}\} \in \sG$} 
-\STATE Identify unique views $\Sx{\sV} = \{x_{v_1}, \cdots, x_{v_m}\}$.
-\STATE $\mH_\sC \gets \mO_{\Sx{\sV}}( \pinv{\mOpp{v_1}{i_i}}, \cdots, \pinv{\mOpp{v_m}{i_m}} )$.
+  \STATE Apply \LearnClique to learn the marginals $\mH_\sC$.
 \ENDFOR
@@
 
-\paragraph{Theoretical properties}
+\paragraph{Consistency and Sample Complexity}
 
-As each step of \LearnMarginals is a consistent estimator, it is
-  straightforward that as whole it is a consistent estimator for $\sG$.
-We describe its asymptotic complexity to see what the dependence on the
-  parameters of the problem are.
+\LearnMarginals~ combines two consistent algorithms, \TensorFactorize and
+  \LearnClique, and is thus consistent itself. Using results from
+  \citet{anandkumar12moments,anandkumar13tensor} we can show that that
+  learning $\mOpp{v_1}{i}$ for the bottleneck $h_i$ with views $x_{v_1},
+  x_{v_2}, x_{v_3}$ has the following sample complexity,
+\begin{align*}
+  \|\mOpphat{v_1}{i} - \mOpp{v_1}{i}\|^2_F &= \frac{1}{\sqrt{n}} O( \frac{k {\pi\oft{i}}_{\max}^2}{\sigma_k(M_{v_1,v_2})^5} ). 
+\end{align*}
 
-\begin{lemma}(Sample complexity for $\sC$)
-
-We require some number of samples to learn the factor for $\sC$.
+We next describe the asymptotic variance for \LearnClique when using the
+  method of moments.
+\begin{lemma}(Asymptotic variance of $Z_\sC$)
+  \label{lem:mom-variance}  
+  The asymptotic variance of $Z_\sC$, $\Sigma^\mom_\sC$ is,
+  \begin{align*}
+    \Sigma^{(\mom)} &= \mOppit{\sV}{\sC} \Sigma_\sV \mOppi{\sV}{\sC},
+  \end{align*}
+  where $\pinv{\mOpp{\sV}{\sC}} \eqdef \pinv{\mOpp{v_1}{i_1}} \otimes
+    \cdots \otimes \pinv{\mOpp{v_m}{i_m}}$, a $k^m \times d^m$ matrix, and 
+\begin{align*}
+  \Sigma_\sV &= \diag  M_\sV  - M_\sV M_\sV^T,
+\end{align*}
+  is the variance of the observations $M_\sV$.
 \end{lemma}
 \begin{proof}
-
-For the sake of simplicity, let us assume that our dataset of $n$
-  samples is partitioned into two sets.
-  The first $\bytwo{n}$ samples are used to estimate $\mOpphat{v}{i}$ for
-  every $h_i \in H$ and $x_v \in \sV_{h}$ in step 1 of \algorithmref{directed}.
-\citet{anandkumar13tensor} show that 
+We have for each clique,
 \begin{align*}
-  \|\mOpphat{v}{i} - \mOpp{v}{i}\|_F &= O( \frac{k \pi_{\max}^2}{\sigma_k(M_2)^5} ). 
+  Z_{\sC} &= \mO_\sV(\mOppi{v_1}{i_1},\cdots,\mOppi{v_m}{i_m}) \\
 \end{align*}
 
-The next $\bytwo{n}$ samples are used to compute the moments $M_\sV$ for
-  every clique in step 2.
-
-We have for each clique,
-\begin{align*}
-  Z_{\sC} &= \mO_\sV(\pinv{\mOpp{v_1}{i_1}},\cdots,\pinv{\mOpp{v_m}{i_m}}) \\
-\end{align*}
-
-Choosing to represent $Z_\sC$ as a vector and $\mOpp{\cdot}{\cdot}$ as a matrix,
-let $\pinv{\mOpp{\sV}{\sC}} \eqdef \pinv{\mOpp{v_1}{i_1}} \otimes \cdots \otimes \pinv{\mOpp{v_m}{i_m}}$,
+Choosing to represent $Z_\sC$ and $M_\sV$ as vectors and
+$\mOppi{v_1}{i_1} \otimes \cdots \otimes \mOppi{v_m}{i_m}$ as the
+matrix, we can rewrite this representation as,
 \begin{align*}
@@
 \end{align*}
-where $\Sigma_\sV$ is the variance. Because $M_\sV$ describes the discrete marginal distribution, we have 
+where $\Sigma_\sV$ is the variance of the observations. 
+Finally, we have that $Z_\sC$ goes as
 \begin{align*}
-  \Sigma_\sV &= \diag  M_\sV  - M_\sV M_\sV^T,
+  \sqrt{n}(\hat Z_\sC - Z_\sC) \convind \sN(0, \pinvt{\mOpp{\sV}{\sC}} \Sigma_\sV \pinv{\mOpp{\sV}{\sC}}).
 \end{align*}
+\end{proof}
 
-Thus, $Z_\sC$ goes as
-\begin{align*}
-  \sqrt{n}(\hat Z_\sV - Z_\sV) \convind \sN(0, \pinvt{\mOpp{\sV}{\sC}} \Sigma_\sV \pinv{\mOpp{\sV}{\sC}}).
-\end{align*}
-
-\todo{Convert from $Z_\sC$ to the actual CPTs}.
-
-\end{proof}
+Note that extending these results to finite sample bounds can be done
+  via a straightforward application of perturbation bounds.

### 2014-01-31 doc/icml2014/discussion.tex
Updated draft

\section{Discussion}
\label{sec:discussion}

In this paper, we've described an algorithm that can consistently
  estimate parameters for the class of directed and undirected graphical
  models for which each hidden variable has at least three conditionally
  independent observed variables (\propertyref{bottleneck}).
This condition is sufficiently general to capture a variety of popular
  models like latent trees and grids (like a pairwise latent Markov
  random field).
The undirected model setting applies to log-linear models, allowing
  for discriminative featurization; furthermore, in practice, the
  bottleneck condition can be enforced on a model if the observed
  features can be split into two uncorrelated sets.

The key property we exploited was the presence of exclusive views
  (\propertyref{exclusive-views}) with learnable parameters.
In some sense, the presence of these views helps tie down the hidden
  variables to the data. 
While other authors, e.g \citet{anandkumar12lda, anandkumar12linear,
  halpern13noisyor} are able to learn parameters for models outside this
  family, they do so with some knowledge of the correlation structure
  between variables. 
In a pre-processing step, they ``subtract'' out the correlations and
  recover a graph that satisfies the bottleneck property
  \propertyref{bottleneck}.
Thus, while \propertyref{bottleneck} is much stronger than
  \propertyref{exclusive-views}, we expect that without any knowledge of
  the correlation between variables, \propertyref{bottleneck} is
  necessary for identifiability.

### 2014-01-31 doc/icml2014/intro.tex
Updated draft

@@
 % 1. Latent variable models are good.
-Latent variable models offer a succinct representation of a rich model
-family. 
+Latent variable graphical models offer a succinct representation of
+a rich model family. 
 % 2. Learning them is hard.
 Despite their success across many fields
-\cite{quattoni04crf,haghighi06prototype,liang06discrimative,kirkpatrick10painless},
+\cite{quattoni04crf,beaumont2004bayesian,haghighi06prototype,liang06discrimative,kirkpatrick10painless},
 learning these models remains a difficult problem due to the
@@
 factorize the moments into parameters, providing strong theoretical guarantees.
-However, these methods are not as universally applicable as EM, applying to a limited set of models.
+However, these methods apply to a limited set of models, and are thus
+not as universally applicable as EM.
 
@@
 In this work, we exploit the spectral method to learn moment constraints
-on the observed variables and show how these constraints lead to consistent parameter estimates for a broad class of graphical models.
+  on the observed variables and show how these constraints lead to consistent parameter estimates 
+  for any graphical model with the property that there exist at least
+  three ``views'' or conditionally independent observed variables for
+  every hidden variable (\propertyref{bottleneck}).
 
@@
 Our approach is illustrated in \figureref{approach}. 
-First, we identify conditionally independent observed variables $x_1,
-  x_2, x_3$ for each hidden variable $h$ (bottlenecks). 
-We use the tensor factorization algorithm of
+First, we identify the views $x_1, x_2, x_3$ for each hidden variable $h$. 
+We then use the tensor factorization algorithm of
   \citet{anandkumar12moments,anandkumar13tensor} to identify the
   conditional moments.
-Given these factors, we show that for every clique of hidden variables
-  $\sC$, the piecewise likelihood of the observed variables is convex in
-  the marginal distribution. 
+Given these factors, we show that there exists a set of observed
+  variables for every clique such that the likelihood is convex in the
+  parameters of that clique. 
 This guarantees that we can recover the true marginal distributions
@@
   consistently recover the true parameters of the model.
-In \sectionref{piecewise}, we detail our algorithm and describe the
+In \sectionref{directed}, we detail our algorithm and describe the
   settings and technical assumptions in which this approach is guaranteed
@@
   grid model depicted in \figureref{approach} under the condition that
-  $O$ is full-rank.
+  $O$ and $T$ is full-rank.
 
-For undirected log-linear latent variable models, this approach allows us to 
-  recover the marginals for each clique, given which the learning
-  problem is convex (\sectionref{undirected}).
+Our approach also applies to undirected log-linear latent variable
+  models, where we use the recovered marginals to make the learning
+  problem convex (\sectionref{undirected}).
 Finally, our approach gracefully extends to more general model families
@@
   EM.
-Empirically, we find that this approach leads to better solutions
-  (\sectionref{experiments}).

### 2014-01-31 doc/icml2014/related.tex
Updated draft

@@
 \section{Related work}
+\label{sec:related}
 
-Anandkumar - linear bayesian nets
+Parameter estimation for latent variable models using the method of
+  moments has received much attention in last few years.
+Our work uses the \TensorFactorize algorithm presented by
+  \citet{anandkumar13tensor}, though \citet{anandkumar12moments} first
+  studied the consistent parameter estimation algorithm for the
+  three-view mixture model.
+%\todo{What about Mossel and Roch?}
+Several authors have extended this work to settings that do not meet the
+  criterion \propertyref{bottleneck}, but rely on knowing more about
+  the underlying correlation structure of the problem.
+\citet{anandkumar12lda} are able to recover parameters for correlated
+  discrete mixtures, e.g. latent Dirichlet allocation; their approach
+  follows that of the three-view mixture model, with a pre-processing step
+  that subtracts the correlation between the views.
+\citet{anandkumar12linear} study a special family of Bayesian networks
+  in which the observed variables are linearly related to a sparse set of
+  hidden variables. Once again, the key step is to identify the
+  correlations between observed variables and exclude them from the views.
+\citet{halpern13noisyor} propose an algorithm to recover factors for
+  a more densely connected bipartite noisy-or network; they also use
+  bottlenecks to learn parameters, but exploit properties of the model
+  to negate correlations from learned edges, allowing for more
+  bottlenecks to be discovered. 
+In comparison, our work applies to a general family of models, including
+  latent tree models, directed and undirected grids, with no assumptions
+  on the correlation structure aside from independences encoded in the
+  graph.
 
-Le song - trees
-
-David sontag - noisy or
+Another prominent line of work in the method of moments community has
+  focussed on recovering parameters for a transformation that admits an {\em observable
+  representation}, i.e. one whose marginals can be directly computed from the data.
+\citet{song2011spectral} propose an algorithm to learn an observable
+  operator representations for latent tree graphical models, like the
+  one in \figureref{examples-tree}, assuming \propertyref{bottleneck}. 
+Their approach is similar to ours, but restricted to trees with observed
+  variables at the leaves.
+\citet{parikh12spectral} extend this approach to general graphical
+  models in which \propertyref{bottleneck} holds, but consider the latent
+  junction tree representation. 
+Consequently, the observable representations have size atleast that of
+  a node in the junction tree, which will be of the order of the treewidth
+  of the graph. 
+Our algorithm only constructs moments of the order of the cliques in the
+  graph, which can be exponentially smaller. 
+For example, an $n\times n$ grid model has a treewidth of $n$, but each
+  clique is of size at most $2$ in the undirected case and $3$ in the
+  directed case.
+In some sense, an observable operator re-formulation of our algorithm
+  subsumes this work.

### 2014-01-31 doc/icml2014/undirected.tex
Updated draft

@@
 
-As before, let $\sG$ be a graphical model with variables $V$, with $X
-  \subseteq V$ being observed and $H = V \setminus X$ hidden.
+As before, let $\sG$ be a graphical model with observed variables $X$ and hidden variables $H$.
 We assume that the undirected graphical model is a log-linear latent
@@
 \begin{align*}
-  p_\theta(\vx, \vh) &= \exp( \theta^T\phi(x,z) - A(\theta) ),
+  p_\theta(\vx, \vh) &= \exp( \sum_\sC \theta_\sC^T \phi_\sC(\vx,\vh) - A(\theta) ),
 \end{align*}
-where $\phi(x,z)$ is a feature vector with components $\phi_\sC(x,z)$
-  for every clique and $A(\theta) = \sum_{\vx, \vh}  \exp(
-  \theta^T\phi(x,z) )$ is the log-partition function.
-While $\phi_\sC(x,z)$ can be an arbitrary featurization, we assume
-  without loss of generality that $\phi_\sC(x,z)$ are the indicator
-  features $\BI[x,z]$\footnote{Note that $\BI[x,z]$ are sufficient
+where $\phi_\sC(\vx,\vh)$ is a feature vector for the clique $\sC$ and
+  $A(\theta) = \sum_{\vx, \vh}  \exp( \sum_\sC \theta_\sC^T \phi_\sC(\vx,\vh) )$ is the
+  log-partition function.
+While $\phi_\sC(x,h)$ can be an arbitrary featurization, we assume
+  without loss of generality that $\phi_\sC(x,h)$ are the indicator
+  features $\BI[x,h]$\footnote{Note that $\BI[x,h]$ are sufficient
   statistics for the model, and hence we do not lose any statistical
-  efficiency in using them}.
+  efficiency in using them.}.
 
@@
       &= \arg\max_{\theta} \sum_{\vx \in \sD} \log p_\theta(\vx) \nonumber \\
-      &= \arg\max_{\theta} \sum_{\vx \in \sD} \log \sum_{\vh} p_\theta(\vx,\vh) \nonumber \\
-      &= \arg\max_{\theta} \sum_{\vx \in \sD} \log \sum_{\vh} \exp( \theta^T\phi(\vx,\vh) - A(\theta) ). \label{eqn:obj-ml}
+      &= \arg\max_{\theta} \sum_{\vx \in \sD} \log \sum_{\vh \in \sH} p_\theta(\vx,\vh) \nonumber \\
+      &= \arg\max_{\theta} \sum_{\vx \in \sD} \log \sum_{\vh \in \sH} \exp( \theta^T\phi(\vx,\vh) - A(\theta) ). \label{eqn:obj-ml}
 \end{align}
 Note that this problem is not concave in $\theta$. 
-
-An alternate objective is to optimize the following lower-bound on the log-likelihood, 
+However, given the moments $\E[ \phi_\sC(\vx,\vh) ] = Z_\sC$, the
+  problem simplifies to,
 \begin{align}
-  \sL_p(q,\theta) &= \sum_{\vx \in \sD} \sum_{\vh} q(\vh \given \vx) \log\frac{p_\theta(\vx,\vh)}{q(\vh \given \vx)} \\
-    &= \KL( q(\vh \given \vx) || p_{\theta}(\vh \given \vx)) \label{eqn:obj-kl}
+  \hat \theta 
+  &= \arg\max_{\theta} \sum_{\sC \in \sG} \theta_\sC^T Z_\sC - A(\theta), \label{eqn:obj}
 \end{align}
-
-Importantly, given $q$, the above objective is convex in $\theta$ and vice versa.
-It can be shown\cite{} that the local-optima of \equationref{obj-kl}
-are also local-optima of \equationref{obj-ml}. Usually, one
-iteratively optimizes $q$ and $p_\theta$ from some initial parameters
-$\theta_0$ to arrive at a local optimum.
-
-This reduces to,
-\begin{align}
-  \sL_p(\theta) &= \theta^T Z - A(\theta),
-\end{align}
-which is a convex problem.
-
-However, from the previous section, for any model satisfying
-  \property{exclusive-views}, we have a consistent estimator for the
-  marginals $q \equiv Z$. 
-Therefore, we can optimize for $\theta$ given this estimate of $q$,
-  $\hat q$.
+which is concave in $\theta$.
+In \sectionref{directed}, we showed that \LearnMarginals is a consistent
+  estimator for $Z_\sC$ for any model satisfying \propertyref{exclusive-views}. 
+The result is that the output of \algorithmref{undirected}, $\hat\theta$, is a consistent estimator for $\theta^*$
+  when the model $p_\theta(\theta)$ is identifiable.
 
@@
   \begin{algorithmic}
-    \REQUIRE A graphical model $\sG$ satisfying \textbf{(P1)}, \textbf{(P2)}, data $\sD$
-    \ENSURE A consistent estimate $\hat \theta$ of the parameters of the model.
+    \REQUIRE A graphical model $\sG$ satisfying \propertyref{bottleneck}, data $\sD$
+    \ENSURE A consistent estimate of parameters $\hat \theta$
       \STATE Apply \LearnMarginals to $\sG$ to learn clique marginals $Z_\sC$.
-      \STATE Optimize \equationref{obj-kl} with respect to $\theta$.
+      \STATE Optimize \equationref{obj} with respect to $\theta$.
   \end{algorithmic}
@@
 
-This approach is outlined in \algorithmref{undirected}.
-
-\todo{Do we have to worry about the fact that this is constrained?}
-\todo{What about model un-identifiability?}
-
 \paragraph{Learning with partial moment constraints}
 
-In more general models, \propertyref{exclusive-views} may not hold for
-  every clique in the graph, however we can still estimate the marginals
-  $Z_\sC$ for cliques in which it does.
-In this section, we will describe a natural method to use these
-  constraints to simplify the parameter search problem.
+In more general models, \propertyref{exclusive-views} may only hold for
+  some subset of the cliques $\sP$ in the graph.
+To utilize these partial moment constraints, we propose using posterior
+  regularization \citep{graca08em} to constrain expectation-maximization.
+While the optimization problem \eqref{eqn:obj-ml} is non-convex, we
+  expect that the measurements aid recover of parameters, and indeed see
+  this is true for experiments on simulated data.
+\todo{Include figure}
 
-Constrain the KL to be in $Q$.
+Posterior regularization optimizes the following objective,
+\begin{align*}
+  \hat \theta &= \arg\min_{\theta} \arg\min_{q \in \sQ} \KL( q(\vx,\vh) \| p_{\theta}(\vx, \vh) ),
+\end{align*}
+where $q$ is a marginal distribution that belongs in constrained set
+$\sQ$. In our case, $\sQ$ is constrained to agree with the partial
+moments we recovered using \LearnMarginals, i.e. 
+\begin{align*}
+  \sQ &= \{ q(\vx,\vh) \given \forall \sC \in \sP ~ \E_{q(\vx,\vh)}[\phi_\sC(\vx, \vh)] = Z_\sC \}.
+\end{align*}
+
+In practice, we would like to regularize for error in our
+  estimates of $Z_\sC$; to do so, we use the measurements model of
+  \citet{liang09measurements} which defines a Bayesian model for the noise
+  in the marginals $Z_\sC$. The resultant optimization problem is,
+\begin{align*}
+  \hat\theta 
+  &= \arg\min_\theta - \sum_{x \in \sD} \log p_\theta(x) + \half \eta_\theta \| \theta \|^2 \\
+  &\quad + \half \eta_\beta \| Z_\sC - \sum_{\vx \in \sD} \E_{\theta}[\phi_\sC(\vx,\vh)] \|^2,
+\end{align*}
+where $\eta_\theta$ and $\eta_\beta$ are regularization parameters.
 
@@
 
-Image segmentation
+\begin{figure}
+  \label{fig:examples-mrf}
+  \centering
+  \subimport{figures/}{mrf.tikz}
+  \caption{Markov Random Field with 2 views}
+\end{figure}
 
-Separate features into views
+Consider the example in \figureref{examples-mrf}, which is a pairwise Markov
+  random field with two observed variables for every hidden variable. 
+In practice, if two features are known to be independent, then they can
+  be separated into independent observed variables.
+Let the parameters of the model be $O \eqdef \Phi(x^a_{i,j}, h_{i,j})
+  = \Phi(x^b_{i,j}, h_{i,j})$ and $T \eqdef \Phi(h_{i,j}, h_{i \pm 1,j \pm 1})$. 
+Note that as this is an undirected model, the clique marginals do not
+  correspond to the parameters of the model.
 
-\todo{Some challenge example requiring measurements em}
+To recover the parameters of this model, note that every hidden
+  variable, $h_{ij} \in \sH$ is a bottleneck, with views
+  $x^{a}_{ij},x^{b}_{ij}$ and any other observed variable in the graph. 
+Thus, \LearnMarginals will be able to learn the marginals of each clique
+  and \LearnLogLinear will give us a consistent estimate of the
+  parameters.
+
+However, the objective \equationref{obj} requires us to compute the
+  partition function, $A(\theta)$ and its gradient. 
+This scales with the treewidth of the model, which is the number
+  of rows/columns of the grid in this example, and hence can be
+  intractable.
+A more computationally efficient, but still consistent, alternative is
+to optimize the pseudo-likelihood \todo{how do I write this out}.

### 2014-01-31 doc/icml2014/directed.tex
Updated figures and some details and notation

@@
     $Z_\sC$ be the marginal distribution for that clique. 
-  Then every mode-$i$ unfolding of $Z_\sC$ has full column rank.
+  Then every mode-$i$ unfolding of $Z_\sC$ has full column rank.\verify
 \end{assumption}
@@
   the observed variables are leaves with exactly one parent. 
+There is a one-to-one correspondence between the parameters of this
+  transformed model and the original one.
 \end{lemma}
@@
 
-  Let $x$ be an observed variable with parents $h_1, \cdots, h_p$ and
-  children $h_{p+1}, \cdots, h_c$. Replace $x$ with a phantom hidden
-  variable $h_x$, and make $x$ the child of $h_x$ (\figureref{reduction}). The domain of $h_x$
-  has the same as $x$, and $\mOpp{x}{h_x} = I$.
+  \providecommand{\hp}{\ensuremath{h_{\textrm phantom}}}
 
-  The parameters for $h_x$ are the same as $x$.
+  Let $x_v$ be an observed variable with parents $P$ and children $C$.
+  Consider the following transformation.
+  Replace $x_v$ with a phantom hidden variable \hp with the same
+  parents $P$ and children $C \union \{x_{v_1}, x_{v_2}, x_{v_3}\}$,
+  where $x_{v_1},x_{v_2},x_{v_3}$ are three copies of $x$
+  (\figureref{reduction}).
+
+  The domain of $\hp$ is the same as $x$
+    \footnote{Though we have assumed that all hidden variables share the
+      same domain, $[k]$, our work generalizes to hidden domains of
+      arbitrary sizes, provided \assumptionref{full-rank-plus} holds.}
+    and $\mOpp{v}{\textrm{phantom}} = I$.
+
+  Then, for any clique $\sC \contains \hp$, the parameters in the
+    original graphical model can be obtained by substituting $\hp$ with
+    $x_v$.
 \end{proof}

### 2014-01-31 doc/icml2014/intro.tex
Updated figures and some details and notation

@@
   \centering
-\begin{tikzpicture}
-
-% Import grid.
-%\node[scale=0.9] (step0) at (0,0) {\input{figures/grid.tikz}};
-\node[scale=0.9] (step1) at (0cm,0cm) {\input{figures/grid-outline.tikz}};
-\node[scale=0.75, below=0.5 cm of step1.south] { \textbf{Step 1:} Identify bottlenecks};
-\node[scale=0.9] (step2) at (5cm,0cm) {\input{figures/factors.tikz}};
-\node[scale=0.75, below=0.5 cm of step2.south] { \textbf{Step 2:} Optimize piecewise likelihoods};
-
-\draw[-latex] (step1) -- (step2);
-
-\end{tikzpicture}
+  \subimport{figures/}{approach.tikz}
   \caption{Overview of approach}

### 2014-01-31 doc/icml2014/directed.tex
Finish approach figure
Added bottleneck outlines for all graphs

--- 2014-01-31 doc/icml2014/directed.tex
@@
   \centering
-  \input{figures/grid.tikz}
-  \caption{A directed grid model.\todo{Annotate features}}
+  \input{figures/grid-outline.tikz}
+  \caption{A directed grid model.}
   \label{fig:grid}
@@
   \begin{figure}
-    \label{fig:reduction}
     \centering
@@
     \caption{Reduction to canonical form.}
+    \label{fig:reduction}
   \end{figure}
 
-  Let $x$ be an observed variable with parents $h_1, \cdots, h_p$ and
-  children $h_{p+1}, \cdots, h_c$. Replace $x$ with a phantom hidden
-  variable $h_x$, and make $x$ the child of $h_x$ (\figureref{reduction}). The domain of $h_x$
-  has the same as $x$, and $\mOpp{x}{h_x} = I$.
+  %\providecommand{\hp}{\ensuremath{h_{\textrm{phantom}}}}
+  \providecommand{\hp}{\ensuremath{h_p}}
 
-  The parameters for $h_x$ are the same as $x$.
+  Let $x_v$ be an observed variable with parents $P$ and children $C$.
+  Consider the following transformation.
+  Replace $x_v$ with a phantom hidden variable \hp with the same 
+  domain\footnote{Though we have assumed that all hidden variables share the
+      same domain, $[k]$, our work generalizes to hidden domains of
+      arbitrary sizes, provided \assumptionref{full-rank-plus} holds.},
+  parents $P$ and children $C \union \{x_{v_1}, x_{v_2}, x_{v_3}\}$,
+  where $x_{v_1},x_{v_2},x_{v_3}$ are three copies of $x$
+  (\figureref{reduction}). Also, set $\mOpp{v}{p} = I$.
+
+  Then, there is a one-to-one correspondence between every value of
+  $\hp$ and $x_v$. Consequently, for any clique $\sC \contains \hp$, the
+  parameters in the original graphical model can be obtained by
+  substituting $\hp$ with $x_v$.
 \end{proof}

### 2014-01-31 doc/icml2014/directed.tex
Added some figures and made some changes to sample complexity stuff

@@
 Applying \TensorFactorize to every bottleneck gives us candidate views
-  for every hidden variable. 
-To proceed, we need to show two things; (i) under what assumptions
-  does \assumptionref{full-rank} hold for each bottleneck and (ii)
-  that the candidate views actually imply that \propertyref{exclusive-views}
-  holds for every clique in the graph.
-
-To address the first, we will need that the following assumption on the
-  parameters of the model:
+  for every hidden variable, provided \assumptionref{full-rank} holds for each bottleneck;
 \begin{assumption} 
   \label{asm:full-rank-plus}
-  For every clique $\sC = \{h_{i_1}, \cdots, h_{i_m}\} \in \sG$, let
-    $Z_\sC$ be the marginal distribution for that clique. 
-  Then every mode-$i$ unfolding of $Z_\sC$ has full column rank.\verify
+  For every hidden variable $h_i$, $\pi\oft{i} \succ 0$ and
+    $\mOpp{v}{i}$ has full column rank.
 \end{assumption}
-Intuitively, the assumption guarantees that the marginal probability of
-  any hidden state is non-zero.
 
-Before we establish the second condition, we describe a reduction to
+Next, we must show that these candidate views actually imply that
+  \propertyref{exclusive-views} holds for every clique in the graph.
+Before we establish this condition, we describe a reduction to
   a canonical form that will allow us to handle the case where an observed
@@
 \LearnMarginals~ combines two consistent algorithms, \TensorFactorize and
-  \LearnClique, and is thus consistent itself. Using results from
-  \citet{anandkumar12moments,anandkumar13tensor} we can show that that
-  learning $\mOpp{v_1}{i}$ for the bottleneck $h_i$ with views $x_{v_1},
-  x_{v_2}, x_{v_3}$ has the following sample complexity,
-\begin{align*}
-  \|\mOpphat{v_1}{i} - \mOpp{v_1}{i}\|^2_F &= \frac{1}{\sqrt{n}} O( \frac{k {\pi\oft{i}}_{\max}^2}{\sigma_k(M_{v_1,v_2})^5} ). 
-\end{align*}
-
+  \LearnClique, and is thus consistent itself. 
 We next describe the asymptotic variance for \LearnClique when using the

### 2014-01-31 doc/icml2014/abstract.tex
Updated figures and some details and notation
Percy's changes

--- 2014-01-27 doc/icml2014/abstract.tex
@@
 \begin{abstract}
+
+Recent work in method of moments provide consistent estimates
+  for latent-variable models, avoiding local optima issues,
+  but these methods can only be applied to certain types of graphical models.
+In this work, we show that the method of moments in conjunction
+  with a composite marginal likelihood objective
+  yields consistent parameter estimates for a much broader
+  class of directed and undirected graphical models,
+  including loopy graphs with high treewidth.
+%In particular, our algorithm applies to any graphical model wherein each
+  %hidden variable has at least three conditionally independent observed
+  %variables.
+  %that sidestep local optima issues for a general class of
+  %directed and undirected graphical models.
+%Parameter estimation in latent-variable models based on EM
+  %is prone to local optima.
+  %due to the non-convexity of the marginal log-likelihood. 
+%In this paper, we extend recent advances in the method of moments 
+  %to provide consistent parameter estimates
+  %that sidestep local optima issues for a general class of
+  %directed and undirected graphical models.
+%In particular, our algorithm applies to any graphical model wherein each
+  %hidden variable has at least three conditionally independent observed
+  %variables.
+%This family includes loopy graphs with high treewidth.
+Specifically, we use tensor factorization to reveal partial information
+about the latent variables, rendering the otherwise non-convex negative
+log-likelihood convex.
+%We use the method of moments to tie each hidden to some observed
+  %variables, and show that the rest of the problem can be cast as convex
+  %optimization of the likelihood.
+Our approach gracefully extends to models outside our class by
+  incorporating the partial information via posterior regulraization.
 \end{abstract}
 
-\paragraph{Errata}
+\paragraph{Last updated:} \today, \currenttime
+
+\paragraph{Errata / Incomplete}
 \begin{enumerate}
-  \item The likelihoods as defined are concave, not convex.
-  \item Define parameters $\theta$ more uniformly.
+  \item Use \texttt{\textbackslash left(}.
+  \item Use $\ell$ instead of $L$.
+  \item convex and negative log likelihood.
 \end{enumerate}

### 2014-01-31 doc/icml2014/background.tex
Percy's changes

@@
 Let $\Pa(a) \subset X \union H$ be the parents of a variable $a$ in a directed
-  model, and $\sN(a) \subset X \union H$ be the neighbours of $a$ in an
+  model, and $\sN(a) \subset X \union H$ be the neighbors of $a$ in an
   undirected model.
@@
   $h_1, h_2, \cdots, h_M \in H$ to denote hidden variables.
-For ease of exposition, we assume that $x_i \in \{1, \cdots, d\}$ for
-  all $i$, and that $h_i \in \{1, \cdots, k\}$ for all $i$.
-Let $\sX$ be the joint domain of the observed variables, i.e. $\sX \eqdef
-  \underbrace{d \times d \cdots \times d}_{L ~\textrm{times}}$ and
-let $\sH$ be the joint domain of the hidden variables, i.e. $\sH \eqdef
-  \underbrace{k \times k \cdots \times k}_{M ~\textrm{times}}$.
+For an integer $n$, let $[n] = \{ 1, \dots, n \}$.
+For ease of exposition, we assume that $x_i \in [d]$ for
+all $i$, and that $h_i \in [k]$ for all $i$.
+Let $\sX$ be the joint domain of the observed variables, i.e. $\sX \eqdef [d]^L$ and
+let $\sH$ be the joint domain of the hidden variables, i.e. $\sH \eqdef [k]^M$.
 
-We assume that the model $\sG$ is parameterized at every clique $\sC$.
-If $\sG$ is directed, the parameters correspond to
-  the conditional probability tables for each clique.
-If $\sG$ is undirected, the parameters correspond to the clique
-  potentials.
+% PL: don't know what this means
+%We assume that the model $\sG$ is parameterized at every clique $\sC$.
+If $\sG$ is directed, the model parameters correspond to
+the conditional probability tables for each clique $(\Pa(a), a) \in \sG$.
+If $\sG$ is undirected, we assume a log-linear parametrization of each clique $\sC \in \sG$.
 
@@
     \sL &\eqdef \max_{\theta} \sum_{\vx \in \sD} \log p_\theta(\vx) \\
-        &=      \max_{\theta} \sum_{\vx \in \sD} \log \sum_{\vh \in \sH} p_\theta(\vx,\vh)
+        &=      \max_{\theta} \sum_{\vx \in \sD} \log \sum_{\vh \in \sH} p_\theta(\vx,\vh).
   \end{align*}
-  which is statistically efficient but computationally intractable
-  because we must marginalize over latent variables $\vh$.
-In practice, one uses gradient-based optimization procedures (e.g., EM
+  Maximum likelihood is statistically efficient,
+  but in general computationally intractable
+  because marginalizing over latent variables $\vh$ yields a non-convex objective.
+In practice, one uses local optimization procedures (e.g., EM
   or L-BFGS) on the marginal likelihood, which can get stuck in local
@@
 For a tensor $T \in \Re^{d \times \ldots \times d}$ and a vector $\vec
-  i = (i_1, \cdots, i_l)$, $T[\vec i] = T_{i_1 \ldots i_l}$.
+  i = (i_1, \cdots, i_l)$, define the projection $T[\vec i] = T_{i_1 \ldots i_l}$.
 Similarly, for a set $\sC$ of size $k$ and vector $\vec x \in \Re^d$, we
   will use the notation $\vec x_\sC$ to get a vector $\Re^k$ which selects
-  the appropriate elements of $\vec x$
+  the appropriate elements of $\vec x$.
 
-We use $\otimes$ to denote the tensor product, e.g if $u \in \Re^k,
-  v \in \Re^d$, then $u \otimes v \in \Re^{k \times d}$.
-For a $\ell$-th order tensor $T \in \Re^{d \times \ldots \times
-  d}$ and vectors $x_1, \cdots, x_\ell \in \Re^{d},$, we define 
+We use $\otimes$ to denote the tensor product: if $u \in \Re^d,
+  v \in \Re^k$, then $u \otimes v \in \Re^{d \times k}$.
+For an $\ell$-th order tensor $T \in \Re^{d \times \ldots \times
+  d}$ and vectors $x_1, \cdots, x_\ell \in \Re^{d}$, define 
+  the partial application:
 \begin{align*}
   T(x_1, \cdots, x_\ell) 
-      &= \sum_{i_1 = 1}^{d} \cdots \sum_{i_\ell = 1}^{d} 
+      %&= \sum_{i_1 = 1}^{d} \cdots \sum_{i_\ell = 1}^{d} 
+  &= \sum_{\bi}
             T[i_1, i_2, \ldots, i_l] x_1[i_1] \cdots x_\ell[i_\ell].
@@
   T(M_1, \cdots, M_\ell)[\vec j]
-      &= \sum_{i_1 = 1}^{d} \cdots \sum_{i_\ell = 1}^{d} 
-            T[\vec i] {M_1}[i_1,j_1] \cdots M_\ell[i_\ell, j_\ell],
+      %&= \sum_{i_1 = 1}^{d} \cdots \sum_{i_\ell = 1}^{d} 
+  &= \sum_{\bi}
+            T[\vec i] {M_1}[i_1,j_1] \cdots M_\ell[i_\ell, j_\ell].
 \end{align*}
-where $\vec i = (i_1, \cdots, i_\ell)$ and $\vec j = (i_1, \cdots, j_\ell)$
+%where $\vec i = (i_1, \cdots, i_\ell)$ and $\vec j = (i_1, \cdots, j_\ell)$
 
-To denote the moments of observed variables, we will use the notation
+To denote the moments, we will use the notation:
 \begin{align*}
@@
   d \times d},\ldots$.
-For the moments of hidden variables, we will use 
+For the latent marginals, we will use 
 \begin{align*}
@@
 
-Recently, \citet{anandkumar12moments} presented a consistent algorithm
+\citet{anandkumar12moments} presented a consistent estimator
   using the method of moments to learn the parameters of a three-view
   mixture model (\figureref{three-view}), an important case of directed
-  graphical models, that sidesteps the local optima problem altogether. 
-In the model, we have three observed variables $x_1, x_2, x_3$
-  that are independent conditioned on a latent variable $h$.  
+  graphical models. %, that sidesteps the local optima problem altogether. 
+In that model, we have three observed variables $x_1, x_2, x_3$
+  that are independent conditioned on a latent variable $h_1$.  
 The parameters of this model are the mixture probabilities $\pi\oft 1 \eqdef
@@
   The key idea is that the observed moments $\mO_{12}, \mO_{23}, \mO_{13}$ and
-  $\mO_{123}$ admit a factorization in terms of parameters;
+  $\mO_{123}$ admit a factorization in terms of parameters:
 \begin{align*}
-  \mO_{ij} &= \sum_h {\pi \oft 1}[h] {\mOppt{i}{1}}[h] \otimes {\mOppt{j}{1}}[h] \quad \forall i,j \in \{1,2,3\} \\
-  \mO_{123} &= \sum_h {\pi \oft 1}[h] {\mOppt{i}{1}}[h] \otimes {\mOppt{2}{1}}[h] \otimes  {\mOppt{2}{1}}[h].
+  \mO_{ij} &= \sum_h {\pi \oft 1}[h] {\mOpp{i}{1}}[h] \otimes {\mOpp{j}{1}}[h] \quad \forall i,j \in \{1,2,3\} \\
+  \mO_{123} &= \sum_h {\pi \oft 1}[h] {\mOpp{1}{1}}[h] \otimes {\mOpp{2}{1}}[h] \otimes  {\mOpp{3}{1}}[h].
 \end{align*}
@@
 \citet{anandkumar13tensor} presented a robust tensor power method, which
-  we call $\TensorFactorize$, which converts consistent estimates of the observed moments 
-  into consistent estimates of the parameters, $\pi\oft 1, \mOpp{1}{1},
+  we call $\TensorFactorize$, which converts estimates of the observed moments 
+  into estimates of the parameters, $\pi\oft 1, \mOpp{1}{1},
   \mOpp{2}{1},\mOpp{3}{1}$ (up to permutation of the columns), provided
@@
 of the whitened $\mO_{123}$; unwhitening yields the columns of $\mOpp{3}{1}$.
-  $\mOpp{1}{1}$ and $\mOpp{2}{1}$ can then be recovered with some algebra.
+  $\mOpp{1}{1}$ and $\mOpp{2}{1}$ can then be recovered with a few matrix operations.
+
+The tensor factorization method attacks the heart of the non-convexity
+  in latent-variable models.  In the next two sections,
+  we will see how to build on top of this result to estimate directed (Section~\ref{sec:directed})
+  and undirected (Section~\ref{sec:undirected}) models.

### 2014-01-31 doc/icml2014/directed.tex
Percy's changes

@@
-\section{Consistent Parameter Estimation in Directed Models}
+% DONE: too wordy
+%\section{Consistent Parameter Estimation in Directed Models}
+\section{Directed models}
 \label{sec:directed}
 
-The tensor factorization method attacks the heart of the non-convexity
-  in latent variable models.
-  Once we recover the conditional moments $\mOpp{v}{j} \eqdef \Pr(x_v
-  | h_j)$ for some $v, j$, we present a systematic approach to learn the
-  conditional probability tables for every clique for a class of directed
-  graphical models.
-We begin by illustrating our approach with an example, followed by
-  describing the algorithm in full generality.
+[TODO: need to say we're going to do this for the infinite data case,
+and then study bounds]
+
+[TODO: try to be consistent: either use latent or hidden]
+
+% DONE: need more setup of what the goal is
+%The tensor factorization method attacks the heart of the non-convexity
+  %in latent variable models.
+  In this section, we will develop a consistent parameter estimate for a class of directed graphical models.
+  In directed models, the conditional moments $\mOpp{v}{i} \eqdef \Pr(x_v \mid h_i)$
+  are only a subset of the parameters.
+  %Once we recover the conditional moments $\mOpp{v}{j} \eqdef \Pr(x_v
+  %| h_j)$ for some $v, j$, we present a systematic approach to learn the
+  %conditional probability tables for every clique for a class of directed
+  %graphical models.
+  Let us see how we can estimate the remaining parameters with an example (Section~\ref{sec:directedExample}).
+  Section~\ref{sec:directedGeneral} will describe the algorithm in full generality.
 For ease of exposition, we will describe our algorithm entirely in the
@@
 \subsection{Example: Directed grid model}
+\label{sec:directedExample}
 
@@
 Consider the directed grid model shown in \figureref{grid} which
-  captures all the different dependency structures possible in
+  % DONE: all the different dependency => too strong
+  captures the local dependency structures possible in
   a Bayesian network.
@@
 The parameters of this model are the conditional probability tables
-  $\pi \eqdef \Pr(h_1), T \eqdef \Pr(h_2 | h_1) = \Pr(h_3 | h_1),
-  V \eqdef \Pr(h_4 | h_2, h_3)$ and $O \eqdef \Pr(x^a_i | h_i)
-  =  \Pr(x^b_i | h_i)$. 
+$\pi \eqdef \Pr(h_1) \in \Re^k, T \eqdef \Pr(h_2 | h_1) = \Pr(h_3 | h_1) \in \Re^{k \times k},
+V \eqdef \Pr(h_4 | h_2, h_3) \in \Re^{k \times k \times k}$ and $O \eqdef \Pr(x^a_i | h_i)
+=  \Pr(x^b_i | h_i) \in \Re^{d \times k}$. 
 Finally, we assume $O$ and $T$ have full column rank.
@@
   \TensorFactorize, as reviewed in \sectionref{setup}, to recover
-  the $O$.
+  $O$.
 
 \paragraph{Estimating $\pi$}
-The moments of $x^a_1$,$\mO_1 \eqdef \Pr(x^a_1)$ are directly related to
-  $\pi$; $\mO_1 = O \pi$. 
+The moments of $x^a_1$, $\mO_1 \eqdef \Pr(x^a_1)$ are directly related to
+  $\pi$ by $\mO_1 = O \pi$. 
 By \assumptionref{full-rank}, $O$ has full column rank and thus can be
-  inverted to recover $\pi$; $\pi = \pinv O \mO_1$.
+  inverted to recover $\pi$: $\pi = \pinv O \mO_1$.
 
@@
 Similarly, we can write down the moments of $x^a_1, x^a_2$, $\mO_{12}
-  \eqdef \Pr(x^a_1, x^a_2)$ in terms of the pairwise marginals $\mH_{12}
-  \eqdef \Pr(h_1, h_2)$ and solve for $\mH_{12}$;
+  \eqdef \Pr(x^a_1, x^a_2)$, in terms of the latent marginals $\mH_{12}
+  \eqdef \Pr(h_1, h_2)$ and solve for $\mH_{12}$:
 \begin{align*}
-  \mO_{12} &= O \mH_{12} O^T \\
-  \mH_{12} &= \pinv O \mO_{12} \pinvt O.
+\mO_{12} = O \mH_{12} O^\top \quad\Rightarrow\quad
+  \mH_{12} = \pinv O \mO_{12} \pinvt O.
 \end{align*}
-$T$ can be recovered from the $\mH_{12}$ by suitable normalization.
+$T$ can be recovered from the $\mH_{12}$ by suitable renormalization.
 
 \paragraph{Estimating $V$}
-Finally, we can estimate $V$ from the third-order moments $\mO_{234} \eqdef \Pr(x^a_2, x^a_3, x^a_4)$ as
-follows,
+Finally, we can estimate $V$ by renormalizing the third-order moments $\mO_{234} \eqdef \Pr(x^a_2, x^a_3, x^a_4)$,
+which can be obtained analogously:
 \begin{align*}
-  \mO_{234} &= \mH_{234}(O, O, O) \\
-  \mH_{234} &= \mO_{234}(\pinv O, \pinv O, \pinv O).
+  \mO_{234} = \mH_{234}(O, O, O) \quad\Rightarrow\quad
+  \mH_{234} = \mO_{234}(\pinv O, \pinv O, \pinv O).
 \end{align*}
@@
 \subsection{General algorithm}
+\label{sec:directedGeneral}
 
-Fundamentally, our algorithm takes advantage of having a view for every
-  hidden variable in a clique to help identify its hidden state. We define this condition as follows,
+Note that in each of the above cases,
+each hidden variable $h_i$ of a clique has its own view which is conditionally independent
+given the other variables in the clique.
+%Fundamentally, algorithm takes advantage of having a view for every
+  %hidden variable in a clique to help identify its hidden state.
+We capture this general property as follows:
 \begin{property}(Exclusive views)
   \label{prop:exclusive-views}
-A clique $\sC$ of $\sG$ is said to posses exclusive views if for every hidden variable $h_i \in \sC$,
+  A clique $\sC$ of $\sG$ is said to possess \textbf{exclusive views property} if for every hidden variable $h_i \in \sC$,
   there is some observed variable, $x_{v}$, which is conditionally
-  independent of the rest of the clique given $h_i$ and whose conditional
-  moment $\mOpp{v}{i} \eqdef \Pr(x_{v} | h_i)$ is known. 
+  independent of the rest of the clique given $h_i$.
+  and whose conditional
+  moment $\mOpp{v}{i} \eqdef \Pr(x_{v} \mid h_i)$ can be estimated consistently.
 We call $x_v$ the exclusive view of $h_i$ in $\sC$.
@@
 Consider any clique $\sC = \{h_{i_1}, \cdots, h_{i_m}\} \in \sG$ for which this property holds. Let
-  $x_{v_j}$ be the exclusive view for $h_{i_j}$ in $\sC$ and $\sV
-  = \{x_{v_1}, \cdots, x_{v_m}\}$ be the set of views for the clique
-  $\sC$.
-We can write down the moments $\mO_\sV$,
+  $x_{v_j}$ be an exclusive view for $h_{i_j}$ in $\sC$ (not necessarily unique) and $\sV
+  = \{x_{v_1}, \cdots, x_{v_m}\}$ be a set of views for the clique $\sC$.
+We can write down the moments $\mO_\sV$ as follows:
 \begin{align*}
@@
       \Pr(\Sh{\sC}) \Pr(\Sx{\sV} \given \Sh{\sC}) \\
-      &= \sum_{\vh \in sH} \Pr(\Sh{\sC}) 
+      &= \sum_{\vh \in \sH} \Pr(\Sh{\sC}) 
           \Pr(x_{i_1} | h_{v_1}) \cdots \Pr(x_{i_m} | h_{v_m}) \\
@@
 If each $\mOpp{v_j}{i_j}$ has full column rank, then we can recover the
-hidden marginals $Z_\sC$ by inverting the conditional moments,
+hidden marginals $Z_\sC$ by inverting:
 \begin{align*}
@@
 Finally, the conditional probability tables for $\sC$ can easily be gotten from
-  $Z_\sC$ via conditioning.
+  $Z_\sC$ via renormalization.
 \algorithmref{learnclique} summarizes the procedure, \LearnClique.
@@
   \begin{algorithmic}
-    \REQUIRE A $\sC$ with \propertyref{exclusive-views}.
-    \ENSURE The marginal distribution of the clique, $Z_\sC$.
+    \REQUIRE Clique $\sC$ with exclusive views (\propertyref{exclusive-views}).
+    \ENSURE Marginal distribution of the clique $Z_\sC$.
       \STATE Identify unique views $\Sx{\sV} = \{x_{v_1}, \cdots, x_{v_m}\}$.
@@
 
-If we can guarantee that every clique in a graph possesses exclusive
-  views, then we can apply \LearnClique iteratively to learn the entire
-  model.
-In the example, we were able to guarantee \propertyref{exclusive-views}
-  by identifying a hidden variable, $h^1$, with three conditionally
-  independent observed variables, $x^a_1, x^b_1, x^a_2$, followed by
-  using \TensorFactorize to learn the conditional moments.
-We define such a hidden variable to be a bottleneck;
-\begin{definition}(Bottleneck)
-  A hidden variable $h$ is said to be a bottleneck if there exists at
+% DONE: transition the content more gradually (checkpoint: what have we done so far, what next)
+% DONE: refer to properties by name in addition to by number
+We have shown how to estimate the parameters of any clique that possesses the exclusive
+  views property (\propertyref{exclusive-views}).  But which cliques have this property?
+In the example above, we were able to guarantee \propertyref{exclusive-views}
+  by identifying a hidden variable, $h^1$ along with three conditionally
+  independent observed variables, $x^a_1, x^b_1, x^a_2$.
+  %followed by using \TensorFactorize to learn the conditional moments.
+
+We define such a hidden variable to be a bottleneck:
+\begin{definition}[Bottleneck]
+  A hidden variable $h$ is said to be a \textbf{bottleneck} if there exists at
     least three conditionally independent observed variables, or views,
     $x_1, x_2, x_3$. 
-  Let $\sV_{h}$ be the set of views for $h$.
+  Let $\sV_{h}$ denote a set of views for $h$ (not necessarily unique).
 \end{definition}
-Finally, the claim we make is that we can recover parameters for any
-  graphical model where every hidden variable is a bottleneck;
-\begin{property}
+
+% DONE: don't talk about parameter recovery yet, just establish exclusive views
+%Finally, the claim we make is that we can recover parameters for any
+  %graphical model where every hidden variable is a bottleneck:
+We will prove that the following simple property guarantees that all cliques
+have the exclusive views property:
+\begin{property}[Uniformly bottlenecked]
   \label{prop:bottleneck}
@@
 \end{property}
-We note that \propertyref{bottleneck} can be relaxed if some cliques
-  share parameters;
-For example, because we can identify the observation potential from just
-  $x^a_1, x^b_1$ and $x^a_2$, we do not require that $h_2, h_3$ or $h_4$
-  be bottlenecks, and can correspondingly omit the observations $x^b_2,
-  x^b_3$ and $x^b_4$.
-In this case, we only require that the hidden variables in distinct
-  cliques share parameters.
 
-Applying \TensorFactorize to every bottleneck gives us candidate views
-  for every hidden variable, provided \assumptionref{full-rank} holds for each bottleneck;
-\begin{assumption} 
-  \label{asm:full-rank-plus}
-  For every hidden variable $h_i$, $\pi\oft{i} \succ 0$ and
-    $\mOpp{v}{i}$ has full column rank.
-\end{assumption}
+%Applying \TensorFactorize to every bottleneck gives us a set of views
+%  for every hidden variable. 
+%To proceed, we need to check that \
+%%(i) understand the assumptions under which
+%  %\assumptionref{full-rank} holds for each bottleneck, and (ii)
+%  that the candidate views actually imply that \propertyref{exclusive-views}
+%  holds for every clique in the graph.
 
-Next, we must show that these candidate views actually imply that
-  \propertyref{exclusive-views} holds for every clique in the graph.
-Before we establish this condition, we describe a reduction to
-  a canonical form that will allow us to handle the case where an observed
-  variable has multiple parents.
+% PL: this is not obviously correct
+%To address the first, we will need that the following assumption on the
+%  parameters of the model:
+%\begin{assumption} 
+%  \label{asm:full-rank-plus}
+%  For every clique $\sC = \{h_{i_1}, \cdots, h_{i_m}\} \in \sG$, let
+%    $Z_\sC$ be the marginal distribution for that clique. 
+%  Then every mode-$i$ unfolding of $Z_\sC$ has full column rank.\verify
+%\end{assumption}
+%Intuitively, the assumption guarantees that the marginal probability of
+%  any hidden state is non-zero.
 
-\begin{lemma}(Canonical form) 
+Before we establish this relationship,
+we describe a reduction of any graphical model to
+  a canonical form to make our arguments simpler.
+  % DONE: don't say this yet since reader can't appreciate it.
+  %that will allow us to handle the case where an observed
+  %variable has multiple parents.
+
+\begin{lemma}[Canonical form]
   \label{lem:reduction}
@@
 
-  %\providecommand{\hp}{\ensuremath{h_{\textrm{phantom}}}}
-  \providecommand{\hp}{\ensuremath{h_p}}
+  \providecommand{\hp}{\ensuremath{h_\text{\rm new}}}
 
@@
   Consider the following transformation.
-  Replace $x_v$ with a phantom hidden variable \hp with the same 
-  domain\footnote{Though we have assumed that all hidden variables share the
+  Replace $x_v$ with a new hidden variable \hp\ with the same
+  parents $\Pa(x_v)$ and children $\text{Ch}(x_v) \union \{x_{v_1}, x_{v_2}, x_{v_3}\}$,
+  where $x_{v_1},x_{v_2},x_{v_3}$ are three copies of $x$
+  (\figureref{reduction}). 
+  The domain of $\hp$ is the same as $x$,
+    and $\mOpp{v}{\textrm{new}} = I$.\footnote{Though we have assumed that all hidden variables share the
       same domain, $[k]$, our work generalizes to hidden domains of
-      arbitrary sizes, provided \assumptionref{full-rank-plus} holds.},
-  parents $P$ and children $C \union \{x_{v_1}, x_{v_2}, x_{v_3}\}$,
-  where $x_{v_1},x_{v_2},x_{v_3}$ are three copies of $x$
-  (\figureref{reduction}). Also, set $\mOpp{v}{p} = I$.
+      arbitrary sizes.
+      [TODO: add qualifier]
+      %provided \assumptionref{full-rank-plus} holds.
+      }
 
@@
 
-\begin{lemma}(Bottlenecks guarantee exclusive views)
+\begin{lemma}[Bottlenecks guarantee exclusive views]
   Provided \propertyref{bottleneck} holds, every hidden variable
@@
 \begin{proof}
-  By \lemmaref{reduction}, we can assume w.l.o.g. that every observed
+  By \lemmaref{reduction}, we can assume that every observed
   variable is a leaf in the graph, causing any v-structures in the graph
   to be active. This allows us to reason about independence simply by
-  checking whether or not two nodes are connected or not; note that this
+  checking whether or not two nodes are connected or not. Note that this
   proof applies for undirected graphs as well.
@@
   A view $x \in \sV_{h}$ is exclusive for $h$ iff $x \not\in \sV_{h'} ~ \forall h'
-  \in \sC^-$
+  \in \sC^-$.
 
@@
 
+[TODO: somewhere give examples that are not in our class, preferably the simplest one, cite Sontag]
+
 With this lemma in place, we present the full algorithm, \LearnMarginals~
@@
 
-      \FOR{$h_i \in H$} 
+      \FOR{each hidden variable $h_i \in H$} 
         \STATE Apply \TensorFactorize to learn conditional moments
-        $\mOpp{v}{i}$ for every $\sV_h$.
+        $\mOpp{v}{i}$ for every $\sV_{h_i}$.
 
@@
 
+% DONE: move here because not central to story
+\paragraph{Remark.} We note that \propertyref{bottleneck} can be relaxed if some cliques
+  share parameters.
+In the directed grid model, we can identify the conditional moments $O$ from just
+  $x^a_1, x^b_1$ and $x^a_2$.
+  Therefore, we do not require that $h_2, h_3$ or $h_4$
+  be bottlenecks, and can omit the observations $x^b_2, x^b_3$ and $x^b_4$.
+  [TODO: make this a footnote if running out of space]
+% PL: don't understand this comment
+%In this case, we only require that the hidden variables in distinct
+  %cliques share parameters.
+
 \paragraph{Consistency and Sample Complexity}
@@
 \LearnMarginals~ combines two consistent algorithms, \TensorFactorize and
-  \LearnClique, and is thus consistent itself. 
+\LearnClique, [TODO: too much space after LearnClique] and is thus consistent itself.
+[TODO: did they prove this; if so just say "they show", not "we can show"]
+Using results from
+  \citet{anandkumar12moments,anandkumar13tensor} we can show that
+  learning $\mOpp{v_1}{i}$ for the bottleneck $h_i$ with views $x_{v_1},
+  x_{v_2}, x_{v_3}$ has the following sample complexity,
+  [TODO: define hat O]
+  [TODO: max subscript is misaligned]
+\begin{align*}
+  \|\mOpphat{v_1}{i} - \mOpp{v_1}{i}\|^2_F &= \frac{1}{\sqrt{n}} O\left( \frac{k {\pi\oft{i}}_{\max}^2}{\sigma_k(M_{v_1,v_2})^5} \right). 
+\end{align*}
+
+[TODO: call MOM squared]
 We next describe the asymptotic variance for \LearnClique when using the
@@
   \label{lem:mom-variance}  
-  The asymptotic variance of $Z_\sC$, $\Sigma^\mom_\sC$ is,
+  The asymptotic variance of $Z_\sC$ is:
   \begin{align*}
-    \Sigma^{(\mom)} &= \mOppit{\sV}{\sC} \Sigma_\sV \mOppi{\sV}{\sC},
+    \Sigma^{\mom} &= \mOppAlli \Sigma_\sV \mOppAllit,
   \end{align*}
-  where $\pinv{\mOpp{\sV}{\sC}} \eqdef \pinv{\mOpp{v_1}{i_1}} \otimes
-    \cdots \otimes \pinv{\mOpp{v_m}{i_m}}$, a $k^m \times d^m$ matrix, and 
+  where $\pinv{\mOppAll} \eqdef \pinv{\mOpp{v_1}{i_1}} \otimes
+  \cdots \otimes \pinv{\mOpp{v_m}{i_m}}$, a $d^m \times k^m$ matrix, and 
 \begin{align*}
-  \Sigma_\sV &= \diag  M_\sV  - M_\sV M_\sV^T,
+  \Sigma_\sV &= \text{\rm diag}(M_\sV) - M_\sV M_\sV^\top,
 \end{align*}
-  is the variance of the observations $M_\sV$.
+  is the asymptotic variance of the observations $M_\sV$.
 \end{lemma}
@@
 
-Choosing to represent $Z_\sC$ and $M_\sV$ as vectors and
-$\mOppi{v_1}{i_1} \otimes \cdots \otimes \mOppi{v_m}{i_m}$ as the
-matrix, we can rewrite this representation as,
+%Choosing to represent $Z_\sC$ and $M_\sV$ as vectors and
+%$\mOppi{v_1}{i_1} \otimes \cdots \otimes \mOppi{v_m}{i_m}$ as the
+%matrix,
+We can rewrite this representation as:
 \begin{align*}
-  Z_{\sC} &= \pinv{\mOpp{\sV}{\sC}} \mO_\sV.
+  Z_{\sC} &= \pinv{\mOppAll} \mO_\sV.
 \end{align*}
 
-We know the asymptotic distribution of $M_\sV$ goes as,
+From classical asymptotics \cite{vaart98asymptotic},
+we have that the asymptotic variance of $M_\sV$ is:
 \begin{align*}
@@
 \begin{align*}
-  \sqrt{n}(\hat Z_\sC - Z_\sC) \convind \sN(0, \pinvt{\mOpp{\sV}{\sC}} \Sigma_\sV \pinv{\mOpp{\sV}{\sC}}).
+  \sqrt{n}(\hat Z_\sC - Z_\sC) \convind \sN(0, \pinvt{\mOppAll} \Sigma_\sV \pinv{\mOppAll}).
 \end{align*}
@@
 
-Note that extending these results to finite sample bounds can be done
-  via a straightforward application of perturbation bounds.
+% DONE: too brazen to say
+%Note that extending these results to finite sample bounds can be done
+  %via a straightforward application of perturbation bounds.

### 2014-01-31 doc/icml2014/discussion.tex
Percy's changes

@@
   variables to the data. 
-While other authors, e.g \citet{anandkumar12lda, anandkumar12linear,
+While other authors, e.g \citet{anandkumar12lda, anandkumar2013linear,
   halpern13noisyor} are able to learn parameters for models outside this

### 2014-01-31 doc/icml2014/intro.tex
Added some figures and made some changes to sample complexity stuff
Percy's changes

@@
 % 1. Latent variable models are good.
-Latent variable graphical models offer a succinct representation of
-a rich model family. 
+Latent-variable graphical models provide compact representations of data
+and have been employed across many fields [TODO: find a much better/broader set of citations]
+\cite{quattoni04crf,beaumont2004bayesian,haghighi06prototype}.
 % 2. Learning them is hard.
-Despite their success across many fields
-\cite{quattoni04crf,beaumont2004bayesian,haghighi06prototype,liang06discrimative,kirkpatrick10painless},
-learning these models remains a difficult problem due to the
-non-convexity of the likelihood. Local optimization (e.g.
-expectation-maximization) is the standard approach, but is susceptible
-to local optima.
+However, learning these models remains a difficult problem due to the
+non-convexity of the negative log-likelihood.  Local methods such as Expectation Maximization (EM)
+are the norm, but are susceptible to local optima.
 
 % 3. People have approached unsupervised learning with the MoM magic sauce, but the sauce is limited.
-Recently, unsupervised learning techniques based on the method of moments and
-spectral decomposition have offered a refreshing and promising perspective on
-this learning problem \citep{hsu09spectral,anandkumar11tree,anandkumar12moments,anandkumar12lda,hsu12identifiability,balle11transducer,balle12automata}.
-These methods exploit the linear algebraic properties of the model and
+Recently, unsupervised learning techniques based on the spectral method of moments
+have offered a refreshing and promising perspective on
+this learning problem
+\citep{hsu09spectral,anandkumar11tree,anandkumar12moments,anandkumar12lda,hsu12identifiability,balle11transducer,balle12automata}.
+[TODO: make sure we are citing the right people]
+These methods exploit the linear algebraic properties of the model to
 factorize the moments into parameters, providing strong theoretical guarantees.
 However, these methods apply to a limited set of models, and are thus
-not as universally applicable as EM.
+not as broadly applicable as EM.
 
@@
   \subimport{figures/}{approach.tikz}
-  \caption{Overview of approach}
+  \caption{[TODO: need three steps: tensor factorization, optimize composite likelihood, recover model parameters; Overview of approach}
 \end{figure}
 
+% DONE: need to say from the beginning that we have a hybrid approach, not just using method of moments;
+% there are really two contributions: one is the consistent estimate result for class of models,
+% and the other is the general idea of using partial constraints]
+
 % 4. State what we do: exploit moment constraints to make the problem easier.
-In this work, we exploit the spectral method to learn moment constraints
-  on the observed variables and show how these constraints lead to consistent parameter estimates 
-  for any graphical model with the property that there exist at least
-  three ``views'' or conditionally independent observed variables for
-  every hidden variable (\propertyref{bottleneck}).
+In this paper,
+we show that a much broader class of directed and undirected graphical models can be consistently estimated,
+specifically those in which each latent variable has three conditionally
+independent observed variables (``views'')---see \propertyref{bottleneck}.
+The key idea is to leverage the spectral method of moments,
+not to directly provide a consistent parameter estimate as in previous work,
+but as partial constraints on a likelihood-based objective.
+%For a broad class of directed and undirected graphical models
 
-% 5. We get moments from third-order tensors from bottlenecks and factorize them into marginals.
+%In this work, we exploit the spectral method to learn moment constraints
+  %on the observed variables and show how these constraints lead to consistent parameter estimates 
+  %for any graphical model with the property that there exist at least
+  %three ``views'' or conditionally independent observed variables for
+  %every hidden variable (\propertyref{bottleneck}).
+
+% 5. We get moments from third-order tensors from bottlenecks and factorize them into marginals,
+% then get parameters.
+% DONE: try to talk about the approach in terms of three steps
 Our approach is illustrated in \figureref{approach}. 
-First, we identify the views $x_1, x_2, x_3$ for each hidden variable $h$. 
-We then use the tensor factorization algorithm of
-  \citet{anandkumar12moments,anandkumar13tensor} to identify the
-  conditional moments.
-Given these factors, we show that there exists a set of observed
-  variables for every clique such that the likelihood is convex in the
-  parameters of that clique. 
-This guarantees that we can recover the true marginal distributions
-  using EM; finally, with some post-processing, we are able to
-  consistently recover the true parameters of the model.
-In \sectionref{directed}, we detail our algorithm and describe the
-  settings and technical assumptions in which this approach is guaranteed
-  to work.
-As a particular example, we are able to recover parameters for the
-  grid model depicted in \figureref{approach} under the condition that
-  $O$ and $T$ is full-rank.
+First, we identify the views for each hidden variable $h_i$ (for example,
+$x_1^a$, $x_1^b$ and $x_3^a$ are conditionally independent given $h_1$) and use
+the tensor factorization algorithm of
+\citet{anandkumar12moments,anandkumar13tensor} to identify the \emph{conditional
+moments} $\E[x_i^a \mid h_i]$ and $\E[x_i^b \mid h_i]$ for each $i$.
+Second, we optimize a \emph{composite marginal likelihood} to recover the marginals over
+each latent clique (e.g., $\mathbb P(h_2, h_3, h_4)$).
+Normally, such a marginal likelihood objective would be non-convex,
+but given the conditional moments, we obtain a convex objective,
+which can be globally optimized using EM.
+Finally, the latent marginals yield the model parameters
+via simple post-processing in the directed case and convex optimization in the undirected case.
+%Given these factors, we show that there exists a set of observed
+  %variables for every clique such that the likelihood is convex in the
+  %parameters of that clique. 
+%This guarantees that we can recover the true marginal distributions
+  %using EM.
+  %finally, with some post-processing, we are able to
+  %consistently recover the true parameters of the model.
+In \sectionref{directed}, we detail our algorithm and
+show that our algorithm yields consistent parameter estimates
+under non-degenerate rank assumptions on the conditional moments.
+%describe the
+  %settings and technical assumptions in which this approach is guaranteed
+  %to work.
+% PL: hard to talk about $O$ and $T$ at this stage.
+%As a particular example, we are able to recover parameters for the
+  %model depicted in \figureref{approach} if $O$ and $T$ are full-rank.
 
-Our approach also applies to undirected log-linear latent variable
-  models, where we use the recovered marginals to make the learning
-  problem convex (\sectionref{undirected}).
+%Our approach also applies to undirected log-linear latent variable
+  %models, where we use the recovered marginals to make the learning
+  %problem convex (\sectionref{undirected}).
 Finally, our approach gracefully extends to more general model families
   in which we are unable to recover the marginal distributions for all
-  cliques; in such scenarios, we use posterior regularization to constrain
-  EM.
+  hidden cliques; in such scenarios, we use posterior regularization to constrain
+  EM.  We show that empirically, using such constraints alleviates local
+  optima issues.

### 2014-01-31 doc/icml2014/related.tex
Percy's changes

@@
   that subtracts the correlation between the views.
-\citet{anandkumar12linear} study a special family of Bayesian networks
+\citet{anandkumar2013linear} study a special family of Bayesian networks
   in which the observed variables are linearly related to a sparse set of

### 2014-01-31 doc/icml2014/undirected.tex
Updated figures and some details and notation
Percy's changes

@@
 \begin{align*}
-  p_\theta(\vx, \vh) &= \exp( \sum_\sC \theta_\sC^T \phi_\sC(\vx,\vh) - A(\theta) ),
+  p_\theta(\vx, \vh) &= \exp( \sum_\sC \theta_\sC^\top \phi_\sC(\vx,\vh) - A(\theta) ),
 \end{align*}
 where $\phi_\sC(\vx,\vh)$ is a feature vector for the clique $\sC$ and
-  $A(\theta) = \sum_{\vx, \vh}  \exp( \sum_\sC \theta_\sC^T \phi_\sC(\vx,\vh) )$ is the
+  $A(\theta) = \sum_{\vx, \vh}  \exp( \sum_\sC \theta_\sC^\top \phi_\sC(\vx,\vh) )$ is the
   log-partition function.
@@
       &= \arg\max_{\theta} \sum_{\vx \in \sD} \log \sum_{\vh \in \sH} p_\theta(\vx,\vh) \nonumber \\
-      &= \arg\max_{\theta} \sum_{\vx \in \sD} \log \sum_{\vh \in \sH} \exp( \theta^T\phi(\vx,\vh) - A(\theta) ). \label{eqn:obj-ml}
+      &= \arg\max_{\theta} \sum_{\vx \in \sD} \log \sum_{\vh \in \sH} \exp( \theta^\top\phi(\vx,\vh) - A(\theta) ). \label{eqn:obj-ml}
 \end{align}
 Note that this problem is not concave in $\theta$.
+
+\note{Consider the supervised setting.
+As we get infinite amounts of data, we get this simple form.
+As MoM is a consistent estimator for marginals, we can optimize this
+  likelihood and it is consistent.
+}
+
 However, given the moments $\E[ \phi_\sC(\vx,\vh) ] = Z_\sC$, the
@@
   \hat \theta 
-  &= \arg\max_{\theta} \sum_{\sC \in \sG} \theta_\sC^T Z_\sC - A(\theta), \label{eqn:obj}
+  &= \arg\max_{\theta} \sum_{\sC \in \sG} \theta_\sC^\top Z_\sC - A(\theta), \label{eqn:obj}
 \end{align}

### 2014-01-31 doc/icml2014/directed.tex
Convered TODO to \todo

@@
 
-[TODO: need to say we're going to do this for the infinite data case,
-and then study bounds]
+\todo{need to say we're going to do this for the infinite data case,
+and then study bounds}
 
-[TODO: try to be consistent: either use latent or hidden]
+\todo{try to be consistent: either use latent or hidden}
 
@@
       arbitrary sizes.
-      [TODO: add qualifier]
+      \todo{add qualifier}
       %provided \assumptionref{full-rank-plus} holds.
@@
 
-[TODO: somewhere give examples that are not in our class, preferably the simplest one, cite Sontag]
+\todo{somewhere give examples that are not in our class, preferably the simplest one, cite Sontag}
 
@@
   be bottlenecks, and can omit the observations $x^b_2, x^b_3$ and $x^b_4$.
-  [TODO: make this a footnote if running out of space]
+  \todo{make this a footnote if running out of space}
 % PL: don't understand this comment
@@
 \LearnMarginals~ combines two consistent algorithms, \TensorFactorize and
-\LearnClique, [TODO: too much space after LearnClique] and is thus consistent itself.
-[TODO: did they prove this; if so just say "they show", not "we can show"]
+\LearnClique, \todo{too much space after LearnClique} and is thus consistent itself.
+\todo{did they prove this; if so just say "they show", not "we can show"}
 Using results from
@@
   x_{v_2}, x_{v_3}$ has the following sample complexity,
-  [TODO: define hat O]
-  [TODO: max subscript is misaligned]
+  \todo{define hat O}
+  \todo{max subscript is misaligned}
 \begin{align*}
@@
 
-[TODO: call MOM squared]
+\todo{call MOM squared}
 We next describe the asymptotic variance for \LearnClique when using the

### 2014-01-31 doc/icml2014/intro.tex
Convered TODO to \todo

@@
 Latent-variable graphical models provide compact representations of data
-and have been employed across many fields [TODO: find a much better/broader set of citations]
+and have been employed across many fields \todo{find a much better/broader set of citations}
 \cite{quattoni04crf,beaumont2004bayesian,haghighi06prototype}.
@@
 \citep{hsu09spectral,anandkumar11tree,anandkumar12moments,anandkumar12lda,hsu12identifiability,balle11transducer,balle12automata}.
-[TODO: make sure we are citing the right people]
+\todo{make sure we are citing the right people}
 These methods exploit the linear algebraic properties of the model to
@@
   \subimport{figures/}{approach.tikz}
-  \caption{[TODO: need three steps: tensor factorization, optimize composite likelihood, recover model parameters; Overview of approach}
+  \caption{\todo{need three steps: tensor factorization, optimize composite likelihood, recover model parameters;} Overview of approach}
 \end{figure}
@@
   hidden cliques; in such scenarios, we use posterior regularization to constrain
-  EM.  We show that empirically, using such constraints alleviates local
-  optima issues.
+  EM.  
+\todo{Is this possible?:  We show that empirically, using such constraints alleviates local
+optima issues.}

### 2014-01-31 doc/icml2014/directed.tex
Merged percy's chagnes and made some minor fixes

@@
 
-\LearnMarginals~ combines two consistent algorithms, \TensorFactorize and
-\LearnClique, \todo{too much space after LearnClique} and is thus consistent itself.
-\todo{did they prove this; if so just say "they show", not "we can show"}
-Using results from
-  \citet{anandkumar12moments,anandkumar13tensor} we can show that
-  learning $\mOpp{v_1}{i}$ for the bottleneck $h_i$ with views $x_{v_1},
-  x_{v_2}, x_{v_3}$ has the following sample complexity,
-  \todo{define hat O}
-  \todo{max subscript is misaligned}
-\begin{align*}
-  \|\mOpphat{v_1}{i} - \mOpp{v_1}{i}\|^2_F &= \frac{1}{\sqrt{n}} O\left( \frac{k {\pi\oft{i}}_{\max}^2}{\sigma_k(M_{v_1,v_2})^5} \right). 
-\end{align*}
+\LearnMarginals combines two consistent algorithms, \TensorFactorize and
+\LearnCliqueNs, and is thus consistent itself.
+%\todo{did they prove this; if so just say "they show", not "we can show"}
+%Using results from
+%  \citet{anandkumar12moments,anandkumar13tensor} we can show that
+%  learning $\mOpp{v_1}{i}$ for the bottleneck $h_i$ with views $x_{v_1},
+%  x_{v_2}, x_{v_3}$ has the following sample complexity,
+%  \todo{define hat O}
+%  \todo{max subscript is misaligned}
+%\begin{align*}
+%  \|\mOpphat{v_1}{i} - \mOpp{v_1}{i}\|^2_F &= \frac{1}{\sqrt{n}} O\left( \frac{k {\pi\oft{i}}_{\max}^2}{\sigma_k(M_{v_1,v_2})^5} \right). 
+%\end{align*}
 

### 2014-01-31 doc/icml2014/directed.tex
more changes, especially to asymptotic variance

@@
 \label{sec:directed}
+
+\todo{we are using clique to refer to only hidden cliques; be explicit about using "hidden"}
 
@@
   conditionally independent given $h_1$, allowing us to use
-  \TensorFactorize, as reviewed in \sectionref{setup}, to recover
+  $\TensorFactorize$, as reviewed in \sectionref{setup}, to recover
   $O$.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{General algorithm}
@@
 Note that in each of the above cases,
-each hidden variable $h_i$ of a clique has its own view which is conditionally independent
+each hidden variable $h_i$ of the clique has its own view which is conditionally independent
 given the other variables in the clique.
@@
 
-\paragraph{Learning cliques}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\paragraph{Estimating hidden clique marginals}
 
@@
 \begin{algorithm}
-  \caption{\LearnClique}
+  \caption{\LearnClique~(pseudoinverse)}
   \label{algo:learnclique}
@@
     \ENSURE Marginal distribution of the clique $Z_\sC$.
-      \STATE Identify unique views $\Sx{\sV} = \{x_{v_1}, \cdots, x_{v_m}\}$.
-      \STATE $\mH_\sC \gets \mO_{\Sx{\sV}}( \pinv{\mOpp{v_1}{i_i}}, \cdots, \pinv{\mOpp{v_m}{i_m}} )$.
+      \STATE Identify exclusive views $\Sx{\sV} = \{x_{v_1}, \cdots, x_{v_m}\}$.
+      \STATE Return $\mH_\sC \gets \mO_{\Sx{\sV}}( \pinv{\mOpp{v_1}{i_i}}, \cdots, \pinv{\mOpp{v_m}{i_m}} )$.
   \end{algorithmic}
 \end{algorithm}
+Of course, for $Z_{\sC}$ to be identifiable, we need various rank conditions,
+which will be established in \todo{section}.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \paragraph{Structural properties}
@@
 %  holds for every clique in the graph.
-
-% PL: this is not obviously correct
-%To address the first, we will need that the following assumption on the
-%  parameters of the model:
-%\begin{assumption} 
-%  \label{asm:full-rank-plus}
-%  For every clique $\sC = \{h_{i_1}, \cdots, h_{i_m}\} \in \sG$, let
-%    $Z_\sC$ be the marginal distribution for that clique. 
-%  Then every mode-$i$ unfolding of $Z_\sC$ has full column rank.\verify
-%\end{assumption}
-%Intuitively, the assumption guarantees that the marginal probability of
-%  any hidden state is non-zero.
 
@@
 
-\todo{somewhere give examples that are not in our class, preferably the simplest one, cite Sontag}
+\todo{somewhere give examples that are not in our class, preferably the simplest one - noisy-or, cite Sontag}
 
@@
       \FOR{each hidden variable $h_i \in H$} 
-        \STATE Apply \TensorFactorize to learn conditional moments
+        \STATE Apply $\TensorFactorize$ to learn conditional moments
         $\mOpp{v}{i}$ for every $\sV_{h_i}$.
@@
 \FOR{every clique $\sC = \{h_{i_1}, \cdots, h_{i_m}\} \in \sG$} 
-  \STATE Apply \LearnClique to learn the marginals $\mH_\sC$.
+\STATE Apply $\LearnClique$ to learn the marginals $\mH_\sC$.
 \ENDFOR
@@
 
-\paragraph{Consistency and Sample Complexity}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Sample complexity}
 
-\LearnMarginals combines two consistent algorithms, \TensorFactorize and
-\LearnCliqueNs, and is thus consistent itself.
+% PL: "combine" is too vague to be useful
+%\LearnMarginals combines two consistent algorithms, \TensorFactorize and
+%\LearnCliqueNs, and is thus consistent itself.
+
+In this section, we provide formal conditions under which $\LearnMarginals$
+will produce consistent estimates.
+We let
+$\mOpphat{v}{i}$,
+$\hat Z_\sC$,
+and $\hat M_\sV$,
+denote estimators for
+$\mOpp{v}{i}$,
+$Z_\sC$,
+and $M_\sV$,
+respectively.
+
+To apply $\TensorFactorize$, we need that each $\mOpp{v}{i}$ has full column rank and that $\pi^{(i)} \succ 0$ \todo{define $\pi$}.
+But $\mOpp{v}{i}$ is a product of
+tensors on the path from $h_i$ to $x_v$, marginalizing out the rest of the graph.
+To make this property more interpretable, let us assume the following property:\verify
+\begin{assumption} 
+\label{asm:full-rank-plus}
+For every clique $\sC \in \sG$ (including ones involving observed variables),
+every order-1 slice of the marginals $Z_\sC$ has full column rank $k$.
+\end{assumption}
+According to our directed grid example \todo{ref}, with $\sC = \{ h_1, h_2 \}$, then this condition implies
+$\pi = \bP(h_1) \succ 0$ and 
+$T = \bP(h_2 \mid h_1) \succ 0$;
+with $\sC = \{ h_1, x_1 \}$, we get
+$O = \bP(x_1 \mid h_1) \succ 0$.
+
 %\todo{did they prove this; if so just say "they show", not "we can show"}
@@
 
-\todo{call MOM squared}
-We next describe the asymptotic variance for \LearnClique when using the
-  method of moments.
-\begin{lemma}(Asymptotic variance of $Z_\sC$)
+From \cite{anandkumar13tensor} given \assumptionref{full-rank-plus},
+we have that $\mOpphat{v}{i}$ converges to $\mOpp{v}{i}$ at a rate of $n^{-\frac12}$ with a constant that
+depends polynomially on the $k$-th singular value of $\mOpp{v}{i}$.
+Of course, the singular value of $\mOpp{v}{i}$ become exponentially worse as $x_v$ and $h_i$ become farther.
+So while we can obtain consistent estimates, the sample complexity
+is exponential in the mimimum distance from $h_i$ to any three views, as to be expected.
+%(\todo{say something
+%quantitative?  annoying part is that we would have to bring in the bounds, so leave it out probably}).
+
+% DONE: need to setup the asymptotic - it's not fresh in everyone's mind.
+% Easy case
+Next, we need to show that the hidden marginals $\hat Z_\sC$ converge to $Z_\sC$.
+First, $\hat M_\sV$ is just an empirical average of multinomials over the data points.
+Abusing notation slightly, we let $M_\sV$ also denote its $d^{|\sV|}$-dimensional vectorized version.
+We also represent $\mH_\sC$ as
+  a vector in $\Re^{k^m}$, and represent $\mOppAll \eqdef
+  \mOpp{v_1}{i_1} [x_{v_1}] \otimes \cdots \otimes
+  \mOpp{v_m}{i_m} [x_{v_m}]$ as a matrix in $\Re^{d^m \times
+  k^m}$.
+By the central limit theorem, we have:
+$\sqrt{n} (\hat M_\sV - M_\sV) \convind \sN(0, \Sigma_\sV)$,
+where $\Sigma_\sV$ is the \emph{asymptotic variance} of $\hat M_\sV$ \todo{doesn't this need to be multiplied by something?}:
+\begin{align*}
+\Sigma_\sV \eqdef \dM_\sV - M_\sV M_\sV^\top, \quad \dM_\sV \eqdef \text{diag}(M_\sV).
+\end{align*}
+
+% PL: this is not obviously correct
+%To address the first, we will need that the following assumption on the
+%  parameters of the model:
+%\begin{assumption} 
+%  \label{asm:full-rank-plus}
+%  For every clique $\sC = \{h_{i_1}, \cdots, h_{i_m}\} \in \sG$, let
+%    $Z_\sC$ be the marginal distribution for that clique. 
+%  Then every mode-$i$ unfolding of $Z_\sC$ has full column rank.\verify
+%\end{assumption}
+%Intuitively, the assumption guarantees that the marginal probability of
+%  any hidden state is non-zero.
+
+We can now use the delta-method to convert the above result into 
+the asymptotic variance for the pseudoinverse version of $\LearnClique$:
+\begin{lemma}[Asymptotic variance of pseudoinverse estimator for $Z_\sC$]
   \label{lem:mom-variance}  
-  The asymptotic variance of $Z_\sC$ is:
+  Assume $\hat M_{\sV}$ has asymptotic variance $\Sigma_\sV$ defined above.
+  Then the asymptotic variance of $Z_\sC$ is:
   \begin{align*}
@@
   where $\pinv{\mOppAll} \eqdef \pinv{\mOpp{v_1}{i_1}} \otimes
-  \cdots \otimes \pinv{\mOpp{v_m}{i_m}}$, a $d^m \times k^m$ matrix, and 
-\begin{align*}
-  \Sigma_\sV &= \text{\rm diag}(M_\sV) - M_\sV M_\sV^\top,
-\end{align*}
-  is the asymptotic variance of the observations $M_\sV$.
+  \cdots \otimes \pinv{\mOpp{v_m}{i_m}}$, a $d^m \times k^m$ matrix.
 \end{lemma}
 \begin{proof}
-We have for each clique,
-\begin{align*}
-  Z_{\sC} &= \mO_\sV(\mOppi{v_1}{i_1},\cdots,\mOppi{v_m}{i_m}) \\
-\end{align*}
-
+For clique $\sC$, recall we have
+  $Z_{\sC} = \mO_\sV(\mOppi{v_1}{i_1},\cdots,\mOppi{v_m}{i_m})$.
 %Choosing to represent $Z_\sC$ and $M_\sV$ as vectors and
@@
 %matrix,
-We can rewrite this representation as:
-\begin{align*}
-  Z_{\sC} &= \pinv{\mOppAll} \mO_\sV.
-\end{align*}
-
-From classical asymptotics \cite{vaart98asymptotic},
-we have that the asymptotic variance of $M_\sV$ is:
-\begin{align*}
-  \sqrt{n}(\hat M_\sV - M_\sV) \convind \sN(0, \Sigma_\sV),
-\end{align*}
-where $\Sigma_\sV$ is the variance of the observations. 
-Finally, we have that $Z_\sC$ goes as
-\begin{align*}
-  \sqrt{n}(\hat Z_\sC - Z_\sC) \convind \sN(0, \pinvt{\mOppAll} \Sigma_\sV \pinv{\mOppAll}).
-\end{align*}
+We can rewrite this representation more compactly as: $Z_{\sC} = \pinv{\mOppAll} \mO_\sV$.
+By the delta-method \cite{vaart98asymptotic},
+% DONE: need to shorten
+%we have that the asymptotic variance of $M_\sV$ is:
+%\begin{align*}
+%  \sqrt{n}(\hat M_\sV - M_\sV) \convind \sN(0, \Sigma_\sV),
+%\end{align*}
+%where $\Sigma_\sV$ is the variance of the observations. 
+we immediately get:
+$\sqrt{n}(\hat Z_\sC - Z_\sC) \convind \sN(0, \mOppAlli \Sigma_\sV \mOppAllit)$.
 \end{proof}

### 2014-01-31 doc/icml2014/discussion.tex
done with section 4

@@
 \label{sec:discussion}
+
+Parameter estimation for latent variable models using the method of
+  moments has received much attention in last few years.
+Our work uses the \TensorFactorize algorithm presented by
+  \citet{anandkumar13tensor}, though \citet{anandkumar12moments} first
+  studied the consistent parameter estimation algorithm for the
+  three-view mixture model.
+%\todo{What about Mossel and Roch?}
+Several authors have extended this work to settings that do not meet the
+  criterion \propertyref{bottleneck}, but rely on knowing more about
+  the underlying correlation structure of the problem.
+\citet{anandkumar12lda} are able to recover parameters for correlated
+  discrete mixtures, e.g. latent Dirichlet allocation; their approach
+  follows that of the three-view mixture model, with a pre-processing step
+  that subtracts the correlation between the views.
+\citet{anandkumar2013linear} study a special family of Bayesian networks
+  in which the observed variables are linearly related to a sparse set of
+  hidden variables. Once again, the key step is to identify the
+  correlations between observed variables and exclude them from the views.
+\citet{halpern13noisyor} propose an algorithm to recover factors for
+  a more densely connected bipartite noisy-or network; they also use
+  bottlenecks to learn parameters, but exploit properties of the model
+  to negate correlations from learned edges, allowing for more
+  bottlenecks to be discovered. 
+In comparison, our work applies to a general family of models, including
+  latent tree models, directed and undirected grids, with no assumptions
+  on the correlation structure aside from independences encoded in the
+  graph.
+
+Another prominent line of work in the method of moments community has
+  focussed on recovering parameters for a transformation that admits an {\em observable
+  representation}, i.e. one whose marginals can be directly computed from the data.
+\citet{song2011spectral} propose an algorithm to learn an observable
+  operator representations for latent tree graphical models, like the
+  one in \figureref{examples-tree}, assuming \propertyref{bottleneck}. 
+Their approach is similar to ours, but restricted to trees with observed
+  variables at the leaves.
+\citet{parikh12spectral} extend this approach to general graphical
+  models in which \propertyref{bottleneck} holds, but consider the latent
+  junction tree representation. 
+Consequently, the observable representations have size atleast that of
+  a node in the junction tree, which will be of the order of the treewidth
+  of the graph. 
+Our algorithm only constructs moments of the order of the cliques in the
+  graph, which can be exponentially smaller. 
+For example, an $n\times n$ grid model has a treewidth of $n$, but each
+  clique is of size at most $2$ in the undirected case and $3$ in the
+  directed case.
+In some sense, an observable operator re-formulation of our algorithm
+  subsumes this work.
 

### 2014-01-31 doc/icml2014/examples.tex
add examples

\subsection{More examples}

We will now instantiate our algorithm for a couple more examples, illustrated in \figureref{examples}.

\begin{figure}
  \centering
  \subfigure[Hidden Markov Model] {
    \includegraphics[width=0.45\columnwidth]{figures/hmm.pdf}
    \label{fig:examples-hmm}
  }
%  \subfigure[Directed grid model] {
%    \label{fig:examples-grid}
%    \includegraphics{figures/grid.pdf}
%  }
  \subfigure[Tree model] {
    \includegraphics[width=0.45\columnwidth]{figures/tree.pdf}
    \label{fig:examples-tree}
  }
  \caption{Additional examples of models learnable using \LearnMarginals}
  \label{fig:examples}
\end{figure}

\paragraph{Hidden Markov Model}

In this example (\figureref{examples-hmm}), we assume that
  $\Pr(x_i|h_i) = O  ~\forall i$  and that $\Pr(h_{i+1} | h_i)
  = T ~\forall i$ (i.e. we have parameter sharing).
Note that while the first (and last) hidden variables $h_1, h_T$ in the
  sequence are not bottlenecks, they still have exclusive views ($x_1$ and
  $x_T$ respectively) whose parameters we know because they share
  parameters, $O$.
In step 1 of our algorithm, we use the bottleneck $h_2$ with views $x_1,
  x_2, x_3$ and solve $O$.
In step 2 of our algorithm, we can recover $\pi$ by solving for the
  unary clique $\{h_1\}$ and recover $T$ from the clique $\{h_{1},
  h_{2}\}$.

\paragraph{Latent Tree Structure}

In the latent tree structure (\figureref{examples-tree}), let the
  parameters be $\Pr(h_i) = \pi$, $\Pr(h_i | h_1) = T ~i \in \{2,3,4\}$
  and $\Pr(x^a_i | h_i) = \Pr(x^b_i | h_i) = O ~i \in \{2,3,4\}$.
Note that while $h_1$ is not directly connected to an observed variable,
  it is still a bottleneck, with views $x^a_2, x^a_3, x^a_4$.

In step 1, we recover the parameters $O$ from the bottleneck $h_2$ with
  views $\{x^a_2, x^b_2, x^a_3\}$. We also recover the conditional moments
  $\mOpp{2}{1}$, $\mOpp{3}{1}$, $\mOpp{4}{1}$ for $h_1$. 
In step 2, we can recover $\pi$ from the clique $\{h_1\}$, using any
  one of views (they are all exclusive). 
To recover $T$ from the clique $\{h_1, h_2\}$, we use the views $x^a_2$
  (exclusive to $h_2$) and $x^a_3$ (exclusive to $h_3$). Note that while
  $x^a_3$ is also a view for $h_2$, $x^a_3$ is independent of $h_2$ given
  $h_1$.

\paragraph{Aggregating observations}
We describe two practical considerations to use samples more efficiently.
Firstly, if we have multiple exclusive views for a hidden variable,
  intuitively, it is better to aggregate over them. 
For example, consider a hidden variable $h_1$ with multiple exclusive
  views $x_1$ and $x_2$.
With the method of moments perspective, to learn the marginal
  distribution over $h_1$, one must solve the following reconstruction
  problem, 
\begin{align*}
  \hat Z_{h_1} &= \arg\min_{Z_{h_1}} \half \|Z_{h_1}(\mOpp{1}{1}) - M_1 \|^2 + \half \|Z_{h_1}(\mOpp{2}{1}) - M_2 \|^2,
\end{align*}
which does not have a closed form solution. 
In contrast, this naturally fits into the convex optimization framework, where $\hat Z_{h_1}$ will now be,
\begin{align*}
  \hat Z_{h_1} &= \arg\min_{Z_{h_1}} \sum_{\vx \in \sD} \log Z_{h_1}( \mOpp{1}{1}[x_1] \odot \mOpp{1}{1}[x_1] ),
\end{align*}
where $\cdot$ denotes element-wise multiplication.
An important note to make is that \TensorFactorize only returns
  a solution up to permutation; if $\mOpp{1}{1}$ and
  $\mOpp{1}{2}$ do not belong to the same bottleneck (e.g. $x^a_2$ and
  $x^b_2$ in \figureref{examples-tree}), then some
  care must be taken to ensure they have the same labelling.

Secondly, we can exploit parameter sharing by aggregating over
  disjoint sets of observed variables. 
For example, in \figureref{examples-hmm}, we can aggregate the statistics for
  bottlenecks $h_i$ with views $\{x_{i-1}, x_{i}, x_{i+1}\}$ before
  running \TensorFactorize; this will give us a consistent estimate for
  $O$ (as well as $T$).

### 2014-01-31 doc/icml2014/undirected.tex
done with undirected

--- 2014-01-30 doc/icml2014/undirected.tex
@@
-\section{Consistent Parameter Estimation for Undirected Latent Variable Log-linear Models}
+\section{Undirected log-linear models}
 \label{sec:undirected}
 
-In this section, we will show that we can extend the results to
-  undirected graphical models parameterized as log-linear latent variable
-  models.
-We exploit a connection between the moments of the clique potentials and
-  the log-partition function to show that learning parameters is convex
-  given the clique marginals $Z_\sC$. 
-The end result is a consistent estimator for log-linear
-  latent-variable models which satisfy \propertyref{exclusive-views}.
-Finally, we generalize the result to classes when only a subset of the
-  clique marginals can be learned.
+% DONE: need more of a transition from previous sections
+Recall that we used tensor factorization and composite likelihood
+to estimate the latent marginals $Z_\sC = \BP(\vh_\sC)$ and $\mOpp{v}{i} = \BP(x_v \mid h_i)$.
+For directed models, obtaining the local conditional tables from $Z_\sC$ requires
+renormalizing $Z_\sC$.
+For undirected log-linear models, the parameters cannot simply be obtained locally;
+we will have to perform global optimization.
+Fortunately, as we will show, this optimization problem is convex,
+by relating the recovered conditional moments with the sufficient statistics of
+the log-linear model.
 
-As before, let $\sG$ be a graphical model with variables $V$, with $X
-  \subseteq V$ being observed and $H = V \setminus X$ hidden.
-We assume that the undirected graphical model is a log-linear latent
-  variable model with parameters $\theta$,
-\begin{align*}
-  p_\theta(\vx, \vh) &= \exp( \theta^T\phi(x,z) - A(\theta) ),
-\end{align*}
-where $\phi(x,z)$ is a feature vector with components $\phi_\sC(x,z)$
-  for every clique and $A(\theta) = \sum_{\vx, \vh}  \exp(
-  \theta^T\phi(x,z) )$ is the log-partition function.
-While $\phi_\sC(x,z)$ can be an arbitrary featurization, we assume
-  without loss of generality that $\phi_\sC(x,z)$ are the indicator
-  features $\BI[x,z]$\footnote{Note that $\BI[x,z]$ are sufficient
-  statistics for the model, and hence we do not lose any statistical
-  efficiency in using them}.
+% Define model
+As before, let $\sG$ be a graphical model with observed variables $X$ and hidden variables $H$
+(see \figureref{examples-mrf} for an example).
+For each clique $\sC \in \sG$, we have a feature vector $\phi_\sC(\vx_\sC, \vh_\sC)$,
+where $\vx_\sC$ and $\vh_\sC$ are the observed and hidden variables in $\sC$, respectively.
+Let $\phi(\vx, \vh) = \sum_{\sC \in \sG} \theta^\top \phi_\sC(\vx_\sC,\vh_\sC)$ be the global feature vector.
+We consider log-linear models of the following form:
+\begin{align}
+p_\theta(\vx, \vh) &= \exp\left( \theta^\top \phi(\vx,\vh) - A(\theta) \right),
+\end{align}
+where $A(\theta) = \sum_{\vx, \vh}  \exp( \sum_\sC \theta^\top \phi(\vx,\vh) )$ is the log-partition function.
 
-As before, our objective is to find a set of parameters $\hat \theta$
-  that maximize the likelihood of data $\sD$.
+% FIGURE: example
+\begin{figure}
+  \label{fig:examples-mrf}
+  \centering
+  \subimport{figures/}{mrf.tikz}
+  \caption{Example: undirected grid model where each hidden variable has two
+  conditionally independent observations.
+  We derive a consistent estimate.}
+\end{figure}
+
+%As an example, consider an undirected grid model in \figureref{examples-mrf},
+%where each hidden variable has two conditionally independent observations.
+%In practice, if two features are known to be independent, then they can
+  %be separated into independent observed variables.
+%Let the parameters of the model be $O \eqdef \Phi(x^a_{i,j}, h_{i,j})
+  %= \Phi(x^b_{i,j}, h_{i,j})$ and $T \eqdef \Phi(h_{i,j}, h_{i \pm 1,j \pm 1})$. 
+%Note that as this is an undirected model, the clique marginals do not
+  %correspond to the parameters of the model.
+%To recover the parameters of this model, note that every hidden
+%  variable, $h_{ij} \in \sH$ is a bottleneck, with views
+%  $x^{a}_{ij},x^{b}_{ij}$ and any other observed variable in the graph. 
+%Thus, \LearnMarginals will be able to learn the marginals of each clique
+%  and \LearnLogLinear will give us a consistent estimate of the
+%  parameters.
+
+Of course, the log-likelihood is non-concave:
 \begin{align}
-  \hat \theta 
-      &= \arg\max_{\theta} \sum_{\vx \in \sD} \log p_\theta(\vx) \nonumber \\
-      &= \arg\max_{\theta} \sum_{\vx \in \sD} \log \sum_{\vh} p_\theta(\vx,\vh) \nonumber \\
-      &= \arg\max_{\theta} \sum_{\vx \in \sD} \log \sum_{\vh} \exp( \theta^T\phi(\vx,\vh) - A(\theta) ). \label{eqn:obj-ml}
+L_\text{unsup}(\theta) \eqdef \E_{\vx \sim \sD}[\log \sum_{\vh \in \sH} p_\theta(\vx,\vh)].
 \end{align}
-Note that this problem is not concave in $\theta$.
+% PL: this doesn't make sense
+%While $\phi_\sC(x,h)$ can be an arbitrary featurization, we assume
+%  without loss of generality that $\phi_\sC(x,h)$ are the indicator
+%  features $\BI[x,h]$\footnote{Note that $\BI[x,h]$ are sufficient
+%  statistics for the model, and hence we do not lose any statistical
+%  efficiency in using them.}.
 
-An alternate objective is to optimize the following lower-bound on the log-likelihood, 
+%In this section, we will show that we can extend the results to
+  %undirected graphical models parameterized as log-linear latent variable
+  %models.
+On the other hand, 
+if we were able to observe $\vh$, the \emph{supervised} likelihood would be concave:
 \begin{align}
-  \sL_p(q,\theta) &= \sum_{\vx \in \sD} \sum_{\vh} q(\vh \given \vx) \log\frac{p_\theta(\vx,\vh)}{q(\vh \given \vx)} \\
-    &= \KL( q(\vh \given \vx) || p_{\theta}(\vh \given \vx)) \label{eqn:obj-kl}
+\label{eqn:logLinearSupervised}
+L_\text{sup}(\theta) &\eqdef \E_{(\vx,\vh) \sim \sD_\text{sup}}[\log p_\theta(\vx,\vh)] \\
+                     &= \theta^\top \left(\sum_{\sC \in \sG} \E[\phi(\vx_\sC,\vh_\sC)]\right) - A(\theta).
 \end{align}
+While we don't have supervised data,
+note that the method of moments gives us latent marginals $Z_\sC = \BP(\vh_\sC)$.
+Furthermore, given $\mOpp{v}{i}$ and $Z_\sC$ of any $\sC \ni v$, we can recover
+all marginals $Z_\sC = \BP(\vx_\sC, \vh_\sC)$ (with a slight augmentation of notation).
+Finally, the key step is that we can compute all the expected feature vectors from the marginals:
+\begin{align}
+\label{eqn:logLinearFeatures}
+\mu_\sC \eqdef \E[\phi(\vx_\sC,\vh_\sC)] = \sum_{\vx_\sC,\vh_\sC} Z_\sC(\vx_\sC,\vh_\sC) \phi(\vx_\sC,\vh_\sC).
+\end{align}
+Therefore, we can optimize the supervised likelihood objective without actually
+having supervised data.
+Of course, in the finite data regime, the method of moments yields estimates
+$\hat \mu^\text{mom}_\sC$ which approximate the true $\mu_\sC$.
+In supervised learning, we have different estimate $\hat\mu^\text{sup}_\sC$ of $\mu_\sC$ based on an empirical average
+over the data points.
+In the limit of infinite data, both estimators converge to $\mu_\sC$.
 
-Importantly, given $q$, the above objective is convex in $\theta$ and vice versa.
-It can be shown\cite{} that the local-optima of \equationref{obj-kl}
-are also local-optima of \equationref{obj-ml}. Usually, one
-iteratively optimizes $q$ and $p_\theta$ from some initial parameters
-$\theta_0$ to arrive at a local optimum.
+%We exploit the connection between the moments of the clique potentials and
+  %the log-partition function to show that learning parameters is convex
+  %given the clique marginals $Z_\sC$. 
+%The end result is a consistent estimator for log-linear
+  %latent-variable models which satisfy \propertyref{exclusive-views}.
+%Finally, we generalize the result to classes when only a subset of the
+  %clique marginals can be learned.
 
-This reduces to,
-\begin{align}
-  \sL_p(\theta) &= \theta^T Z - A(\theta),
-\end{align}
-which is a convex problem.
+%As before, our objective is to find a set of parameters $\hat \theta$
+%  that maximize the likelihood of data $\sD$.
+%\begin{align}
+%  \hat \theta 
+%      &= \arg\max_{\theta} \sum_{\vx \in \sD} \log p_\theta(\vx) \nonumber \\
+%      &= \arg\max_{\theta} \sum_{\vx \in \sD} \log \sum_{\vh \in \sH} p_\theta(\vx,\vh) \nonumber \\
+%      &= \arg\max_{\theta} \sum_{\vx \in \sD} \log \sum_{\vh \in \sH} \exp( \theta^\top\phi(\vx,\vh) - A(\theta) ). \label{eqn:obj-ml}
+%\end{align}
+%Note that this problem is not concave in $\theta$. 
+%
+%\note{Consider the supervised setting.
+%As we get infinite amounts of data, we get this simple form.
+%As MoM is a consistent estimator for marginals, we can optimize this
+%  likelihood and it is consistent.
+%}
 
-However, from the previous section, for any model satisfying
-  \property{exclusive-views}, we have a consistent estimator for the
-  marginals $q \equiv Z$. 
-Therefore, we can optimize for $\theta$ given this estimate of $q$,
-  $\hat q$.
+%However, given the moments $\E[ \phi_\sC(\vx,\vh) ] = Z_\sC$, the
+%  problem simplifies to,
+%\begin{align}
+%  \hat \theta 
+%  &= \arg\max_{\theta} \sum_{\sC \in \sG} \theta_\sC^\top Z_\sC - A(\theta), \label{eqn:obj}
+%\end{align}
+%which is concave in $\theta$.
+\algorithmref{undirected} summarizes our approach.
+In \sectionref{directed}, we showed that $\LearnMarginals$ is a consistent
+estimator for $Z_\sC$ for any graphical model satisfying the exclusive views property (\propertyref{exclusive-views}). 
+Consequently, the result is that the output of \algorithmref{undirected},
+$\hat\theta$, is a consistent estimator.\footnote{Of course, the log-linear model must be identifiable
+(violated, for example, if two components of $\phi(\vx_\sC, \vh_\sC)$ were identical).
+In general, we can identify the parameters up to kernel of $\nabla^2 A(\theta)$.}
 
@@
   \begin{algorithmic}
-    \REQUIRE A graphical model $\sG$ satisfying \textbf{(P1)}, \textbf{(P2)}, data $\sD$
-    \ENSURE A consistent estimate $\hat \theta$ of the parameters of the model.
-      \STATE Apply \LearnMarginals to $\sG$ to learn clique marginals $Z_\sC$.
-      \STATE Optimize \equationref{obj-kl} with respect to $\theta$.
+    \REQUIRE A graphical model $\sG$ satisfying \propertyref{bottleneck}, data $\sD$
+    \ENSURE A consistent estimate of parameters $\hat \theta$
+      \STATE Apply $\LearnMarginals$ to $\sG$ to learn clique marginals $Z_\sC$.
+      \STATE Compute expected features (\equationref{logLinearFeatures}) and optimize \equationref{logLinearSupervised}.
   \end{algorithmic}
@@
 
-This approach is outlined in \algorithmref{undirected}.
+\paragraph{Pseudolikelihood.}
+Unlike for directed models, in the undirected case,
+graphs with high tree-width present a challenge
+for even fully-supervised maximum likelihood.
+This motivates looking at pseudolikelihood objective of \citep{besag75pseudo},
+which optimizes the likelihood of each variable $a$ given its neighbors $\sN(a)$:
+$\sum_{a \in X \cup H} \log p_\theta(a \mid x_{\sN(a)})$.
+In the fully-supervised setting, pseudolikelihood provides consistent estimates,
+albeit at the cost of statistical efficiency.
+However, if we write the objective, we will see that the log-partition function $A(\theta)$
+is replaced with a sum over $A_a(\theta; x_{\sN(a)}) = \sum_a \exp(\theta^\top\phi(a))$,
+which cannot be computed from $Z_\sC$, since the neighbors $\sN(a)$ in general do not form a clique.
+However, we adapt our earlier algorithm to estimate the marginals over all neighborhoods $Z_{\sN(a)}$.
+Of course, this method is only viable for graphs of low degree ($|\sN(a)|$ is small).
+For example, in the grid model, the degree is only 4.
+Note that neither sample nor computational complexity depends on the treewidth.
 
-\todo{Do we have to worry about the fact that this is constrained?}
-\todo{What about model un-identifiability?}
+\paragraph{Learning with partial moment constraints.}
 
-\paragraph{Learning with partial moment constraints}
+In general, we might only have exclusive views (\propertyref{exclusive-views}) for a subset of the cliques.
+In this case, we can estimate the expected features $\mu_\sC$ for those cliques
+and use posterior regularization \citep{graca08em} or generalized expectation criteria \citep{mann08ge}
+to encourage $\E_{p_\theta}[\phi(\vx_\sC,\vh_\sC)]$ to match $\mu_\sC$.
+The resulting objective functions are still non-convex, but in practice we observed
+that the extra information $\mu_\sC$ helps alleviate local optima.
+%We can then integrate these constraints using posterior regularization \citep{graca08em}:
+%\begin{align}
+%L_\text{pr}(\theta) \eqdef L_\text{unsup}(\theta) + C\min_{q \in \sQ} \text{KL}(q(\vx,\vh) \| p_\theta(\vx,\vh)).
+%\end{align}
+%The main idea is to regularize the parameters towards values that produce the 
+%While the optimization problem \eqref{eqn:obj-ml} is non-convex, we
+  %expect that the measurements aid recover of parameters, and indeed see
+  %this is true for experiments on simulated data.
+%\todo{Include figure}
 
-In more general models, \propertyref{exclusive-views} may not hold for
-  every clique in the graph, however we can still estimate the marginals
-  $Z_\sC$ for cliques in which it does.
-In this section, we will describe a natural method to use these
-  constraints to simplify the parameter search problem.
+%Posterior regularization optimizes the following objective,
+%\begin{align*}
+%  \hat \theta &= \arg\min_{\theta} \arg\min_{q \in \sQ} \KL( q(\vx,\vh) \| p_{\theta}(\vx, \vh) ),
+%\end{align*}
+%where $q$ is a marginal distribution that belongs in constrained set
+%$\sQ$. In our case, $\sQ$ is constrained to agree with the partial
+%moments we recovered using \LearnMarginals, i.e. 
+%\begin{align*}
+%  \sQ &= \{ q(\vx,\vh) \given \forall \sC \in \sP ~ \E_{q(\vx,\vh)}[\phi_\sC(\vx, \vh)] = Z_\sC \}.
+%\end{align*}
 
-Constrain the KL to be in $Q$.
+%In practice, we would like to regularize for error in our
+%  estimates of $Z_\sC$; to do so, we use the measurements model of
+%  \citet{liang09measurements} which defines a Bayesian model for the noise
+%  in the marginals $Z_\sC$. The resultant optimization problem is,
+%\begin{align*}
+%  \hat\theta 
+%  &= \arg\min_\theta - \sum_{x \in \sD} \log p_\theta(x) + \half \eta_\theta \| \theta \|^2 \\
+%  &\quad + \half \eta_\beta \| Z_\sC - \sum_{\vx \in \sD} \E_{\theta}[\phi_\sC(\vx,\vh)] \|^2,
+%\end{align*}
+%where $\eta_\theta$ and $\eta_\beta$ are regularization parameters.
 
-\subsection{Examples}
-
-\paragraph{Latent MRFs}
-
-Image segmentation
-
-Separate features into views
-
-\todo{Some challenge example requiring measurements em}
+%
+%However, the objective \equationref{obj} requires us to compute the
+%  partition function, $A(\theta)$ and its gradient. 
+%This scales with the treewidth of the model, which is the number
+%  of rows/columns of the grid in this example, and hence can be
+%  intractable.
+%A more computationally efficient, but still consistent, alternative is
+%to optimize the pseudo-likelihood \todo{how do I write this out}.

### 2014-01-31 doc/icml2014/abstract.tex
Fixed TODO:infinite samples, TODO:using hidden everywhere but latent-variable

@@
 Specifically, we use tensor factorization to reveal partial information
-about the latent variables, rendering the otherwise non-convex negative
+about the hidden variables, rendering the otherwise non-convex negative
 log-likelihood convex.

### 2014-01-31 doc/icml2014/background.tex
Fixed TODO:infinite samples, TODO:using hidden everywhere but latent-variable

@@
   but in general computationally intractable
-  because marginalizing over latent variables $\vh$ yields a non-convex objective.
+  because marginalizing over hidden variables $\vh$ yields a non-convex objective.
 In practice, one uses local optimization procedures (e.g., EM
@@
   d \times d},\ldots$.
-For the latent marginals, we will use 
+For the hidden marginals, we will use 
 \begin{align*}
@@
 In that model, we have three observed variables $x_1, x_2, x_3$
-  that are independent conditioned on a latent variable $h_1$.  
+  that are independent conditioned on a hidden variable $h_1$.  
 The parameters of this model are the mixture probabilities $\pi\oft 1 \eqdef

### 2014-01-31 doc/icml2014/directed.tex
Fixed TODO:infinite samples, TODO:using hidden everywhere but latent-variable

@@
 \todo{we are using clique to refer to only hidden cliques; be explicit about using "hidden"}
-
-\todo{need to say we're going to do this for the infinite data case,
-and then study bounds}
-
-\todo{try to be consistent: either use latent or hidden}
 
@@
   Section~\ref{sec:directedGeneral} will describe the algorithm in full generality.
-For ease of exposition, we will describe our algorithm entirely in the
-  context of directed graphical models, though it generalizes to factor
-  graphs.
-We also present the algorithm solely in terms of the method of moments; in
+For ease of exposition we make the following simplifications to our presentation:
+(i) we describe our algorithm entirely in the context of directed
+  graphical models, though it generalizes to factor graphs.
+(ii) we derive our algorithm assuming infinite data; in \sectionref{sampleComplexity} we 
+  describe error bounds for the finite sample regime.
+(iii) we present the algorithm solely in terms of the method of moments; in
   \sectionref{piecewise}, we show how we can get a statistically more
@@
 Similarly, we can write down the moments of $x^a_1, x^a_2$, $\mO_{12}
-  \eqdef \Pr(x^a_1, x^a_2)$, in terms of the latent marginals $\mH_{12}
+  \eqdef \Pr(x^a_1, x^a_2)$, in terms of the hidden marginals $\mH_{12}
   \eqdef \Pr(h_1, h_2)$ and solve for $\mH_{12}$:
@@
 \subsection{Sample complexity}
-
+\label{sec:sampleComplexity}
 % PL: "combine" is too vague to be useful

### 2014-01-31 doc/icml2014/intro.tex
Fixed TODO:infinite samples, TODO:using hidden everywhere but latent-variable

@@
 we show that a much broader class of directed and undirected graphical models can be consistently estimated,
-specifically those in which each latent variable has three conditionally
+specifically those in which each hidden variable has three conditionally
 independent observed variables (``views'')---see \propertyref{bottleneck}.
@@
 Second, we optimize a \emph{composite marginal likelihood} to recover the marginals over
-each latent clique (e.g., $\mathbb P(h_2, h_3, h_4)$).
+each hidden clique (e.g., $\mathbb P(h_2, h_3, h_4)$).
 Normally, such a marginal likelihood objective would be non-convex,
@@
 which can be globally optimized using EM.
-Finally, the latent marginals yield the model parameters
+Finally, the hidden marginals yield the model parameters
 via simple post-processing in the directed case and convex optimization in the undirected case.
@@
   EM.  
-\todo{Is this possible?:  We show that empirically, using such constraints alleviates local
-optima issues.}
+%\todo{Is this possible?:  We show that empirically, using such constraints alleviates local
+%optima issues.}

### 2014-01-31 doc/icml2014/undirected.tex
Fixed TODO:infinite samples, TODO:using hidden everywhere but latent-variable

--- 2014-01-31 doc/icml2014/undirected.tex
@@
 In this section, we will show that we can extend the results to
-  undirected graphical models parameterized as log-linear latent variable
+  undirected graphical models parameterized as log-linear latent-variable
   models.
@@
 As before, let $\sG$ be a graphical model with observed variables $X$ and hidden variables $H$.
-We assume that the undirected graphical model is a log-linear latent
-  variable model with parameters $\theta$,
+We assume that the undirected graphical model is a log-linear latent-variable model with parameters $\theta$,
 \begin{align*}

### 2014-01-31 doc/icml2014/directed.tex
move examples

--- 2014-01-31 doc/icml2014/directed.tex
@@
 %\section{Consistent Parameter Estimation in Directed Models}
-\section{Directed models}
+\section{Directed models and exclusive views}
 \label{sec:directed}
@@
 
-\begin{figure}
-  \centering
-  \input{figures/grid-outline.tikz}
-  \caption{A directed grid model.}
-  \label{fig:grid}
-\end{figure}
+%\begin{figure}
+%  \centering
+%  \input{figures/grid-outline.tikz}
+%  \caption{A directed grid model.}
+%  \label{fig:grid}
+%\end{figure}
 
-Consider the directed grid model shown in \figureref{grid} which
+%Consider the directed grid model shown in \figureref{grid} which
+Consider the directed grid model shown in \figureref{approach}(a) which
   % DONE: all the different dependency => too strong
@@
   %via a straightforward application of perturbation bounds.
+
+\begin{figure}
+  \centering
+  \subfigure[Hidden Markov Model] {
+    \includegraphics[width=0.45\columnwidth]{figures/hmm.pdf}
+    \label{fig:examples-hmm}
+  }
+%  \subfigure[Directed grid model] {
+%    \label{fig:examples-grid}
+%    \includegraphics{figures/grid.pdf}
+%  }
+  \subfigure[Tree model] {
+    \includegraphics[width=0.45\columnwidth]{figures/tree.pdf}
+    \label{fig:examples-tree}
+  }
+  \caption{Additional examples of models learnable using \LearnMarginals}
+  \label{fig:examples}
+\end{figure}
+
+\paragraph{Hidden Markov Model}
+
+In this example (\figureref{examples-hmm}), we assume that
+  $\Pr(x_i|h_i) = O  ~\forall i$  and that $\Pr(h_{i+1} | h_i)
+  = T ~\forall i$ (i.e. we have parameter sharing).
+Note that while the first (and last) hidden variables $h_1, h_T$ in the
+  sequence are not bottlenecks, they still have exclusive views ($x_1$ and
+  $x_T$ respectively) whose parameters we know because they share
+  parameters, $O$.
+In step 1 of our algorithm, we use the bottleneck $h_2$ with views $x_1,
+  x_2, x_3$ and solve $O$.
+In step 2 of our algorithm, we can recover $\pi$ by solving for the
+  unary clique $\{h_1\}$ and recover $T$ from the clique $\{h_{1},
+  h_{2}\}$.
+
+\paragraph{Latent Tree Structure}
+
+In the latent tree structure (\figureref{examples-tree}), let the
+  parameters be $\Pr(h_i) = \pi$, $\Pr(h_i | h_1) = T ~i \in \{2,3,4\}$
+  and $\Pr(x^a_i | h_i) = \Pr(x^b_i | h_i) = O ~i \in \{2,3,4\}$.
+Note that while $h_1$ is not directly connected to an observed variable,
+  it is still a bottleneck, with views $x^a_2, x^a_3, x^a_4$.
+
+In step 1, we recover the parameters $O$ from the bottleneck $h_2$ with
+  views $\{x^a_2, x^b_2, x^a_3\}$. We also recover the conditional moments
+  $\mOpp{2}{1}$, $\mOpp{3}{1}$, $\mOpp{4}{1}$ for $h_1$. 
+In step 2, we can recover $\pi$ from the clique $\{h_1\}$, using any
+  one of views (they are all exclusive). 
+To recover $T$ from the clique $\{h_1, h_2\}$, we use the views $x^a_2$
+  (exclusive to $h_2$) and $x^a_3$ (exclusive to $h_3$). Note that while
+  $x^a_3$ is also a view for $h_2$, $x^a_3$ is independent of $h_2$ given
+  $h_1$.

### 2014-01-31 doc/icml2014/examples.tex
move examples

@@
-\subsection{More examples}
+%\subsection{More examples}
 
-We will now instantiate our algorithm for a couple more examples, illustrated in \figureref{examples}.
-
-\begin{figure}
-  \centering
-  \subfigure[Hidden Markov Model] {
-    \includegraphics[width=0.45\columnwidth]{figures/hmm.pdf}
-    \label{fig:examples-hmm}
-  }
-%  \subfigure[Directed grid model] {
-%    \label{fig:examples-grid}
-%    \includegraphics{figures/grid.pdf}
-%  }
-  \subfigure[Tree model] {
-    \includegraphics[width=0.45\columnwidth]{figures/tree.pdf}
-    \label{fig:examples-tree}
-  }
-  \caption{Additional examples of models learnable using \LearnMarginals}
-  \label{fig:examples}
-\end{figure}
-
-\paragraph{Hidden Markov Model}
-
-In this example (\figureref{examples-hmm}), we assume that
-  $\Pr(x_i|h_i) = O  ~\forall i$  and that $\Pr(h_{i+1} | h_i)
-  = T ~\forall i$ (i.e. we have parameter sharing).
-Note that while the first (and last) hidden variables $h_1, h_T$ in the
-  sequence are not bottlenecks, they still have exclusive views ($x_1$ and
-  $x_T$ respectively) whose parameters we know because they share
-  parameters, $O$.
-In step 1 of our algorithm, we use the bottleneck $h_2$ with views $x_1,
-  x_2, x_3$ and solve $O$.
-In step 2 of our algorithm, we can recover $\pi$ by solving for the
-  unary clique $\{h_1\}$ and recover $T$ from the clique $\{h_{1},
-  h_{2}\}$.
-
-\paragraph{Latent Tree Structure}
-
-In the latent tree structure (\figureref{examples-tree}), let the
-  parameters be $\Pr(h_i) = \pi$, $\Pr(h_i | h_1) = T ~i \in \{2,3,4\}$
-  and $\Pr(x^a_i | h_i) = \Pr(x^b_i | h_i) = O ~i \in \{2,3,4\}$.
-Note that while $h_1$ is not directly connected to an observed variable,
-  it is still a bottleneck, with views $x^a_2, x^a_3, x^a_4$.
-
-In step 1, we recover the parameters $O$ from the bottleneck $h_2$ with
-  views $\{x^a_2, x^b_2, x^a_3\}$. We also recover the conditional moments
-  $\mOpp{2}{1}$, $\mOpp{3}{1}$, $\mOpp{4}{1}$ for $h_1$. 
-In step 2, we can recover $\pi$ from the clique $\{h_1\}$, using any
-  one of views (they are all exclusive). 
-To recover $T$ from the clique $\{h_1, h_2\}$, we use the views $x^a_2$
-  (exclusive to $h_2$) and $x^a_3$ (exclusive to $h_3$). Note that while
-  $x^a_3$ is also a view for $h_2$, $x^a_3$ is independent of $h_2$ given
-  $h_1$.
+%We will now instantiate our algorithm for a couple more examples, illustrated in \figureref{examples}.
 

### 2014-01-31 doc/icml2014/undirected.tex

--- 2014-01-31 doc/icml2014/undirected.tex
@@
 Recall that we used tensor factorization and composite likelihood
-to estimate the latent marginals $Z_\sC = \BP(\vh_\sC)$ and $\mOpp{v}{i} = \BP(x_v \mid h_i)$.
+to estimate the hidden marginals $Z_\sC = \BP(\vh_\sC)$ and $\mOpp{v}{i} = \BP(x_v \mid h_i)$.
 For directed models, obtaining the local conditional tables from $Z_\sC$ requires

### 2014-01-31 doc/icml2014/directed.tex
Fixed TODO with asymptotic estimators stuff
Fixed reduction2

@@
 Note that in each of the above cases,
-each hidden variable $h_i$ of the clique has its own view which is conditionally independent
+  for any clique $\sC$, each hidden variable $h_i \in \sC$ has its own
+  view which is conditionally independent
 given the other variables in the clique.
@@
   Assume $\hat M_{\sV}$ has asymptotic variance $\Sigma_\sV$ defined above.
-  Then the asymptotic variance of $Z_\sC$ is:
+  Then the asymptotic variance of $\hat Z_\sC$ is:
   \begin{align*}

### 2014-01-31 doc/icml2014/directed.tex
Fixed a bunch of todos in directed

@@
 
-\todo{we are using clique to refer to only hidden cliques; be explicit about using "hidden"}
+%\todo{we are using clique to refer to only hidden cliques; be explicit about using "hidden"}
 
@@
   for any clique $\sC$, each hidden variable $h_i \in \sC$ has its own
-  view which is conditionally independent
-given the other variables in the clique.
+  view which is conditionally independent given the other hidden variables in
+  the clique.
 %Fundamentally, algorithm takes advantage of having a view for every
@@
   \label{prop:exclusive-views}
-  A clique $\sC$ of $\sG$ is said to possess \textbf{exclusive views property} if for every hidden variable $h_i \in \sC$,
+  A hidden clique $\sC$ of $\sG$ is said to possess \textbf{exclusive views property} if for every hidden variable $h_i \in \sC$,
   there is some observed variable, $x_{v}$, which is conditionally
@@
   distribution of the clique.
-Consider any clique $\sC = \{h_{i_1}, \cdots, h_{i_m}\} \in \sG$ for which this property holds. Let
+Consider any hidden clique $\sC = \{h_{i_1}, \cdots, h_{i_m}\} \in \sG$ for which this property holds. Let
   $x_{v_j}$ be an exclusive view for $h_{i_j}$ in $\sC$ (not necessarily unique) and $\sV
@@
   \begin{algorithmic}
-    \REQUIRE Clique $\sC$ with exclusive views (\propertyref{exclusive-views}).
+    \REQUIRE A hidden clique $\sC$ with exclusive views (\propertyref{exclusive-views}).
     \ENSURE Marginal distribution of the clique $Z_\sC$.
@@
 Of course, for $Z_{\sC}$ to be identifiable, we need various rank conditions,
-which will be established in \todo{section}.
+which will be established in \sectionref{sampleComplexity}.
 
@@
       same domain, $[k]$, our work generalizes to hidden domains of
-      arbitrary sizes.
-      \todo{add qualifier}
-      %provided \assumptionref{full-rank-plus} holds.
+      different sizes, provided \assumptionref{full-rank-plus} is satisfied.
       }
@@
 
-To apply $\TensorFactorize$, we need that each $\mOpp{v}{i}$ has full column rank and that $\pi^{(i)} \succ 0$ \todo{define $\pi$}.
+To apply $\TensorFactorize$, we need that each $\mOpp{v}{i}$ has full
+  column rank and that $\pi^{(i)} \eqdef \Pr(h_i) \succ 0$.
 But $\mOpp{v}{i}$ is a product of
@@
 For every clique $\sC \in \sG$ (including ones involving observed variables),
-every order-1 slice of the marginals $Z_\sC$ has full column rank $k$.
+every slice of the marginals $Z_\sC$ has full column rank $k$.
 \end{assumption}
-According to our directed grid example \todo{ref}, with $\sC = \{ h_1, h_2 \}$, then this condition implies
-$\pi = \bP(h_1) \succ 0$ and 
-$T = \bP(h_2 \mid h_1) \succ 0$;
-with $\sC = \{ h_1, x_1 \}$, we get
-$O = \bP(x_1 \mid h_1) \succ 0$.
+In our directed grid example (\sectionref{directedExample}), for the clique $\sC = \{ h_1, h_2 \}$ this
+  condition implies $\pi = \bP(h_1) \succ 0$ and $T = \bP(h_2 \mid h_1)$ has column rank $k$; 
+for the clique $\sC = \{ h_1, x_1 \}$, it implies $O = \bP(x_1 \mid h_1)$ has column rank $k$
 
@@
 $\sqrt{n} (\hat M_\sV - M_\sV) \convind \sN(0, \Sigma_\sV)$,
-where $\Sigma_\sV$ is the \emph{asymptotic variance} of $\hat M_\sV$ \todo{doesn't this need to be multiplied by something?}:
+where $\Sigma_\sV$ is the \emph{asymptotic variance} of $\hat M_\sV$: 
+% ARUN: I don't think so - \todo{doesn't this need to be multiplied by something?}
 \begin{align*}

### 2014-01-31 doc/icml2014/directed.tex
push

--- 2014-01-31 doc/icml2014/directed.tex
@@
 
+\paragraph{Example: hidden Markov model.}
+
+In the HMM (\figureref{examples-hmm}), the parameters
+are $O \eqdef \Pr(x_i|h_i)$  and that $T \eqdef \Pr(h_{i+1} | h_i)$
+for all $i$ (i.e. we have parameter sharing).
+While the first (and last) hidden variables $h_1, h_T$ in the
+  sequence are not bottlenecks, they still have exclusive views ($x_1$ and
+  $x_T$ respectively) whose parameters we know because they share
+  parameters, $O$.
+%We first use the bottleneck $h_2$ with views $x_1, x_2, x_3$ to estimate $O$,
+%and then recover $\pi$ from $\{h_1\}$ and recover $T$ from the clique $\{h_{1}, h_{2}\}$.
+
+\paragraph{Example: latent tree model.}
+
+In the latent tree model (\figureref{examples-tree}), the parameters
+are $\Pr(h_i) = \pi$, $\Pr(h_i | h_1) = T$ for $i \in \{2,3,4\}$
+and $\Pr(x^a_i | h_i) = \Pr(x^b_i | h_i) = O$ for $i \in \{2,3,4\}$.
+Note that while $h_1$ is not directly connected to an observed variable,
+  it is still a bottleneck, with views $x^a_2, x^a_3, x^a_4$.
+We can recover $T$ from the clique $\{h_1, h_2\}$ by using views $x^a_2$
+  (exclusive to $h_2$) and $x^a_3$ (exclusive to $h_3$).
+  %Note that while
+  %$x^a_3$ is also a view for $h_2$, $x^a_3$ is independent of $h_2$ given
+  %$h_1$.
+
+%In step 1, we recover the parameters $O$ from the bottleneck $h_2$ with
+%  views $\{x^a_2, x^b_2, x^a_3\}$. We also recover the conditional moments
+%  $\mOpp{2}{1}$, $\mOpp{3}{1}$, $\mOpp{4}{1}$ for $h_1$. 
+%In step 2, we can recover $\pi$ from the clique $\{h_1\}$, using any
+%  one of views (they are all exclusive). 
+%To recover $T$ from the clique $\{h_1, h_2\}$, we use the views $x^a_2$
+%  (exclusive to $h_2$) and $x^a_3$ (exclusive to $h_3$). Note that while
+%  $x^a_3$ is also a view for $h_2$, $x^a_3$ is independent of $h_2$ given
+%  $h_1$.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   }
-  \caption{Additional examples of models learnable using \LearnMarginals}
+  \caption{Examples of directed models satisfying the exclusive views property.}
   \label{fig:examples}
 \end{figure}
-
-\paragraph{Hidden Markov Model}
-
-In this example (\figureref{examples-hmm}), we assume that
-  $\Pr(x_i|h_i) = O  ~\forall i$  and that $\Pr(h_{i+1} | h_i)
-  = T ~\forall i$ (i.e. we have parameter sharing).
-Note that while the first (and last) hidden variables $h_1, h_T$ in the
-  sequence are not bottlenecks, they still have exclusive views ($x_1$ and
-  $x_T$ respectively) whose parameters we know because they share
-  parameters, $O$.
-In step 1 of our algorithm, we use the bottleneck $h_2$ with views $x_1,
-  x_2, x_3$ and solve $O$.
-In step 2 of our algorithm, we can recover $\pi$ by solving for the
-  unary clique $\{h_1\}$ and recover $T$ from the clique $\{h_{1},
-  h_{2}\}$.
-
-\paragraph{Latent Tree Structure}
-
-In the latent tree structure (\figureref{examples-tree}), let the
-  parameters be $\Pr(h_i) = \pi$, $\Pr(h_i | h_1) = T ~i \in \{2,3,4\}$
-  and $\Pr(x^a_i | h_i) = \Pr(x^b_i | h_i) = O ~i \in \{2,3,4\}$.
-Note that while $h_1$ is not directly connected to an observed variable,
-  it is still a bottleneck, with views $x^a_2, x^a_3, x^a_4$.
-
-In step 1, we recover the parameters $O$ from the bottleneck $h_2$ with
-  views $\{x^a_2, x^b_2, x^a_3\}$. We also recover the conditional moments
-  $\mOpp{2}{1}$, $\mOpp{3}{1}$, $\mOpp{4}{1}$ for $h_1$. 
-In step 2, we can recover $\pi$ from the clique $\{h_1\}$, using any
-  one of views (they are all exclusive). 
-To recover $T$ from the clique $\{h_1, h_2\}$, we use the views $x^a_2$
-  (exclusive to $h_2$) and $x^a_3$ (exclusive to $h_3$). Note that while
-  $x^a_3$ is also a view for $h_2$, $x^a_3$ is independent of $h_2$ given
-  $h_1$.

### 2014-01-31 doc/icml2014/discussion.tex
push

@@
 
-Parameter estimation for latent variable models using the method of
-  moments has received much attention in last few years.
+% Summarize paper
+For latent-variable models,
+there exists a tension between
+local optimization of likelihood,
+which is broadly applicable but has no theoretical guarantees,
+and the method of moments, which provides consistent estimators
+but are limited to models with special structure.
+The purpose of this work is to show that the two methods
+can be used synergistically to produce consistent estimates
+for a broader class of models.
+The key idea is to (i) perform tensor factorization \citep{anandkumar13tensor} to
+obtain information about each hidden variable in the form of conditional moments;
+(ii) use composite likelihood to estimate the marginal distributions of latent variables;
+and (iii) in the case of undirected models,
+use the marginal distributions to formulate a convex optimization to
+recover the model parameters.
+
+% Bottleneck is key
+Our approach allows us to handle a broad class of directed and undirected
+models in which each hidden variable is a ``bottleneck''---that is,
+it has three conditionally independent observations,
+(although the dependence on the observations could be quite complex as a function of the parameters).
+This bottleneck property of \citep{anandkumar13tensor}
+has been exploited in many other contexts,
+including Latent Dirichlet Allocation \cite{anandkumar12lda},
+parsing \cite{hsu12identifiability},
+noisy-or Bayesian networks \cite{halpern13noisyor},
+mixture of linear regressions \cite{chaganty13regression},
+and others.
+
+%Much recent work on method of moments for estimating latent-variable models
+%has focused on increasing applicability to broader range of models,
+
+%In this paper, we we've described an algorithm that can consistently
+%  estimate parameters for the class of directed and undirected graphical
+%  models for which each hidden variable has at least three conditionally
+%  independent observed variables (\propertyref{bottleneck}).
+
+%This condition is sufficiently general to capture a variety of popular
+  %models like latent trees and grids (like a pairwise latent Markov random field).
+%The undirected model setting applies to log-linear models, allowing
+%  for discriminative featurization; furthermore, in practice, the
+%  bottleneck condition can be enforced on a model if the observed
+%  features can be split into two uncorrelated sets.
+
+% Classes of models
+Notably, the bottleneck property allows us to perform consistent parameter
+estimation in undirected log-linear models with hidden variables with high
+treewidth, which
+
+The key property we exploited was the presence of exclusive views
+  (\propertyref{exclusive-views}) with learnable parameters.
+In some sense, the presence of these views helps tie down the hidden
+  variables to the data. 
+While other authors, e.g \citet{anandkumar12lda, anandkumar2013linear,
+  halpern13noisyor} are able to learn parameters for models outside this
+  family, they do so with some knowledge of the correlation structure
+  between variables. 
+In a pre-processing step, they ``subtract'' out the correlations and
+  recover a graph that satisfies the bottleneck property
+  \propertyref{bottleneck}.
+% TOO BOLD
+%Thus, while \propertyref{bottleneck} is much stronger than
+%  \propertyref{exclusive-views}, we expect that without any knowledge of
+%  the correlation between variables, \propertyref{bottleneck} is
+%  necessary for identifiability.
+
+% Broad interest in learning more models
+The use of method of moments for learning latent-variable models
+%\cite{balle11transducer}
+including 
+Much of the Parameter estimation for latent-variable models using the method of
+  moments has received much attention in recent years,
+aimed at increasing their applicability to broader classes of models.
+
 Our work uses the \TensorFactorize algorithm presented by
-  \citet{anandkumar13tensor}, though \citet{anandkumar12moments} first
+  \citet{anandkumar13tensor}, first
   studied the consistent parameter estimation algorithm for the
   three-view mixture model.
+
 %\todo{What about Mossel and Roch?}
-Several authors have extended this work to settings that do not meet the
-  criterion \propertyref{bottleneck}, but rely on knowing more about
-  the underlying correlation structure of the problem.
+
 \citet{anandkumar12lda} are able to recover parameters for correlated
@@
   bottlenecks to be discovered.
+
+work to settings that do not meet the
+  criterion \propertyref{bottleneck}, but rely on knowing more about
+  the underlying correlation structure of the problem.
+
 In comparison, our work applies to a general family of models, including
@@
 
-In this paper, we've described an algorithm that can consistently
-  estimate parameters for the class of directed and undirected graphical
-  models for which each hidden variable has at least three conditionally
-  independent observed variables (\propertyref{bottleneck}).
-This condition is sufficiently general to capture a variety of popular
-  models like latent trees and grids (like a pairwise latent Markov
-  random field).
-The undirected model setting applies to log-linear models, allowing
-  for discriminative featurization; furthermore, in practice, the
-  bottleneck condition can be enforced on a model if the observed
-  features can be split into two uncorrelated sets.
-
-The key property we exploited was the presence of exclusive views
-  (\propertyref{exclusive-views}) with learnable parameters.
-In some sense, the presence of these views helps tie down the hidden
-  variables to the data. 
-While other authors, e.g \citet{anandkumar12lda, anandkumar2013linear,
-  halpern13noisyor} are able to learn parameters for models outside this
-  family, they do so with some knowledge of the correlation structure
-  between variables. 
-In a pre-processing step, they ``subtract'' out the correlations and
-  recover a graph that satisfies the bottleneck property
-  \propertyref{bottleneck}.
-Thus, while \propertyref{bottleneck} is much stronger than
-  \propertyref{exclusive-views}, we expect that without any knowledge of
-  the correlation between variables, \propertyref{bottleneck} is
-  necessary for identifiability.
+% Identifiability
+Of course, there are models which 
+\cite{kruskal77three}

### 2014-01-31 doc/icml2014/examples.tex
Removed false intuitions about the compositie-likelihood
push

@@
-%\subsection{More examples}
-
 %We will now instantiate our algorithm for a couple more examples, illustrated in \figureref{examples}.
 
-\paragraph{Aggregating observations}
-We describe two practical considerations to use samples more efficiently.
-Firstly, if we have multiple exclusive views for a hidden variable,
-  intuitively, it is better to aggregate over them. 
-For example, consider a hidden variable $h_1$ with multiple exclusive
-  views $x_1$ and $x_2$.
-With the method of moments perspective, to learn the marginal
-  distribution over $h_1$, one must solve the following reconstruction
-  problem, 
-\begin{align*}
-  \hat Z_{h_1} &= \arg\min_{Z_{h_1}} \half \|Z_{h_1}(\mOpp{1}{1}) - M_1 \|^2 + \half \|Z_{h_1}(\mOpp{2}{1}) - M_2 \|^2,
-\end{align*}
-which does not have a closed form solution. 
-In contrast, this naturally fits into the convex optimization framework, where $\hat Z_{h_1}$ will now be,
-\begin{align*}
-  \hat Z_{h_1} &= \arg\min_{Z_{h_1}} \sum_{\vx \in \sD} \log Z_{h_1}( \mOpp{1}{1}[x_1] \odot \mOpp{1}{1}[x_1] ),
-\end{align*}
-where $\cdot$ denotes element-wise multiplication.
-An important note to make is that \TensorFactorize only returns
-  a solution up to permutation; if $\mOpp{1}{1}$ and
-  $\mOpp{1}{2}$ do not belong to the same bottleneck (e.g. $x^a_2$ and
-  $x^b_2$ in \figureref{examples-tree}), then some
-  care must be taken to ensure they have the same labelling.
-
-Secondly, we can exploit parameter sharing by aggregating over
-  disjoint sets of observed variables. 
-For example, in \figureref{examples-hmm}, we can aggregate the statistics for
-  bottlenecks $h_i$ with views $\{x_{i-1}, x_{i}, x_{i+1}\}$ before
-  running \TensorFactorize; this will give us a consistent estimate for
-  $O$ (as well as $T$).
+%\paragraph{Aggregating observations}
+%We describe two practical considerations to use samples more efficiently.
+%Firstly, if we have multiple exclusive views for a hidden variable,
+%  intuitively, it is better to aggregate over them. 
+%For example, consider a hidden variable $h_1$ with multiple exclusive
+%  views $x_1$ and $x_2$.
+%With the method of moments perspective, to learn the marginal
+%  distribution over $h_1$, one must solve the following reconstruction
+%  problem, 
+%\begin{align*}
+%  \hat Z_{h_1} &= \arg\min_{Z_{h_1}} \half \|Z_{h_1}(\mOpp{1}{1}) - M_1 \|^2 + \half \|Z_{h_1}(\mOpp{2}{1}) - M_2 \|^2,
+%\end{align*}
+%which does not have a closed form solution. 
+%In contrast, this naturally fits into the convex optimization framework, where $\hat Z_{h_1}$ will now be,
+%\begin{align*}
+%  \hat Z_{h_1} &= \arg\min_{Z_{h_1}} \sum_{\vx \in \sD} \log Z_{h_1}( \mOpp{1}{1}[x_1] \odot \mOpp{1}{1}[x_1] ),
+%\end{align*}
+%where $\cdot$ denotes element-wise multiplication.
+%An important note to make is that \TensorFactorize only returns
+%  a solution up to permutation; if $\mOpp{1}{1}$ and
+%  $\mOpp{1}{2}$ do not belong to the same bottleneck (e.g. $x^a_2$ and
+%  $x^b_2$ in \figureref{examples-tree}), then some
+%  care must be taken to ensure they have the same labelling.
+%
+%Secondly, we can exploit parameter sharing by aggregating over
+%  disjoint sets of observed variables. 
+%For example, in \figureref{examples-hmm}, we can aggregate the statistics for
+%  bottlenecks $h_i$ with views $\{x_{i-1}, x_{i}, x_{i+1}\}$ before
+%  running \TensorFactorize; this will give us a consistent estimate for
+%  $O$ (as well as $T$).

### 2014-01-31 doc/icml2014/undirected.tex
push

@@
 $\sum_{a \in X \cup H} \log p_\theta(a \mid x_{\sN(a)})$.
-In the fully-supervised setting, pseudolikelihood provides consistent estimates,
-albeit at the cost of statistical efficiency.
+In the fully-supervised setting, pseudolikelihood provides computationally efficient consistent estimates,
+albeit at the cost of statistical efficiency
+(coincidentally, the same high-level motivation for using method of moments in the first place).
 However, if we write the objective, we will see that the log-partition function $A(\theta)$

### 2014-01-31 doc/icml2014/discussion.tex
finish discussion

@@
 can be used synergistically to produce consistent estimates
-for a broader class of models.
-The key idea is to (i) perform tensor factorization \citep{anandkumar13tensor} to
-obtain information about each hidden variable in the form of conditional moments;
-(ii) use composite likelihood to estimate the marginal distributions of latent variables;
-and (iii) in the case of undirected models,
-use the marginal distributions to formulate a convex optimization to
-recover the model parameters.
+for a broader class of directed and undirected models.
+A notable example are undirected log-linear models with hidden variables of
+high treewidth.
+% Classes of models
+
+% Repeat
+%Our approach can be summarized: (i) perform tensor factorization \citep{anandkumar13tensor} to
+%obtain information about each hidden variable in the form of conditional moments;
+%In some sense, the presence of these views helps tie down the hidden
+%  variables to the data. 
+%(ii) use composite likelihood to estimate the marginal distributions of latent variables;
+%and (iii) in the case of undirected models,
+%use the marginal distributions to formulate a convex optimization to
+%recover the model parameters.
 
 % Bottleneck is key
-Our approach allows us to handle a broad class of directed and undirected
+Our approach provides consistent estimates for
 models in which each hidden variable is a ``bottleneck''---that is,
@@
 and others.
+We also introduce the idea of \emph{exclusive views} (\propertyref{exclusive-views}),
+the key property that allows us to learn the correlation structure between hidden variables.
 
@@
 
-% Classes of models
-Notably, the bottleneck property allows us to perform consistent parameter
-estimation in undirected log-linear models with hidden variables with high
-treewidth, which
-
-The key property we exploited was the presence of exclusive views
-  (\propertyref{exclusive-views}) with learnable parameters.
-In some sense, the presence of these views helps tie down the hidden
-  variables to the data. 
-While other authors, e.g \citet{anandkumar12lda, anandkumar2013linear,
-  halpern13noisyor} are able to learn parameters for models outside this
+% Correlation structure
+Other authors, e.g \citet{anandkumar12lda, anandkumar2013linear,
+  halpern13noisyor}, learn parameters for models outside this
   family, they do so with some knowledge of the correlation structure
-  between variables. 
+  between hidden variables. 
 In a pre-processing step, they ``subtract'' out the correlations and
   recover a graph that satisfies the bottleneck property
-  \propertyref{bottleneck}.
+  (\propertyref{bottleneck}).
 % TOO BOLD
@@
 %  necessary for identifiability.
-
-% Broad interest in learning more models
-The use of method of moments for learning latent-variable models
-%\cite{balle11transducer}
-including 
-Much of the Parameter estimation for latent-variable models using the method of
-  moments has received much attention in recent years,
-aimed at increasing their applicability to broader classes of models.
-
-Our work uses the \TensorFactorize algorithm presented by
-  \citet{anandkumar13tensor}, first
-  studied the consistent parameter estimation algorithm for the
-  three-view mixture model.
-
+%Our work uses the \TensorFactorize algorithm presented by
+%  \citet{anandkumar13tensor}, first
+%  studied the consistent parameter estimation algorithm for the
+%  three-view mixture model.
+%
 %\todo{What about Mossel and Roch?}
-
+For example,
 \citet{anandkumar12lda} are able to recover parameters for correlated
@@
   bottlenecks to be discovered. 
+Our work makes a different tradeoff: we do not assume any special correlation structure,
+but do rely on every hidden variable having an exclusive view so that we can
+learn the underlying structure.
 
-work to settings that do not meet the
-  criterion \propertyref{bottleneck}, but rely on knowing more about
-  the underlying correlation structure of the problem.
+%work to settings that do not meet the
+  %criterion \propertyref{bottleneck}, but rely on knowing more about
+  %the underlying correlation structure of the problem.
+%In comparison, our work applies to a general family of models, including
+%  latent tree models, directed and undirected grids, with no assumptions
+%  on the correlation structure aside from independences encoded in the
+%  graph.
 
-In comparison, our work applies to a general family of models, including
-  latent tree models, directed and undirected grids, with no assumptions
-  on the correlation structure aside from independences encoded in the
-  graph.
-
+% Observable operator
 Another prominent line of work in the method of moments community has
-  focussed on recovering parameters for a transformation that admits an {\em observable
-  representation}, i.e. one whose marginals can be directly computed from the data.
+  focused on recovering an {\em observable operator
+  representation} of the observations.
+  These methods which allows for prediction of new observations, but does not
+  identify the actual parameters of the model, which makes them difficult to use
+  in conjunction with likelihood-based models which are driven by parameters.
 \citet{song2011spectral} propose an algorithm to learn an observable
   operator representations for latent tree graphical models, like the
-  one in \figureref{examples-tree}, assuming \propertyref{bottleneck}. 
-Their approach is similar to ours, but restricted to trees with observed
-  variables at the leaves.
+  one in \figureref{examples-tree}, assuming the bottleneck property (\propertyref{bottleneck}). 
+Their approach is similar to our first step of learning conditional moments,
+  but only consider trees.
 \citet{parikh12spectral} extend this approach to general graphical
@@
   junction tree representation. 
-Consequently, the observable representations have size atleast that of
-  a node in the junction tree, which will be of the order of the treewidth
-  of the graph. 
-Our algorithm only constructs moments of the order of the cliques in the
-  graph, which can be exponentially smaller. 
+  Consequently, the size of the observable representations is exponential in
+  the treewidth.
+  In contrast, our algorithm only constructs moments of the order of size of the cliques
+  (and neighborhoods for pseudolikelihood), which can be much smaller.
 For example, an $n\times n$ grid model has a treewidth of $n$, but each
-  clique is of size at most $2$ in the undirected case and $3$ in the
-  directed case.
-In some sense, an observable operator re-formulation of our algorithm
-  subsumes this work.
+  clique is of size at most $2$ and the degree is at most $2$.
+%In some sense, an observable operator re-formulation of our algorithm
+  %subsumes this work.
 
 % Identifiability
-Of course, there are models which 
-\cite{kruskal77three}
+An interesting direction of study is to examine the necessity of the bottleneck
+property.  Certainly, to some extent, three views are necessary to ensure
+identifiability \cite{kruskal77three}, but requiring \emph{each} hidden variable to be
+a bottleneck is still stronger than we would like.  However, we hope that judiciously
+leveraging likelihood based methods in conjunction with the method of moments,
+as we have done in this paper, can generate new hybrid techniques for estimating
+an even richer class of latent-variable models.

### 2014-01-31 doc/icml2014/directed.tex
Udpdated examples (non-example) section

@@
 
-\todo{somewhere give examples that are not in our class, preferably the simplest one - noisy-or, cite Sontag}
-
 With this lemma in place, we present the full algorithm, \LearnMarginals~
@@
 
+\begin{figure}
+  \centering
+  \subfigure[Hidden Markov Model] {
+    \includegraphics[width=0.45\columnwidth]{figures/hmm.pdf}
+    \label{fig:examples-hmm}
+  }
+%  \subfigure[Directed grid model] {
+%    \label{fig:examples-grid}
+%    \includegraphics{figures/grid.pdf}
+%  }
+  \subfigure[Tree model] {
+    \includegraphics[width=0.45\columnwidth]{figures/tree.pdf}
+    \label{fig:examples-tree}
+  }
+  \subfigure[Non-example: a noisy-or network] {
+    \includegraphics[height=5em]{figures/non-example.pdf}
+    \label{fig:examples-noisy-or}
+  }
+  \caption{}
+  \label{fig:examples}
+\end{figure}
+
 \paragraph{Example: hidden Markov model.}
@@
 %  $h_1$.
+
+\paragraph{Non-examples}
+\label{sec:non-example}
+
+The simplest example of a model not identifiable by our algorithm is
+  a mixture model with a single view, i.e. \figureref{three-view} with
+  just the variables $h_1$ and $x_1$.
+However, without further assumptions the model itself is not
+  identifiable\footnote{To see this, observe that the probability-mass
+  can be arbitrarily exchanged between $\Pr(h_1)$ and $\Pr(x_1 | h_1)$}.
+A class of binary-valued noisy-or networks which do not possess the
+  bottleneck property are be shown to be identifiable by
+  \citet{halpern13noisyor}; 
+\figureref{examples-noisy-or} presents one example in this class.
 
@@
   %via a straightforward application of perturbation bounds.
-
-\begin{figure}
-  \centering
-  \subfigure[Hidden Markov Model] {
-    \includegraphics[width=0.45\columnwidth]{figures/hmm.pdf}
-    \label{fig:examples-hmm}
-  }
-%  \subfigure[Directed grid model] {
-%    \label{fig:examples-grid}
-%    \includegraphics{figures/grid.pdf}
-%  }
-  \subfigure[Tree model] {
-    \includegraphics[width=0.45\columnwidth]{figures/tree.pdf}
-    \label{fig:examples-tree}
-  }
-  \caption{Examples of directed models satisfying the exclusive views property.}
-  \label{fig:examples}
-\end{figure}

### 2014-01-31 doc/icml2014/background.tex
move examples
second pass

@@
 For an integer $n$, let $[n] = \{ 1, \dots, n \}$.
-For ease of exposition, we assume that $x_i \in [d]$ for
-all $i$, and that $h_i \in [k]$ for all $i$.
+For ease of exposition, we assume that $x_v \in [d]$ for
+all $v$, and that $h_i \in [k]$ for all $i$.
 Let $\sX$ be the joint domain of the observed variables, i.e. $\sX \eqdef [d]^L$ and
@@
   \begin{align*}
-    \sL &\eqdef \max_{\theta} \sum_{\vx \in \sD} \log p_\theta(\vx) \\
+    L_\text{unsup}(\theta) &\eqdef \max_{\theta} \sum_{\vx \in \sD} \log p_\theta(\vx) \\
         &=      \max_{\theta} \sum_{\vx \in \sD} \log \sum_{\vh \in \sH} p_\theta(\vx,\vh).
@@
   optima.
+We will later return to likelihoods, but let us first
+describe a method of moments approach for parameter estimation.
+The starting point is moments of the marginal distribution $\BP(\vx)$,
+and a bit of notation is necessary.
 
@@
 For a tensor $T \in \Re^{d \times \ldots \times d}$ and a vector $\vec
-  i = (i_1, \cdots, i_l)$, define the projection $T[\vec i] = T_{i_1 \ldots i_l}$.
+  i = (i_1, \cdots, i_\ell)$, define the projection $T[\vec i] = T_{i_1 \ldots i_\ell}$.
 Similarly, for a set $\sC$ of size $k$ and vector $\vec x \in \Re^d$, we
@@
   &= \sum_{\bi}
-            T[i_1, i_2, \ldots, i_l] x_1[i_1] \cdots x_\ell[i_\ell].
+            T[i_1, i_2, \ldots, i_\ell] x_1[i_1] \cdots x_\ell[i_\ell].
 \end{align*}
@@
 
-To denote the moments, we will use the notation:
+We consider up to third-order moments:
 \begin{align*}
@@
   \mO_{ij} &\eqdef \Pr(x_i, x_j), &
-  \mO_{ijk} &\eqdef \Pr(x_i, x_j, x_k),
+  \mO_{ijk} &\eqdef \Pr(x_i, x_j, x_k).
 \end{align*}
-and so on.
-We will view $\mO_i, \mO_{ij}, \mO_{ijk}, \cdots$ as tensors of
-  order 1, 2, 3, \ldots in $\Re^d, \Re^{d\times d}, \Re^{d \times
-  d \times d},\ldots$.
-For the hidden marginals, we will use 
+We will view $\mO_i, \mO_{ij}, \mO_{ijk}$ as tensors of
+  orders $1, 2, 3$ in $\Re^d, \Re^{d\times d}, \Re^{d \times d \times d}$.
+Next, define the \emph{hidden marginals}:
 \begin{align*}
@@
 \end{align*}
+These are tensors of
+  orders $1, 2, 3$ in $\Re^k, \Re^{k\times k}, \Re^{k \times k \times k}$.
 
@@
   \input{figures/three-view.tikz}
-  \caption{A three-view mixture model}
+  \caption{The canonical three-view mixture model can be estimated consistently using tensor factorization \citep{anandkumar13tensor}.}
 \end{figure}
@@
 The parameters of this model are the mixture probabilities $\pi\oft 1 \eqdef
-  \Pr(h) \in \Delta_{k-1}$ and the conditional moments $\mOpp{i}{1} \eqdef
-  \Pr(x_i \given h) \in \Re^{d \times k}$ for $i \in \{1,2,3\}$.
+  \Pr(h_1) \in \Delta_{k-1}$ and the conditional moments $\mOpp{i}{1} \eqdef
+  \Pr(x_i \given h_1) \in \Re^{d \times k}$ for $i \in \{1,2,3\}$.
 
@@
 \begin{align*}
-  \mO_{ij} &= \sum_h {\pi \oft 1}[h] {\mOpp{i}{1}}[h] \otimes {\mOpp{j}{1}}[h] \quad \forall i,j \in \{1,2,3\} \\
-  \mO_{123} &= \sum_h {\pi \oft 1}[h] {\mOpp{1}{1}}[h] \otimes {\mOpp{2}{1}}[h] \otimes  {\mOpp{3}{1}}[h].
+  \mO_{ij} &= \sum_h {\pi \oft 1}[h] {\mOpp{i}{1}}[\cdot,h] \otimes {\mOpp{j}{1}}[\cdot,h] \quad \forall i,j \in \{1,2,3\} \\
+  \mO_{123} &= \sum_h {\pi \oft 1}[h] {\mOpp{1}{1}}[\cdot,h] \otimes {\mOpp{2}{1}}[\cdot,h] \otimes  {\mOpp{3}{1}}[\cdot,h].
 \end{align*}

### 2014-01-31 doc/icml2014/directed.tex
second pass

@@
   In this section, we will develop a consistent parameter estimate for a class of directed graphical models.
-  In directed models, the conditional moments $\mOpp{v}{i} \eqdef \Pr(x_v \mid h_i)$
-  are only a subset of the parameters.
+  In general directed models, the conditional moments $\mOpp{v}{i} \eqdef \Pr(x_v \mid h_i)$
+  do not correspond directly to the underlying parameters.
   %Once we recover the conditional moments $\mOpp{v}{j} \eqdef \Pr(x_v
@@
   %graphical models.
-  Let us see how we can estimate the remaining parameters with an example (Section~\ref{sec:directedExample}).
+  Let us see how we can estimate the parameters with an example (Section~\ref{sec:directedExample}).
   Section~\ref{sec:directedGeneral} will describe the algorithm in full generality.
@@
 (i) we describe our algorithm entirely in the context of directed
-  graphical models, though it generalizes to factor graphs.
+  graphical models, though it generalizes to factor graphs;
 (ii) we derive our algorithm assuming infinite data; in \sectionref{sampleComplexity} we 
-  describe error bounds for the finite sample regime.
-(iii) we present the algorithm solely in terms of the method of moments; in
-  \sectionref{piecewise}, we show how we can get a statistically more
-  efficient estimator using convex optimization.
+  describe error bounds for the finite sample regime; and
+(iii) we present the algorithm solely in terms of linear algebra operations; in
+  \sectionref{piecewise}, we show how we can get a more statistically
+  efficient estimator using composite likelihood.
 

### 2014-01-31 doc/icml2014/intro.tex

@@
 Latent-variable graphical models provide compact representations of data
-and have been employed across many fields \todo{find a much better/broader set of citations}
-\cite{quattoni04crf,beaumont2004bayesian,haghighi06prototype}.
+and have been employed across many fields % \todo{find a much better/broader set of citations}
+\cite{ghahramani1999variational,jaakkola2011variational,blei03lda,quattoni04crf,beaumont2004bayesian,haghighi06prototype}.
 % 2. Learning them is hard.
@@
 this learning problem
-\citep{hsu09spectral,anandkumar11tree,anandkumar12moments,anandkumar12lda,hsu12identifiability,balle11transducer,balle12automata}.
-\todo{make sure we are citing the right people}
+\citep{mossel2005learning,
+hsu09spectral,
+bailly2010spectral,
+song2011spectral,
+anandkumar11tree,anandkumar12moments,anandkumar12lda,
+hsu12identifiability,
+balle12automata}.
+%\todo{make sure we are citing the right people}
 These methods exploit the linear algebraic properties of the model to

### 2014-01-31 doc/icml2014/directed.tex
Minor changes - Added note on experiment, etc.

--- 2014-01-31 doc/icml2014/directed.tex
@@
   The domain of $\hp$ is the same as $x$,
-    and $\mOpp{v}{\textrm{new}} = I$.\footnote{Though we have assumed that all hidden variables share the
+    and $\mOpp{v}{\textrm{new}} = I$.\footnote{
+      Though we have assumed that all hidden variables share the
       same domain, $[k]$, our work generalizes to hidden domains of
@@
   be bottlenecks, and can omit the observations $x^b_2, x^b_3$ and $x^b_4$.
-  \todo{make this a footnote if running out of space}
+%  \todo{make this a footnote if running out of space}
 % PL: don't understand this comment
@@
 tensors on the path from $h_i$ to $x_v$, marginalizing out the rest of the graph.
-To make this property more interpretable, let us assume the following property:\verify
+To make this property more interpretable, let us assume the following property:%\verify
 \begin{assumption} 

### 2014-01-31 doc/icml2014/directed.tex
changes

@@
 %Consider the directed grid model shown in \figureref{grid} which
-Consider the directed grid model shown in \figureref{approach}(a) which
+Consider the directed grid model shown in \figureref{approach}(a).
   % DONE: all the different dependency => too strong
-  captures the local dependency structures possible in
-  a Bayesian network.
+  %captures the local dependency structures possible in
+  %a Bayesian network.
 The model has eight observed variables $x^a_1, x^b_1 \cdots, x^a_4, x^b_4$ and four
@@
 =  \Pr(x^b_i | h_i) \in \Re^{d \times k}$. 
-Finally, we assume $O$ and $T$ have full column rank.
+Assume $O$ and $T$ have full column rank.
 
@@
 Note that the observed variables $x^a_1, x^b_1, x^a_2$ are
-  conditionally independent given $h_1$, allowing us to use
-  $\TensorFactorize$, as reviewed in \sectionref{setup}, to recover
-  $O$.
+  conditionally independent given $h_1$; we can use
+  $\TensorFactorize$ (\sectionref{setup}) to recover $O$.
 
@@
   \label{prop:exclusive-views}
-  A hidden clique $\sC$ of $\sG$ is said to possess \textbf{exclusive views property} if for every hidden variable $h_i \in \sC$,
+  A hidden clique $\sC$ of $\sG$ is said to possess the \textbf{exclusive views
+  property} if for every hidden variable $h_i \in \sC$,
   there is some observed variable, $x_{v}$, which is conditionally
-  independent of the rest of the clique given $h_i$.
+  independent of the rest of the clique given $h_i$,
   and whose conditional
@@
   &\eqdef \Pr(\Sx{\sV}) \\
-  &= \sum_{\vh \in \sH}
-      \Pr(\Sh{\sC}) \Pr(\Sx{\sV} \given \Sh{\sC}) \\
+  %&= \sum_{\vh \in \sH}
+      %\Pr(\Sh{\sC}) \Pr(\Sx{\sV} \given \Sh{\sC}) \\
       &= \sum_{\vh \in \sH} \Pr(\Sh{\sC}) 
@@
   \begin{algorithmic}
-    \REQUIRE A hidden clique $\sC$ with exclusive views (\propertyref{exclusive-views}).
+    \REQUIRE Hidden clique $\sC$ with exclusive views (\propertyref{exclusive-views}).
     \ENSURE Marginal distribution of the clique $Z_\sC$.
@@
 We have shown how to estimate the parameters of any clique that possesses the exclusive
-  views property (\propertyref{exclusive-views}). But which cliques have this property?
+  views property (\propertyref{exclusive-views}). But in a general graph, which cliques have this property?
 In the example above, we were able to guarantee \propertyref{exclusive-views}
-  by identifying a hidden variable, $h^1$ along with three conditionally
+  by identifying a hidden variable, $h_1$ along with three conditionally
   independent observed variables, $x^a_1, x^b_1, x^a_2$.
   %followed by using \TensorFactorize to learn the conditional moments.
-
 We define such a hidden variable to be a bottleneck:
@@
   \centering
-  \subfigure[Hidden Markov Model] {
+  \subfigure[Hidden Markov model] {
     \includegraphics[width=0.45\columnwidth]{figures/hmm.pdf}
@@
   }
-  \subfigure[Non-example: a noisy-or network] {
+  \subfigure[Noisy-or model] {
     \includegraphics[height=5em]{figures/non-example.pdf}
@@
   }
-  \caption{}
+  \caption{(a) and (b): graphs that satisfy the exclusive views property; (c) does not.}
   \label{fig:examples}
@@
 are $O \eqdef \Pr(x_i|h_i)$  and that $T \eqdef \Pr(h_{i+1} | h_i)$
-for all $i$ (i.e. we have parameter sharing).
-While the first (and last) hidden variables $h_1, h_T$ in the
+for all $i$. % (i.e. we have parameter sharing).
+While the first and last hidden variables $h_1, h_T$ in the
   sequence are not bottlenecks, they still have exclusive views ($x_1$ and
@@
 
-The simplest example of a model not identifiable by our algorithm is
+The simplest example of a model that cannot be recovered by our algorithm is
   a mixture model with a single view, i.e. \figureref{three-view} with
@@
   can be arbitrarily exchanged between $\Pr(h_1)$ and $\Pr(x_1 | h_1)$}.
-A class of binary-valued noisy-or networks which do not possess the
-  bottleneck property are be shown to be identifiable by
-  \citet{halpern13noisyor}; 
-\figureref{examples-noisy-or} presents one example in this class.
+A class of binary-valued noisy-or networks
+(see \figureref{examples-noisy-or} for an example)
+which do not possess the
+  bottleneck property are nonetheless identifiable;
+  see \citet{halpern13noisyor} for an algorithm.
 
@@
 For every clique $\sC \in \sG$ (including ones involving observed variables),
-every slice of the marginals $Z_\sC$ has full column rank $k$.
+every $R^{k \times k}$ slice of the marginals $Z_\sC$ has full column rank $k$.
 \end{assumption}
 In our directed grid example (\sectionref{directedExample}), for the clique $\sC = \{ h_1, h_2 \}$ this
-  condition implies $\pi = \bP(h_1) \succ 0$ and $T = \bP(h_2 \mid h_1)$ has column rank $k$; 
-for the clique $\sC = \{ h_1, x_1 \}$, it implies $O = \bP(x_1 \mid h_1)$ has column rank $k$
+  condition implies $\pi = \Pr(h_1) \succ 0$ and $T = \Pr(h_2 \mid h_1)$ has column rank $k$; 
+for the clique $\sC = \{ h_1, x_1 \}$, it implies $O = \Pr(x_1 \mid h_1)$ has column rank $k$
 
@@
 
-From \cite{anandkumar13tensor} given \assumptionref{full-rank-plus},
+From \citet{anandkumar13tensor} given \assumptionref{full-rank-plus},
 we have that $\mOpphat{v}{i}$ converges to $\mOpp{v}{i}$ at a rate of $n^{-\frac12}$ with a constant that
 depends polynomially on the $k$-th singular value of $\mOpp{v}{i}$.
-Of course, the singular value of $\mOpp{v}{i}$ become exponentially worse as $x_v$ and $h_i$ become farther.
-So while we can obtain consistent estimates, the sample complexity
-is exponential in the mimimum distance from $h_i$ to any three views, as to be expected.
+%Of course, the singular value of $\mOpp{v}{i}$ become exponentially worse as $x_v$ and $h_i$ become farther.
+%So while we can obtain consistent estimates, the sample complexity
+%is exponential in the mimimum distance from $h_i$ to any three views, as to be expected.
 %(\todo{say something
 %quantitative?  annoying part is that we would have to bring in the bounds, so leave it out probably}).
-
 % DONE: need to setup the asymptotic - it's not fresh in everyone's mind.
@@
   a vector in $\Re^{k^m}$, and represent $\mOppAll \eqdef
-  \mOpp{v_1}{i_1} [x_{v_1}] \otimes \cdots \otimes
-  \mOpp{v_m}{i_m} [x_{v_m}]$ as a matrix in $\Re^{d^m \times
+  \mOpp{v_1}{i_1} \otimes \cdots \otimes
+  \mOpp{v_m}{i_m}$ as a matrix in $\Re^{d^m \times
   k^m}$.
@@
   \begin{align*}
-    \Sigma^{\mom} &= \mOppAlli \Sigma_\sV \mOppAllit,
+    \Sigma^{\mom} &= \mOppAlli \Sigma_\sV \mOppAllit.
   \end{align*}
-  where $\pinv{\mOppAll} \eqdef \pinv{\mOpp{v_1}{i_1}} \otimes
-  \cdots \otimes \pinv{\mOpp{v_m}{i_m}}$, a $d^m \times k^m$ matrix.
+  %where $\pinv{\mOppAll} \eqdef \pinv{\mOpp{v_1}{i_1}} \otimes
+  %\cdots \otimes \pinv{\mOpp{v_m}{i_m}}$, a $d^m \times k^m$ matrix.
 \end{lemma}

### 2014-01-31 doc/icml2014/directed.tex
Minor type fixes

--- 2014-01-31 doc/icml2014/directed.tex
@@
 % DONE: move here because not central to story
-\paragraph{Remark.} We note that \propertyref{bottleneck} can be relaxed if some cliques
-  share parameters.
-In the directed grid model, we can identify the conditional moments $O$ from just
-  $x^a_1, x^b_1$ and $x^a_2$.
-  Therefore, we do not require that $h_2, h_3$ or $h_4$
-  be bottlenecks, and can omit the observations $x^b_2, x^b_3$ and $x^b_4$.
-%  \todo{make this a footnote if running out of space}
-% PL: don't understand this comment
-%In this case, we only require that the hidden variables in distinct
-  %cliques share parameters.
+%\paragraph{Remark.} We note that \propertyref{bottleneck} can be relaxed if some cliques
+%  share parameters.
+%In the directed grid model, we can identify the conditional moments $O$ from just
+%  $x^a_1, x^b_1$ and $x^a_2$.
+%  Therefore, we do not require that $h_2, h_3$ or $h_4$
+%  be bottlenecks, and can omit the observations $x^b_2, x^b_3$ and $x^b_4$.
+%%  \todo{make this a footnote if running out of space}
+%% PL: don't understand this comment
+%%In this case, we only require that the hidden variables in distinct
+%  %cliques share parameters.
 

### 2014-01-31 doc/icml2014/intro.tex
Minor type fixes

@@
 and have been employed across many fields % \todo{find a much better/broader set of citations}
-\cite{ghahramani1999variational,jaakkola2011variational,blei03lda,quattoni04crf,beaumont2004bayesian,haghighi06prototype}.
+\cite{ghahramani1999variational,jaakkola1999variational,blei03lda,quattoni04crf,beaumont2004bayesian,haghighi06prototype}.
 % 2. Learning them is hard.
@@
 factorize the moments into parameters, providing strong theoretical guarantees.
-However, these methods apply to a limited set of models, and are thus
+However, they apply to a limited set of models, and are thus
 not as broadly applicable as EM.
@@
   \subimport{figures/}{approach.tikz}
-  \caption{\todo{need three steps: tensor factorization, optimize composite likelihood, recover model parameters;} Overview of approach}
+  \caption{
+ % \todo{need three steps: tensor factorization, optimize composite likelihood, recover model parameters;}
+  Overview of approach}
 \end{figure}

### 2014-01-31 doc/icml2014/undirected.tex
Minor type fixes

@@
 we will have to perform global optimization.
-Fortunately, as we will show, this optimization problem is convex,
-by relating the recovered conditional moments with the sufficient statistics of
+Fortunately, we will show that this optimization problem is convex
+by relating the recovered conditional moments with the expected features of
 the log-linear model.
@@
   \centering
-  \subimport{figures/}{mrf.tikz}
+  \includegraphics[width=0.6\columnwidth]{figures/mrf.pdf}
+%  \subimport{figures/}{mrf.tikz}
   \caption{Example: undirected grid model where each hidden variable has two

### 2014-01-31 doc/icml2014/discussion.tex
done

@@
 local optimization of likelihood,
-which is broadly applicable but has no theoretical guarantees,
+which is broadly applicable but has no global theoretical guarantees,
 and the method of moments, which provides consistent estimators
@@
 models in which each hidden variable is a ``bottleneck''---that is,
-it has three conditionally independent observations,
+it has three conditionally independent observations
 (although the dependence on the observations could be quite complex as a function of the parameters).
-This bottleneck property of \citep{anandkumar13tensor}
+This bottleneck property of \citet{anandkumar13tensor}
 has been exploited in many other contexts,
 including Latent Dirichlet Allocation \cite{anandkumar12lda},
-parsing \cite{hsu12identifiability},
+parsing models \cite{hsu12identifiability},
 noisy-or Bayesian networks \cite{halpern13noisyor},
@@
 identifiability \cite{kruskal77three}, but requiring \emph{each} hidden variable to be
-a bottleneck is still stronger than we would like.  However, we hope that judiciously
-leveraging likelihood based methods in conjunction with the method of moments,
-as we have done in this paper, can generate new hybrid techniques for estimating
+a bottleneck is still stronger than we would like.  We hope that by judiciously
+leveraging likelihood-based methods in conjunction with the method of moments
+we can generate new hybrid techniques for estimating
 an even richer class of latent-variable models.

### 2014-01-31 doc/icml2014/undirected.tex
done

@@
 $\hat\theta$, is a consistent estimator.\footnote{Of course, the log-linear model must be identifiable
-(violated, for example, if two components of $\phi(\vx_\sC, \vh_\sC)$ were identical).
+(it is violated, for example, if two components of $\phi(\vx_\sC, \vh_\sC)$ were identical).
 In general, we can identify the parameters up to kernel of $\nabla^2 A(\theta)$.}
@@
 Unlike for directed models, in the undirected case,
-graphs with high tree-width present a challenge
-for even fully-supervised maximum likelihood.
-This motivates looking at pseudolikelihood objective of \citep{besag75pseudo},
-which optimizes the likelihood of each variable $a$ given its neighbors $\sN(a)$:
-$\sum_{a \in X \cup H} \log p_\theta(a \mid x_{\sN(a)})$.
+graphs with high treewidth present a challenge
+for even fully-supervised maximum likelihood due to the intractability of the log-partition function.
+This motivates looking at pseudolikelihood objective of \citep{besag75pseudo} for a computationally more
+efficient solution.
+The pseudolikelihood objective sums over the log-probability of each variable $a$ given its neighbors $\sN(a)$:
+$$L_\text{pseudo}(\theta) \eqdef \E_{x \sim \sD}\left[\sum_{a \in X \cup H} \log p_\theta(a \mid x_{\sN(a)})\right].$$
 In the fully-supervised setting, pseudolikelihood provides computationally efficient consistent estimates,
-albeit at the cost of statistical efficiency
-(coincidentally, the same high-level motivation for using method of moments in the first place).
-However, if we write the objective, we will see that the log-partition function $A(\theta)$
-is replaced with a sum over $A_a(\theta; x_{\sN(a)}) = \sum_a \exp(\theta^\top\phi(a))$,
-which cannot be computed from $Z_\sC$, since the neighbors $\sN(a)$ in general do not form a clique.
-However, we adapt our earlier algorithm to estimate the marginals over all neighborhoods $Z_{\sN(a)}$.
-Of course, this method is only viable for graphs of low degree ($|\sN(a)|$ is small).
+albeit at the cost of statistical efficiency.
+Coincidentally, the same high-level motivation for using method of moments in the first place.
+
+However, pseudolikelihood also replaces the log-partition function $A(\theta)$
+with a sum over the variables: $A_a(\theta; x_{\sN(a)}) = \sum_a \exp(\theta^\top\phi(a))$,
+which cannot be computed from just the hidden clique marginals $Z_\sC$,
+since the neighbors $\sN(a)$ in general do not form a clique.
+However, we adapt our earlier algorithm ($\LearnClique$) to estimate the marginals over all neighborhoods $Z_{\sN(a)}$.
+While this method is only viable for graphs of low degree ($|\sN(a)|$ is small),
+it can handle graphs with high treewidth.
 For example, in the grid model, the degree is only 4.

### 2014-01-31 doc/icml2014/directed.tex
Minor fixes to directed and new graph

--- 2014-01-31 doc/icml2014/directed.tex
@@
 \paragraph{Estimating $V$}
-Finally, we can estimate $V$ by renormalizing the third-order moments $\mO_{234} \eqdef \Pr(x^a_2, x^a_3, x^a_4)$,
-which can be obtained analogously:
+Finally, we can estimate $V$ by renormalizing the hidden marginals
+$\mH_{234} \eqdef \Pr(h_2, h_3, h_4)$ from the third-order moments
+$\mO_{234} \eqdef \Pr(x^a_2, x^a_3, x^a_4)$:
 \begin{align*}

### 2014-01-31 doc/icml2014/intro.tex
done

@@
 Recently, unsupervised learning techniques based on the spectral method of moments
-have offered a refreshing and promising perspective on
+have offered a refreshing perspective on
 this learning problem
@@
  % \todo{need three steps: tensor factorization, optimize composite likelihood, recover model parameters;}
-  Overview of approach}
+  Overview of our approach: we first use tensor factorization to learn the conditional moments for each hidden variable.
+  Then we optimize a composite likelihood to recover the hidden clique marginals,
+  given which we can extract the model parameters.
+  }
 \end{figure}
@@
 In this paper,
-we show that a much broader class of directed and undirected graphical models can be consistently estimated,
+we show that a much broader class of discrete directed and undirected graphical models can be consistently estimated,
 specifically those in which each hidden variable has three conditionally
@@
   %problem convex (\sectionref{undirected}).
-Finally, our approach gracefully extends to more general model families
-  in which we are unable to recover the marginal distributions for all
-  hidden cliques; in such scenarios, we use posterior regularization to constrain
-  EM.  
+%Finally, our approach gracefully extends to more general model families
+%  in which we are unable to recover the marginal distributions for all
+%  hidden cliques; in such scenarios, we use posterior regularization to constrain
+%  EM.  
 %\todo{Is this possible?:  We show that empirically, using such constraints alleviates local

### 2014-02-01 doc/icml2014/directed.tex
Many fine changes

@@
 Consider any hidden clique $\sC = \{h_{i_1}, \cdots, h_{i_m}\} \in \sG$ for which this property holds. Let
-  $x_{v_j}$ be an exclusive view for $h_{i_j}$ in $\sC$ (not necessarily unique) and $\sV
+  $x_{v_j}$ be any exclusive view for $h_{i_j}$ in $\sC$ and $\sV
   = \{x_{v_1}, \cdots, x_{v_m}\}$ be a set of views for the clique $\sC$.
@@
 
-  Let $x_v$ be an observed variable with parents $P$ and children $C$.
+  Let $x_v$ be an observed variable with parents $\Pa(x_v)$ and children $\text{Ch}(x_v)$.
   Consider the following transformation.
@@
     via a path that does not pass through $h$, contradicting the
-    statement that $x_1, x_2 \in \sV$.
+    statement that $x_1, x_2 \in \sV_h$.
 
-  Thus it must be the case that $\sV_{h} \subsetneq \Union_{h' \neq
+  Thus, it must be the case that $\sV_{h} \subsetneq \Union_{h' \neq
     h \in \sC} \sV_{h'}$ and hence there exists a unique view for $h$.
 
-  Algorithmically, finding the unique views can be done by subtracting
-    out the remaining set, i.e. $x_{v_h} \in \sV_h \setminus \Union_{h'
-    \in \sC} \sV_{h'}$
+  Algorithmically, finding the unique view for $h$ can be done by
+  subtracting the views of all other variables in the clique from
+  $\sV_h$,i.e. $x_{v_h} \in \sV_h \setminus \Union_{h' \in \sC} \sV_{h'}$
 \end{proof}
 
-With this lemma in place, we present the full algorithm, \LearnMarginals~
+With this lemma in place, we present the full algorithm, $\LearnMarginals$,
 in \algorithmref{directed}.
@@
   \begin{algorithmic}
-    \REQUIRE A graphical model $\sG$ satisfying \propertyref{bottleneck}, data $\sD$
+    \REQUIRE Graphical model $\sG$ satisfying \propertyref{bottleneck}, data $\sD$
     \ENSURE Marginals $Z_\sC$ for every clique $\sC \in \sG$
@@
 for all $i$. % (i.e. we have parameter sharing).
-While the first and last hidden variables $h_1, h_T$ in the
+While the first and last hidden variables $h_1, h_M$ in the
   sequence are not bottlenecks, they still have exclusive views ($x_1$ and
-  $x_T$ respectively) whose parameters we know because they share
+  $x_M$ respectively) whose parameters we know because they share
   parameters, $O$.
@@
 However, without further assumptions the model itself is not
-  identifiable\footnote{To see this, observe that the probability-mass
-  can be arbitrarily exchanged between $\Pr(h_1)$ and $\Pr(x_1 | h_1)$}.
+  identifiable\footnote{To see this, observe that probability-mass
+  can be arbitrarily exchanged between $\Pr(h_1)$ and $\Pr(x_1 | h_1)$ while preserving the marginals $\Pr(x_1)$}.
 A class of binary-valued noisy-or networks
@@
 First, $\hat M_\sV$ is just an empirical average of multinomials over the data points.
-Abusing notation slightly, we let $M_\sV$ also denote its $d^{|\sV|}$-dimensional vectorized version.
+Abusing notation slightly, we let $M_\sV$ also denote its $d^{m}$-dimensional vectorized version.
 We also represent $\mH_\sC$ as

### 2014-02-01 doc/icml2014/intro.tex
Many fine changes

@@
 song2011spectral,
-anandkumar11tree,anandkumar12moments,anandkumar12lda,
+anandkumar11tree,
+anandkumar12lda,
+anandkumar12moments,
 hsu12identifiability,
@@
 the tensor factorization algorithm of
-\citet{anandkumar12moments,anandkumar13tensor} to identify the \emph{conditional
-moments} $\E[x_i^a \mid h_i]$ and $\E[x_i^b \mid h_i]$ for each $i$.
+\citet{anandkumar13tensor} to identify the \emph{conditional
+moments} $\Pr[x_i^a \mid h_i]$ and $\Pr[x_i^b \mid h_i]$ for each $i$.
 Second, we optimize a \emph{composite marginal likelihood} to recover the marginals over

### 2014-02-05 java/experiments/Pairwise.py
Updated experiments

@@ -10,18 +10,23 @@ import scabby
 
 EXPT_NAME = "PairwiseHMMRecovery"
 
-KD_VALUES = [(2,2), (2,3),] # (3,3),]# (3,5), (3,10), (5,10)]
-L = [3, 5]
-#MODES = ["SpectralConvex", "SpectralInitialization"]
-MODES = ["SpectralConvex", "SpectralInitialization"]
+KD_VALUES = [(2,3),]# (3,10), (5,10)]
+N_VALUES = [1e3, 2e3, 5e3, 7e3, 
+            1e4, 2e4, 5e4, 7e4, 
+            1e5, 2e5, 5e5, 7e5, 
+            1e6, 
+            5e7 # equivalent to infinity
+            ]
+MODES = ["EM", "Piecewise", "Spectral"]
 
 def get_settings(args):
+    l = 5
     for mode in MODES:
         for k,d in KD_VALUES:
-            for l in L:
-                for initO in ["true", "false"]:
-                    assert k <= d
-                    for model_seed in xrange( args.instantiations ):
+            assert k <= d
+            for n in N_VALUES:
+                for model_seed in xrange( args.instantiations ):
+                    for data_seed in xrange( args.realizations ):
                         for init_seed in xrange( args.initializations ):
                             yield dict(locals())
 
@@ -38,10 +43,11 @@ def do_run(args):
     cmd = '\
 ./run.sh learning.experiments.PairwiseHMMRecovery\
  -execPoolDir {args.execdir}\
- -mode {mode} -N 1e5 \
- -stateCount {k} -emissionCount {d} -L {l}\
- -initExactO {initO}\
- -paramRandom {model_seed}\
+ -mode {mode} -N {n} \
+ -K {k} -D {d} -L {l}\
+ -smoothMeasurements 1e-4\
+ -paramsRandom {model_seed}\
+ -genRandom {data_seed}\
  -initRandom {init_seed}'
 
     settings = get_settings(args)
@@ -50,7 +56,10 @@ def do_run(args):
         count = len(list(settings))
         print "%d settings split over %d runs (%d each)" %(count, args.njobs, count/args.njobs+1)
     elif args.parallel:
-        scabby.parallel_spawn( args.exptdir, "qstart-short", cmd, args.njobs, settings )
+        if args.nlpsub:
+            scabby.parallel_spawn( args.exptdir, "qstart-short", cmd, args.njobs, settings )
+        else:
[…]

### 2014-02-05 java/experiments/toyGridGraph.py

@@ -14,20 +14,22 @@ import itertools as it
 EXPT_NAME = "ToyGridGraph"
 MODEL = "grid"
 
-KD_VALUES = [(2,2), (2,3), (3,3), (3,5),]# (3,10), (5,10)]
-N_VALUES = [1e3, 2e3, 5e3, 7e3,
-            1e4, 2e4, 5e4, 7e4,
-            1e5, 2e5, 5e5, 7e5,
+KD_VALUES = [(2,2)] #, (2,3), (3,3), (3,5),]# (3,10), (5,10)]
+N_VALUES = [1e3, 5e3, 
+            1e4, 5e4, 
+            1e5, 5e5, 
             5e7 # equivalent to infinity
             ]
 
-MEASUREMENT_PROB_VALUES = [1.0, 0.7, 0.3, 0.0]
+MEASUREMENT_PROB_VALUES = [1.0, 0.0]
 NOISE_VALUES = [0.,] # 1e-1,] #1e-2]
 
 PRECONDITIONG_VALUES = [0.0,]# 1e-3] #1e-2, 1e-3]
+USE_TRANSITIONS = ["true","false"]
 
 def get_true_settings(args):
     model = MODEL
+    useT = "false"
     preconditioning = 0.0
     mode = "TrueMeasurements"
     for k,d in KD_VALUES:
@@ -48,9 +50,10 @@ def get_spectral_settings(args):
         assert k <= d
         for n in N_VALUES:
             for preconditioning in PRECONDITIONG_VALUES:
-                for model_seed in xrange( args.instantiations ):
-                    for initialization_seed in xrange( args.initializations ):
-                        yield dict(locals())
+                for useT in USE_TRANSITIONS:
+                    for model_seed in xrange( args.instantiations ):
+                        for initialization_seed in xrange( args.initializations ):
+                            yield dict(locals())
 
 def do_run(args):
     #random.seed(args.seed)
@@ -74,6 +77,7 @@ def do_run(args):
  -measuredFraction {measured_fraction}\
  -trueMeasurementNoise {measurement_noise}\
  -preconditioning {preconditioning}\
+ -useTransitions {useT}\
  -genNumExamples {n}\
  -SpectralMeasurements.MeasurementsEM.iters 200 -eIters 10000'
 
@@ -82,7 +86,7 @@ def do_run(args):
         count = len(list(settings))
         print "%d settings split over %d runs (%d each)" %(count, args.njobs, count/args.njobs+1)
[…]

### 2014-02-05 java/experiments/toyHMMGraph.py

@@ -13,14 +13,15 @@ import itertools as it
 
 EXPT_NAME = "ToyHMMGraph"
 
-KD_VALUES = [(2,2), (2,3), (3,3), (3,5),]# (3,10), (5,10)]
-N_VALUES = [1e3, 2e3, 5e3, 7e3,
-            1e4, 2e4, 5e4, 7e4,
-            1e5, 2e5, 5e5, 7e5,
+KD_VALUES = [(2,3),]# (3,3), (3,5),]# (3,10), (5,10)]
+N_VALUES = [1e3, 5e3,
+            1e4, 5e4,
+            1e5, 5e5,
+            1e6
             5e7 # equivalent to infinity
             ]
 
-MEASUREMENT_PROB_VALUES = [1.0, 0.7, 0.3, 0.0]
+MEASUREMENT_PROB_VALUES = [1.0, 0.0]
 NOISE_VALUES = [0.,] # 1e-1,] #1e-2]
 
 PRECONDITIONG_VALUES = [0.0,]# 1e-3] #1e-2, 1e-3]

### 2014-02-05 java/src/learning/experiments/PairwiseHMMRecovery.java

@@ -803,7 +803,13 @@ public class PairwiseHMMRecovery implements  Runnable {
     // Initialize model
     begin_track("Generating model");
     HiddenMarkovModel.Parameters trueParams = model.newParams();
-    trueParams.initRandom(paramsRandom, paramsNoise);
+    
+    double sigmak = 0.;
+    while(sigmak < 0.05) {
+        trueParams.initRandom(paramsRandom, paramsNoise);
+        SimpleMatrix sigma = MatrixOps.svd(new SimpleMatrix(trueParams.getO())).getValue1();
+        sigmak = sigma.get(K-1,K-1);
+    }
 
     // Get data
     Counter<Example> data;

### 2014-02-05 doc/icml2014/abstract.tex
push

@@
 
-Recent work in method of moments provide consistent estimates
+Recent work on the method of moments provide consistent estimates
   for latent-variable models, avoiding local optima issues,
-  but these methods can only be applied to certain types of graphical models.
-In this work, we show that the method of moments in conjunction
-  with a composite marginal likelihood objective
+  but these methods are applicable only to certain types of graphical models.
+  On the other hand, pure likelihood objectives are difficult to optimize.
+In this work, we show that using the method of moments in conjunction
+  with composite marginal likelihood
   yields consistent parameter estimates for a much broader
-  class of directed and undirected graphical models,
+  class of discrete directed and undirected graphical models,
   including loopy graphs with high treewidth.
@@
 Specifically, we use tensor factorization to reveal partial information
-about the hidden variables, rendering the otherwise non-convex negative
-log-likelihood convex.
+about the hidden variables.  This renders the otherwise non-convex
+likelihood objective convex, and EM can then be used to attain the global optimum.
 %We use the method of moments to tie each hidden to some observed
@@
   %optimization of the likelihood.
-Our approach gracefully extends to models outside our class by
-  incorporating the partial information via posterior regulraization.
+%Our approach gracefully extends to models outside our class by
+  %incorporating the partial information via posterior regulraization.
 \end{abstract}
 
-\paragraph{Last updated:} \today, \currenttime
-
-\paragraph{Errata / Incomplete}
-\begin{enumerate}
-  \item Use \texttt{\textbackslash left(}.
-  \item Use $\ell$ instead of $L$.
-  \item convex and negative log likelihood.
-\end{enumerate}
+%\paragraph{Last updated:} \today, \currenttime
+%
+%\paragraph{Errata / Incomplete}
+%\begin{enumerate}
+%  \item Use \texttt{\textbackslash left(}.
+%  \item Use $\ell$ instead of $L$.
+%  \item convex and negative log likelihood.
+%\end{enumerate}
+%

### 2014-02-05 doc/icml2014/discussion.tex

@@
   halpern13noisyor}, learn parameters for models outside this
-  family, they do so with some knowledge of the correlation structure
+  but family, but they do so with some knowledge of the correlation structure
   between hidden variables. 
@@
   representation} of the observations.
-  These methods which allows for prediction of new observations, but does not
+  These methods which allows for prediction of new observations, but do not
   identify the actual parameters of the model, which makes them difficult to use
-  in conjunction with likelihood-based models which are driven by parameters.
+  in conjunction with likelihood-based models. % which are driven by parameters.
 \citet{song2011spectral} propose an algorithm to learn an observable
@@
 For example, an $n\times n$ grid model has a treewidth of $n$, but each
-  clique is of size at most $2$ and the degree is at most $2$.
+  clique is of size at most $2$ and the degree is at most $4$.
 %In some sense, an observable operator re-formulation of our algorithm

### 2014-02-05 doc/icml2014/undirected.tex

@@
 \begin{figure}
-  \label{fig:examples-mrf}
   \centering
@@
   We derive a consistent estimate.}
+  \label{fig:examples-mrf}
 \end{figure}
@@
 While we don't have supervised data,
-note that the method of moments gives us latent marginals $Z_\sC = \BP(\vh_\sC)$.
+note that the method of moments gives us hidden marginals $Z_\sC = \BP(\vh_\sC)$.
 Furthermore, given $\mOpp{v}{i}$ and $Z_\sC$ of any $\sC \ni v$, we can recover
@@
 albeit at the cost of statistical efficiency.
-Coincidentally, the same high-level motivation for using method of moments in the first place.
+Coincidentally, this is the same high-level motivation for using method of moments in the first place.
 
@@
   to encourage $\E_{p_\theta}[\phi(\vx_\sC,\vh_\sC)]$ to match $\mu_\sC$.
-The resulting objective functions are still non-convex, but in practice we observed
-that the extra information $\mu_\sC$ helps alleviate local optima.
+The resulting objective functions are still non-convex, but we expect
+  local optima to be alleviated.
+
+%but in practice we observed that the extra information $\mu_\sC$ helps alleviate local optima.
 %We can then integrate these constraints using posterior regularization \citep{graca08em}:

### 2014-02-10 doc/icml2014/directed.tex
Added proof for assumption 2
Updated pseudolikelihood and added something to the sample complexity section

@@
   \label{prop:exclusive-views}
-  A hidden clique $\sC$ of $\sG$ is said to possess the \textbf{exclusive views
-  property} if for every hidden variable $h_i \in \sC$,
-  there is some observed variable, $x_{v}$, which is conditionally
-  independent of the rest of the clique given $h_i$,
-  and whose conditional
-  moment $\mOpp{v}{i} \eqdef \Pr(x_{v} \mid h_i)$ can be estimated consistently.
-We call $x_v$ the exclusive view of $h_i$ in $\sC$.
+A hidden sub-graph $\sG' \subseteq \sG$ (typically a clique $\sC \in
+  \sG$) is said to possess the \textbf{exclusive views property} if for
+  every hidden variable $h_i \in \sG'$, there is some observed variable,
+  $x_{v}$, which is conditionally independent of the rest of the clique
+  given $h_i$, and whose conditional moment $\mOpp{v}{i} \eqdef
+  \Pr(x_{v} \mid h_i)$ can be estimated consistently.
+We call $x_v$ the exclusive view of $h_i$ in $\sG'$.
 \end{property}
+
+For the rest of this section, we will only consider sub-graphs which are
+  cliques. 
+In \sectionref{undirected}, we will show how the generalization
+  to sub-graphs allows us to efficiently estimate the pseudo-likelihood.
+
+% \begin{property}(Exclusive views)
+%   \label{prop:exclusive-views}
+%   A hidden clique $\sC$ of $\sG$ is said to possess the \textbf{exclusive views
+%   property} if for every hidden variable $h_i \in \sC$,
+%   there is some observed variable, $x_{v}$, which is conditionally
+%   independent of the rest of the clique given $h_i$,
+%   and whose conditional
+%   moment $\mOpp{v}{i} \eqdef \Pr(x_{v} \mid h_i)$ can be estimated consistently.
+% We call $x_v$ the exclusive view of $h_i$ in $\sC$.
+% \end{property}
 
@@
 
-\renewcommand{\algorithmicrequire}{\textbf{Input:}}
-\renewcommand{\algorithmicensure}{\textbf{Output:}}
 \begin{algorithm}
@@
   \label{lem:reduction}
-Every directed graphical model can be transformed into one in which
-  the observed variables are leaves with exactly one parent. 
+Every directed (undirected) graphical model can be transformed into one in which
+  the observed variables are leaves with exactly one parent (neighbour). 
 There is a one-to-one correspondence between the parameters of this
@@
   substituting $\hp$ with $x_v$.
+
+  This procedure applies straightforwardly for undirected graphical
+  models, considering the neighbors $\sN(x_v)$ instead of its parents
+  and children.
 \end{proof}
@@
 \begin{lemma}[Bottlenecks guarantee exclusive views]
-  Provided \propertyref{bottleneck} holds, every hidden variable
-    $h_{i_1}, \cdots, h_{i_m}$ in a clique $\sC$ has a exclusive view.
+  \label{lem:bottleneck-views}  
+Let $\sG' = \{ h_{i_1}, \cdots, h_{i_m} \} \subseteq \sG$ be
+  a connected sub-graph of hidden variables. If every hidden variable
+  $h_i \in \sG'$ is a bottleneck, then they each have an exclusive view.
 \end{lemma}
@@
 
-  Let $h$ be a hidden variable in $\sC$, let $\sV_{h}$ be the set of
-    views for it and let $\sC^- \eqdef \sC \setminus \{h\}$.
+  Let $h$ be a hidden variable in $\sG'$, let $\sV_{h}$ be the set of
+  views for it and let ${\sG'}^- \eqdef \sG' \setminus \{h\}$.
   By definition, the views of $h$ are conditionally independent, viz.
@@
   A view $x \in \sV_{h}$ is exclusive for $h$ iff $x \not\in \sV_{h'} ~ \forall h'
-  \in \sC^-$.
+  \in {\sG'}^-$.
 
-  Suppose $h$ had no unique view, then $\sV_{h} \subseteq \Union_{h' \in \sC^-} \sV_{h'}$. 
+  Suppose $h$ had no unique view, then $\sV_{h} \subseteq \Union_{h' \in {\sG'}^-} \sV_{h'}$. 
   However, this would imply that for every $x_1, x_2 \in \sV_{h}$, there
-    exists some $h_1, h_2 \in \sC^-$ (not necessarily distinct) such that
+  exists some $h_1, h_2 \in {\sG'}^-$ (not necessarily distinct) such that
     $x_1$ is connected to $h_1$ and $x_2$ is connected $h_2$.
-  By participation in the clique, either $h_1 = h_2$ or $h_1$ is
+  By participation in the sub-graph, either $h_1 = h_2$ or $h_1$ is
     connected to $h_2$, implying that $x_1$ and $x_2$ are connected
@@
   Algorithmically, finding the unique view for $h$ can be done by
-  subtracting the views of all other variables in the clique from
+  subtracting the views of all other variables in the sub-graph from
   $\sV_h$,i.e. $x_{v_h} \in \sV_h \setminus \Union_{h' \in \sC} \sV_{h'}$
@@
 For every clique $\sC \in \sG$ (including ones involving observed variables),
-every $R^{k \times k}$ slice of the marginals $Z_\sC$ has full column rank $k$.
+  every $\Re^{k \times k}$ slice of the marginals $Z_\sC$ has full column
+  rank $k$ constructed by summing out the remaining indices.
 \end{assumption}
@@
   condition implies $\pi = \Pr(h_1) \succ 0$ and $T = \Pr(h_2 \mid h_1)$ has column rank $k$; 
-for the clique $\sC = \{ h_1, x_1 \}$, it implies $O = \Pr(x_1 \mid h_1)$ has column rank $k$
+for the clique $\sC = \{ h_1, x_1 \}$, it implies $O = \Pr(x_1 \mid h_1)$ has column rank $k$.
 

### 2014-02-10 doc/icml2014/undirected.tex
Updated pseudolikelihood and added something to the sample complexity section

@@
 $$L_\text{pseudo}(\theta) \eqdef \E_{x \sim \sD}\left[\sum_{a \in X \cup H} \log p_\theta(a \mid x_{\sN(a)})\right].$$
-In the fully-supervised setting, pseudolikelihood provides computationally efficient consistent estimates,
+In the fully-supervised setting, it is well-known that pseudolikelihood provides computationally efficient consistent estimates,
 albeit at the cost of statistical efficiency.
@@
 However, pseudolikelihood also replaces the log-partition function $A(\theta)$
-with a sum over the variables: $A_a(\theta; x_{\sN(a)}) = \sum_a \exp(\theta^\top\phi(a))$,
+with a sum over the variables in the neighbourhood: $A_a(\theta; x_{\sN(a)}) = \sum_{a' \in \{a\} \union \sN(a)} \exp(\theta^\top\phi(a'))$,
 which cannot be computed from just the hidden clique marginals $Z_\sC$,
 since the neighbors $\sN(a)$ in general do not form a clique.
-However, we adapt our earlier algorithm ($\LearnClique$) to estimate the marginals over all neighborhoods $Z_{\sN(a)}$.
-While this method is only viable for graphs of low degree ($|\sN(a)|$ is small),
+We adapt our earlier algorithm ($\LearnClique$) to estimate the
+  marginals over all neighborhoods $Z_{\sN(a)}$ by noting that
+  \lemmaref{bottleneck-views} guarantees exclusive views for any connected
+  sub-graph of $\sG$, not just cliques.
+
+While this method is only viable for graphs of low degree, i.e.\@ when $|\sN(a)|$ is small,
 it can handle graphs with high treewidth.
@@
 
-In general, we might only have exclusive views (\propertyref{exclusive-views}) for a subset of the cliques.
+In general, it might be case that we only have exclusive views (\propertyref{exclusive-views}) for a subset of the cliques.
 In this case, we can estimate the expected features $\mu_\sC$ for those cliques
-  and use posterior regularization \citep{graca08em} or generalized expectation criteria \citep{mann08ge}
+  and use posterior regularization \citep{graca08em} or the generalized expectation criteria \citep{mann08ge}
   to encourage $\E_{p_\theta}[\phi(\vx_\sC,\vh_\sC)]$ to match $\mu_\sC$.

### 2014-02-19 doc/icml2014/directed.tex
Updated appendix a bit:

@@
   \label{prop:exclusive-views}
-A hidden sub-graph $\sG' \subseteq \sG$ (typically a clique $\sC \in
-  \sG$) is said to possess the \textbf{exclusive views property} if for
-  every hidden variable $h_i \in \sG'$, there is some observed variable,
+A hidden clique $\sC \in \sG$ is said to possess the \textbf{exclusive views property} if for
+  every hidden variable $h_i \in \sC$, there is some observed variable,
   $x_{v}$, which is conditionally independent of the rest of the clique
@@
   \Pr(x_{v} \mid h_i)$ can be estimated consistently.
-We call $x_v$ the exclusive view of $h_i$ in $\sG'$.
+We call $x_v$ the exclusive view of $h_i$ in $\sC$.
+
+This property also extends to any connected sub-graph $\sG' \in \sG$.
 \end{property}
@@
   a connected sub-graph of hidden variables. If every hidden variable
-  $h_i \in \sG'$ is a bottleneck, then they each have an exclusive view.
+  $h_i \in \sG'$ is a bottleneck, then they have an exclusive view.
 \end{lemma}
@@
   Thus, it must be the case that $\sV_{h} \subsetneq \Union_{h' \neq
-    h \in \sC} \sV_{h'}$ and hence there exists a unique view for $h$.
+    h \in \sG'} \sV_{h'}$ and hence there exists a unique view for $h$.
 
@@
   subtracting the views of all other variables in the sub-graph from
-  $\sV_h$,i.e. $x_{v_h} \in \sV_h \setminus \Union_{h' \in \sC} \sV_{h'}$
+  $\sV_h$,i.e. $x_{v_h} \in \sV_h \setminus \Union_{h' \in \sG'} \sV_{h'}$
 \end{proof}

### 2014-03-18 doc/icml2014/rebuttal.txt
Added rebuttal

Thank you very much for your constructive comments and suggestions; we
will try to respond to your questions below:

# (All reviewers) Regarding empirical comparisons

We are actively looking into studying the empirical performance of our
proposed algorithms.

# (Reviewer 1) Regarding the canonical transformation in Lemma 1

- The parameter space defined after the canonical transformation (Lemma
  1) is only artificially larger in that it just simplifies notation
  without actually changing the underlying learning problem. 
- In particular, when the conditional moments $O^{(v,new}) = P(x_v
  | h_{new})$ are desired, we already know them to be the identity, $I$,
  and do not need to use the tensor factorization method to learn them.
- Furthermore, in the case when the conditional moments for another
  hidden variable (say $h_1$) are desired, e.g. $O^{(v,1}) = P(x_v
  | h_1)$, then only a single $x_v$ need be used. Thus, the
  transformation does not actually increase the size of the
  parameter space.
- Finally, the identity transformation between $x_v$ and $h_new$ lets us
  write $p(x_v | h_1, h_2) = p(h_new | h_1, h_2)$.

# (Reviwer 3) Novelty in the directed graphical model case

We agree that our work presents a natural extension of the work of
Anandkumar et al., however, we believe our method still offers two novel
contributions in the directed case:
- Firstly, we show that directed graphical models with high tree width
  can be _efficiently_ learned using the spectral method. Though Parikh
  et al. (2012) have studied recovering parameters in a similar class of
  graphical models in the observable operator setting, the
  method presented there relies on a junction tree representation that
  scales with the treewidth. The algorithm we present scales with
  the degree of each variable.
- Secondly, we have shown that using composite likelihoods leads to
  asymptotically better estimators; this result applies in the directed
  setting as well.

# (Reviwer 3) Extensions to more complex distributions

Given the three view assumption, our work straightforwardly generalizes
to the case where the observed variables take real values. Following
similar treatment as in Anandkumar et al. 2013a, we would be able to
recover the conditional means of the output variables. Techniques like
those proposed in Anandkumar et al. (2012a) or Hsu and Kakade (2012)
could be extended to recover covariances in limited settings (i.e.
Dirichlet distributions and spherical Gaussians respectively), but the
problem remains hard in the general case.

### 2014-03-18 doc/icml2014/undirected.tex
Minor edits from 2014 comments

@@
 Furthermore, given $\mOpp{v}{i}$ and $Z_\sC$ of any $\sC \ni v$, we can recover
-all marginals $Z_\sC = \BP(\vx_\sC, \vh_\sC)$ (with a slight augmentation of notation).
+all marginals $\BP(\vx_\sC, \vh_\sC) = Z_\sC \cdot \mOpp[\vx_\sV]$\todo{clean up}.
 Finally, the key step is that we can compute all the expected feature vectors from the marginals:
@@
 \label{eqn:logLinearFeatures}
-\mu_\sC \eqdef \E[\phi(\vx_\sC,\vh_\sC)] = \sum_{\vx_\sC,\vh_\sC} Z_\sC(\vx_\sC,\vh_\sC) \phi(\vx_\sC,\vh_\sC).
+\mu_\sC \eqdef \E[\phi(\vx_\sC,\vh_\sC)] = \sum_{\vx_\sC,\vh_\sC} \BP(\vx_\sC,\vh_\sC) \phi(\vx_\sC,\vh_\sC).
 \end{align}
@@
 However, pseudolikelihood also replaces the log-partition function $A(\theta)$
-with a sum over the variables in the neighbourhood: $A_a(\theta; x_{\sN(a)}) = \sum_{a' \in \{a\} \union \sN(a)} \exp(\theta^\top\phi(a'))$,
+with a sum over the variables in the neighbourhood: $A_a(\theta; x_{\sN(a)}) = \sum_{a' \in \{a\} \union \sN(a)} \exp(\theta^\top\phi_{\sN(a)}(a'))$,
 which cannot be computed from just the hidden clique marginals $Z_\sC$,

### 2014-04-01 doc/icml2014/directed.tex
Fixed several minor changes brought up during rebuttals

@@
 %Consider the directed grid model shown in \figureref{grid} which
-Consider the directed grid model shown in \figureref{approach}(a).
+Consider the directed grid model shown in \figureref{approach}.
   % DONE: all the different dependency => too strong
@@
   $x_{v_j}$ be any exclusive view for $h_{i_j}$ in $\sC$ and $\sV
-  = \{x_{v_1}, \cdots, x_{v_m}\}$ be a set of views for the clique $\sC$.
+  = \{x_{v_1}, \cdots, x_{v_m}\}$ be a set of exclusive views for the clique $\sC$.
 We can write down the moments $\mO_\sV$ as follows:
@@
 \end{align*}
-Finally, the conditional probability tables for $\sC$ can easily be gotten from
+Finally, the conditional probability tables for $\sC$ can easily be obtained from
   $Z_\sC$ via renormalization.
@@
 \begin{definition}[Bottleneck]
-  A hidden variable $h$ is said to be a \textbf{bottleneck} if there exists at
-    least three conditionally independent observed variables, or views,
-    $x_1, x_2, x_3$. 
+  A hidden variable $h$ is said to be a \textbf{bottleneck} if there
+  exists at least three observed variables, or views, $x_1, x_2, x_3$
+  that are conditionally independent given $h$. 
   Let $\sV_{h}$ denote a set of views for $h$ (not necessarily unique).
@@
   exists some $h_1, h_2 \in {\sG'}^-$ (not necessarily distinct) such that
-    $x_1$ is connected to $h_1$ and $x_2$ is connected $h_2$.
+    $x_1$ is connected to $h_1$ and $x_2$ is connected to $h_2$.
   By participation in the sub-graph, either $h_1 = h_2$ or $h_1$ is
@@
 
-  Thus, it must be the case that $\sV_{h} \subsetneq \Union_{h' \neq
+  Thus, it must be the case that $\sV_{h} \nsubseteq \Union_{h' \neq
     h \in \sG'} \sV_{h'}$ and hence there exists a unique view for $h$.
@@
 We can recover $T$ from the clique $\{h_1, h_2\}$ by using views $x^a_2$
-  (exclusive to $h_2$) and $x^a_3$ (exclusive to $h_3$).
+  (exclusive to $h_2$) and $x^a_3$ (exclusive to $h_1$).
   %Note that while

### 2014-04-01 doc/icml2014/discussion.tex
Fixed several minor changes brought up during rebuttals

@@
   halpern13noisyor}, learn parameters for models outside this
-  but family, but they do so with some knowledge of the correlation structure
+  family, but they do so with some knowledge of the correlation structure
   between hidden variables. 
@@
   representation} of the observations.
-  These methods which allows for prediction of new observations, but do not
+  These methods allow prediction of new observations, but do not
   identify the actual parameters of the model, which makes them difficult to use

### 2014-04-01 doc/icml2014/undirected.tex
Fixed several minor changes brought up during rebuttals

@@
 Furthermore, given $\mOpp{v}{i}$ and $Z_\sC$ of any $\sC \ni v$, we can recover
-all marginals $\BP(\vx_\sC, \vh_\sC) = Z_\sC \cdot \mOpp[\vx_\sV]$\todo{clean up}.
+all marginals $\BP(\vx_\sC, \vh_\sC) = Z_\sC \cdot \mOpp[\vx_\sV]$.%\todo{clean up}.
 Finally, the key step is that we can compute all the expected feature vectors from the marginals:

### 2014-04-01 java/tests/learning/models/HiddenMarkovModelsTest.java
Updated tolerances on HMM tests

@@ -207,6 +207,33 @@ public class HiddenMarkovModelsTest {
   }
 
   @Test
+  public void testNodeMarginals() {
+    Example ex1 = new Example(new int[] {1,1}) ;
+    double[][] nodeMarginals = new double[][] {
+            {0.8760330578512396, 0.12396694214876035},
+            {0.31210500729217305, 0.687894992707827}
+          };
+    double[][] nodeMarginals_ = model.computeHiddenNodeMarginals(hmm1, ex1);
+    for(int i = 0; i < 2; i++)
+      for(int j = 0; j < 2; j++)
+        Assert.assertTrue(Math.abs(nodeMarginals[i][j] - nodeMarginals_[i][j]) < 1e-3);
+  }
+
+  @Test
+  public void testEdgeMarginals() {
+    Example ex1 = new Example(new int[] {1,1}) ;
+    double[][][] edgeMarginals = new double[][][]{{
+            {0.2975206611570248, 0.5785123966942148},
+            {0.014584346135148274, 0.10938259601361207}
+        }};
+    double[][][] edgeMarginals_ = model.computeEdgeMarginals(hmm1, ex1);
+    for(int t = 0; t < 1; t++)
+      for(int i = 0; i < 2; i++)
+        for(int j = 0; j < 2; j++)
+        Assert.assertTrue(Math.abs(edgeMarginals[t][i][j] - edgeMarginals_[t][i][j]) < 1e-3);
+  }
+
+  @Test
   public void testMarginals() {
     HiddenMarkovModel.Parameters marginals;
 
@@ -218,15 +245,20 @@ public class HiddenMarkovModelsTest {
     Assert.assertTrue(marginals.isValid());
     // By definition
     for(int i = 0; i < marginals.weights.length; i++) {
-      Assert.assertTrue(Math.abs(marginals.weights[i] - hmm1.weights[i]) < 1e-1);
+      Assert.assertTrue(Math.abs(marginals.weights[i] - hmm1.weights[i]) < 1e-3);
     }
 
+    model.L = 2;
     Counter<Example> data = model.getDistribution(hmm1);
     marginals = (HiddenMarkovModel.Parameters) model.getMarginals(hmm1, data);
-    Assert.assertTrue(marginals.isValid());
+    HiddenMarkovModel.Parameters marginals_ = model.getSampleMarginals(data);
     // By definition
+    log(data);
+    log(marginals);
[…]

### 2014-04-01 java/experiments/do_cca.py
Added CCA visualizing EM demo

@@ -1,0 +1,63 @@
+#!/usr/bin/env python2.7
+"""
+Plot something with CCA
+"""
+
+import numpy as np
+import matplotlib.pyplot as plt
+
+from numpy.linalg import inv, norm
+
+def do_command(args):
+    data = np.loadtxt(args.data)
+    X = data[:,:-1]
+    y = data[:,-1]
+    n, d = X.shape
+
+    # Mean average
+    X = X - X.mean(0)
+    y_0 = y - y.mean(0)
+
+    if args.mode == "cca":
+        # Compute covariances
+        S_xx = X.T.dot(X) / (n-1)
+        S_xy = X.T.dot(y_0) / (n-1)
+
+        # C is proportional to S_xx^{-1} S_xy
+        c = inv(S_xx).dot(S_xy)
+        c = c / norm(c)
+    elif args.mode == "random":
+        c = np.random.rand(d)
+    else:
+        raise AttributeError("Invalid mode ", args.mode)
+
+    # Now plot
+    data = np.vstack( (X.dot(c), y) ).T
+    r_data = data[ data.T[0].argsort(), : ]
+
+    true = data[:args.true_rows]
+    em = data[args.true_rows:args.true_rows+args.em_rows,:]
+
+    plt.xlabel("Projected coordinates")
+    plt.ylabel("Log-likelihood")
+    plt.plot(r_data.T[0], r_data.T[1])
+    plt.plot(true.T[0], true.T[1], 'g*', markersize=30)
+    plt.plot(em.T[0], em.T[1], 'r.', markersize=15)
+
+    if args.output is not None and len(args.output) > 0:
+        plt.savefig(args.output)
+    else:
+        plt.show()
+
+if __name__ == "__main__":
+    import argparse
+    parser = argparse.ArgumentParser( description='Plot EM using CCA' )
+    parser.add_argument( '--mode', choices=["cca","random"], help="mode of projection" )
+    parser.add_argument( '--true-rows', type=int, default=1, help="number of true param rows" )
+    parser.add_argument( '--em-rows', type=int, default=100, help="number of EM rows" )
+    parser.add_argument( '--output', type=str, help="Where to save" )
+    parser.add_argument( 'data', type=str, help="Path to file" )
+    parser.set_defaults(func=do_command)
+
+    ARGS = parser.parse_args()
+    ARGS.func(ARGS)

### 2014-04-28 java/src/learning/common/Utils.java
Updated code to java1.8, included Identifiability experiment

@@ -20,6 +20,11 @@ import java.nio.file.Files;
 import java.util.ArrayList;
 import java.util.Collections;
 import java.util.List;
+import java.util.concurrent.atomic.AtomicInteger;
+import java.util.function.Function;
+import java.util.stream.Stream;
+
+import static java.util.stream.Collectors.toList;
 
 /**
  * Various utilities
@@ -201,13 +206,20 @@ public class Utils {
     return stateVectors.toArray(new int[count][L]);
   }
 
-  public static <T> List<Pair<T, Integer>> argsort(List<T> items) {
-    List<Pair<T,Integer>> idx = new ArrayList<>();
-    for( int i = 0; i < items.size(); i++) {
-      idx.add(Pair.with(items.get(i), i));
-    }
-    Collections.sort(idx);
+  public static <T> Stream<Pair<T,Integer>> mapi(Stream<T> stream) {
+    final AtomicInteger i = new AtomicInteger();
+    return stream.map((T elem) -> Pair.with(elem, i.getAndIncrement()));
+  }
 
-    return idx;
+  public static <T extends Comparable> List<Pair<T, Integer>> argsort(List<T> items) {
+    return argsort(items.stream()).collect(toList());
+  }
+  public static <T extends Comparable> Stream<Pair<T, Integer>> argsort(Stream<T> items) {
+    return mapi(items).sorted();
+  }
+
+  public static <T,U extends Comparable<U>> List<T> sort(List<T> items, Function<T,U> key) {
+    return argsort(items.stream().map(key)).map( (Pair<U,Integer> p) -> items.get(p.getValue1())).collect(toList());
   }
 }
+

### 2014-04-28 java/src/learning/linalg/AsymmetricAlternateMinimizer.java

@@ -117,16 +117,9 @@ public class AsymmetricAlternateMinimizer implements TensorFactorizationAlgorith
                  List<Quartet<Double, SimpleMatrix, SimpleMatrix, SimpleMatrix>> fixedPoints) {
 
     // Sort the list by value of T(a,b,c)
-    List<Double> scores = new ArrayList<>();
-    for(Quartet<Double, SimpleMatrix, SimpleMatrix, SimpleMatrix> fixedPoint : fixedPoints ) {
-      scores.add(T.project3(fixedPoint.getValue1(), fixedPoint.getValue2(), fixedPoint.getValue3()));
-    }
-    List<Pair<Double,Integer>> ordering = Utils.argsort(scores);
-    Collections.reverse(ordering);
-    List<Quartet<Double, SimpleMatrix, SimpleMatrix, SimpleMatrix>> candidates = new LinkedList<>();
-    for( Pair<Double, Integer> elem : ordering ) {
-      candidates.add( fixedPoints.get(elem.getValue1()) );
-    }
+    List<Quartet<Double, SimpleMatrix, SimpleMatrix, SimpleMatrix>> candidates =
+            Utils.sort(fixedPoints, (fixedPoint) -> T.project3(fixedPoint.getValue1(), fixedPoint.getValue2(), fixedPoint.getValue3()) );
+    Collections.reverse(candidates);
 
     List<Quartet<Double, SimpleMatrix, SimpleMatrix, SimpleMatrix>> centers = new ArrayList<>();
     while(centers.size() < K && candidates.size() > 0) {

### 2014-04-28 java/src/learning/models/loglinear/Model.java

@@ -174,8 +174,19 @@ public abstract class Model extends ExponentialFamilyModel<Example> {
     return examples;
   }
 
+
   public int getSize(Example ex) {
     return ex.x.length;
   }
+
+    @Override
+    public Example bestLabelling(Params params, Example ex) {
+        Hypergraph<Example> Hp = createHypergraph(ex, fullParams.weights, null, 1.0);
+        Hp.computePosteriors(true);
+        ex = newExample(ex.x);
+        Hp.fetchBestHyperpath(ex);
+        return ex;
+    }
+
 }
 

### 2014-04-28 java/src/learning/models/loglinear/Models.java

@@ -242,6 +242,16 @@ public class Models {
       }
       return examples;
     }
+      @Override
+      public Counter<Example> getFullDistribution(Params params_) {
+          ParamsVec params = (ParamsVec) params_;
+          Counter<Example> examples = new Counter<>();
+          examples.addAll(generateExamples(L));
+          for(Example ex: examples) {
+              examples.set( ex, getProbability(params, ex));
+          }
+          return examples;
+      }
     @Override
     public double updateMoments(Example ex, double count, SimpleMatrix P12, SimpleMatrix P13, SimpleMatrix P32, FullTensor P123) {
       double updates = 0.;
@@ -541,6 +551,7 @@ public class Models {
       }
       return updates;
     }
+
   }
 
 

### 2014-04-28 doc/icml2014/background.tex
changes
Fixed several minor changes brought up during rebuttals
Updated paper (draft to send to David Sontag)

@@
 % PL: don't know what this means
-%We assume that the model $\sG$ is parameterized at every clique $\sC$.
+%We assume that the model $\sG$ is parametrized at every clique $\sC$.
 If $\sG$ is directed, the model parameters correspond to
@@
   \begin{align*}
-    L_\text{unsup}(\theta) &\eqdef \max_{\theta} \sum_{\vx \in \sD} \log p_\theta(\vx) \\
-        &=      \max_{\theta} \sum_{\vx \in \sD} \log \sum_{\vh \in \sH} p_\theta(\vx,\vh).
+    L_\text{unsup}(\theta) &\eqdef \sum_{\vx \in \sD} \log p_\theta(\vx)
+        =  \sum_{\vx \in \sD} \log \sum_{\vh \in \sH} p_\theta(\vx,\vh).
   \end{align*}
@@
 Similarly, for a set $\sC$ of size $k$ and vector $\vec x \in \Re^d$, we
-  will use the notation $\vec x_\sC$ to get a vector $\Re^k$ which selects
+  will use the notation $\vec x_\sC$ to get a vector in $\Re^k$ which selects
   the appropriate elements of $\vec x$.
@@
   orders $1, 2, 3$ in $\Re^d, \Re^{d\times d}, \Re^{d \times d \times d}$.
-Next, define the \emph{hidden marginals}:
+Next, we define the \emph{hidden marginals}:
 \begin{align*}
@@
 The parameters of this model are the mixture probabilities $\pi\oft 1 \eqdef
-  \Pr(h_1) \in \Delta_{k-1}$ and the conditional moments $\mOpp{i}{1} \eqdef
+  \Pr(h_1) \in \Re^k$ and the conditional moments $\mOpp{i}{1} \eqdef
   \Pr(x_i \given h_1) \in \Re^{d \times k}$ for $i \in \{1,2,3\}$.
@@
 \citet{anandkumar13tensor} presented a robust tensor power method, which
-  we call $\TensorFactorize$, which converts estimates of the observed moments 
+  we call $\TensorFactorize$, which transforms estimates of the observed moments 
   into estimates of the parameters, $\pi\oft 1, \mOpp{1}{1},
@@
 In brief, the algorithm first computes a whitening matrix $W \in
-  \Re^{d \times k}$ such that $W^T \mO_{12} W = I_{k \times k}$,
+  \Re^{d \times k}$ such that $W^\top \mO_{12} W = I_{k \times k}$,
   and uses it to transform $\mO_{123}$ into a symmetric orthogonal tensor.  

### 2014-04-28 doc/icml2014/abstract.tex

@@
   for latent-variable models, avoiding local optima issues,
-  but these methods are applicable only to certain types of graphical models.
+  but are applicable only to certain types of graphical models.
   On the other hand, pure likelihood objectives are difficult to optimize.

### 2014-04-28 doc/icml2014/directed.tex

@@
 By \assumptionref{full-rank}, $O$ has full column rank and thus can be
-  inverted to recover $\pi$: $\pi = \pinv O \mO_1$.
+  inverted to recover $\pi$: $\pi = O\pinv  \mO_1$.
 
@@
 \mO_{12} = O \mH_{12} O^\top \quad\Rightarrow\quad
-  \mH_{12} = \pinv O \mO_{12} \pinvt O.
+  \mH_{12} = O\pinv  \mO_{12} O\pinvt .
 \end{align*}
@@
   \mO_{234} = \mH_{234}(O, O, O) \quad\Rightarrow\quad
-  \mH_{234} = \mO_{234}(\pinv O, \pinv O, \pinv O).
+  \mH_{234} = \mO_{234}(O\pinv , O\pinv , O\pinv ).
 \end{align*}
@@
   every hidden variable $h_i \in \sC$, there is some observed variable,
-  $x_{v}$, which is conditionally independent of the rest of the clique
+  $x_{v}$ which is conditionally independent of the rest of the clique
   given $h_i$, and whose conditional moment $\mOpp{v}{i} \eqdef
@@
   distribution of the clique.
-Consider any hidden clique $\sC = \{h_{i_1}, \cdots, h_{i_m}\} \in \sG$ for which this property holds. Let
+Consider any hidden clique $\sC = \{i_1, \cdots, i_m\}$ for which this property holds. Let
   $x_{v_j}$ be any exclusive view for $h_{i_j}$ in $\sC$ and $\sV
-  = \{x_{v_1}, \cdots, x_{v_m}\}$ be a set of exclusive views for the clique $\sC$.
+  = \{v_1, \cdots, v_m\}$ be a set of exclusive views for the clique $\sC$.
 We can write down the moments $\mO_\sV$ as follows:
@@
       &= \sum_{\vh \in \sH} \Pr(\Sh{\sC}) 
-          \Pr(x_{i_1} | h_{v_1}) \cdots \Pr(x_{i_m} | h_{v_m}) \\
+          \Pr(x_{v_1} | h_{i_1}) \cdots \Pr(x_{v_m} | h_{i_m}) \\
     &= Z_{\sC}(\mOpp{v_1}{i_1},\cdots,\mOpp{v_m}{i_m}).
@@
 \begin{align*}
-  Z_{\sC} &= \mO_\sV(\pinv{\mOpp{v_1}{i_1}},\cdots,\pinv{\mOpp{v_m}{i_m}}).
+  Z_{\sC} &= \mO_\sV(\mOpp{v_1}{i_1}\pinv,\cdots,\mOpp{v_m}{i_m}\pinv).
 \end{align*}
@@
       \STATE Identify exclusive views $\Sx{\sV} = \{x_{v_1}, \cdots, x_{v_m}\}$.
-      \STATE Return $\mH_\sC \gets \mO_{\Sx{\sV}}( \pinv{\mOpp{v_1}{i_i}}, \cdots, \pinv{\mOpp{v_m}{i_m}} )$.
+      \STATE Return $\mH_\sC \gets \mO_{\Sx{\sV}}( \mOpp{v_1}{i_i}\pinv, \cdots, {\mOpp{v_m}{i_m}}\pinv )$.
   \end{algorithmic}
@@
   (\figureref{reduction}). 
-  The domain of $\hp$ is the same as $x$,
-    and $\mOpp{v}{\textrm{new}} = I$.\footnote{
+  The domain of $\hp$ is the same as $x_v$\footnote{
       Though we have assumed that all hidden variables share the
@@
       different sizes, provided \assumptionref{full-rank-plus} is satisfied.
-      }
-
+      },
+    and $\mOpp{v}{\textrm{new}} = I$.
   Then, there is a one-to-one correspondence between every value of
@@
   \end{align*}
-  %where $\pinv{\mOppAll} \eqdef \pinv{\mOpp{v_1}{i_1}} \otimes
-  %\cdots \otimes \pinv{\mOpp{v_m}{i_m}}$, a $d^m \times k^m$ matrix.
+  %where ${\mOppAll}\pinv \eqdef {\mOpp{v_1}{i_1}}\pinv \otimes
+  %\cdots \otimes {\mOpp{v_m}{i_m}}\pinv$, a $d^m \times k^m$ matrix.
 \end{lemma}
@@
 %matrix,
-We can rewrite this representation more compactly as: $Z_{\sC} = \pinv{\mOppAll} \mO_\sV$.
+We can rewrite this representation more compactly as: $Z_{\sC} = {\mOppAll}\pinv \mO_\sV$.
 By the delta-method \cite{vaart98asymptotic},

### 2014-04-29 doc/icml2014/directed.tex
Started working on asymptotic efficiency picture

@@
       \STATE Identify exclusive views $\Sx{\sV} = \{x_{v_1}, \cdots, x_{v_m}\}$.
-      \STATE Return $\mH_\sC \gets \mO_{\Sx{\sV}}( \mOpp{v_1}{i_i}\pinv, \cdots, {\mOpp{v_m}{i_m}}\pinv )$.
+      \STATE Return $\mH_\sC \gets \mO_{\sV}( \mOpp{v_1}{i_i}\pinv, \cdots, {\mOpp{v_m}{i_m}}\pinv )$.
   \end{algorithmic}
@@
 depends polynomially on the $k$-th singular value of $\mOpp{v}{i}$.
-%Of course, the singular value of $\mOpp{v}{i}$ become exponentially worse as $x_v$ and $h_i$ become farther.
-%So while we can obtain consistent estimates, the sample complexity
-%is exponential in the mimimum distance from $h_i$ to any three views, as to be expected.
-%(\todo{say something
-%quantitative?  annoying part is that we would have to bring in the bounds, so leave it out probably}).
-% DONE: need to setup the asymptotic - it's not fresh in everyone's mind.
+
+\todo{Present sample complexity bounds for $\mOpp{v}{i}$.}
+\todo{Present sample complexity bounds for $Z_\sC$.}
+
 % Easy case
-Next, we need to show that the hidden marginals $\hat Z_\sC$ converge to $Z_\sC$.
-First, $\hat M_\sV$ is just an empirical average of multinomials over the data points.
-Abusing notation slightly, we let $M_\sV$ also denote its $d^{m}$-dimensional vectorized version.
-We also represent $\mH_\sC$ as
-  a vector in $\Re^{k^m}$, and represent $\mOppAll \eqdef
-  \mOpp{v_1}{i_1} \otimes \cdots \otimes
-  \mOpp{v_m}{i_m}$ as a matrix in $\Re^{d^m \times
-  k^m}$.
-By the central limit theorem, we have:
-$\sqrt{n} (\hat M_\sV - M_\sV) \convind \sN(0, \Sigma_\sV)$,
-where $\Sigma_\sV$ is the \emph{asymptotic variance} of $\hat M_\sV$: 
-% ARUN: I don't think so - \todo{doesn't this need to be multiplied by something?}
-\begin{align*}
-\Sigma_\sV \eqdef \dM_\sV - M_\sV M_\sV^\top, \quad \dM_\sV \eqdef \text{diag}(M_\sV).
-\end{align*}
-
-% PL: this is not obviously correct
-%To address the first, we will need that the following assumption on the
-%  parameters of the model:
-%\begin{assumption} 
-%  \label{asm:full-rank-plus}
-%  For every clique $\sC = \{h_{i_1}, \cdots, h_{i_m}\} \in \sG$, let
-%    $Z_\sC$ be the marginal distribution for that clique. 
-%  Then every mode-$i$ unfolding of $Z_\sC$ has full column rank.\verify
-%\end{assumption}
-%Intuitively, the assumption guarantees that the marginal probability of
-%  any hidden state is non-zero.
-
-We can now use the delta-method to convert the above result into 
-the asymptotic variance for the pseudoinverse version of $\LearnClique$:
-\begin{lemma}[Asymptotic variance of pseudoinverse estimator for $Z_\sC$]
-  \label{lem:mom-variance}  
-  Assume $\hat M_{\sV}$ has asymptotic variance $\Sigma_\sV$ defined above.
-  Then the asymptotic variance of $\hat Z_\sC$ is:
-  \begin{align*}
-    \Sigma^{\mom} &= \mOppAlli \Sigma_\sV \mOppAllit.
-  \end{align*}
-  %where ${\mOppAll}\pinv \eqdef {\mOpp{v_1}{i_1}}\pinv \otimes
-  %\cdots \otimes {\mOpp{v_m}{i_m}}\pinv$, a $d^m \times k^m$ matrix.
-\end{lemma}
-\begin{proof}
-For clique $\sC$, recall we have
-  $Z_{\sC} = \mO_\sV(\mOppi{v_1}{i_1},\cdots,\mOppi{v_m}{i_m})$.
-%Choosing to represent $Z_\sC$ and $M_\sV$ as vectors and
-%$\mOppi{v_1}{i_1} \otimes \cdots \otimes \mOppi{v_m}{i_m}$ as the
-%matrix,
-We can rewrite this representation more compactly as: $Z_{\sC} = {\mOppAll}\pinv \mO_\sV$.
-By the delta-method \cite{vaart98asymptotic},
-% DONE: need to shorten
-%we have that the asymptotic variance of $M_\sV$ is:
-%\begin{align*}
-%  \sqrt{n}(\hat M_\sV - M_\sV) \convind \sN(0, \Sigma_\sV),
-%\end{align*}
-%where $\Sigma_\sV$ is the variance of the observations. 
-we immediately get:
-$\sqrt{n}(\hat Z_\sC - Z_\sC) \convind \sN(0, \mOppAlli \Sigma_\sV \mOppAllit)$.
-\end{proof}
+% Next, we need to show that the hidden marginals $\hat Z_\sC$ converge to $Z_\sC$.
+% First, $\hat M_\sV$ is just an empirical average of multinomials over the data points.
+% Abusing notation slightly, we let $\hat M_\sV$ also denote its
+% $d^{m}$-dimensional vectorized version.
+% We also represent $\mH_\sC$ as
+%   a vector in $\Re^{k^m}$, and represent $\mOppAll \eqdef
+%   \mOpp{v_1}{i_1} \otimes \cdots \otimes
+%   \mOpp{v_m}{i_m}$ as a matrix in $\Re^{d^m \times
+%   k^m}$.
+% By the central limit theorem, we have:
+% $\sqrt{n} (\hat M_\sV - M_\sV) \convind \sN(0, \Sigma_\sV)$,
+% where $\Sigma_\sV$ is the \emph{asymptotic variance} of $\hat M_\sV$: 
+% % ARUN: I don't think so - \todo{doesn't this need to be multiplied by something?}
+% \begin{align*}
+% \Sigma_\sV \eqdef \dM_\sV - M_\sV M_\sV^\top, \quad \dM_\sV \eqdef \text{diag}(M_\sV).
+% \end{align*}
+% 
+% Our next step will be to use the delta-method to convert the above
+% result into the asymptotic variance for the pseudoinverse version of
+% $\LearnClique$. 
+% \begin{lemma}[Asymptotic variance of pseudoinverse estimator for $\tilde Z_\sC$]
+%   \label{lem:mom-variance}  
+%   Assume $\hat M_{\sV}$ has asymptotic variance $\Sigma_\sV$ defined above.
+%   Then the asymptotic variance of $\hat{Z_\sC}$ is:
+%   \begin{align*}
+%     \Sigma^{\mom} &= \mOppAlli \Sigma_\sV \mOppAllit.
+%   \end{align*}
+%   %where ${\mOppAll}\pinv \eqdef {\mOpp{v_1}{i_1}}\pinv \otimes
+%   %\cdots \otimes {\mOpp{v_m}{i_m}}\pinv$, a $d^m \times k^m$ matrix.
+% \end{lemma}
+% \begin{proof}
+% For clique $\sC$, recall we have
+%   $Z_{\sC} = \mO_\sV(\mOppi{v_1}{i_1},\cdots,\mOppi{v_m}{i_m})$.
+% %Choosing to represent $Z_\sC$ and $M_\sV$ as vectors and
+% %$\mOppi{v_1}{i_1} \otimes \cdots \otimes \mOppi{v_m}{i_m}$ as the
+% %matrix,
+% We can rewrite this representation more compactly as: $Z_{\sC} = {\mOppAll}\pinv \mO_\sV$.
+% By the delta-method \cite{vaart98asymptotic},
+% % DONE: need to shorten
+% %we have that the asymptotic variance of $M_\sV$ is:
+% %\begin{align*}
+% %  \sqrt{n}(\hat M_\sV - M_\sV) \convind \sN(0, \Sigma_\sV),
+% %\end{align*}
+% %where $\Sigma_\sV$ is the variance of the observations. 
+% we immediately get:
+% $\sqrt{n}(\hat Z_\sC - Z_\sC) \convind \sN(0, \mOppAlli \Sigma_\sV \mOppAllit)$.
+% \end{proof}
 

### 2014-05-01 doc/icml2014/icml2014-presentation/diagrams.tex
Added icml presentation

\newcommand<>{\drawgen}[1]{%
  \uncover#2{
    \point{start-gen}{#1}
    \node[style=node] (h) at (start-gen) {$h$};
%    \node[left=0.1cm of h] {$h$};
    \node[style=obsnode,below=1cm of h] (x2) {$x_2$};
    \node[style=obsnode,left=0.5cm of x2] (x1) {$x_1$};
    \node[style=obsnode,right=0.5cm of x2] (x3) {$x_3$};
    \draw[-latex] (h) -- (x1);
    \draw[-latex] (h) -- (x2);
    \draw[-latex] (h) -- (x3);
  }
}

\newcommand<>{\drawdisc}[1]{%
  \uncover#2{
    \point{start-disc}{#1}
    \node[style=node] (h) at (start-disc) {};
    \node[right=0.1cm of h] {$h$};

    \node[style=obsnode,left=0.3cm of h] (x) {};
    \node[left=0.1cm of x] {$x$};

    \node[style=obsnode,below=1cm of h] (y) {};
    \node[left=0.1cm of y] {$y$};
    \draw[-latex] (h) -- (y);
    \draw[-latex] (x) -- (y);
  }
}

\newcommand<>{\drawhmm}[1]  {
  \uncover#2{
% Hidden nodes
   \point{start-hmm}{#1}
   \node[style=node, scale=0.8] (h1) at (start-hmm) {$h_1$};
   \node[style=node, scale=0.8, right= 0.5cm of h1] (h2) {$h_2$};
   \node[style=node, scale=0.8, right= 0.5cm of h2] (h3) {$h_3$};
   \node[scale=0.8, right= 0.5cm of h3] (h4) {$\ldots$};

   \draw[-latex] (h1) -- (h2);
   \draw[-latex] (h2) -- (h3);
   \draw[-latex] (h3) -- (h4);

% Observed nodes
   \node[style=obsnode, scale=0.7, below=0.3cm of h1] (x1) {$x_1$};
   \draw[-latex] (h1) -- (x1);

   \node[style=obsnode, scale=0.7, below=0.3cm of h2] (x2) {$x_2$};
   \draw[-latex] (h2) -- (x2);

   \node[style=obsnode, scale=0.7, below=0.3cm of h3] (x3) {$x_3$};
   \draw[-latex] (h3) -- (x3);
   }
}

\newcommand<>{\drawgrid}[1]  {
  \uncover#2{
    \point{start-grid}{#1}
   \node[style=node, scale=0.8] (h1) at (start-grid) {$h_1$};
   \node[style=node, scale=0.8, below left= 0.5cm of h1] (h2) {$h_2$};
   \node[style=node, scale=0.8, below right= 0.5cm of h1] (h3) {$h_3$};
   \node[style=node, scale=0.8, below right= 0.5cm of h2] (h4) {$h_4$};

   \draw[-latex] (h1) -- (h2);
   \draw[-latex] (h1) -- (h3);
   \draw[-latex] (h2) -- (h4);
   \draw[-latex] (h3) -- (h4);

% Observed nodes
   \node[style=obsnode, scale=0.7, above left=0.3cm of h1] (x1a) {$x^a_1$};
   \node[style=obsnode, scale=0.7, above right=0.3cm of h1] (x1b) {$x^b_1$};
   \draw[-latex] (h1) -- (x1a);
   \draw[-latex] (h1) -- (x1b);

   \node[style=obsnode, scale=0.7, above left=0.3cm of h2] (x2a) {$x^a_2$};
   \node[style=obsnode, scale=0.7, below left=0.3cm of h2] (x2b) {$x^b_2$};
   \draw[-latex] (h2) -- (x2a);
   \draw[-latex] (h2) -- (x2b);

   \node[style=obsnode, scale=0.7, above right=0.3cm of h3] (x3a) {$x^a_3$};
   \node[style=obsnode, scale=0.7, below right=0.3cm of h3] (x3b) {$x^b_3$};
   \draw[-latex] (h3) -- (x3a);
   \draw[-latex] (h3) -- (x3b);

   \node[style=obsnode, scale=0.7, below left=0.3cm of  h4] (x4a) {$x^a_4$};
   \node[style=obsnode, scale=0.7, below right=0.3cm of h4] (x4b) {$x^b_4$};

   \draw[-latex] (h4) -- (x4a);
   \draw[-latex] (h4) -- (x4b);
   }
}

\newcommand<>{\drawtree}[1]  {
\uncover#2{
\point{start-tree}{#1}

% Hidden nodes
   \node[style=node, scale=0.8] (h1) at (0,0) {$h_1$};
   \node[style=node, scale=0.8] (h2) at (-1.3cm,-1cm) {$h_2$};
   \node[style=node, scale=0.8] (h3) at (0,   -1cm) {$h_3$};
   \node[style=node, scale=0.8] (h4) at (1.3cm, -1cm) {$h_4$};

   \draw[-latex] (h1) -- (h2);
   \draw[-latex] (h1) -- (h3);
   \draw[-latex] (h1) -- (h4);

% Observed nodes
  \node[style=obsnode, scale=0.7] (x2a) at (-1.6cm, -2.0cm) {$x^a_2$};
  \node[style=obsnode, scale=0.7] (x2b) at (-1.0cm, -2.0cm) {$x^b_2$};
  \draw[-latex] (h2) -- (x2a);
  \draw[-latex] (h2) -- (x2b);

  \node[style=obsnode, scale=0.7] (x3a) at (-0.3cm, -2.0cm) {$x^a_3$};
  \node[style=obsnode, scale=0.7] (x3b) at (0.3cm, -2.0cm) {$x^b_3$};
  \draw[-latex] (h3) -- (x3a);
  \draw[-latex] (h3) -- (x3b);

  \node[style=obsnode, scale=0.7] (x4a) at (1.0cm, -2.0cm) {$x^a_4$};
  \node[style=obsnode, scale=0.7] (x4b) at (1.6cm, -2.0cm) {$x^b_4$};
  \draw[-latex] (h4) -- (x4a);
  \draw[-latex] (h4) -- (x4b);
  }
}

\newcommand{\tensorfactorization}[1]{%
  \point{start-tf}{#1}
  \tikzcube{tensoring}{black,fill=white}{($(start-tf) + (0,0,0)$)}{1}{1}{1};
  \node at ($(tensoring) + (1.0cm,-0.3cm)$) {$=$};
  \tensorfiber{t1}{fill=blue!70}{($(tensoring) + (2.5cm,0.0cm)$)};
  \node at ($(t1) + (1.0cm,-0.3cm)$) {$+$};
  \tensorfiber{t2}{fill=green!70}{($(t1) + (2.5cm,0.0cm)$)};
  \node at ($(t2) + (1.0cm,-0.3cm)$) {$+ \dots + $};
  \tensorfiber{t3}{fill=red!70}{($(t2) + (3.0cm,0.0cm)$)};

  \draw [decorate,decoration={brace,amplitude=10pt,raise=4pt,mirror},yshift=0pt] 
    ($(t1) + (-1cm,-1cm)$) -- ($(t3) + (0.2cm,-1cm)$) node [below,black,midway,yshift=-0.6cm] {$k$};
}

\newcommand{\matrixfactorization}[1]{%
    \point{start-mf}{#1}
    \tikzrect{mat}{black,fill=white}{($(start-mf) + (0,0)$)}{1}{1};
    \node at ($(mat) + (1.0cm,-0.3cm)$) {$=$};
    \matfiber{t1}{fill=blue!70}{($(mat) + (2.5cm,0.0cm)$)};
    \node at ($(t1) + (1.0cm,-0.3cm)$) {$+$};
    \matfiber{t2}{fill=green!70}{($(t1) + (2.5cm,0.0cm)$)};
    \node at ($(t2) + (1.0cm,-0.3cm)$) {$+ \dots + $};
    \matfiber{t3}{fill=red!70}{($(t2) + (3.0cm,0.0cm)$)};
    \draw [decorate,decoration={brace,amplitude=10pt,raise=4pt,mirror},yshift=0pt] 
      ($(t1) + (-1cm,-1cm)$) -- ($(t3) + (0.2cm,-1cm)$) node [below,black,midway,yshift=-0.6cm] {$k$};
}

\newcommand{\vectorfactorization}[1]{%
    \point{start-vf}{#1}
    \tikzrect{vec}{black,fill=white}{($(start-vf) + (0,0)$)}{0.3}{1};
    \node at ($(vec) + (1.0cm,-0.3cm)$) {$=$};
    \tikzrect{t1}{fill=blue!70}{($($(vec) + (2.5cm,0.0cm)$)$)}{0.3}{1};
    \node at ($(t1) + (1.0cm,-0.3cm)$) {$+$};
    \tikzrect{t2}{fill=green!70}{($($(t1) + (2.5cm,0.0cm)$)$)}{0.3}{1};
    \node at ($(t2) + (1.0cm,-0.3cm)$) {$+ \dots + $};
    \tikzrect{t3}{fill=red!70}{($($(t2) + (3cm,0.0cm)$)$)}{0.3}{1};
    \draw [decorate,decoration={brace,amplitude=10pt,raise=4pt,mirror},yshift=0pt] 
      ($(t1) + (-1cm,-1cm)$) -- ($(t3) + (0.2cm,-1cm)$) node [below,black,midway,yshift=-0.6cm] {$k$};
}

\newcommand{\llhood}[2]{%
  \begin{axis}[
      x=1cm,
      y=3cm,
      scale only axis,
      height=8cm,
      width=4cm,
      axis lines*=left,
      xtick=\empty,
      ytick=\empty,
      xlabel=parameters,
      ylabel=$-$log-likelihood
      ]
    \addplot[
        black,
        thick,
        smooth,
        ] file [% Provide data as a table
          ] {data/llhood.table}
     node[pos=0.27] (em1) {}
     node[pos=0.55] (spec) {}
     node[pos=0.61] (mle) {}
     node[pos=0.85] (em2) {}
     node[pos=0.9] (em2-start) {}
     ;

  \end{axis}
}

\newcommand{\mog}[2]{%
    \begin{axis}[
        xshift=#1,
        yshift=#2,
        scale only axis,
        height=3cm,
        width=3cm,
        axis lines*=left,
        xlabel=$x_1$,
        ylabel=$x_2$,
        xtick=\empty,
        ytick=\empty,
        mark options={scale=0.2,line width=0}
        ]
      \addplot+[
          smooth,
          only marks
          ] file [% Provide data as a table
            ] {data/mog-0.table};
      \addplot+[
          smooth,
          only marks
          ] file [% Provide data as a table
            ] {data/mog-1.table};
      \addplot+[
          smooth,
          only marks
          ] file [% Provide data as a table
            ] {data/mog-2.table}
       ;

    \end{axis}
}

\newcommand{\innerpdiag}[2]{%
  \node[scale=2.0] at ($#1 + (-1.3cm,-0.3cm)$) {$\langle$};
  \node at ($#1!0.5!#2 + (-0.3cm,-0.3cm) $) {$,$};
  \node[scale=2.0] at ($#2 + (0.8cm,-0.3cm)$) {$\rangle$};
}
\newcommand{\innerpdiagv}[2]{%
  \node[scale=2.0] at ($#1 + (-0.7cm,-0.45cm)$) {$\langle$};
  \node at ($#1!0.5!#2 + (-0.1cm,-0.45cm) $) {$,$};
  \node[scale=2.0] at ($#2 + (0.4cm,-0.45cm)$) {$\rangle$};
}
\newcommand{\innerpdiagm}[2]{%
  \node[scale=2.0] at ($#1 + (-1.3cm,-0.45cm)$) {$\langle$};
  \node at ($#1!0.5!#2 + (-0.5cm,-0.45cm) $) {$,$};
  \node[scale=2.0] at ($#2 + (0.4cm,-0.45cm)$) {$\rangle$};
}
\newcommand{\innerpdiagt}[2]{%
  \node[scale=2.0] at ($#1 + (-1.3cm,-0.3cm)$) {$\langle$};
  \node at ($#1!0.5!#2 + (-0.3cm,-0.3cm) $) {$,$};
  \node[scale=2.0] at ($#2 + (0.7cm,-0.3cm)$) {$\rangle$};
}

\newcommand{\regressionA}[1]{%
    \point{start-reg-a}{#1};
    \point{start-reg-a-A}{(start-reg-a)};
    \point{start-reg-a-B}{($(start-reg-a) + (1cm,0) $)};

    \tikzrect{A}{black,fill=yellow}{($(start-reg-a-A) + (0,0)$)}{0.3}{1};
    \tikzrect{B}{black,fill=blue!70}{($(start-reg-a-B) + (0,0)$)}{0.3}{1};
    \innerpdiagv{(start-reg-a-A)}{(start-reg-a-B)};
}

\newcommand{\regressionB}[1]{%
    \point{start-reg-b}{#1};
    \point{start-reg-b-A}{(start-reg-b)};
    \point{start-reg-b-B}{($(start-reg-b) + (1.5cm,0) $)};

    \tikzrect{A}{black,fill=yellow}{($(start-reg-b-A) + (0,0)$)}{1}{1};
    \tikzrect{B}{black,fill=blue!70}{($(start-reg-b-B) + (0,0)$)}{1}{1};
    \innerpdiagm{(start-reg-b-A)}{(start-reg-b-B)};
}

\newcommand{\regressionC}[1]{%
    \point{start-reg-c}{#1};
    \point{start-reg-c-A}{(start-reg-c)};
    \point{start-reg-c-B}{($(start-reg-c) + (1.8cm,0) $)};

    \tikzcube{A}{black,fill=yellow}{($(start-reg-c-A) + (0,0)$)}{1}{1}{1};
    \tikzcube{B}{black,fill=blue!70}{($(start-reg-c-B) + (0,0)$)}{1}{1}{1};
    \innerpdiagt{(start-reg-c-A)}{(start-reg-c-B)};
}

\newcommand{\spectralexpertsdiag}{
  % x, y
  % Nodes
  %\node[style=txt,anchor=south west] (data1) at (1cm,1cm) {$\left\{ x, y \right\}_{(x,y) \in \sD}$};
  \node[style=txt] (data2) at (-4cm,1cm) {$\left\{ x\tp{2}, y^2 \right\}_{(x,y) \in \sD}$};
  \node[style=txt,below=1cm of data2.center] (data3) {$\left\{ x\tp{3}, y^3 \right\}_{(x,y) \in \sD}$};

  %\node[style=txt,anchor=south west] (m1) at (2cm,1cm) {$\E\left[\beta_h\right]$};
  \node[style=txt] (m2) at (1cm,1cm) {$M_2$}; %{$\E\left[\beta_h\tp{2}\right]$};
  \node[style=txt,below=1cm of m2.center] (m3) {$M_3$}; %{$\E\left[\beta_h\tp{3}\right]$};

  \node[below=0.5cm of m2.center] (params-pos) {};
  \node[style=txt,right=3.0cm of params-pos] (params) {$\pi, B$};

  % Arrows
  %\draw[-latex] (data1) -- (m1);
  \draw[-latex] (data2) -- (m2);
  \draw[-latex] (data3) -- (m3);

  % Bias influence
  %\draw[-latex,dashed,gray] (m1.east) -- ++(0.1cm,0) -- ++(0,-0.50cm) -- ++(-3.00cm,0) node (bias-1) {} -- ( bias-1 |- m3) -- (m3.west);

  % Mean
  \draw[-latex] (m2.east) -- ++(0.1cm,0) node (tf-1) {} -- (tf-1 |- params-pos) -- node[above,scale=0.5] {tensor factorization} (params);
  \draw[-latex] (m3.east) -- ++(0.1cm,0) node (tf-2) {} -- (tf-2 |- params-pos) -- (params);

  % Box for regression
  \begin{pgfonlayer}{background}
  %\draw[fill=blue,opacity=0.1,dashed] ($(data1.north east) + (-0.1cm,0.1cm)$) rectangle ($(m3.south west) + (0.1cm,-0.1cm)$);
  \draw[fill=blue,opacity=0.1,dashed] ($(data2.north east) + (-0.1cm,0.1cm)$) rectangle ($(m3.south west) + (0.1cm,-0.1cm)$);
  \end{pgfonlayer}

  % tensor factorization
  \begin{pgfonlayer}{background}
  \draw[fill=green,opacity=0.1,dashed] ($(m2.north east) + (-0.1cm,0.1cm)$) rectangle ($(params.south west) + (0.1cm,-1.0cm)$);
  \end{pgfonlayer}
  \node (params-mid) at ($(params-pos) !.5! (params)$) {};
  \point{reg-label}{($(data3.east) !.5! (m3.west) + (0, -1.0cm)$)};
  \point{tf-label}{(params-mid |- reg-label)};
}

\newcommand{\mkmlrplot}[1]{%
\begin{tikzpicture}
\begin{axis}[ 
    height=5cm, width=5cm, 
    axis lines*=left, 
    xlabel=$x$, ylabel=$y$, 
    xtick=\empty, ytick=\empty, 
    mark options={scale=0.3,line width=0},
    xmin=-1.2, xmax=1.2,
    ]
  #1
\end{axis}
\end{tikzpicture}
}

\newcommand{\mlrfull}{%
  \only<1>{%
    \mkmlrplot{%
       \addplot+[blue,           , mark=none] {0.316 + -0.862*x};
       \addplot+[green,          , mark=none] {-0.715 + -0.268*x};
       \addplot+[red,            , mark=none] {-1.076 + 0.595*x};
     }
  }
  \only<2>{%
    \mkmlrplot{%
       \addplot+[blue,                 mark=none] {0.316 + -0.862*x};
       \addplot+[green,                mark=none] {-0.715 + -0.268*x};
       \addplot+[red,  line width=2pt, mark=none] {-1.076 + 0.595*x};
     }
  }
  \only<3>{%
    \mkmlrplot{%
       \addplot+[blue,                 mark=none] {0.316 + -0.862*x};
       \addplot+[green,                mark=none] {-0.715 + -0.268*x};
       \addplot+[red,  line width=2pt, mark=none] {-1.076 + 0.595*x};
       \addplot[smooth, black, only marks] table {data/mlr-1.table};
     }
  }
  \only<4>{%
    \mkmlrplot{%
       \addplot+[blue, line width=2pt, mark=none] {0.316 + -0.862*x};
       \addplot+[green,                mark=none] {-0.715 + -0.268*x};
       \addplot+[red,                  mark=none] {-1.076 + 0.595*x};
       \addplot[smooth, black, only marks] table {data/mlr-1.table};
     }
  }
  \only<5>{%
    \mkmlrplot{%
       \addplot+[blue, line width=2pt, mark=none] {0.316 + -0.862*x};
       \addplot+[green,                mark=none] {-0.715 + -0.268*x};
       \addplot+[red,                  mark=none] {-1.076 + 0.595*x};
       \addplot[smooth, black, only marks] table {data/mlr-2.table};
     }
  }
  \only<6>{%
    \mkmlrplot{%
       \addplot[smooth, black, only marks] table {data/mlr.table};
       \addplot+[blue, line width=2pt, mark=none] {0.316 + -0.862*x};
       \addplot+[green,line width=2pt, mark=none] {-0.715 + -0.268*x};
       \addplot+[red,  line width=2pt, mark=none] {-1.076 + 0.595*x};
     }
  }
  \only<7>{%
    \mkmlrplot{%
       \addplot[smooth, black, only marks] table {data/mlr.table};
     }
  }
}

### 2014-05-01 doc/icml2014/assumption-proof.tex
Added appendix sections for icml paper

\subsection{Recovering conditional moments}
\label{app:assumption-proof}

% Define up front that we will focus on h, x_1, x_2, x_3
In step 1 of \LearnMarginals, we used the bottleneck property of a hidden
  variable $h_i$ to learn conditional moments $\mOpp{v}{i}$ for every
  view $x_v \in \sV_{h_i}$ using \TensorFactorize. 
In order to do so, we require that \assumptionref{full-rank} holds, i.e.
\begin{assumption*}
  Given a bottleneck $h_1$ with views $x_1, x_2, x_3$, the conditional
  moments $\mOpp{1}{1}, \mOpp{2}{1}, \mOpp{3}{1}$ have full column rank
  $k$, and $\pi\oft 1 \succ 0$.
\end{assumption*}

To interpret this assumption better, we would like find an equivalent
condition on the model parameters $Z_\sC$ that imply the above condition
on $\mOpp{v}{i}$. In this section, we will show that
\assumptionref{full-rank} is actually implied by the following condition
(\assumptionref{full-rank-plus});
\begin{assumption*}
For every clique $\sC \in \sG$ (including ones involving observed
variables), every mode-unfolding of the marginals $Z_\sC$ has full
column rank.
\end{assumption*}

\paragraph{A recursive construction of $\mOpp{v}{i}$}

Without loss of generality, let $i = 1$. We will express the conditional
distribution $\mOpp{v}{1} \eqdef \Pr(x_v | h_1)$ recursively through the
parents of $x_v$, its parents, and so on. The approach is analogous to
message passing, and the key idea is to express the same in the form of
a tensor multiplication. \figureref{message-proof} outlines the
procedure.

\begin{figure}
  \label{fig:message-proof}
  \caption{A recursive construction of the conditional moments $\mOpp{v}{i}$.}
\end{figure}

Recall that $x_v$ is the observed variable in consideration; let $h_t$
be its unique parent. Then, 
\begin{align*}
  \mOpp{v}{1} &\eqdef \Pr( x_v \given h_1 )  \\
              &= \sum_{h_t} \Pr( x_v \given h_t ) \Pr( h_t \given h_1 ) \\
              &= \mOpp{v}{t} \mYpp{t}{1},
\end{align*}
where $\mYpp{i}{j} \eqdef \Pr( h_i \given h_j )$. 
More generally, for two sets of hidden variables $C \eqdef \{h_{C_1}
\cdots h_{C_m} \}$ and $C' \eqdef \{h_{C'_1} \cdots h_{C'_m} \}$, 
define $\mYpp{C}{C'} \eqdef \Pr( h_{C_1} \cdots h_{C_m} \given h_{C'_1}
\cdots h_{C'_n} )$. 
With this representation, $\mYpp{C}{C'}$ is a $|C| + |C'|$-th order
tensor.

Going forward, we will construct $\mYpp{v}{1}$ recursively.
Without loss of generality, let $h_1, h_2, \cdots, h_t$ be a topological
  ordering rooted at $h_1$,
and let $\Pa(h)$ be the parents of a hidden variable $h$ in
  this topological ordering.
Then $\mYpp{t}{1}$ can be expressed as,
\begin{align*}
  \mYpp{t}{1} &\eqdef \Pr( h_t \given h_1 )  \\
  &= \sum_{\vh \in H_{\Pa(h_t)}} \Pr( \Pa(h_t) \given h_1 ) \Pr( h_t \given \Pa(h_t) ) \\
  &= \mYpp{ \Pa(h_t) }{1} \times_{\Pa(h_t)} \mYpp{t}{\Pa(h_t)},
\end{align*}
where $A \times_{C} B$ refers to summation along the indices $C$. We
refer to the operation $A \times_C B$ as tensor multiplication.
\appendixref{tensor-multiplication} prooves some key properties of this
operation. 
Note that $\mYpp{t}{\Pa(h_t)}$ is the conditional probability $\Pr(h_t
 | \Pa(h_t))$, which can be easily computed from the hidden marginals
 $Z_\sC$ that contain $h_t$. 
We will describe this procedure in further detail soon.

Now, let's further expand $\mYpp{ \Pa(h_t) }{1}$ in terms of {\em its
  parents}, recursively doing so until we reach $h_1$.
Let $C$ be an intermediate set of hidden variables. Let $h_c \neq h_1$
  be some hidden variable in $C$ which we wish to eliminate. 
Let $C' = C \union \Pa(h_c) \setminus \{h_c\}$ and $\del C = C'
  \setminus (C \union \{h_1\})$, the variables in the interface between $C$ and $C'$.
Then,
\begin{align*}
  \mYpp{C}{1} &\eqdef \Pr( h_C \given h_1 )  \\
  &= \sum_{\vh \in H_{\del C}} \Pr( C' \given h_1 ) \Pr( h_c \given \Pa(h_c) ) \\
  &= \mYpp{ C' }{ 1 } \times_{\del C} \mYpp{c}{\Pa(h_c)},
\end{align*}
Finally, for the base case, $\mYpp{1}{1} = \ones$.
This procedure is algorithmically described in \algorithmref{Y}.

\begin{algorithm}
  \caption{$\mYpp{C}{1}$}
  \label{algo:Y}
  \begin{algorithmic}
    \REQUIRE The root $h_1$, a set of hidden variables $C$.
    \ENSURE The hidden moments distribution $\mYpp{C}{1}$.
    \IF{ $C = \{h_1\}$ }
      \STATE $\mYpp{1}{1} = \ones$.
    \ELSE
      \STATE Let $h_c \neq h_1$ be some hidden variable in $C$.
      \STATE Let $C' = C \union \Pa(h_c) \setminus \{h_c\}$.
      \STATE $\mYpp{C}{1} = \mYpp{C'}{1} \times_{\Pa(h_c)} \mYpp{c}{\Pa(h_c)}$.
    \ENDIF
  \end{algorithmic}
\end{algorithm}

\paragraph{Constructing $\mYpp{c}{\Pa(h_c)}$}

Note that in the process of the algorithm, $\mYpp{t}{1}$ is expressed
entirely as tensor multiplications with terms of the form
$\mYpp{c}{\Pa(h_c)}$ for some hidden variable $h_c$.  Let's finally
derive a concrete expression for $\mYpp{c}{\Pa(h_c)}$ from the
parameters, the hidden marginals $Z_\sC$.

Without loss of generality, let $\sC \eqdef \{c\} \union P$ be a clique
  in the graph $\sG$.
If instead $\sC \subseteq \sC_1 \union \sC_2$ for some $\sC_1, \sC_2 \in
  \sG$, we can simply construct $Z_{\sC}$ from the union of $\sC_1$ and
  $\sC_2$,
\begin{align*}
  Z_\sC &= Z_{\sC_1} \times_{\del \sC} Z_{\sC_2},
\end{align*}
where $\del \sC = (\sC_1 \union \sC_2) \setminus \sC$.

Proceeding,
\begin{align*}
\mYpp{c}{\Pa(h_c)} 
  &\eqdef \Pr(h_c \given \Pa(h_c)) \\
  &= \frac{\Pr(h_c, \Pa(h_c))}{\Pr(\Pa(h_c))} \\
  &= Z_\sC \diag{Z_\sC(\ones, \cdot, \ldots, \cdot)}^{-1}.
\end{align*}

\begin{lemma}[Sufficiency of \assumptionref{full-rank-plus}]
  \label{lem:full-rank-suff}
  Given that \assumptionref{full-rank-plus} holds, then for any hidden
  variable $h_1$ and observed variable $x_v$, $\mPi{1} \succ 0$ and
  $\mOpp{v}{i}$ has full column rank.
\end{lemma}
\begin{proof}
  The important statement that needs to be made is how $\sigma_k(Y)$
  relates to its parents.
\end{proof}

### 2014-05-01 doc/icml2014/tensor-multiplication.tex
Added appendix sections for icml paper

\subsection{Tensor Multiplication}
\label{app:tensor-multiplication}

In this section, we present basic results regarding tensor
multiplication. First, let us review definitions.

Let $\sP \in \Re^{d_1 \times \ldots \times d_m}$ be a $m$-th order tensor and let $\sQ \in \Re^{d'_1 \times \ldots \times d'_n}$ be an $n$-th order tensor, where wlog, $m \ge n$. 
Let $M$ be an index matching set $M = \{ (a_1; b_1),
  (a_2, b_2), \ldots, (a_\ell, b_\ell) \}$, such that the corresopnding
  dimensions agree, i.e. $d_{a_i} = d'_{b_i}$ for all $i$. Let $D_{\ba}
  = [d_{a_1}] \times \cdots \times [d_{a_\ell}]$ and $D'_{\bb}
  = [d'_{a_1}] \times \cdots \times [d'_{a_\ell}]$ be the domains of the
  corresponding indices.
We define the tensor multiplication operation $\sR \eqdef P \times_{M}
  Q$ to be the $(m - \ell + n - \ell)$-th order tensor formed by
  conjoining $\sP$ and $\sR$ followed by summing out the indices in $M$:
\begin{align*}
  \sR[\vi,\vj] &\eqdef \sum_{\bu, \bv \in D_\ba \times D'_\bb} \sP[\vi,\bu] \sQ[\bv ,\vj] ~ \delta_{u_1 v_1} \delta_{u_2 v_2} \ldots \delta_{u_\ell v_\ell}.
\end{align*}

Note that when $m, n = 2$, i.e. $\sP$ and $\sQ$ are matrices, and $M
  = {(2,1)}$, the operation is equivalent to matrix multiplication. 
Furthermore, similar to matrix multiplication, the operator defined above
  is associative and distributive. 
It is also commutative for symmetric tensors.

Given an index set $I \subset [m]$ we
  define an unfolding of a tensor $\sP$ along $I$,
 $\sP\munf{I}$, to be a matrix of dimension $(\prod_{i \in I} d_i)
 \times (\prod_{j \not\in I} d_j)$. $\sP\munf{I}$ has elements
 $\sP\munf{I}[\vi;\vj] = \sP[\pi(\vi\vj)]$, where $\pi$ is an
 appropriate permutation to align the indices of $\vi$ and $\vj$ with
 the ordering of $I$. 
For example, if $\sP$ were a $4$-th order tensor and $I = \{1,3\}$,
  $\sP\munf{13}[i_1,i_2; j_1,j_2] = \sP[i_1,j_1,i_2,j_2]$.
For notational convenience, we will omit the permutation $\pi$ subsequently.

The main result of this section is, similar to matrix multiplication, if
$\sR = \sP \times_{M} \sQ$, then the singular values of mode-unfolding
of $\sR$ are bounded by those of $\sP$ and $\sQ$.

\begin{lemma}[Tensor Multiplication and Singular Values]
  \label{lem:tensor-multiplication}
Let $\sP, \sQ$ be $n$-th order and $m$-th order tensors, $M$ be an
  admissible index mapping and $\sR = \sP \times_M \sQ$. 
Let $I \subset [m - k + n - k]$ be an arbitrary subset of the index set
  of $\sR$,
and let $I_1$, $I_2$ be the respective subsets of $I$ contained in $\sP$
  and $\sQ$.
Then,
\begin{align*}
\sigma_{\max}(\sR\munf{I}) &\le \sigma_{\max}(\sP\munf{I_1}) \sigma_{\max}(\sQ\munf{I_2}) \\
\sigma_{\min}(\sR\munf{I}) &\ge \sigma_{\min}(\sP\munf{I_1}) \sigma_{\min}(\sQ\munf{I_2}).
\end{align*}
\end{lemma}

Note that in the matrix case, when $m = n = 2$, when $R = P Q^\top$, we get the standard result that 
\begin{align}
\sigma_{\max}(R) &\le \sigma_{\max}(P) \sigma_{\max}(Q) \\
\sigma_{\min}(R) &\ge \sigma_{\min}(P) \sigma_{\min}(Q) \label{eqn:matrix-singular}.
\end{align}

\begin{proof}
  Without loss of generality, let $M = \{(1,1), \ldots, (\ell,\ell)\}$,
  and $D = [d_1] \times \cdots \times [d_\ell]$. 
  Then,
  \begin{align*}
    \sR[\vi\vj] &\eqdef \sum_{\vk \in D_\ell} \sP[\vk \vi] \sQ[\vk \vj].
  \end{align*}

Consider an $(I)$-unfolding of $\sR$,
  \begin{align*}
    \sR\munf{I}[\vi,\vj] 
       &= \sum_{\vk} \sP[\vk \vi] \sQ[\vk \vj] \\
       &= \sum_{\vk} \sP\munf{I_1}[\vx_1; \vy_1, \vk] \sQ\munf{I_2}[\vx_2; \vy_2, \vk],
  \end{align*}
  where $\vx_1 \in \Range(I_1)$, $\vy_1 \in \Range([m] \setminus I_1)$, $\vx_2 \in \Range(I_2)$ and $\vy_2  \in \Range([n] \setminus I_2)$.

The key idea is to represent the above as a matrix multiplication.
To do so, we can temporarily ``inflate'' the unfoldings $\sP\munf{I_1}$
  and $\sQ\munf{I_2}$ by taking the outer product of $\sP\munf{I_1}$ with
  $|I_2|$ identity matrices, and vice-versa. Define 
\begin{align*}
  \sP' &\eqdef \sP \otimes I_{d'_1} \otimes \cdots \otimes I_{d'_{|I_2|}} \\
  \sQ' &\eqdef \sQ \otimes I_{d'_1} \otimes \cdots \otimes I_{d'_{|I_1|}}.
\end{align*}

Now, $\sP'$ is a $(m + 2 * |I_2|)$-th order tensor and similarly $\sQ'$
  is a $(n + 2 * |I_1|)$-th order tensor. We choose new index sets
  $I'_1$ and $I'_2$ to include exactly one of these identity indices. 
Then,
\begin{align*}
    \sR\munf{I}[\vi,\vj] 
    &= \sum_{\vy_1, \vy_2 \vk}
    {\sP'}\munf{I'_1}[\vx_1\vy'_2; \vy_1 \vy_2 \vk]  
    {\sQ'}\munf{I'_2}[\vx_2\vy'_1; \vy_1 \vy_2 \vk] \\
    \sR\munf{I} &= {\sP'}\munf{I'_1} {\sQ'}\munf{I'_2}.
\end{align*}

Using lemma \lemmaref{tensor-prod} and the bound on matrix singular
values, \equationref{matrix-singular}, we get,
\begin{align*}
\sigma_{\max}(\sR\munf{I}) 
    &\le \sigma_{\max}({\sP'}\munf{I'_1}) \sigma_{\max}({\sQ'}\munf{I'_2}) \\
      &= \sigma_{\max}(\sP\munf{I_1}) \sigma_{\max}(\sQ\munf{I_2}) \\
    \sigma_{\min}(\sR\munf{I})  
    &\ge \sigma_{\min}({\sP'}\munf{I'_1}) \sigma_{\min}({\sQ'}\munf{I'_2}) \\
      &= \sigma_{\min}(\sP\munf{I_1}) \sigma_{\min}(\sQ\munf{I_2}).
\end{align*}
\end{proof}

\begin{lemma}[Singular values of the tensor products of matrices]
  \label{lem:tensor-prod}
  Let $A = \Un1 \Sigma_1 \Vnt1$ and $B = \Un2 \Sigma_2 \Vnt2$ be two
  matrices and their singular value decompositions, 
  with $\Sigma_1 = \diag(\sigma_{11}, \ldots, \sigma_{1m})$ and $\Sigma_2
  = \diag(\sigma_{21}, \ldots, \sigma_{2n})$. 
  Then, for $C = A \otimes B$, the matrix unfolding $C\munf{13}$ is full
  rank, with singular values $\sigma_{11} \sigma_{21}, \ldots, \sigma_{1m}
  \sigma_{2n}$.
\end{lemma}
\begin{proof}
  An alternative representation of $A$ and $B$ (and consequently $C$) is as follows
  \begin{align*}
    A &= \sum_{i=1}^m \sigma_{1i}~ \un{1}_i \otimes \vn{1}_i \\
    B &= \sum_{j=1}^n \sigma_{2j}~ \un{2}_j \otimes \vn{2}_j \\
    C^{13} &= \sum_{i=1}^m \sum_{j=1}^n \sigma_{1i} \sigma_{2j}~ (\un{1}_i \otimes \un{2}_j)(\vn{1}_i \otimes \vn{2}_j)^\top.
  \end{align*}

  For the unfolding $\{1,3\}$ which pairs $\un{1}$ and $\un{2}$, we have
basis elements $\un{1} \otimes \un{2}$ that are orthogonal to each
other, and thus the above is a valid singular value decomposition,
with singular values $\{ \sigma_{1i} \sigma_{2j} \}_{i=1,j=1}^{i=m,j=m}$. 
\end{proof}

### 2014-05-01 doc/icml2014/icml2014-presentation/diagrams.tex
Updated presentation

@@
 % Hidden nodes
-   \node[style=node, scale=0.8] (h1) at (0,0) {$h_1$};
-   \node[style=node, scale=0.8] (h2) at (-1.3cm,-1cm) {$h_2$};
-   \node[style=node, scale=0.8] (h3) at (0,   -1cm) {$h_3$};
-   \node[style=node, scale=0.8] (h4) at (1.3cm, -1cm) {$h_4$};
+   \node[style=node, scale=0.8] (h1) at ($(start-tree) + (0,0)$) {$h_1$};
+   \node[style=node, scale=0.8] (h2) at ($(start-tree) + (-1.3cm,-1cm)$) {$h_2$};
+   \node[style=node, scale=0.8] (h3) at ($(start-tree) + (0,   -1cm)$) {$h_3$};
+   \node[style=node, scale=0.8] (h4) at ($(start-tree) + (1.3cm, -1cm)$) {$h_4$};
 
@@
 % Observed nodes
-  \node[style=obsnode, scale=0.7] (x2a) at (-1.6cm, -2.0cm) {$x^a_2$};
-  \node[style=obsnode, scale=0.7] (x2b) at (-1.0cm, -2.0cm) {$x^b_2$};
+  \node[style=obsnode, scale=0.7] (x2a) at ($(start-tree) + (-1.6cm, -2.0cm)$) {$x^a_2$};
+  \node[style=obsnode, scale=0.7] (x2b) at ($(start-tree) + (-1.0cm, -2.0cm)$) {$x^b_2$};
   \draw[-latex] (h2) -- (x2a);
@@
 
-  \node[style=obsnode, scale=0.7] (x3a) at (-0.3cm, -2.0cm) {$x^a_3$};
-  \node[style=obsnode, scale=0.7] (x3b) at (0.3cm, -2.0cm) {$x^b_3$};
+  \node[style=obsnode, scale=0.7] (x3a) at ($(start-tree) + (-0.3cm, -2.0cm)$) {$x^a_3$};
+  \node[style=obsnode, scale=0.7] (x3b) at ($(start-tree) + (0.3cm, -2.0cm)$) {$x^b_3$};
   \draw[-latex] (h3) -- (x3a);
@@
 
-  \node[style=obsnode, scale=0.7] (x4a) at (1.0cm, -2.0cm) {$x^a_4$};
-  \node[style=obsnode, scale=0.7] (x4b) at (1.6cm, -2.0cm) {$x^b_4$};
+  \node[style=obsnode, scale=0.7] (x4a) at ($(start-tree) + (1.0cm, -2.0cm)$) {$x^a_4$};
+  \node[style=obsnode, scale=0.7] (x4b) at ($(start-tree) + (1.6cm, -2.0cm)$) {$x^b_4$};
   \draw[-latex] (h4) -- (x4a);
   \draw[-latex] (h4) -- (x4b);
+  }
+}
+
+\newcommand<>{\drawnoisyor}[1]  {
+\uncover#2{
+\point{start-nor}{#1}
+\node[style=node, scale=0.8] (h1) at ($(start-nor) + (-0.5,0)$) {$h_1$};
+   \node[style=node, scale=0.8] (h2) at ($(start-nor) + (+0.5,0)$) {$h_2$};
+   \node[style=obsnode, scale=0.8] (x1) at ($(start-nor) + (-2.0, -1.0)$) {$x_1$};
+   \node[style=obsnode, scale=0.8] (x2) at ($(start-nor) + (-1.0, -1.0)$) {$x_2$};
+   \node[style=obsnode, scale=0.8] (x3) at ($(start-nor) + (0.0, -1.0) $){$x_3$};
+   \node[style=obsnode, scale=0.8] (x4) at ($(start-nor) + (1.0, -1.0) $){$x_4$};
+   \node[style=obsnode, scale=0.8] (x5) at ($(start-nor) + (2.0, -1.0) $){$x_5$};
+
+   \draw[-latex] (h1) -- (x1);
+   \draw[-latex] (h1) -- (x2);
+   \draw[-latex] (h1) -- (x3);
+
+   \draw[-latex] (h2) -- (x2);
+   \draw[-latex] (h2) -- (x3);
+   \draw[-latex] (h2) -- (x4);
+   \draw[-latex] (h2) -- (x5);
    }

### 2014-05-06 doc/icml2014/tensor-multiplication.tex

@@
-\subsection{Tensor Multiplication}
+\subsection{Tensor Operations}
 \label{app:tensor-multiplication}
 
+%In this section, we present basic results regarding two tensor
+%operations we define, tensor multiplication, which is analogous to
+%matrix multiplication, and the diagonal tensor.
 In this section, we present basic results regarding tensor
-multiplication. First, let us review definitions.
+multiplication, an operation analogous to matrix multiplication.
+First, let us define some notation.
 
-Let $\sP \in \Re^{d_1 \times \ldots \times d_m}$ be a $m$-th order tensor and let $\sQ \in \Re^{d'_1 \times \ldots \times d'_n}$ be an $n$-th order tensor, where wlog, $m \ge n$. 
-Let $M$ be an index matching set $M = \{ (a_1; b_1),
+In the following, let $\sP \in \Re^{d_1 \times \ldots \times d_m}$ be
+  a $m$-th order tensor and let $\sQ \in \Re^{d'_1 \times \ldots \times
+  d'_n}$ be an $n$-th order tensor. %, where wlog, $m \ge n$. 
+We use the notation $\Domain(\sP) = [d_1] \times \cdots \times [d_m]$
+  to denote the domain of the indices of $\sP$.
+Let $M$ be an index matching set, i.e. $M = \{ (a_1; b_1),
   (a_2, b_2), \ldots, (a_\ell, b_\ell) \}$, such that the corresopnding
@@
   = [d_{a_1}] \times \cdots \times [d_{a_\ell}]$ and $D'_{\bb}
-  = [d'_{a_1}] \times \cdots \times [d'_{a_\ell}]$ be the domains of the
+  = [d'_{b_1}] \times \cdots \times [d'_{b_\ell}]$ be the domains of the
   corresponding indices.
+
+Finally, for some $\vi \in [d_1] \times \cdots \times [d_\ell]$ and $\vj
+\in [d_{\ell + 1}] \times \cdots \times [d_m]$ and  we use the notation
+$\sP[\vi \cdot \vj]_\pi \eqdef \sP[\pi(i_1, \cdot, i_\ell, j_{\ell+1},
+\cdots j_m)]$ to conjoin indices $\vi$ and $\vj$ up to the permutation
+$\pi$. 
+%For our purposes, the specific ordering of the tensor's indices
+%are not important and are usually referred to by variable naming.
+
+% \begin{definition}[Diagonal Tensor]
+%   The diagonalized tensor $\diag(\sP)$ is a $(2m)$-th order tensor
+%   constructed by taking the tensor product along $I^{\otimes p}$,
+%   \begin{align*}
+%     \diag(\sP)[\vi,\vj] &\eqdef \sP[\vi] \delta[\vi,\vj],
+%   \end{align*}
+%   where $\vi, \vj \in [d_1] \times \cdots \times [d_m]$ and $\delta(\vi,
+%   \vj) = \BI[\vi = \vj]$ is the Kronecker delta function.
+% \end{definition}
+% 
+% Note that when $\sP$ is a vector, the diagonal tensor is a diagonal
+% matrix, with entries given by the vector.
+
+\begin{definition}[Tensor Multiplication]
   We define the tensor multiplication operation $\sR \eqdef P \times_{M}
-  Q$ to be the $(m - \ell + n - \ell)$-th order tensor formed by
-  conjoining $\sP$ and $\sR$ followed by summing out the indices in $M$:
+  Q$ to be the $(m - \ell + n - \ell)$-th order tensor $\sR$ formed by
+  conjoining $\sP$ and $\sQ$ followed by summing out the indices in $M$:
 \begin{align*}
-  \sR[\vi,\vj] &\eqdef \sum_{\bu, \bv \in D_\ba \times D'_\bb} \sP[\vi,\bu] \sQ[\bv ,\vj] ~ \delta_{u_1 v_1} \delta_{u_2 v_2} \ldots \delta_{u_\ell v_\ell}.
+  \sR[\vi\cdot\vj] &\eqdef \sum_{\mathclap{\bu, \bv \in D_\ba \times D'_\bb}} \sP[\vi \cdot \bu]_{\pi_P} \sQ[\vj \cdot \bv]_{\pi_Q} ~ \delta_{u_1 v_1} \delta_{u_2 v_2} \ldots \delta_{u_\ell v_\ell},
 \end{align*}
+  for every $\vi \in \Domain(\sP) \setminus D_a$ and $\vj \in
+  \Domain(\sQ) \setminus D'_b$, where $\pi_P$ and $\pi_Q$ are
+  permutations of the indices that respect the matching $M$. For
+  notational convenience, we will omit $\pi_P$, $\pi_Q$ in the sequel.
+\end{definition}
 
 Note that when $m, n = 2$, i.e. $\sP$ and $\sQ$ are matrices, and $M
-  = {(2,1)}$, the operation is equivalent to matrix multiplication. 
+  = \{(2,1)\}$, the operation is equivalent to matrix multiplication. 
 Furthermore, similar to matrix multiplication, the operator defined above
@@
 
-Given an index set $I \subset [m]$ we
+\paragraph{Unfoldings}
+For a tensor $\sP$, as defined above, 
+  given an index set $I \subset [m]$ we
   define an unfolding of a tensor $\sP$ along $I$,
- $\sP\munf{I}$, to be a matrix of dimension $(\prod_{i \in I} d_i)
+ $\sP\munf{I}$, to be a matrix of dimensions $(\prod_{i \in I} d_i)
  \times (\prod_{j \not\in I} d_j)$. $\sP\munf{I}$ has elements
- $\sP\munf{I}[\vi;\vj] = \sP[\pi(\vi\vj)]$, where $\pi$ is an
+ $\sP\munf{I}[\vi;\vj] = \sP[\vi\cdot\vj]_{\pi}$, where $\pi$ is an
  appropriate permutation to align the indices of $\vi$ and $\vj$ with
@@
 The main result of this section is, similar to matrix multiplication, if
-$\sR = \sP \times_{M} \sQ$, then the singular values of mode-unfolding
+$\sR = \sP \times_{M} \sQ$, then the singular values of unfoldings
 of $\sR$ are bounded by those of $\sP$ and $\sQ$.
@@
   \label{lem:tensor-multiplication}
-Let $\sP, \sQ$ be $n$-th order and $m$-th order tensors, $M$ be an
-  admissible index mapping and $\sR = \sP \times_M \sQ$. 
-Let $I \subset [m - k + n - k]$ be an arbitrary subset of the index set
+Let $\sP, \sQ$ be $m$-th order and $n$-th order tensors respectively, $M$ be an
+  index mapping and $\sR = \sP \times_M \sQ$. 
+Let $I \subset [m - k + n - k]$ be any arbitrary subset of the index set
   of $\sR$,
-and let $I_1$, $I_2$ be the respective subsets of $I$ contained in $\sP$
-  and $\sQ$.
+and let $I_P$, $I_Q$ be the partitioning of $I$ into subsets contained
+  in $\sP$ and $\sQ$ respectively.
 Then,
 \begin{align*}
-\sigma_{\max}(\sR\munf{I}) &\le \sigma_{\max}(\sP\munf{I_1}) \sigma_{\max}(\sQ\munf{I_2}) \\
-\sigma_{\min}(\sR\munf{I}) &\ge \sigma_{\min}(\sP\munf{I_1}) \sigma_{\min}(\sQ\munf{I_2}).
+\sigma_{\max}(\sR\munf{I}) &\le \sigma_{\max}(\sP\munf{I_P}) \sigma_{\max}(\sQ\munf{I_P}) \\
+\sigma_{\min}(\sR\munf{I}) &\ge \sigma_{\min}(\sP\munf{I_Q}) \sigma_{\min}(\sQ\munf{I_Q}).
 \end{align*}
@@
 
-Note that in the matrix case, when $m = n = 2$, when $R = P Q^\top$, we get the standard result that 
+Note that in the matrix case, when $m = n = 2$ and $M=\{(2,1)\}$, i.e.
+when $R = P Q$, we get the standard result that 
 \begin{align}
@@
   Without loss of generality, let $M = \{(1,1), \ldots, (\ell,\ell)\}$,
-  and $D = [d_1] \times \cdots \times [d_\ell]$. 
+  and $D_\ell = [d_1] \times \cdots \times [d_\ell]$. 
   Then, 
   \begin{align*}
-    \sR[\vi\vj] &\eqdef \sum_{\vk \in D_\ell} \sP[\vk \vi] \sQ[\vk \vj].
+    \sR[\vi\cdot\vj] &\eqdef \sum_{\vk \in D_\ell} \sP[\vk\cdot\vi_P\cdot\vj_P] \sQ[\vk\cdot\vi_Q\cdot\vj_Q],
   \end{align*}
+  where $\vi_P, \vi_Q$, are the components of $\vi$ in $\sP$ and $\sQ$
+  respectively and $\vj_P, \vj_Q$ is similarly defined.
 
@@
   \begin{align*}
-    \sR\munf{I}[\vi,\vj] 
-       &= \sum_{\vk} \sP[\vk \vi] \sQ[\vk \vj] \\
-       &= \sum_{\vk} \sP\munf{I_1}[\vx_1; \vy_1, \vk] \sQ\munf{I_2}[\vx_2; \vy_2, \vk],
+    \sR\munf{I}[\vi;\vj] 
+       &= \sum_{\vk \in D_\ell} \sP[\vk\cdot \vi_P \cdot \vj_P] \sQ[\vk\cdot \vi_Q \cdot \vj_Q] \\
+       &= \sum_{\vk \in D_\ell} \sP\munf{I_P}[\vi_P; \vj_P \cdot \vk] \sQ\munf{I_Q}[\vj_Q; \vj_Q \cdot \vk].
   \end{align*}
-  where $\vx_1 \in \Range(I_1)$, $\vy_1 \in \Range([m] \setminus I_1)$, $\vx_2 \in \Range(I_2)$ and $\vy_2  \in \Range([n] \setminus I_2)$.
+Note that the indices that were summed out belong in the complement of $I_P$ and $I_Q$ respectively.
 
-The key idea is to represent the above as a matrix multiplication.
-To do so, we can temporarily ``inflate'' the unfoldings $\sP\munf{I_1}$
-  and $\sQ\munf{I_2}$ by taking the outer product of $\sP\munf{I_1}$ with
-  $|I_2|$ identity matrices, and vice-versa. Define 
+The key idea is to represent the above expression as matrix multiplication.
+To do so, we can temporarily ``inflate'' the unfoldings $\sP\munf{I_P}$
+  and $\sQ\munf{I_Q}$ by taking the outer product of $\sP\munf{I_P}$ with
+  $|I_Q|$ identity matrices, and vice-versa. Define 
 \begin{align*}
-  \sP' &\eqdef \sP \otimes I_{d'_1} \otimes \cdots \otimes I_{d'_{|I_2|}} \\
-  \sQ' &\eqdef \sQ \otimes I_{d'_1} \otimes \cdots \otimes I_{d'_{|I_1|}}.
+  \sP' &\eqdef \sP \otimes I_{d'_1} \otimes \cdots \otimes I_{d'_{|I_Q|}} \\
+  \sQ' &\eqdef \sQ \otimes I_{d'_1} \otimes \cdots \otimes I_{d'_{|I_P|}}.
 \end{align*}
 
-Now, $\sP'$ is a $(m + 2 * |I_2|)$-th order tensor and similarly $\sQ'$
-  is a $(n + 2 * |I_1|)$-th order tensor. We choose new index sets
-  $I'_1$ and $I'_2$ to include exactly one of these identity indices. 
+Now, $\sP'$ is a $(m + 2 |I_Q|)$-th order tensor and similarly $\sQ'$
+  is a $(n + 2 |I_P|)$-th order tensor. 
+For every identity matrix we tensor $\sP$ and $\sQ$ by we include one of
+  its indices to include in the new set $I'_P$ and $I'_Q$ respectively.
+  Note that $\sP\munf{I'_P}$ is a $(I_P + I_Q)\times (m - |I_P| + |I_Q| + \ell)$ matrix, and 
+$\sP\munf{I'_Q}$ is $(I_P + I_Q) \times (n - |I_Q| + |I_P| + \ell)$
 Then,
 \begin{align*}
-    \sR\munf{I}[\vi,\vj] 
-    &= \sum_{\vy_1, \vy_2 \vk}
-    {\sP'}\munf{I'_1}[\vx_1\vy'_2; \vy_1 \vy_2 \vk]  
-    {\sQ'}\munf{I'_2}[\vx_2\vy'_1; \vy_1 \vy_2 \vk] \\
-    \sR\munf{I} &= {\sP'}\munf{I'_1} {\sQ'}\munf{I'_2}.
+    \sR\munf{I}[\vi;\vj] 
+    &= \sum_{\be_P, \be_Q \vk}
+    {\sP'}\munf{I'_P}[\vi_P\be'_Q; \vj_P \be_Q \vk]  
+    {\sQ'}\munf{I'_Q}[\vi_Q\be'_P; \vj_Q \be_P \vk] \\
+    \sR\munf{I} &= {\sP'}\munf{I'_P} {\sQ'}\munf{I'_Q}.
 \end{align*}
@@
 \sigma_{\max}(\sR\munf{I}) 
-    &\le \sigma_{\max}({\sP'}\munf{I'_1}) \sigma_{\max}({\sQ'}\munf{I'_2}) \\
-      &= \sigma_{\max}(\sP\munf{I_1}) \sigma_{\max}(\sQ\munf{I_2}) \\
+    &\le \sigma_{\max}({\sP'}\munf{I'_P}) \sigma_{\max}({\sQ'}\munf{I'_Q}) \\
+      &= \sigma_{\max}(\sP\munf{I_P}) \sigma_{\max}(\sQ\munf{I_Q}) \\
     \sigma_{\min}(\sR\munf{I})  
-    &\ge \sigma_{\min}({\sP'}\munf{I'_1}) \sigma_{\min}({\sQ'}\munf{I'_2}) \\
-      &= \sigma_{\min}(\sP\munf{I_1}) \sigma_{\min}(\sQ\munf{I_2}).
+    &\ge \sigma_{\min}({\sP'}\munf{I'_P}) \sigma_{\min}({\sQ'}\munf{I'_Q}) \\
+      &= \sigma_{\min}(\sP\munf{I_P}) \sigma_{\min}(\sQ\munf{I_Q}).
 \end{align*}
 \end{proof}
+
+For completeness, we also provide a proof for the following standard
+lemma regarding the tensor product of matrices.
 

### 2014-05-06 doc/icml2014/assumption-proof.tex
Complete tensor proofs

@@
 In step 1 of \LearnMarginals, we used the bottleneck property of a hidden
-  variable $h_i$ to learn conditional moments $\mOpp{v}{i}$ for every
-  view $x_v \in \sV_{h_i}$ using \TensorFactorize. 
+  variable $h_i$ to learn conditional moments $\mOpp{v}{i} \eqdef
+  \Pr(x_v | h_i)$ for every view $x_v \in \sV_{h_i}$ using
+  \TensorFactorize. 
 In order to do so, we require that \assumptionref{full-rank} holds, i.e.
@@
 For every clique $\sC \in \sG$ (including ones involving observed
-variables), every mode-unfolding of the marginals $Z_\sC$ has full
-column rank.
+  variables), for every {\em conditioning} of the marginals, every
+  unfolding has full rank. 
 \end{assumption*}
@@
 \begin{figure}
+  \caption{A recursive construction of the conditional moments $\mOpp{v}{i}$.}
   \label{fig:message-proof}
-  \caption{A recursive construction of the conditional moments $\mOpp{v}{i}$.}
 \end{figure}
@@
 Note that $\mYpp{t}{\Pa(h_t)}$ is the conditional probability $\Pr(h_t
- | \Pa(h_t))$, which can be easily computed from the hidden marginals
- $Z_\sC$ that contain $h_t$. 
-We will describe this procedure in further detail soon.
+ | \Pa(h_t))$, which is assumed to be full rank, and can be easily
+ computed from the hidden marginals $Z_\sC$ that contain $h_t$.
 
@@
   be some hidden variable in $C$ which we wish to eliminate. 
-Let $C' = C \union \Pa(h_c) \setminus \{h_c\}$ and $\del C = C'
-  \setminus (C \union \{h_1\})$, the variables in the interface between $C$ and $C'$.
+Let $C' = C \union \Pa(h_c) \setminus \{h_c\}$, the new set of hidden
+  variables containing the parents of $h_c$ instead of $h_c$ and $\del
+  C = C' \setminus (C \union \{h_1\})$, the variables in the interface
+  between $C$ and $C'$.
 Then,
@@
 entirely as tensor multiplications with terms of the form
-$\mYpp{c}{\Pa(h_c)}$ for some hidden variable $h_c$.  Let's finally
-derive a concrete expression for $\mYpp{c}{\Pa(h_c)}$ from the
-parameters, the hidden marginals $Z_\sC$.
+$\mYpp{c}{\Pa(h_c)}$ for some hidden variable $h_c$. 
+$\mYpp{c}{\Pa(h_c)}$ can be easily computed from the clique marginals
+$Z_\sC$.
 
-Without loss of generality, let $\sC \eqdef \{c\} \union P$ be a clique
+Without loss of generality, let $\sC \eqdef \{c\} \union \Pa(h_c)$ be a clique
   in the graph $\sG$.
@@
   &= \frac{\Pr(h_c, \Pa(h_c))}{\Pr(\Pa(h_c))} \\
-  &= Z_\sC \diag{Z_\sC(\ones, \cdot, \ldots, \cdot)}^{-1}.
+  &= Z_\sC \times_{\Pa(h_c)} \diag{Z_\sC(\ones, \cdot, \ldots, \cdot)}^{-1},
 \end{align*}
+where $\diag(\sP)$ is a ``diagonal'' tensor constructed by taking the
+tensor product with $I^{\otimes p}$,
+\begin{align*}
+  \diag(\sP)[\vi,\vj] &\eqdef \sP[\vi] \delta[\vi,\vj].
+\end{align*}
+
+\paragraph{Sample complexity results}
+
+From the algorithm described above, blah.
 
@@
 \begin{proof}
-  The important statement that needs to be made is how $\sigma_k(Y)$
-  relates to its parents.
+
+  - $Y$
+  - Use theorem of tensor multiplication.
 \end{proof}
+
+The sample complexity of estimating $\mOpp{v}{i}$ depends mainly on
+$\sigma_{\min}(\mOpp{v}{i})$.
+
+\begin{lemma}[Singular values of $\mOpp{v}{1}$]
+  \label{lem:mopp-singular-values}
+
+Let $h_t$ be the unique parent of $x_v$, as above.
+Let $h_1 \succ h_2 \cdots \succ h_t$ be a topological ordering of
+  variables according to the graph $\sG$.
+  Then, the smallest singular value of $\mOpp{v}{1}$ is at least the product of the smallest singular values of $\mYpp{c}{\Pa(h_c)}$.
+\begin{align*}
+  \sigma_{\min}(\sP) &\ge \prod_{c \in [t]} \sigma_{\min}(\mYpp{c}{\Pa(h_c)}),
+\end{align*}
+where $h_c$
+
+\end{lemma}
+\begin{proof}
+  - $Y$
+  - Use theorem of tensor multiplication.
+\end{proof}
+
+\begin{lemma}[Sample complexity for $\mOpp{v}{1}$]
+  \label{lem:sample-complexity}
+  The sample complexity to estimate $\mOpp{v}{i}$ via the tensor power method \citep{anandkumar13tensor} is 
+\end{lemma}
+\begin{proof}
+  The result follows directly from Theorem x of \citet{anandkumar13tensor}.
+\end{proof}

### 2014-05-06 doc/icml2014/tensor-multiplication.tex
Complete tensor proofs

@@
 For notational convenience, we will omit the permutation $\pi$ subsequently.
+Let $D_I = \bigtimes_{i \in I} [d_i]$ and  $\bar I \eqdef [m] \setminus
+  I$ to represent the complement of indices. Note that $\sP\munf{I}
+  = (\sP\munf{\bar I})^\top$.
 
@@
 
-\begin{lemma}[Tensor Multiplication and Singular Values]
-  \label{lem:tensor-multiplication}
+\begin{theorem}[Tensor Multiplication and Singular Values]
+  \label{thm:tensor-multiplication}
 Let $\sP, \sQ$ be $m$-th order and $n$-th order tensors respectively, $M$ be an
@@
 \end{align*}
-\end{lemma}
+\end{theorem}
 
@@
        &= \sum_{\vk \in D_\ell} \sP[\vk\cdot \vi_P \cdot \vj_P] \sQ[\vk\cdot \vi_Q \cdot \vj_Q] \\
-       &= \sum_{\vk \in D_\ell} \sP\munf{I_P}[\vi_P; \vj_P \cdot \vk] \sQ\munf{I_Q}[\vj_Q; \vj_Q \cdot \vk].
+       &= \sum_{\vk \in D_\ell} \sP\munf{I_P}[\vi_P; \vj_P \cdot \vk] \sQ\munf{\bar I_Q}[\vj_Q ; \vi_Q \cdot \vk].
   \end{align*}
-Note that the indices that were summed out belong in the complement of $I_P$ and $I_Q$ respectively.
+Note that the indices that were summed out belong in the complement of
+$I_P$ and $I_Q$ respectively.
 
@@
   \sP' &\eqdef \sP \otimes I_{d'_1} \otimes \cdots \otimes I_{d'_{|I_Q|}} \\
-  \sQ' &\eqdef \sQ \otimes I_{d'_1} \otimes \cdots \otimes I_{d'_{|I_P|}}.
+  \sQ' &\eqdef \sQ \otimes I_{d'_1} \otimes \cdots \otimes I_{d'_{|\bar I_P|}}.
 \end{align*}
@@
 Now, $\sP'$ is a $(m + 2 |I_Q|)$-th order tensor and similarly $\sQ'$
-  is a $(n + 2 |I_P|)$-th order tensor. 
-For every identity matrix we tensor $\sP$ and $\sQ$ by we include one of
-  its indices to include in the new set $I'_P$ and $I'_Q$ respectively.
-  Note that $\sP\munf{I'_P}$ is a $(I_P + I_Q)\times (m - |I_P| + |I_Q| + \ell)$ matrix, and 
-$\sP\munf{I'_Q}$ is $(I_P + I_Q) \times (n - |I_Q| + |I_P| + \ell)$
+  is a $(n + 2 |\bar I_P|)$-th order tensor. 
+For every identity matrix we tensor $\sP$ and $\sQ$ by, we include one of
+  its indices in the new set $I'_P$ and $I'_Q$ respectively.
+Note that ${\sP'}\munf{I'_P}$ is a $(I_P + I_Q)\times (|\bar I_P| + \ell
+  + |I_Q|)$ matrix, and ${\sQ'}\munf{\bar I'_Q}$ is $(|\bar I_Q| + |\bar
+  I_P|) \times (|\bar I_Q| + \ell + |\bar I_P|)$ matrix.
 Then,
@@
     \sR\munf{I}[\vi;\vj] 
-    &= \sum_{\be_P, \be_Q \vk}
-    {\sP'}\munf{I'_P}[\vi_P\be'_Q; \vj_P \be_Q \vk]  
-    {\sQ'}\munf{I'_Q}[\vi_Q\be'_P; \vj_Q \be_P \vk] \\
-    \sR\munf{I} &= {\sP'}\munf{I'_P} {\sQ'}\munf{I'_Q}.
+    &= \sum_{\be_P, \be_Q, \vk}
+    {\sP'}\munf{I'_P}[\vi_P \cdot \be'_Q; \vj_P \cdot \vk \cdot \be_Q ]  \\ 
+    &\quad \hphantom{\sum_{\be_P, \be_Q, \vk}}
+    {\sQ'}\munf{I'_Q}[\vi_Q \cdot \be'_P; \vj_Q \cdot \vk \cdot \be_P ] \\
+    \sR\munf{I} &= {\sP'}\munf{I'_P} ({\sQ'}\munf{\bar I'_Q})^\top \\
+                &= {\sP'}\munf{I'_P} {\sQ'}\munf{I'_Q}.
 \end{align*}

### 2014-05-06 doc/icml2014/assumption-proof.tex
Completed assumption proof section

@@
 For every clique $\sC \in \sG$ (including ones involving observed
-  variables), for every {\em conditioning} of the marginals, every
-  unfolding has full rank. 
+  variables) and for every {\em conditioning} of the clique's marginals,
+  every unfolding has full rank. 
 \end{assumption*}
@@
 be its unique parent. Then, 
-\begin{align*}
-  \mOpp{v}{1} &\eqdef \Pr( x_v \given h_1 )  \\
-              &= \sum_{h_t} \Pr( x_v \given h_t ) \Pr( h_t \given h_1 ) \\
-              &= \mOpp{v}{t} \mYpp{t}{1},
-\end{align*}
+\begin{align}
+  \mOpp{v}{1} &\eqdef \Pr( x_v \given h_1 ) \nonumber \\
+              &= \sum_{h_t}  \Pr( h_t \given h_1 ) \Pr( x_v \given h_t ) \nonumber\\
+              &= \mOpp{v}{t} \mYpp{t}{1}, \label{eqn:expanding-O}
+\end{align}
 where $\mYpp{i}{j} \eqdef \Pr( h_i \given h_j )$. 
 More generally, for two sets of hidden variables $C \eqdef \{h_{C_1}
-\cdots h_{C_m} \}$ and $C' \eqdef \{h_{C'_1} \cdots h_{C'_m} \}$, 
+\cdots h_{C_m} \}$ and $C' \eqdef \{h_{C'_1} \cdots h_{C'_n} \}$, 
 define $\mYpp{C}{C'} \eqdef \Pr( h_{C_1} \cdots h_{C_m} \given h_{C'_1}
@@
   &= \sum_{\vh \in H_{\Pa(h_t)}} \Pr( \Pa(h_t) \given h_1 ) \Pr( h_t \given \Pa(h_t) ) \\
-  &= \mYpp{ \Pa(h_t) }{1} \times_{\Pa(h_t)} \mYpp{t}{\Pa(h_t)},
+  &= \mYpp{t}{\Pa(h_t)} \times_{\Pa(h_t)} \mYpp{ \Pa(h_t) }{1},
 \end{align*}
@@
 refer to the operation $A \times_C B$ as tensor multiplication.
-\appendixref{tensor-multiplication} prooves some key properties of this
+\appendixref{tensor-multiplication} proves some key properties of this
 operation.
+
 Note that $\mYpp{t}{\Pa(h_t)}$ is the conditional probability $\Pr(h_t
@@
   between $C$ and $C'$.
+Note that $\Pa(h_c) \subseteq \del C$.
 Then,
-\begin{align*}
-  \mYpp{C}{1} &\eqdef \Pr( h_C \given h_1 )  \\
-  &= \sum_{\vh \in H_{\del C}} \Pr( C' \given h_1 ) \Pr( h_c \given \Pa(h_c) ) \\
-  &= \mYpp{ C' }{ 1 } \times_{\del C} \mYpp{c}{\Pa(h_c)},
-\end{align*}
-Finally, for the base case, $\mYpp{1}{1} = \ones$.
-This procedure is algorithmically described in \algorithmref{Y}.
+\begin{align}
+  \mYpp{C}{1} &\eqdef \Pr( h_c \given h_1 ) \nonumber \\
+  &= \sum_{\vh \in H_{\del C}} \Pr( h_c \given \Pa(h_c) ) \Pr( C' \given h_1 ) \nonumber \\
+  &= \mYpp{c}{\Pa(h_c)}  \times_{\del C} \mYpp{ C' }{ 1 }. \label{eqn:recursive-step}
+\end{align}
+This process is repeated until the base case, $\mYpp{1}{1} = \ones$.
+\algorithmref{Y} summarizes the procedure.
 
@@
 
-Note that in the process of the algorithm, $\mYpp{t}{1}$ is expressed
-entirely as tensor multiplications with terms of the form
-$\mYpp{c}{\Pa(h_c)}$ for some hidden variable $h_c$. 
-$\mYpp{c}{\Pa(h_c)}$ can be easily computed from the clique marginals
-$Z_\sC$.
+The algorithm constructs an expression for $\mYpp{t}{1}$ entirely as
+  tensor multiplications of terms of the form $\mYpp{c}{\Pa(h_c)}$ for
+  some hidden variable $h_c$. 
+The individual $\mYpp{c}{\Pa(h_c)}$ can be easily computed from the
+  clique marginals $Z_\sC$ as follows.
 
@@
 where $\diag(\sP)$ is a ``diagonal'' tensor constructed by taking the
-tensor product with $I^{\otimes p}$,
+tensor product with $I^{\otimes p}$:
 \begin{align*}
@@
 
-From the algorithm described above, blah.
+With an expression for $\mOpp{v}{i}$, we are finally ready to describe
+sufficient conditions for $\mOpp{v}{i}$ to be full rank, along with the
+sample complexity for estimation.
 
@@
   \label{lem:full-rank-suff}
-  Given that \assumptionref{full-rank-plus} holds, then for any hidden
-  variable $h_1$ and observed variable $x_v$, $\mPi{1} \succ 0$ and
+  Given the full rank conditions in \assumptionref{full-rank-plus}, then
+  for any hidden variable $h_i$ and observed variable $x_v$,
   $\mOpp{v}{i}$ has full column rank.
@@
 \begin{proof}
+Again, wlog, let $i=1$. 
+The key step of the proof will be to use the property that the tensor
+  multiplication in \equationref{recursive-step} preserves rank.
 
-  - $Y$
-  - Use theorem of tensor multiplication.
+\theoremref{tensor-multiplication} shows that for any two tensors $A,
+  B$ and an index set $C$, the unfoldings $I$ of $A \times_C B$ are full
+  rank when the projected unfoldings $I_A$ of $A$ and $I_B$ of $B$ are
+  full rank:
+\begin{align*}
+  \sigma_{\min}( (A \times_C B)\munf{I} )
+    &\ge \sigma_{\min}(A\munf{I_A}) \sigma_{\min}(B\munf{I_B}).
+\end{align*}
+
+By the assumption, any unfolding of $\mYpp{c}{\Pa(h_c)}$ is full-rank. 
+From \equationref{recursive-step} and
+  \theoremref{tensor-multiplication}, this implies that $\mYpp{C}{1}$ is
+  full rank for any $C$, including $C = \{h_t\}$.
+
+Thus, $\mYpp{t}{1}$ is a full rank matrix. 
+Again, by assumption,
+  $\mOpp{v}{t}$ is full rank. From \equationref{expanding-O}, we finally
+  get that $\mOpp{v}{1}$ is full rank.
 \end{proof}
 
-The sample complexity of estimating $\mOpp{v}{i}$ depends mainly on
+Next, we turn to the question of the sample complexity of estimating
+$\mOpp{v}{i}$, which depends mainly on smallest singular value,
 $\sigma_{\min}(\mOpp{v}{i})$.
@@
 \begin{align*}
-  \sigma_{\min}(\sP) &\ge \prod_{c \in [t]} \sigma_{\min}(\mYpp{c}{\Pa(h_c)}),
+  \sigma_{\min}(\sP) &\ge \sigma_{\min}(\mOpp{v}{t}) 
+        \prod_{c \in [t]} \sigma_{\min}(\mYpp{c}{\Pa(h_c)}),
 \end{align*}
-where $h_c$
-
+where the parents $\Pa(h_c)$ are decided by the topological ordering. 
 \end{lemma}
 \begin{proof}
-  - $Y$
-  - Use theorem of tensor multiplication.
+  The proof follows directly by application of
+  \theoremref{tensor-multiplication} to \equationref{expanding-O} and
+  \equationref{recursive-step}.
 \end{proof}
 
-\begin{lemma}[Sample complexity for $\mOpp{v}{1}$]
-  \label{lem:sample-complexity}
-  The sample complexity to estimate $\mOpp{v}{i}$ via the tensor power method \citep{anandkumar13tensor} is 
-\end{lemma}
+Finally, using the tensor power method of \citep{anandkumar13tensor}, we
+  get the following result on sample complexity.
+\begin{theorem}[Sample complexity for $\mOpp{v}{i}$]
+  \label{thm:sample-complexity}
+  Wlog, let $v=1$ and $i=1$. Let $x_1, x_2, x_3$ be three views for
+  $h_1$.  
+  If 
+\begin{align*}
+  \|\hat M_{1,2} - M_{1,2}\|_{\op} &\le \epsilon & \|\hat M_{1,2,3} - M_{1,2,3}\|_{\op} &\le \epsilon,
+\end{align*}
+for some $\epsilon < \half$, then
+with probability at least $1 - \delta$,
+\begin{align*}
+  \|\mOpphat{1}{1} - \mOpp{1}{1}\|_F 
+    &\le  \\
+    &
+      O\left( k 
+      \frac{{\pi\oft{1}}_{\max}/{\pi\oft{1}}_{\min}} 
+      {(\sigma_{\min}(\mOpp{1}{1}) \sigma_{\min}(\mOpp{1}{2}))^{5/2}} \right) \epsilon.
+\end{align*}
+
+Furthermore, if $\sigma_{\min}(\mYpp{c}{\Pa(h_c)}) \le \sigma$ for every
+  such
+$\mYpp{c}{\Pa(h_c)}$, we get,
+\begin{align*}
+  \|\mOpphat{1}{1} - \mOpp{1}{1}\|_F 
+    &\le 
+      O\left( k 
+      \frac{{\pi\oft{1}}_{\max}/{\pi\oft{1}}_{\min}} 
+      {\sigma^{5t}} \right) \epsilon,
+\end{align*}
+where $t$ is the length of the topological ordering, $h_1 > \cdots
+> h_t$ as defined above.
+\end{theorem}
 \begin{proof}
-  The result follows directly from Theorem x of \citet{anandkumar13tensor}.
+  The first statement is a result of from Theorem 5.1 of
+  \citet{anandkumar13tensor}. The second statement follows directly from
+  \lemmaref{mopp-singular-values}.
 \end{proof}
+
+\theoremref{sample-complexity} says that the number of samples required
+  to estimate the view $\mOpp{v}{i}$ grows exponentially in the {\em
+  distance} between $x_v$ and $h_i$. 
+Intuitively, this is because the influence from $h_i$ to $x_v$ gets
+  diluted for every latent variable it must factor through.

### 2014-05-06 doc/icml2014/assumption-proof.tex
Updated assumption's proof and made directed section reflect teh same

@@
 \begin{theorem}[Sample complexity for $\mOpp{v}{i}$]
-  \label{thm:sample-complexity}
+  \label{thm:sample-complexity-1}
   Wlog, let $v=1$ and $i=1$. Let $x_1, x_2, x_3$ be three views for
@@
 \begin{proof}
-  The first statement is a result of from Theorem 5.1 of
-  \citet{anandkumar13tensor}. The second statement follows directly from
+  The first statement follows directly from Theorem 5.1 of
+  \citet{anandkumar13tensor} by noting that $\|M_{1,2}\|_\op \le 1$ and
+  $\|M_{1,2,3}\|_\op \le 1$ (as they represent probability
+  distributions). The second statement follows directly from
   \lemmaref{mopp-singular-values}.
 \end{proof}
-
-\theoremref{sample-complexity} says that the number of samples required
-  to estimate the view $\mOpp{v}{i}$ grows exponentially in the {\em
-  distance} between $x_v$ and $h_i$. 
-Intuitively, this is because the influence from $h_i$ to $x_v$ gets
-  diluted for every latent variable it must factor through.

### 2014-05-06 doc/icml2014/background.tex
Updated assumption's proof and made directed section reflect teh same

@@
 \label{asm:full-rank}
-Assume the conditional moments $\mOpp{1}{1}, \mOpp{2}{1}, \mOpp{3}{1}$ have full column rank
+The conditional moments $\mOpp{1}{1}$, $\mOpp{2}{1}$, $\mOpp{3}{1}$ have full column rank
   $k$, and that $\pi\oft 1 \succ 0$.
@@
   in latent-variable models.  In the next two sections,
-  we will see how to build on top of this result to estimate directed (Section~\ref{sec:directed})
+  we will see how to use this result to estimate directed (Section~\ref{sec:directed})
   and undirected (Section~\ref{sec:undirected}) models.

### 2014-05-06 doc/icml2014/directed.tex
Updated assumption's proof and made directed section reflect teh same

@@
   \sectionref{piecewise}, we show how we can get a more statistically
-  efficient estimator using composite likelihood.
+  efficient estimator using composite likelihoods.
 
@@
 Every directed (undirected) graphical model can be transformed into one in which
-  the observed variables are leaves with exactly one parent (neighbour). 
-There is a one-to-one correspondence between the parameters of this
-  transformed model and the original one.
+  the observed variables are leaves with exactly one parent (neighbor). 
+There is a one-to-one correspondence between the parameters of the
+  transformed and original models.
 \end{lemma}
@@
 \begin{proof}
-  By \lemmaref{reduction}, we can assume that every observed
-  variable is a leaf in the graph, causing any v-structures in the graph
-  to be active. This allows us to reason about independence simply by
-  checking whether or not two nodes are connected or not. Note that this
-  proof applies for undirected graphs as well.
-
-  Let $h$ be a hidden variable in $\sG'$, let $\sV_{h}$ be the set of
-  views for it and let ${\sG'}^- \eqdef \sG' \setminus \{h\}$.
-  By definition, the views of $h$ are conditionally independent, viz.
-    every connecting path between two views $x_1, x_2 \in \sV_h$ must pass through
-    $h$.
-  A view $x \in \sV_{h}$ is exclusive for $h$ iff $x \not\in \sV_{h'} ~ \forall h'
-  \in {\sG'}^-$.
-
-  Suppose $h$ had no unique view, then $\sV_{h} \subseteq \Union_{h' \in {\sG'}^-} \sV_{h'}$. 
-  However, this would imply that for every $x_1, x_2 \in \sV_{h}$, there
-  exists some $h_1, h_2 \in {\sG'}^-$ (not necessarily distinct) such that
-    $x_1$ is connected to $h_1$ and $x_2$ is connected to $h_2$.
-  By participation in the sub-graph, either $h_1 = h_2$ or $h_1$ is
-    connected to $h_2$, implying that $x_1$ and $x_2$ are connected
-    via a path that does not pass through $h$, contradicting the
-    statement that $x_1, x_2 \in \sV_h$.
-
-  Thus, it must be the case that $\sV_{h} \nsubseteq \Union_{h' \neq
-    h \in \sG'} \sV_{h'}$ and hence there exists a unique view for $h$.
-
-  Algorithmically, finding the unique view for $h$ can be done by
-  subtracting the views of all other variables in the sub-graph from
-  $\sV_h$,i.e. $x_{v_h} \in \sV_h \setminus \Union_{h' \in \sG'} \sV_{h'}$
+  The independence conditions required by the bottleneck condition
+  effectively guarantee that atleast one of the three views for
+  a bottleneck be exclusive. 
+  \appendixref{exclusive-views} provides a complete proof.
 \end{proof}
@@
   column rank and that $\pi^{(i)} \eqdef \Pr(h_i) \succ 0$.
-But $\mOpp{v}{i}$ is a product of
-tensors on the path from $h_i$ to $x_v$, marginalizing out the rest of the graph.
-To make this property more interpretable, let us assume the following property:%\verify
+But $\mOpp{v}{i}$ is a product of tensors on the path from $h_i$ to
+  $x_v$, marginalizing out the rest of the graph.
+To make this property more interpretable, let us assume the following
+  property which we prove to be sufficient in
+  \appendixref{assumption-proof}:
 \begin{assumption} 
 \label{asm:full-rank-plus}
-For every clique $\sC \in \sG$ (including ones involving observed variables),
-  every $\Re^{k \times k}$ slice of the marginals $Z_\sC$ has full column
-  rank $k$ constructed by summing out the remaining indices.
+The marginal distribution of every latent variable has full support:
+  $\forall{h_i \in H},~ \mPi{i} \succ 0$.
+For every clique $\sC \in \sG$ (including ones with observed variables),
+  and every conditional distribution $\Pr(h_c \given \Pa(h_c))$, where
+  $h_c \in \sC$, every {\em tensor unfolding} has full column rank, $k$.
 \end{assumption}
-In our directed grid example (\sectionref{directedExample}), for the clique $\sC = \{ h_1, h_2 \}$ this
-  condition implies $\pi = \Pr(h_1) \succ 0$ and $T = \Pr(h_2 \mid h_1)$ has column rank $k$; 
-for the clique $\sC = \{ h_1, x_1 \}$, it implies $O = \Pr(x_1 \mid h_1)$ has column rank $k$.
+In our directed grid example (\sectionref{directedExample}), for the
+  clique $\sC = \{ h_1, h_2 \}$ this condition implies $\pi = \Pr(h_1)
+  \succ 0$ and $T = \Pr(h_2 \mid h_1)$ has column rank $k$; for the clique
+  $\sC = \{ h_1, x_1 \}$, it implies $O = \Pr(x_1 \mid h_1)$ has column
+  rank $k$.
+
+Intuitively, we need to ensure that every marginal distribution
+  $\Pr(x_v)$ can be separated into $k$ different conditional distributions
+  $\Pr(x_v | h_i)$. 
+The full rank assumption asks that the conditional distributions are
+  linearly independent; if this were not the case, the conditional
+  distribution for a particular latent state could be imitated by the
+  combination of several other latent states.
 
@@
 
-From \citet{anandkumar13tensor} given \assumptionref{full-rank-plus},
-we have that $\mOpphat{v}{i}$ converges to $\mOpp{v}{i}$ at a rate of $n^{-\frac12}$ with a constant that
-depends polynomially on the $k$-th singular value of $\mOpp{v}{i}$.
+From \citet{anandkumar13tensor} given \assumptionref{full-rank-plus}, we
+  have that $\mOpphat{v}{i}$ converges to $\mOpp{v}{i}$ at a rate of
+  $n^{-\frac12}$ with a constant that depends polynomially on the $k$-th
+  singular value of $\mOpp{v}{i}$:
 
-\todo{Present sample complexity bounds for $\mOpp{v}{i}$.}
-\todo{Present sample complexity bounds for $Z_\sC$.}
+\begin{theorem}[Sample complexity for $\mOpp{v}{i}$]
+  \label{thm:sample-complexity}
+  Wlog, let $v=1$ and $i=1$. Let $x_1, x_2, x_3$ be three views for
+  $h_1$. With probability at least $1 - \delta$,
+\begin{align*}
+  \|\mOpphat{1}{1} - \mOpp{1}{1}\|_F 
+    &\le  \\
+    &
+      O\left(
+      \frac{k \log(1/\delta) {\pi\oft{1}}_{\max}/{\pi\oft{1}}_{\min} } 
+      {\sqrt{n}(\sigma_{\min}(\mOpp{1}{1}) \sigma_{\min}(\mOpp{1}{2}))^{5/2}} \right) \epsilon.
+\end{align*}
+
+If $\sigma_{\min}(\Pr(c \given \Pa(h_c))) \le \sigma$ for every clique
+  $\sC \in \sG$ and hidden variable $h_c \in \sC$,
+\begin{align*}
+  \|\mOpphat{1}{1} - \mOpp{1}{1}\|_F 
+    &\le 
+      O\left( 
+      \frac{k \log(1/\delta) {\pi\oft{1}}_{\max}/{\pi\oft{1}}_{\min}} 
+      {\sqrt{n} \sigma^{5t}} \right) \epsilon,
+\end{align*}
+where $t$ is the length of the shortest topological ordering from $h_1$
+to the unique parent of $x_1$, $h_t$.
+\end{theorem}
+
+\theoremref{sample-complexity}, proved in \appendixref{assumption-proof}
+  says that the number of samples required to estimate the view
+  $\mOpp{v}{i}$ grows exponentially in the {\em distance} between $x_v$
+  and $h_i$. 
+Intuitively, this is expected because the influence from $h_i$ to $x_v$
+  gets diluted for every latent variable it must factor through.
+
+\todo{Present sample complexity bounds for $Z_\sC$?}
 

### 2014-05-06 doc/icml2014/exclusive-views.tex
Updated assumption's proof and made directed section reflect teh same

\subsection{\lemmaref{lem:bottleneck-views}}
\label{app:exclusive-views}

In \sectionref{directed}, we stated that bottlenecks guaranteed
  exclusive views. We provide the proof below.
\begin{lemma*}[Bottlenecks guarantee exclusive views]
  %\label{lem:bottleneck-views}  
Let $\sG' = \{ h_{i_1}, \cdots, h_{i_m} \} \subseteq \sG$ be
  a connected sub-graph of hidden variables. If every hidden variable
  $h_i \in \sG'$ is a bottleneck, then they have an exclusive view.
\end{lemma*}
\begin{proof}
  By \lemmaref{reduction}, we can assume that every observed
  variable is a leaf in the graph, causing any v-structures in the graph
  to be active. This allows us to reason about independence simply by
  checking whether or not two nodes are connected or not. Note that this
  proof applies for undirected graphs as well.

  Let $h$ be a hidden variable in $\sG'$, let $\sV_{h}$ be the set of
  views for it and let ${\sG'}^- \eqdef \sG' \setminus \{h\}$.
  By definition, the views of $h$ are conditionally independent, viz.
    every connecting path between two views $x_1, x_2 \in \sV_h$ must pass through
    $h$.
  A view $x \in \sV_{h}$ is exclusive for $h$ iff $x \not\in \sV_{h'} ~ \forall h'
  \in {\sG'}^-$.

  Suppose $h$ had no unique view, then $\sV_{h} \subseteq \Union_{h' \in {\sG'}^-} \sV_{h'}$. 
  However, this would imply that for every $x_1, x_2 \in \sV_{h}$, there
  exists some $h_1, h_2 \in {\sG'}^-$ (not necessarily distinct) such that
    $x_1$ is connected to $h_1$ and $x_2$ is connected to $h_2$.
  By participation in the sub-graph, either $h_1 = h_2$ or $h_1$ is
    connected to $h_2$, implying that $x_1$ and $x_2$ are connected
    via a path that does not pass through $h$, contradicting the
    statement that $x_1, x_2 \in \sV_h$.

  Thus, it must be the case that $\sV_{h} \nsubseteq \Union_{h' \neq
    h \in \sG'} \sV_{h'}$ and hence there exists a unique view for $h$.

  Algorithmically, finding the unique view for $h$ can be done by
  subtracting the views of all other variables in the sub-graph from
  $\sV_h$,i.e. $x_{v_h} \in \sV_h \setminus \Union_{h' \in \sG'} \sV_{h'}$
\end{proof}

### 2014-05-07 doc/icml2014/assumption-proof.tex
Added figure to assumption proof
Updated piecewise likelihoods expression
Removed false statement about Z -> Y$

@@
 
-\begin{figure}
-  \caption{A recursive construction of the conditional moments $\mOpp{v}{i}$.}
+\begin{figure}[t]
   \label{fig:message-proof}
+  \centering
+  \subimport{figures/}{assumption.tikz}
+  \caption{A recursive construction of the conditional moments $\mOpp{v}{i}$. Shaded regions indicate participating variables in the intermediate distributions.}
 \end{figure}
@@
 
-\paragraph{Constructing $\mYpp{c}{\Pa(h_c)}$}
-
-The algorithm constructs an expression for $\mYpp{t}{1}$ entirely as
-  tensor multiplications of terms of the form $\mYpp{c}{\Pa(h_c)}$ for
-  some hidden variable $h_c$. 
-The individual $\mYpp{c}{\Pa(h_c)}$ can be easily computed from the
-  clique marginals $Z_\sC$ as follows.
-
-Without loss of generality, let $\sC \eqdef \{c\} \union \Pa(h_c)$ be a clique
-  in the graph $\sG$.
-If instead $\sC \subseteq \sC_1 \union \sC_2$ for some $\sC_1, \sC_2 \in
-  \sG$, we can simply construct $Z_{\sC}$ from the union of $\sC_1$ and
-  $\sC_2$,
-\begin{align*}
-  Z_\sC &= Z_{\sC_1} \times_{\del \sC} Z_{\sC_2},
-\end{align*}
-where $\del \sC = (\sC_1 \union \sC_2) \setminus \sC$.
-
-Proceeding,
-\begin{align*}
-\mYpp{c}{\Pa(h_c)} 
-  &\eqdef \Pr(h_c \given \Pa(h_c)) \\
-  &= \frac{\Pr(h_c, \Pa(h_c))}{\Pr(\Pa(h_c))} \\
-  &= Z_\sC \times_{\Pa(h_c)} \diag{Z_\sC(\ones, \cdot, \ldots, \cdot)}^{-1},
-\end{align*}
-where $\diag(\sP)$ is a ``diagonal'' tensor constructed by taking the
-tensor product with $I^{\otimes p}$:
-\begin{align*}
-  \diag(\sP)[\vi,\vj] &\eqdef \sP[\vi] \delta[\vi,\vj].
-\end{align*}
+% \paragraph{Constructing $\mYpp{c}{\Pa(h_c)}$}
+% 
+% The algorithm constructs an expression for $\mYpp{t}{1}$ entirely as
+%   tensor multiplications of terms of the form $\mYpp{c}{\Pa(h_c)}$ for
+%   some hidden variable $h_c$. 
+% The individual $\mYpp{c}{\Pa(h_c)}$ can be easily computed from the
+%   clique marginals $Z_\sC$ as follows.
+% 
+% Without loss of generality, let $\sC \eqdef \{c\} \union \Pa(h_c)$ be a clique
+%   in the graph $\sG$.
+% If instead $\sC \subseteq \sC_1 \union \sC_2$ for some $\sC_1, \sC_2 \in
+%   \sG$, we can simply construct $Z_{\sC}$ from the union of $\sC_1$ and
+%   $\sC_2$,
+% \begin{align*}
+%   Z_\sC &= Z_{\sC_1} \times_{\del \sC} Z_{\sC_2},
+% \end{align*}
+% where $\del \sC = (\sC_1 \union \sC_2) \setminus \sC$.
+% 
+% Proceeding,
+% \begin{align*}
+% \mYpp{c}{\Pa(h_c)} 
+%   &\eqdef \Pr(h_c \given \Pa(h_c)) \\
+%   &= \frac{\Pr(h_c, \Pa(h_c))}{\Pr(\Pa(h_c))} \\
+%   &= Z_\sC \times_{\Pa(h_c)} \diag{Z_\sC(\ones, \cdot, \ldots, \cdot)}^{-1},
+% \end{align*}
+% where $\diag(\sP)$ is a ``diagonal'' tensor constructed by taking the
+% tensor product with $I^{\otimes p}$:
+% \begin{align*}
+%   \diag(\sP)[\vi,\vj] &\eqdef \sP[\vi] \delta[\vi,\vj].
+% \end{align*}
 

### 2014-05-07 doc/icml2014/assumption-proof.tex
Included section about tensor projection and how to get Y in the assumption proof

@@
 \end{assumption*}
+
+% Tour guide
+Going forward, we will describe $\mOpp{v}{i}$ recursively in terms of
+  tensor multiplications of conditional distributions $\Pr(h_c \given
+  \Pa(h_c))$ in a variable elimination ordering.
+We then show that the above assumption implies the unfoldings of the
+  conditional distributions are full rank as well.
+The analysis also tracks singular values, allowing us to present
+  a sample complexity bound that highlights an exponential dependence on
+  the size of the variable elimination set.
 
@@
 
-% \paragraph{Constructing $\mYpp{c}{\Pa(h_c)}$}
-% 
-% The algorithm constructs an expression for $\mYpp{t}{1}$ entirely as
-%   tensor multiplications of terms of the form $\mYpp{c}{\Pa(h_c)}$ for
-%   some hidden variable $h_c$. 
-% The individual $\mYpp{c}{\Pa(h_c)}$ can be easily computed from the
-%   clique marginals $Z_\sC$ as follows.
-% 
-% Without loss of generality, let $\sC \eqdef \{c\} \union \Pa(h_c)$ be a clique
-%   in the graph $\sG$.
-% If instead $\sC \subseteq \sC_1 \union \sC_2$ for some $\sC_1, \sC_2 \in
-%   \sG$, we can simply construct $Z_{\sC}$ from the union of $\sC_1$ and
-%   $\sC_2$,
-% \begin{align*}
-%   Z_\sC &= Z_{\sC_1} \times_{\del \sC} Z_{\sC_2},
-% \end{align*}
-% where $\del \sC = (\sC_1 \union \sC_2) \setminus \sC$.
-% 
-% Proceeding,
-% \begin{align*}
-% \mYpp{c}{\Pa(h_c)} 
-%   &\eqdef \Pr(h_c \given \Pa(h_c)) \\
-%   &= \frac{\Pr(h_c, \Pa(h_c))}{\Pr(\Pa(h_c))} \\
-%   &= Z_\sC \times_{\Pa(h_c)} \diag{Z_\sC(\ones, \cdot, \ldots, \cdot)}^{-1},
-% \end{align*}
-% where $\diag(\sP)$ is a ``diagonal'' tensor constructed by taking the
-% tensor product with $I^{\otimes p}$:
-% \begin{align*}
-%   \diag(\sP)[\vi,\vj] &\eqdef \sP[\vi] \delta[\vi,\vj].
-% \end{align*}
+\paragraph{Rank conditions for $\mYpp{c}{\Pa(h_c)}$}
+
+\algorithmref{Y} constructs an expression for $\mYpp{t}{1}$ entirely as
+  tensor multiplications of terms of the form $\mYpp{c}{\Pa(h_c)}$ for
+  some hidden variable $h_c$. 
+The singular values of unfoldings of individual $\mYpp{c}{\Pa(h_c)}$ can
+  be easily computed from the clique marginals $Z_\sC$ as follows.
+
+\begin{lemma} 
+  Let $\sC_0 \eqdef {h_c} \union \Pa(h_c) \subseteq \sC_1 \union \cdots
+  \union \sC_\ell$ for $\sC_1, \cdots, \sC_\ell \in \sG$. Note that
+  $\sC_0$ need not belong in the graph.
+
+  Then, for the unfolding $\{c\}$, we get 
+  \begin{align*}
+    \sigma_{\min}(\mYpp{c}{\Pa(c)}\munf{I}) &\ge \max_{i\in[\ell]}\{\sigma_{\min}(Z_{\sC_i}\munf{I_i})\},
+  \end{align*}
+  where $i$ is the subset of $I$ present in clique $\sC_i$.
+\end{lemma} 
+\begin{proof}
+Without loss of generality, let $\sC \eqdef \{c\} \union \Pa(h_c)$ be a clique
+  in the graph $\sG$.
+If instead $\sC \subseteq \sC_1 \union \sC_2$ for some $\sC_1, \sC_2 \in
+\sG$, then $Z_{\sC}$ is the marginalization of some $Z_{\sC_1 \union \sC_2}$
+(which may not be a parameter in our model).
+
+\lemmaref{tensor-projection} states that when a tensor, $Z_{\sC_1 \union
+\sC_2}$, is projected by a non-zero vector $v$ (in this case, $\ones$),
+then the singular values of the projection $Z_\sC$ are preserved along
+any unfolding $I$ of $Z_\sC$:
+\begin{align*}
+  \sigma_{\min}(Z_{\sC}\munf{I}) &\ge \sigma_{\min}(Z_{\sC_1 \union \sC_2}\munf{I})v_{\min} \\
+  \sigma_{\min}(Z_{\sC_1 \union \sC_2}\munf{I}) &\ge
+  \sigma_{\min}(Z_{\sC}\munf{I})v_{\max}\inv.
+\end{align*}
+
+In this case, we get that $\sigma_{\min}(Z_{\sC_1 \union \sC_2}\munf{I})
+= \sigma_{\min}(Z_{\sC}\munf{I})v_{\max}\inv$.
+
+Furthermore, assuming that any unfolding of $Z_{\sC_1}$ is full rank, we
+have that $Z_{\sC_1 \union \sC_2}$ is also full rank along any common
+unfolding.
+
+Proceeding,
+\begin{align*}
+\mYpp{c}{\Pa(h_c)} 
+  &\eqdef \Pr(h_c \given \Pa(h_c)) \\
+  &= \frac{\Pr(h_c, \Pa(h_c))}{\Pr(\Pa(h_c))} \\
+  \mYpp{c}{\Pa(h_c)}\munf{I}
+  &= Z_\sC\munf{I} \diag{Z_\sC(\ones, \cdot, \ldots, \cdot)}^{-1},
+\end{align*}
+where $\diag(\sP)$ is a ``diagonal'' tensor constructed by taking the
+tensor product with the identity matrix,
+\begin{align*}
+  \diag(\sP)[\vi,\vj] &\eqdef \sP[\vi] \delta[\vi,\vj].
+\end{align*}
+
+We have from the previous argument that $Z_\sC\munf{I}$ has full column
+rank $k$; furthermore, $Z_\sC(\ones, \cdot, \ldots, \cdot)$ is
+guaranteed to have at least $k$ non-zero elements because its complete
+marginals, $\mPi{i} \succ 0 ~\forall h_i \in \sC$ have full support. 
+Consequently, $\diag(Z_\sC(\ones, \cdot, \ldots, \cdot))$ has column rank $k$.
+\todo{What happens if $\Pr(h_2, h_3) = 0$ at some point?}
+\end{proof}
 

### 2014-05-07 doc/icml2014/tensor-multiplication.tex
Included section about tensor projection and how to get Y in the assumption proof

@@
 %matrix multiplication, and the diagonal tensor.
-In this section, we present basic results regarding tensor
-multiplication, an operation analogous to matrix multiplication.
+In this section, we present some results regarding a few tensor
+operations, most significant of which is tensor multiplication, an
+operation analogous to matrix multiplication.  
 First, let us define some notation.
 
+\paragraph{Notation}
 In the following, let $\sP \in \Re^{d_1 \times \ldots \times d_m}$ be
@@
 \in [d_{\ell + 1}] \times \cdots \times [d_m]$ and  we use the notation
-$\sP[\vi \cdot \vj]_\pi \eqdef \sP[\pi(i_1, \cdot, i_\ell, j_{\ell+1},
+$\sP[\vi \ \vj]_\pi \eqdef \sP[\pi(i_1, \cdots, i_\ell, j_{\ell+1},
 \cdots j_m)]$ to conjoin indices $\vi$ and $\vj$ up to the permutation
 $\pi$.
+
+\paragraph{Unfoldings}
+For a tensor $\sP$, as defined above, 
+  given an index set $I \subset [m]$ we
+  define an unfolding of a tensor $\sP$ along $I$,
+ $\sP\munf{I}$, to be a matrix of dimensions $(\prod_{i \in I} d_i)
+ \times (\prod_{j \not\in I} d_j)$. $\sP\munf{I}$ has elements
+ $\sP\munf{I}[\vi;\vj] = \sP[\vi\ \vj]_{\pi}$, where $\pi$ is an
+ appropriate permutation to align the indices of $\vi$ and $\vj$ with
+ the ordering of $I$. 
+For example, if $\sP$ were a $4$-th order tensor and $I = \{1,3\}$,
+  $\sP\munf{13}[i_1,i_2; j_1,j_2] = \sP[i_1,j_1,i_2,j_2]$.
+For notational convenience, we will omit the permutation $\pi$ subsequently.
+Let $D_I = \bigtimes_{i \in I} [d_i]$ and  $\bar I \eqdef [m] \setminus
+  I$ represent the complement of indices. Note that $\sP\munf{I}
+  = (\sP\munf{\bar I})^\top$.
+
 %For our purposes, the specific ordering of the tensor's indices
@@
 
+\paragraph{Tensor Multiplication}
+First, let us formally define the tensor multiplication operation.
+
 \begin{definition}[Tensor Multiplication]
@@
 \begin{align*}
-  \sR[\vi\cdot\vj] &\eqdef \sum_{\mathclap{\bu, \bv \in D_\ba \times D'_\bb}} \sP[\vi \cdot \bu]_{\pi_P} \sQ[\vj \cdot \bv]_{\pi_Q} ~ \delta_{u_1 v_1} \delta_{u_2 v_2} \ldots \delta_{u_\ell v_\ell},
+  \sR[\vi\ \vj] &\eqdef \sum_{\mathclap{\bu, \bv \in D_\ba \times D'_\bb}} \sP[\vi \  \bu]_{\pi_P} \sQ[\vj \  \bv]_{\pi_Q} ~ \delta_{u_1 v_1} \delta_{u_2 v_2} \ldots \delta_{u_\ell v_\ell},
 \end{align*}
@@
 
-\paragraph{Unfoldings}
-For a tensor $\sP$, as defined above, 
-  given an index set $I \subset [m]$ we
-  define an unfolding of a tensor $\sP$ along $I$,
- $\sP\munf{I}$, to be a matrix of dimensions $(\prod_{i \in I} d_i)
- \times (\prod_{j \not\in I} d_j)$. $\sP\munf{I}$ has elements
- $\sP\munf{I}[\vi;\vj] = \sP[\vi\cdot\vj]_{\pi}$, where $\pi$ is an
- appropriate permutation to align the indices of $\vi$ and $\vj$ with
- the ordering of $I$. 
-For example, if $\sP$ were a $4$-th order tensor and $I = \{1,3\}$,
-  $\sP\munf{13}[i_1,i_2; j_1,j_2] = \sP[i_1,j_1,i_2,j_2]$.
-For notational convenience, we will omit the permutation $\pi$ subsequently.
-Let $D_I = \bigtimes_{i \in I} [d_i]$ and  $\bar I \eqdef [m] \setminus
-  I$ to represent the complement of indices. Note that $\sP\munf{I}
-  = (\sP\munf{\bar I})^\top$.
-
-The main result of this section is, similar to matrix multiplication, if
-$\sR = \sP \times_{M} \sQ$, then the singular values of unfoldings
-of $\sR$ are bounded by those of $\sP$ and $\sQ$.
+The main result of this section is that, similar to matrix
+multiplication, if $\sR = \sP \times_{M} \sQ$, then the singular values
+of unfoldings of $\sR$ are bounded by those of $\sP$ and $\sQ$.
 
@@
   \begin{align*}
-    \sR[\vi\cdot\vj] &\eqdef \sum_{\vk \in D_\ell} \sP[\vk\cdot\vi_P\cdot\vj_P] \sQ[\vk\cdot\vi_Q\cdot\vj_Q],
+    \sR[\vi\ \vj] &\eqdef \sum_{\vk \in D_\ell} \sP[\vk\ \vi_P\ \vj_P] \sQ[\vk\ \vi_Q\ \vj_Q],
   \end{align*}
@@
     \sR\munf{I}[\vi;\vj] 
-       &= \sum_{\vk \in D_\ell} \sP[\vk\cdot \vi_P \cdot \vj_P] \sQ[\vk\cdot \vi_Q \cdot \vj_Q] \\
-       &= \sum_{\vk \in D_\ell} \sP\munf{I_P}[\vi_P; \vj_P \cdot \vk] \sQ\munf{\bar I_Q}[\vj_Q ; \vi_Q \cdot \vk].
+       &= \sum_{\vk \in D_\ell} \sP[\vk\  \vi_P \  \vj_P] \sQ[\vk\  \vi_Q \  \vj_Q] \\
+       &= \sum_{\vk \in D_\ell} \sP\munf{I_P}[\vi_P; \vj_P \  \vk] \sQ\munf{\bar I_Q}[\vj_Q ; \vi_Q \  \vk].
   \end{align*}
@@
     \sR\munf{I}[\vi;\vj] 
-    &= \sum_{\be_P, \be_Q, \vk}
-    {\sP'}\munf{I'_P}[\vi_P \cdot \be'_Q; \vj_P \cdot \vk \cdot \be_Q ]  \\ 
+    &= \sum_{\vj'_P, \vi'_Q, \vk}
+    {\sP'}\munf{I'_P}[\vi_P \  \vi_Q; \vj'_P \  \vk \  \vi'_Q ]  \\ 
     &\quad \hphantom{\sum_{\be_P, \be_Q, \vk}}
-    {\sQ'}\munf{I'_Q}[\vi_Q \cdot \be'_P; \vj_Q \cdot \vk \cdot \be_P ] \\
+    {\sQ'}\munf{I'_Q}[\vj_Q \  \vj_P; \vi'_Q \  \vk \  \vj'_P ] \\
     \sR\munf{I} &= {\sP'}\munf{I'_P} ({\sQ'}\munf{\bar I'_Q})^\top \\
@@
 \end{proof}
+
+\paragraph{Tensor Projections}
+
+We will now study the singular values of a tensor projection. 
+Let $\sP \in \Re^{d_1 \times \ldots \times d_m}$ be a $m$-th order
+  tensor. 
+For some vector $v \succ 0$, let $\sQ$ be the $m-1$-th order tensor
+  formed by projecting one of the axes of $\sP$ by $v$. 
+Without loss of generality, let the projected axes be the last one, $\sQ
+  \eqdef \sP(\cdot, \cdots, \cdot, v)$.
+Then, we have the following lemma linking the singular values of $\sQ$
+  and $\sP$.
+\begin{lemma}[Tensor projection preserves rank]
+  \label{lem:tensor-projection}
+  Let $\sP, \sQ$ and $v$ be defined as above. For any unfolding $I
+  \not\contains v$, 
+\begin{align*}
+  \sigma_{\max}(\sQ\munf{I}) &\le \sigma_{\max}(\sP\munf{I}) v_{\max} \\
+  \sigma_{\min}(\sQ\munf{I}) &\ge \sigma_{\min}(\sP\munf{I}) v_{\min}.
+\end{align*}
+In particular, $\rank(\sP\munf{I}) = \rank(\sQ\munf{I})$.
+\end{lemma}
+\begin{proof}
+  We will show that $\sQ\munf{I}$ is $\sP\munf{I}$ related by
+  a projection operator. When $v \succ 0$, this projection operator has
+  full column rank, entailing the result.
+
+  By definition of $\sQ$, for any $\vx$
+  \begin{align*}
+    \sQ[\vi] &= \sum_{k=1}^{d_m} \sP[\vi, k] v[k].
+  \end{align*}
+  Consider the unfolding along $I$,
+  \begin{align*}
+    \sQ\munf{I}[\vi;\vj] &= \sum_{k=1}^{d_m} \sP\munf{I}[\vi;\vj\ k] v[k].
+  \end{align*}
+
+  We can represent this as matrix multiplication by taking the tensor
+  product of $v$ with the $D_{\bar I \setminus \{m\}} \times D_{\bar
+  I \setminus \{m\}}$ identity matrix, $\bI$. Note that $\bI \otimes v$ is a $
+  D_{\bar I \setminus \{m\}} d_m \times D_{\bar I \setminus \{m\}}
+  = D_{\bar I } \times D_{\bar I \setminus \{m\}}$ matrix.
+  \begin{align*}
+    \sQ\munf{I}[\vi;\vj] 
+      &= \sum_{j'=\ones, k=1}^{\bd, d_m} \sP\munf{I}[\vi;\vj'\ k] (v[k] \delta[\vj';\vj]) \\
+    \sQ\munf{I}
+      &= \sP\munf{I} \bI \otimes v.
+  \end{align*}
+  If $v$ has no non-zero elements, $\bI \otimes v$ has full-column
+  rank, with singular values $v_1, \cdots, v_{d_m}$.
+\end{proof}

### 2014-05-09 doc/icml2014/directed.tex
Commented out assumptions proofs and touched up piecewise section

@@
       same domain, $[k]$, our work generalizes to hidden domains of
-      different sizes, provided \assumptionref{full-rank-plus} is satisfied.
+      different sizes, provided \assumptionref{full-rank} is satisfied.
       },
@@
   column rank and that $\pi^{(i)} \eqdef \Pr(h_i) \succ 0$.
-But $\mOpp{v}{i}$ is a product of tensors on the path from $h_i$ to
-  $x_v$, marginalizing out the rest of the graph.
-To make this property more interpretable, let us assume the following
-  property which we prove to be sufficient in
-  \appendixref{assumption-proof}:
-\begin{assumption} 
-\label{asm:full-rank-plus}
-The marginal distribution of every latent variable has full support:
-  $\forall{h_i \in H},~ \mPi{i} \succ 0$.
-For every clique $\sC \in \sG$ (including ones with observed variables),
-  and every conditional distribution $\Pr(h_c \given \Pa(h_c))$, where
-  $h_c \in \sC$, every {\em tensor unfolding} has full column rank, $k$.
-\end{assumption}
-In our directed grid example (\sectionref{directedExample}), for the
-  clique $\sC = \{ h_1, h_2 \}$ this condition implies $\pi = \Pr(h_1)
-  \succ 0$ and $T = \Pr(h_2 \mid h_1)$ has column rank $k$; for the clique
-  $\sC = \{ h_1, x_1 \}$, it implies $O = \Pr(x_1 \mid h_1)$ has column
-  rank $k$.
+% But $\mOpp{v}{i}$ is a product of tensors on the path from $h_i$ to
+%   $x_v$, marginalizing out the rest of the graph.
+% To make this property more interpretable, let us assume the following
+%   property which we prove to be sufficient in
+%   \appendixref{assumption-proof}:
+% \begin{assumption} 
+% \label{asm:full-rank-plus}
+% The marginal distribution of every latent variable has full support:
+%   $\forall{h_i \in H},~ \mPi{i} \succ 0$.
+% For every clique $\sC \in \sG$ (including ones with observed variables),
+%   and every conditional distribution $\Pr(h_c \given \Pa(h_c))$, where
+%   $h_c \in \sC$, every {\em tensor unfolding} has full column rank, $k$.
+% \end{assumption}
+% In our directed grid example (\sectionref{directedExample}), for the
+%   clique $\sC = \{ h_1, h_2 \}$ this condition implies $\pi = \Pr(h_1)
+%   \succ 0$ and $T = \Pr(h_2 \mid h_1)$ has column rank $k$; for the clique
+%   $\sC = \{ h_1, x_1 \}$, it implies $O = \Pr(x_1 \mid h_1)$ has column
+%   rank $k$.
 
@@
 
-From \citet{anandkumar13tensor} given \assumptionref{full-rank-plus}, we
+From \citet{anandkumar13tensor} given \assumptionref{full-rank}, we
   have that $\mOpphat{v}{i}$ converges to $\mOpp{v}{i}$ at a rate of
@@
   singular value of $\mOpp{v}{i}$:
-
 \begin{theorem}[Sample complexity for $\mOpp{v}{i}$]
@@
 
-If $\sigma_{\min}(\Pr(c \given \Pa(h_c))) \le \sigma$ for every clique
-  $\sC \in \sG$ and hidden variable $h_c \in \sC$,
-\begin{align*}
-  \|\mOpphat{1}{1} - \mOpp{1}{1}\|_F 
-    &\le 
-      O\left( 
-      \frac{k \log(1/\delta) {\pi\oft{1}}_{\max}/{\pi\oft{1}}_{\min}} 
-      {\sqrt{n} \sigma^{5t}} \right) \epsilon,
-\end{align*}
-where $t$ is the length of the shortest topological ordering from $h_1$
-to the unique parent of $x_1$, $h_t$.
+%If $\sigma_{\min}(\Pr(c \given \Pa(h_c))) \le \sigma$ for every clique
+%  $\sC \in \sG$ and hidden variable $h_c \in \sC$,
+%\begin{align*}
+%  \|\mOpphat{1}{1} - \mOpp{1}{1}\|_F 
+%    &\le 
+%      O\left( 
+%      \frac{k \log(1/\delta) {\pi\oft{1}}_{\max}/{\pi\oft{1}}_{\min}} 
+%      {\sqrt{n} \sigma^{5t}} \right) \epsilon,
+%\end{align*}
+%where $t$ is the length of the shortest topological ordering from $h_1$
+%to the unique parent of $x_1$, $h_t$.
 \end{theorem}
 
-\theoremref{sample-complexity}, proved in \appendixref{assumption-proof}
-  says that the number of samples required to estimate the view
-  $\mOpp{v}{i}$ grows exponentially in the {\em distance} between $x_v$
-  and $h_i$. 
+In general, $\sigma_{\min}(\mOpp{v}{i})$ can depend exponentially on the
+  path length between the hidden variable $h_i$ and $x_v$.
 Intuitively, this is expected because the influence from $h_i$ to $x_v$
@@
 
-\todo{Present sample complexity bounds for $Z_\sC$?}
+%\theoremref{sample-complexity}, proved in \appendixref{assumption-proof}
+%  says that the number of samples required to estimate the view
+%  $\mOpp{v}{i}$ grows exponentially in the {\em distance} between $x_v$
+%  and $h_i$.
+
+%\todo{Present sample complexity bounds for $Z_\sC$?}
 

### 2014-05-10 doc/icml2014/background.tex
Various minor corrections to camera-ready

@@
 Let $\Pa(a) \subset X \union H$ be the parents of a variable $a$ in a directed
-  model, and $\sN(a) \subset X \union H$ be the neighbors of $a$ in an
+  model, and 
+let $\sN(a) \subset X \union H$ be the neighbors of $a$ in an
   undirected model.
@@
   of $M$.
-For a tensor $T \in \Re^{d \times \ldots \times d}$ and a vector $\vec
-  i = (i_1, \cdots, i_\ell)$, define the projection $T[\vec i] = T_{i_1 \ldots i_\ell}$.
-Similarly, for a set $\sC$ of size $k$ and vector $\vec x \in \Re^d$, we
-  will use the notation $\vec x_\sC$ to get a vector in $\Re^k$ which selects
-  the appropriate elements of $\vec x$.
+For a tensor $T \in \Re^{d \times \ldots \times d}$ and a vector $\bi
+  = (i_1, \cdots, i_\ell)$, define the projection $T[\bi] = T[i_1 \ldots
+  i_\ell]$.
+Similarly, for a set $\sC$ of size $m$ and vector $\bx \in \Re^d$, we
+  will use the notation $\bx_\sC$ to get a vector in $\Re^m$ which selects
+  the appropriate elements of $\bx$.
 
@@
 For an $\ell$-th order tensor $T \in \Re^{d \times \ldots \times
-  d}$ and vectors $x_1, \cdots, x_\ell \in \Re^{d}$, define 
-  the partial application:
+  d}$ and vectors $\bx_1, \cdots, \bx_\ell \in \Re^{d}$, define 
+  the application:
 \begin{align*}
-  T(x_1, \cdots, x_\ell) 
+  T(\bx_1, \cdots, \bx_\ell) 
       %&= \sum_{i_1 = 1}^{d} \cdots \sum_{i_\ell = 1}^{d} 
-  &= \sum_{\bi}
-            T[i_1, i_2, \ldots, i_\ell] x_1[i_1] \cdots x_\ell[i_\ell].
+&= \sum_{\bi = \ones}^{\vec d}
+            T[i_1, i_2, \ldots, i_\ell] \bx_1[i_1] \cdots \bx_\ell[i_\ell].
 \end{align*}
@@
   &= \sum_{\bi}
-            T[\vec i] {M_1}[i_1,j_1] \cdots M_\ell[i_\ell, j_\ell].
+            T[\vec i] {M_1}[i_1,j_1] \cdots M_\ell[i_\ell, j_\ell],
 \end{align*}
+a tensor in $\Re^{k \times \ldots \times k}$.
 %where $\vec i = (i_1, \cdots, i_\ell)$ and $\vec j = (i_1, \cdots, j_\ell)$
@@
 \begin{align*}
-  \mO_{ij} &= \sum_h {\pi \oft 1}[h] {\mOpp{i}{1}}[\cdot,h] \otimes {\mOpp{j}{1}}[\cdot,h] \quad \forall i,j \in \{1,2,3\} \\
-  \mO_{123} &= \sum_h {\pi \oft 1}[h] {\mOpp{1}{1}}[\cdot,h] \otimes {\mOpp{2}{1}}[\cdot,h] \otimes  {\mOpp{3}{1}}[\cdot,h].
+  \mO_{ij} &= \sum_h {\pi \oft 1}[h] {\mOppt{i}{1}}[h] \otimes {\mOppt{j}{1}}[h] \quad \forall i,j \in \{1,2,3\} \\
+  \mO_{123} &= \sum_h {\pi \oft 1}[h] {\mOppt{1}{1}}[h] \otimes {\mOppt{2}{1}}[h] \otimes  {\mOppt{3}{1}}[h].
 \end{align*}

### 2014-05-10 doc/icml2014/directed.tex
Various minor corrections to camera-ready

@@
   describe error bounds for the finite sample regime; and
-(iii) we present the algorithm solely in terms of linear algebra operations; in
+(iii) we present the algorithm solely in terms of linear algebraic operations; in
   \sectionref{piecewise}, we show how we can get a more statistically
@@
 \end{property}
-
 For the rest of this section, we will only consider sub-graphs which are
@@
   by identifying a hidden variable, $h_1$ along with three conditionally
-  independent observed variables, $x^a_1, x^b_1, x^a_2$.
+  independent observed variables, $x^a_1, x^b_1, x^a_2$ using the tensor
+  factorization method.
   %followed by using \TensorFactorize to learn the conditional moments.
@@
 % DONE: move here because not central to story
-\paragraph{Remark.} We note that \propertyref{bottleneck} can be relaxed if some cliques
+\paragraph{Remarks.} We note that \propertyref{bottleneck} can be relaxed if some cliques
   share parameters.
@@
   be bottlenecks, and can omit the observations $x^b_2, x^b_3$ and $x^b_4$.
+
+The method also extends directly to case in which the observed variables
+  are real valued, while the hidden variables remain discrete. 
+In this scenario, the tensor factorization method of
+  \citet{anandkumar13tensor} recovers the expected conditional means,
+  $\mu_{vi} \eqdef \E[x_v | h_i]$ for an observed variable $x_v$ and
+  hidden variable $h_i$.
+\algorithmref{learnclique} and \algorithmref{directed} still apply and allow
+  us to recover clique marginals $Z_\sC$.
 %  \todo{make this a footnote if running out of space}
@@
   path length between the hidden variable $h_i$ and $x_v$.
-Intuitively, this is expected because the influence from $h_i$ to $x_v$
+This is expected because the influence from $h_i$ to $x_v$
   gets diluted for every latent variable it must factor through.

### 2014-05-10 doc/icml2014/undirected.tex
Various minor corrections to camera-ready

--- 2014-03-18 doc/icml2014/undirected.tex
@@
 Furthermore, given $\mOpp{v}{i}$ and $Z_\sC$ of any $\sC \ni v$, we can recover
-all marginals $\BP(\vx_\sC, \vh_\sC) = Z_\sC \cdot \mOpp[\vx_\sV]$\todo{clean up}.
+all marginals $\BP(\vx_\sC, \vh_\sC) = Z_\sC \cdot \mOppAll[\vx_\sV]$.%\todo{clean up}.
 Finally, the key step is that we can compute all the expected feature vectors from the marginals:
@@
 (it is violated, for example, if two components of $\phi(\vx_\sC, \vh_\sC)$ were identical).
-In general, we can identify the parameters up to kernel of $\nabla^2 A(\theta)$.}
+In general, we can identify the parameters up to the kernel of $\nabla^2 A(\theta)$.}
 
@@
 
-However, pseudolikelihood also replaces the log-partition function $A(\theta)$
+The pseudolikelihood also replaces the log-partition function $A(\theta)$
 with a sum over the variables in the neighbourhood: $A_a(\theta; x_{\sN(a)}) = \sum_{a' \in \{a\} \union \sN(a)} \exp(\theta^\top\phi_{\sN(a)}(a'))$,

### 2014-05-12 doc/icml2014/abstract.tex
Various minor corrections to camera-ready
gotten up to lemma 2

@@
 
-Recent work on the method of moments provide consistent estimates
-  for latent-variable models, avoiding local optima issues,
-  but are applicable only to certain types of graphical models.
-  On the other hand, pure likelihood objectives are difficult to optimize.
+Recent work on the method of moments enable
+  consistent parameter estimation, but
+  only for certain types of latent-variable models. %, avoiding local optima issues,
+  %but are applicable only to certain types of graphical models.
+On the other hand, pure likelihood objectives, though more universally applicable,
+  are difficult to optimize.
 In this work, we show that using the method of moments in conjunction
@@
 %This family includes loopy graphs with high treewidth.
-Specifically, we use tensor factorization to reveal partial information
+Specifically, we first use tensor factorization to reveal information
 about the hidden variables.  This renders the otherwise non-convex
-likelihood objective convex, and EM can then be used to attain the global optimum.
+likelihood objective convex, and EM can then be used to obtain the global optimum.
 %We use the method of moments to tie each hidden to some observed

### 2014-05-12 doc/icml2014/background.tex
gotten up to lemma 2

@@
   hidden variables $H$.
-Let $\Pa(a) \subset X \union H$ be the parents of a variable $a$ in a directed
-  model, and 
-let $\sN(a) \subset X \union H$ be the neighbors of $a$ in an
-  undirected model.
-We will use $x_1, x_2, \cdots, x_L \in X$ to denote observed variables and
-  $h_1, h_2, \cdots, h_M \in H$ to denote hidden variables.
+For a directed model,
+  let $\Pa(a) \subset X \union H$ be the parents of a variable $a$;
+  for an undirected model,
+  let $\sN(a) \subset X \union H$ be the neighbors of $a$.
+  Abusing notation slightly, we use $\sG$ to denote the set of cliques.
+  %(for directed models, these are $\sG = \{ \Pa(a) \union \{a\} : a \in X \union H \}$).
+We use $x_1, x_2, \dots, x_L \in X$ to denote observed variables and
+  $h_1, h_2, \dots, h_M \in H$ to denote hidden variables.
 For an integer $n$, let $[n] = \{ 1, \dots, n \}$.
 For ease of exposition, we assume that $x_v \in [d]$ for
-all $v$, and that $h_i \in [k]$ for all $i$.
-Let $\sX$ be the joint domain of the observed variables, i.e. $\sX \eqdef [d]^L$ and
+all $v \in [L]$, and that $h_i \in [k]$ for all $i \in [M]$.
+Let $\sX$ be the joint domain of the observed variables, i.e. $\sX \eqdef [d]^L$, and
 let $\sH$ be the joint domain of the hidden variables, i.e. $\sH \eqdef [k]^M$.
@@
 %We assume that the model $\sG$ is parametrized at every clique $\sC$.
-If $\sG$ is directed, the model parameters correspond to
-the conditional probability tables for each clique $(\Pa(a), a) \in \sG$.
-If $\sG$ is undirected, we assume a log-linear parametrization of each clique $\sC \in \sG$.
+For directed models,
+$p_\theta(\bx,\bh) = \prod_{a \in X \cup H} p_\theta(a \mid \Pa(a))$,
+where the parameters $\theta$ are the conditional probability tables.
+For undirected models,
+$p_\theta(\bx,\bh) \propto \prod_{\sC \in \sG} \exp(\theta^\top\phi_\sC(\bx_\sC, \bh_\sC))$,
+where $\theta$ are the canonical parameters of an exponential family.
+
+% PL: log-linear parametrization - can't say this for each clique: this is a global notion.
+%If $\sG$ is undirected, we assume a log-linear parametrization of each clique $\sC \in \sG$.
 
@@
 describe a method of moments approach for parameter estimation.
-The starting point is moments of the marginal distribution $\BP(\vx)$,
-and a bit of notation is necessary.
+First, a bit of notation is necessary.
+%The starting point is moments of the marginal distribution $\BP(\vx)$,
+%and a bit of notation is necessary.
 
@@
   of $M$.
-For a tensor $T \in \Re^{d \times \ldots \times d}$ and a vector $\bi
-  = (i_1, \cdots, i_\ell)$, define the projection $T[\bi] = T[i_1 \ldots
-  i_\ell]$.
-Similarly, for a set $\sC$ of size $m$ and vector $\bx \in \Re^d$, we
-  will use the notation $\bx_\sC$ to get a vector in $\Re^m$ which selects
-  the appropriate elements of $\bx$.
+For a tensor $T \in \Re^{d \times \cdots \times d}$ and a vector $\bi
+  = (i_1, \ldots, i_\ell)$, define the projection $T[\bi] = T[i_1, \ldots, i_\ell]$.
+Similarly, for a set $\sC$ of size $m$ and vector $v \in \Re^d$, we
+  will use the notation $v_\sC$ to get a vector in $\Re^m$ which selects
+  the appropriate elements of $v$.
 
@@
 For an $\ell$-th order tensor $T \in \Re^{d \times \ldots \times
-  d}$ and vectors $\bx_1, \cdots, \bx_\ell \in \Re^{d}$, define 
+  d}$ and vectors $v_1, \cdots, v_\ell \in \Re^{d}$, define 
   the application:
 \begin{align*}
-  T(\bx_1, \cdots, \bx_\ell) 
+  T(v_1, \cdots, v_\ell) 
       %&= \sum_{i_1 = 1}^{d} \cdots \sum_{i_\ell = 1}^{d} 
-&= \sum_{\bi = \ones}^{\vec d}
-            T[i_1, i_2, \ldots, i_\ell] \bx_1[i_1] \cdots \bx_\ell[i_\ell].
+% PL: this isn't great notation; reader can figure it out
+%&= \sum_{\bi = \ones}^{\vec d}
+  &= \sum_{\bi}
+            T[\bi] v_1[i_1] \cdots v_\ell[i_\ell].
 \end{align*}
@@
   &= \sum_{\bi}
-            T[\vec i] {M_1}[i_1,j_1] \cdots M_\ell[i_\ell, j_\ell],
+            T[\vec i] {M_1}[i_1,j_1] \cdots M_\ell[i_\ell, j_\ell].
 \end{align*}
-a tensor in $\Re^{k \times \ldots \times k}$.
+%a tensor in $\Re^{k \times \ldots \times k}$. % PL: not true, since [j] is applied
 %where $\vec i = (i_1, \cdots, i_\ell)$ and $\vec j = (i_1, \cdots, j_\ell)$
 
-We consider up to third-order moments:
+%with respect to the true data distribution $p_{\theta^*}(\bx,\bh)$;
+We will use $\Pr(\cdot)$ to denote various moment tensors
+constructed from the true data distribution $p_{\theta^*}(\bx,\bh)$:
+%We consider up to third-order moments:
 \begin{align*}
@@
 \end{align*}
-We will view $\mO_i, \mO_{ij}, \mO_{ijk}$ as tensors of
+Here, $\mO_i, \mO_{ij}, \mO_{ijk}$ are tensors of
   orders $1, 2, 3$ in $\Re^d, \Re^{d\times d}, \Re^{d \times d \times d}$.
@@
   that are independent conditioned on a hidden variable $h_1$.  
-The parameters of this model are the mixture probabilities $\pi\oft 1 \eqdef
+The parameters of this model are the prior probabilities $\pi\oft 1 \eqdef
   \Pr(h_1) \in \Re^k$ and the conditional moments $\mOpp{i}{1} \eqdef
@@
 The conditional moments $\mOpp{1}{1}$, $\mOpp{2}{1}$, $\mOpp{3}{1}$ have full column rank
-  $k$, and that $\pi\oft 1 \succ 0$.
+  $k$, and $\pi\oft 1 \succ 0$.
 \end{assumption}
@@
   \Re^{d \times k}$ such that $W^\top \mO_{12} W = I_{k \times k}$,
-  and uses it to transform $\mO_{123}$ into a symmetric orthogonal tensor.  
+  and uses $W$ to transform $\mO_{123}$ into a symmetric orthogonal tensor.  
 Then a robust tensor power method is applied to extract the eigenvectors
 of the whitened $\mO_{123}$; unwhitening yields the columns of $\mOpp{3}{1}$.
-  $\mOpp{1}{1}$ and $\mOpp{2}{1}$ can then be recovered with a few matrix operations.
+The other parameters can be recovered with a few additional matrix operations.
+%$\mOpp{1}{1}$ and $\mOpp{2}{1}$ can then be recovered with a few matrix operations.
 

### 2014-05-12 doc/icml2014/directed.tex
gotten up to lemma 2

@@
   In general directed models, the conditional moments $\mOpp{v}{i} \eqdef \Pr(x_v \mid h_i)$
-  do not correspond directly to the underlying parameters.
+which we were able to recover from {\TensorFactorize}, are not the underlying parameters.
   %Once we recover the conditional moments $\mOpp{v}{j} \eqdef \Pr(x_v
@@
   %graphical models.
-  Let us see how we can estimate the parameters with an example (Section~\ref{sec:directedExample}).
+  Let us first consider an example (Section~\ref{sec:directedExample});
   Section~\ref{sec:directedGeneral} will describe the algorithm in full generality.
@@
 (i) we describe our algorithm entirely in the context of directed
-  graphical models, though it generalizes to factor graphs;
+  graphical models, though it generalizes to undirected models;
 (ii) we derive our algorithm assuming infinite data; in \sectionref{sampleComplexity} we 
   describe error bounds for the finite sample regime; and
-(iii) we present the algorithm solely in terms of linear algebraic operations; in
-  \sectionref{piecewise}, we show how we can get a more statistically
-  efficient estimator using composite likelihoods.
+(iii) we present the algorithm solely in terms of linear algebraic operations;
+  \sectionref{piecewise} provides a statistically
+  more efficient estimator using composite likelihood.
 
-\subsection{Example: Directed grid model}
+\subsection{Example: directed grid model}
 \label{sec:directedExample}
@@
 The model has eight observed variables $x^a_1, x^b_1 \cdots, x^a_4, x^b_4$ and four
-  hidden variables $h_1, \cdots, h_4$.
+  hidden variables $h_1, \ldots, h_4$.
 The parameters of this model are the conditional probability tables
@@
 =  \Pr(x^b_i | h_i) \in \Re^{d \times k}$. 
-Assume $O$ and $T$ have full column rank.
+%Assume $O$ and $T$ have full column rank. % PL: if we put this in, we have to say something about V, which is too complex.
 
@@
 Note that the observed variables $x^a_1, x^b_1, x^a_2$ are
-  conditionally independent given $h_1$; we can use
-  $\TensorFactorize$ (\sectionref{setup}) to recover $O$.
+  conditionally independent given $h_1$; we can therefore use
+  $\TensorFactorize$ from \sectionref{setup} to recover $O$.
 
@@
 The moments of $x^a_1$, $\mO_1 \eqdef \Pr(x^a_1)$ are directly related to
-  $\pi$ by $\mO_1 = O \pi$. 
-By \assumptionref{full-rank}, $O$ has full column rank and thus can be
-  inverted to recover $\pi$: $\pi = O\pinv  \mO_1$.
+  $\pi$ by a linear system: $\mO_1 = O \pi$. 
+% PL: weird to refer to refer to this assumption from previous section
+% where it's talking about another model.
+%By \assumptionref{full-rank}, $O$ has full column rank and thus can be
+  %inverted to recover $\pi$: $\pi = O\pinv  \mO_1$.
+If $O$ has full column rank, we can recover $\pi$ by taking the pseudoinverse: $\pi = O\pinv  \mO_1$.
 
@@
 \end{align*}
-$T$ can be recovered from the $\mH_{12}$ by suitable renormalization.
+$T$ can be recovered from the $\mH_{12}$ by renormalizing the columns.
 
@@
 
-Note that in each of the above cases,
-  for any clique $\sC$, each hidden variable $h_i \in \sC$ has its own
-  view which is conditionally independent given the other hidden variables in
-  the clique.
+%Let us now generalize the intuitions of the directed grid to general graphs.
+The intuitions of the directed grid generalize readily to general graphs.
+% PL: this prose is redundant with property
+%Note that in each of the above cases,
+  %for any clique $\sC$, each hidden variable $h_i \in \sC$ has its own
+  %view which is conditionally independent given the other hidden variables in
+  %the clique.
 %Fundamentally, algorithm takes advantage of having a view for every
   %hidden variable in a clique to help identify its hidden state.
-We capture this general property as follows:
+The key property required for our algorithm to work is as follows:
+%We capture this general property as follows:
 \begin{property}(Exclusive views)
   \label{prop:exclusive-views}
-A hidden clique $\sC \in \sG$ is said to possess the \textbf{exclusive views property} if for
-  every hidden variable $h_i \in \sC$, there is some observed variable,
+A hidden clique $\sC \in \sG$ is said to possess the \emph{exclusive views property} if for
+  every hidden variable $h_i \in \sC$, (i)
+  there exists some observed variable,
   $x_{v}$ which is conditionally independent of the rest of the clique
-  given $h_i$, and whose conditional moment $\mOpp{v}{i} \eqdef
+  given $h_i$, and (ii) the conditional moments $\mOpp{v}{i} \eqdef
   \Pr(x_{v} \mid h_i)$ can be estimated consistently.
-We call $x_v$ the exclusive view of $h_i$ in $\sC$.
-
-This property also extends to any connected sub-graph $\sG' \in \sG$.
+We call $x_v$ an exclusive view of $h_i$ in $\sC$.
+If every clique has exclusive views, then we say that $\sG$ does as well.
 \end{property}
-For the rest of this section, we will only consider sub-graphs which are
-  cliques. 
-In \sectionref{undirected}, we will show how the generalization
-  to sub-graphs allows us to efficiently estimate the pseudo-likelihood.
+% PL: to save space
+%This property also extends to any connected subgraph $\sG' \in \sG$.
+%For the rest of this section, we will only consider subgraphs which are
+  %cliques. 
+%In \sectionref{undirected}, we will show how the generalization
+  %to subgraphs allows us to efficiently estimate the pseudo-likelihood.
 
@@
   distribution of the clique.
-Consider any hidden clique $\sC = \{i_1, \cdots, i_m\}$ for which this property holds. Let
-  $x_{v_j}$ be any exclusive view for $h_{i_j}$ in $\sC$ and $\sV
-  = \{v_1, \cdots, v_m\}$ be a set of exclusive views for the clique $\sC$.
+Consider any hidden clique $\sC = \{i_1, \ldots, i_m\}$ with the exclusive views property. Let
+  $x_{v_j}$ be any exclusive view for $h_{i_j}$ in $\sC$ and define $\sV
+  = \{v_1, \ldots, v_m\}$. % be a set of exclusive views for the clique $\sC$.
 We can write down the moments $\mO_\sV$ as follows:
@@
       %\Pr(\Sh{\sC}) \Pr(\Sx{\sV} \given \Sh{\sC}) \\
-      &= \sum_{\vh \in \sH} \Pr(\Sh{\sC}) 
+      &= \sum_{\vh_\sC} \Pr(\Sh{\sC}) 
           \Pr(x_{v_1} | h_{i_1}) \cdots \Pr(x_{v_m} | h_{i_m}) \\
@@
 \end{align*}
-Finally, the conditional probability tables for $\sC$ can easily be obtained from
-  $Z_\sC$ via renormalization.
 \algorithmref{learnclique} summarizes the procedure, \LearnClique.
+Given $Z_\sC$, the conditional probability tables for $\sC$ can easily be
+obtained via renormalization (see \sectionref{sampleComplexity}).
 
@@
     \ENSURE Marginal distribution of the clique $Z_\sC$.
-      \STATE Identify exclusive views $\Sx{\sV} = \{x_{v_1}, \cdots, x_{v_m}\}$.
-      \STATE Return $\mH_\sC \gets \mO_{\sV}( \mOpp{v_1}{i_i}\pinv, \cdots, {\mOpp{v_m}{i_m}}\pinv )$.
+      \STATE Identify exclusive views $\Sx{\sV} = \{x_{v_1}, \dots, x_{v_m}\}$.
+      \STATE Return $\mH_\sC \gets \mO_{\sV}( \mOpp{v_1}{i_i}\pinv, \dots, {\mOpp{v_m}{i_m}}\pinv )$.
   \end{algorithmic}
 \end{algorithm}
-Of course, for $Z_{\sC}$ to be identifiable, we need various rank conditions,
-which will be established in \sectionref{sampleComplexity}.
+% PL: sample complexity, not rank
+%Of course, for $Z_{\sC}$ to be identifiable, we need various rank conditions,
+%which will be established in \sectionref{sampleComplexity}.
 
@@
 We have shown how to estimate the parameters of any clique that possesses the exclusive
-  views property (\propertyref{exclusive-views}). But in a general graph, which cliques have this property?
-In the example above, we were able to guarantee \propertyref{exclusive-views}
-  by identifying a hidden variable, $h_1$ along with three conditionally
-  independent observed variables, $x^a_1, x^b_1, x^a_2$ using the tensor
-  factorization method.
+  views property (\propertyref{exclusive-views}).
+  But in a general graph, which cliques have this property?
+  To provide intuition, we will provide more interpretable sufficient conditions
+  and examples.
+
+%In the example above, we were able to guarantee \propertyref{exclusive-views}
+  %by identifying a hidden variable, $h_1$ along with three conditionally
+  %independent observed variables, $x^a_1, x^b_1, x^a_2$ using the tensor
+  %factorization method.
   %followed by using \TensorFactorize to learn the conditional moments.
-We define such a hidden variable to be a bottleneck:
+First, let us revisit the definition of a bottleneck:
 \begin{definition}[Bottleneck]
-  A hidden variable $h$ is said to be a \textbf{bottleneck} if there
-  exists at least three observed variables, or views, $x_1, x_2, x_3$
-  that are conditionally independent given $h$. 
-  Let $\sV_{h}$ denote a set of views for $h$ (not necessarily unique).
+  A hidden variable $h_i$ is said to be a \textbf{bottleneck} if there
+  exists at least three observed variables (views), $x_{v_1}, x_{v_2}, x_{v_3}$
+  that are conditionally independent given $h_i$,
+  and $\mOpp{v_1}{i},\mOpp{v_2}{i},\mOpp{v_3}{i}$ have full column rank $k$.
+  % PL: need to add rank constraint; otherwise, doesn't imply exclusive views
+  Let $\sV_{h_i}$ denote any set of views for $h_i$.
 \end{definition}
@@
   %graphical model where every hidden variable is a bottleneck:
-We will prove that the following simple property guarantees that all cliques
-have the exclusive views property:
+Now define the following property:
 \begin{property}[Uniformly bottlenecked]
   \label{prop:bottleneck}
-  Every hidden variable is a bottleneck.
+  A graphical model $\sG$ is \emph{uniformly bottlenecked} if
+  every hidden variable in $\sG$ is a bottleneck.
 \end{property}
+Note that two hidden variables are allowed to have overlapping views.
+%If every hidden variable is a bottleneck, then we can estimate $\mOpp{v}{i}$
+It is clear that having uniform bottlenecks implies part (ii)
+of the exclusive views property (\propertyref{exclusive-views}) via \TensorFactorize,
+but we will show that part (i) follows as well.
+
+%We will prove that the following simple property guarantees that all cliques
+%have the exclusive views property:
 
@@
 
-Before we establish this relationship,
-we describe a reduction of any graphical model to
+%Before we establish this relationship,
+First, we describe a reduction of any graphical model to
   a canonical form to make our arguments simpler.
@@
   \providecommand{\hp}{\ensuremath{h_\text{\rm new}}}
-
   Let $x_v$ be an observed variable with parents $\Pa(x_v)$ and children $\text{Ch}(x_v)$.
@@
   parents $\Pa(x_v)$ and children $\text{Ch}(x_v) \union \{x_{v_1}, x_{v_2}, x_{v_3}\}$,
-  where $x_{v_1},x_{v_2},x_{v_3}$ are three copies of $x$
+  where $x_{v_1},x_{v_2},x_{v_3}$ are three copies of $x_v$
   (\figureref{reduction}).
-  The domain of $\hp$ is the same as $x_v$\footnote{
-      Though we have assumed that all hidden variables share the
-      same domain, $[k]$, our work generalizes to hidden domains of
-      different sizes, provided \assumptionref{full-rank} is satisfied.
-      },
-    and $\mOpp{v}{\textrm{new}} = I$.
+  Define $\Pr(\hp \mid \Pa(x_v)) = \Pr(x_v \mid \Pa(x_v))$ and
+  $\Pr(x_{v_j} \mid \hp) = I$.
+  %$\Pr(x_{v_j} \mid \hp) = \left[ \begin{array}{c} I_{k \times k} \\ 0_{(d-k) \times k} \end{array} \right]$.
+  %The domain of $\hp$ is the same as $x_{v_1},x_{v_2},x_{v_3}$.
+  % PL: cut to save space
+  %\footnote{
+      %Though we have assumed that all hidden variables share the
+      %same domain, $[k]$, our work generalizes to hidden domains of
+      %different sizes, provided \assumptionref{full-rank} is satisfied.
+      %},
+    %and $\mOpp{v}{\textrm{new}} = I$.
   Then, there is a one-to-one correspondence between every value of
@@
 
-\begin{lemma}[Bottlenecks guarantee exclusive views]
+Now we establish the link between bottlenecks and exclusive views
+(see \appendixref{exclusive-views} for the complete proof):
+\begin{lemma}[Uniformly bottlenecked implies exclusive views]
   \label{lem:bottleneck-views}  
-Let $\sG' = \{ h_{i_1}, \cdots, h_{i_m} \} \subseteq \sG$ be
-  a connected sub-graph of hidden variables. If every hidden variable
-  $h_i \in \sG'$ is a bottleneck, then they have an exclusive view.
+% PL: simplify to cliques (this is also bad notation, since \sG is a set of cliques, not nodes)
+%Let $\sG' = \{ h_{i_1}, \dots, h_{i_m} \} \subseteq \sG$ be
+  %a connected subgraph of hidden variables.
+  %If every hidden variable
+  %$h_i \in \sG'$ is a bottleneck, then $\sG'$ has the exclusive views property.
+If a graphical model $\sG$ is uniformly bottlenecked,
+then $\sG$ has the exclusive views property.
 \end{lemma}
-\begin{proof}
-  The independence conditions required by the bottleneck condition
-  effectively guarantee that atleast one of the three views for
-  a bottleneck be exclusive. 
-  \appendixref{exclusive-views} provides a complete proof.
-\end{proof}
+% PL: save space, doesn't add much
+%\begin{proof}
+%  The independence conditions required by the bottleneck condition
+%  effectively guarantee that at least one of the three views for
+%  a bottleneck be exclusive. 
+%  \appendixref{exclusive-views} provides a complete proof.
+%\end{proof}
 
@@
         \STATE Apply $\TensorFactorize$ to learn conditional moments
-        $\mOpp{v}{i}$ for every $\sV_{h_i}$.
+        $\mOpp{v}{i}$ for every $v \in \sV_{h_i}$.
 
@@
 %      \COMMENT{\textbf{Step 2:} Recover clique potentials from the piecewise likelihood.}
-\FOR{every clique $\sC = \{h_{i_1}, \cdots, h_{i_m}\} \in \sG$} 
+\FOR{every clique $\sC = \{h_{i_1}, \ldots, h_{i_m}\} \in \sG$} 
 \STATE Apply $\LearnClique$ to learn the marginals $\mH_\sC$.

### 2014-05-12 doc/icml2014/exclusive-views.tex
Added figure to assumption proof
gotten up to lemma 2

@@
-\subsection{\lemmaref{lem:bottleneck-views}}
+\subsection{\lemmaref{bottleneck-views}}
 \label{app:exclusive-views}
@@
   exclusive views. We provide the proof below.
+  We generalize the lemma so that it applies
+  on any connected subgraph, not just cliques.
+  Also, the statement is made in terms of subgraphs,
+  not the entire graph.
+
 \begin{lemma*}[Bottlenecks guarantee exclusive views]
   %\label{lem:bottleneck-views}  
-Let $\sG' = \{ h_{i_1}, \cdots, h_{i_m} \} \subseteq \sG$ be
-  a connected sub-graph of hidden variables. If every hidden variable
-  $h_i \in \sG'$ is a bottleneck, then they have an exclusive view.
+Let $\sG' = \{ h_{i_1}, \ldots, h_{i_m} \} \subseteq \sG$ be
+  a connected subgraph of hidden variables. If every hidden variable
+  $h_i \in \sG'$ is a bottleneck, then $\sG'$ has the exclusive view property.
 \end{lemma*}
+
 \begin{proof}
@@
   to be active. This allows us to reason about independence simply by
-  checking whether or not two nodes are connected or not. Note that this
-  proof applies for undirected graphs as well.
+  checking whether or not two nodes are connected or not.
+  Note that this proof applies for undirected graphs as well.
 
@@
     $h$.
-  A view $x \in \sV_{h}$ is exclusive for $h$ iff $x \not\in \sV_{h'} ~ \forall h'
+  A view $x \in \sV_{h}$ is exclusive for $h$ iff $x \not\in \sV_{h'}$ for all $h'
   \in {\sG'}^-$.
@@
     $x_1$ is connected to $h_1$ and $x_2$ is connected to $h_2$.
-  By participation in the sub-graph, either $h_1 = h_2$ or $h_1$ is
+  By participation in the subgraph, either $h_1 = h_2$ or $h_1$ is
     connected to $h_2$, implying that $x_1$ and $x_2$ are connected
@@
   Algorithmically, finding the unique view for $h$ can be done by
-  subtracting the views of all other variables in the sub-graph from
+  subtracting the views of all other variables in the subgraph from
   $\sV_h$,i.e. $x_{v_h} \in \sV_h \setminus \Union_{h' \in \sG'} \sV_{h'}$

### 2014-05-12 doc/icml2014/intro.tex
Various minor corrections to camera-ready
gotten up to lemma 2

@@
 However, learning these models remains a difficult problem due to the
-non-convexity of the negative log-likelihood.  Local methods such as Expectation Maximization (EM)
+non-convexity of the negative log-likelihood.  Local methods such as expectation maximization (EM)
 are the norm, but are susceptible to local optima.
@@
 These methods exploit the linear algebraic properties of the model to
-factorize the moments into parameters, providing strong theoretical guarantees.
+factorize moments of the observed data distribution into parameters,
+providing strong theoretical guarantees.
 However, they apply to a limited set of models, and are thus
@@
  % \todo{need three steps: tensor factorization, optimize composite likelihood, recover model parameters;}
-  Overview of our approach: we first use tensor factorization to learn the conditional moments for each hidden variable.
-  Then we optimize a composite likelihood to recover the hidden clique marginals,
-  given which we can extract the model parameters.
+  Overview of our approach:
+  we first use tensor factorization to learn the \emph{conditional moments}
+  for each hidden variable.
+  Second, we optimize a composite marginal likelihood to recover the \emph{hidden clique marginals},
+  from which we solve for the \emph{model parameters}.
   }
@@
 In this paper,
-we show that a much broader class of discrete directed and undirected graphical models can be consistently estimated,
-specifically those in which each hidden variable has three conditionally
+we show that a much broader class of discrete directed and undirected graphical models can be consistently estimated:
+specifically those in which \emph{each} hidden variable has three conditionally
 independent observed variables (``views'')---see \propertyref{bottleneck}.
-The key idea is to leverage the spectral method of moments,
+Our key idea is to leverage the method of moments,
 not to directly provide a consistent parameter estimate as in previous work,
-but as partial constraints on a likelihood-based objective.
+but as constraints on a likelihood-based objective.
 %For a broad class of directed and undirected graphical models
@@
 % DONE: try to talk about the approach in terms of three steps
-Our approach is illustrated in \figureref{approach}. 
-First, we identify the views for each hidden variable $h_i$ (for example,
+The essence of our approach is illustrated in \figureref{approach}. 
+First, we identify three views for each hidden variable $h_i$ (for example,
 $x_1^a$, $x_1^b$ and $x_3^a$ are conditionally independent given $h_1$) and use
 the tensor factorization algorithm of
-\citet{anandkumar13tensor} to identify the \emph{conditional
-moments} $\Pr[x_i^a \mid h_i]$ and $\Pr[x_i^b \mid h_i]$ for each $i$.
+\citet{anandkumar13tensor} to estimate the \emph{conditional
+moments} $\Pr(x_i^a \mid h_i)$ and $\Pr(x_i^b \mid h_i)$ for each $i$ (\sectionref{directed}).
 Second, we optimize a \emph{composite marginal likelihood} to recover the marginals over
@@
 but given the conditional moments, we obtain a convex objective,
-which can be globally optimized using EM.
-Finally, the hidden marginals yield the model parameters
-via simple post-processing in the directed case and convex optimization in the undirected case.
+which can be globally optimized using EM (\sectionref{piecewise}).
+So far, our method has relied only on the conditional independence
+structure of the model and applies generically to both directed
+and undirected models.
+The final step of turning hidden marginals into model parameters
+requires some specialization.
+In the directed case, this is simple post-processing;
+in the undirected case, we need to solve another convex optimization problem (\sectionref{undirected}).
 %Given these factors, we show that there exists a set of observed
@@
   %consistently recover the true parameters of the model.
-In \sectionref{directed}, we detail our algorithm and
-show that our algorithm yields consistent parameter estimates
-under non-degenerate rank assumptions on the conditional moments.
+%In \sectionref{directed}, we describe our algorithm in more detail and
+%show that our algorithm yields consistent parameter estimates
+%under non-degenerate rank assumptions.
 %describe the

### 2014-05-12 doc/icml2014/directed.tex
minor

@@
 %We capture this general property as follows:
-\begin{property}(Exclusive views)
+\begin{property}[Exclusive views]
   \label{prop:exclusive-views}
@@
 We call $x_v$ an exclusive view of $h_i$ in $\sC$.
-If every clique has exclusive views, then we say that $\sG$ does as well.
+If every clique has the exclusive views property, then we say that $\sG$ does as well.
 \end{property}
@@
   share parameters.
-In the directed grid model, we can identify the conditional moments $O$ from just
-  $x^a_1, x^b_1$ and $x^a_2$.
-  Therefore, we do not require that $h_2, h_3$ or $h_4$
-  be bottlenecks, and can omit the observations $x^b_2, x^b_3$ and $x^b_4$.
+For example, in the directed grid model, we can recover the conditional moments $O$ from
+  $x^a_1, x^b_1$ and $x^a_2$ with $h_1$ as the bottleneck.
+  Therefore, $h_2, h_3$ and $h_4$
+  need not be bottlenecks, and we can omit the observations $x^b_2, x^b_3$ and $x^b_4$.
 
-The method also extends directly to case in which the observed variables
-  are real valued, while the hidden variables remain discrete. 
-In this scenario, the tensor factorization method of
+Our method also extends directly to case in which the observed variables
+  are real-valued, while the hidden variables remain discrete. 
+In this setting, the tensor factorization method of
   \citet{anandkumar13tensor} recovers the expected conditional means,
-  $\mu_{vi} \eqdef \E[x_v | h_i]$ for an observed variable $x_v$ and
+  $\mu_{vi} \eqdef \E(x_v | h_i) \in \Re^{d \times k}$ for each observed variable $x_v$ and
   hidden variable $h_i$.
@@
 In the HMM (\figureref{examples-hmm}), the parameters
-are $O \eqdef \Pr(x_i|h_i)$  and that $T \eqdef \Pr(h_{i+1} | h_i)$
+are $O \eqdef \Pr(x_i|h_i)$ and $T \eqdef \Pr(h_{i+1} | h_i)$
 for all $i$. % (i.e. we have parameter sharing).
@@
   sequence are not bottlenecks, they still have exclusive views ($x_1$ and
-  $x_M$ respectively) whose parameters we know because they share
-  parameters, $O$.
+  $x_M$, respectively)
+  due to parameter sharing.
+  %whose parameters we know because they share parameters, $O$.
 %We first use the bottleneck $h_2$ with views $x_1, x_2, x_3$ to estimate $O$,
@@
 In the latent tree model (\figureref{examples-tree}), the parameters
-are $\Pr(h_i) = \pi$, $\Pr(h_i | h_1) = T$ for $i \in \{2,3,4\}$
-and $\Pr(x^a_i | h_i) = \Pr(x^b_i | h_i) = O$ for $i \in \{2,3,4\}$.
+are $\pi \eqdef \Pr(h_i)$, $T \eqdef \Pr(h_i | h_1)$ for $i \in \{2,3,4\}$,
+and $O \eqdef \Pr(x^a_i | h_i) = \Pr(x^b_i | h_i)$ for $i \in \{2,3,4\}$.
 Note that while $h_1$ is not directly connected to an observed variable,
@@
 
-The simplest example of a model that cannot be recovered by our algorithm is
-  a mixture model with a single view, i.e. \figureref{three-view} with
-  just the variables $h_1$ and $x_1$.
-However, without further assumptions the model itself is not
-  identifiable\footnote{To see this, observe that probability-mass
-  can be arbitrarily exchanged between $\Pr(h_1)$ and $\Pr(x_1 | h_1)$ while preserving the marginals $\Pr(x_1)$}.
-A class of binary-valued noisy-or networks
-(see \figureref{examples-noisy-or} for an example)
-which do not possess the
-  bottleneck property are nonetheless identifiable;
-  see \citet{halpern13noisyor} for an algorithm.
+%In a certain sense, having three views is necessary for identifiability;
+%two views is certainly not enough.
+% PL: perhaps too simple, saving space
+%The simplest example of a model that cannot be recovered by our algorithm is
+%  a mixture model with a single view, i.e. \figureref{three-view} with
+%  just the variables $h_1$ and $x_1$.
+%However, without further assumptions the model itself is not
+%  identifiable\footnote{To see this, observe that probability-mass
+%  can be arbitrarily exchanged between $\Pr(h_1)$ and $\Pr(x_1 | h_1)$ while preserving the marginals $\Pr(x_1)$}.
+%A class of binary-valued noisy-or networks
+%(see \figureref{examples-noisy-or} for an example)
+%which do not possess the
+%  bottleneck property are nonetheless identifiable;
+%  see \citet{halpern13noisyor} for an algorithm.
+%Having three views is in some sense necessary for the uniqueness of tensor factorization.
+There are certainly models which are identifiable but do not have exclusive views.
+For example, \figureref{examples-noisy-or} shows
+  a binary-valued noisy-or network which can be
+  learned by the algorithm of \citet{halpern13noisyor},
+  but does not satisfy the exclusive views property.
 
@@
 To apply $\TensorFactorize$, we need that each $\mOpp{v}{i}$ has full
-  column rank and that $\pi^{(i)} \eqdef \Pr(h_i) \succ 0$.
+  column rank and that each $\pi^{(i)} \eqdef \Pr(h_i) \succ 0$.
 % But $\mOpp{v}{i}$ is a product of tensors on the path from $h_i$ to

### 2014-05-12 doc/icml2014/assumption-proof.tex
Updated sigma_max, sigma_min to be sigma_1, sigma_k

@@
   \begin{align*}
-    \sigma_{\min}(\mYpp{c}{\Pa(c)}\munf{I}) &\ge \max_{i\in[\ell]}\{\sigma_{\min}(Z_{\sC_i}\munf{I_i})\},
+    \sigma_{k}(\mYpp{c}{\Pa(c)}\munf{I}) &\ge \max_{i\in[\ell]}\{\sigma_{k}(Z_{\sC_i}\munf{I_i})\},
   \end{align*}
@@
 \begin{align*}
-  \sigma_{\min}(Z_{\sC}\munf{I}) &\ge \sigma_{\min}(Z_{\sC_1 \union \sC_2}\munf{I})v_{\min} \\
-  \sigma_{\min}(Z_{\sC_1 \union \sC_2}\munf{I}) &\ge
-  \sigma_{\min}(Z_{\sC}\munf{I})v_{\max}\inv.
+  \sigma_{k}(Z_{\sC}\munf{I}) &\ge \sigma_{k}(Z_{\sC_1 \union \sC_2}\munf{I})v_{k} \\
+  \sigma_{k}(Z_{\sC_1 \union \sC_2}\munf{I}) &\ge
+  \sigma_{k}(Z_{\sC}\munf{I})v_{\max}\inv.
 \end{align*}
 
-In this case, we get that $\sigma_{\min}(Z_{\sC_1 \union \sC_2}\munf{I})
-= \sigma_{\min}(Z_{\sC}\munf{I})v_{\max}\inv$.
+In this case, we get that $\sigma_{k}(Z_{\sC_1 \union \sC_2}\munf{I})
+= \sigma_{k}(Z_{\sC}\munf{I})v_{\max}\inv$.
 
@@
 \begin{align*}
-  \sigma_{\min}( (A \times_C B)\munf{I} )
-    &\ge \sigma_{\min}(A\munf{I_A}) \sigma_{\min}(B\munf{I_B}).
+  \sigma_{k}( (A \times_C B)\munf{I} )
+    &\ge \sigma_{k}(A\munf{I_A}) \sigma_{k}(B\munf{I_B}).
 \end{align*}
@@
 $\mOpp{v}{i}$, which depends mainly on smallest singular value,
-$\sigma_{\min}(\mOpp{v}{i})$.
+$\sigma_{k}(\mOpp{v}{i})$.
 
@@
 \begin{align*}
-  \sigma_{\min}(\sP) &\ge \sigma_{\min}(\mOpp{v}{t}) 
-        \prod_{c \in [t]} \sigma_{\min}(\mYpp{c}{\Pa(h_c)}),
+  \sigma_{k}(\sP) &\ge \sigma_{k}(\mOpp{v}{t}) 
+        \prod_{c \in [t]} \sigma_{k}(\mYpp{c}{\Pa(h_c)}),
 \end{align*}
@@
       \frac{{\pi\oft{1}}_{\max}/{\pi\oft{1}}_{\min}} 
-      {(\sigma_{\min}(\mOpp{1}{1}) \sigma_{\min}(\mOpp{1}{2}))^{5/2}} \right) \epsilon.
+      {(\sigma_{k}(\mOpp{1}{1}) \sigma_{k}(\mOpp{1}{2}))^{5/2}} \right) \epsilon.
 \end{align*}
 
-Furthermore, if $\sigma_{\min}(\mYpp{c}{\Pa(h_c)}) \le \sigma$ for every
+Furthermore, if $\sigma_{k}(\mYpp{c}{\Pa(h_c)}) \le \sigma$ for every
   such

### 2014-05-12 doc/icml2014/directed.tex
Updated sigma_max, sigma_min to be sigma_1, sigma_k
Updated uniformly bottlenecked and the non-degeneracy assumption

@@
 \end{property}
+
+Next, we state the non-degeneracy assumptions under which our general
+  algorithm is guaranteed to obtain consistent estimates.
+\begin{assumption}[Non-degeneracy]
+  \label{asm:non-degeneracy}
+  (i) The marginal distribution of each hidden variable $h_i$ have
+  full support: $\mPi{i} \succ 0$. 
+  (ii) The conditional moments $\mOpp{v}{i}$ have full column rank $k$.
+\end{assumption}
+
+Intuitively, we need to ensure that every marginal distribution
+  $\Pr(x_v)$ can be separated into $k$ different conditional distributions
+  $\Pr(x_v | h_i)$. 
+The full rank assumption asks that the conditional distributions are
+  linearly independent; if this were not the case, the conditional
+  distribution for a particular latent state could be imitated by the
+  combination of several other latent states.
+
 % PL: to save space
@@
   \label{prop:bottleneck}
-  A graphical model $\sG$ is \emph{uniformly bottlenecked} if
-  every hidden variable in $\sG$ is a bottleneck.
+  A clique $\sC$ is \emph{uniformly bottlenecked} if
+  every hidden variable in $\sC$ is a bottleneck. 
+  If every clique $\sC \in \sG$ is uniformly bottlenecked, then we say
+  that the graph $\sG$ is uniformly bottlenecked as well.
 \end{property}
@@
 
-To apply $\TensorFactorize$, we need that each $\mOpp{v}{i}$ has full
-  column rank and that each $\pi^{(i)} \eqdef \Pr(h_i) \succ 0$.
 % But $\mOpp{v}{i}$ is a product of tensors on the path from $h_i$ to
@@
 
-Intuitively, we need to ensure that every marginal distribution
-  $\Pr(x_v)$ can be separated into $k$ different conditional distributions
-  $\Pr(x_v | h_i)$. 
-The full rank assumption asks that the conditional distributions are
-  linearly independent; if this were not the case, the conditional
-  distribution for a particular latent state could be imitated by the
-  combination of several other latent states.
-
 %\todo{did they prove this; if so just say "they show", not "we can show"}
@@
 
-From \citet{anandkumar13tensor} given \assumptionref{full-rank}, we
-  have that $\mOpphat{v}{i}$ converges to $\mOpp{v}{i}$ at a rate of
-  $n^{-\frac12}$ with a constant that depends polynomially on the $k$-th
-  singular value of $\mOpp{v}{i}$:
+From \citet{anandkumar13tensor} given the non-degeneracy assumptions
+  \assumptionref{non-degeneracy}, we have that $\mOpphat{v}{i}$
+  converges to $\mOpp{v}{i}$ at a rate of $n^{-\frac12}$ with a constant
+  that depends polynomially on the $k$-th singular value of
+  $\mOpp{v}{i}$:
 \begin{theorem}[Sample complexity for $\mOpp{v}{i}$]
@@
       \frac{k \log(1/\delta) {\pi\oft{1}}_{\max}/{\pi\oft{1}}_{\min} } 
-      {\sqrt{n}(\sigma_{\min}(\mOpp{1}{1}) \sigma_{\min}(\mOpp{1}{2}))^{5/2}} \right) \epsilon.
+      {\sqrt{n}(\sigma_{k}(\mOpp{1}{1}) \sigma_{k}(\mOpp{1}{2}))^{5/2}} \right) \epsilon.
 \end{align*}
@@
 
-In general, $\sigma_{\min}(\mOpp{v}{i})$ can depend exponentially on the
+In general, $\sigma_{k}(\mOpp{v}{i})$ can depend exponentially on the
   path length between the hidden variable $h_i$ and $x_v$.

### 2014-05-12 doc/icml2014/directed.tex
read up to undirected model

@@
   \label{thm:sample-complexity}
-  Wlog, let $v=1$ and $i=1$. Let $x_1, x_2, x_3$ be three views for
+  Without loss of generality, let $v=1, i=1$, and let $x_1, x_2, x_3$ be three views for
   $h_1$. With probability at least $1 - \delta$,
@@
     &
+  \hspace{-0.3in}
       O\left(
-      \frac{k \log(1/\delta) {\pi\oft{1}}_{\max}/{\pi\oft{1}}_{\min} } 
-      {\sqrt{n}(\sigma_{k}(\mOpp{1}{1}) \sigma_{k}(\mOpp{1}{2}))^{5/2}} \right) \epsilon.
+      % PL: remove pi_max because that's upper bounded by 1.
+      \frac{k \log(1/\delta) } 
+      {\sqrt{n} \, \pi\oft{1}_{\min} (\sigma_{k}(\mOpp{1}{1}) \sigma_{k}(\mOpp{1}{2}))^{5/2}} \right) \epsilon,
 \end{align*}
+where $\pi\oft{1}_{\min}$ is the smallest element of $\pi\oft{1}$.
+\todo{what's $\epsilon$?}
 
@@
 
-In general, $\sigma_{k}(\mOpp{v}{i})$ can depend exponentially on the
-  path length between the hidden variable $h_i$ and $x_v$.
-This is expected because the influence from $h_i$ to $x_v$
-  gets diluted for every latent variable it must factor through.
+% PL: we don't have a crisp theorem (because there's multiple paths), so leave it vague.
+Note that $\sigma_{k}(\mOpp{v}{i})$ can become extremely
+small if $h_i$ and $x_v$ are connected via many intermediate hidden variables.
+\todo{cite work - Sontag?}
+%In general, $\sigma_{k}(\mOpp{v}{i})$ will depend exponentially on the
+  %path length between the hidden variable $h_i$ and $x_v$.
+%This is expected because the influence from $h_i$ to $x_v$
+  %gets diluted for every latent variable it must factor through.
 

### 2014-05-12 java/src/learning/linalg/TensorGradientDescent.java
Introduced deterministic seeds everywhere

@@ -15,7 +15,7 @@ public class TensorGradientDescent extends DeflatingTensorFactorizationAlgorithm
   @Option(gloss="Number of iterations to run the power method")
   public int iters = 50;
   @Option(gloss="Random number generator for tensor method and random projections")
-  Random rnd = new Random();
+  Random rnd = new Random(23);
   @OptionSet(name="lbfgs") public LBFGSMaximizer.Options lbfgs = new LBFGSMaximizer.Options();
   @OptionSet(name="backtrack") public BacktrackingLineSearch.Options backtrack = new BacktrackingLineSearch.Options();
 

### 2014-05-12 java/src/learning/linalg/TensorPowerMethod.java

@@ -17,7 +17,7 @@ public class TensorPowerMethod extends DeflatingTensorFactorizationAlgorithm {
   @Option(gloss="Number of attempts in order to find the real maxima")
   public int attempts = 100;
   @Option(gloss="Random number generator for tensor method and random projections")
-  Random rnd = new Random();
+  Random rnd = new Random(33);
 
   public TensorPowerMethod() {
   }

### 2014-05-12 java/src/learning/models/loglinear/Models.java

@@ -242,7 +242,7 @@ public class Models {
       }
       return examples;
     }
-      @Override
+//      @Override
       public Counter<Example> getFullDistribution(Params params_) {
           ParamsVec params = (ParamsVec) params_;
           Counter<Example> examples = new Counter<>();

### 2014-05-12 java/src/learning/spectral/TensorMethod.java

@@ -23,7 +23,7 @@ public class TensorMethod {
   @Option(gloss="Number of attempts to find good eigen-vectors")
   public int attempts = 300;
   @Option(gloss="Random number generator for tensor method and random projections")
-  public Random rnd = new Random();
+  public Random rnd = new Random(44);
   @Option(gloss="Oversampling factor for the random projection")
   public double oversamplingFactor = 2.0;
 

### 2014-05-13 doc/icml2014/undirected.tex
gotten up to lemma 2
new asymptotic analysis comparing pi and cl - something is wrong

@@
 to estimate the hidden marginals $Z_\sC = \BP(\vh_\sC)$ and $\mOpp{v}{i} = \BP(x_v \mid h_i)$.
-For directed models, obtaining the local conditional tables from $Z_\sC$ requires
+For directed models, obtaining the local conditional tables from $Z_\sC$ simply requires
 renormalizing $Z_\sC$.
-For undirected log-linear models, the parameters cannot simply be obtained locally;
-we will have to perform global optimization.
-Fortunately, we will show that this optimization problem is convex
+For undirected log-linear models, the canonical parameters cannot be obtained
+locally.
+Fortunately, we can construct a convex optimization problem
 by relating the recovered conditional moments with the expected features of
@@
 \end{align}
-where $A(\theta) = \sum_{\vx, \vh}  \exp( \sum_\sC \theta^\top \phi(\vx,\vh) )$ is the log-partition function.
+where $A(\theta) = \log \sum_{\vx, \vh}  \exp( \theta^\top \phi(\vx,\vh) )$ is the log-partition function.
 
@@
   \lemmaref{bottleneck-views} guarantees exclusive views for any connected
-  sub-graph of $\sG$, not just cliques.
+  subgraph of $\sG$, not just cliques.
 

### 2014-05-13 doc/icml2014/data/plot-asymp.py
Updated asymptotic diagram

@@ -1,0 +1,48 @@
+#!/usr/bin/env python2.7
+"""
+
+"""
+
+import scabby
+import numpy as np
+import matplotlib.pyplot as plt
+
+def do_command(args):
+    plt.rc('font', family = "Serif", size=20)
+    plt.rc('axes', labelsize = 24)
+    plt.rc('xtick', labelsize = 24)
+    plt.rc('ytick', labelsize = 24)
+    plt.rc('text', usetex = True)
+
+    data = scabby.tab_to_table(scabby.read_tab_file(open('asymp-comparison.table')), 'noise', 'dZ_pi', 'std_dZ_pi', 'dZ_cl', 'std_dZ_cl')
+    print data
+    plt.errorbar( data.T[0], data.T[1], yerr=data.T[2], linestyle='-', marker='o', label='Pseudoinverse')
+    plt.errorbar( data.T[0], data.T[3], yerr=data.T[4], linestyle='-', marker='s', label='Composite-Likelihood')
+
+    plt.ylabel(r'$\|\theta - \hat \theta\|_2$')
+    plt.xlabel('$\epsilon$')
+
+    plt.xscale('log')
+    plt.yscale('log')
+    plt.gca().invert_xaxis()
+
+    plt.legend(loc='lower left')
+
+    plt.tight_layout()
+
+    plt.savefig('asymp.pdf')
+    plt.show()
+    
+
+if __name__ == "__main__":
+    import argparse
+    parser = argparse.ArgumentParser( description='' )
+    #parser.add_argument( '', type=str, help="" )
+    parser.set_defaults(func=do_command)
+
+    #subparsers = parser.add_subparsers()
+    #command_parser = subparsers.add_parser('command', help='' )
+    #command_parser.set_defaults(func=do_command)
+
+    ARGS = parser.parse_args()
+    ARGS.func(ARGS)

### 2014-05-13 doc/icml2014/directed.tex
Minor fix to epsilon todo and uniform distribution todo.
Fixed epsilon related bug
Included new assumptions for conditional independence

@@
 of the exclusive views property (\propertyref{exclusive-views}) via \TensorFactorize,
-but we will show that part (i) follows as well.
+but we will show that part (i) follows as well, given basic assumptions
+on the conditional independence structure of the graphical model.
+
+\begin{assumption}[Conditional Independence]
+  \label{asm:ci}  
+  \
+  \begin{enumerate}
+    \item If, conditioned on a variable $h$, two variables $x_1, x_2$
+      are independent, then $x_1$ and $x_2$ are not both dependent on
+      any other variable $h'$: $x_1 \perp x_2 \given h \implies x_1
+      \perp h' \given h \vee x_2 \perp h' \given h$ for every $h' \in
+      \sC$.
+    \item
+      For every clique $\sC \in \sG$, conditioned on a variable $h \in
+      \sC$, all other variables in the clique are dependent:
+      $h_i \not\perp h_j \given h$ for every $h_i, h_j \in \sC$.
+    \item
+      Given an active path in the unconditioned model, then it remains
+        active when conditioned on $h$. 
+      Let $a, b, c$ be three variables in the graph such that
+      $$ a \not\perp b \wedge b \not\perp c \implies a \not\perp c,$$
+      then,
+      $$ a \not\perp b \given h \wedge b \not\perp c \given h \implies a \not\perp c \given h.$$
+%      Suppose $$x \not\perp h_1 \wedge h_1 \not\perp h_2 \implies
+%      x \not\perp h_2,$$ then $$x \not\perp h_1 \given h \wedge h_1 \not\perp h_2
+%      \given h \implies x \not\perp h_2 \given h$$.
+  \end{enumerate}
+\end{assumption}
+
+%TODO: Describe with example the cliques ruled out.
 
@@
   %$h_i \in \sG'$ is a bottleneck, then $\sG'$ has the exclusive views property.
-If a graphical model $\sG$ is uniformly bottlenecked,
-then $\sG$ has the exclusive views property.
+  If a graphical model $\sG$ satisfies conditional independence
+  assumptions (\assumptionref{ci}) and is uniformly bottlenecked, then
+  $\sG$ has the exclusive views property.
 \end{lemma}
@@
 
-In this section, we provide formal conditions under which $\LearnMarginals$
-will produce consistent estimates.
-We let
-$\mOpphat{v}{i}$,
-$\hat Z_\sC$,
-and $\hat M_\sV$,
-denote estimators for
-$\mOpp{v}{i}$,
-$Z_\sC$,
-and $M_\sV$,
-respectively.
+%In this section, we provide formal conditions under which $\LearnMarginals$
+%will produce consistent estimates.
+%We let
+%$\mOpphat{v}{i}$,
+%$\hat Z_\sC$,
+%and $\hat M_\sV$,
+%denote estimators for
+%$\mOpp{v}{i}$,
+%$Z_\sC$,
+%and $M_\sV$,
+%respectively.
 
@@
 %\end{align*}
-
+It is easy to see that $\LearnMarginals$ has a polynomial sample complexity because it composes two parts that individually have polynomial sample complexity.
 From \citet{anandkumar13tensor} given the non-degeneracy assumptions
@@
   that depends polynomially on the $k$-th singular value of
-  $\mOpp{v}{i}$:
-\begin{theorem}[Sample complexity for $\mOpp{v}{i}$]
-  \label{thm:sample-complexity}
-  Without loss of generality, let $v=1, i=1$, and let $x_1, x_2, x_3$ be three views for
-  $h_1$. With probability at least $1 - \delta$,
-\begin{align*}
-  \|\mOpphat{1}{1} - \mOpp{1}{1}\|_F 
-    &\le  \\
-    &
-  \hspace{-0.3in}
-      O\left(
-      % PL: remove pi_max because that's upper bounded by 1.
-      \frac{k \log(1/\delta) } 
-      {\sqrt{n} \, \pi\oft{1}_{\min} (\sigma_{k}(\mOpp{1}{1}) \sigma_{k}(\mOpp{1}{2}))^{5/2}} \right) \epsilon,
-\end{align*}
-where $\pi\oft{1}_{\min}$ is the smallest element of $\pi\oft{1}$.
-\todo{what's $\epsilon$?}
+  $\mOpp{v}{i}$.
+% PL: we don't have a crisp theorem (because there's multiple paths), so leave it vague.
+Recovering the marginals $Z_\sC$ from the conditional moments via $\LearnClique$ is a linear operation and thus also has polynomial sample complexity. 
+Note that $\sigma_{k}(\mOpp{v}{i})$ can become extremely
+small if $h_i$ and $x_v$ are connected via many intermediate hidden variables.
+\todo{cite work - Sontag?}
+%\begin{theorem}[Sample complexity for $\mOpp{v}{i}$]
+%  \label{thm:sample-complexity}
+%  Without loss of generality, let $v=1, i=1$, and let $x_1, x_2, x_3$ be three views for
+%  $h_1$. With probability at least $1 - \delta$,
+%\begin{align*}
+%  \|\mOpphat{1}{1} - \mOpp{1}{1}\|_F 
+%    &\le  \\
+%    &
+%  \hspace{-0.3in}
+%      O\left(
+%      % PL: remove pi_max because that's upper bounded by 1.
+%      \frac{k \log(1/\delta) } 
+%      {\sqrt{n} \, \pi\oft{1}_{\min} (\sigma_{k}(\mOpp{1}{1}) \sigma_{k}(\mOpp{1}{2}))^{5/2}} \right),
+%\end{align*}
+%where $\pi\oft{1}_{\min}$ is the smallest element of $\pi\oft{1}$.
 
@@
 %to the unique parent of $x_1$, $h_t$.
-\end{theorem}
-
-% PL: we don't have a crisp theorem (because there's multiple paths), so leave it vague.
-Note that $\sigma_{k}(\mOpp{v}{i})$ can become extremely
-small if $h_i$ and $x_v$ are connected via many intermediate hidden variables.
-\todo{cite work - Sontag?}
+%\end{theorem}
 %In general, $\sigma_{k}(\mOpp{v}{i})$ will depend exponentially on the

### 2014-05-13 doc/icml2014/exclusive-views.tex
Included new assumptions for conditional independence

@@
 In \sectionref{directed}, we stated that bottlenecks guaranteed
-  exclusive views. We provide the proof below.
+exclusive views provided conditional independence assumptions \assumptionref{ci}. 
+We provide the proof below.
   We generalize the lemma so that it applies
@@
   %\label{lem:bottleneck-views}  
-Let $\sG' = \{ h_{i_1}, \ldots, h_{i_m} \} \subseteq \sG$ be
-  a connected subgraph of hidden variables. If every hidden variable
-  $h_i \in \sG'$ is a bottleneck, then $\sG'$ has the exclusive view property.
+  Let $\sG$ be a graphical model satisfying conditional independence
+  assumptions (\assumptionref{ci}) and be uniformly bottlenecked, then
+  $\sG$ has the exclusive views property.
+%Let $\sG' = \{ h_{i_1}, \ldots, h_{i_m} \} \subseteq \sG$ be
+%  a connected subgraph of hidden variables. If every hidden variable
+%  $h_i \in \sG'$ is a bottleneck, then $\sG'$ has the exclusive view property.
 \end{lemma*}
@@
 \begin{proof}
-  By \lemmaref{reduction}, we can assume that every observed
-  variable is a leaf in the graph, causing any v-structures in the graph
-  to be active. This allows us to reason about independence simply by
-  checking whether or not two nodes are connected or not.
-  Note that this proof applies for undirected graphs as well.
+Let $h_0$ be a hidden variable in the clique $\sC$ and $x_1, x_2$ be any
+  two conditionally independent observed variables. 
+One of $x_1$ or $x_2$ is an exclusive view for $h_0$ iff it is
+  independent of all other hidden variables in the clique, given $h_0$.
+Intuitively, if this weren't the case, then we would be able to find
+  an active path between $x_1$ and $x_2$ that does not go through $h_0$,
+  violating the conditional independence condition.
 
-  Let $h$ be a hidden variable in $\sG'$, let $\sV_{h}$ be the set of
-  views for it and let ${\sG'}^- \eqdef \sG' \setminus \{h\}$.
-  By definition, the views of $h$ are conditionally independent, viz.
-    every connecting path between two views $x_1, x_2 \in \sV_h$ must pass through
-    $h$.
-  A view $x \in \sV_{h}$ is exclusive for $h$ iff $x \not\in \sV_{h'}$ for all $h'
-  \in {\sG'}^-$.
+% TODO: One line here about how assumptions save us.
+% The conditional independence assumptions (\assumptionref{ci}) exclude
+%   edge cases wherein the active trail still preserves 
+% for them to be both depend on any other
+% variable $h'$ in the clique.
 
-  Suppose $h$ had no unique view, then $\sV_{h} \subseteq \Union_{h' \in {\sG'}^-} \sV_{h'}$. 
-  However, this would imply that for every $x_1, x_2 \in \sV_{h}$, there
-  exists some $h_1, h_2 \in {\sG'}^-$ (not necessarily distinct) such that
-    $x_1$ is connected to $h_1$ and $x_2$ is connected to $h_2$.
-  By participation in the subgraph, either $h_1 = h_2$ or $h_1$ is
-    connected to $h_2$, implying that $x_1$ and $x_2$ are connected
-    via a path that does not pass through $h$, contradicting the
-    statement that $x_1, x_2 \in \sV_h$.
+Suppose neither $x_1$ nor $x_2$ were exclusive views for $h_0$, 
+  then there exists some $h_1, h_2 \in \sC$ (not necessarily distinct)  that $x_1, x_2$ are
+  respectively dependent on:
+  $x_1 \not\perp h_1 \given h_0$ and $x_2 \not\perp h_2 \given h_0$. 
+Both $h_1$ and $h_2$ are part of the same clique $\sC$, thus by
+  \assumptionref{ci}, we have that $h_1 \not\perp h_2 \given h_0$.
 
-  Thus, it must be the case that $\sV_{h} \nsubseteq \Union_{h' \neq
-    h \in \sG'} \sV_{h'}$ and hence there exists a unique view for $h$.
+Since $x_1 \not\perp h_1 \given h$, there exists an active trail from
+$x_1 \to h_1$. Furthermore, because $h_1$ and $h_2$ are in the same
+clique, the active trail extends to $h_2$, i.e. $x_1 \not\perp h_2$.
+However, since $x_1 \to h_1 \to h_2$ is an active trail in the unconditioned graph, i.e. $x_1 \not\perp h_1$; by \assumptionref{ci}, it remains active trail when conditioned on $h$.
 
-  Algorithmically, finding the unique view for $h$ can be done by
-  subtracting the views of all other variables in the subgraph from
-  $\sV_h$,i.e. $x_{v_h} \in \sV_h \setminus \Union_{h' \in \sG'} \sV_{h'}$
+Consequently, $x_1 \not\perp h_2 \given h_0$, which
+contradicts the assumption that as $x_1 \perp x_2 \given h_0$,
+$x_1 \not\perp h_i \given h_0$ and $x_2 \not\perp h_i \given h_0$ for all $h_i \in \sC$.
+
 \end{proof}

### 2014-05-13 doc/icml2014/directed.tex
Updated exclusive views property to be true with stronger assumptions

@@
 of the exclusive views property (\propertyref{exclusive-views}) via \TensorFactorize,
-but we will show that part (i) follows as well, given basic assumptions
-on the conditional independence structure of the graphical model.
+but we will show that part (i) follows as well, given that the graphical model faithfully represents conditional independences:
 
-\begin{assumption}[Conditional Independence]
+\begin{assumption}[Faithful Representation]
   \label{asm:ci}  
-  \
-  \begin{enumerate}
-    \item If, conditioned on a variable $h$, two variables $x_1, x_2$
-      are independent, then $x_1$ and $x_2$ are not both dependent on
-      any other variable $h'$: $x_1 \perp x_2 \given h \implies x_1
-      \perp h' \given h \vee x_2 \perp h' \given h$ for every $h' \in
-      \sC$.
-    \item
-      For every clique $\sC \in \sG$, conditioned on a variable $h \in
-      \sC$, all other variables in the clique are dependent:
-      $h_i \not\perp h_j \given h$ for every $h_i, h_j \in \sC$.
-    \item
-      Given an active path in the unconditioned model, then it remains
-        active when conditioned on $h$. 
-      Let $a, b, c$ be three variables in the graph such that
-      $$ a \not\perp b \wedge b \not\perp c \implies a \not\perp c,$$
-      then,
-      $$ a \not\perp b \given h \wedge b \not\perp c \given h \implies a \not\perp c \given h.$$
-%      Suppose $$x \not\perp h_1 \wedge h_1 \not\perp h_2 \implies
-%      x \not\perp h_2,$$ then $$x \not\perp h_1 \given h \wedge h_1 \not\perp h_2
-%      \given h \implies x \not\perp h_2 \given h$$.
-  \end{enumerate}
+  If there is an active trail from any two variables $a, b$ in the graph
+    $\sG$, then they are not independent, i.e $a \not\perp b$. 
+  If there is an active trail between $a$ and $b$ conditioned on any variable,
+    say $c$, then they are also not independent, i.e. $a\not\perp b \given c$.
 \end{assumption}
-
-%TODO: Describe with example the cliques ruled out.
 
@@
   %$h_i \in \sG'$ is a bottleneck, then $\sG'$ has the exclusive views property.
-  If a graphical model $\sG$ satisfies conditional independence
-  assumptions (\assumptionref{ci}) and is uniformly bottlenecked, then
+  If a graphical model $\sG$ satisfies the faithful representation 
+  assumption (\assumptionref{ci}) and is uniformly bottlenecked, then
   $\sG$ has the exclusive views property.

### 2014-05-13 doc/icml2014/exclusive-views.tex
Updated exclusive views property to be true with stronger assumptions

@@
 In \sectionref{directed}, we stated that bottlenecks guaranteed
-exclusive views provided conditional independence assumptions \assumptionref{ci}. 
+exclusive views provided the graph faithfully represents conditional independences (\assumptionref{ci}). 
 We provide the proof below.
@@
   %\label{lem:bottleneck-views}  
-  Let $\sG$ be a graphical model satisfying conditional independence
-  assumptions (\assumptionref{ci}) and be uniformly bottlenecked, then
-  $\sG$ has the exclusive views property.
+  Let $\sG$ be a graphical model satisfying the faithful representation assumption (\assumptionref{ci}).
+  Let $\sC \in \sG$ be a uniformly bottlenecked clique, then
+  $\sC$ has the exclusive views property, i.e.\ every variable in $\sC$
+  has an exclusive view.
 %Let $\sG' = \{ h_{i_1}, \ldots, h_{i_m} \} \subseteq \sG$ be
@@
 \begin{proof}
-Let $h_0$ be a hidden variable in the clique $\sC$ and $x_1, x_2$ be any
+By the canonical reduction (\lemmaref{reduction}), we can assume without
+  loss of generality that all observed variables are leaves and thus
+  all v-structures are active.
+Combined with the faithful representation assumption, this lets us
+  reason about independences solely in terms of graph connectivity.
+
+Let $h_0$ be a hidden variable in the clique $\sC$ and let $x_1, x_2$ be any
   two conditionally independent observed variables. 
-One of $x_1$ or $x_2$ is an exclusive view for $h_0$ iff it is
-  independent of all other hidden variables in the clique, given $h_0$.
-Intuitively, if this weren't the case, then we would be able to find
-  an active path between $x_1$ and $x_2$ that does not go through $h_0$,
-  violating the conditional independence condition.
+Recall that if $x_1$ is an exclusive view of $h_0$ in $\sC$,
+  then, conditioned on $h_0$, $x_1$ is independent of every other variable $h_i
+  \in \sC$.
+% Since $x_1 \perp x_2 \given h_0$, every path from $x_1$ to $x_2$ must go through $h_0$. 
+Suppose neither $x_1$ nor $x_2$ are exclusive views for $h_0$, 
+  then there exists some $h_1, h_2 \in \sC$ that $x_1, x_2$ are respectively dependent on.
+However, $h_1$ and $h_2$ are connected by participation in the clique; consequently there is a path $x_1 \to h_1 \to h_2 \to x_2$. 
+By \assumptionref{ci}, this implies that $x_1$ and $x_2$ are not
+  independent conditioned on $h_0$, which is a contradiction.
 
-% TODO: One line here about how assumptions save us.
-% The conditional independence assumptions (\assumptionref{ci}) exclude
-%   edge cases wherein the active trail still preserves 
-% for them to be both depend on any other
-% variable $h'$ in the clique.
-
-Suppose neither $x_1$ nor $x_2$ were exclusive views for $h_0$, 
-  then there exists some $h_1, h_2 \in \sC$ (not necessarily distinct)  that $x_1, x_2$ are
-  respectively dependent on:
-  $x_1 \not\perp h_1 \given h_0$ and $x_2 \not\perp h_2 \given h_0$. 
-Both $h_1$ and $h_2$ are part of the same clique $\sC$, thus by
-  \assumptionref{ci}, we have that $h_1 \not\perp h_2 \given h_0$.
-
-Since $x_1 \not\perp h_1 \given h$, there exists an active trail from
-$x_1 \to h_1$. Furthermore, because $h_1$ and $h_2$ are in the same
-clique, the active trail extends to $h_2$, i.e. $x_1 \not\perp h_2$.
-However, since $x_1 \to h_1 \to h_2$ is an active trail in the unconditioned graph, i.e. $x_1 \not\perp h_1$; by \assumptionref{ci}, it remains active trail when conditioned on $h$.
-
-Consequently, $x_1 \not\perp h_2 \given h_0$, which
-contradicts the assumption that as $x_1 \perp x_2 \given h_0$,
-$x_1 \not\perp h_i \given h_0$ and $x_2 \not\perp h_i \given h_0$ for all $h_i \in \sC$.
+Thus, at least one of $x_1$ or $x_2$ is isolated from the rest of the
+variables in the clique and forms an exclusive view for $h_0$ in $\sC$.
 

### 2014-05-13 doc/icml2014/undirected.tex
Updated exclusive views property to be true with stronger assumptions

@@
 
-The pseudolikelihood also replaces the log-partition function $A(\theta)$
-with a sum over the variables in the neighbourhood: $A_a(\theta; x_{\sN(a)}) = \sum_{a' \in \{a\} \union \sN(a)} \exp(\theta^\top\phi_{\sN(a)}(a'))$,
-which cannot be computed from just the hidden clique marginals $Z_\sC$,
-since the neighbors $\sN(a)$ in general do not form a clique.
-We adapt our earlier algorithm ($\LearnClique$) to estimate the
-  marginals over all neighborhoods $Z_{\sN(a)}$ by noting that
-  \lemmaref{bottleneck-views} guarantees exclusive views for any connected
-  subgraph of $\sG$, not just cliques.
+The pseudolikelihood also replaces the log-partition function
+  $A(\theta)$ with a sum over the variables in the neighborhood:
+  $A_a(\theta; x_{\sN(a)}) = \sum_{a' \in \{a\} \union \sN(a)}
+  \exp(\theta^\top\phi_{\sN(a)}(a'))$.
+Note that the neighbors $\sN(a)$ in general do not form a clique, and
+  thus the uniformly bottlenecked assumption alone is not sufficient to
+  guarantee exclusive views for each variable in the subgraph.
+However, when we do have exclusive views for each neighborhood, as in
+  \figureref{examples-mrf}, we can adapt our earlier algorithm
+  ($\LearnClique$) to estimate the marginals over all neighborhoods $Z_{\sN(a)}$.
+
+% Arun: WRONG
+% The pseudolikelihood also replaces the log-partition function $A(\theta)$
+% with a sum over the variables in the neighborhood: $A_a(\theta; x_{\sN(a)}) = \sum_{a' \in \{a\} \union \sN(a)} \exp(\theta^\top\phi_{\sN(a)}(a'))$,
+% which cannot be computed from just the hidden clique marginals $Z_\sC$,
+% since the neighbors $\sN(a)$ in general do not form a clique.
+% We adapt our earlier algorithm ($\LearnClique$) to estimate the
+%   marginals over all neighborhoods $Z_{\sN(a)}$ by noting that
+%   \lemmaref{bottleneck-views} guarantees exclusive views for any connected
+%   subgraph of $\sG$, not just cliques.
 

### 2014-05-13 doc/icml2014/exclusive-views.tex
Minor edits to lemma 2 proof

@@
 \begin{proof}
-By the canonical reduction (\lemmaref{reduction}), we can assume without
-  loss of generality that all observed variables are leaves and thus
-  all v-structures are active.
-Combined with the faithful representation assumption, this lets us
+%By the canonical reduction (\lemmaref{reduction}), we can assume without
+%  loss of generality that all observed variables are leaves and thus
+%  all v-structures are active.
+The faithful representation assumption lets us
   reason about independences solely in terms of graph connectivity.
+\todo{v-structures no longer matter because of the assumption.
+Does this need to be explained?}
 
@@
   then there exists some $h_1, h_2 \in \sC$ that $x_1, x_2$ are respectively dependent on.
-However, $h_1$ and $h_2$ are connected by participation in the clique; consequently there is a path $x_1 \to h_1 \to h_2 \to x_2$. 
+However, $h_1$ and $h_2$ are connected by participation in the clique; consequently there is an active trail $x_1 - h_1 - h_2 - x_2$. 
 By \assumptionref{ci}, this implies that $x_1$ and $x_2$ are not

### 2014-05-14 doc/icml2014/directed.tex
fix lemma 2

@@
 of the exclusive views property (\propertyref{exclusive-views}) via \TensorFactorize,
-but we will show that part (i) follows as well, given that the graphical model faithfully represents conditional independences:
+but we will show that part (i) follows as well.
 
-\begin{assumption}[Faithful Representation]
-  \label{asm:ci}  
-  If there is an active trail from any two variables $a, b$ in the graph
-    $\sG$, then they are not independent, i.e $a \not\perp b$. 
-  If there is an active trail between $a$ and $b$ conditioned on any variable,
-    say $c$, then they are also not independent, i.e. $a\not\perp b \given c$.
-\end{assumption}
+To rule out edge cases, we assume the graphical model only has conditional
+independences given by the graph:
+\begin{property}[Faithful representation]
+  \label{prop:ci}  
+  %If there is an active trail from any two variables $a, b$ in the graph
+    %$\sG$, then they are not independent, i.e $a \not\perp b$. 
+  We say that a graphical model $\sG$ is \emph{faithful}
+  if for any variables $a,b,c$
+  such that there is an active trail\footnote{See \todo{cite textbook} for a definition.} between $a$ and $b$ conditioned on $c$,
+  then $a$ and $b$ are dependent given $c$ (i.e., $a\not\perp b \given c$).
+\end{property}
 
@@
 %Before we establish this relationship,
-First, we describe a reduction of any graphical model to
+We now describe a reduction of any graphical model to
   a canonical form to make our arguments simpler.
@@
   %$h_i \in \sG'$ is a bottleneck, then $\sG'$ has the exclusive views property.
-  If a graphical model $\sG$ satisfies the faithful representation 
-  assumption (\assumptionref{ci}) and is uniformly bottlenecked, then
+  If a graphical model $\sG$ is faithful and uniformly bottlenecked, then
   $\sG$ has the exclusive views property.
 \end{lemma}
-% PL: save space, doesn't add much
-%\begin{proof}
-%  The independence conditions required by the bottleneck condition
-%  effectively guarantee that at least one of the three views for
-%  a bottleneck be exclusive. 
-%  \appendixref{exclusive-views} provides a complete proof.
-%\end{proof}
+\begin{proof}
+We prove the contrapositive: suppose that there is some hidden clique $\sC \in \sG$ with $h_0 \in \sC$
+that does not have exclusive views.  This means each observed variable $x_i$ is conditionally dependent on some
+$h_i \in \sC \backslash \{h_0\}$ given $h_0$.
+Consider any such two $x_1, x_2$.
+Conditioned on $h_0$, there exists an active trail between them (specifically, $x_1 - h_1 - h_2 - x_2$).
+Since $\sG$ is faithful, we have $x_1 \not\perp x_2 \mid h_0$.
+As this holds for any $x_1,x_2$, $h_0$ cannot be a bottleneck.
+\end{proof}
 

### 2014-05-14 doc/icml2014/directed.tex
Updated figures and minor tweaks to directed section

@@
   if for any variables $a,b,c$
-  such that there is an active trail\footnote{See \todo{cite textbook} for a definition.} between $a$ and $b$ conditioned on $c$,
+  such that there is an active trail\footnote{See
+  \citet{koller2009probabilistic} for a definition.} between $a$ and $b$
+  conditioned on $c$,
   then $a$ and $b$ are dependent given $c$ (i.e., $a\not\perp b \given c$).
@@
 
-Now we establish the link between bottlenecks and exclusive views
-(see \appendixref{exclusive-views} for the complete proof):
+Now we establish the link between bottlenecks and exclusive views:
 \begin{lemma}[Uniformly bottlenecked implies exclusive views]
@@
 Since $\sG$ is faithful, we have $x_1 \not\perp x_2 \mid h_0$.
-As this holds for any $x_1,x_2$, $h_0$ cannot be a bottleneck.
+As this holds for any $x_1$ and $x_2$, $h_0$ cannot be a bottleneck.
 \end{proof}
@@
 Note that $\sigma_{k}(\mOpp{v}{i})$ can become extremely
-small if $h_i$ and $x_v$ are connected via many intermediate hidden variables.
-\todo{cite work - Sontag?}
+small if $h_i$ and $x_v$ are connected via many intermediate hidden variables
+\footnote{To see this, consider a chain-structured graphical model: $h_1
+\to h_2 \cdots \to h_t \to x_v$. In this example, if
+$\sigma_k(\Pr(h_{i+1} \given h_{i})) \ge \sigma_k$ for each $i = 1,
+\cdots, t-1$, then $\sigma_k(\mOpp{v}{1})$ can be as bad as
+$\sigma_k^t \sigma_k(\mOpp{v}{t})$.}.
 %\begin{theorem}[Sample complexity for $\mOpp{v}{i}$]

### 2014-05-14 doc/icml2014/data/plot-asymp.py

@@ -14,7 +14,7 @@ def do_command(args):
     plt.rc('ytick', labelsize = 24)
     plt.rc('text', usetex = True)
 
-    data = scabby.tab_to_table(scabby.read_tab_file(open('asymp-comparison.table')), 'noise', 'dZ_pi', 'std_dZ_pi', 'dZ_cl', 'std_dZ_cl')
+    data = scabby.tab_to_table(scabby.read_tab_file(args.fname), 'noise', 'dZ_pi', 'std_dZ_pi', 'dZ_cl', 'std_dZ_cl')
     print data
     plt.errorbar( data.T[0], data.T[1], yerr=data.T[2], linestyle='-', marker='o', label='Pseudoinverse')
     plt.errorbar( data.T[0], data.T[3], yerr=data.T[4], linestyle='-', marker='s', label='Composite-Likelihood')
@@ -30,14 +30,17 @@ def do_command(args):
 
     plt.tight_layout()
 
-    plt.savefig('asymp.pdf')
+    if args.output:
+        plt.savefig(args.output + ".pdf")
+        plt.savefig(args.output + ".png")
     plt.show()
     
 
 if __name__ == "__main__":
     import argparse
     parser = argparse.ArgumentParser( description='' )
-    #parser.add_argument( '', type=str, help="" )
+    parser.add_argument( 'fname', type=file, help="table to plot" )
+    parser.add_argument( 'output', type=str, default=None, help="Where to output" )
     parser.set_defaults(func=do_command)
 
     #subparsers = parser.add_subparsers()

### 2014-05-14 python/experiments/variances.py
Added experiment to see variance of cl vs pw

@@ -1,0 +1,153 @@
+#!/usr/bin/env python2.7
+"""
+A simple script to compare composite likelihood and pseudo-inversion based estimators.
+"""
+
+import ipdb
+
+import numpy as np
+from numpy.linalg import inv, svd, pinv, norm
+from numpy.random import rand, randn
+from numpy import zeros, ones, eye
+import itertools as it
+
+from collections import Counter
+
+np.random.seed(12)
+
+def generate_random_instance(k, d):
+    z = normalize(rand(k))
+    O = column_normalize(rand(d, k))
+    mu = O.dot(z)
+    return z, O, mu
+
+def generate_marginals(pi, O, T):
+    Z1 = pi
+    M1 = O.dot(Z1.T) # d x 1
+    Z12 = T.dot(np.diag(pi))
+    M12 = O.dot(Z12).dot(O.T) # d x d
+
+    return Z1, M1, Z12, M12
+
+def recover_pseudoinverse(Ot, Ondk, mut_):
+    Oti = pinv(Ot)
+    zt_ = Oti.dot(mut_ - Ondk)
+
+    return zt_
+
+def recover_composite_likelihood(O, mu, iters = 100, eps = 1e-5):
+    d, k = O.shape
+    # Initialize pi randomly
+    lhood_ = -np.inf
+    zt = normalize(rand(k))[:-1]
+
+    #ipdb.set_trace()
+
+    for i in xrange(iters):
+        zt_ = zeros(k - 1)
+        lhood = 0
+        for x, prob in enumerate(mu):
+            z_marg = np.hstack( (O[x,:-1] * zt, O[x,-1] * (1 - zt).sum()) )
+            lhood += prob * np.log(z_marg.sum())
+            zt_ += prob * (z_marg / z_marg.sum())[:-1]
+
+        assert zt_.sum() < 1.
+
+        assert lhood - lhood_ > -1e-5
+        lhood_ = lhood
+
+        done = np.allclose(zt, zt_, eps)
+        zt = zt_
+        if done:
+            break
+
+    return zt
+
+def normalize(obj):
+    return obj / obj.sum()
+
+def column_normalize(obj):
+    if len(obj.shape) == 2:
+        m, n = obj.shape
+        return obj / np.repeat( np.atleast_2d(obj.sum(0)), m, axis=0 )
+    else:
+        raise AttributeError("Incorrect shape")
+
+def noise_data(data, eps):
+    for key, val in data.items():
[…]

### 2014-05-14 python/experiments/variances.py
Updated script to compute all sorts of sigmas

@@ -6,18 +6,24 @@ A simple script to compare composite likelihood and pseudo-inversion based estim
 import ipdb
 
 import numpy as np
-from numpy.linalg import inv, svd, pinv, norm
+from numpy.linalg import inv, svd, pinv, norm, eigvals
 from numpy.random import rand, randn
-from numpy import zeros, ones, eye
+from numpy import zeros, ones, eye, trace
 import itertools as it
+import sys
 
 from collections import Counter
 
 np.random.seed(12)
 
-def generate_random_instance(k, d):
+def generate_random_instance(k, d, badO = False):
     z = normalize(rand(k))
-    O = column_normalize(rand(d, k))
+    if badO:
+        O = 0.9 * column_normalize(rand(d-1, k))
+        O = np.vstack((O, 0.1 * np.ones(k)))
+    else:
+        O = column_normalize(rand(d, k))
+    assert np.allclose(O.sum(), k, 1e-4)
     mu = O.dot(z)
     return z, O, mu
 
@@ -35,33 +41,36 @@ def recover_pseudoinverse(Ot, Ondk, mut_):
 
     return zt_
 
-def recover_composite_likelihood(O, mu, iters = 100, eps = 1e-5):
+def recover_composite_likelihood(O, mu, iters = 1000, eps = 1e-4, z0 = None):
     d, k = O.shape
     # Initialize pi randomly
     lhood_ = -np.inf
-    zt = normalize(rand(k))[:-1]
+    if z0 is None or (z0 < 0).any() or (z0 > 1).any():
+        print "random init"
+        z = normalize(rand(k))
+    else:
+        z = np.hstack((z0, 1 - z0.sum()))
 
     #ipdb.set_trace()
 
     for i in xrange(iters):
-        zt_ = zeros(k - 1)
+        z_ = zeros(k)
         lhood = 0
         for x, prob in enumerate(mu):
-            z_marg = np.hstack( (O[x,:-1] * zt, O[x,-1] * (1 - zt).sum()) )
+            z_marg = O[x] * z
             lhood += prob * np.log(z_marg.sum())
-            zt_ += prob * (z_marg / z_marg.sum())[:-1]
-
-        assert zt_.sum() < 1.
[…]

### 2014-05-14 python/experiments/variances.py
More really minor changers to pw

@@ -19,8 +19,8 @@ np.random.seed(12)
 def generate_random_instance(k, d, badO = False):
     z = normalize(rand(k))
     if badO:
-        O = 0.9 * column_normalize(rand(d-1, k))
-        O = np.vstack((O, 0.1 * np.ones(k)))
+        O = 0.5 * column_normalize(rand(d-1, k))
+        O = np.vstack((O, 0.5 * np.ones(k)))
     else:
         O = column_normalize(rand(d, k))
     assert np.allclose(O.sum(), k, 1e-4)
@@ -41,12 +41,12 @@ def recover_pseudoinverse(Ot, Ondk, mut_):
 
     return zt_
 
-def recover_composite_likelihood(O, mu, iters = 1000, eps = 1e-4, z0 = None):
+def recover_composite_likelihood(O, mu, iters = 100000, eps = 1e-6, z0 = None):
     d, k = O.shape
     # Initialize pi randomly
     lhood_ = -np.inf
-    if z0 is None or (z0 < 0).any() or (z0 > 1).any():
-        print "random init"
+    if z0 is None or z0.sum() > 1 or (z0 < 0).any() or (z0 > 1).any():
+        #print "random init"
         z = normalize(rand(k))
     else:
         z = np.hstack((z0, 1 - z0.sum()))
@@ -125,13 +125,19 @@ def do_command(args):
     zt = z[:-1]
     Ondk = O[:-1, -1]
     Ot = O[:-1, :-1] - Ondk.reshape(D-1, 1).dot( ones((1,K-1)) )
-    print "ones residual", norm( ones((D-1,1)).T.dot(Ot) )
+    Oti = pinv(Ot)
+    o = ones((D-1,1))
+    v = (Ot.dot(Oti)).T.dot(o)
+
+    print "O cond", svd(O)[1]
+    print "ones residual", norm( Ot.T.dot(o) ), norm( v )
+    print "mu residual", (Oti.dot(mut)), norm(Oti.dot(mut))
 
     dZ_pi, dZ_cl = [], []
 
     for attempt in xrange(attempts):
         mut_, zt_pi, zt_cl = do_experiment(O, Ot, Ondk, mu, args)
-        if attempt % attempts/20 == 0:
[…]

### 2014-05-16 doc/icml2014/background.tex
major structural changes up to bidependent set

@@
 
-Let $\sG$ be a discrete graphical model with observed variables $X$ and
-  hidden variables $H$.
-For a directed model,
-  let $\Pa(a) \subset X \union H$ be the parents of a variable $a$;
-  for an undirected model,
-  let $\sN(a) \subset X \union H$ be the neighbors of $a$.
-  Abusing notation slightly, we use $\sG$ to denote the set of cliques.
-  %(for directed models, these are $\sG = \{ \Pa(a) \union \{a\} : a \in X \union H \}$).
-We use $x_1, x_2, \dots, x_L \in X$ to denote observed variables and
-  $h_1, h_2, \dots, h_M \in H$ to denote hidden variables.
-For an integer $n$, let $[n] = \{ 1, \dots, n \}$.
-For ease of exposition, we assume that $x_v \in [d]$ for
-all $v \in [L]$, and that $h_i \in [k]$ for all $i \in [M]$.
-Let $\sX$ be the joint domain of the observed variables, i.e. $\sX \eqdef [d]^L$, and
-let $\sH$ be the joint domain of the hidden variables, i.e. $\sH \eqdef [k]^M$.
+Let $\sG$ be a discrete graphical model with
+observed variables $\bx = (x_1, \dots, x_L)$ and
+hidden variables $\bh = (h_{1}, \dots, h_M)$.
+We assume that the domains of the variables
+are $x_v \in [d]$ for all $v \in [L]$
+and $h_i \in [k]$ for all $i \in [M]$,
+where $[n] = \{ 1, \dots, n \}$.
+Let $\sX \eqdef [d]^L$ and $\sH \eqdef [k]^{M}$ be the joint domains of $\bx$ and $\bh$.
 
-% PL: don't know what this means
-%We assume that the model $\sG$ is parametrized at every clique $\sC$.
-For directed models,
-$p_\theta(\bx,\bh) = \prod_{a \in X \cup H} p_\theta(a \mid \Pa(a))$,
-where the parameters $\theta$ are the conditional probability tables.
-For undirected models,
+For undirected models $\sG$,
+let $\sG$ denote a set of cliques, where each clique $\sC \subseteq \bx \cup \bh$ is a subset of nodes.
+The joint distribution is given by an exponential family:
   $p_\theta(\bx,\bh) \propto \prod_{\sC \in \sG} \exp(\theta^\top\phi_\sC(\bx_\sC, \bh_\sC))$,
-where $\theta$ are the canonical parameters of an exponential family.
+  where $\theta$ is the parameter vector,
+  and $\phi_\sC(\bx_\sC, \bh_\sC)$ is the local feature vector
+  which only depends on the observed ($\bx_\sC$) and hidden ($\bh_\sC$) variables in that clique $\sC$.
+  Also define $\sN(a) = \{ b \neq a : \exists \sC \supseteq \{ a, b \}$ be the neighbors of variable $a$.
 
-% PL: log-linear parametrization - can't say this for each clique: this is a global notion.
-%If $\sG$ is undirected, we assume a log-linear parametrization of each clique $\sC \in \sG$.
+For directed models $\sG$,
+  $p_\theta(\bx,\bh) = \prod_{a \in \bx \cup \bh} p_\theta(a \mid \Pa(a))$,
+  where $\Pa(a) \subseteq \bx \union \bh$ are the parents of a variable $a$.
+  The parameters $\theta$ are the conditional probability tables of all variables,
+  and the cliques are $\sG = \{ \{ a \} \union \Pa(a) : a \in \bx \union \bh \}$.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \paragraph{Problem statement}
@@
 In practice, one uses local optimization procedures (e.g., EM
-  or L-BFGS) on the marginal likelihood, which can get stuck in local
+  or L-BFGS) on the marginal likelihood, but these can get stuck in local
   optima.
@@
 First, a bit of notation is necessary.
-%The starting point is moments of the marginal distribution $\BP(\vx)$,
-%and a bit of notation is necessary.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \paragraph{Notation}
@@
   T(v_1, \cdots, v_\ell) 
-      %&= \sum_{i_1 = 1}^{d} \cdots \sum_{i_\ell = 1}^{d} 
-% PL: this isn't great notation; reader can figure it out
-%&= \sum_{\bi = \ones}^{\vec d}
   &= \sum_{\bi}
@@
   T(M_1, \cdots, M_\ell)[\vec j]
-      %&= \sum_{i_1 = 1}^{d} \cdots \sum_{i_\ell = 1}^{d} 
-  &= \sum_{\bi}
-            T[\vec i] {M_1}[i_1,j_1] \cdots M_\ell[i_\ell, j_\ell].
+  &= \sum_{\bi} T[\vec i] {M_1}[i_1,j_1] \cdots M_\ell[i_\ell, j_\ell].
 \end{align*}
@@
 
-\paragraph{Estimating three-view mixture models}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Assumptions}
+
+In this section, we state technical assumptions that hold for the rest of the
+paper, but that we feel are not central to our main ideas.
+The first one ensures that we will encounter all realizations of a hidden variable are possible:
+
+\begin{assumption}[Non-degeneracy]
+  \label{asm:non-degeneracy}
+  The marginal distribution of each hidden variable $h_i$ has full support: $\Pr(h_i) \succ 0$.
+\end{assumption}
+
+Next, we assume the graphical model only has conditional independences given by
+the graph:
+\begin{assumption}[Faithful]
+  \label{prop:ci}  
+  For any hidden variables $a,b,c \in \bh$
+  such that there is an active trail between $a$ and $b$ conditioned on $c$,
+  we have that $a$ and $b$ are dependent given $c$ (i.e., $a\not\perp b \given c$).
+\end{assumption}
+
+Finally, we describe a reduction of any graphical model to
+  a canonical form to make our arguments simpler.
+\begin{lemma}[Canonical form]
+  \label{lem:reduction}
+Every directed (undirected) graphical model can be transformed into one in which
+  the observed variables are leaves with exactly one parent (neighbor). 
+There is a one-to-one correspondence between the parameters of the
+  transformed and original models.
+\end{lemma}
+\begin{proof}
+  \begin{figure}
+    \centering
+    \subimport{figures/}{reduction.tikz}
+    \caption{Reduction to canonical form.}
+    \label{fig:reduction}
+  \end{figure}
+
+  \providecommand{\hp}{\ensuremath{h_\text{\rm new}}}
+  Let $x_v$ be an observed variable with parents $\Pa(x_v)$ and children $\text{Ch}(x_v)$.
+  Consider the following transformation.
+  Replace $x_v$ with a new hidden variable \hp\ with the same
+  parents $\Pa(x_v)$ and children $\text{Ch}(x_v) \union \{x_{v_1}, x_{v_2}, x_{v_3}\}$,
+  where $x_{v_1},x_{v_2},x_{v_3}$ are three copies of $x_v$
+  (\figureref{reduction}).
+  Define $\Pr(\hp \mid \Pa(x_v)) = \Pr(x_v \mid \Pa(x_v))$ and
+  $\Pr(x_{v_j} \mid \hp) = I$.
+  Then, there is a one-to-one correspondence between every value of
+  $\hp$ and $x_v$. Consequently, for any clique $\sC \contains \hp$, the
+  parameters in the original graphical model can be obtained by
+  substituting $\hp$ with $x_v$.
+
+  We apply this procedure for all non-leaf observed variables.
+  This procedure also applies straightforwardly for undirected graphical
+  models, considering the neighbors $\sN(x_v)$ instead of its parents
+  and children.
+\end{proof}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Bottlenecks}
+
+In this section, we describe how tensor factorization
+can be used to reveal information about the hidden
+variables \citep{anandkumar13tensor}.  The key notion is that of a bottleneck:
+
+\begin{definition}[Bottleneck]
+  A hidden variable $h_i$ is said to be a \emph{bottleneck} if there
+  exists three observed variables (views), $x_{v_1}, x_{v_2}, x_{v_3}$
+  that are conditionally independent given $h_i$,
+  and each $\mOpp{v}{i} \eqdef \Pr(x_v \mid h_i) \in \Re^{d \times k}$ has full column rank $k$
+  for each $v \in \{v_1, v_2, v_3\}$.
+  %Let $\sV_{h_i}$ denote any set of views for $h_i$.
+  We say that a subset of hidden variables $S \subseteq \bh$ is a bottleneck
+  if every $h \in S$ is a bottleneck.
+  We say that a graphical model $\sG$ is bottlenecked if all hidden variables
+  are bottlenecks.
+\end{definition}
+For example, in \figureref{approach}, $x_1^a,x_1^b,x_2^a$ are views of the bottleneck $h_1$;
+$x_2^a,x_2^b,x_1^b$ are view of the bottleneck $h_2$.  Therefore, the clique
+$\{h_1,h_2\}$ is a bottleneck.  Note that views are allowed to overlap.
+
+%Now define the following property:
+%\begin{property}[Uniformly bottlenecked]
+%  \label{prop:bottleneck}
+%  every hidden variable in $S$ is a bottleneck. 
+%  If every hidden clique $\sC \in \sG$ is uniformly bottlenecked, then we say
+%  that the graph $\sG$ is uniformly bottlenecked as well.
+%\end{property}
+%If every hidden variable is a bottleneck, then we can estimate $\mOpp{v}{i}$
+
+%Intuitively, we need to ensure that every marginal distribution
+  %$\Pr(x_v)$ can be separated into $k$ different conditional distributions
+  %$\Pr(x_v | h_i)$. 
+The full rank assumption on the conditional moments $\mOpp{v}{i} = \Pr(x_v \mid h_i)$
+ensures that all states of $h_i$ ``behave differently''.
+In particular, the conditional distribution of one state cannot be
+a mixture of that of other states.
+%PL: leave vague
+%asks that the conditional distributions are
+%  linearly independent; if this were not the case, the conditional
+%  distribution for a particular latent state could be imitated by the
+%  combination of several other latent states.
+
+%\begin{assumption}
+%\label{asm:full-rank}
+%The conditional moments $\mOpp{1}{1}$, $\mOpp{2}{1}$, $\mOpp{3}{1}$ have full column rank
+%  $k$, and $\pi\oft 1 \succ 0$.
+%\end{assumption}
 
@@
   \input{figures/three-view.tikz}
-  \caption{The canonical three-view mixture model can be estimated consistently using tensor factorization \citep{anandkumar13tensor}.}
+  \caption{The canonical three-view mixture model can be estimated consistently
+  using tensor factorization \citep{anandkumar13tensor}.}
 \end{figure}
 
-\citet{anandkumar12moments} presented a consistent estimator
-  using the method of moments to learn the parameters of a three-view
-  mixture model (\figureref{three-view}), an important case of directed
-  graphical models. %, that sidesteps the local optima problem altogether. 
-In that model, we have three observed variables $x_1, x_2, x_3$
-  that are independent conditioned on a hidden variable $h_1$.  
-The parameters of this model are the prior probabilities $\pi\oft 1 \eqdef
-  \Pr(h_1) \in \Re^k$ and the conditional moments $\mOpp{i}{1} \eqdef
-  \Pr(x_i \given h_1) \in \Re^{d \times k}$ for $i \in \{1,2,3\}$.
+\citet{anandkumar12moments} provide an efficient tensor factorization algorithm
+for estimating $\Pr(x_v \mid h_i)$:
+  %using the method of moments to learn the parameters of a three-view
+\begin{theorem}[Tensor factorization]
+Let $h_i \in \bh$ be a hidden variable
+with views $x_{v_1},x_{v_2},x_{v_3}$.
+Then there exists an algorithm $\TensorFactorize$ that
+returns consistent estimates $\mOpphat{v}{i}$ for each $v \in \{v_1,v_2,v_3\}$.
+\end{theorem}
 
+%\citet{anandkumar12moments} presented a consistent estimator
+  %using the method of moments to learn the parameters of a three-view
+  %mixture model (\figureref{three-view}), an important case of directed
+  %graphical models. %, that sidesteps the local optima problem altogether. 
+%In that model, we have three observed variables $x_1, x_2, x_3$
+  %that are independent conditioned on a hidden variable $h_1$.  
+%The parameters of this model are the prior probabilities $\pi\oft 1 \eqdef
+  %\Pr(h_1) \in \Re^k$ and the conditional moments $\mOpp{i}{1} \eqdef
+  %\Pr(x_i \given h_1) \in \Re^{d \times k}$ for $i \in \{1,2,3\}$.
+
+To simplify notation, consider the example in \figureref{three-view} where $h_1=1,v_1=1,v_2=2,v_3=3$.
 The key idea is that the observed moments $\mO_{12}, \mO_{23}, \mO_{13}$ and
@@
 \begin{align*}
-  \mO_{ij} &= \sum_h {\pi \oft 1}[h] {\mOppt{i}{1}}[h] \otimes {\mOppt{j}{1}}[h] \quad \forall i,j \in \{1,2,3\} \\
+  \mO_{ij} &= \sum_h {\pi \oft 1}[h] {\mOppt{i}{1}}[h] \otimes {\mOppt{j}{1}}[h] \\ %\quad \forall i,j \in \{1,2,3\} \\
   \mO_{123} &= \sum_h {\pi \oft 1}[h] {\mOppt{1}{1}}[h] \otimes {\mOppt{2}{1}}[h] \otimes  {\mOppt{3}{1}}[h].
 \end{align*}
-
-\citet{anandkumar13tensor} presented a robust tensor power method, which
-  we call $\TensorFactorize$, which transforms estimates of the observed moments 
-  into estimates of the parameters, $\pi\oft 1, \mOpp{1}{1},
-  \mOpp{2}{1},\mOpp{3}{1}$ (up to permutation of the columns), provided
-  the following assumption holds:
-
-\begin{assumption}
-\label{asm:full-rank}
-The conditional moments $\mOpp{1}{1}$, $\mOpp{2}{1}$, $\mOpp{3}{1}$ have full column rank
-  $k$, and $\pi\oft 1 \succ 0$.
-\end{assumption}
-
-In brief, the algorithm first computes a whitening matrix $W \in
+The algorithm first computes a whitening matrix $W \in
   \Re^{d \times k}$ such that $W^\top \mO_{12} W = I_{k \times k}$,
@@
 of the whitened $\mO_{123}$; unwhitening yields the columns of $\mOpp{3}{1}$.
-The other parameters can be recovered with a few additional matrix operations.
-%$\mOpp{1}{1}$ and $\mOpp{2}{1}$ can then be recovered with a few matrix operations.
+The other model parameters can be recovered with a few additional matrix operations.
+%  we call $\TensorFactorize$, which transforms estimates of the observed moments 
+%  into estimates of the parameters, $\pi\oft 1, \mOpp{1}{1},
+%  \mOpp{2}{1},\mOpp{3}{1}$ (up to permutation of the columns), provided
+%  the following assumption holds:
 
+%It is easy to see that $\LearnMarginals$ has a polynomial sample complexity
+%because it composes two parts that individually have polynomial sample
+%complexity.
+%From \citet{anandkumar13tensor} given the non-degeneracy assumptions
+  %\assumptionref{non-degeneracy},
+  We have that $\mOpphat{v}{i}$
+  converges to $\mOpp{v}{i}$ at a rate of $n^{-\frac12}$ with a constant
+  that depends polynomially on the $k$-th singular value of
+  $\mOpp{v}{i}$.
+% PL: we don't have a crisp theorem (because there's multiple paths), so leave it vague.
+%Recovering the marginals $Z_\sC$ from the conditional moments via $\LearnClique$ is a linear operation and thus also has polynomial sample complexity. 
+However, $\sigma_{k}(\mOpp{v}{i})$ can become extremely
+small if $h_i$ and $x_v$ are connected via many intermediate hidden variables.
+
+% No parameters yet
 The tensor factorization method attacks the heart of the non-convexity
-  in latent-variable models.  In the next two sections,
-  we will see how to use this result to estimate directed (Section~\ref{sec:directed})
-  and undirected (Section~\ref{sec:undirected}) models.
+  in latent-variable models, providing some information about the hidden variables
+  in the form of the conditional moments $\mOpp{v}{i} = \Pr(x_v \mid h_i)$.
+  Note that $\TensorFactorize$ only looks at the conditional independence structure
+  of the graphical model, not its parametrization.
+  If $v$ is the single parent of $i$ (e.g., $\Pr(x_1^a \mid h_1)$ in \figureref{approach}),
+  then this conditional moment is a parameter of the model.
+  If not (e.g., $\Pr(x_2^a \mid h_1)$), then this is not a parameter.
+  Furthermore, there are some parameters (e.g., $\Pr(h_4 \mid h_2, h_3)$) which are
+  are still not recovered.
+  In general, there is a gap between the conditional moments
+  and the model parameters,
+  something we will address in the next section.
+
+  %In the next two sections,
+  %we will see how to use this result to estimate directed (Section~\ref{sec:directed})
+  %and undirected (Section~\ref{sec:undirected}) models.

### 2014-05-16 doc/icml2014/directed.tex
major structural changes up to bidependent set

--- 2014-05-14 doc/icml2014/directed.tex
@@
-% DONE: too wordy
-%\section{Consistent Parameter Estimation in Directed Models}
-\section{Directed models and exclusive views}
+\section{Exclusive views}
 \label{sec:directed}
 
-%\todo{we are using clique to refer to only hidden cliques; be explicit about using "hidden"}
-
-% DONE: need more setup of what the goal is
-%The tensor factorization method attacks the heart of the non-convexity
-  %in latent variable models.
-  In this section, we will develop a consistent parameter estimate for a class of directed graphical models.
-  In general directed models, the conditional moments $\mOpp{v}{i} \eqdef \Pr(x_v \mid h_i)$
-which we were able to recover from {\TensorFactorize}, are not the underlying parameters.
+%In this section, we will develop a consistent parameter estimate for a class of directed graphical models.
+%The conditional moments $\mOpp{v}{i} \eqdef \Pr(x_v \mid h_i)$
+%recovered from {\TensorFactorize}, are not the underlying parameters.
   %Once we recover the conditional moments $\mOpp{v}{j} \eqdef \Pr(x_v
@@
   %graphical models.
-  Let us first consider an example (Section~\ref{sec:directedExample});
-  Section~\ref{sec:directedGeneral} will describe the algorithm in full generality.
-For ease of exposition we make the following simplifications to our presentation:
-(i) we describe our algorithm entirely in the context of directed
-  graphical models, though it generalizes to undirected models;
-(ii) we derive our algorithm assuming infinite data; in \sectionref{sampleComplexity} we 
-  describe error bounds for the finite sample regime; and
-(iii) we present the algorithm solely in terms of linear algebraic operations;
-  \sectionref{piecewise} provides a statistically
-  more efficient estimator using composite likelihood.
+In this section, we will use the conditional moments $\mOpp{v}{i} \eqdef
+\Pr(x_v \mid h_i)$ recovered from the previous section
+to compute the marginal distribution of sets of hidden variables $Z_S \eqdef \Pr(\bh_S)$,
+taking us one step closer to the actual model parameters.
+For clarity, we derive our algorithm in the infinite data limit,
+although the actual algorithm works with estimates.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Example: directed grid model}
@@
 
-%\begin{figure}
-%  \centering
-%  \input{figures/grid-outline.tikz}
-%  \caption{A directed grid model.}
-%  \label{fig:grid}
-%\end{figure}
-
-%Consider the directed grid model shown in \figureref{grid} which
-Consider the directed grid model shown in \figureref{approach}.
-  % DONE: all the different dependency => too strong
-  %captures the local dependency structures possible in
-  %a Bayesian network.
-The model has eight observed variables $x^a_1, x^b_1 \cdots, x^a_4, x^b_4$ and four
-  hidden variables $h_1, \ldots, h_4$.
+To gain some intuition, consider the directed grid model from \figureref{approach}.
+%The model has eight observed variables $x^a_1, x^b_1 \cdots, x^a_4, x^b_4$ and four
+  %hidden variables $h_1, \ldots, h_4$.
 The parameters of this model are the conditional probability tables
@@
 =  \Pr(x^b_i | h_i) \in \Re^{d \times k}$.
-%Assume $O$ and $T$ have full column rank. % PL: if we put this in, we have to say something about V, which is too complex.
 
 \paragraph{Estimating $O$}
-Note that the observed variables $x^a_1, x^b_1, x^a_2$ are
-  conditionally independent given $h_1$; we can therefore use
-  $\TensorFactorize$ from \sectionref{setup} to recover $O$.
+Since the observed variables $x^a_1, x^b_1, x^a_2$ are
+  conditionally independent given $h_1$, we can $\TensorFactorize$ from
+  \sectionref{setup} to recover $O$.
 
@@
   $\pi$ by a linear system: $\mO_1 = O \pi$. 
-% PL: weird to refer to refer to this assumption from previous section
-% where it's talking about another model.
-%By \assumptionref{full-rank}, $O$ has full column rank and thus can be
-  %inverted to recover $\pi$: $\pi = O\pinv  \mO_1$.
 If $O$ has full column rank, we can recover $\pi$ by taking the pseudoinverse: $\pi = O\pinv  \mO_1$.
@@
 
-%Let us now generalize the intuitions of the directed grid to general graphs.
+  %Section~\ref{sec:directedGeneral} will describe the algorithm in full generality.
+%For ease of exposition we make the following simplifications to our presentation:
+%(i) we describe our algorithm entirely in the context of directed
+  %graphical models, though it generalizes to undirected models;
+%(iii) we present the algorithm solely in terms of linear algebraic operations;
+  %\sectionref{piecewise} provides a statistically
+  %more efficient estimator using composite likelihood.
+
 The intuitions of the directed grid generalize readily to general graphs.
-% PL: this prose is redundant with property
-%Note that in each of the above cases,
-  %for any clique $\sC$, each hidden variable $h_i \in \sC$ has its own
-  %view which is conditionally independent given the other hidden variables in
-  %the clique.
-%Fundamentally, algorithm takes advantage of having a view for every
-  %hidden variable in a clique to help identify its hidden state.
 The key property required for our algorithm to work is as follows:
-%We capture this general property as follows:
 \begin{property}[Exclusive views]
   \label{prop:exclusive-views}
-A hidden clique $\sC \in \sG$ is said to possess the \emph{exclusive views property} if for
-  every hidden variable $h_i \in \sC$, (i)
-  there exists some observed variable,
-  $x_{v}$ which is conditionally independent of the rest of the clique
-  given $h_i$, and (ii) the conditional moments $\mOpp{v}{i} \eqdef
-  \Pr(x_{v} \mid h_i)$ can be estimated consistently.
-We call $x_v$ an exclusive view of $h_i$ in $\sC$.
-If every clique has the exclusive views property, then we say that $\sG$ does as well.
+Given a graphical model $\sG$,
+  a subset of hidden variables $S \subset \bh$ is said to satisfy the \emph{exclusive views property} if for
+  each $h_i \in S$, the two conditions hold:
+  (i) there exists some observed variable
+  $x_{v}$ which is conditionally independent of the others $S \backslash \{ h_i \}$ given $h_i$,
+  and (ii) the conditional moment matrix $\mOpp{v}{i} \eqdef
+  \Pr(x_{v} \mid h_i)$ have full column rank $k$ and can be recovered.
+We call $x_v$ an exclusive view of $h_i$ in $S$.
+%If every hidden clique has the exclusive views property, then we say that $\sG$ does as well.
 \end{property}
-
-Next, we state the non-degeneracy assumptions under which our general
-  algorithm is guaranteed to obtain consistent estimates.
-\begin{assumption}[Non-degeneracy]
-  \label{asm:non-degeneracy}
-  (i) The marginal distribution of each hidden variable $h_i$ have
-  full support: $\mPi{i} \succ 0$. 
-  (ii) The conditional moments $\mOpp{v}{i}$ have full column rank $k$.
-\end{assumption}
-
-Intuitively, we need to ensure that every marginal distribution
-  $\Pr(x_v)$ can be separated into $k$ different conditional distributions
-  $\Pr(x_v | h_i)$. 
-The full rank assumption asks that the conditional distributions are
-  linearly independent; if this were not the case, the conditional
-  distribution for a particular latent state could be imitated by the
-  combination of several other latent states.
-
-% PL: to save space
-%This property also extends to any connected subgraph $\sG' \in \sG$.
-%For the rest of this section, we will only consider subgraphs which are
-  %cliques. 
-%In \sectionref{undirected}, we will show how the generalization
-  %to subgraphs allows us to efficiently estimate the pseudo-likelihood.
-
-% \begin{property}(Exclusive views)
-%   \label{prop:exclusive-views}
-%   A hidden clique $\sC$ of $\sG$ is said to possess the \textbf{exclusive views
-%   property} if for every hidden variable $h_i \in \sC$,
-%   there is some observed variable, $x_{v}$, which is conditionally
-%   independent of the rest of the clique given $h_i$,
-%   and whose conditional
-%   moment $\mOpp{v}{i} \eqdef \Pr(x_{v} \mid h_i)$ can be estimated consistently.
-% We call $x_v$ the exclusive view of $h_i$ in $\sC$.
-% \end{property}
 
@@
 
-We now show this condition is sufficient to learn the marginal
-  distribution of the clique.
-Consider any hidden clique $\sC = \{i_1, \ldots, i_m\}$ with the exclusive views property. Let
-  $x_{v_j}$ be any exclusive view for $h_{i_j}$ in $\sC$ and define $\sV
-  = \{v_1, \ldots, v_m\}$. % be a set of exclusive views for the clique $\sC$.
-We can write down the moments $\mO_\sV$ as follows:
+We now show that if a subset of hidden variables $S$ has the exclusive views property,
+then we can recover the marginal distribution $\Pr(\bh_S)$.
+Consider any $S = \{i_1, \ldots, i_m\}$ with the exclusive views property. Let
+  $x_{v_j}$ be an exclusive view for $h_{i_j}$ in $S$ and define $\sV
+  = \{v_1, \ldots, v_m\}$. % be a set of exclusive views for the clique $S$.
+By the exclusive views property,
+the marginal over the observed variables $\Pr(\bx_\sV)$
+factorizes according to the marginal over the hidden variables $\Pr(\bh_S)$
+times the conditional moments:
 \begin{align*}
@@
   &\eqdef \Pr(\Sx{\sV}) \\
-  %&= \sum_{\vh \in \sH}
-      %\Pr(\Sh{\sC}) \Pr(\Sx{\sV} \given \Sh{\sC}) \\
-      &= \sum_{\vh_\sC} \Pr(\Sh{\sC}) 
+  &= \sum_{\vh_S} \Pr(\Sh{S}) 
                     \Pr(x_{v_1} | h_{i_1}) \cdots \Pr(x_{v_m} | h_{i_m}) \\
-    &= Z_{\sC}(\mOpp{v_1}{i_1},\cdots,\mOpp{v_m}{i_m}).
+  &= Z_{S}(\mOpp{v_1}{i_1},\dots,\mOpp{v_m}{i_m}) \\
+  &= Z_{S}(\mOppAll),
+\end{align*}
+where $\mOppAll = \mOpp{v_1}{i_1} \otimes \cdots \otimes \mOpp{v_m}{i_m}$ is the tensor product of
+all the conditional moments.
+Vectorizing, we have that
+$Z_S \in \Re^{k^m}$,
+$M_\sV \in \Re^{d^m}$,
+and $\mOppAll \in \Re^{d^m \times k^m}$.
+Since each $\mOpp{v}{i}$ has full column rank $k$,
+$\mOppAll$ has full column rank $k^m$.
+Succinctly, $\mO_\sV$ (which can be estimated directly from data)
+is a linear function of $Z_S$ (what we seek to recover).
+We can solve for the hidden marginals $Z_S$ simply by projecting $\mO_\sV$ onto the column
+space of $\mOppAll$:
+\begin{align*}
+  Z_{S} &= \mO_\sV(\mOpp{v_1}{i_1}\pinv,\cdots,\mOpp{v_m}{i_m}\pinv).
 \end{align*}
 
-If each $\mOpp{v_j}{i_j}$ has full column rank, then we can recover the
-hidden marginals $Z_\sC$ by inverting:
-\begin{align*}
-  Z_{\sC} &= \mO_\sV(\mOpp{v_1}{i_1}\pinv,\cdots,\mOpp{v_m}{i_m}\pinv).
-\end{align*}
 \algorithmref{learnclique} summarizes the procedure, \LearnClique.
-Given $Z_\sC$, the conditional probability tables for $\sC$ can easily be
-obtained via renormalization (see \sectionref{sampleComplexity}).
+Given $Z_S$, the conditional probability tables for $S$ can easily be
+obtained via renormalization.
+\begin{theorem}[Hidden marginals from exclusive views]
+If $S \subseteq \bx$ is a subset of hidden variables with the exclusive views property,
+then \algorithmref{learnclique} recovers the marginals $Z_S = \Pr(\bh_S)$.
+\end{theorem}
+
+%If each $\mOpp{v_j}{i_j}$ has full column rank, then
 
 \begin{algorithm}
-  \caption{\LearnClique~(pseudoinverse)}
+  \caption{\LearnClique~(pseudoinverse version)}
   \label{algo:learnclique}
   \begin{algorithmic}
-    \REQUIRE Hidden clique $\sC$ with exclusive views (\propertyref{exclusive-views}).
-    \ENSURE Marginal distribution of the clique $Z_\sC$.
+    \REQUIRE Hidden subset $S$ with exclusive views (\propertyref{exclusive-views}).
+    \ENSURE Marginal distribution $Z_S$.
       \STATE Identify exclusive views $\Sx{\sV} = \{x_{v_1}, \dots, x_{v_m}\}$.
-      \STATE Return $\mH_\sC \gets \mO_{\sV}( \mOpp{v_1}{i_i}\pinv, \dots, {\mOpp{v_m}{i_m}}\pinv )$.
+      \STATE Return $\mH_S \gets \mO_{\sV}( \mOpp{v_1}{i_i}\pinv, \dots, {\mOpp{v_m}{i_m}}\pinv )$.
   \end{algorithmic}
 \end{algorithm}
-% PL: sample complexity, not rank
-%Of course, for $Z_{\sC}$ to be identifiable, we need various rank conditions,
-%which will be established in \sectionref{sampleComplexity}.
 
@@
 
-% DONE: transition the content more gradually (checkpoint: what have we done so far, what next)
-% DONE: refer to properties by name in addition to by number
 We have shown how to estimate the parameters of any clique that possesses the exclusive
@@
 
-%In the example above, we were able to guarantee \propertyref{exclusive-views}
-  %by identifying a hidden variable, $h_1$ along with three conditionally
-  %independent observed variables, $x^a_1, x^b_1, x^a_2$ using the tensor
-  %factorization method.
-  %followed by using \TensorFactorize to learn the conditional moments.
-First, let us revisit the definition of a bottleneck:
-\begin{definition}[Bottleneck]
-  A hidden variable $h_i$ is said to be a \textbf{bottleneck} if there
-  exists at least three observed variables (views), $x_{v_1}, x_{v_2}, x_{v_3}$
-  that are conditionally independent given $h_i$,
-  and $\mOpp{v_1}{i},\mOpp{v_2}{i},\mOpp{v_3}{i}$ have full column rank $k$.
-  % PL: need to add rank constraint; otherwise, doesn't imply exclusive views
-  Let $\sV_{h_i}$ denote any set of views for $h_i$.
-\end{definition}
-
-% DONE: don't talk about parameter recovery yet, just establish exclusive views
-%Finally, the claim we make is that we can recover parameters for any
-  %graphical model where every hidden variable is a bottleneck:
-Now define the following property:
-\begin{property}[Uniformly bottlenecked]
-  \label{prop:bottleneck}
-  A clique $\sC$ is \emph{uniformly bottlenecked} if
-  every hidden variable in $\sC$ is a bottleneck. 
-  If every clique $\sC \in \sG$ is uniformly bottlenecked, then we say
-  that the graph $\sG$ is uniformly bottlenecked as well.
-\end{property}
-Note that two hidden variables are allowed to have overlapping views.
-%If every hidden variable is a bottleneck, then we can estimate $\mOpp{v}{i}$
 It is clear that having uniform bottlenecks implies part (ii)
@@
 
-To rule out edge cases, we assume the graphical model only has conditional
-independences given by the graph:
-\begin{property}[Faithful representation]
-  \label{prop:ci}  
-  %If there is an active trail from any two variables $a, b$ in the graph
-    %$\sG$, then they are not independent, i.e $a \not\perp b$. 
-  We say that a graphical model $\sG$ is \emph{faithful}
-  if for any variables $a,b,c$
-  such that there is an active trail\footnote{See \todo{cite textbook} for a definition.} between $a$ and $b$ conditioned on $c$,
-  then $a$ and $b$ are dependent given $c$ (i.e., $a\not\perp b \given c$).
-\end{property}
+The exclusive views property 
+Next, we define a (in analogy with biconnected component)
+\begin{definition}[Bidependent set]
+Given a graphical model $\sG$, we say that a subset of nodes $S$ is \emph{bidependent} if
+conditioned on any $a \in S$, there is an active trail between any other two nodes $b,c \in S$.
+\end{definition}
+Intuitively
+Note that all cliques are bidependent, but 
+Bottlenecks provide three views, but to satisfy condition (ii) of \propertyref{exclusive-views},
+only two is required.
 
-%We will prove that the following simple property guarantees that all cliques
-%have the exclusive views property:
-
-%Applying \TensorFactorize to every bottleneck gives us a set of views
-%  for every hidden variable. 
-%To proceed, we need to check that \
-%%(i) understand the assumptions under which
-%  %\assumptionref{full-rank} holds for each bottleneck, and (ii)
-%  that the candidate views actually imply that \propertyref{exclusive-views}
-%  holds for every clique in the graph.
-
-%Before we establish this relationship,
-We now describe a reduction of any graphical model to
-  a canonical form to make our arguments simpler.
-  % DONE: don't say this yet since reader can't appreciate it.
-  %that will allow us to handle the case where an observed
-  %variable has multiple parents.
-
-\begin{lemma}[Canonical form]
-  \label{lem:reduction}
-Every directed (undirected) graphical model can be transformed into one in which
-  the observed variables are leaves with exactly one parent (neighbor). 
-There is a one-to-one correspondence between the parameters of the
-  transformed and original models.
-\end{lemma}
-\begin{proof}
-  \begin{figure}
-    \centering
-    \subimport{figures/}{reduction.tikz}
-    \caption{Reduction to canonical form.}
-    \label{fig:reduction}
-  \end{figure}
-
-  \providecommand{\hp}{\ensuremath{h_\text{\rm new}}}
-  Let $x_v$ be an observed variable with parents $\Pa(x_v)$ and children $\text{Ch}(x_v)$.
-  Consider the following transformation.
-  Replace $x_v$ with a new hidden variable \hp\ with the same
-  parents $\Pa(x_v)$ and children $\text{Ch}(x_v) \union \{x_{v_1}, x_{v_2}, x_{v_3}\}$,
-  where $x_{v_1},x_{v_2},x_{v_3}$ are three copies of $x_v$
-  (\figureref{reduction}).
-  Define $\Pr(\hp \mid \Pa(x_v)) = \Pr(x_v \mid \Pa(x_v))$ and
-  $\Pr(x_{v_j} \mid \hp) = I$.
-  %$\Pr(x_{v_j} \mid \hp) = \left[ \begin{array}{c} I_{k \times k} \\ 0_{(d-k) \times k} \end{array} \right]$.
-  %The domain of $\hp$ is the same as $x_{v_1},x_{v_2},x_{v_3}$.
-  % PL: cut to save space
-  %\footnote{
-      %Though we have assumed that all hidden variables share the
-      %same domain, $[k]$, our work generalizes to hidden domains of
-      %different sizes, provided \assumptionref{full-rank} is satisfied.
-      %},
-    %and $\mOpp{v}{\textrm{new}} = I$.
-  Then, there is a one-to-one correspondence between every value of
-  $\hp$ and $x_v$. Consequently, for any clique $\sC \contains \hp$, the
-  parameters in the original graphical model can be obtained by
-  substituting $\hp$ with $x_v$.
-
-  This procedure applies straightforwardly for undirected graphical
-  models, considering the neighbors $\sN(x_v)$ instead of its parents
-  and children.
-\end{proof}
-
-Now we establish the link between bottlenecks and exclusive views
-(see \appendixref{exclusive-views} for the complete proof):
+Now we establish the link between bottlenecks and exclusive views:
 \begin{lemma}[Uniformly bottlenecked implies exclusive views]
@@
   If a graphical model $\sG$ is faithful and uniformly bottlenecked, then
-  $\sG$ has the exclusive views property.
+  every bidependent set in $\sG$ has the exclusive views property (and hence so does $\sG$).
 \end{lemma}
 \begin{proof}
-We prove the contrapositive: suppose that there is some hidden clique $\sC \in \sG$ with $h_0 \in \sC$
-that does not have exclusive views.  This means each observed variable $x_i$ is conditionally dependent on some
-$h_i \in \sC \backslash \{h_0\}$ given $h_0$.
+We prove the contrapositive: suppose there exists a biconnected subgraph of $\sH \subset \sG$ consisting of hidden nodes
+such that some $h_0 \in \sH$ does not have an exclusive view.
+This means each observed variable $x_i$ in the graph is conditionally dependent on some
+$h_i \in \sH \backslash \{h_0\}$ given $h_0$.
 Consider any such two $x_1, x_2$.
-Conditioned on $h_0$, there exists an active trail between them (specifically, $x_1 - h_1 - h_2 - x_2$).
+Conditioned on $h_0$,
+there is an active trail between $h_1$ and $h_2$ because $\sH$ is biconnected.
+Thus, there is an active trail $x_1 - h_1 - h_2 - x_2$ conditioned on $h_0$.
 Since $\sG$ is faithful, we have $x_1 \not\perp x_2 \mid h_0$.
@@
 
-      \FOR{each hidden variable $h_i \in H$} 
+      \FOR{each hidden variable $h_i \in \bh$} 
         \STATE Apply $\TensorFactorize$ to learn conditional moments
@@
 are $O \eqdef \Pr(x_i|h_i)$ and $T \eqdef \Pr(h_{i+1} | h_i)$
-for all $i$. % (i.e. we have parameter sharing).
+for all $i$.
 While the first and last hidden variables $h_1, h_M$ in the
@@
   due to parameter sharing.
-  %whose parameters we know because they share parameters, $O$.
-%We first use the bottleneck $h_2$ with views $x_1, x_2, x_3$ to estimate $O$,
-%and then recover $\pi$ from $\{h_1\}$ and recover $T$ from the clique $\{h_{1}, h_{2}\}$.
 
@@
   (exclusive to $h_2$) and $x^a_3$ (exclusive to $h_1$).
-  %Note that while
-  %$x^a_3$ is also a view for $h_2$, $x^a_3$ is independent of $h_2$ given
-  %$h_1$.
-
-%In step 1, we recover the parameters $O$ from the bottleneck $h_2$ with
-%  views $\{x^a_2, x^b_2, x^a_3\}$. We also recover the conditional moments
-%  $\mOpp{2}{1}$, $\mOpp{3}{1}$, $\mOpp{4}{1}$ for $h_1$. 
-%In step 2, we can recover $\pi$ from the clique $\{h_1\}$, using any
-%  one of views (they are all exclusive). 
-%To recover $T$ from the clique $\{h_1, h_2\}$, we use the views $x^a_2$
-%  (exclusive to $h_2$) and $x^a_3$ (exclusive to $h_3$). Note that while
-%  $x^a_3$ is also a view for $h_2$, $x^a_3$ is independent of $h_2$ given
-%  $h_1$.
 
@@
 
-%In a certain sense, having three views is necessary for identifiability;
-%two views is certainly not enough.
-% PL: perhaps too simple, saving space
-%The simplest example of a model that cannot be recovered by our algorithm is
-%  a mixture model with a single view, i.e. \figureref{three-view} with
-%  just the variables $h_1$ and $x_1$.
-%However, without further assumptions the model itself is not
-%  identifiable\footnote{To see this, observe that probability-mass
-%  can be arbitrarily exchanged between $\Pr(h_1)$ and $\Pr(x_1 | h_1)$ while preserving the marginals $\Pr(x_1)$}.
-%A class of binary-valued noisy-or networks
-%(see \figureref{examples-noisy-or} for an example)
-%which do not possess the
-%  bottleneck property are nonetheless identifiable;
-%  see \citet{halpern13noisyor} for an algorithm.
-%Having three views is in some sense necessary for the uniqueness of tensor factorization.
 There are certainly models which are identifiable but do not have exclusive views.
@@
   but does not satisfy the exclusive views property.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Sample complexity}
-\label{sec:sampleComplexity}
-% PL: "combine" is too vague to be useful
-%\LearnMarginals combines two consistent algorithms, \TensorFactorize and
-%\LearnCliqueNs, and is thus consistent itself.
-
-%In this section, we provide formal conditions under which $\LearnMarginals$
-%will produce consistent estimates.
-%We let
-%$\mOpphat{v}{i}$,
-%$\hat Z_\sC$,
-%and $\hat M_\sV$,
-%denote estimators for
-%$\mOpp{v}{i}$,
-%$Z_\sC$,
-%and $M_\sV$,
-%respectively.
-
-% But $\mOpp{v}{i}$ is a product of tensors on the path from $h_i$ to
-%   $x_v$, marginalizing out the rest of the graph.
-% To make this property more interpretable, let us assume the following
-%   property which we prove to be sufficient in
-%   \appendixref{assumption-proof}:
-% \begin{assumption} 
-% \label{asm:full-rank-plus}
-% The marginal distribution of every latent variable has full support:
-%   $\forall{h_i \in H},~ \mPi{i} \succ 0$.
-% For every clique $\sC \in \sG$ (including ones with observed variables),
-%   and every conditional distribution $\Pr(h_c \given \Pa(h_c))$, where
-%   $h_c \in \sC$, every {\em tensor unfolding} has full column rank, $k$.
-% \end{assumption}
-% In our directed grid example (\sectionref{directedExample}), for the
-%   clique $\sC = \{ h_1, h_2 \}$ this condition implies $\pi = \Pr(h_1)
-%   \succ 0$ and $T = \Pr(h_2 \mid h_1)$ has column rank $k$; for the clique
-%   $\sC = \{ h_1, x_1 \}$, it implies $O = \Pr(x_1 \mid h_1)$ has column
-%   rank $k$.
-
-%\todo{did they prove this; if so just say "they show", not "we can show"}
-%Using results from
-%  \citet{anandkumar12moments,anandkumar13tensor} we can show that
-%  learning $\mOpp{v_1}{i}$ for the bottleneck $h_i$ with views $x_{v_1},
-%  x_{v_2}, x_{v_3}$ has the following sample complexity,
-%  \todo{define hat O}
-%  \todo{max subscript is misaligned}
-%\begin{align*}
-%  \|\mOpphat{v_1}{i} - \mOpp{v_1}{i}\|^2_F &= \frac{1}{\sqrt{n}} O\left( \frac{k {\pi\oft{i}}_{\max}^2}{\sigma_k(M_{v_1,v_2})^5} \right). 
-%\end{align*}
-It is easy to see that $\LearnMarginals$ has a polynomial sample complexity because it composes two parts that individually have polynomial sample complexity.
-From \citet{anandkumar13tensor} given the non-degeneracy assumptions
-  \assumptionref{non-degeneracy}, we have that $\mOpphat{v}{i}$
-  converges to $\mOpp{v}{i}$ at a rate of $n^{-\frac12}$ with a constant
-  that depends polynomially on the $k$-th singular value of
-  $\mOpp{v}{i}$.
-% PL: we don't have a crisp theorem (because there's multiple paths), so leave it vague.
-Recovering the marginals $Z_\sC$ from the conditional moments via $\LearnClique$ is a linear operation and thus also has polynomial sample complexity. 
-Note that $\sigma_{k}(\mOpp{v}{i})$ can become extremely
-small if $h_i$ and $x_v$ are connected via many intermediate hidden variables.
-\todo{cite work - Sontag?}
-%\begin{theorem}[Sample complexity for $\mOpp{v}{i}$]
-%  \label{thm:sample-complexity}
-%  Without loss of generality, let $v=1, i=1$, and let $x_1, x_2, x_3$ be three views for
-%  $h_1$. With probability at least $1 - \delta$,
-%\begin{align*}
-%  \|\mOpphat{1}{1} - \mOpp{1}{1}\|_F 
-%    &\le  \\
-%    &
-%  \hspace{-0.3in}
-%      O\left(
-%      % PL: remove pi_max because that's upper bounded by 1.
-%      \frac{k \log(1/\delta) } 
-%      {\sqrt{n} \, \pi\oft{1}_{\min} (\sigma_{k}(\mOpp{1}{1}) \sigma_{k}(\mOpp{1}{2}))^{5/2}} \right),
-%\end{align*}
-%where $\pi\oft{1}_{\min}$ is the smallest element of $\pi\oft{1}$.
-
-%If $\sigma_{\min}(\Pr(c \given \Pa(h_c))) \le \sigma$ for every clique
-%  $\sC \in \sG$ and hidden variable $h_c \in \sC$,
-%\begin{align*}
-%  \|\mOpphat{1}{1} - \mOpp{1}{1}\|_F 
-%    &\le 
-%      O\left( 
-%      \frac{k \log(1/\delta) {\pi\oft{1}}_{\max}/{\pi\oft{1}}_{\min}} 
-%      {\sqrt{n} \sigma^{5t}} \right) \epsilon,
-%\end{align*}
-%where $t$ is the length of the shortest topological ordering from $h_1$
-%to the unique parent of $x_1$, $h_t$.
-%\end{theorem}
-%In general, $\sigma_{k}(\mOpp{v}{i})$ will depend exponentially on the
-  %path length between the hidden variable $h_i$ and $x_v$.
-%This is expected because the influence from $h_i$ to $x_v$
-  %gets diluted for every latent variable it must factor through.
-
-%\theoremref{sample-complexity}, proved in \appendixref{assumption-proof}
-%  says that the number of samples required to estimate the view
-%  $\mOpp{v}{i}$ grows exponentially in the {\em distance} between $x_v$
-%  and $h_i$.
-
-%\todo{Present sample complexity bounds for $Z_\sC$?}
-
-% Easy case
-% Next, we need to show that the hidden marginals $\hat Z_\sC$ converge to $Z_\sC$.
-% First, $\hat M_\sV$ is just an empirical average of multinomials over the data points.
-% Abusing notation slightly, we let $\hat M_\sV$ also denote its
-% $d^{m}$-dimensional vectorized version.
-% We also represent $\mH_\sC$ as
-%   a vector in $\Re^{k^m}$, and represent $\mOppAll \eqdef
-%   \mOpp{v_1}{i_1} \otimes \cdots \otimes
-%   \mOpp{v_m}{i_m}$ as a matrix in $\Re^{d^m \times
-%   k^m}$.
-% By the central limit theorem, we have:
-% $\sqrt{n} (\hat M_\sV - M_\sV) \convind \sN(0, \Sigma_\sV)$,
-% where $\Sigma_\sV$ is the \emph{asymptotic variance} of $\hat M_\sV$: 
-% % ARUN: I don't think so - \todo{doesn't this need to be multiplied by something?}
-% \begin{align*}
-% \Sigma_\sV \eqdef \dM_\sV - M_\sV M_\sV^\top, \quad \dM_\sV \eqdef \text{diag}(M_\sV).
-% \end{align*}
-% 
-% Our next step will be to use the delta-method to convert the above
-% result into the asymptotic variance for the pseudoinverse version of
-% $\LearnClique$. 
-% \begin{lemma}[Asymptotic variance of pseudoinverse estimator for $\tilde Z_\sC$]
-%   \label{lem:mom-variance}  
-%   Assume $\hat M_{\sV}$ has asymptotic variance $\Sigma_\sV$ defined above.
-%   Then the asymptotic variance of $\hat{Z_\sC}$ is:
-%   \begin{align*}
-%     \Sigma^{\mom} &= \mOppAlli \Sigma_\sV \mOppAllit.
-%   \end{align*}
-%   %where ${\mOppAll}\pinv \eqdef {\mOpp{v_1}{i_1}}\pinv \otimes
-%   %\cdots \otimes {\mOpp{v_m}{i_m}}\pinv$, a $d^m \times k^m$ matrix.
-% \end{lemma}
-% \begin{proof}
-% For clique $\sC$, recall we have
-%   $Z_{\sC} = \mO_\sV(\mOppi{v_1}{i_1},\cdots,\mOppi{v_m}{i_m})$.
-% %Choosing to represent $Z_\sC$ and $M_\sV$ as vectors and
-% %$\mOppi{v_1}{i_1} \otimes \cdots \otimes \mOppi{v_m}{i_m}$ as the
-% %matrix,
-% We can rewrite this representation more compactly as: $Z_{\sC} = {\mOppAll}\pinv \mO_\sV$.
-% By the delta-method \cite{vaart98asymptotic},
-% % DONE: need to shorten
-% %we have that the asymptotic variance of $M_\sV$ is:
-% %\begin{align*}
-% %  \sqrt{n}(\hat M_\sV - M_\sV) \convind \sN(0, \Sigma_\sV),
-% %\end{align*}
-% %where $\Sigma_\sV$ is the variance of the observations. 
-% we immediately get:
-% $\sqrt{n}(\hat Z_\sC - Z_\sC) \convind \sN(0, \mOppAlli \Sigma_\sV \mOppAllit)$.
-% \end{proof}
-
-% DONE: too brazen to say
-%Note that extending these results to finite sample bounds can be done
-  %via a straightforward application of perturbation bounds.

### 2014-05-16 doc/icml2014/background.tex

@@
   For any hidden variables $a,b,c \in \bh$
-  such that there is an active trail between $a$ and $b$ conditioned on $c$,
+  such that there is an active trail\footnote{See
+  \citet{koller2009probabilistic} for a definition.} between $a$ and $b$ conditioned on $c$,
   we have that $a$ and $b$ are dependent given $c$ (i.e., $a\not\perp b \given c$).
@@
 However, $\sigma_{k}(\mOpp{v}{i})$ can become extremely
-small if $h_i$ and $x_v$ are connected via many intermediate hidden variables.
+small if $h_i$ and $x_v$ are connected via many intermediate hidden
+variables.\footnote{To see this, consider a chain-structured graphical model:
+$h_1 \to h_2 \cdots \to h_t \to x_v$. In this example, if
+$\sigma_k(\Pr(h_{i+1} \given h_{i})) \ge \sigma_k$ for each $i = 1,
+\cdots, t-1$, then $\sigma_k(\mOpp{v}{1})$ can be as bad as
+$\sigma_k^t \sigma_k(\mOpp{v}{t})$.}.
 

### 2014-05-16 python/experiments/variances.py
Merge conflict?

@@ -18,12 +18,13 @@ np.random.seed(12)
 
 def generate_random_instance(k, d, badO = False):
     z = normalize(rand(k))
-    if badO:
-        O = 0.5 * column_normalize(rand(d-1, k))
-        O = np.vstack((O, 0.5 * np.ones(k)))
-    else:
-        O = column_normalize(rand(d, k))
-    assert np.allclose(O.sum(), k, 1e-4)
+    O = np.vstack( (0.5 * eye(2), [0.5, 0.5] ) )
+    # if badO:
+    #     O = 0.9 * column_normalize(rand(d-1, k))
+    #     O = np.vstack((O, 0.1 * np.ones(k)))
+    # else:
+    #     O = column_normalize(rand(d, k))
+    # assert np.allclose(O.sum(), k, 1e-4)
     mu = O.dot(z)
     return z, O, mu
 
@@ -112,6 +113,14 @@ def do_experiment(O, Ot, Ondk, mu, args):
 
     return mut_, zt_pi, zt_cl
 
+def get_lhood_gradient(mu_, Ot, z):
+    """
+    Returns - [Ot; -1^T Ot]^T D^-1 e_x
+    """
+    d_, _ = Ot.shape
+    O = np.vstack( (Ot, -ones(d_).T.dot(Ot)) )
+    mu_z = O.dot(z)
+    return O.dot.diag(mu_z).dot(mu_)
 
 def do_command(args):
     np.random.seed(args.seed)
@@ -130,13 +139,17 @@ def do_command(args):
     v = (Ot.dot(Oti)).T.dot(o)
 
     print "O cond", svd(O)[1]
-    print "ones residual", norm( Ot.T.dot(o) ), norm( v )
     print "mu residual", (Oti.dot(mut)), norm(Oti.dot(mut))
+    print "ones residual", norm( ones((D-1,1)).T.dot(Ot) )
 
     dZ_pi, dZ_cl = [], []
 
     for attempt in xrange(attempts):
         mut_, zt_pi, zt_cl = do_experiment(O, Ot, Ondk, mu, args)
+        print get_lhood_gradient(mut_, Ot, zt_cl)
+        # l'n(z) = l'n(z*) + ln''(z*) (z'-z*)
+        # l'n(z) = O[x]/mu(z)
+        #"z - z* = l''(z*)^inv -l'n(z)"
         if attempt % (attempts/20) == 0:
             sys.stderr.write('.')
             sys.stderr.flush()

### 2014-05-16 doc/icml2014/undirected.tex

@@
 \begin{align}
-L_\text{unsup}(\theta) \eqdef \E_{\vx \sim \sD}[\log \sum_{\vh \in \sH} p_\theta(\vx,\vh)].
+  L_\text{unsup}(\theta) \eqdef \E_{\vx \sim p^*}[\log \sum_{\vh \in \sH} p_\theta(\vx,\vh)],
 \end{align}
+where we are assuming for simplicity that we have infinite data drawn from the true distribution $p^*$.
 % PL: this doesn't make sense
@@
 On the other hand, 
-if we were able to observe $\vh$, the \emph{supervised} likelihood would be concave:
+if we were able to observe $\vh$, we would optimize \emph{supervised} likelihood, which is concave:
 \begin{align}
 \label{eqn:logLinearSupervised}
-L_\text{sup}(\theta) &\eqdef \E_{(\vx,\vh) \sim \sD_\text{sup}}[\log p_\theta(\vx,\vh)] \\
+L_\text{sup}(\theta) &\eqdef \E_{(\vx,\vh) \sim p^*}[\log p_\theta(\vx,\vh)] \\
                      &= \theta^\top \left(\sum_{\sC \in \sG} \E[\phi(\vx_\sC,\vh_\sC)]\right) - A(\theta).
@@
 While we don't have supervised data,
-note that the method of moments gives us hidden marginals $Z_\sC = \BP(\vh_\sC)$.
-Furthermore, given $\mOpp{v}{i}$ and $Z_\sC$ of any $\sC \ni v$, we can recover
-all marginals $\BP(\vx_\sC, \vh_\sC) = Z_\sC \cdot \mOppAll[\vx_\sV]$.%\todo{clean up}.
-Finally, the key step is that we can compute all the expected feature vectors from the marginals:
+recall that the method of moments provides the hidden marginals $Z_\sC = \BP(\vh_\sC)$
+and conditional moments $\mOpp{v}{i} = \Pr(x_v \mid h_i)$.
+Assuming our graphical model is in canonical form (\lemmaref{reduction}),
+every observed variable is connected to exactly one hidden variable.
+Then $Z_\sC$ and $\mOpp{v}{i}$ directly yield marginals for every clique $\BP(\vx_\sC, \vh_\sC)$.
+Given these marginals, we can compute the expected local feature vector:
 \begin{align}
@@
 \end{align}
-Therefore, we can optimize the supervised likelihood objective without actually
-having supervised data.
-Of course, in the finite data regime, the method of moments yields estimates
-$\hat \mu^\text{mom}_\sC$ which approximate the true $\mu_\sC$.
-In supervised learning, we have different estimate $\hat\mu^\text{sup}_\sC$ of $\mu_\sC$ based on an empirical average
-over the data points.
+Note that $\{\mu_\sC\}_{\sC \in \sG}$ are exactly the sufficient statistics of the model (\equationref{logLinearSupervised}),
+so we can optimize the supervised likelihood objective without actually having any supervised data!
+In the finite data regime, the method of moments yields the estimate
+$\hat \mu^\text{mom}_\sC$ which approximates the true $\mu_\sC$.
+In supervised learning, we obtain a different estimate $\hat\mu^\text{sup}_\sC$ of $\mu_\sC$ based on an empirical average
+over data points.
 In the limit of infinite data, both estimators converge to $\mu_\sC$.
@@
 In \sectionref{directed}, we showed that $\LearnMarginals$ is a consistent
-estimator for $Z_\sC$ for any graphical model satisfying the exclusive views property (\propertyref{exclusive-views}). 
-Consequently, the result is that the output of \algorithmref{undirected},
-$\hat\theta$, is a consistent estimator.\footnote{Of course, the log-linear model must be identifiable
+estimator for $Z_\sC$ for any graphical model satisfying the uniformly bottlenecked property (\propertyref{bottleneck}). 
+Consequently, the output of \algorithmref{undirected}
+is a consistent estimator.\footnote{Of course, the log-linear model must be identifiable
 (it is violated, for example, if two components of $\phi(\vx_\sC, \vh_\sC)$ were identical).
@@
     \REQUIRE A graphical model $\sG$ satisfying \propertyref{bottleneck}, data $\sD$
-    \ENSURE A consistent estimate of parameters $\hat \theta$
+    \ENSURE Parameters $\theta$
       \STATE Apply $\LearnMarginals$ to $\sG$ to learn clique marginals $Z_\sC$.
-      \STATE Compute expected features (\equationref{logLinearFeatures}) and optimize \equationref{logLinearSupervised}.
+      \STATE Compute expected features $\{\mu_\sC\}_{\sC \in \sG}$ (\equationref{logLinearFeatures}) and optimize \equationref{logLinearSupervised}.
   \end{algorithmic}
@@
 graphs with high treewidth present a challenge
-for even fully-supervised maximum likelihood due to the intractability of the log-partition function.
+for even fully-supervised maximum likelihood due to the intractability of the log-partition function $A(\theta)$.
+%One can employ various variational approximations of $A(\theta)$ \cite{wainwright08varinf},
 This motivates looking at pseudolikelihood objective of \citep{besag75pseudo} for a computationally more
 efficient solution.
-The pseudolikelihood objective sums over the log-probability of each variable $a$ given its neighbors $\sN(a)$:
-$$L_\text{pseudo}(\theta) \eqdef \E_{x \sim \sD}\left[\sum_{a \in X \cup H} \log p_\theta(a \mid x_{\sN(a)})\right].$$
+The pseudolikelihood objective is a sum over the log-probability of each variable $a$ given its neighbors $\sN(a)$:
+\begin{align}
+  \label{eqn:pseudolikelihood}
+L_\text{pseudo}(\theta) \eqdef \E_{(\bx,\bh) \sim p^*}\left[\sum_{a \in X \cup H} \log p_\theta(a \mid \sN(a))\right].
+\end{align}
 In the fully-supervised setting, it is well-known that pseudolikelihood provides computationally efficient consistent estimates,
@@
 
-The pseudolikelihood also replaces the log-partition function
-  $A(\theta)$ with a sum over the variables in the neighborhood:
-  $A_a(\theta; x_{\sN(a)}) = \sum_{a' \in \{a\} \union \sN(a)}
-  \exp(\theta^\top\phi_{\sN(a)}(a'))$.
-Note that the neighbors $\sN(a)$ in general do not form a clique, and
-  thus the uniformly bottlenecked assumption alone is not sufficient to
-  guarantee exclusive views for each variable in the subgraph.
-However, when we do have exclusive views for each neighborhood, as in
-  \figureref{examples-mrf}, we can adapt our earlier algorithm
-  ($\LearnClique$) to estimate the marginals over all neighborhoods $Z_{\sN(a)}$.
+Let $\phi_{a,\sN(a)}(a, \sN(a)) = \sum_{\sC \ni a} \phi_\sC(\bx_\sC,\bh_\sC)$
+denote the sum over cliques $\sC$ that contain $a$; note that only depends on $a$ and its neighbors $\sN(a)$.
+We can write each conditional log-likelihood as:
+\begin{align}
+p_\theta(a \mid \sN(a)) = \exp(\theta^\top\phi_{a,\sN(a)}(a, \sN(a)) - A_a(\theta; \sN(a))),
+\end{align}
+where the conditional log-partition function $A_a(\theta; \sN(a)) = \log \sum_{a'} \exp(\theta^\top\phi_{a,\sN(a)}(a', \sN(a)))$
+involves summing only over values to a single variable.
+
+If we knew $\mu_{a,\sN(a)} \eqdef \E[\phi_{a,\sN(a)}(a,\sN(a))]$,
+then we would be able to optimize the pseudolikelihood objective again without
+having access to any labeled data.
+%The pseudolikelihood also replaces the log-partition function
+%  $A(\theta)$ with a sum over the variables in the neighborhood:
+%  $A_a(\theta; x_{\sN(a)}) = \sum_{a' \in \{a\} \union \sN(a)}
+%  \exp(\theta^\top\phi_{\sN(a)}(a'))$.
+Note that $\{a\}\cup\sN(a)$ in general do not form a clique.
+Fortunately, \lemmaref{bottleneck-views} holds for all biconnected subgraphs.
+We can break $\{a\}\cup\sN(a)$ into a collection of 
+  % PL: this is not necessary
+  %thus the uniformly bottlenecked assumption alone is not sufficient to
+  %guarantee exclusive views for each variable in that subgraph.
+%However, when we do have exclusive views for each neighborhood, as in
+  \figureref{examples-mrf},
+  Therefore, we can adapt our algorithm ($\LearnClique$) to estimate all marginals of the form $Z_{a,\sN(a)} \eqdef \Pr(a,\sN(a))$,
+  from which we can compute the features $\mu_{a,\sN(a)}$,
+  and optimize \equationref{pseudolikelihood} to obtain a consistent estimate of $\theta$.
 

### 2014-05-17 doc/icml2014/background.tex
intermediate sync

--- 2014-05-16 doc/icml2014/background.tex
@@
   For any hidden variables $a,b,c \in \bh$
-  such that there is an active trail between $a$ and $b$ conditioned on $c$,
+  such that there is an active trail\footnote{See
+  \citet{koller2009probabilistic} for a definition.
+  We do condition on observed variables.} between $a$ and $b$ conditioned on $c$,
   we have that $a$ and $b$ are dependent given $c$ (i.e., $a\not\perp b \given c$).
@@
 \begin{definition}[Bottleneck]
+  \label{def:bottleneck}
   A hidden variable $h_i$ is said to be a \emph{bottleneck} if there
@@
 However, $\sigma_{k}(\mOpp{v}{i})$ can become extremely
-small if $h_i$ and $x_v$ are connected via many intermediate hidden variables.
+small if $h_i$ and $x_v$ are connected via many intermediate hidden
+variables.\footnote{To see this, suppose $h_1$ has a view $x_v$ via a chain:
+$h_1 - h_2 \cdots - h_t - x_v$. In this example, if
+$\sigma_k(\Pr(h_{i+1} \given h_{i})) = a_k$ for each $i = 1,
+\cdots, t-1$, then $\sigma_k(\mOpp{v}{1}) = a_k^t \sigma_k(\mOpp{v}{t})$.}.
 

### 2014-05-17 doc/icml2014/directed.tex
intermediate sync

@@
 The key property required for our algorithm to work is as follows:
-\begin{property}[Exclusive views]
-  \label{prop:exclusive-views}
-Given a graphical model $\sG$,
-  a subset of hidden variables $S \subset \bh$ is said to satisfy the \emph{exclusive views property} if for
-  each $h_i \in S$, the two conditions hold:
+\begin{definition}[Exclusive views]
+  \label{def:exclusive-views}
+Let $S \subseteq \bh$ be a subset of hidden variables.
+We say $h_i \in S$ has an exclusive view $x_v$
+  if the two conditions hold:
   (i) there exists some observed variable
@@
   \Pr(x_{v} \mid h_i)$ have full column rank $k$ and can be recovered.
-We call $x_v$ an exclusive view of $h_i$ in $S$.
+We say that $S$ has the \emph{exclusive views property} if every $h_i \in S$ has an exclusive view.
 %If every hidden clique has the exclusive views property, then we say that $\sG$ does as well.
-\end{property}
+\end{definition}
 
@@
 Since each $\mOpp{v}{i}$ has full column rank $k$,
-$\mOppAll$ has full column rank $k^m$.
+the tensor product $\mOppAll$ has full column rank $k^m$.
 Succinctly, $\mO_\sV$ (which can be estimated directly from data)
@@
   \begin{algorithmic}
-    \REQUIRE Hidden subset $S$ with exclusive views (\propertyref{exclusive-views}).
-    \ENSURE Marginal distribution $Z_S$.
-      \STATE Identify exclusive views $\Sx{\sV} = \{x_{v_1}, \dots, x_{v_m}\}$.
+    \REQUIRE Hidden subset $S = \{ h_{i_1}, \dots, h_{i_m} \}$ with exclusive views $\sV = \{ x_{v_1}, \dots, x_{v_m} \}$ and $\{\mOpp{v_1}{i_i}\pinv\}$ % (\definitionref{exclusive-views})
+    \ENSURE Marginal distribution $Z_S = \Pr(\bh_S)$.
+      %\STATE Identify exclusive views $\Sx{\sV} = \{x_{v_1}, \dots, x_{v_m}\}$.
       \STATE Return $\mH_S \gets \mO_{\sV}( \mOpp{v_1}{i_i}\pinv, \dots, {\mOpp{v_m}{i_m}}\pinv )$.
@@
 We have shown how to estimate the parameters of any clique that possesses the exclusive
-  views property (\propertyref{exclusive-views}).
+  views property (\definitionref{exclusive-views}).
   But in a general graph, which cliques have this property?
@@
 
-It is clear that having uniform bottlenecks implies part (ii)
-of the exclusive views property (\propertyref{exclusive-views}) via \TensorFactorize,
-but we will show that part (i) follows as well.
-
-The exclusive views property 
-Next, we define a (in analogy with biconnected component)
+% Define bidependent
+First, define the notion of a bidependent set (in analogy with biconnected component),
+in which conditionining on one variable does not break the set apart:
 \begin{definition}[Bidependent set]
@@
 \end{definition}
-Intuitively
-Note that all cliques are bidependent, but 
-Bottlenecks provide three views, but to satisfy condition (ii) of \propertyref{exclusive-views},
-only two is required.
+Note that all cliques are bidependent, but bidependent sets can have more conditional independences.
+this will be important for pseudolikliehood in \sectionref{pseudolikelihood}.
 
-Now we establish the link between bottlenecks and exclusive views:
-\begin{lemma}[Uniformly bottlenecked implies exclusive views]
+% bidependent, bottleneck => exclusive views
+Bidependent sets are significant because they guarantee exclusive views if they are bottlenecks:
+%If a subset of hidden nodes $S$ is a bottleneck, it is clear that (ii)
+%of the exclusive views property (\definitionref{exclusive-views}) via \TensorFactorize,
+%but we will show that part (i) follows as well.
+%Now we establish the link between bottlenecks and exclusive views:
+\begin{lemma}[Bottlenecked implies exclusive views]
   \label{lem:bottleneck-views}  
-% PL: simplify to cliques (this is also bad notation, since \sG is a set of cliques, not nodes)
-%Let $\sG' = \{ h_{i_1}, \dots, h_{i_m} \} \subseteq \sG$ be
-  %a connected subgraph of hidden variables.
-  %If every hidden variable
-  %$h_i \in \sG'$ is a bottleneck, then $\sG'$ has the exclusive views property.
-  If a graphical model $\sG$ is faithful and uniformly bottlenecked, then
-  every bidependent set in $\sG$ has the exclusive views property (and hence so does $\sG$).
+  Let $S \subseteq \bh$ be a bidependent subset of hidden variables.
+  If $S$ is a bottleneck, then $S$ has the exclusive views property.
 \end{lemma}
 \begin{proof}
-We prove the contrapositive: suppose there exists a biconnected subgraph of $\sH \subset \sG$ consisting of hidden nodes
-such that some $h_0 \in \sH$ does not have an exclusive view.
-This means each observed variable $x_i$ in the graph is conditionally dependent on some
-$h_i \in \sH \backslash \{h_0\}$ given $h_0$.
-Consider any such two $x_1, x_2$.
-Conditioned on $h_0$,
-there is an active trail between $h_1$ and $h_2$ because $\sH$ is biconnected.
+Let $S$ be a bidependent subset and fix any $h_0 \in S$.
+Since $h_0$ is a bottleneck, it has three conditionally independent views,
+say $x_1, x_2, x_3$ without loss of generality. 
+For condition (i), we will show that at least one of the views is conditionally independent
+of $S \backslash \{ h_0 \}$ given $h_0$.
+For the sake of contradiction, 
+suppose that each observed variable $x_i$ is conditionally dependent on some
+$h_i \in S \backslash \{h_0\}$ given $h_0$, for $i \in \{1, 2, 3\}$.
+Then conditioned on $h_0$,
+there is an active trail between $h_1$ and $h_2$ because $S$ is biconnected.
 Thus, there is an active trail $x_1 - h_1 - h_2 - x_2$ conditioned on $h_0$.
-Since $\sG$ is faithful, we have $x_1 \not\perp x_2 \mid h_0$.
-As this holds for any $x_1,x_2$, $h_0$ cannot be a bottleneck.
+Since the graphical model is faithful by assumption, we have $x_1 \not\perp x_2 \mid h_0$,
+contradicting the fact that $x_1$ and $x_2$ are conditionally independent given $h_0$.
+To show condition (ii), assume, without loss of generality, that $x_1$ is an exclusive view.
+Then we can recover $\mOpp{1}{0} = \Pr(x_1 \mid h_0)$ via $\TensorFactorize$.
 \end{proof}
 
-With this lemma in place, we present the full algorithm, $\LearnMarginals$,
+%Note that bottlenecks provide three views, which is important for condition (ii) of
+%\definitionref{exclusive-views}, only two is required for condition (i).
+With this lemma in place, we present our full algorithm, $\LearnMarginals$,
 in \algorithmref{directed}.
@@
   \begin{algorithmic}
-    \REQUIRE Graphical model $\sG$ satisfying \propertyref{bottleneck}, data $\sD$
-    \ENSURE Marginals $Z_\sC$ for every clique $\sC \in \sG$
-
-      \FOR{each hidden variable $h_i \in \bh$} 
+    \REQUIRE Bidependent and bottlenecked subset of hidden nodes $S \subseteq \bh$
+    \ENSURE Marginals $Z_S = \Pr(\bh_S)$ % for every bidependent subset of hidden variables $S \in \sG$
+    \FOR{each hidden node $h_i \in S$ with exclusive view $x_v$} 
       \STATE Apply $\TensorFactorize$ to learn conditional moments
-        $\mOpp{v}{i}$ for every $v \in \sV_{h_i}$.
-
-%    \COMMENT{Recover observation potentials $O$ using bottlenecks}
+      $\mOpp{v}{i} = \Pr(x_v \mid h_i)$.
     \ENDFOR
-%      \COMMENT{\textbf{Step 2:} Recover clique potentials from the piecewise likelihood.}
-\FOR{every clique $\sC = \{h_{i_1}, \ldots, h_{i_m}\} \in \sG$} 
     \STATE Apply $\LearnClique$ to learn the marginals $\mH_\sC$.
-\ENDFOR
   \end{algorithmic}
@@
 % DONE: move here because not central to story
-\paragraph{Remarks.} We note that \propertyref{bottleneck} can be relaxed if some cliques
+\paragraph{Remarks.}
+Note that even having two independent views for each $h_i \in S$ is sufficient for condition (i)
+  of the exclusive views property, while three is needed for condition (ii).
+The bottleneck property (\definitionref{bottleneck} can also be relaxed if some cliques
   share parameters.

### 2014-05-17 doc/icml2014/undirected.tex
intermediate sync

@@
 
-\paragraph{Pseudolikelihood.}
+\subsection{Pseudolikelihood}
+\label{sec:pseudolikelihood}
+
 Unlike for directed models, in the undirected case,

### 2014-05-17 doc/icml2014/directed.tex
Fixed typos in piecewise

--- 2014-05-16 doc/icml2014/directed.tex
@@
 
-It is clear that having uniform bottlenecks implies part (ii)
-of the exclusive views property (\propertyref{exclusive-views}) via \TensorFactorize,
-but we will show that part (i) follows as well.
+%It is clear that having uniform bottlenecks implies part (ii)
+%of the exclusive views property (\propertyref{exclusive-views}) via \TensorFactorize,
+%but we will show that part (i) follows as well.
 
-The exclusive views property 
-Next, we define a (in analogy with biconnected component)
+First, define the notion of a bidependent set (in analogy with biconnected component),
+in which conditionining on one variable does not make anything independent:
 \begin{definition}[Bidependent set]
@@
 \end{definition}
-Intuitively
-Note that all cliques are bidependent, but 
+Note that all cliques are bidependent, but bidependent can have fewer dependencies.
+
 Bottlenecks provide three views, but to satisfy condition (ii) of \propertyref{exclusive-views},
@@
 Now we establish the link between bottlenecks and exclusive views:
-\begin{lemma}[Uniformly bottlenecked implies exclusive views]
+\begin{lemma}[Bottlenecked implies exclusive views]
   \label{lem:bottleneck-views}  
-% PL: simplify to cliques (this is also bad notation, since \sG is a set of cliques, not nodes)
-%Let $\sG' = \{ h_{i_1}, \dots, h_{i_m} \} \subseteq \sG$ be
-  %a connected subgraph of hidden variables.
-  %If every hidden variable
-  %$h_i \in \sG'$ is a bottleneck, then $\sG'$ has the exclusive views property.
-  If a graphical model $\sG$ is faithful and uniformly bottlenecked, then
-  every bidependent set in $\sG$ has the exclusive views property (and hence so does $\sG$).
+  If a bidependent subset of hidden variables $S$ is a bottleneck,
+  then $S$ has the exclusive views property.
 \end{lemma}
@@
   but does not satisfy the exclusive views property.
+
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \subsection{Sample complexity}
+% \label{sec:sampleComplexity}
+% PL: "combine" is too vague to be useful
+%\LearnMarginals combines two consistent algorithms, \TensorFactorize and
+%\LearnCliqueNs, and is thus consistent itself.
+
+%In this section, we provide formal conditions under which $\LearnMarginals$
+%will produce consistent estimates.
+%We let
+%$\mOpphat{v}{i}$,
+%$\hat Z_\sC$,
+%and $\hat M_\sV$,
+%denote estimators for
+%$\mOpp{v}{i}$,
+%$Z_\sC$,
+%and $M_\sV$,
+%respectively.
+
+% But $\mOpp{v}{i}$ is a product of tensors on the path from $h_i$ to
+%   $x_v$, marginalizing out the rest of the graph.
+% To make this property more interpretable, let us assume the following
+%   property which we prove to be sufficient in
+%   \appendixref{assumption-proof}:
+% \begin{assumption} 
+% \label{asm:full-rank-plus}
+% The marginal distribution of every latent variable has full support:
+%   $\forall{h_i \in H},~ \mPi{i} \succ 0$.
+% For every clique $\sC \in \sG$ (including ones with observed variables),
+%   and every conditional distribution $\Pr(h_c \given \Pa(h_c))$, where
+%   $h_c \in \sC$, every {\em tensor unfolding} has full column rank, $k$.
+% \end{assumption}
+% In our directed grid example (\sectionref{directedExample}), for the
+%   clique $\sC = \{ h_1, h_2 \}$ this condition implies $\pi = \Pr(h_1)
+%   \succ 0$ and $T = \Pr(h_2 \mid h_1)$ has column rank $k$; for the clique
+%   $\sC = \{ h_1, x_1 \}$, it implies $O = \Pr(x_1 \mid h_1)$ has column
+%   rank $k$.
+
+%\todo{did they prove this; if so just say "they show", not "we can show"}
+%Using results from
+%  \citet{anandkumar12moments,anandkumar13tensor} we can show that
+%  learning $\mOpp{v_1}{i}$ for the bottleneck $h_i$ with views $x_{v_1},
+%  x_{v_2}, x_{v_3}$ has the following sample complexity,
+%  \todo{define hat O}
+%  \todo{max subscript is misaligned}
+%\begin{align*}
+%  \|\mOpphat{v_1}{i} - \mOpp{v_1}{i}\|^2_F &= \frac{1}{\sqrt{n}} O\left( \frac{k {\pi\oft{i}}_{\max}^2}{\sigma_k(M_{v_1,v_2})^5} \right). 
+%\end{align*}
+%It is easy to see that $\LearnMarginals$ has a polynomial sample complexity because it composes two parts that individually have polynomial sample complexity.
+%From \citet{anandkumar13tensor} given the non-degeneracy assumptions
+%  \assumptionref{non-degeneracy}, we have that $\mOpphat{v}{i}$
+%  converges to $\mOpp{v}{i}$ at a rate of $n^{-\frac12}$ with a constant
+%  that depends polynomially on the $k$-th singular value of
+%  $\mOpp{v}{i}$.
+% PL: we don't have a crisp theorem (because there's multiple paths), so leave it vague.
+%Recovering the marginals $Z_\sC$ from the conditional moments via $\LearnClique$ is a linear operation and thus also has polynomial sample complexity. 
+%Note that $\sigma_{k}(\mOpp{v}{i})$ can become extremely
+%small if $h_i$ and $x_v$ are connected via many intermediate hidden variables
+%\footnote{To see this, consider a chain-structured graphical model: $h_1
+%\to h_2 \cdots \to h_t \to x_v$. In this example, if
+%$\sigma_k(\Pr(h_{i+1} \given h_{i})) \ge \sigma_k$ for each $i = 1,
+%\cdots, t-1$, then $\sigma_k(\mOpp{v}{1})$ can be as bad as
+%$\sigma_k^t \sigma_k(\mOpp{v}{t})$.}.
+%\begin{theorem}[Sample complexity for $\mOpp{v}{i}$]
+%  \label{thm:sample-complexity}
+%  Without loss of generality, let $v=1, i=1$, and let $x_1, x_2, x_3$ be three views for
+%  $h_1$. With probability at least $1 - \delta$,
+%\begin{align*}
+%  \|\mOpphat{1}{1} - \mOpp{1}{1}\|_F 
+%    &\le  \\
+%    &
+%  \hspace{-0.3in}
+%      O\left(
+%      % PL: remove pi_max because that's upper bounded by 1.
+%      \frac{k \log(1/\delta) } 
+%      {\sqrt{n} \, \pi\oft{1}_{\min} (\sigma_{k}(\mOpp{1}{1}) \sigma_{k}(\mOpp{1}{2}))^{5/2}} \right),
+%\end{align*}
+%where $\pi\oft{1}_{\min}$ is the smallest element of $\pi\oft{1}$.
+
+%If $\sigma_{\min}(\Pr(c \given \Pa(h_c))) \le \sigma$ for every clique
+%  $\sC \in \sG$ and hidden variable $h_c \in \sC$,
+%\begin{align*}
+%  \|\mOpphat{1}{1} - \mOpp{1}{1}\|_F 
+%    &\le 
+%      O\left( 
+%      \frac{k \log(1/\delta) {\pi\oft{1}}_{\max}/{\pi\oft{1}}_{\min}} 
+%      {\sqrt{n} \sigma^{5t}} \right) \epsilon,
+%\end{align*}
+%where $t$ is the length of the shortest topological ordering from $h_1$
+%to the unique parent of $x_1$, $h_t$.
+%\end{theorem}
+%In general, $\sigma_{k}(\mOpp{v}{i})$ will depend exponentially on the
+  %path length between the hidden variable $h_i$ and $x_v$.
+%This is expected because the influence from $h_i$ to $x_v$
+  %gets diluted for every latent variable it must factor through.
+
+%\theoremref{sample-complexity}, proved in \appendixref{assumption-proof}
+%  says that the number of samples required to estimate the view
+%  $\mOpp{v}{i}$ grows exponentially in the {\em distance} between $x_v$
+%  and $h_i$.
+
+%\todo{Present sample complexity bounds for $Z_\sC$?}
+
+% Easy case
+% Next, we need to show that the hidden marginals $\hat Z_\sC$ converge to $Z_\sC$.
+% First, $\hat M_\sV$ is just an empirical average of multinomials over the data points.
+% Abusing notation slightly, we let $\hat M_\sV$ also denote its
+% $d^{m}$-dimensional vectorized version.
+% We also represent $\mH_\sC$ as
+%   a vector in $\Re^{k^m}$, and represent $\mOppAll \eqdef
+%   \mOpp{v_1}{i_1} \otimes \cdots \otimes
+%   \mOpp{v_m}{i_m}$ as a matrix in $\Re^{d^m \times
+%   k^m}$.
+% By the central limit theorem, we have:
+% $\sqrt{n} (\hat M_\sV - M_\sV) \convind \sN(0, \Sigma_\sV)$,
+% where $\Sigma_\sV$ is the \emph{asymptotic variance} of $\hat M_\sV$: 
+% % ARUN: I don't think so - \todo{doesn't this need to be multiplied by something?}
+% \begin{align*}
+% \Sigma_\sV \eqdef \dM_\sV - M_\sV M_\sV^\top, \quad \dM_\sV \eqdef \text{diag}(M_\sV).
+% \end{align*}
+% 
+% Our next step will be to use the delta-method to convert the above
+% result into the asymptotic variance for the pseudoinverse version of
+% $\LearnClique$. 
+% \begin{lemma}[Asymptotic variance of pseudoinverse estimator for $\tilde Z_\sC$]
+%   \label{lem:mom-variance}  
+%   Assume $\hat M_{\sV}$ has asymptotic variance $\Sigma_\sV$ defined above.
+%   Then the asymptotic variance of $\hat{Z_\sC}$ is:
+%   \begin{align*}
+%     \Sigma^{\mom} &= \mOppAlli \Sigma_\sV \mOppAllit.
+%   \end{align*}
+%   %where ${\mOppAll}\pinv \eqdef {\mOpp{v_1}{i_1}}\pinv \otimes
+%   %\cdots \otimes {\mOpp{v_m}{i_m}}\pinv$, a $d^m \times k^m$ matrix.
+% \end{lemma}
+% \begin{proof}
+% For clique $\sC$, recall we have
+%   $Z_{\sC} = \mO_\sV(\mOppi{v_1}{i_1},\cdots,\mOppi{v_m}{i_m})$.
+% %Choosing to represent $Z_\sC$ and $M_\sV$ as vectors and
+% %$\mOppi{v_1}{i_1} \otimes \cdots \otimes \mOppi{v_m}{i_m}$ as the
+% %matrix,
+% We can rewrite this representation more compactly as: $Z_{\sC} = {\mOppAll}\pinv \mO_\sV$.
+% By the delta-method \cite{vaart98asymptotic},
+% % DONE: need to shorten
+% %we have that the asymptotic variance of $M_\sV$ is:
+% %\begin{align*}
+% %  \sqrt{n}(\hat M_\sV - M_\sV) \convind \sN(0, \Sigma_\sV),
+% %\end{align*}
+% %where $\Sigma_\sV$ is the variance of the observations. 
+% we immediately get:
+% $\sqrt{n}(\hat Z_\sC - Z_\sC) \convind \sN(0, \mOppAlli \Sigma_\sV \mOppAllit)$.
+% \end{proof}
+
+% DONE: too brazen to say
+%Note that extending these results to finite sample bounds can be done
+  %via a straightforward application of perturbation bounds.

### 2014-05-17 doc/icml2014/background.tex
changes

@@
 where $[n] = \{ 1, \dots, n \}$.
-Let $\sX \eqdef [d]^L$ and $\sH \eqdef [k]^{M}$ be the joint domains of $\bx$ and $\bh$.
+Let $\sX \eqdef [d]^L$ and $\sH \eqdef [k]^{M}$ be the joint domains of $\bx$ and $\bh$, respectively.
 
@@
   and $\phi_\sC(\bx_\sC, \bh_\sC)$ is the local feature vector
-  which only depends on the observed ($\bx_\sC$) and hidden ($\bh_\sC$) variables in that clique $\sC$.
-  Also define $\sN(a) = \{ b \neq a : \exists \sC \supseteq \{ a, b \}$ be the neighbors of variable $a$.
+  which only depends on the observed ($\bx_\sC$) and hidden ($\bh_\sC$) variables in clique $\sC$.
+  Also define $\sN(a) = \{ b \neq a : \exists \sC \supseteq \{ a, b \} \}$ be the neighbors of variable $a$.
 
 For directed models $\sG$,
-  $p_\theta(\bx,\bh) = \prod_{a \in \bx \cup \bh} p_\theta(a \mid \Pa(a))$,
+  define $p_\theta(\bx,\bh) = \prod_{a \in \bx \cup \bh} p_\theta(a \mid \Pa(a))$,
   where $\Pa(a) \subseteq \bx \union \bh$ are the parents of a variable $a$.
-  The parameters $\theta$ are the conditional probability tables of all variables,
+  The parameters $\theta$ are the conditional probability tables of each variable,
   and the cliques are $\sG = \{ \{ a \} \union \Pa(a) : a \in \bx \union \bh \}$.
@@
 We are given $n$ i.i.d.~examples of the observed variables $\sD
-  = (\vx^{(1)}, \dots, \vx^{(n)})$ where each $\vx^{(i)} \sim p_{\theta^*}$ for
+  = (\vx^{(1)}, \dots, \vx^{(n)})$, where each $\vx^{(i)} \sim p_{\theta^*}$ for
   some true parameters $\theta^*$.
@@
 % Maximum likelihood
-The standard estimation procedure is maximum (marginal) likelihood,
+The standard estimation procedure is maximum likelihood:
   \begin{align*}
@@
 describe a method of moments approach for parameter estimation.
-First, a bit of notation is necessary.
+To do this, let's introduce some notation.
 
@@
 
-We use the notation $[\cdot]$ to indicate indexing; for example $M[i]$
-  is the $i$-th row of a matrix $M$ and $M[i,j]$ is the $i,j$-th element
+We use the notation $[\cdot]$ to indicate indexing; for example, $M[i]$
+  is the $i$-th row of a matrix $M$ and $M[i,j]$ is the $(i,j)$-th element
   of $M$.
@@
   = (i_1, \ldots, i_\ell)$, define the projection $T[\bi] = T[i_1, \ldots, i_\ell]$.
-Similarly, for a set $\sC$ of size $m$ and vector $v \in \Re^d$, we
-  will use the notation $v_\sC$ to get a vector in $\Re^m$ which selects
-  the appropriate elements of $v$.
+%Similarly, for a set $\sC$ of size $m$ and vector $v \in \Re^d$, we
+  %will use the notation $v_\sC$ to get a vector in $\Re^m$ which selects
+  %the appropriate elements of $v$.
 
@@
 \end{align*}
-Analogously, for a matrices $M_1 \in \Re^{d \times k}, \cdots,
-  M_\ell \in \Re^{d \times k}$, we define
+Analogously, for matrices $M_1 \in \Re^{d \times k}, \cdots,
+  M_\ell \in \Re^{d \times k}$:
 \begin{align*}
@@
   orders $1, 2, 3$ in $\Re^k, \Re^{k\times k}, \Re^{k \times k \times k}$.
+  Finally, we define \emph{conditional moments} $\mOpp{v}{i} \eqdef \Pr(x_v \mid h_i) \in \Re^{d \times k}$
+  for each $v \in [L]$ and $i \in [M]$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\paragraph{Approach overview}
+%
+%\algorithmref{overview} summarizes our approach.
+%
+%\begin{algorithm}
+%  \caption{\LearnMarginals}
+%  \label{algo:overview}
+%  \begin{algorithmic}
+%    \REQUIRE Graphical model $\sG$, data $\sD$
+%    \ENSURE Parameters $\theta$
+%    \For{each clique $\sN$}
+%    \FOR{each hidden node $h_i \in S$ with exclusive view $x_v$} 
+%      \STATE Apply $\TensorFactorize$ to learn conditional moments
+%      $\mOpp{v}{i} = \Pr(x_v \mid h_i)$.
+%    \ENDFOR
+%    \STATE Apply $\LearnClique$ to learn the marginals $\mH_\sC$.
+%  \end{algorithmic}
+%\end{algorithm}
 
@@
 paper, but that we feel are not central to our main ideas.
-The first one ensures that we will encounter all realizations of a hidden variable are possible:
+The first one ensures that all realizations of each hidden variable are possible:
 
@@
 \begin{assumption}[Faithful]
-  \label{prop:ci}  
+  \label{asm:faithful}  
   For any hidden variables $a,b,c \in \bh$
-  such that there is an active trail\footnote{See
+  such that an active trail\footnote{See
   \citet{koller2009probabilistic} for a definition.
-  We do condition on observed variables.} between $a$ and $b$ conditioned on $c$,
-  we have that $a$ and $b$ are dependent given $c$ (i.e., $a\not\perp b \given c$).
+  We do not condition on observed variables.} connects $a$ and $b$ conditioned on $c$,
+  we have that $a$ and $b$ are dependent given $c$. % (i.e., $a\not\perp b \given c$).
 \end{assumption}
 
-Finally, we describe a reduction of any graphical model to
-  a canonical form to make our arguments simpler.
-\begin{lemma}[Canonical form]
+Finally, we assume the graphical model is in a canonical form
+in which all observed variables are leaves:
+\begin{assumption}[Canonical form]
+  \label{asm:canonical}
+  For each observed variable $x_v$, there exists exactly one $\sC \in \sG$
+  such that $\sC = \{ x_v, x_{v'} \}$ for some other node $x_{v'}$.
+\end{assumption}
+The following lemma shows that this is not a real assumption (see the appendix for the proof):
+\begin{lemma}[Reduction to canonical form]
 \label{lem:reduction}
-Every directed (undirected) graphical model can be transformed into one in which
-  the observed variables are leaves with exactly one parent (neighbor). 
-There is a one-to-one correspondence between the parameters of the
-  transformed and original models.
+Every graphical model can be transformed into canonical
+form (via \figureref{reduction}).
+There is a one-to-one correspondence between the parameters of the transformed
+and original models.
 \end{lemma}
-\begin{proof}
-  \begin{figure}
-    \centering
-    \subimport{figures/}{reduction.tikz}
-    \caption{Reduction to canonical form.}
-    \label{fig:reduction}
-  \end{figure}
-
-  \providecommand{\hp}{\ensuremath{h_\text{\rm new}}}
-  Let $x_v$ be an observed variable with parents $\Pa(x_v)$ and children $\text{Ch}(x_v)$.
-  Consider the following transformation.
-  Replace $x_v$ with a new hidden variable \hp\ with the same
-  parents $\Pa(x_v)$ and children $\text{Ch}(x_v) \union \{x_{v_1}, x_{v_2}, x_{v_3}\}$,
-  where $x_{v_1},x_{v_2},x_{v_3}$ are three copies of $x_v$
-  (\figureref{reduction}).
-  Define $\Pr(\hp \mid \Pa(x_v)) = \Pr(x_v \mid \Pa(x_v))$ and
-  $\Pr(x_{v_j} \mid \hp) = I$.
-  Then, there is a one-to-one correspondence between every value of
-  $\hp$ and $x_v$. Consequently, for any clique $\sC \contains \hp$, the
-  parameters in the original graphical model can be obtained by
-  substituting $\hp$ with $x_v$.
-
-  We apply this procedure for all non-leaf observed variables.
-  This procedure also applies straightforwardly for undirected graphical
-  models, considering the neighbors $\sN(x_v)$ instead of its parents
-  and children.
-\end{proof}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Bottlenecks}
+\section{Bottlenecks}
+\label{sec:bottlenecks}
 
-In this section, we describe how tensor factorization
-can be used to reveal information about the hidden
-variables \citep{anandkumar13tensor}.  The key notion is that of a bottleneck:
+We start by trying to reveal some information about the hidden variables
+that will be used by subsequent sections.
+Specifically, we review how
+the tensor factorization method of \citet{anandkumar13tensor}
+can be used to recover the conditional moments
+$\mOpp{v}{i} = \Pr(x_v \mid h_i)$.
+The key notion is that of a bottleneck:
 
@@
 
-%Now define the following property:
-%\begin{property}[Uniformly bottlenecked]
-%  \label{prop:bottleneck}
-%  every hidden variable in $S$ is a bottleneck. 
-%  If every hidden clique $\sC \in \sG$ is uniformly bottlenecked, then we say
-%  that the graph $\sG$ is uniformly bottlenecked as well.
-%\end{property}
-%If every hidden variable is a bottleneck, then we can estimate $\mOpp{v}{i}$
-
 %Intuitively, we need to ensure that every marginal distribution
@@
 a mixture of that of other states.
-%PL: leave vague
-%asks that the conditional distributions are
-%  linearly independent; if this were not the case, the conditional
-%  distribution for a particular latent state could be imitated by the
-%  combination of several other latent states.
-
-%\begin{assumption}
-%\label{asm:full-rank}
-%The conditional moments $\mOpp{1}{1}$, $\mOpp{2}{1}$, $\mOpp{3}{1}$ have full column rank
-%  $k$, and $\pi\oft 1 \succ 0$.
-%\end{assumption}
 
@@
 \end{theorem}
-
-%\citet{anandkumar12moments} presented a consistent estimator
-  %using the method of moments to learn the parameters of a three-view
-  %mixture model (\figureref{three-view}), an important case of directed
-  %graphical models. %, that sidesteps the local optima problem altogether. 
-%In that model, we have three observed variables $x_1, x_2, x_3$
-  %that are independent conditioned on a hidden variable $h_1$.  
-%The parameters of this model are the prior probabilities $\pi\oft 1 \eqdef
-  %\Pr(h_1) \in \Re^k$ and the conditional moments $\mOpp{i}{1} \eqdef
-  %\Pr(x_i \given h_1) \in \Re^{d \times k}$ for $i \in \{1,2,3\}$.
 
@@
 $\sigma_k(\Pr(h_{i+1} \given h_{i})) = a_k$ for each $i = 1,
+DEFINE
 \cdots, t-1$, then $\sigma_k(\mOpp{v}{1}) = a_k^t \sigma_k(\mOpp{v}{t})$.}.

### 2014-05-17 doc/icml2014/directed.tex
changes

--- 2014-05-17 doc/icml2014/directed.tex
@@
-\section{Exclusive views}
-\label{sec:directed}
+\section{Recovering hidden marginals}
+\label{sec:exclusiveViews}
 
@@
   %graphical models.
-In this section, we will use the conditional moments $\mOpp{v}{i} \eqdef
-\Pr(x_v \mid h_i)$ recovered from the previous section
-to compute the marginal distribution of sets of hidden variables $Z_S \eqdef \Pr(\bh_S)$,
-taking us one step closer to the actual model parameters.
-For clarity, we derive our algorithm in the infinite data limit,
-although the actual algorithm works with estimates.
+Having recovered conditional moments $\mOpp{v}{i} \eqdef \Pr(x_v \mid h_i)$,
+we now seek to compute the marginal distribution of sets of hidden variables
+$Z_S \eqdef \Pr(\bh_S)$.
+%taking us one step closer to the actual model parameters.
+%For clarity, we derive our algorithm in the infinite data limit,
+%although the actual algorithm works with estimates.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{General algorithm}
-\label{sec:directedGeneral}
+\subsection{Exclusive views}
+\label{sec:general}
 
@@
 
-%If each $\mOpp{v_j}{i_j}$ has full column rank, then
-
 \begin{algorithm}
@@
   \begin{algorithmic}
-    \REQUIRE Hidden subset $S = \{ h_{i_1}, \dots, h_{i_m} \}$ with exclusive views $\sV = \{ x_{v_1}, \dots, x_{v_m} \}$ and $\{\mOpp{v_1}{i_i}\pinv\}$ % (\definitionref{exclusive-views})
+    \REQUIRE Hidden subset $S = \{ h_{i_1}, \dots, h_{i_m} \}$ with exclusive views $\sV = \{ x_{v_1}, \dots, x_{v_m} \}$
+    and $\{\mOpp{v_1}{i_i}\pinv\}$ % (\definitionref{exclusive-views})
     \ENSURE Marginal distribution $Z_S = \Pr(\bh_S)$.
@@
 in \algorithmref{directed}.
-
-\begin{algorithm}
-  \caption{\LearnMarginals}
-  \label{algo:directed}
-  \begin{algorithmic}
-    \REQUIRE Bidependent and bottlenecked subset of hidden nodes $S \subseteq \bh$
-    \ENSURE Marginals $Z_S = \Pr(\bh_S)$ % for every bidependent subset of hidden variables $S \in \sG$
-    \FOR{each hidden node $h_i \in S$ with exclusive view $x_v$} 
-      \STATE Apply $\TensorFactorize$ to learn conditional moments
-      $\mOpp{v}{i} = \Pr(x_v \mid h_i)$.
-    \ENDFOR
-    \STATE Apply $\LearnClique$ to learn the marginals $\mH_\sC$.
-  \end{algorithmic}
-\end{algorithm}
 

### 2014-05-17 doc/icml2014/intro.tex
major structural changes up to bidependent set
changes

@@
 balle12automata}.
-%\todo{make sure we are citing the right people}
 These methods exploit the linear algebraic properties of the model to
@@
   \caption{
- % \todo{need three steps: tensor factorization, optimize composite likelihood, recover model parameters;}
   Overview of our approach:
@@
 In this paper,
-we show that a much broader class of discrete directed and undirected graphical models can be consistently estimated:
+we show that a much broader class of discrete directed and undirected graphical
+models can be consistently estimated:
 specifically those in which \emph{each} hidden variable has three conditionally
-independent observed variables (``views'')---see \propertyref{bottleneck}.
+independent observed variables (``views'').
 Our key idea is to leverage the method of moments,
@@
 but as constraints on a likelihood-based objective.
+Notably, our method applies to latent undirected log-linear models with high treewidth.
 %For a broad class of directed and undirected graphical models
-
-%In this work, we exploit the spectral method to learn moment constraints
-  %on the observed variables and show how these constraints lead to consistent parameter estimates 
-  %for any graphical model with the property that there exist at least
-  %three ``views'' or conditionally independent observed variables for
-  %every hidden variable (\propertyref{bottleneck}).
 
@@
 % DONE: try to talk about the approach in terms of three steps
-The essence of our approach is illustrated in \figureref{approach}. 
+The essence of our approach is illustrated in \figureref{approach},
+which contains three steps.
 First, we identify three views for each hidden variable $h_i$ (for example,
@@
 \citet{anandkumar13tensor} to estimate the \emph{conditional
-moments} $\Pr(x_i^a \mid h_i)$ and $\Pr(x_i^b \mid h_i)$ for each $i$ (\sectionref{directed}).
+moments} $\Pr(x_i^a \mid h_i)$ and $\Pr(x_i^b \mid h_i)$ for each $i$ (\sectionref{bottlenecks}).
 Second, we optimize a \emph{composite marginal likelihood} to recover the marginals over
-each hidden clique (e.g., $\mathbb P(h_2, h_3, h_4)$).
+subsets of hidden nodes (e.g., $\mathbb P(h_2, h_3, h_4)$).
 Normally, such a marginal likelihood objective would be non-convex,
 but given the conditional moments, we obtain a convex objective,
-which can be globally optimized using EM (\sectionref{piecewise}).
+which can be globally optimized using EM
+(see Sections~\ref{sec:exclusiveViews} and \ref{sec:piecewise}).
 So far, our method has relied only on the conditional independence
@@
 requires some specialization.
-In the directed case, this is simple post-processing;
-in the undirected case, we need to solve another convex optimization problem (\sectionref{undirected}).
-%Given these factors, we show that there exists a set of observed
-  %variables for every clique such that the likelihood is convex in the
-  %parameters of that clique. 
-%This guarantees that we can recover the true marginal distributions
-  %using EM.
-  %finally, with some post-processing, we are able to
-  %consistently recover the true parameters of the model.
-%In \sectionref{directed}, we describe our algorithm in more detail and
-%show that our algorithm yields consistent parameter estimates
-%under non-degenerate rank assumptions.
-%describe the
-  %settings and technical assumptions in which this approach is guaranteed
-  %to work.
-% PL: hard to talk about $O$ and $T$ at this stage.
-%As a particular example, we are able to recover parameters for the
-  %model depicted in \figureref{approach} if $O$ and $T$ are full-rank.
-
-%Our approach also applies to undirected log-linear latent variable
-  %models, where we use the recovered marginals to make the learning
-  %problem convex (\sectionref{undirected}).
-%Finally, our approach gracefully extends to more general model families
-%  in which we are unable to recover the marginal distributions for all
-%  hidden cliques; in such scenarios, we use posterior regularization to constrain
-%  EM.  
-%\todo{Is this possible?:  We show that empirically, using such constraints alleviates local
-%optima issues.}
+In the directed case, this is simple normalization;
+in the undirected case, we need to solve another convex optimization problem
+(\sectionref{undirected}).

### 2014-05-17 doc/icml2014/background.tex
Updated intro figure, added exclusive views figure
some changes near beginning - bottlenecks, etc.

@@
 
+Finally, for clarity, we will derive our algorithms using exact moments of the true
+distribution $p_{\theta^*}$.  In practice, we would use moments estimated from data $\sD$.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 can be used to recover the conditional moments
-$\mOpp{v}{i} = \Pr(x_v \mid h_i)$.
+$\mOpp{v}{i} \eqdef \Pr(x_v \mid h_i)$.
 The key notion is that of a bottleneck:
@@
   \label{def:bottleneck}
-  A hidden variable $h_i$ is said to be a \emph{bottleneck} if there
-  exists three observed variables (views), $x_{v_1}, x_{v_2}, x_{v_3}$
+  A hidden variable $h_i$ is said to be a \emph{bottleneck} if (i) there
+  exists three observed variables (views), $x_{v_1}, x_{v_2}, x_{v_3}$,
   that are conditionally independent given $h_i$,
-  and each $\mOpp{v}{i} \eqdef \Pr(x_v \mid h_i) \in \Re^{d \times k}$ has full column rank $k$
+  and (ii) each $\mOpp{v}{i} \eqdef \Pr(x_v \mid h_i) \in \Re^{d \times k}$ has full column rank $k$
   for each $v \in \{v_1, v_2, v_3\}$.
@@
   if every $h \in S$ is a bottleneck.
-  We say that a graphical model $\sG$ is bottlenecked if all hidden variables
+  We say that a graphical model $\sG$ is bottlenecked if all its hidden variables
   are bottlenecks.
 \end{definition}
-For example, in \figureref{approach}, $x_1^a,x_1^b,x_2^a$ are views of the bottleneck $h_1$;
-$x_2^a,x_2^b,x_1^b$ are view of the bottleneck $h_2$.  Therefore, the clique
+For example, in \figureref{approach}, $x_1^a,x_1^b,x_2^a$ are views of the bottleneck $h_1$,
+and $x_2^a,x_2^b,x_1^b$ are views of the bottleneck $h_2$.  Therefore, the clique
 $\{h_1,h_2\}$ is a bottleneck.  Note that views are allowed to overlap.
@@
 The full rank assumption on the conditional moments $\mOpp{v}{i} = \Pr(x_v \mid h_i)$
-ensures that all states of $h_i$ ``behave differently''.
+ensures that all states of $h_i$ ``behave differently.''
 In particular, the conditional distribution of one state cannot be
@@
 
-\begin{figure}[t]
-  \label{fig:three-view}
-  \centering
-  \input{figures/three-view.tikz}
-  \caption{The canonical three-view mixture model can be estimated consistently
-  using tensor factorization \citep{anandkumar13tensor}.}
-\end{figure}
+%\begin{figure}[t]
+%  \label{fig:three-view}
+%  \centering
+%  \input{figures/three-view.tikz}
+%  \caption{The canonical three-view mixture model can be estimated consistently
+%  using tensor factorization \citep{anandkumar13tensor}.}
+%\end{figure}
 
@@
 \begin{theorem}[Tensor factorization]
-Let $h_i \in \bh$ be a hidden variable
-with views $x_{v_1},x_{v_2},x_{v_3}$.
+Let $h_i \in \bh$ be a bottleneck with views $x_{v_1},x_{v_2},x_{v_3}$.
 Then there exists an algorithm $\TensorFactorize$ that
-returns consistent estimates $\mOpphat{v}{i}$ for each $v \in \{v_1,v_2,v_3\}$.
+returns consistent estimates of $\mOpp{v}{i}$ for each $v \in \{v_1,v_2,v_3\}$.
 \end{theorem}
@@
 To simplify notation, consider the example in \figureref{three-view} where $h_1=1,v_1=1,v_2=2,v_3=3$.
-The key idea is that the observed moments $\mO_{12}, \mO_{23}, \mO_{13}$ and
-  $\mO_{123}$ admit a factorization in terms of parameters:
+The observed moments $\mO_{12}, \mO_{23}, \mO_{13}$ and $\mO_{123}$ can be factorized as follows:
 \begin{align*}
-  \mO_{ij} &= \sum_h {\pi \oft 1}[h] {\mOppt{i}{1}}[h] \otimes {\mOppt{j}{1}}[h] \\ %\quad \forall i,j \in \{1,2,3\} \\
+  \mO_{vv'} &= \sum_h {\pi \oft 1}[h] {\mOppt{v}{1}}[h] \otimes {\mOppt{v'}{1}}[h] \\ %\quad \forall i,j \in \{1,2,3\} \\
   \mO_{123} &= \sum_h {\pi \oft 1}[h] {\mOppt{1}{1}}[h] \otimes {\mOppt{2}{1}}[h] \otimes  {\mOppt{3}{1}}[h].
 \end{align*}
-The algorithm first computes a whitening matrix $W \in
+The $\TensorFactorize$ algorithm first computes a whitening matrix $W \in
   \Re^{d \times k}$ such that $W^\top \mO_{12} W = I_{k \times k}$,
   and uses $W$ to transform $\mO_{123}$ into a symmetric orthogonal tensor.  
-Then a robust tensor power method is applied to extract the eigenvectors
-of the whitened $\mO_{123}$; unwhitening yields the columns of $\mOpp{3}{1}$.
-The other model parameters can be recovered with a few additional matrix operations.
+Then a robust tensor power method is used to extract the eigenvectors
+of the whitened $\mO_{123}$; unwhitening yields the columns of $\mOpp{3}{1}$
+(up to permutation and scaling).
+The other conditional moments can be recovered similarly.
 %  we call $\TensorFactorize$, which transforms estimates of the observed moments 
@@
   %\assumptionref{non-degeneracy},
-  We have that $\mOpphat{v}{i}$
-  converges to $\mOpp{v}{i}$ at a rate of $n^{-\frac12}$ with a constant
-  that depends polynomially on the $k$-th singular value of
-  $\mOpp{v}{i}$.
+  The resulting estimate of %$\mOpphat{v}{i}$
+  $\mOpp{v}{i}$ based on $n$ data points converges at a rate of $n^{-\frac12}$ with a constant
+  that depends polynomially on $\sigma_k(\mOpp{v}{i})^{-1}$,
+  the inverse of the $k$-th largest singular value of $\mOpp{v}{i}$.
 % PL: we don't have a crisp theorem (because there's multiple paths), so leave it vague.
 %Recovering the marginals $Z_\sC$ from the conditional moments via $\LearnClique$ is a linear operation and thus also has polynomial sample complexity. 
-However, $\sigma_{k}(\mOpp{v}{i})$ can become extremely
+Note that $\sigma_{k}(\mOpp{v}{i})$ can become quite
 small if $h_i$ and $x_v$ are connected via many intermediate hidden
@@
 $\sigma_k(\Pr(h_{i+1} \given h_{i})) = a_k$ for each $i = 1,
-DEFINE
-\cdots, t-1$, then $\sigma_k(\mOpp{v}{1}) = a_k^t \sigma_k(\mOpp{v}{t})$.}.
+\cdots, t-1$, then $\sigma_k(\mOpp{v}{1}) = a_k^t \sigma_k(\mOpp{v}{t})$.}
 
@@
   in the form of the conditional moments $\mOpp{v}{i} = \Pr(x_v \mid h_i)$.
-  Note that $\TensorFactorize$ only looks at the conditional independence structure
+  Note that $\TensorFactorize$ only examines the conditional independence structure
   of the graphical model, not its parametrization.
-  If $v$ is the single parent of $i$ (e.g., $\Pr(x_1^a \mid h_1)$ in \figureref{approach}),
-  then this conditional moment is a parameter of the model.
-  If not (e.g., $\Pr(x_2^a \mid h_1)$), then this is not a parameter.
-  Furthermore, there are some parameters (e.g., $\Pr(h_4 \mid h_2, h_3)$) which are
-  are still not recovered.
+  If $i$ is the single parent of $v$ (e.g., $\Pr(x_1^a \mid h_1)$ in \figureref{approach}),
+  then this conditional moment is a parameter of the model,
+  but this is in general not the case (e.g., $\Pr(x_2^a \mid h_1)$).
+  Furthermore, there are other parameters (e.g., $\Pr(h_4 \mid h_2, h_3)$) which we do
+  not have a handle on yet.
   In general, there is a gap between the conditional moments
   and the model parameters,
-  something we will address in the next section.
+  something we will address in the next two sections.
 

### 2014-05-17 doc/icml2014/directed.tex
Updated intro figure, added exclusive views figure
some changes near beginning - bottlenecks, etc.
terninology change

@@
 \label{sec:exclusiveViews}
+
+\begin{figure}
+  \centering
+  \subfigure[Three-view mixture model] {
+    \includegraphics[width=0.45\columnwidth]{figures/three-view.pdf}
+    \label{fig:three-view}
+  }
+  \subfigure[Exclusive views]{
+    \includegraphics[width=0.45\columnwidth]{figures/exclusive-views.pdf}
+    \label{fig:exclusive-views}
+  }
+
+  \caption{(a) The canonical three-view mixture model can be estimated consistently
+  using tensor factorization \citep{anandkumar13tensor}.
+  (b) An example of exclusive views for a bi-dependent subset $S$.
+  }
+\end{figure}
 
@@
 %taking us one step closer to the actual model parameters.
-%For clarity, we derive our algorithm in the infinite data limit,
-%although the actual algorithm works with estimates.
 
@@
 \label{sec:general}
-
-  %Section~\ref{sec:directedGeneral} will describe the algorithm in full generality.
-%For ease of exposition we make the following simplifications to our presentation:
-%(i) we describe our algorithm entirely in the context of directed
-  %graphical models, though it generalizes to undirected models;
-%(iii) we present the algorithm solely in terms of linear algebraic operations;
-  %\sectionref{piecewise} provides a statistically
-  %more efficient estimator using composite likelihood.
 
@@
 
-\algorithmref{learnclique} summarizes the procedure, \LearnClique.
+\algorithmref{learnclique} summarizes the procedure, \LearnMarginals.
 Given $Z_S$, the conditional probability tables for $S$ can easily be
@@
 \begin{algorithm}
-  \caption{\LearnClique~(pseudoinverse version)}
-  \label{algo:learnclique}
+  \caption{\LearnMarginals~(pseudoinverse version)}
+  \label{algo:learnMarginals}
   \begin{algorithmic}
     \REQUIRE Hidden subset $S = \{ h_{i_1}, \dots, h_{i_m} \}$ with exclusive views $\sV = \{ x_{v_1}, \dots, x_{v_m} \}$
-    and $\{\mOpp{v_1}{i_i}\pinv\}$ % (\definitionref{exclusive-views})
-    \ENSURE Marginal distribution $Z_S = \Pr(\bh_S)$.
+    and conditional moments $\mOpp{v_j}{i_j} = \Pr(x_{v_i} \mid h_{i_j})$
+    \ENSURE Marginals $Z_S = \Pr(\bh_S)$.
       %\STATE Identify exclusive views $\Sx{\sV} = \{x_{v_1}, \dots, x_{v_m}\}$.

### 2014-05-17 doc/icml2014/undirected.tex
terninology change

@@
   \figureref{examples-mrf},
-  Therefore, we can adapt our algorithm ($\LearnClique$) to estimate all marginals of the form $Z_{a,\sN(a)} \eqdef \Pr(a,\sN(a))$,
+  Therefore, we can adapt our algorithm ($\LearnMarginals$) to estimate all marginals of the form $Z_{a,\sN(a)} \eqdef \Pr(a,\sN(a))$,
   from which we can compute the features $\mu_{a,\sN(a)}$,

### 2014-05-17 doc/icml2014/discussion.tex
Fixed refeerenes to bottleneck property and exclusive views

@@
 and others.
-We also introduce the idea of \emph{exclusive views} (\propertyref{exclusive-views}),
+We also introduce the idea of \emph{exclusive views} (\definitionref{exclusive-views}),
 the key property that allows us to learn the correlation structure between hidden variables.
@@
 In a pre-processing step, they ``subtract'' out the correlations and
-  recover a graph that satisfies the bottleneck property
-  (\propertyref{bottleneck}).
+  recover a graph that is bottlenecked.
 % TOO BOLD
@@
   operator representations for latent tree graphical models, like the
-  one in \figureref{examples-tree}, assuming the bottleneck property (\propertyref{bottleneck}). 
+  one in \figureref{examples-tree}, assuming the graph is bottlenecked. 
 Their approach is similar to our first step of learning conditional moments,
@@
 \citet{parikh12spectral} extend this approach to general graphical
-  models in which \propertyref{bottleneck} holds, but consider the latent
+  models which are bottlenecked, but consider the latent
   junction tree representation. 

### 2014-05-17 doc/icml2014/undirected.tex
Fixed refeerenes to bottleneck property and exclusive views

--- 2014-05-17 doc/icml2014/undirected.tex
@@
 In \sectionref{directed}, we showed that $\LearnMarginals$ is a consistent
-estimator for $Z_\sC$ for any graphical model satisfying the uniformly bottlenecked property (\propertyref{bottleneck}). 
+estimator for $Z_\sC$ for any bottlenecked clique in the graphical model. 
 Consequently, the output of \algorithmref{undirected}
@@
 \begin{algorithm}
-  \caption{\LearnLogLinear}
+  \caption{\LearnLogLinear (for undirected models)}
   \label{algo:undirected}
   \begin{algorithmic}
-    \REQUIRE A graphical model $\sG$ satisfying \propertyref{bottleneck}, data $\sD$
+    \REQUIRE Conditional moments $\mOpp{v}{i}$ and clique marginals $Z_\sC$.
     \ENSURE Parameters $\theta$
-      \STATE Apply $\LearnMarginals$ to $\sG$ to learn clique marginals $Z_\sC$.
-      \STATE Compute expected features $\{\mu_\sC\}_{\sC \in \sG}$ (\equationref{logLinearFeatures}) and optimize \equationref{logLinearSupervised}.
+      \STATE Compute expected features $\{\mu_\sC\}_{\sC \in \sG}$ (\equationref{logLinearFeatures}).
+      \STATE Optimize \equationref{logLinearSupervised}.
     \end{algorithmic}
@@
 
-In general, it might be case that we only have exclusive views (\propertyref{exclusive-views}) for a subset of the cliques.
+In general, it might be case that we only have exclusive views for a subset of the cliques.
 In this case, we can estimate the expected features $\mu_\sC$ for those cliques

### 2014-05-17 doc/icml2014/intro.tex

@@
   Overview of our approach:
-  we first use tensor factorization to learn the \emph{conditional moments}
-  for each hidden variable.
-  Second, we optimize a composite marginal likelihood to recover the \emph{hidden clique marginals},
-  from which we solve for the \emph{model parameters}.
+  (i) we use tensor factorization to learn the \emph{conditional moments}
+  for each hidden variable;
+  (ii) we optimize a composite likelihood to recover the hidden \emph{marginals};
+  and (iii) we optimize another likelihood objective to the \emph{model parameters}.
+  Both likelihood objectives are convex.
   }

### 2014-05-17 doc/icml2014/directed.tex
Updated figures
Updated references to learnclique, directed in the exclusive-views section

@@
   \centering
-  \subfigure[Three-view mixture model] {
-    \includegraphics[width=0.45\columnwidth]{figures/three-view.pdf}
+  \subfigure[Bottleneck] {
+    \includegraphics[width=0.35\columnwidth]{figures/three-view.pdf}
     \label{fig:three-view}
@@
 
-  \caption{(a) The canonical three-view mixture model can be estimated consistently
+  \caption{(a) A bottleneck can be estimated consistently
   using tensor factorization \citep{anandkumar13tensor}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph{Estimating hidden clique marginals}
+\paragraph{Estimating hidden subset marginals}
 
@@
 
-\algorithmref{learnclique} summarizes the procedure, \LearnMarginals.
+\algorithmref{learnMarginals} summarizes the procedure, \LearnMarginals.
 Given $Z_S$, the conditional probability tables for $S$ can easily be
@@
 If $S \subseteq \bx$ is a subset of hidden variables with the exclusive views property,
-then \algorithmref{learnclique} recovers the marginals $Z_S = \Pr(\bh_S)$.
+then \algorithmref{learnMarginals} recovers the marginals $Z_S = \Pr(\bh_S)$.
 \end{theorem}
@@
 Note that all cliques are bidependent, but bidependent sets can have more conditional independences.
-this will be important for pseudolikliehood in \sectionref{pseudolikelihood}.
+This will be important for pseudolikliehood in \sectionref{pseudolikelihood}.
 
@@
 %\definitionref{exclusive-views}, only two is required for condition (i).
-With this lemma in place, we present our full algorithm, $\LearnMarginals$,
-in \algorithmref{directed}.
+% With this lemma in place, we present our full algorithm, $\LearnMarginals$,
+% in \algorithmref{learnMarginals}.
 
@@
   hidden variable $h_i$.
-\algorithmref{learnclique} and \algorithmref{directed} still apply and allow
-  us to recover clique marginals $Z_\sC$.
+\algorithmref{learnMarginals} % and \algorithmref{directed} 
+still applies and allow us to recover clique marginals $Z_\sC$.
 %  \todo{make this a footnote if running out of space}

### 2014-05-17 doc/icml2014/directed.tex
simplify directed grid example

@@
 \section{Recovering hidden marginals}
-\label{sec:exclusiveViews}
+\label{sec:hiddenMarginals}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Example: directed grid model}
-\label{sec:directedExample}
+%\subsection{Example: directed grid model}
+%\label{sec:directedExample}
+
+\paragraph{Example}
+
+%The model has eight observed variables $x^a_1, x^b_1 \cdots, x^a_4, x^b_4$ and four
+  %hidden variables $h_1, \ldots, h_4$.
+%The parameters of this model are the conditional probability tables
+%$\pi \eqdef \Pr(h_1) \in \Re^k, T \eqdef \Pr(h_2 | h_1) = \Pr(h_3 | h_1) \in \Re^{k \times k},
+%V \eqdef \Pr(h_4 | h_2, h_3) \in \Re^{k \times k \times k}$ and $O \eqdef \Pr(x^a_i | h_i)
+%=  \Pr(x^b_i | h_i) \in \Re^{d \times k}$.
+
+%\paragraph{Estimating $O$}
+%Since the observed variables $x^a_1, x^b_1, x^a_2$ are
+%  conditionally independent given $h_1$, we can $\TensorFactorize$ from
+%  \sectionref{setup} to recover $O$.
+
+%\paragraph{Estimating $\pi$}
+%The moments of $x^a_1$, $\mO_1 \eqdef \Pr(x^a_1)$ are directly related to
+%  $\pi$ by a linear system: $\mO_1 = O \pi$. 
+%If $O$ has full column rank, we can recover $\pi$ by taking the pseudoinverse: $\pi = O\pinv  \mO_1$.
+
+%\paragraph{Estimating $T$}
 
 To gain some intuition, consider the directed grid model from \figureref{approach}.
-%The model has eight observed variables $x^a_1, x^b_1 \cdots, x^a_4, x^b_4$ and four
-  %hidden variables $h_1, \ldots, h_4$.
-The parameters of this model are the conditional probability tables
-$\pi \eqdef \Pr(h_1) \in \Re^k, T \eqdef \Pr(h_2 | h_1) = \Pr(h_3 | h_1) \in \Re^{k \times k},
-V \eqdef \Pr(h_4 | h_2, h_3) \in \Re^{k \times k \times k}$ and $O \eqdef \Pr(x^a_i | h_i)
-=  \Pr(x^b_i | h_i) \in \Re^{d \times k}$.
+We can express the observed marginals $\mO_{12} \eqdef \Pr(x^a_1, x^a_2) \in \Re^{d \times d}$
+as a linear function of the hidden marginals $\mH_{12} \eqdef \Pr(h_1, h_2) \in \Re^{k \times k}$,
+where the linear coefficients are based on the conditional moments
+$\mOpp{1}{1},\mOpp{2}{2} \in \Re^{d \times k}$:
+%down the moments of $x^a_1, x^a_2$,
+\begin{align*}
+\mO_{12} = \mOpp{1}{1} \mH_{12} \mOppt{2}{2}.
+\end{align*}
+We can then solve for $\mH_{12}$ by matrix inversion:
+\begin{align*}
+\mH_{12} = \mOppi{1}{1} \mO_{12} \mOppit{2}{2}.
+\end{align*}
+%$T$ can be recovered from the $\mH_{12}$ by renormalizing the columns.
 
-\paragraph{Estimating $O$}
-Since the observed variables $x^a_1, x^b_1, x^a_2$ are
-  conditionally independent given $h_1$, we can $\TensorFactorize$ from
-  \sectionref{setup} to recover $O$.
-
-\paragraph{Estimating $\pi$}
-The moments of $x^a_1$, $\mO_1 \eqdef \Pr(x^a_1)$ are directly related to
-  $\pi$ by a linear system: $\mO_1 = O \pi$. 
-If $O$ has full column rank, we can recover $\pi$ by taking the pseudoinverse: $\pi = O\pinv  \mO_1$.
-
-\paragraph{Estimating $T$}
-Similarly, we can write down the moments of $x^a_1, x^a_2$, $\mO_{12}
-  \eqdef \Pr(x^a_1, x^a_2)$, in terms of the hidden marginals $\mH_{12}
-  \eqdef \Pr(h_1, h_2)$ and solve for $\mH_{12}$:
-\begin{align*}
-\mO_{12} = O \mH_{12} O^\top \quad\Rightarrow\quad
-  \mH_{12} = O\pinv  \mO_{12} O\pinvt .
-\end{align*}
-$T$ can be recovered from the $\mH_{12}$ by renormalizing the columns.
-
-\paragraph{Estimating $V$}
-Finally, we can estimate $V$ by renormalizing the hidden marginals
-$\mH_{234} \eqdef \Pr(h_2, h_3, h_4)$ from the third-order moments
-$\mO_{234} \eqdef \Pr(x^a_2, x^a_3, x^a_4)$:
-\begin{align*}
-  \mO_{234} = \mH_{234}(O, O, O) \quad\Rightarrow\quad
-  \mH_{234} = \mO_{234}(O\pinv , O\pinv , O\pinv ).
-\end{align*}
+%\paragraph{Estimating $V$}
+%Finally, we can estimate $V$ by renormalizing the hidden marginals
+%$\mH_{234} \eqdef \Pr(h_2, h_3, h_4)$ from the third-order moments
+%$\mO_{234} \eqdef \Pr(x^a_2, x^a_3, x^a_4)$:
+%\begin{align*}
+%  \mO_{234} = \mH_{234}(O, O, O) \quad\Rightarrow\quad
+%  \mH_{234} = \mO_{234}(O\pinv , O\pinv , O\pinv ).
+%\end{align*}
 
@@
 
-The intuitions of the directed grid generalize readily to general graphs.
-The key property required for our algorithm to work is as follows:
+%The intuitions of the directed grid generalize readily to general graphs.
+%The key property required for our algorithm to work is as follows:
+For which subsets of hidden nodes can we recover the marginals?  The following
+definition offers an characterization:
 \begin{definition}[Exclusive views]
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph{Estimating hidden subset marginals}
+\paragraph{Estimating hidden marginals}
 

### 2014-05-17 doc/icml2014/intro.tex
simplify directed grid example

@@
 which can be globally optimized using EM
-(see Sections~\ref{sec:exclusiveViews} and \ref{sec:piecewise}).
+(see Sections~\ref{sec:hiddenMarginals} and \ref{sec:piecewise}).
 So far, our method has relied only on the conditional independence

### 2014-05-17 doc/icml2014/undirected.tex
Updated figures
Fixed minor typos and stuff in undirected section

--- 2014-05-17 doc/icml2014/undirected.tex
@@
 % Define model
-As before, let $\sG$ be a graphical model with observed variables $X$ and hidden variables $H$
+As before, let $\sG$ be a graphical model with observed variables $\bx$ and hidden variables $\bh$
 (see \figureref{examples-mrf} for an example).
@@
 We consider log-linear models of the following form:
-\begin{align}
+\begin{align*}
 p_\theta(\vx, \vh) &= \exp\left( \theta^\top \phi(\vx,\vh) - A(\theta) \right),
-\end{align}
+\end{align*}
 where $A(\theta) = \log \sum_{\vx, \vh}  \exp( \theta^\top \phi(\vx,\vh) )$ is the log-partition function.
@@
 Of course, the log-likelihood is non-concave:
-\begin{align}
+\begin{align*}
   L_\text{unsup}(\theta) \eqdef \E_{\vx \sim p^*}[\log \sum_{\vh \in \sH} p_\theta(\vx,\vh)],
-\end{align}
+\end{align*}
 where we are assuming for simplicity that we have infinite data drawn from the true distribution $p^*$.
@@
 On the other hand, 
-if we were able to observe $\vh$, we would optimize \emph{supervised} likelihood, which is concave:
+if we were able to observe $\vh$, we would optimize the \emph{supervised} likelihood, which is concave:
 \begin{align}
+L_\text{sup}(\theta) &\eqdef \E_{(\vx,\vh) \sim p^*}[\log p_\theta(\vx,\vh)] \nonumber \\
+                     &= \theta^\top \left(\sum_{\sC \in \sG} \E[\phi(\vx_\sC,\vh_\sC)]\right) - A(\theta).
 \label{eqn:logLinearSupervised}
-L_\text{sup}(\theta) &\eqdef \E_{(\vx,\vh) \sim p^*}[\log p_\theta(\vx,\vh)] \\
-                     &= \theta^\top \left(\sum_{\sC \in \sG} \E[\phi(\vx_\sC,\vh_\sC)]\right) - A(\theta).
 \end{align}
@@
 \algorithmref{undirected} summarizes our approach.
-In \sectionref{directed}, we showed that $\LearnMarginals$ is a consistent
+In \sectionref{exclusiveViews}, we showed that $\LearnMarginals$ is a consistent
 estimator for $Z_\sC$ for any bottlenecked clique in the graphical model. 
-Consequently, the output of \algorithmref{undirected}
-is a consistent estimator.\footnote{Of course, the log-linear model must be identifiable
-(it is violated, for example, if two components of $\phi(\vx_\sC, \vh_\sC)$ were identical).
+Consequently, \algorithmref{undirected}
+is also a consistent estimator.\footnote{Of course, the log-linear model must be identifiable.
+Identifability is violated, for example, if two components of $\phi(\vx_\sC, \vh_\sC)$ were identical.
 In general, we can identify the parameters up to the kernel of $\nabla^2 A(\theta)$.}
@@
 \begin{algorithm}
-  \caption{\LearnLogLinear (for undirected models)}
+  \caption{\LearnParameters (for undirected models)}
   \label{algo:undirected}
@@
 Let $\phi_{a,\sN(a)}(a, \sN(a)) = \sum_{\sC \ni a} \phi_\sC(\bx_\sC,\bh_\sC)$
-denote the sum over cliques $\sC$ that contain $a$; note that only depends on $a$ and its neighbors $\sN(a)$.
+denote the sum over cliques $\sC$ that contain $a$; note that $\phi_{a, \sN(a)}$ only depends on $a$ and its neighbors $\sN(a)$.
 We can write each conditional log-likelihood as:
-\begin{align}
+\begin{align*}
 p_\theta(a \mid \sN(a)) = \exp(\theta^\top\phi_{a,\sN(a)}(a, \sN(a)) - A_a(\theta; \sN(a))),
-\end{align}
+\end{align*}
 where the conditional log-partition function $A_a(\theta; \sN(a)) = \log \sum_{a'} \exp(\theta^\top\phi_{a,\sN(a)}(a', \sN(a)))$
-involves summing only over values to a single variable.
+involves marginalizing only over a single variable.
 
@@
 %  \exp(\theta^\top\phi_{\sN(a)}(a'))$.
-Note that $\{a\}\cup\sN(a)$ in general do not form a clique.
+Note that $\{a\}\cup\sN(a)$ in general does not form a clique.
 Fortunately, \lemmaref{bottleneck-views} holds for all biconnected subgraphs.
-We can break $\{a\}\cup\sN(a)$ into a collection of 
+We can break $\{a\}\cup\sN(a)$ into a collection of biconnected subgraphs 
   % PL: this is not necessary
@@
 %However, when we do have exclusive views for each neighborhood, as in
-  \figureref{examples-mrf},
-  Therefore, we can adapt our algorithm ($\LearnClique$) to estimate all marginals of the form $Z_{a,\sN(a)} \eqdef \Pr(a,\sN(a))$,
+ (e.g. \figureref{examples-mrf}).
+  Therefore, we can adapt our algorithm ($\LearnMarginals$) to estimate all marginals of the form $Z_{a,\sN(a)} \eqdef \Pr(a,\sN(a))$,
   from which we can compute the features $\mu_{a,\sN(a)}$,

### 2014-05-17 doc/icml2014/background.tex
Updated bottlenecked definition
terninology change
finished going through recovering hidden marginals

@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\paragraph{Approach overview}
-%
-%\algorithmref{overview} summarizes our approach.
-%
-%\begin{algorithm}
-%  \caption{\LearnMarginals}
-%  \label{algo:overview}
-%  \begin{algorithmic}
-%    \REQUIRE Graphical model $\sG$, data $\sD$
-%    \ENSURE Parameters $\theta$
-%    \For{each clique $\sN$}
-%    \FOR{each hidden node $h_i \in S$ with exclusive view $x_v$} 
-%      \STATE Apply $\TensorFactorize$ to learn conditional moments
-%      $\mOpp{v}{i} = \Pr(x_v \mid h_i)$.
-%    \ENDFOR
-%    \STATE Apply $\LearnClique$ to learn the marginals $\mH_\sC$.
-%  \end{algorithmic}
-%\end{algorithm}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   %Let $\sV_{h_i}$ denote any set of views for $h_i$.
-  We say that a subset of hidden variables $S \subseteq \bh$ is a bottleneck
+  We say that a subset of hidden variables $S \subseteq \bh$ is bottlenecked
   if every $h \in S$ is a bottleneck.
@@
 and $x_2^a,x_2^b,x_1^b$ are views of the bottleneck $h_2$.  Therefore, the clique
-$\{h_1,h_2\}$ is a bottleneck.  Note that views are allowed to overlap.
+$\{h_1,h_2\}$ is bottlenecked.  Note that views are allowed to overlap.
 
@@
 of the whitened $\mO_{123}$; unwhitening yields the columns of $\mOpp{3}{1}$
-(up to permutation and scaling).
+(up to permutation).
 The other conditional moments can be recovered similarly.
@@
 % PL: we don't have a crisp theorem (because there's multiple paths), so leave it vague.
-%Recovering the marginals $Z_\sC$ from the conditional moments via $\LearnClique$ is a linear operation and thus also has polynomial sample complexity. 
 Note that $\sigma_{k}(\mOpp{v}{i})$ can become quite

### 2014-05-17 doc/icml2014/directed.tex
Fixed exclusive views
finished going through recovering hidden marginals

@@
 
-  \caption{(a) A bottleneck can be estimated consistently
+  \caption{
+    \label{fig:bottleneckExclusiveViews}
+  (a) A bottleneck can be estimated consistently
   using tensor factorization \citep{anandkumar13tensor}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph{Structural properties}
+\paragraph{Relationship to bottlenecks}
 
-We have shown how to estimate the parameters of any clique that possesses the exclusive
-  views property (\definitionref{exclusive-views}).
-  But in a general graph, which cliques have this property?
-  To provide intuition, we will provide more interpretable sufficient conditions
-  and examples.
+%We have shown how to estimate the parameters of any clique that possesses the exclusive
+  %views property (\definitionref{exclusive-views}).
+  %But in a general graph, which cliques have this property?
+  %To provide intuition, we will provide more interpretable sufficient conditions
+  %and examples.
 
-% Define bidependent
-First, define the notion of a bidependent set (in analogy with biconnected component),
-in which conditionining on one variable does not break the set apart:
+The bottleneck property allow recovery of conditional moments,
+and the exclusive views property allow recovery of hidden marginals.
+But we will now show that the latter property is in fact implied by the former
+property for special sets of hidden variables, which we call bidependent
+sets (in analogy with biconnected components),
+in which conditioning on one variable does not break the set apart:
 \begin{definition}[Bidependent set]
-Given a graphical model $\sG$, we say that a subset of nodes $S$ is \emph{bidependent} if
+We say that a subset of nodes $S$ is \emph{bidependent} if
 conditioned on any $a \in S$, there is an active trail between any other two nodes $b,c \in S$.
 \end{definition}
-Note that all cliques are bidependent, but bidependent sets can have more conditional independences.
-This will be important for pseudolikliehood in \sectionref{pseudolikelihood}.
+Note that all cliques are bidependent, but bidependent sets can have more conditional independences
+(e.g., $\{h_1,h_2,h_3\}$ in \figureref{bottleneckExclusiveViews}(b)).
+This will be important for pseudolikelihood in \sectionref{pseudolikelihood}.
 
 % bidependent, bottleneck => exclusive views
-Bidependent sets are significant because they guarantee exclusive views if they are bottlenecks:
+Bidependent sets are significant because they guarantee exclusive views if they
+are bottlenecked:
 %If a subset of hidden nodes $S$ is a bottleneck, it is clear that (ii)
@@
   Let $S \subseteq \bh$ be a bidependent subset of hidden variables.
-  If $S$ is a bottleneck, then $S$ has the exclusive views property.
+  If $S$ is bottlenecked, then $S$ has the exclusive views property.
 \end{lemma}
@@
 there is an active trail between $h_1$ and $h_2$ because $S$ is biconnected.
-Thus, there is an active trail $x_1 - h_1 - h_2 - x_2$ conditioned on $h_0$.
+This means there is also an active trail $x_1 - h_1 - h_2 - x_2$ conditioned on $h_0$.
 Since the graphical model is faithful by assumption, we have $x_1 \not\perp x_2 \mid h_0$,
@@
 \paragraph{Remarks.}
-Note that even having two independent views for each $h_i \in S$ is sufficient for condition (i)
+Note that having only two independent views for each $h_i \in S$ is sufficient for condition (i)
   of the exclusive views property, while three is needed for condition (ii).
 The bottleneck property (\definitionref{bottleneck} can also be relaxed if some cliques
-  share parameters.
-For example, in the directed grid model, we can recover the conditional moments $O$ from
-  $x^a_1, x^b_1$ and $x^a_2$ with $h_1$ as the bottleneck.
-  Therefore, $h_2, h_3$ and $h_4$
-  need not be bottlenecks, and we can omit the observations $x^b_2, x^b_3$ and $x^b_4$.
+  share parameters (see examples below).
+%For example, in the directed grid model, we can recover the conditional moments $O$ from
+  %$x^a_1, x^b_1$ and $x^a_2$ with $h_1$ as the bottleneck.
+  %Therefore, $h_2, h_3$ and $h_4$
+  %need not be bottlenecks, and we can omit the observations $x^b_2, x^b_3$ and $x^b_4$.
 
-Our method also extends directly to case in which the observed variables
-  are real-valued, while the hidden variables remain discrete. 
-In this setting, the tensor factorization method of
-  \citet{anandkumar13tensor} recovers the expected conditional means,
-  $\mu_{vi} \eqdef \E(x_v | h_i) \in \Re^{d \times k}$ for each observed variable $x_v$ and
-  hidden variable $h_i$.
-\algorithmref{learnMarginals} % and \algorithmref{directed} 
-still applies and allow us to recover clique marginals $Z_\sC$.
+Our method extends naturally to the case in which the observed variables
+  are real-valued $x_v \in \Re^d$, as long as the hidden variables remain discrete. 
+In this setting, the conditional moments
+  $\mOpp{v}{i} \eqdef \E(x_v | h_i) \in \Re^{d \times k}$
+  would no longer be distributions but general rank $k$ matrices.
+  %for each observed variable $x_v$ and
+  %hidden variable $h_i$.
+%\algorithmref{learnMarginals} % and \algorithmref{directed} 
+%still applies and allow us to recover clique marginals $Z_\sC$.
 %  \todo{make this a footnote if running out of space}
@@
   }
-  \caption{(a) and (b): graphs that satisfy the exclusive views property; (c) does not.}
+  \caption{(a) and (b): graphical models that satisfy the exclusive views property;
+  (c) a graphical model that does not.}
   \label{fig:examples}
@@
 
-In the HMM (\figureref{examples-hmm}), the parameters
-are $O \eqdef \Pr(x_i|h_i)$ and $T \eqdef \Pr(h_{i+1} | h_i)$
-for all $i$.
-While the first and last hidden variables $h_1, h_M$ in the
-  sequence are not bottlenecks, they still have exclusive views ($x_1$ and
-  $x_M$, respectively)
-  due to parameter sharing.
+In the HMM (\figureref{examples-hmm}),
+$h_2$ is a bottleneck, so we can recover $O \eqdef \Pr(x_2 \mid h_2)$.
+%the parameters
+%are $O \eqdef \Pr(x_i|h_i)$ and $T \eqdef \Pr(h_{i+1} | h_i)$
+%for all $i$.
+While the first hidden variable $h_1$
+is not a bottleneck,
+it still has an exclusive view $x_1$ with respect to the clique $\{h_1,h_2\}$,
+assuming parameter sharing across emissions ($\Pr(x_1 \mid h_1) = O$).
 
@@
 
-In the latent tree model (\figureref{examples-tree}), the parameters
-are $\pi \eqdef \Pr(h_i)$, $T \eqdef \Pr(h_i | h_1)$ for $i \in \{2,3,4\}$,
-and $O \eqdef \Pr(x^a_i | h_i) = \Pr(x^b_i | h_i)$ for $i \in \{2,3,4\}$.
-Note that while $h_1$ is not directly connected to an observed variable,
-  it is still a bottleneck, with views $x^a_2, x^a_3, x^a_4$.
-We can recover $T$ from the clique $\{h_1, h_2\}$ by using views $x^a_2$
-  (exclusive to $h_2$) and $x^a_3$ (exclusive to $h_1$).
+In the latent tree model (\figureref{examples-tree}),
+%the parameters
+%are $\pi \eqdef \Pr(h_i)$, $T \eqdef \Pr(h_i | h_1)$ for $i \in \{2,3,4\}$,
+%and $O \eqdef \Pr(x^a_i | h_i) = \Pr(x^b_i | h_i)$ for $i \in \{2,3,4\}$.
+$h_1$ is not directly connected to an observed variable,
+but it is still a bottleneck, with views $x^a_2, x^a_3, x^a_4$, for example.
+The clique $\{h_1,h_2\}$ has exclusive views $\{x_2^a,x_3^a\}$.
+%We can recover $T$ from the clique $\{h_1, h_2\}$ by using views $x^a_2$
+  %(exclusive to $h_2$) and $x^a_3$ (exclusive to $h_1$).
 
-\paragraph{Non-examples}
+\paragraph{Non-example}
 \label{sec:non-example}
 
-There are certainly models which are identifiable but do not have exclusive views.
-For example, \figureref{examples-noisy-or} shows
-  a binary-valued noisy-or network which can be
-  learned by the algorithm of \citet{halpern13noisyor},
-  but does not satisfy the exclusive views property.
+%There are certainly models which are identifiable but do not have exclusive views.
+In \figureref{examples-noisy-or},
+  $h_1$ does not have exclusive views.
+  Without parameter sharing, the techniques in this paper
+  are insufficient.
+In the special case where the graphical model represents a binary-valued noisy-or network,
+  we can use the algorithm of \citet{halpern13noisyor},
+  which first learns $h_2$ and subtracts off its influence,
+  thereby making $h_1$ a bottleneck.

### 2014-05-17 doc/icml2014/background.tex
coarse pass over composite likelihood

@@
 The standard estimation procedure is maximum likelihood:
-  \begin{align*}
+  \begin{align}
+    \label{eqn:unsup}
     L_\text{unsup}(\theta) &\eqdef \sum_{\vx \in \sD} \log p_\theta(\vx)
         =  \sum_{\vx \in \sD} \log \sum_{\vh \in \sH} p_\theta(\vx,\vh).
-  \end{align*}
+  \end{align}
   Maximum likelihood is statistically efficient,

### 2014-05-17 doc/icml2014/directed.tex
coarse pass over composite likelihood

@@
   }
+  \hspace{2em}
   \subfigure[Exclusive views]{
-    \includegraphics[width=0.45\columnwidth]{figures/exclusive-views.pdf}
+    \includegraphics[width=0.35\columnwidth]{figures/exclusive-views.pdf}
     \label{fig:exclusive-views}
@@
   using tensor factorization \citep{anandkumar13tensor}.
-  (b) An example of exclusive views for a bi-dependent subset $S$.
+  (b) An example of exclusive views for a bidependent subset $S$.
   }
@@
 \begin{algorithm}
-  \caption{\LearnMarginals~(pseudoinverse version)}
+  \caption{\LearnMarginals~(pseudoinverse)}
   \label{algo:learnMarginals}
@@
     \REQUIRE Hidden subset $S = \{ h_{i_1}, \dots, h_{i_m} \}$ with exclusive views $\sV = \{ x_{v_1}, \dots, x_{v_m} \}$
-    and conditional moments $\mOpp{v_j}{i_j} = \Pr(x_{v_i} \mid h_{i_j})$
+    and conditional moments $\mOpp{v_j}{i_j} = \Pr(x_{v_i} \mid h_{i_j})$.
     \ENSURE Marginals $Z_S = \Pr(\bh_S)$.

### 2014-05-17 doc/icml2014/background.tex
Minor fixes to background. Added statement about hidden variable relabelling in exclusive views section

@@
   which only depends on the observed ($\bx_\sC$) and hidden ($\bh_\sC$) variables in clique $\sC$.
-  Also define $\sN(a) = \{ b \neq a : \exists \sC \supseteq \{ a, b \} \}$ be the neighbors of variable $a$.
+  Also define $\sN(a) = \{ b \neq a : \exists \sC \supseteq \{ a, b \} \}$ to be the neighbors of variable $a$.
 
@@
   For each observed variable $x_v$, there exists exactly one $\sC \in \sG$
-  such that $\sC = \{ x_v, x_{v'} \}$ for some other node $x_{v'}$.
+  such that $\sC = \{ x_v, h_{i} \}$ for some hidden node $h_{i}$.
 \end{assumption}
@@
 Then there exists an algorithm $\TensorFactorize$ that
-returns consistent estimates of $\mOpp{v}{i}$ for each $v \in \{v_1,v_2,v_3\}$.
+returns consistent estimates of $\mOpp{v}{i}$ for each $v \in \{v_1,v_2,v_3\}$ up to permutation.
 \end{theorem}

### 2014-05-17 doc/icml2014/directed.tex
Minor fixes to background. Added statement about hidden variable relabelling in exclusive views section

@@
 We now show that if a subset of hidden variables $S$ has the exclusive views property,
-then we can recover the marginal distribution $\Pr(\bh_S)$.
+then we can recover the marginal distribution $\Pr(\bh_S)$ up to relabeling of the hidden variables.\footnote{The relabeling for each hidden variables is determined by the permutation in $\TensorFactorize$.}
 Consider any $S = \{i_1, \ldots, i_m\}$ with the exclusive views property. Let

### 2014-05-17 doc/icml2014/background.tex
Minor fixes (| -> given)

@@
   %$\Pr(x_v)$ can be separated into $k$ different conditional distributions
-  %$\Pr(x_v | h_i)$. 
+  %$\Pr(x_v \given h_i)$. 
 The full rank assumption on the conditional moments $\mOpp{v}{i} = \Pr(x_v \mid h_i)$
@@
 Then there exists an algorithm $\TensorFactorize$ that
-returns consistent estimates of $\mOpp{v}{i}$ for each $v \in \{v_1,v_2,v_3\}$ up to permutation.
+returns consistent estimates of $\mOpp{v}{i}$ for each $v \in \{v_1,v_2,v_3\}$ up to relabeling of the hidden variables.
 \end{theorem}

### 2014-05-17 doc/icml2014/directed.tex
Minor fixes (| -> given)

--- 2014-05-17 doc/icml2014/directed.tex
@@
   %Once we recover the conditional moments $\mOpp{v}{j} \eqdef \Pr(x_v
-  %| h_j)$ for some $v, j$, we present a systematic approach to learn the
+  %\given h_j)$ for some $v, j$, we present a systematic approach to learn the
   %conditional probability tables for every clique for a class of directed
@@
 For which subsets of hidden nodes can we recover the marginals?  The following
-definition offers an characterization:
+definition offers a characterization:
 \begin{definition}[Exclusive views]
@@
   and (ii) the conditional moment matrix $\mOpp{v}{i} \eqdef
-  \Pr(x_{v} \mid h_i)$ have full column rank $k$ and can be recovered.
+  \Pr(x_{v} \mid h_i)$ has full column rank $k$ and can be recovered.
 We say that $S$ has the \emph{exclusive views property} if every $h_i \in S$ has an exclusive view.
@@
   &= \sum_{\vh_S} \Pr(\Sh{S}) 
-                    \Pr(x_{v_1} | h_{i_1}) \cdots \Pr(x_{v_m} | h_{i_m}) \\
+                    \Pr(x_{v_1} \given h_{i_1}) \cdots \Pr(x_{v_m} \given h_{i_m}) \\
   &= Z_{S}(\mOpp{v_1}{i_1},\dots,\mOpp{v_m}{i_m}) \\
@@
 If $S \subseteq \bx$ is a subset of hidden variables with the exclusive views property,
-then \algorithmref{learnMarginals} recovers the marginals $Z_S = \Pr(\bh_S)$.
+then \algorithmref{learnMarginals} recovers the marginals $Z_S
+= \Pr(\bh_S)$ up to a global relabeling of the hidden variables
+determined by the labeling in $\TensorFactorize$.
 \end{theorem}
@@
 
-The bottleneck property allow recovery of conditional moments,
-and the exclusive views property allow recovery of hidden marginals.
+The bottleneck property allows recovery of conditional moments,
+and the exclusive views property allows recovery of hidden marginals.
 But we will now show that the latter property is in fact implied by the former
@@
 In this setting, the conditional moments
-  $\mOpp{v}{i} \eqdef \E(x_v | h_i) \in \Re^{d \times k}$
+  $\mOpp{v}{i} \eqdef \E(x_v \given h_i) \in \Re^{d \times k}$
   would no longer be distributions but general rank $k$ matrices.

### 2014-05-17 doc/icml2014/undirected.tex
Minor fixes (| -> given)

@@
 \algorithmref{undirected} summarizes our approach.
-In \sectionref{exclusiveViews}, we showed that $\LearnMarginals$ is a consistent
+In \sectionref{hiddenMarginals}, we showed that $\LearnMarginals$ is a consistent
 estimator for $Z_\sC$ for any bottlenecked clique in the graphical model. 

### 2014-05-17 doc/icml2014/discussion.tex
Ran through aspell

@@
   between hidden variables. 
-In a pre-processing step, they ``subtract'' out the correlations and
+In a preprocessing step, they ``subtract'' out the correlations and
   recover a graph that is bottlenecked.
@@
   discrete mixtures, e.g. latent Dirichlet allocation; their approach
-  follows that of the three-view mixture model, with a pre-processing step
+  follows that of the three-view mixture model, with a preprocessing step
   that subtracts the correlation between the views.

### 2014-05-17 doc/icml2014/undirected.tex
undirected draft finished

--- 2014-05-17 doc/icml2014/undirected.tex
@@
-\section{Undirected log-linear models}
+\section{Recovering parameters}
 \label{sec:undirected}
 
-% DONE: need more of a transition from previous sections
-Recall that we used tensor factorization and composite likelihood
-to estimate the hidden marginals $Z_\sC = \BP(\vh_\sC)$ and $\mOpp{v}{i} = \BP(x_v \mid h_i)$.
-For directed models, obtaining the local conditional tables from $Z_\sC$ simply requires
-renormalizing $Z_\sC$.
-For undirected log-linear models, the canonical parameters cannot be obtained
-locally.
-Fortunately, we can construct a convex optimization problem
-by relating the recovered conditional moments with the expected features of
-the log-linear model.
+% Transition
+We have thus far shown how to recover the conditional moments
+$\mOpp{v}{i} = \BP(x_v \mid h_i)$ for each exclusive view $x_v$ of each hidden variable $h_i$,
+as well as the hidden marginals $Z_S = \BP(\vh_S)$ for each bidependent subset of hidden variables $S$.
+Now all that remains to be done is to recover the parameters.
+
+Since our graphical model is in canonical form (\assumptionref{canonical}),
+all cliques $\sC \in \sG$
+either consist of hidden variables $\vh_\sC$ or are of the form $\{x_v,h_i\}$.
+The key observation is that the clique marginals are actually sufficient
+statistics of the model $p_\theta$.
+How we turn these clique marginals $\{ \Pr(\bx_\sC,\bh_\sC) \}_{\sC \in \sG}$
+into parameters $\theta$ depends on the exact model parametrization.
+
+For directed models, the parameters are simply the local conditional tables
+$p_\theta(a \mid \Pa(a))$ for each clique $\sC = \{a\} \cup \Pa(a)$.
+These conditional distributions can be obtained by simply normalizing $Z_\sC$
+for each assignment of $\Pa(a)$.
+
+%by relating the recovered conditional moments with the expected features of
+%the log-linear model.
 
 % Define model
-As before, let $\sG$ be a graphical model with observed variables $\bx$ and hidden variables $\bh$
-(see \figureref{examples-mrf} for an example).
-For each clique $\sC \in \sG$, we have a feature vector $\phi_\sC(\vx_\sC, \vh_\sC)$,
-where $\vx_\sC$ and $\vh_\sC$ are the observed and hidden variables in $\sC$, respectively.
-Let $\phi(\vx, \vh) = \sum_{\sC \in \sG} \theta^\top \phi_\sC(\vx_\sC,\vh_\sC)$ be the global feature vector.
-We consider log-linear models of the following form:
-\begin{align*}
-p_\theta(\vx, \vh) &= \exp\left( \theta^\top \phi(\vx,\vh) - A(\theta) \right),
-\end{align*}
-where $A(\theta) = \log \sum_{\vx, \vh}  \exp( \theta^\top \phi(\vx,\vh) )$ is the log-partition function.
+%As before, let $\sG$ be a graphical model with observed variables $\bx$ and hidden variables $\bh$
+%(see \figureref{examples-mrf} for an example).
+%For each clique $\sC \in \sG$, we have a feature vector $\phi_\sC(\vx_\sC, \vh_\sC)$,
+%where $\vx_\sC$ and $\vh_\sC$ are the observed and hidden variables in $\sC$, respectively.
+%Let $\phi(\vx, \vh) = \sum_{\sC \in \sG} \theta^\top \phi_\sC(\vx_\sC,\vh_\sC)$ be the global feature vector.
+%We consider log-linear models of the following form:
+%\begin{align*}
+%p_\theta(\vx, \vh) &= \exp\left( \theta^\top \phi(\vx,\vh) - A(\theta) \right),
+%\end{align*}
+%where $A(\theta) = \log \sum_{\vx, \vh}  \exp( \theta^\top \phi(\vx,\vh) )$ is the log-partition function.
 
@@
   conditionally independent observations.
-  We derive a consistent estimate.}
+  This model has high treewidth,
+  but we can estimate it efficiently using pseudolikelihood.}
   \label{fig:examples-mrf}
@@
 
-%As an example, consider an undirected grid model in \figureref{examples-mrf},
-%where each hidden variable has two conditionally independent observations.
-%In practice, if two features are known to be independent, then they can
-  %be separated into independent observed variables.
-%Let the parameters of the model be $O \eqdef \Phi(x^a_{i,j}, h_{i,j})
-  %= \Phi(x^b_{i,j}, h_{i,j})$ and $T \eqdef \Phi(h_{i,j}, h_{i \pm 1,j \pm 1})$. 
-%Note that as this is an undirected model, the clique marginals do not
-  %correspond to the parameters of the model.
-%To recover the parameters of this model, note that every hidden
-%  variable, $h_{ij} \in \sH$ is a bottleneck, with views
-%  $x^{a}_{ij},x^{b}_{ij}$ and any other observed variable in the graph. 
-%Thus, \LearnMarginals will be able to learn the marginals of each clique
-%  and \LearnLogLinear will give us a consistent estimate of the
-%  parameters.
-
-Of course, the log-likelihood is non-concave:
-\begin{align*}
-  L_\text{unsup}(\theta) \eqdef \E_{\vx \sim p^*}[\log \sum_{\vh \in \sH} p_\theta(\vx,\vh)],
-\end{align*}
-where we are assuming for simplicity that we have infinite data drawn from the true distribution $p^*$.
-% PL: this doesn't make sense
-%While $\phi_\sC(x,h)$ can be an arbitrary featurization, we assume
-%  without loss of generality that $\phi_\sC(x,h)$ are the indicator
-%  features $\BI[x,h]$\footnote{Note that $\BI[x,h]$ are sufficient
-%  statistics for the model, and hence we do not lose any statistical
-%  efficiency in using them.}.
-
-%In this section, we will show that we can extend the results to
-  %undirected graphical models parameterized as log-linear latent variable
-  %models.
-On the other hand, 
-if we were able to observe $\vh$, we would optimize the \emph{supervised} likelihood, which is concave:
+%Of course, the log-likelihood is non-concave:
+%\begin{align*}
+  %L_\text{unsup}(\theta) \eqdef \E_{\vx \sim p_{\theta^*}}[\log \sum_{\vh \in \sH} p_\theta(\vx,\vh)],
+%\end{align*}
+%where we are assuming for simplicity that we have infinite data drawn from the true distribution $p_{\theta^*}$.
+For undirected log-linear models, the canonical parameters $\theta$ cannot be obtained locally,
+but we can construct a global convex optimization problem to solve for $\theta$.
+Suppose we were able to observe $\vh$.
+Then we could optimize the \emph{supervised} likelihood, which is concave:
 \begin{align}
-L_\text{sup}(\theta) &\eqdef \E_{(\vx,\vh) \sim p^*}[\log p_\theta(\vx,\vh)] \nonumber \\
-                     &= \theta^\top \left(\sum_{\sC \in \sG} \E[\phi(\vx_\sC,\vh_\sC)]\right) - A(\theta).
+  L_\text{sup}(\theta)
+  & \eqdef \E_{(\vx,\vh) \sim p_{\theta^*}}[\log p_\theta(\vx,\vh)] \nonumber \\
+  & = \theta^\top \left(\sum_{\sC \in \sG} \E[\phi(\vx_\sC,\vh_\sC)]\right) - A(\theta).
 \label{eqn:logLinearSupervised}
 \end{align}
-While we don't have supervised data,
-recall that the method of moments provides the hidden marginals $Z_\sC = \BP(\vh_\sC)$
-and conditional moments $\mOpp{v}{i} = \Pr(x_v \mid h_i)$.
-Assuming our graphical model is in canonical form (\lemmaref{reduction}),
-every observed variable is connected to exactly one hidden variable.
-Then $Z_\sC$ and $\mOpp{v}{i}$ directly yield marginals for every clique $\BP(\vx_\sC, \vh_\sC)$.
-Given these marginals, we can compute the expected local feature vector:
+Of course we don't have supervised data,
+but we do have the marginals $\Pr(\vx_\sC,\vh_\sC)$,
+from which we can easily compute the expected features:
+%recall that the method of moments provides the hidden marginals $Z_\sC = \BP(\vh_\sC)$
+%and conditional moments $\mOpp{v}{i} = \Pr(x_v \mid h_i)$.
+%Assuming our graphical model is in canonical form (\lemmaref{reduction}),
+%every observed variable is connected to exactly one hidden variable.
+%Then $Z_\sC$ and $\mOpp{v}{i}$ directly yield marginals for every clique $\BP(\vx_\sC, \vh_\sC)$.
+%Given these marginals, we can compute the expected local feature vector:
 \begin{align}
@@
 \end{align}
-Note that $\{\mu_\sC\}_{\sC \in \sG}$ are exactly the sufficient statistics of the model (\equationref{logLinearSupervised}),
-so we can optimize the supervised likelihood objective without actually having any supervised data!
+%Note that $\{\mu_\sC\}_{\sC \in \sG}$ are exactly the sufficient statistics of the model (\equationref{logLinearSupervised}),
+Therefore, we can optimize the supervised likelihood objective without actually
+having any supervised data!
 In the finite data regime, the method of moments yields the estimate
@@
 
-%We exploit the connection between the moments of the clique potentials and
-  %the log-partition function to show that learning parameters is convex
-  %given the clique marginals $Z_\sC$. 
-%The end result is a consistent estimator for log-linear
-  %latent-variable models which satisfy \propertyref{exclusive-views}.
-%Finally, we generalize the result to classes when only a subset of the
-  %clique marginals can be learned.
+%\algorithmref{undirected} summarizes our approach.
+%In \sectionref{exclusiveViews}, we showed that $\LearnMarginals$ is a consistent
+%estimator for $Z_\sC$ for any bottlenecked clique in the graphical model. 
+%Consequently, \algorithmref{undirected}
+%is also a consistent estimator.\footnote{Of course, the log-linear model must be identifiable.
+%Identifability is violated, for example, if two components of $\phi(\vx_\sC, \vh_\sC)$ were identical.
+%In general, we can identify the parameters up to the kernel of $\nabla^2 A(\theta)$.}
 
-%As before, our objective is to find a set of parameters $\hat \theta$
-%  that maximize the likelihood of data $\sD$.
-%\begin{align}
-%  \hat \theta 
-%      &= \arg\max_{\theta} \sum_{\vx \in \sD} \log p_\theta(\vx) \nonumber \\
-%      &= \arg\max_{\theta} \sum_{\vx \in \sD} \log \sum_{\vh \in \sH} p_\theta(\vx,\vh) \nonumber \\
-%      &= \arg\max_{\theta} \sum_{\vx \in \sD} \log \sum_{\vh \in \sH} \exp( \theta^\top\phi(\vx,\vh) - A(\theta) ). \label{eqn:obj-ml}
-%\end{align}
-%Note that this problem is not concave in $\theta$. 
-%
-%\note{Consider the supervised setting.
-%As we get infinite amounts of data, we get this simple form.
-%As MoM is a consistent estimator for marginals, we can optimize this
-%  likelihood and it is consistent.
-%}
-
-%However, given the moments $\E[ \phi_\sC(\vx,\vh) ] = Z_\sC$, the
-%  problem simplifies to,
-%\begin{align}
-%  \hat \theta 
-%  &= \arg\max_{\theta} \sum_{\sC \in \sG} \theta_\sC^\top Z_\sC - A(\theta), \label{eqn:obj}
-%\end{align}
-%which is concave in $\theta$.
-\algorithmref{undirected} summarizes our approach.
-In \sectionref{exclusiveViews}, we showed that $\LearnMarginals$ is a consistent
-estimator for $Z_\sC$ for any bottlenecked clique in the graphical model. 
-Consequently, \algorithmref{undirected}
-is also a consistent estimator.\footnote{Of course, the log-linear model must be identifiable.
-Identifability is violated, for example, if two components of $\phi(\vx_\sC, \vh_\sC)$ were identical.
-In general, we can identify the parameters up to the kernel of $\nabla^2 A(\theta)$.}
-
-\renewcommand{\algorithmicrequire}{\textbf{Input:}}
-\renewcommand{\algorithmicensure}{\textbf{Output:}}
+%\renewcommand{\algorithmicrequire}{\textbf{Input:}}
+%\renewcommand{\algorithmicensure}{\textbf{Output:}}
 \begin{algorithm}
-  \caption{\LearnParameters (for undirected models)}
+  \caption{\LearnParameters}
   \label{algo:undirected}
   \begin{algorithmic}
-    \REQUIRE Conditional moments $\mOpp{v}{i}$ and clique marginals $Z_\sC$.
-    \ENSURE Parameters $\theta$
+    \REQUIRE Conditional moments $\mOpp{v}{i} = \Pr(x_v \mid h_i)$
+             and hidden marginals $Z_S = \Pr(\bh_S)$.
+    \ENSURE Parameters $\theta$.
+    \IF{$\sG$ is directed} \STATE Normalize each $\Pr(a,\Pa(a))$. \ENDIF
+    \IF{$\sG$ is undirected with low treewidth}
       \STATE Compute expected features $\{\mu_\sC\}_{\sC \in \sG}$ (\equationref{logLinearFeatures}).
-      \STATE Optimize \equationref{logLinearSupervised}.
+      \STATE Optimize full likelihood (\equationref{logLinearSupervised}).
+    \ENDIF
+    \IF{$\sG$ is undirected with high treewidth}
+      \STATE Compute expected features $\{\mu_S\}_{S \in \sB}$ (\equationref{logLinearFeatures}).
+      \STATE Optimize pseudolikelihood (\equationref{pseudolikelihood}).
+    \ENDIF
     \end{algorithmic}
@@
 
+% Learning with partial moment constraints
+\paragraph{Remark} If we have exclusive views for only a subset of the cliques,
+we can still obtain the expected features $\mu_\sC$ for those cliques
+  and use posterior regularization \citep{graca08em}, measurements \citep{liang09measurements}, or generalized
+  expectation criteria \citep{mann08ge}
+  to encourage $\E_{p_\theta}[\phi(\vx_\sC,\vh_\sC)]$ to match $\mu_\sC$.
+The resulting objective functions would be non-convex, but we expect
+  local optima to be less of an issue.
+  %How to characterize this precisely is an interesting direction for future work.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Pseudolikelihood}
@@
 
-Unlike for directed models, in the undirected case,
-graphs with high treewidth present a challenge
-for even fully-supervised maximum likelihood due to the intractability of the log-partition function $A(\theta)$.
-%One can employ various variational approximations of $A(\theta)$ \cite{wainwright08varinf},
-This motivates looking at pseudolikelihood objective of \citep{besag75pseudo} for a computationally more
-efficient solution.
-The pseudolikelihood objective is a sum over the log-probability of each variable $a$ given its neighbors $\sN(a)$:
+While we now have a complete algorithm for estimating directed and undirected
+models, optimizing the full likelihood (\equationref{logLinearSupervised})
+can still be computationally intractable for undirected models with high treewidth
+due to the intractability of the log-partition function $A(\theta)$.
+One can employ various variational approximations of $A(\theta)$ \cite{wainwright08varinf},
+but these generally lead to inconsistent estimates of $\theta$.
+We thus turn to an older idea of pseudolikelihood \citep{besag75pseudo}.
+The pseudolikelihood objective is a sum over the log-probability of each
+variable $a$ given its neighbors $\sN(a)$:
 \begin{align}
   \label{eqn:pseudolikelihood}
-L_\text{pseudo}(\theta) \eqdef \E_{(\bx,\bh) \sim p^*}\left[\sum_{a \in X \cup H} \log p_\theta(a \mid \sN(a))\right].
+  L_\text{pseudo}(\theta) \eqdef \E_{(\bx,\bh) \sim p_{\theta^*}}\left[\sum_{a \in X \cup H} \log p_\theta(a \mid \sN(a))\right].
 \end{align}
-In the fully-supervised setting, it is well-known that pseudolikelihood provides computationally efficient consistent estimates,
-albeit at the cost of statistical efficiency.
-Coincidentally, this is the same high-level motivation for using method of moments in the first place.
+In the fully-supervised setting, it is well-known that pseudolikelihood
+provides consistent estimates which are computationally efficient
+but less statistically efficiency.
+(Coincidentally, this is the same high-level motivation for using method of
+moments in the first place.)
 
@@
 denote the sum over cliques $\sC$ that contain $a$; note that $\phi_{a, \sN(a)}$ only depends on $a$ and its neighbors $\sN(a)$.
-We can write each conditional log-likelihood as:
+We can write each conditional log-likelihood from \equationref{pseudolikelihood} as:
 \begin{align*}
@@
 where the conditional log-partition function $A_a(\theta; \sN(a)) = \log \sum_{a'} \exp(\theta^\top\phi_{a,\sN(a)}(a', \sN(a)))$
-involves marginalizing only over a single variable.
+involves marginalizing only over the single variable $a$.
 
@@
 having access to any labeled data.
-%The pseudolikelihood also replaces the log-partition function
-%  $A(\theta)$ with a sum over the variables in the neighborhood:
-%  $A_a(\theta; x_{\sN(a)}) = \sum_{a' \in \{a\} \union \sN(a)}
-%  \exp(\theta^\top\phi_{\sN(a)}(a'))$.
-Note that $\{a\}\cup\sN(a)$ in general does not form a clique.
-Fortunately, \lemmaref{bottleneck-views} holds for all biconnected subgraphs.
-We can break $\{a\}\cup\sN(a)$ into a collection of biconnected subgraphs 
-  % PL: this is not necessary
-  %thus the uniformly bottlenecked assumption alone is not sufficient to
-  %guarantee exclusive views for each variable in that subgraph.
-%However, when we do have exclusive views for each neighborhood, as in
- (e.g. \figureref{examples-mrf}).
-  Therefore, we can adapt our algorithm ($\LearnMarginals$) to estimate all marginals of the form $Z_{a,\sN(a)} \eqdef \Pr(a,\sN(a))$,
-  from which we can compute the features $\mu_{a,\sN(a)}$,
-  and optimize \equationref{pseudolikelihood} to obtain a consistent estimate of $\theta$.
+However, $\{a\}\cup\sN(a)$ does not always have exclusive views
+(for example, consider $a = h_1$ and $\sN(a) = \{h_2,h_3,h_4\}$ in \figureref{examples-tree}).
 
-% Arun: WRONG
-% The pseudolikelihood also replaces the log-partition function $A(\theta)$
-% with a sum over the variables in the neighborhood: $A_a(\theta; x_{\sN(a)}) = \sum_{a' \in \{a\} \union \sN(a)} \exp(\theta^\top\phi_{\sN(a)}(a'))$,
-% which cannot be computed from just the hidden clique marginals $Z_\sC$,
-% since the neighbors $\sN(a)$ in general do not form a clique.
-% We adapt our earlier algorithm ($\LearnClique$) to estimate the
-%   marginals over all neighborhoods $Z_{\sN(a)}$ by noting that
-%   \lemmaref{bottleneck-views} guarantees exclusive views for any connected
-%   subgraph of $\sG$, not just cliques.
+However, we can decompose $\{a\}\cup\sN(a)$ as follows:
+conditioning on $a$ partitions $\sN(a)$ into independent subsets;
+let $\sB(a)$ be the collection of these subsets,
+which we will call \emph{sub-neighborhoods}.
+For example, $\sB(h_1) = \{ \{h_2\},\{h_3\},\{h_4\} \}$ in \figureref{examples-tree}
+and $\sB(h_{2,2}) = \{ \{ h_{1,2}, h_{2,3}, h_{3,2}, h_{2,1} \} \}$ contains a single sub-neighborhood.
 
-While this method is only viable for graphs of low degree, i.e.\@ when $|\sN(a)|$ is small,
-it can handle graphs with high treewidth.
-For example, in the grid model, the degree is only 4.
-Note that neither sample nor computational complexity depends on the treewidth.
+A key observation is that for each $B \in \sB(a)$, each $\{a\} \cup B$ is bidependent:
+conditioning on $a$ does not introduce new independencies within $B$ by construction of $\sB(a)$,
+and conditioning on any $b \in B$ does not either since every other $b' \in B \backslash \{b\}$
+is connected to $a$.
+Assuming $\sG$ is bottlenecked, by \lemmaref{bottleneck-views},
+we have that $\{a\} \cup B$ has exclusive views.
+Hence, we can recover $\Pr(a,B)$ for each $a$ and $B \in \sB(a)$.
+Based on conditional independence of the sub-neighborhoods $B$ given $a$,
+we have that $\Pr(a, \sN(a)) = \Pr(a) \prod_{B \in \sB(a)} \Pr(B \mid a)$.
+This allows us to estimate the expected features $\mu_{a,\sN(a)}$ and use these quantities
+in the optimization of the pseudolikelihood objective.
 
-\paragraph{Learning with partial moment constraints.}
+%  Therefore, we can adapt our algorithm ($\LearnMarginals$) to estimate all
+%  marginals of the form $Z_{a,\sN(a)} \eqdef \Pr(a,\sN(a))$,
+%  from which we can compute the features $\mu_{a,\sN(a)}$,
+%  and optimize \equationref{pseudolikelihood} to obtain a consistent estimate of $\theta$.
 
-In general, it might be case that we only have exclusive views for a subset of the cliques.
-In this case, we can estimate the expected features $\mu_\sC$ for those cliques
-  and use posterior regularization \citep{graca08em} or the generalized expectation criteria \citep{mann08ge}
-  to encourage $\E_{p_\theta}[\phi(\vx_\sC,\vh_\sC)]$ to match $\mu_\sC$.
-The resulting objective functions are still non-convex, but we expect
-  local optima to be alleviated.
-
-%but in practice we observed that the extra information $\mu_\sC$ helps alleviate local optima.
-%We can then integrate these constraints using posterior regularization \citep{graca08em}:
-%\begin{align}
-%L_\text{pr}(\theta) \eqdef L_\text{unsup}(\theta) + C\min_{q \in \sQ} \text{KL}(q(\vx,\vh) \| p_\theta(\vx,\vh)).
-%\end{align}
-%The main idea is to regularize the parameters towards values that produce the 
-%While the optimization problem \eqref{eqn:obj-ml} is non-convex, we
-  %expect that the measurements aid recover of parameters, and indeed see
-  %this is true for experiments on simulated data.
-%\todo{Include figure}
-
-%Posterior regularization optimizes the following objective,
-%\begin{align*}
-%  \hat \theta &= \arg\min_{\theta} \arg\min_{q \in \sQ} \KL( q(\vx,\vh) \| p_{\theta}(\vx, \vh) ),
-%\end{align*}
-%where $q$ is a marginal distribution that belongs in constrained set
-%$\sQ$. In our case, $\sQ$ is constrained to agree with the partial
-%moments we recovered using \LearnMarginals, i.e. 
-%\begin{align*}
-%  \sQ &= \{ q(\vx,\vh) \given \forall \sC \in \sP ~ \E_{q(\vx,\vh)}[\phi_\sC(\vx, \vh)] = Z_\sC \}.
-%\end{align*}
-
-%In practice, we would like to regularize for error in our
-%  estimates of $Z_\sC$; to do so, we use the measurements model of
-%  \citet{liang09measurements} which defines a Bayesian model for the noise
-%  in the marginals $Z_\sC$. The resultant optimization problem is,
-%\begin{align*}
-%  \hat\theta 
-%  &= \arg\min_\theta - \sum_{x \in \sD} \log p_\theta(x) + \half \eta_\theta \| \theta \|^2 \\
-%  &\quad + \half \eta_\beta \| Z_\sC - \sum_{\vx \in \sD} \E_{\theta}[\phi_\sC(\vx,\vh)] \|^2,
-%\end{align*}
-%where $\eta_\theta$ and $\eta_\beta$ are regularization parameters.
-
-%
-%However, the objective \equationref{obj} requires us to compute the
-%  partition function, $A(\theta)$ and its gradient. 
-%This scales with the treewidth of the model, which is the number
-%  of rows/columns of the grid in this example, and hence can be
-%  intractable.
-%A more computationally efficient, but still consistent, alternative is
-%to optimize the pseudo-likelihood \todo{how do I write this out}.
+Note that our pseudolikelihood-based approach
+does depend exponentially on the size of the sub-neighborhoods,
+which could be larger than the largest clique size.
+Therefore,
+for this approach to be effective,
+each node essentially should have low degree or locally exhibits a lot of
+conditional independence.
+%While this method is only viable for graphs with small sub-neighboof low
+%degree, i.e.\@ when $|\sN(a)|$ is small,
+On the positive side, we can handle graphical models with high treewidth;
+neither sample nor computational complexity depends on the treewidth.
+For example, in the grid model, the treewidth is linear in the size of the grid,
+but it is the degree (only 4 in this case) that governs the computational complexity.

### 2014-05-17 doc/icml2014/directed.tex

@@
   %Once we recover the conditional moments $\mOpp{v}{j} \eqdef \Pr(x_v
-  %\given h_j)$ for some $v, j$, we present a systematic approach to learn the
+  %| h_j)$ for some $v, j$, we present a systematic approach to learn the
   %conditional probability tables for every clique for a class of directed
@@
   &= \sum_{\vh_S} \Pr(\Sh{S}) 
-                    \Pr(x_{v_1} \given h_{i_1}) \cdots \Pr(x_{v_m} \given h_{i_m}) \\
+                    \Pr(x_{v_1} | h_{i_1}) \cdots \Pr(x_{v_m} | h_{i_m}) \\
   &= Z_{S}(\mOpp{v_1}{i_1},\dots,\mOpp{v_m}{i_m}) \\
@@
 In this setting, the conditional moments
-  $\mOpp{v}{i} \eqdef \E(x_v \given h_i) \in \Re^{d \times k}$
+  $\mOpp{v}{i} \eqdef \E(x_v | h_i) \in \Re^{d \times k}$
   would no longer be distributions but general rank $k$ matrices.

### 2014-05-17 doc/icml2014/discussion.tex
finish discussion

@@
 For latent-variable models,
-there exists a tension between
+there has been tension between
 local optimization of likelihood,
-which is broadly applicable but has no global theoretical guarantees,
-and the method of moments, which provides consistent estimators
+which is broadly applicable but offers no global theoretical guarantees,
+and the spectral method of moments, which provides consistent estimators
 but are limited to models with special structure.
@@
 for a broader class of directed and undirected models.
-A notable example are undirected log-linear models with hidden variables of
-high treewidth.
-% Classes of models
+%for example, undirected log-linear models with high treewidth.
 
-% Repeat
-%Our approach can be summarized: (i) perform tensor factorization \citep{anandkumar13tensor} to
-%obtain information about each hidden variable in the form of conditional moments;
-%In some sense, the presence of these views helps tie down the hidden
-%  variables to the data. 
-%(ii) use composite likelihood to estimate the marginal distributions of latent variables;
-%and (iii) in the case of undirected models,
-%use the marginal distributions to formulate a convex optimization to
-%recover the model parameters.
-
-% Bottleneck is key
+% Bottleneck is key: previous work is preprocessing
 Our approach provides consistent estimates for
-models in which each hidden variable is a ``bottleneck''---that is,
-it has three conditionally independent observations
-(although the dependence on the observations could be quite complex as a function of the parameters).
+a family of models in which each hidden variable is a \emph{bottleneck}---that is,
+it has three conditionally independent observations.
 This bottleneck property of \citet{anandkumar13tensor}
@@
 including Latent Dirichlet Allocation \cite{anandkumar12lda},
-parsing models \cite{hsu12identifiability},
+mixture of spherical Gaussians \cite{hsu13spherical},
+probabilistic grammars \cite{hsu12identifiability},
 noisy-or Bayesian networks \cite{halpern13noisyor},
@@
 and others.
-We also introduce the idea of \emph{exclusive views} (\definitionref{exclusive-views}),
-the key property that allows us to learn the correlation structure between hidden variables.
+Each of these methods can be viewed as ``preprocessing'' the given model into a
+form that exposes the bottleneck or tensor factorization structure.
+The model parameters correspond directly to the solution of the factorization.
+
+% Our work: post-processing
+In contrast, the bottlenecks in our graphical models are given by assumption,
+but the conditional distribution of the observations given the bottleneck
+can be quite complex.
+Our work can therefore be viewed as ``postprocessing'',
+where the conditional moments recovered from tensor factorization
+are used to further obtain the hidden marginals and eventually the parameters.
+Along the way, we developed the notion of exclusive views and bidependent sets,
+which characterize conditions under which the conditional moments can reveal
+the dependency structure between hidden variables.
+We also made use of custom likelihood functions which were constructed to be
+easy to optimize.
+
+%We also show that bottlenecks lead to a notion of \emph{exclusive views}
+%(\definitionref{exclusive-views}), the key property that allows us to learn the
+%correlation structure between hidden variables.
 
@@
 % Correlation structure
-Other authors, e.g \citet{anandkumar12lda, anandkumar2013linear,
-  halpern13noisyor}, learn parameters for models outside this
-  family, but they do so with some knowledge of the correlation structure
-  between hidden variables. 
-In a preprocessing step, they ``subtract'' out the correlations and
-  recover a graph that is bottlenecked.
+%Other authors \citep{anandkumar12lda, anandkumar2013linear,
+%  halpern13noisyor} learn parameters for models outside our
+%  family, but they do so with additional knowledge of the correlation structure
+%  between hidden variables, allowing them to ``subtract'' out these correlations and
+%  obtain a graph that is bottlenecked.
 % TOO BOLD
@@
 %\todo{What about Mossel and Roch?}
-For example,
-\citet{anandkumar12lda} are able to recover parameters for correlated
-  discrete mixtures, e.g. latent Dirichlet allocation; their approach
-  follows that of the three-view mixture model, with a preprocessing step
-  that subtracts the correlation between the views.
-\citet{anandkumar2013linear} study a special family of Bayesian networks
-  in which the observed variables are linearly related to a sparse set of
-  hidden variables. Once again, the key step is to identify the
-  correlations between observed variables and exclude them from the views.
-\citet{halpern13noisyor} propose an algorithm to recover factors for
-  a more densely connected bipartite noisy-or network; they also use
-  bottlenecks to learn parameters, but exploit properties of the model
-  to negate correlations from learned edges, allowing for more
-  bottlenecks to be discovered. 
-Our work makes a different tradeoff: we do not assume any special correlation structure,
-but do rely on every hidden variable having an exclusive view so that we can
-learn the underlying structure.
+%For example,
+%\citet{anandkumar12lda} are able to recover parameters for correlated
+%  discrete mixtures, e.g. latent Dirichlet allocation; their approach
+%  follows that of the three-view mixture model, with a pre-processing step
+%  that subtracts the correlation between the views.
+%\citet{anandkumar2013linear} study a special family of Bayesian networks
+%  in which the observed variables are linearly related to a sparse set of
+%  hidden variables. Once again, the key step is to identify the
+%  correlations between observed variables and exclude them from the views.
+%\citet{halpern13noisyor} propose an algorithm to recover factors for
+%  a more densely connected bipartite noisy-or network; they also use
+%  bottlenecks to learn parameters, but exploit properties of the model
+%  to negate correlations from learned edges, allowing for more
+%  bottlenecks to be discovered. 
+%Our work makes a different tradeoff: we do not assume any special correlation structure,
+%but do rely on every hidden variable having an exclusive view so that we can
+%learn the underlying structure.
 
@@
 Another prominent line of work in the method of moments community has
-  focused on recovering an {\em observable operator
-  representation} of the observations.
+  focused on recovering {\em observable operator
+  representations} \citep{jaeger2000observable,hsu09spectral,bailly2010spectral,balle12automata}.
   These methods allow prediction of new observations, but do not
-  identify the actual parameters of the model, which makes them difficult to use
+  recover the actual parameters of the model, making them difficult to use
   in conjunction with likelihood-based models. % which are driven by parameters.
-\citet{song2011spectral} propose an algorithm to learn an observable
+\citet{song2011spectral} proposed an algorithm to learn observable
   operator representations for latent tree graphical models, like the
@@
 Their approach is similar to our first step of learning conditional moments,
-  but only consider trees.
-\citet{parikh12spectral} extend this approach to general graphical
-  models which are bottlenecked, but consider the latent
-  junction tree representation. 
+  but they only consider trees.
+\citet{parikh12spectral} extended this approach to general graphical
+  models which are bottlenecked using a latent junction tree representation. 
   Consequently, the size of the observable representations is exponential in
@@
   In contrast, our algorithm only constructs moments of the order of size of the cliques
-  (and neighborhoods for pseudolikelihood), which can be much smaller.
+  (and sub-neighborhoods for pseudolikelihood), which can be much smaller.
 For example, an $n\times n$ grid model has a treewidth of $n$, but each
@@
 % Identifiability
-An interesting direction of study is to examine the necessity of the bottleneck
-property.  Certainly, to some extent, three views are necessary to ensure
+An interesting direction is to examine the necessity of the bottleneck
+property.  Certainly, more than two views is necessary to ensure
 identifiability \cite{kruskal77three}, but requiring \emph{each} hidden variable to be
-a bottleneck is still stronger than we would like.  We hope that by judiciously
-leveraging likelihood-based methods in conjunction with the method of moments
+a bottleneck is stronger than we would like.  We hope that by judiciously
+leveraging likelihood-based methods in conjunction with the method of moments,
 we can generate new hybrid techniques for estimating

### 2014-05-17 doc/icml2014/directed.tex

--- 2014-05-17 doc/icml2014/directed.tex
@@
 In the special case where the graphical model represents a binary-valued noisy-or network,
-  we can use the algorithm of \citet{halpern13noisyor},
+  we can use the algorithm of \citet{halpern2013unsupervised},
   which first learns $h_2$ and subtracts off its influence,

### 2014-05-17 doc/icml2014/discussion.tex

--- 2014-05-17 doc/icml2014/discussion.tex
@@
 parsing models \cite{hsu12identifiability},
-noisy-or Bayesian networks \cite{halpern13noisyor},
+noisy-or Bayesian networks \cite{halpern2013unsupervised},
 mixture of linear regressions \cite{chaganty13regression},
@@
 Other authors, e.g \citet{anandkumar12lda, anandkumar2013linear,
-  halpern13noisyor}, learn parameters for models outside this
+  halpern2013unsupervised}, learn parameters for models outside this
   family, but they do so with some knowledge of the correlation structure
@@
   correlations between observed variables and exclude them from the views.
-\citet{halpern13noisyor} propose an algorithm to recover factors for
+\citet{halpern2013unsupervised} propose an algorithm to recover factors for
   a more densely connected bipartite noisy-or network; they also use

### 2014-05-17 doc/icml2014/undirected.tex

@@
     \ENSURE Parameters $\theta$.
-    \IF{$\sG$ is directed} \STATE Normalize each $\Pr(a,\Pa(a))$. \ENDIF
-    \IF{$\sG$ is undirected with low treewidth}
+    \IF{$\sG$ is directed} \STATE Normalize each $\Pr(a,\Pa(a))$. 
+    \ELSIF{$\sG$ is undirected with low treewidth}
       \STATE Compute expected features $\{\mu_\sC\}_{\sC \in \sG}$ (\equationref{logLinearFeatures}).
       \STATE Optimize full likelihood (\equationref{logLinearSupervised}).
-    \ENDIF
-    \IF{$\sG$ is undirected with high treewidth}
-      \STATE Compute expected features $\{\mu_S\}_{S \in \sB}$ (\equationref{logLinearFeatures}).
+    \ELSIF{$\sG$ is undirected with high treewidth}
+      \STATE Compute expected features $\{\mu_S\}_{S \in \sB}$ (\equationref{neighborhoodFeatures}).
       \STATE Optimize pseudolikelihood (\equationref{pseudolikelihood}).
@@
 
-If we knew $\mu_{a,\sN(a)} \eqdef \E[\phi_{a,\sN(a)}(a,\sN(a))]$,
+If we knew the marginals for each neighborhood,
+\begin{align}
+  \mu_{a,\sN(a)} &\eqdef \E[\phi_{a,\sN(a)}(a,\sN(a))], \label{eqn:neighborhoodFeatures}
+\end{align}
 then we would be able to optimize the pseudolikelihood objective again without

### 2014-05-17 doc/icml2014/discussion.tex

--- 2014-05-17 doc/icml2014/discussion.tex
@@
 probabilistic grammars \cite{hsu12identifiability},
-noisy-or Bayesian networks \cite{halpern13noisyor},
+noisy-or Bayesian networks \cite{halpern2013unsupervised},
 mixture of linear regressions \cite{chaganty13regression},

### 2014-05-18 doc/icml2014/undirected.tex
Updated undirected section

@@
 provides consistent estimates which are computationally efficient
-but less statistically efficiency.
-(Coincidentally, this is the same high-level motivation for using method of
-moments in the first place.)
+but less statistically efficiency.\footnote{
+Coincidentally, this is the same high-level motivation for using method of
+moments in the first place.}
 
@@
 having access to any labeled data.
-However, $\{a\}\cup\sN(a)$ does not always have exclusive views
-(for example, consider $a = h_1$ and $\sN(a) = \{h_2,h_3,h_4\}$ in \figureref{examples-tree}).
+Unfortunately, $\{a\}\cup\sN(a)$ does not always have exclusive views.
+For example, consider $a = h_1$ and $\sN(a) = \{h_2,h_3,h_4\}$ in \figureref{examples-tree}.
 
@@
 For example, $\sB(h_1) = \{ \{h_2\},\{h_3\},\{h_4\} \}$ in \figureref{examples-tree}
-and $\sB(h_{2,2}) = \{ \{ h_{1,2}, h_{2,3}, h_{3,2}, h_{2,1} \} \}$ contains a single sub-neighborhood.
+and $\sB(h_{2,2}) = \{ \{ h_{1,2}, h_{2,3}, h_{3,2}, h_{2,1} \} \}$ contains a single sub-neighborhood in \figureref{examples-mrf}.
 
-A key observation is that for each $B \in \sB(a)$, each $\{a\} \cup B$ is bidependent:
+A key observation is that for each sub-neighborhood $B \in \sB(a)$, each $\{a\} \cup B$ is bidependent:
 conditioning on $a$ does not introduce new independencies within $B$ by construction of $\sB(a)$,
@@
 is connected to $a$.
-Assuming $\sG$ is bottlenecked, by \lemmaref{bottleneck-views},
+Assuming $\sG$ is bottlenecked, by \lemmaref{bottleneck-views}
 we have that $\{a\} \cup B$ has exclusive views.
@@
 for this approach to be effective,
-each node essentially should have low degree or locally exhibits a lot of
+each node essentially should have low degree or locally exhibit a lot of
 conditional independence.

### 2014-05-18 doc/icml2014/abstract.tex
minor edits

@@
   consistent parameter estimation, but
-  only for certain types of latent-variable models. %, avoiding local optima issues,
-  %but are applicable only to certain types of graphical models.
+  only for certain types of latent-variable models.
 On the other hand, pure likelihood objectives, though more universally applicable,
@@
 In this work, we show that using the method of moments in conjunction
-  with composite marginal likelihood
+  with composite likelihood
   yields consistent parameter estimates for a much broader
@@
   including loopy graphs with high treewidth.
-%In particular, our algorithm applies to any graphical model wherein each
-  %hidden variable has at least three conditionally independent observed
-  %variables.
-  %that sidestep local optima issues for a general class of
-  %directed and undirected graphical models.
-%Parameter estimation in latent-variable models based on EM
-  %is prone to local optima.
-  %due to the non-convexity of the marginal log-likelihood. 
-%In this paper, we extend recent advances in the method of moments 
-  %to provide consistent parameter estimates
-  %that sidestep local optima issues for a general class of
-  %directed and undirected graphical models.
-%In particular, our algorithm applies to any graphical model wherein each
-  %hidden variable has at least three conditionally independent observed
-  %variables.
-%This family includes loopy graphs with high treewidth.
-Specifically, we first use tensor factorization to reveal information
-about the hidden variables.  This renders the otherwise non-convex
-likelihood objective convex, and EM can then be used to obtain the global optimum.
-%We use the method of moments to tie each hidden to some observed
-  %variables, and show that the rest of the problem can be cast as convex
-  %optimization of the likelihood.
-%Our approach gracefully extends to models outside our class by
-  %incorporating the partial information via posterior regulraization.
+Specifically, we use tensor factorization to reveal information
+about the hidden variables.
+This allows us to construct convex likelihoods which can be globally optimized
+to recover the parameters.
 \end{abstract} 
+%This renders the otherwise non-convex
+%likelihood objective convex, and EM can then be used to obtain the global optimum.
 

### 2014-05-18 doc/icml2014/discussion.tex
minor edits

@@
   (and sub-neighborhoods for pseudolikelihood), which can be much smaller.
-For example, an $n\times n$ grid model has a treewidth of $n$, but each
-  clique is of size at most $2$ and the degree is at most $4$.
 %In some sense, an observable operator re-formulation of our algorithm
@@
 An interesting direction is to examine the necessity of the bottleneck
-property.  Certainly, more than two views is necessary to ensure
+property.  Certainly, three views is in general needed to ensure
 identifiability \cite{kruskal77three}, but requiring \emph{each} hidden variable to be
-a bottleneck is stronger than we would like.  We hope that by judiciously
+a bottleneck is stronger than what we would like.  We hope that by judiciously
 leveraging likelihood-based methods in conjunction with the method of moments,
 we can generate new hybrid techniques for estimating
-an even richer class of latent-variable models.
+even richer classes of latent-variable models.

### 2014-05-18 doc/icml2014/undirected.tex
minor edits

@@
     \ENSURE Parameters $\theta$.
-    \IF{$\sG$ is directed} \STATE Normalize each $\Pr(a,\Pa(a))$. 
+    \IF{$\sG$ is directed} \STATE Normalize $\Pr(a,\Pa(a))$ for $a \in \bx \cup \bh$.
     \ELSIF{$\sG$ is undirected with low treewidth}
-      \STATE Compute expected features $\{\mu_\sC\}_{\sC \in \sG}$ (\equationref{logLinearFeatures}).
+      \STATE Compute features $\mu_\sC$ for $\sC \in \sG$ (\equationref{logLinearFeatures}).
       \STATE Optimize full likelihood (\equationref{logLinearSupervised}).
     \ELSIF{$\sG$ is undirected with high treewidth}
-      \STATE Compute expected features $\{\mu_S\}_{S \in \sB}$ (\equationref{neighborhoodFeatures}).
+      %\STATE Compute expected features $\mu_{\{a\} \cup B}$ for $a \in \bh, B \in \sB(a)$ (\equationref{neighborhoodFeatures}).
+      \STATE Compute features $\mu_{\{a\} \cup \sN(a)}$ for $a \in \bh$ (\equationref{neighborhoodFeatures}).
       \STATE Optimize pseudolikelihood (\equationref{pseudolikelihood}).
@@
   \label{eqn:pseudolikelihood}
-  L_\text{pseudo}(\theta) \eqdef \E_{(\bx,\bh) \sim p_{\theta^*}}\left[\sum_{a \in X \cup H} \log p_\theta(a \mid \sN(a))\right].
+  L_\text{pseudo}(\theta) \eqdef \E_{(\bx,\bh) \sim p_{\theta^*}}\left[\sum_{a \in \bx \cup \bh} \log p_\theta(a \mid \sN(a))\right].
 \end{align}
@@
 we have that $\Pr(a, \sN(a)) = \Pr(a) \prod_{B \in \sB(a)} \Pr(B \mid a)$.
-This allows us to estimate the expected features $\mu_{a,\sN(a)}$ and use these quantities
+This allows us to compute the expected features $\mu_{a,\sN(a)}$ and use them
 in the optimization of the pseudolikelihood objective.
@@
 On the positive side, we can handle graphical models with high treewidth;
-neither sample nor computational complexity depends on the treewidth.
-For example, in the grid model, the treewidth is linear in the size of the grid,
-but it is the degree (only 4 in this case) that governs the computational complexity.
+neither sample nor computational complexity necessarily depends on the treewidth.
+%For example, for an $n \times n$ grid model, the treewidth is $n$,
+%but it is the degree (only 4 in this case) that governs the computational complexity.
+For example, an $n\times n$ grid model has a treewidth of $n$, but each
+  clique is of size at most $2$ and the degree is at most $4$.

### 2014-05-18 doc/icml2014/undirected.tex
Minor change in phi notation - undirected

--- 2014-05-18 doc/icml2014/undirected.tex
@@
 \end{align*}
-where the conditional log-partition function $A_a(\theta; \sN(a)) = \log \sum_{a'} \exp(\theta^\top\phi_{a,\sN(a)}(a', \sN(a)))$
+where the conditional log-partition function $A_a(\theta; \sN(a)) = \log \sum_{\alpha \in [k]} \exp(\theta^\top\phi_{a,\sN(a)}(\alpha, \sN(a)))$
 involves marginalizing only over the single variable $a$.

### 2014-05-18 doc/icml2014/directed.tex

@@
     \label{fig:bottleneckExclusiveViews}
-  (a) A bottleneck can be estimated consistently
-  using tensor factorization \citep{anandkumar13tensor}.
-  (b) An example of exclusive views for a bidependent subset $S$.
+  (a) A bottleneck $h_1$ has three conditionally independent views $x_1,x_2,x_3$.
+  %using tensor factorization \citep{anandkumar13tensor}.
+  (b) A bidependent subset $S$ has exclusive views $\{x_1,x_2,x_3,x_4\}$.
   }
@@
 then we can recover the marginal distribution $\Pr(\bh_S)$.
-Consider any $S = \{i_1, \ldots, i_m\}$ with the exclusive views property. Let
+Consider any $S = \{h_{i_1}, \ldots, h_{i_m}\}$ with the exclusive views property. Let
   $x_{v_j}$ be an exclusive view for $h_{i_j}$ in $S$ and define $\sV
-  = \{v_1, \ldots, v_m\}$. % be a set of exclusive views for the clique $S$.
+  = \{x_{v_1}, \ldots, x_{v_m}\}$. % be a set of exclusive views for the clique $S$.
 By the exclusive views property,
@@
 is a linear function of $Z_S$ (what we seek to recover).
-We can solve for the hidden marginals $Z_S$ simply by projecting $\mO_\sV$ onto the column
-space of $\mOppAll$:
+We can solve for the hidden marginals $Z_S$ simply by multiplying $\mO_\sV$ by the pseudoinverse of
+$\mOppAll$:
 \begin{align*}
@@
     \REQUIRE Hidden subset $S = \{ h_{i_1}, \dots, h_{i_m} \}$ with exclusive views $\sV = \{ x_{v_1}, \dots, x_{v_m} \}$
-    and conditional moments $\mOpp{v_j}{i_j} = \Pr(x_{v_i} \mid h_{i_j})$.
+    and conditional moments $\mOpp{v_j}{i_j} = \Pr(x_{v_j} \mid h_{i_j})$.
     \ENSURE Marginals $Z_S = \Pr(\bh_S)$.
       %\STATE Identify exclusive views $\Sx{\sV} = \{x_{v_1}, \dots, x_{v_m}\}$.
-      \STATE Return $\mH_S \gets \mO_{\sV}( \mOpp{v_1}{i_i}\pinv, \dots, {\mOpp{v_m}{i_m}}\pinv )$.
+      \STATE Return $\mH_S \gets \mO_{\sV}( \mOpp{v_1}{i_1}\pinv, \dots, {\mOpp{v_m}{i_m}}\pinv )$.
   \end{algorithmic}
@@
 (e.g., $\{h_1,h_2,h_3\}$ in \figureref{bottleneckExclusiveViews}(b)).
-This will be important for pseudolikelihood in \sectionref{pseudolikelihood}.
+This will be important in \sectionref{pseudolikelihood}.
 
@@
   of the exclusive views property, while three is needed for condition (ii).
-The bottleneck property (\definitionref{bottleneck} can also be relaxed if some cliques
+The bottleneck property (\definitionref{bottleneck}) can also be relaxed if some cliques
   share parameters (see examples below).
@@
 Our method extends naturally to the case in which the observed variables
-  are real-valued $x_v \in \Re^d$, as long as the hidden variables remain discrete. 
+  are real-valued ($x_v \in \Re^d$), as long as the hidden variables remain discrete. 
 In this setting, the conditional moments

### 2014-05-18 doc/icml2014/background.tex
more change

@@
 Every graphical model can be transformed into canonical
-form (via \figureref{reduction}).
+form. % (via \figureref{reduction}).
 There is a one-to-one correspondence between the parameters of the transformed
@@
 
-% No parameters yet
 The tensor factorization method attacks the heart of the non-convexity
@@
   of the graphical model, not its parametrization.
+
+% No parameters yet
   If $i$ is the single parent of $v$ (e.g., $\Pr(x_1^a \mid h_1)$ in \figureref{approach}),
@@
   and the model parameters,
-  something we will address in the next two sections.
+  which we will address in the next two sections.
 

### 2014-05-18 doc/icml2014/undirected.tex
more change

--- 2014-05-18 doc/icml2014/undirected.tex
@@
 does depend exponentially on the size of the sub-neighborhoods,
-which could be larger than the largest clique size.
+which could be exceed the largest clique size.
 Therefore,
-for this approach to be effective,
+%for our approach to be effective,
 each node essentially should have low degree or locally exhibit a lot of
@@
 %but it is the degree (only 4 in this case) that governs the computational complexity.
-For example, an $n\times n$ grid model has a treewidth of $n$, but each
-  clique is of size at most $2$ and the degree is at most $4$.
+For example, an $n\times n$ grid model has a treewidth of $n$, but the degree is at most $4$.

### 2014-05-18 doc/icml2014/background.tex
Fixed a bunch of bugs

@@
   exists three observed variables (views), $x_{v_1}, x_{v_2}, x_{v_3}$,
-  that are conditionally independent given $h_i$,
+  that are conditionally independent given $h_i$ (\figureref{bottleneck}),
   and (ii) each $\mOpp{v}{i} \eqdef \Pr(x_v \mid h_i) \in \Re^{d \times k}$ has full column rank $k$

### 2014-05-18 doc/icml2014/discussion.tex
Fixed a bunch of bugs

@@
 has been exploited in many other contexts,
-including Latent Dirichlet Allocation \cite{anandkumar12lda},
+including latent Dirichlet allocation \cite{anandkumar12lda},
 mixture of spherical Gaussians \cite{hsu13spherical},

### 2014-06-18 doc/nips2013/rebuttal.txt

Thank you for your comments; we will try to respond to your questions below.

# Major Comments

## Regarding Novelty and Section 4 (Examples)

Several reviewers remarked that our development of an algorithm for parameter estimation in latent variable models and the examples we used in Section 4 were derivative from [8,13]. While we certainly build on their work, the contribution of our work is to present an alternative view point to parameter estimation using the method of moments wherein the moments act as constraints for an objective as opposed viewing the problem as one of solving a polynomial. This perspective allows us to open up the space of possible estimation procedures in a very practical sense. We demonstrate this in section 4 by describing how to estimate parameters for various MRF models, models which cannot actually be handled using prior work.

in particular, our presentation of the HMM might be better appreciated if one realizes that it is actually an undirected MRF with general parameter tying. This allows arbitrary coarse features (e.g. all words ending in '-ing') to be included in the model, which has been shown to be extremely effective for several unsupervised language learning tasks in [25]. Further more,
the moment equations for the grid and tall mixture models are considerably complex and it is not clear how the techniques employed in [8,13] could be used to solve them.

% In the HMM example, we were able to use all of the data to get fairly reliable estimates of the emission probabilities ($O$) and consequently avoid several nasty local optimal, as the experiments show.

[25] Painless Unsupervised Learning with Features; Taylor Berg-Kirkpatrick, Alexandre Bouchard-Côté, John DeNero, Dan Klein; NAACL 2010.

## Strong Convexity of Log-Partition Function (Reviewer 3)
If the log-partition function is not strongly convex, standard results show that the model is not identifiable and no procedure can work. Our algorithm will constrain the parameters to the equivalence class of correct parameters and allows regularization, etc. to break ties if desired. In effect, our method does not require the assumption that the log-partition function be strongly convex, but when it is, it finds an exact fit in the exponential family.

## Regarding how the HMMs and mixture models were generated (Reviewer 3)
We generated parameters for both the mixture model and for the
HMMs by generating the model parameter vector $\theta$ with uniform
random entries in [-1,1]. Note that normalization is taken care of by
the partition function.

## Regarding Assumptions 1 and 2 (Reviewer 3)
The assumption that $B$ have full column rank (Assumption 1) is a standard assumption [8,11,13] and is reasonable in practice. From a practical stand-point, if $B$ is rank deficient, then it is indicative that we are over-estimating the number of latent states, and should choose a smaller $k$.
Assumption 2 is certainly strong, but can still be useful when in a large network, we can find a small subset of sources from which we have 3 conditionally independent views (e.g. as in [26])

[26] Unsupervised Learning of Noisy-Or Bayesian Networks; Yoni Halpern, David Sontag; UAI 2013.

## Regarding Section 5 (Reviewer 6)

The size of the composite domain is indeed the product over the domains of the latent variables. With regard to the sensitivity of the unshuffling procedure, the estimates $A_2$ that appear in the regularization term have a sample complexity that is polynomial in the separation between factors; this is not too different from other separation terms that appear in the method of moments algorithms in [8,13].

### 2014-06-18 doc/icml2014/assumption-proof.tex

@@
   variable $h_i$ to learn conditional moments $\mOpp{v}{i} \eqdef
-  \Pr(x_v | h_i)$ for every view $x_v \in \sV_{h_i}$ using
+  \Pr(x_v \given h_i)$ for every view $x_v \in \sV_{h_i}$ using
   \TensorFactorize. 
@@
 Without loss of generality, let $i = 1$. We will express the conditional
-distribution $\mOpp{v}{1} \eqdef \Pr(x_v | h_1)$ recursively through the
+distribution $\mOpp{v}{1} \eqdef \Pr(x_v \given h_1)$ recursively through the
 parents of $x_v$, its parents, and so on. The approach is analogous to
@@
 Note that $\mYpp{t}{\Pa(h_t)}$ is the conditional probability $\Pr(h_t
- | \Pa(h_t))$, which is assumed to be full rank, and can be easily
+ \given \Pa(h_t))$, which is assumed to be full rank, and can be easily
  computed from the hidden marginals $Z_\sC$ that contain $h_t$.

### 2014-06-18 doc/icml2014/directed.tex

@@
     \includegraphics[width=0.35\columnwidth]{figures/three-view.pdf}
-    \label{fig:three-view}
+    \label{fig:bottleneck}
   }
@@
   (i) there exists some observed variable
-  $x_{v}$ which is conditionally independent of the others $S \backslash \{ h_i \}$ given $h_i$,
+  $x_{v}$ which is conditionally independent of the others $S \backslash \{ h_i \}$ given $h_i$ (\figureref{exclusive-views}),
   and (ii) the conditional moment matrix $\mOpp{v}{i} \eqdef
@@
 = \Pr(\bh_S)$ up to a global relabeling of the hidden variables
-determined by the labeling in $\TensorFactorize$.
+determined by the labeling from $\TensorFactorize$.
 \end{theorem}

### 2014-06-18 doc/icml2014/measurements.tex

@@
 Suppose that some of the variables are observed ($x = \{ z_j \}_{j \in O}$, where $O \subset V$) 
-and the rest are latent ($h = \{ z_j \}_{j \in H}$, where $H = V \backslash O$).
+and the rest are hidden ($h = \{ z_j \}_{j \in H}$, where $H = V \backslash O$).
 Let $C \subset 2^V$ be a collection of subsets of variables,

### 2014-06-18 doc/icml2014/tensor-multiplication.tex
Fixing bugs in piecewise
Updates to the pw proof
Updated sigma_max, sigma_min to be sigma_1, sigma_k

@@
 Let $M$ be an index matching set, i.e. $M = \{ (a_1; b_1),
-  (a_2, b_2), \ldots, (a_\ell, b_\ell) \}$, such that the corresopnding
+  (a_2, b_2), \ldots, (a_\ell, b_\ell) \}$, such that the corresponding
   dimensions agree, i.e. $d_{a_i} = d'_{b_i}$ for all $i$. Let $D_{\ba}
@@
   is associative and distributive.
-It is also commutative for symmetric tensors.
 
@@
 \begin{align*}
-\sigma_{\max}(\sR\munf{I}) &\le \sigma_{\max}(\sP\munf{I_P}) \sigma_{\max}(\sQ\munf{I_P}) \\
-\sigma_{\min}(\sR\munf{I}) &\ge \sigma_{\min}(\sP\munf{I_Q}) \sigma_{\min}(\sQ\munf{I_Q}).
+\sigma_{1}(\sR\munf{I}) &\le \sigma_{1}(\sP\munf{I_P}) \sigma_{1}(\sQ\munf{I_P}) \\
+\sigma_{k}(\sR\munf{I}) &\ge \sigma_{k}(\sP\munf{I_Q}) \sigma_{k}(\sQ\munf{I_Q}).
 \end{align*}
@@
 \begin{align}
-\sigma_{\max}(R) &\le \sigma_{\max}(P) \sigma_{\max}(Q) \\
-\sigma_{\min}(R) &\ge \sigma_{\min}(P) \sigma_{\min}(Q) \label{eqn:matrix-singular}.
+\sigma_{1}(R) &\le \sigma_{1}(P) \sigma_{1}(Q) \\
+\sigma_{k}(R) &\ge \sigma_{k}(P) \sigma_{k}(Q) \label{eqn:matrix-singular}.
 \end{align}
@@
   and $\sQ\munf{I_Q}$ by taking the outer product of $\sP\munf{I_P}$ with
-  $|I_Q|$ identity matrices, and vice-versa. Define 
+  $|I_Q|$ identity matrices, and vice versa. Define 
 \begin{align*}
@@
 \begin{align*}
-\sigma_{\max}(\sR\munf{I}) 
-    &\le \sigma_{\max}({\sP'}\munf{I'_P}) \sigma_{\max}({\sQ'}\munf{I'_Q}) \\
-      &= \sigma_{\max}(\sP\munf{I_P}) \sigma_{\max}(\sQ\munf{I_Q}) \\
-    \sigma_{\min}(\sR\munf{I})  
-    &\ge \sigma_{\min}({\sP'}\munf{I'_P}) \sigma_{\min}({\sQ'}\munf{I'_Q}) \\
-      &= \sigma_{\min}(\sP\munf{I_P}) \sigma_{\min}(\sQ\munf{I_Q}).
+\sigma_{1}(\sR\munf{I}) 
+    &\le \sigma_{1}({\sP'}\munf{I'_P}) \sigma_{1}({\sQ'}\munf{I'_Q}) \\
+      &= \sigma_{1}(\sP\munf{I_P}) \sigma_{1}(\sQ\munf{I_Q}) \\
+    \sigma_{k}(\sR\munf{I})  
+    &\ge \sigma_{k}({\sP'}\munf{I'_P}) \sigma_{k}({\sQ'}\munf{I'_Q}) \\
+      &= \sigma_{k}(\sP\munf{I_P}) \sigma_{k}(\sQ\munf{I_Q}).
 \end{align*}
@@
 \begin{align*}
-  \sigma_{\max}(\sQ\munf{I}) &\le \sigma_{\max}(\sP\munf{I}) v_{\max} \\
-  \sigma_{\min}(\sQ\munf{I}) &\ge \sigma_{\min}(\sP\munf{I}) v_{\min}.
+  \sigma_{k}(\sP\munf{I}) &\ge \frac{\sigma_{k}(\sQ\munf{I})}{v_{\max}}.
 \end{align*}
-In particular, $\rank(\sP\munf{I}) = \rank(\sQ\munf{I})$.
+In particular, if $\sQ\munf{I}$ has full row/column rank, $\sR\munf{I}$
+  does too.
 \end{lemma}
@@
     \sQ\munf{I}[\vi;\vj] 
-      &= \sum_{j'=\ones, k=1}^{\bd, d_m} \sP\munf{I}[\vi;\vj'\ k] (v[k] \delta[\vj';\vj]) \\
+    &= \sum_{j'=\ones, k=1}^{\bd, d_m} \sP\munf{I}[\vi;\vj'\ k] \underbrace{(v[k] \delta[\vj' \vj])}_{v'[k \vj'; \vj]} \\
     \sQ\munf{I}
@@
   rank, with singular values $v_1, \cdots, v_{d_m}$. 
+  Matrix multiplication can not increase the rank of the matrix, and
+  thus $\rank(\sP\munf{I}) \ge \rank(\sQ\munf{I})$.
+
 \end{proof}

### 2014-06-19 doc/icml2014/icml2014-presentation/diagrams.tex

@@
     \draw[-latex] (h) -- (x3);
+  }
+}
+
+\newcommand<>{\drawgensquiggle}[1]{%
+  \uncover#2{
+    \point{start-gen}{#1}
+    \node[style=node] (h) at (start-gen) {$h$};
+%    \node[left=0.1cm of h] {$h$};
+    \node[style=obsnode,below=1cm of h] (x2) {$x_2$};
+    \node[style=obsnode,left=0.5cm of x2] (x1) {$x_1$};
+    \node[style=obsnode,right=0.5cm of x2] (x3) {$x_3$};
+    \draw[-latex,squiggle] (h) -- (x1);
+    \draw[-latex,squiggle] (h) -- (x2);
+    \draw[-latex,squiggle] (h) -- (x3);
   }
@@
 
+   \draw[-latex] (h2) -- (x1);
    \draw[-latex] (h2) -- (x2);
@@
    \draw[-latex] (h2) -- (x5);
+   }
+}
+
+\newcommand<>{\drawbridge}[1]  {
+  \uncover#2{
+    \point{start-bridge}{#1};
+    % Hidden nodes
+   \node[style=node, scale=0.8] (h1) at (start-bridge) {$h_1$};
+   \node[style=node, scale=0.8, right= 1.0cm of h1] (h2) {$h_2$};
+   \draw[-latex] (h1) -- node[above] (h1h2) {} (h2);
+
+% Observed nodes
+   \node[style=obsnode, scale=0.6, below left=0.3cm of h1] (x1a) {$x_1^a$};
+   \node[style=obsnode, scale=0.6, below=0.3cm of h1] (x1b) {$x_1^b$};
+   \node[style=obsnode, scale=0.6, below=0.3cm of h2] (x2a) {$x_2^a$};
+   \node[style=obsnode, scale=0.6, below right=0.3cm of h2] (x2b) {$x_2^b$};
+   \draw[-latex] (h1) -- (x1a);
+   \draw[-latex] (h1) -- (x1b);
+   \draw[-latex] (h2) -- (x2a);
+   \draw[-latex] (h2) -- (x2b);
+   }
+}
+
+\newcommand<>{\drawubridge}[1]  {
+  \uncover#2{
+    \point{start-ubridge}{#1};
+    % Hidden nodes
+   \node[style=node, scale=0.8] (h1) at (start-ubridge) {$h_1$};
+   \node[style=node, scale=0.8, right= 1.0cm of h1] (h2) {$h_2$};
+   \draw[-] (h1) -- node[above] (h1h2) {} (h2);
+
+% Observed nodes
+   \node[style=obsnode, scale=0.6, below left=0.3cm of h1] (x1a) {$x_1^a$};
+   \node[style=obsnode, scale=0.6, below=0.3cm of h1] (x1b) {$x_1^b$};
+   \node[style=obsnode, scale=0.6, below=0.3cm of h2] (x2a) {$x_2^a$};
+   \node[style=obsnode, scale=0.6, below right=0.3cm of h2] (x2b) {$x_2^b$};
+   \draw[-] (h1) -- (x1a);
+   \draw[-] (h1) -- (x1b);
+   \draw[-] (h2) -- (x2a);
+   \draw[-] (h2) -- (x2b);
    }
@@
   \node at ($(t1) + (1.0cm,-0.3cm)$) {$+$};
-  \tensorfiber{t2}{fill=green!70}{($(t1) + (2.5cm,0.0cm)$)};
+  \tensorfiber{t2}{fill=DarkGreen!70}{($(t1) + (2.5cm,0.0cm)$)};
   \node at ($(t2) + (1.0cm,-0.3cm)$) {$+ \dots + $};
@@
     \node at ($(t1) + (1.0cm,-0.3cm)$) {$+$};
-    \matfiber{t2}{fill=green!70}{($(t1) + (2.5cm,0.0cm)$)};
+    \matfiber{t2}{fill=DarkGreen!70}{($(t1) + (2.5cm,0.0cm)$)};
     \node at ($(t2) + (1.0cm,-0.3cm)$) {$+ \dots + $};
@@
     \node at ($(t1) + (1.0cm,-0.3cm)$) {$+$};
-    \tikzrect{t2}{fill=green!70}{($($(t1) + (2.5cm,0.0cm)$)$)}{0.3}{1};
+    \tikzrect{t2}{fill=DarkGreen!70}{($($(t1) + (2.5cm,0.0cm)$)$)}{0.3}{1};
     \node at ($(t2) + (1.0cm,-0.3cm)$) {$+ \dots + $};

### 2014-06-21 doc/icml2014/icml2014-presentation/diagrams.tex
Updated figures, used an outline slide, and updatd setup

@@
     \point{start-gen}{#1}
-    \node[style=node] (h) at (start-gen) {$h$};
+    \node[style=node] (h) at (start-gen) {$h_1$};
 %    \node[left=0.1cm of h] {$h$};
@@
 
+\newcommand<>{\drawgensquigglescale}[2]{%
+  \uncover#3{
+    \point{start-gen}{#1}
+    \node[style=node,scale=#2] (h) at (start-gen) {$h_1$};
+%    \node[left=0.1cm of h] {$h$};
+    \node[style=obsnode,below=1cm of h,scale=#2] (x2) {$x_2$};
+    \node[style=obsnode,left=0.5cm of x2,scale=#2] (x1) {$x_1$};
+    \node[style=obsnode,right=0.5cm of x2,scale=#2] (x3) {$x_3$};
+    \draw[-latex,squiggle,scale=#2] (h) -- (x1);
+    \draw[-latex,squiggle,scale=#2] (h) -- (x2);
+    \draw[-latex,squiggle,scale=#2] (h) -- (x3);
+  }
+}
+
 \newcommand<>{\drawdisc}[1]{%
@@
     \node[style=node] (h) at (start-disc) {};
-    \node[right=0.1cm of h] {$h$};
+    \node[right=0.1cm of h] {$h_1$};
 
@@
 }
+
+\newcommand{\initialoverview}[1]{
+  \point{start-initoverview}{#1};
+  \node (obs-mom) at (start-initoverview) {$M \eqdef \Pr(\bx)$};
+  \node (params) at ($(obs-mom) + (2cm, -2cm) + (-2cm, -2cm) + (0, -2cm)$) {$\theta$};
+
+  \node[right=0.1cm of obs-mom] {Observed moments};
+  \node[right=0.1cm of params] {Parameters};
+
+  \draw[-latex] (obs-mom) -- (params);
+}
+
+\newcommand{\overview}[1]{
+  \point{start-overview}{#1};
+  \node (obs-mom) at (start-overview) {$M \eqdef \Pr(\bx)$};
+  \node (cond-mom) at ($(obs-mom) + (2cm, -2cm)$) {$\mOpp{1}{1} \eqdef \Pr(\bx | \bh)$};
+  \node (hid-marg) at ($(cond-mom) + (-2cm, -2cm)$) {$Z \eqdef \Pr(\bh)$};
+  \node (params) at ($(hid-marg) + (0, -2cm)$) {$\theta$};
+
+  \node[right=0.1cm of obs-mom] {Observed moments};
+  \node[right=0.1cm of cond-mom] {Conditional moments};
+  \node[right=0.1cm of hid-marg] {Hidden marginals};
+  \node[right=0.1cm of params] {Parameters};
+
+  \draw[-latex] (obs-mom) -- node[right] (obs-edge) {} (cond-mom);
+  \draw[-latex] (cond-mom) -- node[right] (cond-edge) {} (hid-marg);
+  \draw[-latex] (obs-mom) -- node[right] (hid-edge) {} (hid-marg);
+  \draw[-latex] (hid-marg) -- node[right] (params-edge) {}    (params);
+}

### 2014-06-23 doc/icml2014/icml2014-poster/column2.tex
Added poster

\begin{block}{Algorithm Overview}
  \begin{figure}
  \centering
\begin{tikzpicture}[scale=1]
  \overview{(0cm,0cm)};
  \node[right=0.5cm of obs-edge] {\small 1. Solve bottlenecks};
  \node[right=0.5cm of cond-edge] {\objw{10cm}{\small 
  2a. Pseudoinverse \\
  2b. Composite likelihood
  }};
  \node[right=0.5cm of params-edge] {\obj{\small 
  3a. Renormalization \\
  3b. Convex optimization
  }};
\end{tikzpicture}
\end{figure}
\end{block}

\begin{block}{2. Exclusive views and bottlenecked graphs}
  \splitcolumn{%
    \begin{itemize}
      \item A variable $h_i \in S \subseteq \bh$ has an {\bf exclusive
        view} $x_v$ if
      \begin{itemize}
        \item there exists some observed variable $x_{v}$ which is
          conditionally independent of the others $S \backslash \{ h_i \}$
          given $h_i$.
        \item the conditional moments $\mOpp{v}{i} \eqdef \Pr(x_{v} \mid
          h_i)$ are full column rank $k$ and can be recovered.
      \end{itemize}
      \item {\bf Theorem:} Every bottlenecked bidependent set (think
        a clique) has exclusive views.
    \end{itemize}
  }{%
  \begin{figure}
    \centering
      \includegraphics[width=0.95\textwidth,height=5cm,keepaspectratio]{figures/exclusive-views.pdf}
%      \caption{Exclusive views}
  \end{figure}

    \begin{itemize}
    \item Given {\em exclusive views}, $\Pr(x \given h)$,
      learning cliques is solving a linear equation.
      \begin{align*}
        M &= Z(\mOpp{1}{1}, \cdots, \mOpp{m}{m}) \\
        Z &= M(\mOppi{1}{1}, \cdots, \mOppi{m}{m}).
      \end{align*}
    \end{itemize}
  }
\end{block}
\begin{block}{2b. Composite likelihood for hidden marginals}
  \splitcolumn{

  \begin{figure}
    \centering
    \begin{tikzpicture}
      \point{stuff}{(2cm,2.5cm)};
      \drawhmm{(stuff)};
      \begin{pgfonlayer}{background}
      \draw[draw=black,fill=green!70,rounded corners,line width=1pt, dotted] 
                      ($(h1.west) + (180:0.3cm)$) -- 
                      ($(h1.north) + (90:0.3cm)$) -- 
                      ($(h2.north) + (90:0.3cm)$) -- 
                      ($(h2.east) + (0:0.3cm)$) -- 
                      ($(x2.east) + (0:0.3cm)$) -- 
                      ($(x2.south) + (-90:0.3cm)$) -- 
                      ($(x1.south) + (-90:0.3cm)$) -- 
                      ($(x1.west) + (180:0.3cm)$) -- 
                      cycle;
      \end{pgfonlayer}
    \end{tikzpicture}
  \end{figure}

  \begin{itemize}
    \item By fixing the conditional mom., the {\em composite
      likelihood} of a subset of variables is {\em convex}.
    \item Estimation with composite likelihoods is consistent
      (\cite{lindsay88composite}).
    \item Asymptotically, the composite likelihood estimator is {\em more
      efficient}.
  \end{itemize}

  }{%
  \begin{figure}
    \centering
    \includegraphics[width=\textwidth,height=6cm,keepaspectratio]{figures/asymp-k2d5-bordered.pdf}
    \caption{Comparison of the pseudoinverse and composite likelihood
    estimators on a HMM with $k=2$ hidden and $d=5$ observed values.
    }
  \end{figure}
  }
\end{block}

\begin{block}{3. Recovering parameters}
  \splitcolumn{%
  \begin{itemize}
    \item In directed models: CPTs can be recovered by normalization. 
    \item In undirected mdoels: $Z_{12}$ can be used to estimate {\em
      feature marginals} and can be solved by {\em convex optimization},
      similar to supervised learning.
    \item Can be used with pseudo-likelihoods to efficiently
      learn high-treewidth models.
  \end{itemize}
  }{% 
    \small

      \begin{align*}
        \Pr(h_2 \given h_1) &= \frac{\Pr(h_1, h_2)}{\sum_{h_2} \Pr(h_1, h_2)}.
        \end{align*}
\vfill
      \begin{align*}
          \sL(\theta) &= \theta^\top \mu_\sC + A(\theta) \\
          \mu_\sC &= 
              \sum_{\sC \in \sG} 
              \E_{(\bx,\bh) \sim \sD_\text{sup}}[\phi(\bx_\sC,\bh_\sC)] \\
                  &= \sum_{\bx_\sC, \bh_\sC} \underbrace{\mathmg{\Pr(\bx_\sC \given \bh_\sC)}}_{\textmg{cond. moments}} 
          \underbrace{\mathmb{\Pr(\bh_\sC)}}_{\textmb{hid. marg.}} \phi(\bx_\sC,\bh_\sC).
          \end{align*}
  }
\end{block}

\begin{block}{Conclusion}
  \splitcolumn{
  \begin{itemize}
    \item We provide an algorithm for {\bf any bottlenecked
      graphical model}, including {\em high treewidth} and {\em undirected graphical models}.
    \item Combining moment estimates with likelihood objectives
      provides a new perspective balancing efficiency and accuracy.
  \end{itemize}
  }{
  \centering
  \includegraphics[width=\textwidth,height=5cm,keepaspectratio]{figures/mrf.pdf}
  }
\end{block}
